### 2015-06-18 var_select.rmd
started var_select markdown file

---
title: "Variable Selection"
author: "Amy Cook"
date: "Thursday, June 18, 2015"
output: word_document
---

```{r, echo=FALSE, include=FALSE}
setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
library('car')
library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
```

```{r load_data, echo=FALSE}
all6a<- read.csv('all6a.csv')[,-1]
all6a$Start.Date<- as.Date(all6a$Start.Date)
all7<- read.csv('all7.csv')[,-1]
all7$Start.Date<- as.Date(all7$Start.Date)
```

Check for normality of return variable return per dollar
Outliers were deleted first

```{r, echo=FALSE}
# try deleting very high return.pdol values greater than 3
test<- all7 %>% filter(return.pdol <=3 & return.pdol> -2)
qqPlot(test$return.pdol)
qqPlot(log(test$return.pdol+.97))
#have a look at the shape of return.pdol as bar graph
ggplot(test, aes(x=return.pdol)) + geom_histogram(aes(y=..density..), binwidth=.1)
ggplot(test, aes(x=log(test$return.pdol+.97))) + geom_histogram(aes(y=..density..), binwidth=.1)
```

definietly don't want to use log - worse.
Let's look at standard deviation, mean, median, skewness, kurtosis
```{r}
sd(test$return.pdol)
mean(test$return.pdol)
median(test$return.pdol)
#mode
names(sort(-table(test$return.pdol)))[1]
skewness(test$return.pdol)
skewness(log(test$return.pdol+.97))
kurtosis(test$return.pdol)
```

skewness only 0.92!! for unlogged
skewness is -2.6 for logged data

kurtosis equals 2.85 - very peakedy - ie not flat. normal distribution is 0.

Create new data set that has outliers deleted, all7a

```{r}
### delete outliers! and variables that aren't by milestone
all7a<- all7 %>% filter(return.pdol <=3 & return.pdol> -2)
all7a<- all7 %>% select(-Paperless,-Innovation, -JobInvCount,-job.first.inv.email.Sent, 
                        -Total.Costs..AUD., -Stage, -Folders, -Total.Fee, -Disburse, -Subcon.fee,
                        -Job.Size, -Tot.Invoiced, -charge, -cost, -Dis.subcon, -hours,-profit,-balance,
                        -Role)

```

# Variable Selection

It would be good to be able to narrow down the variables to the correlated or strongly affecting ones. This improves accuracy and speed of the final machine learning algoriths.

Anova and Linear regression will give us an idea of variable importance. Look at coefficients

In order to get Anova and linear regression to work on this data set, start with the variables that are complete or almost complete. Then can add additional variables one by one. The function below does this automatically.

For lm, the output is ordered from abs(highest coefficient)

```{r, echo=TRUE}
BT.lin<- function(type= 'lm', data= all7a, add.var = 'Job.Type.Primary') {
        formula = paste("return.pdol ~ inv.mlsto + Discipline + client.count + Business +
                              Biz.size + Biz.type + Year + Num.days +
                              no.users + pc.contracttech + client.neginv + client.numinv + client.totinv +
                              pc.director + pc.midtech + pc.midpro + pc.gradpro + pc.seniortech + pc.seniorpro +
                              pc.pro + mean.peeps + hours.perday + num.inv + mean.inv + num.neginv + client.meaninv +
                              code.director + ProjEng.Pos", add.var, sep="+")
        if (type== 'aov') {
            model= aov(as.formula(formula)              
                      ,data= data)
            return(summary(model)) 
        }
        else{
                model=lm(as.formula(formula)                     
                        ,data= data)
                df<- as.data.frame(summary(model)$coefficients)
                df$var<-rownames(summary(model)$coefficients)
                colnames(df)[names(df) %in% 'Pr(>|t|)']<-'Pval'
                return(df %>% arrange(-abs(Estimate)))

        }

}

BT.lin(type= 'lm', data= all7a[1:1313,], add.var = 'Job.Type.Primary') %>% slice(1:20)
BT.lin(type= 'lm', data= all7a, add.var = 'Job.Type.Primary') %>% slice(1:10)
BT.lin(type= 'lm', data= all7a[1314:2283,], add.var = 'Job.Type.Primary') %>% slice(1:10)
BT.lin(type= 'aov', data= all7a, add.var = 'Job.Type.Primary')
BT.lin(type= 'aov', data= all7a[1314:2283,], add.var = 'Job.Type.Primary')

```

Most influential variables are:
* Business type
* director code
* Discipline
* Biz.type

Excluded variables include:
* Post.Code
* Billing.Type
* Job.Source
* Job.Detail.Primary
* Job.Detail.Secondary
* Job.Type.Primary
* Job.Type.Secondary
* State
* Type
* no.employees
* contact.count
* JD.Primary
* JD.Second
* dist
* timespan (won't know to begin with)
* Num.disc
* Inv.freq
* client.invfreq

### 2015-06-18 1st_stats.R

@@ -230,11 +230,11 @@ matrix4<- spm(~ inv.vs.cost + pc.tech + mean.peeps + hours.perday
 #lets write a function for running these anova and lm models
 
 BT.lin<- function(type= 'lm', data= all7a, add.var = 'Job.Type.Primary') {
-        formula = paste("inv.vs.cost~ inv.mlsto + Discipline + client.count + Business+
-                              Biz.size + Biz.type + 
-                              no.users + pc.contracttech +
-                              pc.director + pc.midtech + pc.midpro+ pc.gradpro+ pc.seniortech+ pc.seniorpro+
-                              pc.pro + mean.peeps + hours.perday + 
+        formula = paste("return.pdol ~ inv.mlsto + Discipline + client.count + Business +
+                              Biz.size + Biz.type + Year + Num.days +
+                              no.users + pc.contracttech + client.neginv + client.numinv + client.totinv +
+                              pc.director + pc.midtech + pc.midpro + pc.gradpro + pc.seniortech + pc.seniorpro +
+                              pc.pro + mean.peeps + hours.perday + num.inv + mean.inv + num.neginv + client.meaninv +
                               code.director + ProjEng.Pos", add.var, sep="+")
         if (type== 'aov') {
             model= aov(as.formula(formula)              

### 2015-06-19 var_select.rmd
var_select finished numeric vars

@@
 ```{r, echo=FALSE, include=FALSE}
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
+library('knitr')
+# library('knitr', lib = 'C:/Progra~1/R/R-3.1.2/library')
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
+# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+# library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
+# library('car', lib = 'C:/Progra~1/R/R-3.2.0/library')
+# library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
+# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.0/library')
+
+library('ggplot2')
+library("dplyr")
+library("plyr")
+library('magrittr')
+library('reshape2')
+library("rpart")
 library('car')
-library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
+library('e1071')
+library('corrgram')
 ```
@@
 
-```{r, echo=FALSE}
+```{r}
 # try deleting very high return.pdol values greater than 3
@@
 ```{r}
+
 ### delete outliers! and variables that aren't by milestone
@@
 
-```{r, echo=TRUE}
+```{r, echo=FALSE}
 BT.lin<- function(type= 'lm', data= all7a, add.var = 'Job.Type.Primary') {
@@
 }
+```
 
-BT.lin(type= 'lm', data= all7a[1:1313,], add.var = 'Job.Type.Primary') %>% slice(1:20)
-BT.lin(type= 'lm', data= all7a, add.var = 'Job.Type.Primary') %>% slice(1:10)
+```{r}
+
+# BT.lin(type= 'lm', data= all7a[1:1313,], add.var = 'Job.Type.Primary') %>% slice(1:20)
+BT.lin(type= 'lm', data= all7a, add.var = 'Job.Type.Primary') %>% slice(1:20)
 BT.lin(type= 'lm', data= all7a[1314:2283,], add.var = 'Job.Type.Primary') %>% slice(1:10)
@@
 Most influential variables are:
+
 * Business type
@@
 Excluded variables include:
+
 * Post.Code
@@
 * client.invfreq
+
+# Correlation Analysis - Numeric variables with 'return.pdol' variable
+
+First run Corrgram  - corrgram function
+Run as type 'spearman' - non linear
+
+```{r, echo=FALSE}
+
+#### CORRELATION ANALYSIS
+
+# str(all7a)
+num.all<- all7a %>% select(client.count, no.employees, dist, charge.ph, no.users, pc.contracttech,
+                           pc.director, pc.gradpro, pc.midpro, pc.midtech, pc.seniorpro, pc.seniortech, 
+                           pc.unknown, pc.pro, pc.tech, Year, balance.mlsto,
+                           timespan, Num.disc, Num.days, mean.peeps, hours.perday, hrs.mlsto, cost.mlsto, 
+                           dis.sc.mlsto, 
+                           inv.mlsto, return.pdol,
+                           num.inv, mean.inv, Inv.freq, num.neginv, client.meaninv, client.invfreq, client.neginv, 
+                           client.numinv,
+                           client.totinv)
+corrgram(num.all %>% select(client.count, charge.ph, no.users, pc.director, pc.gradpro, pc.midpro, 
+                            pc.midtech, pc.contracttech, return.pdol,
+                            pc.seniorpro, pc.seniortech, balance.mlsto,
+                            pc.unknown, pc.pro, pc.tech, Year, Num.days, mean.peeps, hours.perday, hrs.mlsto, 
+                            cost.mlsto,
+                            dis.sc.mlsto, inv.mlsto, num.inv, mean.inv, client.meaninv, client.totinv,
+                            timespan) %>% filter(complete.cases(.)), order=T, 
+         lower.panel = panel.shade,
+         upper.panel = panel.pie,
+         text.panel = panel.txt,
+         cor.method = 'spearman')
+# summary(num.all)
+# num.all[rowSums(is.na(num.all))>0,]
+```
+
+Now output cor.test results for all numeric variables using both pearson (linear) and kendall (non-linear) to compare.
+Kendall does not have problems with ties
+
+```{r, echo=FALSE}
+
+#correlation by each variable against return.perdol individually - numeric variables only
+
+cor.all<- c("client.count", "no.users", "pc.director", "pc.gradpro", "pc.midpro", 
+            "pc.midtech", "pc.contracttech", "return.pdol","Inv.freq", 'num.neginv','client.invfreq','client.neginv',
+            "pc.seniorpro", "pc.seniortech", "no.employees", "dist", "Num.disc",'client.numinv',
+            "pc.unknown", "pc.pro", "pc.tech", "Year", "Num.days", "mean.peeps", "hours.perday", "hrs.mlsto", 
+            "cost.mlsto",
+            "dis.sc.mlsto", "inv.mlsto", "num.inv", "mean.inv", "client.meaninv", "client.totinv",
+            "timespan")
+
+summ<- data.frame("test"=NULL,
+                  'var' = NULL,
+                  'corr' = NULL,
+                  'p.val' = NULL)
+
+for(i in 1:length(cor.all)) {
+
+        a = cor.test(all7a$return.pdol, all7a[,cor.all[i]], method = c('pearson'))
+
+        b = cor.test(all7a$return.pdol, all7a[,cor.all[i]], method = c('kendall'))
+
+        inter.summ = data.frame('test'= c('pearson','kendall'),
+                                'var' = c(cor.all[i], cor.all[i]),
+                                'corr' = c(a$estimate, b$estimate),
+                                'p.val' = c(a$p.val, b$p.val))
+
+        summ = rbind(summ,inter.summ)
+}
+
+summ$p.val <- format(summ$p.val, scientific=TRUE, digits = 3)
+summ$p.val <- as.numeric(summ$p.val)
+summ$corr <- round(summ$corr, 3)
+
+```
+
+Output ordered by p value
+```{r, echo=TRUE}
+
+summ %>% arrange(test, p.val)
+
+```
+
+Output ordered by correlation value - only 'significant' results to 5%
+
+```{r}
+
+summ %>% filter(p.val<= 0.05) %>% arrange(test, -abs(corr))
+
+```
+
+From this list it can be concluded what the most influential variables are. 
+For the purposes of return.pdol prediction, cost.mlsto, hrs.mlsto, Num.days, dis.sc.mlsto,
+num.neginv, Year (won't help in the future) cannot be known.
+
+This leaves the following important variables:
+
+```{r, echo=FALSE}
+kable(
+        summ %>% filter(p.val<= 0.05, test== 'kendall') %>% arrange(test, -abs(corr)) %>% 
+                select(var, p.val, corr) %>%
+                filter(var!= 'cost.mlsto', var!= 'hrs.mlsto', var!= 'Num.days', var!= 'dis.sc.mlsto',
+                       var!= 'num.neginv',var!= 'Year', var!= 'return.pdol')
+         )
+
+```
+
+# Strength of association with Categorical variables

### 2015-06-19 var_select.rmd
added cforest

@@
 ```{r, echo=FALSE, include=FALSE}
-library('knitr')
-# library('knitr', lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library('knitr')
+library('knitr', lib = 'C:/Progra~1/R/R-3.1.2/library')
 # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
 opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
-# library('car', lib = 'C:/Progra~1/R/R-3.2.0/library')
-# library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
-# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.0/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.0/library')
+library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
+library('corrgram', lib = 'C:/Progra~1/R/R-3.2.0/library')
+library('party', lib = 'C:/Progra~1/R/R-3.2.0/library')
 
-library('ggplot2')
-library("dplyr")
-library("plyr")
-library('magrittr')
-library('reshape2')
-library("rpart")
-library('car')
-library('e1071')
-library('corrgram')
+# library('ggplot2')
+# library("dplyr")
+# library("plyr")
+# library('magrittr')
+# library('reshape2')
+# library("rpart")
+# library('car')
+# library('e1071')
+# library('corrgram')
+# library('party')
 ```
@@
 # Strength of association with Categorical variables
+
+Use cforest from party package - (Strobl etc, 2007)
+- must be subsampling without replacement
+- subsampling size 0.632 times recommended because about 63.2% of data end up in bootstrap sample
+
+```{r}
+
+all7b<- all7a %>% select(return.pdol, no.users,timespan,mean.peeps,hours.perday,Num.disc,pc.pro,
+                num.inv, pc.gradpro , inv.mlsto , client.numinv, client.totinv, client.meaninv , 
+                client.neginv , mean.inv , pc.director , code.client , Post.Code , Billing.Type , 
+                Discipline , Job.Detail.Primary , Job.Detail.Secondary , Job.Type.Primary , 
+                Job.Type.Secondary , State , Type , Business , Biz.size , Biz.type , JD.Primary ,
+                JD.Second , code.director , code.ProjEng , ProjEng.Pos , code.Eng2 , Eng2.Pos)
+
+var.imp<- cforest(return.pdol~.,
+        data= all7b
+        )
+
+```

### 2015-06-22 var_select.rmd
correct pc.pro by most

--- 2015-06-19 var_select.rmd
@@
 ```{r, echo=FALSE, include=FALSE}
-library('knitr')
+
 # library('knitr', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
 # library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
@@
 # library('corrgram', lib = 'C:/Progra~1/R/R-3.2.0/library')
+# library('party', lib = 'C:/Progra~1/R/R-3.2.0/library')
 
+library('knitr')
 library('ggplot2')
@@
 library('corrgram')
+library('party')
+
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
 ```
@@
                            inv.mlsto, return.pdol,
-                           num.inv, mean.inv, Inv.freq, num.neginv, client.meaninv, client.invfreq, client.neginv, 
+                           num.inv, mean.inv, Inv.freq, num.neginv, client.meaninv, client.invfreq, 
+                           client.neginv, 
                            client.numinv,
@@
 # Strength of association with Categorical variables
+
+Use cforest from party package - (Strobl etc, 2007)
+
+* must be subsampling without replacement
+* subsampling size 0.632 times recommended because about 63.2% of data end up in bootstrap sample
+* cforest defauly is unbiased - for unbiased variable importance as per Strobl etc 2007
+* since variables are potentially correlated to one another, need to set conditional permutation importance varimp(obj, conditional=TRUE)
+    * This iterates each variable against correlated variables in a controlled way. Ie the target variable is assessed against subgroups of the data where a correlated variable is constant.
+
+For cforest, NA values are not handled. Therefore create a 'core' dataframe consisting of the variables which have very low NA entries (say max 10) and delete all NA rows or impute.
+
+State- impute QLD
+Business - 22 NA values - impute 'unknown'
+Biz.size - 38 NA's - find median, impute?
+Biz.type - 18 NA's impute???
+charge.ph - delete
+charge.pc - delete
+Start.Date - delete
+End.Date - delete
+no.users - 4 NA values - impute
+pc.usersss - 7 NA's - what to do. 0 for all? averages for all?
+Num.disc - 84 NA's - impute 1?
+num.inv and mean.inv- 1 NA?
+
+```{r core_creation}
+#impute NA values in some variables to be in core, call this all7b
+all7b<- all7a
+all7b[is.na(all7b$State),]$State <- 'QLD'
+#RSL Care client - now have details
+all7b[all7b$code.client %in% 'C2835',]$no.employees <- 1001
+all7b[all7b$code.client %in% 'C2835',]$Biz.type <- 'NFP'
+all7b[all7b$code.client %in% 'C2835',]$Business <- 'developer/real estate'
+all7b[all7b$code.client %in% 'C2758',]$Biz.size <- 'national'
+all7b[all7b$code.client %in% 'C2758',]$no.employees <- 2
+all7b[all7b$code.client %in% 'C2758',]$Biz.size <- 'local'
+all7b[all7b$code.client %in% 'C2758',]$Business <- 'business'
+all7b[all7b$code.client %in% 'C2326',]$Business <- 'health'
+all7b[all7b$code.client %in% 'C2338',]$Biz.size <- 'local'
+all7b[all7b$code.client %in% 'C2338',]$Business <- 'developer/real estate'
+all7b[all7b$code.client %in% 'C2338',]$Biz.type <- 'private'
+all7b[all7b$code.client %in% 'C2882',]$Biz.size <- 'local'
+all7b[all7b$code.client %in% 'C2882',]$Business <- 'builder'
+all7b[all7b$code.client %in% 'C2882',]$Biz.type <- 'private'
+all7b[all7b$code.client %in% 'C2882',]$no.employees <- 3
+#all remaining NA Business will be 'unknown'
+all7b$Business<- as.character(all7b$Business)
+all7b[is.na(all7b$Business),]$Business <- 'unknown'
+all7b$Business<- as.factor(all7b$Business)
+
+#impute Biz.size to be 'local' by default - this is most popular category
+all7b[is.na(all7b$Biz.size),]$Biz.size <- 'local'
+#impute Biz.type to be 'private' by default - this is most popular category
+#client which is 'department of national parks' shall be 'gov'
+all7b[is.na(all7b$Biz.type),]$Biz.type <- 'private'
+all7b[all7b$code.client %in% 'C2503',]$Biz.type <- 'gov'
+#delete charge.ph and charge.pc, Start.Date and End.Date
+all7b<- all7b %>% select(-charge.ph, -charge.pc, -Start.Date, -End.Date)
+#impute NA no.users as 2 = median and mode
+all7b[is.na(all7b$no.users),]$no.users <- 2
+#impute NA num.disc as 1 - overwhelmingly the most common
+all7b[is.na(all7b$Num.disc),]$Num.disc <- 1
+#num.inv and mean.inv both have a single NA value
+# this consisted of single large -ve invoice. I think we should delete this case 
+all7b<- all7b %>% filter(!is.na(num.inv))
+# write.csv(all7b,'all7b.csv')
+
+#create core data frame - variables with very few NA's and delete these NA's
+
+core<- all7b %>% select(code.client , Discipline ,  client.count , inv.mlsto , Business ,
+        Biz.size , Biz.type , no.users , Year , timespan , Num.disc , Num.days , mean.peeps,
+        pc.contracttech , hours.perday , profit.mlsto , balance.mlsto , hrs.mlsto , 
+        cost.mlsto , dis.sc.mlsto , inv.mlsto , return.pdol , code.contact ,
+        pc.director , pc.midtech , pc.midpro , pc.gradpro , pc.seniortech , pc.seniorpro ,
+        pc.pro , code.director , code.ProjEng , ProjEng.Pos , num.inv , mean.inv , num.neginv , 
+        client.meaninv , client.neginv , client.numinv , client.totinv)
+
+var.imp<- cforest(return.pdol~.,
+        data= all7b
+        )
+
+```

### 2015-06-22 1st_stats.R

@@ -14,9 +14,9 @@ all6a$Start.Date<- as.Date(all6a$Start.Date)
 #sort by hours first
 all6a<- all6a %>% arrange(mlsto, -hours)
 dup<- all6a[,names(all6a) %in% c('mlsto')]
-all7<-all6a[!(all6a$Tot.Invoiced==0 | duplicated(dup)),]
+all7<-all6a[!(all6a$inv.mlsto==0 | duplicated(dup)),]
 
-# write.csv(all7, 'all7.csv')
+write.csv(all7, 'all7.csv')
 all7<- read.csv('all7.csv')[,-1]
 all7$Start.Date<- as.Date(all7$Start.Date)
 

### 2015-06-22 de_identify.R

@@ -181,7 +181,8 @@ write.csv(all5a, 'all6.csv')
 #merge new coded invoicing engineered variables from invoices_eng.csv'
 all6<- read.csv('all6.csv')[,-1]
 inv.eng<- read.csv('invoices_eng.csv')[,-1]
-all6a<- merge(all6, inv.eng %>% select(code.jobnum, num.inv, mean.inv, Inv.freq, num.neginv), by='code.jobnum', all.x=T)
+all6a<- merge(all6, inv.eng %>% select(mlsto, num.inv, mean.inv, Inv.freq, num.neginv) %>% unique, 
+              by='mlsto', all.x=T)
 all6a<- merge(all6a, inv.eng %>% select(code.client, client.meaninv, client.invfreq, client.neginv, client.numinv, client.totinv) %>% unique,
               by='code.client', all.x=T)
 test<- merge(all6 %>% select(mlsto), inv.eng %>% select(mlsto), by= 'mlsto', all.x=TRUE, all.y=TRUE)

### 2015-06-22 hours_cleaning.R

@@ -211,11 +211,29 @@ hours1<- read.csv('hours_clean.csv')[,-1]
 
 ############ ENGINEER MORE VARIABLES  #####################
 
-# number of diff people who worked on a job
-V1<- hours1[!(hours1$user.code %in% c('disburse', 'Admin')), names(hours1) %in% c('Job.num', 'user.code')]
-V1 <- V1 %>% arrange(Job.num) %>%
+#calculate percentage of hours by each position for each MILESTONE, not JOB NUMBER - 2015-06-22
+#merge mlsto by Job.Number - use file all5a
+
+#create new column for milestone prefix
+hours1<- hours1 %>% arrange(Job.num)
+temp1<- hours1[1:77588,]
+#string split to create another column with the first part of a job number. For the first half of jobs just the first bit indicates it could be invoiced together ie 2002.001 by itself
+temp1$mlsto<-sapply(temp1$Job.num,FUN=function(x){
+        substr(x,1,8)})
+#do the same with second half of data
+temp2<- hours1[77589:120508,]
+temp2$mlsto<-sapply(temp2$Job.num,FUN=function(x){
+        substr(x,1,10)})
+temp2$mlsto<- ifelse(temp2$mlsto == '2011.072.3', temp2$Job.num, temp2$mlsto)
+#rbind temp1 and temp2 back together to make temp
+hours1<- rbind(temp1, temp2)
+hours1$mlsto<- as.factor(hours1$mlsto)
+
+# number of diff people who worked on a milestone
+V1<- hours1[!(hours1$user.code %in% c('disburse', 'Admin')), names(hours1) %in% c('mlsto', 'user.code')]
+V1 <- V1 %>% arrange(mlsto) %>%
         unique() 
-V1a<- ddply(V1, .(Job.num), nrow)
+V1a<- ddply(V1, .(mlsto), nrow)
 #rename columns
 colnames(V1a)[names(V1a) %in% 'V1']<-'no.users'
 V1a %>% arrange(-no.users) %>% head()
@@ -225,11 +243,11 @@ V1a %>% arrange(-no.users) %>% head()
 # variables: % hours by each position
 
 V2<- hours1[!(hours1$user.code %in% c('disburse', 'Admin')), 
[…]

### 2015-06-24 var_select.rmd
revised core df to exclude client and contact codes

@@
 
-# library('knitr', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
-# library('car', lib = 'C:/Progra~1/R/R-3.2.0/library')
-# library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
-# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.0/library')
-# library('party', lib = 'C:/Progra~1/R/R-3.2.0/library')
+library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
 
-library('knitr')
-library('ggplot2')
-library("dplyr")
-library("plyr")
-library('magrittr')
-library('reshape2')
-library("rpart")
-library('car')
-library('e1071')
-library('corrgram')
-library('party')
+# library('knitr')
+# library('ggplot2')
+# library("dplyr")
+# library("plyr")
+# library('magrittr')
+# library('reshape2')
+# library("rpart")
+# library('car')
+# library('e1071')
+# library('corrgram')
+# library('party')
 
 # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
 opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
@@
 
+# write.csv(all7a,'all7a.csv')
+all7a<- read.csv('all7a.csv')
 ```
@@
 
-```{r core_creation}
+```{r core_creation, include=FALSE}
 #impute NA values in some variables to be in core, call this all7b
@@
 all7b<- all7b %>% filter(!is.na(num.inv))
+# 7 NA cases with significant dis.sc.mlsto - ie done by contractor 
+pc<- c('pc.contracttech','pc.director', 'pc.midtech', 'pc.midpro', 'pc.gradpro', 'pc.seniortech', 'pc.seniorpro', 'pc.pro', 'pc.tech')
+
+for (i in 1:length(pc)){
+        all7b[is.na(all7b[,pc[i]]),][,pc[i]] <-0
+}
+
+#change unknown Num.disc to be 1
+all7b[is.na(all7b$Num.disc),]$Num.disc <- 1
 # write.csv(all7b,'all7b.csv')
+
+```
+
+Start running cforests
+Very slow, start with 100 rows only, using only core variables
+
+```{r, echo=FALSE}
+all7b<- read.csv('all7b.csv')
 
@@
 
-var.imp<- cforest(return.pdol~.,
-        data= all7b
+mini.core2 <- all7b %>% select(code.client , Discipline ,  client.count , inv.mlsto , Business ,
+        Biz.size , Biz.type , no.users , timespan, Num.disc, 
+        return.pdol, code.contact,
+        pc.director,
+        pc.pro , code.director , ProjEng.Pos ,
+        client.meaninv , client.neginv , client.numinv , client.totinv)
+
+mini.core3 <- all7b %>%
+        select(Discipline , inv.mlsto , Business,
+        Biz.size , Biz.type , no.users , timespan, Num.disc, 
+        return.pdol, pc.director,
+        pc.pro , code.director , ProjEng.Pos,
+        client.meaninv , client.neginv , client.numinv , client.totinv)
+
+```
+
+Try to make all cforests random selection of 200 rows - this time post 2009
+
+```{r, echo=FALSE}
+
+Try running a cforest over night with full dataset - don't include 
+```
+
+```{r, echo=FALSE}
+
+system.time(
+all.cfor2<- cforest(return.pdol~.,
+        data= mini.core3
+        )
 )
 
+var.imp4<- varimp(all.cfor) %>% as.data.frame
+colnames(var.imp4)[names(var.imp4) %in% '.']<-'var.imp'
+var.imp4$var <- rownames(var.imp4)
+#reorder levels for bar plot
+var.imp4 <- within(var.imp4, 
+                    var <- factor(var,levels=
+                                            var.imp4[order(-abs(var.imp4$var.imp)),]$var))
+
+q<- ggplot(data=var.imp4, aes(x=var, y=var.imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
+q + geom_bar(stat='identity', position = 'dodge')
+
+save(all.cfor,file= 'C:/Users/n9232371/Documents/github/consult/object/all_cfor.RData')
+load('C:/Users/n9232371/Documents/github/consult/object/all_cfor.RData')
 ```
+
+From analysing all jobs including 2009 and beyond, the most important variables are:
+`levels(var.imp4$var)[1:10] %>% as.data.frame`
+
+The least important are 
+`levels(var.imp4$var)[11:19]`
+
+Let's compare this to
+
+* all data from 2002 to 2014 - currently running in another R window
+* ANOVA - all data because fast
+* random forest
+
+## Anova:
+
+```{r anova_varimp, echo=FALSE}
+#use mini.core2- all years
+system.time(
+        aov.all<- aov(return.pdol~.,
+                      data= mini.core3)
+        )
+
+summary(aov.all)
+```
+
+Therefore most significant variables are:
+
+* no.users - number of people on the job - <<.1%
+* code.director - <<.1%
+* timespan - 4%
+* Num.disc - 5.6%
+* Discipline - 5.65%
+
+worst: Business, ProjEng.Pos, client.numinv,
+
+Contrasts to cforest results include
+
+* code.client pval=1 - this is likely because code.contact and code.client are highly correlated
+* inv.mlsto pval= .721 - surprising
+* client.numinv pval = 0.923 - surprisingly different
+
+# random forests
+
+First optimise parameters mtry and ntree - exclude code.client and code.contact for now as they have way too many categories
+
+```{r randomForest, echo=FALSE}
+#random forest can have max 53 categories in a variable
+#code.client has the most categories - 664
+
+# find optimal mtry
+
+#create train/test sets
+
+train_ind <- sample(seq_len(2355), size = 1570)
+
+train <- mini.core4[train_ind, ]
+test <- mini.core4[-train_ind, ]
+oob.err<-double(13)
+test.err<-double(13)
+for(mtry in 1:13){
+        fit<-randomForest(return.pdol ~., data=train, mtry=mtry, ntree=400)
+        oob.err[mtry]<-fit$mse[400]
+        test.err[mtry]<- mean((predict(fit,test)-test$return.pdol)^2)
+        cat(mtry," ")
+}
+
+qplot(1:mtry,test.err,geom=c("point","line"),color="pink")+
+        geom_line(aes(1:mtry,oob.err,colour="blue"))+
+        geom_point(aes(1:mtry,oob.err, colour="blue"))
+
+#COMPLETELY FLAT - MAY AS WELL USE MTRY = 5????? NOW VARY NUMBER OF TREES
+
+trees<- c(300,500,700,1000,1500)
+oob.err<-double(5)
+test.err<-double(5)
+for(ntree in 1:5){
+        fit<-randomForest(return.pdol ~., data=train, mtry=5, ntree=trees[ntree])
+        oob.err[ntree]<-fit$mse[trees[ntree]]
+        test.err[ntree]<- mean((predict(fit,test)-test$return.pdol)^2)
+        cat(ntree," ")
+}
+
+qplot(1:ntree,test.err,geom=c("point","line"),color="pink")+
+        geom_line(aes(1:ntree,oob.err,colour="blue"))+
+        geom_point(aes(1:ntree,oob.err, colour="blue"))
+
+#again, doesn't matter!! line is dead flat. i'm not having much confidence in the predictiveness of this model. 
+```
+
+Extract important variables from randomForest loop
+
+423 of the 664 client companies and 485 of the 670 client contacts appear ONLY once in the entire data set.
+While they may be great predictors what do we do in the majority of cases where the contact is new!?
+
+```{r, echo=FALSE}
+set.seed(47)
+
+#try running without code.client or code.contact
+multi.forest1<-function(data=mini.core3, nforest=1:10){
+        sample=double(50)
+        imp = NULL
+        for (i in 1:length(nforest)){
+
+                #run tree
+                assign(paste("Forest",i,sep=""),randomForest(return.pdol~.,
+                                                             data=data,
+                                                             ntree=400,mtry=5, importance=TRUE),
+                       envir=.GlobalEnv)
+
+                #put together importance rank for each variable
+
+                rand.imp= importance(paste("Forest",i,sep="") %>% get, type=1, scale=FALSE) %>% as.data.frame
+                colnames(rand.imp)[names(rand.imp) %in% '%IncMSE']= 'IncMSE'
+                rand.imp$var= rownames(rand.imp)
+                rand.imp= rand.imp %>% arrange(-abs(IncMSE))
+                rand.imp$rank= 1:nrow(rand.imp)
+
+                #rbind all variables and ranks on top of one another for each Forest
+
+                imp = rbind(imp, rand.imp %>% select(var,rank))
+
+                cat(i," ")
+
+        }
+
+        #output average rank of each variable
+        rank = ddply(imp, .(var), summarise, m.rank =mean(rank)) %>% arrange(m.rank)
+        return(rank)
+
+}
+
+rand.rank<- multi.forest1(data=mini.core2 %>% select(-code.contact, -code.client), nforest=1:10)
+
+kable(rand.rank)
+
+```
+
+See how variable selection goes with random forest first, then probably just run cforest again without code.client and contact.code

### 2015-06-26 var_select.rmd
cforests

@@
 mini.core3 <- all7b %>%
-        select(Discipline , inv.mlsto , Business,
+        select(Discipline , inv.mlsto , Business, client.count,
         Biz.size , Biz.type , no.users , timespan, Num.disc, 
@@
 
-```{r, echo=FALSE}
-
 Try running a cforest over night with full dataset - don't include
-```
 
@@
 
-var.imp4<- varimp(all.cfor) %>% as.data.frame
-colnames(var.imp4)[names(var.imp4) %in% '.']<-'var.imp'
-var.imp4$var <- rownames(var.imp4)
+var.imp<- varimp(all.cfor2) %>% as.data.frame
+colnames(var.imp)[names(var.imp) %in% '.']<-'imp'
+var.imp$var <- rownames(var.imp)
 #reorder levels for bar plot
-var.imp4 <- within(var.imp4, 
+var.imp <- within(var.imp, 
                     var <- factor(var,levels=
-                                            var.imp4[order(-abs(var.imp4$var.imp)),]$var))
+                                            var.imp[order(-abs(var.imp$imp)),]$var))
 
-q<- ggplot(data=var.imp4, aes(x=var, y=var.imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
-q + geom_bar(stat='identity', position = 'dodge')
+q<- ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
+q + geom_bar(stat='identity')
 
-save(all.cfor,file= 'C:/Users/n9232371/Documents/github/consult/object/all_cfor.RData')
-load('C:/Users/n9232371/Documents/github/consult/object/all_cfor.RData')
+save(all.cfor2,file= 'C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
+load('C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
+
 ```
@@
 From analysing all jobs including 2009 and beyond, the most important variables are:
-`levels(var.imp4$var)[1:10] %>% as.data.frame`
+`levels(var.imp$var) %>% as.data.frame %>% slice(1:8)`
 
 The least important are 
-`levels(var.imp4$var)[11:19]`
+`levels(var.imp4$var) %>% as.data.frame %>% slice(9:16)`
 
@@
 
-See how variable selection goes with random forest first, then probably just run cforest again without code.client and contact.code
+Keep:
+
+* inv.mlsto
+* timespan
+* no.users
+* Discipline
+* pc.pro
+* client.totinv
+* client.numinv
+* pc.director
+* code.director
+* Num.disc
+* Business
+* Biz.type
+
+Dump
+
+* client.neginv
+* client.count
+* ProjEng.Pos
+* Biz.size
+* client.meaninv

### 2015-06-28 var_select.rmd
added client.age variable

--- 2015-06-24 var_select.rmd
@@
 
-Try running a cforest over night with full dataset - don't include 
-```
-
-```{r, echo=FALSE}
-
 system.time(
@@
 
+# New Variable - client age
+
+```{r, echo=FALSE}
+#read csv file with all the client details - want to merge this with client df
+details<- read.csv('17marchclients.csv',
+                   na.strings = c("", " ", "NA"))
+head(details)
+#delete client.contact, client.count, and contact.count columns
+details<- details[,c(1,5:9)]
+#rename columns
+names(details)[2]<- 'Business'
+names(details)[4]<- 'Biz.size'
+names(details)[5]<- 'Biz.type'
+names(details)[6]<- 'Biz.age'
+
+#delete rows with 4 NA values
+details<-details[rowSums(is.na(details))!=5,]
+details<-details[rowSums(is.na(details))!=4,]
+
+#add row for client code
+code.client<- read.csv('code_client2.csv')[,-1]
+details<- merge(details, code.client, by.x ='client', by.y ='client2', all.x=TRUE, all.y=FALSE)
+details<- transform(details, client.age = 2014- Biz.age)
+
+#merge client.age column of details to all7b by code.client
+all7b<- merge(all7b, details %>% select(code, client.age), by.x = 'code.client', by.y = 'code', all.x=TRUE, all.y=FALSE)
+
+```
+
 See how variable selection goes with random forest first, then probably just run cforest again without code.client and contact.code
+
+Use cforest to test remaining variables (with significant NA components)
+
+These include:
+
+* code.client
+* Post.Code
+* Billing.Type
+* Job.Source
+* Job.Detail.Primary
+* Job.Detail.Secondary
+* Job.Type.Primary
+* Job.Type.Secondary
+* Type
+* Project.Value
+* no.employees
+* client age????
+* contact.count
+* JD.Primary
+* JD.Second
+* dist
+* code.contact
+* client.age
+
+First eliminate some of the job.detail.secondary - that kind of thing
+Then see if any have matching NA rows
+Then run cforest on each variable individually

### 2015-06-29 var_select.rmd

--- 2015-06-26 var_select.rmd
@@
 
+# New Variable - client age
+
+```{r, echo=FALSE}
+#read csv file with all the client details - want to merge this with client df
+details<- read.csv('17marchclients.csv',
+                   na.strings = c("", " ", "NA"))
+head(details)
+#delete client.contact, client.count, and contact.count columns
+details<- details[,c(1,5:9)]
+#rename columns
+names(details)[2]<- 'Business'
+names(details)[4]<- 'Biz.size'
+names(details)[5]<- 'Biz.type'
+names(details)[6]<- 'Biz.age'
+
+#delete rows with 4 NA values
+details<-details[rowSums(is.na(details))!=5,]
+details<-details[rowSums(is.na(details))!=4,]
+
+#add row for client code
+code.client<- read.csv('code_client2.csv')[,-1]
+details<- merge(details, code.client, by.x ='client', by.y ='client2', all.x=TRUE, all.y=FALSE)
+details<- transform(details, client.age = 2014- Biz.age)
+
+#merge client.age column of details to all7b by code.client
+all7b<- merge(all7b, details %>% select(code, client.age), by.x = 'code.client', by.y = 'code', all.x=TRUE, all.y=FALSE)
+
+```
+
+See how variable selection goes with random forest first, then probably just run cforest again without code.client and contact.code
+
+Use cforest to test remaining variables (with significant NA components)
+
+These include:
+
+* code.client
+* Post.Code
+* Billing.Type
+* Job.Source
+* Job.Detail.Primary
+* Job.Detail.Secondary
+* Job.Type.Primary
+* Job.Type.Secondary
+* Type
+* Project.Value
+* no.employees
+* client age????
+* contact.count
+* JD.Primary
+* JD.Second
+* dist
+* code.contact
+* client.age
+
+First eliminate some of the job.detail.secondary - that kind of thing
+Then see if any have matching NA rows
+Then run cforest on each variable individually
+
 Keep:

### 2015-06-30 var_select.rmd
checked var importance of vars with high NA
wrote a function

@@
 
+#redo client.count and contact.count according to this reduced dataset!!
+client.count<- ddply(all7b, .(code.client), nrow)
+all7b<- merge(all7b, client.count, by='code.client', all.x=TRUE)
+all7b %>% select(code.client, client.count, V1) %>% slice(1000:1020)
+all7b<- transform(all7b, diff= client.count - V1)
+# all7b %>% select(code.client, client.count, V1, diff) %>% unique %>% arrange(-diff) %>% head
+#above line tells us that our most popular clients are also the ones with the most lost jobs - fine
+all7b$client.count<- all7b$V1
+all7b<- all7b %>% select(-diff, -V1)
+
 #change unknown Num.disc to be 1
@@
 ```{r, echo=FALSE}
-all7b<- read.csv('all7b.csv')
+all7b<- read.csv('all7b.csv')[,-1]
 
@@
 
-save(all.cfor2,file= 'C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
-load('C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
+save(all.cfor2,file= 'C:/Users/n9232371/Documents/Consultbusiness/data/objects/all_cfor2.RData')
+# load('C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
 
@@
 
+Keep:
+
+* inv.mlsto
+* timespan
+* no.users
+* Discipline
+* pc.pro
+* client.totinv
+* client.numinv
+* pc.director
+* code.director
+* Num.disc
+* Business
+* Biz.type
+
+Dump
+
+* client.neginv
+* client.count
+* ProjEng.Pos
+* Biz.size
+* client.meaninv
+
 # New Variable - client age
@@
 all7b<- merge(all7b, details %>% select(code, client.age), by.x = 'code.client', by.y = 'code', all.x=TRUE, all.y=FALSE)
+write.csv(all7b, 'all7c.csv')
+all7c<- read.csv('all7c.csv')[,-1]
 
@@
 * no.employees
-* client age????
+* client.age
 * contact.count
@@
 
-Keep:
+```{r eliminate sme variables, echo=FALSe}
+#between Job.Detail.Primary, Job.Detail.Secondary, Job.Type.Primary, Job.Type.Secondary, Type, JD.Primary, JD.Second
+# need to delete some
 
-* inv.mlsto
-* timespan
-* no.users
-* Discipline
-* pc.pro
-* client.totinv
-* client.numinv
-* pc.director
-* code.director
-* Num.disc
-* Business
-* Biz.type
+#Job.Detail.Primary - 1A, 1B, 1C and so on. 1215 NA values - delete
+#JD.Primary - 1A, 1B, 1C and so on. 628 NA values
+#Job.Detail.Secondary - 1A, 1B, 1C and so on. 2090 NA values - most don't make sense (ARch housing with industrial buidling) 
+#                                                            - industrial bldg with land subdivision - checked out, all wrong!
+#        - some do ie 2C Sustainable design with 2B water-sensitive urban design (WSUD) - use!
+#        - some do  ie 3C integrated water management and WSUD and 4C alternative sources of water supply - use!
 
-Dump
+#Job.Type.Primary - eg 1. Building Structures, 628 NA values, quite useful - guessing that it matches JD.Primary NA values
 
-* client.neginv
-* client.count
-* ProjEng.Pos
-* Biz.size
-* client.meaninv
+#Job.Type.Secondary - confusing
+#                   - 2. LandInfrastucture could be copied into land subdivision for Job.Detail.Secondary
+all7c$Job.Detail.Secondary <- as.character(all7c$Job.Detail.Secondary)
+all7c$Job.Detail.Secondary<- ifelse(all7c$Job.Detail.Secondary=='2E Land Subdivision' & all7c$Job.Type.Secondary=='2. LandInfrastructure',
+                                    '2. LandInfrastructure', all7c$Job.Detail.Secondary)
+all7c$Job.Detail.Secondary<- ifelse(is.na(all7c$Job.Detail.Secondary) & all7c$Job.Type.Secondary=='2. LandInfrastructure',
+                                    '2. LandInfrastructure', all7c$Job.Detail.Secondary)
+all7c$Job.Detail.Secondary <- as.factor(all7c$Job.Detail.Secondary)
+
+#                   - copy '4. Integrated water management' as 'water management' into JD.Second
+all7c$JD.Second <- as.character(all7c$JD.Second)
+all7c$JD.Second<- ifelse(is.na(all7c$JD.Second) & all7c$Job.Type.Secondary=='4. IntegratedWaterManagement',
+                                    'water management', all7c$JD.Second)
+all7c$JD.Second <- as.factor(all7c$JD.Second)
+
+# Type - short description - quite good - 1206 NA's - all 2010 or earlier
+
+#JD.Second - more detailed version of JD.Primary- 1277 NA's - across all years - keep
+
+all7c<- all7c %>% select(-Job.Detail.Primary, -Job.Type.Secondary)
+
+write.csv(all7c, 'all7c.csv')
+all7c<- read.csv('all7c.csv')[,-1]
+
+```
+
+now left with
+
+* code.client
+* Post.Code
+* Billing.Type
+* Job.Source
+* Job.Detail.Secondary
+* Job.Type.Primary
+* Type
+* Project.Value
+* no.employees
+* client.age
+* contact.count
+* JD.Primary
+* JD.Second
+* dist
+* code.contact
+* client.age
+
+Do any of these have significantly overlapping NA values???
+
+First need to change code.client and code.contact. Need to assign NA to any code.client or code.contact with less than 3 counts
+
+```{r NA into code.client, echo=FALSE}
+
+#all code.clients with counts less than 3 shall be 'NA'
+client.check<- ddply(all7c,.(code.client), nrow)
+all7c<- merge(all7c, client.check, by='code.client', all.x=TRUE)
+all7c$code.client<- as.character(all7c$code.client)
+all7c$code.client<- ifelse(all7c$V1>2, all7c$code.client, NA)
+all7c$code.client<- as.factor(all7c$code.client)
+all7c<- all7c %>% select(-V1)
+
+#all code.contacts with counts less than 3 shall be 'NA'
+#code for 'unknown' which should be NA
+all7c[all7c$code.contact %in% 'CN1983',]$code.contact <- NA
+contact.check<- ddply(all7c,.(code.contact), nrow)
+contact.check<- contact.check %>% filter(V1>2)
+all7c$code.contact<- as.character(all7c$code.contact)
+all7c$code.contact<- ifelse(all7c$code.contact %in% contact.check$code.contact, all7c$code.contact,
+                            NA)
+all7c$code.contact<- as.factor(all7c$code.contact)
+
+#redo contact.count
+cont.count<- ddply(all7c,.(code.contact), nrow)
+cont.count[is.na(cont.count$code.contact),]$V1<- NA
+all7c<- merge(all7c, cont.count, by='code.contact', all.x=TRUE)
+all7c$contact.count<- all7c$V1
+all7c<- all7c %>% select(-V1)
+
+#Project.Value - anything with '0' should be 'NA'
+all7c[all7c$Project.Value == 0,]$Project.Value<- NA
+#check out what the ratios of inv.mlsto/Project.Value are
+temp<- all7c %>% select(Project.Value, inv.mlsto,mlsto) %>% filter(!is.na(Project.Value))
+temp<- transform(temp, ratio = round(inv.mlsto/Project.Value*100,2))
+temp<- temp %>% filter(ratio< 16)
+all7c$Project.Value<- ifelse(all7c$mlsto %in% temp$mlsto, all7c$Project.Value, NA)
+write.csv(all7c, 'all7c.csv')
+
+```
+
+```{r NA overlap, echo=FALSE}
+
+#overlapping NA values for variables with many NA's
+# first lets look at Job.Detail.Secondary, Job.Type.Primary, Type, JD.Primary, JD.Second
+experiment<- all7c
+experiment<- experiment %>% select(code.client, Post.Code, Billing.Type, Job.Source, Job.Detail.Secondary, Job.Type.Primary, Type, Project.Value,
+                                   no.employees, client.age, contact.count, JD.Primary, JD.Second, dist, code.contact, client.age)
+experiment<- sapply(experiment, function(x) is.na(x))
+NA.plot<- melt(experiment)
+NA.plot$mlsto<- rep(all7c$mlsto, ncol(experiment))
+#reorder levels for ggplot by total number of NA TRUE
+# table(NA.plot$Var2, NA.plot$value)[,'TRUE']
+NA.plot <- within(NA.plot, 
+                         Var2 <- factor(Var2, 
+                                               levels=names(sort(table(NA.plot$Var2, NA.plot$value)[,'TRUE'], 
+                                                                 decreasing=FALSE))))
+
+ggplot(NA.plot, aes(x=Var2, y=mlsto)) +
+        geom_tile(aes(fill = value)) + 
+        theme(axis.text.x=element_text(angle=45,hjust=1))
+
+```
+
+Yes,
+
+* dist and Post.Code
+* Job.Type.Primary and JD.Primary
+* code.contact and contact.count
+
+For now, we will treat the rest as having 'different
+
+Run 12 cforests with the core variables!!
+
+```{r cforests for NAs, echo=FALSE}
+
+#function to create a subset dataframe based on extra variable
+#then runs cforest, and saves variable importance ggplot in a folder
+
+vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client'
+            'client.age', 'contact.count', 'JD.Second', 'client.age', 'Billing.Type', c('dist','Post.Code'),
+            c('code.contact','contact.count'), c('Job.Type.Primary','JD.Primary'))
+cfor.plotsave<- function(extra= vars){
+        for(i in 1:length(vars)){
+
+                core.var= all7c[,names(all7c) %in% c('inv.mlsto', 'timespan', 'no.users', 
+                                                     'Discipline', 'pc.pro', 
+                        'client.totinv', 'client.numinv','pc.director', 'code.director', 'Num.disc', 
+                        'Business', 'Biz.type', 'return.pdol', extra[[i]])]
+                core.var = core.var[!is.na(core.var[,extra[[i]][1]]),]
+
+                #run cforest
+                cfor.var1= cforest(return.pdol~.,
+                                   data= core.var)
+                #calculate variable importances
+                var.imp= varimp(cfor.var1) %>% as.data.frame
+                colnames(var.imp)[names(var.imp) %in% '.']= 'imp'
+                var.imp$var = rownames(var.imp)
+                #reorder levels for bar plot
+                var.imp<- within(var.imp, 
+                    var <- factor(var,levels=var.imp[order(-abs(var.imp$imp)),]$var))
+
+                q= ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
+                r=q + geom_bar(stat='identity') + 
+                        labs(title = extra[[i]])
+                #save plot
+                ggsave(plot=r,
+                       filename=paste('var',i,'.png',sep=''),
+                       path= 'C:/Users/n9232371/Documents/Consultbusiness/barplots/cforests/')
+
+                cat(i," ")
+
+        }
+
+}
+
+cfor.plotsave(extra=vars)
+
+```

### 2015-07-01 var_select.rmd
deleted outliers and re-ran a bunch of stuff

@@
 library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
 library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
 library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
@@
 
-Start running cforests
-Very slow, start with 100 rows only, using only core variables
-
-```{r, echo=FALSE}
-all7b<- read.csv('all7b.csv')[,-1]
-
-#create core data frame - variables with very few NA's and delete these NA's
-
-core<- all7b %>% select(code.client , Discipline ,  client.count , inv.mlsto , Business ,
-        Biz.size , Biz.type , no.users , Year , timespan , Num.disc , Num.days , mean.peeps,
-        pc.contracttech , hours.perday , profit.mlsto , balance.mlsto , hrs.mlsto , 
-        cost.mlsto , dis.sc.mlsto , inv.mlsto , return.pdol , code.contact ,
-        pc.director , pc.midtech , pc.midpro , pc.gradpro , pc.seniortech , pc.seniorpro ,
-        pc.pro , code.director , code.ProjEng , ProjEng.Pos , num.inv , mean.inv , num.neginv , 
-        client.meaninv , client.neginv , client.numinv , client.totinv)
-
-mini.core2 <- all7b %>% select(code.client , Discipline ,  client.count , inv.mlsto , Business ,
-        Biz.size , Biz.type , no.users , timespan, Num.disc, 
-        return.pdol, code.contact,
-        pc.director,
-        pc.pro , code.director , ProjEng.Pos ,
-        client.meaninv , client.neginv , client.numinv , client.totinv)
-
-mini.core3 <- all7b %>%
-        select(Discipline , inv.mlsto , Business, client.count,
-        Biz.size , Biz.type , no.users , timespan, Num.disc, 
-        return.pdol, pc.director,
-        pc.pro , code.director , ProjEng.Pos,
-        client.meaninv , client.neginv , client.numinv , client.totinv)
-
-```
-
-Try to make all cforests random selection of 200 rows - this time post 2009
-
-Try running a cforest over night with full dataset - don't include
-
-```{r, echo=FALSE}
-
-system.time(
-all.cfor2<- cforest(return.pdol~.,
-        data= mini.core3
-        )
-)
-
-var.imp<- varimp(all.cfor2) %>% as.data.frame
-colnames(var.imp)[names(var.imp) %in% '.']<-'imp'
-var.imp$var <- rownames(var.imp)
-#reorder levels for bar plot
-var.imp <- within(var.imp, 
-                    var <- factor(var,levels=
-                                            var.imp[order(-abs(var.imp$imp)),]$var))
-
-q<- ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
-q + geom_bar(stat='identity')
-
-save(all.cfor2,file= 'C:/Users/n9232371/Documents/Consultbusiness/data/objects/all_cfor2.RData')
-# load('C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
-
-```
-
-From analysing all jobs including 2009 and beyond, the most important variables are:
-`levels(var.imp$var) %>% as.data.frame %>% slice(1:8)`
-
-The least important are 
-`levels(var.imp4$var) %>% as.data.frame %>% slice(9:16)`
-
-Let's compare this to
-
-* all data from 2002 to 2014 - currently running in another R window
-* ANOVA - all data because fast
-* random forest
-
-## Anova:
-
-```{r anova_varimp, echo=FALSE}
-#use mini.core2- all years
-system.time(
-        aov.all<- aov(return.pdol~.,
-                      data= mini.core3)
-        )
-
-summary(aov.all)
-```
-
-Therefore most significant variables are:
-
-* no.users - number of people on the job - <<.1%
-* code.director - <<.1%
-* timespan - 4%
-* Num.disc - 5.6%
-* Discipline - 5.65%
-
-worst: Business, ProjEng.Pos, client.numinv,
-
-Contrasts to cforest results include
-
-* code.client pval=1 - this is likely because code.contact and code.client are highly correlated
-* inv.mlsto pval= .721 - surprising
-* client.numinv pval = 0.923 - surprisingly different
-
-# random forests
-
-First optimise parameters mtry and ntree - exclude code.client and code.contact for now as they have way too many categories
-
-```{r randomForest, echo=FALSE}
-#random forest can have max 53 categories in a variable
-#code.client has the most categories - 664
-
-# find optimal mtry
-
-#create train/test sets
-
-train_ind <- sample(seq_len(2355), size = 1570)
-
-train <- mini.core4[train_ind, ]
-test <- mini.core4[-train_ind, ]
-oob.err<-double(13)
-test.err<-double(13)
-for(mtry in 1:13){
-        fit<-randomForest(return.pdol ~., data=train, mtry=mtry, ntree=400)
-        oob.err[mtry]<-fit$mse[400]
-        test.err[mtry]<- mean((predict(fit,test)-test$return.pdol)^2)
-        cat(mtry," ")
-}
-
-qplot(1:mtry,test.err,geom=c("point","line"),color="pink")+
-        geom_line(aes(1:mtry,oob.err,colour="blue"))+
-        geom_point(aes(1:mtry,oob.err, colour="blue"))
-
-#COMPLETELY FLAT - MAY AS WELL USE MTRY = 5????? NOW VARY NUMBER OF TREES
-
-trees<- c(300,500,700,1000,1500)
-oob.err<-double(5)
-test.err<-double(5)
-for(ntree in 1:5){
-        fit<-randomForest(return.pdol ~., data=train, mtry=5, ntree=trees[ntree])
-        oob.err[ntree]<-fit$mse[trees[ntree]]
-        test.err[ntree]<- mean((predict(fit,test)-test$return.pdol)^2)
-        cat(ntree," ")
-}
-
-qplot(1:ntree,test.err,geom=c("point","line"),color="pink")+
-        geom_line(aes(1:ntree,oob.err,colour="blue"))+
-        geom_point(aes(1:ntree,oob.err, colour="blue"))
-
-#again, doesn't matter!! line is dead flat. i'm not having much confidence in the predictiveness of this model. 
-```
-
-Extract important variables from randomForest loop
-
-423 of the 664 client companies and 485 of the 670 client contacts appear ONLY once in the entire data set.
-While they may be great predictors what do we do in the majority of cases where the contact is new!?
-
-```{r, echo=FALSE}
-set.seed(47)
-
-#try running without code.client or code.contact
-multi.forest1<-function(data=mini.core3, nforest=1:10){
-        sample=double(50)
-        imp = NULL
-        for (i in 1:length(nforest)){
-
-                #run tree
-                assign(paste("Forest",i,sep=""),randomForest(return.pdol~.,
-                                                             data=data,
-                                                             ntree=400,mtry=5, importance=TRUE),
-                       envir=.GlobalEnv)
-
-                #put together importance rank for each variable
-
-                rand.imp= importance(paste("Forest",i,sep="") %>% get, type=1, scale=FALSE) %>% as.data.frame
-                colnames(rand.imp)[names(rand.imp) %in% '%IncMSE']= 'IncMSE'
-                rand.imp$var= rownames(rand.imp)
-                rand.imp= rand.imp %>% arrange(-abs(IncMSE))
-                rand.imp$rank= 1:nrow(rand.imp)
-
-                #rbind all variables and ranks on top of one another for each Forest
-
-                imp = rbind(imp, rand.imp %>% select(var,rank))
-
-                cat(i," ")
-
-        }
-
-        #output average rank of each variable
-        rank = ddply(imp, .(var), summarise, m.rank =mean(rank)) %>% arrange(m.rank)
-        return(rank)
-
-}
-
-rand.rank<- multi.forest1(data=mini.core2 %>% select(-code.contact, -code.client), nforest=1:10)
-
-kable(rand.rank)
-
-```
-
-Keep:
-
-* inv.mlsto
-* timespan
-* no.users
-* Discipline
-* pc.pro
-* client.totinv
-* client.numinv
-* pc.director
-* code.director
-* Num.disc
-* Business
-* Biz.type
-
-Dump
-
-* client.neginv
-* client.count
-* ProjEng.Pos
-* Biz.size
-* client.meaninv
-
 # New Variable - client age
@@
 
-now left with
-
-* code.client
-* Post.Code
-* Billing.Type
-* Job.Source
-* Job.Detail.Secondary
-* Job.Type.Primary
-* Type
-* Project.Value
-* no.employees
-* client.age
-* contact.count
-* JD.Primary
-* JD.Second
-* dist
-* code.contact
-* client.age
-
-Do any of these have significantly overlapping NA values???
-
 First need to change code.client and code.contact. Need to assign NA to any code.client or code.contact with less than 3 counts
@@
 ```{r NA into code.client, echo=FALSE}
+#Job.Source - 'twitter' should be NA
+all7c[all7c$Job.Source %in% 'Twitter',]$Job.Source<- NA
+
+#Billing.Type - 'Estimate' and 'FeeProposal' and 'NoCharge' should be NA
+all7c[all7c$Billing.Type %in% c('Estimate','FeeProposal','NoCharge'),]$Billing.Type<- NA
+
+#code.ProjEng - 'Unknown' should be NA
+all7c[all7c$code.ProjEng %in% c('Unknown'),]$code.ProjEng<- NA
+
+#ProjEng.Pos - 'Unknown' should be NA
+all7c[all7c$ProjEng.Pos %in% c('Unknown'),]$ProjEng.Pos<- NA
+
+#code.Eng2 - 'Unknown' should be NA
+all7c[all7c$code.Eng2 %in% c('Unknown'),]$code.Eng2<- NA
 
@@
 all7c$Project.Value<- ifelse(all7c$mlsto %in% temp$mlsto, all7c$Project.Value, NA)
+
+#delete outliers : return.pdol>30 GONE
+all7c %>% filter(return.pdol>3) %>% select(return.pdol, inv.mlsto, hrs.mlsto, dis.sc.mlsto,mlsto, cost.mlsto) %>% arrange(-return.pdol)
+all7c<- all7c %>% filter(!return.pdol>3)
+
+#drop levels in each variable
+all.vars<- colnames(all7c)
+for (i in 1:length(all.vars)){
+        if(all7c[,all.vars[i]] %>% class == 'factor'){
+        all7c[,all.vars[i]]<- droplevels(all7c[,all.vars[i]])
+        }
+
+}
+
 write.csv(all7c, 'all7c.csv')
 
+all7c<- read.csv('all7c.csv')[,-1]
+all7c$Post.Code<- as.factor(all7c$Post.Code)
+
 ```
+
+Start running cforests on core list of variables
+
+```{r, echo=FALSE}
+
+#create core data frame - variables with very few NA's and delete these NA's
+
+core<- all7c %>% select(code.client , Discipline ,  client.count , inv.mlsto , Business ,
+        Biz.size , Biz.type , no.users , Year , timespan , Num.disc , Num.days , mean.peeps,
+        pc.contracttech , hours.perday , profit.mlsto , balance.mlsto , hrs.mlsto , 
+        cost.mlsto , dis.sc.mlsto , inv.mlsto , return.pdol , code.contact ,
+        pc.director , pc.midtech , pc.midpro , pc.gradpro , pc.seniortech , pc.seniorpro ,
+        pc.pro , code.director , code.ProjEng , ProjEng.Pos , num.inv , mean.inv , num.neginv , 
+        client.meaninv , client.neginv , client.numinv , client.totinv)
+
+mini.core2 <- all7c %>% select(code.client , Discipline ,  client.count , inv.mlsto , Business ,
+        Biz.size , Biz.type , no.users , timespan, Num.disc, 
+        return.pdol, code.contact,
+        pc.director,
+        pc.pro , code.director , ProjEng.Pos , code.ProjEng,
+        client.meaninv , client.neginv , client.numinv , client.totinv)
+
+mini.core3 <- all7c %>%
+        select(Discipline , inv.mlsto , Business, client.count,
+        Biz.size , Biz.type , no.users , timespan, Num.disc, 
+        return.pdol, pc.director,
+        pc.pro , code.director , 
+        client.meaninv , client.neginv , client.numinv , client.totinv)
+
+```
+
+Try running a cforest over night with full dataset - don't include
+
+```{r, echo=FALSE}
+
+system.time(
+all.cfor2<- cforest(return.pdol~.,
+        data= mini.core3
+        )
+)
+
+var.imp<- varimp(all.cfor2) %>% as.data.frame
+colnames(var.imp)[names(var.imp) %in% '.']<-'imp'
+var.imp$var <- rownames(var.imp)
+#reorder levels for bar plot
+var.imp <- within(var.imp, 
+                    var <- factor(var,levels=
+                                            var.imp[order(-abs(var.imp$imp)),]$var))
+
+q<- ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
+q + geom_bar(stat='identity')
+
+save(all.cfor2,file= 'C:/Users/n9232371/Documents/Consultbusiness/data/objects/all_cfor2.RData')
+# load('C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
+
+```
+
+From analysing all jobs including 2009 and beyond, the most important variables are:
+`levels(var.imp$var) %>% as.data.frame %>% slice(1:8)`
+
+The least important are 
+`levels(var.imp4$var) %>% as.data.frame %>% slice(9:16)`
+
+Let's compare this to
+
+* all data from 2002 to 2014 - currently running in another R window
+* ANOVA - all data because fast
+* random forest
+
+## Anova:
+
+```{r anova_varimp, echo=FALSE}
+#use mini.core2- all years
+
+aov.all<- aov(return.pdol~.,
+              data= mini.core3)
+
+summary(aov.all)
+```
+
+Therefore most significant variables are:
+
+* no.users - number of people on the job - << 1%
+* timespan - << 1%
+* Discipline - <<1%
+* inv.mlsto - 8.9%
+* Business << 1%
+* Biz.type - 2%
+* pc.pro - .13%
+
+worst: Biz.size, client.numinv, client.neginv, client.meaninv, client.totinv, pc.director, Num.disc, Biz.size
+
+Contrasts to cforest results include
+
+* Num.disc
+* pc.director
+* Biz.size
+* Biz.type = cforest rated this very low
+
+# random forests
+
+First optimise parameters mtry and ntree - exclude code.client and code.contact for now as they have way too many categories
+
+```{r randomForest, echo=FALSE}
+#random forest can have max 53 categories in a variable
+#code.client has the most categories - 664
+
+# find optimal mtry
+
+#create train/test sets
+
+train_ind <- sample(seq_len(nrow(mini.core3)), size = 1546)
+
+train <- mini.core3[train_ind, ]
+test <- mini.core3[-train_ind, ]
+oob.err<-double(13)
+test.err<-double(13)
+for(mtry in 1:13){
+        fit<-randomForest(return.pdol ~., data=train, mtry=mtry, ntree=400)
+        oob.err[mtry]<-fit$mse[400]
+        test.err[mtry]<- mean((predict(fit,test)-test$return.pdol)^2)
+        cat(mtry," ")
+}
+
+qplot(1:mtry,test.err,geom=c("point","line"),color="pink")+
+        geom_line(aes(1:mtry,oob.err,colour="blue"))+
+        geom_point(aes(1:mtry,oob.err, colour="blue"))
+
+#USE MTRY = 5, NOW VARY NUMBER OF TREES
+
+trees<- c(300,500,700,1000,1500)
+oob.err<-double(5)
+test.err<-double(5)
+for(ntree in 1:5){
+        fit<-randomForest(return.pdol ~., data=train, mtry=5, ntree=trees[ntree])
+        oob.err[ntree]<-fit$mse[trees[ntree]]
+        test.err[ntree]<- mean((predict(fit,test)-test$return.pdol)^2)
+        cat(ntree," ")
+}
+
+qplot(1:ntree,test.err,geom=c("point","line"),color="pink")+
+        geom_line(aes(1:ntree,oob.err,colour="blue"))+
+        geom_point(aes(1:ntree,oob.err, colour="blue"))
+
+#use 500 trees
+```
+
+Extract important variables from randomForest loop
+
+423 of the 664 client companies and 485 of the 670 client contacts appear ONLY once in the entire data set.
+While they may be great predictors what do we do in the majority of cases where the contact is new!?
+
+```{r, echo=FALSE}
+set.seed(47)
+
+#try running without code.client or code.contact
+multi.forest1<-function(data=mini.core3, nforest=1:10){
+        sample=double(50)
+        imp = NULL
+        for (i in 1:length(nforest)){
+
+                #run tree
+                assign(paste("Forest",i,sep=""),randomForest(return.pdol~.,
+                                                             data=data,
+                                                             ntree=400,mtry=5, importance=TRUE),
+                       envir=.GlobalEnv)
+
+                #put together importance rank for each variable
+
+                rand.imp= importance(paste("Forest",i,sep="") %>% get, type=1, scale=FALSE) %>% as.data.frame
+                colnames(rand.imp)[names(rand.imp) %in% '%IncMSE']= 'IncMSE'
+                rand.imp$var= rownames(rand.imp)
+                rand.imp= rand.imp %>% arrange(-abs(IncMSE))
+                rand.imp$rank= 1:nrow(rand.imp)
+
+                #rbind all variables and ranks on top of one another for each Forest
+
+                imp = rbind(imp, rand.imp %>% select(var,rank))
+
+                cat(i," ")
+
+        }
+
+        #output average rank of each variable
+        rank = ddply(imp, .(var), summarise, m.rank =mean(rank)) %>% arrange(m.rank)
+        return(rank)
+
+}
+
+rand.rank<- multi.forest1(data=mini.core3, nforest=1:10)
+
+kable(rand.rank)
+
+```
+
+* no.users - number of people on the job - << 1%
+* timespan - << 1%
+* Discipline - <<1%
+* inv.mlsto - 8.9%
+* Business << 1%
+* Biz.type - 2%
+* pc.pro - .13%
+
+Keep:
+
+* inv.mlsto
+* timespan
+* no.users
+* Discipline
+* pc.pro
+* client.totinv
+* pc.director
+* code.director
+* Num.disc
+* Business
+* Biz.type
+* Biz.size
+
+Dump
+
+* client.neginv
+* client.count
+* ProjEng.Pos
+* client.numinv
+* client.meaninv
+
+#### YOU ARE HEREE #####
+
+now left with
+
+* ProjEng.Pos
+* code.ProjEng
+* code.client
+* Post.Code
+* Billing.Type
+* Job.Source
+* Job.Detail.Secondary
+* Job.Type.Primary
+* Type
+* Project.Value
+* no.employees
+* client.age
+* contact.count
+* JD.Primary
+* JD.Second
+* dist
+* code.contact
+* client.age
+
+Do any of these have significantly overlapping NA values???
 
@@
 
-vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client'
-            'client.age', 'contact.count', 'JD.Second', 'client.age', 'Billing.Type', c('dist','Post.Code'),
+vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client',
+            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', c('dist','Post.Code'),
             c('code.contact','contact.count'), c('Job.Type.Primary','JD.Primary'))
-cfor.plotsave<- function(extra= vars){
+cfor.plotsave<- function(extra= vars, name= 'vara'){
         for(i in 1:length(vars)){
@@
                 ggsave(plot=r,
-                       filename=paste('var',i,'.png',sep=''),
+                       filename=paste(name,i,'.png',sep=''),
                        path= 'C:/Users/n9232371/Documents/Consultbusiness/barplots/cforests/')
@@
 
-cfor.plotsave(extra=vars)
+cfor.plotsave(extra=vars, name='vara')
+cfor.plotsave(extra=vars, name='varb')
+
+all7c %>% select(Job.Detail.Secondary) %>% summary
 
 ```
+
+From the cforests, the extra variables worth keeping are
+
+* code.contact - ranks 2nd after timespan/2nd
+* code.client (ranks 4th/4th)
+* Project.Value (ranks 3rd/4th)
+* no.employees - ranks 7th - before code.director  / 3rd last 
+* JD.Second - ranks 6th - before code.director and Num.disc / 5th
+* Billing.Type - ranks 3rd last but I can't see how this is not important/last
+* Job.Type.Primary - ranks 7th - 6 nice broad categories/4th last
+
+Throwing out:
+* Job.Source - ranked last / 3rd last
+* Job.Detail.Secondary - too patchy I believe / 3rd last
+* Type - nice and broad but didn't seem to work - ranked 4th last / 2nd last
+* client.age - 3rd last/last
+* contact.count - 4th last/3rd last
+* dist - 5th last/last
+* Post.Code - 6th last or 7th last 
+* contact.count - 3rd last/4th last 
+* JD.Primary - 3rd last :(/2nd last
+
+```{r run anova on all, echo=FALSE}
+vars.vec<- c('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client',
+            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'dist','Post.Code',
+            'code.contact','contact.count', 'Job.Type.Primary','JD.Primary')
+
+aov.loop<- function(extra=vars.vec){
+        for (i in 1:length(extra)){
+                core.var= all7c[,names(all7c) %in% c('inv.mlsto', 'timespan', 'no.users', 
+                                                     'Discipline', 'pc.pro', 
+                        'client.totinv', 'client.numinv','pc.director', 'code.director', 'Num.disc', 
+                        'Business', 'Biz.type', 'return.pdol', extra[i])]
+                core.var = core.var[!is.na(core.var[,extra[i]]),]
+
+                formula = paste("return.pdol ~ inv.mlsto + Discipline + Business+
+                        Biz.type + Num.disc +
+                        no.users + client.numinv +
+                        pc.director + client.totinv + 
+                        pc.pro + timespan +
+                        code.director +", extra[i], sep="+")
+
+                model= aov(as.formula(formula)              
+                           ,data= core.var)
+
+                assign(paste("summ.",extra[i],sep=""), model, envir=.GlobalEnv)
+
+                cat(i," ")
+                }
+        }
+
+aov.loop(extra=vars.vec)
+
+```
+
+From Anova, the only new variables that are significant are:
+
+* Project.Value - 17%
+* Post.Code - 11.4%
+* code.contact - .000013%
+
+Visualise all categorical variables as boxplots vs return.pdol
+
+```{r Tukey Ancova visualisation, echo=FALSE}
+
+all.vars<- colnames(all7c %>% select(-mlsto))
+
+auto.BP<- function(colnames=all.vars, yvar = 'return.pdol', df= all7c, name='box'){
+
+        #make all boxplot dots match colours
+        update_geom_defaults("point", list(colour = NULL))
+
+        for (i in 1:length(colnames)){
+
+                #only perform plot if variable is a factor
+                #if variable is numeric or an integer, break into 8 bins
+                if(df[,colnames[i]] %>% class %in% c('factor')) {
+
+                medians= ddply(all7c,.(all7c[,colnames[i]]), summarise, med=median(return.pdol))
+
+                df[,colnames[i]] <- factor(df[,colnames[i]], 
+                                               levels= medians[order(-medians$med),][,1] )
+
+                r= ggplot(df[!is.na(df[,colnames[i]]),],
+                       aes_string(x=colnames[i], y=yvar, colour = colnames[i])) + 
+                        geom_boxplot(alpha = 0, size = .2, outlier.size = .5)+
+                        geom_jitter(alpha= 0.5, size=.2) +
+                        labs(title = paste(colnames[i], "vs.", yvar, sep=' '),
+                        x= colnames[i],
+                        y = yvar) + 
+                        theme(legend.position = 'none',
+                              text = element_text(size =9.5),
+                                axis.text.x= element_text(angle=45, hjust=1))+
+                        stat_summary(fun.y = mean, geom= 'point', shape = 1, size = .5)
+
+                #save plot
+                ggsave(plot=r,
+                       filename=paste(name,i,'.png',sep=''),
+                       path= 'C:/Users/n9232371/Documents/Consultbusiness/barplots/boxplots/')
+
+                cat(i," ")  
+                }
+
+}
+}
+
+auto.BP(colnames=all.vars, yvar = 'return.pdol', df= all7c, name='box')
+
+```
+
+colnames(all7c)

### 2015-07-02 cluster.rmd
finished var_select

---
title: "Clustering"
author: "Amy Cook"
date: "Thursday, July 02, 2015"
output: word_document
---

Cluster exploration!!

time to have a look at if some of these variables with many many categories can be simplified

Remaining categories include:

Full:
* inv.mlsto
* timespan
* no.users
* Discipline
* pc.pro
* client.totinv
* code.director
* Num.disc
* Business
* Biz.type

Many NA:
* code.contact
* code.client
* Project.Value
* JD.Second
* Billing.Type
* Post.Code 
* code.ProjEng
* Billing.Type
* majority.pos

### 2015-07-02 var_select.rmd
finished var_select

@@
 all7b[is.na(all7b$Num.disc),]$Num.disc <- 1
+
+#create new variable - tech.pos
+all7b %>% select(pc.contracttech, pc.midtech, pc.seniortech) %>% slice(1000:1020)
+test<- melt(all7b %>% select(mlsto, pc.contracttech, pc.midtech, pc.seniortech))
+test<- ddply(test, .(mlsto), max= max(value))
+#delete any rows where percentage is less than 10
+test<- test[!test$value<10,] %>% arrange(mlsto,-value)
+#now I just want the first row for every milestone - ie tech person that did the MOST
+test2<- ddply(test, .(mlsto), function(x) head(x,1))
+#now delete pc from the front of all the variables
+test2$variable<- sapply(test2$variable,FUN=function(x){
+        substr(x,4,15)})
+test2$variable<- as.factor(test2$variable)
+test2$value.bin<- cut(test2$value,9)
+table(test2$variable, test2$value.bin)
+mosaicplot(test2$variable ~ test2$value.bin, color=brewer.pal(9,"Spectral"),
+           cex.axis = .5)
+all7b<- merge(all7b, test2 %>% select(mlsto, variable), by='mlsto', all.x=TRUE, all.y=FALSE)
+colnames(all7b)[names(all7b) %in% 'variable']<-'pos.tech'
+
+#create new variable - majority.pos
+all7b %>% select(pc.contracttech, pc.midtech, pc.seniortech, pc.director, pc.seniorpro, pc.midpro) %>% slice(1000:1020)
+testa<- melt(all7b %>% select(mlsto, pc.contracttech, pc.midtech, pc.seniortech, pc.director, pc.gradpro,
+                             pc.midpro, pc.seniorpro, pc.unknown))
+#delete any rows where percentage is less than 10
+testa<- testa[!testa$value<10,] %>% arrange(mlsto,-value)
+#now I just want the first row for every milestone - ie tech person that did the MOST
+testa2<- ddply(testa, .(mlsto), function(x) head(x,1))
+#now delete pc from the front of all the variables
+testa2$variable<- sapply(testa2$variable,FUN=function(x){
+        substr(x,4,15)})
+testa2$variable<- as.factor(testa2$variable)
+#create bins for mosaic plot
+testa2$value.bin<- cut(testa2$value,5)
+table(testa2$variable, testa2$value.bin)
+#reorder levels for mosaic plot
+testa2$variable <- factor(testa2$variable, levels=names(sort(table(testa2$variable), 
+                                                                 decreasing=TRUE)))
+mosaicplot(testa2$variable ~ testa2$value.bin, color=brewer.pal(9,"Spectral"),
+           cex.axis = .6)
+all7b<- merge(all7b, testa2 %>% select(mlsto, variable), by='mlsto', all.x=TRUE, all.y=FALSE)
+colnames(all7b)[names(all7b) %in% 'variable']<-'majority.pos'
+all7b[all7b$majority.pos %in% 'unknown',]$majority.pos<- NA
+
 # write.csv(all7b,'all7b.csv')
@@
 all7b<- merge(all7b, details %>% select(code, client.age), by.x = 'code.client', by.y = 'code', all.x=TRUE, all.y=FALSE)
-write.csv(all7b, 'all7c.csv')
+# write.csv(all7b, 'all7c.csv')
 all7c<- read.csv('all7c.csv')[,-1]
@@
 #Billing.Type - 'Estimate' and 'FeeProposal' and 'NoCharge' should be NA
-all7c[all7c$Billing.Type %in% c('Estimate','FeeProposal','NoCharge'),]$Billing.Type<- NA
+all7c[all7c$Billing.Type %in% c('Estimate','FeeProposal','No Charge'),]$Billing.Type<- NA
 
@@
         Biz.size , Biz.type , no.users , timespan, Num.disc, 
-        return.pdol, pc.director,
+        return.pdol,
         pc.pro , code.director , 
@@
 * client.age
+* majority.pos
 
@@
 experiment<- all7c
-experiment<- experiment %>% select(code.client, Post.Code, Billing.Type, Job.Source, Job.Detail.Secondary, Job.Type.Primary, Type, Project.Value,
-                                   no.employees, client.age, contact.count, JD.Primary, JD.Second, dist, code.contact, client.age)
+experiment<- experiment %>% select(code.client, Post.Code, Billing.Type, Job.Source, Job.Detail.Secondary, 
+                                   Job.Type.Primary, Type, Project.Value,
+                                   no.employees, client.age, contact.count, JD.Primary, JD.Second, dist, code.contact,
+                                   client.age, ProjEng.Pos, code.ProjEng, majority.pos )
 experiment<- sapply(experiment, function(x) is.na(x))
@@
 * code.contact and contact.count
+* ProjEng.Pos and code.ProjEng
 
@@
 vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client',
-            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', c('dist','Post.Code'),
-            c('code.contact','contact.count'), c('Job.Type.Primary','JD.Primary'))
-cfor.plotsave<- function(extra= vars, name= 'vara'){
+            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'pos.tech', 'majority.pos',
+            c('dist','Post.Code'),
+            c('code.contact','contact.count'), c('Job.Type.Primary','JD.Primary'),
+            c('ProjEng.Pos', 'code.ProjEng'))
+
+cfor.plotsave<- function(extra= vars, name= 'vare'){
         for(i in 1:length(vars)){
@@
                                                      'Discipline', 'pc.pro', 
-                        'client.totinv', 'client.numinv','pc.director', 'code.director', 'Num.disc', 
-                        'Business', 'Biz.type', 'return.pdol', extra[[i]])]
+                        'client.totinv', 'pc.director', 'code.director', 'Num.disc', 
+                        'Business', 'Biz.type', 'Biz.size', 'return.pdol', extra[[i]])]
                 core.var = core.var[!is.na(core.var[,extra[[i]][1]]),]
@@
 cfor.plotsave(extra=vars, name='varb')
+cfor.plotsave(extra=vars, name='varc')
+cfor.plotsave(extra=vars, name='vard')
+cfor.plotsave(extra=list('majority.pos'), name='vare')
 
@@
 
-* code.contact - ranks 2nd after timespan/2nd
+* code.contact - ranks 4th after no.users, timespan, pc.pro
 * code.client (ranks 4th/4th)
-* Project.Value (ranks 3rd/4th)
-* no.employees - ranks 7th - before code.director  / 3rd last 
-* JD.Second - ranks 6th - before code.director and Num.disc / 5th
-* Billing.Type - ranks 3rd last but I can't see how this is not important/last
-* Job.Type.Primary - ranks 7th - 6 nice broad categories/4th last
+* Project.Value (ranks 4th/5th)
+* JD.Second - ranks 6/6th - before code.director and Num.disc / 5th
+* Billing.Type - ranks 3rd last/4th last but I can't see how this is not important/last
+* Post.Code - 4th
+* code.ProjEng - 8th
+* majority.pos
 
 Throwing out:
-* Job.Source - ranked last / 3rd last
-* Job.Detail.Secondary - too patchy I believe / 3rd last
-* Type - nice and broad but didn't seem to work - ranked 4th last / 2nd last
-* client.age - 3rd last/last
-* contact.count - 4th last/3rd last
-* dist - 5th last/last
+* Job.Source - ranked 4th last/3rd last
+* Job.Detail.Secondary - too patchy I believe / 5th last/3rd last
+* Type - nice and broad but didn't seem to work - ranked 5th last / 3rd last
+* client.age - last/last
+* contact.count - 4th last/last
+* dist - 3rd last/2nd last
 * Post.Code - 6th last or 7th last 
-* contact.count - 3rd last/4th last 
-* JD.Primary - 3rd last :(/2nd last
+* contact.count - last/2nd lsat 
+* ProjEng.Pos - 3rd last/ 4th last
+* no.employees - 3rd last / 2nd last
+* Job.Type.Primary - 3rd last
+* JD.Primary - 8th/9th,ie 7th last not too bad - too similar to JD.Second and JD.Second performs better
 
@@
             'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'dist','Post.Code',
-            'code.contact','contact.count', 'Job.Type.Primary','JD.Primary')
+            'code.contact','contact.count', 'Job.Type.Primary','JD.Primary','ProjEng.Pos', 'code.ProjEng', 'pos.tech',
+            'majority.pos')
 
@@
                                                      'Discipline', 'pc.pro', 
-                        'client.totinv', 'client.numinv','pc.director', 'code.director', 'Num.disc', 
-                        'Business', 'Biz.type', 'return.pdol', extra[i])]
+                        'client.totinv', 'Biz.type','code.director', 'Num.disc', 
+                        'Business', 'return.pdol', extra[i])]
                 core.var = core.var[!is.na(core.var[,extra[i]]),]
@@
                         Biz.type + Num.disc +
-                        no.users + client.numinv +
-                        pc.director + client.totinv + 
+                        no.users + 
+                        client.totinv + 
                         pc.pro + timespan +
-                        code.director +", extra[i], sep="+")
+                        code.director", extra[i], sep="+")
 
@@
 
-                assign(paste("summ.",extra[i],sep=""), model, envir=.GlobalEnv)
+                print(extra[i])
+                print(model %>% summary)
 
@@
 
+#try adding pc.director and pc.seniorpro to make pc.seniorpro.d - didn't go that well
+
+# all7c<- transform(all7c, pc.seniorpro.d = pc.seniorpro+pc.director)
+
 ```
@@
 * Project.Value - 17%
-* Post.Code - 11.4%
-* code.contact - .000013%
+* Post.Code - 23%
+* code.contact - 14.4%
+* code.client - .25%
+* JD.Second - 1.25%
+* dist - 17%
+* JD.Primary - 9.8% - delete, confusing with JD.Second
+* ProjEng.Pos - 1.5%
+* code.ProjEng - .9%
+* majority.pos - 3.7%
 
@@
 
-```{r Tukey Ancova visualisation, echo=FALSE}
-
-all.vars<- colnames(all7c %>% select(-mlsto))
-
-auto.BP<- function(colnames=all.vars, yvar = 'return.pdol', df= all7c, name='box'){
+```{r boxplots of each variable, echo=FALSE}
+
+auto.BP<- function(yvar = 'return.pdol', df= all7c, name='box'){
+
+        colnames= colnames(df %>% select(-mlsto))
 
@@
                        aes_string(x=colnames[i], y=yvar, colour = colnames[i])) + 
-                        geom_boxplot(alpha = 0, size = .2, outlier.size = .5)+
-                        geom_jitter(alpha= 0.5, size=.2) +
+                        geom_boxplot(alpha = 0, size = .5, outlier.size = .5)+
+                        geom_jitter(alpha= 0.5, size=.5) +
                         labs(title = paste(colnames[i], "vs.", yvar, sep=' '),
@@
                                 axis.text.x= element_text(angle=45, hjust=1))+
-                        stat_summary(fun.y = mean, geom= 'point', shape = 1, size = .5)
+                        stat_summary(fun.y = mean, geom= 'point', shape = 1, size = .6)
 
@@
 
-auto.BP(colnames=all.vars, yvar = 'return.pdol', df= all7c, name='box')
-
-```
-
-colnames(all7c)
+auto.BP(yvar = 'return.pdol', df= all7c %>% select(mlsto,return.pdol, majority.pos), name='box_maj')
+
+```
+
+Visualise all numerical variables as boxplots vs return.pdol
+
+```{r boxplots numeric vars, echo=FALSE}
+
+autonum.BP<- function(colnames=all.vars, yvar = 'return.pdol', df= all7c, name='box', breaks=4){
+
+        for (i in 1:length(colnames)){
+
+                #only perform plot if variable is a factor
+                #if variable is numeric or an integer, break into 8 bins
+
+                #how many total cols
+                        tot.num = df %>% ncol
+                        #how many numeric variables?
+                        num = df[,!sapply(df, is.factor)] %>% ncol
+
+                if(df[,colnames[i]] %>% class %in% c('numeric', 'integer')) {
+#                         print('yes')}
+
+                        #cut numeric columns into new variables with 8 categories, these are called 'cuti' 
+                        df[, paste0('cut',i)] <- cut_number(df[,colnames[i]] %>% as.numeric, breaks)
+#                         df[, paste0('cut',i)] <- cut_number(df[,colnames[i]] %>% as.numeric, 8)
+
+#                 medians= ddply(df,
+#                                .(df[,paste0('cut',i)]), summarise, med=median(return.pdol))
+
+#                 df[,paste0('cut',i)] <- factor(df[,paste0('cut',i)], 
+#                                                levels= medians[order(-medians$med),][,1] )
+
+                r= ggplot(df[!is.na(df[,paste0('cut',i)]),],
+                       aes_string(x=paste0('cut',i), y=yvar, colour = paste0('cut',i))) + 
+                        geom_boxplot(alpha = 0, size = .8, outlier.size = 1)+
+                        geom_jitter(alpha= 0.4, size=1) +
+                        labs(title = paste(colnames[i], "vs.", yvar, sep=' '),
+                        x= colnames[i],
+                        y = yvar) + 
+                        theme(legend.position = 'none',
+                              text = element_text(size =12),
+                                axis.text.x= element_text(angle=45, hjust=1))+
+                        stat_summary(fun.y = mean, geom= 'point', shape = 2, size = 1)
+
+                #save plot
+                ggsave(plot=r,
+                       filename=paste(name,i,'.png',sep=''),
+                       path= 'C:/Users/n9232371/Documents/Consultbusiness/barplots/boxplots/')
+
+                cat(i," ")  
+                }
+
+}
+}
+
+df<- all7c %>% select(-X, -No.Blue.Files, -contact.count, -no.users, -pc.contracttech,
+                      -pc.director, -pc.gradpro, -pc.midpro, -pc.midtech,-pc.seniorpro, -pc.seniortech,
+                      -pc.unknown, -pc.pro, -pc.tech, -Year, -Num.disc, -mean.peeps,
+                      -dis.sc.mlsto, -num.inv, -num.neginv, -client.neginv, -client.count)
+all.vars<- colnames(df)
+autonum.BP(colnames=all.vars, yvar = 'return.pdol', df= df, name='boxnum', breaks=8)
+
+```
+
+Visualise all percent variables, cut into 5, ie 0-20%, 20-40% etc
+
+```{r boxplots numeric vars, echo=FALSE}
+
+autopc.BP<- function(colnames=all.vars, yvar = 'return.pdol', df= df1, name='box', breaks=5){
+
+        for (i in 1:length(colnames)){
+
+                #how many total cols
+                        tot.num = df %>% ncol
+                        #how many numeric variables?
+                        num = df[,!sapply(df, is.factor)] %>% ncol
+
+                if(df[,colnames[i]] %>% class %in% c('numeric', 'integer')) {
+#                         print('yes')}
+
+                        #remove rows which =0
+                        df2<- df[!(df[,colnames[i]] == 0),]
+                        #cut numeric columns into new variables with 8 categories, these are called 'cuti' 
+                        df2[, paste0('cut',i)] <- cut(df2[,colnames[i]] %>% as.numeric, breaks)
+#                         df[, paste0('cut',i)] <- cut_number(df[,colnames[i]] %>% as.numeric, 8)
+
+#                 medians= ddply(df,
+#                                .(df[,paste0('cut',i)]), summarise, med=median(return.pdol))
+
+#                 df[,paste0('cut',i)] <- factor(df[,paste0('cut',i)], 
+#                                                levels= medians[order(-medians$med),][,1] )
+
+                r= ggplot(df2[!is.na(df2[,paste0('cut',i)]),],
+                       aes_string(x=paste0('cut',i), y=yvar, colour = paste0('cut',i))) + 
+                        geom_boxplot(alpha = 0, size = .8, outlier.size = 1)+
+                        geom_jitter(alpha= 0.4, size=1) +
+                        labs(title = paste(colnames[i], "vs.", yvar, sep=' '),
+                        x= colnames[i],
+                        y = yvar) + 
+                        theme(legend.position = 'none',
+                              text = element_text(size =12),
+                                axis.text.x= element_text(angle=45, hjust=1))+
+                        stat_summary(fun.y = mean, geom= 'point', shape = 2, size = 1)
+
+                #save plot
+                ggsave(plot=r,
+                       filename=paste(name,i,'.png',sep=''),
+                       path= 'C:/Users/n9232371/Documents/Consultbusiness/barplots/boxplots/')
+
+                cat(i," ")  
+                }
+
+}
+}
+
+df1<- all7c %>% select(pc.contracttech, return.pdol,
+                      pc.director, pc.gradpro, pc.midpro, pc.midtech,pc.seniorpro, pc.seniortech,
+                      pc.unknown, pc.pro, pc.tech, Year, mean.peeps, dis.sc.mlsto)
+all.vars<- colnames(df1)
+autopc.BP(colnames=all.vars, yvar = 'return.pdol', df= df1, name='boxpc', breaks=5)
+
+#test - added pc.seniorpro and pc.director together - didn't go that well
+# trial<- c('pc.seniorpro.d')
+# autopc.BP(colnames=trial, yvar = 'return.pdol', df= all7c %>% select(return.pdol,pc.seniorpro.d), name='boxpca', 
+#           breaks=5)
+
+```
+
+Plot integers which don't need cutting
+
+```{r, echo=FALSE}
+
+autoint.BP<- function(colnames=all.vars, yvar = 'return.pdol', df= all7c, name='boxint'){
+
+        for (i in 2:length(colnames)){
+
+                df[,colnames[i]]<- as.factor(df[,colnames[i]])
+
+                r= ggplot(df[!is.na(df[,colnames[i]]),],
+                       aes_string(x=colnames[i], y= yvar, colour = colnames[i] )) + 
+                        geom_boxplot(alpha = 0, size = .8, outlier.size = 1)+
+                        geom_jitter(alpha= 0.4, size=1) +
+                        labs(title = paste(colnames[i], "vs.", yvar, sep=' '),
+                        x= colnames[i],
+                        y = yvar) + 
+                        theme(legend.position = 'none',
+                              text = element_text(size =12),
+                                axis.text.x= element_text(angle=45, hjust=1))+
+                        stat_summary(fun.y = mean, geom= 'point', shape = 2, size = 1)
+
+                #save plot
+                ggsave(plot=r,
+                       filename=paste(name,i,'.png',sep=''),
+                       path= 'C:/Users/n9232371/Documents/Consultbusiness/barplots/boxplots/')
+
+                cat(i," ")  
+                }
+
+}
+
+df<- all7c %>% select(return.pdol, No.Blue.Files, contact.count, no.users, Num.disc,
+                      num.inv, num.neginv, client.neginv, client.count)
+all.vars<- colnames(df)
+autoint.BP(colnames=all.vars, yvar = 'return.pdol', df= df, name='boxint')
+
+```

### 2015-07-07 final write up/draft.rmd
began thesis draft

---
title: "draft 1"
author: "Amy Cook"
date: "Tuesday, July 07, 2015"
output: word_document
---

Originality Statement

'I hereby declare that this submission .... look up QUT'

Signed .................
Date....................

# Abstract

# Contents

# List of Figures

# List of Tables

# Glossary

B2B- Business to Business

# Acknowledgements

# Publications

# Introduction - 2 pages

* vague interesting few sentences - context

Predicting how long it will take a person to do something is a notoriously tricky task. People make mental calculations of these sort every day, guessing how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish. These mental calculations can be particularly inaccurate if the individual has never performed the task before. These errors are trivial in everyday activities, but they can have more meaning for businesses that sell their time as their chief product. For example, consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time the constultant spent on the problem. The delivery of 'expert advice' is generally in the form of a document (which could be anything from a report to hundreds of drawings) and can range from an hours advice to thousands of hours work on a substantial project. Before a project commences, the client and consultant must agree on a fee for the expert advice, or a fee structure. The nature of how a fee is structured can be as creative as the engaged parties wish, however a couple of simple examples are as follows: the client agrees to pay for the consultants time by the hour until the task is completed, or the consultant provides a fixed fee to complete the project in full, regardless of their hours spent. The nature of the fee structure brings an element of financial risk to one party which must be mitigated.

This thesis focuses on the risk a consulting company undertakes when setting a fixed fee for a project before their role has commenced. Clients commonly ask for competitive fixed quotes from several consultants before settling with one or negotiating further. The way a consulting manager decides on their fixed fee varies greatly from industry to industry and company to company. Typically, consulting manager has extensive experience in the type of project he is quoting and can use a combination of intuition, ratios of the entire project cost, and looking up a couple of past similar projects. Such a method often results in a wide range of profitabilities for projects depending on the ability of the manager to predict the amount of time the project will demand from the consulting team. This thesis examines a case study consulting company that has recorded internal historical project performance data since 2002. The study tests whether a statistical predictive model based on past project data can predict the profitability of a project based on characteristics available at the beginning of a project. This thesis asks further whether the prediction can impact a consulting manager's determination of a fixed fee.

# Chapter 1 Problem Description

In this chapter the research motivations are explained along with the goals of the project and the specific research hypothesis. Finally the contributions of this body of work are outlined along with the structure of the thesis.

## 1.1 Research Motivations

* thousands of consulting companies globally
* range from single consultant to 10's of 1000's of employees
* employees broad range of people from fresh graduates, to founders with decades of experience
* internal environment: worker feelings of morale and security can be heavily influenced by success and profitability of their projects. Not to mention possibilities for pay rises and promotions which all come together to create a positive thriving culture. 
* harsher economic climate, particularly important to get competitive edge
* find a source explaining how companies create their fixed fees - intuition, experience
* find a source saying how badly it goes!
* in the case study (and is typical of smaller organisations??), the CRM is an untapped source of information storing
    * employee hours
    * client information/characteristics
    * invoicing
    * employee costs
    * employee charge out rates
    * project information
* currently the case study company uses a widely popular CRM, where they can output information but cannot perform modelling or analysis:
    * look up one job at a time
    * basic summaries of overall hours spent for the month

### 1.1.1 Limitations

* looking at a single case study company in a specific industry
* Managers commonly price jobs based on the budget for the entire project, of which the consultant is a minor part. An example of this is a consultant who typically charges 1% of the total project cost
    * this data has not been historically recorded. Can start being recorded and would be extremely useful
    * even better to have expected cost of entire project at start of project, as well as actual final cost of project.
* A fee is based off detailed information that is unique to each project, for example the size of a building in terms of storeys and square meterage. This detailed information is not available in the data, and is too complex and varied for an algorithm to digest. Therefore it is not reasonable for an algorithm to predict the fee of a project at this stage, but much more achievable to predict the profitability of a project (ie return per dollar) based on historical data. This profitability prediction relates to the profitability of the project if the fee is generated using the same method the company has used in its historical data.
* Although the aim is to predict the profitability of a project, the model aims to change managers' fee setting decisions if the algorithm predicts low profit or a loss based on historical data.

## Thesis aims

For the reasons explained above, the aim of this research is as follows:

**General Aim** The aim of this project is to use statistical techniques to model decisions for consulting businesses. Research will focus on a case study Engineering consulting company that sells their expert advice (in the currency of time) to business clients. The business decisions include predicting project fees before project engagement and client analysis using existing invoicing and hourly employee-cost data. Several statistical and machine learning techniques will be tested, compared and refined.

**Hypothesis 1** 
k-means clustering methods can identify groups of similar projects with significant differences in profitability.

Unsupervised learning to identify clusters is a successful marketing technique used by retail companies to target groups of similar customers with similar habits and preferences. In this case of consulting companies, clusters may have multiple uses. New projects may have characteristics that slot into an 'unprofitable' cluster and the decision is clear to not engage in the project. Alternatively, if past projects and accompanying clients are found to fall into a very profitable cluster, consulting managers may focus services on these clients and reach out to these clients in particular for new projects.

**Hypothesis 2**
A statistical or machine learning model based on historical project data can predict the profitability of a new project with sufficient confidence to influence a project manager's fee proposal.

## 1.2 Thesis contributions - Gaps in literature

* a gap in the study of predicting internal performance in businesses with business clients. The literature has assessed performance of movies, customer analytics but rarely employee performance or business clients. This is particularly relevant to consulting companies that tackle complicated discrete projects with set timelines. Statistical and machine learning models to predict the correct fee for projects according to project type, client, and other characteristics could reduce risk for a range of consulting businesses in setting initial fees.

* A gap in machine learning predictive models which cater for variables with large amounts of missing data. the nature of consulting jobs is quite complex and historically, not all information about a project is possible to obtain. Variables with missing data are alternatively treated as optional bonus variables in a multi step model as opposed to missing information which must be imputed or deletion of the entire case that contains missing information.

* the application of clustering methods to business to business decisions.

## 1.3 Thesis structure

lit review
Background info
present the case study
Present the models you've developed
summary of findings, implications for use within industry, limitations

# Chapter 2 Literature Review - 25 pages for pHd

This chapter provides an overview of the literature available about current business to business (B2B) decision strategy as well as methods in machine learning and statistical business prediction models. In particular, the literature on B2B decision strategy is centred around setting fixed consulting fees and methods of targeting new clients or projects. The most prominent statistical and machine-learning models developed for business applications shall be reviewed along with models applied to the most closely related business cases to this project.

## 2.1 Business to Business Decision Strategy

### 2.1.1 Client and project analysis

* clustering
* how do businesses currently review their projects or clients

### 2.1.2 Fixed fees

* fixed fee problem in the IT consulting world - developed new techniques?
* managers currently go by gut feeling or look up a couple of past examples
* difficult to combine insights from similar jobs/clients

### 2.1.3 Gap
* no mobilisation of CRM data into sophisticated models in B2B world

## 2.2 Statistical and machine learning methods applied to business problems

### 2.2.1 Most simple and frequent models

* K means
* anova
* in construction journals, commonly applied linear regression
* do these acknowledge their data is not normally distributed?

### 2.2.2 Complex models

* SVM, Neural networks, random forest
* data must not be normal
* most successful?

### 2.2.3 ensemble methods

* papers
* netflix?
* multi step to deal with missing data

### 2.2.4 similar applications

* Employee churn - anything similar?
* stepped method business application?
* Gap

## 2.3 Conclusion

summary of 
* how are B2B decisions currently made
* what are the most popular/effective methods
* which ensemble methods may suit our missing data situation
* Gap in literature/contribution

# Chapter 3 Introduction of Case study

## 3.1 Obtaining dat

### 3.1.1 How data was recorded

* years of employees entering hours each day
* project managers or admin entering client details

### 3.1.2 How data was extracted - CRM

* queried direct from CRM
* data de-identification

## 3.2 Summary of variables

* categorical
* numerical
* significant cleaning
    * cost sometimes entered in as hours. ie $24 entered as 24 hours
* engineered variables
    * grep through project names to create categories - categories developed with case study manager
    * return per dollar = (invoiced - cost)/cost
    * hours analysed to produce: employee position that performed majority of the hours
        * % of hours completed by majority position
    * timespan based on timesheet entries as well as number of days with hours recorded
    * number of workers on project based on timesheet entries
    * client characteristics such as mean total amount invoiced for a repeat client

* amount of missingness

## 3.3 Variable selection

* combination of cforest, anova, random forest

### 2015-07-07 cluster.rmd
began clustering

@@
 ---
-title: "Clustering"
-author: "Amy Cook"
-date: "Thursday, July 02, 2015"
+title: "clusters"
 output: word_document
 ---
+
+```{r, echo=FALSE, include=FALSE}
+
+library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
+
+# library('knitr')
+# library('ggplot2')
+# library("dplyr")
+# library("plyr")
+# library('magrittr')
+# library('reshape2')
+# library("rpart")
+# library('car')
+# library('e1071')
+# library('corrgram')
+# library('party')
+# library('dendextend')
+# 
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
+
+all7c<- read.csv('all7c.csv')[,-1]
+all7c$Post.Code<- as.factor(all7c$Post.Code)
+
+```
 
@@
 Full:
+
+inv.mlsto
+timespan
+no.users
+Discipline
+pc.pro
+client.totinv
+code.director
+Num.disc
+Business
+Biz.type
+
+Many NA:
+
+code.contact
+code.client
+Project.Value
+JD.Second
+Billing.Type
+Post.Code
+code.ProjEng
+Billing.Type
+majority.pos
+
+Create reduced dataset and have a think about what you want to cluster against each other
+
+```{r final_var, echo=FALSE}
+
+all7d<- all7c %>% select(inv.mlsto, timespan, no.users, Discipline, pc.pro, client.totinv, code.director, Num.disc, Business,
+                         Biz.type, code.contact, pc.majpos,
+                         code.client, Project.Value, JD.Second, Billing.Type, Post.Code, code.ProjEng, Billing.Type, majority.pos)
+
+```
+
+What kind of clusters are you looking for? Within the numeric variables:
+
+```{r, echo=FALSE}
+all7c[,!sapply(all7c, is.factor)] %>% colnames
+all7d[,!sapply(all7d, is.factor)] %>% colnames
+
+```
+
+Numeric variables in reduced dataset are: inv.mlsto, timespan, no.users, pc.pro, client.totinv, Num.disc, Project.Value, pc.majpos
+
+Numeric variables worth considering in full dataset: Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto, cost.mlsto,
+
+So, lets choose:
+
 * inv.mlsto
@@
 * no.users
-* Discipline
 * pc.pro
 * client.totinv
-* code.director
 * Num.disc
-* Business
-* Biz.type
+* Num.days
+* mean.peeps
+* hours.perday
+* balance.mlsto
+* hrs.mlsto
+* cost.mlsto
 
-Many NA:
-* code.contact
-* code.client
+Many NA - do second
+* pc.majpos
 * Project.Value
-* JD.Second
-* Billing.Type
-* Post.Code 
-* code.ProjEng
-* Billing.Type
-* majority.pos
+
+Before clustering, need to standardise variables, and have a look at 'normality' of other variables
+
+```{r var_check, echo=FALSE}
+
+#Q-Q plot of each variable
+clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
+                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
+                         cost.mlsto, pc.majpos, return.pdol)
+
+#get rid of zeroes for log transform
+
+clust[clust$timespan ==0,]$timespan <- 1
+clust[clust$pc.pro ==0,]$pc.pro <- 5
+clust[clust$hours.perday ==0,]$hours.perday <- .1
+clust[clust$hours.perday ==0,]$hours.perday <- .1
+clust[clust$balance.mlsto ==0,]$balance.mlsto <- 1
+clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
+clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1
+clust[clust$return.pdol ==0,]$return.pdol <- .01
+
+for(i in 1:ncol(clust)){
+qqPlot(clust[,i],
+       main=colnames(clust)[i])
+}
+
+#qqplots for log of each variable
+for(i in 2:ncol(clust)+1){
+qqPlot(clust[,i] %>% log,
+       main=colnames(clust)[i])
+}
+
+#qqplots for sqrt of each variable
+for(i in 2:ncol(clust[,-1])+1){
+qqPlot(clust[,i] %>% sqrt,
+       main=colnames(clust)[i])
+}
+
+#density plots - plain variables
+for(i in 1:ncol(clust)){
+print(
+        ggplot(clust, aes_string(x=colnames(clust[i]))) + geom_histogram(aes(y=..density..))
+        )
+}
+#density plots - log transformed variables
+clust.log<- log(clust)
+for(i in 1:ncol(clust.log)){
+        print(
+                ggplot(clust.log, aes_string(x=colnames(clust.log[i]))) + geom_histogram(aes(y=..density..))
+                )
+}
+
+#plot sqrt of timespan
+clust$timespan <- clust$timespan %>% sqrt
+ggplot(clust, aes_string(x='timespan')) + geom_histogram(aes(y=..density..))
+
+```
+
+taking the log worked for:
+* cost.mlsto, hrs.mlsto, hours.perday, Num.days, client.totinv, no.users, inv.mlsto
+* sqrt of timespan
+
+so we end up with a numeric data set with the above variables log transformed.
+Then scale each by subtracting the mean of each variable and then dividing by the sd of each variable.
+
+```{r clust final, echo=FALSE}
+clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
+                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
+                         cost.mlsto, return.pdol)
+
+#get rid of zeroes for log transform
+
+clust[clust$timespan ==0,]$timespan <- 1
+clust[clust$hours.perday ==0,]$hours.perday <- .1
+clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
+clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1
+
+#take log of all variables to be logged in a for loop
+logged<- c('cost.mlsto', 'hrs.mlsto', 'hours.perday', 'Num.days', 'client.totinv',
+           'no.users', 'inv.mlsto')
+for(i in 1:length(logged)){
+        clust[,logged[i]]<- clust[,logged[i]] %>% log
+}
+clust$timespan<- sqrt(clust$timespan)
+
+#now to standardise
+clust<- scale(clust[,-1]) %>% as.data.frame
+```
+
+Now to do hierarchical clustering
+
+```{r hierarch, echo=FALSE}
+
+#Ward's clustering
+
+x<- dist(clust, method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+
+rect.hclust(fit, k=5, border='red')
+#5 clusters looks pretty good
+
+#try horizontal dend with branches coloured for 8 clusters - very slow
+# dend<- fit %>% as.dendrogram
+# dend %>% color_branches(k=8) %>% plot(horiz=T)
+# rect.dendrogram(dend, k=8, horiz=T)
+
+#how many points in each cluster?
+clust$groups<- cutree(fit,k=5)
+table(clust$groups)
+
+```
+
+Now look at properties of groups using k means clustering
+
+```{r kmeans, echo=FALSE}
+
+k1<- kmeans(clust %>% select(-groups), 5)
+k1.centres<- k1$centers %>% as.data.frame
+k1.centres$group<- rownames(k1.centres)
+k1.melt<- melt(k1.centres)
+k1.melt$group<- k1.melt$group %>% as.factor
+ggplot(k1.melt, aes(x=variable, y= value, fill=variable)) + geom_bar(stat='identity', position='dodge')+
+        facet_wrap(~group) +
+        theme(axis.text.x=element_text(angle=45,hjust=1))
+
+#try plotting means of actual values
+
+all7c$groups<- k1$cluster
+kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
+                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
+                         cost.mlsto, return.pdol, groups)
+all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
+all7c.plot$groups<- rownames(all7c.plot)
+k2.melt<- melt(all7c.plot)
+
+#summary means of all numeric variables in 5 clusters
+clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
+tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
+bal<- sum(all7c$balance.mlsto)
+hrs<-  sum(all7c$hrs.mlsto)
+clust.summary$pc.jobs<- round( tab[,2]*100, 1)
+clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
+        select(2)
+clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
+        select(2)
+
+clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]
+
+```

### 2015-07-07 var_select.rmd
began clustering

@@
 # library('party')
-
+# 
 # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
@@
 
+#create new variable - pc.majpos - percentage done by majority position
+all7b<- merge(all7b, testa2 %>% select(mlsto, value), by='mlsto', all.x=T, all.y=F)
+#if majority.pos is unknown, then value must be unknown - NA
+all7b$value<- ifelse(is.na(all7b$majority.pos), NA, all7b$value)
+colnames(all7b)[names(all7b) %in% 'value']<-'pc.majpos'
+
 # write.csv(all7b,'all7b.csv')
@@
                                    no.employees, client.age, contact.count, JD.Primary, JD.Second, dist, code.contact,
-                                   client.age, ProjEng.Pos, code.ProjEng, majority.pos )
+                                   client.age, ProjEng.Pos, code.ProjEng, majority.pos, pc.majpos )
 experiment<- sapply(experiment, function(x) is.na(x))
@@
 
-vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client',
-            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'pos.tech', 'majority.pos',
+vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees',
+            'code.client',
+            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'pos.tech', 
+            'majority.pos', 'pc.majpos',
             c('dist','Post.Code'),
@@
 * majority.pos
+* pc.majpos
 
@@
             'code.contact','contact.count', 'Job.Type.Primary','JD.Primary','ProjEng.Pos', 'code.ProjEng', 'pos.tech',
-            'majority.pos')
+            'majority.pos', 'pc.majpos')
 

### 2015-07-08 cluster.rmd
more clust

@@
 # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
+# opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
 
@@
 majority.pos
+pc.majpos
 
@@
 
-#Q-Q plot of each variable
-clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
-                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
-                         cost.mlsto, pc.majpos, return.pdol)
-
-#get rid of zeroes for log transform
-
-clust[clust$timespan ==0,]$timespan <- 1
-clust[clust$pc.pro ==0,]$pc.pro <- 5
-clust[clust$hours.perday ==0,]$hours.perday <- .1
-clust[clust$hours.perday ==0,]$hours.perday <- .1
-clust[clust$balance.mlsto ==0,]$balance.mlsto <- 1
-clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
-clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1
-clust[clust$return.pdol ==0,]$return.pdol <- .01
-
-for(i in 1:ncol(clust)){
-qqPlot(clust[,i],
-       main=colnames(clust)[i])
-}
-
-#qqplots for log of each variable
-for(i in 2:ncol(clust)+1){
-qqPlot(clust[,i] %>% log,
-       main=colnames(clust)[i])
-}
-
-#qqplots for sqrt of each variable
-for(i in 2:ncol(clust[,-1])+1){
-qqPlot(clust[,i] %>% sqrt,
-       main=colnames(clust)[i])
-}
-
-#density plots - plain variables
-for(i in 1:ncol(clust)){
-print(
-        ggplot(clust, aes_string(x=colnames(clust[i]))) + geom_histogram(aes(y=..density..))
-        )
-}
-#density plots - log transformed variables
-clust.log<- log(clust)
-for(i in 1:ncol(clust.log)){
-        print(
-                ggplot(clust.log, aes_string(x=colnames(clust.log[i]))) + geom_histogram(aes(y=..density..))
-                )
-}
-
-#plot sqrt of timespan
-clust$timespan <- clust$timespan %>% sqrt
-ggplot(clust, aes_string(x='timespan')) + geom_histogram(aes(y=..density..))
+# #Q-Q plot of each variable
+# clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
+#                          Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
+#                          cost.mlsto, pc.majpos, return.pdol)
+# 
+# #get rid of zeroes for log transform
+# 
+# clust[clust$timespan ==0,]$timespan <- 1
+# clust[clust$pc.pro ==0,]$pc.pro <- 5
+# clust[clust$hours.perday ==0,]$hours.perday <- .1
+# clust[clust$hours.perday ==0,]$hours.perday <- .1
+# clust[clust$balance.mlsto ==0,]$balance.mlsto <- 1
+# clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
+# clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1
+# clust[clust$return.pdol ==0,]$return.pdol <- .01
+# 
+# for(i in 1:ncol(clust)){
+# qqPlot(clust[,i],
+#        main=colnames(clust)[i])
+# }
+# 
+# #qqplots for log of each variable
+# for(i in 2:ncol(clust)+1){
+# qqPlot(clust[,i] %>% log,
+#        main=colnames(clust)[i])
+# }
+# 
+# #qqplots for sqrt of each variable
+# for(i in 2:ncol(clust[,-1])+1){
+# qqPlot(clust[,i] %>% sqrt,
+#        main=colnames(clust)[i])
+# }
+# 
+# #density plots - plain variables
+# for(i in 1:ncol(clust)){
+# print(
+#         ggplot(clust, aes_string(x=colnames(clust[i]))) + geom_histogram(aes(y=..density..))
+#         )
+# }
+# #density plots - log transformed variables
+# clust.log<- log(clust)
+# for(i in 1:ncol(clust.log)){
+#         print(
+#                 ggplot(clust.log, aes_string(x=colnames(clust.log[i]))) + geom_histogram(aes(y=..density..))
+#                 )
+# }
+# 
+# #plot sqrt of timespan
+# clust$timespan <- clust$timespan %>% sqrt
+# ggplot(clust, aes_string(x='timespan')) + geom_histogram(aes(y=..density..))
 
@@
 
-clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]
+clust.summary<- clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]
+
+# add medians
+all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))
+
+all7c.med<- all7c.med[, c(2,13,11,12,14,10,3,8,7,4,5,9,6)]
 
 ```
+
+Very interesting, now want to optimise number of clusters- refer wards tutorial
+
+```{r optimise clustnum,echo=FALSE}
+library('fpc', lib='C:/Progra~1/R/R-3.2.0/library')
+library('vegan', lib = 'C:/Progra~1/R/R-3.2.0/library')
+
+#try clustering a low number of numeric variables, trialling iffy ones to see which ones are good for clustering
+clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
+                                      hrs.mlsto), 2, 20, iter= 100)
+plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
+# client.totinv indicating about 4 clusters
+# Num.days - good, indicating 18 custers
+# hrs.mlsto - ok, heavy dip, good at about 17 clusters
+# pc.pro - 5 groups good, still iffy
+# mean.peeps - good at about 4 clusters
+# Num.disc - ok, inidicating 12 groups
+# timespan: not good for clustering
+# no.users: not good for clustering
+# hours.perday: no good for clustering
+
+## try out all iffy variables together, eliminate duds
+clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
+                                      client.totinv, Num.days, hrs.mlsto, pc.pro, mean.peeps,
+                                      Num.disc), 2, 20, iter= 100)
+plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
+# all together, no good
+# smaller subset of variables I am most interested in that still works for clustering:
+
+clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
+                                      hrs.mlsto
+                                      ), 2, 20, iter= 100)
+plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
+
+# 17 clusters
+# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
+# then calinksi (index of cluster homogeneity and distance where higher number is better)
+# calinski: between cluster distance/within cluster distance
+ind.plot<- melt(clust.km$results)
+ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
+        geom_line( ) +
+        theme(axis.text.x=element_text(angle=45,hjust=1))
+ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
+        geom_line( )+
+        theme(axis.text.x=element_text(angle=45,hjust=1))
+# 17 clusters should be good
+
+```
+
+Analysis shows that inv.mlsto, balance.mlsto, return.pdol, and hrs.mlsto should be used for clustering and that 17 clusters is ideal.
+I'm not completely happy with the clustering.
+I think I should now add a new variable: inv.mlsto x return.pdol
+This will give a 'success' scale, weighing large jobs even with a smaller return.pdol as quite good and tiny jobs with a good return.pdol as average to low.
+
+this means I will need to check normality, scale the variable before adding it to clust
+
+```{r new var, echo=FALSE}
+#new variable for all7c - success
+# success= return.pdolxinv.mlsto
+
+all7c<- transform(all7c, suc= inv.mlsto*return.pdol)
+all7c$suc %>% summary
+
+all7c %>% arrange(-suc) %>% head
+
+qqPlot(all7c$suc,
+       main='success')
+
+qqPlot((all7c$suc+102700) %>% log)
+
+#can't log because of negatives
+
+#density plots - plain variables
+
+ggplot(all7c, aes_string(x='suc')) + geom_histogram(aes(y=..density..))
+
+#density plots - log transformed variables
+
+trial<- (all7c$suc) %>% as.data.frame
+ggplot(trial, aes_string(x='.')) + geom_histogram(aes(y=..density..))
+
+### CAN WE TAKE LOG OF ABSOLUTE VALUES AND THEN MAKE NUMBERS THAT WERE PREVIOUSLY NEG NEG AGAIN?
+
+clust$suc<- scale(all7c$suc)
+
+```
+
+Repeat clustering with new suc
+
+```{r new successvar, echo=FALSE}
+
+#investigate new variable a little
+
+clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
+                                      suc, pc.pro, hrs.mlsto
+                                      ), 2, 20, iter= 100)
+plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
+
+# 10 clusters
+# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
+# then calinksi (index of cluster homogeneity and distance where higher number is better)
+# calinski: between cluster distance/within cluster distance
+ind.plot<- melt(clust2.km$results)
+ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
+        geom_line( ) +
+        theme(axis.text.x=element_text(angle=45,hjust=1))
+ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
+        geom_line( )+
+        theme(axis.text.x=element_text(angle=45,hjust=1))
+# 17 clusters should be good
+```

### 2015-07-08 combine.R

@@ -13,6 +13,7 @@ library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
 
 all4e<- read.csv('all4e.csv')[,-1]
 str(all4e)
+all4e<- all4e %>% filter(!(Job.Number == '2010.041.1' & dist == 1362.8))
 #delete Start.Date, tot.TS.Charge, tot.TS.cost, tot.TS.hours, Year
 all4e<- all4e %>% select(-Start.Date, -Tot.TS.Cost, -Tot.TS.Charge, -TS.Hours, -Year, -Profit, -Tot.Invoiced)
 
@@ -33,26 +34,26 @@ str(all5)
 all5<- transform(all5, profit= Tot.Invoiced-charge-Dis.subcon)
 #balance column - subtract real costs
 all5<- transform(all5, balance= Tot.Invoiced-cost-Dis.subcon)
-all5<- all5[-c(2068:2070),]
 write.csv(all5, 'all5.csv')
 all5<- read.csv('all5.csv')[,-1]
+all5<- all5 %>% filter(!Job.Number == '2003.24')
 
 #Need to sum up all the profits and balances for jobs with milestones
 # first 
 #create subset of just the things i want to add up - profits, balance, hours in each milestone
 temp<- all5 %>% select(Job.Number, hours, profit, balance, cost, Dis.subcon, Tot.Invoiced)
 temp$Job.Number<- as.character(temp$Job.Number)
-temp1<- temp[1:2426,]
+temp1<- temp[1:2425,]
 #string split to create another column with the first part of a job number. For the first half of jobs just the first bit indicates it could be invoiced together ie 2002.001 by itself
 temp1$mlsto<-sapply(temp1$Job.Number,FUN=function(x){
-        paste(strsplit(x,split='[.]')[[1]][1], strsplit(x,split='[.]')[[1]][2], sep=".")})
+        substr(x,1,8)})
 #Then ddply using this shortened job number to group. Add up all balance, profit hours in shortened job number. MErge back with original data set.
 temp1<- merge(temp1, ddply(temp1, .(mlsto), summarise, profit.mlsto= sum(profit), balance.mlsto = sum(balance), hrs.mlsto = sum(hours), cost.mlsto=sum(cost),
                            dis.sc.mlsto = sum(Dis.subcon),  inv.mlsto=sum(Tot.Invoiced)),
[…]

### 2015-07-08 hours_cleaning.R

@@ -205,171 +205,7 @@ hours1[hours1$Work.Date %in% '4/12/1912', names(hours1) %in% 'Work.Date'] <- '12
 #delete User column
 hours1<- hours1 %>% select(-User, -input.by)
 write.csv(hours1, 'hours_clean.csv')
-
-hours1<- read.csv('hours_clean.csv')[,-1]
-
-
-############ ENGINEER MORE VARIABLES  #####################
-
-#calculate percentage of hours by each position for each MILESTONE, not JOB NUMBER - 2015-06-22
-#merge mlsto by Job.Number - use file all5a
-
-#create new column for milestone prefix
-hours1<- hours1 %>% arrange(Job.num)
-temp1<- hours1[1:77588,]
-#string split to create another column with the first part of a job number. For the first half of jobs just the first bit indicates it could be invoiced together ie 2002.001 by itself
-temp1$mlsto<-sapply(temp1$Job.num,FUN=function(x){
-        substr(x,1,8)})
-#do the same with second half of data
-temp2<- hours1[77589:120508,]
-temp2$mlsto<-sapply(temp2$Job.num,FUN=function(x){
-        substr(x,1,10)})
-temp2$mlsto<- ifelse(temp2$mlsto == '2011.072.3', temp2$Job.num, temp2$mlsto)
-#rbind temp1 and temp2 back together to make temp
-hours1<- rbind(temp1, temp2)
-hours1$mlsto<- as.factor(hours1$mlsto)
-
-# number of diff people who worked on a milestone
-V1<- hours1[!(hours1$user.code %in% c('disburse', 'Admin')), names(hours1) %in% c('mlsto', 'user.code')]
-V1 <- V1 %>% arrange(mlsto) %>%
-        unique() 
-V1a<- ddply(V1, .(mlsto), nrow)
-#rename columns
-colnames(V1a)[names(V1a) %in% 'V1']<-'no.users'
-V1a %>% arrange(-no.users) %>% head()
-
-
-
-# variables: % hours by each position
-
-V2<- hours1[!(hours1$user.code %in% c('disburse', 'Admin')), 
-            names(hours1) %in% c('mlsto', 'Position', 'Hours' )]
-
-#reshape data - sum number of hours for each positon, then convert into percentage of total hours
-
[…]

### 2015-07-09 cluster.rmd
10 clusters

@@
 
-qqPlot(all7c$suc,
+qqPlot(all7c$return.pdol,
        main='success')
 
-qqPlot((all7c$suc+102700) %>% log)
+qqPlot(sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)}))
+qqPlot((all7c$suc) %>% math.cbrt)
 
@@
 
-ggplot(all7c, aes_string(x='suc')) + geom_histogram(aes(y=..density..))
+ggplot(all7c, aes_string(x='inv.vs.cost')) + geom_histogram(aes(y=..density..))
 
@@
 
-trial<- (all7c$suc) %>% as.data.frame
-ggplot(trial, aes_string(x='.')) + geom_histogram(aes(y=..density..))
+trial<- (sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)})) %>% as.data.frame
+colnames(trial)[names(trial) %in% '.']<-'var'
+ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
+skewness(trial$var)
+kurtosis(trial$var)
+mean(trial$var)
+sd(trial$var)
+trial$scale.var<- scale(trial$var)
+ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
+# ggplot still looks bimodal after scaling
 
-### CAN WE TAKE LOG OF ABSOLUTE VALUES AND THEN MAKE NUMBERS THAT WERE PREVIOUSLY NEG NEG AGAIN?
+## conclusion - lets take neglog of suc and scale it, add it to clust
 
-clust$suc<- scale(all7c$suc)
+clust$suc<- sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)}) %>% scale()
+
+## should we make return.pdol bimodal??
+math.cbrt <- function(x) {
+        sign(x) * abs(x)^(1/3)
+}
+qqPlot(all7c$return.pdol %>% math.cbrt,
+       main='success')
+trial<- (all7c$return.pdol %>% math.cbrt) %>% as.data.frame
+colnames(trial)[names(trial) %in% '.']<-'var'
+ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
+skewness(trial$var)
+kurtosis(trial$var)
+trial$scale.var<- scale(trial$var)
+ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
+#take cube root of return.pdol and scale it to clust
+
+clust$return.pdol<- all7c$return.pdol %>% math.cbrt %>% scale()
 
@@
 
-Repeat clustering with new suc
+Repeat clustering with new suc and revised bi-modal return.pdol
 
@@
 clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
-                                      suc, pc.pro, hrs.mlsto
-                                      ), 2, 20, iter= 100)
+                                      suc, hrs.mlsto), 2, 15, iter= 100)
 plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
@@
         theme(axis.text.x=element_text(angle=45,hjust=1))
-# 17 clusters should be good
+
+# 10 clusters should be good
+
+# have a look at dendrogram with 10 clusters
+
+x<- dist(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 
+         method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+
+rect.hclust(fit, k=10, border='red')
+
+#looks good!
+
 ```
+
+now want 10 clusters, based on k means optimisation, have a look at average characteristics
+
+```{r kmeans, echo=FALSE}
+
+k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 10)
+
+#try plotting means of actual values
+
+all7c$groups<- k1$cluster
+kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
+                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
+                         cost.mlsto, return.pdol, suc, groups)
+all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
+all7c.plot$groups<- rownames(all7c.plot)
+k2.melt<- melt(all7c.plot)
+
+#summary means of all numeric variables in 5 clusters
+clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
+tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
+bal<- sum(all7c$balance.mlsto)
+hrs<-  sum(all7c$hrs.mlsto)
+clust.summary$pc.jobs<- round( tab[,2]*100, 1)
+clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
+        select(2)
+clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
+        select(2)
+clust.summary$job.count<- table(all7c$groups) %>% as.data.frame %>% select(Freq)
+
+clust.summary<- clust.summary %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto, hrs.mlsto, hours.perday, timespan, Num.days, Num.disc, no.users, pc.pro, mean.peeps, client.totinv, pc.jobs, bal.pc, hrs.pc)
+
+# add medians
+all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))
+
+all7c.med<- all7c.med %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto, hrs.mlsto, hours.perday, timespan, Num.days, Num.disc, no.users, pc.pro, mean.peeps, client.totinv)
+
+```

### 2015-07-10 cluster.rmd
finished clustering - 11 clusters

@@
 # opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
-
-all7c<- read.csv('all7c.csv')[,-1]
+opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
+all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
 all7c$Post.Code<- as.factor(all7c$Post.Code)
@@
 
-```{r final_var, echo=FALSE}
+```{r final_var, include=FALSE}
 
-all7d<- all7c %>% select(inv.mlsto, timespan, no.users, Discipline, pc.pro, client.totinv, code.director, Num.disc, Business,
+all7d<- all7c %>% select(inv.mlsto, timespan, no.users, Discipline, pc.pro, client.totinv, code.director, 
+                         Num.disc, Business,
                          Biz.type, code.contact, pc.majpos,
-                         code.client, Project.Value, JD.Second, Billing.Type, Post.Code, code.ProjEng, Billing.Type, majority.pos)
+                         code.client, Project.Value, JD.Second, Billing.Type, Post.Code, code.ProjEng, 
+                         Billing.Type, majority.pos)
 
@@
 
-```{r, echo=FALSE}
+```{r, include=FALSE}
 all7c[,!sapply(all7c, is.factor)] %>% colnames
@@
 
-```{r var_check, echo=FALSE}
+```{r var_check, include=FALSE}
 
@@
 
-```{r clust final, echo=FALSE}
+```{r clust final, include=FALSE}
 clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
@@
 #now to standardise
-clust<- scale(clust[,-1]) %>% as.data.frame
+clust<- cbind(scale(clust[,-1]) %>% as.data.frame, 'mlsto'=clust$mlsto)
 ```
@@
 
-```{r hierarch, echo=FALSE}
+```{r hierarch, include=FALSE}
 
@@
 
-x<- dist(clust, method = 'euclidean')
+x<- dist(clust %>% select(-mlsto), method = 'euclidean')
 fit<- hclust(x, method = 'ward.D2')
@@
 
-```{r kmeans, echo=FALSE}
+```{r kmeans, include=FALSE}
 
-k1<- kmeans(clust %>% select(-groups), 5)
-k1.centres<- k1$centers %>% as.data.frame
-k1.centres$group<- rownames(k1.centres)
-k1.melt<- melt(k1.centres)
-k1.melt$group<- k1.melt$group %>% as.factor
-ggplot(k1.melt, aes(x=variable, y= value, fill=variable)) + geom_bar(stat='identity', position='dodge')+
-        facet_wrap(~group) +
-        theme(axis.text.x=element_text(angle=45,hjust=1))
-
-#try plotting means of actual values
-
-all7c$groups<- k1$cluster
-kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
-                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
-                         cost.mlsto, return.pdol, groups)
-all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
-all7c.plot$groups<- rownames(all7c.plot)
-k2.melt<- melt(all7c.plot)
-
-#summary means of all numeric variables in 5 clusters
-clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
-tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
-bal<- sum(all7c$balance.mlsto)
-hrs<-  sum(all7c$hrs.mlsto)
-clust.summary$pc.jobs<- round( tab[,2]*100, 1)
-clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
-        select(2)
-clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
-        select(2)
-
-clust.summary<- clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]
-
-# add medians
-all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))
-
-all7c.med<- all7c.med[, c(2,13,11,12,14,10,3,8,7,4,5,9,6)]
+# k1<- kmeans(clust %>% select(-groups, -mlsto), 5)
+# k1.centres<- k1$centers %>% as.data.frame
+# k1.centres$group<- rownames(k1.centres)
+# k1.melt<- melt(k1.centres)
+# k1.melt$group<- k1.melt$group %>% as.factor
+# ggplot(k1.melt, aes(x=variable, y= value, fill=variable)) + geom_bar(stat='identity', position='dodge')+
+#         facet_wrap(~group) +
+#         theme(axis.text.x=element_text(angle=45,hjust=1))
+# 
+# #try plotting means of actual values
+# 
+# all7c$groups<- k1$cluster
+# kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
+#                          Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
+#                          cost.mlsto, return.pdol, groups)
+# all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
+# all7c.plot$groups<- rownames(all7c.plot)
+# k2.melt<- melt(all7c.plot)
+# 
+# #summary means of all numeric variables in 5 clusters
+# clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
+# tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
+# bal<- sum(all7c$balance.mlsto)
+# hrs<-  sum(all7c$hrs.mlsto)
+# clust.summary$pc.jobs<- round( tab[,2]*100, 1)
+# clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
+#         select(2)
+# clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
+#         select(2)
+# 
+# clust.summary<- clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]
+# 
+# # add medians
+# all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))
+# 
+# all7c.med<- all7c.med[, c(2,13,11,12,14,10,3,8,7,4,5,9,6)]
 
@@
 
-```{r optimise clustnum,echo=FALSE}
+```{r optimise clustnum,include=FALSE}
 library('fpc', lib='C:/Progra~1/R/R-3.2.0/library')
@@
 ## try out all iffy variables together, eliminate duds
-clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
-                                      client.totinv, Num.days, hrs.mlsto, pc.pro, mean.peeps,
-                                      Num.disc), 2, 20, iter= 100)
-plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
-# all together, no good
-# smaller subset of variables I am most interested in that still works for clustering:
-
-clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
-                                      hrs.mlsto
-                                      ), 2, 20, iter= 100)
-plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
-
-# 17 clusters
-# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
-# then calinksi (index of cluster homogeneity and distance where higher number is better)
-# calinski: between cluster distance/within cluster distance
-ind.plot<- melt(clust.km$results)
-ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
-        geom_line( ) +
-        theme(axis.text.x=element_text(angle=45,hjust=1))
-ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
-        geom_line( )+
-        theme(axis.text.x=element_text(angle=45,hjust=1))
+# clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
+#                                       client.totinv, Num.days, hrs.mlsto, pc.pro, mean.peeps,
+#                                       Num.disc), 2, 20, iter= 100)
+# plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
+# # all together, no good
+# # smaller subset of variables I am most interested in that still works for clustering:
+# 
+# clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
+#                                       hrs.mlsto
+#                                       ), 2, 20, iter= 100)
+# plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
+# 
+# # 17 clusters
+# # Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
+# # then calinksi (index of cluster homogeneity and distance where higher number is better)
+# # calinski: between cluster distance/within cluster distance
+# ind.plot<- melt(clust.km$results)
+# ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
+#         geom_line( ) +
+#         theme(axis.text.x=element_text(angle=45,hjust=1))
+# ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
+#         geom_line( )+
+#         theme(axis.text.x=element_text(angle=45,hjust=1))
 # 17 clusters should be good
@@
 
-```{r new var, echo=FALSE}
+```{r new var, include=FALSE}
 #new variable for all7c - success
@@
 all7c<- transform(all7c, suc= inv.mlsto*return.pdol)
-all7c$suc %>% summary
+# all7c$suc %>% summary
 
-all7c %>% arrange(-suc) %>% head
+# all7c %>% arrange(-suc) %>% head
 
@@
 qqPlot(sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)}))
-qqPlot((all7c$suc) %>% math.cbrt)
-
-#can't log because of negatives
 
@@
 
-# 10 clusters
+# 11 clusters
 # Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
@@
 # calinski: between cluster distance/within cluster distance
-ind.plot<- melt(clust2.km$results)
+ind.plot<- melt(clust.km$results)
 ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
@@
 
-# 10 clusters should be good
+# 11 clusters should be good
 
-# have a look at dendrogram with 10 clusters
+# have a look at dendrogram with 11 clusters
 
@@
 
-rect.hclust(fit, k=10, border='red')
+rect.hclust(fit, k=11, border='red')
 
@@
 
-now want 10 clusters, based on k means optimisation, have a look at average characteristics
+now want 11 clusters, based on k means optimisation, have a look at average characteristics
 
-```{r kmeans, echo=FALSE}
+```{r, echo=FALSE}
 
-k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 10)
+k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 11)
 
@@
 
-clust.summary<- clust.summary %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto, hrs.mlsto, hours.perday, timespan, Num.days, Num.disc, no.users, pc.pro, mean.peeps, client.totinv, pc.jobs, bal.pc, hrs.pc)
+clust.summary<- clust.summary %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto,
+                                         hrs.mlsto, hours.perday, timespan, Num.days, Num.disc,
+                                         no.users, pc.pro, mean.peeps, client.totinv, pc.jobs,
+                                         bal.pc, hrs.pc, job.count)
 
@@
 
+clust.summary<- cbind(all7c.med %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto,
+                                     hrs.mlsto, hours.perday, timespan, Num.days, Num.disc,
+                                     no.users, pc.pro, mean.peeps, client.totinv),
+                      clust.summary %>% select(pc.jobs,bal.pc, hrs.pc, job.count))
+
+clust.summary$bal<- ddply(all7c, .(groups), summarise, bal= round(sum(balance.mlsto), 1)) %>% 
+        select(2)
+
 ```

### 2015-07-10 cluster.rmd
cluster refine

@@
 
-clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
-                                      suc, hrs.mlsto), 2, 15, iter= 100)
+clust.km<- cascadeKM(clust %>% select(inv.mlsto,suc), 2, 20, iter= 100)
 plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
@@
 
-rect.hclust(fit, k=11, border='red')
+rect.hclust(fit, k=15, border='red')
 
@@
 
-k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 11)
+k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 15)
 
@@
 # add medians
-all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))
+all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median)) %>% round(2)
 
@@
 
+#investigate clusters
+all7c %>% filter(groups == 5) %>% select(inv.mlsto, suc, pc.majpos, majority.pos,
+                                         code.client, client.meaninv, code.director, code.ProjEng,
+                                         return.pdol, balance.mlsto, Year, timespan, pc.pro,
+                                         pc.director, JD.Second, Business, client.count, Discipline,
+                                         code.client, Billing.Type) %>% summary
+
 ```
+
+Visualise confidence intervals in terms of 'success'
+
+```{r CI, echo=FALSE}
+library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
+# principal component analysis
+#convert cluster groups to integer
+
+all7c$groups<- as.integer(all7c$groups)
+p<- prcomp(~suc+ inv.mlsto + balance.mlsto + cost.mlsto + return.pdol, data=all7c, centre=T)
+# summary(p)
+biplot(p, xlabs = as.character(all7c$groups))
+#putlers in-house biplot shows centroids
+bpCent(p, clsAsgn = all7c$groups, centroids=TRUE, xlabs = as.character(all7c$groups))
+
+#now look at 3D - triplot - using prcomp
+#scores stored as 'x' variable
+p$x
+plot3d(p$x[,1:3], col=as.character(all7c$groups))
+text3d(p$x[,1:3], texts = as.character(all7c$groups), color=as.character(all7c$groups), adj=2)
+text3d(p$rotation[,1:3], text = rownames(p$rotation), color= 'red')
+coords <- NULL
+for (i in 1:nrow(p$rotation)) {
+        coords <- rbind(coords, rbind(c(0,0,0), p$rotation[i,1:3]))
+}
+lines3d(coords, col="red", lwd=4)
+
+```

### 2015-07-13 cluster.rmd
clustering

--- 2015-07-10 cluster.rmd
@@
 sd(trial$var)
-trial$scale.var<- scale(trial$var)
+trial$scale.var<- trial$var/sd(trial$var)
 ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
@@
 
-clust$suc<- sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)}) %>% scale()
+clust$suc<- sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)})
+clust$suc<- clust$suc/sd(clust$suc)
 
@@
 kurtosis(trial$var)
-trial$scale.var<- scale(trial$var)
+trial$scale.var<- trial$var/sd(trial$var)
 ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
@@
 
-clust$return.pdol<- all7c$return.pdol %>% math.cbrt %>% scale()
+clust$return.pdol<- math.cbrt(all7c$return.pdol)
+clust$return.pdol<- all7c$return.pdol/sd(all7c$return.pdol)
+
+## should we make balance.mlsto bimodal??
+
+qqPlot(all7c$balance.mlsto,
+       main='balance')
+qqPlot(sapply(all7c$balance.mlsto, function(x){sign(x)*log(abs(x) + 1)}),
+       main='balance')
+trial<- (sapply(all7c$balance.mlsto, function(x){sign(x)*log(abs(x) + 1)})) %>% as.data.frame
+colnames(trial)[names(trial) %in% '.']<-'var'
+ggplot(trial, aes_string(x='var')) + geom_histogram(aes(y=..density..))
+skewness(trial$var)
+kurtosis(trial$var)
+trial$scale.var<- trial$var/sd(trial$var)
+ggplot(trial, aes_string(x='scale.var')) + geom_histogram(aes(y=..density..))
+#take cube root of return.pdol and scale it to clust
+
+clust$balance.mlsto<- sapply(all7c$balance.mlsto, function(x){sign(x)*log(abs(x) + 1)})
+clust$balance.mlsto<- clust$balance.mlsto/sd(clust$balance.mlsto)
 
@@
 
-clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
-                                      suc, hrs.mlsto), 2, 15, iter= 100)
+clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto), 2, 30, iter= 100)
 plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
@@
 
-# 11 clusters should be good
+# 16 clusters should be good
 
-# have a look at dendrogram with 11 clusters
+# have a look at dendrogram with 16 clusters
 
-x<- dist(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 
+x<- dist(clust %>% select(inv.mlsto, balance.mlsto), 
          method = 'euclidean')
@@
 
-rect.hclust(fit, k=11, border='red')
+rect.hclust(fit, k=16, border='red')
 
@@
 
-now want 11 clusters, based on k means optimisation, have a look at average characteristics
+now want 16 clusters, based on k means optimisation, have a look at average characteristics
 
@@
 
-k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 11)
+k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto), 16)
 
@@
 
-all7c$groups<- k1$cluster
+all7c$group.num<- k1$cluster
 kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                          Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
-                         cost.mlsto, return.pdol, suc, groups)
-all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
-all7c.plot$groups<- rownames(all7c.plot)
+                         cost.mlsto, return.pdol, suc, group.num)
+all7c.plot<- ddply(kmeans.plot, .(group.num), numcolwise(mean))
+all7c.plot$group.num<- rownames(all7c.plot)
 k2.melt<- melt(all7c.plot)
@@
 #summary means of all numeric variables in 5 clusters
-clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
-tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
+clust.summary<- all7c.plot %>% select(-group.num) %>% round(2)
+tab<- table(all7c$group.num) %>% prop.table %>% as.data.frame
 bal<- sum(all7c$balance.mlsto)
@@
 clust.summary$pc.jobs<- round( tab[,2]*100, 1)
-clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
+clust.summary$bal.pc<- ddply(all7c, .(group.num), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
         select(2)
-clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
+clust.summary$hrs.pc<- ddply(all7c, .(group.num), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
         select(2)
-clust.summary$job.count<- table(all7c$groups) %>% as.data.frame %>% select(Freq)
+clust.summary$job.count<- table(all7c$group.num) %>% as.data.frame %>% select(Freq)
 
@@
 # add medians
-all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))
+all7c.med<- ddply(kmeans.plot, .(group.num), numcolwise(median)) %>% round(2)
 
@@
 
-clust.summary$bal<- ddply(all7c, .(groups), summarise, bal= round(sum(balance.mlsto), 1)) %>% 
+clust.summary$bal<- ddply(all7c, .(group.num), summarise, bal= round(sum(balance.mlsto), 1)) %>% 
         select(2)
 
+#investigate clusters
+all7c[all7c$group.num == 12,] %>% select(inv.mlsto, suc, pc.majpos, majority.pos,
+                                         code.client, client.meaninv, code.director, code.ProjEng,
+                                         return.pdol, balance.mlsto, Year, timespan, pc.pro,
+                                         pc.director, JD.Second, Business, client.count, Discipline,
+                                         code.client, Billing.Type) %>% summary
+
 ```
+
+Visualise confidence intervals in terms of 'success'
+
+```{r CI, echo=FALSE}
+library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
+# principal component analysis
+
+#add clust columns to all7c
+all7c$clust.inv<- clust$inv.mlsto
+all7c$clust.bal<- clust$balance.mlsto
+all7c$clust.cost<- clust$cost.mlsto
+all7c$clust.return<- clust$return.pdol
+all7c$clust.hours<- clust$hrs.mlsto
+all7c$clust.suc<- clust$suc
+
+#convert cluster groups to integer
+
+all7c$group.num<- as.integer(all7c$group.num)
+p<- prcomp(~clust.inv + clust.bal+ clust.cost, data=all7c, centre=T)
+# summary(p)
+biplot(p, xlabs = as.character(all7c$group.num))
+#putlers in-house biplot shows centroids
+bpCent(p, clsAsgn = all7c$group.num, centroids=TRUE, xlabs = as.character(all7c$Business))
+
+##try ggbiplot
+library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
+#visualise clusters
+q<- ggbiplot(p, obs.scale = 1, var.scale = 1, 
+              groups = all7c$group.num %>% as.factor, ellipse = TRUE,
+             labels = all7c$Business,
+              circle = TRUE)
+q <- q + scale_color_discrete(name = '')
+q <- q + theme(legend.direction = 'vertical', 
+               legend.position = 'right')
+q
+
+#visualise factors within clusters
+g<- ggbiplot(p, obs.scale = 1, var.scale = 1, 
+              groups = all7c$Business %>% as.factor(), ellipse = TRUE, 
+              circle = TRUE)
+g <- g + scale_color_discrete(name = '')
+g <- g + theme(legend.direction = 'vertical', 
+               legend.position = 'right')
+g
+
+#confusion matrix of groups vs. categorical variables
+
+#Business
+table(all7c$group.num, all7c$Business)
+#notes
+# only single NFP - combine to institution
+all7c[all7c$Business %in% 'NFP',]$Business<- 'institution'
+#combine person and friend
+all7c[all7c$Business %in% 'friend',]$Business<- 'person'
+#fabricator labelled manufactorer/supplier
+all7c[all7c$Business %in% 'fabricator',]$Business<- 'manufacturer/supplier'
+all7c[all7c$Business %in% 'membrane fabricator',]$Business<- 'manufacturer/supplier'
+#combine remedial building services to building services
+all7c[all7c$Business %in% 'remedial building service',]$Business<- 'building services'
+#combine gov and council
+all7c[all7c$Business %in% 'council',]$Business<- 'gov'
+all7c$Business<- droplevels(all7c$Business)
+
+```
+
+View counts of each category in each group as a heatmap
+
+Order groups by profitability and size

### 2015-07-13 cluster.rmd
heatmaps

@@
 
-library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+# library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
+# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
+# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
+# library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
+# library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
 
-# library('knitr')
-# library('ggplot2')
-# library("dplyr")
-# library("plyr")
-# library('magrittr')
-# library('reshape2')
-# library("rpart")
-# library('car')
-# library('e1071')
-# library('corrgram')
-# library('party')
-# library('dendextend')
+# library('RcmdrPlugin.BCA')
+library(devtools)
+library('ggbiplot')
+library('fpc')
+library('vegan')
+library('knitr')
+library('ggplot2')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library("rpart")
+library('car')
+library('e1071')
+library('corrgram')
+library('party')
+library('dendextend')
 # 
@@
 opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
-all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
+# all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
+all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
 all7c$Post.Code<- as.factor(all7c$Post.Code)
@@
 
-x<- dist(clust %>% select(-mlsto), method = 'euclidean')
-fit<- hclust(x, method = 'ward.D2')
-plot(fit)
-
-rect.hclust(fit, k=5, border='red')
+# x<- dist(clust %>% select(-mlsto), method = 'euclidean')
+# fit<- hclust(x, method = 'ward.D2')
+# plot(fit)
+# 
+# rect.hclust(fit, k=5, border='red')
 #5 clusters looks pretty good
@@
 #how many points in each cluster?
-clust$groups<- cutree(fit,k=5)
-table(clust$groups)
+# clust$groups<- cutree(fit,k=5)
+# table(clust$groups)
 
@@
 ```{r optimise clustnum,include=FALSE}
-library('fpc', lib='C:/Progra~1/R/R-3.2.0/library')
-library('vegan', lib = 'C:/Progra~1/R/R-3.2.0/library')
 
 #try clustering a low number of numeric variables, trialling iffy ones to see which ones are good for clustering
-clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
-                                      hrs.mlsto), 2, 20, iter= 100)
-plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
+# clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
+#                                       hrs.mlsto), 2, 20, iter= 100)
+# plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
 # client.totinv indicating about 4 clusters
@@
 
-k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto), 16)
+k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto), 15)
 
@@
 ```{r CI, echo=FALSE}
-library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
+
 # principal component analysis
@@
 #putlers in-house biplot shows centroids
-bpCent(p, clsAsgn = all7c$group.num, centroids=TRUE, xlabs = as.character(all7c$Business))
+# bpCent(p, clsAsgn = all7c$group.num, centroids=TRUE, xlabs = as.character(all7c$Business))
 
 ##try ggbiplot
-library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
+
 #visualise clusters
@@
               groups = all7c$group.num %>% as.factor, ellipse = TRUE,
-             labels = all7c$Business,
+             labels = all7c$group.num,
               circle = TRUE)
@@
 #confusion matrix of groups vs. categorical variables
+#reorder cluster tag levels from most amount and neg to best
+all7c$group.num<- as.factor(all7c$group.num)
+levs<- data.frame('clust' = rownames(clust.summary), 'bal' = clust.summary$balance.mlsto)
+all7c$group.num <- factor(all7c$group.num, levels=levs[order(levs$bal),]$clust)
 
@@
 
+#Biz.type
+table(all7c$group.num, all7c$Biz.type)
+#JD.Second table
+table(all7c$group.num, all7c$JD.Second)
+#combine JD.Second categories
+all7c$JD.Second<- as.character(all7c$JD.Second)
+#combine waste water and sewerage
+all7c[all7c$JD.Second %in% 'sewerage',]$JD.Second<- 'waste water'
+#combine re-use (water) and recycled water
+all7c[all7c$JD.Second %in% 're-use (water)',]$JD.Second<- 'recycled water'
+#combine water cycle management and water management and water strategy and flood studies
+all7c[all7c$JD.Second %in% c('water cycle management', 'water strategy', 'flood studies'),]$JD.Second<-
+        'water management'
+#combine apartments and townhouses and community housing
+all7c[all7c$JD.Second %in% c('townhouses','community housing','apartments'),]$JD.Second<- 'multi-dwelling'
+#combine childcare and community
+all7c[all7c$JD.Second %in% c('childcare', 'religious'),]$JD.Second<- 'community'
+#combine commercial shop and dining
+all7c[all7c$JD.Second %in% c('commercial shop','dining'),]$JD.Second<- 'dining/retail'
+all7c$JD.Second<- as.factor(all7c$JD.Second)
+all7c$JD.Second<- droplevels(all7c$JD.Second)
+
+#Biz.type
+table(all7c$group.num, all7c$pc.majpos)
 ```
@@
 Order groups by profitability and size
+
+```{r heatmaps, echo=FALSE}
+#function for heatmap
+
+heatmap<- function(colour = 'deeppink3', clust = 'group.num', var = 'JD.Second', df=all7c, percentof_clust=FALSE){
+
+        #order var levels by highest amount invoiced total
+        a = ddply(df, c(var), summarise, summ = sum(inv.mlsto))
+        df[,var] = factor(df[,var], levels= a[order(-a$summ),][,var])
+
+        if(percentof_clust== FALSE){
+                #table with percentages within each category
+                tbl= round(prop.table(table(df[,clust],df[,var]),1)*100,2) %>% as.data.frame
+                p = ggplot(tbl, aes(x=Var2, y=Var1)) +
+                        geom_tile(aes(fill=Freq), colour = 'white') +
+                        scale_fill_gradient(low='white', high=colour, name = 'percent') + 
+                        labs(x='', y='', title= 'percentage of jobs in each cluster') +
+                        theme(axis.ticks=element_blank(),
+                              axis.text.x=element_text(angle=45,hjust=1))+
+                        scale_x_discrete(expand = c(0, 0)) + scale_y_discrete(expand = c(0, 0)) 
+        }
+        else{
+                #table with percentages within each cluster
+                tbl= round(prop.table(table(df[,clust],df[,var]),2)*100,2) %>% as.data.frame
+                p = ggplot(tbl, aes(x=Var2, y=Var1)) +
+                        geom_tile(aes(fill=Freq), colour = 'white') +
+                        scale_fill_gradient(low='white', high=colour, name = 'percent') + 
+                        labs(x='Profit measure', y='', title= 'percentage of jobs in each X-axis category') +
+                        theme(axis.ticks=element_blank(),
+                              axis.text.x=element_text(angle=45,hjust=1))+
+                        scale_x_discrete(expand = c(0, 0)) + scale_y_discrete(expand = c(0, 0))       
+        }
+        print(p)
+}
+
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'JD.Second', df=all7c, percentof_clust=FALSE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'JD.Second', df=all7c, percentof_clust=TRUE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Business', df=all7c, percentof_clust=FALSE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Business', df=all7c, percentof_clust=TRUE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Num.disc', df=all7c, percentof_clust=FALSE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'majority.pos', df=all7c, percentof_clust=FALSE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'majority.pos', df=all7c, percentof_clust=TRUE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Billing.Type', df=all7c, percentof_clust=TRUE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Discipline', df=all7c, percentof_clust=TRUE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Discipline', df=all7c, percentof_clust=FALSE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.client', df=all7c, percentof_clust=FALSE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.client', df=all7c, percentof_clust=TRUE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.contact', df=all7c, percentof_clust=FALSE)
+heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.contact', df=all7c, percentof_clust=TRUE)
+
+```

### 2015-07-16 cluster.rmd
knn start

@@
 library('dendextend')
+library('FNN')
+
 # 
@@
 JD.Second
-Billing.Type
 Post.Code
@@
 
-all7d<- all7c %>% select(inv.mlsto, timespan, no.users, Discipline, pc.pro, client.totinv, code.director, 
+all7d<- all7c %>% select(inv.mlsto, timespan, no.users, Discipline, pc.pro, 
+                         client.totinv, code.director, 
                          Num.disc, Business,
                          Biz.type, code.contact, pc.majpos,
-                         code.client, Project.Value, JD.Second, Billing.Type, Post.Code, code.ProjEng, 
+                         code.client, Project.Value, JD.Second, Post.Code, code.ProjEng, 
                          Billing.Type, majority.pos)
@@
 ```
+
+Say we are trying to retrieve the nearest neighbours.
+There will be two types of distances:
+
+* numerical nearest neighbours
+* categorical nearest neighbours
+
+Which numerical categories would we want to include in the nearest neighbour search?
+
+No NA:
+* inv.mlsto
+* timespan
+* no.users
+* Num.disc
+* pc.pro
+* client.totinv
+Discipline
+code.director
+Business
+Biz.type
+
+Many NA:
+
+* pc.majpos
+* Project.Value
+code.contact
+code.client
+JD.Second
+Post.Code
+code.ProjEng
+Billing.Type
+majority.pos

### 2015-07-17 cluster.rmd
knn start

@@
 
-# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
-# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
-# library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
-# library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
+library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
+library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
+library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
 
 # library('RcmdrPlugin.BCA')
-library(devtools)
-library('ggbiplot')
-library('fpc')
-library('vegan')
-library('knitr')
-library('ggplot2')
-library("plyr")
-library("dplyr")
-library('magrittr')
-library('reshape2')
-library("rpart")
-library('car')
-library('e1071')
-library('corrgram')
-library('party')
-library('dendextend')
-library('FNN')
+# library(devtools)
+# library('ggbiplot')
+# library('fpc')
+# library('vegan')
+# library('knitr')
+# library('ggplot2')
+# library("plyr")
+# library("dplyr")
+# library('magrittr')
+# library('reshape2')
+# library("rpart")
+# library('car')
+# library('e1071')
+# library('corrgram')
+# library('party')
+# library('dendextend')
+# library('FNN')
 
@@
 # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
 # opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
 opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
-# all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
-all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
+all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
+# all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
 all7c$Post.Code<- as.factor(all7c$Post.Code)
@@
 majority.pos
+
+Retrieve nearest neighbour
+
+```{r knn, echo=FALSE}
+
+```

### 2015-07-20 cluster.rmd
first inn

--- 2015-07-16 cluster.rmd
@@
 
-# library('RcmdrPlugin.BCA')
+library('RcmdrPlugin.BCA')
 library(devtools)
@@
 library('FNN')
+library('caret')
 
@@
 # opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
-opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
+# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
 # all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
@@
 
-all7d<- all7c %>% select(inv.mlsto, timespan, no.users, Discipline, pc.pro, 
+all7d<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, Discipline, pc.pro, 
                          client.totinv, code.director, 
@@
                          code.client, Project.Value, JD.Second, Post.Code, code.ProjEng, 
-                         Billing.Type, majority.pos)
+                         Billing.Type, majority.pos, return.pdol, cost.mlsto, hrs.mlsto)
 
@@
 majority.pos
+
+Retrieve nearest neighbour - numeric first
+
+```{r knn, echo=FALSE}
+#pretend you are predicting .. post code
+
+inv.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3){
+        inv.knn = knn(df[,predict], df[new.cases,predict],
+                df$pc.pro, k=k)
+        indices<- attr(inv.knn,  "nn.index")
+        #checkout the nearest neigbours!
+        df %>% slice(c(new.cases,indices[1,])) %>% print
+
+        }
+inv.knn(df= all7d, predict= 'inv.mlsto', new.cases= 2000, k=3)
+
+```

### 2015-07-21 cluster.rmd
transferred Role info to JD.Second

@@
 
-# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
-# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
-# library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
-# library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
+library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
+library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
+library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
+library('FNN', lib='C:/Progra~1/R/R-3.2.1/library')
+library('caret', lib='C:/Progra~1/R/R-3.2.1/library')
 
 library('RcmdrPlugin.BCA')
-library(devtools)
-library('ggbiplot')
-library('fpc')
-library('vegan')
-library('knitr')
-library('ggplot2')
-library("plyr")
-library("dplyr")
-library('magrittr')
-library('reshape2')
-library("rpart")
-library('car')
-library('e1071')
-library('corrgram')
-library('party')
-library('dendextend')
-library('FNN')
-library('caret')
+# library(devtools)
+# library('ggbiplot')
+# library('fpc')
+# library('vegan')
+# library('knitr')
+# library('ggplot2')
+# library("plyr")
+# library("dplyr")
+# library('magrittr')
+# library('reshape2')
+# library("rpart")
+# library('car')
+# library('e1071')
+# library('corrgram')
+# library('party')
+# library('dendextend')
+# library('FNN')
+# library('caret')
 
@@
 # opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
-# all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
-all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
+all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
+# all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
 all7c$Post.Code<- as.factor(all7c$Post.Code)
@@
 }
+
 inv.knn(df= all7d, predict= 'inv.mlsto', new.cases= 2000, k=3)
@@
 ```
+
+Retrieve nearest neighbour - categorical
+Need to create binary matrix!!
+
+```{r, echo=FALSE}
+# want to use the following cat variables for knn:
+# Discipline, Business, Biz.type, code.contact, code.client, JD.Second, Billing.Type
+# in order to use model.matrix, need to convert NA's to 'unknown' factor.
+
+#trial model.matrix function
+vars.vec<- c('Discipline','Business','Biz.type','code.contact', 'code.client', 'JD.Second', 'Billing.Type')
+
+binary.build<- function(vars= vars.vec, df= all7d){
+        options("contrasts")
+        #create final df bin.knn
+        bin.knn= rep(0, nrow(df))
+        #loop through each variable in vars vector
+        for(i in 1:2){
+                temp = data.frame(df[, vars[i]])
+                colnames(temp)<- c(vars[i])
+                formula = paste('~',vars[i],'-1')
+                a= model.matrix( as.formula(formula), temp)
+                a= a[match(rownames(temp), rownames(a)),]
+                bin.knn= cbind(bin.knn,a)
+        }
+
+        print(bin.knn)
+}
+
+binary.build(vars= vars.vec, df= all7d) %>% head
+```

### 2015-07-21 var_select.rmd
more clust
transferred Role info to JD.Second

@@
 ```{r load_data, echo=FALSE}
-all6a<- read.csv('all6a.csv')[,-1]
-all6a$Start.Date<- as.Date(all6a$Start.Date)
 all7<- read.csv('all7.csv')[,-1]
@@
 # write.csv(all7a,'all7a.csv')
-all7a<- read.csv('all7a.csv')
+all7a<- read.csv('all7a.csv')[,-1]
 ```
@@
 all7b<- merge(all7b, details %>% select(code, client.age), by.x = 'code.client', by.y = 'code', all.x=TRUE, all.y=FALSE)
+
 # write.csv(all7b, 'all7c.csv')
@@
 
-write.csv(all7c, 'all7c.csv')
+# write.csv(all7c, 'all7c.csv')
 all7c<- read.csv('all7c.csv')[,-1]

### 2015-07-21 1st_stats.R

@@ -7,14 +7,14 @@ library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
 library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
 library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
 
-all6a<- read.csv('all6a.csv')[,-1]
-all6a$Start.Date<- as.Date(all6a$Start.Date)
+all6b<- read.csv('all6b.csv')[,-1]
+all6b$Start.Date<- as.Date(all6b$Start.Date)
 
 #delete milestone rows! sort by highest hours - choose this discipline.!
 #sort by hours first
-all6a<- all6a %>% arrange(mlsto, -hours)
-dup<- all6a[,names(all6a) %in% c('mlsto')]
-all7<-all6a[!(all6a$inv.mlsto==0 | duplicated(dup)),]
+all6b<- all6b %>% arrange(mlsto, -hours)
+dup<- all6b[,names(all6b) %in% c('mlsto')]
+all7<-all6b[!(all6b$inv.mlsto==0 | duplicated(dup)),]
 
 write.csv(all7, 'all7.csv')
 all7<- read.csv('all7.csv')[,-1]

### 2015-07-21 de_identify.R

@@ -190,3 +190,124 @@ write.csv(all6a,'all6a.csv')
 
 
 
+### grep through Role and alter JD.Second!!!!!!!!!!!!!!!!!!!!!!!!!!!
+
+all6a$Role %>% head
+#look at all entries in Role with less than 10 characters
+check<- sapply(all6a$Role %>% as.character, FUN=function(x) {nchar(x)}) %>% as.data.frame
+check$Role<- all6a$Role
+colnames(check)[names(check) %in% '.']<-'char'
+check %>% unique %>% filter(char==1)
+check %>% unique %>% filter(char==5)
+check$Discipline<- all6a$Discipline
+check$JD.Second<- all6a$JD.Second
+check$mlsto<- all6a$mlsto
+check<- check %>% arrange(mlsto)
+check$Job.Name<- all5a$Job.Name
+
+#anything with nchar<=4 should be 'NA'
+RN<- check[check$char<=4,] %>% rownames
+all6a[RN,]$Role <- NA
+all6a$Role <- droplevels(all6a$Role)
+
+#now lets view remaining Role entries
+all6a %>% filter(!is.na(Role)) %>% select(Role, JD.Second) %>% head
+
+#create data frame of category key words (that I made manually by sscrolling through all the data) 
+cat<- read.csv('Catgrep.csv', na.strings= "")
+cat<- t(cat)
+# delete columns with 29 NA values, rename columns using first row
+colnames(cat)<- cat[1,]
+cat<-cat[-c(1,2),]
+cat<-cat[ ,colSums(is.na(cat))!=nrow(cat)]
+#delete white space before and after
+trim <- function (x) gsub("^\\s+|\\s+$", "", x)
+cat<- trim(cat)
+
+#create data frame of category key words (that I made manually by sscrolling through all the data) 
+sec<- read.csv('second_detail.csv', na.strings= "")
+sec<- t(sec)
+# delete columns with 29 NA values, rename columns using first row
+colnames(sec)<- sec[1,]
+sec<-sec[-c(1),]
+sec<-sec[ ,colSums(is.na(sec))!=nrow(cat)]
+#delete white space before and after
+trim <- function (x) gsub("^\\s+|\\s+$", "", x)
+sec<- trim(sec)
+
+#create trimmed vectors for each column to feed into grepping
+grep.svec<-function(df=cat){
+        for (i in 1:ncol(df)){
+                tempvec= c(df[,i])
[…]

### 2015-07-22 cluster.rmd
knn

@@
 pc.majpos
-
-Create reduced dataset and have a think about what you want to cluster against each other
-
-```{r final_var, include=FALSE}
-
-all7d<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, Discipline, pc.pro, 
-                         client.totinv, code.director, 
-                         Num.disc, Business,
-                         Biz.type, code.contact, pc.majpos,
-                         code.client, Project.Value, JD.Second, Post.Code, code.ProjEng, 
-                         Billing.Type, majority.pos, return.pdol, cost.mlsto, hrs.mlsto)
-
-```
 
@@
 all7c$JD.Second<- as.character(all7c$JD.Second)
-#combine waste water and sewerage
-all7c[all7c$JD.Second %in% 'sewerage',]$JD.Second<- 'waste water'
 #combine re-use (water) and recycled water
@@
 #combine water cycle management and water management and water strategy and flood studies
-all7c[all7c$JD.Second %in% c('water cycle management', 'water strategy', 'flood studies'),]$JD.Second<-
+all7c[all7c$JD.Second %in% c('water cycle management', 'water strategy'),]$JD.Second<-
         'water management'
@@
 #Biz.type
-table(all7c$group.num, all7c$pc.majpos)
+table(all7c$group.num, all7c$Biz.type)
 ```
@@
 
+Create reduced dataset and have a think about what you want to cluster against each other
+
+```{r final_var, include=FALSE}
+
+all7d<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, Discipline, pc.pro, 
+                         client.totinv, code.director, 
+                         Num.disc, Business,
+                         Biz.type, code.contact, pc.majpos,
+                         code.client, Project.Value, JD.Second, Post.Code, code.ProjEng, 
+                         Billing.Type, majority.pos, return.pdol, cost.mlsto, hrs.mlsto)
+
+```
+
 Retrieve nearest neighbour - numeric first
@@
         #create final df bin.knn
-        bin.knn= rep(0, nrow(df))
+        bin.knn= NULL
         #loop through each variable in vars vector
-        for(i in 1:2){
+        for(i in 1:length(vars.vec)){
                 temp = data.frame(df[, vars[i]])
@@
 
-        print(bin.knn)
+        return(bin.knn)
 }
 
-binary.build(vars= vars.vec, df= all7d) %>% head
+knn.binary<- binary.build(vars= vars.vec, df= all7d)
+knn.binary<- as.data.frame(knn.binary)
+
 ```
+
+Have created binary matrix
+Turn into dataframe and then perform jaccard distance nearest neighbour
+
+```{r knn, echo=FALSE}
+#pretend you are predicting .. post code
+
+# First turn new case into giant binary dataframe using function from above
+new.case<- binary.build(vars= vars.vec, df= knn.binary %>% slice(2000))
+
+inv.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3){
+        inv.knn = knn(df[,predict], df[new.cases,predict],
+                df$pc.pro, k=k)
+        indices<- attr(inv.knn,  "nn.index")
+        #checkout the nearest neigbours!
+        df %>% slice(c(new.cases,indices[1,])) %>% print
+
+}
+
+inv.knn(df= all7d, predict= 'inv.mlsto', new.cases= 2000, k=3)
+
+```

### 2015-07-22 de_identify.R

@@ -187,12 +187,14 @@ all6a<- merge(all6a, inv.eng %>% select(code.client, client.meaninv, client.invf
               by='code.client', all.x=T)
 test<- merge(all6 %>% select(mlsto), inv.eng %>% select(mlsto), by= 'mlsto', all.x=TRUE, all.y=TRUE)
 write.csv(all6a,'all6a.csv')
+all6a<- read.csv('all6a.csv')
 
 
 
 ### grep through Role and alter JD.Second!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
 all6a$Role %>% head
+all6a$Role<- as.character(all6a$Role)
 #look at all entries in Role with less than 10 characters
 check<- sapply(all6a$Role %>% as.character, FUN=function(x) {nchar(x)}) %>% as.data.frame
 check$Role<- all6a$Role
@@ -202,7 +204,6 @@ check %>% unique %>% filter(char==5)
 check$Discipline<- all6a$Discipline
 check$JD.Second<- all6a$JD.Second
 check$mlsto<- all6a$mlsto
-check<- check %>% arrange(mlsto)
 check$Job.Name<- all5a$Job.Name
 
 #anything with nchar<=4 should be 'NA'
@@ -212,6 +213,8 @@ all6a$Role <- droplevels(all6a$Role)
 
 #now lets view remaining Role entries
 all6a %>% filter(!is.na(Role)) %>% select(Role, JD.Second) %>% head
+all6a$Role<- as.factor(all6a$Role)
+
 
 #create data frame of category key words (that I made manually by sscrolling through all the data) 
 cat<- read.csv('Catgrep.csv', na.strings= "")
@@ -278,7 +281,8 @@ all6a$Role.JD<- grepster(nstruc.list, nstruc.title, c('Civil','Environmental Pla
 
 #worked!
 #now compare when JD.Second differs from Role.JD
-all6a %>% filter(!is.na(Role.JD), !is.na(JD.Second),!(Role.JD==JD.Second)) %>% select(Role, Role.JD, JD.Second, Discipline) %>% View()
+all6a %>% filter(!is.na(Role.JD), !is.na(JD.Second),!(Role.JD==JD.Second)) %>% select(Role, Role.JD, JD.Second, Discipline,mlsto) %>% View()
+
 #2014.371.3 - keep JD.Second
 #2014.117.3 - keep JD.Second
 #2014.414.3 - keep JD.Second

### 2015-07-22 cluster.rmd
again

@@
                          code.client, Project.Value, JD.Second, Post.Code, code.ProjEng, 
-                         Billing.Type, majority.pos, return.pdol, cost.mlsto, hrs.mlsto)
+                         Billing.Type, majority.pos, return.pdol, cost.mlsto, hrs.mlsto, balance.mlsto)
+all7d$mlsto<- as.character(all7d$mlsto)
 
@@
         #checkout the nearest neigbours!
-        df %>% slice(c(new.cases,indices[1,])) %>% print
+        df %>% slice(c(new.cases, indices[1,])) %>% return
 
@@
 
-inv.knn(df= all7d, predict= 'inv.mlsto', new.cases= 2000, k=3)
+a<-inv.knn(df= all7d, predict= 'inv.mlsto', new.cases= 2000, k=3)
 
@@
 
+#NA values don't work in knn function so turn all NA into zeroes
+knn.binary[is.na(knn.binary)]<- 0
+knn.binary$pc.pro<- all7d$pc.pro
+
 ```
@@
 
-inv.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3){
+cat.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3, result.df= all7d){
         inv.knn = knn(df[,predict], df[new.cases,predict],
@@
         indices<- attr(inv.knn,  "nn.index")
+        dists<- attr(inv.knn, 'nn.dist') %>% t %>% as.data.frame %>% round(2)
         #checkout the nearest neigbours!
-        df %>% slice(c(new.cases,indices[1,])) %>% print
+        final = cbind(result.df %>% slice(c(indices[1,])), dists)
+        colnames(final)[names(final) %in% 'V1']<-'knn.dist'
+        return(final)
 
@@
 
-inv.knn(df= all7d, predict= 'inv.mlsto', new.cases= 2000, k=3)
+b<- cat.knn(df= knn.binary, predict= colnames(knn.binary), new.cases= 60, k=5, result.df= all7d)
 
 ```
+
+Combine two knn results
+
+```{r combine_knn, echo=FALSE}
+
+b<- cat.knn(df= knn.binary, predict= colnames(knn.binary), new.cases= 60, k=20, result.df= all7d)
+c<- inv.knn(df= b, predict= 'inv.mlsto', new.cases= 1, k=4) %>% slice(-1)
+c
+
+```

### 2015-07-23 cluster.rmd
starting knn app

@@
 
+write.csv(all7d,'all7d.csv')
+
 ```
@@
 inv.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3){
-        inv.knn = knn(df[,predict], df[new.cases,predict],
+        invo.knn = knn(df[,predict], df[new.cases,predict],
                 df$pc.pro, k=k)
-        indices<- attr(inv.knn,  "nn.index")
+        indices<- attr(invo.knn,  "nn.index")
         #checkout the nearest neigbours!
@@
 knn.binary$pc.pro<- all7d$pc.pro
+knn.binary$mlsto<- all7d$mlsto
 
@@
 
-cat.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3, result.df= all7d){
-        inv.knn = knn(df[,predict], df[new.cases,predict],
-                df$pc.pro, k=k)
-        indices<- attr(inv.knn,  "nn.index")
-        dists<- attr(inv.knn, 'nn.dist') %>% t %>% as.data.frame %>% round(2)
+cat.knn<- function(df= all7d, nn.by= 'inv.mlsto', predict='pc.pro', case= c(1000), k=3, result.df= all7d){
+        invo.knn = knn(df[,nn.by], df[case,nn.by],
+                df[,predict], k=k)
+        indices= attr(invo.knn,  "nn.index")
+        dists= attr(invo.knn, 'nn.dist') %>% t %>% as.data.frame %>% round(2)
         #checkout the nearest neigbours!
@@
         colnames(final)[names(final) %in% 'V1']<-'knn.dist'
+
+        #final actions
+        assign('index', indices, envir=.GlobalEnv)
         return(final)
-
 }
 
-b<- cat.knn(df= knn.binary, predict= colnames(knn.binary), new.cases= 60, k=5, result.df= all7d)
+bin<- knn.binary %>% select(-mlsto)
+b<- cat.knn(df= bin, nn.by= colnames(bin)[-length(colnames(bin))], predict = 'pc.pro', case= 60, k=5, result.df= all7d)
 
@@
 ```{r combine_knn, echo=FALSE}
-
-b<- cat.knn(df= knn.binary, predict= colnames(knn.binary), new.cases= 60, k=20, result.df= all7d)
-c<- inv.knn(df= b, predict= 'inv.mlsto', new.cases= 1, k=4) %>% slice(-1)
+bin<- knn.binary %>% select(-mlsto)
+b<- cat.knn(df= bin, nn.by= colnames(bin)[!grepl('pc.pro', colnames(bin))], predict = 'pc.pro', case= 80, k=10, result.df= all7d)
+c<- inv.knn(df= b, predict= 'inv.mlsto', new.cases= 1, k=5) %>% slice(-1)
 c
+ #step 2 for user - option to sort by code.client?
+d<- b[b[,'code.contact'] %in% b[,'code.contact'][1], ]
+d
+e<- inv.knn(df= d, predict= 'inv.mlsto', new.cases= 1, k=5) %>% slice(-1)
+e
 

### 2015-07-23 final_app/server.R

@@ -25,6 +25,71 @@ df2 <- within(df2,
               probability <- factor(probability, 
                                     levels=c('95%','75%','50%')))
 
+#### CREATE TABLE DATA FRAME FOR NEAREST NEIGHBOURS ####
+
+all7d<- read.csv('all7d.csv')
+all7d$mlsto<- as.character(all7d$mlsto)
+
+# test.case<- all7d %>% select(Discipline, Business, Biz.type, code.contact, code.client, JD.Second, Billing.Type)
+
+#function for retrieving nearest neighbours by inv.mlsto:
+inv.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3){
+        invo.knn = knn(df[,predict], df[new.cases,predict],
+                       df$pc.pro, k=k)
+        indices<- attr(invo.knn,  "nn.index")
+        #checkout the nearest neigbours!
+        df %>% slice(c(new.cases, indices[1,])) %>% return
+        
+}
+
+#function for retrieving nearest neighbours by category - must first build binary matrix
+#call binary matrix 'knn.binary'
+vars.vec<- c('Discipline','Business','Biz.type','code.contact', 'code.client', 'JD.Second', 'Billing.Type')
+
+binary.build<- function(vars= vars.vec, df= all7d){
+        options("contrasts")
+        #create final df bin.knn
+        bin.knn= NULL
+        #loop through each variable in vars vector
+        for(i in 1:length(vars.vec)){
+                temp = data.frame(df[, vars[i]])
+                colnames(temp)<- c(vars[i])
+                formula = paste('~',vars[i],'-1')
+                a= model.matrix( as.formula(formula), temp)
+                a= a[match(rownames(temp), rownames(a)),]
+                bin.knn= cbind(bin.knn,a)
+        }
+        
+        return(bin.knn)
+}
+
+knn.binary<- binary.build(vars= vars.vec, df= all7d)
+knn.binary<- as.data.frame(knn.binary)
+
+#NA values don't work in knn function so turn all NA into zeroes
+knn.binary[is.na(knn.binary)]<- 0
+knn.binary$pc.pro<- all7d$pc.pro
+knn.binary$mlsto<- all7d$mlsto
+
+bin<- knn.binary %>% select(-mlsto)
+
+#knn function for categorical retrieving:
+
[…]

### 2015-07-24 final write up/draft.rmd
writeup

@@
 
-## 1.1 Research Motivations
+## 1.1 Research Motivations - 4.5 pages
 
-* thousands of consulting companies globally
-* range from single consultant to 10's of 1000's of employees
-* employees broad range of people from fresh graduates, to founders with decades of experience
-* internal environment: worker feelings of morale and security can be heavily influenced by success and profitability of their projects.
-Not to mention possibilities for pay rises and promotions which all come together to create a positive thriving culture.
-* harsher economic climate, particularly important to get competitive edge
-* find a source explaining how companies create their fixed fees - intuition, experience
-* find a source saying how badly it goes!
+description of problem:
+
+Project decision makers have struggled with forecasting project costs accurately for decades, across many industries.
+A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% (Flyvbjerg, 2006).
+A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were over 200% over budget on average (Flyvbjerg and Budzier, 2011).
+Complex longer term projects produce outcomes ranging from railways, houses, websites, and movies which involve professionals across a variety of industries.
+Depending on the industry, consulting companies are enlisted to complete portions of a project, and some of the financial risk of the overall project may be transferred to the consultant.
+This is achieved by requiring a fixed flat fee from the consultant and a committment to perform their role for the nominated price.
+In a similar way to the aforementioned infrastructure and IT projects, consultants are susceptible to underestimating the amount of their time that will be spent in their role, and must wear the cost of extra employee hours.
+For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and thriving.
+
+The ramifications of unprofitable projects in a private business are multifold.
+Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery.
+Employees also experience stress attempting to complete the project within a disappearing budget and may perform their work at a lower quality in exchange for speed.
+This is unfortunate if the reality is that the project fee was under estimated.
+Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would improve a business' standing.
+The financial risk taken on fixed flat fee projects (and disappointing results) may also discourage business' from taking different risks with more potential for return, such as expanding the business into a new area.
+Clearly, failure to predict fixed fee project costs has the potential to negatively affect a business' growth opportunities and staff morale.
+
+In some industries, consulting projects are often very different from one project to the next, covering complex subjectmatter with a range of subconstultants and stakeholders involved.
+This often means the consultant takes on financial risk by predicting the number of hours they will spend on a project and committing to this value until project completion.
+Depending on the situation, consultants may be able to negotiate fee variations, however this may not be beneficial in the long term.
+A fee variation has a negative impact on the consultant's client relationship.
+
+Why its not easily solvable:
+* often can't change fixed fees - client relationships and reputation. people talk.
+
+case study:
 * in the case study (and is typical of smaller organisations??), the CRM is an untapped source of information storing
@@
     * basic summaries of overall hours spent for the month
+
+REference class forecasting
+* find a source explaining how companies create their fixed fees - intuition, experience
+* predicting cost of construction from other perspectives - builders.
+
+Other industries
+* relates to other customer prediction problems - but this is for clients.
 
@@
 * Although the aim is to predict the profitability of a project, the model aims to change managers' fee setting decisions if the algorithm predicts low profit or a loss based on historical data. 
+* external factors forcing price down in some cases - competiton. clients
 

### 2015-07-26 cluster.rmd
app dev

@@
 
-library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
-library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
-library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
-library('FNN', lib='C:/Progra~1/R/R-3.2.1/library')
-library('caret', lib='C:/Progra~1/R/R-3.2.1/library')
+# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+# library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
+# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
+# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
+# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
+# library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
+# library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
+# library('FNN', lib='C:/Progra~1/R/R-3.2.1/library')
+# library('caret', lib='C:/Progra~1/R/R-3.2.1/library')
 
-library('RcmdrPlugin.BCA')
-# library(devtools)
-# library('ggbiplot')
-# library('fpc')
-# library('vegan')
-# library('knitr')
-# library('ggplot2')
-# library("plyr")
-# library("dplyr")
-# library('magrittr')
-# library('reshape2')
-# library("rpart")
-# library('car')
-# library('e1071')
-# library('corrgram')
-# library('party')
-# library('dendextend')
-# library('FNN')
-# library('caret')
+library(devtools)
+library('ggbiplot')
+library('fpc')
+library('vegan')
+library('knitr')
+library('ggplot2')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library("rpart")
+library('car')
+library('e1071')
+library('corrgram')
+library('party')
+library('dendextend')
+library('FNN')
+library('caret')
 
@@
 # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
+opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
 # opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
-all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
-# all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
+# all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
+all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
 all7c$Post.Code<- as.factor(all7c$Post.Code)

### 2015-07-26 final_app/server.R

@@ -1,12 +1,19 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
 # library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
 # library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
 # library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
 # library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
 # library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shiny')
+# library('shiny',lib='C:/Progra~1/R/R-3.1.3/library')
 # library('shinythemes',lib='C:/Progra~1/R/R-3.2.0/library')
+# library('ggplot2')
+library("dplyr")
+library("plyr")
+library('magrittr')
+library('reshape2')
+library('shiny')
+library('shinythemes')
 
 
 all7<- read.csv('all7.csv')[,-1]
@@ -118,9 +125,36 @@ shinyServer(
                             predict = 'pc.pro', case= input$case, k=10, result.df= all7d)
                 })
                 
-#                 c<- reactive({
-#                         inv.knn(df= b, predict= 'inv.mlsto', new.cases= 1, k=input$ks) %>% slice(-1)
+                c<- reactive({
+                        inv.knn(df= b(), predict= 'inv.mlsto', new.cases= 1, k=input$ks) %>% slice(-1)
+                })
+                
+#                 e<- reactive({
+#                         b()[,input$column, drop=FALSE]
 #                 })
+
+
+                e= reactive({
+                        if(input$column == 'NA'){
+                                test= c()
+                        }
+                        else {
+                                f=b()
+#                                 g= f[,input$column,drop=FALSE] %in% f[,input$column,drop=FALSE][1]
+                                test= f[f[,input$column] %in% f[,input$column][1],,drop=FALSE]
+                                test= inv.knn(df= test, predict= 'inv.mlsto', new.cases= 1, k= input$ks) %>% slice(-1)
+                        }    
+                })
+
+#                 else{
+#                         d<- reactive({b()[b()[,input$column] %in% b()[,input$column][1], ]
+#                         })
+#                         
[…]

### 2015-07-26 final_app/ui.R

@@ -170,7 +170,7 @@ shinyUI(fluidPage(
                                                            ),
                                                     
                                                     column(6,
-                                                           selectInput("Column", "Narrow search by:",
+                                                           selectInput("column", "Narrow search by:",
                                                                        list('none'='NA',
                                                                             "Client" = "code.client", 
                                                                             "Contact" = "code.contact", 

### 2015-07-30 cluster.rmd
not sure

@@
 
-all7d<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, Discipline, pc.pro, 
+all7d<- all7c %>% select(mlsto, Year, inv.mlsto, timespan, no.users, Discipline, pc.pro, 
                          client.totinv, code.director, 
@@
 
+head(all7d)
+
 ```

### 2015-07-30 var_select.rmd
knn
not sure

@@
 library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('RColorBrewer',lib='C:/Progra~1/R/R-3.2.1/library')
 
@@
 #delete charge.ph and charge.pc, Start.Date and End.Date
-all7b<- all7b %>% select(-charge.ph, -charge.pc, -Start.Date, -End.Date)
+all7b<- all7b %>% select(-charge.ph, -charge.pc, -End.Date)
 #impute NA no.users as 2 = median and mode
@@
 # write.csv(all7b,'all7b.csv')
+all7b<- read.csv('all7b.csv')[,-1]
 
@@
 #read csv file with all the client details - want to merge this with client df
-details<- read.csv('17marchclients.csv',
-                   na.strings = c("", " ", "NA"))
-head(details)
-#delete client.contact, client.count, and contact.count columns
-details<- details[,c(1,5:9)]
-#rename columns
-names(details)[2]<- 'Business'
-names(details)[4]<- 'Biz.size'
-names(details)[5]<- 'Biz.type'
-names(details)[6]<- 'Biz.age'
-
-#delete rows with 4 NA values
-details<-details[rowSums(is.na(details))!=5,]
-details<-details[rowSums(is.na(details))!=4,]
-
-#add row for client code
-code.client<- read.csv('code_client2.csv')[,-1]
-details<- merge(details, code.client, by.x ='client', by.y ='client2', all.x=TRUE, all.y=FALSE)
-details<- transform(details, client.age = 2014- Biz.age)
-
-#merge client.age column of details to all7b by code.client
-all7b<- merge(all7b, details %>% select(code, client.age), by.x = 'code.client', by.y = 'code', all.x=TRUE, all.y=FALSE)
-
-# write.csv(all7b, 'all7c.csv')
+# details<- read.csv('17marchclients.csv',
+#                    na.strings = c("", " ", "NA"))
+# head(details)
+# #delete client.contact, client.count, and contact.count columns
+# details<- details[,c(1,5:9)]
+# #rename columns
+# names(details)[2]<- 'Business'
+# names(details)[4]<- 'Biz.size'
+# names(details)[5]<- 'Biz.type'
+# names(details)[6]<- 'Biz.age'
+# 
+# #delete rows with 4 NA values
+# details<-details[rowSums(is.na(details))!=5,]
+# details<-details[rowSums(is.na(details))!=4,]
+# 
+# #add row for client code
+# code.client<- read.csv('code_client2.csv')[,-1]
+# details<- merge(details, code.client, by.x ='client', by.y ='client2', all.x=TRUE, all.y=FALSE)
+# details<- transform(details, client.age = 2014- Biz.age)
+# 
+# #merge client.age column of details to all7b by code.client
+# all7b<- merge(all7b, details %>% select(code, client.age), by.x = 'code.client', by.y = 'code', all.x=TRUE, all.y=FALSE)
+
+write.csv(all7b, 'all7c.csv')
 all7c<- read.csv('all7c.csv')[,-1]
@@
 # write.csv(all7c, 'all7c.csv')
-all7c<- read.csv('all7c.csv')[,-1]
-
-```
-
-First need to change code.client and code.contact. Need to assign NA to any code.client or code.contact with less than 3 counts
+# all7c<- read.csv('all7c.csv')[,-1]
+
+```
+
+First need to change code.client and code.contact. Need to assign NA to any code.client or code.contact with less than 2 counts
 
@@
 all7c$code.client<- as.character(all7c$code.client)
-all7c$code.client<- ifelse(all7c$V1>2, all7c$code.client, NA)
+all7c$code.client<- ifelse(all7c$V1>1, all7c$code.client, NA)
 all7c$code.client<- as.factor(all7c$code.client)
@@
 contact.check<- ddply(all7c,.(code.contact), nrow)
-contact.check<- contact.check %>% filter(V1>2)
+contact.check<- contact.check %>% filter(V1>1)
 all7c$code.contact<- as.character(all7c$code.contact)

### 2015-07-30 final_app/server.R

@@ -1,23 +1,24 @@
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shiny',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shinythemes',lib='C:/Progra~1/R/R-3.2.0/library')
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+library('shiny',lib='C:/Progra~1/R/R-3.1.3/library')
+library('shinythemes',lib='C:/Progra~1/R/R-3.2.0/library')
+library('FNN',lib='C:/Progra~1/R/R-3.2.1/library')
 # library('ggplot2')
-library("dplyr")
-library("plyr")
-library('magrittr')
-library('reshape2')
-library('shiny')
-library('shinythemes')
+# library("dplyr")
+# library("plyr")
+# library('magrittr')
+# library('reshape2')
+# library('shiny')
+# library('shinythemes')
 
 
-all7<- read.csv('all7.csv')[,-1]
-all7$Start.Date<- as.Date(all7$Start.Date)
+all7<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7.csv')[,-1]
+ all7$Start.Date<- as.Date(all7$Start.Date)
 #delete anything more than 4 for inv.vs.cost
 #delete anything less than .2 for inv.vs.cost
[…]

### 2015-07-30 final_app/ui.R

@@ -176,7 +176,8 @@ shinyUI(fluidPage(
                                                                             "Contact" = "code.contact", 
                                                                             "Billing Type" = "Billing.Type",
                                                                             "Discipline" = "Discipline",
-                                                                            'Business' = 'Business'
+                                                                            'Business' = 'Business',
+                                                                            'Job Type'='JD.Second'
                                                                        )),
                                                            submitButton(text="Update")
                                                            )
@@ -184,7 +185,15 @@ shinyUI(fluidPage(
                                                     )),
                                             
                                             fluidRow(
-                                                    tableOutput('knn.table')
+                                                    h4('Job Details'),
+                                                    tableOutput('knn.table'),
+                                                    h4('Finances'),
+                                                    tableOutput('knn.table1'),
+                                                    h4('Client Details'),
+                                                    tableOutput('knn.table2'),
+                                                    h4('Staff Details'),
+                                                    tableOutput('knn.table3')
+                                                    
                                                     )
                                    )
           

### 2015-08-02 final write up/draft.rmd
wkend work

--- 2015-07-07 final write up/draft.rmd
@@
 
+****
+
 # Introduction - 2 pages
@@
 
+****
+
 # Chapter 1 Problem Description
@@
 
-## 1.1 Research Motivations
+## 1.1 Research Motivations - 4.5 pages
 
-* thousands of consulting companies globally
-* range from single consultant to 10's of 1000's of employees
-* employees broad range of people from fresh graduates, to founders with decades of experience
-* internal environment: worker feelings of morale and security can be heavily influenced by success and profitability of their projects.
-Not to mention possibilities for pay rises and promotions which all come together to create a positive thriving culture.
-* harsher economic climate, particularly important to get competitive edge
-* find a source explaining how companies create their fixed fees - intuition, experience
-* find a source saying how badly it goes!
+description of problem:
+
+Project decision makers have struggled with forecasting project costs accurately for decades, across many industries.
+A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% (Flyvbjerg, 2006).
+A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average (Flyvbjerg and Budzier, 2011).
+Complex longer term projects are necessary to produce  railways, houses, websites, and movies which involve professionals across a many industries.
+Depending on the industry, consulting companies are enlisted to complete portions of a project, and some of the financial risk of the overall project may be transferred to the consultant.
+This is achieved by requiring a fixed flat fee from the consultant, meaning they have committed to perform their role for the nominated price.
+In a similar way to infrastructure and IT projects, consultants are susceptible to underestimating the amount of their time that will be spent in their role, and must wear the cost of extra employee hours spent.
+For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and succeeding.
+
+The ramifications of unprofitable projects in a private business are multifold.
+Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery.
+Employees also experience stress attempting to complete the project within a disappearing budget and may perform lower quality work in exchange for speed.
+This is unfortunate if the reality is that the project fee was under estimated.
+Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would improve a business' standing.
+The financial risk taken on fixed flat fee projects (and disappointing results) may also discourage business' from taking different risks with better potential pay-offs, such as expanding the business into a new area.
+Clearly, failure to predict fixed fee project costs can negatively affect a business' growth opportunities and staff morale.
+
+Why its not easily solvable:
+
+The challenge for consultants to reduce the risk of a project going over budget, in terms of hours spent, is not easily solved.
+If a job does go over budget, the business experiences reduced profits or  even losses as they are contractually bound to complete the work.
+The reason this task is challenging is that complex projects are always different from each other, and similar work may take significantly different amounts of time.
+For example, an ad filmed in a remote location for chocolate will cost significantly more than an ad filmed on an exisitng set, even though the product structure is almost identical.
+A fixed product price, which you could apply to repetive sales such as appliances, is inappropriate.
+One option is charging for consultants' time by the hour; a zero risk fee structure.
+However in many industries this is not competitive enough.
+Other consultancies may be willing to offer a fixed flat fee for the same work, reducing financial risk for the client, and luring the client away.
+Another fee strategy may be to track worker hours, and stop work once the consulting fee has been depleted.
+Then, negotiate further fees, or variations, with the client and persuade them to accept.
+This also satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective.
+However, if the project has been following a predictable process, consultants often prefer to wear losses or low profits to preserve the relationship with their client.
+Consultant's may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry over creating financial friction.
+Clearly, there is no straightforward solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects.
+In some indusries, fixed flat fee structures are expected and business' must conform to survive.
+
+Solutions:
+
+Previous research has been done on the prediction of final costs of a complex project with varying success.
+In the prediction of IT software development, formal estimation methods have existed for many years, however expert judgment is still the dominant method and available evidence does not suggest the models improve on expert judgment (Jorgensen and Shepperd, 2007).
+Construction industry research has used models ranging from neural networks to case based reasoning to regression to predict final building costs.
+However, often expert judgment was not mobilised, and the models were not comprehensive enough to be used in industry.
+Manager's experience and intuition is still the dominant method for construction cost calculation.
+
+- research has gone into targeting common repeated points in a projects timeline where things can go awry. reference
+- example: find research, IT projects? nothing in engineering or film?
+- Although no solutions are available, the use of reference class forecasting is being pioneered in england for large government infrastructure
+- method for fee pricing has remained the same for decades - detailed analysis of project, add up time based on tasks, allow buffer time for error.
+
+Reference class forecasting
+* predicting cost of construction from other perspectives - builders.
+
+Other industries
+* predicition of building construction costs - statistical analysis
+
+Potential solution:
+* stat analysis like reference class forecasting and building construction except analysis performed internally using the company's historical data on past projects.
+
+case study:
+
+This project undertakes a case study on a consulting company in the construction industry. Specifically, the
+
 * in the case study (and is typical of smaller organisations??), the CRM is an untapped source of information storing

### 2015-08-04 final write up/draft.rmd
draft update

@@
 A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average (Flyvbjerg and Budzier, 2011).
-Complex longer term projects are necessary to produce  railways, houses, websites, and movies which involve professionals across a many industries.
-Depending on the industry, consulting companies are enlisted to complete portions of a project, and some of the financial risk of the overall project may be transferred to the consultant.
-This is achieved by requiring a fixed flat fee from the consultant, meaning they have committed to perform their role for the nominated price.
+These statistics cast a concerning glimpse at the financial story playing out behind the scenes of these projects which can cause job losses, ruin private businesses, and burn through government balancesheets.
+Complex long term projects are necessary to produce  railways, houses, websites, and movies.
+Depending on the industry, the the financial risk of completing a project on budget is transferred to consulting companies who are enlisted to complete portions of a project.
+The risk is transferred by requiring a fixed flat fee from the consultant, meaning they commit to complete their role for a nominated price.
 In a similar way to infrastructure and IT projects, consultants are susceptible to underestimating the amount of their time that will be spent in their role, and must wear the cost of extra employee hours spent.
@@
 Employees also experience stress attempting to complete the project within a disappearing budget and may perform lower quality work in exchange for speed.
-This is unfortunate if the reality is that the project fee was under estimated.
+This is unfortunate if in reality, the project fee was under estimated.
 Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would improve a business' standing.
-The financial risk taken on fixed flat fee projects (and disappointing results) may also discourage business' from taking different risks with better potential pay-offs, such as expanding the business into a new area.
-Clearly, failure to predict fixed fee project costs can negatively affect a business' growth opportunities and staff morale.
+The financial risk taken on fixed flat fee projects may also discourage business' from taking different risks with better potential pay-offs, such as expanding the business into a new area.
+Clearly, predicting fixed fee project costs more accurately could positively affect a business' growth opportunities and staff morale.
 
@@
 The challenge for consultants to reduce the risk of a project going over budget, in terms of hours spent, is not easily solved.
-If a job does go over budget, the business experiences reduced profits or  even losses as they are contractually bound to complete the work.
-The reason this task is challenging is that complex projects are always different from each other, and similar work may take significantly different amounts of time.
+The reason is that complex projects are always different from each other, and similar work may take significantly different amounts of time.
 For example, an ad filmed in a remote location for chocolate will cost significantly more than an ad filmed on an exisitng set, even though the product structure is almost identical.
-A fixed product price, which you could apply to repetive sales such as appliances, is inappropriate.
-One option is charging for consultants' time by the hour; a zero risk fee structure.
+Therefore, a fixed product price, which you could apply to products such as appliances, is inappropriate for film projects.
+One option is charging for consultants' time and the hiring of materials by the hour; a zero risk fee structure.
 However in many industries this is not competitive enough.
@@
 This also satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective.
-However, if the project has been following a predictable process, consultants often prefer to wear losses or low profits to preserve the relationship with their client.
+However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or low profits to preserve the relationship with their client.
 Consultant's may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry over creating financial friction.
@@
 Solutions:
+
+Little research has been done into surveying the pricing methods of consulting projects, however
+
+- current practise: literature available on IT and construction projects
+        - detailed analysis of project
+        - comparison to previous projects by experts - case based reasoning
+        - intuition
+        - some predictive statistics based on past data 
+        - reviewing - trying to find places where things keep going wrong
+
+- research: dozens of papers using statistical and machine learning to predict project costs
+        - construc
+        - IT
+        - film costs?
+
+- has any of this research been mobilised into practise, if not, why?
+        - reference class forecasting - pioneered in british government large gov infrastructure
 
@@
 
-- research has gone into targeting common repeated points in a projects timeline where things can go awry. reference
-- example: find research, IT projects? nothing in engineering or film?
-- Although no solutions are available, the use of reference class forecasting is being pioneered in england for large government infrastructure
 - method for fee pricing has remained the same for decades - detailed analysis of project, add up time based on tasks, allow buffer time for error.
-
-Reference class forecasting
-* predicting cost of construction from other perspectives - builders.
-
-Other industries
-* predicition of building construction costs - statistical analysis
 

### 2015-08-04 combine.R

@@ -70,5 +70,5 @@ all5<- merge(all5, temp[,names(temp) %in% c('mlsto','Job.Number','profit.mlsto',
 all5 <- transform(all5, return.pdol= balance.mlsto/(cost.mlsto+dis.sc.mlsto))
 all5 <- transform(all5, inv.vs.cost= inv.mlsto/(cost.mlsto+dis.sc.mlsto))
 
-all5<- write.csv(all5, 'all5a.csv')
+ all5<- write.csv(all5, 'all5a.csv')
 # go to de_identify.R

### 2015-08-05 final write up/draft.rmd
motivation for research writing

@@
 output: word_document
+bibliography: Library.bib
 ---
@@
 
-Solutions:
+current practise:
 
-Little research has been done into surveying the pricing methods of consulting projects, however
+Limited research is available on the cost estimation techniques performed by industry for complex projects, however two surveys were performed in 2000: one on construction projects, the other on IT projects.
+The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practises.
+They found that the main method for cost estimation was breaking the project into detailed parts and adding up the cost of each item.
+The next two most popular methods were comparison with similar projects based on documented facts, and comparison with similar projects based on personal experience.
+These can all be classified as experience based models [@Akintoye2000].
+A survey by Moores and Edwards in 1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project.
+In the same survey, 91% of software companies cited cost estimation as a problem.
+Although limited recent literature is available on industry cost estimation practises, historically it can be observed that for complex construction and software projects, cost predictions are chiefly made with a mixture of experience, project comparisons, and an analysis of the project details and tasks.
+It is also acknowledged that cost estimation is a problem and hypothesised that inexperience and lack of time are chief contributors [@Akintoye2000].
 
-- current practise: literature available on IT and construction projects
-        - detailed analysis of project
-        - comparison to previous projects by experts - case based reasoning
-        - intuition
-        - some predictive statistics based on past data 
-        - reviewing - trying to find places where things keep going wrong
+theory about problem with current practise:
+
+@Lovallo2003 tackled the psychology behind the high failure rates of executives in predicting costs of projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
+Their theory stems from Kahneman's work on decision making that won him the nobel prize for economics in 2002.
+The research argues that humans' natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
+A manager sees challenges in a project as something that can be overcome by the team's high skill level, and downplay or ignore the risk of problems that are out of the team's control.
+It is for this reason that it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
+Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].
+@Lovallo2003 call the practise of analysing a project based on project details and its unique complexities the 'inside view'.
+Research has shown that if people are forced to make predictions about their skills after being exposed to an 'outside view', their predictions are significantly more accurate.
+This outside view can be applied to complex projects, and involves ignoring the details of the current project and instead analysing the outcomes of several projects of a similar description.
+This is called reference class forecasting [@Lovallo2003].
+One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
+The correlation for the current project can then be estimated via a statistical model which is used to adjust the forecasted cost made by detailed analysis (inside view) [@Flyvbjerg2011].
+The traditional way to think about a project is to focus on the project details.
+Even though gathering data from similar projects could significantly improve a cost prediction, it is rarely thought of by managers [@Flyvbjerg2011].
+
+The implementation of reference class forecasting began in project management for the first time in 2004, was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, Australia.
+An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].
+
+![This method was implemented by the British government for a rail project which is still under completion](C:/Users/n9232371/Documents/github/consult/final write up/UKrail.JPG)
+
+The visualisation presents a powerful communication tool for influencing decision makers and improving forecast cost accuracy, however access to credible data to a sufficient number of projects can be a challenge [@Flyvbjerg2011].
+Overall, the theory of optimism bias is a credible explanation for the problem of persistant underestimation of project costs.
+The accompanying solution of reference class forecasting presents a promising solution given enough accurate data.
 
@@
 
-- has any of this research been mobilised into practise, if not, why?
         - reference class forecasting - pioneered in british government large gov infrastructure
@@
 Manager's experience and intuition is still the dominant method for construction cost calculation.
-
-- method for fee pricing has remained the same for decades - detailed analysis of project, add up time based on tasks, allow buffer time for error.
 

### 2015-08-15 finalwriteup/draft.rmd
big dump started regression

--- 2015-08-05 final write up/draft.rmd
@@
 
-![This method was implemented by the British government for a rail project which is still under completion](C:/Users/n9232371/Documents/github/consult/final write up/UKrail.JPG)
+![This method was implemented by the British government for a rail project which is still under completion](UKrail.JPG)
 
@@
 
-- research: dozens of papers using statistical and machine learning to predict project costs
-        - construc
-        - IT
-        - film costs?
+previous research:
 
-        - reference class forecasting - pioneered in british government large gov infrastructure
+Given this theory of reference class forecasting was first clearly published in 2003, the idea of predicting project costs from previous data has been the subect of research for many years before that and continues to this day.
+Although research results have often been promising, industry uptake of the idea is not yet established.
+
+First an overview of the methods used will be presented followed by published theories for the lack of industry uptake.
+
+***need to really do more research on this!!***
+
+A substantial amount of research on project costprediction has been dedicated to the fields of software development and construction projects in particular.
+These analyses have generally used project information from 15 to 20 large projects.
+A single dataset may consist of projects from around the world, and each project a different client and project team.
+The mathematical methods used to predict project costs vary from statistical regressions to deep-learning neural networks.
+Most studies report neural networks are more accurate than regression
 
@@
 Manager's experience and intuition is still the dominant method for construction cost calculation.
+
+*** ^^^more research^^^ ***
 
@@
 
-* K means
 * anova
-* in construction journals, commonly applied linear regression
+
+* linear regression
+    * pros
+
+    * cons
         * do these acknowledge their data is not normally distributed?

### 2015-08-15 combine.R

@@ -47,6 +47,7 @@ temp1<- temp[1:2425,]
 #string split to create another column with the first part of a job number. For the first half of jobs just the first bit indicates it could be invoiced together ie 2002.001 by itself
 temp1$mlsto<-sapply(temp1$Job.Number,FUN=function(x){
         substr(x,1,8)})
+temp1$mlsto<- ifelse(temp1$mlsto %in% c('2004.056','2010.236'), temp1$Job.Number, temp1$mlsto)
 #Then ddply using this shortened job number to group. Add up all balance, profit hours in shortened job number. MErge back with original data set.
 temp1<- merge(temp1, ddply(temp1, .(mlsto), summarise, profit.mlsto= sum(profit), balance.mlsto = sum(balance), hrs.mlsto = sum(hours), cost.mlsto=sum(cost),
                            dis.sc.mlsto = sum(Dis.subcon),  inv.mlsto=sum(Tot.Invoiced)),
@@ -56,7 +57,7 @@ temp1<- merge(temp1, ddply(temp1, .(mlsto), summarise, profit.mlsto= sum(profit)
 temp2<- temp[2426:4168,]
 temp2$mlsto<-sapply(temp2$Job.Number,FUN=function(x){
         substr(x,1,10)})
-temp2$mlsto<- ifelse(temp2$mlsto == '2011.072.3', temp2$Job.Number, temp2$mlsto)
+temp2$mlsto<- ifelse(temp2$mlsto %in% c('2011.072.3','2011.063.3','2012.453.3'), temp2$Job.Number, temp2$mlsto)
 temp2<- merge(temp2, ddply(temp2, .(mlsto), summarise, profit.mlsto= sum(profit), balance.mlsto = sum(balance), hrs.mlsto = sum(hours), cost.mlsto=sum(cost),
                            dis.sc.mlsto= sum(Dis.subcon), inv.mlsto=sum(Tot.Invoiced)),
               by.x= 'mlsto', by.y= 'mlsto', all.x=TRUE, all.y=FALSE)

### 2015-08-15 de_identify.R

@@ -281,7 +281,7 @@ all6a$Role.JD<- grepster(nstruc.list, nstruc.title, c('Civil','Environmental Pla
 
 #worked!
 #now compare when JD.Second differs from Role.JD
-all6a %>% filter(!is.na(Role.JD), !is.na(JD.Second),!(Role.JD==JD.Second)) %>% select(Role, Role.JD, JD.Second, Discipline,mlsto) %>% View()
+# all6a %>% filter(!is.na(Role.JD), !is.na(JD.Second),!(Role.JD==JD.Second)) %>% select(Role, Role.JD, JD.Second, Discipline,mlsto) %>% View()
 
 #2014.371.3 - keep JD.Second
 #2014.117.3 - keep JD.Second

### 2015-08-15 invoice_eng.R

@@ -35,11 +35,12 @@ temp1<- temp[1:7680,]
 #string split to create another column with the first part of a job number. For the first half of jobs just the first bit indicates it could be invoiced together ie 2002.001 by itself
 temp1$mlsto<-sapply(temp1$Job.Number,FUN=function(x){
         paste(strsplit(x,split='[.]')[[1]][1], strsplit(x,split='[.]')[[1]][2], sep=".")})
+temp1$mlsto<- ifelse(temp1$mlsto %in% c('2004.056','2010.236'), temp1$Job.Number, temp1$mlsto)
 #do the same with second half of data
 temp2<- temp[7681:11627,]
 temp2$mlsto<-sapply(temp2$Job.Number,FUN=function(x){
         substr(x,1,10)})
-temp2$mlsto<- ifelse(temp2$mlsto == '2011.072.3', temp2$Job.Number, temp2$mlsto)
+temp2$mlsto<- ifelse(temp2$mlsto %in% c('2011.072.3','2011.063.3','2012.453.3'), temp2$Job.Number, temp2$mlsto)
 #rbind temp1 and temp2 back together to make temp
 temp<- rbind(temp1, temp2)
 

### 2015-08-19 Regression.rmd
regression markdown

---
title: "Regression"
author: "Amy Cook"
date: "August 15, 2015"
output: word_document
---

```{r, echo=FALSE, include=FALSE}

# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
# library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')

library(devtools)
library('ggbiplot')
library('fpc')
library('vegan')
library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('corrgram')

# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
# all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
all7c$Post.Code<- as.factor(all7c$Post.Code)

```

This file is to do a multi-step regression to predict RETURN PER DOLLAR.
The reason for multistepping is to allow for multiple variables with lots of missing NA values.

Variables to include:

Full:

inv.mlsto
timespan
no.users
Discipline
pc.pro
client.totinv
code.director
Num.disc
Business
Biz.type

Many NA:

code.contact
code.client
Project.Value
JD.Second
Post.Code
code.ProjEng
Billing.Type
majority.pos
pc.majpos

The 'Many NA' variables will be split into groups:

* code.contact: 1652 NA
* code.client: 422 NA
* Project.Value: 2333 NA
* JD.Second: 1057 NA
* Post.Code: 356 NA
* code.ProjEng: 47 NA
* Billing.Type: 1314 NA
* majority.pos: 377 NA
* pc.majpos: 377 NA

```{r echo=FALSE}
some <- all7c %>% select(return.pdol, mlsto, inv.mlsto, timespan,no.users,Discipline,pc.pro,
                         client.totinv,code.director,Num.disc,Business,Biz.type,
                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,majority.pos,
                         pc.majpos)

#heatmap of any similar NA value ranges

experiment<- sapply(some, function(x) is.na(x))
NA.plot<- melt(experiment)
NA.plot$mlsto<- rep(all7c$mlsto, ncol(experiment))
#reorder levels for ggplot by total number of NA TRUE
# table(NA.plot$Var2, NA.plot$value)[,'TRUE']
NA.plot <- within(NA.plot, 
                         Var2 <- factor(Var2, 
                                               levels=names(sort(table(NA.plot$Var2, NA.plot$value)[,'TRUE'], 
                                                                 decreasing=FALSE))))

ggplot(NA.plot, aes(x=Var2, y=mlsto)) +
        geom_tile(aes(fill = value)) + 
        theme(axis.text.x=element_text(angle=45,hjust=1))

```

### 2015-08-19 final_app/server.R

@@ -156,7 +156,8 @@ shinyServer(
                                               fill='blue', linetype='blank', width=0.25)+
                                 geom_errorbar(aes(ymax=150, ymin=150, x=id, y=min), width=0.25, colour='navyblue', size=1.5) +
                                 geom_errorbar(aes(ymax=100, ymin=100, x=id, y=min), width=0.25, colour='red') +
-                                scale_y_continuous(breaks = seq(80, 170, by = 10), limits=c(80,170)) +
+                                scale_y_continuous(breaks = seq(80, 170, by = 10)) +
+                                ylim(80,170) +
                                 labs(y="Final Spend divided by Quote (%)", x="New Project", title= "Prediction of Final Spend vs Initial Quote") +
                                 theme(axis.text.y = element_blank(),
                                       plot.title = element_text(size=18, face='plain'),

### 2015-08-21 Regression.rmd
getting somewhere with regression

@@
 
-# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('caret',lib='C:/Progra~1/R/R-3.2.1/library')
 # library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
 # library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
@@
 # opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
-# all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
-all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
-all7c$Post.Code<- as.factor(all7c$Post.Code)
+all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
+all7d<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
+all7d$Post.Code<- as.factor(all7d$Post.Code)
 
@@
 Biz.type
+code.ProjEng
 
@@
 Post.Code
-code.ProjEng
 Billing.Type
@@
 * Post.Code: 356 NA
-* code.ProjEng: 47 NA
 * Billing.Type: 1314 NA
@@
 ```{r echo=FALSE}
-some <- all7c %>% select(return.pdol, mlsto, inv.mlsto, timespan,no.users,Discipline,pc.pro,
+some <- all7d %>% select(return.pdol, mlsto, inv.mlsto, timespan,no.users,Discipline,pc.pro,
                          client.totinv,code.director,Num.disc,Business,Biz.type,
@@
 NA.plot<- melt(experiment)
-NA.plot$mlsto<- rep(all7c$mlsto, ncol(experiment))
+NA.plot$mlsto<- rep(all7d$mlsto, ncol(experiment))
 #reorder levels for ggplot by total number of NA TRUE
@@
 ```
+
+Will address each of the variables with many nA separately. 8 of these variables totally. run a separate lm for each of these
+
+Need to transform some numeric variables:
+
+Full:
+
+return.pdol - cube root
+inv.mlsto - log
+timespan - cube root
+pc.pro - leave alone
+client.totinv - log
+
+Many NA:
+
+Project.Value - log
+pc.majpos - log
+
+```{r, echo=FALSE}
+
+#write a function for testing transformations
+
+trans<- function(var= 'timespan', df= all7d, transform = 'cbrt'){
+        #initial q plot
+        init<- qqPlot(df[,var], 
+                       main = 'untouched variable')
+        print(init)
+
+        #make trial dataframe
+        trial = sapply(df[,var], function(x){
+                if(transform == 'log'){
+                        log(x)
+                } else
+
+                if(transform == 'neglog')
+                        {
+                        sign(x)*log(abs(x) + 1)
+                } else
+
+                if(transform == 'sqrt')
+                {
+                        sqrt(x)
+                } else
+                        if(transform == 'cbrt')
+                {
+                        sign(x) * abs(x)^(1/3)
+                } else
+                { x^2 }
+
+                }) %>% 
+                as.data.frame
+
+        colnames(trial)[names(trial) %in% '.']<-'trans.var'
+
+        #qplot for transformation
+        qqtrans<- qqPlot(trial$trans.var, 
+                         main = 'transformed')
+        print(qqtrans)
+
+        ##move on to density plots
+
+        #original variable
+        dens<- ggplot(df, aes_string(x=var)) + geom_histogram(aes(y=..density..))
+        print(dens)
+
+        #transformed variable
+        dens.trans<- ggplot(trial, aes_string(x='trans.var')) + geom_histogram(aes(y=..density..))
+        print(dens.trans)
+}
+
+trans(var= 'timespan', df= all7d, transform = 'cbrt')
+trans(var= 'inv.mlsto', df= all7d, transform = 'log')
+trans(var= 'return.pdol', df= all7d, transform = 'cbrt')
+trans(var= 'pc.majpos', df= all7d, transform = 'log')
+trans(var= 'client.totinv', df= all7d, transform = 'log')
+trans(var= 'Project.Value', df= all7d, transform = 'log')
+
+```
+
+Have now decided on transformations for numeric variables.
+
+Create all7d ready for linear regression analysis
+
+```{r, echo=FALSE}
+
+math.cbrt <- function(x) {
+        sign(x) * abs(x)^(1/3)
+}
+
+all7d$timespan.cbrt<- math.cbrt(all7d$timespan)
+all7d$inv.mlsto.log <- log(all7d$inv.mlsto)
+all7d$client.totinv.log <- log(all7d$client.totinv)
+all7d$Project.Value.log <- log(all7d$Project.Value)
+all7d$pc.majpos.log <- log(all7d$pc.majpos)
+
+```
+
+Root mean square error as initial check. Write code which transforms response variable back to original format by cubing, then finds RMSE.
+
+```{r, echo=FALSE}
+
+#predict Petal.Width using lm, find RMSE
+#answer will be return.pdol
+#guess will be response of prediction which is cube root
+RMSE<- function(answer, guess){
+        error= answer-guess
+        root.error= sqrt(mean(error^2))
+        print(root.error)
+}
+
+```
+
+Now fit full variables to linear regression
+
+```{r, echo=FALSE}
+
+full = lm(return.pdol ~ 
+                  inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + client.totinv.log + code.director + 
+                  Num.disc + Business + Biz.type + code.ProjEng,
+          data = all7d)
+summary(full)
+
+#have a look at the residuals
+
+res<- resid(full)
+plot(all7d$inv.mlsto.log, res)
+abline(0,0)
+
+plot(density(res))
+qqPlot(res)
+
+#compare some of the answers
+ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full, all7d, interval= 'predict'))
+
+RMSE(ans$answer, ans$full.lm.fit)
+
+```
+
+RMSE is 0.4768 which isn't great in terms of predicting return per dollar.
+R^2 is 0.1905
+
+Let's try deleting some full variables
+
+```{r, echo=FALSE}
+full.2<- update(full, .~. -client.totinv.log)
+summary(full.2)
+full.3<- update(full.2, .~. -code.director)
+summary(full.3)
+full.4<- update(full.3, .~. -Num.disc)
+summary(full.4)
+full.5<- update(full.4, .~. -code.ProjEng)
+summary(full.5)
+full.6<- update(full.5, .~. -Biz.type + Discipline)
+summary(full.6)
+
+#RMSE check
+
+ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full.6, all7d, interval= 'predict'))
+
+RMSE(ans$answer, ans$full.lm.fit)
+
+```
+
+Feels better. That leaves us with
+
+* inv.mlsto.log
+* Discipline
+* timespan.cbrt
+* no.users
+* pc.pro
+* Discipline
+* Business
+
+RMSE = 0.48266
+
+Going back to full variables, lets try pairwise multiplying..
+
+```{r, echo=FALSE}
+#try diff pairs
+
+pairs.lm<- lm(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
+                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
+              data=all7d)
+
+summary(pairs.lm)
+pairs<- predict(pairs.lm, all7d)
+ans<- data.frame(answer= all7d$return.pdol, pairs.lm = pairs)
+
+RMSE(ans$answer, ans$pairs.lm)
+
+#down to 0.4759
+
+```
+
+Let's try adding more variables:
+
+```{r, echo=FALSE}
+#try adding code.client
+
+addone<- function(var= 'JD.Second', df= all7d){
+        reduced = df[!(is.na(df[,var])),]
+        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business", var, sep=" + ")
+        add.one = lm(as.formula(formula), data = reduced)
+
+        print(summary(add.one))
+
+        ans= data.frame(answer= reduced$return.pdol, addone.lm = predict(add.one, reduced, interval= 'predict'))
+
+        print(head(ans))
+
+        print(
+        paste("RMSE=", RMSE(ans$answer, ans$addone.lm.fit) %>% round(4), sep= " ")
+        )
+
+}
+
+#without additional variables, RMSE sits at 0.4759
+
+addone(var='JD.Second', df=all7d)
+#RMSE = .4709
+addone(var='Post.Code', df=all7d)
+#RMSE = .4353
+addone(var='majority.pos', df=all7d)
+#RMSE = .4541
+addone(var='pc.majpos.log', df=all7d)
+#RMSE = .4553
+addone(var='code.client', df=all7d)
+#RMSE = .4394
+addone(var='Billing.Type', df=all7d)
+#RMSE = .4579
+addone(var='code.contact', df=all7d)
+#RMSE = .3972
+
+#dont use post code ever. 
+#JD.Second doesn't help that much. acutally makes RMSE worse.
+
+```
+
+Delete Post.Code option
+
+Billing.Type should be paired with inv.mlsto and maybe Discipline
+
+maybe client should be paired with inv.mlsto
+maybe JD.Second should be paired with inv.mlsto and code.client and timespan and pc.pro and no.users?? which ever one does something.
+
+majority.pos pair with inv.mlsto
+pc.majpos pair with inv.mlsto
+
+Write a function that lists the extra variable you're trying and write out the interaction.
+
+```{r, echo=FALSE}
+
+addmult<- function(var= c('JD.Second', 'code.client'), combos= c('JD.Second*inv.mlsto.log'), df= all7d){
+        reduced = df[complete.cases(df[,var]),]
+        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business")
+
+        for(i in 1:length(var)){
+                formula = paste(formula, var[i], sep= " + ")
+        }
+
+        for(i in 1:length(combos)){
+                formula = paste(formula, combos[i], sep= " + ")
+        }
+
+        add.one = lm(as.formula(formula), data = reduced)
+
+        print(summary(add.one))
+
+        ans= data.frame(answer= reduced$return.pdol, addone.lm = predict(add.one, reduced, interval= 'predict'))
+
+        print(head(ans))
+
+        print(
+        paste("RMSE=", RMSE(ans$answer, ans$addone.lm.fit) %>% round(4), sep= " ")
+        )
+
+}
+
+#try Billing.Type with inv.mlsto.log
+addmult(var= c('Billing.Type'), combos= c('Billing.Type*inv.mlsto.log'), df= all7d)
+# no
+
+#try code.client*inv.mlsto.log
+addmult(var= c('code.client'), combos= c('code.client*inv.mlsto.log'), df= all7d)
+#made a difference, brought RMSE down to .4076 :)!
+
+#try JD.Second with inv.mlsto.log (no), code.client (too intense), timespan (no), pc.pro (no), no.users (yes)?
+addmult(var= c('JD.Second'), combos= c('JD.Second*no.users'), df= all7d)
+#add JD.Second*no.users
+
+addmult(var= c('majority.pos'), combos= c('majority.pos*inv.mlsto.log'), df= all7d)
+#no
+
+addmult(var= c('pc.majpos.log'), combos= c('pc.majpos.log*inv.mlsto.log'), df= all7d)
+#no
+
+```
+
+It is worthwhile to include:
+
+* code.client by inv.mlsto.log because brought down RMSE from 0.4394 (code.client) to 0.4076 (code.client + code.client by inv.mlsto.log)
+* JD.Second by no.users. Brought down RMSE from 0.4709 to 0.4612...
+
+Now look at cross validation for full variable set:
+
+```{r, echo=FALSE}
+
+#using caret package
+tc <- trainControl("cv",10,savePred=F)
+fit<- train(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
+                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
+              data=all7d,
+            method = 'lm', trControl = tc)
+
+#RMSE = 0.49246
+#compared to non-cross validated
+# RMSE = 0.4759
+```
+
+Not really trying to tune anything via cross validation, however RMSE increased from 0.4759 to 0.49246.
+
+Not THAT bad.
+
+Now, how to get variable importance from lm..
+
+```{r, echo=FALSE}
+
+#variable importance for full variables only - pairs.lm
+#use varImp from caret package - uses t-test statistic
+
+imp<- varImp(pairs.lm, scale=FALSE)
+imp$variable<- rownames(imp)
+imp<- imp %>% select(variable, Overall)
+imp %>% arrange(-Overall) %>% head()
+
+sum(imp$Overall)
+
+```
+
+Trial some scenarios:
+
+Run full variables plus
+
+Make a function to output:
+
+* var importances
+* RMSE
+
+based on a list of variables that you want to use
+
+```{r, echo=FALSE}
+
+sample<- all7d[1000,]
+sample<- sample[,!is.na(sample)]
+vars<- names(sample) 
+vars<- vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc','Biz.type','Post.Code',
+               'code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto','balance.mlsto','client.totinv.log','pc.majpos')]
+vars
+
+RMSE<- function(answer, guess){
+        error= answer-guess
+        root.error= sqrt(mean(error^2))
+        # print(root.error)
+}
+
+final.output<- function(df=all7d, samp = 800) {
+        #create sample df
+        sample = df[samp,]
+        sample = sample[,!is.na(sample)]
+        vars= names(sample)
+        vars = vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc',
+                                  'Biz.type','Post.Code','code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto',
+                                  'balance.mlsto','client.totinv.log','pc.majpos')]
+        variables = vars
+
+        print(
+                list("sample case", t(sample %>% select(-Year, -timespan.cbrt, -inv.mlsto.log, -client.totinv.log, -pc.majpos.log)))
+        )
+
+        #make reduced dataset with complete cases for all variables we want to use
+        reduced = df[complete.cases(df[,variables]),]
+
+        print(
+                paste(nrow(reduced), "cases in full model")
+        )
+
+        formula = paste("return.pdol ~ ")
+
+        for(i in 1:length(variables)){
+                formula = paste(formula, variables[i], sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'inv.mlsto.log' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'code.client' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*code.client', sep= " + ")
+        }
+
+        if('JD.Second' %in% variables &
+           'no.users' %in% variables){
+                formula = paste(formula, 'JD.Second*no.users', sep= " + ")
+        }
+
+        final = aov(as.formula(formula), data = reduced)
+
+        fit= summary(final)
+
+        ans= data.frame(answer= reduced$return.pdol, final.lm = predict(final, reduced, interval= 'predict'))
+
+        # print(head(ans))
+
+        #print variable importances for reduced lm just run
+
+        imp= data.frame('vars'= rownames(fit[[1]]), 'F.val' = fit[[1]]$`F value`)
+        #remove white space from levels in imp$vars
+        levels(imp$vars)<- sub("\\s+$", "", levels(imp$vars))
+        #remove Residuals row
+        imp= imp[!(imp$vars %in% 'Residuals'),]
+
+        #print variable importances
+        print(
+                list("variable importances",
+                     imp %>% arrange(-F.val) %>% head())
+        )
+
+        #print RMSE for cases that match certain categories only
+        #first subset predictions to cases that match the category
+
+        reduced$predict= ans$final.lm.fit
+
+        cats= c('code.client', 'code.contact', 'JD.Second', 'Business', 'Discipline', 'majority.pos')
+
+        for(i in 1:length(cats)){
+                if(cats[i] %in% vars){
+                        cats.red = reduced[reduced[,cats[i]] %in% sample[,cats[i]],]
+                        cat.RMSE = RMSE(cats.red$return.pdol, cats.red$predict) %>% round(4)
+                        print(
+                                paste(cats[i], "RMSE =", cat.RMSE, sep= " "))
+                        print(
+                                paste("Number of cases =", nrow(cats.red), sep= " "))
+                        print(
+                                head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto','return.pdol','predict',cats[i])])
+                        )
+                        }
+        }
+
+        final.RMSE= RMSE(ans$answer, ans$final.lm.fit) %>% round(4)
+
+        print(
+        paste("Overall predicted RMSE=", final.RMSE , sep= " ")
+        )
+
+        predict.rpdol = predict(final, sample) %>% round(3)
+        predict.rpdol = predict.rpdol[[1]]
+        predict.cost = sample$inv.mlsto/(1+predict.rpdol) %>% round(1)
+        print(
+               paste("return per dollar prediction", predict.rpdol, sep = " ")
+        )
+
+        print(
+               paste("answer", sample$return.pdol %>% round(3), sep = " ")
+        )
+
+        print(
+                paste("Invoiced= $", sample$inv.mlsto, ",   predicted cost= $", predict.cost %>% round(1), sep="")
+        )
+
+        print(
+                paste("Actual cost= $", sample$cost.mlsto, sep="")
+        )
+
+}
+
+final.output(df=all7d, samp = 500)
+
+```

### 2015-08-21 cluster.rmd
getting somewhere with regression

@@
 
-# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
 # library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
 # library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
 # library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
@@
 
-Create reduced dataset and have a think about what you want to cluster against each other
+Create reduced dataset and have a think about what you want to cluster against each otherim
 

### 2015-08-21 var_select.rmd
big dump started regression
getting somewhere with regression

@@
 num.inv and mean.inv- 1 NA?
+code.ProjEng - impute: all from pre 2005. All projEng can = director for that job
+code.director - impute only 'unknown' to be S121 - by far the majority and fits job description
 
@@
 
+#impute code.director Unknown to be S168
+all7b[all7b$code.director == 'Unknown',]$code.director<-'S168'
+all7b$code.director<- droplevels(all7b$code.director)
+#impute code.ProjEng to be Director codes for that job
+all7b[is.na(all7b$code.ProjEng),]$code.ProjEng<- all7b[is.na(all7b$code.ProjEng),]$code.director
+all7b$code.ProjEng<- droplevels(all7b$code.ProjEng)
+
 #redo client.count and contact.count according to this reduced dataset!!
@@
 
-# write.csv(all7b,'all7b.csv')
+write.csv(all7b,'all7b.csv')
 all7b<- read.csv('all7b.csv')[,-1]

### 2015-08-25 rand_forest.rmd
random forest start

--- 2015-08-21 Regression.rmd
@@
 library(devtools)
-library('ggbiplot')
-library('fpc')
-library('vegan')
 library('knitr')
@@
 library('corrgram')
+library('randomForest')
+library('caret')
 
@@
 
-```{r echo=FALSE}
-some <- all7d %>% select(return.pdol, mlsto, inv.mlsto, timespan,no.users,Discipline,pc.pro,
-                         client.totinv,code.director,Num.disc,Business,Biz.type,
-                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,majority.pos,
-                         pc.majpos)
+Will address each of the variables with many NA separately. 8 of these variables totally. run a separate lm for each of these
 
-#heatmap of any similar NA value ranges
+No need to transform numeric variables - not required for trees
 
-experiment<- sapply(some, function(x) is.na(x))
-NA.plot<- melt(experiment)
-NA.plot$mlsto<- rep(all7d$mlsto, ncol(experiment))
-#reorder levels for ggplot by total number of NA TRUE
-# table(NA.plot$Var2, NA.plot$value)[,'TRUE']
-NA.plot <- within(NA.plot, 
-                         Var2 <- factor(Var2, 
-                                               levels=names(sort(table(NA.plot$Var2, NA.plot$value)[,'TRUE'], 
-                                                                 decreasing=FALSE))))
-
-ggplot(NA.plot, aes(x=Var2, y=mlsto)) +
-        geom_tile(aes(fill = value)) + 
-        theme(axis.text.x=element_text(angle=45,hjust=1))
-
-```
-
-Will address each of the variables with many nA separately. 8 of these variables totally. run a separate lm for each of these
-
-Need to transform some numeric variables:
-
-Full:
-
-return.pdol - cube root
-inv.mlsto - log
-timespan - cube root
-pc.pro - leave alone
-client.totinv - log
-
-Many NA:
-
-Project.Value - log
-pc.majpos - log
-
-```{r, echo=FALSE}
-
-#write a function for testing transformations
-
-trans<- function(var= 'timespan', df= all7d, transform = 'cbrt'){
-        #initial q plot
-        init<- qqPlot(df[,var], 
-                       main = 'untouched variable')
-        print(init)
-
-        #make trial dataframe
-        trial = sapply(df[,var], function(x){
-                if(transform == 'log'){
-                        log(x)
-                } else
-
-                if(transform == 'neglog')
-                        {
-                        sign(x)*log(abs(x) + 1)
-                } else
-
-                if(transform == 'sqrt')
-                {
-                        sqrt(x)
-                } else
-                        if(transform == 'cbrt')
-                {
-                        sign(x) * abs(x)^(1/3)
-                } else
-                { x^2 }
-
-                }) %>% 
-                as.data.frame
-
-        colnames(trial)[names(trial) %in% '.']<-'trans.var'
-
-        #qplot for transformation
-        qqtrans<- qqPlot(trial$trans.var, 
-                         main = 'transformed')
-        print(qqtrans)
-
-        ##move on to density plots
-
-        #original variable
-        dens<- ggplot(df, aes_string(x=var)) + geom_histogram(aes(y=..density..))
-        print(dens)
-
-        #transformed variable
-        dens.trans<- ggplot(trial, aes_string(x='trans.var')) + geom_histogram(aes(y=..density..))
-        print(dens.trans)
-}
-
-trans(var= 'timespan', df= all7d, transform = 'cbrt')
-trans(var= 'inv.mlsto', df= all7d, transform = 'log')
-trans(var= 'return.pdol', df= all7d, transform = 'cbrt')
-trans(var= 'pc.majpos', df= all7d, transform = 'log')
-trans(var= 'client.totinv', df= all7d, transform = 'log')
-trans(var= 'Project.Value', df= all7d, transform = 'log')
-
-```
-
-Have now decided on transformations for numeric variables.
-
-Create all7d ready for linear regression analysis
-
-```{r, echo=FALSE}
-
-math.cbrt <- function(x) {
-        sign(x) * abs(x)^(1/3)
-}
-
-all7d$timespan.cbrt<- math.cbrt(all7d$timespan)
-all7d$inv.mlsto.log <- log(all7d$inv.mlsto)
-all7d$client.totinv.log <- log(all7d$client.totinv)
-all7d$Project.Value.log <- log(all7d$Project.Value)
-all7d$pc.majpos.log <- log(all7d$pc.majpos)
-
-```
-
-Root mean square error as initial check. Write code which transforms response variable back to original format by cubing, then finds RMSE.
+Root mean square error as initial check.
 
@@
 
-Now fit full variables to linear regression
+Now fit full variables to random forest, tune mtry and number of trees
+
+Try rpart first
 
@@
 
-full = lm(return.pdol ~ 
-                  inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + client.totinv.log + code.director + 
-                  Num.disc + Business + Biz.type + code.ProjEng,
-          data = all7d)
-summary(full)
+train_ind <- sample(seq_len(nrow(all7d)), size = nrow(all7d)*2/3)
 
-#have a look at the residuals
+train <- all7d[train_ind, ]
+test <- all7d[-train_ind, ]
 
-res<- resid(full)
-plot(all7d$inv.mlsto.log, res)
-abline(0,0)
+#create tree with training data only
+tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
+summary(tree.all7d)
 
-plot(density(res))
-qqPlot(res)
+#have a look at cp values, automatic cross validation in rpart
+printcp(tree.all7d)
+plotcp(tree.all7d)
+#min cp = 0.013 will give simplest tree within 1 std of x-val rel error
+tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.011))
 
-#compare some of the answers
-ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full, all7d, interval= 'predict'))
+split.fun <- function(x, labs, digits, varlen, faclen){
+  # replace commas with spaces (needed for strwrap)
+  labs <- gsub(",", " ", labs)
+  for(i in 1:length(labs)) {
+    # split labs[i] into multiple lines
+    labs[i] <- paste(strwrap(labs[i], width=25), collapse="\n")
+    }
+  labs
+  }
 
-RMSE(ans$answer, ans$full.lm.fit)
+prp(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
+#make prediction with current tree
+tree.pred<-predict(tree.all7d, test)
+#calculate RMSE for test data
+RMSE(test$return.pdol, tree.pred)
+# 51.4 cents, already close to regression..
+
+#try to make a rainbow tree
+heat.tree <- function(tree, low.is.green=FALSE, ...) { # dots args passed to prp
+  y <- tree$frame$yval
+  if(low.is.green)
+    y <- -y
+  max <- max(y)
+  min <- min(y)
+  cols <- rainbow(99, end=.36)[
+    ifelse(y > y[1], (y-y[1]) * (99-50) / (max-y[1]) + 50,
+           (y-min) * (50-1) / (y[1]-min) + 1)]
+  prp(tree, branch.col=cols, box.col=cols, ...)
+  }
+
+heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
 
@@
 
-RMSE is 0.4768 which isn't great in terms of predicting return per dollar.
-R^2 is 0.1905
+RMSE is 0.514 which is worse than regression in predicting return per dollar.
 
-Let's try deleting some full variables
+Decision trees not affected by too many variables - leave all in
+
+Decision trees cannot handle pairwise interactions
+
+Let's try 10 fold cross validation and have a look at the error.
+Use caret package
 
 ```{r, echo=FALSE}
-full.2<- update(full, .~. -client.totinv.log)
-summary(full.2)
-full.3<- update(full.2, .~. -code.director)
-summary(full.3)
-full.4<- update(full.3, .~. -Num.disc)
-summary(full.4)
-full.5<- update(full.4, .~. -code.ProjEng)
-summary(full.5)
-full.6<- update(full.5, .~. -Biz.type + Discipline)
-summary(full.6)
+#10 fold cross validation using caret package
+#specify 10 fold
+tc <- trainControl("cv",10)
+#keep cp static at 0.011
+rpart.grid <- expand.grid(.cp=0.011)
+train.rpart <- train( return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type + code.ProjEng, data=all7d, method="rpart",trControl=tc,
+                  tuneGrid=rpart.grid)
 
-#RMSE check
-
-ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full.6, all7d, interval= 'predict'))
-
-RMSE(ans$answer, ans$full.lm.fit)
-
-```
-
-Feels better. That leaves us with
-
-* inv.mlsto.log
-* Discipline
-* timespan.cbrt
-* no.users
-* pc.pro
-* Discipline
-* Business
-
-RMSE = 0.48266
-
-Going back to full variables, lets try pairwise multiplying..
-
-```{r, echo=FALSE}
-#try diff pairs
-
-pairs.lm<- lm(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
-                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
-              data=all7d)
-
-summary(pairs.lm)
-pairs<- predict(pairs.lm, all7d)
-ans<- data.frame(answer= all7d$return.pdol, pairs.lm = pairs)
-
-RMSE(ans$answer, ans$pairs.lm)
-
-#down to 0.4759
+#RMSE = 0.50386
+# Rsquared = 0.0844, quite bad
 
@@
         reduced = df[!(is.na(df[,var])),]
-        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business", var, sep=" + ")
-        add.one = lm(as.formula(formula), data = reduced)
+        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type + code.ProjEng", var, sep=" + ")
+        add.one = rpart(as.formula(formula), data = reduced)
 
-        print(summary(add.one))
+        #plot decision tree
+        print(
+          heat.tree(add.one, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
+          )
 
-        ans= data.frame(answer= reduced$return.pdol, addone.lm = predict(add.one, reduced, interval= 'predict'))
-
-        print(head(ans))
+        #perform 10 fold xval using caret package
+        tc = trainControl("cv",10)
+        #keep cp static at 0.011
+        rpart.grid = expand.grid(.cp=0.011)
+        train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
+                  tuneGrid=rpart.grid)
 
         print(
-        paste("RMSE=", RMSE(ans$answer, ans$addone.lm.fit) %>% round(4), sep= " ")
+        train.rpart$results
         )

### 2015-08-26 cluster.rmd
update randoom forest

@@
 
+Adjust inv.mlsto, cost.mlsto, balance.mlsto, for inflation!!!!
+
+```{r, echo=FALSE}
+
+library('quantmod', lib= 'C:/Progra~1/R/R-3.2.2/library')
+getSymbols("CPIAUCSL", src='FRED') #Consumer Price Index for All Urban Consumers: All Items
+#[1] "CPIAUCSL"
+tail(CPIAUCSL)
+
+# make an df object of costs
+p<- data.frame('Year' = all7d$Year, 'inv.mlsto'= all7d$inv.mlsto, 'balance.mlsto' = all7d$balance.mlsto,
+               'cost.mlsto' = all7d$cost.mlsto, 'mlsto' = all7d$mlsto)
+
+#calculate annual averave inflation since 2004
+avg.cpi <- apply.yearly(CPIAUCSL, mean)
+#ratio using 2014 as base year
+cf <- avg.cpi/as.numeric(avg.cpi['2014'])
+#create new cf column for date and extract year
+cf<- as.data.frame(cf)
+cf$Date<- rownames(cf)
+cf$Date<- cf$Date %>% as.Date()
+cf$Date<- format(cf$Date,"%Y")
+p<- merge(p, cf, by.x='Year', by.y= 'Date')
+#divide all dollar values by CPIAUCSL column
+p<- transform(p, adj.inv.mlsto = inv.mlsto/CPIAUCSL,
+          adj.balance.mlsto = balance.mlsto/CPIAUCSL,
+          adj.cost.mlsto = cost.mlsto/CPIAUCSL)
+#check
+s<- sample(1:nrow(all7d),10)
+p[s,]
+#replace existing all7d columns
+all7d<- merge(all7d, p %>% select(mlsto, adj.inv.mlsto, adj.balance.mlsto, adj.cost.mlsto) , by='mlsto')
+all7d<- all7d %>% select(-inv.mlsto, - balance.mlsto, -cost.mlsto)
+colnames(all7d)[names(all7d) %in% "adj.inv.mlsto"] <- "inv.mlsto"
+colnames(all7d)[names(all7d) %in% "adj.balance.mlsto"] <- "balance.mlsto"
+colnames(all7d)[names(all7d) %in% "adj.cost.mlsto"] <- "cost.mlsto"
+
+write.csv(all7d, 'all7d.csv')
+
+```
+
 Retrieve nearest neighbour - numeric first

### 2015-08-26 rand_forest.rmd
update randoom forest

@@
 
-library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('caret',lib='C:/Progra~1/R/R-3.2.1/library')
-# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
-# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
-# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
+library('knitr', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.2/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.2/library')
+library('caret',lib='C:/Progra~1/R/R-3.2.2/library')
+library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
+library('rpart',lib='C:/Progra~1/R/R-3.2.2/library')
+library('rpart.plot',lib='C:/Progra~1/R/R-3.2.2/library')
+# library("rpart",lib = 'C:/Program Files/R/R-3.2.2/library')
+library('car', lib = 'C:/Progra~1/R/R-3.2.2/library')
+# library('e1071', lib = 'C:/Progra~1/R/R-3.2.2/library')
+# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.2/library')
+# library('party', lib = 'C:/Progra~1/R/R-3.2.2/library')
+# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.2/library')
+# library('fpc', lib='C:/Progra~1/R/R-3.2.2/library')
+# library('vegan', lib = 'C:/Progra~1/R/R-3.2.2/library')
+# library(devtools,lib='C:/Progra~1/R/R-3.2.2/library')
 
@@
 # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
 opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
@@
                   Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
-summary(tree.all7d)
+# summary(tree.all7d)
 
@@
                      client.totinv + code.director + 
-                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.011))
+                  Num.disc + Business + Biz.type + code.ProjEng, data=train,
+                  control = rpart.control(cp= 0.013))
 
@@
 RMSE(test$return.pdol, tree.pred)
-# 51.4 cents, already close to regression..
+# 49.8 cents, already close to regression..
 
@@
 
-RMSE is 0.514 which is worse than regression in predicting return per dollar.
+RMSE is 0.498 which is worse than regression in predicting return per dollar.
 
@@
 
-#RMSE = 0.50386
-# Rsquared = 0.0844, quite bad
+#RMSE = 0.505
+# Rsquared = 0.095, quite bad
 
@@
 
+        #select cp
+        cp.new = add.one$cptable[which.min(add.one$cptable[,"xerror"]),"CP"]-0.0002
+        add.one = rpart(as.formula(formula), data = reduced,
+                        control = rpart.control(cp= cp.new))
+
         #plot decision tree
@@
         #keep cp static at 0.011
-        rpart.grid = expand.grid(.cp=0.011)
+        rpart.grid = expand.grid(.cp=cp.new)
         train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
@@
 addone(var='JD.Second', df=all7d)
-#RMSE = .4709
+#RMSE = .528
 addone(var='Post.Code', df=all7d)
-#RMSE = .4353
+#RMSE = .499
 addone(var='majority.pos', df=all7d)
-#RMSE = .4541
-addone(var='pc.majpos.log', df=all7d)
-#RMSE = .4553
+#RMSE = .49
+addone(var='pc.majpos', df=all7d)
+#RMSE = .486
 addone(var='code.client', df=all7d)
-#RMSE = .4394
+#RMSE = .534
 addone(var='Billing.Type', df=all7d)
-#RMSE = .4579
+#RMSE = .502
 addone(var='code.contact', df=all7d)
-#RMSE = .3972
+#RMSE = .535
 
-#dont use post code ever. 
-#JD.Second doesn't help that much. acutally makes RMSE worse.
+# none of the extra variables actually appear in the revised trees except JD.Second
+# can only assume the different subsets are what is causing the variation in RMSE
 
@@
 
-Delete Post.Code option
-
-Billing.Type should be paired with inv.mlsto and maybe Discipline
-
-maybe client should be paired with inv.mlsto
-maybe JD.Second should be paired with inv.mlsto and code.client and timespan and pc.pro and no.users?? which ever one does something.
-
-majority.pos pair with inv.mlsto
-pc.majpos pair with inv.mlsto
-
-Write a function that lists the extra variable you're trying and write out the interaction.
+Write a function that can plot a tree
+Limit Year
+choose additional variables to include.
 
@@
 * JD.Second by no.users. Brought down RMSE from 0.4709 to 0.4612...
-
-Now look at cross validation for full variable set:
-
-```{r, echo=FALSE}
-
-#using caret package
-tc <- trainControl("cv",10,savePred=F)
-fit<- train(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
-                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
-              data=all7d,
-            method = 'lm', trControl = tc)
-
-#RMSE = 0.49246
-#compared to non-cross validated
-# RMSE = 0.4759
-```
-
-Not really trying to tune anything via cross validation, however RMSE increased from 0.4759 to 0.49246.
-
-Not THAT bad.
 

### 2015-08-27 rand_forest.rmd
tried boosted trees

@@
 
-heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
+heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)
 
@@
 
-addone<- function(var= 'JD.Second', df= all7d){
-        reduced = df[!(is.na(df[,var])),]
+addone<- function(var= 'JD.Second', min.Year = 2008, df= all7d){
+        reduced = df[!(is.na(df[,var])) & df$Year > min.Year,]
         formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                      client.totinv + code.director + 
-                  Num.disc + Business + Biz.type + code.ProjEng", var, sep=" + ")
+                  Num.disc + Business + Biz.type", var, sep=" + ")
         add.one = rpart(as.formula(formula), data = reduced)
@@
         )
-
 }
@@
 
-addone(var='JD.Second', df=all7d)
+addone(var='JD.Second', min.Year = 2008, df=all7d)
 #RMSE = .528
-addone(var='Post.Code', df=all7d)
+addone(var='Post.Code', min.Year = 2008, df=all7d)
 #RMSE = .499
-addone(var='majority.pos', df=all7d)
+addone(var='majority.pos', min.Year = 2008, df=all7d)
 #RMSE = .49
-addone(var='pc.majpos', df=all7d)
+addone(var='pc.majpos', min.Year = 2008, df=all7d)
 #RMSE = .486
-addone(var='code.client', df=all7d)
+addone(var='code.client', min.Year = 2008, df=all7d)
 #RMSE = .534
-addone(var='Billing.Type', df=all7d)
+addone(var='Billing.Type', min.Year = 2008, df=all7d)
 #RMSE = .502
-addone(var='code.contact', df=all7d)
+addone(var='code.contact', min.Year = 2008, df=all7d)
 #RMSE = .535
@@
 
-addmult<- function(var= c('JD.Second', 'code.client'), combos= c('JD.Second*inv.mlsto.log'), df= all7d){
-        reduced = df[complete.cases(df[,var]),]
-        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business")
+addmult<- function(var= c('JD.Second', 'code.client'), min.Year = 2008, df= all7d){
+        reduced = df[complete.cases(df[,var]) & df$Year>min.Year,]
+        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type")
 
@@
 
-        for(i in 1:length(combos)){
-                formula = paste(formula, combos[i], sep= " + ")
-        }
+        add.mult = rpart(as.formula(formula), data = reduced)
 
-        add.one = lm(as.formula(formula), data = reduced)
+        #select cp
+        cp.new = add.mult$cptable[which.min(add.mult$cptable[,"xerror"]),"CP"]-0.0002
+        add.mult = rpart(as.formula(formula), data = reduced,
+                        control = rpart.control(cp= cp.new))
 
-        print(summary(add.one))
+        #plot decision tree
+        print(
+          heat.tree(add.mult, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)
+          )
 
-        ans= data.frame(answer= reduced$return.pdol, addone.lm = predict(add.one, reduced, interval= 'predict'))
-
-        print(head(ans))
+        #perform 10 fold xval using caret package
+        tc = trainControl("cv",10)
+        #keep cp static at 0.011
+        rpart.grid = expand.grid(.cp=cp.new)
+        train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
+                  tuneGrid=rpart.grid)
 
         print(
-        paste("RMSE=", RMSE(ans$answer, ans$addone.lm.fit) %>% round(4), sep= " ")
+        train.rpart$results
         )
@@
 #try Billing.Type with inv.mlsto.log
-addmult(var= c('Billing.Type'), combos= c('Billing.Type*inv.mlsto.log'), df= all7d)
-# no
-
-#try code.client*inv.mlsto.log
-addmult(var= c('code.client'), combos= c('code.client*inv.mlsto.log'), df= all7d)
-#made a difference, brought RMSE down to .4076 :)!
-
-#try JD.Second with inv.mlsto.log (no), code.client (too intense), timespan (no), pc.pro (no), no.users (yes)?
-addmult(var= c('JD.Second'), combos= c('JD.Second*no.users'), df= all7d)
-#add JD.Second*no.users
-
-addmult(var= c('majority.pos'), combos= c('majority.pos*inv.mlsto.log'), df= all7d)
-#no
-
-addmult(var= c('pc.majpos.log'), combos= c('pc.majpos.log*inv.mlsto.log'), df= all7d)
-#no
+## YES! print this one!! 
+addmult(var= c('pc.majpos', 'majority.pos'), min.Year = 2008, df= all7d)
+#RMSE at 48.6 cents
 
@@
 
-It is worthwhile to include:
+Adding aditional variables is making everything worse.
+Try randomForest
 
-* code.client by inv.mlsto.log because brought down RMSE from 0.4394 (code.client) to 0.4076 (code.client + code.client by inv.mlsto.log)
-* JD.Second by no.users. Brought down RMSE from 0.4709 to 0.4612...
+```{r, echo=FALSE}
+#perform 10 fold xval using caret package
+tc = trainControl("cv",5)
+rf.grid = expand.grid(mtry= 1:10)
+min.form= "return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type"
+train.rf = train(as.formula(min.form), data=all7d, method="rf",trControl=tc,
+                    tuneGrid=rf.grid)
+#result - use mtry= 5
 
-Now, how to get variable importance from lm..
+#make first single randomForest
+first.rf<- randomForest(as.formula(min.form),
+                  data=all7d,mtry=1,ntree=500,importance=TRUE)
+
+#already have made train and test set
+#try to predict medv
+rf.all<-randomForest(as.formula(min.form),data=all7d,mtry=1,ntree=500,importance=TRUE)
+
+#create vectors full of '0's
+oob.err<-double(7)
+test.err<-double(7)
+for(mtry in 1:7){
+        fit<-randomForest(as.formula(min.form),data=train, mtry=ntry, ntree=500)
+        oob.err[mtry]<-fit$mse[mtry]
+        test.err[mtry]<- mean((predict(fit,test)-test$return.pdol)^2) %>% sqrt()
+        cat(mtry," ")
+}
+
+qplot(1:mtry,test.err,geom=c("point","line"),color="pink")+
+        geom_line(aes(1:mtry,oob.err),colour="blue")+
+        geom_point(aes(1:mtry,oob.err),colour="blue")
+
+#mtry = 3 is optimal! test RMSE of 48.2 cents ....
+oob.err<-double(5)
+test.err<-double(5)
+ntree = c(300,400,500,600,700)
+for(i in 1:5){
+        fit<-randomForest(as.formula(min.form),data=train, mtry=3, ntree=ntree[i])
+        oob.err[i]<-fit$mse[i]
+        test.err[i]<- mean((predict(fit,test)-test$return.pdol)^2) %>% sqrt()
+        cat(i," ")
+}
+
+qplot(ntree,test.err,geom=c("point","line"),color="pink")+
+        geom_line(aes(ntree,oob.err),colour="blue")+
+        geom_point(aes(ntree,oob.err),colour="blue")
+#doesn't really matter, use 400!
+
+#final forest with full variables
+
+full.forest<- randomForest(as.formula(min.form),data=all7d,mtry=5,ntree=500,importance=TRUE)
+plot(full.forest)
+importance(full.forest)
+varImpPlot(full.forest)
+
+full.forest<- randomForest(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type +code.client, data= all7d[complete.cases(all7d$code.client),],
+                  mtry=5,ntree=500,importance=TRUE)
+
+```
+
+Try adding variables
 
@@
 
-#variable importance for full variables only - pairs.lm
-#use varImp from caret package - uses t-test statistic
+addmult<- function(var= c('JD.Second', 'code.client'), min.Year = 2008, df= all7d){
+        reduced = df[complete.cases(df[,var]) & df$Year>min.Year,]
+        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+                     client.totinv + code.director + 
+                  Num.disc + Business + Biz.type")
 
-imp<- varImp(pairs.lm, scale=FALSE)
-imp$variable<- rownames(imp)
-imp<- imp %>% select(variable, Overall)
-imp %>% arrange(-Overall) %>% head()
+        for(i in 1:length(var)){
+                formula = paste(formula, var[i], sep= " + ")
+        }
 
-sum(imp$Overall)
+        add.mult = randomForest(as.formula(formula),mtry=5, ntree=500, data = reduced)
+
+        #perform 10 fold xval using caret package
+        tc = trainControl("cv",5)
+        #keep cp static at 0.011
+        rf.grid = expand.grid(mtry= 5:6)
+        train.rf = train(as.formula(formula), data=df, method="rf",trControl=tc,
+                  tuneGrid=rf.grid)
+
+        print(
+        train.rf
+        )
+
+}
+
+#try Billing.Type with inv.mlsto.log
+## YES! print this one!! 
+addmult(var= c('pc.majpos', 'majority.pos'), min.Year = 2008, df= all7d)
+#RMSE at 48.6 cents
 

### 2015-08-28 Regression.rmd
done with random forest

@@
 # RMSE = 0.4759
+
+#try using caret package on code.client
+fit<- train(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
+                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business + code.client,
+              data=all7d,
+            method = 'lm', trControl = tc)
+
 ```

### 2015-08-28 rand_forest.rmd
done with random forest

@@
 
+addmult(var= c('pc.majpos', 'majority.pos','JD.Second'), min.Year = 2008, df= all7d)
+
+test<- all7d[complete.cases(all7d[,c('code.client','pc.majpos')]) &
+                     all7d$Year>2008 & all7d$Discipline == 'Structural',]
+test$code.client<- droplevels(test$code.client)
+test %>% str
+
 ```
@@
 
-addmult<- function(var= c('JD.Second', 'code.client'), min.Year = 2008, df= all7d){
-        reduced = df[complete.cases(df[,var]) & df$Year>min.Year,]
-        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+addmult<- function(var= c('pc.majpos','majority.pos', 'code.client', 'code.contact'), min.Year = 2008, df= all7d, 
+                   Disc= "Structural"){
+        #subset data frame based on complete cases with additional variables
+        reduced = df[complete.cases(df[,var]) & df$Year>min.Year & df$Discipline == Disc,]
+
+        if('code.client' %in% var){
+        sample = factor(sample(unique(reduced$code.client),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.client %in% sample,]
+
+        reduced$code.client= factor(reduced$code.client)
+        }
+
+        if('code.contact' %in% var){
+        sample1 = factor(sample(unique(reduced$code.contact),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.contact %in% sample1,]
+
+        reduced$code.contact= factor(reduced$code.contact)
+        }
+
+        #make formula
+        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + pc.pro + 
                      client.totinv + code.director + 
@@
 
-        add.mult = randomForest(as.formula(formula),mtry=5, ntree=500, data = reduced)
+        #make forest
+        red.forest= randomForest(as.formula(formula), data=reduced, mtry=5,ntree=500,importance=TRUE)
+
+        print(
+                varImpPlot(red.forest)
+        )
 
@@
         #keep cp static at 0.011
-        rf.grid = expand.grid(mtry= 5:6)
+        rf.grid = expand.grid(mtry= 5)
         train.rf = train(as.formula(formula), data=df, method="rf",trControl=tc,
@@
 
-#try Billing.Type with inv.mlsto.log
-## YES! print this one!! 
-addmult(var= c('pc.majpos', 'majority.pos'), min.Year = 2008, df= all7d)
-#RMSE at 48.6 cents
+addmult(var= c('pc.majpos','majority.pos', 'code.client'), min.Year = 2008, df= all7d, 
+                   Disc= "Structural")
+
+#with code.client i'm still getting 48.9 cents RMSE! that sucks!!!! with linear regression i was getting
+
+addmult(var= c('pc.majpos','majority.pos','code.contact'), min.Year = 2008, df= all7d, 
+                   Disc= "Structural")
+#50.6 cents
+#importance levels of code.client and code.contact are up there but they're not THAT high. like midway out of all options.
 

### 2015-09-04 Regression.rmd
further work on regression.rmd

@@
 
-final.output<- function(df=all7d, samp = 800) {
+final.output<- function(df=all7d, samp = 600) {
         #create sample df
@@
         print(
-                list("sample case", t(sample %>% select(-Year, -timespan.cbrt, -inv.mlsto.log, -client.totinv.log, -pc.majpos.log)))
+                list("sample case", t(sample %>% select(-Year, -timespan.cbrt, -inv.mlsto.log, -client.totinv.log)))
         )
@@
         #make reduced dataset with complete cases for all variables we want to use
-        reduced = df[complete.cases(df[,variables]),]
+        reduced = df[-samp,]
+        reduced = reduced[complete.cases(reduced[,variables]),]
 
@@
 
-        final.RMSE= RMSE(ans$answer, ans$final.lm.fit) %>% round(4)
+        #random RMSE
+        r= mean(reduced$return.pdol)
+        random.RMSE= RMSE(reduced$return.pdol, r, print=T) %>% round(3)
+
+        print(
+        paste("Random Guess RMSE=", random.RMSE , sep= " ")
+        )
+
+        #final RMSE for all rows in reduced in a regression
+        final.RMSE= RMSE(ans$answer, ans$final.lm.fit) %>% round(3)
 
@@
         print(
-                paste("Invoiced= $", sample$inv.mlsto, ",   predicted cost= $", predict.cost %>% round(1), sep="")
+                paste("Invoiced= $", sample$inv.mlsto %>% round(2), ",   predicted cost= $", predict.cost %>% round(1), sep="")
         )
@@
         print(
-                paste("Actual cost= $", sample$cost.mlsto, sep="")
+                paste("Actual cost= $", sample$cost.mlsto %>% round(2), sep="")
         )
@@
 
-final.output(df=all7d, samp = 500)
+final.output(df=all7d, samp = 800)
 
 ```
+
+Check what RMSE would be for random guess
+
+```{r, echo=F}
+
+m<- mean(all7d$return.pdol)
+
+RMSE<- function(answer, guess, print= TRUE){
+        error= answer-guess
+        root.error= sqrt(mean(error^2))
+        if(print == TRUE){
+        print(root.error)
+        }
+}
+
+RMSE(all7d$return.pdol, m, print=T)
+
+```
+
+RMSE from random guess is 53 cents.
+
+# delete any code.client or contact. client which has less than 3 entries!
+# need to revisit how i got my inv.mlsto, balance.mlsto, cost.mlsto for job 2013.301.2

### 2015-09-04 rand_forest.rmd
further work on regression.rmd

@@
 # opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
-all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
-all7d<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
-all7d$Post.Code<- as.factor(all7d$Post.Code)
+all7d.tree<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
+all7d.tree<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
+all7d.tree$Post.Code<- as.factor(all7d$Post.Code)
 
@@
 
-train_ind <- sample(seq_len(nrow(all7d)), size = nrow(all7d)*2/3)
+train_ind <- sample(seq_len(nrow(all7d.tree)), size = nrow(all7d.tree)*2/3)
 
-train <- all7d[train_ind, ]
-test <- all7d[-train_ind, ]
+train <- all7d.tree[train_ind, ]
+test <- all7d.tree[-train_ind, ]
 
 #create tree with training data only
-tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+tree.all7d.tree<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                      client.totinv + code.director + 
                   Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
-# summary(tree.all7d)
+# summary(tree.all7d.tree)
 
 #have a look at cp values, automatic cross validation in rpart
-printcp(tree.all7d)
-plotcp(tree.all7d)
+printcp(tree.all7d.tree)
+plotcp(tree.all7d.tree)
 #min cp = 0.013 will give simplest tree within 1 std of x-val rel error
-tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
+tree.all7d.tree<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                      client.totinv + code.director + 
@@
 
-prp(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
+prp(tree.all7d.tree, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
 #make prediction with current tree
-tree.pred<-predict(tree.all7d, test)
+tree.pred<-predict(tree.all7d.tree, test)
 #calculate RMSE for test data
@@
 
-heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)
+heat.tree(tree.all7d.tree, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)
 

### 2015-09-04 final_app/server.R

@@ -1,13 +1,13 @@
 setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
 # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library('shiny',lib='C:/Progra~1/R/R-3.1.3/library')
-library('shinythemes',lib='C:/Progra~1/R/R-3.2.0/library')
-library('FNN',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.2/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.2/library')
+library('shiny',lib='C:/Progra~1/R/R-3.2.2/library')
+library('shinythemes',lib='C:/Progra~1/R/R-3.2.2/library')
+library('FNN',lib='C:/Progra~1/R/R-3.2.2/library')
 # library('ggplot2')
 # library("dplyr")
 # library("plyr")
@@ -18,7 +18,7 @@ library('FNN',lib='C:/Progra~1/R/R-3.2.1/library')
 
 
 all7<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7.csv')[,-1]
- all7$Start.Date<- as.Date(all7$Start.Date)
+all7$Start.Date<- as.Date(all7$Start.Date)
 #delete anything more than 4 for inv.vs.cost
 #delete anything less than .2 for inv.vs.cost
 all7<- all7 %>% filter(!(inv.vs.cost<0.2), !(inv.vs.cost>4))

### 2015-10-07 Regression.rmd
BN prep

@@
                          client.totinv,code.director,Num.disc,Business,Biz.type,
-                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,majority.pos,
+                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,
+                         majority.pos,
                          pc.majpos)
@@
 
-RMSE<- function(answer, guess){
+#make new dataframe for future use with normalised variables and variables that you want to use only
+all8a<- all7d %>% select(no.users, Discipline, pc.pro, Business, code.client, JD.Second, timespan.cbrt,
+                         inv.mlsto.log,client.totinv.log, pc.majpos.log, code.contact, Billing.Type,
+                         majority.pos, return.pdol, mlsto, Year)
+write.csv(all8a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')
+
+RMSE<- function(answer, guess, print= TRUE){
         error= answer-guess
         root.error= sqrt(mean(error^2))
-        # print(root.error)
+        if(print == TRUE){
+        print(root.error)
+        }
 }
 
-final.output<- function(df=all7d, samp = 600) {
+final.output<- function(df=all7d, samp = 800) {
         #create sample df
@@
                                   'Biz.type','Post.Code','code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto',
+                                  'dis.sc.mlsto',
                                   'balance.mlsto','client.totinv.log','pc.majpos')]
@@
 
-        cats= c('code.client', 'code.contact', 'JD.Second', 'Business', 'Discipline', 'majority.pos')
+        cats= c('code.client', 'code.contact', 'JD.Second', 'Business', 'no.users', 'majority.pos')
 
@@
                         }
+        }
+
+        num.cats= c('timespan.cbrt', 'inv.mlsto.log', 'pc.pro')
+
+        for(i in 1:length(num.cats)){
+                if(num.cats[i] %in% vars){
+                        span = 0.05*(max(reduced[,num.cats][i])-min(reduced[,num.cats[i]]))
+                        range = c(df[samp,num.cats[i]]-span, 
+                                  df[samp,num.cats[i]]+span)
+                        cats.red = reduced[reduced[,num.cats[i]]<range[2] & reduced[,num.cats[i]]>range[1],]
+                        cat.RMSE = RMSE(cats.red$return.pdol, cats.red$predict) %>% round(4)
+                        print(
+                                paste(num.cats[i], "RMSE =", cat.RMSE, sep= " "))
+                        print(
+                                paste("Number of cases =", nrow(cats.red), sep= " "))
+                        print(
+                                if(num.cats[i] == 'pc.pro'){
+                                        head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto','return.pdol','predict',
+                                                 num.cats[i])])} 
+                                else if(num.cats[i] == 'timespan.cbrt'){
+                                                head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto',
+                                                                 'return.pdol','predict', 'timespan')])}
+                                else {
+                                        head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto',
+                                                                 'return.pdol','predict')])
+                                }
+
+                        )}
         }
@@
 
-RMSE<- function(answer, guess, print= TRUE){
-        error= answer-guess
-        root.error= sqrt(mean(error^2))
-        if(print == TRUE){
-        print(root.error)
-        }
-}
-
 RMSE(all7d$return.pdol, m, print=T)

### 2015-10-07 cluster.rmd
BN prep

--- 2015-08-21 cluster.rmd
@@
 # library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
-# library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
 # library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
@@
                          Num.disc, Business,
-                         Biz.type, code.contact, pc.majpos,
+                         Biz.type, code.contact, pc.majpos, dis.sc.mlsto
                          code.client, Project.Value, JD.Second, Post.Code, code.ProjEng, 
@@
 write.csv(all7d,'all7d.csv')
+
+```
+
+Adjust inv.mlsto, cost.mlsto, balance.mlsto, for inflation!!!!
+
+```{r, echo=FALSE}
+
+library('quantmod', lib= 'C:/Progra~1/R/R-3.2.2/library')
+getSymbols("CPIAUCSL", src='FRED') #Consumer Price Index for All Urban Consumers: All Items
+#[1] "CPIAUCSL"
+tail(CPIAUCSL)
+
+# make an df object of costs
+p<- data.frame('Year' = all7d$Year, 'inv.mlsto'= all7d$inv.mlsto, 'balance.mlsto' = all7d$balance.mlsto,
+               'cost.mlsto' = all7d$cost.mlsto, 'mlsto' = all7d$mlsto, 'dis.sc.mlsto'=all7d$dis.sc.mlsto)
+
+#calculate annual averave inflation since 2004
+avg.cpi <- apply.yearly(CPIAUCSL, mean)
+#ratio using 2014 as base year
+cf <- avg.cpi/as.numeric(avg.cpi['2014'])
+#create new cf column for date and extract year
+cf<- as.data.frame(cf)
+cf$Date<- rownames(cf)
+cf$Date<- cf$Date %>% as.Date()
+cf$Date<- format(cf$Date,"%Y")
+p<- merge(p, cf, by.x='Year', by.y= 'Date')
+#divide all dollar values by CPIAUCSL column
+p<- transform(p, adj.inv.mlsto = inv.mlsto/CPIAUCSL,
+          adj.balance.mlsto = balance.mlsto/CPIAUCSL,
+          adj.cost.mlsto = cost.mlsto/CPIAUCSL,
+          adj.dis.sc = dis.sc.mlsto/CPIAUCSL)
+#check
+s<- sample(1:nrow(all7d),10)
+p[s,]
+#replace existing all7d columns
+all7d<- merge(all7d, p %>% select(mlsto, adj.inv.mlsto, adj.balance.mlsto, adj.cost.mlsto, adj.dis.sc), by='mlsto')
+all7d<- all7d %>% select(-inv.mlsto, - balance.mlsto, -cost.mlsto, -dis.sc.mlsto)
+colnames(all7d)[names(all7d) %in% "adj.inv.mlsto"] <- "inv.mlsto"
+colnames(all7d)[names(all7d) %in% "adj.balance.mlsto"] <- "balance.mlsto"
+colnames(all7d)[names(all7d) %in% "adj.cost.mlsto"] <- "cost.mlsto"
+colnames(all7d)[names(all7d) %in% "adj.dis.sc"] <- "dis.sc.mlsto"
+
+write.csv(all7d, 'all7d.csv')
 

### 2015-10-07 final_app/server.R

@@ -28,7 +28,8 @@ df <- read.table(text = " id  min  max
     Sp1     150          151     ", header=TRUE)
 
 #make a new dataframe with information about crossbar sizes
-df2<- data.frame(id=c(rep("Sp1",3)), ymin= c(130, 139,146), ymax=c(170, 158, 155), probability=c('95%','75%','50%'))
+df2<- data.frame(id=c(rep("Sp1",3)), ymin= c(19000, 22000,24000),
+                 ymax=c(33000, 30000, 28000), probability=c('95%','75%','50%'))
 df2 <- within(df2, 
               probability <- factor(probability, 
                                     levels=c('95%','75%','50%')))
@@ -154,17 +155,17 @@ shinyServer(
                         s= ggplot(df[1,]) +
                                 geom_crossbar(data=df2, aes(ymin = ymin, ymax = ymax, x = id, y=ymin, alpha=probability),
                                               fill='blue', linetype='blank', width=0.25)+
-                                geom_errorbar(aes(ymax=150, ymin=150, x=id, y=min), width=0.25, colour='navyblue', size=1.5) +
-                                geom_errorbar(aes(ymax=100, ymin=100, x=id, y=min), width=0.25, colour='red') +
-                                scale_y_continuous(breaks = seq(80, 170, by = 10)) +
-                                ylim(80,170) +
-                                labs(y="Final Spend divided by Quote (%)", x="New Project", title= "Prediction of Final Spend vs Initial Quote") +
+                                geom_errorbar(aes(ymax=26000, ymin=26000, x=id, y=min), width=0.25, colour='navyblue', size=1.5) +
+                                geom_errorbar(aes(ymax=20000, ymin=20000, x=id, y=min), width=0.25, colour='red', linetype = 2) +
+                                scale_y_continuous(breaks = seq(15000, 40000, by = 5000)) +
+                                ylim(15000,40000) +
+                                labs(y="Final Spend ($)", x="New Project", title= "Prediction of Final Spend") +
                                 theme(axis.text.y = element_blank(),
                                       plot.title = element_text(size=18, face='plain'),
                                       axis.title.y = element_text(size = 16),
                                       axis.title.x = element_text(size=16),
                                       axis.text = element_text(size = 16)) +
-                                annotate("text", x = 1.3, y = 101, label = "Final Spend = Initial Quote", colour='red', size=4)+
+                                annotate("text", x = 1.3, y = 20000, label = "Final Spend = Initial Quote", colour='red', size=4)+
[…]

### 2015-10-07 final_app/ui.R

@@ -32,13 +32,14 @@ shinyUI(fluidPage(
                 
                 column(3,
                        
-                       selectInput("jobtype", "Primary Job Type:",
-                                   list("Building Structures" = "1. BuildingStructures", 
-                                        "Land Infrastructure" = "2. LandInfrastructure", 
-                                        "Environmental Planning" = "3. EnvironmentalPlanning",
-                                        "Integrated Water Management" = "4. IntegratedWaterManagement",
-                                        "Special Structures"= "5. Special Structures",
-                                        "Other" = "Other")),
+                       selectizeInput(
+                               'JD.Second', 'Job Type:', 
+                               choices = levels(all7$JD.Second),
+                               options = list(
+                                       placeholder = 'Please start typing',
+                                       onInitialize = I('function() { this.setValue(""); }'),
+                                       maxItems=1
+                               )),
                        
                        selectizeInput(
                                'Business', 'Client Industry:', 
@@ -83,7 +84,7 @@ shinyUI(fluidPage(
                 
                 column(3,
                        
-                       selectInput("bulk.pos", "Bulk of Work by:",
+                       selectInput("majority.pos", "Bulk of Work by:",
                                    list("Director" = "Director", 
                                         "Senior Professional" = "Senior Professional", 
                                         "Mid Professional" = "Mid Professional",
@@ -101,14 +102,9 @@ shinyUI(fluidPage(
                 
                 column(3,
                        
-                       selectizeInput(
-                               'JD.Second', 'Secondary Job Type:', 
-                               choices = levels(all7$JD.Second),
-                               options = list(
-                                       placeholder = 'Please start typing',
-                                       onInitialize = I('function() { this.setValue(""); }'),
-                                       maxItems=1
-                               )),
+                       sliderInput(
+                               'pc.pro', '% Hours by Professional:', 
+                               min=0, max=100, value=50, step=5, round=0),
                        
                        sliderInput("inv.mlsto", "Approximate Fee:",
                                    min=500, max=250000, value=20000, step=500, round=2
@@ -145,8 +141,8 @@ shinyUI(fluidPage(
                                         "Public" = "public"
                                         )),
                        
-                       sliderInput("pc.pro", "% Hours by Professional:",
-                                   min=0, max=100, value=50, step=5, round=0
+                       sliderInput("timespan", "Timespan (days):",
+                                   min=0, max=365, value=60, step=20, round=0
                        ),
                        
                        submitButton(text="Calculate")                 

### 2015-10-12 finalwriteup/draft.rmd
BN prep

@@
 * combination of cforest, anova, random forest
+
+# Chapter ? Method
+
+##Bayesian Network
+
+* could not handle _any_ numeric variables, had to discretise all numeric vars
+    * used normalised numeric variables
+    * I did this by making a hierarchical dendrogram to visualise the clusters
+    * then I chose the number of clusters, drew the red rectangle
+    * summarised the clusters to find max and min values for the variable within each cluster. Rounded this value up
+    * manually discretised variable using rounded values
+
+* first trialled complete cases of number of users, Discipline, percent pro, Business, client code, job detail, majority position, timespan, invoiced amount, percent hours by majority position, and return per dollar.
+This gave 907 complete cases to work with.
+
+*using Genie Smile.
+   * noticed that because most of the cases (31%) fall under a return per dollar of between ___ and ____, results are heavily biased towards that category.
+This could be a problem in the random forest model I made and possibly the regression model.
+   * Need to implement a package that oversamples from the minority classes to create a balanced range of return per dollar classes.
+       * after doing some reading about ROSE and SMOTE package (which employs oversampling) readers suggest oversampling is best for very very skewed data.
+Otherwise assigning weighted costs for misclassification to functions is better.
+       * Another alternative to oversampling with multiple classes is to use one function to classify as in the majority class or not.
+and then another function to figure out _which_ minority class the case would belong to.
+       * never mind can just re-bin return per dollar so that classes are equal sizes!!!

### 2015-10-12 BNprep.R

@@ -6,7 +6,7 @@ library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
 library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
 library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
 library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('hclust',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
 
 # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
 all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')[,-1]
@@ -146,7 +146,7 @@ for(i in 1:k.choose){
 #create clusters based on summary:
 
 all8a$b.pcmajpos <- man.clust(first.vec=all8a$pc.majpos.log, 
-                                   breaks.man=c(3.1,3.9,4.1,4.3,4.5,4.6))
+                                   breaks.man=c(3.1,3.9,4.1,4.3,4.5,4.7))
 
 summary(all8a$b.pcmajpos)
 
@@ -173,10 +173,13 @@ for(i in 1:k.choose){
 
 #create clusters based on summary:
 
-all8a$b.rpdol <- man.clust(first.vec=all8a$return.pdol, 
-                                   breaks.man=c(-1,-.5,0,.15,0.3,0.6,0.85,1.3,1.76,3))
-
-summary(all8a$b.rpdol)
+#experiment with cut_number function because we want even sized bins for prediction
+all8a$b.rpdol<- cut_number(all8a$return.pdol,8) #this works fine! comment out cluster method below
+# summary(all8a$b.rpdol)
+# all8a$b.rpdol <- man.clust(first.vec=all8a$return.pdol, 
+#                                    breaks.man=c(-1,-.5,0,.15,0.3,0.6,0.85,1.3,1.76,3))
+# 
+# summary(all8a$b.rpdol)
 
 ####
 ##users
@@ -203,7 +206,19 @@ summary(all8b)
 
 write.csv(all8b,'C:/Users/n9232371/Documents/Consultbusiness/data/all8b.csv' )
 
-#what happens to missing values if i look at post 2008?
-all8b<- merge(all8b, all8a %>% select(Year, mlsto), by='mlsto')
[…]

### 2015-10-19 boost_tree.rmd
started boosting

---
title: "Boosted"
author: "Amy Cook"
date: "October 19, 2015"
output: word_document
---

```{r, load library and packages, include= FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')

# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
```

Read in data. Change return.pdol to 0,1 variable

```{r, echo=FALSE}
all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')[,-1]

all8a$b.rpdol<- all8a$return.pdol 
all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 0, 1)

summary(all8a$b.rpdol)

all9a<- all8a
```

Create test and train sample

```{r, echo=FALSE}
set.seed(300)
sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
train<- all9a[sample,]
test<- all9a[-sample,]

```

Try building a boosted tree wihtout any training

```{r, echo=FALSE}
boost.1 <- gbm(b.rpdol~.- mlsto - return.pdol-Year-code.contact - code.client, 
               distribution = "bernoulli", data=train,
          n.trees = 10000,
          shrinkage = 0.001,
          interaction.depth = 1,
          n.minobsinnode = 10,
          bag.fraction = 0.5)

print(boost.1)
summary(boost.1)

#look at first tree
pretty.gbm.tree(boost.1,9000)
boost.1$var.names

#view OOB error plot
gbm.perf(boost.1, method="OOB", plot.it=TRUE)
#error still rapidly reducing at 10000 trees

#partial dependence plots:
i=8
a<- plot.gbm(boost.1, i, type="response")
a
#Discipline, Business, inv.mlsto.log, client.totinv.log, Billing.Type dont matter?

```

First tune shrinkage and plot. Get rid of year because in future, won't have historical data on year..

```{r, echo=FALSE}
formula = "b.rpdol~.- mlsto - return.pdol-Year-code.client-code.contact"

errors<- double(4)
shrink.seq<-c(.001,.005,.01,.05)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
for(i in 1:4){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees=5000,
                 shrinkage=shrink.seq[i],interaction.depth=3)
        pred<- predict(fit, test, n.tree=5000, type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=shrink.seq, "error" = errors[1:4])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

```

Try caret method of tuning

```{r, echo=FALSE, include=FALSE}
# must turn outcome into two level factor
# all9a$f.rpdol<- as.factor(all9a$b.rpdol)
# set.seed(100)
# sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
# train<- all9a[sample,]
# test<- all9a[-sample,]
# 
# # set seeds
# set.seed(123)
# seeds <- vector(mode = "list", length = 51)
# for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
# ## For the last model:
# seeds[[51]] <- sample.int(1000, 1)
# 
# # first tune shrinkage and number of trees
# cv.Control <- trainControl(method = "cv",
#                            number = 10,
#                            seeds= seeds)
# gbmGrid <- expand.grid(shrinkage = c(.001,.005),
#                        n.trees = c(2000), interaction.depth = 2,
#                        n.minobsinnode = 10)
# set.seed(2)
# gbmFit <- train(f.rpdol~.- mlsto - return.pdol-Year, data= train,
#                  method = "gbm", trControl = cv.Control, verbose = FALSE,
#                  bag.fraction = 0.5, tuneGrid = gbmGrid,
#                 metric = 'kappa')
# 
# plot(gbmFit)
```

Caret package is not working, tune manually:

```{r, echo=FALSE}

#shrinkage should be 0.001, now find number of trees
errors<- double(5)
ntree.seq<-c(4000,8000,10000,12000,14000)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
for(i in 1:5){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= ntree.seq[i],
                 shrinkage=0.001,interaction.depth=3)
        pred<- predict(fit, test, n.tree=ntree.seq[i], type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=ntree.seq, "error" = errors[1:5])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

#10000 trees seems best?

#----------------------------------------------------------------------

#shrinkage should be 0.001, 10000 trees, now find interaction.depth
errors<- double(5)
int.seq<-c(1,2,4,6,8)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
for(i in 1:5){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 10000,
                 shrinkage=0.001,interaction.depth=int.seq[i])
        pred<- predict(fit, test, n.tree=10000, type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=int.seq, "error" = errors[1:5])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

## use interaction.depth = 4?

#------------------------------------------------
# now tune min obs in node
errors<- double(4)
obs.seq<-c(10,20,40,50)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
for(i in 1:4){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 10000,
                 shrinkage=0.001,interaction.depth=4,
                 n.minobsinnode = obs.seq[i])
        pred<- predict(fit, test, n.tree=10000, type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=obs.seq, "error" = errors[1:4])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

### 2015-10-19 boost_prep.R

@@ -1,0 +1,29 @@
+
+## prep for boosted trees
+
+#discretise numeric variables and change all 'NA' to blank
+
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
+
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')[,-1]
+head(all8a)
+
+#try to predict profit/loss for return per dollar!
+###
+#return.pdol
+# split into profit/loss
+
+all8a$b.rpdol<- all8a$return.pdol 
+all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, "loss", "profit")
+all8a$b.rpdol<- as.factor(all8a$b.rpdol)
+
+summary(all8a$b.rpdol)
+
+#build boosted tree
+
+

### 2015-10-20 boost_tree.rmd
more boosting

@@
 library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')
 
@@
 all8a$b.rpdol<- all8a$return.pdol 
-all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 0, 1)
+all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 1, 0)
 
@@
 all9a<- all8a
+write.csv(all9a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')
+
 ```
@@
 
+Try deleting code.client variable
+
 ```{r, echo=FALSE}
-formula = "b.rpdol~.- mlsto - return.pdol-Year-code.client-code.contact"
+
+#tune which variables should be included!!
+vars.excl = c("JD.Second", "Business", "JD.Second-Business","")
 
 errors<- double(4)
-shrink.seq<-c(.001,.005,.01,.05)
 conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                    '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
+auc<- double(4)
+colours <- c("red","green","blue", "pink")
+
+for(i in 1:4){
+        if(vars.excl[i] %in% ""){
+              formula<- "b.rpdol~.- mlsto - return.pdol-Year - code.contact - code.client"  
+        } else{
+        formula<- paste("b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact",
+                        "-", vars.excl[i], sep=" ")
+        }
+
+        fit<-gbm(as.formula(formula), 
+                 data=train, distribution="bernoulli", n.trees=5000,
+                 shrinkage=0.001,
+                 interaction.depth=3)
+        pred<- predict(fit, test, n.tree=5000, type="response")
+        errors[i]<- sum(temp)/nrow(test)
+
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres=T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
+        }
+
+        auc[i]<- auc(pred.p)
+
+        #confusion matrix
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred1' = pred1, 'ans' = test$b.rpdol)
+        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred1, temp2$ans))
+        cat(i," ")
+}
+
+conf.matrix
+auc
+
+#create new data frame 'shrink' to place trees and errors vectors together, then plot:
+shrink<- data.frame("shrinkage"=shrink.seq, "auc" = auc[1:4])
+ggplot(shrink, aes(x=shrinkage, y= auc)) + geom_line() + geom_point()+ geom_line()
+
+#best to not have JD.Second OR Business!!!  :(
+```
+
+Tune shrinkage value:
+```{r, echo=FALSE}
+
+formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business"
+
+errors<- double(4)
+shrink.seq<-c(0.0005, .001,.005, 0.01)
+conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
+                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
+auc<- double(4)
+colours <- c("red", "yellow", "green", "blue")
+
 for(i in 1:4){
@@
         pred<- predict(fit, test, n.tree=5000, type="response")
-        pred<- ifelse(pred>=0.5,1,0)
-        temp<- ifelse(pred == test$b.rpdol, 0,1)
         errors[i]<- sum(temp)/nrow(test)
 
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres=T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
+        }
+
+        auc[i]<- auc(pred.p)
+
         #confusion matrix
-        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred1' = pred1, 'ans' = test$b.rpdol)
+        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred1, temp2$ans))
+        cat(i," ")
+}
+
+conf.matrix
+auc
+
+#create new data frame 'shrink' to place trees and errors vectors together, then plot:
+shrink<- data.frame("shrinkage"=shrink.seq, "auc" = auc[1:4])
+ggplot(shrink, aes(x=shrinkage, y= auc)) + geom_line() + geom_point()+ geom_line()
+
+#best case auc = 0.756 with shrinkage = 0.001
+
+```
+
+```{r, echo=FALSE}
+
+#shrinkage should be 0.001, now find number of trees
+errors<- double(5)
+ntree.seq<-c(2000,3000,4000,6000,8000)
+conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
+                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
+                   '5' = data.frame(c(0,0), c(0,0)))
+auc<- double(5)
+colours <- c("red", "yellow", "green", "blue", "purple")
+
+for(i in 1:5){
+        fit<-gbm(as.formula(formula), 
+                 data=train, distribution="bernoulli", n.trees= ntree.seq[i],
+                 shrinkage=0.001,interaction.depth=3)
+        pred<- predict(fit, test, n.tree=ntree.seq[i], type="response")
+        errors[i]<- sum(temp)/nrow(test)
+
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres=T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
+        }
+
+        auc[i]<- auc(pred.p)
+
+        #confusion matrix
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
         conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
@@
 conf.matrix
-errors
+auc
 
 #create new data frame 'shrink' to place trees and errors vectors together, then plot:
-shrink<- data.frame("shrinkage"=shrink.seq, "error" = errors[1:4])
-ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()
+shrink<- data.frame("shrinkage"=ntree.seq, "auc" = auc[1:5])
+ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()
 
-```
+#4000 trees seems best?
 
-Try caret method of tuning
+#----------------------------------------------------------------------
 
-```{r, echo=FALSE, include=FALSE}
-# must turn outcome into two level factor
-# all9a$f.rpdol<- as.factor(all9a$b.rpdol)
-# set.seed(100)
-# sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
-# train<- all9a[sample,]
-# test<- all9a[-sample,]
-# 
-# # set seeds
-# set.seed(123)
-# seeds <- vector(mode = "list", length = 51)
-# for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
-# ## For the last model:
-# seeds[[51]] <- sample.int(1000, 1)
-# 
-# # first tune shrinkage and number of trees
-# cv.Control <- trainControl(method = "cv",
-#                            number = 10,
-#                            seeds= seeds)
-# gbmGrid <- expand.grid(shrinkage = c(.001,.005),
-#                        n.trees = c(2000), interaction.depth = 2,
-#                        n.minobsinnode = 10)
-# set.seed(2)
-# gbmFit <- train(f.rpdol~.- mlsto - return.pdol-Year, data= train,
-#                  method = "gbm", trControl = cv.Control, verbose = FALSE,
-#                  bag.fraction = 0.5, tuneGrid = gbmGrid,
-#                 metric = 'kappa')
-# 
-# plot(gbmFit)
-```
-
-Caret package is not working, tune manually:
-
-```{r, echo=FALSE}
-
-#shrinkage should be 0.001, now find number of trees
+#shrinkage should be 0.001, 4000 trees, now find interaction.depth
 errors<- double(5)
-ntree.seq<-c(4000,8000,10000,12000,14000)
+int.seq<-c(1,2,3,4,6)
 conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
@@
                    '5' = data.frame(c(0,0), c(0,0)))
+auc<- double(5)
+colours <- c("red", "yellow", "green", "blue", "purple")
+
 for(i in 1:5){
         fit<-gbm(as.formula(formula), 
-                 data=train, distribution="bernoulli", n.trees= ntree.seq[i],
-                 shrinkage=0.001,interaction.depth=3)
-        pred<- predict(fit, test, n.tree=ntree.seq[i], type="response")
-        pred<- ifelse(pred>=0.5,1,0)
-        temp<- ifelse(pred == test$b.rpdol, 0,1)
+                 data=train, distribution="bernoulli", n.trees= 4000,
+                 shrinkage=0.001,interaction.depth=int.seq[i])
+        pred<- predict(fit, test, n.tree=4000, type="response")
         errors[i]<- sum(temp)/nrow(test)
 
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres=T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
+        }
+
+        auc[i]<- auc(pred.p)
+
         #confusion matrix
-        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
         conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
@@
 conf.matrix
-errors
+auc
 
 #create new data frame 'shrink' to place trees and errors vectors together, then plot:
-shrink<- data.frame("shrinkage"=ntree.seq, "error" = errors[1:5])
-ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()
+shrink<- data.frame("shrinkage"=int.seq, "auc" = auc[1:5])
+ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()
 
-#10000 trees seems best?
+## use interaction.depth = 3
 
-#----------------------------------------------------------------------
+#------------------------------------------------
+# now tune min obs in node
 
-#shrinkage should be 0.001, 10000 trees, now find interaction.depth
-errors<- double(5)
-int.seq<-c(1,2,4,6,8)
+errors<- double(4)
+obs.seq<-c(10,20,40,50)
 conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
-                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
-                   '5' = data.frame(c(0,0), c(0,0)))
-for(i in 1:5){
+                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
+auc<- double(4)
+colours <- c("red", "yellow", "green", "blue")
+
+for(i in 1:4){
         fit<-gbm(as.formula(formula), 
-                 data=train, distribution="bernoulli", n.trees= 10000,
-                 shrinkage=0.001,interaction.depth=int.seq[i])
-        pred<- predict(fit, test, n.tree=10000, type="response")
-        pred<- ifelse(pred>=0.5,1,0)
-        temp<- ifelse(pred == test$b.rpdol, 0,1)
-        errors[i]<- sum(temp)/nrow(test)
+                 data=train, distribution="bernoulli", n.trees= 4000,
+                 shrinkage=0.001,interaction.depth=3,
+                 n.minobsinnode = obs.seq[i])
+        pred<- predict(fit, test, n.tree=4000, type="response")
+
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres = T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = T)
+        }
+
+        auc[i]<- auc(pred.p)
 
         #confusion matrix
-        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
         conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
@@
 conf.matrix
-errors
+auc
 
 #create new data frame 'shrink' to place trees and errors vectors together, then plot:
-shrink<- data.frame("shrinkage"=int.seq, "error" = errors[1:5])
-ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()
+shrink<- data.frame("shrinkage"=obs.seq, "auc" = auc[1:4])
+ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()
 
-## use interaction.depth = 4?
+#go with 10
 
-#------------------------------------------------
-# now tune min obs in node
-errors<- double(4)
-obs.seq<-c(10,20,40,50)
+```
+
+So paramters are ....
+
+    * shrinkage = 0.001
+    * exclude vars = JD.Second, Business, code.client, Year, code.contact
+    * n.trees = 4000
+    * interaction.depth = 3
+    * minobsinnode = 10
+
+to obtain AUC = 0.785
+
+But can we improve the AUC by bringing back some of the missing variables in a smart way??
+
+For example, bring back code.client, but only if clients have more than say 10 jobs?
+
+```{r, echo=FALSE}
+
+#new variable for clients with more than 10 jobs
+#list of clients with more than 10 jobs
+all9b<- all9a
+client.count<- ddply(all9b, .(code.client), nrow)
+client.count<- arrange(client.count, -V1)
+
+client.count2<- client.count[client.count$V1>20 & !is.na(client.count$code.client),]
+
+all9b$code.client<- as.character(all9b$code.client)
+all9b$reduc.client<- ifelse(all9b$code.client %in% client.count2$code.client,
+                            all9b$code.client,
+                            NA)
+all9b$code.client<- as.factor(all9b$code.client)
+all9b$reduc.client<- as.factor(all9b$reduc.client)
+
+#see if ROC improves
+
+formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business"
+
+client.lim <- c(20,30,40,60,70)
 conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
-                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
-for(i in 1:4){
+                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
+                   '5' = data.frame(c(0,0), c(0,0)))
+auc<- double(5)
+colours <- c("red", "yellow", "green", "blue", "purple")
+
+for(i in 1:5){
+
+        client.count2<- client.count[client.count$V1> client.lim[i] &
+                                             !is.na(client.count$code.client),]
+        all9b$code.client<- as.character(all9b$code.client)
+        all9b$reduc.client<- ifelse(all9b$code.client %in% client.count2$code.client,
+                            all9b$code.client,
+                            NA)
+        all9b$code.client<- as.factor(all9b$code.client)
+        all9b$reduc.client<- as.factor(all9b$reduc.client)
+
+        #see if ROC improves
+
+        #redo train and test
+        set.seed = 300
+        sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
+        train<- all9a[sample,]
+        test<- all9a[-sample,]
+
+        #make boosted tree
         fit<-gbm(as.formula(formula), 
-                 data=train, distribution="bernoulli", n.trees= 10000,
-                 shrinkage=0.001,interaction.depth=4,
-                 n.minobsinnode = obs.seq[i])
-        pred<- predict(fit, test, n.tree=10000, type="response")
-        pred<- ifelse(pred>=0.5,1,0)
-        temp<- ifelse(pred == test$b.rpdol, 0,1)
-        errors[i]<- sum(temp)/nrow(test)
+                 data=train, distribution="bernoulli", n.trees= 4000,
+                 shrinkage=0.001,interaction.depth=3,
+                 n.minobsinnode = 10)
+        pred<- predict(fit, test, n.tree=4000, type="response")
+
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres = T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = T)
+        }
+
+        auc[i]<- auc(pred.p)
 
         #confusion matrix
-        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
         conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
         cat(i," ")
+
 }
@@
 conf.matrix
-errors
+auc
 
 #create new data frame 'shrink' to place trees and errors vectors together, then plot:
-shrink<- data.frame("shrinkage"=obs.seq, "error" = errors[1:4])
-ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()
+shrink<- data.frame("shrinkage"=client.lim, "auc" = auc[1:5])
+ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()
+
+#best I can do is 0.783
+```
+
+Make a function to do the same with code.contact, JD.Second, and Business????

### 2015-10-20 finalwriteup/draft.rmd
more boosting

--- 2015-08-15 finalwriteup/draft.rmd
@@
 * combination of cforest, anova, random forest
+
+# Chapter ? Method
+
+## Regression
+    * tried predicting return per dollar as a numeric outcome.
+        * proved very difficult. Find error rate compared to just guessing the mean. Not far off.
+
+##Bayesian Network
+
+* could not handle _any_ numeric variables, had to discretise all numeric vars
+    * used normalised numeric variables
+    * I did this by making a hierarchical dendrogram to visualise the clusters
+    * then I chose the number of clusters, drew the red rectangle
+    * summarised the clusters to find max and min values for the variable within each cluster. Rounded this value up
+    * manually discretised variable using rounded values
+
+* first trialled complete cases of number of users, Discipline, percent pro, Business, client code, job detail, majority position, timespan, invoiced amount, percent hours by majority position, and return per dollar.
+This gave 907 complete cases to work with.
+
+*using Genie Smile.
+   * turned return per dollar into profit/loss. This gave 182 loss cases and 725 profit cases. A good distribution for working with bayesian networks. Ie not too biased
+   * a good, simple problem to start with
+
+* establish a baseline. Using ROC, area under the curve method
+    * to test our classifiers we need to beat random chance. Ie AUC = 0.5
+    * will also establish the AUC for using one variable only! See table below:
+       *Note all using 10 fold cross validation
+
+Single Variable  | AUC from ROC chart | Num losses correct
+---------------- | ------------------ | ------------------
+Random chance    | 0.5                | ?                 
+percent pro      | 0.618              | 0
+percent maj pos  | 0.616              | 0
+number of users  | 0.661              | 0
+timespan         | 0.600              | 0
+amount invoiced  | 0.535              | 0
+Discipline       | 0.514              | 0
+Majority pos     | 0.605              | 3
+Job detail       | 0.605              | 1
+client code      | 0.567              | 25
+client Business  | 0.524              | 5
+
+* try using the following variables:
+
+```{r, echo=FALSE, include=FALSE}
+
+read.csv(all8c,'C:/Users/n9232371/Documents/Consultbusiness/data/all8c.csv' )
+
+vars<- names(all8c)
+vars<- vars[-c("Business", "JD_Second", "code_client")]
+
+```
+
+* number of users
+* Discipline
+* percent hours by a professional
+* position that completed the majority of hours
+* timespan
+* amount invoiced
+* percent of hours completed by majority position
+
+The results are as follows:
+
+AUC = 0.705
+Number of losses predicted correctly: 124
+
+This is our 'base' model. Now try adding one complex variable at a time:
+
+Added Variable   | AUC from ROC chart | Num losses correct (tp) | Num losses wrong (fp)
+---------------- | ------------------ | ----------------------- | --------------------
+none/base        | 0.710              | 122                     | 272
+client code      | 0.624              | 171                     | 528
+Busines          | 0.667              | 144                     | 381
+job detail       | 0.653              | 150                     | 438
+
+## Boosted Trees
+
+### Advantages
+
+    * Wonderful in that NA's don't interfere with prediciton process
+        * explain..
+        * can comfortably include the entire dataset without having to worry about imputing
+
+tuned which variables to include:
+    * exclude JD.Second, Business, code.client, code.contact
+    * tuned parameters:
+        * shrinkage = 0.001
+        * n.trees = 4000
+        * interaction.depth = 3
+        * min.nobs = 10
+    * manage to get AUC up to 0.785!!
+
+can we improve this by bringing huge variables back in a smarter way?

### 2015-10-20 BNprep.R

@@ -212,16 +212,25 @@ all8b[is.na(all8b)]<- ""
 all8b[]<- lapply(all8b, as.factor)
 summary(all8b)
 
+#remove spaces, '.' and '/' from column titles
+names(all8b)<- gsub("\\.|\\/|\\s", "_", names(all8b))
+
+#remove spaces, '.' and '/' from all variable levels
+all8b[]<- sapply(all8b, function(x) {gsub("\\.|\\/|\\s", "_", x)} )
+all8b[]<- lapply(all8b, factor)
+
 write.csv(all8b,'C:/Users/n9232371/Documents/Consultbusiness/data/all8b.csv' )
 
 #look at complete cases for no.users, discipline, pc.pro, business, code.client,
 #JD.Second, majority.pos, timespan.cbrt, inv.mlsto.log, pc.majpos.log
 
 all8b<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8b.csv',
-                 na.strings = "")
-all8c<- all8b %>% select(b.no.users,Discipline,b.pc.pro, Business, code.client,
-                         JD.Second, majority.pos, b.timespan.cbrt, 
-                         b.inv.log, b.pcmajpos, b.rpdol)
+                 na.strings = "")[,-1]
+all8c<- all8b %>% select(b_no_users, Discipline, b_pc_pro, Business, code_client,
+                         JD_Second, majority_pos, b_timespan_cbrt,
+                         b_inv_log, b_pcmajpos, b_rpdol)
+
+
 all8c<- all8c[complete.cases(all8c),]
 dim(all8c)
 

### 2015-10-21 boost_tree.rmd
more boosting

@@
 all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 1, 0)
+all8a$timespan.cbrt <- ifelse(all8a$timespan.cbrt ==0, 1, all8a$timespan.cbrt)
 
@@
 write.csv(all9a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')
+all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]
 
@@
 ```{r, echo=FALSE}
-set.seed(300)
+set.seed(200)
 sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
@@
 #shrinkage should be 0.001, now find number of trees
-errors<- double(5)
+
 ntree.seq<-c(2000,3000,4000,6000,8000)
@@
 
-#4000 trees seems best?
+#6000 trees seems best?
 
@@
 #shrinkage should be 0.001, 4000 trees, now find interaction.depth
-errors<- double(5)
+
 int.seq<-c(1,2,3,4,6)
@@
         fit<-gbm(as.formula(formula), 
-                 data=train, distribution="bernoulli", n.trees= 4000,
+                 data=train, distribution="bernoulli", n.trees= 6000,
                  shrinkage=0.001,interaction.depth=int.seq[i])
-        pred<- predict(fit, test, n.tree=4000, type="response")
-        errors[i]<- sum(temp)/nrow(test)
+        pred<- predict(fit, test, n.tree=6000, type="response")
 
@@
 
-errors<- double(4)
 obs.seq<-c(10,20,40,50)
@@
         fit<-gbm(as.formula(formula), 
-                 data=train, distribution="bernoulli", n.trees= 4000,
+                 data=train, distribution="bernoulli", n.trees= 6000,
                  shrinkage=0.001,interaction.depth=3,
                  n.minobsinnode = obs.seq[i])
-        pred<- predict(fit, test, n.tree=4000, type="response")
+        pred<- predict(fit, test, n.tree=6000, type="response")
 
@@
 
-#go with 10
+#go with 10????!
 
@@
 to obtain AUC = 0.785
+
+Can we change timespan to discretised without affecting AUC??
+
+```{r, echo=FALSE}
+
+formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business - timespan.cbrt"
+
+conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
+                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
+auc<- double(4)
+colours <- c("red", "yellow", "green", "blue")
+timeclust.seq<- c(4:7)
+
+x<- dist(all9a$timespan.cbrt, method = 'euclidean')
+h.fit<- hclust(x, method = 'ward.D2')
+
+for(i in 1:4){
+        #create timespan cluster number column
+        all9a$b.timespan.cbrt<- as.factor(cutree(h.fit, k= timeclust.seq[i]))
+
+        set.seed(300)
+        sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
+        train<- all9a[sample,]
+        test<- all9a[-sample,]
+
+        fit<-gbm(as.formula(formula), 
+                 data=train, distribution="bernoulli", n.trees= 4000,
+                 shrinkage=0.001,interaction.depth=3,
+                 n.minobsinnode = 10)
+        pred<- predict(fit, test, n.tree=4000, type="response")
+
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres = T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = T)
+        }
+
+        auc[i]<- auc(pred.p)
+
+        #confusion matrix
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
+        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
+        cat(i," ")
+}
+
+conf.matrix
+auc
+
+#create new data frame 'shrink' to place trees and errors vectors together, then plot:
+shrink<- data.frame("shrinkage"= timeclust.seq, "auc" = auc[1:4])
+ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()
+
+#takes it down to 0.781 with 5 clusters, not too bad for the ability to use timespan clusters.
+#takes it down to 0.783 with 6 clusters
+
+```
+
+6 timespan clusters seems to give quite good performance :) 0.783. Lets see what those clusters are:
+
+```{r, echo=FALSE}
+
+#timespan.cbrt
+first.vec<- all9a$timespan.cbrt
+x<- dist(first.vec, method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+k.choose=6
+rect.hclust(fit, k = k.choose)
+
+#assign a cluster number to all rows in data frame
+all9a$b.timespan.cbrt<- as.factor(cutree(fit, k= k.choose))
+
+for(i in 1:k.choose){
+        print(
+        summary(all9a %>% select(b.timespan.cbrt, timespan.cbrt) %>% 
+                        filter(b.timespan.cbrt ==i))
+        )
+}
+
+#clusters roughly break down to 1 day, 3 wks, 2.5m, 9m, 1.5yr, 3 yr, over 3 yr
+# in cbrt days this is equivalent to breaks of 0,1,2,7589, 4.2358, 6.5, 8.183, 10.307, 14.1
+#make new clusters:
+all9a$b.timespan.cbrt<- cut(all9a$timespan.cbrt, 
+                            breaks= c(0,2.7589, 4.2358, 6.5, 8.183, 10.307, 14.1), 
+                            include.lowest=TRUE, 
+                            labels= c("1d-3wk","3wk-2.5m", "2.5m-9m","9m-1.5y","1.5y-3y",">3y"))
+
+set.seed(300)
+sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
+train<- all9a[sample,]
+test<- all9a[-sample,]
+
+fit<-gbm(as.formula(formula), 
+        data=train, distribution="bernoulli", n.trees= 4000,
+        shrinkage=0.001,interaction.depth=3,
+        n.minobsinnode = 10)
+pred<- predict(fit, test, n.tree=4000, type="response")
+
+#pROC
+pred.p<- roc(test$b.rpdol, pred)
+plot.roc(pred.p, col= "pink", print.thres = T)
+auc(pred.p)
+
+#reduces AUC to 0.780 :(
+#go through tuning process again?
+
+all9a<- all9a %>% select(-timespan.cbrt)
+```
 
@@
         #redo train and test
-        set.seed = 300
+        set.seed = 200
         sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)

### 2015-10-21 BNprep.R

@@ -39,6 +39,8 @@ all8a$b.pc.pro<- cut(all8a$pc.pro, breaks=c(0,15,40,75,95,100), include.lowest=T
 
 ###
 #timespan.cbrt
+# all timespan == 0 should be timespan ==1. Checked all7a and they all have at least 0.5 hour booked, max 18 hours.
+all8a$timespan.cbrt <- ifelse(all8a$timespan.cbrt ==0, 1, all8a$timespan.cbrt)
 first.vec<- all8a$timespan.cbrt
 x<- dist(first.vec, method = 'euclidean')
 fit<- hclust(x, method = 'ward.D2')

### 2015-10-21 financial.R

@@ -1,0 +1,85 @@
+---
+title: "Financial_Analysis"
+author: "Amy Cook"
+date: "October 21, 2015"
+output: word_document
+---
+
+#Assess financial implications for a predictive model
+
+We typically get a confusion matrix built from probabilities
+
+
+```{r_loadpackages, echo=FALSE, include=FALSE}
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')
+
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+```
+
+Load data
+
+```{r, echo=FALSE}
+all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]
+all9a$b.timespan.cbrt<- cut(all9a$timespan.cbrt, 
+                            breaks= c(0,2.7589, 4.2358, 6.5, 8.183, 10.307, 14.1), 
+                            include.lowest=TRUE, 
+                            labels= c("1d-3wk","3wk-2.5m", "2.5m-9m","9m-1.5y","1.5y-3y",">3y"))
+
+set.seed(300)
+sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
+train<- all9a[sample,]
+test<- all9a[-sample,]
+```
+
+Run boosted tree
+
+```{r, echo=FALSE}
+
+formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business - timespan.cbrt"
+
+fit<-gbm(as.formula(formula), 
+         data=train, distribution="bernoulli", n.trees= 6000,
+         shrinkage=0.001,interaction.depth=3,
+         n.minobsinnode = 10)
+pred<- predict(fit, test, n.tree=4000, type="response")
+ans<- test$b.rpdol
+
+
+```
+
+Create vector of profit points
+
[…]

### 2015-10-23 boost_tree.rmd
finishing boosted

@@
 library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')
+library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
 
@@
 all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 1, 0)
+all8a$pc.pro<- ifelse(all8a$pc.pro == 0 & is.na(all8a$pc.majpos.log), NA, all8a$pc.pro)
 all8a$timespan.cbrt <- ifelse(all8a$timespan.cbrt ==0, 1, all8a$timespan.cbrt)
@@
 
+Attempt at caret package with gbm
+
+```{r, echo=FALSE}
+
+# must turn outcome into two level factor
+all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]
+all9a$f.rpdol<- as.factor(all9a$b.rpdol)
+set.seed(100)
+sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
+train<- all9a[sample,]
+test<- all9a[-sample,]
+
+# set seeds
+set.seed(123)
+seeds <- vector(mode = "list", length = 51)
+for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
+## For the last model:
+seeds[[51]] <- sample.int(1000, 1)
+
+# first tune shrinkage and number of trees
+train2<- train
+train2$b.rpdol<- as.factor(train2$b.rpdol)
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="0"] <- "profit"
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="1"] <- "loss"
+
+formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log +
+client.totinv.log + pc.majpos.log + Billing.Type + majority.pos"
+
+cv.Control <- trainControl(method = "cv",
+                           number = 10,
+                           seeds= seeds
+                           , summaryFunction=twoClassSummary,
+                           classProbs=TRUE
+                           )
+gbmGrid <- expand.grid(shrinkage = c(0.001),
+                       n.trees = c(10000), interaction.depth = c(1,2,3,4,5),
+                       n.minobsinnode = c(10,20,30))
+set.seed(2)
+gbmFit <- train(as.formula(formula), data= train2,
+                 method = "gbm", trControl = cv.Control, verbose = FALSE,
+                 bag.fraction = 0.5, tuneGrid = gbmGrid,
+                metric = 'ROC',
+                na.action = na.pass)
+
+plot(gbmFit)
+plot(gbmFit, plotType = 'level')
+
+#conclusion:
+# min num in terminal node = 10
+# max tree depth = 4
+# n.trees = 10000
+# shrinkage = 0.001
+
+```
+
+After caret tuning, find AUC for test data set for set.seed = 100
+
+```{r, echo = FALSE}
+set.seed(300)
+sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
+train<- all9a[sample,]
+test<- all9a[-sample,]
+
+formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + pc.majpos.log + Billing.Type + majority.pos"
+
+fit<-gbm(as.formula(formula), 
+                 data=train, distribution="bernoulli", n.trees= 10000,
+                 shrinkage=0.001,interaction.depth=4,
+                 n.minobsinnode = 10)
+pred<- predict(fit, test, n.tree=10000, type="response")
+
+#pROC
+pred.p<- roc(test$b.rpdol, pred)
+plot.roc(pred.p, col= colours[i], print.thres = T)
+
+#auc = 0.7788
+
+```
+
 But can we improve the AUC by bringing back some of the missing variables in a smart way??
@@
 
-formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business"
+formula<- "b.rpdol~.- mlsto - return.pdol- Year- code.client - code.contact- JD.Second - Business-timespan.cbrt"
 
-client.lim <- c(20,30,40,60,70)
+client.lim <- c(20,30,40,60,70, 90)
 conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                    '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
-                   '5' = data.frame(c(0,0), c(0,0)))
-auc<- double(5)
-colours <- c("red", "yellow", "green", "blue", "purple")
+                   '5' = data.frame(c(0,0), c(0,0)), '6' = data.frame(c(0,0), c(0,0)))
+auc<- rep(list(rep(0,6)),6)
+colours <- c("red", "yellow", "green", "blue", "purple", "pink")
 
-for(i in 1:5){
+#set up train and test
+        set.seed = 100
+        sample<- sample(1:nrow(all9a), .75*nrow(all9a), replace=F)
+        train<- all9a[sample,]
+        test<- all9a[-sample,]
+        #set up 10 fold cv using caret function
+        folds <- createFolds(train$b.rpdol, k=10, list=T, returnTrain=F)
+        #now have folds[[1]] through folds[[5]] list of rows to exclude
+
+for(i in 1:6){
 
@@
 
-        #redo train and test
-        set.seed = 200
+        for (j in 1:10){
+                #make boosted tree
+                fit<-gbm(as.formula(formula), 
+                         data=train[-folds[[j]],], distribution="bernoulli", n.trees= 10000,
+                         shrinkage=0.001,interaction.depth=4,
+                         n.minobsinnode = 10)
+
+                pred<- predict(fit, train[folds[[j]], ], n.tree=10000, type="response")
+
+                        #pROC
+                        pred.p<- roc(train[folds[[j]],'b.rpdol'], pred)
+
+                        if(i==1 & j==1){
+                                plot.roc(pred.p, col= colours[i], print.thres = F)
+
+                        } else{
+                                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = F)
+                        }
+
+                        auc[[i]][j]<- auc(pred.p)
+
+                        #confusion matrix
+#                         conf.mat<- coords(pred.p, "best")
+#                         pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+#                         temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
+#                         conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
+#                         cat(i," ")  
+                        cat(j, " ")
+        }
+
+        print(i)        
+}
+
+conf.matrix
+auc
+
+#create new data frame 'shrink' to place trees and errors vectors together, then plot:
+shrink<- data.frame("shrinkage"=client.lim, "auc" = auc[[1:5]][6])
+means<- sapply(auc, function(x) tail(x,1))
+ggplot(shrink, aes(x=shrinkage, y=means)) + geom_line() + geom_point()
+a<- ggplot(melt(auc), aes(x= factor(L1), y= value)) + geom_boxplot()
+
+#save AUC for each 10 fold cross validation iteration, where client count cut off varied from 
+client.cutoff<- melt(auc)
+client.cutoff$L1<- as.factor(client.cutoff$L1)
+levels(client.cutoff$L1)<- c("20","30","40","60","70", "90")
+colnames(client.cutoff)[names(client.cutoff) %in% 'L1']<-'min.client.count'
+colnames(client.cutoff)[names(client.cutoff) %in% 'value']<-'AUC'
+write.csv(client.cutoff, 'C:/Users/n9232371/Documents/github/consult/finalwriteup/images/boosted tree/client_cutoff.csv')
+
+a<- ggplot(client.cutoff, aes(x= min.client.count, y= AUC)) + geom_boxplot()
+ggsave('C:/Users/n9232371/Documents/github/consult/finalwriteup/images/boosted tree/clientcode.jpg', a)
+
+#client id useless for boosted trees!!
+
+```
+
+What about JD.Second?
+
+```{r,echo=FALSE}
+
+#compact JD.Second into less categories
+all9a$JD.Second<- as.character(all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("heritage|expert", all9a$JD.Second), 
+                          "heritage_expert", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("wharf|bridges", all9a$JD.Second), 
+                          "wharf_bridge", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("membrane|art|awning|facade", all9a$JD.Second), 
+                          "art_facade_awn_memb", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("hosp|emerg|health|carpark", all9a$JD.Second), 
+                          "hosp_health_carpark", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("sign|product", all9a$JD.Second), 
+                          "sign_product", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("hotel|office|dining", all9a$JD.Second), 
+                          "hotel_office_dining", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("report|investigation", all9a$JD.Second), 
+                          "report", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("waste water|management", all9a$JD.Second), 
+                          "waste_wat_manage", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("univ|commun|school|exten", all9a$JD.Second), 
+                          "edu_exten_community", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("multi|house", all9a$JD.Second), 
+                          "residential", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("flood|harvest|recycled", all9a$JD.Second), 
+                          "flood_h2o_harvest", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("bldg|subdiv|sewer", all9a$JD.Second), 
+                          "civBldg_subdiv_sewer", all9a$JD.Second)
+all9a$JD.Second <- ifelse(grepl("aged|erosion", all9a$JD.Second), 
+                          NA, all9a$JD.Second)
+
+all9a$JD.Second<- as.factor(all9a$JD.Second)
+
+#now see if improvement with JD.Second
+
+var.seq<-c("no.users + JD.Second", 
+           "no.users")
+
+conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)))
+auc<- double(2)
+colours <- c("red", "yellow")
+
+#make fresh random sample of data
+set.seed(1000)
 sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
@@
 
-        #make boosted tree
+for(i in 1:2){
+
+        formula<- paste("b.rpdol~ Discipline + pc.pro + inv.mlsto.log + pc.majpos.log + majority.pos + b.timespan.cbrt + client.totinv.log + Billing.Type", var.seq[i], sep= " + ")
+
         fit<-gbm(as.formula(formula), 
-                 data=train, distribution="bernoulli", n.trees= 4000,
-                 shrinkage=0.001,interaction.depth=3,
+                 data=train, distribution="bernoulli", n.trees= 10000,
+                 shrinkage=0.001,interaction.depth=4,
                  n.minobsinnode = 10)
-        pred<- predict(fit, test, n.tree=4000, type="response")
+        pred<- predict(fit, test, n.tree=10000, type="response")
 
@@
         cat(i," ")
-
 }
@@
 #create new data frame 'shrink' to place trees and errors vectors together, then plot:
-shrink<- data.frame("shrinkage"=client.lim, "auc" = auc[1:5])
+shrink<- data.frame("shrinkage"=1:2, "auc" = auc[1:2])
 ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()
 
-#best I can do is 0.783
+# generally slightly better without JD.Second :(
+
 ```
 
-Make a function to do the same with code.contact, JD.Second, and Business????
+What about Business?

### 2015-10-23 Business_compact.R

@@ -1,0 +1,62 @@
+# re-jig Business
+
+str(all9a)
+
+B.invest<- table(all9a$b.rpdol, all9a$Business) %>% t %>% as.data.frame.matrix()
+colnames(B.invest) <- c("profit", "loss")
+B.invest<- transform(B.invest, ratio = round(loss/profit, 2))
+B.invest$JD <- rownames(B.invest) 
+B.invest<- arrange(B.invest, ratio) %>% select(JD, profit, loss, ratio)
+
+#group:
+
+# university, school
+# lawyer/body corp
+# signs/manufacturer
+# environment
+# roads/rail/gov
+# artist/landscape arch
+# hospital/institution/health
+# utilities
+# engineer/water
+# business/building services/internal
+# builder
+# architect
+# town planner/developer
+# resources
+# person
+
+# unknown should be NA
+
+
+#code to compact:
+
+all9a$test<- as.character(all9a$Business)
+all9a$test <- ifelse(grepl("heritage|expert", all9a$Business), 
+                     "heritage_expert", all9a$test)
+all9a$test <- ifelse(grepl("wharf|bridges", all9a$Business), 
+                     "wharf_bridge", all9a$test)
+all9a$test <- ifelse(grepl("membrane|art|awning|facade", all9a$Business), 
+                     "art_facade_awn_memb", all9a$test)
+all9a$test <- ifelse(grepl("hosp|emerg|health|carpark", all9a$Business), 
+                     "hosp_health_carpark", all9a$test)
+all9a$test <- ifelse(grepl("sign|product", all9a$Business), 
+                     "sign_product", all9a$test)
+all9a$test <- ifelse(grepl("hotel|office|dining", all9a$Business), 
+                     "hotel_office_dining", all9a$test)
+all9a$test <- ifelse(grepl("report|investigation", all9a$Business), 
+                     "report", all9a$test)
+all9a$test <- ifelse(grepl("waste water|management", all9a$Business), 
+                     "waste_wat_manage", all9a$test)
+all9a$test <- ifelse(grepl("univ|commun|school|exten", all9a$Business), 
+                     "edu_exten_community", all9a$test)
+all9a$test <- ifelse(grepl("multi|house", all9a$Business), 
+                     "residential", all9a$test)
+all9a$test <- ifelse(grepl("flood|harvest|recycled", all9a$Business), 
+                     "flood_h2o_harvest", all9a$test)
[…]

### 2015-10-23 JD.Second_compact.R

@@ -1,0 +1,70 @@
+
+# re-jog JD.Second and Business
+
+str(all9a)
+
+JD.invest<- table(all9a$b.rpdol, all9a$JD.Second) %>% t %>% as.data.frame.matrix()
+colnames(JD.invest) <- c("profit", "loss")
+JD.invest<- transform(JD.invest, ratio = round(loss/profit, 2))
+JD.invest$JD <- rownames(JD.invest) 
+JD.invest<- arrange(JD.invest, ratio) %>% select(JD, profit, loss, ratio)
+
+#group:
+
+# heritage/ expert opinion
+# wharf/port, bridges
+# membrane shade structure/art/awning/facade
+# hospital/emergency services/health/carpark
+# refurb/renovation
+# sign, product development
+# hotel/resort, offices, dining/retail
+# report/investigation
+# waste water/water management
+# university/school/community/extension
+# multi-dwelling/house
+# industrial
+# flood studies, water harvesting, recycled water
+# parks and open spaces
+# civil works to bldg, Subdivision, sewerage
+
+
+
+# delete - aged care, erosion and sediment
+
+#code to compact:
+
+all9a$test<- as.character(all9a$JD.Second)
+all9a$test <- ifelse(grepl("heritage|expert", all9a$JD.Second), 
+                          "heritage_expert", all9a$test)
+all9a$test <- ifelse(grepl("wharf|bridges", all9a$JD.Second), 
+                          "wharf_bridge", all9a$test)
+all9a$test <- ifelse(grepl("membrane|art|awning|facade", all9a$JD.Second), 
+                          "art_facade_awn_memb", all9a$test)
+all9a$test <- ifelse(grepl("hosp|emerg|health|carpark", all9a$JD.Second), 
+                          "hosp_health_carpark", all9a$test)
+all9a$test <- ifelse(grepl("sign|product", all9a$JD.Second), 
+                          "sign_product", all9a$test)
+all9a$test <- ifelse(grepl("hotel|office|dining", all9a$JD.Second), 
+                          "hotel_office_dining", all9a$test)
+all9a$test <- ifelse(grepl("report|investigation", all9a$JD.Second), 
+                          "report", all9a$test)
+all9a$test <- ifelse(grepl("waste water|management", all9a$JD.Second), 
+                          "waste_wat_manage", all9a$test)
+all9a$test <- ifelse(grepl("univ|commun|school|exten", all9a$JD.Second), 
[…]

### 2015-10-26 boost_tree.rmd
moving on from boosted

@@
 What about Business?
+
+```{r,echo=FALSE}
+
+#compact JD.Second into less categories
+
+all9a$Business<- as.character(all9a$Business)
+all9a$Business <- ifelse(grepl("lawyer|body", all9a$Business), 
+                     "lawyer_bodyC", all9a$Business)
+all9a$Business <- ifelse(grepl("univ|school", all9a$Business), 
+                     "uni_school", all9a$Business)
+all9a$Business <- ifelse(grepl("sign|manuf", all9a$Business), 
+                     "sign_manufac", all9a$Business)
+all9a$Business <- ifelse(grepl("environ|water", all9a$Business), 
+                     "enviro_water", all9a$Business)
+all9a$Business <- ifelse(grepl("road|gov|util", all9a$Business), 
+                     "road_rail_gov_util", all9a$Business)
+all9a$Business <- ifelse(grepl("artist|landscape", all9a$Business), 
+                     "artist_landarch", all9a$Business)
+all9a$Business <- ifelse(grepl("hospit|institut|health", all9a$Business), 
+                     "hosp_institut_health", all9a$Business)
+all9a$Business <- ifelse(grepl("business|building|internal", all9a$Business), 
+                     "biz_bldgserv_internal", all9a$Business)
+
+all9a$Business<- as.factor(all9a$Business)
+
+write.csv(all9a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv') # c for compacted variables
+all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
+
+#now see if improvement with JD.Second
+
+var.seq<-c("no.users + Business", 
+           "no.users")
+
+conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)))
+auc<- double(2)
+colours <- c("red", "yellow")
+
+#make fresh random sample of data
+set.seed(200)
+sample<- sample(1:nrow(all9c), 2/3*nrow(all9c), replace=F)
+train<- all9c[sample,]
+test<- all9c[-sample,]
+
+for(i in 1:1){
+
+        formula<- paste("b.rpdol~ Discipline + pc.pro + inv.mlsto.log + pc.majpos.log + majority.pos + b.timespan.cbrt + client.totinv.log + Billing.Type", var.seq[i], sep= " + ")
+
+        fit<-gbm(as.formula(formula), 
+                 data=train, distribution="bernoulli", n.trees= 10000,
+                 shrinkage=0.001,interaction.depth=4,
+                 n.minobsinnode = 20)
+        pred<- predict(fit, test, n.tree=10000, type="response")
+
+        #pROC
+        pred.p<- roc(test$b.rpdol, pred)
+
+        if(i==1){
+                plot.roc(pred.p, col= colours[i], print.thres = T)
+
+        } else{
+                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = T)
+        }
+
+        auc[i]<- auc(pred.p)
+
+        #confusion matrix
+        conf.mat<- coords(pred.p, "best")
+        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
+        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
+        cat(i," ")
+}
+
+conf.matrix
+auc
+
+#create new data frame 'shrink' to place trees and errors vectors together, then plot:
+shrink<- data.frame("shrinkage"=1:2, "auc" = auc[1:2])
+ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()
+
+# Business helps!!!!!!!!
+
+```
+
+Tune again with caret package, INCLUDING Business category.
+
+```{r, echo=FALSE}
+
+# set seeds
+set.seed(123)
+seeds <- vector(mode = "list", length = 51)
+for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
+## For the last model:
+seeds[[51]] <- sample.int(1000, 1)
+
+#create test train
+sample<- sample(1:nrow(all9c), 2/3*nrow(all9c), replace=F)
+train<- all9c[sample,]
+test<- all9c[-sample,]
+
+# first tune shrinkage and number of trees
+train2<- train
+train2$b.rpdol<- as.factor(train2$b.rpdol)
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="0"] <- "profit"
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="1"] <- "loss"
+
+formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log +
+client.totinv.log + pc.majpos.log + Billing.Type + majority.pos + Business"
+
+cv.Control <- trainControl(method = "cv",
+                           number = 5,
+                           seeds= seeds
+                           , summaryFunction=twoClassSummary,
+                           classProbs=TRUE
+                           )
+gbmGrid <- expand.grid(shrinkage = c(0.001),
+                       n.trees = c(10000), interaction.depth = c(1,2,3,4,5),
+                       n.minobsinnode = c(10,20,30,40))
+set.seed(2)
+gbmFit <- train(as.formula(formula), data= train2,
+                 method = "gbm", trControl = cv.Control, verbose = FALSE,
+                 bag.fraction = 0.5, tuneGrid = gbmGrid,
+                metric = 'ROC',
+                na.action = na.pass)
+
+plot(gbmFit)
+plot(gbmFit, plotType = 'level')
+
+# therefore:
+# shrinkage = 0.001
+# no trees = 10000
+# min terminal node size = 20
+# max tree depth = 4
+
+# AUC approx 0.770 in 5 fold CV
+
+```
+
+Does the boosted tree perform better with a subset of the data based on later years?
+
+```{r, echo=FALSE}
+
+formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log +
+client.totinv.log + pc.majpos.log + majority.pos + Business"
+
+Year.lim <- c(2000, 2004, 2006, 2008, 2009, 2010)
+conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
+                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
+                   '5' = data.frame(c(0,0), c(0,0)), '6' = data.frame(c(0,0), c(0,0)))
+auc<- rep(list(rep(0,8)),6)
+colours <- c("red", "yellow", "green", "blue", "purple", "pink")
+
+for(i in 1:6){
+
+        #set up train and test
+        set.seed = 100
+        subset = all9c %>% filter(Year> Year.lim[i])
+        sample<- sample(1:nrow(subset), .75*nrow(subset), replace=F)
+        train<- all9c[sample,]
+        test<- all9c[-sample,]
+        #set up 10 fold cv using caret function
+        folds <- createFolds(train$b.rpdol, k=8, list=T, returnTrain=F)
+        #now have folds[[1]] through folds[[5]] list of rows to exclude
+
+        #see if ROC improves
+
+        for (j in 1:8){
+                #make boosted tree
+                fit<-gbm(as.formula(formula), 
+                         data=train[-folds[[j]],], distribution="bernoulli", n.trees= 10000,
+                         shrinkage=0.001,interaction.depth=4,
+                         n.minobsinnode = 20)
+
+                pred<- predict(fit, train[folds[[j]], ], n.tree=10000, type="response")
+
+                        #pROC
+                        pred.p<- roc(train[folds[[j]],'b.rpdol'], pred)
+
+                        if(i==1 & j==1){
+                                plot.roc(pred.p, col= colours[i], print.thres = F)
+
+                        } else{
+                                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = F)
+                        }
+
+                        auc[[i]][j]<- auc(pred.p)
+
+                        #confusion matrix
+#                         conf.mat<- coords(pred.p, "best")
+#                         pred1<- ifelse(pred>= conf.mat[1], 1, 0)
+#                         temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
+#                         conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
+#                         cat(i," ")  
+                        cat(j, " ")
+        }
+
+        print(i)        
+}
+
+conf.matrix
+auc
+saveRDS(auc, 'C:/Users/n9232371/Documents/github/consult/finalwriteup/images/boosted tree/Year_vary_cv.rds')
+
+a<- ggplot(melt(auc), aes(x= factor(L1), y= value)) + geom_boxplot()
+
+# All years are required! Gives the best result
+```

### 2015-10-26 Business_compact.R

@@ -13,50 +13,44 @@ B.invest<- arrange(B.invest, ratio) %>% select(JD, profit, loss, ratio)
 # university, school
 # lawyer/body corp
 # signs/manufacturer
-# environment
-# roads/rail/gov
+# environment/water
+# roads/rail/gov/util
 # artist/landscape arch
 # hospital/institution/health
 # utilities
-# engineer/water
+# engineer
+# water
 # business/building services/internal
 # builder
 # architect
-# town planner/developer
+# townplanner
+# developer/realestate
 # resources
 # person
 
-# unknown should be NA
 
 
 #code to compact:
 
 all9a$test<- as.character(all9a$Business)
-all9a$test <- ifelse(grepl("heritage|expert", all9a$Business), 
-                     "heritage_expert", all9a$test)
-all9a$test <- ifelse(grepl("wharf|bridges", all9a$Business), 
-                     "wharf_bridge", all9a$test)
-all9a$test <- ifelse(grepl("membrane|art|awning|facade", all9a$Business), 
-                     "art_facade_awn_memb", all9a$test)
-all9a$test <- ifelse(grepl("hosp|emerg|health|carpark", all9a$Business), 
-                     "hosp_health_carpark", all9a$test)
-all9a$test <- ifelse(grepl("sign|product", all9a$Business), 
-                     "sign_product", all9a$test)
-all9a$test <- ifelse(grepl("hotel|office|dining", all9a$Business), 
-                     "hotel_office_dining", all9a$test)
-all9a$test <- ifelse(grepl("report|investigation", all9a$Business), 
-                     "report", all9a$test)
-all9a$test <- ifelse(grepl("waste water|management", all9a$Business), 
-                     "waste_wat_manage", all9a$test)
-all9a$test <- ifelse(grepl("univ|commun|school|exten", all9a$Business), 
-                     "edu_exten_community", all9a$test)
-all9a$test <- ifelse(grepl("multi|house", all9a$Business), 
-                     "residential", all9a$test)
-all9a$test <- ifelse(grepl("flood|harvest|recycled", all9a$Business), 
-                     "flood_h2o_harvest", all9a$test)
-all9a$test <- ifelse(grepl("bldg|subdiv|sewer", all9a$Business), 
-                     "civBldg_subdiv_sewer", all9a$test)
-all9a$test <- ifelse(grepl("aged|erosion", all9a$Business), 
-                     NA, all9a$test)
[…]

### 2015-10-30 boost_tree.rmd
finish forest and log

@@
                             labels= c("1d-3wk","3wk-2.5m", "2.5m-9m","9m-1.5y","1.5y-3y",">3y"))
+write.csv(all9a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')
 
@@
 
-for(i in 1:1){
+for(i in 1:2){
 
@@
                          data=train[-folds[[j]],], distribution="bernoulli", n.trees= 10000,
-                         shrinkage=0.001,interaction.depth=4,
                          n.minobsinnode = 20)

### 2015-10-30 financial.rmd
finish forest and log

---
title: "Financial_Analysis"
author: "Amy Cook"
date: "October 21, 2015"
output: word_document
---

#Assess financial implications for a predictive model

We typically get a confusion matrix built from probabilities

```{r_loadpackages, echo=FALSE, include=FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')

# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
```

Load data

```{r, echo=FALSE}
test.pred <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/test_review.csv')
all7d <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')
test.pred<- merge(test.pred, all7d %>% select(mlsto, balance.mlsto), by = "mlsto")

```

Create vector of profit points

```{r, echo=FALSE}

threshold<- seq(0,1, by=.05)
fin.prof<- rep(0, length(threshold))

for(i in 1:length(threshold)){
        prof<- data.frame("balance" = test.pred$balance.mlsto, 
                  "pred" = test.pred$pred,
                  "ans" = test.pred$b.rpdol)
        #pred = probabilities
        prof$conf <- rep(0, nrow(prof))
        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 1, 'TP', prof$conf)
        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
        prof$profit<- rep(0, nrow(prof))
        prof$profit<- ifelse(prof$conf %in% "TP", -prof$balance, prof$profit)
        prof$profit<- ifelse(prof$conf %in% "FP", -prof$balance, prof$profit)

        fin.prof[i] = sum(prof$profit)

}

prof.plot<- data.frame("threshold" = threshold, "final.profit" = fin.prof)

ggplot(prof.plot, aes(x=threshold, y=final.profit)) + geom_point() + geom_line()

# if say no to all jobs with higher than prob of 0.5, save $500,000

head(prof)

####
# SUMMARY STATS
####

# would dismiss 56 jobs
prof %>% filter(pred>=0.5) %>% nrow

# would dismiss 19 profitable jobs
prof %>% filter(pred >= 0.5 & balance > 0) %>% nrow
#out of 607
prof %>% filter(balance > 0) %>% nrow

# would dismiss 37 losing jobs
prof %>% filter(pred >= 0.5 & balance < 0) %>% nrow
#out of 135
prof %>% filter(balance < 0) %>% nrow

#without algorithm, balance would be $1,475,642
sum(prof$balance)

# without any losing jobs, balance would be $2,516,246
prof %>% filter(balance > 0) %>% sum

# without any profitable jobs, balance would be -$1,040,322
prof %>% filter(balance < 0) %>% sum

#how much real profit would you be dismissing? $82631
prof.dismis<- prof %>% filter(pred >= 0.5 & balance > 0) %>% select(balance) %>% sum
#how much real loss would you be avoiding? $566,481.6
loss.dismis<- prof %>% filter(pred >= 0.5 & balance < 0) %>% select(balance) %>% sum
# total benefit = $483,850
algor.prof<- -(loss.dismis + prof.dismis)

# total balance would be $1,959,492
sum(prof$balance) + algor.prof

# a 32.8% improvement
algor.prof/sum(prof$balance)

```

### 2015-10-30 rf_log_naiive_auc.rmd
finish forest and log

---
title: "test_imputed"
author: "Amy Cook"
date: "October 27, 2015"
output: word_document
---

Test on random forest, logistic regression and naiive bayes :)
Compare:

* subset 9c data
* rf imputed
* rfMICEd imputed
* subset 9a data (uncompressed JD.Second and Business)
* rf imputed with 9a
* rfMICE imputed with 9a

Out of the best from above, can experiment with adding in client and contact.

Load 3 datasets:

```{r, echo=FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
library('e1071',lib='C:/Progra~1/R/R-3.2.2/library')
library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')

setwd("C:/Users/n9232371/Documents/Consultbusiness/data")

all10mice<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')[,-1]
all10rf<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10rf.csv')[,-1]
all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]

#VARS::::
# no.users, Discipline, pc.pro, Business, code.client,JD.Second, majority.pos, b.timespan.cbrt,
# inv.mlsto.log, client.totinv.log, pc.majpos.log, code.contact, Billing.Type, mlsto, Year, b.rpdol

all10a<- all9c %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol)
all10a<- all10a[complete.cases(all10a),]
# 2280 cases

all10b<- all9c %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10b<- all10b[complete.cases(all10b),]
# 1987 cases

all10c<- all9c %>% select(no.users, Discipline, pc.pro, Business, JD.Second,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10c<- all10c[complete.cases(all10c),]
# 1106 cases

#make 3 new data sets from all9a instead of all9c. Ie long winded Business and JD.Second levels
all10d<- all9a %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol)
all10d<- all10d[complete.cases(all10d),]
# 2280 cases

all10e<- all9a %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10e<- all10e[complete.cases(all10e),]
# 1987 cases

all10f<- all9a %>% select(no.users, Discipline, pc.pro, Business, JD.Second,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
all10f<- all10f[complete.cases(all10f),]
# 1106 cases

```

First tune random forest parameters with caret package.
Try tuning with all 3 datasets

```{r, echo=FALSE}

# set seeds
set.seed(123)
seeds <- vector(mode = "list", length = 51)
for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
## For the last model:
seeds[[51]] <- sample.int(1000, 1)

d.sets <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")

# must turn outcome into two level factor
for( i in 1:length(d.sets)){
        df = get(d.sets[i])
        df$f.rpdol<- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol)=="0"] <- "profit"
        levels(df$f.rpdol)[levels(df$f.rpdol)=="1"] <- "loss"
        assign(d.sets[i], df, envir = .GlobalEnv)
}

#function for running caret..

caret.all <- function(df = all10rf){
        set.seed(100)
        sample<- sample(1:nrow(df), 2/3*nrow(df), replace=F)
        train<- df[sample,]
        test<- df[-sample,]

        if(nrow(df) == 2364){
                formula<- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type"

        } else{
                if(nrow(df) <= 2300){
                        formula<- "f.rpdol ~ . -Year - mlsto -b.rpdol"
                }
        }

        cv.Control <- trainControl(method = "cv",
                           number = 10,
                           seeds= seeds
                           , summaryFunction= twoClassSummary,
                           classProbs=TRUE
                           )
        gbmGrid <- expand.grid(mtry = c(3,4,5,6,7))

        set.seed(2)
        gbmFit <- train(as.formula(formula), data= train,
                method = "rf", trControl = cv.Control, verbose = FALSE,
                tuneGrid = gbmGrid,
                metric = 'ROC', allowParallel = T, ntree= 1000)

        plot(gbmFit)

}

caret.all(df = all10rf) # mtry = 4, AUC = 0.7545 
caret.all(df = all10mice) # mtry = 5, AUC = 0.757
caret.all(df = all10a) # mtry = 3, AUC = 0.719
caret.all(df = all10b) # mtry = 4, AUC = 0.771
caret.all(df = all10c) # mtry =4, AUC = 0.7735

# all10c is performing best! let's make a function that runs each df through the test set?
df.list <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")

test.roc <- function(dfs = df.list, seed = 100, mtrys = c(4,5,3,4,4), method = "log",
                     extra.var = NULL, mult = 1){

        auc<- rep(list(rep(0, length(dfs))), length(dfs))
        colours <- c("red", "yellow", "green", "blue", "purple")

        for(i in 1:length(dfs)){
                df = get(df.list[i])

                # mlsto is spinning errors, get rid of
                df = df[, !(names(df) %in% 'mlsto')]
                set.seed(seed)

                # create f.rpdol
                df$f.rpdol <- as.factor(df$b.rpdol)
                levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
                levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"

                #set up 10 fold cv using caret function
                folds <- createFolds(df$f.rpdol, k = 10, list = T, returnTrain = F)
                #now have folds[[1]] through folds[[5]] list of rows to exclude

                if(mult > 1){
                        for(k in 2:mult){
                                set.seed(sample(1:1000,1))
                                folds.temp = createFolds(df$f.rpdol, k = 10, list = T, returnTrain = F)
                                folds = append(folds, folds.temp)
                        }
                }

                #if extra.var not found in names(df), change to NULL
                if(!is.null(extra.var)){
                        if(!(extra.var %in% names(df))) extra.var = NULL
                }

                #define formula for algorith methods, either rf, log, boost

                if(nrow(df) == 2364){
                        if(is.null(extra.var)){
                                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

                        } else {
                                formula<- paste("f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second", "+", extra.var, sep = " ")
                        }

                } else{
                if(nrow(df) <= 2300){
                        formula<- "f.rpdol ~ . -Year -b.rpdol"
                }
                }

                if(method %in% 'boost') {
                        formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"

                }

                #for loop for each fold

                for (j in 1:length(folds)) {

                        #turn predictors and response into 'ordered' variables for roc() to work
                        test<- df[folds[[j]],]
                        levels(test$f.rpdol) <- c('profit', 'loss')
                        test$f.rpdol <- ordered(test$f.rpdol)

                        if(method %in% 'rf'){
                                fit <- randomForest(as.formula(formula), data = df[-folds[[j]],], 
                                            mtry = mtrys[i], ntree = 1000)
                                pred <- predict(fit, test, type = "prob")
                                pred <- pred[,2]

                        } else {
                                if(method %in% 'log'){
                                        fit<- glm(as.formula(formula), 
                                                  data = df[-folds[[j]],], family = binomial())
                                        pred <- predict(fit, test,
                                                        type = "response")
                                } else {
                                        fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
                                                   distribution = "bernoulli", n.trees = 10000,
                                                   shrinkage = 0.001, interaction.depth = 4,
                                                   n.minobsinnode = 20)

                                                pred <-predict(fit, test, 
                                                               n.tree = 10000, type = "response")
                                                }
                        }

                        #pROC
                        pred.p <- roc(test$f.rpdol, pred)

                        if (i == 1 & j== 1) {
                        plot.roc(pred.p, col = colours[i], print.thres = F)

                        } else{
                        plot.roc(
                        pred.p, col = colours[i], add = TRUE , print.thres = F)
                        }

                        auc[[i]][j]<- auc(pred.p)

                        #print j value
                        cat(j, " ")

                        }

                print(i)

        }

        assign("auc.func", auc, envir = .GlobalEnv)
        assign("formula", formula, envir = .GlobalEnv)

}

df.list <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")
test.roc(dfs = df.list, seed = 100, mtrys = c(4,5,3,4,4), method = "rf")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

test.roc(dfs = df.list, seed = 400, mtrys = c(4,5,3,4,4), method = "rf")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[1:10], auc$value[11:20])

# for random forest, the all10rf performs second/third best and most consistently, about AUC = 0.765
```

Now run logistic regression

```{r, echo= FALSE}

df.list <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")
test.roc(dfs = df.list, seed = 400, mtrys = c(4,5,3,4,4), method = "log")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[1:10], auc$value[11:20])

# logistic regression does well! with all10mice, AUC = 0.775

```

Try naiive bayes - try full NA dataset, all9c

```{r, echo=FALSE}

formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type + JD.Second"

df = all10mice

set.seed(sample(1:1000,1))
sample<- sample(1:nrow(df), 2/3*nrow(df), replace=F)
train<- df[sample,]
test<- df[-sample,]

fit<- naiveBayes( as.formula(formula), train)
pred <- predict(fit, test, type = 'raw')
pred<- pred[,2]

pred.p <- roc(test$b.rpdol, pred)
plot.roc(pred.p, col = colours[i], print.thres = T)

#naiive bayes can do about 0.71 at best!! logistic regression, random forest and  boosted trees are
# definitely doing something!!!

```

Test if data sets subset from all9a (full original Business and JD.Second) instead of all9c perform better

```{r, echo=FALSE}

#create factor profit/loss variable
d.sets<- c("all10d", "all10e", "all10f")

for( i in 1:length(d.sets)){
        df = get(d.sets[i])
        df$f.rpdol<- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol)=="0"] <- "profit"
        levels(df$f.rpdol)[levels(df$f.rpdol)=="1"] <- "loss"
        assign(d.sets[i], df, envir = .GlobalEnv)
}

#perform 10 fold cv with same parameters as 10a, 10b, 10c
df.list <- c( "all10a", "all10b", "all10c", "all10d", "all10e", "all10f")
test.roc(dfs = df.list, seed = 100, mtrys = c(3,4,4,3,4,4), method = "rf")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#try log reg
test.roc(dfs = df.list, seed = 100, mtrys = c(3,4,4,3,4,4), method = "log")
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")
#performed much better with compacted categories

```

Experiment without JD.Second:

```{r, echo=FALSE}
df.list <- c("all10rf", "all10mice")
test.roc(dfs = df.list, seed = 100, mtrys = c(4,5), method = "log")
with.JDauc<- melt(auc.func)

#remove 'JD.Second' from formula and re-run test.roc
test.roc(dfs = df.list, seed = 100, mtrys = c(4,5), method = "log")
witho.JDauc<- melt(auc.func)

all.JDauc<- rbind(with.JDauc, witho.JDauc)
all.JDauc[21:30,]$L1<- 3
all.JDauc[31:40,]$L1<- 4

ggplot(all.JDauc, aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
t.test(all.JDauc$value[1:10], all.JDauc$value[21:30])
t.test(all.JDauc$value[11:20], all.JDauc$value[31:40])

#JD.Second really doesn't matter, pretty much just noise, meh keep it in.

```

See what effect including clients has on random forest and log.
It did not help boosted trees.
Experiment on all10mice dataset.

Create five experimental datasets:

1. all10mice.c1: complete cases of all10mice once added code.client
2. all10mice.c2: only include clients with over 20 count. rest become 'unknown'
3. all10mice.c3: only include clients with over 40 count. rest become 'unknown'
4. all10mice.c4: only include clients with over 60 count. rest become 'unknown'
5. all10mice.c5: only include clients with over 70 count. rest become 'unknown'

```{r, echo=FALSE}
#create data frame of client codes and their counts, as well as contacts
client.count<- ddply(all10mice, .(code.client), nrow)
client.count<- arrange(client.count, -V1)
contact.count<- ddply(all10mice, .(code.contact), nrow)
contact.count<- arrange(contact.count, -V1)

#all10mice.c1
client.count2<- client.count[1:53,]
all10mice.c1<- all10mice %>% filter(code.client %in% client.count2$code.client)
all10mice.c1<- all10mice.c1 %>% select(no.users, Discipline, pc.pro, Business,
                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
                         mlsto, Year, b.rpdol, JD.Second, majority.pos, pc.majpos.log,
                         Billing.Type, code.client, f.rpdol)
all10mice.c1 <- all10mice.c1[complete.cases(all10mice.c1),]
all10mice.c1$code.client<- droplevels((all10mice.c1$code.client))
all10mice.c1$reduc.client<- all10mice.c1$code.client

#four new data frames for different client thresholds

all10mice.c2 <- all10mice
all10mice.c3 <- all10mice
all10mice.c4 <- all10mice
all10mice.c5 <- all10mice

add.clientvar<- function(df = c('all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5'), 
                         thresh = c(20, 40, 60, 70), thresh.contact = c(5,8,13,20)) {

        for (i in 1:length(thresh)) {

                df.temp = get(df[i])
                threshold = thresh[i]

                #client code first
                client.count2 <- client.count[client.count$V1 > threshold &
                                                      !is.na(client.count$code.client),]

                df.temp$code.client <- as.character(df.temp$code.client)
                df.temp$reduc.client <- ifelse(df.temp$code.client %in% client.count2$code.client,
                                             df.temp$code.client,
                                             'unknown')
                df.temp$code.client <- as.factor(df.temp$code.client)
                df.temp$reduc.client <- as.factor(df.temp$reduc.client)

                #contact code second
                contact.count2 <- contact.count[contact.count$V1 > thresh.contact[i] &
                                                      !is.na(contact.count$code.contact),]

                df.temp$code.contact <- as.character(df.temp$code.contact)
                df.temp$reduc.contact <- ifelse(df.temp$code.contact %in% contact.count2$code.contact,
                                             df.temp$code.contact,
                                             'unknown')
                df.temp$code.contact <- as.factor(df.temp$code.contact)
                df.temp$reduc.contact <- as.factor(df.temp$reduc.contact)

                assign(df[i], df.temp, envir = .GlobalEnv)

                }
        }

add.clientvar(df = c('all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5'), 
                         thresh = c(20, 40, 60, 70), thresh.contact = c(5,8,13,20))

#now have 5 datasets to compare, all derived from client count and all10mice
# want to run each of these through 10 fold cross validation to see if we get better/worse results

#tune first
caret.all(df = all10mice.c1)

#10 fold test/train cv for code.client
df.list<- c('all10mice', 'all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5')
test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "rf",
         extra.var = 'reduc.client')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

test.roc(dfs = df.list, seed = 100, mtrys = rep(5,5), method = "log",
         extra.var = 'reduc.client')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#10 fold test/train cv for code.contact
df.list<- c('all10mice', 'all10mice.c2', 'all10mice.c3', 'all10mice.c4', 'all10mice.c5')
test.roc(dfs = df.list, seed = 100, mtrys = rep(5,5), method = "rf",
         extra.var = 'reduc.contact')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

test.roc(dfs = df.list, seed = 100, mtrys = rep(5,5), method = "log",
         extra.var = 'reduc.contact')
ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[1:10], auc$value[11:20])

# including client.code really doesnt matter.
# including contact.code seems to make it marginally worse
# leave.. code.contact out, leave code.client in maybe.

#save all10mice.c4 for BN trial
write.csv(all10mice.c4, 'C:/Users/n9232371/Documents/Consultbusiness/data/all10mice_c4.csv')

```

Compare

* logistic regression
* random forest
* boosted tree

```{r, echo = FALSE}

all10mice.c4 <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice_c4.csv')
df.list<- c('all10mice')

#logistic regression with all10mice
test.roc(dfs = 'all10mice.c4', seed = 100, mtrys = c(5), method = "log",
         extra.var = NULL, mult = 3)
auc.log.noBil = auc.func
auc.log = auc.func
auc.log.nocli = auc.func

#random forest with all10mice
test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "rf",
         extra.var = NULL, mult = 3)
auc.rf = auc.func

#boosted tree with all10mice
test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "boost",
         extra.var = NULL, mult = 3)
auc.boost = auc.func

auc.func<- list(auc.log, auc.rf, auc.boost, auc.log.noBil, auc.log.nocli)

ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
        stat_summary(fun.y="mean", geom="point", colour = "red")

#significant difference??
auc<- melt(auc.func)
t.test(auc$value[61:90], auc$value[91:120])

# no significant difference. I'm leaning towards boosted tree.
```

Look at variable importance

```{r, echo= FALSE}
formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

#random forest
fit.rf <-randomForest(as.formula(formula), data = all10mice,
        mtry = 5, ntree = 1000, importance = T)
imp<- as.data.frame(importance(fit.rf))
imp<- imp[order(imp$MeanDecreaseAccuracy,decreasing=TRUE), ]

#logistic regression
fit.log <- glm(as.formula(formula),
           data = all10mice, family = binomial())

summary(fit.log)

#boosted tree
formula<- formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"

fit.boost <-  gbm(as.formula(formula), data = all10mice,
        distribution = "bernoulli", n.trees = 10000,
        shrinkage = 0.001, interaction.depth = 4,
        n.minobsinnode = 20)

summary(fit.boost)
saveRDS(fit.boost, 'C:/Users/n9232371/Documents/Consultbusiness/data/fit_boost.rds')
fit.boost<- readRDS('C:/Users/n9232371/Documents/Consultbusiness/data/fit_boost.rds')

#partial dependence plots
fit.boost$var.names
plot.gbm(fit.boost, 9, type="response")
biz<- plot.gbm(fit.boost, 7, type="response")

##
### cforest
formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

system.time(
fit.cforest<- cforest(as.formula(formula),
        data= all10mice
        )
)

var.imp<- varimp(fit.cforest) %>% as.data.frame
colnames(var.imp)[names(var.imp) %in% '.']<-'imp'
var.imp$var <- rownames(var.imp)
#reorder levels for bar plot
var.imp <- within(var.imp, 
                    var <- factor(var,levels=
                                            var.imp[order(-abs(var.imp$imp)),]$var))

q<- ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
a<- q + geom_bar(stat='identity')

ggsave(a, filename="cfor_all10mice.png", path=
               'C:/Users/n9232371/Documents/github/consult/finalwriteup/images/cforest_all10mice')

```

Look at results

```{r, echo = F}

#random forest
set.seed(200)
sample<- sample(1:nrow(all10mice), 2/3*nrow(all10mice), replace=F)
train<- all10mice[sample,]
test<- all10mice[-sample,]

formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

#random forest
fit.rf <-randomForest(as.formula(formula), data = train,
        mtry = 5, ntree = 1000, importance = T)
pred.rf <- predict(fit.rf, test, type = "prob")
imp<- as.data.frame(importance(fit.rf))
imp<- imp[order(imp$MeanDecreaseAccuracy,decreasing=TRUE), ]

#logistic regression
fit.log <- glm(as.formula(formula),
           data = train, family = binomial())
pred.log <- predict(fit.log, test,
                type = "response")
summary(fit.log)

#boosted tree
formula<- formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"

fit.boost <-  gbm(as.formula(formula), data = train,
        distribution = "bernoulli", n.trees = 10000,
        shrinkage = 0.001, interaction.depth = 4,
        n.minobsinnode = 20)

pred.boost <- predict(fit.boost, test, n.tree = 10000, type = "response")
pred.p <- roc(test$f.rpdol, pred.boost)

plot.roc(pred.p, col = 'red', print.thres = T)
summary(fit.boost)

# look at test data and predictions
test.review <- test 
test.review$pred <- pred.boost 
test.review<- arrange(test.review, -pred)

write.csv(test.review, 'C:/Users/n9232371/Documents/Consultbusiness/data/test_review.csv')
```

### 2015-10-30 BNprep_all9c.R

@@ -1,0 +1,117 @@
+## prep for bayesian network
+
+#discretise numeric variables and change all 'NA' to blank
+
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
+
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
+head(all9c)
+
+# write a function to
+# discretise numeric columns!!!! pc.pro, timespan.cbrt, inv.mlsto.log, client.totinv.log,
+# pc.majpos.log, return.pdol
+
+discretise<- function(df = all9c, extra.var = ('code.client', 'code.contact')){
+        
+        #pc.pro
+        df[df$pc.pro>100 & !is.na(df$pc.pro),'pc.pro']<- 100
+        df$b.pc.pro<- cut(df$pc.pro, breaks=c(0,15,40,75,95,100), include.lowest=TRUE,
+                             labels= c("p0_15", "p15_40", "p40_75", "p75_95", "p100"))
+        
+        ###
+        #timespan.cbrt
+        #done already
+        
+        #create clusters based on summary:
+        man.clust<- function(first.vec=all9c$timespan.cbrt, breaks.man=c(0,4.16,6.558,15),
+                             labs = c("t0_70", "t70_280", "t280_")){
+                final = cut(first.vec, breaks=breaks.man, include.lowest=TRUE, labels= labs)
+                return(final)
+        }
+        
+        ###
+        #inv.mlsto.log
+        
+        all9c$b.inv.log<- man.clust(first.vec=all9c$inv.mlsto.log, 
+                                    breaks.man=c(4.5, 7.81, 9, 11, 14.4),
+                                    labs = c("i_0_2_5k", "i_2_5k_8_1k", "i_8_1k_60k", "i_60k_1_6m"))
+        
+        ###
+        #client.totinv.log
+        
+        all9c$b.client.totinv <- man.clust(first.vec=all9c$client.totinv.log, 
[…]

### 2015-10-30 financial.R

@@ -1,85 +1,0 @@
----
-title: "Financial_Analysis"
-author: "Amy Cook"
-date: "October 21, 2015"
-output: word_document
----
-
-#Assess financial implications for a predictive model
-
-We typically get a confusion matrix built from probabilities
-
-
-```{r_loadpackages, echo=FALSE, include=FALSE}
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')
-
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-```
-
-Load data
-
-```{r, echo=FALSE}
-all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]
-all9a$b.timespan.cbrt<- cut(all9a$timespan.cbrt, 
-                            breaks= c(0,2.7589, 4.2358, 6.5, 8.183, 10.307, 14.1), 
-                            include.lowest=TRUE, 
-                            labels= c("1d-3wk","3wk-2.5m", "2.5m-9m","9m-1.5y","1.5y-3y",">3y"))
-
-set.seed(300)
-sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
-train<- all9a[sample,]
-test<- all9a[-sample,]
-```
-
-Run boosted tree
-
-```{r, echo=FALSE}
-
-formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business - timespan.cbrt"
-
-fit<-gbm(as.formula(formula), 
-         data=train, distribution="bernoulli", n.trees= 6000,
-         shrinkage=0.001,interaction.depth=3,
-         n.minobsinnode = 10)
-pred<- predict(fit, test, n.tree=4000, type="response")
-ans<- test$b.rpdol
-
-
-```
-
-Create vector of profit points
-
[…]

### 2015-10-30 imputingall9c.R

@@ -9,42 +9,46 @@ library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
 library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
 library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
 library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
+library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
+library('CALIBERrfimpute', lib = 'C:/Progra~1/R/R-3.2.2/library')
 
-library('randomForest', lib = 'C:/Progra~1/R/R-3.2.2/library')
 all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
 all9c$b.rpdol<- as.factor(all9c$b.rpdol)
 
-ab<- rfImpute(b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log +
-                 client.totinv.log + pc.majpos.log + majority.pos + Business + JD.Second, data= all9c,
+all10rf<- rfImpute(b.rpdol ~ Discipline + pc.pro + timespan.cbrt + no.users + inv.mlsto.log +
+                 client.totinv.log + pc.majpos.log + majority.pos + Business + JD.Second + Billing.Type,
+                 data= all9c,
               ntree= 500)
+all10rf<- cbind(all10rf, all9c[,names(all9c) %in% c('code.client', 'code.contact', 'mlsto',
+                                                    'b.timespan.cbrt', 'b.rpdol')])
 
-experiment<- mice.impute.rf(all9c$pc.pro, ry = 
-                            x= all9c[,names(all9c) %in% c('no.users', 'Discipline', 'Business', "timespan.cbrt",
-                                                      "inv.mlsto.log", 'client.totinv.log', 'return.pdol','b.rpdol',
-                                                      'Year')], 
-                            ntree= 500)
-
-testing<- mice(all9c, meth = 'rf', ntree =10)
 
 ## test if it helps boosted trees
 
 #convert predictor variable back to numeric 0,1
 
-levels(ab$b.rpdol)<- c("0", "1")
-ab$b.rpdol<- as.character(ab$b.rpdol)
-ab$b.rpdol<- as.integer(ab$b.rpdol)
+levels(all10rf$b.rpdol)<- c("0", "1")
[…]

### 2015-11-09 finalwriteup/draft.rmd
save early nov

@@
 A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average (Flyvbjerg and Budzier, 2011).
-These statistics cast a concerning glimpse at the financial story playing out behind the scenes of these projects which can cause job losses, ruin private businesses, and burn through government balancesheets.
+These statistics cast a concerning glimpse at the financial story playing out behind the scenes of these projects which can cause job losses, ruin private businesses, and churn through government balancesheets.
 Complex long term projects are necessary to produce  railways, houses, websites, and movies.
-Depending on the industry, the the financial risk of completing a project on budget is transferred to consulting companies who are enlisted to complete portions of a project.
+This research focuses on the financial risk that is transferred to consulting companies who are enlisted to complete portions of these complex projects.
 The risk is transferred by requiring a fixed flat fee from the consultant, meaning they commit to complete their role for a nominated price.
@@
 Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery.
-Employees also experience stress attempting to complete the project within a disappearing budget and may perform lower quality work in exchange for speed.
+Employees also experience stress attempting to complete the project within a disappearing budget and may produce lower quality work in exchange for speed.
 This is unfortunate if in reality, the project fee was under estimated.
@@
 The financial risk taken on fixed flat fee projects may also discourage business' from taking different risks with better potential pay-offs, such as expanding the business into a new area.
-Clearly, predicting fixed fee project costs more accurately could positively affect a business' growth opportunities and staff morale.
+Logically, predicting fixed fee project costs more accurately could positively affect a business' growth opportunities and staff morale.
 
@@
 The challenge for consultants to reduce the risk of a project going over budget, in terms of hours spent, is not easily solved.
-The reason is that complex projects are always different from each other, and similar work may take significantly different amounts of time.
+Complex projects are always different from each other, and similar work may take significantly different amounts of time.
 For example, an ad filmed in a remote location for chocolate will cost significantly more than an ad filmed on an exisitng set, even though the product structure is almost identical.
@@
 However in many industries this is not competitive enough.
-Other consultancies may be willing to offer a fixed flat fee for the same work, reducing financial risk for the client, and luring the client away.
+Other consultancies may be willing to offer a fixed flat fee for the same work, reducing financial risk for the client, and luring the client to them.
 Another fee strategy may be to track worker hours, and stop work once the consulting fee has been depleted.
@@
 This also satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective.
-However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or low profits to preserve the relationship with their client.
+However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or endure very marginable profits to preserve the relationship with their client.
 Consultant's may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry over creating financial friction.
@@
 They found that the main method for cost estimation was breaking the project into detailed parts and adding up the cost of each item.
-The next two most popular methods were comparison with similar projects based on documented facts, and comparison with similar projects based on personal experience.
+The next two most popular methods were 'comparison with similar projects based on documented facts', and 'comparison with similar projects based on personal experience'.
 These can all be classified as experience based models [@Akintoye2000].
@@
 Their theory stems from Kahneman's work on decision making that won him the nobel prize for economics in 2002.
-The research argues that humans' natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
-A manager sees challenges in a project as something that can be overcome by the team's high skill level, and downplay or ignore the risk of problems that are out of the team's control.
+His research argues that humans' natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
+A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplay or ignore the risk of problems that are out of the team's control.
 It is for this reason that it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
@@
 This outside view can be applied to complex projects, and involves ignoring the details of the current project and instead analysing the outcomes of several projects of a similar description.
-This is called reference class forecasting [@Lovallo2003].
+Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003].
 One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
-The correlation for the current project can then be estimated via a statistical model which is used to adjust the forecasted cost made by detailed analysis (inside view) [@Flyvbjerg2011].
+The correlation for the current project can then be estimated via a statistical model which is used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011].
 The traditional way to think about a project is to focus on the project details.
@@
 
-The implementation of reference class forecasting began in project management for the first time in 2004, was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, Australia.
+The implementation of reference class forecasting began in project management for the first time in 2004, was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, and Australia.
 An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].
@@
 
-Given this theory of reference class forecasting was first clearly published in 2003, the idea of predicting project costs from previous data has been the subect of research for many years before that and continues to this day.
+The theory of reference class forecasting was first clearly published in 2003, however the idea of predicting project costs from previous data has been the subject of research for many years before that and continues to this day.
 Although research results have often been promising, industry uptake of the idea is not yet established.
 
-First an overview of the methods used will be presented followed by published theories for the lack of industry uptake.
+### 1.1.1 Overview of Past Reference Class Forecasting Methods and relevant Industry Uptake
+
+In the following section, an overview of the historical attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.
 
@@
 
-A substantial amount of research on project costprediction has been dedicated to the fields of software development and construction projects in particular.
+#### 1.1.1.1 Existing Models
+
+A substantial amount of research on project cost prediction has been dedicated to the fields of software development and construction projects in particular.
 These analyses have generally used project information from 15 to 20 large projects.
 A single dataset may consist of projects from around the world, and each project a different client and project team.
-The mathematical methods used to predict project costs vary from statistical regressions to deep-learning neural networks.
-Most studies report neural networks are more accurate than regression
+The mathematical methods used to predict project costs vary from statistical regressions to deep-learning neural networks [@Love2005][@Kim2004].
+Most studies report neural networks are more accurate than regression, and generally the studies report positive predictive powers of their developed models.
 
-Previous research has been done on the prediction of final costs of a complex project with varying success.
-In the prediction of IT software development, formal estimation methods have existed for many years, however expert judgment is still the dominant method and available evidence does not suggest the models improve on expert judgment (Jorgensen and Shepperd, 2007).
-Construction industry research has used models ranging from neural networks to case based reasoning to regression to predict final building costs.
-However, often expert judgment was not mobilised, and the models were not comprehensive enough to be used in industry.
-Manager's experience and intuition is still the dominant method for construction cost calculation.
+A McKinsey study outlined a case in the film industry where a motion picture company used reference class forecasting of movie project success (cases were weighted by similarity) to decide which movies should be heavily promoted.
+This model improved forecasts on financial return by 135% relative to single project analogies [@Flyvbjerg2014].
 
-*** ^^^more research^^^ ***
+Statistical research projects which trial reference forecasting have shown promising results in IT, construction, and film projects.
+However, there is little evidence that the methods have successfully transitioned into industry.
 
-Potential solution:
-* stat analysis like reference class forecasting and building construction except analysis performed internally using the company's historical data on past projects.
+#### 1.1.1.2 Industry Uptake of Existing Models and Current Practises
 
-case study:
+Surveys on current industry practise for cost estimation are scarce, however, a survey of the construction industry in the UK in 2000 found that a manager's experience and intuition is still the dominant method for construction cost calculation [@Akintoye2000].
+A study on the IT industry by @Jorgensen2007 cited that the parametric models were not comprehensive enough to be used in industry and also found that expert judgment was not mobilised enough in the process of developing these tools.
 
-This project undertakes a case study on a consulting company in the construction industry. Specifically, the
+Although reference class models have not gained traction in either the software or construction industry (where most of the research to date has focussed), it is worth outlining their current practises in cost estimation and the reasons behind these.
 
-* in the case study (and is typical of smaller organisations??), the CRM is an untapped source of information storing
-    * employee hours
+In the construction industry, infrastructure and building projects follow a traditional and well established process, where contractual norms have developed over many decades [@Badenfelt2011].
+The industry is so large, broad and well established that it has inertia in the way that projects are estimated, and contractually bound.
+The contractor (builder) first reviews the drawings and if the plans are not detailed enough, they offer a temporary budget price.
+Once they feel that the drawings are sufficiently detailed, they will offer a fixed price [@Badenfelt2011].
+However, as the job progresses, if any work is added that was not in the detailed drawings, they maintain the right to charge a variation [@Badenfelt2011].
+This method is clear and variations are easily justifiable to the client, however cost, labour and time estimates are still regularly under calculated, even from detailed drawings.
+And for large infrastructure projects, small percentage miscalculations can be exceedingly expensive.
+
+Contractor's practise of pricing from detailed drawings and charging for variations does not translate to other consultants in the construction industry, such as engineers and architects.
+Consultants must devlop the detailed plans for the contractors, and must therefore offer fixed price contracts based on drawings with little detail.
+It is much harder for a consultant to charge a variation for a portion of their time spent on something that was not included in the original simple plans.
+This is despite the real and common instance of budget blow outs due to overspent time [@Harris1999].
+One method to provide opportunity for a variation is stating a minimum percentage fee of the final building cost, which can overrule the original fixed price contract.
+If a consultant has spent time beyond their fixed price budget,this may account for some or in the best case, all, of their overspent time.
+
+Construction projects generally are completed more on time and on budget when compared to IT projects as the greatest cost is the project is building materials [@Badenfelt2011].
+Unlike time, building materials can be accurately calculated from detailed design drawings.
+Again, this omits data from consultants who are part of the building industry but charge based on their time only.
+
+In contrast to construction, IT projects are a relatively new practise where alternative contractural arangements have been trialled.
+Through experimentation with contracts, the Agile method was developed.
+In summary, the method treats both cost and time as fixed quantities for a project, and if any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
+Furthermore, the project outcomes are continually adjusted ad revised based on frequent progress review points that assesses unexpected problems.
+This agile attitude has become ingrained in many IT practises and this flexible upfront arrangement with clients often forges better long term relationships [Cockburn2001].
+
+It is not completely clear why industry uptake of cost estimation tools using the outside view has been weak, and it may be the case that in modern industries such as in the IT industry, problems associated with cost estimation can be innovatively designed out of the contract.
+Possible reasons that parametric reference class forecasting models developed by researchers have not been incorporated into industry practise include:
+
+* lack of appropriate data
+    * if a project is unique, such as a national infrastructure project, it many not feel appropriate to gather sufficient data points from other countries and time periods.
+* lack of understanding from industry decision makers - the prediction process was not explained well in marketing of the tools
+* models were not developed in collaboration with industry
+* does not feel intuitive price from an 'outside view' of a project which disregards small project details
+* insufficient time allocated to pricing that makes change burdensome
+
+In fact, a study by @Moores1992 concluded that a lack of framework to support the outside view model use and a failure in marketing were the two most common explanations for lack of uptake of cost estimating tools which use statistical techniques in predicting project costs.
+This is unfortunate as almost all research indicates an outside view analysis of a job substantially improves cost estimation accuracy for fixed price projects.
+
+### 1.1.2 Case Study
+
+This research will explore the potential for reference class forecasting to improve cost estimation in the context of an engineering consulting company in the construction industry.
+In this field, labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm.
+In recent years, the size of losses from unprofitable projects in the case study could equal up to 25-30% of the profits from profitable projects.
+This demonstrates substantial room for improvement and the possiblity of multitude positive flow on effects.
+
+Cost estimation is currently performed by first carefully reviewing preliminary drawings for a project where time costs will be calculated using a manager's personal experience with similar projects.
+This will often be cross-checked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
+No formal mathematical method of comparison to similar projects is performed.
+This is the case for a number of possible reasons:
+
+* lack of time to interrogate the CRM
+* clunky data availability offered by the CRM
+* lack of awareness of projects performed by other managers that may have been similar.
+
+Reference class forecasting in this study differs from previous research, in that statistical analysis will be performed on internal past company project data that has been progressively recorded in the company's Customer Relationship Management (CRM) software which was purchased.
+This presents an opportunity to evaluate reference class forecasting in a unique situation, where the 'reference' projects are internal and over two thousand cases are available.
+
+The CRM data currently stands as an untapped source of information within the case study organisation. It stores a rich variety of data including:
+
+* employee timesheet hours with dates
+* other project costs (taxis, printing)
 * client information/characteristics
-    * invoicing
+* client identification code
+* invoiced amounts for each project and dates
 * employee costs
 * employee charge out rates
-    * project information
-* currently the case study company uses a widely popular CRM, where they can output information but cannot perform modelling or analysis:
-    * look up one job at a time
-    * basic summaries of overall hours spent for the month
+* project description
 
-### 1.1.1 Limitations
+A statistical model predicting a measure of project financial success will be developed using the available project variables and trialling numerous machine learning algorithms and statistical methods.
+The company uses well established CRM software which readily provides simple output statistics such as project time-cost vs invoiced amount summaries.
+However, the analytical capabilities were limited to:
 
-* looking at a single case study company in a specific industry
-* Managers commonly price jobs based on the budget for the entire project, of which the consultant is a minor part.
-An example of this is a consultant who typically charges 1% of the total project cost
-    * this data has not been historically recorded. Can start being recorded and would be extremely useful
-    * even better to have expected cost of entire project at start of project, as well as actual final cost of project.
-* A fee is based off detailed information that is unique to each project, for example the size of a building in terms of storeys and square meterage.
-This detailed information is not available in the data, and is too complex and varied for an algorithm to digest.
-Therefore it is not reasonable for an algorithm to predict the fee of a project at this stage, but much more achievable to predict the profitability of a project (ie return per dollar) based on historical data.
-This profitability prediction relates to the profitability of the project if the fee is generated using the same method the company has used in its historical data.
-* Although the aim is to predict the profitability of a project, the model aims to change managers' fee setting decisions if the algorithm predicts low profit or a loss based on historical data. 
-* external factors forcing price down in some cases - competiton. clients
+* simple scatter plots and bar charts of the raw data
+* summaries such as overall hours spent vs invoiced amount for each project
+
+The company has a wealth of data but limited means to extract insight.
+This study will exploit the CRM data by performing more sophisticated statistical analysis with the intention of building a predictive model for cost estimation.
+The benefits of this study in comparison to previous cases are that thousands of past data points are available, when previous studies used on average 15-20 cases and at most 170.
+This improves the potential for accurate predictions.
+If successful, the model also has higher potential for managerial uptake as the algorithm will be directly built and trained on company data.
+Managers can interrogate 'similar' data points and relate to the actual projects or speak to a colleague who was involved.
+This has the potential to give the 'outside view' stronger influence on the final decision.
+
+### 1.1.3 Case Study Limitations
+
+The case study provides a framework to test the value of a consulting company's CRM data in assisting project cost estimates.
+However, it is important to be aware of limitations when reviewing outcomes and framing the problem.
+The most obvious limitation is that we are disection a single company in a specific industry, which limits the ability to state conclusions could apply to a broad range of cases.
+This can be overcome with future studies.
+
+The data set from this case study does not include information that would improve project cost predictions.
+For example, managers commonly price jobs based on the budget for the entire project, of which the consultant is a minor part.
+The manager might typically moderate his fee by say 1% of the entire project cost.
+The value of the entire project has typically not been recorded although there is a space in the project software form for this value.
+It would be even more beneficial for managers to record the expected total project cost at the beginning of a project versus the conclusion of a completed construction project.
+Another piece of data that was not recorded was external factors influencing the fixed price quote including competition from similar firms that the client played against the case study company.
+
+In addition to this data that was not typically entered by workers, the dataset also excludes detailed information about the projects which are crucial to the final pricing.
+This includes details such as preliminary drawings describing the project (square meterage, number of storeys), and project summaries provided by the client to enable a quote.
+This kind of detailed information will never be recorded in a CRM and in many ways contradicts the purpose of the 'outside view' which intends to present similar projects without being over-influenced by the finer details of a project.
+Given the lack of detailed information, it is not reasonable for an algorithm to predict the specifc fee of a project.
+It is much more effective to attempt predict the profitability of a project (ie return per dollar) based on the historical data.
+This profitability prediction relates to the profitability of the project if the fee was generated using the established method the company used in its historical cases.
+
+Finally, CRM data is entered by employees manually as time passes, jobs are invoiced, and projects are finally completed.
+Because the data is input by humans, it is susceptible to input error.
+Many of errors were detectable during data cleaning, however it is possible that undetectable human errors exist in the data.
+
+Although this case study presents an opportunity to test the predictive power held within the organisations' existing CRM data, limitations exist which must be understood and built into the framework for predictions.
+These include missing data that could have been entered for each project, but historically has not been required, missing data that is very detailed and will never be available, and the potential for a small degree of human error to be embedded within the data.
+
+### 1.1.4 Research Motivation Summary
+
+Fixed price cost estimation for complex projects is a difficult task and when performed incorrectly, can have devastating affects on businesses and governments.
+Many theories have delved into human nature and our tendency to optimistically assess our capabilities with respect to a given task.
+In this case, the optimism often results in negative business consequences affecting employee morale and the capacity of a company to flourish and grow.
+Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project, by statistically comparing a new project to numerous similar projects and their characteristics.
+It has been shown that these models improve the cost estimation, however industry uptake has not been successful.
+It has proven to be a challenging task to persuade managers to use this so called reference class models in practise and remove themselves from additions based on intricate details of the completed project.
+This may be due to a number of reasons including lack of understanding, time, and lack of collaboration with industry in development of these models.
 
@@
 ```{r, echo=FALSE, include=FALSE}
+library('magrittr',lib='C:/Progra~1/R/R-3.2.2/library')
 
-read.csv(all8c,'C:/Users/n9232371/Documents/Consultbusiness/data/all8c.csv' )
+all8c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8c.csv' )
 
 vars<- names(all8c)
-vars<- vars[-c("Business", "JD_Second", "code_client")]
+vars<- vars[!vars %in% c("Business", "JD_Second", "code_client")]
 

### 2015-11-09 financial.rmd
save early nov

@@
 library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('randomForest', lib = 'C:/Progra~1/R/R-3.2.2/library')
 library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')
+library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
 
@@
 #without algorithm, balance would be $1,475,642
-sum(prof$balance)
+orig.bal<- sum(prof$balance)
 
 # without any losing jobs, balance would be $2,516,246
-prof %>% filter(balance > 0) %>% sum
+orig.prof<- prof %>% filter(balance > 0) %>% sum
 
 # without any profitable jobs, balance would be -$1,040,322
-prof %>% filter(balance < 0) %>% sum
+orig.loss<- prof %>% filter(balance < 0) %>% sum
 
@@
 ```
+
+Write function for making 10 boosted trees and checking the predictions/savings for each!
+
+```{r, echo=FALSE}
+
+all10mice<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')[,-1]
+
+profit.cv<- function(df.string = "all10mice", method = "log", seed = 300, k.folds = 8){
+
+        results = data.frame('orig.bal' = rep(0, k.folds), 'orig.prof' = rep(0, k.folds), 
+                             'orig.loss' = rep(0, k.folds),
+                             'prof.dismis' = rep(0, k.folds), 'loss.dismis' = rep(0, k.folds), 
+                             'algor.prof' = rep(0, k.folds),
+                             'improv' = rep(0, k.folds), 'thresh' = rep(0, k.folds),
+                             'auc' = rep(0, k.folds))
+        colours <- c("red", "yellow", "green", "blue", "purple")
+
+        df = get(df.string)
+        set.seed(seed)
+
+        # create f.rpdol
+        df$f.rpdol <- as.factor(df$b.rpdol)
+        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <-"profit"
+        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <-"loss"
+
+        #set up 10 fold cv using caret function
+        folds <- createFolds(df$f.rpdol, k = k.folds, list = T, returnTrain = F)
+        #now have folds[[1]] through folds[[10]] list of rows to exclude
+
+        if(method %in% 'boost') {
+                        formula = "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
+        } else {
+                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
+        }
+
+        for (j in 1:length(folds)) {
+
+                        #turn predictors and response into 'ordered' variables for roc() to work
+                        test<- df[folds[[j]],]
+                        levels(test$f.rpdol) <- c('profit', 'loss')
+                        test$f.rpdol <- ordered(test$f.rpdol)
+
+                        if(method %in% 'rf'){
+                                fit <- randomForest(as.formula(formula), data = df[-folds[[j]],], 
+                                            mtry = 5, ntree = 2000)
+                                pred <- predict(fit, test, type = "prob")
+                                pred <- pred[,2]
+
+                        } else {
+                                if(method %in% 'log'){
+                                        fit<- glm(as.formula(formula), 
+                                                  data = df[-folds[[j]],], family = binomial())
+                                        pred <- predict(fit, test,
+                                                        type = "response")
+                                } else {
+                                        fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
+                                                   distribution = "bernoulli", n.trees = 10000,
+                                                   shrinkage = 0.001, interaction.depth = 4,
+                                                   n.minobsinnode = 20)
+
+                                                pred <-predict(fit, test, 
+                                                               n.tree = 10000, type = "response")
+                                                }
+                        }
+
+                        #pROC
+                        pred.p <- roc(test$f.rpdol, pred)
+
+                        results$auc[j] <- auc(pred.p)
+
+                        ##
+                        #create full test df
+                        test$pred<- pred
+                        test.pred <- merge(test, all7d %>% select(mlsto, balance.mlsto), by = "mlsto")
+
+                        #create vector of profit points
+
+                        threshold = seq(0,1, by = .05)
+                        fin.prof = rep(0, length(threshold))
+
+                        for (i in 1:length(threshold)) {
+                                prof <- data.frame("balance" = test.pred$balance.mlsto,
+                                "pred" = test.pred$pred,
+                                "ans" = test.pred$b.rpdol)
+
+                                #pred = probabilities
+                                prof$conf <- rep(0, nrow(prof))
+                                prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 1, 'TP', prof$conf)
+                                prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
+                                prof$profit <- rep(0, nrow(prof))
+                                prof$profit <- ifelse(prof$conf %in% "TP",-prof$balance, prof$profit)
+                                prof$profit <- ifelse(prof$conf %in% "FP",-prof$balance, prof$profit)
+
+                                fin.prof[i] = sum(prof$profit)
+
+                                }
+
+                        prof.plot<- data.frame("threshold" = threshold, "final.profit" = fin.prof)
+
+                        print(
+                        ggplot(prof.plot, aes(x=threshold, y=final.profit)) + geom_point() + geom_line()
+                        )
+
+                        best.thresh = arrange(prof.plot, -final.profit)[1,1]
+                        results$thresh[j] = best.thresh
+
+                        ####
+                        # SUMMARY STATS
+                        ####
+
+                        #without algorithm, balance would be $1,475,642
+                        results$orig.bal[j] = sum(prof$balance)
+                        # without any losing jobs, balance would be $2,516,246
+                        results$orig.prof[j] = prof %>% filter(balance > 0) %>% sum
+                        # without any profitable jobs, balance would be -$1,040,322
+                        results$orig.loss[j] = prof %>% filter(balance < 0) %>% sum
+                        #how much real profit would you be dismissing? $82631
+                        temp<- prof %>% filter(pred >= best.thresh & balance > 0)
+                        results$prof.dismis[j] = ifelse(nrow(temp)>0, temp %>% select(balance) %>% sum,
+                                                        0)
+                        #how much real loss would you be avoiding? $566,481.6
+                        results$loss.dismis[j] = prof %>% filter(pred >= best.thresh & balance < 0) %>% 
+                                select(balance) %>% sum
+
+                        #print j value
+                        cat(j, " ")
+
+        }
+
+        # total benefit = $483,850
+        results = transform(results, algor.prof = -(loss.dismis + prof.dismis))
+        results = transform(results, improv = algor.prof/orig.bal)
+
+        return(results)
+
+}
+
+results.1 <- profit.cv(df.string = "all10mice", method = "log", seed = 400, k.folds = 8)
+results.2 <- profit.cv(df.string = "all10mice", method = "boost", seed = 400, k.folds = 8)
+results.3 <- profit.cv(df.string = "all10mice", method = "rf", seed = 300, k.folds = 3)
+
+```

### 2015-11-09 rf_log_naiive_auc.rmd
save early nov

@@
                 folds <- createFolds(df$f.rpdol, k = 10, list = T, returnTrain = F)
-                #now have folds[[1]] through folds[[5]] list of rows to exclude
+                #now have folds[[1]] through folds[[10]] list of rows to exclude
 

### 2015-11-10 finalwriteup/draft.rmd
small update

@@
 This may be due to a number of reasons including lack of understanding, time, and lack of collaboration with industry in development of these models.
+This research examines the potential for a company's internally generated CRM data to be used in creating a reference class forecasting model.
+This has the potential to influence the managers' decisions in the company more however its limitations must be understood in order to create a good model.
 
-## Thesis aims
+## 2.0 Thesis aims
 
@@
 
-**General Aim** The aim of this project is to use statistical techniques to model decisions for consulting businesses.
-Research will focus on a case study Engineering consulting company that sells their expert advice (in the currency of time) to business clients.
-The business decisions include predicting project fees before project engagement and client analysis using existing invoicing and hourly employee-cost data.
+**General Aim** The aim of this project is to use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
+Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
+The project outcomes are intended to assist the business in  predicting project costs before project engagement and other business analyses such as client analysis.
 Several statistical and machine learning techniques will be tested, compared and refined.
@@
 **Hypothesis 1**
-k-means clustering methods can identify groups of similar projects with significant differences in profitability.
-
-Unsupervised learning to identify clusters is a successful marketing technique used by retail companies to target groups of similar customers with similar habits and preferences.
-In this case of consulting companies, clusters may have multiple uses.
-New projects may have characteristics that slot into an 'unprofitable' cluster and the decision is clear to not engage in the project.
-Alternatively, if past projects and accompanying clients are found to fall into a very profitable cluster, consulting managers may focus services on these clients and reach out to these clients in particular for new projects.
+A statistical or machine learning model based on historical project data can predict the profitability of a new project.
 
 **Hypothesis 2**
-A statistical or machine learning model based on historical project data can predict the profitability of a new project with sufficient confidence to influence a project manager's fee proposal.
+The predictive model built from Hypothesis 1 can be proven to have a positive impact on the bottom line of the case study business.
 
@@
 
-This chapter provides an overview of the literature available about current business to business (B2B) decision strategy as well as methods in machine learning and statistical business prediction models.
-In particular, the literature on B2B decision strategy is centred around setting fixed consulting fees and methods of targeting new clients or projects.
+This chapter provides an overview of the literature available about current business to business (B2B) cost estimation methods, use cases of CRM data, as well as machine learning algorithms and statistical business prediction models.
+In particular, the literature on B2B decision strategy is centred around setting fixed consulting fees and methods of managing effective consulting projects.
 The most prominent statistical and machine-learning models developed for business applications shall be reviewed along with models applied to the most closely related business cases to this project.
 
-## 2.1 Business to Business Decision Strategy
+## 2.1 Fixed Cost Estimation Methods
 
-### 2.1.1 Client and project analysis
+### 2.1.1 Contract Structures
 
-* clustering
-* how do businesses currently review their projects or clients
+* research into contract structures to prevent loss
+* benefits of flexible type contracts
+* Gap: mobilising when to implement flexible contracts - takes away from client relationship but jobs that make losses are pointless and eventually deteoriate relationships
 
-### 2.1.2 Fixed fees
+### 2.1.2 Previous work on Regression and Neural Networks?
 
-* fixed fee problem in the IT consulting world - developed new techniques?
 * managers currently go by gut feeling or look up a couple of past examples
-* difficult to combine insights from similar jobs/clients
+* difficult to combine insights from similar jobs/clients, so complex
 
@@
 
-## 2.2 Statistical and machine learning methods applied to business problems
+## 2.2 Use cases of CRM data
 
-### 2.2.1 Most simple and frequent models
+* no idea
+* gap - my project obviously
+
+## 2.3 Statistical and machine learning methods applied to business problems
+
+### 2.3.1 Most simple and frequent models
 
@@
 
-### 2.2.2 Complex models
+### 2.3.2 Complex models
 
@@
 
-### 2.2.3 ensemble methods
+### 2.3.3 ensemble methods
 
@@
 
-### 2.2.4 similar applications
+### 2.3.4 similar applications
 
@@
 
-## 2.3 Conclusion
+## 2.4 Conclusion
 

### 2015-11-20 finalwriteup/draft.rmd
final update

@@
 the nature of consulting jobs is quite complex and historically, not all information about a project is possible to obtain.
-Variables with missing data are alternatively treated as optional bonus variables in a multi step model as opposed to missing information which must be imputed or deletion of the entire case that contains missing information.
+Variables with missing data are alternatively treated as optional bonus variables in a multi-step model as opposed to missing information which must be imputed or deletion of the entire case that contains missing information.
 
@@
 This chapter provides an overview of the literature available about current business to business (B2B) cost estimation methods, use cases of CRM data, as well as machine learning algorithms and statistical business prediction models.
-In particular, the literature on B2B decision strategy is centred around setting fixed consulting fees and methods of managing effective consulting projects.
 The most prominent statistical and machine-learning models developed for business applications shall be reviewed along with models applied to the most closely related business cases to this project.
+Gaps in past research will also be highlighted.
 
@@
 
-### 2.1.1 Contract Structures
+* Previous methods to address problem of cost estimation
 
-* research into contract structures to prevent loss
-* benefits of flexible type contracts
-* Gap: mobilising when to implement flexible contracts - takes away from client relationship but jobs that make losses are pointless and eventually deteoriate relationships
+    *
 
@@
 * Gap
+
+**
 

### 2015-11-20 financial.rmd
final update

@@
 
-profit.cv<- function(df.string = "all10mice", method = "log", seed = 300, k.folds = 8){
+profit.cv<- function(df.string = "all10mice", method = "boost", seed = 300, k.folds = 3){
 
@@
                         pred.p <- roc(test$f.rpdol, pred)
+                        # plot.roc(pred.p, col= colours[i], print.thres=T)
 

### 2015-11-20 prob_review.rmd
final update

--- 2015-11-20 financial.rmd
@@
 ---
-title: "Financial_Analysis"
+title: "result_review.rmd"
 author: "Amy Cook"
-date: "October 21, 2015"
+date: "November 19, 2015"
 output: word_document
 ---
-
-#Assess financial implications for a predictive model
-
-We typically get a confusion matrix built from probabilities
 
@@
 ```{r, echo=FALSE}
-test.pred <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/test_review.csv')
+
 all7d <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')
-test.pred<- merge(test.pred, all7d %>% select(mlsto, balance.mlsto), by = "mlsto")
 
@@
 
-Create vector of profit points
-
-```{r, echo=FALSE}
-
-threshold<- seq(0,1, by=.05)
-fin.prof<- rep(0, length(threshold))
-
-for(i in 1:length(threshold)){
-        prof<- data.frame("balance" = test.pred$balance.mlsto, 
-                  "pred" = test.pred$pred,
-                  "ans" = test.pred$b.rpdol)
-        #pred = probabilities
-        prof$conf <- rep(0, nrow(prof))
-        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 1, 'TP', prof$conf)
-        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
-        prof$profit<- rep(0, nrow(prof))
-        prof$profit<- ifelse(prof$conf %in% "TP", -prof$balance, prof$profit)
-        prof$profit<- ifelse(prof$conf %in% "FP", -prof$balance, prof$profit)
-
-        fin.prof[i] = sum(prof$profit)
-
-}
-
-prof.plot<- data.frame("threshold" = threshold, "final.profit" = fin.prof)
-
-ggplot(prof.plot, aes(x=threshold, y=final.profit)) + geom_point() + geom_line()
-
-# if say no to all jobs with higher than prob of 0.5, save $500,000
-
-head(prof)
-
-####
-# SUMMARY STATS
-####
-
-# would dismiss 56 jobs
-prof %>% filter(pred>=0.5) %>% nrow
-
-# would dismiss 19 profitable jobs
-prof %>% filter(pred >= 0.5 & balance > 0) %>% nrow
-#out of 607
-prof %>% filter(balance > 0) %>% nrow
-
-# would dismiss 37 losing jobs
-prof %>% filter(pred >= 0.5 & balance < 0) %>% nrow
-#out of 135
-prof %>% filter(balance < 0) %>% nrow
-
-#without algorithm, balance would be $1,475,642
-orig.bal<- sum(prof$balance)
-
-# without any losing jobs, balance would be $2,516,246
-orig.prof<- prof %>% filter(balance > 0) %>% sum
-
-# without any profitable jobs, balance would be -$1,040,322
-orig.loss<- prof %>% filter(balance < 0) %>% sum
-
-#how much real profit would you be dismissing? $82631
-prof.dismis<- prof %>% filter(pred >= 0.5 & balance > 0) %>% select(balance) %>% sum
-#how much real loss would you be avoiding? $566,481.6
-loss.dismis<- prof %>% filter(pred >= 0.5 & balance < 0) %>% select(balance) %>% sum
-# total benefit = $483,850
-algor.prof<- -(loss.dismis + prof.dismis)
-
-# total balance would be $1,959,492
-sum(prof$balance) + algor.prof
-
-# a 32.8% improvement
-algor.prof/sum(prof$balance)
-
-```
+Assemble full data set with predictions
 
@@
 
-profit.cv<- function(df.string = "all10mice", method = "boost", seed = 300, k.folds = 3){
+predict.all<- function(df.string = "all10mice", method = "boost", seed = 300, k.folds = 3){
 
-        results = data.frame('orig.bal' = rep(0, k.folds), 'orig.prof' = rep(0, k.folds), 
-                             'orig.loss' = rep(0, k.folds),
-                             'prof.dismis' = rep(0, k.folds), 'loss.dismis' = rep(0, k.folds), 
-                             'algor.prof' = rep(0, k.folds),
-                             'improv' = rep(0, k.folds), 'thresh' = rep(0, k.folds),
-                             'auc' = rep(0, k.folds))
-        colours <- c("red", "yellow", "green", "blue", "purple")
+        #create dataset for final results to tag onto
+        all10full = NULL
 
@@
 
-                        results$auc[j] <- auc(pred.p)
+                        auc(pred.p)
 
@@
 
-                        #create vector of profit points
-
-                        threshold = seq(0,1, by = .05)
-                        fin.prof = rep(0, length(threshold))
-
-                        for (i in 1:length(threshold)) {
-                                prof <- data.frame("balance" = test.pred$balance.mlsto,
-                                "pred" = test.pred$pred,
-                                "ans" = test.pred$b.rpdol)
-
-                                #pred = probabilities
-                                prof$conf <- rep(0, nrow(prof))
-                                prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 1, 'TP', prof$conf)
-                                prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
-                                prof$profit <- rep(0, nrow(prof))
-                                prof$profit <- ifelse(prof$conf %in% "TP",-prof$balance, prof$profit)
-                                prof$profit <- ifelse(prof$conf %in% "FP",-prof$balance, prof$profit)
-
-                                fin.prof[i] = sum(prof$profit)
-
-                                }
-
-                        prof.plot<- data.frame("threshold" = threshold, "final.profit" = fin.prof)
-
-                        print(
-                        ggplot(prof.plot, aes(x=threshold, y=final.profit)) + geom_point() + geom_line()
-                        )
-
-                        best.thresh = arrange(prof.plot, -final.profit)[1,1]
-                        results$thresh[j] = best.thresh
-
-                        ####
-                        # SUMMARY STATS
-                        ####
-
-                        #without algorithm, balance would be $1,475,642
-                        results$orig.bal[j] = sum(prof$balance)
-                        # without any losing jobs, balance would be $2,516,246
-                        results$orig.prof[j] = prof %>% filter(balance > 0) %>% sum
-                        # without any profitable jobs, balance would be -$1,040,322
-                        results$orig.loss[j] = prof %>% filter(balance < 0) %>% sum
-                        #how much real profit would you be dismissing? $82631
-                        temp<- prof %>% filter(pred >= best.thresh & balance > 0)
-                        results$prof.dismis[j] = ifelse(nrow(temp)>0, temp %>% select(balance) %>% sum,
-                                                        0)
-                        #how much real loss would you be avoiding? $566,481.6
-                        results$loss.dismis[j] = prof %>% filter(pred >= best.thresh & balance < 0) %>% 
-                                select(balance) %>% sum
+                        all10full <- rbind(all10full, test.pred)
 
@@
 
-        # total benefit = $483,850
-        results = transform(results, algor.prof = -(loss.dismis + prof.dismis))
-        results = transform(results, improv = algor.prof/orig.bal)
-
-        return(results)
+        return(all10full)
 
@@
 
-results.1 <- profit.cv(df.string = "all10mice", method = "log", seed = 400, k.folds = 8)
-results.2 <- profit.cv(df.string = "all10mice", method = "boost", seed = 400, k.folds = 8)
-results.3 <- profit.cv(df.string = "all10mice", method = "rf", seed = 300, k.folds = 3)
+results.1 <- predict.all(df.string = "all10mice", method = "boost", seed = 400, k.folds = 8)
 

### 2015-11-23 finalwriteup/draft.rmd
monday

@@
 
-## 2.1 Fixed Cost Estimation Methods
+## 2.1 Project Cost Estimation Methods
 
-* Previous methods to address problem of cost estimation
+This project will tackle the problem of estimating fixed fees for complex consulting projects.
+It is therefore imperitive to understand what research has been performed to date and how this project contributes to the body of work.
+The task of estimating fixed fees can be likened to the terms 'cost estimation' or 'effort estimation' in the literature.
+The bulk of research to date has been performed with project data in either the construction industry or software development.
+There is also a body of research on predicting returns on movies in the film industry.
+The research into cost estimation in each of these three industries is reviewed below, detailing different methods and outcomes.
+Finally a gap in the literature will be highlighted which this research aims to fill.
 
-    *
+### 2.1.1 Cost Estimation in the Construction Industry
+
+Research into cost estimation in the construction industry chiefly focuses on predicting the final building cost of construction.
+Various studies have surveyed popular methods for cost estimation while others have  investigated new methods such as building information modelling (BIM) analyisis, case based reasoning, and predictive algorithms.
+The cost estimation methods of detailed analysis (inside view), case based reasoning, and predictive algorithms shall be reviewed.
+
+**Detailed Analysis**
+
+Detailed analysis refers to the process of an engineer or builder in the construction team carefully reviewing construction drawings and sum the cost of materials, labour, machinery hire, overheads, profit etc.
+[@Akintoye2000].
+@Akintoye2000's study based in the UK found that this is the most popular method for pricing amongst contractors.
+The second most popular method was to perform an additional manual comparison with similar projects and make judgments based on personal experience.
+Note that this does not involve a mathematical model.
+
+@Elfaki2014 discusses how much detailed cost estimates for cosntruction projects can vary from estimator (engineer) to estimator, which contributes to the lack of accuracy in predciting final cost of a project.
+He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and therefore estimate is prone to subjectivity.
+@Shane2009 theorise that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues.
+It is therefore almost impossible to achieve an accurate cost prediction manually [@Elfaki2014].
+
+Groundbreaking methods into calculating construction costs by @Ma2014 mobilise the data from building information modelling, although this method is still technically a detailed analysis of the costs.
+The idea is to quantify the cost of building a structure directly from a three dimensional model created by the design consultants such as engineers and architects.
+@Ma2014's first trial was to automate the cost calculation of structure composed of reinforced concrete only.
+His chief development was to program his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
+From the construction technique and material, a cost can be determined for each element.
+This was successful and Ma aim's to further his work so that more construction information can be intelligently obtained from BIM models.
+A system like Ma's would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could be prone to innaccurate estimates in a similar way to a detailed analysis by a person.
+The detailed view does not 'step back' and take into account set backs in construction that may have affected similar projects in the past.
+There is potential for this to be built into the same system but this is still in early stages of development.
+
+Survey information on the latest techniques for cost estimation is limited, however what is available indicates that cost estimation through detailed analysis (inside view) remains the most prevalent method.
+This is despite construction having a long history of projects that have run over time and over budget using the same cost estimation technique [@Shane2009].
+There is even current research on how to automate this way of cost estimation using BIM.
+The method of detailed sums shall be compared to the results of studies using parametric algorithms and case based reasoning ('outside view') in the following sections.
+
+**Statistical and Machine Learning Models**
+
+As the problem of cost estimation in construction projects is well recognised, dozens of studies have modelled the financial performance of these projects using machine learning and statistical models.
+The aim is to test whether a mathematical comparison of many similar projects, using meaningful descriptive variables, can predict the cost of construction.
+This section will first review the previous studies and most popular algorithms and finally summarise industry uptake of the methods.
+
+@Elfaki2014 completed a study on cost estimation techniques and research from 2004 to 2014.
+It was found that artificial neural networks and support vector machines were the most common machine learning techniques used.
+He highlighted that machine learning techniques deal well with uncertainty, however lack technical justification.
+In a way, this fulfils the purpose of the 'outside view', which is to step back from the detailed analysis and perform a comparison with similar projects.
+
+@Shin2015 was the first
+
+parametric estimating is not popular with contractors. this could be because:
+The reasons include: (a) lack of familiarity with the techniques; (b) the degree of sophistication involved in the techniques is unwarranted for project performance; (c) time plus lack of information and knowledge; (d) doubts whether these techniques are applicable to the construction industry; (e) most construction projects are seldom large enough to warrant the use of these techniques or research into them; (f) they require availability of sound data to ensure con®dence;
+
+Recent developments in cost esti- mating methods and tools that consider risks and variability in cost estimates, such as the use of range estimating and parametric estimating techniques, have not been adopted by contractors.
+
+At the conclusion of the papers, few compare the performance against the original method????
+
+* case based reasoning
 

### 2015-12-16 finalwriteup/draft.rmd
thesis update

@@
 The bulk of research to date has been performed with project data in either the construction industry or software development.
-There is also a body of research on predicting returns on movies in the film industry.
-The research into cost estimation in each of these three industries is reviewed below, detailing different methods and outcomes.
+The research into cost estimation in each of these industries is reviewed below, detailing different methods and outcomes.
 Finally a gap in the literature will be highlighted which this research aims to fill.
@@
 
-Detailed analysis refers to the process of an engineer or builder in the construction team carefully reviewing construction drawings and sum the cost of materials, labour, machinery hire, overheads, profit etc.
+Detailed analysis refers to the process of an engineer or builder in the construction team carefully reviewing construction drawings to sum the cost of materials, labour, machinery hire, overheads, profit etc.
 [@Akintoye2000].
@@
 @Elfaki2014 discusses how much detailed cost estimates for cosntruction projects can vary from estimator (engineer) to estimator, which contributes to the lack of accuracy in predciting final cost of a project.
-He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and therefore estimate is prone to subjectivity.
-@Shane2009 theorise that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues.
-It is therefore almost impossible to achieve an accurate cost prediction manually [@Elfaki2014].
+He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and, in turn, their estimate is prone to subjectivity.
+@Shane2009 theorises that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues that it is almost impossible to achieve an accurate cost prediction manually.
 
 Groundbreaking methods into calculating construction costs by @Ma2014 mobilise the data from building information modelling, although this method is still technically a detailed analysis of the costs.
-The idea is to quantify the cost of building a structure directly from a three dimensional model created by the design consultants such as engineers and architects.
-@Ma2014's first trial was to automate the cost calculation of structure composed of reinforced concrete only.
+The idea is to quantify the cost of building a structure directly from a three dimensional model created by the design consultants (engineers and architects).
+@Ma2014's first trial was to automate the cost calculation of a reinforced concrete structure.
 His chief development was to program his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
@@
 
-Survey information on the latest techniques for cost estimation is limited, however what is available indicates that cost estimation through detailed analysis (inside view) remains the most prevalent method.
-This is despite construction having a long history of projects that have run over time and over budget using the same cost estimation technique [@Shane2009].
+Surveyed information on the latest techniques for cost estimation is limited, however what is available indicates that cost estimation through detailed analysis (inside view) remains the most prevalent method.
+This is despite construction having a long history of projects running over time and over budget using the same cost estimation technique [@Shane2009].
 There is even current research on how to automate this way of cost estimation using BIM.
@@
 
-As the problem of cost estimation in construction projects is well recognised, dozens of studies have modelled the financial performance of these projects using machine learning and statistical models.
+As the problem of cost estimation in construction projects is well recognised, dozens of studies have built prediction models of these projects using machine learning and statistical models.
 The aim is to test whether a mathematical comparison of many similar projects, using meaningful descriptive variables, can predict the cost of construction.
-This section will first review the previous studies and most popular algorithms and finally summarise industry uptake of the methods.
+This section will first review the previous studies and most accurate algorithms and finally summarise industry uptake of the methods.
 
@@
 It was found that artificial neural networks and support vector machines were the most common machine learning techniques used.
-He highlighted that machine learning techniques deal well with uncertainty, however lack technical justification.
-In a way, this fulfils the purpose of the 'outside view', which is to step back from the detailed analysis and perform a comparison with similar projects.
+He highlighted that these two machine learning techniques deal well with uncertainty, however lack technical justification to the decision maker (known as a black box).
+Neural networks are also time consuming to train, and must be re-trained and re-tested with each additional piece of data [@Kim2004].
+Nevertheless, neural networks and SVM's recieved a significant amount of attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
 
-@Shin2015 was the first
+@Shin2015 pioneered the application of boosted trees to cost estimation in construction projects.
+This is surprising given the rapid uptake of boosted trees over the past decade.
+In the study, data from 234 school building construction projects in Korea was used.
+Boosted trees were compared to neural networks, which are one of the most popular algorithms to apply to the problem.
+Boosted trees were found to predict costs more slightly accurately than neural networks, but not significantly (p value < 0.5).
+Even though the two algorithms performed similarly, the output from boosted tree models provide insight into the structure of the model such as variable importance and variable partial dependence plots.
+This additional insight may mean boosted trees are preferable over neural networks because of their potential to engage decision makers.
 
-parametric estimating is not popular with contractors. this could be because:
-The reasons include: (a) lack of familiarity with the techniques; (b) the degree of sophistication involved in the techniques is unwarranted for project performance; (c) time plus lack of information and knowledge; (d) doubts whether these techniques are applicable to the construction industry; (e) most construction projects are seldom large enough to warrant the use of these techniques or research into them; (f) they require availability of sound data to ensure con®dence;
+A notable amount of literature has studied the predictive power of multiple linear regression in the construction cost estimation problem.
+Often, regression is the only model assessed, without comparison to other methods such as neural networks, which first started appearing in literature in the 1990's [@Kim2004].
+However, even post-1990, many studies focused on regression only.
+This may be because the technique is straightforward, easy to use and widely available in statistical packages [@Chan2005].
+Out of those that have compared regression to neural networks, the studies have shown that neural networks outperform regression, however other studies show they are approximately equal [@Kim2004][@Attalla2003].
+@Dissanayaka1999 found that neural networks outperformed multiple linear regression models, however regression models could be performed first to determine variable importance and condense the number of variables.
+It can be argued that boosted trees could perform as well as neural networks while providing additional insights (even more so than regression models) such as variable importance and partial dependence of variables against cost.
 
-Recent developments in cost esti- mating methods and tools that consider risks and variability in cost estimates, such as the use of range estimating and parametric estimating techniques, have not been adopted by contractors.
+Despite neural networks typically outperforming regression models, it is intersting to note some of the outcomes of regression studies in cost estimation.
+Different studies tend to produce a wide range of variables that contribute to different degrees in the models.
+This is likey due to the nature of construction projects which are multidisciplinary and involve a range of different parties such as the clients, consultants, contractors and suppliers, each of which can have a range of characteristics and degrees of input [@Chan2005].
+@Chan2005's study on 87 building projects in Singapore included special, complex projects.
+Therefore, variables such as contractor's specialised skills, who the client was (public vs private), the client level of experience, and the contractor's financial management highly influenced final cost of the project.
+Other studies found variables such as project complexity, duration, team experience, information availablility, site requirements, and labour climate to contribute most [@Akintoye2000][@Trost2003][@Pinto1988].
+As these studies obtained information from questionaires, the available variables also largely depended on the questions in the questionaires, and therefore each study produces a range of different important variables.
+Despite the differences in qualitative outcomes, the studies all generally report their mean absolute error or a similar metric from the tested models.
+This is encouraging, however there is a gap in the literature in how to translate the improved errors from the model into an argument for industry to uptake the model as an improvement on their bottom line.
 
-At the conclusion of the papers, few compare the performance against the original method????
+Surveys into construction industry uptake of these algorithmic models are sparse, however a survey by @Akintoye2000 showed that parametric estimating methods have not been adopted by contractors, despite the research into this area over 20 years prior.
+The majority of contractors use experience based methods.
+@Akintoye2000 hypothesises that this could be due to a lack of familiarity and knowledge about the techniques, doubts whether these techniques are applicable to the construction industry, and the availability of sound data to ensure confidence.
+This presents a clear gap in the literature on how to analyse the models in terms of the goals and limitations that a contractor may experience.
 
-* case based reasoning
+The literature on cost estimation for building projects do not break down the cost estimation task into smaller components, even though a construction project is completed by dozens of consultants and sub-contractors.
+Many of the consultants, such as Engineers and Architects, produce their own cost estimates for their effort that contributes to the project.
+Their contracts are often fixed price, with minimal room for movement, which means they wear the risk of over-spending to meet fixed cost contracts [@Harris1999].
+Furthermore, the existing studies collect data from a wide range of projects that have certain characteristics in common, such as country, or building type and information is obtained through surveys.
+This leaves a gap in research exploring a single contractor's history of projects, which would be available in their company database.
 
-### 2.1.2 Previous work on Regression and Neural Networks?
+**case based reasoning**
 
-* managers currently go by gut feeling or look up a couple of past examples
-* difficult to combine insights from similar jobs/clients, so complex
+Case based reasoning (CBR) is a method of estimating by using the results of similar cases.
+The idea is that in the case of a new project, similar projects are chosen, before generalisations about the data are made.
+This contrasts to algorithmic models that are built from the entire set of data [@Elfaki2014].
+Research into the application of case based reasoning to construction cost estimation began in the 1980's [@Kim2004].
+The general procedure involves:
 
-### 2.1.3 Gap
-* no mobilisation of CRM data into sophisticated models in B2B world
+1. Storing a collection of projects with key variable values
+2.
+Once a new project or case arrives, similar cases are retrieved.
+This can be achieved either algorithmically with a distance function applied to the variables, or manually by users reviewing past cases.
+The nearest neighbour algorithm has been used to find the distance between cases in previous studies.
+It is programmed to calculate the euclidean distance in n-dimensional space between cases, where each variable is a dimension [@Kumar2007].
+3.
+The cost of the new project is estimated by extrapolating characteristics of the similar cases to the new case.
+This can also be done either algorithmically or manually by an experienced decision maker.
 
-## 2.2 Use cases of CRM data
+A study by @Kim2004 compared CBR to neural networks and regression in the construction cost estimation problem.
+For CBR, the ESTEEM software package was used which algorithmically calculated the similarity of variables, weighted the variables using gradient descent, and deduced the cost using the most similar cases (the specific method for this last step was not explained).
+The discussion compared mean absolute error rates of the three methods and showed that CBR was more accurate than regression, but less accurate than neural networks in cost estimation.
+Despite neural networks outperforming CBR, CBR maintained significant advantages over neural networks which have been referenced in other studies as well.
 
-* no idea
-* gap - my project obviously
+Advantages of CBR include algorithmic efficiency as well as user engagement.
+In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to neural networks, which must be slowly re-trained and re-tuned.
+New cases also did not need to have every variable complete which can occur often in real, messy data.
+Neural networks on the other hand require complete data.
+Another significant advantage of CBR is the ability for users to review chosen similar cases and make sense of the predicted cost, as opposed to neural networks which are a black box [@Kim2004][@Elfaki2014][@Kumar2007].
+A disadvantage of CBR is the accuracy can be highly dependent on the number of selected cases [@Elfaki2014].
+Overall, CBR may provide prediction accuracies between regression and neural networks but its results can be justified by the user - a valuable asset.
+
+CBR can be described as a systematic method of expert judgment, where the decision maker manually compares similar projects from his or her experience [@Shepperd1996].
+CBR that is performed manually through personal experience, without the assistance of algorithms, was the second most popular method for construction estimation in @Akintoye2000's survey in the UK.
+This indicates it is quite an intuitive method for decision maker's to cross check their detailed cost estimate, although an algorithmic version of CBR has not been taken up by industry.
+
+**summary**
+
+In summary, research into cost estimation in the construction industry has been undertaken using a variety of methods ranging from detailed manual analysis, algorithmic models, and case based reasoning.
+Neural networks generally outperformed all other systematic methods, such as regression and case based reasoning, but had a significant disadvantage of being a black box.
+Alternative methods such as boosted trees have the potential to perform as well as neural networks and provide insight into the structure of the algorithm.
+This is true for CBR as well, although it has not been shown to predict as accurately as neural networks.
+There exists a gap in the research in ways to intelligently combine CBR with machine learning methods that could predict well, provide insight to the structure, and engage the user in reviewing similar projects.
+Also, as stated previously, there exists a gap in the application of cost estimating in the construction industry to smaller components of the project, such as consulting and contracting companies who each face a similar problem in estimating their fixed fee.
+These companies could mobilise information in their internal project databases, as opposed to available literature which surveys dozens of companies involved in selected construction projects.
+
+### 2.1.1 Effort Estimation in the IT Industry
+
+Like the construction industry, the software industry faces the challenging task of estimating the cost of a project before full engagement with their client.
+However, unlike construction projects, the major cost associated with the project is *effort* as opposed to the cost of building materials.
+A parallel can be drawn between software project effort estimation research and consultants effort in the construction industry, which was a gap in the research of the previous section.
+
+Effort estimation is indeed a problem, and was demonstrated by a review of 6 surveys of IT projects between 1984 and 1994 by @Molokken2003.
+It revealed that 60-80% of IT projects encounter effort and/or schedule overruns, where the average overrun was by 30 to 40%.
+A survey of software managers asked whether they saw effort estimation as a problem, and 91% responded 'yes', while only 9% answered 'no' [@Moores1992].
+The most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%), which falls squarely within the theory of optimism bias, or the 'inside view' [@Flyvbjerg2011].
+Given this widespread problem in the industry, plenty of research has been dedicated to the problem of effort prediction, with methods similar to the construction industry literature: detailsed analysis, algorithmic modelling, and case based reasoning.
+
+**Detailed Analysis**
+
+Similar to the construction industry, expert judgment or detailed analysis, is the most widely practised method for effort estimation [@Shepperd1996][@Molokken2003].
+This is despite many years of research being dedicated to developing algorithmic models and industry tools that, in the research context, outperform expert judgment.
+A survey by @Moores1992 found that software tools assisting a software project were in significant use, but only the project planning tools, not the cost estimating tools.
+This suggests a mental link for project managers to price jobs based on *details* of the software project as opposed to taking an outside view.
+Another study by @Heemstra1992 found that there was no evidence estimation accuracy improved when estimation tools were used.
+@Bergeron1992 similarly found jobs that used algorithmic models were actually associated with less accurate estimtes.
+However, this trend may be coincidental due to a lack of cases where estimation tools were used [@Molokken2003].
+The scarcity of real project evidence that algorithmic estimation tools improve estimations and an unintuitive mental jump could be reasons contributing to detailed expert analysis remaining the most widespread technique for effort estimation.
+
+**Algorithmic methods**
+
+Algorithmic methods to predict effort for IT project costs has been a popular subject of research.
+Regression models have been the most substantially tested followed by neural networks [@Finnie1997].
+This section will assess the details and performance of the algorithmic models, important variables, applications and finally gaps in the literature.
+
+Multiple studies have shown that neural networks definitively outperform regression models in effort estimation, although regression is the most popular method in the literature [@Finnie1997][@Pai2013][@Matson1993].
+Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the models accuracy: size [@Shepperd1996][@Finnie1997][@Pai2013].
+Size can refer to the expected number of lines of code or function point.
+Function point refers to the amount of business functionality expected from an information system product.
+This is easier to guess correctly at the beginning of a project, and therefore leads to more accurate predictions [@Finnie1997].
+One study used a single variable approach (size) with a linear coefficient and an exponential coefficient as follows:
+
+$effort = \alpha*size^{\beta}$
+$\alpha = productivity coefficient$
+$\beta = economies of scale coefficient$
+$size = estimated lines of code$
+
+This model was compared to case based reasoning, and case based reasoning outperformed it [@Shepperd1996].
+
+There are disadvantages to algorithmic methods, similar to the construction industry.
+First, there are often not enough cases to create a good model in the software industry, particularly if the cases must be from within the company [@Finnie1997][@Pai2013].
+A study by @Mendes2004 demonstrated using 67 web projects that cross-company models were significantly less accurate than a within company model, so it is in the company's interest to create an in-house model.
+This means that effort estimation tools should use in-house data and must be adapted by statistical experts to each company [@Shepperd1996].
+
+Algorithmic estimation tools have not been successfully adopted by industry.
+This may be because no model has proved to be outstandingly successful at consistently predicting required effort [@Finnie1997].
+Also, neural networks (the most accurate algorithm) lack the ability to capably explain its results and instill trust in decision makers [@Finnie1997].
+
+The literature heavily focuses on effort estimation for the software industry.
+This leaves a gap in studies for effort estimation for consulting companies in other industries such as the construction industry.
+Construction projects do not have a simple 'project size' or function point variable and would need to utilise other variables available.
+
+**Case Based Reasoning**
+
+In the IT industry, applications of case based reasoning to effort estimation have been researched in a similar manner to the construction industry.
+The procedure is described as a algorithmicallyfinding similar projects to a new case using measurements of euclidean distance in n-dimensional space, where each dimension corresponds to a variable [@Shepperd1996].
+The similar cases are then used to predict effort for the new case, and in @Shepperd1996's study, linear regression was used for this final step.
+
+In contrast to the construction industry, the literature in software effort estimation found CBR to perform approximately equally to algorithmic methods such as neural networks [@Shepperd1996][@Finnie1997].
+It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects and provides insight into the prediction that neural networks do not [@Finnie1997].
+Research into CBR in the IT industry found similar results to the construction industry except that CBR performed as well as algorithmic methods.
+
+**Summary**
+
+Research into effort estimation in the software industry has followed a similar path to the cost prediction of construction projects.
+Expert judgment via detailed analysis is still the predominant method for effort estimation.
+Case based reasoning and algorithmic methods such as neural networks have found some success in effort estimation (at similar levels) but have not been successfully adopted by industry.
+It was noted that case based reasoning resonated more with decision makers because its effort prediction could be justified by the model.
+This visibility of the model prediction structure is very important for future work if the results are to be translated to industry.
+Effort estimation remains a problem, however agile methods of project management and effort management are changing the contractual approach of consulting in the software industry [@Badenfelt2011].
+This innovative method keeps delivery times short (and therefore predictable effort) and defines projects as something with a fixed cost and time.
+This means if a change occurs, other less important elements must be discarded maintain the fixed amount.
+Software development is a relatively new field in comparison to the construction industry and is able to adopt this new model relatively quickly across the industry.
+However, effort estimation for consulting companies in the construction industry are faced with a long history of tradition that is much harder to tackle.
+This leaves a gap in research in algorithimic models that predict effort for consulting companies in the construction industry.
+From previous literature, the models should allow decision makers to interrogate the structure of the prediction and be based on internal company data.
+
+## 2.2 Statistical and machine learning methods
+
+### 2.2.3 Introduction
+
+### 2.2.4 Linear Regression/logistic regression
+
+### 2.2.5 Naiive Bayes
+
+### 2.2.6 Ensemble Trees
+
+### 2.2.7 Bayesian Networks
+
+### 2.2.7 SVM's and Neural Networks
+
+### 2.2.8 Summary
 
@@
 ### 2.3.1 Most simple and frequent models
-
-* anova
-
-* linear regression
-    * pros
-
-    * cons
-        * do these acknowledge their data is not normally distributed?
 

### 2015-12-16 rf_log_naiive_auc.rmd
thesis update

@@
 fit.boost$var.names
-plot.gbm(fit.boost, 9, type="response")
+plot.gbm(fit.boost, 3, type="response")
 biz<- plot.gbm(fit.boost, 7, type="response")
@@
 pred.boost <- predict(fit.boost, test, n.tree = 10000, type = "response")
-pred.p <- roc(test$f.rpdol, pred.boost)
+pred.p <- roc(test$b.rpdol, pred.boost)
 

### 2015-12-22 finalwriteup/draft.rmd
more lit review

@@
 output: word_document
-bibliography: Library.bib
+bibliography: Library2.bib
 ---
@@
 
-### 2.2.3 Introduction
+### 2.2.1 Introduction
 
-### 2.2.4 Linear Regression/logistic regression
+The aim of this project is to use statistical techniques to model the profitability of projects.
+In order to optimise the model, several statistical and machine learning techniques will be tested, and compared and it is therefore prudent to review the range of statistical and machine learning techniques that have been successfully applied in the literature.
+These vary from simple methods such as Linear Regression to complex learning such as Neural Networks.
+The previous section highlighted the use of linear regression, neural networks, SVM and in one case boosted trees for prediction, however other business problems have utilised a wider range of methods.
+These include Naiive Bayes, Random Forests, and machine learned Bayesian Networks.
+The following section will present each method's advantages and disadvatanges and an example of a past successful application.
 
-### 2.2.5 Naiive Bayes
+### 2.2.2 Linear Regression
 
-### 2.2.6 Ensemble Trees
+Linear regression is one of the simplest and most popular statistical prediction methods [@Hastie2009].
+An equation for a line is learned, which is defined by a constant and a coefficient times the value of each variable.
+The coefficients are chosen to minimise the residual sum of squares (least squared) [@Hastie2009].
+The reason this method is popular is that it is simple, provide stable predictions, and can be adapted for categorical variables [@Seng2010].
+The integration of categorical variables is called Anova, and the algorithm compares group mean variances within the categorical variable.
 
-### 2.2.7 Bayesian Networks
+There are several drawbacks to linear regression.
+For example assumptions must be made about the structure of the data.
+Predictor varialbes are assumed to be linearly related to the response variable and the varialbes are normally distributed.
+Many real world phenomena do not correspond to these assumptions which can make it difficult, sometimes impossible, to produce useful results [@Hastie2009][@Putler2012].
+Furthermore, response variables must be continous numeric variables and if two or more predictor variables are correlated, the values of their coefficients are highly unstable.
+On the other hand, correlated variables do not impact the predictive accuracy of the model.
+A final drawback is that the variable coefficients cannot be used to compare variable importance unless the varialbes have all been standardised to one another [@Putler2012].
+This limits the amount of insight that can be drawn from a linear regression model.
 
-### 2.2.7 SVM's and Neural Networks
+Despite the limitations of linear regression models, they have widespread business applications due to their simplicity and ease of application.
+Regression is used to predict risk in the finance and insurance industries, predict who to target in marketing exercises, and predict consumption spending in the field of economics [@Harrell2013].
+Due to its simplicity and widespread application, linear regression is a good starting point for an analysis and benchmark for performance against complex models.
 
-### 2.2.8 Summary
+### 2.2.3 Logistic Regression
+
+The method behind linear regression can be adapted to predict a binary response variable (1 or 0).
+A predictive linear equation is computed so that the predictor variables are combined with coefficients to give an output between 0 and 1.
+This is then transformed into the log odds to mediate extreme variable input values.
+The result is a probability an input case will result in a 0 or 1 (i.e.
+success or failure) [@Moore1989].
+Similar pros and cons exist for logistic regression as linear regression, and it too can be used as a benchmark to compare other binary predictive models due to its speed and simplicity.
+Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survive/death [@Moore1989].
+
+### 2.2.4 Naive Bayes
+
+[@Provost2013]
+
+Naive Bayes prediction method makes conditional independence assumptions about the predictor variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
+Specifically, it is assumed that each variable is conditionally independent given the respose class alone, without considering other predictor variables.
+Mathematically this means that each variable's probability contribution can be simply calculated by looking at the probability of a response class given a predictor value by itself.
+The predictor variables probability estimates are combined for each possible response class, and the response class with the highest probability is chosen [@Provost2013].
+
+The advantages of this method are that the conditional independence assumption enables very fast calculations and predictions.
+The method can perform very well for real world tasks because the assumption of independence does not damage predictions significantly.
+This is because if multiple varialbes are related, the variables independently still direct the prediction in the correct direction.
+The correlated variables will double or triple their emphasis on the predicted response variable, however this does not necessarily impact classification accuracy as the final step is to simply choose the class with the *highest* probability [@Provost2013].
+This is fine for ranking.
+On the other hand, the output probabilities are not realisitic, and the values themselves should not be used [@Caruana2006].
+Another disadvantage is that in order to calculate the class probability from a numeric variable, a distribution must be assumed (often gaussian) which the numeric data may not fit into neatly [@Caruana2006].
+Finally, Naive Bayes classifers are known to not perform well in binary classification [@Caruana2006].
+
+A real world success story for Naive Bayes classifiers hinges on their ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
+Naive Bayes has been used in complex spam detection systems where new spam emails or toxic text themes can be quickly added to the filtering model.
+The Naive Bayes method provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
+
+### 2.2.5 Decision Trees/ Ensemble Trees
+
+There are several brands of basic decision tree algorithms including ID3, C4.5, CART, and CHAID with CART and C4.5 being the most popular (Kabra & Bichkar, 2011).
+Ravi Kumar and Ravi (2007) recommends the CART algorithm as it is capable of solving classification and regression problems (regression trees predict an outcome that is within a range of real numbers) whereas the remaining decision trees solve classification problems only.
+Overall, simple decision trees have the risk of over-fitting the data but have favourable characteristics such as providing rules that can act as warning flags in real-life scenarios.
+In addition their results do not depend heavily on the skill of the data analyst (Kabra & Bichkar, 2011)(Putler & Krider, 2012).
+
+Three examples of these ensemble decision tree methods are bagging, boosting, and random forests.
+Bagging creates multiple trees by sampling the training data set in different ways and then combining tree results (Breiman, 1996).
+The boosted decision tree approach is a more advanced version of bagging which resamples the data in order to build multiple trees.
+The trees are then weighted, where trees with weaker predictive performance are weighted more heavily (boosted).
+This allows the algorithm to resample and strengthen weak areas (Kotsiantis, 2013).
+Finally, the random forests approach has many similarities to bagging in that multiple trees are sampled from the same training data.
+However, when creating the tree for each sample dataset, a random subset of attributes is used to determine each split.
+This prevents the trees from become too correlated (Breiman, 2001a).
+Bayesian CART (BCART) models are another way of solving the over-fitting problem found in decision trees.
+These models use the prior (initial set of CART trees) to induce a posterior distribution that will guide the search for the correct CART model output.
+This includes a guide to the size and shape of the tree.
+The best CART models can be selected according to their posterior probabilities which indicate fit to the data (Chipman, George, McCulloch, Zhang & Al, 1998).
+
+### 2.2.5 Bayesian Networks
+
+### 2.2.6 SVM's and Neural Networks
+
+Neural networks were inspired by biological neural networks of the human nervous system and require a lot of training data.
+They are very flexible in the relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
+A disadvantage is that although the results are accurate, the algorithm is a black box meaning the data is input and the results output without providing the user insight into the methodology.
+. A notable disadvantage for neural networks is that they provide a black box prediction, meaning the data is input and the results output without providing the user insight into the methodology.
+
+### 2.2.7 Summary
 
@@
 
-### 2.3.1 Most simple and frequent models
+The aim of part two of this research project is to model strategic internal decisions within an Engineering consulting business using statistical and machine learning techniques.
+A large body of relevant research is available that addresses business profitability and prediction for the purpose of decision-making.
+The majority of the literature focuses on the development of prediction methods in the context of applied topics.
+Popular topics include predicting stock fluctuation, customer churn analysis, fraud prediction, customer classification, market segment analysis, product success prediction, and recommendation systems (Seng & Chen, 2010).
+This review will discuss the use and merits of statistical methods that were employed in the literature.
+First, simple statistical methods that were used for data exploration will be addressed, followed by advanced statistical and machine learning prediction techniques and why some methods were reported to perform better than others.
+Finally, a case study that applied customer-churn prediction techniques to internal employee churn will be discussed.
+
+### 2.3.1 Simple models
+
+* linear regression - example
+* clustering - example
+* naiive bayes - example
+* single decision trees?
 
@@
 
-* SVM, Neural networks, random forest
-* data must not be normal
-* most successful?
+Simple theoretical models such as linear regression are informative but often provide unsatisfactory accuracy for modelling real life data which is messy and does not necessarily align with the rules of statistical theory (Breiman, 2001b).
+Algorithmic machine-learning predictive models have developed rapidly in recent years because they adapt to the data and are not bound by theoretical models.
+These machine-learning models can consistently fit real-life data with more accuracy than theoretical models and have become popular (Breiman, 2001b).
+Ravi Kumar and Ravi (2007) performed a detailed review of statistical and machine learning techniques applied over 37 years in the context of bankruptcy prediction in banks.
+(financial business prediction).
+The most widely used model was neural networks, however logistic and linear regression, decision trees, support vector machines (SVM’s), discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
+Supporting Breiman’s (2001b) statements, it was found that DA and standard regression techniques were not preferred due to their low accuracy.
+The overall assessment was that SVM’s outperformed Neural networks (back propagation neural networks BPNNs were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
+Although SVM’s performed the best, they are often extremely complex and slow, requiring a high memory (Ravi Kumar & Ravi, 2007).
+The reason for this is that SVM’s create complicated curved 3D decision boundaries called kernels.
+However, statisticians experienced in predictive machine learning algorithms generally acknowledge that SVM’s yield the highest prediction accuracy compared to other machine learning algorithms (Davenport & Harris, 2006).
 
-### 2.3.3 ensemble methods
+As SVM’s are often computationally too expensive and advanced for a smaller scale business application, it is useful to compare neural networks and decision trees which are simpler than SVM’s but are also capable of achieving very accurate predictive models.
+Ravi Kumar and Ravi (2007) found that neural networks and decision trees were both capable of out-performing the other, depending on the context – which is limited to bankruptcy prediction in Kumar’s study.
+Both have advantages and disadvantages to consider (refer section 2.2), however the most notable drawback for neural networks is the difficulty in interpreting the algorithms' explanation of human behaviour – a valuable feature in business problems (Tsai & Chiou, 2009).
 
-* papers
-* netflix?
-* multi step to deal with missing data
+Decision trees on the other hand output a flowchart-like tree where each split represents an attribute of the data (Kabra & Bichkar, 2011).
+Trees are known to be intuitive to understand and have been argued to function similarly to how the human mind thinks (Carrizosa, Martin-Barragan, & Morales, 2010).
+This makes the application popular for medicine as it assimilates with the thinking process of a doctor (Seng & Chen, 2010) and it is likely the same benefit can be applied to business managers.
+More complex ensemble decision trees can overcome over-fitting and achieve better predictive results by combining more than one decision tree (Breiman, 1996).
 
-### 2.3.4 similar applications
+Ensemble techniques also refer to combinations of two completely different algorithms and can often outperform individual techniques (Ravi Kumar & Ravi, 2007).
+For example, combining the contrasting advantages of neural networks and decision trees is a worthwhile ensemble technique.
+A paper studying financial earnings management prediction combined these two methods by running them one after another.
+The neural network was run first to create a high rate of prediction accuracy via its complex non-linear learning properties (Tsai, 2009).
+Then, to resolve the lack of explanatory qualities that neural networks are notorious for, decision trees were employed.
+The 81% of cases that were correctly predicted by the neural network were used to generate decision trees, and in turn useful decision rules (Tsai & Chiou, 2009).
+This strategy is a promising way to benefit from the strengths of well performing complementary techniques.
+
+### 2.3.3 similar applications
 
 * Employee churn - anything similar?
-* stepped method business application?
-* Gap
+The literature is limited around predictive models that assist with internal business decisions dealing with employees, teams, and delivering projects – most of the studies focus on businesses that sell products or on the overall performance of large businesses.
+An exception to this is Saradhi and Vijaya’s (2011) study on employee churn, where ‘churn’ refers to the number of individuals moving out of a group within a certain time.
+Saradhi and Vijaya applied popular customer churn predictive models to employee churn - a novel application that focussed internally on employees rather than externally on customers.
+The project chose three machine learning classification algorithms that are normally used in customer churn.
+Saradhi and Vijaya (2011) proposed that the variables and model behaviour of employee churn could be related to customer churn.
+The associated costs of losing customers and finding new customers can also be correlated to the costs of losing staff and hiring new staff.
+Naive Bayes, Random forests (an advanced decision tree) and SVM’s were built and it was expected that random forests would perform best using knowledge from past literature (Coussement & Van den Poel, 2008).
+Naïve Bayes is an algorithm based of simple design that classifies based on the likelihood of each variable independently of the presence of other variables.
+A comparison study showed that Naïve Bayes is outperformed by random forests which match Saradhi and Vijaya’s (2011) prediction (Caruana & Niculescu-Mizil, 2006).
+The SVM model actually far out performed random forests and Naive Bayes by achieving 81% vs 51% and 55% correct predictions respectively.
+This was attributed to the ability to introduce class penalties in SVMs whereas the other two methods are limited by class imbalance (Saradhi & Vijaya, 2011).
+This means that SVM’s had a superior method of balancing the importance of the variables in the data set which was skewed by a heavy majority of employees who kept their job versus the minority who left (25% churn).
+Their study proved that predictive techniques used for businesses that sell products could be translated to predicting internal behaviour of employees.
 
-**
+### 2.3.4 Gap/conclusion
+
+The limited literature on analysis similar to Sardhi and Vijaya’s Employee Churn study (2011) highlights a gap in the study of predicting internal performance in businesses.
+This is particularly relevant to consulting companies that tackle complicated discrete projects with set timelines.
+In some industries, for example engineering consulting, the business provides clients with a fixed fee to complete the assigned project.
+This initial fee creates an element of risk as the business attempts to predict the number of employee hours to be spent.
+Statistical and machine learning models to predict the correct fee for projects according to project type, client, and other characteristics could reduce risk for a range of consulting businesses in setting initial fees.
 

### 2016-01-01 finalwriteup/draft.rmd
finished lit review

--- 2015-12-16 finalwriteup/draft.rmd
@@
 
-### 2.2.3 Introduction
+### 2.2.1 Introduction
 
-### 2.2.4 Linear Regression/logistic regression
+The aim of this project is to use statistical techniques to model the profitability of projects.
+In order to optimise the model, several statistical and machine learning techniques will be tested, and compared and it is therefore prudent to review the range of statistical and machine learning techniques that have been successfully applied in the literature.
+These vary from simple methods such as Linear Regression to complex, deep learning such as Neural Networks.
+The previous section highlighted the use of linear regression, neural networks, SVM and in one case boosted trees for prediction, however other business problems have utilised a wider range of methods.
+These include Naiive Bayes, Random Forests, and machine learned Bayesian Networks.
+The following section will present each method's advantages and disadvatanges and an example of a past successful application.
 
-### 2.2.5 Naiive Bayes
+### 2.2.2 Linear Regression
 
-### 2.2.6 Ensemble Trees
+Linear regression is one of the simplest and most popular statistical prediction methods [@Hastie2009].
+An equation for a line is learned, which is defined by a constant and a coefficient times the value of each variable.
+The coefficients are chosen to minimise the residual sum of squares (least squared) [@Hastie2009].
+The reason this method is popular is that it is simple, provide stable predictions, and can be adapted for categorical variables [@Seng2010].
+The integration of categorical variables is called Anova, and the algorithm compares group mean variances within the categorical variable.
 
-### 2.2.7 Bayesian Networks
+There are several drawbacks to linear regression.
+For example assumptions must be made about the structure of the data.
+Predictor varialbes are assumed to be linearly related to the response variable and the varialbes are normally distributed.
+Many real world phenomena do not correspond to these assumptions which can make it difficult, sometimes impossible, to produce useful results [@Hastie2009][@Putler2012].
+Furthermore, response variables must be continous numeric variables and if two or more predictor variables are correlated, the values of their coefficients are highly unstable.
+On the other hand, correlated variables do not impact the predictive accuracy of the model.
+A final drawback is that the variable coefficients cannot be used to compare variable importance unless the varialbes have all been standardised to one another [@Putler2012].
+This limits the amount of insight that can be drawn from a linear regression model.
 
-### 2.2.7 SVM's and Neural Networks
+Despite the limitations of linear regression models, they have widespread business applications due to their simplicity and ease of application.
+Regression is used to predict risk in the finance and insurance industries, predict who to target in marketing exercises, and predict consumption spending in the field of economics [@Harrell2013].
+Due to its simplicity and widespread application, linear regression is a good starting point for an analysis and benchmark for performance against complex models.
 
-### 2.2.8 Summary
+### 2.2.3 Logistic Regression
+
+The method behind linear regression can be adapted to predict a binary response variable (1 or 0).
+A predictive linear equation is computed so that the predictor variables are combined with coefficients to give an output between 0 and 1.
+This is then transformed into the log odds to mediate extreme variable input values.
+The result is a probability an input case will result in a 0 or 1 (i.e.
+success or failure) [@Moore1989].
+Similar pros and cons exist for logistic regression as linear regression, and it too can be used as a benchmark to compare other binary predictive models due to its speed and simplicity.
+Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survive/death [@Moore1989].
+
+### 2.2.4 Naive Bayes
+
+Naive Bayes prediction method makes conditional independence assumptions about the predictor variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
+Specifically, it is assumed that each variable is conditionally independent given the respose class alone, without considering other predictor variables.
+Mathematically this means that each variable's probability contribution can be simply calculated by looking at the probability of a response class given a predictor value by itself.
+The predictor variables probability estimates are combined for each possible response class, and the response class with the highest probability is chosen [@Provost2013].
+
+The advantages of this method are that the conditional independence assumption enables very fast calculations and predictions.
+The method can perform very well for real world tasks because the assumption of independence does not damage predictions significantly.
+This is because if multiple varialbes are related, the variables independently still direct the prediction in the correct direction.
+The correlated variables will double or triple their emphasis on the predicted response variable, however this does not necessarily impact classification accuracy as the final step is to simply choose the class with the *highest* probability [@Provost2013].
+This is fine for ranking.
+On the other hand, the output probabilities are not realisitic, and the values themselves should not be used [@Caruana2006].
+Another disadvantage is that in order to calculate the class probability from a numeric variable, a distribution must be assumed (often gaussian) which the numeric data may not fit into neatly [@Caruana2006].
+Finally, Naive Bayes classifers are known to not perform well in binary classification [@Caruana2006].
+
+A real world success story for Naive Bayes classifiers hinges on their ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
+Naive Bayes has been used in complex spam detection systems where new spam emails or toxic text themes can be quickly added to the filtering model.
+The Naive Bayes method provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
+
+### 2.2.5 Decision Trees/ Ensemble Trees
+
+Decision trees are one of the simplest and most intuitive machine learning methods.
+There are several brands of basic decision tree algorithms including ID3, C4.5, CART, and CHAID with CART and C4.5 being the most popular [@Kabra2001].
+Ravi Kumar and Ravi [@Kumar2007] recommend the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
+Single decision trees have the risk of over-fitting the data and low predictive accuracy but have favourable characteristics such as providing intuitive rules that a decision maker can follow in real-life scenarios.
+In addition their results do not depend heavily on the skill of the data analyst [@Kabra2001][@Putler2012].
+
+To tackle the problem of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
+Three examples of these ensemble decision tree methods are bagging, boosting, and random forests.
+Bagging creates multiple trees by sampling a different training data set for each tree and then combining tree results [@Breiman1996].
+Random forests is an advanced form of bagging in that multiple trees are sampled from the same training data.
+However, when creating the tree for each sample dataset, a random subset of attributes is used to determine each split.
+This prevents the trees from becoming too correlated [@Breiman2001].
+The boosted decision tree approach combines gradient descent with a series of decision trees.
+Each tree is limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) from the preceding tree.
+The limited depth of each tree prevents overfitting at each stage and the combined result of up to thousands of trees is very powerful [@Elith2008].
+
+Ensemble decision trees have a different foundation to more traditional statistical prediction methods such as regression.
+For this reason, there are some advantages and disadvantages.
+The progression of binary splits means non-linear feature interactions can be captured and the predictor variables do not need to be transformed as no assumptions are made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
+Other advantages are that the ensemble methods are very fast in comparison to SVM's and neural networks but perform just as well and provide insights such as variable importance and variable relationships [@Sealfon2012].
+Random forests in particular has the benefit of not being affected by noise in the data as well as the capacity to handle a large number of variables [@Sealfon2012].
+Caruana [@Caruana2006] tested boosted trees, random forests, neural networks, SVM's, logistic regression and naive bayes on 11 binary classification problems and found that boosted trees performed best, followed by random forests.
+Clearly, ensemble tree methods can compete with high level machine learning algorithms.
+
+Along with the many advantages of trees and ensemble trees, it is important to be aware of their limitations.
+As trees are not built on a probabilistic framework, results can not be provided in this framework.
+For example, confidence intervals for predictions are not available for standard ensemble methods [@Louppe2014].
+Also, variable importance tables provided by random forests and boosted trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
+The methods can also be prone to overfitting if not handled carefully.
+Ensemble trees have many advantages particularly predictive power, but it is important to recognise that the best method for a problem often depends on the type of problem [@Louppe2014][@Caruana2006].
+
+### 2.2.5 Bayesian Networks
+
+A bayesian network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
+The model is visually represented by a DAG (directed acyclic graph) and is capable of linking the conditional dependency between any variable to another variable or via other variables which are represetned by arcs [@Heckerman1998].
+This kind of conditional relationship is bayesian.
+Bayesian networks have found success in combining deterministic models with observation data as well as expert knowledge as the statistical relationships between variables can be determined from different places and manually entered into a single model [@Kragt2009].
+
+Bayesian networks are excellent at representing knowledge from different sources and types, and visuallising them in a way that communicates results well to decision makers [@Kragt2009].
+Therefore bayesian networks become an excellent decision support tool.
+Network relatitonships can be machine learned, however this is not widely included in the suite of machine learning methods and is referred to more as a way to model complex networks.
+Drawbacks of bayesian networks include their inability to handle continuous numeric variables (which must be discretised).
+This means that the varialbe relationships are categorical [@Kragt2009].
+Popular applications of bayesian networks are modelling uncertainty in natural resource management and modelling complex business network structures such as airports [@Wu2013].
+
+### 2.2.6 SVM's and Neural Networks
+
+Neural networks were inspired by biological neural networks of the human nervous system and require a lot of training data.
+They are very flexible in the relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
+A disadvantage is that although the results are accurate, the algorithm is a black box - meaning the data is input and the results output without providing the user insight into the methodology.
+Neural networks can perform high level tasks as a result of deep learning such as hand writing recognition, vehicle control, face recognition, and cancer detection [@Haykin2004].
+
+The simplest type of Support vector machines (SVM's) calculate the linear boundary between two categories in feature space so that the category of future cases may be determined depending on which side of the boundary they fall.
+The location of the boundary is determined by the position of the widest clear margin that can be drawn between the two categories, allowing a certain number of misclassifications.
+This concept can be adapted to non linear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
+Advantages of SVM's include the ability to model non linear relationships in data, and that no assumptions about the distributions of the data must be made as the model is non-probabilistic.
+They are also known to be robust over different samples and perform well in high-dimensional feature space [@Auria2008].
+On the other hand, the output is in the form of distance to the boundary (as opposed to probability) and results are not transparent (similar to neural networks) [@Caruana2006].
+Also, SVM's are known to be very slow to train and therefore often not suitable for industry purposes [@Auria2008].
+Because SVM's perform well in high-dimensional space and can form non-linear rules they have been applied to protein classification (medicine) and text and image recognition [@Byun2002].
+
+### 2.2.7 Summary
+
+This section reviewed the advantages and disadvantages of statistical and machine learning techniques that have been successfully applied to business problems.
+These include linear regression, logistic regression, naive bayes, decision trees and ensemble trees, bayesian networks, neural networks, and SVM's.
+Linear/logistic regression and naive bayes are computationally very fast with a simple concept that make broad assumptions about the data, but work well as baseline models against which to compare complex models.
+In certain cases these simpler models are significantly outperformed by complex models however it is important to compare both.
+Neural networks and SVM's are excellent machine learning predictors which can perform well with messy data and no assumptions, however they are slow and do not provide insight into the results of analysis.
+This is not appropriate for the profitability prediction problem in this project because communication of the reasoning behind predictions is crucial for uptake of the model by decision makers.
+
+On the other hand, ensemble tree methods can perform as well as neural networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
+Bayesian networks are excellent for presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally not been included in the machine learning literature as much as the other methods.
+This may be due to low awareness of learned bayesian networks or lower performance.
+In conclusion, simple models such as linear/logistic regression and naive bayes are good baseline models to implement, while ensemble tree methods and bayesian networks may handle messy data better than the simpler models while still providing insight to decision makers.
 
@@
 
-### 2.3.1 Most simple and frequent models
+The aim of this research project is to create a predictive model for the profitability of a company's consulting projects using their internal CRM data.
+The literature pertaining closely to the cost prediction topic has been reviewed in section 2.1, however it is relevant to broaden the assessessment of the literature on advanced machine learning techniques that have been applied to general business problems.
+The simpler statistical methods have been covered already, so the focus of this sections is on the topic of machine learning prediction methods for business decision-makers.
+Popular applied topics include predicting stock fluctuation, customer churn analysis, fraud prediction, customer classification, market segment analysis, product success prediction, and recommendation systems (Seng & Chen, 2010).
+This section will first discuss the use and merits of the employed more advanced machine learning methods, followed by a case study of employee-churn prediction.
 
-### 2.3.2 Complex models
+### 2.3.2 Summary of Advanced Statistical and Machine Learning Methods Applied to Business Problems
 
-* SVM, Neural networks, random forest
-* data must not be normal
-* most successful?
+According to Breiman [@Breiman2001b], simple theoretical models such as linear regression are informative but often provide unsatisfactory accuracy for modelling real life data which is messy and does not necessarily align with the rules of statistical theory.
+As this project will address a problem with a large quantity of messy data, it is worthwhile reviewing the algorithmic machine-learning predictive models that have developed rapidly in recent years.
+They have found success because they are able to sometimes fit real-life data with more accuracy than theoretical models [@Breiman2001b].
+Ravi Kumar and Ravi [@Kumar2007] performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
+The most widely used model was neural networks, however logistic and linear regression, decision trees, SVM’s, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
+Supporting Breiman’s [@Breiman2001b] statements, it was found that DA and linear regression techniques were not preferred due to their low accuracy.
+The overall assessment was that SVM’s outperformed Neural networks (back propagation neural networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
+Although SVM’s performed the best, as discussed, they are often extremely complex and slow, requiring a high memory [@Kumar2007].
+In support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that statisticians experienced in predictive machine learning algorithms generally acknowledge that SVM’s yield the highest prediction accuracy compared to other machine learning algorithms.
 
-### 2.3.3 ensemble methods
+As SVM’s are often computationally too expensive and advanced for a smaller scale business applications, such as this research project, it is useful to compare neural networks and decision trees which are simpler than SVM’s but are also capable of achieving very accurate predictive models.
+Ravi Kumar and Ravi [@Kumar2007] found that neural networks and decision trees were both capable of out-performing the other, depending on the context – which is limited to bankruptcy prediction in Kumar’s study.
+Both have advantages and disadvantages to consider (refer section 2.2), however the most notable drawback for neural networks is the difficulty in interpreting the algorithms' explanation of human behaviour – a valuable feature in business problems [@Tsai2009].
+On the other hand, trees are known to be intuitive to understand and have been argued to function similarly to how the human mind thinks [@Carrizosa2010].
 
-* papers
-* netflix?
-* multi step to deal with missing data
+Ravi Kumar and Ravi's [@Kumar2007] study concludes that ensemble techniques, which refers to combinations of two completely different algorithms, can often outperform individual techniques.
+For example, combining the contrasting advantages of neural networks and decision trees is a worthwhile ensemble technique.
+A paper studying financial earnings management prediction combined these two methods by running them one after another.
+The neural network was run first to create a high rate of prediction accuracy via its complex non-linear learning properties [@Tsai2009].
+Then, to resolve the lack of explanatory qualities that neural networks are notorious for, decision trees were employed.
+The 81% of cases that were correctly predicted by the neural network were used to generate decision trees, and in turn useful decision rules [@Tsai2009].
+This strategy is a promising way to benefit from the strengths of well performing complementary techniques.
 
-### 2.3.4 similar applications
+### 2.3.3 Employee Churn Case Study
 
-* Employee churn - anything similar?
-* stepped method business application?
-* Gap
+The literature on predictive business models is generally limited to studies on product sales, customer behaviour, or overall performance of large businesses.
+There are far fewer studies that assist with internal business decisions dealing with employees, teams, and delivering projects as is the case with this project.
+An exception is Saradhi and Vijaya’s [@Saradhi2011] study on employee churn, where ‘churn’ refers to the number of individuals moving out of a group within a certain time.
+Saradhi and Vijaya applied popular customer churn predictive models to employee churn - a novel application that focussed internally on employees rather than externally on customers.
+The project chose three machine learning classification algorithms that are normally used in customer churn and it was proposed that the variables and model behaviour of employee churn could be related to customer churn.
+The associated costs of losing customers and finding new customers can also be correlated to the costs of losing staff and hiring new staff.
 
-**
+Naive Bayes, Random forests (an ensemble decision tree method) and SVM’s were built.
+A comparison study showed that Naïve Bayes is outperformed by random forests which match Saradhi and Vijaya’s prediction [@Caruana2006].
+The SVM model actually far out performed random forests and Naive Bayes by achieving 81% vs 51% and 55% correct predictions respectively.
+This was attributed to the ability to introduce class penalties in SVMs whereas the other two methods are limited by class imbalance [@Saradhi2011].
+This means that SVM’s had a superior method of balancing the importance of the variables in the data set which was skewed by a heavy majority of employees who kept their job versus the minority who left (25% churn).
+It should be noted that the boosted tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVMs.
+Particularly because boosted trees provide insight into the predictive model which could aid uptake of the model with decisoin makers.
+However, it is noteworthy that their study proved that predictive techniques used on customres could be translated to predicting internal behaviour of employees.
+
+### 2.3.4 Gap/conclusion
+
+The limited literature on analysis similar to Sardhi and Vijaya’s Employee Churn study [@Saradhi2011] highlights a gap in the application of predictive techniques to model internal performance in businesses.
+This is particularly relevant to consulting companies that tackle complicated discrete projects with set timelines.
+In this project, internal consulting project data will be used to predict a new project performance in terms of internal employee structure as well as client characteristics.
+A review of the literature in machine learning applied to business problems revealed that SVM's have found to be the most accurate technique followed by neural networks and ensemble tree methods.
+As discussed in section 2.2, it was also noted that neural networks and SVM's do not provide insight into the model's predictions to the decision maker.
+A gap exists in applying these cutting edge machine learning techniques to internal business decision-making.
 
@@
 
-summary of 
-* how are B2B decisions currently made
-* what are the most popular/effective methods
-* which ensemble methods may suit our missing data situation
-* Gap in literature/contribution
+The literature pertaining to the aim of this research project, cost prediction using internal consulting CRM data, has been extensively reviewed.
+The problem of cost prediction has been addressed in many studies beforehand which have generally been limited to prediction of construction costs for buildings and infrastructure and effort estimation for IT projects.
+The prediction models consisted of case based reasoning, linear regression, neural networks and in one or two instances an ensemble tree or SVM.
+This leaves a gap in the literature in effort estimation for consulting companies in the construction industry as well as trialling a broader range of prediction methods.
+
+An exhaustive list of machine learning and statistical methods were reviewed theoretically as well as practically in the broader scope of business problem prediction.
+It was determined that bayesian networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as neural networks while providing the additional benefit of insights into the reasoning behind model predictions.
 

### 2016-01-30 finalwriteup/draft.rmd
began methods

@@
 
-# Chapter 3 Introduction of Case study
+# Chapter 3 Methods - 2 to 8 pages
 
-## 3.1 Obtaining dat
+The aim of this project is to determine whether the profitability of projects for consulting businesses can be predicted using statistical and machine learning techniques.
+As stated in the literature review, a gap in research exists for testing whether internal CRM data from a company in the construction industry could predict project profitability.
+This section will explain the chosen methodology for testing this aim using a case study Engineering consulting company that charges for their time spent on project delivery.
+Once the predicition potential has been assessed, a secondary aim is to determine the positive impact the predictive model could have on the overall profitability of the company.
 
-### 3.1.1 How data was recorded
+**justify the process by which the research questions were answered**
+**argue and justify each decision that was taken to arrive at how to organise reserach**
 
-* years of employees entering hours each day
-* project managers or admin entering client details
+## 3.1 Obtaining data
 
-### 3.1.2 How data was extracted - CRM
+An assessment of the literature revealed that the predictive power of internal timesheet data has not yet been tested yet for project profitability for consulting companies in the construction industry.
+In addition, this is the sole source of project data available for these types of businesses besides interviewing project managers and reviewing the account records of invoicing versus employee costs.
+As the internal CRM for the case study company stores a detailed account of time spent on each project along and client information for twelve years, it was clear that this source of data could reveal the most and was most suitable to statistical analysis.
+The complete twelve years of project timesheets and invoices was extracted via the CRM software interface, describing 4169 projects.
+The projects varied from total invoiced amounts of $500 to over $1,000,000 across four internal disciplines in the company.
+At the time of data extraction, it was not clear whether the full twelve years were necessary and relevant to predicting current project profitability but this was to be determined later in the analysis.
 
-* queried direct from CRM
-* data de-identification
+Once the data was extracted from the CRM, an employee from the case study company performed the task of de-identification.
+This is an important step before research commences as the privacy of project employees, clients, and project names must be removed for ethical reasons.
+In the case of employee names, professional titles replaced the names, and for client codes replaced client names.
 
-## 3.2 Summary of variables
+## 3.2 Cleaning and variable engineering
 
-* categorical
-* numerical
-* significant cleaning
-    * cost sometimes entered in as hours. ie $24 entered as 24 hours
-* engineered variables
-    * grep through project names to create categories - categories developed with case study manager
-    * return per dollar = (invoiced - cost)/cost
-    * hours analysed to produce: employee position that performed majority of the hours
-        * % of hours completed by majority position
-    * timespan based on timesheet entries as well as number of days with hours recorded
-    * number of workers on project based on timesheet entries
-    * client characteristics such as mean total amount invoiced for a repeat client
+The obtained data contained a rich source of project information and a lengthy process of cleaning the data and engineering useful variables was required.
+The data was initially in three types of datasets extracted from the same CRM software:
 
-* amount of missingness
+* invoicing data
+* timesheet data
+* project summary data
+
+These three sources needed to be compiled into a single dataset that detailed one project per row as this is the structure required for predictive analytic methods when predicting the overall profitability of a project.
+Before compilation could begin however, thorough cleaning is required where the data was plotted and statistically assessed so that outliers could be visually or statistically discovered.
+Outliers were then investigated for data entry errors.
+Many such errors were encountered.
+One type of error for example was a user entering a 123 kilometer drive in a car as 123 hours.
+Once these errors were detected, they were discussed with the company directors and appropriate corrective action was taken.
+
+### 3.2.1 Variable Engineering
+
+Although dozens of variables were available from the initial dataset, it was prudent to engineer further descriptive variables, particularly in the invoicing dataset and timesheet dataset where the information needed to be compiled into a single row per project.
+All variables would eventually be tested for variable importance and predictive power with respect to project profitability.
+Examples of these engineered variables include:
+
+Timesheet Data
+* percent of hours performed by each professional role for the entire project
+* timespan of the entered project hours
+* percent of hours performed by 'profesional' employees as opposed to 'technical'
+* total cost of employee hours per project
+* total cost of external subcontractors or disbursements per project
+* total number of users that entered hours on each project
+* mean hours per day entered on a project
+* number of disciplines active in a project
+
+Invoicing data
+* total amount invoiced and renumerated per project
+* mean invoice size per project
+* mean invoice per client
+* invoice frequency per project
+* client invoice frequency
+
+Project Data
+* text analysis of project descriptions detected a list of key words that could classify projects into 16 categories.
+This key word analysis was done in conjunction with a company employee and the resulting classifications were reviewed.
+* number of projects completed with each client and client contact
+
+Once the three data sources were combined, further variables were engineered:
+* project profit
+* return per dollar for each project (how much was spent on expnses for every dollar earned?)
+
+Additional variables included in the project summary dataset were client information such as industry, discipline, job description, and post code.
+Once numerous potentially important variables were engineered, the most important ones can be narrowed down to reduce noise in the predictive models.
 
@@
 * combination of cforest, anova, random forest
+
+## 3.4 Model selection
+
+## 3.5 Model comparison, Model averaging
+
+- need for statistical measurements
+
+## 3.6 Bottom line translation
+
+## 3.7 Conclusion/Link to next chapter
 

### 2016-04-06 boost_tree.rmd
wrote some stuff to go in the final write up

@@
 #partial dependence plots:
-i=8
+i=1
 a<- plot.gbm(boost.1, i, type="response")

### 2016-04-06 finalwriteup/jam-in.rmd
wrote some stuff to go in the final write up

---
title: "filler"
author: "Amy Cook"
date: "April 6, 2016"
output: word_document
---

This R markdown document is filler for the draft i have written up at home.

I was up to Method. Just finished conditional forest? and maybe had moved on to teh next part?

Write outline for next two parts: Results and Discussion and Conclusion

# Results - over 5 pages, up to 20 pages

Guidelines:
* Outlines what you found out in relation to your research questions or hypotheses, presented in figures and in written text.
* Results contain the facts of your research. Often you will include a brief comment on the significance of key results, with the expectation that more generalised comments about results will be made in the Discussion section.

Which results to discuss
* restate research questions and hypotheses. can we predict profitability? ofocurse you can make a prediction, 
real question might be is it worthwhile for the business to engage in this activity? we will look at this as a bottom line analysis. acknowledge that it may be worthwhile for other reasons such as confidence and transparency in historic projects.
* important variables: cforest diagram vs anova vs random forest
    * which variables were chosen for machine learning analysis
* first tried random forest predicting return per dollar
    * fail, why - for discussion
* moved on to predicting yes or no profitability.
    * tried: random forest, logistic regression, gradient boosted trees, naiive bayes (baseline), bayesian networks
    * boosted trees could handle the missing data, whereas for all? other methods I had to narrow down data to complete data
    * used MICE package - explain what this is in method.
    * comparison of predictive accuracy of all methods. non are significantly better than the other.

* tried model averaging - comparison to other methods.

* bottom line analysis - run 1000 times to get confidence interval around the line.
    * should the business adopt the model?

* to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the case based reasoning approach

# Discussion - 10 to 20 pages

The Discussion section:

comments on your results;
explains what your results mean;
interprets your results in a wider context; indicates which results were expected or unexpected;
provides explanations for unexpected results.
The Discussion should also relate your specific results to previous research or theory. You should point out what the limitations were of your study, and note any questions that remain unanswered. The Discussion CAN also include Conclusions/Future Research.

* cannot predict return per dollar. may be that this is much to specific for the volume of data available, may be that the project variation is not captured by the type of data available
* simpler task of predicting profitability or not does provide predictive accuracy well above random chance and also above our base measure - naiive bayes, but maybe not significantly? question of whether the 0.05 significance is appropriate
* how many trials do I need to do to make it significant?!
* best models are gradient boosted trees and logistic regression. why? can hypothesis but not really
    * this shows the profitability can be predicted with some success, but real question is, is it financially viable?
* depending what the bottom line plot shows - yes or no. 
* compare to past research. problems with past research and does this solve any of those?
    * poor user uptake - future work?
    * lack of trust - nearest neighbour must be presented with results - case bsaed reasoning
        * results from boosted tree shows partial dependence plots - insight.
    * never been applied to a consultant within construction industry. has been applied lots to software
        * still has the potential to help with the way the industry is structured - hinges on user uptake.

# Conclusion

Future work
* user testing on trust levels of prediction with nearest neighbour

### 2016-04-06 final_app/server.R

@@ -1,13 +1,13 @@
 setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
 # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.2/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.2/library')
-library('shiny',lib='C:/Progra~1/R/R-3.2.2/library')
-library('shinythemes',lib='C:/Progra~1/R/R-3.2.2/library')
-library('FNN',lib='C:/Progra~1/R/R-3.2.2/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.3/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.3/library')
+library('shiny',lib='C:/Progra~1/R/R-3.2.3/library')
+library('shinythemes',lib='C:/Progra~1/R/R-3.2.3/library')
+library('FNN',lib='C:/Progra~1/R/R-3.2.3/library')
 # library('ggplot2')
 # library("dplyr")
 # library("plyr")
@@ -28,15 +28,15 @@ df <- read.table(text = " id  min  max
     Sp1     150          151     ", header=TRUE)
 
 #make a new dataframe with information about crossbar sizes
-df2<- data.frame(id=c(rep("Sp1",3)), ymin= c(19000, 22000,24000),
-                 ymax=c(33000, 30000, 28000), probability=c('95%','75%','50%'))
+df2<- data.frame(id=c(rep("Sp1",1)), ymin= c(4), y.mean = c(13),
+                 ymax=c(22), confidence=c('90%'))
 df2 <- within(df2, 
-              probability <- factor(probability, 
-                                    levels=c('95%','75%','50%')))
+              confidence <- factor(confidence, 
+                                    levels=c('90%')))
 
[…]

### 2016-04-10 finalwriteup/draft.rmd
finished variable selection

@@
 * project profit
-* return per dollar for each project (how much was spent on expnses for every dollar earned?)
+* return per dollar for each project (how much was spent on expenses for every dollar earned?)
 
-Additional variables included in the project summary dataset were client information such as industry, discipline, job description, and post code.
-Once numerous potentially important variables were engineered, the most important ones can be narrowed down to reduce noise in the predictive models.
+Besides the engineered variables, additional variables included in the project summary dataset were client industry, discipline, job description, and post code.
+Once numerous potentially important variables were engineered, the most important ones could be narrowed down.
+This improves the accuracy of a model because unnecessary or irrelevant variables add noise to the prediction of target values.
+Variable selection also increases computational efficiency by reducing the number of calculations and improves understanding of the prediction structure [@Weisberg2005].
 
@@
 
-* combination of cforest, anova, random forest
+### 3.3.1 Outlier deletion
+
+Before determining which variables contribute most to the target value, return per dollar, outliers were deleted.
+Specifically, cases with extreme return per dollar values were investigated as special scenarios may have contributed to these values and the model is not intended to predict special scenarios.
+Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles as intially practised by Tukey, the inventor of the box plot [@Tukey1977].
+The outlier projects determined by this method were then reviewed with an employee from the case study company.
+Based on their domain knowledge and assessment of the outlier cases, cut-off values for the range of return per dollar values were determined.
+Outliers were projects with return per dollar values greater than 3 or below -2.
+
+### 3.3.2 Variable selection Methods
+
+Three methods of variable selection were trialled and compared to test whether the different methods highlighted different variables as important.
+These were Analysis of Variance (ANOVA), conditional inferece forests (cforests), and random forests.
+It was expected that random forests would highlight variables that had more categories as more important which is a well known bias; it is also a bias that cforests have overcome [@Strobl2007].
+ANOVA is based on linear theory which contrasts conditional forest and random forests more free-form structure.
+Therefore, it is important to compare the contrasting methods' assessment of variable importance.
+
+#### 3.3.2.1 ANOVA
+
+ANOVA is closely related to linear fit models but incorporates the analysis of differences in group means (i.e.
+categorical variables) [@Lunney1970].
+The method is simple and fast but assumes variables are normally distributed and linearlly related.
+The variables must therefore be normalised before analysis, however it cannot be assumed that each variable can be perfectly normalised.
+A disadvantage of this method is that linear relationships and normal variables do not necessarily represent real world data in the best way [@Breiman2001b].
+
+In order to compare variable importance, the output of the model is assessed which includes the linear coefficients for each variable and the p-values for the coefficients [@Markham2016].
+The magnitude of the coefficient does not indicate importance relative to the other variables because the magintude is directly dependent on the values within the variable.
+This leaves the p-values which are the result of a test for whether there is a significant relationship between the variable and the target variable.
+This is a good indication of whether the variable has a relationship with the target variable, however doesnt necessarily rank importance.
+If a variable is measured more precisely, it will have a smaller p value, whereas a variable measured roughly will have a higher p value.
+This does not necessarily make the more precisely measure variable more important than the other.
+The p test does however distinguish between variables that have a statistical relationship to the target variable and ones that do not.
+With regards to p-values it is also important to keep in mind that, with a dozen variables, chance alone can produce a variable with a pvalue < 0.001 7% of the time [@Rice1989].
+
+The p-value output from ANOVA models give a good indication of which variables have a statistically significant relationship with the target variable.
+However, the values should not be used to rank the variables against one another, and significant p-values can occur by chance.
+
+#### 3.3.2.2 Random Forest
+
+The random forest algorithm can produce a permutation variable importance for each variable as part of its output.
+Variable importance is represented as a score, and to do this a permutation and calcuations are performed on each variable.
+More specifically, consider a single covariate.
+The values of this variable in the out of bag sample are randomly permuted (reordered) for each tree - a technique that mimics the absence of that variable.
+The trees in the random forest are then run again.
+Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted to the original random forest.
+If the accuracy of the premuted forest is much less than the original forest, then that variable is important.
+Importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the random forest [@Breiman2005].
+
+The random forest variable importance provides valuable insight because it addresses the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables.
+It handles uncorrelated variables well if the samples for the trees are subsampled without replacement [@Strobl2007].
+However, a severe disadvantage of random forest variable importance is that it is not reliable for variable selection where potential predictor variables vary in their scale of measurement or their number of categories.
+The importance of correlated predictors is overestimated and the algorithm tends to select variables that have many possible splits or many missing values [@Strobl2007].
+Overall, the random forest permutation importance is useful as it is able to compare variables in a machine learning environment and can include multivariate interactions, but the current random forest importance measure overestimates the importance of variables with many categories and some numeric variables.
+
+#### 3.3.2.3 Cforests
+
+Cforests are an alterantive ensemble tree method to random forests that overcome the aforementioned shortfalls of random forests such as overestimating the importance of variables with many categories and numeric variables.
+It is built from ctrees that are based on a conditional inference framework.
+The key difference between ctrees and standard decision trees are that a significance test procedure for splitting instead of an purity meausre such as the Gini coefficient, which is centered around information gain.
+
+The details behind the significance test procedure of a ctree are as follows.
+Each variable is permuted in every possible way, and a correlation value is calucated between the tested variable and target variable, for each permutation.
+The unchanged variable correlation is then compared with the correlation values for all permutations of that variable.
+From this, a p-value for the true correlation value can be calculated.
+The predictor variable with the lowest p-value is selected as the splitting variable [@Hothorn2006].
+Using an ensemble forest of these trees, permutation is then calculated in the same fashion as random forests, via permuting a variable, re-running the forest and comparing the decrease in accuracy.
+Several sources recommend cforests over other machine learning varialbe importance methods due to its unbiased qualities [@Strobl2007][@Strobl2009][@Hothorn2006].
+
+#### 3.3.3 Variable Selection Summary
+
+Limiting the predictive model to a concise set of meaningful variables reduces noise and improves predictions.
+Less variables means that a simpler model is being used for prediction which is easier to understand for stakeholders [@Weisberg2005].
+For these reasons, a subset of the most important variables were chosen before modelling began.
+This entailed first eliminating outliers, then comparing important variables from ANOVA as well as random forests and cforests.
+The literature implies that cforests would suit the data for this project as it is unbiased and not limited to linear theory [@Strobl2007].
 
 ## 3.4 Model selection
+
+*
 
@@
 ## 3.7 Conclusion/Link to next chapter
-
-# Chapter ? Method
 

### 2016-04-10 de_identify.R

@@ -315,3 +315,5 @@ all6a<- all6a %>% select(-Role.JD)
 all6a$JD.Second<- as.factor(all6a$JD.Second)
 
 write.csv(all6a,'all6b.csv')
+
+#go to 1st_stats.R

### 2016-04-10 finalwriteup/draft.rmd
added jam in

@@
 can we improve this by bringing huge variables back in a smarter way?
+
+# Results - over 5 pages, up to 20 pages
+
+Guidelines:
+* Outlines what you found out in relation to your research questions or hypotheses, presented in figures and in written text.
+* Results contain the facts of your research.
+Often you will include a brief comment on the significance of key results, with the expectation that more generalised comments about results will be made in the Discussion section.
+
+Which results to discuss
+* restate research questions and hypotheses. can we predict profitability? ofocurse you can make a prediction, 
+real question might be is it worthwhile for the business to engage in this activity?
+we will look at this as a bottom line analysis.
+acknowledge that it may be worthwhile for other reasons such as confidence and transparency in historic projects.
+* important variables: cforest diagram vs anova vs random forest
+    * which variables were chosen for machine learning analysis
+* first tried random forest predicting return per dollar
+    * fail, why - for discussion
+* moved on to predicting yes or no profitability.
+    * tried: random forest, logistic regression, gradient boosted trees, naiive bayes (baseline), bayesian networks
+    * boosted trees could handle the missing data, whereas for all? other methods I had to narrow down data to complete data
+    * used MICE package - explain what this is in method.
+    * comparison of predictive accuracy of all methods. non are significantly better than the other.
+
+* tried model averaging - comparison to other methods.
+
+* bottom line analysis - run 1000 times to get confidence interval around the line.
+    * should the business adopt the model?
+
+* to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the case based reasoning approach
+
+# Discussion - 10 to 20 pages
+
+The Discussion section:
+
+comments on your results;
+explains what your results mean;
+interprets your results in a wider context; indicates which results were expected or unexpected;
+provides explanations for unexpected results.
+The Discussion should also relate your specific results to previous research or theory.
+You should point out what the limitations were of your study, and note any questions that remain unanswered.
+The Discussion CAN also include Conclusions/Future Research.
+
+* cannot predict return per dollar. may be that this is much to specific for the volume of data available, may be that the project variation is not captured by the type of data available
+* simpler task of predicting profitability or not does provide predictive accuracy well above random chance and also above our base measure - naiive bayes, but maybe not significantly?
+question of whether the 0.05 significance is appropriate
+* how many trials do I need to do to make it significant?!
+* best models are gradient boosted trees and logistic regression. why? can hypothesis but not really
+    * this shows the profitability can be predicted with some success, but real question is, is it financially viable?
+* depending what the bottom line plot shows - yes or no. 
+* compare to past research. problems with past research and does this solve any of those?
+    * poor user uptake - future work?
+    * lack of trust - nearest neighbour must be presented with results - case bsaed reasoning
+        * results from boosted tree shows partial dependence plots - insight.
+    * never been applied to a consultant within construction industry. has been applied lots to software
+        * still has the potential to help with the way the industry is structured - hinges on user uptake.
+
+# Conclusion
+
+Future work
+* user testing on trust levels of prediction with nearest neighbour

### 2016-04-27 finalwriteup/draft.rmd
further work on method

@@
 output: word_document
-bibliography: Library2.bib
+bibliography: Library.bib
 ---
@@
 Timesheet Data
+
 * percent of hours performed by each professional role for the entire project
@@
 Invoicing data
+
 * total amount invoiced and renumerated per project
@@
 Project Data
+
 * text analysis of project descriptions detected a list of key words that could classify projects into 16 categories.
@@
 Once the three data sources were combined, further variables were engineered:
+
 * project profit

### 2016-05-13 finalwriteup/draft.rmd
finishing method

@@
 Using an ensemble forest of these trees, permutation is then calculated in the same fashion as random forests, via permuting a variable, re-running the forest and comparing the decrease in accuracy.
-Several sources recommend cforests over other machine learning varialbe importance methods due to its unbiased qualities [@Strobl2007][@Strobl2009][@Hothorn2006].
+Several sources recommend cforests over other machine learning variable importance methods due to its unbiased qualities [@Strobl2007][@Strobl2009][@Hothorn2006].
 
@@
 
-*
+* as stated in lit review, there is a need for testing more sophisticated ML algorithms such as boosted trees, random forest and bayesian networks on effort estimation.
+ML algorithms MUST have a way of explaining results.
+this excludes SVm and neural networks.
+
+* naiive bayes - good baseline
+* bayesian network - bad
+* random forest 
+* logistic regression - baseline
+* boosted tree
 
@@
 
-- need for statistical measurements
+* used AUC for comparison statistic bc... 
+* run each method 10 times? statistically significant difference. 10 fold cross validation
+* if no statistical difference, run more times -20, 50 75, 100.
+* select best wto or three methods and trial model averaging
+* improve AUC by model averaging
+* choose best method: either a single method or combined
 
@@
 
+* although predictive power is important, for the method to be integrated into the real world, the affect on a business' bottom line must be tested.
+* This can be done using a lift chart - a chart that plots the amount of profit the company earns on the y axis vs the probability cut off point on the x axis.
+For this project, the x-axis represents the point where directors would say 'no' to  a job with a probability for loss greater than that point.
+There will be an optimal point where saying no to a job above a certain probability level will mean the company earns more profits, becasue they are saying 'no' to jobs that are likely to make a loss.
+Using the figure of anticipated increase in profit, it will be easier to judge howe much the company would benefit from integrating the algorithm into decision making.
+* the performance of the lift chart will vary depending on the training and test set, so this will be done with 20 fold cross validation, so teh resulting line can be plotted with a confdience interval.
+A lift chart will be explained in more detail in the results section.
+
 ## 3.7 Conclusion/Link to next chapter
+
+* obtained data: 4169 jobs from a single case study company over 12 years. Data was de-dentified by an employee of the company
+* the data was all from the company CRM. There were three chief sources:
+    * invoicing data
+    * timesheet data
+    * project summary data
+* the data was cleaned from mistakes which were generally discovered as outliers 
+* many new variables were engineered from the data, often by summarising time series data into a value that is relevant to a final data set of a single job per row
+* variable selection: dozens of variables available which were condensed to the most important 10-12.
+variable importance methods were ANOVA, random forest, and cforests.
+the overlapping important variables were chosen first as well as some others
+* 5 models were tested: baseiline: log reg, naiive bayes.
+* sophisticated ML: boosted reg, random frest, bayesian networks
+* best were comapred to a model averaged version
+* Analysis of impact on bottom line for business if model is integrated into decision making
+* link: results
+
+# Chapter 4: Results
 

### 2016-05-20 Regression.rmd
regression report

@@
 
-library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('caret',lib='C:/Progra~1/R/R-3.2.1/library')
+library('knitr')
+library('ggplot2')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library('caret')
 # library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('car')
 # library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
@@
 
+new analysis for thesis write up - include cross validation and delete all the printing
+
+```{r, echo=FALSE}
+
+sample<- all7d[1000,]
+sample<- sample[,!is.na(sample)]
+vars<- names(sample) 
+vars<- vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc','Biz.type','Post.Code',
+               'code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto','balance.mlsto','client.totinv.log','pc.majpos')]
+vars
+
+#make new dataframe for future use with normalised variables and variables that you want to use only
+all8a<- all7d %>% select(no.users, Discipline, pc.pro, Business, code.client, JD.Second, timespan.cbrt,
+                         inv.mlsto.log,client.totinv.log, pc.majpos.log, code.contact, Billing.Type,
+                         majority.pos, return.pdol, mlsto, Year)
+write.csv(all8a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')
+
+RMSE<- function(answer, guess){
+        error= answer-guess
+        root.error= sqrt(mean(error^2))
+        return(root.error)
+}
+
+final.output<- function(df=all8a, no.tests = 50, core = TRUE) {
+
+        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)
+
+        for(i in 1:no.tests){
+
+        if(core == FALSE){
+        samp = sample.int(nrow(df), 1)                        
+        #create sample df - use all variables that are complete from a sample row of data
+        sample = df[samp,]
+        sample = sample[,!is.na(sample)]
+        vars= names(sample)
+        vars = vars[! vars %in% c('mlsto','Year','return.pdol')]
+        variables = vars
+        } else{
+                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
+                              'no.users', 'pc.pro', 'Business')
+        }
+
+        #make reduced dataset with complete cases for all variables we want to use
+        reduced = df[complete.cases(df[,variables]),]
+
+        formula = paste("return.pdol ~ ")
+
+        for(j in 1:length(variables)){
+                formula = paste(formula, variables[j], sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'inv.mlsto.log' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'code.client' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*code.client', sep= " + ")
+        }
+
+        if('JD.Second' %in% variables &
+           'no.users' %in% variables){
+                formula = paste(formula, 'JD.Second*no.users', sep= " + ")
+        }
+
+        samp_size = floor(0.75*nrow(reduced))
+        create.train = base::sample.int(nrow(reduced), samp_size)
+        train.df = reduced[create.train,]
+        test.df = reduced[-create.train,]
+
+        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
+        for(k in seq_along(cats)){
+                if(cats[k] %in% vars){
+                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
+        }
+        }
+
+        final = aov(as.formula(formula), data = train.df)
+
+        fit= summary(final)
+
+        ans = data.frame(answer= test.df$return.pdol, final.lm = predict(final, test.df))
+
+        #print RMSE for cases that match certain categories only
+        #first subset predictions to cases that match the category
+
+        test.df$predict= ans$final.lm
+
+        #random RMSE
+        r= mean(reduced$return.pdol)
+        random.RMSE= RMSE(reduced$return.pdol, r) %>% round(3)
+
+        #final RMSE for all rows in reduced in a regression
+        final.RMSE= RMSE(ans$answer, ans$final.lm) %>% round(3)
+
+        summary.tests[i,1]= random.RMSE
+        summary.tests[i,2]= final.RMSE
+        summary.tests[i,3]= nrow(test.df)
+        }
+        summary.tests = as.data.frame(summary.tests)
+        names(summary.tests) = c('random.RMSE', 'final.RMSE', 'test.nrow')
+        return(summary.tests)
+}
+
+results = final.output(df=all8a, no.tests = 100, core = FALSE)
+results = results %>% mutate(diff = random.RMSE - final.RMSE)
+results = results %>% filter(diff>-25)
+results.core = final.output(df=all8a, no.tests = 50, core = TRUE)
+results.core = results.core %>% mutate(diff = random.RMSE - final.RMSE)
+saveRDS(results, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/reg_lm.rds")
+saveRDS(results.core, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/reg_lm_core.rds")
+
+results %>% ggplot(., aes(x = test.nrow, y = final.RMSE)) + geom_point()
+
+#histogram of difference
+results %>% ggplot(., aes(x =diff)) + geom_histogram()
+
+# t- test
+t_test = t.test(diff, alternative = "greater", mu = 0)
+str(t_test)
+t_test$p.value
+
+```
+
 Check what RMSE would be for random guess

### 2016-05-20 finalwriteup/draft.rmd
regression report

--- 2016-04-27 finalwriteup/draft.rmd
@@
 date: "Tuesday, July 07, 2015"
-output: word_document
+output:
+  html_document: default
+  word_document:
+    reference_docx: mystyles.docx
 bibliography: Library.bib
 ---
+
+```{r, echo=FALSE, include=FALSE}
+
+library('knitr')
+library('ggplot2')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library("rpart")
+library('car')
+library('e1071')
+library('party')
+library('randomForest')
+library('RColorBrewer')
+library('pixiedust')
+
+# 
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
+opts_chunk$set(fig.width=8, dpi = 300)
+```
 
@@
 
-**General Aim** The aim of this project is to use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
+**General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
 Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
@@
 
+When running variable importance tests, the dependent variable must be clearly defined.
+For this problem, the dependent variable can be numeric or categorical: whether a project was profitable (binary classification) or *how* profitable a project was (regression).
+The regression model was chosen as a variable important for predicting the degree of profitability would also predict profitable vs not profitable.
+
 #### 3.3.2.1 ANOVA
@@
 
-*
+Previous studies predicting project profitability in software and construction projects were largely limited to case-based reasoning, regression and neural networks.
+There is a need to test more sophisticated machine learning algorithms that are as powerful as neural networks but provide insight into the reasoning behind predictions.
+Ensemble trees (boosted trees and random forests) and bayesian networks fit these criteria.
+Regression and naiive bayes models were also included as simple baseline models.
+Complex models should be measured against simple models that can be computed at a fraction of the cost.
+The complete list of models for this study are:
+
+* Bayesian network
+* Random forest
+* Gradient boosted trees
+* Regression - baseline model
+* Naiive bayes - baseline model
+
+Some models have limitations and required additional data processing steps such as normalising numeric variables (Bayesian Networks and Regression) and discretising continuous variables (Bayesian Networks).
+Discretising numeric variables was performed by generating a hierarchical dendrogram of each variable to visualise the clusters.
+Between four and six clusters were chosen and summarised to find the maximum and minimum values within each cluster.
+Then, each job was assigned a cluster label to replace a numeric variable.
+It was decided that for timespans, the discretised variable should be applied to all models because it is easier for a manager to predict a time category than the number of days.
+For example, a small job could be confidently assigned to less than three weeks and a large job could be assigned to 1.5 - 3 years.
+
+All methods except Gradient Boosted trees could not handle missing values.
+Preliminary runs of each method used subsets of the data without missing values.
+Then, imputing methods were trialled with gradient boosted trees since boosted trees can handle missing values.
+The MICE random forest method has been proven to work well with complex data sets [@Shah2014].
+It works by first performing a standard random forest imputation of missing values which are treated as 'place holders'.
+Then, the imputations for one variable are deleted and the remaining full variables are used to impute the single variable using random forest imputation.
+This is repeated for each variable and their 'place-holder' imputations are replaced by an imputation targeted at the single variable.
+This cycle is repeated five times by default [@Azur2011].
+If the imputed dataset gave similar predictive results using boosted trees as unimputed data using boosted trees, the imputed data must be reasonable.
+It is then trialled on the remaining methods and compared to un-imputed trials.
 
@@
 
-- need for statistical measurements
-
-## 3.6 Bottom line translation
-
-## 3.7 Conclusion/Link to next chapter
-
-## Regression
+To compare the models, mean squared error statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used for binary classification.
+AUC is a better statistic than classification accuracy in comparing algorithms when the probability of the outcome is relevant to the problem.
+This is true for problems where the decision can depend on whether the probability is above a certain threshold [@Huang2005].
+
+An ROC graph plots the false positive rate on the x-axis and true positive rate on the y-axis where:
+
+$$TPR = \frac{Count of TP}{Count of Positives}$$
+
+$$FPR = \frac{Count of FP}{Count of Negatives}$$
+
+After a model has been created using a training set, the test set is run through the model, but in order to classify the test set, a probability threshold must be chosen.
+That is, the model outputs a probability to classify a case as 0 or 1.
+The probability is not a true statistical probability, but instead a measure of the  certainty of the model.
+If a probability threshold is chosen to be 0.6, each case in the test set can then be classified as 0 or 1 depending on whether the probability calculated by the model is greater or less than 0.6.
+Once the classifications have been made, the true positive rate and false positive rate can be calculated for that model and threshold.
+This is plotted as a single point on the ROC curve, say point C.
+Refer below:
+
+![](images/method/ROC1.JPG)
+
+[@Provost2013]
+
+A model that is perfectly classified would have all positives correctly classified (1.0 True Postiive Rate) and no incorrectly classified positives (0.0 False Positive Rate).
+If a model has a 1.0 TPR and 1.0 FPR (top right corner) this means it is correctly classifying all positives at the expense of incorrectly classifying all negatives as positive.
+It is 'dumbly' classifying all cases as positive.
+if a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previosu example.
+There is an 80% chance any case will classified positive.
+Therefore, models that lie on the diagonal line classify cases as well as random chance.
+
+To plot the curve in an ROC, points are plotted for results on the training set using numerous threshold probability values between 0 and 1.
+Once enough points are plotted, a curve can be drawn.
+The closer the curve reaches the top left hand corner, the closer the algorithm is to perfectly predicting positive and negative cases.
+This would give an area under the curve (AUC) of 1.
+AUC of between 0.5 and 1 means the model is performing better than random chance for all possible probability thresholds.
+
+![](images/method/ROC2.JPG)
+
+[@Provost2013]
+
+In order to compare whether models perform significantly better than one another, the AUC or mean squared error results from multiple-fold cross validations were used as the data points.
+Once the best two or three models were determined, these were combined using model averaging.
+Finally, the best model is determined between the single models and model averaged model.
+
+## 3.6 Bottom Line Translation
+
+Although predictive power of the final model is important, for the method to be integrated into a company's decisions, the affect on a business' bottom line should be tested.
+This was done using a profit curve - a chart that plots the amount of profit the company earns on the y axis vs the probability threshold on the x axis.
+
+A simple approach was taken for this analysis, where projects with a probability to be a loss-making job greater than the threshold were rejected entirely.
+Therefore all profits and losses from jobs above threshold were discounted.
+If the threshold is zero, all jobs are rejected and the profit is $0.
+If the threshold is 1.0, all jobs are accepted and the profit is the same as the profit the company actually experienced since the data is a sample of historic projects.
+The aim was to find the optimal threshold point where saying no to a job above that level will mean result in higher profits, becasue jobs that are likely to make a loss are being rejected.
+This chart will clarify what percentage of profit increase the company can expect by integrating the algorithm into decision making.
+
+A profit curve is made from a single train/test instance of the data.
+In order to understand the uncertainty around the profit curve, hundred-fold train/test splits were run and results calculated for each model and threshold point.
+This resulted in a statistical confidence interval around each threshold point which can be represented graphically.
+The final profit curve and surrounding confidence interval will indicate whether the algorithm makes a compelling case for integration into the company's decision making process.
+
+## 3.7 Method Conclusion
+
+In summary, to determine whether a 
+* obtained data: 4169 jobs from a single case study company over 12 years. Data was de-dentified by an employee of the company
+* the data was all from the company CRM. There were three chief sources:
+    * invoicing data
+    * timesheet data
+    * project summary data
+* the data was cleaned from mistakes which were generally discovered as outliers 
+* many new variables were engineered from the data, often by summarising time series data into a value that is relevant to a final data set of a single job per row
+* variable selection: dozens of variables available which were condensed to the most important 10-12.
+variable importance methods were ANOVA, random forest, and cforests.
+the overlapping important variables were chosen first as well as some others
+* 5 models were tested: baseiline: log reg, naiive bayes.
+* sophisticated ML: boosted reg, random frest, bayesian networks
+* best were comapred to a model averaged version
+* Analysis of impact on bottom line for business if model is integrated into decision making
+* link: results
+
+# Chapter 4: Results
+
+This section presents key charts and tables of results that progressively answer the research question: can a statistical or machine learning model based on historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
+
+## Variable Selection
+
+To create the best possible model, only important variables were selected to reduce the affects of noisy independent variables.
+There are currently 46 independent variables for the single dependent variable, return per dollar which is the measure of profitability for a project.
+
+### Regression
+
+Regression requires complete datasets.
+Therefore, analysis was done with a core set of 13 complete variables and a single additional incomplete variable.
+That means a model was made for each incomplete variable, resulting in 21 separate ANOVA models.
+The p-values for the F statistic of each variable is plotted in the boxplot below.
+
+```{r, echo=FALSE, fig.height=5}
+
+plot.pvals = readRDS("AOV_varimp_plot.rds")
+
+ggplot(plot.pvals, aes(x = mod.vars, y = p.val)) + geom_boxplot() +
+        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+        geom_hline(yintercept = 0.05, colour = 'red', linetype = 2) +
+        scale_y_continuous(breaks = c(0, 0.05, .25, .5, .75, 1)) +
+        labs(y = "P value of F statistic", x = "Independent Variables",
+             title = "Summary of P-values from Anova tests")
+
+```
+
+The ANOVA model indicates the following variables have a median p-value below 0.05 and therefore significantly contribute to the rejection of the null hypothesis - the variables do not correlate linearly with the dependent variable, return per dollar.
+
+* number of employees that contributed to the project (no.users)
+* Percent of hours completed by a professional as opposed to technical employee (pc.pro)
+* Business category of Client (Business) 
+* Discipline
+* Number of internal disciplines involved in the project (Num.disc)
+* Timespan of the project (timespan)
+* Total amount invoiced for the project (inv.mlsto)
+
+Four of the added incomplete variables (categorised as add.variable) were significant in their ANOVA models:
+
+* Client id (code.client)
+* Project category (JD.Second)
+* Position of the main professional working on the project, ie mid-level, senior (ProjEng.Pos)
+* id of main professional working on the project (code.ProjEng)
+
+### Random Forest
+
+The random forest algorithm imputes missing values, however this was not desirable for the early stage of variable selection - before modelling had commenced.
+Instead, a subset of complete data was used made up of variables with at least 2300 complete cases.
+Additionally, random forests cannot handle categorical variables with more than 53 categories which eliminated more variables.
+The resulting 'core' dataset contained 16 variables and 2364 cases.
+Because each run of the algoritm is slightly different, 10 algorithms were run, which each output the ranking of variables by importance.
+The mean ranking of each variable is summarised below:
+
+```{r, echo = FALSE}
+
+rf.rank = readRDS("RF_var_rank_subbed.rds")
+kable(rf.rank, format = 'markdown')
+
+```
+
+### cForest
+
+Cforests were also run with a core of 13 complete variables, and incomplete variables were given added to the core variables individually in separate runs.
+Variables with an unlimited number of c ategories can be included in cforests which is an advantage over randomForests.
+Cforests compute a variable importance ranking which is a number relating to the reduction in error the variable provides.
+The variable importance rankings for the core 13 variables are as follows:
+
+```{r, echo = FALSE}
+
+cfor.varimp = readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/cfor_varimp.rds")
+
+ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        geom_bar(stat='identity') +
+        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance rankings of Complete Variables")
+
+```
+
+Some incomplete variables that were added one at a time ranked in the top 6 most important variables in their model:
+
+* Percent of Hours completed by the main employee on the project (pc.majpos)
+* Client id (code.client)
+* Client contact id (code.contact)
+* Project category (JD.Second)
+
+### Variable selection summary
+
+Based on the results of the three methods for variable importance, 11 independent variables were selected for the prediction modelling:
+
+* Percent of hours completed by a professional level employee (pc.pro) -all models
+* Project timespan (timespan) - all models
+* Number of employees on Project (no.users) - all models
+* Amount Invoiced for Project (inv.mlsto) - all models
+* Business category of Client (Business) - moderately for all models
+* Project category (JD.Second) - ANOVA and cForest
+* Total Amount Invoiced from Past Client Jobs (clint.totinv) - randomForest
+* Discipline - ANOVA
+* position of main employee on the project (majority.pos) - newly created to cater for findings in ANOVA
+* Percent of Hours completed by the main employee on the project (pc.majpos) - cForest
+* Billing Type (hourly vs. fixed fee) - advice of Case Study partner
+
+## Model Results
+
+Once the independent variables were narrowed down to a succinct list, the prediction models were tested.
+First prediction return per dollar as a regression problem was trialled followed by prediction profit or loss as a binary classification problem.
+
+### Regression Model
+
+#### ANOVA
+
+In ANOVA, linear relationships are tested and normal distributions are assumed for numeric variables. Therefore, where appropriate, numeric variables were norma
+
+* normalised variables for linear model:
+    * log of amount invoiced, total invoiced for client in past projects, % of work done by majority position
+    * cube root of timespan
+
+* trials with different variable interactions resulted in three interactions that lowered RMSE
+
+* Baseline check: check RMSE from predictions against RMSE if use mean
+
+* ran ANOVA 50 times with core important and largley complete variables
+        * amount invoiced
+        * Discipline
+        * timespan
+        * team size
+        * pc by professional
+        * Business category
+* subtract RMSE from model predictions from RMSE using mean return per dollar. If model effective, RMSE is lower for model predictions, and difference is greater than 0. Histogram of difference:
+
+```{r, echo = FALSE}
+
+results.core = readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/reg_lm_core.rds")
+
+```
+
+Student t test to test whether mean difference is significantly above 0 gave p-value of ___. Reject null hypothesis that difference is lses than or equal to zero. Difference is ___.
+
+#you are here
+
+* continue with the bagging method for ANOVA
+* then randomForest results
+
+#### Linear Regression and Random Forest
     * tried predicting return per dollar as a numeric outcome.
         * proved very difficult. Find error rate compared to just guessing the mean. Not far off.
-
-##Bayesian Network
+    * table of results for random forest and linear regression - MSE
+
+### Classification Models
+
+* text on required imputing
+
+* chart showing AUC's for all methods on one plot : )
+* table showing significance p value numbers
+
+* other results from methods?? 
+* Bayesian network
+* Random forest
+* Gradient boosted trees
+* Regression - baseline model
+* Naiive bayes - baseline model
+
+### Bayesian Network
 
@@
 
-* cannot predict return per dollar. may be that this is much to specific for the volume of data available, may be that the project variation is not captured by the type of data available
+## Variable Importance
+
+The variable importance results from the three types of models: ANOVA regression, randomForest, and cForest were largely similar. All three models ranked five variables highly:
+
+* percent of hours by professional
+* project timespan
+* number of employees in project team
+* total amount invoiced
+* business category of client which was positive
+
+This was not necessarily expected particularly between ANOVA and the ensemble tree methods as ANOVA measures linear relationships as opposed to non-linear ensemble trees.
+Random forests favour variables that are numeric or have many categories and this held true for the analysis.
+The 6 highest ranked variables were numeric and one categorical variable, Business category of client, with 28 categories.
+The two lowest ranking variables had only 4 categories.
+It would be expected that cForests would rank the categorical variables with only 4 categories more fairly which was the case with Discipline.
+It was ranked mid way.
+Business category of the client and Broad business type however, were ranked similarly in randomforests.
+
+The nature of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, general client characteristics and id, project timespan, and project category.
+It was
+expected that Billing Type would play a more important role in predicting project profitability as it describes the nature fees are charged.
+For this reason, and with advice from the case study business, the Billing Type was temporarily kept in the model.
+
+For construction projects and IT software projects, the literature indicated size was the most important variable predicting effort.
+Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996][@Finnie1997][@Pai2013].
+In this case study, profitability is the dependent variable, not total effort and the size variable plays a different role.
+Instead of contributing to predicting the total effort, it is describing whether the size of a project correlates with how profitable the project is.
+The closest estimate of size in this case study is the invoiced amount which does show up as very important.
+The exact invoiced amount cannot be known before a project begins, as this would solve the problem of profitability.
+Therefore, the invoiced amount was binned into categories that a manager could choose as an estimate.
+The other important variables (including number of employees on a project, client business type and percent of hours by certain types of employees) have not been strongly represented in the literature to date.
+
+## Regression
+
+* cannot predict return per dollar. may be that this is much too specific for the volume of data available, may be that the project variation is not captured by the type of data available
 * simpler task of predicting profitability or not does provide predictive accuracy well above random chance and also above our base measure - naiive bayes, but maybe not significantly?

### 2016-05-20 var_select.rmd
regression report

@@
 
-library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
-library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('RColorBrewer',lib='C:/Progra~1/R/R-3.2.1/library')
+library('knitr')
+library('ggplot2')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library("rpart")
+library('car')
+library('e1071')
+library('corrgram')
+library('party')
+library('randomForest')
+library('RColorBrewer')
 
@@
             model= aov(as.formula(formula)              
-                      ,data= data)
+                      , na.action = na.omit, data= data)
             return(summary(model)) 
@@
                 model=lm(as.formula(formula)                     
-                        ,data= data)
+                        , na.action = na.omit, data= data)
                 df<- as.data.frame(summary(model)$coefficients)
@@
                 colnames(df)[names(df) %in% 'Pr(>|t|)']<-'Pval'
-                return(df %>% arrange(-abs(Estimate)))
+                return(df %>% arrange(Pval))
 
@@
 BT.lin(type= 'lm', data= all7a, add.var = 'Job.Type.Primary') %>% slice(1:20)
-BT.lin(type= 'lm', data= all7a[1314:2283,], add.var = 'Job.Type.Primary') %>% slice(1:10)
 BT.lin(type= 'aov', data= all7a, add.var = 'Job.Type.Primary')
-BT.lin(type= 'aov', data= all7a[1314:2283,], add.var = 'Job.Type.Primary')
+BT.lin(type= 'aov', data= all7a, add.var = 'Billing.Type')
+
+#loop to extract p values from anova of all additional variables.
+
+#incomplete variabless:
+incomp.vars = c('Post.Code', 'Billing.Type', 'Job.Source', 'Job.Detail.Primary', 'Job.Detail.Secondary', 'Job.Type.Primary',
+                'Job.Type.Secondary', 'State', 'Type', 'no.employees', 'contact.count', 'JD.Primary', 'JD.Second', 'dist',
+                'timespan', 'Num.disc', 'Inv.freq', 'client.invfreq')
+plot.pvals = NULL
+
+for(i in seq_along(incomp.vars)){
+        formula = paste("return.pdol ~ inv.mlsto + Discipline + client.count + Business +
+                              Biz.size + Biz.type + Year + Num.days +
+                              no.users + pc.contracttech + client.neginv + client.numinv + client.totinv +
+                              pc.director + pc.midtech + pc.midpro + pc.gradpro + pc.seniortech + pc.seniorpro +
+                              pc.pro + mean.peeps + hours.perday + num.inv + mean.inv + num.neginv + client.meaninv +
+                              code.director + ProjEng.Pos", incomp.vars[i], sep="+")
+
+        model= aov(as.formula(formula)              
+                      , na.action = na.omit, data= data)
+
+        #create columns for results and plotting
+        mod.vars = summary(model)[[1]] %>% rownames %>% .[-29]
+        mod.vars = sub("\\s+$", "", mod.vars)
+        mod.vars = sub(incomp.vars[i], "add.variable", mod.vars)
+        p.val = summary(model)[[1]]$'Pr(>F)' %>% .[-29]
+        add.variable = rep(incomp.vars[i], 28)
+
+        plot.pvals = rbind(plot.pvals, 
+                data.frame(mod.vars, p.val, add.variable)
+        )
+
+}
+
+#order levels in plot.pvals by median
+order.vars = ddply(plot.pvals, .(mod.vars), summarise, med = median(p.val))
+order.vars = arrange(order.vars, med)
+plot.pvals <- within(plot.pvals, 
+                         mod.vars <- factor(mod.vars, 
+                                               levels=order.vars$mod.vars))
+saveRDS(plot.pvals, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/AOV_varimp_plot.rds")
+
+gg <- ggplot(plot.pvals, aes(x = mod.vars, y = p.val)) + geom_boxplot() +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) + 
+        geom_hline(yintercept = 0.05, colour = 'red', linetype = 2) +
+        scale_y_continuous(breaks = c(0,0.05,.25,.5,.75,1)) +
+        labs(y = "P value of F statistic", x = "Independent Variables", 
+             title = "Summary of P-values from Anova tests")
+#which added variable was sgnificant?
+plot.pvals %>% filter(p.val <=0.05, mod.vars == "add.variable")
+# timespan
 
@@
 * Biz.type
+* hours.perday - 0.000455
+* code.director - 0.00118
+* pc.midtech - 9.11e-07
+* no.users = 0.000691
+* Num.days = 0.039480
 
@@
 
-write.csv(all7b,'all7b.csv')
+# write.csv(all7b,'all7b.csv')
 all7b<- read.csv('all7b.csv')[,-1]
@@
 
+saveRDS(var.imp, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/cfor_varimp.rds")
+
 q<- ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
@@
 save(all.cfor2,file= 'C:/Users/n9232371/Documents/Consultbusiness/data/objects/all_cfor2.RData')
-# load('C:/Users/n9232371/Documents/github/consult/object/all_cfor2.RData')
+load('C:/Users/n9232371/Documents/Consultbusiness/data/objects/all_cfor2.RData')
 
@@
 rand.rank<- multi.forest1(data=mini.core3, nforest=1:10)
+rand.rank$var <- sub("inv.mlsto", "Amount Invoiced for Project", rand.rank$var)
+rand.rank$var <- sub("no.users", "Number of employees on Project", rand.rank$var)
+rand.rank$var <- sub("client.totinv", "Total Amount Invoiced from the Client - Past Jobs", rand.rank$var)
+rand.rank$var <- sub("pc.pro", "% of Hours Completed by a Professional-level Employee", rand.rank$var)
+rand.rank$var <- sub("Business", "Business Category of Client", rand.rank$var)
+rand.rank$var <- sub("client.numinv", "Mean Number of Invoices Submitted to Client - past Jobs", rand.rank$var)
+rand.rank$var <- sub("client.meaninv", "Mean Invoice Size for Client - Past Jobs", rand.rank$var)
+rand.rank$var <- sub("client.count", "Number of Past Jobs with Client", rand.rank$var)
+rand.rank$var <- sub("code.director", "Director", rand.rank$var)
+rand.rank$var <- sub("Num.disc", "Number of Internal Disciplines Involved", rand.rank$var)
+rand.rank$var <- sub("client.neginv", "Number Bad Debt Client Invoices - Past Jobs", rand.rank$var)
+rand.rank$var <- sub("Biz.size", "Client Size in Number of Employees", rand.rank$var)
+rand.rank$var <- sub("Biz.type", "Broad Client Business Category (Gov, Private etc)", rand.rank$var)
+rand.rank = rename(rand.rank, Variable_Description = var, Rank = m.rank)
+
+saveRDS(rand.rank, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/RF_var_rank.rds")
+saveRDS(rand.rank, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/RF_var_rank_subbed.rds")
+rand.rank <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/RF_var_rank.rds")
+rf.rank <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/RF_var_rank_subbed.rds")
 
@@
 cfor.plotsave<- function(extra= vars, name= 'vare'){
-        for(i in 1:length(vars)){
+        for(i in 1:length(extra)){
 
@@
 cfor.plotsave(extra=list('majority.pos'), name='vare')
+cfor.plotsave(extra=list('pc.majpos'), name='varf')
 
@@
             'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'dist','Post.Code',
-            'code.contact','contact.count', 'Job.Type.Primary','JD.Primary','ProjEng.Pos', 'code.ProjEng', 'pos.tech',
-            'majority.pos', 'pc.majpos')
+            'code.contact','contact.count', 'Job.Type.Primary','JD.Primary','ProjEng.Pos', 'code.ProjEng',
+            'pos.tech', 'majority.pos', 'pc.majpos')
 
 aov.loop<- function(extra=vars.vec){
-        for (i in 1:length(extra)){
-                core.var= all7c[,names(all7c) %in% c('inv.mlsto', 'timespan', 'no.users', 
-                                                     'Discipline', 'pc.pro', 
-                        'client.totinv', 'Biz.type','code.director', 'Num.disc', 
-                        'Business', 'return.pdol', extra[i])]
-                core.var = core.var[!is.na(core.var[,extra[i]]),]
-
-                formula = paste("return.pdol ~ inv.mlsto + Discipline + Business+
-                        Biz.type + Num.disc +
-                        no.users + 
-                        client.totinv + 
-                        pc.pro + timespan +
-                        code.director", extra[i], sep="+")
-
-                model= aov(as.formula(formula)              
-                           ,data= core.var)
-
-                print(extra[i])
-                print(model %>% summary)
-
-                cat(i," ")
-                }
-        }
+        plot.pvals = NULL
+        for (i in 1:length(extra)) {
+                core.var = all7c[, names(all7c) %in% c('inv.mlsto','timespan','no.users','Discipline',
+                                                       'pc.pro','client.totinv','Biz.type',
+                                                       'code.director','Num.disc','Business',
+                                                       'Biz.size', 'pc.director',
+                                                       'return.pdol', extra[i])]
+
+                core.var = core.var[!is.na(core.var[, extra[i]]),]
+
+                formula = paste("return.pdol ~ inv.mlsto + Discipline + Business+Biz.type + Num.disc +
+                no.users + client.totinv + pc.pro + timespan + code.director +
+                Biz.size + pc.director", extra[i], sep = "+")
+
+                model = aov(as.formula(formula),data = core.var,na.action = na.omit)
+
+                #create columns for results and plotting
+                mod.vars = summary(model)[[1]] %>% rownames %>% .[-14]
+                mod.vars = sub("\\s+$", "", mod.vars)
+                mod.vars = sub(extra[i], "add.variable", mod.vars)
+                p.val = summary(model)[[1]]$'Pr(>F)' %>% .[-14]
+                add.variable = rep(extra[i], 13)
+
+                plot.pvals = rbind(plot.pvals, data.frame(mod.vars, p.val, add.variable))
+        }
+
+        return(plot.pvals)
+}
+
+plot.pvals = aov.loop(extra=vars.vec)
+
+#order levels in plot.pvals by median
+order.vars = ddply(plot.pvals, .(mod.vars), summarise, med = median(p.val))
+order.vars = arrange(order.vars, med)
+plot.pvals <- within(plot.pvals, 
+                         mod.vars <- factor(mod.vars, 
+                                               levels=order.vars$mod.vars))
+saveRDS(plot.pvals, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/AOV_varimp_plot.rds")
+
+gg <- ggplot(plot.pvals, aes(x = mod.vars, y = p.val)) + geom_boxplot() +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) + 
+        geom_hline(yintercept = 0.05, colour = 'red', linetype = 2) +
+        scale_y_continuous(breaks = c(0,0.05,.25,.5,.75,1)) +
+        labs(y = "P value of F statistic", x = "Independent Variables", 
+             title = "Summary of P-values from Anova tests")
+#which added variable was sgnificant?
+plot.pvals %>% filter(p.val <=0.05, mod.vars == "add.variable")
+plot.pvals %>% filter(add.variable == "pc.majpos")
+# timespan
 

### 2016-05-20 imputingall9c.R

@@ -1,6 +1,6 @@
 # impute missing values
 
-library('mice', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('mice')
 library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
 library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
 library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
@@ -90,7 +90,7 @@ auc
 
 
 
-#experiment with MICE package
+#experiment with MICE package - should not include return per dollar during imputation!
 
 new <- mice(all9c[,c(1:4,6:10,12:14,16)], meth = 'rf')
 

### 2016-05-20 finalwriteup/draft.rmd
minor update to regression report

@@
 
-In ANOVA, linear relationships are tested and normal distributions are assumed for numeric variables. Therefore, where appropriate, numeric variables were norma
-
-* normalised variables for linear model:
-    * log of amount invoiced, total invoiced for client in past projects, % of work done by majority position
-    * cube root of timespan
-
-* trials with different variable interactions resulted in three interactions that lowered RMSE
-
-* Baseline check: check RMSE from predictions against RMSE if use mean
-
-* ran ANOVA 50 times with core important and largley complete variables
-        * amount invoiced
-        * Discipline
-        * timespan
-        * team size
-        * pc by professional
-        * Business category
-* subtract RMSE from model predictions from RMSE using mean return per dollar. If model effective, RMSE is lower for model predictions, and difference is greater than 0. Histogram of difference:
+In ANOVA, linear relationships are tested and normal distributions are assumed for numeric variables.
+Therefore, where appropriate, numeric variables were normalised via log or cube root transformations.
+Trials of various variable interactions revealed that interactions between:
+
+* total amount invoiced * percent of job by professional level employee
+* total amount invoiced * client id
+* project category * number of employees on the job
+
+were statistically significant.
+Two types of ANOVA trials were run.
+The first used only a 'core' list of 6 variables that were largely complete and any relevant interaction terms.
+The second method was developed to get around the problem that when all variables are combined, only 15% of the projects have complete data.
+Therefore the ANOVA test would eliminate 85% of the data and imputing the data was not desirable at an early stage due to the amount of missing categorical data.
+The second method would choose a project (row in the dataset) randomly, and this case would have certain columns complete.
+The data was then filtered for all jobs that had at least the same columns complete, and the ANOVA test was run on this reduced set.
+This method allowed variables with a larger proportion of missing data to be analysed while not attempting to analyse all variables at once.
+
+To assess the results of the ANOVA models, the root mean squared error (RMSE) of the return per dollar predictions were compared against the RMSE if all predictions were simply the mean of return per dollar for all projects.
+Then, the RMSE's from model predictions were subtracted from RMSE's using mean return per dollar.
+If the models were effective, RMSE should be lower for model predictions, and the difference would be greater than 0.
+Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only:
 

### 2016-05-23 Regression.rmd
minor update to regression report

@@
 library('caret')
+library('pwr')
 # library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
@@
         if('inv.mlsto.log' %in% variables &
-           'inv.mlsto.log' %in% variables){
+           'pc.pro' %in% variables){
                 formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
@@
 
+#how many tests required? power calculation
+d.calc = mean(results.core$diff)/sd(results.core$diff)
+pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05, power = .8, type = c("one.sample"), alternative = 'greater')
+# 3 tests required for results.core
+
 results = final.output(df=all8a, no.tests = 100, core = FALSE)
@@
 # t- test
-t_test = t.test(diff, alternative = "greater", mu = 0)
+t_test = t.test(results$diff, alternative = "greater", mu = 0)
 str(t_test)

### 2016-05-23 finalwriteup/draft.rmd

@@
 output:
-  html_document: default
   word_document:
     reference_docx: mystyles.docx
+  html_document: default
 bibliography: Library.bib
@@
 library('RColorBrewer')
-library('pixiedust')
-
-# 
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+library('pwr')
+
+# detect OS
+if(.Platform$OS.type == 'windows'){
         # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
         opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
-opts_chunk$set(fig.width=8, dpi = 300)
+} else{
+        # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+        opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/report_data')
+}
+
+opts_chunk$set(fig.width=8, fig.depth = 5, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
 ```
@@
 The mathematical methods used to predict project costs vary from statistical regressions to deep-learning neural networks [@Love2005][@Kim2004].
-Most studies report neural networks are more accurate than regression, and generally the studies report positive predictive powers of their developed models.
+Most studies report neural networks are more accurate than linear regression, and generally the studies report positive predictive powers of their developed models.
 
@@
 A notable amount of literature has studied the predictive power of multiple linear regression in the construction cost estimation problem.
-Often, regression is the only model assessed, without comparison to other methods such as neural networks, which first started appearing in literature in the 1990's [@Kim2004].
-However, even post-1990, many studies focused on regression only.
+Often, linear regression is the only model assessed, without comparison to other methods such as neural networks, which first started appearing in literature in the 1990's [@Kim2004].
+However, even post-1990, many studies focused on linear regression only.
 This may be because the technique is straightforward, easy to use and widely available in statistical packages [@Chan2005].
@@
 
-Despite neural networks typically outperforming regression models, it is intersting to note some of the outcomes of regression studies in cost estimation.
+Despite neural networks typically outperforming linear regression models, it is intersting to note some of the outcomes of regression studies in cost estimation.
 Different studies tend to produce a wide range of variables that contribute to different degrees in the models.
@@
 
-A study by @Kim2004 compared CBR to neural networks and regression in the construction cost estimation problem.
+A study by @Kim2004 compared CBR to neural networks and linear regression in the construction cost estimation problem.
 For CBR, the ESTEEM software package was used which algorithmically calculated the similarity of variables, weighted the variables using gradient descent, and deduced the cost using the most similar cases (the specific method for this last step was not explained).
@@
 
-cfor.varimp = readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/cfor_varimp.rds")
+cfor.varimp = readRDS("cfor_varimp.rds")
 
@@
 
-* total amount invoiced * percent of job by professional level employee
-* total amount invoiced * client id
-* project category * number of employees on the job
+* total amount invoiced 
+* percent of job by professional level employee
+* total amount invoiced 
+* client id
+* project category 
+* number of employees on the job
 
@@
 
+```{r insert_data}
+# load function for determining number of required samples for 0.8 power
+n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
+        d.calc = abs(mean(data))/sd(data)
+        pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05, power = .8, type = c(type.calc), alternative = alt)
+        return(ceiling(pow_pow$n))
+}
+
+results.core = readRDS("reg_lm_core.rds")
+# n.samples(data = results.core$diff)
+
+results.A2 = readRDS("reg_lm.rds")
+results.A2 <- dplyr::filter(results.A2, diff>-15)
+# n.samples(data = results.A2$diff)
+
+```
+
 To assess the results of the ANOVA models, the root mean squared error (RMSE) of the return per dollar predictions were compared against the RMSE if all predictions were simply the mean of return per dollar for all projects.
@@
 If the models were effective, RMSE should be lower for model predictions, and the difference would be greater than 0.
-Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only:
+Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only.
+Note, only `r n.samples(data = results.A2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
 
@@
 
-results.core = readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/reg_lm_core.rds")
+#plot histogram to show differences
+results.core %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models- core variables only")
 
@@
 
-Student t test to test whether mean difference is significantly above 0 gave p-value of ___. Reject null hypothesis that difference is lses than or equal to zero. Difference is ___.
-
-#you are here
-
-* continue with the bagging method for ANOVA
-* then randomForest results
-
-#### Linear Regression and Random Forest
-    * tried predicting return per dollar as a numeric outcome.
-        * proved very difficult. Find error rate compared to just guessing the mean. Not far off.
-    * table of results for random forest and linear regression - MSE
+A student t test indicates the difference is significantly above 0 with a p-value of `r t.test(results.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
+Therefore, the null hypothesis that the difference is less than or equal to zero can be rejected.
+In other words, the ANOVA model improved estimates of return per dollar over using the mean return per dollar to a statistically significant degree.
+However, the mean difference is only $`r mean(results.core$diff) %>% format(digits = 1)`.
+Furthermore, the mean error in return per dollar is $`r mean(results.core$final.RMSE) %>% format(digits = 2)` which is unacceptably high.
+
+Results from 100 models using method 2, which captures incomplete variables, are shown below.
+
+```{r, echo = FALSE}
+
+#plot histogram to show differences
+results.A2 %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models - all variables sampled")
+
+```
+
+A student t test indicates the difference is not significantly above 0 which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
+This method clearly performed poorly in comparison to the core variables.
+A reason for this could be that the mean number of cases in the dataset was `r results.core$test.nrow[1]*4` for the core variables but averaged only `r (mean(results.A2$test.nrow)*4) %>% round(0)` for the sampled set of all variables.
+Regression results from ANOVA models were poor - randomForest was trialled next.
+
+#### randomForest
+
+```{r load_rf_results}
+results.rf.core = readRDS("reg_rf_core.rds")
+# n.samples(data = results.rf.core$diff)
+
+results.rf2 = readRDS("reg_rf.rds")
+# n.samples(data = results.rf2$diff)
+
+```
+
+Numeric predictions of return per dollar were attempted first with core variables that were largely complete followed by a method of sampling a mix of core and less complete variables in every run.
+The randomForest parameters were first tuned using the caret package and found optimal values mtry = 5 and ntree = 500 [@caret].
+A density plot of 50 models of the core variables using different train/test sets is shown below.
+The x-axis measures the decrease in RMSE obtained by using the randomForest model over simply the mean of all project return per dollar values.
+Note, only `r n.samples(data = results.rf2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
+
+```{r, echo = FALSE}
+
+#plot histogram to show differences
+results.rf.core %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 randomForest models- core variables only")
+
+```
+
+This plot shows a mean return per dollar RMSE improvement of $`r mean(results.core$diff) %>% format(digits = 1)` which is the same as the ANOVA model.
+The p-value for the null hypothesis that the difference is less than or equal to zero is `r t.test(results.rf.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)` and is rejected.
+Results from 100 randomForest models using method 2 are shown below:
+
+```{r, echo = FALSE}
+
+#plot histogram to show differences
+results.rf2 %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models - all variables sampled")
+
+```
+
+The p-value for these models and the afore-mentioned null hypothesis is `r t.test(results.rf2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
+The null hypothesis is rejected which contrasts the parallel ANOVA test.
+On the right hand side, of the graph above, the RMSE is improved by almost $0.20 but also worsens by over $0.10 per dollar.
+A density plot of the RMSE for the models is shown below:
+
+```{r, echo = FALSE}
+
+#plot histogram of RMSE's
+results.rf2 %>% ggplot(., aes(x =final.RMSE, y = ..density..))+ geom_histogram() + geom_density() +
+        labs(x = "RMSE", title = "Density Plot of RMSE for 50 randomForest models - all variables sampled")
+
+```
+
+Although the randomForest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for return per dollar was wide and the mean was high ($`r mean(results.rf2$final.RMSE) %>% round(2)`).
+
+The two regression models predicting return per dollar resulted in high mean RMSE's ranging from $`r mean(results.rf.core$final.RMSE) %>% round(2)` to $`r mean(results.A2$final.RMSE) %>% round(2)`.
+Because of the poor results, further development of regression models was postponed in favour of investiation of binary classification models.
+The classification is a simpler task than regression because it requires less precision.
 
@@
 
-* text on required imputing
-
-* chart showing AUC's for all methods on one plot : )
-* table showing significance p value numbers
+Structure:
+
+* boosted trees - auc with a few iterations - cross val
+
+* AUC for random forest, everything else without imputing?? thats what i did
+    * boxplot showing AUC for all methods. how many runs, 20?
+* only take best performing models and try with imputed data
+
+* imputing - boosted trees still fine
+    * apply to other methods
+
+* boxplot chart showing AUC's for all methods on one plot : )
+
+* table showing significance p value numbers - in comparison to what? the best performing one?
 
@@
 ```{r, echo=FALSE, include=FALSE}
-library('magrittr',lib='C:/Progra~1/R/R-3.2.2/library')
-
+library('magrittr')
+
+if(.Platform$OS.type == 'windows'){
         all8c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8c.csv' )
+} else{
+        all8c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all8c.csv')
+}
 
@@
 
-* cannot predict return per dollar. may be that this is much too specific for the volume of data available, may be that the project variation is not captured by the type of data available
-* simpler task of predicting profitability or not does provide predictive accuracy well above random chance and also above our base measure - naiive bayes, but maybe not significantly?
-question of whether the 0.05 significance is appropriate
-* how many trials do I need to do to make it significant?!
+Both ANOVA and randomForest models predicted at best within an average of $0.48 for the dependent variable, return per dollar using either core variables or a sample of core and less complete variables.
+RandomForest predicts much better than ANOVA with a mix of core (almost complete) and incomplete variables while randomForest and ANOVA perform similarly with the core variables only.
+There is much less data available for training the models that include less complete variables so it is interesting that randomForest performed much better (RMSE of $0.50 for randomForest vs.
+$2 for ANOVA).
+The main difference between ensemble tree methods and ANOVA is that ANOVA assumes linear relationships between variables (some of which are normalised) whereas trees are capable of capturing non-linear relationships.
+It could be the case that the less complete variables had non-linear relationships with the dependent variable.
+RandomForest is also an ensemble method while ANOVA is not.
+Each randomForest is the average of 500 single trees, whereas each ANOVA model is simply a single model.
+50 randomForest models produces 500 times more models than 50 ANOVA models so this averaging effect may be effectively eliminating outliers for better overall performance.
+
+The RMSE can be evaluated in terms of the end-user's perspective (business manager who is proposing fees for new projects).
+There is a record of how previous projects have performed in terms of return per dollar.
+Refer below:
+
+```{r}
+if(.Platform$OS.type == 'windows'){
+        all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv' )
+} else{
+        all8a<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all8a.csv')
+}
+ggplot(all8a, aes(x = return.pdol, y = ..density..)) + geom_histogram() + geom_density() +
+        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects")
+
+```
+
+The 95% confidence interval for this data is $`r (mean(all8a$return.pdol) - qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)` to $`r (mean(all8a$return.pdol) + qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)`.
+If a project returns $0 for every dollar spent (return per dollar = $0), the project has earned only enough to cover the cost of doing the project (salary costs and business costs).
+If the return per dollar is -$0.50, the project has earned enough to cover half of the costs of doing the job.
+Finally, if the return per dollar is $1, the project has earned double what it cost to perform.
+If predictions have an average return per dollar error of $0.50, this represents an error in revenue of half what the project cost to complete.
+Clearly, this is not informative enough to bring into business decision making processes.
+
+Further regression methods could have been persued, however randomForest is a high-performing algorithm and was far from providing usable results.
+It is likely the case that either not enough data was available or that the given variables do not capably explain the level of return per dollar.
+Each row of data is a project, and takes the company months, sometimes years, to create a single row of data.
+Therefore, more data cannot be accessed or easily created.
+It was decided categorical predictions should be pursued next as predicting whether a job will be profitable or not is a simpler task and could more effectively leverage the limited data.
+
+Case studies in the literature tend to predict total cost/effort which is a regression problem (on a numeric scale) [@Kim2004][@Chan2005][@Akintoye2000].
+This is likely because the literature dealt with large projects which impact governments and large companies, and a simple profit or loss prediction is not useful.
+A regression prediction is not feasible for the data in this project, but a profit/loss binary prediction may be useful for this company because they complete hundreds of small projects a year.
+If binary classification is successful, the method the prediction contributes to company decisions must then be explored.
+
+## Binary Classification
+
 * best models are gradient boosted trees and logistic regression. why? can hypothesis but not really

### 2016-05-25 boost_tree.rmd
Aleysha commit

@@
 ```
+
+Try discretising total amount invoiced for the job.
+
+Try tuning with caret package including Business and all years
+
+```{r, echo=FALSE}
+
+#decide how many clusters
+x<- dist(all9c$inv.mlsto.log, method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+
+# optimise number of clusters
+clusts = c(4, 5,7,9)
+for(j in seq_along(clusts)){
+
+all9c$b.inv.log<- as.factor(cutree(fit, k= clusts[j]))
+
+# set seeds
+set.seed(123)
+seeds <- vector(mode = "list", length = 51)
+for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
+## For the last model:
+seeds[[51]] <- sample.int(1000, 1)
+
+#create test train
+sample<- sample(1:nrow(all9c), 2/3*nrow(all9c), replace=F)
+train<- all9c[sample,]
+test<- all9c[-sample,]
+
+# first tune shrinkage and number of trees
+train2<- train
+train2$b.rpdol<- as.factor(train2$b.rpdol)
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="0"] <- "profit"
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="1"] <- "loss"
+
+formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log +
+client.totinv.log + pc.majpos.log + Billing.Type + majority.pos + Business"
+
+cv.Control <- trainControl(method = "cv",
+                           number = 5,
+                           seeds= seeds
+                           , summaryFunction=twoClassSummary,
+                           classProbs=TRUE
+                           )
+gbmGrid <- expand.grid(shrinkage = c(0.001),
+                       n.trees = c(10000), interaction.depth = c(5),
+                       n.minobsinnode = c(20))
+set.seed(2)
+gbmFit <- train(as.formula(formula), data= train2,
+                 method = "gbm", trControl = cv.Control, verbose = FALSE,
+                 bag.fraction = 0.5, tuneGrid = gbmGrid,
+                metric = 'ROC',
+                na.action = na.pass)
+
+assign(paste("gbmFit",j,sep = ""), gbmFit, envir = .GlobalEnv)
+}
+
+AUCs = c(gbmFit4$results$ROC, gbmFit1$results$ROC, gbmFit2$results$ROC, gbmFit3$results$ROC)
+plot = data.frame(no.clusts = clusts, AUC = AUCs)
+ggplot(plot, aes(x = no.clusts %>% as.numeric, y = AUC)) + geom_line()
+
+# therefore:
+# shrinkage = 0.001
+# no trees = 10000
+# min terminal node size = 20
+# max tree depth = 5
+
+# AUC approx 0.748 for 5 clusters of invoiced amount for job = best
+#
+
+```
+
+Tune boosted tree for 5 clusters of job invoice size
+
+```{r, echo=FALSE}
+
+#decide how many clusters
+x<- dist(all9c$inv.mlsto.log, method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+rect.hclust(fit, k = 5)
+
+#assign a cluster number to all rows in data frame
+all9c$b.inv.log<- as.factor(cutree(fit, k=5))
+head(all9c$b.inv.log)
+
+key = all9c %>% group_by(b.inv.log) %>% summarise(min = min(inv.mlsto.log) %>% exp() %>% round(0), max = max(inv.mlsto.log) %>% exp() %>% round(0))
+
+arrange(key, min)
+key$min %>% round(-2)
+breaks = c(100,600,2500,8000,60000, 1800000)
+breaks = sort(breaks)
+log(breaks)
+
+#create clusters based on summary:
+man.clust<- function(first.vec=all9c$timespan.cbrt, breaks.man=c(0,4.16,6.558,15),
+                     labs = c("t0_70", "t70_280", "t280_")){
+        final = cut(first.vec, breaks=breaks.man, include.lowest=TRUE, labels= labs)
+        return(final)
+}
+
+#inv.mlsto.log
+
+all9c$b.inv.log <- man.clust(first.vec=all9c$inv.mlsto.log, 
+                            breaks.man= log(breaks),
+                            labs = c("100-600","600-2.5k","2.5k-8k","8k-60k","60k-1.8m"
+                            ))
+
+# set seeds
+set.seed(123)
+seeds <- vector(mode = "list", length = 51)
+for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
+## For the last model:
+seeds[[51]] <- sample.int(1000, 1)
+
+#create test train
+sample<- sample(1:nrow(all9c), 2/3*nrow(all9c), replace=F)
+train<- all9c[sample,]
+test<- all9c[-sample,]
+
+# first tune shrinkage and number of trees
+train2<- train
+train2$b.rpdol<- as.factor(train2$b.rpdol)
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="0"] <- "profit"
+levels(train2$b.rpdol)[levels(train2$b.rpdol)=="1"] <- "loss"
+
+formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log +
+client.totinv.log + pc.majpos.log + Billing.Type + majority.pos + Business"
+
+cv.Control <- trainControl(method = "cv",
+                           number = 5,
+                           seeds= seeds
+                           , summaryFunction=twoClassSummary,
+                           classProbs=TRUE
+                           )
+gbmGrid <- expand.grid(shrinkage = c(0.001),
+                       n.trees = c(10000), interaction.depth = c(3,4,5,6),
+                       n.minobsinnode = c(20))
+set.seed(2)
+gbmFit <- train(as.formula(formula), data= train2,
+                 method = "gbm", trControl = cv.Control, verbose = FALSE,
+                 bag.fraction = 0.5, tuneGrid = gbmGrid,
+                metric = 'ROC',
+                na.action = na.pass)
+
+plot(gbmFit)
+
+# interaction depth = 5 is best
+
+write.csv(all9c, 'C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')
+write.csv(all9c, 'C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all9c.csv')
+
+```

### 2016-05-25 finalwriteup/draft.rmd
finished 10 fold runs

@@
 opts_chunk$set(fig.width=8, fig.depth = 5, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
+getwd()
 ```
@@
 n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
+        if(type.calc == 'one.sample'){
                 d.calc = abs(mean(data))/sd(data)
@@
                 return(ceiling(pow_pow$n))
+        }
+
+        if(type.calc == 'two.sample'){
+                d.calc = abs(mean(data[[1]]) - mean(data[[2]]))/sqrt(mean(c(var(data[[1]]), var(data[[2]]))))
+                pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05, power = .8, type = c(type.calc), alternative = 'two.sided')
+                return(ceiling(pow_pow$n))
+        }
+
 }
@@
 
-Structure:
-
-* boosted trees - auc with a few iterations - cross val
+#### Initial Trial of five Methods
+
+To begin analysis, all prediction methods were trialled using as much data as possible for each method.
+Boosted trees and Naiive bayes algorithm are able to process data with missing values, so all data could be used.
+Logistic regression, randomForest and Bayesian Networks require complete datasets so a subset of complete data was used for these.
+It was expected that Boosted trees would perform best as they are powerful and can use incomplete data.
+For this reason, 4 x 5 fold runs of each method were performed (20 models total for each method) and statistically compared against boosted trees.
+A power calculation was used to calculate the number of runs required for a statistical power of 0.8.
+
+The models were compared using the Area under the curve (AUC) test statistic calculated from the Reciever Operating Curve (ROC) as this gives a good indication of model performance under binary classification [@pROC][@Putler2012].
+An example plot of five ROC's resulting from five fold cross validation models using the boosted tree algorithm is shown below:
+
+```{r}
+# image below generated from file rf_log_naiive_auc.rmd
+# lines 683 to 689
+
+```
+
+![](images/boosted tree/5_fold_ROC.png)
+
+[@pROC]
+
+The boosted tre model was first tuned using the caret package to optimise the following parameters [@caret]:
+
+* shrinkage = 0.001
+* no trees = 10000
+* min terminal node size = 20
+* max tree depth = 5
+
+In this example, the average area under the curve (AUC) was 0.763 indicating that on the whole, taking into account all possible thresholds, the model is predicting profit/loss better than random chance but not perfectly.
+20 models and their resulting AUC statistics from each method were initially determined.
+These were used to calculate how many models were required to achieve a minimum statistical power of 0.8 when compared to the boosted tree results.
+The comparison of logistic regression required 100 models, while randomForest required over 700 because the AUC distributions were close.
+It was decided that randomForest and boosted trees would likely not be distinguishable, and 100 models of each method were created via performing 10 fold cross validation 10 times.
+The violin plot below summarises the AUC values from each method.
+
+```{r reduced_violin}
+
+reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
+
+ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
+        geom_jitter(alpha = 0.2, shape = 16) +
+        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
+        scale_fill_manual(values = topo.colors(4)) +
+        scale_colour_manual(values = topo.colors(4)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of each Method")
+
+```
 
@@
 
-* other results from methods?? 
 * Bayesian network
@@
 
-### Bayesian Network
+#### Bayesian Network
 

### 2016-05-25 rf_log_naiive_auc.rmd
regression report
finished 10 fold runs

@@
 all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
-all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
+all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]
 
@@
 all10a<- all10a[complete.cases(all10a),]
-# 2280 cases
+# 2280 cases - 10 vars
 
@@
 all10b<- all10b[complete.cases(all10b),]
-# 1987 cases
+# 1987 cases - 12 vars
 
@@
 all10c<- all10c[complete.cases(all10c),]
-# 1106 cases
+# 1106 cases - 13 vars
 
@@
 
-test.roc <- function(dfs = df.list, seed = 100, mtrys = c(4,5,3,4,4), method = "log",
+test.roc <- function(dfs = df.list, seed = 100, mtrys = c(4,5,3,4,4), method = "log", x.folds = 5,
                      extra.var = NULL, mult = 1){
@@
         auc<- rep(list(rep(0, length(dfs))), length(dfs))
-        colours <- c("red", "yellow", "green", "blue", "purple")
+        colours <- topo.colors(x.folds*mult)
 
@@
                 #set up 10 fold cv using caret function
-                folds <- createFolds(df$f.rpdol, k = 10, list = T, returnTrain = F)
+                folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
                 #now have folds[[1]] through folds[[10]] list of rows to exclude
@@
                                 set.seed(sample(1:1000,1))
-                                folds.temp = createFolds(df$f.rpdol, k = 10, list = T, returnTrain = F)
+                                folds.temp = createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
                                 folds = append(folds, folds.temp)
@@
                         if(is.null(extra.var)){
-                                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
+                                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
 
                         } else {
-                                formula<- paste("f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second", "+", extra.var, sep = " ")
+                                formula<- paste("f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second", "+", extra.var, sep = " ")
                         }
@@
                 if(nrow(df) <= 2300){
-                        formula<- "f.rpdol ~ . -Year -b.rpdol"
+                        # include code.client, code.contact, Billing.Type
+                        formula<- "f.rpdol ~ . -Year -b.rpdol -inv.mlsto.log - return.pdol - timespan.cbrt - mlsto"
                 }
@@
                 if(method %in% 'boost') {
-                        formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
+                        #no JD.Second
+                        formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
 
@@
                                                    distribution = "bernoulli", n.trees = 10000,
-                                                   shrinkage = 0.001, interaction.depth = 4,
+                                                   shrinkage = 0.001, interaction.depth = 5,
                                                    n.minobsinnode = 20)
@@
                         #pROC
-                        pred.p <- roc(test$f.rpdol, pred)
+                        pred.p <- pROC::roc(test$f.rpdol, pred)
 
@@
                         plot.roc(
-                        pred.p, col = colours[i], add = TRUE , print.thres = F)
+                        pred.p, col = colours[j], add = TRUE , print.thres = F)
                         }
@@
 #boosted tree
-formula<- formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
+formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
 
@@
 ```
+
+Create plot for final report
+Boosted tree
+
+```{r}
+df.list<- c('all9c')
+test.roc(dfs = df.list, seed = 100, mtrys = c(5), method = "boost",
+         extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
+        stat_summary(fun.y="mean", geom="point", colour = "red")
+
+```

### 2016-05-25 imputingall9c.R

@@ -92,17 +92,18 @@ auc
 
 #experiment with MICE package - should not include return per dollar during imputation!
 
-new <- mice(all9c[,c(1:4,6:10,12:14,16)], meth = 'rf')
+new <- mice(all9c[,c(1:4,6,8:10,12:14,16,18,19)], meth = 'rf')
 
 all10mice <- complete(new)
 all10mice<- cbind(all10mice, all9c[,names(all9c) %in% c('code.client', 'code.contact', 'mlsto',
-                                                        'b.timespan.cbrt', 'b.rpdol')])
+                                                        'b.rpdol')])
 
 ## -------------------------------------------------------------------------------------------------
 
 # now have two imputed datasets all10rf and all10mice
 
 write.csv(all10mice, 'C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')
+write.csv(all10mice, '~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
 summary(all10rf)
 summary(all10mice)
 

### 2016-05-26 finalwriteup/draft.rmd
trying to fix

--- 2016-05-23 finalwriteup/draft.rmd
@@
 
-Structure:
+#### Initial Trial of five Methods
+
+To begin analysis, all prediction methods were trialled using as much data as possible for each method.
+Boosted trees and Naiive bayes algorithm are able to process data with missing values, so all data could be used.
+Logistic regression, randomForest and Bayesian Networks require complete datasets so a subset of complete data was used for these.
+It was expected that Boosted trees would perform best as they are powerful and can use incomplete data.
+For this reason, 4 x 5 fold runs of each method were performed (20 models total for each method) and statistically compared against boosted trees.
+A power calculation was used to calculate the number of runs required for a statistical power of 0.8.
+
+The models were compared using the Area under the curve (AUC) test statistic calculated from the Reciever Operating Curve (ROC) as this gives a good indication of model performance under binary classification [@pROC][@Putler2012].
+An example plot of five ROC's resulting from five fold cross validation models using the boosted tree algorithm is shown below:
+
+```{r}
+# image below generated from file rf_log_naiive_auc.rmd
+# lines 683 to 689
+
+```
+
+![](images/boosted tree/5_fold_ROC.png)
+
+[@pROC]
+
+* Boosted trees first tuned using caret package [@caret]:
+* shrinkage = 0.001
+* no trees = 10000
+* min terminal node size = 20
+* max tree depth = 5
 

### 2016-05-26 model_averaging.rmd
backwards at work

---
title: "Model Averaging"
author: "Amy Cook"
date: "May 26, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('caret')
library('pROC')
library('e1071')
library('randomForest')
library('party')

setwd("C:/Users/n9232371/Documents/Consultbusiness/data")

all10mice<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')[,-1]

```

Model averaging how to do it>?
Best models are ..

* boosted tree
* logisitic regression
* random forest

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2016-05-26 rf_log_naiive_auc.rmd
backwards at work

@@
 ```{r, echo=FALSE}
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
-library('e1071',lib='C:/Progra~1/R/R-3.2.2/library')
-library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
-library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library('ggplot2')
+library('gbm')
+library('caret')
+library('pROC')
+library('e1071')
+library('randomForest')
+library('party')
 
@@
 ```
+
+Checking p value for BN
+
+```{r}
+reduc.totAUC <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/reduc_df_auc.rds")
+BN = read.csv("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/BN_AUCs.csv")
+
+n.samples(data = list(reduc.totAUC[[2]][1:20], BN$AUC), type.calc = 'two.sample', alt = 'two.sided')
+
+# only 5 samples required according to power calculation
+
+```

### 2016-05-27 finalwriteup/draft.rmd
backwards at work
re-syncing mac draft

--- 2016-05-25 finalwriteup/draft.rmd
@@
 The violin plot below summarises the AUC values from each method.
+The 'violins' are coloured according whether the distributions significantly vary to boosted trees using a critical value of 0.05.
 
@@
 
-* AUC for random forest, everything else without imputing?? thats what i did
-    * boxplot showing AUC for all methods. how many runs, 20?
-* only take best performing models and try with imputed data
-
-* imputing - boosted trees still fine
-    * apply to other methods
-
-* boxplot chart showing AUC's for all methods on one plot : )
-
-* table showing significance p value numbers - in comparison to what? the best performing one?
-
-* Bayesian network
-* Random forest
-* Gradient boosted trees
-* Regression - baseline model
-* Naiive bayes - baseline model
+The performance of logistic regression and random forest algorithms cannot be statistically differntiated from boosted trees.
+The same process was repeated using an imputed dataset.
+Imputation was done using the MICE package and allowed all algorithms to be trialled with a full dataset.
+Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for random forest unless over 450000 models were made.
+
+```{r mice_violin}
+
+reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
+
+ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
+        geom_jitter(alpha = 0.2, shape = 16) +
+        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
+        scale_fill_manual(values = topo.colors(4)) +
+        scale_colour_manual(values = topo.colors(4)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of each Method")
+
+```
+
+```{r all_violin}
+
+reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
+
+ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
+        geom_jitter(alpha = 0.2, shape = 16) +
+        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
+        scale_fill_manual(values = topo.colors(4)) +
+        scale_colour_manual(values = topo.colors(4)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of each Method")
+
+```
+
+Imputed data tightens the distribution of randomForest substantially but slightly spreads the distriution of Boosted trees and logistic regression.
 

### 2016-05-27 rf_log_naiive_auc.rmd
re-syncing mac draft

@@
 setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
 
@@
 all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
+all9c<- read.csv('all9c.csv')[,-1]
 all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]
@@
 all10a<- all9c %>% select(no.users, Discipline, pc.pro, Business,
-                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
+                         b.timespan.cbrt, b.inv.log, client.totinv.log, 
                          mlsto, Year, b.rpdol)
@@
 all10b<- all9c %>% select(no.users, Discipline, pc.pro, Business,
-                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
+                         b.timespan.cbrt, b.inv.log, client.totinv.log, 
                          mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
@@
 all10c<- all9c %>% select(no.users, Discipline, pc.pro, Business, JD.Second,
-                         b.timespan.cbrt, inv.mlsto.log, client.totinv.log, 
+                         b.timespan.cbrt, b.inv.log, client.totinv.log, 
                          mlsto, Year, b.rpdol, majority.pos, pc.majpos.log)
@@
         for(i in 1:length(dfs)){
-                df = get(df.list[i])
+                df = get(dfs[i])
 
@@
                         # include code.client, code.contact, Billing.Type
-                        formula<- "f.rpdol ~ . -Year -b.rpdol -inv.mlsto.log - return.pdol - timespan.cbrt - mlsto"
+                        formula<- "f.rpdol ~ . -Year -b.rpdol "
                 }
@@
                         formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
+
+                }
+
+                if(method %in% 'naiive') {
+                        formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type + JD.Second"
 
@@
 
-                        } else {
+                        }
+
                         if(method %in% 'log'){
@@
                                                 type = "response")
-                                } else {
+                        } 
+                        if(method %in% 'boost'){
                                 fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
@@
 
-                                                pred <-predict(fit, test, 
-                                                               n.tree = 10000, type = "response")
+                                pred <- predict(fit, test, n.tree = 10000, type = "response")
                         }
+
+                        if(method %in% 'naiive'){
+                                fit<- naiveBayes( as.formula(formula), df[-folds[[j]],])
+                                pred <- predict(fit, test, type = 'raw')
+                                pred<- pred[,2]
                         }
@@
                         auc[[i]][j]<- auc(pred.p)
+                        names(auc)[[i]] = method
 
@@
 df.list <- c("all10rf", "all10mice", "all10a", "all10b", "all10c")
-test.roc(dfs = df.list, seed = 400, mtrys = c(4,5,3,4,4), method = "log")
+test.roc(dfs = df.list, seed = 400, mtrys = c(4), method = "log",
+         x.folds = 5, extra.var = NULL, mult = 1)
+names(auc.func) = df.list
 ggplot(melt(auc.func), aes(x= factor(L1), y= value)) + geom_boxplot() + geom_point() +
@@
 
-```{r, echo=FALSE}
+```{r naiive bayes, echo=FALSE}
 
@@
 df = all10mice
+df = all9c
 
@@
 
-Checking p value for BN
+Create loop through each method - 20 models - compile AUC results
 
 ```{r}
-reduc.totAUC <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/reduc_df_auc.rds")
-BN = read.csv("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/BN_AUCs.csv")
+test.roc(dfs = c('all10a'), seed = 100, mtrys = c(5), method = "log",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc = auc.func
 
-n.samples(data = list(reduc.totAUC[[2]][1:20], BN$AUC), type.calc = 'two.sample', alt = 'two.sided')
+test.roc(dfs = c('all9c'), seed = 100, mtrys = c(5), method = "boost",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc = append(tot.auc,auc.func)
 
-# only 5 samples required according to power calculation
+test.roc(dfs = c('all10c'), seed = 100, mtrys = c(4), method = "rf",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc = append(tot.auc,auc.func)
+
+test.roc(dfs = c('all9c'), seed = 100, mtrys = c(4), method = "naiive",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc = append(tot.auc,auc.func)
+
+saveRDS(tot.auc,"~/Documents/github/consult/finalwriteup/report_data/reduc_df_auc.rds")
+
+ggplot(melt(tot.auc), aes(x= factor(L1), y= value)) + geom_violin() + geom_jitter() +
+        stat_summary(fun.y="mean", geom="point", colour = "red")
+
+# n.samples function is defined in 'draft.rmd'
+index = c(1,3,4)
+ns = NULL
+for(i in seq_along(index)){
+        temp = n.samples(data = tot.auc[c(2, index[i])], type.calc = 'two.sample', alt = 'two.sided')
+        ns[i] = temp
+}
+ns
 
 ```
+
+Re-run with MICE imputed data
+
+```{r}
+
+all10mice_f = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
+
+test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc.m = auc.func
+
+test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "boost",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc.m = append(tot.auc.m,auc.func)
+
+test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(4), method = "rf",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc.m = append(tot.auc.m,auc.func)
+
+test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(4), method = "naiive",
+         extra.var = NULL, mult = 10, x.folds = 10)
+tot.auc.m = append(tot.auc.m,auc.func)
+
+saveRDS(tot.auc.m,"~/Documents/github/consult/finalwriteup/report_data/mice_df_auc.rds")
+
+ggplot(melt(tot.auc.m), aes(x= factor(L1), y= value)) + geom_violin() + geom_jitter() +
+        stat_summary(fun.y="mean", geom="point", colour = "red")
+
+# n.samples function is defined in 'draft.rmd'
+index = c(1,3,4)
+ns = NULL
+for(i in seq_along(index)){
+        temp = n.samples(data = tot.auc.m[c(2, index[i])], type.calc = 'two.sample', alt = 'two.sided')
+        ns[i] = temp
+}
+ns
+
+```

### 2016-05-27 finalwriteup/draft.rmd

@@
 opts_chunk$set(fig.width=8, fig.depth = 5, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
-getwd()
 ```
@@
 n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
-        if(type.calc == 'one.sample'){
         d.calc = abs(mean(data))/sd(data)
@@
         return(ceiling(pow_pow$n))
-        }
-
-        if(type.calc == 'two.sample'){
-                d.calc = abs(mean(data[[1]]) - mean(data[[2]]))/sqrt(mean(c(var(data[[1]]), var(data[[2]]))))
-                pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05, power = .8, type = c(type.calc), alternative = 'two.sided')
-                return(ceiling(pow_pow$n))
-        }
-
 }
@@
 
-The boosted tre model was first tuned using the caret package to optimise the following parameters [@caret]:
-
+* Boosted trees first tuned using caret package [@caret]:
 * shrinkage = 0.001

### 2016-05-27 finalwriteup/draft.rmd

@@
 The violin plot below summarises the AUC values from each method.
-The 'violins' are coloured according whether the distributions significantly vary to boosted trees using a critical value of 0.05.
 
@@
 
-reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
+reduc.totAUC <- readRDS("reduc_df_auc.rds")
 
@@
 
-The performance of logistic regression and random forest algorithms cannot be statistically differntiated from boosted trees.
-The same process was repeated using an imputed dataset.
-Imputation was done using the MICE package and allowed all algorithms to be trialled with a full dataset.
-Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for random forest unless over 450000 models were made.
-
-```{r mice_violin}
-
-reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
-
-ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
-        geom_jitter(alpha = 0.2, shape = 16) +
-        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
-        scale_fill_manual(values = topo.colors(4)) +
-        scale_colour_manual(values = topo.colors(4)) +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method")
-
-```
-
-```{r all_violin}
-
-reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
-
-ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
-        geom_jitter(alpha = 0.2, shape = 16) +
-        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
-        scale_fill_manual(values = topo.colors(4)) +
-        scale_colour_manual(values = topo.colors(4)) +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method")
-
-```
-
-Imputed data tightens the distribution of randomForest substantially but slightly spreads the distriution of Boosted trees and logistic regression.
+* boosted trees - auc with a few iterations - cross val
+
+* AUC for random forest, everything else without imputing?? thats what i did
+    * boxplot showing AUC for all methods. how many runs, 20?
+* only take best performing models and try with imputed data
+
+* imputing - boosted trees still fine
+    * apply to other methods
+
+* boxplot chart showing AUC's for all methods on one plot : )
+
+* table showing significance p value numbers - in comparison to what? the best performing one?
+
+* Bayesian network
+* Random forest
+* Gradient boosted trees
+* Regression - baseline model
+* Naiive bayes - baseline model
 

### 2016-05-30 finalwriteup/draft.rmd
fixed lost work

--- 2016-05-27 finalwriteup/draft.rmd
@@
 n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
+        if(type.calc == "one.sample"){
               d.calc = abs(mean(data))/sd(data)
-        pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05, power = .8, type = c(type.calc), alternative = alt)
+              pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
+                                   power =.8, type = c(type.calc), 
+                                   alternative = alt)
               return(ceiling(pow_pow$n))  
+        }
+
+        if(type.calc == 'two.sample'){
+                d.calc = abs(mean(data[[1]])-mean(data[[2]]))/
+                        (sqrt((sd(data[[1]])^2+sd(data[[2]])^2)/2))
+                pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
+                                   power =.8, type = c(type.calc), 
+                                   alternative = 'two.sided')
+                return(ceiling(pow_pow$n)) 
+        }
+
 }
@@
 
-In this example, the average area under the curve (AUC) was 0.763 indicating that on the whole, taking into account all possible thresholds, the model is predicting profit/loss better than random chance but not perfectly.
+In this example, the average area under the curve (AUC) was 0.763 indicating that on the whole, taking into account all possible thresholds, the model is predicting profit/loss better than random chance (AUC = 0.5) but not perfectly (AUC = 1).
 20 models and their resulting AUC statistics from each method were initially determined.
@@
 The comparison of logistic regression required 100 models, while randomForest required over 700 because the AUC distributions were close.
-It was decided that randomForest and boosted trees would likely not be distinguishable, and 100 models of each method were created via performing 10 fold cross validation 10 times.
+Bayesian Networks and Naiive Bayes did not perform as well and the power calculation revealed only 5 and 8  models were required to achieve a power of 0.8 respectively.
+Bayesian Networks were then dropped as Naiive Bayes perfomed slightly better and was fast and simple enough to work well as a baseline comparison to the higher performing methods.
+
+It was finally decided that randomForest and boosted trees would likely not be distinguishable since the power calculations indicate 700 models are required.
+hence 100 models of each method were created as required for logistic regression.
+10 fold cross validation was performed 10 times.
 The violin plot below summarises the AUC values from each method.
-The 'violins' are coloured according whether the distributions significantly vary to boosted trees using a critical value of 0.05.
+The 'violins' are coloured according to whether the distributions significantly vary to boosted trees using a critical value of 0.05.
 
@@
 
-reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
-
-ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
-        geom_jitter(alpha = 0.2, shape = 16) +
-        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
-        scale_fill_manual(values = topo.colors(4)) +
-        scale_colour_manual(values = topo.colors(4)) +
+reduc.totAUC <- readRDS("reduc_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(reduc.totAUC[[2]], reduc.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+reduc.totAUC = melt(reduc.totAUC)
+reduc.totAUC$sigdiff = ifelse(reduc.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different \nfrom boosted trees") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different \nfrom boosted trees") +
         labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method")
+             title = "AUC Distribution from 100 models of each Method") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naiive Bayes", "rf" = "Random Forest"))
 
@@
 Imputation was done using the MICE package and allowed all algorithms to be trialled with a full dataset.
-Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for random forest unless over 450000 models were made.
+Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for random forest unless over 450,000 models were made.
 
@@
 
-reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
-
-ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
-        geom_jitter(alpha = 0.2, shape = 16) +
-        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
-        scale_fill_manual(values = topo.colors(4)) +
-        scale_colour_manual(values = topo.colors(4)) +
+mice.totAUC <- readRDS("mice_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+mice.totAUC = melt(mice.totAUC)
+mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different \nfrom boosted trees") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different \nfrom boosted trees") +
         labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method")
+             title = "AUC Distribution from 100 models of each Method - Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naiive Bayes", "rf" = "Random Forest"))
 
@@
 
-reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
-
-ggplot(melt(reduc.totAUC), aes(x= factor(L1), y= value)) + geom_violin(aes(fill = L1, colour = L1)) + 
-        geom_jitter(alpha = 0.2, shape = 16) +
-        stat_summary(fun.y="mean", geom="point", colour = "darkblue", shape =3) +
-        scale_fill_manual(values = topo.colors(4)) +
-        scale_colour_manual(values = topo.colors(4)) +
+reduc.totAUC$Data_Source = "Original"
+mice.totAUC$Data_Source = "Imputed"
+combine.micer <- rbind(reduc.totAUC, mice.totAUC)
+
+combine.micer$Data_Source <- factor(combine.micer$Data_Source, 
+                                    levels=c("Original", "Imputed"))
+
+ggplot(combine.micer, aes(x= factor(L1), y= value)) +
+        geom_violin(aes(fill = Data_Source, colour = Data_Source)) +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Data Source") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Data Source") +
         labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method")
+             title = "Comparison of Original Data vs. Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naiive Bayes", "rf" = "Random Forest"))
 
@@
 
-Imputed data tightens the distribution of randomForest substantially but slightly spreads the distriution of Boosted trees and logistic regression.
-
-#### Bayesian Network
-
-* could not handle _any_ numeric variables, had to discretise all numeric vars
-    * used normalised numeric variables
-    * I did this by making a hierarchical dendrogram to visualise the clusters
-    * then I chose the number of clusters, drew the red rectangle
-    * summarised the clusters to find max and min values for the variable within each cluster. Rounded this value up
-    * manually discretised variable using rounded values
-
-* first trialled complete cases of number of users, Discipline, percent pro, Business, client code, job detail, majority position, timespan, invoiced amount, percent hours by majority position, and return per dollar.
-This gave 907 complete cases to work with.
-
-*using Genie Smile.
-   * turned return per dollar into profit/loss. This gave 182 loss cases and 725 profit cases. A good distribution for working with bayesian networks. Ie not too biased
-   * a good, simple problem to start with
-
-* establish a baseline. Using ROC, area under the curve method
-    * to test our classifiers we need to beat random chance. Ie AUC = 0.5
-    * will also establish the AUC for using one variable only! See table below:
-       *Note all using 10 fold cross validation
-
-Single Variable  | AUC from ROC chart | Num losses correct
----------------- | ------------------ | ------------------
-Random chance    | 0.5                | ?                 
-percent pro      | 0.618              | 0
-percent maj pos  | 0.616              | 0
-number of users  | 0.661              | 0
-timespan         | 0.600              | 0
-amount invoiced  | 0.535              | 0
-Discipline       | 0.514              | 0
-Majority pos     | 0.605              | 3
-Job detail       | 0.605              | 1
-client code      | 0.567              | 25
-client Business  | 0.524              | 5
-
-* try using the following variables:
-
-```{r, echo=FALSE, include=FALSE}
-library('magrittr')
-
-if(.Platform$OS.type == 'windows'){
-        all8c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8c.csv' )
-} else{
-        all8c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all8c.csv')
-}
-
-vars<- names(all8c)
-vars<- vars[!vars %in% c("Business", "JD_Second", "code_client")]
-
-```
-
-* number of users
-* Discipline
-* percent hours by a professional
-* position that completed the majority of hours
-* timespan
-* amount invoiced
-* percent of hours completed by majority position
-
-The results are as follows:
-
-AUC = 0.705
-Number of losses predicted correctly: 124
-
-This is our 'base' model. Now try adding one complex variable at a time:
-
-Added Variable   | AUC from ROC chart | Num losses correct (tp) | Num losses wrong (fp)
----------------- | ------------------ | ----------------------- | --------------------
-none/base        | 0.710              | 122                     | 272
-client code      | 0.624              | 171                     | 528
-Busines          | 0.667              | 144                     | 381
-job detail       | 0.653              | 150                     | 438
-
-## Boosted Trees
-
-### Advantages
-
-    * Wonderful in that NA's don't interfere with prediciton process
-        * explain..
-        * can comfortably include the entire dataset without having to worry about imputing
-
-tuned which variables to include:
-    * exclude JD.Second, Business, code.client, code.contact
-    * tuned parameters:
-        * shrinkage = 0.001
-        * n.trees = 4000
-        * interaction.depth = 3
-        * min.nobs = 10
-    * manage to get AUC up to 0.785!!
-
-can we improve this by bringing huge variables back in a smarter way?
+Imputed data tightens the distribution of randomForest substantially but slightly spreads the distribution of Boosted trees and logistic regression.
+Since the methods using imputed data have a similar performance, slightly tighter distributions of results and use more data, models using imputed data was used when developing blended models of all methods.
 
@@
 acknowledge that it may be worthwhile for other reasons such as confidence and transparency in historic projects.
-* important variables: cforest diagram vs anova vs random forest
-    * which variables were chosen for machine learning analysis
-* first tried random forest predicting return per dollar
-    * fail, why - for discussion
-* moved on to predicting yes or no profitability.
-    * tried: random forest, logistic regression, gradient boosted trees, naiive bayes (baseline), bayesian networks
-    * boosted trees could handle the missing data, whereas for all? other methods I had to narrow down data to complete data
-    * used MICE package - explain what this is in method.
-    * comparison of predictive accuracy of all methods. non are significantly better than the other.
 
@@
 
-* best models are gradient boosted trees and logistic regression. why? can hypothesis but not really
+comments on your results;
+explains what your results mean;
+interprets your results in a wider context; indicates which results were expected or unexpected;
+provides explanations for unexpected results.
+The Discussion should also relate your specific results to previous research or theory.
+You should point out what the limitations were of your study, and note any questions that remain unanswered.
+The Discussion CAN also include Conclusions/Future Research.
+
+* best models are gradient boosted trees and logistic regression and random forest. why? can hypothesis but not really
+* envouraging results - mean sits around 0.775. real question is bottom line analysis
+* was this expected? didnt expect logistic regression to perform that well
+* imputed data worked almost the same, slightly better for randomforests?
+    * wasn't sure what to expect so .. i dunno
+* previous research on logistic regression working as well?
+* limitations: didn't try all possible variable combinations
+* try and improve by blending the three top performing models
+    * this will eliminate some of the benefits ofhaving a single model such as 
+        * partial dependency plots from boosted tree
+            * these will be investigated later
+
 * this shows the profitability can be predicted with some success, but real question is, is it financially viable?

### 2016-05-30 finalwriteup/draft.rmd
started binary result discussion

@@
 
+Testing gradient boosted trees, random forests, naiive bayes, logistic regression and bayesian networks in binary classification showed that boosted trees, logistic regression and random forests performed best (according to AUC).
+Bayesian Networks are not particularly popular and require discretised variables so it was expected that this algorithm may not perform as well as highly regarded methods such as boosted trees and random forests.
+Naiive bayes and logistic regression were included as baseline models and it was expected that the more complex models would outperform these.
+Therefore it was surprising that results from 100 logistic regression models was not statistically lower than boosted trees.
+Literature??
+
+logistic regression:
+- one decision boundary - does not have to be parallel to variable axes
+- less prone to overfitting
+    - handles many variables better
+- weaker predictive power than ensemble trees (An Empirical Comparison of Supervised Learning Algorithms)
+
+ensemble trees:
+- can be hundreds of decision boundaries but they have to be parallel to variable axes
+- may not be enough data to learn complex decision rules better than a simple model
+
+Boosted trees, random forests and logistic regression all have AUC means around 0.775.
+The tests were repeated using imputed data and again the means sat around 0.775.
+The variables were not particularly well suited for imputing so this result is logical and could have even reduced the AUC.
+Many of the imputed variables were categorical and difficult to impute (such as job category, percent of job completed by a professional level employee, and the position of the majority worker on a job).
+The benefit of having an imputed dataset that performs well enough is that each model produces predictions for the full set of data.
+Beforehand, each model was limited to a complete set of data which implies a reduced data size.
+With complete model predictions, the high performing models can be blended to test whether better predictions are possible.
+Blending models brings about limitations which is discussed in the next section.
+
+This individual models perform well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a potential job will be profitable or loss-making.
+Whether the model is worth implementing in the work place is dependent on how the model affects decisions and the degree the algorithm will improve the 'bottom line' profits for the business.
+This will be reviewed in the following section.
+
+## Blended Model
+
 comments on your results;
@@
 The Discussion CAN also include Conclusions/Future Research.
-
-* best models are gradient boosted trees and logistic regression and random forest. why? can hypothesis but not really
-* envouraging results - mean sits around 0.775. real question is bottom line analysis
-* was this expected? didnt expect logistic regression to perform that well
-* imputed data worked almost the same, slightly better for randomforests?
-    * wasn't sure what to expect so .. i dunno
-* previous research on logistic regression working as well?
-* limitations: didn't try all possible variable combinations
-* try and improve by blending the three top performing models
-    * this will eliminate some of the benefits ofhaving a single model such as 
-        * partial dependency plots from boosted tree
-            * these will be investigated later
 

### 2016-05-31 finalwriteup/draft.rmd
started model blending

@@
 Testing gradient boosted trees, random forests, naiive bayes, logistic regression and bayesian networks in binary classification showed that boosted trees, logistic regression and random forests performed best (according to AUC).
-Bayesian Networks are not particularly popular and require discretised variables so it was expected that this algorithm may not perform as well as highly regarded methods such as boosted trees and random forests.
+Bayesian Networks are not particularly popular and require discretised variables so it was expected that this algorithm would not perform as well as highly regarded methods such as boosted trees and random forests.
 Naiive bayes and logistic regression were included as baseline models and it was expected that the more complex models would outperform these.
-Therefore it was surprising that results from 100 logistic regression models was not statistically lower than boosted trees.
-Literature??
-
-logistic regression:
-- one decision boundary - does not have to be parallel to variable axes
-- less prone to overfitting
-    - handles many variables better
-- weaker predictive power than ensemble trees (An Empirical Comparison of Supervised Learning Algorithms)
-
-ensemble trees:
-- can be hundreds of decision boundaries but they have to be parallel to variable axes
-- may not be enough data to learn complex decision rules better than a simple model
+Therefore it was surprising that results from 100 logistic regression models were not statistically lower than boosted trees.
+
+A possible explanation for this according to the literature is that there was not enough data for the ensemble trees to learn the complex decision rules at which they excel.
+Trees tend to overfit the patterns in a smaller training set.
+Logistic regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
+This may explain why logistic regression is performing as well as enseble trees with the case study's smaller dataset with many variables.
 

### 2016-05-31 model_blend.rmd
started model blending

---
title: "Model Averaging"
author: "Amy Cook"
date: "May 26, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('caret')
library('pROC')
library('e1071')
library('randomForest')
library('party')

setwd("C:/Users/n9232371/Documents/Consultbusiness/data")

if(.Platform$OS.type == 'windows'){
        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
} else{
        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
}

```

Model averaging how to do it>?
Best models are ..

* boosted tree
* logisitic regression
* random forest

# For each method, create a complete dataset of predictions created from ten fold cross validation
```{r}

#create test and train set: 0.75/0.25
set.seed = 100
train = sample_frac(all10mice, 0.75, replace = FALSE)
test = setdiff(all10mice, train)

#perform 10-fold cross validation on train set to create complete train set of predictions for each method
full.predict <- function(df = train, method = "log", x.folds = 5, manual.seed = 100){

        # create f.rpdol
        df$f.rpdol <- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"

        #set up 10 fold cv using caret function
        set.seed(manual.seed)
        folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
        #now have folds[[1]] through folds[[10]] list of rows to exclude

        #create NULL dataframe with row names from folds
        method.pred <- data.frame('rownames' = unlist(folds))
        temp.pred = NULL

        #define formula
        if(method %in% c('log','rf')){
                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
        }

        if(method == 'boost') {
                #no JD.Second
                formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
        }

        #loop over each fold, and compile predictions
        for (j in 1:length(folds)) {

                #turn predictors and response into 'ordered' variables for roc() to work
                test<- df[folds[[j]],]
                levels(test$f.rpdol) <- c('profit', 'loss')
                test$f.rpdol <- ordered(test$f.rpdol)

                if(method %in% 'rf'){
                        fit <- randomForest(as.formula(formula), data = df[-folds[[j]],], 
                                    mtry = 4, ntree = 1000)
                        pred <- predict(fit, test, type = "prob")
                        pred <- pred[,2]

                }

                if(method %in% 'log'){
                        fit<- glm(as.formula(formula), 
                                  data = df[-folds[[j]],], family = binomial())
                        pred <- predict(fit, test,
                                        type = "response")
                } 
                if(method %in% 'boost'){
                        fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
                                   distribution = "bernoulli", n.trees = 10000,
                                   shrinkage = 0.001, interaction.depth = 5,
                                   n.minobsinnode = 20)

                        pred <- predict(fit, test, n.tree = 10000, type = "response")
                }

                temp.pred <- c(temp.pred, pred)

                #print j value
                cat(j, " ")

        }

        method.pred[, method] = temp.pred
        return(method.pred)

}

pred.log = full.predict(df = train, method = 'log', x.folds = 5, manual.seed = 100)
pred.rf = full.predict(df = train, method = 'rf', x.folds = 5, manual.seed = 100)
pred.boost = full.predict(df = train, method = 'boost', x.folds = 5, manual.seed = 100)

#merge all by rownames
pred.full = Reduce(full_join, list(pred.log, pred.rf, pred.boost))
gather.pred = gather(pred.full, key = method, value = pred.val,
                     -rownames)
order.rows = gather.pred %>% group_by(rownames) %>% 
        summarise(mean = mean(pred.val)) %>%
        arrange(mean) %>% ungroup
gather.pred$rownames <- factor(gather.pred$rownames,
                               levels= order.rows$rownames)

#plot spread of predictions

ggplot(gather.pred, aes(x = rownames, y = pred.val)) + geom_point(aes(colour = method))

ggplot(gather.pred, aes(x = method, y = pred.val)) + geom_boxplot(aes(fill = method), colour = "darkgray")

#randomForest seems to predict a bit higher than the other two methods. should I standardise it a bit??

```

Now create complete train set

```{r}
train.blend = full_join(train %>% mutate(rown = 1:nrow(train)), pred.full, 
                        by = c('rown'='rownames'))
train.blend <- train.blend %>% select(-rown)

```

# Try Feature weighted linear stacking

```{r cars}
#first write formula
# create f.rpdol
train.blend$f.rpdol <- as.factor(train.blend$b.rpdol)
levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "0"] <- "profit"
levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "1"] <- "loss"

formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + log + rf + boost"

#need to add interaction terms

```

# Try random forest

```{r pressure, echo=FALSE}
plot(pressure)
```

# Try boosted trees
```{r pressure, echo=FALSE}
plot(pressure)
```

### 2016-06-01 model_blend.rmd
started model blending

@@
         all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
+        source("C:/Users/n9232371/Documents/github/consult/functions.R")
 } else{
@@
 
-#create test and train set: 0.75/0.25
-set.seed = 100
-train = sample_frac(all10mice, 0.75, replace = FALSE)
-test = setdiff(all10mice, train)
+## new variables for FWLS and methods
+# create f.rpdol
+all10mice$f.rpdol <- as.factor(all10mice$b.rpdol)
+levels(all10mice$f.rpdol)[levels(all10mice$f.rpdol) == "0"] <- "profit"
+levels(all10mice$f.rpdol)[levels(all10mice$f.rpdol) == "1"] <- "loss"
 
 #perform 10-fold cross validation on train set to create complete train set of predictions for each method
-full.predict <- function(df = train, method = "log", x.folds = 5, manual.seed = 100){
-
-        # create f.rpdol
-        df$f.rpdol <- as.factor(df$b.rpdol)
-        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
-        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"
+full.predict <- function(df = all10mice, method = "log", x.folds = 5, manual.seed = 100){
 
@@
 
-pred.log = full.predict(df = train, method = 'log', x.folds = 5, manual.seed = 100)
-pred.rf = full.predict(df = train, method = 'rf', x.folds = 5, manual.seed = 100)
-pred.boost = full.predict(df = train, method = 'boost', x.folds = 5, manual.seed = 100)
+pred.log = full.predict(df = all10mice, method = 'log', x.folds = 5, manual.seed = 100)
+pred.rf = full.predict(df = all10mice, method = 'rf', x.folds = 5, manual.seed = 100)
+pred.boost = full.predict(df = all10mice, method = 'boost', x.folds = 5, manual.seed = 100)
 
@@
 
-ggplot(gather.pred, aes(x = rownames, y = pred.val)) + geom_point(aes(colour = method))
+ggplot(gather.pred, aes(x = rownames, y = pred.val, group = method)) + geom_point(aes(colour = method)) +
+        geom_line(aes(colour = method))
 
@@
 ```{r}
-train.blend = full_join(train %>% mutate(rown = 1:nrow(train)), pred.full, 
+
+all10.blend = full_join(all10mice %>% mutate(rown = 1:nrow(all10mice)), pred.full, 
                         by = c('rown'='rownames'))
-train.blend <- train.blend %>% select(-rown)
+all10.blend <- all10.blend %>% select(-rown)
+
+# need to normalise method responses
+ggplot(all10.blend, aes(x= sign(log) * abs(log)^(1/3))) + geom_histogram()
+ggplot(all10.blend, aes(x= rf^0.5)) + geom_histogram()
+ggplot(all10.blend, aes(x= log(boost))) + geom_histogram()
+
+#cube root of log
+all10.blend = mutate(all10.blend, norm.log = sign(log) * abs(log)^(1/3))
+all10.blend = mutate(all10.blend, norm.rf = rf^0.5)
+all10.blend = mutate(all10.blend, norm.boost = log(boost))
+
+ggplot(all10.blend, aes(x= norm.log)) + geom_histogram()
+
+#create test and train set: 0.75/0.25
+set.seed(100)
+train = sample_frac(all10.blend, 0.75, replace = FALSE)
+test = setdiff(all10.blend, train)
 
@@
 #first write formula
-# create f.rpdol
-train.blend$f.rpdol <- as.factor(train.blend$b.rpdol)
-levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "0"] <- "profit"
-levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "1"] <- "loss"
 
-formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + log + rf + boost"
+formula.log <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + norm.log + norm.rf + norm.boost"
 
-#need to add interaction terms
+formula.rf <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + norm.log + norm.rf + norm.boost"
+
+formula.boost <- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + Billing.Type + norm.log + norm.rf + norm.boost"
+
+#formula to add interaction terms for logistic regression
+metafeatures = strsplit(formula.log, split = c("\\s\\~\\s|\\s\\+\\s"))[[1]]
+methods = metafeatures[13:15]
+metafeatures = metafeatures[2:12]
+
+for(j in seq_along(methods)){
+
+        for(m in seq_along(metafeatures)){
+
+                formula.log <- paste(formula.log, " + ", metafeatures[m], "*", methods[j], sep = "")
+
+        }
+}
+
+#add terms to allow for constants [@Sill2009] - already done as everything is included as a single variable without interaction
+
+#run linear regression
+find.auc <- function(method = "log") {
+
+        if (method == "log") {
+                fit <- glm(as.formula(formula.log),
+                               data = train,
+                               family = binomial())
+
+                pred <- predict(fit, test, type = "response")
+
+                formula.check = "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
+
+                check.fit <- glm(as.formula(formula.check),
+                               data = train,
+                               family = binomial())
+
+                pred.c <- predict(check.fit, test, type = "response")
+        }
+
+        if (method == "rf") {
+        fit <- randomForest( as.formula(formula.rf), data = train,
+                mtry = 4, ntree = 1000, importance = TRUE
+                )
+        pred <- predict(fit, test, type = "prob")
+        pred <- pred[, 2]
+
+        formula.check = "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
+
+        check.fit <- randomForest( as.formula(formula.check), data = train,
+                mtry = 4, ntree = 1000
+                )
+        pred.c <- predict(check.fit, test, type = "prob")
+        pred.c <- pred.c[, 2]
+        }
+
+        if (method == "boost") {
+                fit <- gbm(as.formula(formula.boost), data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 3,
+                                   n.minobsinnode = 40)
+
+                pred <- predict(fit, test, n.tree = 6000, type = "response")
+
+                formula.check = "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + Billing.Type"
+
+                check.fit <- gbm(as.formula(formula.check), data = train,
+                        distribution = "bernoulli", n.trees = 10000,
+                        shrinkage = 0.001, interaction.depth = 5,
+                        n.minobsinnode = 20)
+                pred.c <- predict(check.fit, test, n.tree = 6000, type = "response")
+        }
+
+        #assign fit to global
+        assign("fit", fit, envir = .GlobalEnv)
+
+        #pROC
+        pred.p <- pROC::roc(test$f.rpdol, pred)
+
+        pROC::plot.roc(pred.p, col = "red", print.thres = F)
+
+        print(pROC::auc(pred.p))
+
+        #print check
+        pred.pc <- pROC::roc(test$f.rpdol, pred.c)
+        pROC::plot.roc(pred.pc, col = "blue", add = TRUE , print.thres = F)
+
+        #print AUC's
+        print(
+                c("blend" = pROC::auc(pred.p), "check" = pROC::auc(pred.pc))
+        )
+
+}
+
+find.auc(method = "log")
+#function saves model as 'fit'
+summary(fit)
+#not many of the interactions had significant relationships
 
@@
 ```{r pressure, echo=FALSE}
-plot(pressure)
+
+#use caret to tune 
+caret.all(df = train, formula = formula.rf, sample.frac = 0.75, seed = 100)
+
+find.auc(method = "rf")
+importance(fit, type = 1) #permutation importance
+#indicates norm.log is more important
+
 ```
@@
 ```{r pressure, echo=FALSE}
-plot(pressure)
+#tune parameters
+
+# shrinkage = 0.0005
+# n.trees = 6000
+# interaction.depth = 3
+# n.minobsinnode = 40
+
+find.auc(method = "boost")
+summary(fit)
+# norm.log was by far the most important.
+
 ```
+
+Hm none of the blends are doing better than models alone.
+Try comparing how similar the predictions are
+
+```{r}
+
+#plain model averaging
+
+#how to normalise the prediction variables
+all10.blend = mutate(all10.blend, model.av = (log + rf +boost)/3)
+
+pred.p <- pROC::roc(all10.blend$f.rpdol, all10.blend$model.av)
+
+pROC::plot.roc(pred.p, col = "red", print.thres = F)
+pROC::plot.roc(pred.p, col = "blue", print.thres = F, add = T)
+
+#hmm model averaging actually makes a bit of an improvement, weird.
+
+# what about boosted tree with only 3 variables?
+# or log regression wiht only 3 variables?
+
+```
+
+Try majority vote

### 2016-06-01 functions.R

@@ -1,0 +1,51 @@
+#function for running caret..
+
+caret.all <- function(method = "boost", df = train, formula = formula.rf, sample.frac = 0.75, seed = 100){
+        # set seeds
+        set.seed(seed)
+        seeds <- vector(mode = "list", length = 51)
+        for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
+        ## For the last model:
+        seeds[[51]] <- sample.int(1000, 1)
+        
+        train.c = sample_frac(df, sample.frac, replace = FALSE)
+        test.c = setdiff(df, train.c)
+        
+        cv.Control <- trainControl(method = "cv",
+                                   number = 5,
+                                   seeds= seeds
+                                   , summaryFunction= twoClassSummary,
+                                   classProbs=TRUE
+                                   )
+        
+        if(method == "rf"){
+                mtry.Grid <- expand.grid(mtry = c(3,4,5,6,7))
+                
+                set.seed(2)
+                gbmFit <- caret::train(as.formula(formula), data= train.c,
+                                       method = "rf", trControl = cv.Control, verbose = FALSE,
+                                       tuneGrid = mtry.Grid, ntree = 1000,
+                                       metric = 'ROC', allowParallel = T)
+        }
+        
+        if(method == "boost"){
+                Grid <- expand.grid(shrinkage = c(.0005),
+                                         n.trees = c(6000),
+                                    interaction.depth = c(3,4,5),
+                                    n.minobsinnode = c(10,20,30,40))
+                
+                # , interaction.depth = c(1,2,3,4,5),
+                # n.minobsinnode = c(10,20,30)
+                
+                set.seed(2)
+                gbmFit <- caret::train(as.formula(formula), data= train.c,
+                                method = "gbm", trControl = cv.Control, verbose = FALSE,
+                                bag.fraction = 0.5, tuneGrid = Grid,
+                                metric = 'ROC',
+                                na.action = na.pass)
+        }
+        
+        
+        plot(gbmFit)
+        
+}

### 2016-06-05 finalwriteup/draft.rmd
almost finished model blending

@@
         # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+        # setwd("~/Documents/github/consult/finalwriteup/report_data")
         opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/report_data')
@@
 
+#### Blended Models
+
+Blended models combine the results of each well-performing individual model to create averaged or 'blended' predictions.
+In this case, the randomForest, boosted trees, and logistic regression models were chosen for blending.
+The simplest way of combing models is to average the probability results (a number between 0 and 1) of all three models.
+Feature Weight Linear Stacking (FWLS), randomForest and boosted trees were all tested as blending methods where the predictions from each of the three single models becomes an additional variable in a new model.
+Each model blending method has its own method of weighting the outcomes from the individual models as a function of the values in the original model.
+For example, for a certain discipline in the case study company, randomForests may predict the outcome best so this is weighted higher for those cases.
+For this reason, a variable importance study was conducted again to simplify the models.
+
+##### Variable Selection: Blended Model
+
+*Variable Importance: cforest*
+
+```{r}
+
+cfor.varimp = readRDS("report_data/cfor_blend_varimp.rds")
+
+ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) + 
+        geom_bar(stat='identity') + 
+                        labs(title = "Cforest variable importance: blend")
+
+```
+
+*Variable Importance: Logistic Regression*
+
+```{r}
+
+log.varimp = readRDS("report_data/logblend.rds")
+
+log.varimp = summary(log.varimp)$coefficients %>% as.data.frame
+colnames(log.varimp)[names(log.varimp) %in% 'Pr(>|z|)']<-'P_value'
+log.varimp = add_rownames(log.varimp)
+log.varimp = log.varimp %>% filter(P_value <= 0.05) %>% 
+        select(rowname, Estimate, P_value) %>% arrange(P_value)
+
+# order levels
+log.varimp$rowname <- factor(log.varimp$rowname,
+                               levels= log.varimp$rowname)
+
+#plot p values
+# ggplot(data=log.varimp, aes(x=rowname, y=P_value)) +
+#         theme(axis.text.x=element_text(angle=45,hjust=1)) + 
+#         geom_bar(stat='identity') +
+#         labs(title = "LogReg variable importance: blend")
+
+```
+
+Most significant variables (P value for coefficient < 0.05) are:
+
+Interaction term between:
+
+* Discipline * results of logistic regression model
+* Discipline * results of boosted tree model
+* number of employees on a job * results of logistic regression model
+* Client Business type * results of boosted tree model
+
+Single variables that aren't in interaction terms:
+
+* Total amount invoiced for the job
+* Timespan
+* Project category
+
+*Variable Importance: randomForest*
+
+```{r}
+
+rf.varimp = readRDS("report_data/rfblend.rds")
+rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
+rf.varimp = add_rownames(rf.varimp)
+rf.varimp = arrange(rf.varimp, desc(MeanDecreaseAccuracy))
+# order levels
+rf.varimp$rowname <- factor(rf.varimp$rowname,
+                               levels= rf.varimp$rowname)
+
+#plot p values
+ggplot(data=rf.varimp, aes(x=rowname, y=MeanDecreaseAccuracy)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        geom_bar(stat='identity') +
+        labs(title = "randomForest variable importance: blend")
+
+```
+
+Permutation importance by randomForest rates all three prediction models as most important followed by number of employees on a project, percent of hours done by the chief employee on the project, timespan, and project category.
+
+*Variable Importance: boosted trees*
+
+```{r}
+
+boost.varimp = readRDS("report_data/boostblend.rds")
+boost.varimp = add_rownames(boost.varimp)
+# order levels
+boost.varimp$rowname <- factor(boost.varimp$rowname,
+                               levels= boost.varimp$rowname)
+
+#plot p values
+ggplot(data=boost.varimp, aes(x=rowname, y=rel.inf)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        geom_bar(stat='identity') +
+        labs(title = "Boosted Trees variable importance: blend")
+
+```
+Boosted trees ranks all three individual prediction models highest followed by Project category, Client Business type, and timespan.
+
+*Variable Importance Summary*
+
+Upon reviewing important variables ranked by each method of blended models, the following variables were deemed important by at least two methods and shall be used in the comparison of blended models:
+
+* all three results of prediction models, particularly logistic regression
+* project category
+* Client Business type
+* Total Amount invoiced for a project
+* Number of employees on a project
+* Timespan
+
+#### Comparison of Blended Models
+
+Four model blending techniques were compared using the 8 variables determined above.
+These are FWLS, randomForest, boosted trees, simple averaging and were compared against logistic regression.
+Initially 20 train/test runs were performed for each method.
+
 # Results - over 5 pages, up to 20 pages

### 2016-06-05 model_blend.rmd
almost finished model blending

@@
 library("dplyr")
+library('tidyr')
 library('magrittr')
@@
         all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
+        source("/Users/yam/Documents/github/consult/functions.R")
+
 }
@@
 
+write.csv(all10.blend, '~/OneDrive/shared files/Bligh Tanner/masters/data/all10blend.csv')
+
 ggplot(all10.blend, aes(x= norm.log)) + geom_histogram()
@@
 
-formula.boost <- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + Billing.Type + norm.log + norm.rf + norm.boost"
+formula.boost <- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + Billing.Type + norm.log + norm.rf + norm.boost + JD.Second"
 
@@
 for(j in seq_along(methods)){
-
         for(m in seq_along(metafeatures)){
-
                 formula.log <- paste(formula.log, " + ", metafeatures[m], "*", methods[j], sep = "")
-
         }
 }
-
 #add terms to allow for constants [@Sill2009] - already done as everything is included as a single variable without interaction
@@
         assign("fit", fit, envir = .GlobalEnv)
+        saveRDS(fit, paste("/Users/yam/Documents/github/consult/finalwriteup/report_data/",
+                           method,"blend.rds", sep = ""))
 
@@
 #use caret to tune 
-caret.all(df = train, formula = formula.rf, sample.frac = 0.75, seed = 100)
+caret.all(method = "rf", df = train, formula = formula.rf, sample.frac = 0.75, seed = 100)
 
@@
 summary(fit)
+saveRDS(summary(fit), paste("/Users/yam/Documents/github/consult/finalwriteup/report_data/",
+                           method,"blend.rds", sep = ""))
 # norm.log was by far the most important.
@@
 
-Try majority vote
+Variable importance from cforest
+```{r}
+
+cfor.plotsave(df= all10.blend, name= 'cfor_blend', formular = formula.rf)
+
+```
+
+20 runs of each method. compare against logistic regression
+8 variables have been selected via variable importance analysis
+
+```{r}
+
+formula.log <- "f.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost"
+
+formula.rf <- "f.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost"
+
+formula.boost <- "b.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost"
+
+formula.sboost <- "b.rpdol ~ norm.log + norm.rf + norm.boost"
+
+#formula to add interaction terms for logistic regression
+metafeatures = strsplit(formula.log, split = c("\\s\\~\\s|\\s\\+\\s"))[[1]]
+methods = metafeatures[7:9]
+metafeatures = metafeatures[2:6]
+
+for(j in seq_along(methods)){
+        for(m in seq_along(metafeatures)){
+                formula.log <- paste(formula.log, " + ", metafeatures[m], "*", methods[j], sep = "")
+        }
+}
+
+# tune boosted trees and rf
+caret.all(df = train, formula = formula.rf, sample.frac = 0.75, seed = 100)
+#mtry = 3
+caret.all(df = train, formula = formula.sboost, sample.frac = 0.75, seed = 100)
+# interaction.depth = 6, nminobs = 10, shrink = 0.0005, trees = 6000
+caret.all(df = train, formula = formula.boost, sample.frac = 0.75, seed = 100)
+# interaction.depth = , nminobs = , shrink = 0.0005, trees = 6000
+
+iter.auc <- function(mult = 4, x.folds = 5, 
+                     methods = c("orig.log","log","rf","boost","average",
+                                 "simp.log","simp.boost"),
+                     seedy = 100, df = all10.blend) {
+
+        #create test and train set: 0.75/0.25
+        set.seed(seedy)
+        auc <- rep(list(rep(0, mult*x.folds)), length(methods))
+        names(auc) = methods
+
+        #create model averaged column
+        df = mutate(df, model.av = (log + rf +boost)/3)
+
+        #set up 10 fold cv using caret function
+        folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
+        #now have folds[[1]] through folds[[10]] list of rows to exclude
+        if(mult > 1){
+                for(k in 2:mult){                                
+                    set.seed(sample(1:1000,1))
+                    folds.temp = createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain=F)
+                    folds = append(folds, folds.temp)
+                }
+        }
+
+        for(i in seq_along(folds)){
+
+        train = df[-folds[[i]],]
+        test = df[folds[[i]],]
+
+        if ("log" %in% methods) {
+                print("log")
+                fit <- glm(as.formula(formula.log),
+                               data = train,
+                               family = binomial())
+
+                pred <- predict(fit, test, type = "response")
+                auc[["log"]][i] <- pROC::roc(test$f.rpdol, pred) %>% auc() %>% round(3)
+        }
+
+        if ("simp.log" %in% methods) {
+                print("simp.log")
+                fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost,
+                               data = train,
+                               family = binomial())
+
+                pred <- predict(fit, test, type = "response")
+                auc[["simp.log"]][i] <- pROC::roc(test$f.rpdol, pred) %>% auc() %>% round(3)
+        }
+
+        if ("rf" %in% methods) {
+                print("rf")
+        fit <- randomForest( as.formula(formula.rf), data = train,
+                mtry = 3, ntree = 1000, importance = TRUE
+                )
+        pred <- predict(fit, test, type = "prob")
+        pred <- pred[, 2]
+
+        auc[["rf"]][i] <- pROC::roc(test$f.rpdol, pred) %>% auc() %>% round(3)
+
+        }
+
+        if ("simp.boost" %in% methods) {
+                print("simp.boost")
+                fit <- gbm(b.rpdol ~ norm.log + norm.rf + norm.boost, data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 6,
+                                   n.minobsinnode = 10)
+                pred <- predict(fit, test, n.tree = 6000, type = "response")
+                auc[["simp.boost"]][i] <- pROC::roc(test$f.rpdol, pred) %>% auc() %>% round(3)
+
+        }
+
+        if ("boost" %in% methods) {
+                print("boost")
+                fit <- gbm(as.formula(formula.boost), data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 3,
+                                   n.minobsinnode = 10)
+                pred <- predict(fit, test, n.tree = 6000, type = "response")
+                auc[["boost"]][i] <- pROC::roc(test$f.rpdol, pred) %>% auc() %>% round(3)
+
+        }
+
+        if ("orig.log" %in% methods) {
+                print("orig.log")
+                pred <- test$log
+                auc[["orig.log"]][i] <- pROC::roc(test$f.rpdol, pred) %>% auc() %>% round(3)
+
+        }
+
+        if ("average" %in% methods) {
+                print("average")
+                pred <- test$model.av
+                auc[["average"]][i] <- pROC::roc(test$f.rpdol, pred) %>% auc() %>% round(3)
+        }
+
+        #print AUC's
+        cat(i, " ")
+
+        }
+        #assign fit to global
+        assign("auc.list", auc, envir = .GlobalEnv)
+        # saveRDS(auc, paste("/Users/yam/Documents/github/consult/finalwriteup/report_data/",
+        #                    "auc_blend_20.rds", sep = ""))
+
+}
+
+iter.auc(mult = 10, x.folds = 10, seedy = 100, df = all10.blend,
+         methods = c("orig.log","log","average", "boost","rf", "simp.log","simp.boost"))
+
+auc.list
+saveRDS(auc.list, "/Users/yam/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
+
+ggplot(melt(auc.list), aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = L1, colour = L1)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 20 models of each Method") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression", "average" = "Simple Average",
+                         "orig.log" = "Orignal LogReg", "rf" = "Random Forest"))
+```

### 2016-06-05 functions.R

@@ -29,9 +29,12 @@ caret.all <- function(method = "boost", df = train, formula = formula.rf, sample
         }
         
         if(method == "boost"){
+                train.c$b.rpdol<- as.factor(train.c$b.rpdol)
+                levels(train.c$b.rpdol)[levels(train.c$b.rpdol)=="0"] <- "profit"
+                levels(train.c$b.rpdol)[levels(train.c$b.rpdol)=="1"] <- "loss"
                 Grid <- expand.grid(shrinkage = c(.0005),
                                          n.trees = c(6000),
-                                    interaction.depth = c(3,4,5),
+                                    interaction.depth = c(3,4,5,6),
                                     n.minobsinnode = c(10,20,30,40))
                 
                 # , interaction.depth = c(1,2,3,4,5),
@@ -45,7 +48,36 @@ caret.all <- function(method = "boost", df = train, formula = formula.rf, sample
                                 na.action = na.pass)
         }
         
-        
         plot(gbmFit)
+        return(gbmFit)
+        
+}
+
+cfor.plotsave<- function(df= vars, name= 'cfor_blend', formular = formula.rf){
+                
+                #run cforest
+                cfor.var1= cforest(as.formula(formular),
+                                   data= df)
+                #calculate variable importances
+                var.imp= varimp(cfor.var1) %>% as.data.frame
+                colnames(var.imp)[names(var.imp) %in% '.']= 'imp'
+                var.imp$var = rownames(var.imp)
+                #reorder levels for bar plot
+                var.imp<- within(var.imp, 
+                                 var <- factor(var,levels=var.imp[order(-abs(var.imp$imp)),]$var))
+                
+                q= ggplot(data=var.imp, aes(x=var, y=imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
+                r=q + geom_bar(stat='identity') + 
+                        labs(title = "Cforest variable importance: blend")
+                #save plot
+                # ggsave(plot=r,
+                #        filename=paste(name,i,'.png',sep=''),
+                #        path= 'C:/Users/n9232371/Documents/Consultbusiness/barplots/cforests/')
+                
+                print(r)
+                
+                #save datafraem for cforest plot
+                saveRDS(var.imp, "/Users/yam/Documents/github/consult/finalwriteup/report_data/cfor_blend_varimp.rds")
+                
         
 }

### 2016-06-06 finalwriteup/draft.rmd
finished model blending

@@
 
-opts_chunk$set(fig.width=8, fig.depth = 5, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
+opts_chunk$set(fig.width=10, fig.depth = 7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
 ```
@@
 
-Four model blending techniques were compared using the 8 variables determined above.
-These are FWLS, randomForest, boosted trees, simple averaging and were compared against logistic regression.
-Initially 20 train/test runs were performed for each method.
+Three model blending techniques were compared using the 8 variables determined above.
+These are FWLS, randomForest, and boosted trees.
+Furthermore, 3 simple model blending techniques were tested that used only the predictions of the three individual models as variables.
+These were averaging, a simple boosted tree, and a simple logistic regression.
+All six methods were compared against the original logistic regression predictions.
+Initially 100 train/test runs were performed for each method by running 20 iterations of 5-fold cross validation.
+The AUC results of these iterations are displayed below:
+
+```{r}
+
+auc.list <- readRDS("report_data/blend_auc1.rds")
+means = lapply(auc.list, mean) %>% unlist
+greater = means[means> means["orig.log"]] %>% names
+auc.melt = melt(auc.list)
+auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)
+
+ggplot(auc.melt, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = greater, colour = greater)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 Blended Models of each Method") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression", 
+                         "average" = "Simple Average",
+                         "orig.log" = "Orignal Logistic Regression", 
+                         "rf" = "Random Forest",
+                         "simp.boost" = "Simple Boosted Trees",
+                         "simp.log" = "Simple Logistic Regression")) +
+        scale_fill_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC greater \nthan Original Logistic \nRegression") +
+        scale_colour_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC greater \nthan Original Logistic \nRegression")
+
+```
+
+The plot shows that 4 methods could have a statistically significantly higher mean that the original logistic regression model: the simple blended logistic regression, the simple average, and the two boosted tree models.
+To achieve a power of 0.8, a two sample power analysis was performed on the 100 AUC results from each blended model against the AUC results from the original logistic regression.
+The boosted models required over 1000 samples while the simple averaged and simple logistic regression models required only 120.
+Therefore, 50 more iterations of the simple averaged and simple logistic models were performed since the boosted models results were not as promising.
+
+```{r}
+#load results from 150 iterations
+auc.150 = readRDS("report_data/blend_auc150.rds")
+
+```
+
+The results of the 150 iterations are shown in the plot below.
+Simple model averaging results in a p-value of `t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logisitc regression blended model had a two-sample t.test p-value of `t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
+
+```{r}
+
+index = c(2:3)
+pvals = data.frame("method" = names(auc.150)[c(2:3)], 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(auc.150[[1]], auc.150[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value<0.05) %>% .$method
+auc.150 = melt(auc.150)
+auc.150$signif = ifelse(auc.150$L1 %in% signif, TRUE, FALSE)
+
+ggplot(auc.150, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = signif, colour = signif)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y = "mean", geom="point", colour = "gray60", shape =3) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 150 Blended Models of Best Methods") +
+        scale_x_discrete(labels = c( 
+                         "average" = "Simple Average",
+                         "orig.log" = "Orignal Logistic Regression",
+                         "simp.log" = "Simple Logistic Regression")) +
+        scale_fill_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC significantly \ngreater than Original \nLogistic Regression") +
+        scale_colour_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC significantly \ngreater than Original \nLogistic Regression")
+
+```
+
+Simple logistic regression and averaging the the results of the three individual models (logistic regression, boosted trees and random forests) give virtually the same distribution of results.
+As averaging the results of the models is simpler than logistic regression, this was adopted as the best model.
 
@@
 
-The 95% confidence interval for this data is $`r (mean(all8a$return.pdol) - qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)` to $`r (mean(all8a$return.pdol) + qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)`.
+The 95% confidence interval for this data is $`r (mean(all8a$return.pdol)- qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)` to $`r (mean(all8a$return.pdol) + qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)`.
 If a project returns $0 for every dollar spent (return per dollar = $0), the project has earned only enough to cover the cost of doing the project (salary costs and business costs).
@@
 
-## Blended Model
+## Blended Models
+
+Trials of different methods of blending model results demonstrated that again, the simplest methods worked best with the data.
+In this case, averaging the results of the best three single models (logistic regression, random forests, and boosted trees) or taking a logistic regression of the three models outperformed more complex boosted trees, random forests, and logistic regression blends that facilitated interaction of the individual model results and original variables (meta-features).
+
+It was plausible that simple blending models would perform better than complex methods because the data was not big enough for complex models to learn intricate patterns.
+This was observed in the single method models when logistic regression performed as well as random forests and boosted trees.
+The literature does not extensively cover blended machine learning methods however their success in the 2009 Netflix competition generated some publications.
+The Netflix dataset comprised of almost 3,000,000 observations and the size of the datacould have facilitated the success of complex blending methods such as FWLS [@Sill2009].
+That being the case, it was still surprising that simple averaging of probabilities from the individual models performed as well as blending results using logistic regression.
+It was expected that logistic regression would have an advantage over model averaging due to its ability to weight the prediction variables while maintaining a simple model structure.
+
+The logistic regression blending model did indeed weight predictions from the original models differently.
+If the predicted probability from the original logistic regression model increased by 0.1, the odds of the model prediciting a loss making job increased by 37.4%.
+Similarly when probability predictions from the random forests model increased by 0.1, the odds of the blended model predicting a loss increased by 25.8%.
+On the other hand, for the same increase in the original boosted tree model, the odds of the blended logistic regression model predicting a loss increased by only 2.6%.
+This contrasted with the alternative simple model that averaged the probabilities from each original model (i.e.
+weighted them equally) but still managed to produce similar AUC results.
+It was not clear why this was the case, however both simple models were carried forward to analyse their respective bottom line impacts on the company's profits which could reveal performance differences.
+
+## Bottom Line Analysis
 
@@
 
-* this shows the profitability can be predicted with some success, but real question is, is it financially viable?
-* depending what the bottom line plot shows - yes or no. 
 * compare to past research. problems with past research and does this solve any of those?
@@
 
+* this shows the profitability can be predicted with some success, but real question is, is it financially viable?
+* depending what the bottom line plot shows - yes or no.
+
 # Conclusion

### 2016-06-06 model_blend.rmd
finished model blending

@@
 levels(all10mice$f.rpdol)[levels(all10mice$f.rpdol) == "1"] <- "loss"
+# first level denotes failure
+# models predict probability of success
 
@@
 
+#setwd in onedrive data
 write.csv(all10.blend, '~/OneDrive/shared files/Bligh Tanner/masters/data/all10blend.csv')
+# write.csv(all10.blend, 'all10blend2.csv')
+# all10.blend<- read.csv('all10blend2.csv')
+# ensure levels in correct order:
+all10.blend$f.rpdol =  factor(all10.blend$f.rpdol, levels= c("profit", "loss"))
 
@@
                 print("simp.log")
-                fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost,
+                fit <- glm(f.rpdol ~ log + rf + boost,
                                data = train,
@@
 
-iter.auc(mult = 10, x.folds = 10, seedy = 100, df = all10.blend,
+iter.auc(mult = 20, x.folds = 5, seedy = 100, df = all10.blend,
          methods = c("orig.log","log","average", "boost","rf", "simp.log","simp.boost"))
+# iter.auc(mult = 10, x.folds = 10, seedy = 100, df = all10.blend,
+#          methods = c("simp.log"))
 
-auc.list
-saveRDS(auc.list, "/Users/yam/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
+# simp.log.notrans <- auc.list
+# saveRDS(auc.list, "/Users/yam/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
+auc.list <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
+# auc.list <- append(auc.list, simp.log.notrans)
+# names(auc.list)[[8]] = "simp.log.notrans"
+#don't worry about simp.log.notrans  - transformation of model predictions doesn't improve results at all
 
@@
 ```
+
+Now to check power for each of the tests against original log
+
+```{r}
+index = c(2:7)
+ns = data.frame("method" = names(auc.list)[c(2:7)], "n"= NA)
+for(i in seq_along(index)){
+        temp = n.samples(data = auc.list[c(1, index[i])], type.calc = 'two.sample', alt = 'two.sided')
+        ns[i,2] = temp
+        ns[i,1] = names(auc.list)[index[i]]
+}
+ns
+
+```
+
+According to power test, need 150 tests to tell if simp.log or average is better than previosu logistic regression methods.
+Other methods require thousands (ie not much different): boost, simp.boost
+Other methods look worse: FLWS, rf
+
+```{r}
+
+iter.auc(mult = 30, x.folds = 5, seedy = 100, df = all10.blend,
+         methods = c("orig.log","average", "simp.log"))
+auc.150 = auc.list
+saveRDS(auc.150, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/blend_auc150.rds")
+
+ggplot(melt(auc.150), aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = L1, colour = L1)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 20 models of each Method") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression", "average" = "Simple Average",
+                         "orig.log" = "Orignal LogReg", "rf" = "Random Forest"))
+
+# n test for power = 0.8
+
+index = c(2:3)
+ns = data.frame("method" = names(auc.list)[c(2:3)], "n"= NA)
+for(i in seq_along(index)){
+        temp = n.samples(data = auc.150[c(1, index[i])], type.calc = 'two.sample', alt = 'two.sided')
+        ns[i,2] = temp
+        ns[i,1] = names(auc.list)[index[i]]
+}
+ns
+
+index = c(2:3)
+pvals = data.frame("method" = names(auc.list)[c(2:3)], 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(auc.150[[1]], auc.150[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+```
+
+Why does logistic regression perform the same as simple model averaging? lets check the coeffcients of the model.
+
+```{r}
+formula.slog <- "f.rpdol ~ log + rf + boost"
+
+fit.log = glm(as.formula(formula.slog),
+              data = train,
+              family = binomial())
+
+pred <- predict(fit.log, test, type = "response")
+pred.compare <- predict(fit.log, test, type = "terms")
+summary(fit.log)
+summary(fit.log)$coefficients %>% as.data.frame
+
+```

### 2016-06-06 functions.R

@@ -81,3 +81,29 @@ cfor.plotsave<- function(df= vars, name= 'cfor_blend', formular = formula.rf){
                 
         
 }
+
+
+# check number of samples required to achieve power of 0.8
+
+n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
+        if(type.calc == "one.sample"){
+                d.calc = abs(mean(data))/sd(data)
+                pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
+                                     power =.8, type = c(type.calc), 
+                                     alternative = alt)
+                return(ceiling(pow_pow$n))  
+        }
+        
+        if(type.calc == 'two.sample'){
+                d.calc = abs(mean(data[[1]])-mean(data[[2]]))/
+                        (sqrt((sd(data[[1]])^2+sd(data[[2]])^2)/2))
+                pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
+                                     power =.8, type = c(type.calc), 
+                                     alternative = 'two.sided')
+                return(ceiling(pow_pow$n)) 
+        }
+        
+}
+
+
+

### 2016-06-09 finalwriteup/draft.rmd
report writing profit curve

@@
 
+### Bottom Line Analysis
+
+The first research question, whether machine learning and statistical techniques can predict project profitability based on historic data, has been answered.
+The measure of profitability could not be modelled but simplifying the profitability prediction to a binary (profit/loss) model was successful.
+To answer the second research question, whether the model would improve the business' profitability, a bottom line analysis was performed.
+
+The models output a 'probability' between 0 and 1 that each project will be a loss-making job (where probability = 1).
+This leaves the question, at what probability would a decision-maker round the probability up to 1 or down to 0?
+And what business decision would then be made?
+To find the rounding threshold, an experimental business-scenario was tested.
+At a given threshold, all projects calculated above the threshold were considered too risky, and were rejected.
+All profits and losses from these projects were forfeited, and the profits from the remaining jobs (below the threshold) were summed to give a total profit.
+
+Using this logic, if the threshold was 0, all projects were rejected and total profit would equal 0.
+If the threshold was 1, all projects were accepted and the total profits would be the same as historical figures.
+Probability results from the two blended models as well as three individual models were tested and because the models require a training and testing set, the process was repeated 100 times with 20 x 5 fold cross validation models to achieve a statistical power of 0.8.
+The plot below illustrates the results of each model at each threshold with a local polynomial regression line of best fit (loess) and a 95% confidence interval around the mean.
+
+```{r profit_curve, fig.depth = 10}
+finance.plot = readRDS("report_data/plot_finance.rds" )
+
+ggplot(finance.plot, aes(x= threshold, y= profit.ratio, colour = method)) + 
+        geom_point(alpha = 0.1, shape = 16) + 
+        stat_smooth() +
+        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
+        theme(legend.position = "bottom") + 
+        scale_colour_discrete(labels = c("average" = "Blend - Average",
+                                       "orig.boost"= "Blend - Log Reg",
+                                       "orig.log" = "Logistic Regression",
+                                       "orig.rf" = "Random Forest",
+                                       "simp.log" = "Boosted Trees"))+
+        guides(col = guide_legend(nrow = 2)) +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Profit Curve for Blended and Individual Models")
+
+```
+
+The two blended models outperformed the individual models as shown in this table describing each models' optimal thresholds:
+
+```{r profit_table}
+
+finance.plot$method = as.factor(finance.plot$method)
+# want mean and sd for each method at each threshold
+
+fin.summary = finance.plot %>% group_by(threshold, method) %>% 
+        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
+                            sd_prof_ratio = round(sd(.), 1)),
+                       profit.ratio)
+
+#for each method find max mean
+
+best.fin = fin.summary %>% group_by(method) %>% 
+        filter(mean_profit_ratio == max(mean_profit_ratio)) %>% 
+        arrange(desc(mean_profit_ratio))
+kable(best.fin)
+
+best.mean = best.fin %>% filter(method == "average") %>% .$mean_profit_ratio
+best.sd = best.fin %>% filter(method == "average") %>% .$sd_prof_ratio
+
+```
+
+The average of the three individual models performed best with the highest mean profit ratio at `r best.mean`% and the narrowed standard deviation at `r best.sd`%.
+This means that for the blended average model, if all jobs above the probability threshold 0.55 were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profit data.
+
+For clarity, the plot for the average blended model is shown below:
+
+```{r av_profit_curve}
+ggplot(finance.plot %>% filter(method == "average"), 
+       aes(x= threshold, y= profit.ratio, colour = method)) + 
+        geom_point(alpha = 0.1, shape = 16) + 
+        stat_smooth() +
+        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
+        theme(legend.position = "none") +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Profit Curve for Blended Average Model")
+
+```
+
 # Results - over 5 pages, up to 20 pages
@@
 acknowledge that it may be worthwhile for other reasons such as confidence and transparency in historic projects.
-
-* tried model averaging - comparison to other methods.
 
@@
 comments on your results;
+* blended models significantly (?) outperformed individual models
+*
+
 explains what your results mean;
+
+* If the company's profits for one year were $1,000,000, the mean yearly increased profit would be `r (best.mean-100)/100*1000000` by using the best model and rejecting all jobs with a probability higher than `r best.fin %>% filter(method == 'average') %>% .$threshold`.
+
 interprets your results in a wider context; indicates which results were expected or unexpected;
+it was expected that it would show financial benefits and the results look reasonable - not particuarly high or low.
+it was also expected that blended model would outperform the single models.
+Interesting that simple model averaging ended up signficantly (?) outperformiing blended lgoistic regression model in profit curve.
+
 provides explanations for unexpected results.
+why did it end up outperforming? no idea.
+
 The Discussion should also relate your specific results to previous research or theory. 
-You should point out what the limitations were of your study, and note any questions that remain unanswered.
-The Discussion CAN also include Conclusions/Future Research.
-
-* compare to past research. problems with past research and does this solve any of those?
+* problems with past research and does this solve any of those?
     * poor user uptake - future work?
@@
 
-* this shows the profitability can be predicted with some success, but real question is, is it financially viable?
-* depending what the bottom line plot shows - yes or no.
+You should point out what the limitations were of your study, and note any questions that remain unanswered. 
+* not sure how easy number of users is to predict - could do a study
+* could do a study on trust levels
+
+The Discussion CAN also include Conclusions/Future Research.
+could be a viable business move worth integrating
 

### 2016-06-09 financial.rmd
report writing profit curve

@@
 
-```{r_loadpackages, echo=FALSE, include=FALSE}
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('randomForest', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
+```{r loadpackages, echo=FALSE, include=FALSE}
+
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library('ggplot2')
+library('gbm')
+library('randomForest')
+library('caret')
+library('pROC')
+library('testthat')
+library('purrr')
+library(tidyr)
+library(broom)
 
@@
 ```{r, echo=FALSE}
+
 test.pred <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/test_review.csv')
@@
         prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
+        # if model marks it as a positive (greater than threshold), eliminate any balance
         prof$profit<- rep(0, nrow(prof))
@@
 ```
+
+Need a new function for final analysis! comparing two blending methods
+Can i get rid of boosted tree model? that would speed things up a lot
+
+Need to try simple average and log reg without boosted trees. Reduces spread but doesnt improve mean. i'd like to keep the wide spread because you reach greater heights (and lower lows)
+
+```{r}
+if(.Platform$OS.type == 'windows'){
+        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
+        source("C:/Users/n9232371/Documents/github/consult/functions.R")
+} else{
+        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
+        source("/Users/yam/Documents/github/consult/functions.R")
+
+}
+
+profit <- function(df = all10mice, method = c("average","simp.log","orig.log","orig.rf","orig.boost"), 
+                   seed = 300, 
+                   k.folds = 5, # number of folds in initial model.blend generation
+                   max.iter = 5, #max number of iterations from initial model.blend generation
+                   k.folds.2 = 5, # number of folds for second step blended model development
+                   tot.tests = 50 # total number of results to create
+                   ){
+
+        length.k.folds = ceiling(tot.tests/k.folds.2)*k.folds.2
+
+        set.seed(seed)
+
+        # create f.rpdol
+        df$f.rpdol <- as.factor(df$b.rpdol)
+        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <-"profit"
+        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <-"loss"
+
+        # testthat check
+        expect_equal(levels(df$f.rpdol), c("profit", "loss"))
+
+        #loop through however many versions of df.blend are required
+        compile.results = NULL
+
+        for(m in 1:ceiling(tot.tests/max.iter)){
+
+        print(
+                paste("m =", m, "of", ceiling(tot.tests/max.iter),"\n", sep = ' ') %>% cat()
+        )
+
+        # create complete set
+        df.blend = assemble(methods = c("log","rf","boost"), df.assem = df, x.folds.assem = 5,
+                     manual.seed.assem = sample(1:1000,1))
+
+        df.blend = mutate(df.blend, norm.log = sign(log) * abs(log)^(1/3))
+        df.blend = mutate(df.blend, norm.rf = rf^0.5)
+        df.blend = mutate(df.blend, norm.boost = log(boost))
+
+        saveit <- iter.auc(mult = max.iter/k.folds.2, x.folds.iter = k.folds.2,
+                           methods = c("orig.log","average","simp.log", "orig.rf","orig.boost"),
+                           df.iter = df.blend)
+
+        dontknow = map(saveit, function(x) spread(x, iteration, pred))
+
+        if(m == 1){
+                compile.results = dontknow
+        } else {
+                compile.results = map2(compile.results, dontknow, function(x,y) merge(x, y, by = 'index'))
+        }
+
+        }
+
+        # saveRDS(compile.results, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
+        # compile.results <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
+
+        ####
+        # now have a suite of predictions - merge with original all10.blend$balance.mlsto
+        df.blend$index = rownames(df.blend) %>% as.numeric
+        # load balances
+        all6 <-read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+        all6 <- all6 %>% select(mlsto, balance.mlsto) %>% distinct()
+        all6$mlsto <- as.character(all6$mlsto)
+        df.blend = left_join(df.blend, all6)
+        compile.results <- map(compile.results, function(x) 
+                left_join(x, df.blend %>% select(index, balance.mlsto, b.rpdol)))
+
+        #create vector of profit points
+        threshold = seq(0, 1, by = .05)
+
+        # set up list of blank dataframes for profit calcs for each iteration: prof.results
+        colnames = names(compile.results[[1]])[grepl('iter', names(compile.results[[1]]))]
+        new.df = matrix(NA, nrow = length(threshold), ncol = ceiling(tot.tests/max.iter) + 1) %>% as.data.frame
+        colnames(new.df) <- c('threshold', colnames)
+        new.df$threshold = threshold 
+        prof.results = list(new.df, new.df, new.df, new.df, new.df)
+        names(prof.results) = names(compile.results)
+
+        # try eliminating outliers of balance over $200,000
+        # comp.results.2 <- lapply(compile.results, function(x) x = x %>% filter(abs(balance.mlsto)<200000))
+
+        prof.results <- map2(compile.results, prof.results, function(x,y) {
+
+                for(n in 2:ncol(y)){
+
+                        for (p in 1:length(threshold)) {
+
+                        # sum the balance.mlsto of all jobs that fall below the cut off threshold
+                        # this equals the total profits (or losses!) across all jobs that were completed
+                        # given that you reject any job over the threshold
+
+                        prof <- x[x[,n] < y$threshold[p], 'balance.mlsto']
+
+                        y[p, n] = sum(prof)
+
+                }
+
+                }
+
+                return(y)
+
+        })
+
+        prof.results <- map2(prof.results, names(prof.results), function(x,y) {
+                x$method = y
+                return(x)})
+
+        assign(prof.results, "prof.results", envir = .GlobalEnv)
+
+}
+
+# gather results for plotting
+plot.results = rbind_all(prof.results)
+plot.results = gather(plot.results, key = iteration, value = profit, -method, - threshold)
+
+# censor actual profit amount as fraction of historical profit level
+orig.prof = plot.results %>% filter(threshold == 1) %>% .$profit %>% .[1]
+plot.results = mutate(plot.results, profit.ratio = (profit/orig.prof*100) %>% round(1))
+
+# saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance.rds" )
+
+ggplot(plot.results, aes(x= threshold, y= profit.ratio, colour = method)) + 
+        geom_point(alpha = 0.1, shape = 16) + 
+        stat_smooth() +
+        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2)
+
+# why is it so shit? calc average prob for each iteration
+mean.prob = apply(compile.results[['simp.log']], 1, function(x) mean(x[2:11])) 
+analysis = data.frame(index = 1:length(mean.prob), mean.prob = mean.prob, 
+                      bal = compile.results[['simp.log']]$balance.mlsto)
+analysis %>% arrange(mean.prob) %>% head(10)
+ggplot(analysis, aes(x = mean.prob, y = bal)) + geom_point()
+
+```
+
+Now need to check how mnay experiments to do.. power calc.
+
+```{r}
+
+#power calc - where should I do it from?
+
+```
+
+Check what happens when you increase the amount you invoiced by a certain percentage
+
+```{r}
+
+# what if you find for jobs over a certain threshold, increase fee by ?% still do all jobs
+compile.results <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
+####
+# now have a suite of predictions - merge with original all10.blend$balance.mlsto
+df.blend$index = rownames(df.blend) %>% as.numeric
+# load balances
+all6 <-read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+all6 <- all6 %>% select(mlsto, balance.mlsto, inv.mlsto, cost.mlsto) %>% distinct()
+all6$mlsto <- as.character(all6$mlsto)
+df.blend = left_join(df.blend, all6)
+compile.results <- map(compile.results, function(x) {
+        left_join(x, df.blend %>% select(index, balance.mlsto, b.rpdol, inv.mlsto, cost.mlsto))})
+
+#create vector of profit points
+threshold = seq(0, 1, by = .05)
+
+# set up list of blank dataframes for profit calcs for each iteration: prof.results
+colnames = names(compile.results[[1]])[grepl('iter', names(compile.results[[1]]))]
+new.df = matrix(NA, nrow = length(threshold), ncol = ceiling(tot.tests/max.iter) + 1) %>% as.data.frame
+colnames(new.df) <- c('threshold', colnames)
+new.df$threshold = threshold 
+prof.results = list(new.df, new.df, new.df, new.df, new.df)
+names(prof.results) = names(compile.results)
+
+prof.results <- map2(compile.results, prof.results, function(x,y) {
+
+        for(n in 2:ncol(y)){
+
+                for (p in 1:length(threshold)) {
+
+                # sum the balance.mlsto of all jobs - you do them all
+                # then sum a 10% increase in what you charge for all jobs above a certain threshold
+
+                        prof <- x[, 'balance.mlsto']
+                        increase <- x[x[,n] > y$threshold[p], 'inv.mlsto']*0.1
+
+                        y[p, n] = sum(prof) + sum(increase)
+
+                }
+
+        }
+                return(y)
+})
+
+prof.results <- map2(prof.results, names(prof.results), function(x,y) {
+        x$method = y
+        return(x)})
+
+assign(prof.results, "prof.results", envir = .GlobalEnv)
+
+# gather results for plotting
+plot.results = rbind_all(prof.results)
+plot.results = gather(plot.results, key = iteration, value = profit, -method, - threshold)
+
+ggplot(plot.results, aes(x= threshold, y= profit, colour = method)) + 
+        geom_point(alpha = 0.1, shape = 16) + 
+        stat_smooth()
+
+# doesn't reveal much - better off doing a simple check once decided threshold.
+
+# I like simp.log best.
+
+simp.l <- compile.results[['simp.log']]
+simp.l$mean.iter <- rowMeans(simp.l[,2:11])
+
+# How many jobs are above threshold of 0.65?
+simp.l %>% filter(mean.iter>0.65) %>% nrow
+# 74 - not many
+
+#how much was lost above 0.65?
+initial = simp.l %>% filter(mean.iter>0.65) %>% .$balance.mlsto %>% sum
+# $600000
+# what if increased invoice amount by 0.1?
+costs = simp.l %>% filter(mean.iter>0.65) %>% .$cost.mlsto %>% sum
+simp.l %>% filter(mean.iter>0.65) %>% mutate(frac = inv.mlsto*0.27) %>% .$frac %>% sum
+
+# would have to increase all the jobs which rated above 0.65 by 27% to break even.
+# would have to increase all the jobs which rated above 0.65 by 46% to generate a 15% profit.
+# this is probably impractical - hard to compete or argue for it. could also go for Hourly rate billing
+
+profit.15 = costs*.15
+
+```

### 2016-06-09 model_blend.rmd
report writing profit curve

@@
 
+formula.slog <- "f.rpdol ~ norm.log + norm.rf"
+
 formula.rf <- "f.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost"
@@
         #create model averaged column
-        df = mutate(df, model.av = (log + rf +boost)/3)
+        df = mutate(df, model.av = (log + rf + boost)/3)
 
@@
          methods = c("orig.log","log","average", "boost","rf", "simp.log","simp.boost"))
-# iter.auc(mult = 10, x.folds = 10, seedy = 100, df = all10.blend,
-#          methods = c("simp.log"))
-
-# simp.log.notrans <- auc.list
+# iter.auc(mult = 20, x.folds = 5, seedy = 100, df = all10.blend,
+#          methods = c("simp.log", "average"))
+# 
+# no.boost <- auc.list
 # saveRDS(auc.list, "/Users/yam/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
 auc.list <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
-# auc.list <- append(auc.list, simp.log.notrans)
-# names(auc.list)[[8]] = "simp.log.notrans"
+# auc.list <- append(auc.list, no.boost)
+# names(auc.list)[c(8,9)] = c("simp.log.noboost","average.noboost")
 #don't worry about simp.log.notrans  - transformation of model predictions doesn't improve results at all

### 2016-06-10 finalwriteup/draft.rmd
almost finished finance discussion

@@
 
-mice.totAUC <- readRDS("mice_df_auc.rds")
-
-#significant difference??
-index = c(1,3,4)
-pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
-for(i in seq_along(index)){
-        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
-                format(digits = 2) %>% as.numeric
-}
-
-signif = pvals %>% filter(p_value < 0.05) %>% .$method
-mice.totAUC = melt(mice.totAUC)
-mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)
-
-ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
-        scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different \nfrom boosted trees") +
-        scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different \nfrom boosted trees") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method - Imputed Data") +
-        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
-                         "log" = "Logistic Regression",
-                         "naiive" = "Naiive Bayes", "rf" = "Random Forest"))
+mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
 
@@
         scale_colour_discrete(labels = c("average" = "Blend - Average",
-                                       "orig.boost"= "Blend - Log Reg",
+                                       "orig.boost"= "Boosted Trees",
                                        "orig.log" = "Logistic Regression",
                                        "orig.rf" = "Random Forest",
-                                       "simp.log" = "Boosted Trees"))+
+                                       "simp.log" = "Blend - Log Reg"))+
         guides(col = guide_legend(nrow = 2)) +
@@
 best.sd = best.fin %>% filter(method == "average") %>% .$sd_prof_ratio
+best.t = best.fin %>% filter(method == "average") %>% .$threshold
 
@@
 
-comments on your results;
-* blended models significantly (?) outperformed individual models
-*
-
-explains what your results mean;
-
-* If the company's profits for one year were $1,000,000, the mean yearly increased profit would be `r (best.mean-100)/100*1000000` by using the best model and rejecting all jobs with a probability higher than `r best.fin %>% filter(method == 'average') %>% .$threshold`.
-
-interprets your results in a wider context; indicates which results were expected or unexpected;
-it was expected that it would show financial benefits and the results look reasonable - not particuarly high or low.
-it was also expected that blended model would outperform the single models.
-Interesting that simple model averaging ended up signficantly (?) outperformiing blended lgoistic regression model in profit curve.
-
-provides explanations for unexpected results.
-why did it end up outperforming? no idea.
-
-The Discussion should also relate your specific results to previous research or theory. 
+```{r load_winner_results}
+
+simp.l <- readRDS("report_data/finance_average_model.rds")
+
+```
+
+The blended models significantly (?) outperformed individual models which was expected given the two blended models' AUC results were significantly higher than the original logistic regression results (which were in line with random forests' and boosted trees' results).
+The simple average of logistic regression, random forests, and boosted trees results produced a better profit curve than a logistic regression of the same three variables.
+The model produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`.
+To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r (best.mean-100)/100*1000000` by rejecting all jobs with a probability higher than `r best.t`.
+In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.
+The improvements in profit sound promising, however it is important to consider possible limitations in the data which is discussed in the next section.
+
+Because of the positive AUC results, it was expected the models would translate into financial benefits for the company.
+The results look reasonable, profit gains may be high enough to trigger further cost-benefit analyses for introducing the model into business decisons.
+It was surprising that, given the similar AUC results, simple model averaging signficantly (?) outperformed the blended logistic regression model's profit curve.
+This may be because ...
+ROC curves.
+
+*** you are here ****
+
+Previous research indicated problems with user uptake once models were developed and a lack of trust in the model output [@Akintoye2000].
+
 * problems with past research and does this solve any of those?
     * poor user uptake - future work?
-    * lack of trust - nearest neighbour must be presented with results - case bsaed reasoning
+    * lack of trust - nearest neighbour must be presented with results - case based reasoning
     * results from boosted tree shows partial dependence plots - insight.
@@
 
-The Discussion CAN also include Conclusions/Future Research.
-could be a viable business move worth integrating
-
 # Conclusion
@@
 * user testing on trust levels of prediction with nearest neighbour
+
+Conclusions
+could be a viable business move worth integrating
+
+Suggestions
+
+Could collect data for future projects: expected team size, expected time frame.
+
+* trialled ensemble tree methods on cost estimation (in this case project proftability)
+* attempted to predict project profitability using a single consulting company's CRM data in the construction industry
+* assessed tangible profit outcomes of machine learning models

### 2016-06-10 financial.rmd
almost finished finance discussion

@@
 # why is it so shit? calc average prob for each iteration
-mean.prob = apply(compile.results[['simp.log']], 1, function(x) mean(x[2:11])) 
+mean.prob = apply(compile.results[['average']], 1, function(x) mean(x[2:11])) 
 analysis = data.frame(index = 1:length(mean.prob), mean.prob = mean.prob, 
@@
 
-simp.l <- compile.results[['simp.log']]
+simp.l <- compile.results[['average']]
 simp.l$mean.iter <- rowMeans(simp.l[,2:11])
+
+saveRDS(simp.l, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_average_model.rds")
 

### 2016-06-13 finalwriteup/draft.rmd
finished rough draft for paul

@@
 
-opts_chunk$set(fig.width=10, fig.depth = 7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
+opts_chunk$set(fig.width=8, fig.depth = 8, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
 ```
@@
 
-B2B- Business to Business
+B2B - Business to Business
+CRM - Customer Relationship Management.
+Refers to a common type of software used by businesses to record client and project details as well as employee timesheet records.
+Generally accessable to each employee in a company.
 
@@
 # Introduction - 2 pages
-
-* vague interesting few sentences - context
 
@@
 
-****
-
-# Chapter 1 Problem Description
+# Problem Description
 
@@
 
-## 1.1 Research Motivations - 4.5 pages
-
-description of problem:
+## Research Motivation - 4.5 pages
+
+**Problem Description:**
 
@@
 
-Why its not easily solvable:
+**Why its Not Easily Solvable:**
 
@@
 
-current practise:
+**Current Practise:**
 
@@
 It is also acknowledged that cost estimation is a problem and hypothesised that inexperience and lack of time are chief contributors [@Akintoye2000].
-
-theory about problem with current practise:
 
@@
 
-previous research:
-
 The theory of reference class forecasting was first clearly published in 2003, however the idea of predicting project costs from previous data has been the subject of research for many years before that and continues to this day.
@@
 
-### 1.1.1 Overview of Past Reference Class Forecasting Methods and relevant Industry Uptake
+### Overview of Past Reference Class Forecasting Methods and relevant Industry Uptake
 
@@
 
-***need to really do more research on this!!***
-
-#### 1.1.1.1 Existing Models
+#### Existing Models
 
@@
 
-#### 1.1.1.2 Industry Uptake of Existing Models and Current Practises
+#### Industry Uptake of Existing Models and Current Practises
 
@@
 
-### 1.1.2 Case Study
+### Case Study
 
@@
 
-### 1.1.3 Case Study Limitations
+### Case Study Limitations
 
@@
 
-### 1.1.4 Research Motivation Summary
+### Research Motivation Summary
 
@@
 
-## 2.0 Thesis aims
+## Thesis aims
 
@@
 
-## 1.2 Thesis contributions - Gaps in literature
+## Thesis Contributions - Gaps in literature
 
@@
 
-* the application of clustering methods to business to business decisions.
-
-## 1.3 Thesis structure
-
-lit review
-Background info
-present the case study
-Present the models you've developed
-summary of findings, implications for use within industry, limitations
-
-# Chapter 2 Literature Review - 25 pages for pHd
+# Literature Review - 25 pages for pHd
 
@@
 
-## 2.1 Project Cost Estimation Methods
+## Project Cost Estimation Methods
 
@@
 
-### 2.1.1 Cost Estimation in the Construction Industry
+### Cost Estimation in the Construction Industry
 
@@
 
-**summary**
+**Summary**
 
@@
 
-### 2.1.1 Effort Estimation in the IT Industry
+### Effort Estimation in the IT Industry
 
@@
 
-## 2.2 Statistical and machine learning methods
-
-### 2.2.1 Introduction
+## Statistical and machine learning methods
+
+### Introduction
 
@@
 
-### 2.2.2 Linear Regression
+### Linear Regression
 
@@
 
-### 2.2.3 Logistic Regression
+### Logistic Regression
 
@@
 
-### 2.2.4 Naive Bayes
+### Naive Bayes
 
@@
 
-### 2.2.5 Decision Trees/ Ensemble Trees
+### Decision Trees/ Ensemble Trees
 
@@
 
-### 2.2.5 Bayesian Networks
+### Bayesian Networks
 
@@
 
-### 2.2.6 SVM's and Neural Networks
+### SVM's and Neural Networks
 
@@
 
-### 2.2.7 Summary
+### Summary
 
@@
 
-## 2.3 Statistical and machine learning methods applied to business problems
+## Statistical and Machine Learning Methods Applied to Business Problems
 
@@
 
-### 2.3.2 Summary of Advanced Statistical and Machine Learning Methods Applied to Business Problems
+### Summary of Advanced Methods
 
@@
 
-### 2.3.3 Employee Churn Case Study
+### Employee Churn Case Study
 
@@
 
-### 2.3.4 Gap/conclusion
+### Gap
 
@@
 
-## 2.4 Conclusion
+## Conclusion
 
@@
 
-# Chapter 3 Methods - 2 to 8 pages
+# Chapter 3 Method - 2 to 8 pages
 
@@
 
-**justify the process by which the research questions were answered**
-**argue and justify each decision that was taken to arrive at how to organise reserach**
-
-## 3.1 Obtaining data
+## Obtaining data
 
@@
 
-## 3.2 Cleaning and variable engineering
+## Cleaning and variable engineering
 
@@
 
-* invoicing data
-* timesheet data
-* project summary data
+* Invoicing data
+* Timesheet data
+* Project summary data
 
@@
 
-### 3.2.1 Variable Engineering
+### Variable Engineering
 
@@
 
-## 3.3 Variable selection
-
-### 3.3.1 Outlier deletion
+## Variable Selection
+
+### Outlier Deletion
 
@@
 
-### 3.3.2 Variable selection Methods
+### Variable Selection Methods
 
@@
 
-#### 3.3.2.1 ANOVA
+#### ANOVA
 
@@
 
-#### 3.3.2.2 Random Forest
+#### Random Forest
 
@@
 
-#### 3.3.2.3 Cforests
+#### Cforests
 
@@
 
-#### 3.3.3 Variable Selection Summary
+#### Variable Selection Summary
 
@@
 
-## 3.4 Model selection
+## Model selection
 
@@
 
-## 3.5 Model comparison, Model averaging
+## Model Comparison, Model Averaging
 
@@
 
-## 3.6 Bottom Line Translation
+## Bottom Line Translation
 
@@
 
-## 3.7 Method Conclusion
-
-In summary, to determine whether a 
-* obtained data: 4169 jobs from a single case study company over 12 years. Data was de-dentified by an employee of the company
-* the data was all from the company CRM. There were three chief sources:
-    * invoicing data
-    * timesheet data
-    * project summary data
-* the data was cleaned from mistakes which were generally discovered as outliers 
-* many new variables were engineered from the data, often by summarising time series data into a value that is relevant to a final data set of a single job per row
-* variable selection: dozens of variables available which were condensed to the most important 10-12.
+```{r}
+## Method Conclusion
+
+# In summary, to determine whether a 
+# * obtained data: 4169 jobs from a single case study company over 12 years. Data was de-dentified by an employee of the company
+# * the data was all from the company CRM. There were three chief sources:
+#     * invoicing data
+#     * timesheet data
+#     * project summary data
+# * the data was cleaned from mistakes which were generally discovered as outliers 
+# * many new variables were engineered from the data, often by summarising time series data into a value that is relevant to a final data set of a single job per row
+# * variable selection: dozens of variables available which were condensed to the most important 10-12.
 variable importance methods were ANOVA, random forest, and cforests.
 the overlapping important variables were chosen first as well as some others
-* 5 models were tested: baseiline: log reg, naiive bayes.
-* sophisticated ML: boosted reg, random frest, bayesian networks
-* best were comapred to a model averaged version
-* Analysis of impact on bottom line for business if model is integrated into decision making
-* link: results
-
-# Chapter 4: Results
+# * 5 models were tested: baseiline: log reg, naiive bayes.
+# * sophisticated ML: boosted reg, random frest, bayesian networks
+# * best were comapred to a model averaged version
+# * Analysis of impact on bottom line for business if model is integrated into decision making
+# * link: results
+```
+
+# Results
 
@@
 
-### cForest
+### Cforest
 
@@
 
-### Variable selection summary
+### Variable Selection Summary
 
@@
 
-mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
-
+mice.totAUC <- readRDS("mice_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+mice.totAUC = melt(mice.totAUC)
+mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different \nfrom boosted trees") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different \nfrom boosted trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of each Method - Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naiive Bayes", "rf" = "Random Forest"))
 ```
@@
 
-cfor.varimp = readRDS("report_data/cfor_blend_varimp.rds")
+cfor.varimp = readRDS("cfor_blend_varimp.rds")
 
@@
 
-log.varimp = readRDS("report_data/logblend.rds")
+log.varimp = readRDS("logblend.rds")
 
@@
 
-rf.varimp = readRDS("report_data/rfblend.rds")
+rf.varimp = readRDS("rfblend.rds")
 rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
@@
 
-boost.varimp = readRDS("report_data/boostblend.rds")
+boost.varimp = readRDS("boostblend.rds")
 boost.varimp = add_rownames(boost.varimp)
@@
 
-```{r}
-
-auc.list <- readRDS("report_data/blend_auc1.rds")
+```{r, fig.height=10}
+
+auc.list <- readRDS("blend_auc1.rds")
 means = lapply(auc.list, mean) %>% unlist
@@
 #load results from 150 iterations
-auc.150 = readRDS("report_data/blend_auc150.rds")
+auc.150 = readRDS("blend_auc150.rds")
 
@@
 The results of the 150 iterations are shown in the plot below.
-Simple model averaging results in a p-value of `t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logisitc regression blended model had a two-sample t.test p-value of `t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
-
-```{r}
+Simple model averaging results in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logisitc regression blended model had a two-sample t.test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
+
+```{r, fig.width = 8, fig.height=10}
 
@@
 ```{r profit_curve, fig.depth = 10}
-finance.plot = readRDS("report_data/plot_finance.rds" )
+finance.plot = readRDS("plot_finance.rds" )
 
@@
 
-```{r av_profit_curve}
+```{r av_profit_curve, fig.height=10}
 ggplot(finance.plot %>% filter(method == "average"), 
@@
 
+```{r}
 # Results - over 5 pages, up to 20 pages
-
-Guidelines:
-* Outlines what you found out in relation to your research questions or hypotheses, presented in figures and in written text.
-* Results contain the facts of your research.
+# 
+# Guidelines:
+# * Outlines what you found out in relation to your research questions or hypotheses, presented in figures and in written text.
+# * Results contain the facts of your research.
 Often you will include a brief comment on the significance of key results, with the expectation that more generalised comments about results will be made in the Discussion section.
-
-Which results to discuss
-* restate research questions and hypotheses. can we predict profitability? ofocurse you can make a prediction, 
-real question might be is it worthwhile for the business to engage in this activity?
+# 
+# Which results to discuss
+# * restate research questions and hypotheses. can we predict profitability? ofocurse you can make a prediction, 
+# real question might be is it worthwhile for the business to engage in this activity?
 we will look at this as a bottom line analysis.
 acknowledge that it may be worthwhile for other reasons such as confidence and transparency in historic projects.
-
-* bottom line analysis - run 1000 times to get confidence interval around the line.
-    * should the business adopt the model?
-
-* to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the case based reasoning approach
+# 
+# * bottom line analysis - run 1000 times to get confidence interval around the line.
+#     * should the business adopt the model?
+#     
+# * to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the case based reasoning approach
+```
 
@@
 
-The Discussion section:
-
-comments on your results;
-explains what your results mean;
-interprets your results in a wider context; indicates which results were expected or unexpected;
-provides explanations for unexpected results.
-The Discussion should also relate your specific results to previous research or theory.
+```{r}
+# The Discussion section:
+# 
+# comments on your results;
+# explains what your results mean;
+# interprets your results in a wider context; indicates which results were expected or unexpected;
+# provides explanations for unexpected results.
+# The Discussion should also relate your specific results to previous research or theory.
 You should point out what the limitations were of your study, and note any questions that remain unanswered.
 The Discussion CAN also include Conclusions/Future Research.
+```
 
@@
 
-simp.l <- readRDS("report_data/finance_average_model.rds")
+simp.l <- readRDS("finance_average_model.rds")
 
@@
 It was surprising that, given the similar AUC results, simple model averaging signficantly (?) outperformed the blended logistic regression model's profit curve.
+This means that model averaging predicted projects with probabilities closer to 1 (loss-making jobs) with more accuracy than blended logistic regression.
 This may be because ...
@@
 
-*** you are here ****
-
-Previous research indicated problems with user uptake once models were developed and a lack of trust in the model output [@Akintoye2000].
-
-* problems with past research and does this solve any of those?
-    * poor user uptake - future work?
-    * lack of trust - nearest neighbour must be presented with results - case based reasoning
-    * results from boosted tree shows partial dependence plots - insight.
-    * never been applied to a consultant within construction industry. has been applied lots to software
-    * still has the potential to help with the way the industry is structured - hinges on user uptake.
-
-You should point out what the limitations were of your study, and note any questions that remain unanswered. 
-* not sure how easy number of users is to predict - could do a study
-* could do a study on trust levels
+### Limitations and Future Work
+
+One of the unknown factors in translating this research into business practises is how well users can guess the project variables.
+The data used to develop the model was obtained from the company's records of how projects played out and there are no records of initially expected values of variables.
+For example, if a project lasted 2 years and it was initially expected it would only take 1.5 years, the only data the company retains is that the project lasted two years.
+This begs the question, how well could a user predict the variables before beginning a project?
+Uncertainty has been managed to a degree by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e.
+"1 day - 3 weeks", "3 weeks - 2.5 months", etc).
+However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables.
+It would be interesting to test how well business managers could guess these four variables at the start of a project.
+
+The process of guessing varibles at the beginning of a project and then waiting for the project to complete is a long checking process.
+However, it would be worthwhile integrating this data into the system's pre-existing data collecting processes.
+When a new project is entered into the database, it could be required that the project manager guesses the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent to be completed by professionals.
+This data itself could feed into a revised version of the prediction algorithm.
+
+Another limitation for these types of algorithms in general, as identified in the literature, is user uptake of the models and a lack of trust in the output [@Akintoye2000].
+It was for this reason the case study only tested models that provided insight into variable relationships (ensemble trees and logistic regression) as opposed to neural networks and SVM's that operate as 'black box' predictors.
+It is envisioned that a future user interface for this algorithm would display variable relationships such as this partial dependency plot produced by a boosted tree model that displays the probability of losing money vs.
+the team size:
+
+![](images/boosted tree/partial_dep_plots/no_users.png)
+
+Running a nearest-neighbour algorithm concurrently with the probability algorithm could provide a cross check of the algorithm's results and engage the user's past experiences in similar projects.
+This could work by presenting a graphical summary of the 4 or 5 most-similar past projects to the new project being assessed by the manager.
+Memories of how these similar projects played out would be triggered as well as communication with other manager's that led the similar projects.
+This would allow the user to connect specific past experience with the probability model output.
+Further development and testing is required to assess how much users trust the algorithm output and how trustworthiness may be improved.
+
+![](images/app/NN_wharf.png)
 
@@
 
-Future work
-* user testing on trust levels of prediction with nearest neighbour
-
-Conclusions
-could be a viable business move worth integrating
-
-Suggestions
-
-Could collect data for future projects: expected team size, expected time frame.
-
-* trialled ensemble tree methods on cost estimation (in this case project proftability)
-* attempted to predict project profitability using a single consulting company's CRM data in the construction industry
-* assessed tangible profit outcomes of machine learning models
+The aim of this project was to predict the profitability of projects for consulting businesses using their internal CRM data.
+Specifically, a statistical or machine learning model was created based on historical project data.
+The second aim was to test whether the model would have a positive impact on the overall profitability of a case study business.
+These research aims were fulfilled and the prediction model was refined to a binary classification problem of whether a project would be profitable or loss-making.
+This suited the limited data size best which was also highly complex.
+After testing several machine-learning and statistical models, a blended model of logistic regression, random forests, and boosted tree models predicted project profitability best.
+When applied to improving overall profitability of the business, the blended model which simply averaged the raw probability results from the three aforementioned models outperformed the individual models as well as the logistic regression blended  model.
+Final results showed the model could improve profitability of the business by between `r (-1.96*best.sd + best.mean) %>% round(1)`% and `r (1.96*best.sd + best.mean) %>% round(1)`% (the 95% confidence interval) by rejecting projects that have a 'probability' of being a loss-making job over `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).
+
+These results make a significant contribution to literature in this area as they provide a clear framework for how a business can realise tangible profit improvements using a set of algorithms and project data that is passively collected by businesses.
+Previous studies verified algorithms' prediction accuracy in cost estimation but stopped short of how the algorithms could be implemented in every day decision making and the affect on budgets.
+Additionally, this research tested ensemble tree methods and furthermore blended models on the problem of cost estimation.
+Past research had largely experimented with regression models and to some extent neural networks and case based reasoning.
+The advantage of ensemble tree methods is that they provide insight into the model's reasoning while having comparable predictive power to neural networks.
+Having said that, a limitation to the study was that there was not sufficient data to utilise the strength of ensemble tree methods and logistic regression performed simiarly.
+However, the final highest performing model was a blend of regression and two ensemble tree methods.
+
+There are several aspects of this work that should be investigated further including how well managers can estimate project input variables before a project is won.
+Particularly, estimations of the timespan, team size, total invoiced amount, and percentage of hours completed by professionals should be tested.
+These initial estimates could also be included in the each project's data entry for future analysis.
+Finally, research into user uptake of the model should be assessed.
+It is likely that in order to establish a manager's trust with the algorithm, additional model output will be required such as partial dependency plots from the boosted trees.
+Also, a 'nearest neighbour' algorithm could be run concurrently to show the manager outcomes of past similar projects, which should align with the model's probabiltiy prediction.
+Overall, this project has successfully demonstrated a blended statistical and machine learning model has the potential to aid manager's in rejecting loss-making projects to improve their business's profitability.

### 2016-06-14 finalwriteup/draft.rmd
started review of draft

@@
 
-****
+Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.
 
@@
 Predicting how long it will take a person to do something is a notoriously tricky task.
-People make mental calculations of these sort every day, guessing how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish.
-These mental calculations can be particularly inaccurate if the individual has never performed the task before.
-These errors are trivial in everyday activities, but they can have more meaning for businesses that sell their time as their chief product.
-For example, consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time the constultant spent on the problem.
+People make mental calculations of these sort every day- guessing how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish.
+As most people have experienced, these estimates can be particularly inaccurate if the individual is attempting the task for the first time.
+These errors are trivial in everyday activities, but they have more meaning for businesses that sell their time as their chief revenue.
+Consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time a constultant spent on the problem.
 The delivery of 'expert advice' is generally in the form of a document (which could be anything from a report to hundreds of drawings) and can range from an hours advice to thousands of hours work on a substantial project.
-Before a project commences, the client and consultant must agree on a fee for the expert advice, or a fee structure.
-The nature of how a fee is structured can be as creative as the engaged parties wish, however a couple of simple examples are as follows: the client agrees to pay for the consultants time by the hour until the task is completed, or the consultant provides a fixed fee to complete the project in full, regardless of their hours spent.
-The nature of the fee structure brings an element of financial risk to one party which must be mitigated.
-
-This thesis focuses on the risk a consulting company undertakes when setting a fixed fee for a project before their role has commenced.
-Clients commonly ask for competitive fixed quotes from several consultants before settling with one or negotiating further.
-The way a consulting manager decides on their fixed fee varies greatly from industry to industry and company to company.
-Typically, consulting manager has extensive experience in the type of project he is quoting and can use a combination of intuition, ratios of the entire project cost, and looking up a couple of past similar projects.
-Such a method often results in a wide range of profitabilities for projects depending on the ability of the manager to predict the amount of time the project will demand from the consulting team.
-This thesis examines a case study consulting company that has recorded internal historical project performance data since 2002.
-The study tests whether a statistical predictive model based on past project data can predict the profitability of a project based on characteristics available at the beginning of a project.
-This thesis asks further whether the prediction can impact a consulting manager's determination of a fixed fee.
+Before a project commences, the client and consultant must agree on a fee for the expert advice, or at least a fee structure.
+The nature of how a fee is formulated can be as creative as the engaged parties wish, however a couple of simple examples are as follows: the client agrees to pay for the consultants time by the hour until the task is complete, or the consultant provides a fixed fee to complete the project in full, regardless of their hours spent.
+Both contracts assign financial risk to one party which they must then mitigate to the best of their abilities.
+
+This thesis focuses on the risk taken upon by a consulting company when agreeing to a fixed fee.
+Clients commonly collect fixed quotes from several consultants before settling with one or negotiating further so fees must be competitive.
+The way a consulting manager calculates their fixed fee varies from industry to industry and even company to company.
+Typically a consulting manager has experience in the type of project he is quoting and can use a combination of intuition, rules of thumb (such as a fraction of the entire project cost), and comparison to past projects.
+Such methods can result in a range of profit margins which ultimately depend on how well the manager predicts the amount of time a project demands.
+This thesis examines the problem of cost prediction via a case study on a company's internal project performance data for the previous twelve years.
+The study tests whether a statistical model based on historical project data can predict the profitability of a project using characteristics available at the beginning of a project.
+Furthermore, this thesis investigates how the model's predictions could impact a consulting manager's determination of a fixed fee and the resulting impact on the company's bottom line.
 
@@
 
-In this chapter the research motivations are explained along with the goals of the project and the specific research hypothesis.
+In this chapter the research motivations are explained along with the goals of the project and the specific research hypotheses.
 Finally the contributions of this body of work are outlined along with the structure of the thesis.
@@
 
-Project decision makers have struggled with forecasting project costs accurately for decades, across many industries.
-A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% (Flyvbjerg, 2006).
-A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average (Flyvbjerg and Budzier, 2011).
-These statistics cast a concerning glimpse at the financial story playing out behind the scenes of these projects which can cause job losses, ruin private businesses, and churn through government balancesheets.
-Complex long term projects are necessary to produce  railways, houses, websites, and movies.
-This research focuses on the financial risk that is transferred to consulting companies who are enlisted to complete portions of these complex projects.
-The risk is transferred by requiring a fixed flat fee from the consultant, meaning they commit to complete their role for a nominated price.
-In a similar way to infrastructure and IT projects, consultants are susceptible to underestimating the amount of their time that will be spent in their role, and must wear the cost of extra employee hours spent.
+Project managers have struggled with forecasting project costs accurately for decades, across many industries.
+A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007].
+A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average [@Flyvbjerg2011].
+These statistics reveal a concerning glimpse at projects' hidden financial stories which can cause job losses, ruin private businesses, and churn through government balancesheets.
+Complex long term projects are necessary for modern society producing infrastructure, houses, websites, and movies.
+This research focuses on the companies who complete portions of these complex projects and the financial risk that is contractually transferred to them.
+Specifically, the risk of the project owner is transferred to their sub-consultant by mandating fixed flat fee structures to complete a given amount of work.
+The subconsultant commits to completing their role in a project for a nominated price, regardless of whether the work takes less or more time than intially anticipated.
+In a similar way to infrastructure and IT project budgets, consultants are susceptible to underestimating the amount of time a project will demand, and eventually wear the cost of extra employee hours spent.
 For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and succeeding.
 
-The ramifications of unprofitable projects in a private business are multifold.
+The ramifications of unprofitable projects are multifold.
 Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery.
 Employees also experience stress attempting to complete the project within a disappearing budget and may produce lower quality work in exchange for speed.
-This is unfortunate if in reality, the project fee was under estimated.
-Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would improve a business' standing.
-The financial risk taken on fixed flat fee projects may also discourage business' from taking different risks with better potential pay-offs, such as expanding the business into a new area.
-Logically, predicting fixed fee project costs more accurately could positively affect a business' growth opportunities and staff morale.
-
-**Why its Not Easily Solvable:**
-
-The challenge for consultants to reduce the risk of a project going over budget, in terms of hours spent, is not easily solved.
-Complex projects are always different from each other, and similar work may take significantly different amounts of time.
-For example, an ad filmed in a remote location for chocolate will cost significantly more than an ad filmed on an exisitng set, even though the product structure is almost identical.
-Therefore, a fixed product price, which you could apply to products such as appliances, is inappropriate for film projects.
-One option is charging for consultants' time and the hiring of materials by the hour; a zero risk fee structure.
-However in many industries this is not competitive enough.
-Other consultancies may be willing to offer a fixed flat fee for the same work, reducing financial risk for the client, and luring the client to them.
-Another fee strategy may be to track worker hours, and stop work once the consulting fee has been depleted.
+This experience is unfortunate if in reality, the project fee was under-estimated.
+Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would foster growth and futher profits.
+The financial risk taken on fixed flat fee projects may also discourage business' from taking other, more thought-out risks, such as expanding the business into a new area.
+Logically, predicting fixed fee project costs more accurately could positively affect business growth opportunities and staff morale.
+
+**Why It's Not Easily Solvable:**
+
+The challenge for consultants to reduce budget blowouts, in terms of hours spent, is not easily solved.
+Complex projects always differ from one another, and similar work may demand significantly different amounts of time.
+For example, an ad filmed in a remote location will cost significantly more than an ad filmed on an existing set, even though the effect of the final advert could be identical.
+Therefore, a constant fixed price to produce an advert is clearly inappropriate for film projects, but some kind of fixed price is often demanded by the client before a project is awarded.
+Charging clients by the hour is not competitive enough in many industries.
+Other consultancies are often willing to offer a fixed flat fee for the same work, luring the client with reduced financial risk on their part.
+
+Another strategy to mitigate risks could be to track employee hours, and stop work once the fixed consulting fee has been depleted.
 Then, negotiate further fees, or variations, with the client and persuade them to accept.
-This also satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective.
-However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or endure very marginable profits to preserve the relationship with their client.
-Consultant's may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry over creating financial friction.
-Clearly, there is no straightforward solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects.
-In some indusries, fixed flat fee structures are expected and business' must conform to survive.
-
-**Current Practise:**
-
-Limited research is available on the cost estimation techniques performed by industry for complex projects, however two surveys were performed in 2000: one on construction projects, the other on IT projects.
+This satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective.
+However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or endure marginable profits to preserve the relationship with their client.
+Consultants may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
+Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects.
+In many industries, the reality is fixed flat fee structures are expected and business' must conform to survive.
+
+**Current Practice:**
+
+Limited research is available on complex-project cost estimation techniques performed by industry.
+However, two surveys are available from 2000: one on construction projects, the other on IT projects.
 The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practises.
@@
 These can all be classified as experience based models [@Akintoye2000].
+
 A survey by Moores and Edwards in 1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project.
@@
 
+**Cost Estimation and Human Nature**
+
 @Lovallo2003 tackled the psychology behind the high failure rates of executives in predicting costs of projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
-Their theory stems from Kahneman's work on decision making that won him the nobel prize for economics in 2002.
-His research argues that humans' natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
-A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplay or ignore the risk of problems that are out of the team's control.
+Their theory stems from Kahneman's work on decision-making that won him the nobel prize for economics in 2002.
+His research argues that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
+A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplays or ignores the risk of problems that are out of the team's control.
 It is for this reason that it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
 Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].
+
 @Lovallo2003 call the practise of analysing a project based on project details and its unique complexities the 'inside view'.
-Research has shown that if people are forced to make predictions about their skills after being exposed to an 'outside view', their predictions are significantly more accurate.
-This outside view can be applied to complex projects, and involves ignoring the details of the current project and instead analysing the outcomes of several projects of a similar description.
+This constrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details.
+For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate.
+According to @Locallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project are ignored in favour of analysing outcomes of several similar projects.
 Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003].
 One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
-The correlation for the current project can then be estimated via a statistical model which is used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011].
-The traditional way to think about a project is to focus on the project details.
-Even though gathering data from similar projects could significantly improve a cost prediction, it is rarely thought of by managers [@Flyvbjerg2011].
-
-The implementation of reference class forecasting began in project management for the first time in 2004, was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, and Australia.
+The correlation for the current project can then be estimated via a statistical model which is then used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011].
+Managers rarely analyse data from similar projects even though it could significantly improve cost predictions [@Flyvbjerg2011].
+This tendency led to a formalised introduction of 'outside view' methods into some institutions.
+
+The implementation of reference class forecasting began in project management for the first time in 2004.
+It was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, and Australia.
 An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].
@@
 
-The theory of reference class forecasting was first clearly published in 2003, however the idea of predicting project costs from previous data has been the subject of research for many years before that and continues to this day.
+The theory of reference class forecasting was first definitively published in 2003, however the idea of predicting project costs from previous data has been the subject of research for many preceding years and continues to this day.
 Although research results have often been promising, industry uptake of the idea is not yet established.
 
-### Overview of Past Reference Class Forecasting Methods and relevant Industry Uptake
-
-In the following section, an overview of the historical attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.
+### Further Reference Class Forecasting Models and Industry Uptake
+
+In the following section, an overview of several historical attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.
 
@@
 The null hypothesis is rejected which contrasts the parallel ANOVA test.
-On the right hand side, of the graph above, the RMSE is improved by almost $0.20 but also worsens by over $0.10 per dollar.
+On the right hand side, of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar.
 A density plot of the RMSE for the models is shown below:
@@
 
-The two regression models predicting return per dollar resulted in high mean RMSE's ranging from $`r mean(results.rf.core$final.RMSE) %>% round(2)` to $`r mean(results.A2$final.RMSE) %>% round(2)`.
+The two regression models predicting return per dollar resulted in high mean RMSE's ranging from $`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
 Because of the poor results, further development of regression models was postponed in favour of investiation of binary classification models.
@@
 
+```{r power_profit_curve}
+
+# How to achieve statistical power of 0.8?
+finance.plot = readRDS("report_data/plot_finance.rds" )
+fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
+        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
+                            sd_prof_ratio = round(sd(.), 1)),
+                       profit.ratio)
+
+#for each method find max mean 
+best.fin.1 = fin.summary.1 %>% group_by(method) %>% 
+        filter(mean_profit_ratio == max(mean_profit_ratio)) %>% 
+        arrange(desc(mean_profit_ratio))
+
+#two sample power test between each method and its % data points at its threshold
+finance.plot$threshold <- as.factor(finance.plot$threshold)
+power.fin = finance.plot %>% filter(threshold == '0.55' & method == "average"|
+                                    threshold == '0.6' & method == "orig.boost"|
+                                    threshold == '0.75' & method == "orig.log"|
+                                    threshold == '0.55' & method == "orig.rf"|
+                                    threshold == '0.6' & method == "simp.log")
+
+# test each sample against "average" because it's the best
+# n.samples function is defined in 'draft.rmd'
+index = c("orig.boost","orig.log", "orig.rf","simp.log")
+ns = NULL
+for(i in seq_along(index)){
+        temp = n.samples(data = list(power.fin %>% filter(method == index[i]) %>% .$profit.ratio, power.fin %>% filter(method == 'average') %>% .$profit.ratio), type.calc = 'two.sample', alt = 'two.sided')
+        ns[i] = temp
+}
+ns
+
+# orig.boost needs 6, orig.log needs 4, orig.rf needs 6 and simp.log needs 64 : ) do 100!
+
+```
+
 Using this logic, if the threshold was 0, all projects were rejected and total profit would equal 0.
 If the threshold was 1, all projects were accepted and the total profits would be the same as historical figures.
-Probability results from the two blended models as well as three individual models were tested and because the models require a training and testing set, the process was repeated 100 times with 20 x 5 fold cross validation models to achieve a statistical power of 0.8.
+Probability results from the two blended models as well as three individual models were tested and because the models require a training and testing set, the process was repeated 100 times with 20 x 5 fold cross validation models to achieve a statistical power of 0.8 (calculated after an initial run of 10 models).
 The plot below illustrates the results of each model at each threshold with a local polynomial regression line of best fit (loess) and a 95% confidence interval around the mean.
@@
 ```{r profit_curve, fig.depth = 10}
-finance.plot = readRDS("plot_finance.rds" )
+
+finance.plot = readRDS("report_data/plot_finance.rds" )
 
@@
 In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.
-The improvements in profit sound promising, however it is important to consider possible limitations in the data which is discussed in the next section.
+
+Alternative strategies exist for how to deal with projects that are marked with a probability of loss greater than `r best.t`.
+For example, contracts for these projects could be changed to hourly rate contracts without exceptions.
+This means that each hour booked on the project will definitely be charged to the client for each stage of the project.
+Another example would be to increase fixed fees for projects above the threshold by a nominal percentage.
+However, this method seems less practical as the case study projects above `r best.t` would need to increase fees by 46% in order to generate a 15% profit.
+A proposed fee increase of that magnitude may not be as acceptable as an hourly rate contract depending on the project and client.
+Contracts could be adjusted in many other creative ways to reduce risk in high probability projects, and the threshold of `r best.t` was optimised based on the accept/reject strategy.
+Different courses of actions could be developed for specific clients or project sizes via analysis of threshold scenarios.
 
@@
 
+The improvements in profit sound promising, however it is important to consider possible limitations in the data which is discussed in the next section.
 One of the unknown factors in translating this research into business practises is how well users can guess the project variables.

### 2016-06-14 finalwriteup/draft.rmd
minor edit on draft

@@
 
-In the following section, an overview of several historical attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.
+In the following section, an overview of several attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.
 
@@
 
-A substantial amount of research on project cost prediction has been dedicated to the fields of software development and construction projects in particular.
+A substantial amount of research on project cost prediction has been dedicated to the fields of software development and construction projects, along with some work in the film industry.
 These analyses have generally used project information from 15 to 20 large projects.
-A single dataset may consist of projects from around the world, and each project a different client and project team.
+A single dataset may consist of projects completed in different countries with different clients and project teams.
 The mathematical methods used to predict project costs vary from statistical regressions to deep-learning neural networks [@Love2005][@Kim2004].
-Most studies report neural networks are more accurate than linear regression, and generally the studies report positive predictive powers of their developed models.
-
-A McKinsey study outlined a case in the film industry where a motion picture company used reference class forecasting of movie project success (cases were weighted by similarity) to decide which movies should be heavily promoted.
-This model improved forecasts on financial return by 135% relative to single project analogies [@Flyvbjerg2014].
-
-Statistical research projects which trial reference forecasting have shown promising results in IT, construction, and film projects.
-However, there is little evidence that the methods have successfully transitioned into industry.
+Most studies report their predictive models perform well and that generally neural networks are more accurate than linear regression.
+
+A McKinsey study outlined a case in the film industry where a motion picture company used reference class forecasting of movie revenue (cases were weighted by similarity) to decide which movies should be heavily promoted.
+This model improved forecasts on financial return by 135% relative to comparison of a movie to a single past case [@Flyvbjerg2014].
+
+Statistical research projects on reference forecasting have shown promising results in IT, construction, and film projects. However, persuading industry to adopt these methods is a separate challenge.
+
+** you are here **
 
@@
 
-Surveys on current industry practise for cost estimation are scarce, however, a survey of the construction industry in the UK in 2000 found that a manager's experience and intuition is still the dominant method for construction cost calculation [@Akintoye2000].
+Surveys on current industry practise in cost estimation are scarce; a survey of the construction industry in the UK in 2000 found that a manager's experience and intuition is still the dominant method for construction cost calculation [@Akintoye2000].
 A study on the IT industry by @Jorgensen2007 cited that the parametric models were not comprehensive enough to be used in industry and also found that expert judgment was not mobilised enough in the process of developing these tools.

### 2016-06-15 finalwriteup/draft.rmd
reviewed motivation

@@
 B2B - Business to Business 
+CBR - Case based reasoning. A method of estimating by using the results of similar cases.
 CRM - Customer Relationship Management.
@@
 Generally accessable to each employee in a company.
+IT - Information technology
+BIM - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
 
@@
 
-Statistical research projects on reference forecasting have shown promising results in IT, construction, and film projects. However, persuading industry to adopt these methods is a separate challenge.
-
-** you are here **
-
-#### Industry Uptake of Existing Models and Current Practises
-
-Surveys on current industry practise in cost estimation are scarce; a survey of the construction industry in the UK in 2000 found that a manager's experience and intuition is still the dominant method for construction cost calculation [@Akintoye2000].
-A study on the IT industry by @Jorgensen2007 cited that the parametric models were not comprehensive enough to be used in industry and also found that expert judgment was not mobilised enough in the process of developing these tools.
-
-Although reference class models have not gained traction in either the software or construction industry (where most of the research to date has focussed), it is worth outlining their current practises in cost estimation and the reasons behind these.
-
-In the construction industry, infrastructure and building projects follow a traditional and well established process, where contractual norms have developed over many decades [@Badenfelt2011].
-The industry is so large, broad and well established that it has inertia in the way that projects are estimated, and contractually bound.
+Statistical research projects on reference forecasting have shown promising results in IT, construction, and film projects.
+However, persuading industry to adopt these methods into everyday business is a separate challenge.
+
+#### Industry Uptake of Existing Models and Current Practices
+
+A survey of the construction industry in the UK in 2000 found that a manager's experience and intuition is still the dominant method for construction cost calculation [@Akintoye2000].
+Surveys on current industry practise in cost estimation are scarce, however another study on the IT industry by @Jorgensen2007 argued that previously developed parametric models were not comprehensive enough to be used in industry and that expert judgment was not mobilised enough in the process of developing these tools.
+
+Although the surveys reveal reference class models have not gained traction in either the software or construction industry, it is worth outlining their current practices in cost estimation, fee structures, and the reasons behind these.
+In the construction industry, infrastructure and building projects follow a traditional process, where contractual norms have developed over many decades [@Badenfelt2011].
+The industry is so large, broad and well established that it has inertia in the way projects are estimated, and contractually bound.
 The contractor (builder) first reviews the drawings and if the plans are not detailed enough, they offer a temporary budget price.
 Once they feel that the drawings are sufficiently detailed, they will offer a fixed price [@Badenfelt2011].
-However, as the job progresses, if any work is added that was not in the detailed drawings, they maintain the right to charge a variation [@Badenfelt2011].
-This method is clear and variations are easily justifiable to the client, however cost, labour and time estimates are still regularly under calculated, even from detailed drawings.
-And for large infrastructure projects, small percentage miscalculations can be exceedingly expensive.
-
-Contractor's practise of pricing from detailed drawings and charging for variations does not translate to other consultants in the construction industry, such as engineers and architects.
-Consultants must devlop the detailed plans for the contractors, and must therefore offer fixed price contracts based on drawings with little detail.
-It is much harder for a consultant to charge a variation for a portion of their time spent on something that was not included in the original simple plans.
+As the job progresses however, if additional work on top of the detailed drawings is required, the builder maintains the right to charge a variation [@Badenfelt2011].
+A dual source of income therefore exists, the initial proposed fee for construction along with variations due to additional materials (and time required to install the materials) that were not explicitly specified on drawings.
+Variations fees are generally calculated after completion the work, but the initial construction fee contains risk.
+Material costs are forseeable but labour and time estimates can still be under calculated, even from detailed drawings.
+For large infrastructure projects, small percentage miscalculations can be exceedingly expensive.
+
+The builders' procedure of calculating a fixed price from detailed drawings and charging for variations does not translate to other consultants in the construction industry involved with the design of structures, such as engineers and architects.
+These consultants develop the detailed plans for the builders, and must therefore offer fixed price contracts based on low-detail preliminary drawings that sum up the client's expectations.
+It is also much harder for a consultant to charge variations for portions of their time spent on items that were not included in the original simple plans.
+Time is less tangible than physical additional materials required, and often is shrugged off by consultants for the sake of client relationships.
 This is despite the real and common instance of budget blow outs due to overspent time [@Harris1999].
-One method to provide opportunity for a variation is stating a minimum percentage fee of the final building cost, which can overrule the original fixed price contract.
-If a consultant has spent time beyond their fixed price budget,this may account for some or in the best case, all, of their overspent time.
+One method of creating a variation opportunity is by stipulating a percentage fee of the final building cost that can overrule the original fixed price contract.
+If a consultant has spent time beyond their fixed price budget, this may account for some or in the best case, all, of their overspent time.
 
@@
 In contrast to construction, IT projects are a relatively new practise where alternative contractural arangements have been trialled.
-Through experimentation with contracts, the Agile method was developed.
-In summary, the method treats both cost and time as fixed quantities for a project, and if any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
+Through experimentation with contracts, the Agile movement was developed.
+In summary, the method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
 Furthermore, the project outcomes are continually adjusted ad revised based on frequent progress review points that assesses unexpected problems.
-This agile attitude has become ingrained in many IT practises and this flexible upfront arrangement with clients often forges better long term relationships [Cockburn2001].
-
-It is not completely clear why industry uptake of cost estimation tools using the outside view has been weak, and it may be the case that in modern industries such as in the IT industry, problems associated with cost estimation can be innovatively designed out of the contract.
-Possible reasons that parametric reference class forecasting models developed by researchers have not been incorporated into industry practise include:
+This agile attitude has become ingrained in many IT practices and the flexible upfront arrangement with clients often forges better long term relationships [Cockburn2001].
+
+It is not completely clear why industry uptake of cost estimation tools using the outside view has been weak, and it may be the case that in modern industries such as the IT industry, problems associated with cost estimation can be innovatively designed out of the contract.
+Possible reasons that parametric reference class forecasting models developed by researchers have not been incorporated into industry practice include:
 
@@
 * models were not developed in collaboration with industry
-* does not feel intuitive price from an 'outside view' of a project which disregards small project details
+* does not feel intuitive to price from an 'outside view' of a project which disregards small project details
 * insufficient time allocated to pricing that makes change burdensome
 
-In fact, a study by @Moores1992 concluded that a lack of framework to support the outside view model use and a failure in marketing were the two most common explanations for lack of uptake of cost estimating tools which use statistical techniques in predicting project costs.
-This is unfortunate as almost all research indicates an outside view analysis of a job substantially improves cost estimation accuracy for fixed price projects.
+In fact, a study by @Moores1992 concluded that a lack of framework to support use of the outside view model and a failure in marketing were the two most common explanations for lack of uptake of parametric cost estimating tools.
+This is unfortunate as almost all research in the area indicates an outside view analysis of a job substantially improves cost estimation accuracy for fixed price projects [@Moores1992].
 
@@
 
-This research will explore the potential for reference class forecasting to improve cost estimation in the context of an engineering consulting company in the construction industry.
+This research explored the potential for reference class forecasting to improve cost estimation in the context of an engineering consulting company in the construction industry.
 In this field, labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm.
-In recent years, the size of losses from unprofitable projects in the case study could equal up to 25-30% of the profits from profitable projects.
+In recent years, the size of losses from unprofitable projects in the case study equalled up to 25-30% of the profits from profitable projects.
 This demonstrates substantial room for improvement and the possiblity of multitude positive flow on effects.
@@
 Cost estimation is currently performed by first carefully reviewing preliminary drawings for a project where time costs will be calculated using a manager's personal experience with similar projects.
-This will often be cross-checked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
+This is often cross-checked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
 No formal mathematical method of comparison to similar projects is performed.
@@
 
-* lack of time to interrogate the CRM
-* clunky data availability offered by the CRM
-* lack of awareness of projects performed by other managers that may have been similar.
-
-Reference class forecasting in this study differs from previous research, in that statistical analysis will be performed on internal past company project data that has been progressively recorded in the company's Customer Relationship Management (CRM) software which was purchased.
-This presents an opportunity to evaluate reference class forecasting in a unique situation, where the 'reference' projects are internal and over two thousand cases are available.
-
-The CRM data currently stands as an untapped source of information within the case study organisation. It stores a rich variety of data including:
+* lack of time to interrogate past project data
+* clunky data availability offered by the current database software
+* lack of awareness of similar projects performed by other managers
+
+A Customer Relationship Management software package (CRM) was employed by this company to collect and store project data.
+It is available to all employees over the company intranet, and each employee completes daily timesheets alotting their hours to certain projects.
+Additionally, information is recorded against each project as well as each client, client contact, and each invoice.
+
+Reference class forecasting in this study differs from previous research in that statistical analysis was performed on internal past company project data that had been progressively recorded in the company's CRM.
+This presented an opportunity to evaluate reference class forecasting in a unique situation, where the 'reference' projects were all internal and a comparatively large number of cases (over 2,000 projects) were available.
+
+The CRM data stood as an untapped source of information within the case study organisation. It stores a rich variety of data including:
 
@@
 
-A statistical model predicting a measure of project financial success will be developed using the available project variables and trialling numerous machine learning algorithms and statistical methods.
-The company uses well established CRM software which readily provides simple output statistics such as project time-cost vs invoiced amount summaries.
-However, the analytical capabilities were limited to:
+A statistical or machine learning model predicting a measure of project financial success was developed using the available project information and by trialling numerous machine learning algorithms and statistical methods.
+The company uses well established CRM software which readily provides simple output statistics, however, analytical capabilities were limited to:
 
 * simple scatter plots and bar charts of the raw data
-* summaries such as overall hours spent vs invoiced amount for each project
-
-The company has a wealth of data but limited means to extract insight.
+* summaries such as overall hours spent vs. invoiced amount for each project
+
+The company had a wealth of data but limited means to extract insight.
 This study will exploit the CRM data by performing more sophisticated statistical analysis with the intention of building a predictive model for cost estimation.
-The benefits of this study in comparison to previous cases are that thousands of past data points are available, when previous studies used on average 15-20 cases and at most 170.
+The benefits of this study in comparison to previous cases are that thousands of past data points are available, when previous studies used on average 15-20 cases and at most 300 [@Finnie1997][@Pai2013][@Shin2015].
 This improves the potential for accurate predictions.
@@
 
-The case study provides a framework to test the value of a consulting company's CRM data in assisting project cost estimates.
-However, it is important to be aware of limitations when reviewing outcomes and framing the problem.
-The most obvious limitation is that we are disection a single company in a specific industry, which limits the ability to state conclusions could apply to a broad range of cases.
+The case study provided a framework to test the value of a consulting company's CRM data in assisting project cost estimates.
+However, it was important to be aware of limitations when reviewing outcomes and framing the problem.
+The most obvious limitation is that a single company is being disected in a specific industry, which limits the ability to state conclusions applicable to a broad range of cases.
 This can be overcome with future studies.
-
-The data set from this case study does not include information that would improve project cost predictions.
+Other constraining factors include various forms of missing data and human data entry error.
+
+The data set from this case study includes scarce information on the overall cost of entire construction projects.
 For example, managers commonly price jobs based on the budget for the entire project, of which the consultant is a minor part.
 The manager might typically moderate his fee by say 1% of the entire project cost.
-The value of the entire project has typically not been recorded although there is a space in the project software form for this value.
-It would be even more beneficial for managers to record the expected total project cost at the beginning of a project versus the conclusion of a completed construction project.
-Another piece of data that was not recorded was external factors influencing the fixed price quote including competition from similar firms that the client played against the case study company.
-
-In addition to this data that was not typically entered by workers, the dataset also excludes detailed information about the projects which are crucial to the final pricing.
-This includes details such as preliminary drawings describing the project (square meterage, number of storeys), and project summaries provided by the client to enable a quote.
-This kind of detailed information will never be recorded in a CRM and in many ways contradicts the purpose of the 'outside view' which intends to present similar projects without being over-influenced by the finer details of a project.
-Given the lack of detailed information, it is not reasonable for an algorithm to predict the specifc fee of a project.
-It is much more effective to attempt predict the profitability of a project (ie return per dollar) based on the historical data.
-This profitability prediction relates to the profitability of the project if the fee was generated using the established method the company used in its historical cases.
+The value of the entire project has typically not been recorded although there is an allotted space in the project database for this.
+It would be even more beneficial for managers to record the expected total project cost at the beginning of a project versus the concluding completed cost.
+Another piece of data not recorded was the external factors influencing the fixed price quote including competition from similar firms that clients may have played against the case study company.
+
+In addition to the missing data mentioned above that could be entered by workers in the future, the database also excludes the detailed information about projects that contribute to final pricing.
+Such details include the preliminary drawings describing the project (square meterage, number of storeys), and project summaries provided by the client to enable a quote.
+This level of information will never be recorded in a CRM and in many ways contradicts the purpose of the 'outside view', the intention of which is to present similar projects without being over-influenced by the finer details.
+Given this lack of detailed information, it was not reasonable for an algorithm to predict the numeric fee for a project.
+It appeared much more effective to predict the profitability of a project (ie return per dollar) based on available data.
+This profitability is a ratio between amount invoiced and business costs.
+The algorithm therefore would predict the ratio of these two values based on fees and costs generated using the established method of fee proposals and time expenditure in historical cases.
 
 Finally, CRM data is entered by employees manually as time passes, jobs are invoiced, and projects are finally completed.
-Because the data is input by humans, it is susceptible to input error.
-Many of errors were detectable during data cleaning, however it is possible that undetectable human errors exist in the data.
-
-Although this case study presents an opportunity to test the predictive power held within the organisations' existing CRM data, limitations exist which must be understood and built into the framework for predictions.
+Because the data is input by people, it is susceptible to input error.
+Many errors were detectable during data cleaning, however it is possible that a low number of undetectable human errors exist in the data and marginally influence predictions.
+
+Although this case study presented an opportunity to test the predictive power held within the organisations' existing CRM data, limitations exist which must be understood and built into the framework of the problem.
 These include missing data that could have been entered for each project, but historically has not been required, missing data that is very detailed and will never be available, and the potential for a small degree of human error to be embedded within the data.
@@
 In this case, the optimism often results in negative business consequences affecting employee morale and the capacity of a company to flourish and grow.
-Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project, by statistically comparing a new project to numerous similar projects and their characteristics.
+Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project, by statistically comparing a new project to a collection of similar projects and their characteristics.
 It has been shown that these models improve the cost estimation, however industry uptake has not been successful.
-It has proven to be a challenging task to persuade managers to use this so called reference class models in practise and remove themselves from additions based on intricate details of the completed project.
+It has proven to be a challenging task persuading managers to use reference class forecasting models in practise and remove themselves from summaries based on intricate details of the completed project.
 This may be due to a number of reasons including lack of understanding, time, and lack of collaboration with industry in development of these models.
 This research examines the potential for a company's internally generated CRM data to be used in creating a reference class forecasting model.
-This has the potential to influence the managers' decisions in the company more however its limitations must be understood in order to create a good model.
+Familiarity with the data has the potential to influence managers' decisions in the company more, however its limitations must be understood in order to create a valid model.
 
@@
 Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
+
 The project outcomes are intended to assist the business in  predicting project costs before project engagement and other business analyses such as client analysis.
@@
 **Hypothesis 2**
-The predictive model built from Hypothesis 1 can be proven to have a positive impact on the bottom line of the case study business.
-
-## Thesis Contributions - Gaps in literature
-
-* a gap in the study of predicting internal performance in businesses with business clients.
-The literature has assessed performance of movies, customer analytics but rarely employee performance or business clients.
-This is particularly relevant to consulting companies that tackle complicated discrete projects with set timelines.
-Statistical and machine learning models to predict the correct fee for projects according to project type, client, and other characteristics could reduce risk for a range of consulting businesses in setting initial fees.
-
-* A gap in machine learning predictive models which cater for variables with large amounts of missing data.
-the nature of consulting jobs is quite complex and historically, not all information about a project is possible to obtain.
-Variables with missing data are alternatively treated as optional bonus variables in a multi-step model as opposed to missing information which must be imputed or deletion of the entire case that contains missing information.
+The predictive model built from Hypothesis 1 can be shown to have a positive impact on the bottom line of the case study business.
+
+## Thesis Contributions
+
+This work contributes to the body of research in numerous ways.
+There is a gap in the literature of applying statistical and machine learning methods to internally generated business data to facilitate internal business decisions.
+Studies have modelled the commercial performance of movies, customer churn, the share market, and bankruptcies but few have modelled more day-to-day data that a business generates which could actually assist day-to-day decisions.
+Improving business decisions would reduce risk and improve profits in scenarios that could be applicable in thousands of businesses around the world.
+
+In the case of this study, the problem of cost estimation is tackled.
+This issue has been worked on extensively in past research, however the problem has rarely been addressed using a single company's internal data.
+In the literature, dozens of projects from other companys and even countries are collected and analysed.
+This thesis uses the internal CRM data from a company which contains over 2,000 past job cases - far more than the usual cost estimation study.
+Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry that estimates and must stand by their fixed fees before a project is won.
+This differs from other studies estimating actual construction costs of projects and is likened more to software development effort estimation although the nature of the work differs considerably.
+
+In addition to addressing cost estimation with a different kind of dataset and different industry sector, ensemble tree methods will be applied which have been minimally tested on this problem.
+To date, linear regression, neural networks, case based reasoning, and SVM's have been used to predict cost/effort even though ensemble tree methods can perform as well as neural networks in some cases and generally outperform linear regression.
+Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots provide insight into the model's calculations.
+This is in contrast to neural networks and support vector machines (SVM's) which are 'black-box' predictors.
+Providing insight to a user is a major benefit in this case as past research has demonstrated issues with user trust and uptake of models, even if the models were successful.
+It is anticipated that using internal data only also improves user trust and creates tangible meaning for users.
+Blending multiple machine learning and statistical models was trialled which has not yet been applied to the cost estimation problem.
+
+Finally, this study takes the model one step further than previous studies and analyses concrete profit improvements for the company if the model output were to be used in certain decision-making scenarios.
+A few possibilities are presented as ways to take model results into account with business decisions and where possible, final profit calcuations are presented.
+It is anticipated that this kind of thought experiment would further push industry to seriously consider the worth in adopting statistical cost estimation models.
+
+## Thesis Structure
+
+The remainder of this thesis has the following structure: the next chapter is a thorough review of the literature relating to this thesis to date.
+This covers work solving the cost-estimation problem, studies on statistical and machine-learning models in business applications, and a review of effective statistical and machine learning methods.
+
+The following chapter, Chapter 4, describes the method followed throughout the research process.
+It includes first how the data was obtained, the lengthy cleaning process, followed by variable importance analysis, variable selection, trials of selected predictive algorithms and problem constructs.
+The cost estimation problem was attempted first as a regression problem predicting return per dollar, followed by a simpler binary classification problem predicitng profit or loss.
+Once the best methods were selected, they were blended in numerous ways, using both simple averaging techniques and sophisticated machine-learning algorithms.
+These were compared against individual models and the best constructs were selected.
+Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision making scenarios and further applications were considered.
+
+Chapter 5 presents the results of the series of analyses described above and directly relates them to the overall aim and hypotheses of the project.
+This is in the format of tables and plots where results are briefly discussed.
+The examination of results is taken deeper as well as in a broader context in Chapter 6: Discussion.
+Here the nature of the results are debated in terms of how surprising they were, their impact on the research hypothesis and limitations as well as suggested future work.
+The final Chapter 7: Conclusion summarises the  findings of the project, how the research answered the hypotheses and how the work could be viably applied in industry.
 
@@
 
-This chapter provides an overview of the literature available about current business to business (B2B) cost estimation methods, use cases of CRM data, as well as machine learning algorithms and statistical business prediction models.
-The most prominent statistical and machine-learning models developed for business applications shall be reviewed along with models applied to the most closely related business cases to this project.
+This chapter provides an overview of the literature available in current cost estimation methods, use cases of CRM data, as well as machine learning algorithms and statistical business prediction models.
+The most prominent statistical and machine-learning models developed for business applications shall be reviewed along with models applied to business cases that can be related to this project.
 Gaps in past research will also be highlighted.
@@
 
-This project will tackle the problem of estimating fixed fees for complex consulting projects.
-It is therefore imperitive to understand what research has been performed to date and how this project contributes to the body of work.
+The application for this project is the task of estimating fixed fees for complex consulting projects.
+It is therefore prudent to understand what research has been performed to date and how this project contributes to the body of work.
 The task of estimating fixed fees can be likened to the terms 'cost estimation' or 'effort estimation' in the literature.
 The bulk of research to date has been performed with project data in either the construction industry or software development.
-The research into cost estimation in each of these industries is reviewed below, detailing different methods and outcomes.
+Research into cost estimation in each of these industries is reviewed below, detailing different methods and outcomes.
 Finally a gap in the literature will be highlighted which this research aims to fill.
@@
 
-Research into cost estimation in the construction industry chiefly focuses on predicting the final building cost of construction.
-Various studies have surveyed popular methods for cost estimation while others have  investigated new methods such as building information modelling (BIM) analyisis, case based reasoning, and predictive algorithms.
-The cost estimation methods of detailed analysis (inside view), case based reasoning, and predictive algorithms shall be reviewed.
+Research on cost estimation in the construction industry has primarily focused on predicting the final building cost of construction.
+Various studies have surveyed methods for cost estimation while others have investigated new methods such as building information modelling (BIM) analyisis, case based reasoning, and predictive algorithms.
+The cost estimation methods of detailed analysis (inside view), case based reasoning, and predictive algorithms are reviewed in this section.
 
@@
 
-Detailed analysis refers to the process of an engineer or builder in the construction team carefully reviewing construction drawings to sum the cost of materials, labour, machinery hire, overheads, profit etc.
+Detailed analysis refers to the process of an engineer or builder in the construction team carefully reviewing construction drawings to sum the cost of materials, labour, machinery hire, overheads, and profit etc.
 [@Akintoye2000].
 @Akintoye2000's study based in the UK found that this is the most popular method for pricing amongst contractors.
-The second most popular method was to perform an additional manual comparison with similar projects and make judgments based on personal experience.
+The second most popular method was to carry out an additional comparison with similar projects and make judgments based on personal experience.
 Note that this does not involve a mathematical model.
 
-@Elfaki2014 discusses how much detailed cost estimates for cosntruction projects can vary from estimator (engineer) to estimator, which contributes to the lack of accuracy in predciting final cost of a project.
+@Elfaki2014 discusses how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator, which contributes to the lack of accuracy in predciting final cost of a project.
 He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and, in turn, their estimate is prone to subjectivity.
-@Shane2009 theorises that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues that it is almost impossible to achieve an accurate cost prediction manually.
-
-Groundbreaking methods into calculating construction costs by @Ma2014 mobilise the data from building information modelling, although this method is still technically a detailed analysis of the costs.
+@Shane2009 theorises that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues, that it is almost impossible to achieve an accurate cost prediction manually.
+
+Groundbreaking methods into calculating construction costs by @Ma2014 mobilise the data from BIM models, although this method is still technically a detailed analysis of the costs.
 The idea is to quantify the cost of building a structure directly from a three dimensional model created by the design consultants (engineers and architects).
 @Ma2014's first trial was to automate the cost calculation of a reinforced concrete structure.
-His chief development was to program his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
-From the construction technique and material, a cost can be determined for each element.
-This was successful and Ma aim's to further his work so that more construction information can be intelligently obtained from BIM models.
-A system like Ma's would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could be prone to innaccurate estimates in a similar way to a detailed analysis by a person.
+His chief contribution was to program his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
+From the construction technique and material, a cost can be derived for each element.
+This was successful and @Ma2014 aims to further his work so that more construction information can be intelligently obtained from BIM models.
+A system like @Ma2014's would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could be prone to innaccurate estimates similarly to detailed analysis by a person.
 The detailed view does not 'step back' and take into account set backs in construction that may have affected similar projects in the past.
-There is potential for this to be built into the same system but this is still in early stages of development.
+There is potential for this innovation to be combined with an outside-view system but @Ma2014's is still in early stages of development.
 
@@
 
-As the problem of cost estimation in construction projects is well recognised, dozens of studies have built prediction models of these projects using machine learning and statistical models.
+As the problem of cost estimation in construction projects is well recognised, dozens of studies have built predictive models of these projects using machine learning and statistical models.
 The aim is to test whether a mathematical comparison of many similar projects, using meaningful descriptive variables, can predict the cost of construction.
@@
 @Elfaki2014 completed a study on cost estimation techniques and research from 2004 to 2014.
-It was found that artificial neural networks and support vector machines were the most common machine learning techniques used.
+They found that artificial neural networks and SVM's were the most common machine learning techniques used.
 He highlighted that these two machine learning techniques deal well with uncertainty, however lack technical justification to the decision maker (known as a black box).
@@
 @Shin2015 pioneered the application of boosted trees to cost estimation in construction projects.
-This is surprising given the rapid uptake of boosted trees over the past decade.
-In the study, data from 234 school building construction projects in Korea was used.
-Boosted trees were compared to neural networks, which are one of the most popular algorithms to apply to the problem.
-Boosted trees were found to predict costs more slightly accurately than neural networks, but not significantly (p value < 0.5).
+This is surprisingly late given the rapid uptake of boosted trees over the past decade.
+In the study, data from 234 school building construction projects in Korea were used.
+Boosted trees were compared to neural networks, which are one of the most popular algorithms to apply to this problem.
+Boosted trees were found to predict costs slightly more accurately than neural networks, but not statstically significantly (p value < 0.5).
 Even though the two algorithms performed similarly, the output from boosted tree models provide insight into the structure of the model such as variable importance and variable partial dependence plots.
@@
 This may be because the technique is straightforward, easy to use and widely available in statistical packages [@Chan2005].
-Out of those that have compared regression to neural networks, the studies have shown that neural networks outperform regression, however other studies show they are approximately equal [@Kim2004][@Attalla2003].
+Out of those that have compared regression to neural networks, the studies have shown that neural networks outperform regression, however other studies established they are approximately equal [@Kim2004][@Attalla2003].
 @Dissanayaka1999 found that neural networks outperformed multiple linear regression models, however regression models could be performed first to determine variable importance and condense the number of variables.
-It can be argued that boosted trees could perform as well as neural networks while providing additional insights (even more so than regression models) such as variable importance and partial dependence of variables against cost.
-
-Despite neural networks typically outperforming linear regression models, it is intersting to note some of the outcomes of regression studies in cost estimation.
-Different studies tend to produce a wide range of variables that contribute to different degrees in the models.
-This is likey due to the nature of construction projects which are multidisciplinary and involve a range of different parties such as the clients, consultants, contractors and suppliers, each of which can have a range of characteristics and degrees of input [@Chan2005].
+It can be argued that boosted trees could perform as well as neural networks while providing additional insights (even more so than regression models) such as variable importance and partial dependence of variables against the independent variable, cost.
+
+Despite neural networks typically outperforming linear regression models, it is interesting to note some of the outcomes of regression studies in cost estimation.
+Different studies tended to produce a wide range of variables that contribute at fluctuating levels in the models.
+This is likely because construction projects are multidisciplinary and involve a range of different parties such as the clients, consultants, contractors and suppliers, each of which contribute varying levels of input [@Chan2005].
 @Chan2005's study on 87 building projects in Singapore included special, complex projects.
-Therefore, variables such as contractor's specialised skills, who the client was (public vs private), the client level of experience, and the contractor's financial management highly influenced final cost of the project.
+Therefore, variables such as contractor's specialised skills, who the client was (public vs.
+private), the client level of experience, and the contractor's financial management highly influenced final cost of the project.
 Other studies found variables such as project complexity, duration, team experience, information availablility, site requirements, and labour climate to contribute most [@Akintoye2000][@Trost2003][@Pinto1988].
-As these studies obtained information from questionaires, the available variables also largely depended on the questions in the questionaires, and therefore each study produces a range of different important variables.
-Despite the differences in qualitative outcomes, the studies all generally report their mean absolute error or a similar metric from the tested models.
-This is encouraging, however there is a gap in the literature in how to translate the improved errors from the model into an argument for industry to uptake the model as an improvement on their bottom line.
-
-Surveys into construction industry uptake of these algorithmic models are sparse, however a survey by @Akintoye2000 showed that parametric estimating methods have not been adopted by contractors, despite the research into this area over 20 years prior.
+As these studies obtained information from questionaires, the available variables also largely depended on the questions in the questionaires, and therefore each study produced a range of different important variables.
+Despite the differences in qualitative outcomes, the studies all generally reported their mean absolute error or a similar metric from the tested models.
+Similar result structures are useful, however there is a gap in the literature in how to translate the improved errors from the model into a well justified argument for industry to reduce losses by integrating a parametric model into their business.
+
+Surveys into construction industry uptake of these algorithmic models are sparse, however a survey by @Akintoye2000 showed that parametric estimating methods have not been adopted by contractors, despite the research into this area over the 20 prior years.
 The majority of contractors use experience based methods.
 @Akintoye2000 hypothesises that this could be due to a lack of familiarity and knowledge about the techniques, doubts whether these techniques are applicable to the construction industry, and the availability of sound data to ensure confidence.
-This presents a clear gap in the literature on how to analyse the models in terms of the goals and limitations that a contractor may experience.
-
-The literature on cost estimation for building projects do not break down the cost estimation task into smaller components, even though a construction project is completed by dozens of consultants and sub-contractors.
-Many of the consultants, such as Engineers and Architects, produce their own cost estimates for their effort that contributes to the project.
-Their contracts are often fixed price, with minimal room for movement, which means they wear the risk of over-spending to meet fixed cost contracts [@Harris1999].
-Furthermore, the existing studies collect data from a wide range of projects that have certain characteristics in common, such as country, or building type and information is obtained through surveys.
+This presents a clear gap in the literature on how to assess the models in terms of the end user's goals and limitations.
+
+The literature on cost estimation for building projects does not break down the cost estimation task into smaller components, even though a construction project is completed by dozens of consultants and sub-contractors.
+Many of the consultants, such as engineers and architects, produce their own cost estimates for their effort that contributes to the project.
+Their contracts are often fixed price, with minimal room for movement, which means they wear the risk of over-spending their time to meet fixed cost contracts [@Harris1999].
+Furthermore, the existing studies collect data from a wide range of projects that have certain characteristics in common, such as country or building type, and information is obtained through surveys.
 This leaves a gap in research exploring a single contractor's history of projects, which would be available in their company database.
 
-**case based reasoning**
-
-Case based reasoning (CBR) is a method of estimating by using the results of similar cases.
+**Case Based Reasoning**
+
+Case based reasoning (CBR) is a method of estimation that uses  the results of similar cases.
 The idea is that in the case of a new project, similar projects are chosen, before generalisations about the data are made.

### 2016-06-17 finalwriteup/draft.rmd
reviewed lit review started method review

@@
 
+ANOVA - Anaylis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
 B2B - Business to Business 
@@
 Generally accessable to each employee in a company.
+DAG - directed acyclic graph.
+A graph consisting of vertices connected by edges, where each edge directs a vertex to another vertex.
+The arrangement of the edges is such that a vertex cannot follow a path that loops back to itself.
 IT - Information technology
@@
 In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to neural networks, which must be slowly re-trained and re-tuned.
-New cases also did not need to have every variable complete which can occur often in real, messy data.
+New cases also did not need to have every variable completed which can occur often in real, messy data.
 Neural networks on the other hand require complete data.
 Another significant advantage of CBR is the ability for users to review chosen similar cases and make sense of the predicted cost, as opposed to neural networks which are a black box [@Kim2004][@Elfaki2014][@Kumar2007].
-A disadvantage of CBR is the accuracy can be highly dependent on the number of selected cases [@Elfaki2014].
-Overall, CBR may provide prediction accuracies between regression and neural networks but its results can be justified by the user - a valuable asset.
+One disadvantage is that accuracy can be highly dependent on the number of selected cases [@Elfaki2014].
+Overall, CBR may provide prediction accuracies between regression and neural networks but the results can be justified by the user - a valuable asset.
 
@@
 CBR that is performed manually through personal experience, without the assistance of algorithms, was the second most popular method for construction estimation in @Akintoye2000's survey in the UK.
-This indicates it is quite an intuitive method for decision maker's to cross check their detailed cost estimate, although an algorithmic version of CBR has not been taken up by industry.
+This indicates it is quite an intuitive method that allows decision makers to cross check their detailed cost estimates, although an algorithmic version of CBR has not found success in industry.
 
@@
 
-In summary, research into cost estimation in the construction industry has been undertaken using a variety of methods ranging from detailed manual analysis, algorithmic models, and case based reasoning.
-Neural networks generally outperformed all other systematic methods, such as regression and case based reasoning, but had a significant disadvantage of being a black box.
+In summary, research into cost estimation in the construction industry has trialled a variety of methods ranging from detailed manual analysis, algorithmic models, and case based reasoning.
+Neural networks generally outperformed all other systematic methods, such as regression and case based reasoning, but had the significant disadvantage of being a black box.
 Alternative methods such as boosted trees have the potential to perform as well as neural networks and provide insight into the structure of the algorithm.
@@
 Also, as stated previously, there exists a gap in the application of cost estimating in the construction industry to smaller components of the project, such as consulting and contracting companies who each face a similar problem in estimating their fixed fee.
-These companies could mobilise information in their internal project databases, as opposed to available literature which surveys dozens of companies involved in selected construction projects.
+These companies could mobilise information in their internal project databases, as opposed to the current body of literature that used data from dozens of companies.
 
@@
 
-Like the construction industry, the software industry faces the challenging task of estimating the cost of a project before full engagement with their client.
+Like the construction industry, the software industry faces the challenging task of estimating the cost of a project before formal engagement by their client.
 However, unlike construction projects, the major cost associated with the project is *effort* as opposed to the cost of building materials.
-A parallel can be drawn between software project effort estimation research and consultants effort in the construction industry, which was a gap in the research of the previous section.
+A parallel can be drawn between software development effort estimation and consultants' effort estimation in the construction industry - which was highlighted as a gap in the research of the previous section.
 
 Effort estimation is indeed a problem, and was demonstrated by a review of 6 surveys of IT projects between 1984 and 1994 by @Molokken2003.
-It revealed that 60-80% of IT projects encounter effort and/or schedule overruns, where the average overrun was by 30 to 40%.
+It revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
 A survey of software managers asked whether they saw effort estimation as a problem, and 91% responded 'yes', while only 9% answered 'no' [@Moores1992].
 The most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%), which falls squarely within the theory of optimism bias, or the 'inside view' [@Flyvbjerg2011].
-Given this widespread problem in the industry, plenty of research has been dedicated to the problem of effort prediction, with methods similar to the construction industry literature: detailsed analysis, algorithmic modelling, and case based reasoning.
+Given this widespread problem in the industry, plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailsed analysis, algorithmic modelling, and case based reasoning.
 
@@
 The scarcity of real project evidence that algorithmic estimation tools improve estimations and an unintuitive mental jump could be reasons contributing to detailed expert analysis remaining the most widespread technique for effort estimation.
+This trend mirrors what was found in the construction industry.
 
@@
 Algorithmic methods to predict effort for IT project costs has been a popular subject of research.
-Regression models have been the most substantially tested followed by neural networks [@Finnie1997].
 This section will assess the details and performance of the algorithmic models, important variables, applications and finally gaps in the literature.
@@
 Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the models accuracy: size [@Shepperd1996][@Finnie1997][@Pai2013].
-Size can refer to the expected number of lines of code or function point.
+Size can refer to the expected number of lines of code in the software package or function point.
 Function point refers to the amount of business functionality expected from an information system product.
-This is easier to guess correctly at the beginning of a project, and therefore leads to more accurate predictions [@Finnie1997].
+This is easier to guess correctly at the beginning of a project than lines of code, and therefore leads to more accurate predictions [@Finnie1997].
 One study used a single variable approach (size) with a linear coefficient and an exponential coefficient as follows:
 
-$effort = \alpha*size^{\beta}$
-$\alpha = productivity coefficient$
-$\beta = economies of scale coefficient$
-$size = estimated lines of code$
-
-This model was compared to case based reasoning, and case based reasoning outperformed it [@Shepperd1996].
+$$
+\begin{aligned}
+ effort &= \alpha*size^{\beta} \\
+ \alpha &= productivity \ coefficient \\
+ \beta &= economies \ of\ scale\ coefficient \\
+ size &= estimated \ lines \ of \ code
+\end{aligned}
+$$
+
+This model was compared to CBR, which outperformed the model [@Shepperd1996].
 
@@
 
-Algorithmic estimation tools have not been successfully adopted by industry.
+Algorithmic estimation tools have not been successfully adopted by the IT industry.
 This may be because no model has proved to be outstandingly successful at consistently predicting required effort [@Finnie1997].
-Also, neural networks (the most accurate algorithm) lack the ability to capably explain its results and instill trust in decision makers [@Finnie1997].
-
-The literature heavily focuses on effort estimation for the software industry.
-This leaves a gap in studies for effort estimation for consulting companies in other industries such as the construction industry.
-Construction projects do not have a simple 'project size' or function point variable and would need to utilise other variables available.
+Also, neural networks (the most accurate algorithm) lack the ability to capably explain its results and instill trust in decision makers which is another parallel to research from the construction industry [@Finnie1997].
+
+As alluded to previosuly, studies on effort estimation as opposed to project cost estimation are concentrated in the software industry.
+This leaves a gap in studies for effort estimation in other industries such as the construction industry.
+Infrastructure projects do not have a function point variable and would need to utilise other information such as expected total project cost (which is notoriously hard to predict based on the previous section), client characteristics, and other project characteristics.
 
@@
 
-In the IT industry, applications of case based reasoning to effort estimation have been researched in a similar manner to the construction industry.
-The procedure is described as a algorithmicallyfinding similar projects to a new case using measurements of euclidean distance in n-dimensional space, where each dimension corresponds to a variable [@Shepperd1996].
-The similar cases are then used to predict effort for the new case, and in @Shepperd1996's study, linear regression was used for this final step.
-
-In contrast to the construction industry, the literature in software effort estimation found CBR to perform approximately equally to algorithmic methods such as neural networks [@Shepperd1996][@Finnie1997].
+In the IT industry, CBR applications have been researched using similar methodologies to the construction industry.
+The procedure is described as algorithmically finding similar projects using measurements of euclidean distance in n-dimensional space, where each dimension corresponds to a variable [@Shepperd1996].
+Then, the similar cases are used to predict effort for the new case.
+In @Shepperd1996's study, linear regression was employed.
+
+The literature found CBR to perform approximately equally to algorithmic methods such as neural networks which contrasts findings in the construction industry [@Shepperd1996][@Finnie1997].
 It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects and provides insight into the prediction that neural networks do not [@Finnie1997].
-Research into CBR in the IT industry found similar results to the construction industry except that CBR performed as well as algorithmic methods.
 
@@
 
-Research into effort estimation in the software industry has followed a similar path to the cost prediction of construction projects.
-Expert judgment via detailed analysis is still the predominant method for effort estimation.
+Research into effort estimation in the software industry has followed a similar path to cost prediction studies on construction projects.
+Expert judgment via detailed analysis is still the predominant method for effort/cost estimation.
 Case based reasoning and algorithmic methods such as neural networks have found some success in effort estimation (at similar levels) but have not been successfully adopted by industry.
-It was noted that case based reasoning resonated more with decision makers because its effort prediction could be justified by the model.
-This visibility of the model prediction structure is very important for future work if the results are to be translated to industry.
+It was noted that case based reasoning resonated more with decision makers because its predcitions could be better justified by the model.
+The visibility of model structure is very important for future work if the results are to be translated to industry.
 Effort estimation remains a problem, however agile methods of project management and effort management are changing the contractual approach of consulting in the software industry [@Badenfelt2011].
-This innovative method keeps delivery times short (and therefore predictable effort) and defines projects as something with a fixed cost and time.
-This means if a change occurs, other less important elements must be discarded maintain the fixed amount.
-Software development is a relatively new field in comparison to the construction industry and is able to adopt this new model relatively quickly across the industry.
-However, effort estimation for consulting companies in the construction industry are faced with a long history of tradition that is much harder to tackle.
-This leaves a gap in research in algorithimic models that predict effort for consulting companies in the construction industry.
-From previous literature, the models should allow decision makers to interrogate the structure of the prediction and be based on internal company data.
-
-## Statistical and machine learning methods
+This innovative method keeps delivery times short (and therefore more predictable effort) and defines projects as something with a fixed cost and time.
+Therefore, if a change occurs, other less important elements must be discarded to maintain the fixed amount.
+Software development is a relatively new field in comparison to the construction industry and was able to quickly adopt this new model.
+In contrast, effort estimation in the construction industry faces a long tradition of fee structures that is difficult to reinvent.
+This leaves a gap in research of algorithimic models that predict effort (and not construction cost) for consulting companies in the construction industry.
+From findings in previous literature, the models should reveal their predictive structure to decision makers and be based on internal company data.
+Review of the literature as a whole reveals a lack in analysis of overall profit improvement a model could provide to a business.
+This could pesuade industry adoption and provide a framework around how models should affect decisions.
+
+## Statistical and Machine Learning Methods
 
@@
 
-The aim of this project is to use statistical techniques to model the profitability of projects.
-In order to optimise the model, several statistical and machine learning techniques will be tested, and compared and it is therefore prudent to review the range of statistical and machine learning techniques that have been successfully applied in the literature.
-These vary from simple methods such as Linear Regression to complex, deep learning such as Neural Networks.
-The previous section highlighted the use of linear regression, neural networks, SVM and in one case boosted trees for prediction, however other business problems have utilised a wider range of methods.
+AS stated, the aim of this project is to use statistical and machine learning techniques to model the profitability of projects.
+Therefore, in order to optimise the model, several statistical and machine learning techniques will be tested and it is prudent to review the range of algorithms that have been successfully applied in the literature.
+These vary from simple methods such as Linear Regression to complex, deep learning Neural Networks.
+The previous section highlighted the use of linear regression, neural networks, SVMs and in one case boosted trees for prediction, however research on other business problems utilised a wider range of methods.
 These include Naiive Bayes, Random Forests, and machine learned Bayesian Networks.
-The following section will present each method's advantages and disadvatanges and an example of a past successful application.
+The following section presents each method's advantages and disadvatanges and an example of a past successful application.
 
@@
 The coefficients are chosen to minimise the residual sum of squares (least squared) [@Hastie2009].
-The reason this method is popular is that it is simple, provide stable predictions, and can be adapted for categorical variables [@Seng2010].
-The integration of categorical variables is called Anova, and the algorithm compares group mean variances within the categorical variable.
+The reason this method is popular is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
+The integration of categorical variables is called Anaylis of Variance (ANOVA) linear regression, and the algorithm compares group mean variances within the categorical variables.
 
@@
 For example assumptions must be made about the structure of the data.
-Predictor varialbes are assumed to be linearly related to the response variable and the varialbes are normally distributed.
+Predictor variables are assumed to be linearly related to the response variable and the variables to be normally distributed.
 Many real world phenomena do not correspond to these assumptions which can make it difficult, sometimes impossible, to produce useful results [@Hastie2009][@Putler2012].
@@
 On the other hand, correlated variables do not impact the predictive accuracy of the model.
-A final drawback is that the variable coefficients cannot be used to compare variable importance unless the varialbes have all been standardised to one another [@Putler2012].
+A final drawback is that the variable coefficients cannot be used to compare variable importance unless the variables have all been standardised to one another [@Putler2012].
 This limits the amount of insight that can be drawn from a linear regression model.
 
-Despite the limitations of linear regression models, they have widespread business applications due to their simplicity and ease of application.
+Despite limitations of linear regression models, they have widespread business applications due to their simplicity and ease of application.
 Regression is used to predict risk in the finance and insurance industries, predict who to target in marketing exercises, and predict consumption spending in the field of economics [@Harrell2013].
-Due to its simplicity and widespread application, linear regression is a good starting point for an analysis and benchmark for performance against complex models.
+Its simplicity and widespread application make linear regression a good starting point for analysis and can serve as a performance benchmark against complex models.
 
@@
 The method behind linear regression can be adapted to predict a binary response variable (1 or 0).
-A predictive linear equation is computed so that the predictor variables are combined with coefficients to give an output between 0 and 1.
-This is then transformed into the log odds to mediate extreme variable input values.
-The result is a probability an input case will result in a 0 or 1 (i.e.
+A linear equation is fit to the log of the odds of the binary response variable against each of the explanatory variables.
+The response variable (log odds) in the linear equation can then be translated back to odds by taking the inverse of log, the exponential.
+Finally the odds is converted to probability by the following equation:
+
+$probability = \frac{odds}{1 + odds}$
+[@Lowry2016]
+
+This translates the original log odds response variable to a probability S curve bounded by 0 and 1. An example of this translation is graphically shown below:
+
+![](images/lit_review/logreg.JPG)
+[@Lowry2016]
+
+The result of this linear fit and logistic translation is a model that predicts the probability (between 0 and 1) that an input case will result in a 0 or 1 (i.e.
 success or failure) [@Moore1989].
-Similar pros and cons exist for logistic regression as linear regression, and it too can be used as a benchmark to compare other binary predictive models due to its speed and simplicity.
-Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survive/death [@Moore1989].
+Similar pros and cons exist for logistic regression as linear regression, and it can also be used as a benchmark to compare other binary predictive models due to its speed and simplicity.
+Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survival/death [@Moore1989].
 
@@
 
-Naive Bayes prediction method makes conditional independence assumptions about the predictor variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
-Specifically, it is assumed that each variable is conditionally independent given the respose class alone, without considering other predictor variables.
-Mathematically this means that each variable's probability contribution can be simply calculated by looking at the probability of a response class given a predictor value by itself.
-The predictor variables probability estimates are combined for each possible response class, and the response class with the highest probability is chosen [@Provost2013].
+The naive bayes method works by making conditional independence assumptions about the explanatory variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
+More specifically, it is assumed that each variable is conditionally independent given the respose class, without considering other predictor variables.
+Mathematically this means that each variable's probability contribution can be simply calculated by looking at the probability of a response class given a single predictor value.
+A probability of each response class can be calculated for each explanatory variable.
+These are then combined for each possible response class using Bayes theorem, and the response class with the highest probability is chosen [@Provost2013].
 
@@
 The method can perform very well for real world tasks because the assumption of independence does not damage predictions significantly.
-This is because if multiple varialbes are related, the variables independently still direct the prediction in the correct direction.
+This is because if multiple variables are related, the variables independently still direct the prediction in the correct direction.
 The correlated variables will double or triple their emphasis on the predicted response variable, however this does not necessarily impact classification accuracy as the final step is to simply choose the class with the *highest* probability [@Provost2013].
@@
 
-A real world success story for Naive Bayes classifiers hinges on their ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
-Naive Bayes has been used in complex spam detection systems where new spam emails or toxic text themes can be quickly added to the filtering model.
+A real world success story for Naive Bayes classifiers hinged on their ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
+Naive Bayes was used in a complex spam detection system where new spam emails or toxic text themes could be quickly added to the filtering model.
 The Naive Bayes method provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
@@
 There are several brands of basic decision tree algorithms including ID3, C4.5, CART, and CHAID with CART and C4.5 being the most popular [@Kabra2001].
-Ravi Kumar and Ravi [@Kumar2007] recommend the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
-Single decision trees have the risk of over-fitting the data and low predictive accuracy but have favourable characteristics such as providing intuitive rules that a decision maker can follow in real-life scenarios.
-In addition their results do not depend heavily on the skill of the data analyst [@Kabra2001][@Putler2012].
-
-To tackle the problem of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
-Three examples of these ensemble decision tree methods are bagging, boosting, and random forests.
-Bagging creates multiple trees by sampling a different training data set for each tree and then combining tree results [@Breiman1996].
-Random forests is an advanced form of bagging in that multiple trees are sampled from the same training data.
-However, when creating the tree for each sample dataset, a random subset of attributes is used to determine each split.
-This prevents the trees from becoming too correlated [@Breiman2001].
-The boosted decision tree approach combines gradient descent with a series of decision trees.
-Each tree is limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) from the preceding tree.
+Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
+Single decision trees tend to overfit the data and provide low predictive accuracy but have favourable characteristics such as creating intuitive rules that a decision maker can follow in real-life scenarios.
+In addition, their results do not depend heavily on the skill of the data analyst [@Kabra2001][@Putler2012].
+
+To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
+Three examples of these ensemble decision tree methods are boostrap aggregatiing (bagging), boosting, and random forests.
+Bagging creates multiple trees by sampling a different training data set (*with* replacement) for each tree and then combining tree results [@Breiman1996].
+Random forests is an advanced form of bagging and builds on the idea of training multiple trees from the same data by sampling bootstrapped training sets with replacement.
+However, when creating each tree, a random subset of attributes (variables) are considered at each split.
+The reduced subset of attributes is resampled for each split in the tree.
+This allows dominant variables to be supressed for a fraction of the splits, allowing the algorithm to explore signals in weaker variables.
+It also prevents the trees from becoming too correlated [@Breiman2001].
+The boosted decision tree approach applies gradient descent theory to a series of decision trees.
+The trees are limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) of the preceding tree.
 The limited depth of each tree prevents overfitting at each stage and the combined result of up to thousands of trees is very powerful [@Elith2008].
 
-Ensemble decision trees have a different foundation to more traditional statistical prediction methods such as regression.
+Ensemble decision trees have a different foundation to traditional statistical prediction methods such as regression.
 For this reason, there are some advantages and disadvantages.
-The progression of binary splits means non-linear feature interactions can be captured and the predictor variables do not need to be transformed as no assumptions are made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
+Decision trees use a progression of binary splits which means the data is first split by one variable at a certain value, then these subsets are split by another variable and so on in order to partition the response variable values into groups with similar values.
+The aim of each split is to achieve the most 'pure' split of the response values in a greedy fashion.
+If this process is visualised graphically, it is clear that non-linear relationships can be captured.
+
+![](images/lit_review/dec_tree.JPG)
+[@Jeevan2015]
+
+In contrast to linear regression, the predictor variables do not need to be transformed as no assumptions are made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
 Other advantages are that the ensemble methods are very fast in comparison to SVM's and neural networks but perform just as well and provide insights such as variable importance and variable relationships [@Sealfon2012].
-Random forests in particular has the benefit of not being affected by noise in the data as well as the capacity to handle a large number of variables [@Sealfon2012].
+Random forests in particular have the benefit being unaffected by noise in the data as well as the capacity to handle a large number of variables [@Sealfon2012].
 Caruana [@Caruana2006] tested boosted trees, random forests, neural networks, SVM's, logistic regression and naive bayes on 11 binary classification problems and found that boosted trees performed best, followed by random forests.
-Clearly, ensemble tree methods can compete with high level machine learning algorithms.
-
-Along with the many advantages of trees and ensemble trees, it is important to be aware of their limitations.
-As trees are not built on a probabilistic framework, results can not be provided in this framework.
-For example, confidence intervals for predictions are not available for standard ensemble methods [@Louppe2014].
+This demonstrates ensemble tree methods are capable of competing with high level machine learning algorithms.
+
+Ensemble trees have numerous advantages however it is important to be aware of their limitations.
+Trees are not built on a probabilistic framework, and therefore their results can not be provided in this framework.
+For example, statistical confidence intervals for predictions are not available for standard ensemble methods [@Louppe2014].
+This may or may not be required depending on the nature of the problem.
 Also, variable importance tables provided by random forests and boosted trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
-The methods can also be prone to overfitting if not handled carefully.
+The methods can also be prone to overfitting if not tuned carefully.
 Ensemble trees have many advantages particularly predictive power, but it is important to recognise that the best method for a problem often depends on the type of problem [@Louppe2014][@Caruana2006].
@@
 A bayesian network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
-The model is visually represented by a DAG (directed acyclic graph) and is capable of linking the conditional dependency between any variable to another variable or via other variables which are represetned by arcs [@Heckerman1998].
-This kind of conditional relationship is bayesian.
-Bayesian networks have found success in combining deterministic models with observation data as well as expert knowledge as the statistical relationships between variables can be determined from different places and manually entered into a single model [@Kragt2009].
-
-Bayesian networks are excellent at representing knowledge from different sources and types, and visuallising them in a way that communicates results well to decision makers [@Kragt2009].
-Therefore bayesian networks become an excellent decision support tool.
-Network relatitonships can be machine learned, however this is not widely included in the suite of machine learning methods and is referred to more as a way to model complex networks.
-Drawbacks of bayesian networks include their inability to handle continuous numeric variables (which must be discretised).
-This means that the varialbe relationships are categorical [@Kragt2009].
+The model is visually represented by a DAG (directed acyclic graph) and is capable of linking the conditional dependency between any variable to another variable or via other variables.
+The links are graphically represented by arcs or edges [@Heckerman1998].
+This kind of conditional relationship is bayesian, where the probabilities in one node are conditional upon values in nodes directed to it as well as preceding nodes.
+Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.
+The statistical relationships between variables can be determined from separate sources and manually combined into a single model [@Kragt2009].
+
+Bayesian networks are excellent at representing knowledge from different origins, and then visually communicating results well to decision makers [@Kragt2009].
+Therefore bayesian networks have become an excellent decision support tool.
+Network relationships can be machine learned, however this is not widely included as part of the suite of machine learning methods and is usually recommeneded for modelling complex networks.
+Drawbacks of bayesian networks include their inability to handle continuous numeric variables (all continuous variables must be discretised).
+This ensures all variable relationships are categorical [@Kragt2009].
 Popular applications of bayesian networks are modelling uncertainty in natural resource management and modelling complex business network structures such as airports [@Wu2013].
@@
 
-Neural networks were inspired by biological neural networks of the human nervous system and require a lot of training data.
-They are very flexible in the relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
-A disadvantage is that although the results are accurate, the algorithm is a black box - meaning the data is input and the results output without providing the user insight into the methodology.
+Neural networks were inspired by biological neural networks of the human nervous system and require a lot of training data relative to other methods.
+They are flexible in the range of relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
+A disadvantage is that although the results are accurate, the algorithm is a black box - meaning the data is input and the results output without providing the user with insight into the methodology.
 Neural networks can perform high level tasks as a result of deep learning such as hand writing recognition, vehicle control, face recognition, and cancer detection [@Haykin2004].
 
-The simplest type of Support vector machines (SVM's) calculate the linear boundary between two categories in feature space so that the category of future cases may be determined depending on which side of the boundary they fall.
-The location of the boundary is determined by the position of the widest clear margin that can be drawn between the two categories, allowing a certain number of misclassifications.
+The simplest type of SVM's calculate the linear boundary between two categories in feature space so that the category of future cases may be determined depending on which side of the boundary they fall.
+The location of the boundary is determined by the position of the widest clear margin that can be drawn between data points from the two categories, while allowing a certain number of misclassifications.
+The simple plot below demonstrates this concept and has no misclassifications:
+
+![](images/lit_review/svm.JPG)
+[@Provost2013]
+
 This concept can be adapted to non linear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
@@
 This section reviewed the advantages and disadvantages of statistical and machine learning techniques that have been successfully applied to business problems.
-These include linear regression, logistic regression, naive bayes, decision trees and ensemble trees, bayesian networks, neural networks, and SVM's.
+These included linear regression, logistic regression, naive bayes, decision trees and ensemble trees, bayesian networks, neural networks, and SVM's.
 Linear/logistic regression and naive bayes are computationally very fast with a simple concept that make broad assumptions about the data, but work well as baseline models against which to compare complex models.
 In certain cases these simpler models are significantly outperformed by complex models however it is important to compare both.
-Neural networks and SVM's are excellent machine learning predictors which can perform well with messy data and no assumptions, however they are slow and do not provide insight into the results of analysis.
-This is not appropriate for the profitability prediction problem in this project because communication of the reasoning behind predictions is crucial for uptake of the model by decision makers.
-
-On the other hand, ensemble tree methods can perform as well as neural networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
-Bayesian networks are excellent for presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally not been included in the machine learning literature as much as the other methods.
+Neural networks and SVM's are excellent machine learning predictors which can perform well with messy data and no distributional assumptions, however they are slow and do not provide insight into the results of analysis.
+This is not appropriate for the effort estimation problem in this project because communication of the reasoning behind predictions is crucial for uptake of the model by decision makers.
+
+On the other hand, ensemble tree methods are capable of performing as well as neural networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
+Bayesian networks are excellent for presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally been excluded from the machine learning literature in comparison to the other methods.
 This may be due to low awareness of learned bayesian networks or lower performance.
-In conclusion, simple models such as linear/logistic regression and naive bayes are good baseline models to implement, while ensemble tree methods and bayesian networks may handle messy data better than the simpler models while still providing insight to decision makers.
+In conclusion, simple models such as linear/logistic regression and naive bayes are good baseline models to implement, while ensemble tree methods and bayesian networks may handle messy 'real world' data better than the simpler models while still providing insight to decision makers.
 
@@
 The aim of this research project is to create a predictive model for the profitability of a company's consulting projects using their internal CRM data.
-The literature pertaining closely to the cost prediction topic has been reviewed in section 2.1, however it is relevant to broaden the assessessment of the literature on advanced machine learning techniques that have been applied to general business problems.
-The simpler statistical methods have been covered already, so the focus of this sections is on the topic of machine learning prediction methods for business decision-makers.
+The literature pertaining closely to the cost prediction topic has been reviewed in the first section of this chapter, however it is relevant to broaden the critique of the literature on advanced machine learning techniques that have been applied to general business problems.
+The simpler statistical methods have been covered already, so the focus of this section are the machine learning prediction methods developed for business decisions.
 Popular applied topics include predicting stock fluctuation, customer churn analysis, fraud prediction, customer classification, market segment analysis, product success prediction, and recommendation systems (Seng & Chen, 2010).
-This section will first discuss the use and merits of the employed more advanced machine learning methods, followed by a case study of employee-churn prediction.
+This section will first discuss the use and merits of the advanced machine learning methods applied to these problems, followed by a case study of employee-churn prediction.
 
@@
 
-According to Breiman [@Breiman2001b], simple theoretical models such as linear regression are informative but often provide unsatisfactory accuracy for modelling real life data which is messy and does not necessarily align with the rules of statistical theory.
-As this project will address a problem with a large quantity of messy data, it is worthwhile reviewing the algorithmic machine-learning predictive models that have developed rapidly in recent years.
-They have found success because they are able to sometimes fit real-life data with more accuracy than theoretical models [@Breiman2001b].
+According to Breiman [@Breiman2001b], simple theoretical models such as linear regression are informative but often provide unsatisfactory accuracy for modelling real life data.
+This data is often messy and does not necessarily align with the rules of statistical theory.
+As this project has a large quantity of messy data, it is worthwhile reviewing the algorithmic machine-learning predictive models that have developed in recent years.
+They have found success because they were able to fit real-life data with more accuracy than theoretical models [@Breiman2001b].
+
 Ravi Kumar and Ravi [@Kumar2007] performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
-The most widely used model was neural networks, however logistic and linear regression, decision trees, SVM’s, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
-Supporting Breiman’s [@Breiman2001b] statements, it was found that DA and linear regression techniques were not preferred due to their low accuracy.
-The overall assessment was that SVM’s outperformed Neural networks (back propagation neural networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
-Although SVM’s performed the best, as discussed, they are often extremely complex and slow, requiring a high memory [@Kumar2007].
-In support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that statisticians experienced in predictive machine learning algorithms generally acknowledge that SVM’s yield the highest prediction accuracy compared to other machine learning algorithms.
-
-As SVM’s are often computationally too expensive and advanced for a smaller scale business applications, such as this research project, it is useful to compare neural networks and decision trees which are simpler than SVM’s but are also capable of achieving very accurate predictive models.
-Ravi Kumar and Ravi [@Kumar2007] found that neural networks and decision trees were both capable of out-performing the other, depending on the context – which is limited to bankruptcy prediction in Kumar’s study.
-Both have advantages and disadvantages to consider (refer section 2.2), however the most notable drawback for neural networks is the difficulty in interpreting the algorithms' explanation of human behaviour – a valuable feature in business problems [@Tsai2009].
+The most widely used model was neural networks, however logistic and linear regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
+Supporting @Breiman2001b's statements, it was found that DA and linear regression techniques were not preferred due to their low accuracy.
+The overall assessment was that SVM's outperformed neural networks (back propagation neural networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
+Although SVM's performed the best, as discussed, they are often extremely complex and slow, requiring a great deal of memory [@Kumar2007].
+In futher support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms generally agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
+
+As SVM's are often computationally too expensive and advanced for smaller scale business applications, which this research project would fall into, it is useful to further compare neural networks and decision trees which are simpler than SVM's but are also capable of achieving very accurate predictive models.
+Ravi Kumar and Ravi [@Kumar2007] found that neural networks and decision trees were both capable of out-performing the other, depending on the context - which is limited to bankruptcy prediction in Kumar's study.
+Both have advantages and disadvantages to consider, however as derived in the previous section, @Tsai2009's study on earnings management prediction found the most notable drawback for neural networks is the difficulty in interpreting the reasoning behind the algorithm.
+Earning's management is influenced by managers' choices and insight into the model structure could provide explanations for human behaviour - a valuable output in business problems.
 On the other hand, trees are known to be intuitive to understand and have been argued to function similarly to how the human mind thinks [@Carrizosa2010].
@@
 For example, combining the contrasting advantages of neural networks and decision trees is a worthwhile ensemble technique.
-A paper studying financial earnings management prediction combined these two methods by running them one after another.
-The neural network was run first to create a high rate of prediction accuracy via its complex non-linear learning properties [@Tsai2009].
+@Tsai2009's study eventually combined these two methods after trialling each one.
+They were run successively and the neural network ran first because it had a higher rate of prediction accuracy via its complex non-linear learning properties.
 Then, to resolve the lack of explanatory qualities that neural networks are notorious for, decision trees were employed.
@@
 
-The literature on predictive business models is generally limited to studies on product sales, customer behaviour, or overall performance of large businesses.
-There are far fewer studies that assist with internal business decisions dealing with employees, teams, and delivering projects as is the case with this project.
-An exception is Saradhi and Vijaya’s [@Saradhi2011] study on employee churn, where ‘churn’ refers to the number of individuals moving out of a group within a certain time.
-Saradhi and Vijaya applied popular customer churn predictive models to employee churn - a novel application that focussed internally on employees rather than externally on customers.
-The project chose three machine learning classification algorithms that are normally used in customer churn and it was proposed that the variables and model behaviour of employee churn could be related to customer churn.
-The associated costs of losing customers and finding new customers can also be correlated to the costs of losing staff and hiring new staff.
-
-Naive Bayes, Random forests (an ensemble decision tree method) and SVM’s were built.
-A comparison study showed that Naïve Bayes is outperformed by random forests which match Saradhi and Vijaya’s prediction [@Caruana2006].
-The SVM model actually far out performed random forests and Naive Bayes by achieving 81% vs 51% and 55% correct predictions respectively.
-This was attributed to the ability to introduce class penalties in SVMs whereas the other two methods are limited by class imbalance [@Saradhi2011].
-This means that SVM’s had a superior method of balancing the importance of the variables in the data set which was skewed by a heavy majority of employees who kept their job versus the minority who left (25% churn).
+The literature on predictive business models is generally limited to applications on product sales, customer behaviour, or overall performance of large businesses.
+There are far fewer studies that assist with internal business decisions dealing with employees, teams, and delivering projects which is what this research project aims to achieve.
+An exception is @Saradhi2011's study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
+@Saradhi2011 applied popular customer churn predictive models to employee churn - a novel application that focussed internally on employees rather than externally on customers.
+The project tested three machine learning classification algorithms that were commonly applied to customer churn.
+The reasoning behind this was that the variables and model behaviour of employee churn could be related to customer churn.
+The associated costs of losing customers and finding new customers can also be connected to the costs of losing staff and hiring new staff.
+
+Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
+All three models performed at similar levels for overall accuracy at around 80% (the general ranking was random forests, followed by SVM then Naive Bayes).
+However, when the true positive rates (accuracy in  predicting of employee resignations) of the models were compared, the SVM model far out performed random forests and Naive Bayes by achieving 81% TP vs.
+51% and 55% respectively.
+This was attributed to the ability of SVM's to incorporate class penalties whereas the other two methods were limited by the class imbalance problem [@Saradhi2011].
+This means that SVM's had an effective method of increasing the importance of 'positive' cases in the data set which was skewed towards the majority of employees who kept their job versus the minority who left (25% churn).
 It should be noted that the boosted tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVMs.
-Particularly because boosted trees provide insight into the predictive model which could aid uptake of the model with decisoin makers.
-However, it is noteworthy that their study proved that predictive techniques used on customres could be translated to predicting internal behaviour of employees.
+Furthermore, boosted trees provide insight into the predictive model which could build trust in the company's decision makers.
+Nevertheless, it is noteworthy that their study proved employee churn models for customers could be translated to predicting internal churn of employees.
+
+Another idea @Saradhi2011 adopted from customer churn methods was the value models.
+Company's calcuate customer lifetime values (CLV's) in order to structure which potentially churning customer's should be acted upon.
+Then the optimal number of customers to reach out to (reaching out incurs a cost) that prevents the most financial loss from churn can be determined.
+@Saradhi2011 developed a system for determining the value of each employee in terms of the importance of the projects they were on and their monthly chargeability.
+This allowed them to rank the employees identified as 'high risk of churn' by value and provides a clear direction for manager's to act upon.
+Manager's can then start brainstorming what they can do to prevent high value, high risk of churn employees from leaving.
+This extension of the study provided a comprehensive analysis of how business manager's could adopt their findings to improve their business operations.
+It was a valuable addition which is absent from most cost estimation and effort estimation research.
 
@@
 
-The limited literature on analysis similar to Sardhi and Vijaya’s Employee Churn study [@Saradhi2011] highlights a gap in the application of predictive techniques to model internal performance in businesses.
+The limited literature on analysis similar to @Saradhi2011's Employee Churn study highlights a gap in the application of predictive techniques to model internal performance in businesses.
 This is particularly relevant to consulting companies that tackle complicated discrete projects with set timelines.
-In this project, internal consulting project data will be used to predict a new project performance in terms of internal employee structure as well as client characteristics.
-A review of the literature in machine learning applied to business problems revealed that SVM's have found to be the most accurate technique followed by neural networks and ensemble tree methods.
-As discussed in section 2.2, it was also noted that neural networks and SVM's do not provide insight into the model's predictions to the decision maker.
-A gap exists in applying these cutting edge machine learning techniques to internal business decision-making.
+In this research, internal consulting project data will be used to predict a new project's profitability in terms of internal employee structure as well as client characteristics.
+A review of the literature in machine learning applied to business problems revealed that SVM's were found to be the most accurate technique followed by neural networks and ensemble tree methods.
+As discussed in the previous section, neural networks and SVM's do not provide insight into the predictions to the decision maker.
+A gap exists in applying these cutting edge machine learning techniques to the wide range of internal decisions that businesses make.
 
@@
 The literature pertaining to the aim of this research project, cost prediction using internal consulting CRM data, has been extensively reviewed.
-The problem of cost prediction has been addressed in many studies beforehand which have generally been limited to prediction of construction costs for buildings and infrastructure and effort estimation for IT projects.
+The problem of cost prediction has been addressed in many studies beforehand which have generally been limited to prediction of construction costs for buildings and infrastructure as well as effort estimation for IT projects.
 The prediction models consisted of case based reasoning, linear regression, neural networks and in one or two instances an ensemble tree or SVM.
-This leaves a gap in the literature in effort estimation for consulting companies in the construction industry as well as trialling a broader range of prediction methods.
-
-An exhaustive list of machine learning and statistical methods were reviewed theoretically as well as practically in the broader scope of business problem prediction.
+This leaves a gap in the literature in effort estimation for consulting companies in the construction industry in combination with trialling a broader range of prediction methods.
+
+An exhaustive list of machine learning and statistical models were reviewed theoretically as well as practically in the broader scope of business problem prediction.
 It was determined that bayesian networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as neural networks while providing the additional benefit of insights into the reasoning behind model predictions.
-
-# Chapter 3 Method - 2 to 8 pages
+These insights are very important as they address the lack of industry uptake - a major issue identified by numerous studies in both industries.
+Other considerations and strategies will be developed to address this issue as well.
+
+# Method - 2 to 8 pages
 
 The aim of this project is to determine whether the profitability of projects for consulting businesses can be predicted using statistical and machine learning techniques.
-As stated in the literature review, a gap in research exists for testing whether internal CRM data from a company in the construction industry could predict project profitability.
-This section will explain the chosen methodology for testing this aim using a case study Engineering consulting company that charges for their time spent on project delivery.
-Once the predicition potential has been assessed, a secondary aim is to determine the positive impact the predictive model could have on the overall profitability of the company.
+As stated in the literature review, a couple of gaps exist in this research space including working with internal CRM data from a consulting company in the construction industry, trialling ensemble tree methods against the more popular regression and neural networks, and analysing the impact model results could have on business decisions.
+This section will explain the chosen methodology for fulfilling the research aim using a case study engineering consulting company that charges for their time spent on project delivery.
 
@@
 
-An assessment of the literature revealed that the predictive power of internal timesheet data has not yet been tested yet for project profitability for consulting companies in the construction industry.
-In addition, this is the sole source of project data available for these types of businesses besides interviewing project managers and reviewing the account records of invoicing versus employee costs.
-As the internal CRM for the case study company stores a detailed account of time spent on each project along and client information for twelve years, it was clear that this source of data could reveal the most and was most suitable to statistical analysis.
-The complete twelve years of project timesheets and invoices was extracted via the CRM software interface, describing 4169 projects.
-The projects varied from total invoiced amounts of $500 to over $1,000,000 across four internal disciplines in the company.
-At the time of data extraction, it was not clear whether the full twelve years were necessary and relevant to predicting current project profitability but this was to be determined later in the analysis.
+An assessment of the literature revealed that the capacity of internal timesheet data (extracted from a company's CRM software) to predict project profitability has not yet been tested in the context of consulting companies in the construction industry.
+Project data can be extracted from the CRM, interviewing project team members, emails, accounting records, and project drawings and calculations.
+As the internal CRM data in this case study houses a detailed account of time spent on each project (timesheet data) along with client information and invoicing records for the preceding twelve years, it was clear that this source of data was not only the most accessible and suitable for statistical analysis, but had the potential to reveal information not typically investigated by the company.
+
+The complete twelve years of project timesheets and invoices was extracted via the CRM software interface, which described 4169 projects.
+The projects varied from total invoiced amounts of \$500 to over \$1,000,000 and were divided between four internal disciplines in the company.
+During data extraction, it was not clear whether the full twelve years were necessary and relevant to predicting future project profitability but this was determined later in the analysis.
 
@@
 This is an important step before research commences as the privacy of project employees, clients, and project names must be removed for ethical reasons.
-In the case of employee names, professional titles replaced the names, and for client codes replaced client names.
+In the case of employee names, professional titles replaced the names, and client codes replaced client names.
+Project names and addresses were removed, and each project was identified by a number.
+The resulting de-identified datasets were three .csv files of different data structures which are described in the next section.
 
@@
 
-The obtained data contained a rich source of project information and a lengthy process of cleaning the data and engineering useful variables was required.
-The data was initially in three types of datasets extracted from the same CRM software:
-
-* Invoicing data
-* Timesheet data
-* Project summary data
-
-These three sources needed to be compiled into a single dataset that detailed one project per row as this is the structure required for predictive analytic methods when predicting the overall profitability of a project.
-Before compilation could begin however, thorough cleaning is required where the data was plotted and statistically assessed so that outliers could be visually or statistically discovered.
+The data contained a rich source of project information and a lengthy process of cleaning the data followed by engineering potentially useful variables was then required.
+The three initial datasets described:
+
+* Invoicing records: dated invoiced amounts and sizes for each project and whether the invoice had been paid or written off
+* Timesheet entries: the number of hours spent on each project and on which day.
+Entries were input by individual employees, but names for each entry had been replaced with employee positions (such as mid-level technical)
+* Project summary data: information describing each project such as client code, client contact code, discipline, subject, post code, director code, and suburb
+
+These three sources needed to be compiled into a single dataset that detailed one project per row as the model was to predict the overall profitability of a project.
+Before compilation could begin however, thorough cleaning was required where the data was plotted and statistically assessed so that outliers could be visually or statistically discovered.
 Outliers were then investigated for data entry errors.
 Many such errors were encountered.
-One type of error for example was a user entering a 123 kilometer drive in a car as 123 hours.
+For example, a common glitch in the data collection process translated a user entering a 123 kilometer drive in a car as 123 hours.
+Or an expense printing claim of $9.50 would be translated to 9.5 hours spent on that job.
 Once these errors were detected, they were discussed with the company directors and appropriate corrective action was taken.
+Mistakes in data entry were often able to be converted into correct data by notes that a user had entered into the system.
+It is possible that some data entry errors were not detected and this is a problem that must be dealt with in implementation of the model as well.
 
@@
 
-Although dozens of variables were available from the initial dataset, it was prudent to engineer further descriptive variables, particularly in the invoicing dataset and timesheet dataset where the information needed to be compiled into a single row per project.
+Although dozens of variables were available from the initial dataset, it was advantageous to engineer further descriptive variables to trial in the models.
+This was particularly true for the invoicing dataset and timesheet dataset where a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
 All variables would eventually be tested for variable importance and predictive power with respect to project profitability.
@@
 
-Timesheet Data
-
-* percent of hours performed by each professional role for the entire project
+**Timesheet Data**
+
+* percent of hours performed by each professional role over the course of a project
 * timespan of the entered project hours
-* percent of hours performed by 'profesional' employees as opposed to 'technical'
+* percent of hours performed by 'profesional' employees as opposed to 'technical' employees
+* position of the employee who completed the most hours on each project
+* percent of hours done by the majority contributor to a project
 * total cost of employee hours per project
@@
 * total number of users that entered hours on each project
-* mean hours per day entered on a project
+* mean number of hours per day entered on a project
 * number of disciplines active in a project
 
-Invoicing data
+**Invoicing data**
 
 * total amount invoiced and renumerated per project
-* mean invoice size per project
-* mean invoice per client
-* invoice frequency per project
-* client invoice frequency
-
-Project Data
+* mean invoice size per project (for example if a project had four invoices totalling \$10,000, the mean invoice size would be \$2,500)
+* mean invoice per client, i.e.
+the average invoice size across all invoices sent to a client.
+This gives an indication of whether the client generally does fast paced big jobs which would result in large monthly invoices for example or small jobs with small monthly invoiced amounts.
+* invoice frequency per project. This indicates how drawn out a project was.
+* client invoice frequency which gives an indication of how much work the company does for that client.
+
+**Project Data**
 
 * text analysis of project descriptions detected a list of key words that could classify projects into 16 categories.
-This key word analysis was done in conjunction with a company employee and the resulting classifications were reviewed.
+This key word analysis was done in conjunction with a company employee and the resulting classifications were reviewed by the employee to ensure accuracy.
 * number of projects completed with each client and client contact
 
-Once the three data sources were combined, further variables were engineered:
-
-* project profit
-* return per dollar for each project (how much was spent on expenses for every dollar earned?)
-
-Besides the engineered variables, additional variables included in the project summary dataset were client industry, discipline, job description, and post code.
-Once numerous potentially important variables were engineered, the most important ones could be narrowed down.
-This improves the accuracy of a model because unnecessary or irrelevant variables add noise to the prediction of target values.
-Variable selection also increases computational efficiency by reducing the number of calculations and improves understanding of the prediction structure [@Weisberg2005].
+With the engineered variables, all three data sources could be represented in the format of a single row per project.
+They were hten combined and further variables were engineered using combinations of variables across the three data sets:
+
+* profit
+
+$project\ profit = total\ invoiced\ amount - cost\ from\ recorded\ employee\ hours$
+
+* return per dollar as the measure of profitability
+
+$return\ per\ dollar = \frac{project\ profit}{cost\ from\ hours}$
+
+Besides the engineered variables, additional variables included in the project summary dataset were client industry, internal company discipline, job description, and post code.
+Once numerous potentially important variables were engineered, the most important ones could then be narrowed down.
+This practice improves the accuracy of a model because unnecessary or irrelevant variables add noise to the prediction of target values.
+Variable selection also increases computational efficiency by reducing the number of calculations and enhances understanding of the prediction structure [@Weisberg2005].
 
@@
 
+Once
+
 ### Outlier Deletion
 
-Before determining which variables contribute most to the target value, return per dollar, outliers were deleted.
+Before determining which variables contribute most to the response variable, return per dollar, outliers were deleted.
 Specifically, cases with extreme return per dollar values were investigated as special scenarios may have contributed to these values and the model is not intended to predict special scenarios.
@@
 
-plot.pvals = readRDS("AOV_varimp_plot.rds")
+plot.pvals = readRDS("report_data/AOV_varimp_plot.rds")
 
@@
 # How to achieve statistical power of 0.8?
-finance.plot = readRDS("report_data/plot_finance.rds" )
+finance.plot = readRDS("report_data/plot_finance_100.rds" )
 fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 

### 2016-06-17 finalwriteup/trial_code.rmd
reviewed lit review started method review

---
title: "trials"
author: "Amy Cook"
date: "June 16, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$effort = \alpha*size^{\beta}$  
$\alpha = productivity \ coefficient$  
$\beta = economies \ of\ scale\ coefficient$  
$size = estimated \ lines \ of \ code$

$$
\begin{aligned}
 effort &= \alpha*size^{\beta} \\
 \alpha &= productivity \ coefficient \\
 \beta &= economies \ of\ scale\ coefficient \\
 size &= estimated \ lines \ of \ code
\end{aligned}
$$

* $project\ profit = total\ invoiced\ amount - cost\ from\ recorded\ employee\ hours$
* $return\ per\ dollar = \frac{project\ profit}{cost\ from\ hours}$

### 2016-06-17 financial.rmd
reviewed lit review started method review

@@
                    k.folds.2 = 5, # number of folds for second step blended model development
-                   tot.tests = 50 # total number of results to create
+                   tot.tests = 500 # total number of results to create
                    ){
@@
 
-        assign(prof.results, "prof.results", envir = .GlobalEnv)
+        assign("prof.results", prof.results, envir = .GlobalEnv)
 
 }
+
+#function below assigns results to 'prof.results' object in global env
+# profit(df = all10mice, method = c("average","simp.log","orig.log","orig.rf","orig.boost"), 
+#                    seed = 300, 
+#                    k.folds = 5, # number of folds in initial model.blend generation
+#                    max.iter = 5, #max number of iterations from initial model.blend generation
+#                    k.folds.2 = 5, # number of folds for second step blended model development
+#                    tot.tests = 500 # total number of results to create
+#                    )
 
@@
 
-# saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance.rds" )
+# saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100.rds" )
 

### 2016-06-19 finalwriteup/draft.rmd
started review of variable selection method

@@
 
-Once
+Once variables were engineered, the value each variable could contribute to the predictive model was assessed.
+As previously stated, variable selection can enhance model accuracy by reducing noise from irrelevant variables and increase simplicity which in turn improves interpretability for the user.
+To assess variable importance, models predicting 'return per dollar' were built using all available explanatory variables.
+The models then output which variables contributed significantly or most improved results.
+Different types of predictive models have different methods of calculating variable importance and since it was not known which model would best predict return per dollar, a few models' variable outputs were reviewed.
+These included ANOVA linear regression and two ensemble tree methods.
+Before models assessing variable importance were built, outlier cases were deleted to prevent distortion of valuable explanatory variables.
 
@@
 
-Before determining which variables contribute most to the response variable, return per dollar, outliers were deleted.
-Specifically, cases with extreme return per dollar values were investigated as special scenarios may have contributed to these values and the model is not intended to predict special scenarios.
-Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles as intially practised by Tukey, the inventor of the box plot [@Tukey1977].
-The outlier projects determined by this method were then reviewed with an employee from the case study company.
-Based on their domain knowledge and assessment of the outlier cases, cut-off values for the range of return per dollar values were determined.
-Outliers were projects with return per dollar values greater than 3 or below -2.
+Case study projects with extreme 'return per dollar' values were likely the result of special scenarios which the model is not intended to predict.
+For example, many projects appeared wildly profitable but on closer inspection these were all very small jobs.
+After discussion with the case study company, it was concluded that for those outliers, the small jobs most likely required so little time that busy employees did not bother writing down their hours for that job.
+Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles.
+This rule was intially practised by Tukey, the inventor of the box plot [@Tukey1977].
+All outlier projects determined by this method were reviewed with an employee from the case study company and based on their domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined.
+In summary, projects with 'return per dollar' values greater than 3 or below -2 were removed.
 
@@
 
-Three methods of variable selection were trialled and compared to test whether the different methods highlighted different variables as important.
-These were Analysis of Variance (ANOVA), conditional inferece forests (cforests), and random forests.
-It was expected that random forests would highlight variables that had more categories as more important which is a well known bias; it is also a bias that cforests have overcome [@Strobl2007].
-ANOVA is based on linear theory which contrasts conditional forest and random forests more free-form structure.
-Therefore, it is important to compare the contrasting methods' assessment of variable importance.
+Three methods which output variable importance were trialled and compared to test whether they highlighted different variables as valuable.
+This was anticipated as each method models data using distinct theoretical foundations.
+The tested methods were ANOVA linear regression, conditional inference forests (cforests), and random forests.
+It was expected that random forests would favour variables that had more categories, a well known bias; which is also a bias that cforests have overcome [@Strobl2007].
+ANOVA is based on linear theory which contrasts conditional forest and random forest's free-form structure.
+Therefore, it is benefical to compare the methods' assessments of variable importance.
 
 When running variable importance tests, the dependent variable must be clearly defined.
-For this problem, the dependent variable can be numeric or categorical: whether a project was profitable (binary classification) or *how* profitable a project was (regression).
-The regression model was chosen as a variable important for predicting the degree of profitability would also predict profitable vs not profitable.
+For this problem, the response variable could be continuous or categorical i.e.
+*how* profitable a project was (continous) or whether a project was profitable or not (binary classification).
+The continuous response variable was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
 

### 2016-06-20 finalwriteup/draft.rmd
almost finished method review

@@
 BIM - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
+ROC - Reciever Operating Characteristic
+AUC - Area under the curve. In this project, the area refers to the area under an ROC curve
+TPR - True positive rate
+FPR - False positive rate
+TP - true positive
+FP - false positive
+DA - Discriminant analysis
+CLV - Customer Lifetime Value
+IQR - Interquartile range
+ESS - error sum of squares
+RMSE - root mean squared error
 
@@
 The method is simple and fast but assumes variables are normally distributed and linearlly related.
-The variables must therefore be normalised before analysis, however it cannot be assumed that each variable can be perfectly normalised.
-A disadvantage of this method is that linear relationships and normal variables do not necessarily represent real world data in the best way [@Breiman2001b].
-
-In order to compare variable importance, the output of the model is assessed which includes the linear coefficients for each variable and the p-values for the coefficients [@Markham2016].
-The magnitude of the coefficient does not indicate importance relative to the other variables because the magintude is directly dependent on the values within the variable.
-This leaves the p-values which are the result of a test for whether there is a significant relationship between the variable and the target variable.
-This is a good indication of whether the variable has a relationship with the target variable, however doesnt necessarily rank importance.
-If a variable is measured more precisely, it will have a smaller p value, whereas a variable measured roughly will have a higher p value.
-This does not necessarily make the more precisely measure variable more important than the other.
-The p test does however distinguish between variables that have a statistical relationship to the target variable and ones that do not.
-With regards to p-values it is also important to keep in mind that, with a dozen variables, chance alone can produce a variable with a pvalue < 0.001 7% of the time [@Rice1989].
+The variables must therefore be normalised before analysis to achieve a statistical distribution as close to gaussian as possible.
+A disadvantage of this method is that linear relationships and normally distributed variables do not necessarily represent real world data accurately [@Breiman2001b].
+
+In order to compare variable importance, two types of output from the model are assessed which include the linear coefficients for each variable and the p-values for the coefficients [@Markham2016].
+The magnitude of a coefficient does not indicate importance relative to the other variables because its value is directly reflects the values within the variable.
+The sign of the coefficient however indicates whether the variable is directly or inversely related to the response variable.
+A p-values for each variable coefficient is output as the result of a significance test for whether the coefficient equals zero.
+Therefore, if the null hypothesis is rejected (p value < 0.05), it indicates that changes in the explanatory variable are related to changes in the response variable.
+P-values are a good indication of whether variables have a relationship with the target variable, however they do not necessarily rank importance.
+If a variable is measured more precisely, it will have a smaller p-value, whereas a variable measured roughly will have a higher p-value.
+This does not necessarily make the more precisely measure variable is more important than the other.
+It is also important to keep in mind that, with a dozen variables, chance alone can produce a variable with a p-value < 0.001 7% of the time [@Rice1989].
 
@@
 The random forest algorithm can produce a permutation variable importance for each variable as part of its output.
-Variable importance is represented as a score, and to do this a permutation and calcuations are performed on each variable.
-More specifically, consider a single covariate.
-The values of this variable in the out of bag sample are randomly permuted (reordered) for each tree - a technique that mimics the absence of that variable.
+Variable importance is represented as a score, which is calculated by performing permutations on each variable in the test data for each tree and re-running the test data through the trees.
+More specifically, consider a single variable.
+The values of this variable in the out-of-bag sample (the test data that the random forest algorithm partitions before building each tree) are randomly permuted (reordered) for each tree.
+The purpose of permuting is to mimic the absence of that variable.
 The trees in the random forest are then run again.
 Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted to the original random forest.
-If the accuracy of the premuted forest is much less than the original forest, then that variable is important.
-Importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the random forest [@Breiman2005].
-
-The random forest variable importance provides valuable insight because it addresses the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables.
-It handles uncorrelated variables well if the samples for the trees are subsampled without replacement [@Strobl2007].
-However, a severe disadvantage of random forest variable importance is that it is not reliable for variable selection where potential predictor variables vary in their scale of measurement or their number of categories.
-The importance of correlated predictors is overestimated and the algorithm tends to select variables that have many possible splits or many missing values [@Strobl2007].
+If the accuracy of the premuted forest is much less than the original forest, then that variable was important.
+Permutation importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the random forest [@Breiman2005].
+
+The random forest variable importance provides valuable insight as
+it addresses the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables [@Strobl2007].
+However, a severe disadvantage of random forest variable importance is that it is not reliable for variable selection where explanatory variables vary in their scale of measurement or their count of categories.
+Also, the importance of correlated predictors is overestimated and the algorithm tends to favour variables that have many possible splits or many missing values [@Strobl2007].
+The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably, however this does not necessarily indicate the variable impacts the response variable most.
 Overall, the random forest permutation importance is useful as it is able to compare variables in a machine learning environment and can include multivariate interactions, but the current random forest importance measure overestimates the importance of variables with many categories and some numeric variables.
@@
 
-Cforests are an alterantive ensemble tree method to random forests that overcome the aforementioned shortfalls of random forests such as overestimating the importance of variables with many categories and numeric variables.
-It is built from ctrees that are based on a conditional inference framework.
-The key difference between ctrees and standard decision trees are that a significance test procedure for splitting instead of an purity meausre such as the Gini coefficient, which is centered around information gain.
-
-The details behind the significance test procedure of a ctree are as follows.
-Each variable is permuted in every possible way, and a correlation value is calucated between the tested variable and target variable, for each permutation.
+Cforests are an alterantive ensemble tree method to random forests that overcome the aforementioned shortfalls of random forest variable importance measures.
+It is built from ctrees which are decision trees based on a conditional inference framework.
+The key difference between ctrees and standard decision trees is that a significance test is used for splitting instead of a purity measure such as the Gini coefficient, which is centered around information gain.
+
+The procedure behind the significance test at each split in a ctree is as follows.
+To determine which explanatory variable should be used at each split, each variable is permuted in every possible way, and a correlation value is calculated between the tested variable and response variable, for each permutation.
 The unchanged variable correlation is then compared with the correlation values for all permutations of that variable.
-From this, a p-value for the true correlation value can be calculated.
-The predictor variable with the lowest p-value is selected as the splitting variable [@Hothorn2006].
-Using an ensemble forest of these trees, permutation is then calculated in the same fashion as random forests, via permuting a variable, re-running the forest and comparing the decrease in accuracy.
-Several sources recommend cforests over other machine learning variable importance methods due to its unbiased qualities [@Strobl2007][@Strobl2009][@Hothorn2006].
+From this, a p-value for the true correlation value compared to the permuted correlation values can be calculated.
+The predictor variable with the lowest p-value is then selected as the splitting variable [@Hothorn2006].
+Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as random forests, via permuting a variable, re-running the forest and comparing the decrease in response variable accuracy.
+Several sources recommend cforests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as random forests (as discussed) as well as boosted trees [@Strobl2007][@Strobl2009][@Hothorn2006].
 
@@
 Limiting the predictive model to a concise set of meaningful variables reduces noise and improves predictions.
-Less variables means that a simpler model is being used for prediction which is easier to understand for stakeholders [@Weisberg2005].
-For these reasons, a subset of the most important variables were chosen before modelling began.
-This entailed first eliminating outliers, then comparing important variables from ANOVA as well as random forests and cforests.
-The literature implies that cforests would suit the data for this project as it is unbiased and not limited to linear theory [@Strobl2007].
+Less variables means that a simpler model is being used for prediction which in turn is easier for stakeholders to understand [@Weisberg2005].
+For these reasons, a subset of the most important variables were chosen before modelling which entailed first eliminating outliers, then comparing important variables from ANOVA as well as random forests and cforests.
+It can be inferred from the literature that cforests would suit the case study data best as it is unbiased and not limited to linear theory [@Strobl2007].
 
@@
 
-Previous studies predicting project profitability in software and construction projects were largely limited to case-based reasoning, regression and neural networks.
-There is a need to test more sophisticated machine learning algorithms that are as powerful as neural networks but provide insight into the reasoning behind predictions.
+Previous studies predicting project profitability in software and construction projects were predominantly limited to case-based reasoning, regression and neural networks.
+There is a need to test other sophisticated machine learning algorithms that are as powerful as neural networks but provide insight into the reasoning behind predictions.
 Ensemble trees (boosted trees and random forests) and bayesian networks fit these criteria.
 Regression and naiive bayes models were also included as simple baseline models.
-Complex models should be measured against simple models that can be computed at a fraction of the cost.
-The complete list of models for this study are:
-
+Complex models should be measured against simple models that can be built at a fraction of the computational cost.
+The complete list of models tested in this study are:
+
+* Regression - baseline model
+* Naiive bayes - baseline model
 * Bayesian network
@@
 * Gradient boosted trees
-* Regression - baseline model
-* Naiive bayes - baseline model
 
@@
 Between four and six clusters were chosen and summarised to find the maximum and minimum values within each cluster.
-Then, each job was assigned a cluster label to replace a numeric variable.
-It was decided that for timespans, the discretised variable should be applied to all models because it is easier for a manager to predict a time category than the number of days.
+For example, the diagram below illustrates the hierarchical dendrogram for timespan with 6 clusters boxed.
+
+![](images/method/dendrogram.png)
+
+The number of clusters was determined both visually from the dendrogram and experimentally.
+Visually, the height of the dendrogram 'branches' indicates the change in 'tightness' of the data points to their cluster centroids as the number of clusters increases.
+Tightness in the case of Ward's method is the sum of squared distances of each data point to the centroid of its respective cluster (ESS).
+For example, the height of the top horizontal bar indicates the ESS for one cluster and the height of the second from top horizontal bar is the ESS for two clusters.
+The hieght of the vertical branch between the two bars indicates the improvement in ESS by adding a cluster [@Putler2012].
+As more clusters are added, decreases in ESS are less significant and the aim is to choose the number of clusters at an optimal level in this process.
+After initial comparisons of prediction methods had been trialled, the number of clusters for each discretised variable was re-assessed by trialling one more or one less cluster and comparing model output.
+
+Once the number of clusters had been decided, each case was assigned a cluster label to replace its numeric variable.
+It was decided that for timespans and invoiced amounts, the discretised variables should be applied to all models (not just bayesian networks) because manager's must guess these values initially when operating the model.
+It is easier for a manager to predict a timespan category than the exact number of days a project will last.
 For example, a small job could be confidently assigned to less than three weeks and a large job could be assigned to 1.5 - 3 years.
 
-All methods except Gradient Boosted trees could not handle missing values.
-Preliminary runs of each method used subsets of the data without missing values.
-Then, imputing methods were trialled with gradient boosted trees since boosted trees can handle missing values.
-The MICE random forest method has been proven to work well with complex data sets [@Shah2014].
-It works by first performing a standard random forest imputation of missing values which are treated as 'place holders'.
-Then, the imputations for one variable are deleted and the remaining full variables are used to impute the single variable using random forest imputation.
+The revised discretised categories for timespan and total amount invoiced were as follows:
+
+| Timespan Categories  | Total Invoiced Amount Categories |
+|----------------------|----------------------------------|
+| 1 day - 3 weeks      | \$100 - \$600                    |
+| 3 weeks - 2.5 months | \$600 - \$2,500                  |
+| 2.5 - 9 months       | \$2,500 - \$8,000                |
+| 9 months - 1.5 years | \$8,000 - \$60,000               |
+| 1.5 - 3 years        | \$60,000 - \$1.8m                |
+| more than 3 years    |                                  |
+
+### Missing Data Imputation
+
+All methods except gradient boosted trees could not handle missing data.
+Therefore, preliminary runs of each method used subsets of the data which had complete data.
+Then, imputation methods were trialled against gradient boosted tree results since boosted trees can handle missing values and should perform equally well with imputed data.
+
+The MICE random forest method was chosen because it has been proven to work well with complex data sets [@Shah2014].
+It first performs a standard random forest imputation of missing values which are treated as initial 'place holders'.
+Random forest imputation is a complex process which works by first doing a rough imputation of the missing values.
+Then a ten-tree forest is run to calculate proximities between all cases.
+Proximities between two cases represent the number of trees in a forest where the two cases were assigned to the same end node in a tree.
+Once proximities are calculated, missing values for a variable (say variable m) are filled by an average of the other complete m values weighted by the proximity of the case representing each m value.
+
+After the initial random forest imputation is complete, the MICE process continues.
+The imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable again using random forest imputation.
 This is repeated for each variable and their 'place-holder' imputations are replaced by an imputation targeted at the single variable.
 This cycle is repeated five times by default [@Azur2011].
-If the imputed dataset gave similar predictive results using boosted trees as unimputed data using boosted trees, the imputed data must be reasonable.
-It is then trialled on the remaining methods and compared to un-imputed trials.
-
-## Model Comparison, Model Averaging
-
-To compare the models, mean squared error statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used for binary classification.
-AUC is a better statistic than classification accuracy in comparing algorithms when the probability of the outcome is relevant to the problem.
-This is true for problems where the decision can depend on whether the probability is above a certain threshold [@Huang2005].
-
-An ROC graph plots the false positive rate on the x-axis and true positive rate on the y-axis where:
+
+The final imputed dataset was fed into a boosted tree algorithm and if similar predictive results were obtained using boosted trees imputed data and unimputed data, the imputed data must be reasonable.
+The imputed dataset was then trialled on the remaining methods and compared to un-imputed trials.
+
+## Model Comparison
+
+To compare the models, root mean squared error (RMSE) statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to measure binary classification.
+For binary classification, AUC is a more useful statistic than classification accuracy when output is available as a probability which can be applied to the problem.
+This is pertinent when it is useful to rank the outcomes and trial probability thresholds that will optimise the rates of true positives and true negatives [@Huang2005].
+Note the 'probability' outputs from the models are not true statistical probabilities, but a score the model has assigned to a case using its own measure of certainty.
+
+An ROC graph visualises the curve from which an AUC score is calculated.
+Its two axes are the false positive rate (FPR) on the x-axis and true positive rate (TPR) on the y-axis.
+In this problem a 'positive' is a loss-making job so a true positive (TP) is a case when the model predicts a job will be loss making, and it did indeed lose money.
+A false positive (FP) is when the model predicts a job will be loss making but in reality it was profitable.
+These values are calculated as follows:
 
@@
 
-After a model has been created using a training set, the test set is run through the model, but in order to classify the test set, a probability threshold must be chosen.
-That is, the model outputs a probability to classify a case as 0 or 1.
-The probability is not a true statistical probability, but instead a measure of the  certainty of the model.
-If a probability threshold is chosen to be 0.6, each case in the test set can then be classified as 0 or 1 depending on whether the probability calculated by the model is greater or less than 0.6.
+Only predictions made on the test data should be entered into the equations above, therefore after a model must be first created using a training set, then the test data is input.
+From the test data, the model outputs probability values between 0 and 1 for each case, which cannot be entered into the equations above.
+In order to classify each case in the test set as 0 or 1, a probability threshold must be chosen.
+That is, if a probability threshold is chosen to be 0.6, each case in the test set can then be classified as 0 or 1 depending on whether the probability calculated by the model is greater or less than 0.6.
 Once the classifications have been made, the true positive rate and false positive rate can be calculated for that model and threshold.
 This is plotted as a single point on the ROC curve, say point C.
-Refer below:
+Refer the diagram below:
 
@@
 It is 'dumbly' classifying all cases as positive.
-if a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previosu example.
-There is an 80% chance any case will classified positive.
+if a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previous example.
+There is an 80% chance *any* case will be classified as positive.
 Therefore, models that lie on the diagonal line classify cases as well as random chance.
 
-To plot the curve in an ROC, points are plotted for results on the training set using numerous threshold probability values between 0 and 1.
+To plot the curve in an ROC, a series of points and their coordinates are first calculated.
+This is done by choosing numerous threshold probability values between 0 and 1, then applying each threshold point to the probability outcomes in the test set, and calculating the TPR and FPR for that threshold point.
 Once enough points are plotted, a curve can be drawn.
-The closer the curve reaches the top left hand corner, the closer the algorithm is to perfectly predicting positive and negative cases.
-This would give an area under the curve (AUC) of 1.
-AUC of between 0.5 and 1 means the model is performing better than random chance for all possible probability thresholds.
+The closer the curve reaches toward the top left hand corner, the closer the algorithm is to perfectly predicting positive and negative cases at some optimal threshold.
+This would give an AUC of 1.
+An AUC of between 0.5 and 1 means the model is performing better than random chance (as previously stated the diagonal line represents a model randomly assigning positives to all cases at a certain rate).
 
@@
 
-In order to compare whether models perform significantly better than one another, the AUC or mean squared error results from multiple-fold cross validations were used as the data points.
-Once the best two or three models were determined, these were combined using model averaging.
-Finally, the best model is determined between the single models and model averaged model.
-
-## Bottom Line Translation
-
-Although predictive power of the final model is important, for the method to be integrated into a company's decisions, the affect on a business' bottom line should be tested.
-This was done using a profit curve - a chart that plots the amount of profit the company earns on the y axis vs the probability threshold on the x axis.
+In order to compare which models performed significantly better than others, an adequate sample size of results statistics was required.
+Multiple models could be made by using different data in the training vs.
+testing sets, each providing a resulting test statistic (RMSE or AUC).
+5 fold cross validation was used, which created 5 models, then the division of 5 folds was repeated using a different random seed to create another set of 5 training/testing data sets.
+Initially, 20 models of each method were created in this fashion.
+Then a two-sample power calculation was run using the two sets of 20 results to determine the sample size to achieve a statistical power of 0.8.
+Once the number of required samples was determined, more models were built via 5 fold cross validation to obtain additional results statistics.
+The highest performing models were advanced to the next stage where various methods of combining, or blending, model results were tested.
+
+## Model Blending
+
+- simple averaigng, log reg, FWLS, boosted trees random forests
+- repeat 20 times - power - create more samples
+- carry all forward for bottom line translation anyway
+
+## Profit Curve
+
+Although predictive power of the final model is important, for the method to be integrated into a company's decisions, the affect on a business' bottom line was tested.
+This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs.
+the probability threshold on the x-axis.
 
 A simple approach was taken for this analysis, where projects with a probability to be a loss-making job greater than the threshold were rejected entirely.
-Therefore all profits and losses from jobs above threshold were discounted.
-If the threshold is zero, all jobs are rejected and the profit is $0.
-If the threshold is 1.0, all jobs are accepted and the profit is the same as the profit the company actually experienced since the data is a sample of historic projects.
-The aim was to find the optimal threshold point where saying no to a job above that level will mean result in higher profits, becasue jobs that are likely to make a loss are being rejected.
-This chart will clarify what percentage of profit increase the company can expect by integrating the algorithm into decision making.
-
-A profit curve is made from a single train/test instance of the data.
-In order to understand the uncertainty around the profit curve, hundred-fold train/test splits were run and results calculated for each model and threshold point.
-This resulted in a statistical confidence interval around each threshold point which can be represented graphically.
-The final profit curve and surrounding confidence interval will indicate whether the algorithm makes a compelling case for integration into the company's decision making process.
+Therefore all profits and losses from jobs above the threshold were discounted.
+If the threshold was zero, all jobs were rejected and the profit would be $0.
+If the threshold was 1.0, all jobs were accepted and the profit would be the same as the profit the company actually experienced since the data is a sample of historic projects.
+The aim was to find the optimal threshold point where saying 'no' to a job above that level would result in higher profits, because jobs that were likely to make a loss were being rejected.
+This chart will clarified what percentage of profit increase the company could expect by integrating the algorithm into decision making.
+It also provides a clear way to implement the algorithm and promote industry uptake of the research.
+
+Finally, since a profit curve is made from a single training/testing instance of the data, the curve varies with different divisions of the data.
+Therefore, in order to understand the uncertainty around the profit curve and to determine which blended or individual model performed statistically better than others, a large sample size of curve was required.
+Again, 20 curves of each method were built and a power calculation was performed to achieve a power of 0.8.
+This determined the required sample size which was achieved by repeating 5-fold training/testing splits.
+This enabled the highest points of each curve to be statistically compared to one another.
+A 95% confidence interval also be determined around the highest point on each curve.
+The final expected increase in profit and the percentage of projects to be rejected presented a clear scenario that the case study business manager's could assess in terms of their business strategy.
 

### 2016-06-21 finalwriteup/draft.rmd
finished blended model method

@@
 RMSE - root mean squared error
+FWLS - feature weighted linear stacking
 
@@
 Once the number of required samples was determined, more models were built via 5 fold cross validation to obtain additional results statistics.
-The highest performing models were advanced to the next stage where various methods of combining, or blending, model results were tested.
+The highest performing models were advanced to the next stage where various methods of combining, or blending, methods were tested.
 
@@
 
-- simple averaigng, log reg, FWLS, boosted trees random forests
-- repeat 20 times - power - create more samples
-- carry all forward for bottom line translation anyway
+Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009].
+In this research the term 'blending' is used to avoid confusion between ensemble tree methods.
+The idea behind model blending is to combine predictions from several methods with different theoretical foundations so that the strengths of each method can be combined.
+Historically, this was called model averaging in statistics and since the 1990's research showed that averaging results from different methods provided better predictive accuracy than any single model [@Madigan1994].
+New methods of model blending were developed for the Netflix Prize that applied sophisiticated machine learning techniques to big, messy data.
+The prize-winning solution was a complex blend of sub-blends that incorporated results of individual models as variables along with original variables.
+When original variables are used in blending they are called meta-features [@Sill2009].
+The multiple layers of blending gave incrememntal improvements in predictive accuracy, but the computational cost of the complex layers did not justify benefits for Netflix in practice and a simplified model was adopted.
+Therefore, in this project a single layer of blending was trialled.
+
+Five blending methods, ranging from simple to complex were tested on the top performing individual models.
+These included simple averaging of the individual model results, building a regression model using the individual model results only, feature weighted linear stacking (FWLS), random forests, and boosted trees.
+The first two models are simply the average or weighted average of the three individual models.
+The last three methods take advantage of potentially meaningful interactions between an individual model's results and meta-features (the original variables).
+For example, if the regression predicted profitability better than the other models for Water projects, the model would take advantage of this interaction.
+Regression predictions would be weighted higher than the other model predictions for Water projects.
+
+A simplified explanation of FWLS is as a linear regression where meta-features as well as model predictions from individual models are included as variables.
+Then, each meta-feature is interacted with each column of model results [@Sill2009].
+A linear regression is performed to weight each term's contribution to the final prediction.
+Random forests and boosted trees are also used to blend meta-features with output from several models.
+Feature interaction is performed passively due to the nature of how trees are built.
+A split in a node is conditional upon the preceding split based on another variable and so on.
+
+The five blended models were built on training and testing sets multiple times in a similar fashion to the individual models.
+A single training set for a blended model must be built from the test results of the individual models.
+Therefore, five fold cross validation was performed using each individual model to create a full dataset of test results.
+This was then added to the original dataset, and training and testing data could be partitioned to test the blended models.
+A maximum of five blended models were created from a complete set of test results, then a new set of test results were created for the next five blended models.
+This ensured a good mix of training and test data was employed.
+
+To begin, 20 models of each blending method were created.
+Then, a two-sample power calculation was performed to achieve statistical power of 0.8.
+The required number of samples (models) were then developed so that the highest performing blended model could be tested against the others for statistical significance.
+The next step was to assess the blended models' performance on overall profits for the company.
+This analysis differed from predicitve model accuracy, so several blended models as well as individual models were carried forward for this analysis.
 
@@
 
-Although predictive power of the final model is important, for the method to be integrated into a company's decisions, the affect on a business' bottom line was tested.
+Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the affect on a business' bottom line was tested.
 This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs.

### 2016-06-21 finalwriteup/draft.rmd
finished var select linreg

--- 2016-06-20 finalwriteup/draft.rmd
@@
 
-- simple averaigng, log reg, FWLS, boosted trees random forests
-- repeat 20 times - power - create more samples
-- carry all forward for bottom line translation anyway
+Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009].
+In this document, the term 'blending' will be used to avoid confusion with ensemble tree methods.
+The idea behind model blending is to combine predictions from several models derived from different methods that have different theoretical foundations.
+In this way, the strengths of each method can be combined.
+Historically, in statistics this was called model averaging and since the 1990's research has shown that averaging results from different methods provides better predictive accuracy than any single model [@Madigan1994].
+New methods of model blending were developed for the Netflix Prize that applied sophisiticated machine learning techniques to big, messy data.
+The prize-winning solution was a complex blend of sub-blends which interacted results of individual models with the original variables.
+When original variables are used in blending they are called meta-features [@Sill2009].
+The multiple layers of blending gave incremental improvements in predictive accuracy, but the computational cost of all the complex layers did not justify the benefits for Netflix in practice and a simplified model was adopted.
+Therefore, in this project just a single layer of blending was trialled.
+
+Five blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
+These included simple averaging of the individual model results, building a regression model using the individual model results only, feature weighted linear stacking (FWLS), random forests, and boosted trees.
+The first two models are simply the average or weighted average of the three individual model predictions.
+The last three methods take advantage of potentially meaningful interactions between an individual model's predictions and meta-features (the original variables).
+For example, if the boosted tree model predicted profitability better than the other models for Water projects, the blending model would take advantage of this interaction.
+Boosted tree predictions would be weighted higher than the other model predictions for Water projects.
+
+A simplified explanation of FWLS is as a linear regression where meta-features as well as model predictions from individual models are included as explanatory variables.
+Then, each meta-feature is interacted with each set of model results [@Sill2009].
+A linear regression is performed to weight each term's contribution to the final predictive accuracy.
+Random forests and boosted trees were also used to blend meta-features with output from the best models.
+Feature interaction is performed passively due to the nature of how trees are built.
+A split in a node determined by one variable is conditional upon the preceding split which was based on another variable and so on.
+
+The five blended model types were built on training and testing data sets multiple times in a similar fashion to the individual models.
+A single training set for a blended model must be built from the test results of the individual models.
+Therefore, five fold cross validation was performed using each individual model to create a full dataset as a compilation of test data results.
+The complete test data predictions were then added to the complete original dataset, and training and testing data could be partitioned to test the blended models.
+A maximum of five blended models were created from a complete set of test results (built from the individual models), then a new set of test results were created for the next five blended models.
+This ensured the blended models were trained on a thorough mix of the complete data set.
+
+20 models of each blending method were initially created.
+Then, a two-sample power calculation was performed to calculate the number of samples required to achieve statistical power of 0.8.
+The specified number of models were then developed so that the highest performing blended model could be tested against the others for statistical significance.
+The next step was to assess the blended models' performance in terms of improving overall profits for the company.
+This analysis differed from accuracy in predicting profitability, so several blended models as well as individual models were carried forward for this analysis.
 
@@
 
-This section presents key charts and tables of results that progressively answer the research question: can a statistical or machine learning model based on historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
+This chapter presents key charts and tables of results that progressively answer the research question: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
 
@@
 
-To create the best possible model, only important variables were selected to reduce the affects of noisy independent variables.
-There are currently 46 independent variables for the single dependent variable, return per dollar which is the measure of profitability for a project.
-
-### Regression
-
-Regression requires complete datasets.
-Therefore, analysis was done with a core set of 13 complete variables and a single additional incomplete variable.
-That means a model was made for each incomplete variable, resulting in 21 separate ANOVA models.
-The p-values for the F statistic of each variable is plotted in the boxplot below.
+To create the best possible model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
+Initially there were 34 explanatory variables for the single response variable, return per dollar, i.e.
+the chosen measure of profitability for a project.
+By the end of variable selection, only 11 explanatory variables remained.
+As stated in the previous chapter, three predictive methods were selected which output variable importance scores which differ due to their unique theoretical foundations.
+These were linear regression, random forests, and cforests.
+
+### Linear Regression
+
+Linear Regression must be fed complete datasets but the case study data set only had a core set of 13 complete variables (including the response variable).
+If an incomplete variable was added to the core variables, the 'complete' data set would shrink to the size of the incomplete variable.
+If another incomplete variable was added, the complete set of data would shrink again.
+Therefore, to make the most of the available data, each incomplete variable was added to the core variables one at a time, where a linear regression model was built for each.
+That meant a model was made for each incomplete variable, resulting in 21 separate ANOVA models.
+The p-values for the F statistic of each variable is plotted below.
+The core variables recieved p-values in all 21 models, while 'add.variable' represents p-values for the additional incomplete variable in each model.
 
@@
         labs(y = "P value of F statistic", x = "Independent Variables",
-             title = "Summary of P-values from Anova tests")
-
-```
-
-The ANOVA model indicates the following variables have a median p-value below 0.05 and therefore significantly contribute to the rejection of the null hypothesis - the variables do not correlate linearly with the dependent variable, return per dollar.
-
-* number of employees that contributed to the project (no.users)
+             title = "Summary of F-statistic P-values from ANOVA models")
+
+```
+
+The ANOVA models indicated that the following variables have a median p-value below 0.05 and therefore significantly contribute to the rejection of the null hypothesis:
+
+* number of employees on the project team (no.users)
 * Percent of hours completed by a professional as opposed to technical employee (pc.pro)
-* Business category of Client (Business) 
-* Discipline
+* Business category of the client (Business) 
+* Internal discipline (Discipline)
 * Number of internal disciplines involved in the project (Num.disc)
-* Timespan of the project (timespan)
+* Timespan of the project from first hours to final hours entered in timesheets (timespan)
 * Total amount invoiced for the project (inv.mlsto)
 
-Four of the added incomplete variables (categorised as add.variable) were significant in their ANOVA models:
+Four of the added incomplete variables (represented as 'add.variable') had F-statistic p-values below 0.05 in their ANOVA models:
 
 * Client id (code.client)
-* Project category (JD.Second)
-* Position of the main professional working on the project, ie mid-level, senior (ProjEng.Pos)
-* id of main professional working on the project (code.ProjEng)
-
-### Random Forest
+* Internal project category (JD.Second)
+* Position of the main professional working on the project, ie mid-level, senior etc. (ProjEng.Pos)
+* Id of the main professional working on the project (code.ProjEng)
+
+The variables listed above had p-values below 0.05, meaning the null hypothesis, that the continuous explanatory variables have no linear relationship with the reponse variable or that categorical variables have the same mean response values for each group, could be rejected.
+The important variable relationships found in linear ANOVA regression shall be compared to output from random forests and cforests below.
+
+### Random Forests
 
@@
 * Project category (JD.Second) - ANOVA and cForest
-* Total Amount Invoiced from Past Client Jobs (clint.totinv) - randomForest
+* Total Amount Invoiced from Past Client Jobs (client.totinv) - randomForest
 * Discipline - ANOVA

### 2016-06-22 finalwriteup/draft.rmd
finished reviewing results to blended and added html

@@
 output:
+  html_document: default
   word_document:
     reference_docx: mystyles.docx
-  html_document: default
 bibliography: Library.bib
@@
 
-plot.pvals = readRDS("report_data/AOV_varimp_plot.rds")
+plot.pvals = readRDS("AOV_varimp_plot.rds")
 
@@
 
-The random forest algorithm imputes missing values, however this was not desirable for the early stage of variable selection - before modelling had commenced.
-Instead, a subset of complete data was used made up of variables with at least 2300 complete cases.
-Additionally, random forests cannot handle categorical variables with more than 53 categories which eliminated more variables.
-The resulting 'core' dataset contained 16 variables and 2364 cases.
-Because each run of the algoritm is slightly different, 10 algorithms were run, which each output the ranking of variables by importance.
-The mean ranking of each variable is summarised below:
+The randomForest algorithm can process datasets with missing values, however it does so by automatically imputing missing values [@randomForest].
+This was not desirable in the early stage of variable selection.
+Instead, the completeness of the variables were assessed and it was decided that all variables with at least 2300 overlapping complete cases would be chosen.
+Additionally, the random forest algorithm cannot process categorical variables with more than 53 categories.
+This eliminated more variables and the resulting 'core' dataset contained 15 variables and 2364 cases.
+Because of the random forest algorithm's known biases and limitations with numbers of categories, a briefer approach to variable importance analysis was taken in comparison to regression and cforests.
+
+Each tree in a random forest is created with a bootstrapped training sample of the data.
+The random sampling means the results of each forest are slightly different.
+To cater for the variation in results, 10 forests were run and the mean rank of each variable was recorded.
+Variable importance output from a single forest is simply a ranked list of the variables where rank 1 indicates the most important variable:
 
@@
 
+#dataset created in var_select.rmd
 rf.rank = readRDS("RF_var_rank_subbed.rds")
@@
 
-Cforests were also run with a core of 13 complete variables, and incomplete variables were given added to the core variables individually in separate runs.
-Variables with an unlimited number of c ategories can be included in cforests which is an advantage over randomForests.
-Cforests compute a variable importance ranking which is a number relating to the reduction in error the variable provides.
-The variable importance rankings for the core 13 variables are as follows:
-
-```{r, echo = FALSE}
+Cforests were also run with the core 15 complete variables with 2364 complete cases.
+Incomplete variables were added to the core variables individually, each with their own separate run.
+Variables with an unlimited number of categories can be included in the cForest function which is an advantage over randomForests [@Hothorn2006; Strobl2007; Strobl2008].
+Cforests compute a variable importance ranking which is a number relating to the reduction in error the variable provides (similar method to the permutation importance in random forests).
+The variable importance rankings for the core 15 variables are as follows:
+
+```{r, echo = FALSE, fig.depth = 5}
 
 cfor.varimp = readRDS("cfor_varimp.rds")
+# how did pc.director sneak in there?
+cfor.varimp = cfor.varimp %>% filter(! var == "pc.director")
 
@@
         geom_bar(stat='identity') +
-        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance rankings of Complete Variables")
-
-```
-
-Some incomplete variables that were added one at a time ranked in the top 6 most important variables in their model:
+        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance Rankings of Complete Variables")
+
+```
+
+Four incomplete variables that were added indpendently ranked in the top 6 most important variables in their model:
 
@@
 
+Once several methods of variable importance analysis have been completed, the separate results were rationalised and final variables were selected for the predictive models.
+
 ### Variable Selection Summary
 
-Based on the results of the three methods for variable importance, 11 independent variables were selected for the prediction modelling:
-
-* Percent of hours completed by a professional level employee (pc.pro) -all models
-* Project timespan (timespan) - all models
-* Number of employees on Project (no.users) - all models
-* Amount Invoiced for Project (inv.mlsto) - all models
-* Business category of Client (Business) - moderately for all models
-* Project category (JD.Second) - ANOVA and cForest
-* Total Amount Invoiced from Past Client Jobs (client.totinv) - randomForest
-* Discipline - ANOVA
-* position of main employee on the project (majority.pos) - newly created to cater for findings in ANOVA
-* Percent of Hours completed by the main employee on the project (pc.majpos) - cForest
-* Billing Type (hourly vs. fixed fee) - advice of Case Study partner
+Based on the results of the three methods for variable importance, 11 independent variables were selected for predictive modelling.
+They are listed in the table below along with the methods that highlighted each variable as important (randomForest = RF, cForest = CF):
+
+| Variable Description                                                 | ANOVA | RF | CF |
+|----------------------------------------------------------------------|-------|----|----|
+| % of Hours Completed by a Professional-level Employee                | x     | x  | x  |
+| Timespan                                                             | x     | x  | x  |
+| Number of Employees on Project                                       | x     | x  | x  |
+| Amount Invoiced for Project                                          | x     | x  | x  |
+| Business Category of Client                                          | x     | x  | x  |
+| Project Category                                                     | x     |    | x  |
+| Total Amount Invoiced from the Client - Past Jobs                    |       | x  |    |
+| Discipline                                                           | x     |    |    |
+| Position of Main Project Employee - new variable from ANOVA findings | x     |    |    |
+| % of Hours by Main Project Employee                                  |       |    | x  |
+| Billing Type (requested by case study company)                       |       |    |    |
+
+The variables above were used to build the predictive models and the results are summarised in the following two sections.
 
@@
 
-Once the independent variables were narrowed down to a succinct list, the prediction models were tested.
-First prediction return per dollar as a regression problem was trialled followed by prediction profit or loss as a binary classification problem.
-
-### Regression Model
+Once the explanatory variables were narrowed down to a succinct list, the predictive models were built and tested.
+First, prediction of return per dollar as a regression problem was attempted followed by prediction of profit or loss as a binary classification problem.
+linear regression, random forests and boosted tree methods were applied to both problems and had varying levels of success.
+
+### Regression Models
 
@@
 
-In ANOVA, linear relationships are tested and normal distributions are assumed for numeric variables.
-Therefore, where appropriate, numeric variables were normalised via log or cube root transformations.
+In ANOVA, categorical and continuous explanatory variables are treated differently.
+For categorical variables, differences in the mean values of the response variable in each group are tested.
+In contrast, linear relationships are examined between continuous explanatory variables and the response variable.
+Since linear relationships are being tested, continous variables are assumed to be normally distributed.
+Therefore, where appropriate, numeric variables were normalised via logarithmic or cube root transformations.
 Trials of various variable interactions revealed that interactions between:
 
-* total amount invoiced 
-* percent of job by professional level employee
-* total amount invoiced 
-* client id
-* project category 
-* number of employees on the job
+* total amount invoiced * percent of job by professional level employee
+* total amount invoiced * client id
+* project category * number of employees on the job
 
 were statistically significant.
-Two types of ANOVA trials were run.
+Two types of ANOVA models were run.
 The first used only a 'core' list of 6 variables that were largely complete and any relevant interaction terms.
-The second method was developed to get around the problem that when all variables are combined, only 15% of the projects have complete data.
-Therefore the ANOVA test would eliminate 85% of the data and imputing the data was not desirable at an early stage due to the amount of missing categorical data.
-The second method would choose a project (row in the dataset) randomly, and this case would have certain columns complete.
-The data was then filtered for all jobs that had at least the same columns complete, and the ANOVA test was run on this reduced set.
-This method allowed variables with a larger proportion of missing data to be analysed while not attempting to analyse all variables at once.
+The second method was developed to overcome the problem of missing data.
+When all variables were combined, only 15% of the project cases had complete data and therefore the ANOVA test would eliminate 85% of the data upon execution.
+Imputing the missing data was not desirable at that stage due to the volume of missing categorical data and an inability to ensure the imputed data was not hampering results.
+The second method was a more complicated procedure that required dozens of models but allowed variables with missing data to be included.
+First, a project was randomly selected (a row in the dataset) which would have certain variables complete.
+The data was then filtered for all projects that had at least the same columns complete, and an ANOVA model was built on this reduced data set.
+This method allowed variables with a larger proportion of missing data to be included while not attempting to analyse all variables at once.
 
@@
 
-To assess the results of the ANOVA models, the root mean squared error (RMSE) of the return per dollar predictions were compared against the RMSE if all predictions were simply the mean of return per dollar for all projects.
-Then, the RMSE's from model predictions were subtracted from RMSE's using mean return per dollar.
-If the models were effective, RMSE should be lower for model predictions, and the difference would be greater than 0.
-Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only.
+The results of the ANOVA models were assessed using the root mean squared error (RMSE) statistic of the return per dollar predictions.
+These were compared against a baseline RMSE which was the RMSE of the predicted return per dollar values against the mean return per dollar of all projects.
+Then, the RMSE from each model was subtracted from the RMSE using mean return per dollar.
+If the models were effective, the RMSE should be lower for ANOVA model predictions, so the difference would be greater than 0.
+Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only (method 1 using ANOVA).
 Note, only `r n.samples(data = results.A2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
@@
 
-A student t test indicates the difference is significantly above 0 with a p-value of `r t.test(results.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
+A student t-test indicates the difference is significantly above 0 with a p-value of `r t.test(results.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
 Therefore, the null hypothesis that the difference is less than or equal to zero can be rejected.
 In other words, the ANOVA model improved estimates of return per dollar over using the mean return per dollar to a statistically significant degree.
-However, the mean difference is only $`r mean(results.core$diff) %>% format(digits = 1)`.
-Furthermore, the mean error in return per dollar is $`r mean(results.core$final.RMSE) %>% format(digits = 2)` which is unacceptably high.
+However, the mean difference is only \$`r mean(results.core$diff) %>% format(digits = 1)`.
+Furthermore, the mean RMSE in return per dollar is \$`r mean(results.core$final.RMSE) %>% format(digits = 2)` which is unacceptably high.
+Almost 50 cents is a large mean error when the value is estimating the profit from \$1 spent.
 
@@
 
-A student t test indicates the difference is not significantly above 0 which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
+A student t-test indicates the difference is not significantly above 0 which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
 This method clearly performed poorly in comparison to the core variables.
 A reason for this could be that the mean number of cases in the dataset was `r results.core$test.nrow[1]*4` for the core variables but averaged only `r (mean(results.A2$test.nrow)*4) %>% round(0)` for the sampled set of all variables.
-Regression results from ANOVA models were poor - randomForest was trialled next.
-
-#### randomForest
+Regression results from ANOVA models were poor - the randomForest algorithm was trialled next.
+
+#### Random Forest
 
@@
 
-Numeric predictions of return per dollar were attempted first with core variables that were largely complete followed by a method of sampling a mix of core and less complete variables in every run.
-The randomForest parameters were first tuned using the caret package and found optimal values mtry = 5 and ntree = 500 [@caret].
-A density plot of 50 models of the core variables using different train/test sets is shown below.
-The x-axis measures the decrease in RMSE obtained by using the randomForest model over simply the mean of all project return per dollar values.
+Numeric predictions of return per dollar were attempted first with core variables that were largely complete followed by a method of sampling a mix of core and less complete variables in every run (similar to the two ANOVA methods).
+The randomForest parameters were first tuned using the caret package and optimal values mtry = 5 (the number of randomly selected explanatory variables to consider at each split) and ntree = 500 (the number of decision trees to ensemble) were determined [@caret].
+A density plot of 50 models of the core variables using different training/test sets is shown below.
+The plot was created using the same procedure used in the ANOVA models where the x-axis measures the RMSE obtained by using the randomForest model subtracted from the RMSE using the mean of all project return per dollar values.
 Note, only `r n.samples(data = results.rf2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
@@
 
-This plot shows a mean return per dollar RMSE improvement of $`r mean(results.core$diff) %>% format(digits = 1)` which is the same as the ANOVA model.
-The p-value for the null hypothesis that the difference is less than or equal to zero is `r t.test(results.rf.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)` and is rejected.
+This plot shows a mean return per dollar RMSE improvement of \$`r mean(results.core$diff) %>% format(digits = 1)` which is the same as the ANOVA model.
+The p-value for the null hypothesis that the difference is less than or equal to zero is `r t.test(results.rf.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)` and could be rejected.
 Results from 100 randomForest models using method 2 are shown below:
@@
 The p-value for these models and the afore-mentioned null hypothesis is `r t.test(results.rf2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
-The null hypothesis is rejected which contrasts the parallel ANOVA test.
-On the right hand side, of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar.
+The null hypothesis was rejected which contrasted the matching ANOVA test.
+On the right hand tail of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar on the left tail.
 A density plot of the RMSE for the models is shown below:
@@
 
-Although the randomForest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for return per dollar was wide and the mean was high ($`r mean(results.rf2$final.RMSE) %>% round(2)`).
-
-The two regression models predicting return per dollar resulted in high mean RMSE's ranging from $`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
-Because of the poor results, further development of regression models was postponed in favour of investiation of binary classification models.
-The classification is a simpler task than regression because it requires less precision.
-
-### Classification Models
-
-#### Initial Trial of five Methods
-
-To begin analysis, all prediction methods were trialled using as much data as possible for each method.
-Boosted trees and Naiive bayes algorithm are able to process data with missing values, so all data could be used.
-Logistic regression, randomForest and Bayesian Networks require complete datasets so a subset of complete data was used for these.
-It was expected that Boosted trees would perform best as they are powerful and can use incomplete data.
-For this reason, 4 x 5 fold runs of each method were performed (20 models total for each method) and statistically compared against boosted trees.
-A power calculation was used to calculate the number of runs required for a statistical power of 0.8.
-
-The models were compared using the Area under the curve (AUC) test statistic calculated from the Reciever Operating Curve (ROC) as this gives a good indication of model performance under binary classification [@pROC][@Putler2012].
-An example plot of five ROC's resulting from five fold cross validation models using the boosted tree algorithm is shown below:
+Although the randomForest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for return per dollar was wide and the mean RMSE was high (\$`r mean(results.rf2$final.RMSE) %>% round(2)`).
+
+The two regression models predicting return per dollar resulted in high mean RMSE's ranging from \$`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
+Because of the poor results, further development of regression models was deffered in favour of investiation of binary classification models.
+Binary classification is a simpler task than regression because it requires less precision.
+The challenge is how to gainfully apply the results to the business scenario.
+
+### Binary Classification Models
+
+The cost estimation problem analysed in the literature review had been converted to a profitability estimation problem and was finally converted to a binary profit/loss prediction.
+The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable (or not).
+The original return per dollar response variable was transformed by converting aLl 'return per dollar' values less than or equal to zero to 'loss making' jobs and 'return per dollar' values greater than 1 became 'profitable' jobs.
+The following section summarises results from the trialled binary classification methods.
+
+#### Initial Trial of Five Methods
+
+The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were intially trialled without imputing data, using the maximum amount of data possible depending on the method.
+Boosted Trees and Naiive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
+Logistic Regression, Random Forest and Bayesian Networks require datasets without missing values so a subset of complete data was used for these.
+It was expected that Boosted Trees would perform best as they are known to have powerful predictive capabilities and can use incomplete data.
+For this reason, 4 x 5-fold runs of each method were performed (20 models each) and statistically compared against Boosted Trees' results.
+A power calculation was then used to calculate the number of runs required for a statistical power of 0.8.
+
+The models were compared using the AUC test statistic calculated from the ROC as this gives a good indication of model performance under binary classification [@pROC][@Putler2012].
+An example plot of five ROC's built from 5-fold cross validation using the boosted tree algorithm is shown below.
+
+Before the plot could be made, the Boosted Tree algorithm parameters were tuned to optimise the AUC outcomes.
+The caret package was used for this and the tuned parameters are summarised below [@caret]:
+
+* shrinkage = 0.001
+* number of trees = 10,000
+* minimum terminal node size = 20
+* maximum tree depth = 5
 
@@
 
-* Boosted trees first tuned using caret package [@caret]:
-* shrinkage = 0.001
-* no trees = 10000
-* min terminal node size = 20
-* max tree depth = 5
-
-In this example, the average area under the curve (AUC) was 0.763 indicating that on the whole, taking into account all possible thresholds, the model is predicting profit/loss better than random chance (AUC = 0.5) but not perfectly (AUC = 1).
+In the plot above, the average AUC was 0.763 which indicated that altogether, taking into account a range of possible thresholds, the model was predicting profit/loss better than random chance (AUC = 0.5) but not perfectly (AUC = 1).
 20 models and their resulting AUC statistics from each method were initially determined.
@@
 The comparison of logistic regression required 100 models, while randomForest required over 700 because the AUC distributions were close.
-Bayesian Networks and Naiive Bayes did not perform as well and the power calculation revealed only 5 and 8  models were required to achieve a power of 0.8 respectively.
-Bayesian Networks were then dropped as Naiive Bayes perfomed slightly better and was fast and simple enough to work well as a baseline comparison to the higher performing methods.
-
-It was finally decided that randomForest and boosted trees would likely not be distinguishable since the power calculations indicate 700 models are required.
-hence 100 models of each method were created as required for logistic regression.
-10 fold cross validation was performed 10 times.
-The violin plot below summarises the AUC values from each method.
+Bayesian Networks and Naiive Bayes did not perform as well and the power calculation revealed only 5 and 8 models were required respectively.
+Bayesian Networks were then dropped as Naiive Bayes perfomed slightly better and were faster and simple enough to work as a baseline comparison to the higher-performing methods.
+
+It was finally decided that randomForest and Boosted Trees would likely not be distinguishable since the power calculation indicated that 700 models were required.
+Hence, 100 models of each method were built as calculated for logistic regression.
+10-fold cross validation was performed 10 times.
+The violin plot below summarises the AUC values produced by each method.
 The 'violins' are coloured according to whether the distributions significantly vary to boosted trees using a critical value of 0.05.
@@
 
-The performance of logistic regression and random forest algorithms cannot be statistically differntiated from boosted trees.
-The same process was repeated using an imputed dataset.
-Imputation was done using the MICE package and allowed all algorithms to be trialled with a full dataset.
+The AUC performance of Logistic Regression and the randomForest algorithms cannot be statistically differentiated from Boosted Trees.
+
+Next, data imputation methods were trialled which would make the complete data set available to Logistic Regression and randomForests.
+It was possible this would improve AUC results and the same process as above could be repeated.
+Imputation was done using the MICE method with randomForest imputation [@mice].
 Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for random forest unless over 450,000 models were made.
@@
 
+Boosted Trees, Logistic Regression, and randomForest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed boosted trees.
+The results from each method using reduced data sets (to cater for missing values) alongside results using imputed complete data are shown below.
+
 ```{r all_violin}
@@
 
-Imputed data tightens the distribution of randomForest substantially but slightly spreads the distribution of Boosted trees and logistic regression.
-Since the methods using imputed data have a similar performance, slightly tighter distributions of results and use more data, models using imputed data was used when developing blended models of all methods.
+Results using imputed data tightened the distribution of randomForest AUC output substantially but slightly increased the distribution spread of Boosted Trees and Logistic Regression.
+Since the methods using imputed data performed similarly, substantially tightened the distribution of randomForest results, and made use of more data, the imputed data models were propogated forward to the development of blended models.
 
@@
 # How to achieve statistical power of 0.8?
-finance.plot = readRDS("report_data/plot_finance_100.rds" )
+finance.plot = readRDS("plot_finance_100.rds" )
 fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
@@
 
-finance.plot = readRDS("report_data/plot_finance.rds" )
+finance.plot = readRDS("plot_finance.rds" )
 
@@
 Overall, this project has successfully demonstrated a blended statistical and machine learning model has the potential to aid manager's in rejecting loss-making projects to improve their business's profitability.
+
+# References

### 2016-06-22 finalwriteup/trial_code.rmd
finished reviewing results to blended and added html

@@
 * $return\ per\ dollar = \frac{project\ profit}{cost\ from\ hours}$
+
+* total amount invoiced * percent of job by professional level employee
+* total amount invoiced * client id
+* project category * number of employees on the job
+
+| Variable Description                                                 | ANOVA | RF | CF |
+|----------------------------------------------------------------------|-------|----|----|
+| % of Hours Completed by a Professional-level Employee                | x     | x  | x  |
+| Timespan                                                             | x     | x  | x  |
+| Number of Employees on Project                                       | x     | x  | x  |
+| Amount Invoiced for Project                                          | x     | x  | x  |
+| Business Category of Client                                          | x     | x  | x  |
+| Project Category                                                     | x     |    | x  |
+| Total Amount Invoiced from the Client - Past Jobs                    |       | x  |    |
+| Discipline                                                           | x     |    |    |
+| Position of Main Project Employee - new variable from ANOVA findings | x     |    |    |
+| % of Hours by Main Project Employee                                  |       |    | x  |
+| Billing Type (requested by case study company)                       |       |    |    |

### 2016-06-22 var_select.rmd
finished reviewing results to blended and added html

@@
 rand.rank$var <- sub("pc.pro", "% of Hours Completed by a Professional-level Employee", rand.rank$var)
+rand.rank$var <- sub("timespan", "Timespan", rand.rank$var)
 rand.rank$var <- sub("Business", "Business Category of Client", rand.rank$var)
@@
 rand.rank$var <- sub("Biz.type", "Broad Client Business Category (Gov, Private etc)", rand.rank$var)
-rand.rank = rename(rand.rank, Variable_Description = var, Rank = m.rank)
+rand.rank = rename(rand.rank, Variable_Description = var, Mean_Rank = m.rank)
+rand.rank = rename(rand.rank, Mean_Rank = Rank)
 

### 2016-06-23 finalwriteup/draft.rmd
almost finished results review

@@
 
-Five blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
-These included simple averaging of the individual model results, building a regression model using the individual model results only, feature weighted linear stacking (FWLS), random forests, and boosted trees.
-The first two models are simply the average or weighted average of the three individual model predictions.
+Six blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
+These included simple averaging of the individual model results, building a regression model using the individual model results only, a boosted tree model using individual model results only, feature weighted linear stacking (FWLS), random forests, and boosted trees.
+The first three models are simply the average, weighted average, of the three individual model predictions.
 The last three methods take advantage of potentially meaningful interactions between an individual model's predictions and meta-features (the original variables).
@@
 
-Blended models combine the results of each well-performing individual model to create averaged or 'blended' predictions.
-In this case, the randomForest, boosted trees, and logistic regression models were chosen for blending.
-The simplest way of combing models is to average the probability results (a number between 0 and 1) of all three models.
-Feature Weight Linear Stacking (FWLS), randomForest and boosted trees were all tested as blending methods where the predictions from each of the three single models becomes an additional variable in a new model.
-Each model blending method has its own method of weighting the outcomes from the individual models as a function of the values in the original model.
-For example, for a certain discipline in the case study company, randomForests may predict the outcome best so this is weighted higher for those cases.
-For this reason, a variable importance study was conducted again to simplify the models.
+Blended models combine the predictions from each higher-performing individual model to create averaged or 'blended' predictions.
+In this case, the randomForest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialled ranging from simple to complex.
+The simplest way was to average the probability output (a number between 0 and 1) of all three models.
+The next two simplest methods were to build a logistic regression and boosted tree from the three model outputs only.
+Feature Weighted Linear Stacking (FWLS), randomForest and boosted trees were also tested as blending methods where the predictions from each model became additional variables to the original explanatory variables (meta-features).
+The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature conditions could be weighted as such.
+Some meta-features may have been dispensable in the blending process, so in order to maintain simple comprehensible models, a variable importance study was conducted again.
 
@@
 
+Variable assessments were completed and compared using cForests, Logistic Regression, randomForests and Boosted Trees.
+Cforests have the favourable characteristic of being unbiased while the remaining methods were used for the blending process.
+The results are summarised in the sections below.
+
 *Variable Importance: cforest*
@@
 
-cfor.varimp = readRDS("cfor_blend_varimp.rds")
+cfor.varimp = readRDS("report_data/cfor_blend_varimp.rds")
 
@@
 
+The cForest highlighted the results from the Logistic Regression and Boosted Trees as the most important variables.
+This was followed by the results from the randomForest model and the project team size, project category, total invoiced amount (category) and timespan (category).
+
 *Variable Importance: Logistic Regression*
@@
 
-log.varimp = readRDS("logblend.rds")
+log.varimp = readRDS("report_data/logblend.rds")
 
@@
 
-Most significant variables (P value for coefficient < 0.05) are:
+The logistic regression found the F-statistics from the following variables were significant in rejecting the null hypothesis (p-value < 0.05):
 
@@
 * Discipline * results of boosted tree model
-* number of employees on a job * results of logistic regression model
+* team size for project * results of logistic regression model
 * Client Business type * results of boosted tree model
 
-Single variables that aren't in interaction terms:
+Single variables that were not also in interaction terms:
 
@@
 
-rf.varimp = readRDS("rfblend.rds")
+rf.varimp = readRDS("report_data/rfblend.rds")
 rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
@@
 
-Permutation importance by randomForest rates all three prediction models as most important followed by number of employees on a project, percent of hours done by the chief employee on the project, timespan, and project category.
-
-*Variable Importance: boosted trees*
+Permutation importance by randomForest rated all three prediction models as most important followed by team size for the project, percent of hours done by the chief employee on the project, timespan (category), project category, and client business type.
+
+*Variable Importance: Boosted Trees*
 
@@
 
-boost.varimp = readRDS("boostblend.rds")
+boost.varimp = readRDS("report_data/boostblend.rds")
 boost.varimp = add_rownames(boost.varimp)
@@
 ```
-Boosted trees ranks all three individual prediction models highest followed by Project category, Client Business type, and timespan.
+Boosted Trees ranked output from all three individual prediction models highest (Logistic Regression ranked highest) followed by Project category, Client Business type, and timespan (category).
 
@@
 
-Upon reviewing important variables ranked by each method of blended models, the following variables were deemed important by at least two methods and shall be used in the comparison of blended models:
-
-* all three results of prediction models, particularly logistic regression
-* project category
-* Client Business type
-* Total Amount invoiced for a project
-* Number of employees on a project
+Upon reviewing important variables ranked by cForest and each method of model blending, the following variables were deemed important by at least two methods and were used in the development of blended models:
+
+* All three results of prediction models, particularly logistic regression
+* Project category
+* Client business type
+* Total amount invoiced for a project (category)
+* Team size on a project
 * Timespan
@@
 
-Three model blending techniques were compared using the 8 variables determined above.
-These are FWLS, randomForest, and boosted trees.
-Furthermore, 3 simple model blending techniques were tested that used only the predictions of the three individual models as variables.
-These were averaging, a simple boosted tree, and a simple logistic regression.
-All six methods were compared against the original logistic regression predictions.
-Initially 100 train/test runs were performed for each method by running 20 iterations of 5-fold cross validation.
+Only three of the model blending techniques used the 8 variables determined above.
+These were FWLS, randomForest, and Boosted Trees.
+The remaining 3 techniques used only the predictions of the three individual models.
+All six methods were compared against the original logistic regression model.
+Initially 100 train/test runs were performed for each blending method by running 20 iterations of 5-fold cross validation.
 The AUC results of these iterations are displayed below:
 
-```{r, fig.height=10}
-
-auc.list <- readRDS("blend_auc1.rds")
+```{r, fig.height=8}
+
+auc.list <- readRDS("report_data/blend_auc1.rds")
 means = lapply(auc.list, mean) %>% unlist
@@
              title = "AUC Distribution from 100 Blended Models of each Method") +
-        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
-                         "log" = "Logistic Regression", 
-                         "average" = "Simple Average",
-                         "orig.log" = "Orignal Logistic Regression", 
-                         "rf" = "Random Forest",
-                         "simp.boost" = "Simple Boosted Trees",
-                         "simp.log" = "Simple Logistic Regression")) +
+        scale_x_discrete(labels = c("boost" = "BT Blend", 
+                         "log" = "FWLS", 
+                         "average" = "Simple Average Blend",
+                         "orig.log" = "Orignal LR", 
+                         "rf" = "RF Blend",
+                         "simp.boost" = "Simple BT Blend",
+                         "simp.log" = "Simple LR Blend")) +
         scale_fill_manual(values = c('navajowhite','sienna1'),
@@
 
-The plot shows that 4 methods could have a statistically significantly higher mean that the original logistic regression model: the simple blended logistic regression, the simple average, and the two boosted tree models.
-To achieve a power of 0.8, a two sample power analysis was performed on the 100 AUC results from each blended model against the AUC results from the original logistic regression.
+The above plot shows that 4 methods had a higher mean AUC than the original logistic regression model:
+
+* the simple blended logistic regression 
+* the simple average 
+* and the two boosted tree models.
+
+To achieve a power of 0.8, a two-sample power analysis was performed on the 100 AUC results from each blended model against the AUC results from the original logistic regression.
 The boosted models required over 1000 samples while the simple averaged and simple logistic regression models required only 120.
-Therefore, 50 more iterations of the simple averaged and simple logistic models were performed since the boosted models results were not as promising.
+Therefore, 50 more iterations of the simple averaged and simple logistic models were performed.
+The boosted tree models would not achieve a statistical power of 0.8 with only 150 samples when compared to the original logistic regression and were therefore dropped from the next stage of analysis.
 
@@
 #load results from 150 iterations
-auc.150 = readRDS("blend_auc150.rds")
+auc.150 = readRDS("report_data/blend_auc150.rds")
 
@@
 The results of the 150 iterations are shown in the plot below.
-Simple model averaging results in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logisitc regression blended model had a two-sample t.test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
+Simple model averaging resulted in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logisitc regression blended model had a two-sample t-test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
 
@@
         scale_x_discrete(labels = c( 
-                         "average" = "Simple Average",
-                         "orig.log" = "Orignal Logistic Regression",
-                         "simp.log" = "Simple Logistic Regression")) +
+                         "average" = "Simple Average Blend",
+                         "orig.log" = "Orignal LR",
+                         "simp.log" = "Simple LR Blend")) +
         scale_fill_manual(values = c('navajowhite','sienna1'),
@@
 
-Simple logistic regression and averaging the the results of the three individual models (logistic regression, boosted trees and random forests) give virtually the same distribution of results.
-As averaging the results of the models is simpler than logistic regression, this was adopted as the best model.
+The simple Logistic Regression blend and simple averaging of the individual models (logistic regression, boosted trees and random forests) gave virtually the same distribution of results, where both AUC distributions were signficantly greater than the original logistic regression model.
+The p-values from the two-sample t-tests were `r pvals$p_value[1]` and `r pvals$p_value[2]` for simple average and simple Logistic Regression methods respectively.
+It was encouraging that blended models produced higher AUC values than the original logistic regression model, however for the case study business, the real measure of success is the model's impact on the business' overall profits.
+It was not guaranteed higher AUC values directly related to improvements in profit so several methods, including original individual models as well as blended methods, were progressed to the profit analysis.
 

### 2016-06-23 financial.rmd
almost finished results review

@@
 
-profit <- function(df = all10mice, method = c("average","simp.log","orig.log","orig.rf","orig.boost"), 
+profit <- function(df = all10mice, method = c("average","simp.log","simp.boost","orig.log","orig.rf","orig.boost"), 
                    seed = 300, 
@@
 
+        saveRDS(df.blend, "C:/Users/n9232371/Documents/github/consult/dfblend.rds")
+
         saveit <- iter.auc(mult = max.iter/k.folds.2, x.folds.iter = k.folds.2,
-                           methods = c("orig.log","average","simp.log", "orig.rf","orig.boost"),
+                           methods = c("orig.log","average","simp.boost", "simp.log", "orig.rf","orig.boost"),
                            df.iter = df.blend)

### 2016-06-23 functions.R

@@ -204,7 +204,7 @@ assemble <- function(methods = c("log", "rf", "boost"), df.assem = all10mice, x.
 
 
 iter.auc <- function(mult =2, x.folds.iter = 5, 
-                     methods = c("orig.log","average",
+                     methods = c("orig.log","average","simp.boost",
                                  "simp.log", "orig.rf", "orig.boost"),
                      df.iter = all10.blend) {
         
@@ -241,6 +241,21 @@ iter.auc <- function(mult =2, x.folds.iter = 5,
                         
                         pred <- predict(fit, test, type = "response")
                         pred.list[["simp.log"]] = rbind(pred.list[["simp.log"]],
+                                                        data.frame("index" = folds[[k]],
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))
+                        
+                }
+                
+                if ("simp.boost" %in% methods) {
+                        print("simp.boost")
+                        fit <- gbm(b.rpdol ~ norm.log + norm.rf + norm.boost, data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 6,
+                                   n.minobsinnode = 10)
+                        
+                        pred <- predict(fit, test, n.tree = 6000, type = "response")
+                        pred.list[["simp.boost"]] = rbind(pred.list[["simp.boost"]],
                                                         data.frame("index" = folds[[k]],
                                                                    "pred" = pred,
                                                                    "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))

### 2016-06-27 TPR_FPR.rmd
started TP FP

---
title: "R Notebook"
output: html_notebook
---

```{r}

library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('randomForest')
library('caret')
library('pROC')
library('testthat')
library('purrr')
library(tidyr)
library('broom')
library('stringr')

# detect OS
if(.Platform$OS.type == 'windows'){
        # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
} else{
        # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
        # setwd("~/Documents/github/consult/finalwriteup/report_data")
        opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/')
}

```

Want to calculate TPR, FPR etc for an 'average' blended model. Find an average model from plot_finance_100_3 data frame

```{r}

best.fin<- readRDS("~/Documents/github/consult/finalwriteup/report_data/best_fin.rds")
plot.fin <- readRDS("~/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds")
head(plot.fin)
?inner_join
# join best thresholds to plot.fin
plot.fin <- inner_join(plot.fin, best.fin[,c('threshold','method','mean_profit_ratio')], by = "method")
# filter for thresholds == best. thresholds
plot.fin <- plot.fin %>% filter(threshold.x == threshold.y)
plot.fin <- plot.fin %>% filter(abs(profit.ratio- mean_profit_ratio)<=.5)
#subset first one from each method
iter.key <- plot.fin %>% mutate(diff = abs(profit.ratio - mean_profit_ratio)) %>%
        group_by(method) %>% arrange(diff) %>% slice(1) %>% ungroup()

#iter.key lists the iterations I want to extract for each method from compile.results (finance_results)
#may need to compile all my .. compile.results
CR.av_simplog<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_av_simplog.rds")
CR.comprf<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_comprf.rds")
CR.FWLS<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_FWLS.rds")
CR.simpboost<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_simpboost.rds")
CR.compboost<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_FWLScomp_boost.rds")
CR.compboost[['FWLS']] <- NULL
ans <- CR.compboost[['comp.boost']] %>% .$b.rpdol
CR.compboost[['comp.boost']] <- CR.compboost[['comp.boost']] %>% select(-balance.mlsto, -b.rpdol) 
#missing simp.boost for now
# append all lists together

CR <- append(CR.av_simplog, c(CR.comprf, CR.compboost, CR.simpboost, CR.FWLS))
#flatten list
CR <- bind_rows(CR, .id = "id") %>% as_data_frame
# create new dataframe of CR, filtered by iter.key
# use map_df
key.CR <- iter.key %>% filter(method %in% c('average','comp.boost','comp.rf','FWLS','simp.log', 'simp.boost'))
new.CR <- data.frame(row = 1:2364)
for(i in 1:nrow(key.CR)){
        col = CR %>% filter(id == key.CR$method[i]) %>% select_(key.CR$iteration[i])
        names(col) <- key.CR$method[i]
        new.CR = cbind(new.CR, col)
}

#now want to cbind optimal original results, stored in 'joined'
joined <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/joined.rds")
key.oCR <- iter.key %>% filter(grepl('orig', method))
key.oCR$index <- c(6, 91, 4)
key.oCR$method <- gsub("orig.","", key.oCR$method)
for(i in 1:nrow(key.oCR)){
        col = joined[[key.oCR$index[i]]] %>% select_(key.oCR$method[i])
        names(col) <- paste("orig.", key.oCR$method[i], sep = "")
        new.CR = cbind(new.CR, col)
}

#now have all mean outputs for each method
# want to create a confusion matrix for each
# need to append answers.. 
new.CR <- cbind(new.CR, b.rpdol = ans)
conf.CR <- new.CR

#push all probabilities one way or the other
for(p in seq_along(iter.key$method)){
        vec = conf.CR[,iter.key$method[p]]
        vec <- ifelse(vec < iter.key[iter.key$method == iter.key$method[p],]$threshold.x, 0, 1)
        conf.CR[,iter.key$method[p]] <- vec
}

```
Make confusion matricies

```{r}

# sub all 1's for loss and 0's for profit
conf.CR <- apply(conf.CR, 2, function(x) ifelse(x < 0.5, "profit", "loss")) %>% as.data.frame
conf.CR <- apply(conf.CR, 2, function(x) x = factor(names(x), 
                                               levels=c("profit","loss"))) %>% as.data.frame %>% head

#generate list of tables
methods = iter.key$method
confs = vector(mode = 'list', length(methods))
for(m in seq_along(methods)){
        confs[[m]] = (table(conf.CR[, methods[m]], conf.CR$b.rpdol) %>% prop.table(2)*100) %>% round(1)
}

conf.plot <- as.data.frame(confs) %>% melt
conf.plot <- conf.plot[,17:20]
names(conf.plot)[1:4] <- c("model.pred","actual","method","pc")
conf.plot$method <- gsub('.Freq',"",conf.plot$method)

saveRDS(conf.plot, "~/Documents/github/consult/finalwriteup/report_data/conf_plot.rds")

```

### 2016-06-27 finalwriteup/draft.rmd
started TP FP

@@
 
-![This method was implemented by the British government for a rail project which is still under completion](UKrail.JPG)
+![This method was implemented by the British government for a rail project which is still under completion](images/lit_review/UKrail.JPG)
 
@@
 It includes first how the data was obtained, the lengthy cleaning process, followed by variable importance analysis, variable selection, trials of selected predictive algorithms and problem constructs.
-The cost estimation problem was attempted first as a regression problem predicting return per dollar, followed by a simpler binary classification problem predicitng profit or loss.
+The cost estimation problem was attempted first as a regression problem predicting return per dollar, followed by a simpler binary classification problem predicting profit or loss.
 Once the best methods were selected, they were blended in numerous ways, using both simple averaging techniques and sophisticated machine-learning algorithms.
@@
 
-Chapter 5 presents the results of the series of analyses described above and directly relates them to the overall aim and hypotheses of the project.
-This is in the format of tables and plots where results are briefly discussed.
-The examination of results is taken deeper as well as in a broader context in Chapter 6: Discussion.
-Here the nature of the results are debated in terms of how surprising they were, their impact on the research hypothesis and limitations as well as suggested future work.
-The final Chapter 7: Conclusion summarises the  findings of the project, how the research answered the hypotheses and how the work could be viably applied in industry.
+The stages of research outlined in the method were divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion:
+
+* Chapter 5: Variable Selection
+* Chapter 6: Regression Models
+* Chapter 7: Binary Classification Models
+* Chapter 8: Extended Research including model blending, overall profit analysis, research limitations and future work
+
+The results section in each chapter presents key charts and tables that progressively answer the research question: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
+Then, a discussion finalises each chapter, where the results are examined deeply and analysed in a broader context.
+The nature of the results are also debated in terms of how surprising they were, their impact on the research hypothesis and limitations as well as suggested future work.
+The final chapter, the Conclusion, summarises the findings of the project, how the research answered the hypotheses and how the work could be viably applied in industry.
 
@@
 
-```{r}
 ## Method Conclusion
 
-# In summary, to determine whether a 
-# * obtained data: 4169 jobs from a single case study company over 12 years. Data was de-dentified by an employee of the company
-# * the data was all from the company CRM. There were three chief sources:
-#     * invoicing data
-#     * timesheet data
-#     * project summary data
-# * the data was cleaned from mistakes which were generally discovered as outliers 
-# * many new variables were engineered from the data, often by summarising time series data into a value that is relevant to a final data set of a single job per row
-# * variable selection: dozens of variables available which were condensed to the most important 10-12.
+<!-- In summary, to determine whether a  -->
+<!-- * obtained data: 4169 jobs from a single case study company over 12 years. Data was de-dentified by an employee of the company -->
+<!-- * the data was all from the company CRM. There were three chief sources: -->
+<!--     * invoicing data -->
+<!--     * timesheet data -->
+<!--     * project summary data -->
+<!-- * the data was cleaned from mistakes which were generally discovered as outliers  -->
+<!-- * many new variables were engineered from the data, often by summarising time series data into a value that is relevant to a final data set of a single job per row -->
+<!-- * variable selection: dozens of variables available which were condensed to the most important 10-12.
 variable importance methods were ANOVA, random forest, and cforests.
-the overlapping important variables were chosen first as well as some others
-# * 5 models were tested: baseiline: log reg, naiive bayes.
-# * sophisticated ML: boosted reg, random frest, bayesian networks
-# * best were comapred to a model averaged version
-# * Analysis of impact on bottom line for business if model is integrated into decision making
-# * link: results
-```
-
-# Results
-
-This chapter presents key charts and tables of results that progressively answer the research question: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
-
-## Variable Selection
+the overlapping important variables were chosen first as well as some others -->
+<!-- * 5 models were tested: baseiline: log reg, naiive bayes. -->
+<!-- * sophisticated ML: boosted reg, random frest, bayesian networks -->
+<!-- * best were comapred to a model averaged version -->
+<!-- * Analysis of impact on bottom line for business if model is integrated into decision making -->
+<!-- * link: results -->
+
+# Chapter _ Variable Selection
 
@@
 
-### Linear Regression
+## Linear Regression
 
@@
 
-### Random Forests
+## Random Forests
 
@@
 
-### Cforest
+## Cforest
 
@@
 
-Once several methods of variable importance analysis have been completed, the separate results were rationalised and final variables were selected for the predictive models.
-
-### Variable Selection Summary
-
-Based on the results of the three methods for variable importance, 11 independent variables were selected for predictive modelling.
+Once several methods of variable importance analysis had been completed, the separate results were rationalised and final variables were selected for the predictive models.
+
+## Variable Selection Results Summary
+
+Based on the results of the three methods for variable importance, 11 explanatory variables were selected for predictive modelling.
 They are listed in the table below along with the methods that highlighted each variable as important (randomForest = RF, cForest = CF):
@@
 
-The variables above were used to build the predictive models and the results are summarised in the following two sections.
-
-## Model Results
-
-Once the explanatory variables were narrowed down to a succinct list, the predictive models were built and tested.
-First, prediction of return per dollar as a regression problem was attempted followed by prediction of profit or loss as a binary classification problem.
-linear regression, random forests and boosted tree methods were applied to both problems and had varying levels of success.
-
-### Regression Models
-
-#### ANOVA
+## Variable Selection Discussion
+
+The variable importance results from the three types of models: ANOVA regression, randomForest, and cForest showed more similarities than differences. All three models ranked five variables highly:
+
+* percent of hours by a professional
+* project timespan
+* number of employees in project team
+* total amount invoiced
+* business category of client
+
+This was not necessarily expected, particularly between ANOVA and the ensemble tree methods as ANOVA measures linear relationships in constrast to non-linear ensemble trees.
+Random forests favour variables that are numeric or have many categories and this held true for the analysis.
+The 6 highest-ranked variables were numeric with one categorical variable, Business category of client, which had 28 categories.
+The two lowest ranking variables had only 4 categories.
+It was expected that cForests would rank the categorical variables with only 4 categories more fairly.
+This was the case with Discipline and it recieved a mid-level ranking.
+Business category of the client and Broad business type however, were ranked similarly to randomForests.
+
+The qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, general client characteristics, project timespan, and project category.
+It was expected that Billing Type would play a more important role in predicting profitability as it describes the structure of the incoming fees.
+For this reason, and with advice from the case study business, the Billing Type was kept in the model.
+
+For construction projects and IT software projects, the literature indicated size was the most important variable predicting effort.
+Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996][@Finnie1997][@Pai2013].
+In this case study, profitability is the dependent variable, not total effort and the size variable plays a different role.
+Instead of contributing to predicting the total effort, it is describing whether the size of a project correlates with how profitable the project is.
+The closest estimate of size in this case study is the invoiced amount variable which did rank as very important.
+The exact invoiced amount is difficult to guess before a project begins, and is part of the problem this study is addressing.
+Therefore, the invoiced amount was binned into categories that a manager could much more confidently choose as an estimate.
+The other variables indicated as important in this study  (including the number of employees on a project, client business type and percent of hours by certain types of employees) have not been strongly represented in the literature to date.
+This is because the variables were easily calculated from internal CRM data but would not be easily obtained externally from projects across many companies or regions.
+
+The 11 explanatory variables from table __ were progressed to the next stage where they were used to build predictive models.
+The outcomes of the regression and categorical models are summarised and discussed in the following two chapters.
+
+# Chapter _ Regression Models
+
+Prediction of 'return per dollar' as a regression problem was attempted first followed by prediction of profit or loss as a binary classification problem.
+Linear Regression, Random Forests and Boosted Tree methods were applied to both problems with varying success.
+
+## ANOVA
 
@@
 
-were statistically significant.
-Two types of ANOVA models were run.
-The first used only a 'core' list of 6 variables that were largely complete and any relevant interaction terms.
+were statistically significant. Two types of ANOVA models were run. The first used only a 'core' list of 6 variables that were largely complete and any relevant interaction terms.
+
+* Total amount invoiced 
+* Discipline 
+* Timespan
+* Team size
+* % of hours completed by a professional level employee
+* Client business category
+
 The second method was developed to overcome the problem of missing data.
@@
 
-results.A2 = readRDS("reg_lm.rds")
+results.A2 = readRDS("report_data/reg_lm.rds")
 results.A2 <- dplyr::filter(results.A2, diff>-15)
@@
 
-#### Random Forest
+## Random Forest
 
 ```{r load_rf_results}
-results.rf.core = readRDS("reg_rf_core.rds")
+results.rf.core = readRDS("report_data/reg_rf_core.rds")
 # n.samples(data = results.rf.core$diff)
 
-results.rf2 = readRDS("reg_rf.rds")
+results.rf2 = readRDS("report_data/reg_rf.rds")
 # n.samples(data = results.rf2$diff)
@@
 
-The two regression models predicting return per dollar resulted in high mean RMSE's ranging from \$`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
-Because of the poor results, further development of regression models was deffered in favour of investiation of binary classification models.
+The two regression models predicting 'return per dollar' resulted in high mean RMSE's ranging from \$`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
+The next section discusses possible reasons for the poor results and the decision to halt progression of regression models.
+
+## Regression Models Discussion
+
+Both ANOVA and randomForest models produced a similar lowest RMSE value, but performed differently when directly comparing models built on the same list of variables.
+Both methods predicted at best within an average of \$0.48 for the response variable, 'return per dollar', using either core variables or a sample of core and less complete variables.
+Random Forest predictictions were much more accurate than ANOVA's with a mix of core (almost complete) and less variables while randomForest and ANOVA performed similarly with the core variables only.
+There was much less data available for training the models that used less-complete variables so it was interesting that randomForest performed considerably better (RMSE of \$0.50 for randomForest vs.
+\$2 for ANOVA).
+The main difference between ensemble tree methods and ANOVA is that ANOVA assumes linear relationships between variables (some of which are normalised) whereas trees are capable of capturing non-linear relationships.
+It is possible the less complete variables had non-linear relationships with the response variable.
+RandomForest is also an ensemble method while ANOVA is not.
+Each randomForest is the average of 500 single trees, whereas each ANOVA model is a single model only.
+50 randomForest models produce 500 times more models than 50 ANOVA models so the averaging effect may have eliminated outliers for better overall performance.
+
+If the best RMSE achievable at that point was \$0.48, what does that imply in terms of the business decision?
+The magnitude of the RMSE can be evaluated from the end-user's perspective (the manager who is proposing fees for new projects).
+To give a feel for typical project performance, refer the plot below of historical 'return per dollar' values:
+
+```{r}
+if(.Platform$OS.type == 'windows'){
+        all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv' )
+} else{
+        all8a<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all8a.csv')
+}
+ggplot(all8a, aes(x = return.pdol, y = ..density..)) + geom_histogram() + geom_density() +
+        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects")
+
+```
+
+The 95% confidence interval for this data is \$`r (mean(all8a$return.pdol)- qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)` to \$`r (mean(all8a$return.pdol) + qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)`.
+If a project returns \$0 for every dollar spent (return per dollar = \$0), the project has earned only enough to cover the cost of doing the project (salary costs and business costs).
+If the return per dollar is -\$0.50, the project has earned enough to cover half of the costs of doing the job.
+Finally, if the return per dollar is \$1, the project earned double what it cost to perform.
+Using this reasoning, if model predictions have an average error of \$.50 for return per dollar, this represents an error in revenue of half what the project cost to complete.
+Unfortunately, this is clearly not informative enough to bring into the business decision-making process.
+
+Further regression methods could have been persued, however randomForest is a high-performing algorithm and was far from providing usable results.
+It is likely that either not enough data was available or that the given variables did not capably explain all the factors affecting return per dollar.
+Each row of data is a project, which can take the company from a few weeks to sometimes years to complete.
+Therefore, more data cannot be easily accessed or created.
+It was decided categorical predictions should be pursued next as predicting whether a job will be profitable or not is a simpler task and could more effectively leverage the limited data.
+
+This conclusion differs to case studies in the literature which tend to predict total cost/effort as a regression problem [@Kim2004][@Chan2005][@Akintoye2000].
+This is likely because the literature dealt with substantial projects which are rarer and impact governments and large companies.
+A simple profit or loss prediction would not be useful.
+The case study company on the other hand may find a profit/loss binary prediction useful because they complete hundreds of small projects a year.
+If binary classification is successful, applications of the predictions must be explored along with contributions to the company decisions and profits.
+
+In summary, because of the poor results when predicting return per dollar, further development of regression models was deferred in favour of binary classification models.
 Binary classification is a simpler task than regression because it requires less precision.
@@
 
-### Binary Classification Models
+# Chapter _ Binary Classification Models
 
 The cost estimation problem analysed in the literature review had been converted to a profitability estimation problem and was finally converted to a binary profit/loss prediction.
-The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable (or not).
-The original return per dollar response variable was transformed by converting aLl 'return per dollar' values less than or equal to zero to 'loss making' jobs and 'return per dollar' values greater than 1 became 'profitable' jobs.
-The following section summarises results from the trialled binary classification methods.
-
-#### Initial Trial of Five Methods
-
-The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were intially trialled without imputing data, using the maximum amount of data possible depending on the method.
+The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable project (or not).
+The original 'return per dollar' response variable was transformed by converting all 'return per dollar' values less than or equal to zero to 'loss making' jobs and 'return per dollar' values greater than 1 to 'profitable' jobs.
+The following section summarises results from the trialled methods.
+
+## Results from Five Methods
+
+The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were intially developed without imputing data, using the maximum amount of data possible depending on the method.
 Boosted Trees and Naiive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
-Logistic Regression, Random Forest and Bayesian Networks require datasets without missing values so a subset of complete data was used for these.
+Logistic Regression, Random Forest and Bayesian Networks require datasets without missing values so a subset of complete data was used.
 It was expected that Boosted Trees would perform best as they are known to have powerful predictive capabilities and can use incomplete data.
 For this reason, 4 x 5-fold runs of each method were performed (20 models each) and statistically compared against Boosted Trees' results.
-A power calculation was then used to calculate the number of runs required for a statistical power of 0.8.
+A power calculation was then used to calculate the number of runs required to achieve a statistical power of 0.8.
 
@@
 
-Before the plot could be made, the Boosted Tree algorithm parameters were tuned to optimise the AUC outcomes.
-The caret package was used for this and the tuned parameters are summarised below [@caret]:
+Before the plot could be made, the Boosted Tree algorithm parameters were tuned to optimise the AUC outcomes. The caret package was used to tune the following parameters [@caret]:
 
@@
 
-```{r}
-# image below generated from file rf_log_naiive_auc.rmd
-# lines 683 to 689
-
-```
+<!-- image below generated from file rf_log_naiive_auc.rmd -->
+<!-- lines 683 to 689 -->
 
@@
 
-reduc.totAUC <- readRDS("reduc_df_auc.rds")
+reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
 
@@
 
-Results using imputed data tightened the distribution of randomForest AUC output substantially but slightly increased the distribution spread of Boosted Trees and Logistic Regression.
-Since the methods using imputed data performed similarly, substantially tightened the distribution of randomForest results, and made use of more data, the imputed data models were propogated forward to the development of blended models.
-
-#### Blended Models
+## Binary Classification Discussion
+
+Testing Boosted Trees, Random Forests, Naive Bayes, Logistic Regression and Bayesian Networks on the binary classification problem showed that Boosted Trees, Logistic Regression and Random Forests performed best (according to AUC).
+Bayesian Networks are not particularly popular and require discretised variables so it was expected that this algorithm would not perform as well as highly regarded methods such as Boosted Trees and Random Forests.
+Naive Bayes and Logistic Regression were included as baseline models and it was expected that the more complex models would outperform these.
+Therefore it was surprising that results from 100 Logistic Regression models were not statistically lower than Boosted Trees.
+
+A possible explanation for this, according to the literature, is that there was not enough data for the ensemble trees to learn the complex decision rules at which they excel.
+Trees tend to overfit the patterns in a smaller training set.
+Logistic regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
+This may explain Logistic Regression's comparitively high performance on the case study's smaller dataset with many variables.
+
+Boosted trees, Random Forests and Logistic Regression all had mean AUC values around 0.775.
+The tests were repeated using imputed data and again the means sat around 0.775.
+The variables were not particularly well suited for imputing so this result is logical and could have even diminished the mean AUC.
+For example, many of the imputed variables were categorical such as job category, percent of job completed by a professional level employee, and the position of the majority worker on a job.
+The benefit of having an imputed dataset that performs acceptably well is that each model is fed the full data set and therefore produces predictions for the full data.
+Beforehand, randomForest and Logistic Regression models were limited to a complete set of original data which implied a reduced data set.
+With complete model predictions, the high performing models could be blended to test whether better predictions were possible.
+For this reason, the imputed data models were propogated forward to the development of blended models.
+
+In summary, the individual binary classification models performed well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a new job will be profitable or loss-making.
+Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model actually affects decisions in practise.
+Results from model blending and overall profit impact are presented and discussed in the next chapter.
+
+# Chapter _ Extended Analysis
+
+## Blended Models
 
@@
 In this case, the randomForest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialled ranging from simple to complex.
-The simplest way was to average the probability output (a number between 0 and 1) of all three models.
-The next two simplest methods were to build a logistic regression and boosted tree from the three model outputs only.
-Feature Weighted Linear Stacking (FWLS), randomForest and boosted trees were also tested as blending methods where the predictions from each model became additional variables to the original explanatory variables (meta-features).
+The simplest way averaged the probability output (a number between 0 and 1) of all three models.
+The next two simplest methods consisted of building a Logistic Regression and Boosted Trees model from the three individual model outputs only.
+Feature Weighted Linear Stacking (FWLS), randomForest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (meta-features).
 The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature conditions could be weighted as such.
@@
 
-##### Variable Selection: Blended Model
+### Variable Selection: Blended Model
 
@@
 
-The logistic regression found the F-statistics from the following variables were significant in rejecting the null hypothesis (p-value < 0.05):
+An ANOVA regression found the F-statistics from the following variables were significant in rejecting the null hypothesis for p-value < 0.10:
 
@@
 
-* Discipline * results of logistic regression model
-* Discipline * results of boosted tree model
-* team size for project * results of logistic regression model
+* Timespan * results of logistic regression model
+* Total amount invoiced * results of logistic regression model
+* Total amount invoiced * results of randomForest model
 * Client Business type * results of boosted tree model
@@
 * Timespan
+* Team size
+* Client business category
 * Project category
+* Results from each individual model
 
@@
 
-#### Comparison of Blended Models
+Random Forests and Boosted Trees passively interact explanatory variables, but for logistic regression, these must be manually input.
+Only interaction terms with statistically significant F-statistics from the ANOVA test were included in order to reduce noise.
+
+### Comparison of Blended Models
 
@@
 
-auc.list <- readRDS("report_data/blend_auc1.rds")
+auc.list <- readRDS("report_data/blend_auc2.rds")
 means = lapply(auc.list, mean) %>% unlist
@@
 
-```{r, fig.width = 8, fig.height=10}
+```{r, fig.width = 8, fig.height=5}
 
@@
 
-### Bottom Line Analysis
+## Profit Analysis
 
@@
 The measure of profitability could not be modelled but simplifying the profitability prediction to a binary (profit/loss) model was successful.
-To answer the second research question, whether the model would improve the business' profitability, a bottom line analysis was performed.
-
-The models output a 'probability' between 0 and 1 that each project will be a loss-making job (where probability = 1).
+To answer the second research question, whether the model would improve the business' profitability, a profit analysis was performed.
+
+The models output a 'probability' between 0 and 1 that each project will be a loss-making job (probability = 1 indicates a loss-making job).
 This leaves the question, at what probability would a decision-maker round the probability up to 1 or down to 0?
 And what business decision would then be made?
-To find the rounding threshold, an experimental business-scenario was tested.
+To find the threshold point for rounding, an experimental business-scenario was tested.
 At a given threshold, all projects calculated above the threshold were considered too risky, and were rejected.
-All profits and losses from these projects were forfeited, and the profits from the remaining jobs (below the threshold) were summed to give a total profit.
+All profits and losses from these projects were forfeited, while the profits from the remaining jobs (below the threshold) were summed to give a revised total profit.
 
@@
 # How to achieve statistical power of 0.8?
-finance.plot = readRDS("plot_finance_100.rds" )
+finance.plot = readRDS("report_data/plot_finance.rds" )
 fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
@@
 }
-ns
-
-# orig.boost needs 6, orig.log needs 4, orig.rf needs 6 and simp.log needs 64 : ) do 100!
-
-```
-
-Using this logic, if the threshold was 0, all projects were rejected and total profit would equal 0.
+# ns
+
+# orig.boost needs 6, orig.log needs 4, orig.rf needs 9 and simp.log needs 64 : ) do 100!
+
+```
+
+Using this logic, if the threshold was 0, all projects were rejected and the total profit would equal \$0.
 If the threshold was 1, all projects were accepted and the total profits would be the same as historical figures.
-Probability results from the two blended models as well as three individual models were tested and because the models require a training and testing set, the process was repeated 100 times with 20 x 5 fold cross validation models to achieve a statistical power of 0.8 (calculated after an initial run of 10 models).
-The plot below illustrates the results of each model at each threshold with a local polynomial regression line of best fit (loess) and a 95% confidence interval around the mean.
+This profit calculation was made on a range of thresholds between 0 and 1 at 0.05 increments.
+The total profits were plotted for each threshold and joined to make a profit curve where at an optimal threshold, the total profits were highest (refer plot below).
+
+Profit curves for the three simple blended models as well as the three individual models were calculated.
+However, the total profit coordinates were based on probability outputs in test datasets.
+This means that the profit curve for the simple logistic regression blend (for example) would be different for different training/testing data divisions.
+I.e.
+the total profits at each point in the profit curve depends on the model parameters that were calculated using the training data and which projects were analysed in the test data.
+Therefore, the process of training a blended model, then calculating the profit curve from the test data was repeated multiple times to create a distribution of profit curves for each method (blended and individual).
+
+To start, 10 curves for each method were calculated (2 x 5-fold cross validation).
+Two-sampled power calculations (for a power of 0.8) were performed between the results of each method at their optimal threshold against the best performing method (simple average blend).
+For example, the optimal threshold for the simple average blended model was 0.55 while the optimal threshold for the original boosted tree model was 0.60.
+The 10 'total profit' values at these two thresholds were compared in the two-sample power calculation to determine the number of samples required to achieve a power of 0.8.
+The maximum required number of samples was 64 so 100 models were of each method were built (20 x 5 fold cross validation).
+
+The plot below illustrates the results of each model at each threshold with a line joining the mean value at each threshold for each method.
+The grey ribbon illustrates the 95% confidence interval around the mean at each threshold for the 100 models.
 
@@
 
-finance.plot = readRDS("plot_finance.rds" )
+finance.plot = readRDS("report_data/plot_finance_100_3.rds" )
+finance.plot$method <- factor(finance.plot$method, 
+                                    levels=c("average", "simp.log","simp.boost",
+                                             "comp.boost","comp.rf", "FWLS","orig.log",
+                                             "orig.rf","orig.boost"))
 
 ggplot(finance.plot, aes(x= threshold, y= profit.ratio, colour = method)) + 
-        geom_point(alpha = 0.1, shape = 16) + 
-        stat_smooth() +
+        # geom_point(alpha = 0.005, shape = 16, size = 1) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) +
+        stat_summary(fun.y = mean, 
+                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
+                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
+                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
+        stat_summary(fun.y = mean, geom="line") +
         scale_y_continuous(breaks = seq(0,110, by = 10)) + 
         scale_x_continuous(breaks = seq(0,1, by = .1)) +  
-        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
-        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
-        theme(legend.position = "bottom") + 
-        scale_colour_discrete(labels = c("average" = "Blend - Average",
-                                       "orig.boost"= "Boosted Trees",
-                                       "orig.log" = "Logistic Regression",
-                                       "orig.rf" = "Random Forest",
-                                       "simp.log" = "Blend - Log Reg"))+
-        guides(col = guide_legend(nrow = 2)) +
-        labs(x = "Probability Threshold", 
-             y = "Ratio of Model Profit to Historical Profit (%)",
-             title = "Profit Curve for Blended and Individual Models")
-
-```
-
-The two blended models outperformed the individual models as shown in this table describing each models' optimal thresholds:
-
-```{r profit_table}
-
-finance.plot$method = as.factor(finance.plot$method)
-# want mean and sd for each method at each threshold
-
-fin.summary = finance.plot %>% group_by(threshold, method) %>% 
-        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
-                            sd_prof_ratio = round(sd(.), 1)),
-                       profit.ratio)
-
-#for each method find max mean
-
-best.fin = fin.summary %>% group_by(method) %>% 
-        filter(mean_profit_ratio == max(mean_profit_ratio)) %>% 
-        arrange(desc(mean_profit_ratio))
-kable(best.fin)
-
-best.mean = best.fin %>% filter(method == "average") %>% .$mean_profit_ratio
-best.sd = best.fin %>% filter(method == "average") %>% .$sd_prof_ratio
-best.t = best.fin %>% filter(method == "average") %>% .$threshold
-
-```
-
-The average of the three individual models performed best with the highest mean profit ratio at `r best.mean`% and the narrowed standard deviation at `r best.sd`%.
-This means that for the blended average model, if all jobs above the probability threshold 0.55 were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profit data.
-
-For clarity, the plot for the average blended model is shown below:
-
-```{r av_profit_curve, fig.height=10}
-ggplot(finance.plot %>% filter(method == "average"), 
-       aes(x= threshold, y= profit.ratio, colour = method)) + 
-        geom_point(alpha = 0.1, shape = 16) + 
-        stat_smooth() +
-        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
-        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
-        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
-        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
         theme(legend.position = "none") +
@@
              y = "Ratio of Model Profit to Historical Profit (%)",
-             title = "Profit Curve for Blended Average Model")
+             title = "Profit Curve for Blended and Individual Models") + 
+        facet_wrap(~method, 
+                   labeller = labeller( method = c("average" = "Simple Average Blend",
+                                       "orig.boost"= "Orig. Boosted Trees",
+                                       "orig.log" = "Orig. Logistic Regression",
+                                       "orig.rf" = "Orig. Random Forest",
+                                       "simp.log" = "Simple Log. Reg. Blend",
+                                       "simp.boost" = "Simple Boosted Tree Blend",
+                                       "comp.boost"= "Complex Boosted Tree Blend",
+                                       "comp.rf"= "Complex randomForest Blend",
+                                       "FWLS" = "FWLS"
+                                       )))
+
+```
+
+The two blended models outperformed the individual models as shown in the plot above with higher profit ratios as well as tighter confidence intervals.
+The table below denotes each models' optimal thresholds and the corresponding increase in total profits:
+
+```{r profit_table}
+
+finance.plot$method = as.factor(finance.plot$method)
+# want mean and sd for each method at each threshold
+
+fin.summary = finance.plot %>% group_by(threshold, method) %>% 
+        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
+                            sd_prof_ratio = round(sd(.), 2)),
+                       profit.ratio)
+#for each method find max mean
+
+best.fin = fin.summary %>% group_by(method) %>% 
+        filter(mean_profit_ratio == max(mean_profit_ratio)) %>% 
+        arrange(desc(mean_profit_ratio), sd_prof_ratio) %>% 
+        slice(1) %>% ungroup() %>% 
+        arrange(desc(mean_profit_ratio), sd_prof_ratio)
+
+best.fin <- best.fin %>% mutate(lower_bound_95CI = (mean_profit_ratio - 1.96*sd_prof_ratio)%>% round(1))
+best.fin <- best.fin %>% mutate(upper_bound_95CI = (mean_profit_ratio + 1.96*sd_prof_ratio)%>% round(1))
+
+kable(best.fin)
+# saveRDS(best.fin, "report_data/best_fin.rds")
+
+best.mean = best.fin %>% filter(method == "simp.log") %>% .$mean_profit_ratio
+best.sd = best.fin %>% filter(method == "simp.log") %>% .$sd_prof_ratio
+best.t = best.fin %>% filter(method == "simp.log") %>% .$threshold
+
+```
+
+The simple logistic regression blend performed best with the highest mean profit ratio at `r best.mean`% with a standard deviation of `r best.sd`%.
+This means that for the simple blended logistic regression model, if all jobs above the probability threshold `r best.fin %>% filter(method == "simp.log") %>% .$threshold` were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profit data.
+
+For clarity, the plot for the simple blended logistic regression model is shown below:
+
+```{r av_profit_curve}
+ggplot(finance.plot %>% filter(method == "simp.log"), 
+       aes(x= threshold, y= profit.ratio, colour = method)) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
+        geom_point(alpha = 0.1, shape = 16) +
+        stat_summary(fun.y = mean, 
+                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
+                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
+                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
+        stat_summary(fun.y = mean, geom="line") +
+        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
+        theme(legend.position = "none") +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Profit Curve for Simple Logistic Regression Blended Model")
 
@@
 
-# Discussion - 10 to 20 pages
-
-```{r}
-# The Discussion section:
-# 
-# comments on your results;
-# explains what your results mean;
-# interprets your results in a wider context; indicates which results were expected or unexpected;
-# provides explanations for unexpected results.
-# The Discussion should also relate your specific results to previous research or theory.
-You should point out what the limitations were of your study, and note any questions that remain unanswered.
-The Discussion CAN also include Conclusions/Future Research.
-```
-
-## Variable Importance
-
-The variable importance results from the three types of models: ANOVA regression, randomForest, and cForest were largely similar. All three models ranked five variables highly:
-
-* percent of hours by professional
-* project timespan
-* number of employees in project team
-* total amount invoiced
-* business category of client which was positive
-
-This was not necessarily expected particularly between ANOVA and the ensemble tree methods as ANOVA measures linear relationships as opposed to non-linear ensemble trees.
-Random forests favour variables that are numeric or have many categories and this held true for the analysis.
-The 6 highest ranked variables were numeric and one categorical variable, Business category of client, with 28 categories.
-The two lowest ranking variables had only 4 categories.
-It would be expected that cForests would rank the categorical variables with only 4 categories more fairly which was the case with Discipline.
-It was ranked mid way.
-Business category of the client and Broad business type however, were ranked similarly in randomforests.
-
-The nature of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, general client characteristics and id, project timespan, and project category.
-It was
-expected that Billing Type would play a more important role in predicting project profitability as it describes the nature fees are charged.
-For this reason, and with advice from the case study business, the Billing Type was temporarily kept in the model.
-
-For construction projects and IT software projects, the literature indicated size was the most important variable predicting effort.
-Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996][@Finnie1997][@Pai2013].
-In this case study, profitability is the dependent variable, not total effort and the size variable plays a different role.
-Instead of contributing to predicting the total effort, it is describing whether the size of a project correlates with how profitable the project is.
-The closest estimate of size in this case study is the invoiced amount which does show up as very important.
-The exact invoiced amount cannot be known before a project begins, as this would solve the problem of profitability.
-Therefore, the invoiced amount was binned into categories that a manager could choose as an estimate.
-The other important variables (including number of employees on a project, client business type and percent of hours by certain types of employees) have not been strongly represented in the literature to date.
-
-## Regression
-
-Both ANOVA and randomForest models predicted at best within an average of $0.48 for the dependent variable, return per dollar using either core variables or a sample of core and less complete variables.
-RandomForest predicts much better than ANOVA with a mix of core (almost complete) and incomplete variables while randomForest and ANOVA perform similarly with the core variables only.
-There is much less data available for training the models that include less complete variables so it is interesting that randomForest performed much better (RMSE of $0.50 for randomForest vs.
-$2 for ANOVA).
-The main difference between ensemble tree methods and ANOVA is that ANOVA assumes linear relationships between variables (some of which are normalised) whereas trees are capable of capturing non-linear relationships.
-It could be the case that the less complete variables had non-linear relationships with the dependent variable.
-RandomForest is also an ensemble method while ANOVA is not.
-Each randomForest is the average of 500 single trees, whereas each ANOVA model is simply a single model.
-50 randomForest models produces 500 times more models than 50 ANOVA models so this averaging effect may be effectively eliminating outliers for better overall performance.
-
-The RMSE can be evaluated in terms of the end-user's perspective (business manager who is proposing fees for new projects).
-There is a record of how previous projects have performed in terms of return per dollar.
-Refer below:
-
-```{r}
-if(.Platform$OS.type == 'windows'){
-        all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv' )
-} else{
-        all8a<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all8a.csv')
-}
-ggplot(all8a, aes(x = return.pdol, y = ..density..)) + geom_histogram() + geom_density() +
-        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects")
-
-```
-
-The 95% confidence interval for this data is $`r (mean(all8a$return.pdol)- qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)` to $`r (mean(all8a$return.pdol) + qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)`.
-If a project returns $0 for every dollar spent (return per dollar = $0), the project has earned only enough to cover the cost of doing the project (salary costs and business costs).
-If the return per dollar is -$0.50, the project has earned enough to cover half of the costs of doing the job.
-Finally, if the return per dollar is $1, the project has earned double what it cost to perform.
-If predictions have an average return per dollar error of $0.50, this represents an error in revenue of half what the project cost to complete.
-Clearly, this is not informative enough to bring into business decision making processes.
-
-Further regression methods could have been persued, however randomForest is a high-performing algorithm and was far from providing usable results.
-It is likely the case that either not enough data was available or that the given variables do not capably explain the level of return per dollar.
-Each row of data is a project, and takes the company months, sometimes years, to create a single row of data.
-Therefore, more data cannot be accessed or easily created.
-It was decided categorical predictions should be pursued next as predicting whether a job will be profitable or not is a simpler task and could more effectively leverage the limited data.
-
-Case studies in the literature tend to predict total cost/effort which is a regression problem (on a numeric scale) [@Kim2004][@Chan2005][@Akintoye2000].
-This is likely because the literature dealt with large projects which impact governments and large companies, and a simple profit or loss prediction is not useful.
-A regression prediction is not feasible for the data in this project, but a profit/loss binary prediction may be useful for this company because they complete hundreds of small projects a year.
-If binary classification is successful, the method the prediction contributes to company decisions must then be explored.
-
-## Binary Classification
-
-Testing gradient boosted trees, random forests, naiive bayes, logistic regression and bayesian networks in binary classification showed that boosted trees, logistic regression and random forests performed best (according to AUC).
-Bayesian Networks are not particularly popular and require discretised variables so it was expected that this algorithm would not perform as well as highly regarded methods such as boosted trees and random forests.
-Naiive bayes and logistic regression were included as baseline models and it was expected that the more complex models would outperform these.
-Therefore it was surprising that results from 100 logistic regression models were not statistically lower than boosted trees.
-
-A possible explanation for this according to the literature is that there was not enough data for the ensemble trees to learn the complex decision rules at which they excel.
-Trees tend to overfit the patterns in a smaller training set.
-Logistic regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
-This may explain why logistic regression is performing as well as enseble trees with the case study's smaller dataset with many variables.
-
-Boosted trees, random forests and logistic regression all have AUC means around 0.775.
-The tests were repeated using imputed data and again the means sat around 0.775.
-The variables were not particularly well suited for imputing so this result is logical and could have even reduced the AUC.
-Many of the imputed variables were categorical and difficult to impute (such as job category, percent of job completed by a professional level employee, and the position of the majority worker on a job).
-The benefit of having an imputed dataset that performs well enough is that each model produces predictions for the full set of data.
-Beforehand, each model was limited to a complete set of data which implies a reduced data size.
-With complete model predictions, the high performing models can be blended to test whether better predictions are possible.
-Blending models brings about limitations which is discussed in the next section.
-
-This individual models perform well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a potential job will be profitable or loss-making.
-Whether the model is worth implementing in the work place is dependent on how the model affects decisions and the degree the algorithm will improve the 'bottom line' profits for the business.
-This will be reviewed in the following section.
-
-## Blended Models
+## Extended Analysis Discussion
+
+### Blended Models
 
 Trials of different methods of blending model results demonstrated that again, the simplest methods worked best with the data.
-In this case, averaging the results of the best three single models (logistic regression, random forests, and boosted trees) or taking a logistic regression of the three models outperformed more complex boosted trees, random forests, and logistic regression blends that facilitated interaction of the individual model results and original variables (meta-features).
+In this case, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
 
 It was plausible that simple blending models would perform better than complex methods because the data was not big enough for complex models to learn intricate patterns.
-This was observed in the single method models when logistic regression performed as well as random forests and boosted trees.
+This was observed in the single method models (previous chapter) when Logistic Regression performed as well as Random Forests and Boosted Trees.
 The literature does not extensively cover blended machine learning methods however their success in the 2009 Netflix competition generated some publications.
-The Netflix dataset comprised of almost 3,000,000 observations and the size of the datacould have facilitated the success of complex blending methods such as FWLS [@Sill2009].
-That being the case, it was still surprising that simple averaging of probabilities from the individual models performed as well as blending results using logistic regression.
-It was expected that logistic regression would have an advantage over model averaging due to its ability to weight the prediction variables while maintaining a simple model structure.
-
-The logistic regression blending model did indeed weight predictions from the original models differently.
-If the predicted probability from the original logistic regression model increased by 0.1, the odds of the model prediciting a loss making job increased by 37.4%.
-Similarly when probability predictions from the random forests model increased by 0.1, the odds of the blended model predicting a loss increased by 25.8%.
-On the other hand, for the same increase in the original boosted tree model, the odds of the blended logistic regression model predicting a loss increased by only 2.6%.
+The Netflix dataset comprised of almost 3,000,000 observations, so the size of the data could have facilitated the success of complex blending methods such as FWLS [@Sill2009].
+That being the case, it was still surprising that simple averaging of probabilities from the individual models performed as well as blending results using Logistic Regression.
+It was expected that Logistic Regression would have an advantage over model averaging due to its ability to weight the prediction variables while maintaining a simple model structure.
+
+The Logistic Regression blending model did indeed weight predictions from the original models differently.
+If the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model prediciting a loss making job increased by 37.4%.
+Similarly when probability predictions from the Random Forests model increased by 0.1, the odds of the blended model predicting a loss increased by 25.8%.
+On the other hand, for the same increase in the original Boosted Tree model, the odds of the blended Logistic Regression model predicting a loss increased by only 2.6%.
 This contrasted with the alternative simple model that averaged the probabilities from each original model (i.e.
 weighted them equally) but still managed to produce similar AUC results.
-It was not clear why this was the case, however both simple models were carried forward to analyse their respective bottom line impacts on the company's profits which could reveal performance differences.
-
-## Bottom Line Analysis
+It was not clear why this was the case, however both simple models were compared again in the analysis of the impact each model could have on the company's overall profits.
+The analysis of overall profits did not directly relate to AUC measurements, so a range of models were carried forward for comparison.
+These included the three original individual models, three simple blended models and two complex blended models.
+
+### Profit Curve Analysis - Discussion
 
@@
 
-simp.l <- readRDS("finance_average_model.rds")
-
-```
-
-The blended models significantly (?) outperformed individual models which was expected given the two blended models' AUC results were significantly higher than the original logistic regression results (which were in line with random forests' and boosted trees' results).
-The simple average of logistic regression, random forests, and boosted trees results produced a better profit curve than a logistic regression of the same three variables.
-The model produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`.
-To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r (best.mean-100)/100*1000000` by rejecting all jobs with a probability higher than `r best.t`.
+simp.l <- readRDS("report_data/finance_average_model.rds")
+
+```
+
+**Highest Profit Curves**
+
+The simple average blend and logistic regression blended models outperformed the complex blended models, which is rational in terms of theit AUC values, which were significantly higher than the others.
+The simple average of logistic regression, random forests, and boosted trees results produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression that blended the same three variables.
+The models produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t` for the blended Logistic Regression and `r best.fin %>% filter(method == 'average') %>% .$threshold` for simple averaging.
+To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r (best.mean-100)/100*1000000` by rejecting all jobs with a probability higher than `r best.t` (blended logistic regression).
 In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.
 
+**Comparison of All Profit Curves**
+
+The tested profit curve models can be divided into three categories: original individual methods, simple blends of the individual methods, and complex blends of the individual methods that employed interactions with original features.
+All original methods did achieve improvements of profit at their optimal thresholds, but did not necessarily consistently improve profits across all cross-folded models.
+The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
+That is, it cannot be said with 95% confidence that the original methods' would produce a profit higher than historical profits if projects higher than the optimal thresholds (jobs that are likely to be loss-making) were rejected.
+Clearly this is not helpful for the case-study business to adopt.
+
+The complex blended models had mixed results.
+FWLS and randomForests performed more poorly than the individual methods while Boosted Trees produced a marginally higher profitability ratio (`r best.fin %>% filter(method == "comp.boost") %>% .$mean_profit_ratio`) but the lower bound of its confidence interval was `r best.fin %>% filter(method == "comp.boost") %>% .$lower_bound_95CI`, i.e.
+barely above original profits.
+
+The simple blended models performed more favourably, as already discussed, particularly the simple average and Logistic Regression.
+The Boosted Tree's lower bound confidence interval was `r best.fin %>% filter(method == "simp.boost") %>% .$lower_bound_95CI`% at its optimal threshold point.
+This is above 100%, so it is highly likely the algorithm would improve profits, but the margin was not high in comparison to the other two simple methods.
+
+It is not clear why simple linear models and the averaging model outperformed ensemble tree blending methods.
+It may be that there was not enough data for the ensemble trees to adequately learn the complex series of rules they develop, although this should be compensated for in the tuning phase.
+It was also surprising FWLS was ineffective given the success it had in the Netflix Prize competition [@Sill2009].
+The only differences between FWLS and the simple Logistic Regression (that performed best) were two additional explanatory variables and four interaction terms.
+The number of variables was not high for the amount of data.
+According to @Peduzzi1996, 10 events per explanatory variable or more avoids the risk of biased estimation of variable coefficients in logistic regression.
+The data contained 315 events per explanatory variable.
+Nevertheless, the additional variables must have added misleading noise to the model.
+The performance of the models so far has been concerned with AUC and profit, however insight into the differences in model performance may be gained by reviewing TP and FP statistics.
+
+**Comparison of Confusion Matrix Statistics**
+
+```{r TP_plot}
+conf.plot <- readRDS("report_data/conf_plot.rds")
+
+#if model.pred = loss, this is TP
+conf.plot$model.pred <- ifelse(conf.plot$actual == "loss" & 
+                                       conf.plot$model.pred == "loss", "TP", "FN")
+# reorder levels in method to suit TP 
+conf.plot$method <- factor(conf.plot$method, 
+                           levels= conf.plot %>% filter(actual == "loss", model.pred == "TP") %>%
+                                   arrange(desc(pc)) %>% .$method)
+
+ggplot(conf.plot %>% filter(actual == "loss"), 
+       aes(x = method, y = pc, fill = model.pred)) +
+        geom_bar(stat = "identity", position = "stack", alpha = 0.7) +
+        scale_fill_manual(values = c('dodgerblue2','red2'),
+                          labels = c("FN: predicted as profit", "TP: predicted as loss"),
+                          name = "Confusion \nMatrix Reference") +
+        labs(x = "Method (Original or Blended)", y = "Percent of All Loss-Making Projects (%)",
+             title = "Breakdown of True Positives and False Negatives for Loss-Making Jobs") +
+        scale_x_discrete(labels = c("average" = "Simple Av",
+                                      "orig.boost"= "Orig. BT",
+                                      "orig.log" = "Orig. LR",
+                                      "orig.rf" = "Orig. RF",
+                                      "simp.log" = "Simple LR",
+                                      "simp.boost" = "Simple BT",
+                                      "comp.boost"= "Complex BT",
+                                      "comp.rf"= "Complex RF",
+                                      "FWLS" = "FWLS")) +
+        theme(axis.text.x=element_text(angle=30, hjust=1))
+
+```
+
+```{r TN_plot}
+
+conf.plot <- readRDS("report_data/conf_plot.rds")
+
+#if model.pred = loss, this is TP
+conf.plot$model.pred <- ifelse(conf.plot$actual == "profit" & 
+                                       conf.plot$model.pred == "profit", "TN", "FP")
+# reorder levels in method to suit TP 
+conf.plot$method <- factor(conf.plot$method, 
+                           levels= conf.plot %>% filter(actual == "profit", model.pred == "TN") %>%
+                                   arrange(desc(pc)) %>% .$method)
+
+conf.plot$model.pred <- factor(conf.plot$model.pred, 
+                           levels= c("TN","FP"))
+levels(conf.plot$model.pred)
+ggplot(conf.plot %>% filter(actual == "profit"), 
+       aes(x = method, y = pc, fill = model.pred)) +
+        geom_bar(stat = "identity", position = "stack", alpha = 0.7) +
+        scale_fill_manual(values = c('dodgerblue2','red2'),
+                          labels = c("TN: predicted as profit", "FP: predicted as loss"),
+                          name = "Confusion \nMatrix Reference") +
+        labs(x = "Method (Original or Blended)", y = "Percent of All Profitable Projects (%)",
+             title = "Breakdown of True Negatives and False Positives for Profitable Jobs") +
+        scale_x_discrete(labels = c("average" = "Simple Av",
+                                      "orig.boost"= "Orig. BT",
+                                      "orig.log" = "Orig. LR",
+                                      "orig.rf" = "Orig. RF",
+                                      "simp.log" = "Simple LR",
+                                      "simp.boost" = "Simple BT",
+                                      "comp.boost"= "Complex BT",
+                                      "comp.rf"= "Complex RF",
+                                      "FWLS" = "FWLS")) +
+        theme(axis.text.x=element_text(angle=30, hjust=1))
+
+```
+
+**Alternate Interpretations of Blended Model Results**
+
 Alternative strategies exist for how to deal with projects that are marked with a probability of loss greater than `r best.t`.
-For example, contracts for these projects could be changed to hourly rate contracts without exceptions.
+For example, contracts for these projects could be changed to hourly rate contracts without exception.
 This means that each hour booked on the project will definitely be charged to the client for each stage of the project.
@@
 However, this method seems less practical as the case study projects above `r best.t` would need to increase fees by 46% in order to generate a 15% profit.
-A proposed fee increase of that magnitude may not be as acceptable as an hourly rate contract depending on the project and client.
-Contracts could be adjusted in many other creative ways to reduce risk in high probability projects, and the threshold of `r best.t` was optimised based on the accept/reject strategy.
-Different courses of actions could be developed for specific clients or project sizes via analysis of threshold scenarios.
-
-Because of the positive AUC results, it was expected the models would translate into financial benefits for the company.
-The results look reasonable, profit gains may be high enough to trigger further cost-benefit analyses for introducing the model into business decisons.
-It was surprising that, given the similar AUC results, simple model averaging signficantly (?) outperformed the blended logistic regression model's profit curve.
-This means that model averaging predicted projects with probabilities closer to 1 (loss-making jobs) with more accuracy than blended logistic regression.
-This may be because ...
-ROC curves.
-
-### Limitations and Future Work
-
-The improvements in profit sound promising, however it is important to consider possible limitations in the data which is discussed in the next section.
-One of the unknown factors in translating this research into business practises is how well users can guess the project variables.
-The data used to develop the model was obtained from the company's records of how projects played out and there are no records of initially expected values of variables.
-For example, if a project lasted 2 years and it was initially expected it would only take 1.5 years, the only data the company retains is that the project lasted two years.
+A proposed fee increase of that magnitude may not be as palatable as an hourly rate contract depending on the project and client.
+Contracts could be adjusted in many other creative ways to reduce risk in high probability projects.
+The threshold of `r best.t` was optimised based on the accept/reject strategy but different courses of actions could be developed for specific clients or project sizes via similar analyses of threshold scenarios.
+
+To conclude, because of the promising AUC results, it is logical that the models would translate into financial benefits for the company.
+The `r (best.mean-100)`% improvement in profits looks reasonable, and may be high enough to trigger further cost-benefit analyses before finally introducing the model into business decisons.
+Further analyses are recommended before implementation of the model and are described in the section below.
+
+## Limitations and Future Work
+
+The improvements in profit sound promising, however it is important to consider possible limitations in the data and implementation process, which in turn motivates future research.
+These include how accurately users of the algorithm would be capable of guessing input values before onset of a project and successful user uptake of the model, which was identified as a problem in the literature.
+
+First, succesfully estimating project variables is critical in translating this research into tangible business benefits.
+The research data was obtained from the company's records of *completed* projects and initial expected values of variables were not documented.
+For example, if a project lasted 2 years and it was initially expected to take 1.5 years, the only data the company retains is that it lasted two years.
 This begs the question, how well could a user predict the variables before beginning a project?
-Uncertainty has been managed to a degree by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e.
+Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e.
 "1 day - 3 weeks", "3 weeks - 2.5 months", etc).
-However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables.
+However, two remaining categories: team size and percent of project completed by a professional level employees were left as continuous variables.
 It would be interesting to test how well business managers could guess these four variables at the start of a project.
 
-The process of guessing varibles at the beginning of a project and then waiting for the project to complete is a long checking process.
-However, it would be worthwhile integrating this data into the system's pre-existing data collecting processes.
+The process of guessing varibles at the beginning of a project and then comparing estimates upon project completion is a long checking process.
+However, it would be worthwhile integrating this practice into the system's pre-existing data-collecting processes.
 When a new project is entered into the database, it could be required that the project manager guesses the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent to be completed by professionals.
@@
 It was for this reason the case study only tested models that provided insight into variable relationships (ensemble trees and logistic regression) as opposed to neural networks and SVM's that operate as 'black box' predictors.
-It is envisioned that a future user interface for this algorithm would display variable relationships such as this partial dependency plot produced by a boosted tree model that displays the probability of losing money vs.
+It is envisioned that a future user interface for this algorithm would display variable relationships such as this partial dependency plot produced by a boosted tree model.
+It displays the probability of a loss-making job vs.
 the team size:
@@
 ![](images/app/NN_wharf.png)
+
+<!-- The Discussion section: -->
+
+<!-- comments on your results; -->
+<!-- explains what your results mean; -->
+<!-- interprets your results in a wider context; indicates which results were expected or unexpected; -->
+<!-- provides explanations for unexpected results. -->
+<!-- The Discussion should also relate your specific results to previous research or theory.
+You should point out what the limitations were of your study, and note any questions that remain unanswered.
+The Discussion CAN also include Conclusions/Future Research.
+-->
 

### 2016-06-27 financial.rmd
started TP FP

--- 2016-06-10 financial.rmd
@@
 
-profit <- function(df = all10mice, method = c("average","simp.log","orig.log","orig.rf","orig.boost"), 
-                   seed = 300, 
+profit <- function(df = all10mice, method = c("average","simp.log","simp.boost","orig.log","orig.rf","orig.boost"), 
+                   seed = 400, 
                    k.folds = 5, # number of folds in initial model.blend generation
@@
 
+        #store results of three original models in list for each iteration
+        joined <- assemble(methods = c("log", "rf", "boost"), df.assem = df, x.folds.assem = 5,
+                    iterations = ceiling(tot.tests/k.folds.2))
+
+        # saveRDS(joined, "/Users/yam/Documents/github/consult/finalwriteup/report_data/joined.rds")
+        # run blended models on each of the 100 outputs from original models stored
+        # in joined list. Store as new list 'compile.results' for each blending method
         for(m in 1:ceiling(tot.tests/max.iter)){
@@
 
-        # create complete set
-        df.blend = assemble(methods = c("log","rf","boost"), df.assem = df, x.folds.assem = 5,
-                     manual.seed.assem = sample(1:1000,1))
+        # create complete set using mth list item in joined
+        df.blend = full_join(df %>% mutate(rown = 1:nrow(df)), joined[[m]], 
+                                by = c('rown'='rownames'))
+        df.blend <- df.blend %>% select(-rown)
+
+        # df.blend = assemble(methods = c("log","rf","boost"), df.assem = df, x.folds.assem = 5)
 
@@
 
+        # saveRDS(df.blend, "C:/Users/n9232371/Documents/github/consult/dfblend.rds")
+        # df.blend <- readRDS("/Users/yam/Documents/github/consult/dfblend.rds")
+
         saveit <- iter.auc(mult = max.iter/k.folds.2, x.folds.iter = k.folds.2,
-                           methods = c("orig.log","average","simp.log", "orig.rf","orig.boost"),
+                           methods = c(
+                                  # "simp.log","average",
+                                  # "orig.log","orig.rf","orig.boost",
+                                  # "comp.rf","comp.boost","FWLS",
+                                   "simp.boost"
+                                  ),
                            df.iter = df.blend)
@@
 
-        # saveRDS(compile.results, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
+        # saveRDS(compile.results, "/Users/yam/Documents/github/consult/finalwriteup/report_data/finance_results_simpboost.rds")
         # compile.results <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
@@
         all6 <-read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+        #iOS
+        all6 <-read.csv('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
         all6 <- all6 %>% select(mlsto, balance.mlsto) %>% distinct()
@@
         new.df$threshold = threshold 
-        prof.results = list(new.df, new.df, new.df, new.df, new.df)
+        prof.results = list(new.df)
         names(prof.results) = names(compile.results)
@@
 
-        assign(prof.results, "prof.results", envir = .GlobalEnv)
+        assign("prof.results", prof.results, envir = .GlobalEnv)
 
 }
+
+#function below assigns results to 'prof.results' object in global env
+profit(df = all10mice, method = c("average","simp.log",'simp.boost', "orig.log","orig.rf","orig.boost"),
+                   seed = 300,
+                   k.folds = 5, # number of folds in initial model.blend generation
+                   max.iter = 5, #max number of iterations from initial model.blend generation
+                   k.folds.2 = 5, # number of folds for second step blended model development
+                   tot.tests = 500 # total number of results to create
+                   )
 
@@
 
-# saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance.rds" )
+saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds" )
+saveRDS(plot.results,"/Users/yam/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds" )
+
+bind <- bind %>% filter(!method %in% 'FWLS')
+bind <- rbind(bind, plot.results)
 
@@
 #power calc - where should I do it from?
+
+```
+
+Check for significant difference between samples
+
+```{r t_test}
+plot.results
+
+index = c(2:8)
+pvals = data.frame("method" = levels(plot.results$method %>% as.factor)[c(2:8)], 
+                   "threshold" = 
+                   'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(auc.150[[1]], 
+                                  auc.150[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
 
@@
 
-simp.l <- compile.results[['average']]
-simp.l$mean.iter <- rowMeans(simp.l[,2:11])
+simp.l <- compile.results[['simp.log']]
+simp.l$mean.iter <- rowMeans(simp.l[,2:101])
 
-saveRDS(simp.l, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_average_model.rds")
+saveRDS(simp.l, "/Users/yam/Documents/github/consult/finalwriteup/report_data/finance_average_model.rds")
 

### 2016-06-27 model_blend.rmd
started TP FP

@@
 #setwd in onedrive data
-write.csv(all10.blend, '~/OneDrive/shared files/Bligh Tanner/masters/data/all10blend.csv')
+write.csv(all10.blend, '~/OneDrive/shared files/Bligh Tanner/masters/data/all10blend2.csv')
+# all10.blend <- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10blend2.csv')[,-1]
 # write.csv(all10.blend, 'all10blend2.csv')
@@
 
+Now narrow down which interaction terms to include
+```{r}
+fit2 <- aov(b.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + no.users*norm.log + b.inv.log*norm.log + Business*norm.log + JD.Second*norm.log + b.timespan.cbrt*norm.rf + no.users*norm.rf + b.inv.log*norm.rf + Business*norm.rf + JD.Second*norm.rf + b.timespan.cbrt*norm.boost + no.users*norm.boost + b.inv.log*norm.boost + Business*norm.boost + JD.Second*norm.boost + Discipline*norm.boost + Discipline*norm.log + Discipline*norm.rf,
+                                   data = all10.blend)
+
+```
+
 ```{r}
 
-formula.log <- "f.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost"
+formula.log <- "f.rpdol ~ norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + b.inv.log*norm.log + b.inv.log*norm.rf + Business*norm.boost"
 
@@
 formula.sboost <- "b.rpdol ~ norm.log + norm.rf + norm.boost"
-
-#formula to add interaction terms for logistic regression
-metafeatures = strsplit(formula.log, split = c("\\s\\~\\s|\\s\\+\\s"))[[1]]
-methods = metafeatures[7:9]
-metafeatures = metafeatures[2:6]
-
-for(j in seq_along(methods)){
-        for(m in seq_along(metafeatures)){
-                formula.log <- paste(formula.log, " + ", metafeatures[m], "*", methods[j], sep = "")
-        }
-}
 
@@
 iter.auc(mult = 20, x.folds = 5, seedy = 100, df = all10.blend,
-         methods = c("orig.log","log","average", "boost","rf", "simp.log","simp.boost"))
+         methods = c("orig.log","log","average", "comp.boost","comp.rf", "simp.log","simp.boost"))
+iter.auc(mult = 20, x.folds = 5, seedy = 100, df = all10.blend,
+         methods = c("log"))
 # iter.auc(mult = 20, x.folds = 5, seedy = 100, df = all10.blend,
@@
 # no.boost <- auc.list
-# saveRDS(auc.list, "/Users/yam/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
-auc.list <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
+saveRDS(auc.list, "/Users/yam/Documents/github/consult/finalwriteup/report_data/blend_auc.rds")
+auc.list <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/blend_auc1.rds")
 # auc.list <- append(auc.list, no.boost)

### 2016-06-27 functions.R

@@ -110,10 +110,9 @@ n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = '
 # perform 5-fold cross validation on train set to create complete train set of predictions for a method
 # output is a vector of predictions for the method
 
-full.predict <- function(df.pred = all10mice, method = "log", x.folds = 5, manual.seed = 100){
+full.predict <- function(df.pred = all10mice, method = "log", x.folds = 5){
         
         #set up 10 fold cv using caret function
-        set.seed(manual.seed)
         folds <- createFolds(df.pred$f.rpdol, k = x.folds, list = T, returnTrain = F)
         #now have folds[[1]] through folds[[10]] list of rows to exclude
         
@@ -176,36 +175,42 @@ full.predict <- function(df.pred = all10mice, method = "log", x.folds = 5, manua
         
 }
 
-#assemble dataset with individual model results as columns
+# assemble dataset with individual model results as columns
+# stores columns of original methods results in a list - 'iterations' long
 assemble <- function(methods = c("log", "rf", "boost"), df.assem = all10mice, x.folds.assem = 5,
-                     manual.seed.assem = 100){
+                     iterations = 100){
         
-        joined = NULL
+        joined = vector("list", iterations)
         
+        for(n in 1:iterations){
+        
+                print(
+                        paste("n =", n, "of", iterations,"\n", sep = ' ') %>% cat()
+                )
+                
         #create empty vectors  - one for each method
         for(i in seq_along(methods)){
-                temp = full.predict(df = df.assem, method = methods[i], x.folds = x.folds.assem,
-                                    manual.seed = manual.seed.assem)
-                if(is.null(joined)){
-                        joined = temp
+                temp = full.predict(df.pred = df.assem, method = methods[i], x.folds = x.folds.assem)
+                if(is.null(joined[[n]])){
+                        joined[[n]] = temp
                 } else{
-                        joined = full_join(joined, temp)
+                        joined[[n]] = full_join(joined[[n]], temp)
                 }
                 
         }
+                
+        }
         
-        df.blend = full_join(df.assem %>% mutate(rown = 1:nrow(df.assem)), joined, 
-                                by = c('rown'='rownames'))
-        df.blend <- df.blend %>% select(-rown)
-        
-        return(df.blend)
+        return(joined)
         
 }
 
+# testing <- assemble(methods = c("log", "rf", "boost"), df.assem = df, x.folds.assem = 5,
[…]

### 2016-06-28 TPR_FPR.rmd
finished TP

@@
 if(.Platform$OS.type == 'windows'){
-        # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
+        setwd("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/")
+        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/")
 } else{
@@
 ```{r}
-
+if(.Platform$OS.type == 'windows'){
+best.fin<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/best_fin.rds")
+plot.fin <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds")
+} else{
 best.fin<- readRDS("~/Documents/github/consult/finalwriteup/report_data/best_fin.rds")
 plot.fin <- readRDS("~/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds")
+}
+
 head(plot.fin)
@@
 #may need to compile all my .. compile.results
+if(.Platform$OS.type == 'windows'){
+CR.av_simplog<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_av_simplog.rds")
+CR.comprf<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_comprf.rds")
+CR.FWLS<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_FWLS.rds")
+CR.simpboost<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_simpboost.rds")
+CR.compboost<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_FWLScomp_boost.rds")
+joined <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/joined.rds")
+
+} else{
 CR.av_simplog<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_av_simplog.rds")
@@
 CR.compboost<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_FWLScomp_boost.rds")
+joined <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/joined.rds")
+
+}
+
 CR.compboost[['FWLS']] <- NULL
@@
 #now want to cbind optimal original results, stored in 'joined'
-joined <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/joined.rds")
 key.oCR <- iter.key %>% filter(grepl('orig', method))
@@
 ```
+
+Investigate which jobs were deemed unprofitable
+Use simple logistic regression blend
+
+```{r}
+if(.Platform$OS.type == 'windows'){
+        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
+        all6 <-read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+
+} else{
+        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
+        all6 <-read.csv('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+
+}
+
+#iOS
+all6 <- all6 %>% select(mlsto, balance.mlsto) %>% distinct()
+all6$mlsto <- as.character(all6$mlsto)
+df.blend = left_join(df.blend, all6)
+
+#merge simp.log
+head(new.CR)
+all10mice<- cbind(all10mice, new.CR %>% select(simp.log))
+
+```

### 2016-06-28 finalwriteup/draft.rmd
finished TP

@@
 Nevertheless, the additional variables must have added misleading noise to the model.
-The performance of the models so far has been concerned with AUC and profit, however insight into the differences in model performance may be gained by reviewing TP and FP statistics.
+The performance of the models so far has been concerned with AUC and profit, however insight into the differences in model performance may be gained by reviewing confusion matrix statistics such as TPR and FPR.
 
 **Comparison of Confusion Matrix Statistics**
+
+To investigate method performance further, confusion matrix statistics were employed.
+For each method, 100 full data sets of results were generated via 5-fold cross validation.
+Each data set of probability results differed to the next because of the random division into the five-fold data sets used for training and testing.
+Nine full data sets of results was chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
+These were deemed a 'typical' set of predictions for each method.
+Then, the confusion matrix for each set of results at their respective thresholds could be calculated.
+The plot below shows the breakdown of predictions by each method on actual loss-making projects.
+If the project was predicted correctly as loss-making, this was a true positive, and the job was rejected (in the fabricated scenario).
+If it was predicted incorrectly as a profit-making job, this was a false negative, and the project was completed, resulting in a loss for the business.
 
@@
         geom_bar(stat = "identity", position = "stack", alpha = 0.7) +
-        scale_fill_manual(values = c('dodgerblue2','red2'),
+        scale_fill_manual(values = c('red2','dodgerblue2'),
                           labels = c("FN: predicted as profit", "TP: predicted as loss"),
@@
 
+Choosing the right probability threshold point is about finding the right balance point in an imperfect model.
+In this case, projects above a threshold point will include profitable and loss-making jobs.
+The optimal point is where the savings from rejecting loss-making jobs (TP) outweighs the lost profits from rejecting profitable jobs (FP) the most.
+All models were capable of different TP and FN ratios at alternate threshold points.
+
+The plot above shows the Complex Boosted Tree Blend has the highest TP rate followed by the Simple Averaged Blend and the Simple Logistic Regression Blend.
+All of these TP rates are below 25% which showed it was not worth catching a large percentage of loss-making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
+This implies the models were not able to very cleanly divide loss-making jobs from profitable jobs.
+
+A summary of predictive success on actual profitable jobs is shown below at each method's optimal threshold point.
+If a job was predicted to be profitable, and was therefore accepted, it was classified TN.
+If a job was loss-making but labelled profitable, the business would complete the job and make a loss (FN).
+
 ```{r TN_plot}
@@
                            levels= c("TN","FP"))
-levels(conf.plot$model.pred)
+
 ggplot(conf.plot %>% filter(actual == "profit"), 
@@
 
+The complex boosted tree method had by far the highest TP rate, but also had a high FP rate.
+The increased number of FP (i.e.
+rejected profitable jobs) offset the more accurate rejection of loss-making jobs so much that the simple averaged blend and simple logisitc regression blend outperformed it.
+The afore-mentioned simple blends performed almost identically, with a higher than average TP rate and average FP rate incomparison to other methods.
+Again, it is surprising FWLS did not produce a better profit curve as its TPR was almost on-par with the simple blends and has a better FP rate than the simple blends.
+The additional variables in FWLS seem to have changed the probabilities of jobs in a way that has reordered them less effectively.
+Perhaps in terms of the size of the losses and profits.
+
+All individual original methods used substantially lower TP rates at their optimal thresholds, yet still had higher FP rates than the other methods.
+This demonstrates how above their threshold points, there was a more equal ratio of actual profitable to actual loss-making jobs in comparison to the blended methods.
+Blending the individual methods has made use of their individual strengths and seems to have clustered the probabilities of the actual loss making jobs more accurately towards 1, and the profitable jobs towards 0.
+This allowed for higher TP rates with lower FP rates at the blended methods' optimal thresholds.
+
 **Alternate Interpretations of Blended Model Results**

### 2016-06-28 financial.rmd
finished TP

@@
 
-results.1 <- profit.cv(df.string = "all10mice", method = "log", seed = 400, k.folds = 8)
-results.2 <- profit.cv(df.string = "all10mice", method = "boost", seed = 400, k.folds = 8)
+results.1 <- profit.cv(df.string = "all10mice", method = "log", seed = 400, k.folds = 5)
+results.2 <- profit.cv(df.string = "all10mice", method = "boost", seed = 400, k.folds =5)
 results.3 <- profit.cv(df.string = "all10mice", method = "rf", seed = 300, k.folds = 3)
@@
         # load balances
+        if(.Platform$OS.type == 'windows'){
         all6 <-read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+        } else{
+        all6 <-read.csv('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+        }
         #iOS
-        all6 <-read.csv('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
         all6 <- all6 %>% select(mlsto, balance.mlsto) %>% distinct()

### 2016-06-28 TPR_FPR.rmd
save TPR before changes

@@
 library('stringr')
+library('knitr')
 
@@
 best.fin<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/best_fin.rds")
-plot.fin <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds")
+plot.fin <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100_4.rds")
 } else{
 best.fin<- readRDS("~/Documents/github/consult/finalwriteup/report_data/best_fin.rds")
-plot.fin <- readRDS("~/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds")
+plot.fin <- readRDS("~/Documents/github/consult/finalwriteup/report_data/plot_finance_100_4.rds")
 }
@@
 CR.compboost<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_FWLScomp_boost.rds")
+# CR<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
 joined <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/joined.rds")
@@
 
-saveRDS(conf.plot, "~/Documents/github/consult/finalwriteup/report_data/conf_plot.rds")
+# saveRDS(conf.plot, "~/Documents/github/consult/finalwriteup/report_data/conf_plot.rds")
 
@@
 all6$mlsto <- as.character(all6$mlsto)
-df.blend = left_join(df.blend, all6)
+all10mice = left_join(all10mice, all6)
+head(all10mice)
 
@@
 all10mice<- cbind(all10mice, new.CR %>% select(simp.log))
+all10mice$f.rpdol <- as.factor(all10mice$b.rpdol)
+levels(all10mice$f.rpdol)[levels(all10mice$f.rpdol) == "0"] <-"profit"
+levels(all10mice$f.rpdol)[levels(all10mice$f.rpdol) == "1"] <-"loss"
+
+#for simp.log 0.6 is the threshold, so summarise everything above 0.6
+all10mice %>% filter(simp.log >= 0.6) %>% summary
+reject <- all10mice %>% filter(simp.log >= 0.6)
+reject$compare <- "reject"
+
+#want to understnad the comparison of perentages between full dataset and rpob>0.6 dataset
+all10mice$compare <- "full"
+all10mice <- rbind(all10mice, reject)
+
+saveRDS(all10mice, '/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/reject_plot.rds')
+
+#create % column for factor variable
+# count number of cases for each
+
+cols = c(1:11,13,14,18,19,22)
+
+#wot function saves plots to folder in onedrive
+wot <- function( compare.cols = compare.cols){
+
+        for (i in seq_along(compare.cols)){
+        print(i)
+
+        if(class(all10mice[,compare.cols[i]]) %in% c('numeric','integer')){
+                p <- ggplot(all10mice, aes_string(x = names(all10mice)[compare.cols[i]])) + 
+                        geom_density() +
+                        facet_wrap(~compare)
+        }
+
+        if(class(all10mice[,compare.cols[i]]) %in% c('factor')){
+                p <- ggplot(all10mice, aes_string(x = names(all10mice)[compare.cols[i]])) + 
+                        geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +
+                        facet_wrap(~compare) + 
+                        theme(axis.text.x=element_text(angle=30, hjust=1)) +
+                        scale_y_continuous(labels = scales::percent)
+
+        }
+                ggsave(p, filename = paste(names(all10mice)[compare.cols[i]], 'compare.png', sep = ""), 
+                       path = '/Users/yam/OneDrive/shared files/Bligh Tanner/masters/barplots/compare_simplog/', 
+                       width = 5, height = 4)
+}}
+
+wot(compare.cols = cols)
+
+#plot of profit size broken up by job
+head(reject)
+
+ggplot(reject, aes(x = f.rpdol, y = abs(balance.mlsto), fill = mlsto)) +
+        geom_bar(stat = "identity") + 
+        theme(legend.position="none") +
+        scale_fill_manual(values = rep(c('blue', 'yellow'), 54)) + 
+        scale_y_continuous(labels = scales::dollar)
+```
+
+```{r}
+
+groups <- (rbinom(32, n = 5, prob = 0.4))
+tapply(groups, groups, sum) #- is almost the same as
+table(groups)
+
+ind <- list(c(1, 2, 2), c("A", "A", "B"))
+table(ind)
+tapply(1:3, ind) #-> the split vector
+tapply(1:3, ind, sum)
 

### 2016-06-29 finalwriteup/draft.rmd
re-ran models for 2009.238

@@
 
+## Blended Models Discussion
+
+Trials of different methods of blending model results demonstrated that again, the simplest methods worked best with the data.
+In this case, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
+It was plausible that simple blending models would perform better than complex methods because the data was not big enough for complex models to learn intricate patterns.
+This was observed in the single method models (previous chapter) when Logistic Regression performed as well as Random Forests and Boosted Trees.
+The literature does not extensively cover blended machine learning methods however their success in the 2009 Netflix competition generated some publications.
+The Netflix data set comprised of almost 3,000,000 observations, so the size of the data could have enabled the success of complex blending methods such as FWLS [@Sill2009].
+That being the case, it was still surprising that simple averaging of probabilities from the individual models performed almost as well as blending results using Logistic Regression.
+It was expected that Logistic Regression would have an advantage over model averaging due to its ability to weight the prediction variables while maintaining a simple model structure.
+
+The Logistic Regression blending model did indeed weight predictions from the original models differently.
+If the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model prediciting a loss making job increased by 37.4%.
+Similarly when probability predictions from the Random Forests model increased by 0.1, the odds of the blended model predicting a loss increased by 25.8%.
+On the other hand, for the same increase in the original Boosted Tree model, the odds of the blended Logistic Regression model predicting a loss increased by only 2.6%.
+This contrasted with the alternative simple model that averaged the probabilities from each original model (i.e.
+weighted them equally) but still managed to produce similar AUC results.
+It was not clear why this was the case, however both simple models were compared again in the analysis of the impact each model could have on the company's overall profits.
+The analysis of overall profits did not directly relate to AUC measurements, so a range of models were carried forward for comparison.
+These included the three original individual models, three simple blended models and three complex blended models.
+
 ## Profit Analysis
@@
 Profit curves for the three simple blended models as well as the three individual models were calculated.
-However, the total profit coordinates were based on probability outputs in test datasets.
+However, the total profit coordinates were based on probability outputs in test data sets.
 This means that the profit curve for the simple logistic regression blend (for example) would be different for different training/testing data divisions.
@@
 The plot below illustrates the results of each model at each threshold with a line joining the mean value at each threshold for each method.
-The grey ribbon illustrates the 95% confidence interval around the mean at each threshold for the 100 models.
+The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold for the 100 models.
 
@@
 
-finance.plot = readRDS("report_data/plot_finance_100_3.rds" )
+finance.plot = readRDS("report_data/plot_finance_100_4.rds" )
 finance.plot$method <- factor(finance.plot$method, 
@@
 
-The two blended models outperformed the individual models as shown in the plot above with higher profit ratios as well as tighter confidence intervals.
+Two blended models clearly outperformed the individual models as shown in the plot above with higher profit ratios as well as tighter confidence intervals.
 The table below denotes each models' optimal thresholds and the corresponding increase in total profits:
@@
 The simple logistic regression blend performed best with the highest mean profit ratio at `r best.mean`% with a standard deviation of `r best.sd`%.
-This means that for the simple blended logistic regression model, if all jobs above the probability threshold `r best.fin %>% filter(method == "simp.log") %>% .$threshold` were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profit data.
+This means that for the simple blended logistic regression model, if all jobs above the probability threshold `r best.t` were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profits.
 
@@
 
-## Extended Analysis Discussion
-
-### Blended Models
-
-Trials of different methods of blending model results demonstrated that again, the simplest methods worked best with the data.
-In this case, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
-
-It was plausible that simple blending models would perform better than complex methods because the data was not big enough for complex models to learn intricate patterns.
-This was observed in the single method models (previous chapter) when Logistic Regression performed as well as Random Forests and Boosted Trees.
-The literature does not extensively cover blended machine learning methods however their success in the 2009 Netflix competition generated some publications.
-The Netflix dataset comprised of almost 3,000,000 observations, so the size of the data could have facilitated the success of complex blending methods such as FWLS [@Sill2009].
-That being the case, it was still surprising that simple averaging of probabilities from the individual models performed as well as blending results using Logistic Regression.
-It was expected that Logistic Regression would have an advantage over model averaging due to its ability to weight the prediction variables while maintaining a simple model structure.
-
-The Logistic Regression blending model did indeed weight predictions from the original models differently.
-If the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model prediciting a loss making job increased by 37.4%.
-Similarly when probability predictions from the Random Forests model increased by 0.1, the odds of the blended model predicting a loss increased by 25.8%.
-On the other hand, for the same increase in the original Boosted Tree model, the odds of the blended Logistic Regression model predicting a loss increased by only 2.6%.
-This contrasted with the alternative simple model that averaged the probabilities from each original model (i.e.
-weighted them equally) but still managed to produce similar AUC results.
-It was not clear why this was the case, however both simple models were compared again in the analysis of the impact each model could have on the company's overall profits.
-The analysis of overall profits did not directly relate to AUC measurements, so a range of models were carried forward for comparison.
-These included the three original individual models, three simple blended models and two complex blended models.
-
-### Profit Curve Analysis - Discussion
+## Profit Curve Analysis - Discussion
 
@@
 
-The simple average blend and logistic regression blended models outperformed the complex blended models, which is rational in terms of theit AUC values, which were significantly higher than the others.
-The simple average of logistic regression, random forests, and boosted trees results produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression that blended the same three variables.
-The models produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t` for the blended Logistic Regression and `r best.fin %>% filter(method == 'average') %>% .$threshold` for simple averaging.
-To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r (best.mean-100)/100*1000000` by rejecting all jobs with a probability higher than `r best.t` (blended logistic regression).
+The simple average blend and logistic regression blended models outperformed the complex blended models, which is rational because their AUC values were significantly higher than the others.
+The simple average of Logistic Regression, Random Forests, and Boosted Trees results produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression that blended the same three variables.
+The blended Simple Logistic Regression produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`.
+To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r comma_format() ((best.mean-100)/100*1000000)` by rejecting all jobs with a probability higher than `r best.t`.
 In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.
@@
 FWLS and randomForests performed more poorly than the individual methods while Boosted Trees produced a marginally higher profitability ratio (`r best.fin %>% filter(method == "comp.boost") %>% .$mean_profit_ratio`) but the lower bound of its confidence interval was `r best.fin %>% filter(method == "comp.boost") %>% .$lower_bound_95CI`, i.e.
-barely above original profits.
+not above original profits.
 
@@
 The Boosted Tree's lower bound confidence interval was `r best.fin %>% filter(method == "simp.boost") %>% .$lower_bound_95CI`% at its optimal threshold point.
-This is above 100%, so it is highly likely the algorithm would improve profits, but the margin was not high in comparison to the other two simple methods.
+This is not above 100%, and again does not make a convincing argument for the case-study to adopt it.
 
 It is not clear why simple linear models and the averaging model outperformed ensemble tree blending methods.
-It may be that there was not enough data for the ensemble trees to adequately learn the complex series of rules they develop, although this should be compensated for in the tuning phase.
+It may be that there was not enough data for the ensemble trees to adequately learn the complex series of rules they develop as previously stated.
 It was also surprising FWLS was ineffective given the success it had in the Netflix Prize competition [@Sill2009].
@@
 According to @Peduzzi1996, 10 events per explanatory variable or more avoids the risk of biased estimation of variable coefficients in logistic regression.
-The data contained 315 events per explanatory variable.
+The data contained 315 events per explanatory variable and did not pose that risk.
 Nevertheless, the additional variables must have added misleading noise to the model.
@@
 
-To investigate method performance further, confusion matrix statistics were employed.
+To investigate method performance further, confusion matrix statistics were scrutinised.
 For each method, 100 full data sets of results were generated via 5-fold cross validation.
@@
 In this case, projects above a threshold point will include profitable and loss-making jobs.
-The optimal point is where the savings from rejecting loss-making jobs (TP) outweighs the lost profits from rejecting profitable jobs (FP) the most.
+The optimal point is where the savings from rejecting loss-making jobs (TP) outweigh the lost profits from rejecting profitable jobs (FP) the most.
 All models were capable of different TP and FN ratios at alternate threshold points.
 
-The plot above shows the Complex Boosted Tree Blend has the highest TP rate followed by the Simple Averaged Blend and the Simple Logistic Regression Blend.
-All of these TP rates are below 25% which showed it was not worth catching a large percentage of loss-making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
+The plot above shows the Complex Boosted Tree Blend has the highest TP rate followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
+All of these TP rates are below 25% which shows it was not worth catching a large percentage of loss-making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
 This implies the models were not able to very cleanly divide loss-making jobs from profitable jobs.
 
-A summary of predictive success on actual profitable jobs is shown below at each method's optimal threshold point.
+A summary of predicted results on actual profitable jobs is shown below at each method's optimal threshold point.
 If a job was predicted to be profitable, and was therefore accepted, it was classified TN.
@@
 
-The complex boosted tree method had by far the highest TP rate, but also had a high FP rate.
+The complex Boosted Tree method had the highest TP rate, but also had a high FP rate.
 The increased number of FP (i.e.
 rejected profitable jobs) offset the more accurate rejection of loss-making jobs so much that the simple averaged blend and simple logisitc regression blend outperformed it.
-The afore-mentioned simple blends performed almost identically, with a higher than average TP rate and average FP rate incomparison to other methods.
-Again, it is surprising FWLS did not produce a better profit curve as its TPR was almost on-par with the simple blends and has a better FP rate than the simple blends.
-The additional variables in FWLS seem to have changed the probabilities of jobs in a way that has reordered them less effectively.
-Perhaps in terms of the size of the losses and profits.
-
-All individual original methods used substantially lower TP rates at their optimal thresholds, yet still had higher FP rates than the other methods.
-This demonstrates how above their threshold points, there was a more equal ratio of actual profitable to actual loss-making jobs in comparison to the blended methods.
+The afore-mentioned simple blends performed similarly, with a higher than average TP rate and average FP rate in comparison to other methods.
+The complex Random Forest blend and FLWS both had lower TP rates than the simple blends but higher TN rates.
+In the end, the improved TN rate did not outweight the lower TP rate, i.e.
+the higher rate of accepting profitable jobs did not outweight the lower rate of accurately rejecting loss-making jobs.
+
+All individual original methods used substantially lower TP rates at their optimal thresholds, while their TN rates were spread out.
+This means, as the trialled threshold values moved, and the TP rate improved, the increasing lost profits from FP's was too great to justify a TP rate higher than 4.6%.
 Blending the individual methods has made use of their individual strengths and seems to have clustered the probabilities of the actual loss making jobs more accurately towards 1, and the profitable jobs towards 0.
 This allowed for higher TP rates with lower FP rates at the blended methods' optimal thresholds.
+
+**Review of Jobs Rejected by Simple Logistic Regression**
+
+All below analyses for results of a typical Simple Blended Logistic Regression Model
+
+* ratio of profitable to unprofitable
+
+```{r}
+reject.plot <- readRDS('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/reject_plot.rds')
+
+ggplot(reject.plot, aes(x = f.rpdol)) +
+        geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) + 
+        scale_y_continuous(labels = scales::percent) +
+        labs(x = "Project Performance", 
+             y = "Relative Frequency",
+             title = "Ratio of Profitable vs. Loss-Making Projects")
+
+```
+
+* size of projects in profitable and unprofitable divisions
+
+```{r}
+ggplot(reject.plot %>% filter(compare == "reject"), 
+             aes(x = f.rpdol, y = abs(balance.mlsto), fill = mlsto)) +
+        geom_bar(stat = "identity") + 
+        theme(legend.position="none") +
+        scale_fill_manual(values = rep(c('dodgerblue4', 'darkolivegreen1'), 59)) + 
+        scale_y_continuous(labels = scales::dollar) +
+        labs(x = "Project Performance", 
+             y = "Cumulative Profits/Losses",
+             title = "Magnitude of Profits/Losses from Rejected Projects") +
+        geom_text(data = NULL, x = .5, y = 720000, size = 3, 
+                  colour = "gray60", hjust = 0,
+                  label = "Alternate Colours \nIndicate Different \nProjects")
+
+```
+
+* some plots of variables that vary quite a bit
+    * pc.pro
+
+```{r}
+ggplot(reject.plot, 
+             aes(x = pc.pro)) +
+        geom_density() + 
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) +
+        labs(x = "% of Hours Completed by Professional Level Employee", 
+             y = "Density",
+             title = "Density of Percentage of Hours by Professional Level Employees") 
+```
+
+    * timespan kind of interesting: no 1d - 3wk or 3wk to 2.5 m
+
+```{r}
+ggplot(reject.plot, 
+             aes(x = b.timespan.cbrt)) +
+        geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) +
+        labs(x = "Timespan", 
+             y = "Relative Frequency",
+             title = "Relative Frequency of Project Timespans") + 
+        theme(axis.text.x=element_text(angle=30, hjust=1))
+
+```
+    * return per dollar quite nice
+
+```{r}
+ggplot(reject.plot, 
+             aes(x = return.pdol)) +
+        geom_density() + 
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) +
+        labs(x = "Return per Dollar", 
+             y = "Density",
+             title = "Density Plots of Return per Dollar") +
+        scale_x_continuous(labels = scales::dollar)
+```
+
+Not in a plot, but state:
+    * majority position: mid tech and senior tech disproportionately high
+    * JD.Second: 16 categories originally.
+        * subdivision and sewerage disproportionately high
 

### 2016-06-29 financial.rmd
re-ran models for 2009.238

@@
 
+# #clean up some JD.SEcond categories!
+# summary(all10mice$JD.Second)
+# # all Subdivision sohould be under civBldg_subdiv_sewer
+# all10mice$JD.Second <- as.character(all10mice$JD.Second)
+# all10mice$JD.Second <- ifelse(all10mice$JD.Second == "Subdivision", "civBldg_subdiv_sewer",
+#                               all10mice$JD.Second)
+# all10mice$JD.Second <- as.factor(all10mice$JD.Second)
+# # mlsto 20090.238 should be flood harvest
+# all10mice[all10mice$mlsto == "2009.238",]$JD.Second <- 'flood_h2o_harvest'
+# write.csv(all10mice, '~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
+
 profit <- function(df = all10mice, method = c("average","simp.log","simp.boost","orig.log","orig.rf","orig.boost"), 
@@
 
-        # saveRDS(joined, "/Users/yam/Documents/github/consult/finalwriteup/report_data/joined.rds")
+        # saveRDS(joined, "/Users/yam/Documents/github/consult/finalwriteup/report_data/joined2.rds")
         # run blended models on each of the 100 outputs from original models stored
@@
                            methods = c(
-                                  # "simp.log","average",
-                                  # "orig.log","orig.rf","orig.boost",
-                                  # "comp.rf","comp.boost","FWLS",
+                                  "simp.log","average",
+                                  "orig.log","orig.rf","orig.boost",
+                                  "comp.rf","comp.boost","FWLS",
                                    "simp.boost"
@@
 
-        # saveRDS(compile.results, "/Users/yam/Documents/github/consult/finalwriteup/report_data/finance_results_simpboost.rds")
+        # saveRDS(compile.results, "/Users/yam/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
         # compile.results <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
@@
         new.df$threshold = threshold 
-        prof.results = list(new.df)
+        prof.results = list(new.df,new.df,new.df,new.df,new.df,new.df,new.df,new.df,new.df)
         names(prof.results) = names(compile.results)
@@
 #function below assigns results to 'prof.results' object in global env
-profit(df = all10mice, method = c("average","simp.log",'simp.boost', "orig.log","orig.rf","orig.boost"),
+profit(df = all10mice, method = c("average","simp.log",'simp.boost','FWLS','comp.boost','comp.rf', "orig.log","orig.rf","orig.boost"),
                    seed = 300,
@@
 
-saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds" )
-saveRDS(plot.results,"/Users/yam/Documents/github/consult/finalwriteup/report_data/plot_finance_100_3.rds" )
+saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100_4.rds" )
+saveRDS(plot.results,"/Users/yam/Documents/github/consult/finalwriteup/report_data/plot_finance_100_4.rds" )
 

### 2016-06-29 finalwriteup/draft.rmd
finished rejected jobs analysis

@@
 
-**Highest Profit Curves**
+### Highest Profit Curves
 
@@
 
-**Comparison of All Profit Curves**
+### Comparison of All Profit Curves
 
@@
 
-**Comparison of Confusion Matrix Statistics**
+### Comparison of Confusion Matrix Statistics
 
@@
 This means, as the trialled threshold values moved, and the TP rate improved, the increasing lost profits from FP's was too great to justify a TP rate higher than 4.6%.
-Blending the individual methods has made use of their individual strengths and seems to have clustered the probabilities of the actual loss making jobs more accurately towards 1, and the profitable jobs towards 0.
+Blending the individual methods has made use of their unique strengths and seems to have clustered the probabilities of the actual loss-making jobs more accurately towards 1, and the profitable jobs towards 0.
 This allowed for higher TP rates with lower FP rates at the blended methods' optimal thresholds.
 
-**Review of Jobs Rejected by Simple Logistic Regression**
-
-All below analyses for results of a typical Simple Blended Logistic Regression Model
-
-* ratio of profitable to unprofitable
+### Review of Jobs Rejected by Simple Logistic Regression
+
+The profit curves and TP/FP rates give a good indication of overall performance of the methods, however an analysis of the characteristics of the predicted loss-making projects can provide more detailed insight.
+
+All below plots and comments summarise results from one Simple Blended Logistic Regression Model.
+The results from this model were chosen because it represented the mean for improvement in profits output by this blending method.
+The plot below illustrates how the ratio of profitable to unprofitable jobs compare in the full project population vs.
+the rejected projects.
 
 ```{r}
-reject.plot <- readRDS('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/reject_plot.rds')
+reject.plot <- readRDS('report_data/reject_plot.rds')
 
@@
 
-* size of projects in profitable and unprofitable divisions
+Originally, 25% of projects were loss-making.
+The model rejects a population of projects which are over 60% loss-making.
+This represents a significant swing in proportions, but is not perfect.
+The class imbalance towards profitable jobs in the initial dataset may have weakened the model's ability to classify losses.
+@Saradhi2011 took advantage of SVM's ability to adjust class penalties to address class imbalance, however SVM's were avoided because of their black box predictions in this work and the remaining methods cannot accommodate class penalties.
+
+The plot below illustrates the monetary size of profits and losses from rejected projects.
+The different coloured bands indicate different projects and it can be observed that the magnitude of rejected profits and losses vary greatly.
+They range from profits/losses of just \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% tail(1) %>% round(0)` to losses of \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% head(1) %>% round(-3)`.
 
 ```{r}
-ggplot(reject.plot %>% filter(compare == "reject"), 
+ggplot(reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))), 
              aes(x = f.rpdol, y = abs(balance.mlsto), fill = mlsto)) +
@@
 
-* some plots of variables that vary quite a bit
-    * pc.pro
+The variable 'percentage of hours completed by a professional level employee' had an interesting contrast in the statistical distribution of all projects vs. the rejected projects (refer plot below).
 
@@
 
-    * timespan kind of interesting: no 1d - 3wk or 3wk to 2.5 m
+The algorithm tended to accept projects which were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
+This could either indicate that professional level employees complete projects more efficiently or that the types of projects that do not require technical employees are more profitable.
+
+The distribution of timespans of rejected projects differed significantly to the timespan distribution of all projects.
+The chart below shows that proportionally very few short jobs were rejectd, i.e.
+projects that lasted between 1 day and 2.5 months.
+These projects are likely small and may be easier to estimate fixed fees.
+A much higher relative proportion of projects lasting 1.5 - 3 years were rejected.
+They may face an opposite challenge to the small projects.
+On the other hand, large projects can provide long term security for a small business and it is unlikely the company would want to reject them.
+It may make more sense to adjust the contract structure of large projects that have been highlighted as a risk.
+Alternate uses of the algorithm results are discussed in the next section.
 
@@
 ```
-    * return per dollar quite nice
+
+The statistical distribution of return per dollar in the rejected projects are skewed towards negative returns.
+This means the algorithm has identified a proportionally much higher percentage of loss-making projects which is not surprising given previous results.
 
@@
 
-Not in a plot, but state:
-    * majority position: mid tech and senior tech disproportionately high
-    * JD.Second: 16 categories originally.
-        * subdivision and sewerage disproportionately high
-
-**Alternate Interpretations of Blended Model Results**
+There were a couple more revealing differences between the full set of projects and the rejected projects.
+The first was that rejected projects had disproportionately more projects with mid-level and senior-level technical employees as the main worker on a project.
+The second was that a comparatively high percentage of rejected projects were subdivisions, sewerage, or civil building works.
+
+The detailed analysis of rejected project characteristics gave insight into the characteristics of projects that are likely to be loss-making as well as the accuracy of the algorithm.
+While it is interesting to observe characeristics of rejected projects, a complex combination of all the features contribute to the final probability of loss prediction.
+
+### Alternate Interpretations of Blended Model Results
 
@@
 
+### User Interface and Engagement
+
+* Highlighted as a major problem in previosu cost estimation studies
+    * need to (1) communicate results usefully (2) trust
+* User actions will vary, options listed in previous section: reject project, adjust contract structure
+    * Communicate confidence
+    * future work: probability: bayesian?
+    * simple yes/no, easy, or on a scale of probability 0.6 to 1, but output is not statistically driven. score driven by machine learning algorithms.
+    * future work: which is preferable user chooses
+
+* Shiny App: engage user trust by two strategies
+    * nearest neighbours
+    * model output
+
+Another limitation for these types of algorithms in general, as identified in the literature, is user uptake of the models and a lack of trust in the output [@Akintoye2000].
+It was for this reason the case study only tested models that provided insight into variable relationships (ensemble trees and logistic regression) as opposed to neural networks and SVM's that operate as 'black box' predictors.
+It is envisioned that a future user interface for this algorithm would display variable relationships such as this partial dependency plot produced by a boosted tree model.
+It displays the probability of a loss-making job vs.
+the team size:
+
+![](images/boosted tree/partial_dep_plots/no_users.png)
+
+Running a nearest-neighbour algorithm concurrently with the probability algorithm could provide a cross check of the algorithm's results and engage the user's past experiences in similar projects.
+This could work by presenting a graphical summary of the 4 or 5 most-similar past projects to the new project being assessed by the manager.
+Memories of how these similar projects played out would be triggered as well as communication with other manager's that led the similar projects.
+This would allow the user to connect specific past experience with the probability model output.
+Further development and testing is required to assess how much users trust the algorithm output and how trustworthiness may be improved.
+
+![](images/app/NN_wharf.png)
+
 ## Limitations and Future Work
@@
 
-Another limitation for these types of algorithms in general, as identified in the literature, is user uptake of the models and a lack of trust in the output [@Akintoye2000].
-It was for this reason the case study only tested models that provided insight into variable relationships (ensemble trees and logistic regression) as opposed to neural networks and SVM's that operate as 'black box' predictors.
-It is envisioned that a future user interface for this algorithm would display variable relationships such as this partial dependency plot produced by a boosted tree model.
-It displays the probability of a loss-making job vs.
-the team size:
-
-![](images/boosted tree/partial_dep_plots/no_users.png)
-
-Running a nearest-neighbour algorithm concurrently with the probability algorithm could provide a cross check of the algorithm's results and engage the user's past experiences in similar projects.
-This could work by presenting a graphical summary of the 4 or 5 most-similar past projects to the new project being assessed by the manager.
-Memories of how these similar projects played out would be triggered as well as communication with other manager's that led the similar projects.
-This would allow the user to connect specific past experience with the probability model output.
-Further development and testing is required to assess how much users trust the algorithm output and how trustworthiness may be improved.
-
-![](images/app/NN_wharf.png)
+**updating the model!**
+A lot of data cleaning in past data
+need to build a pipeline of cleaning - only 11 variables now, not too hard
+* re run algorithm monthly? does not take long, maybe 1 minute.
+* repeat validation every few months
 

### 2016-06-30 TPR_FPR.rmd
finished shiny app section

@@
 head(plot.fin)
-?inner_join
 # join best thresholds to plot.fin
@@
 if(.Platform$OS.type == 'windows'){
-CR.av_simplog<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_av_simplog.rds")
-CR.comprf<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_comprf.rds")
-CR.FWLS<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_FWLS.rds")
-CR.simpboost<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_simpboost.rds")
-CR.compboost<- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results_FWLScomp_boost.rds")
+CR <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
 joined <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/joined.rds")
@@
 } else{
-CR.av_simplog<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_av_simplog.rds")
-CR.comprf<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_comprf.rds")
-CR.FWLS<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_FWLS.rds")
-CR.simpboost<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_simpboost.rds")
-CR.compboost<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results_FWLScomp_boost.rds")
-# CR<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
+CR<- readRDS("~/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
 joined <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/joined.rds")
@@
 
-CR.compboost[['FWLS']] <- NULL
-ans <- CR.compboost[['comp.boost']] %>% .$b.rpdol
-CR.compboost[['comp.boost']] <- CR.compboost[['comp.boost']] %>% select(-balance.mlsto, -b.rpdol) 
-#missing simp.boost for now
-# append all lists together
-
-CR <- append(CR.av_simplog, c(CR.comprf, CR.compboost, CR.simpboost, CR.FWLS))
 #flatten list
@@
 key.CR <- iter.key %>% filter(method %in% c('average','comp.boost','comp.rf','FWLS','simp.log', 'simp.boost'))
-new.CR <- data.frame(row = 1:2364)
+new.CR <- data.frame(CR[1:nrow(joined[[1]]),c('balance.mlsto','b.rpdol')])
 for(i in 1:nrow(key.CR)){
@@
 key.oCR <- iter.key %>% filter(grepl('orig', method))
-key.oCR$index <- c(6, 91, 4)
+key.oCR$index <- sapply(key.oCR$iteration, grep, colnames(CR)) - 2
 key.oCR$method <- gsub("orig.","", key.oCR$method)
@@
 # need to append answers.. 
-new.CR <- cbind(new.CR, b.rpdol = ans)
 conf.CR <- new.CR
@@
 # sub all 1's for loss and 0's for profit
-conf.CR <- apply(conf.CR, 2, function(x) ifelse(x < 0.5, "profit", "loss")) %>% as.data.frame
-conf.CR <- apply(conf.CR, 2, function(x) x = factor(names(x), 
-                                               levels=c("profit","loss"))) %>% as.data.frame %>% head
+conf.CR[,2:11] <- apply(conf.CR[,2:11], 2, function(x) ifelse(x < 0.5, "profit", "loss")) %>% as.data.frame
 
@@
 }
+names(confs) <- methods
 
@@
 
-saveRDS(all10mice, '/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/reject_plot.rds')
+# saveRDS(all10mice, '/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/reject_plot.rds')
 
@@
         theme(legend.position="none") +
-        scale_fill_manual(values = rep(c('blue', 'yellow'), 54)) + 
+        scale_fill_manual(values = rep(c('blue', 'yellow'), 59)) + 
         scale_y_continuous(labels = scales::dollar)
@@
 ```
+
+Making Shiny visualisation
+What is the spread in probability values like?
+
+```{r}
+CR$mean <- apply(CR[,3:102], 1, mean) 
+CR$sd <- apply(CR[,3:102], 1, sd) 
+CR <- mutate(CR, conf_95 = sd*1.96*2)
+summary(CR$conf_95)
+# median conf interval for a probability is 0.11, mean is 0.137, just say 0.12 spread
+ggplot(CR, aes(y = conf_95, x = mean)) + geom_point()
+#find mean conf_95 per bin of mean
+CR$mean_bins <- cut(CR$mean, 10)
+ddply(CR, .(mean_bins), summarise, mean_conf = median(conf_95))
+
+# probability outputs for log.reg
+ggplot(new.CR, aes(y = simp.log, x = "potato")) + geom_boxplot()
+max(new.CR$simp.log)
+exp <- new.CR
+all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
+exp <- cbind(exp, return.pdol = all10mice$return.pdol)
+ggplot(exp, aes(x = simp.log, y = return.pdol)) + geom_point()
+
+```

### 2016-06-30 finalwriteup/draft.rmd
finished shiny app section

@@
 The Logistic Regression blending model did indeed weight predictions from the original models differently.
-If the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model prediciting a loss making job increased by 37.4%.
-Similarly when probability predictions from the Random Forests model increased by 0.1, the odds of the blended model predicting a loss increased by 25.8%.
-On the other hand, for the same increase in the original Boosted Tree model, the odds of the blended Logistic Regression model predicting a loss increased by only 2.6%.
-This contrasted with the alternative simple model that averaged the probabilities from each original model (i.e.
+A summary of the resulting coefficients from an example blended Logistic Regression model is shown in the table below.
+
+| Model Results as Variable | Coefficient (Log Odds) | Odds  | Increase in Variable Input | Increases in Odds |
+|---------------------------|------------------------|-------|----------------------------|-------------------|
+| Logistic Regression       | 1.7223                 | 5.60  | 0.1                        | 56%               |
+| Random Forest             | 1.4753                 | 4.37  | 0.1                        | 43.7%             |
+| Boosted Trees             | 2.7943                 | 16.35 | 0.1                        | 163.5%            |
+
+As shown in the table, if the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model prediciting a loss-making job increased by 56.0%.
+The weightings contrasted the other simple model that averaged the original probabilities (i.e.
 weighted them equally) but still managed to produce similar AUC results.
@@
 
-To conclude, because of the promising AUC results, it is logical that the models would translate into financial benefits for the company.
+To conclude, because of the promising AUC results, it was logical that the models translated into financial benefits for the company.
 The `r (best.mean-100)`% improvement in profits looks reasonable, and may be high enough to trigger further cost-benefit analyses before finally introducing the model into business decisons.
-Further analyses are recommended before implementation of the model and are described in the section below.
+Development of the user interface as well as some further data analyses are recommended before implementation of the model and are described in the 'User Interface and Engagement' and 'Limitations and Future Work' sections below.
 
@@
 
-* Highlighted as a major problem in previosu cost estimation studies
-    * need to (1) communicate results usefully (2) trust
-* User actions will vary, options listed in previous section: reject project, adjust contract structure
-    * Communicate confidence
-    * future work: probability: bayesian?
-    * simple yes/no, easy, or on a scale of probability 0.6 to 1, but output is not statistically driven. score driven by machine learning algorithms.
-    * future work: which is preferable user chooses
-
-* Shiny App: engage user trust by two strategies
-    * nearest neighbours
-    * model output
-
-Another limitation for these types of algorithms in general, as identified in the literature, is user uptake of the models and a lack of trust in the output [@Akintoye2000].
-It was for this reason the case study only tested models that provided insight into variable relationships (ensemble trees and logistic regression) as opposed to neural networks and SVM's that operate as 'black box' predictors.
-It is envisioned that a future user interface for this algorithm would display variable relationships such as this partial dependency plot produced by a boosted tree model.
-It displays the probability of a loss-making job vs.
-the team size:
-
-![](images/boosted tree/partial_dep_plots/no_users.png)
-
-Running a nearest-neighbour algorithm concurrently with the probability algorithm could provide a cross check of the algorithm's results and engage the user's past experiences in similar projects.
-This could work by presenting a graphical summary of the 4 or 5 most-similar past projects to the new project being assessed by the manager.
-Memories of how these similar projects played out would be triggered as well as communication with other manager's that led the similar projects.
-This would allow the user to connect specific past experience with the probability model output.
-Further development and testing is required to assess how much users trust the algorithm output and how trustworthiness may be improved.
-
-![](images/app/NN_wharf.png)
+Industry's lack of uptake of cost estimation models was marked as a significant issue in the literature [@Akintoye2000].
+Decision makers did not sufficiently trust the model output which seemed disconnected to their experiences.
+In many studies, case-based reasoning had more traction than more accurate models because users' could justify the output.
+Clearly this is an important aspect of the study to address.
+The user interface should do two things well:
+
+* Communicate the model results practically and clearly
+* Create trust via a connection to the user's experiences
+
+A prototype user interface was developed using the Shiny package in R where a strategy was developed for how the input options and results would be communicated [@shiny].
+
+![](images/app/front_page.png)
+
+The manager selects the charactersitics that describe their new job in the top panel.
+Then once the 'Calculate' button is clicked, the blended Logistic Regression model works behind the scenes to generate a probability score between 0 and 1 where 1 represents a loss-making job.
+This output is not intuitive to understand for a manager because the probability score does not represent a statistical probability, rather a score the algorithm generated using its own measures.
+The optimal threshold dictated that scores below 0.6 should be classified as profitable jobs and scores above 0.6 are loss-making.
+The scores were re-scaled to a more intuitive numbering system where scores between 0 and 0.6 were scaled between 0 and 10, and scores higher than 0.6 were scaled between 0 and -10.
+The shaded confidence interval was calculated from the distributed spread of probability outputs from the 100 models calculated in the Profit Curve Analysis.
+The 100 models gave 100 probability outputs for each project, and the 95% confidence interval for these outputs was calculated.
+A typical spread for the 95% confidence interval is shaded in the graphic which is relative to the 'bin' that the score fell into (i.e.
+6-8).
+This provides an intuitive feel for the results the model's uncertainty.
+
+This graphic is an initial suggestion for how to communicate model results to industry.
+Many variations are possible which could depend on how the business wants to act on the output; for example whether they plan to reject risky projects outright or adjust the fee structure.
+Additional methods built on top of the current model may be able to produce statistical probabilities instead of probability scores.
+This may be achieved using bayesian methods such as bayesian additive regression trees, however it is uncertain whether this would improve practical effectiveness.
+
+Two strategies were adopted to facilitate user trust in the model's output: illustration of the model structure and a nearest neighbour graph.
+Algorithms were originally selected that could reveal their predictive structure by either variable importance and/or variable relationships.
+The first tab in the application displays several variable dependency plots extracted from a typical boosted tree model.
+Understanding the plot would require some training.
+The y-axis represents the probability a project will be loss-making after taking into account average effects from variables other than the x-axis variable.
+
+![](images/app/team_size_dep.png) 
+![](images/app/timespan_dep.png)
+
+Secondly, the nearest neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
+The aim was to provide the user with a cross check of the algorithm's results by finding projects with the most similar features to the one currently being assessed.
+This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'nearest neighbours'.
+Some screen shots of the nearest neigbhour graphic are presented below.
+
+![](images/app/NN_top.png) 
+![](images/app/NN_emp.png) 
+![](images/app/NN_client.png)
+
+Tables detailing specific details of the most similar projects are also provided below the chart:
+
+![](images/app/NN_table_top.png) 
+![](images/app/NN_table.png)
+
+The similar projects are calculated using a different method from the blended algorithm but it is important to have a form of cross-checking and engage the user's own experiences.
+Further development and testing is required to assess how intuitive the results are for users to interpret, how much users trust the algorithm output given the variable plots and nearest neighbour charts, and finally how trustworthiness may be further improved.
 
@@
 
-The improvements in profit sound promising, however it is important to consider possible limitations in the data and implementation process, which in turn motivates future research.
+The `r (best.mean-100)`% improvements in profit sound promising, however it is important to consider possible limitations in the data and implementation process, which in turn motivates future research.
 These include how accurately users of the algorithm would be capable of guessing input values before onset of a project and successful user uptake of the model, which was identified as a problem in the literature.

### 2016-06-30 final_app/server.R

@@ -1,13 +1,22 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.3/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.3/library')
-library('shiny',lib='C:/Progra~1/R/R-3.2.3/library')
-library('shinythemes',lib='C:/Progra~1/R/R-3.2.3/library')
-library('FNN',lib='C:/Progra~1/R/R-3.2.3/library')
+if(.Platform$OS.type == 'windows'){
+        setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+        all7<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7.csv')[,-1]
+        all7d<- read.csv('all7d.csv')[,-1]
+        
+} else{
+        setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+        all7<- read.csv('all7.csv')[,-1]
+        all7d<- read.csv('all7d.csv')[,-1]
+        
+}
+library('ggplot2')
+library("dplyr")
+library("plyr")
+library('magrittr')
+library('reshape2')
+library('shiny')
+library('shinythemes')
+library('FNN')
 # library('ggplot2')
 # library("dplyr")
 # library("plyr")
@@ -17,7 +26,6 @@ library('FNN',lib='C:/Progra~1/R/R-3.2.3/library')
 # library('shinythemes')
 
 
-all7<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7.csv')[,-1]
 all7$Start.Date<- as.Date(all7$Start.Date)
 #delete anything more than 4 for inv.vs.cost
 #delete anything less than .2 for inv.vs.cost
@@ -28,15 +36,16 @@ df <- read.table(text = " id  min  max
     Sp1     150          151     ", header=TRUE)
 
 #make a new dataframe with information about crossbar sizes
-df2<- data.frame(id=c(rep("Sp1",1)), ymin= c(4), y.mean = c(13),
-                 ymax=c(22), confidence=c('90%'))
[…]

### 2016-06-30 final_app/ui.R

@@ -107,26 +107,39 @@ shinyUI(fluidPage(
                                         "Senior Technical" = "Senior Technical",
                                         "Mid Technical" = "Mid Technical")),
                        
-                       sliderInput(
-                               'pc.pro', '% Hours by Main Employee:', 
-                               min=0, max=100, value=50, step=.5, round=0),
+                       
 
-                       sliderInput("no.user", "Team Size:",
-                                   min=1, max=6, value=1, step=1, round=0
+                       selectInput("timespan", "Timespan (days):",
+                                   list("1 day - 3 wks" = "1d-3wk", 
+                                        "3 wks - 2.5 months" = "3wk-2.5m", 
+                                        "2.5 - 9 months" = "2.5m-9m",
+                                        "9 months - 1.5 yrs" = "9m-1.5y",
+                                        "1.5 -3 years" = "1.5y-3y",
+                                        "More than 3 years" = ">3y")
+                       ),
+                       
+                       selectInput("inv.mlsto", "Approximate Fee:",
+                                   list("$100 - $600" = "100-600", 
+                                        "$600 - $2,500" = "600-2.5k", 
+                                        "$2,500 - $8,000" = "2.5k-8k",
+                                        "$8,000 - $60,000" = "9m-1.5y",
+                                        "$60,000 - $1.8m" = "60k-1.8m")
                        )
+                       
                             
                        ),
                 
                 
                 column(3,
                        
-                       sliderInput("timespan", "Timespan (days):",
-                                   min=0, max=1000, value=60, step=20, round=0
+                       sliderInput("no.user", "Team Size:",
+                                   min=1, max=6, value=1, step=1, round=0
                        ),
                        
-                       sliderInput("inv.mlsto", "Approximate Fee:",
-                                   min=500, max=250000, value=20000, step=500, round=2
-                                   ),
+                       sliderInput(
+                               'pc.pro', '% Hours by Main Employee:', 
+                               min=0, max=100, value=50, step=.5, round=0),
+                       
                        
                        submitButton(text="Calculate")
                        

### 2016-06-30 finalwriteup/draft.rmd
started conclusion

@@
 library('pwr')
+library('scales')
 
@@
 
-The project outcomes are intended to assist the business in  predicting project costs before project engagement and other business analyses such as client analysis.
+The project outcomes are intended to assist the business in  predicting project costs before project engagement.
 Several statistical and machine learning techniques will be tested, compared and refined.
@@
 
-results.A2 = readRDS("report_data/reg_lm.rds")
+results.A2 = readRDS("reg_lm.rds")
 results.A2 <- dplyr::filter(results.A2, diff>-15)
@@
 ```{r load_rf_results}
-results.rf.core = readRDS("report_data/reg_rf_core.rds")
+results.rf.core = readRDS("reg_rf_core.rds")
 # n.samples(data = results.rf.core$diff)
 
-results.rf2 = readRDS("report_data/reg_rf.rds")
+results.rf2 = readRDS("reg_rf.rds")
 # n.samples(data = results.rf2$diff)
@@
 
-```{r, fig.height=8}
+```{r blend_violin, fig.height=8}
 
@@
         pvals$p_value[i] = t.test(auc.150[[1]], auc.150[[index[i]]])$p.value %>%
-                format(digits = 2) %>% as.numeric
-}
+                format(digits = 2) %>% as.numeric }
 
@@
 # How to achieve statistical power of 0.8?
-finance.plot = readRDS("report_data/plot_finance.rds" )
+finance.plot = readRDS("plot_finance.rds" )
 fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
@@
 
-finance.plot = readRDS("report_data/plot_finance_100_4.rds" )
+finance.plot = readRDS("plot_finance_100_4.rds" )
 finance.plot$method <- factor(finance.plot$method, 
@@
 kable(best.fin)
-# saveRDS(best.fin, "report_data/best_fin.rds")
+# saveRDS(best.fin, "best_fin.rds")
 
@@
 ```{r TP_plot}
-conf.plot <- readRDS("report_data/conf_plot.rds")
+conf.plot <- readRDS("conf_plot.rds")
 
@@
 
-conf.plot <- readRDS("report_data/conf_plot.rds")
+conf.plot <- readRDS("conf_plot.rds")
 
@@
 ```{r}
-reject.plot <- readRDS('report_data/reject_plot.rds')
+reject.plot <- readRDS('reject_plot.rds')
 
@@
 
-The aim of this project was to predict the profitability of projects for consulting businesses using their internal CRM data.
-Specifically, a statistical or machine learning model was created based on historical project data.
-The second aim was to test whether the model would have a positive impact on the overall profitability of a case study business.
-These research aims were fulfilled and the prediction model was refined to a binary classification problem of whether a project would be profitable or loss-making.
-This suited the limited data size best which was also highly complex.
+**General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
+Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
+
+The project outcomes are intended to assist the business in  predicting project costs before project engagement.
+Several statistical and machine learning techniques will be tested, compared and refined.
+
+**Hypothesis 1**
+A statistical or machine learning model based on historical project data can predict the profitability of a new project.
+
+**Hypothesis 2**
+The predictive model built from Hypothesis 1 can be shown to have a positive impact on the bottom line of the case study business.
+
+The objectives of this research consisted of a general aim and two research hypotheses, all of which have been fulfilled and tested.
+The findings made a significant contribution to the present body of literature and highlighted areas of future work that would refine the outcomes.
+
+The general aim was to use statistical techniques to model the profitability of projects for a case study consulting business using their internal CRM data.
+This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitablility and as a binary classification problem predicting simply profit or loss.
+Several statistical and machine learning approaches were applied to both problems including naive bayes, bayesian networks, linear regression, random forests, gradient boosted trees and multiple methods of blending the individual models.
+
+Hypothesis 1 stated a statistical or machine learning model based on the CRM data could predict the profitability of a new project.
+Testing this hypothesis revealed that the degree of profitability (the regression problem) could not be accurately predicted but whether a job would be profitable or not (binary classification) could be predicted to a level significantly better than random assignment (AUC = 0.76).
+This was achieved by 3 individual methods while various techniques that blended the individual methods improved results further (AUC = 0.77).
+It is likely the simpler task of binary classification suited the complex yet small data set better.
+
+Lastly, hypothesis 2 stated the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom l ine) of the case study business.
+This theory was tested on a range of predictive models because their AUC scores were not directly linked to profitability once the models were applied to a business scenario.The models output a probability, so a business scenario was tested where a probability threshold was chosen, and if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
+If a project was accepted, the profits and losses were summed while if a project was rejected the profits were forfeited.
+If a model performed perfectly, all losses
+
 After testing several machine-learning and statistical models, a blended model of logistic regression, random forests, and boosted tree models predicted project profitability best.

### 2016-07-03 finalwriteup/draft.rmd
for duncan

@@
 
-# Glossary
-
-ANOVA - Anaylis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
-B2B - Business to Business 
-CBR - Case based reasoning. A method of estimating by using the results of similar cases.
-CRM - Customer Relationship Management.
+# Acronyms
+
+**ANOVA** - Anaylis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
+
+**B2B** - Business to Business. Commerce transactions between two businesses.
+
+**CBR** - Case based reasoning. A method of estimating by using the results of similar cases.
+
+**CRM** - Customer Relationship Management.
 Refers to a common type of software used by businesses to record client and project details as well as employee timesheet records.
 Generally accessable to each employee in a company.
-DAG - directed acyclic graph.
+
+**DAG** - Directed acyclic graph.
 A graph consisting of vertices connected by edges, where each edge directs a vertex to another vertex.
 The arrangement of the edges is such that a vertex cannot follow a path that loops back to itself.
-IT - Information technology
-BIM - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
-ROC - Reciever Operating Characteristic
-AUC - Area under the curve. In this project, the area refers to the area under an ROC curve
-TPR - True positive rate
-FPR - False positive rate
-TP - true positive
-FP - false positive
-DA - Discriminant analysis
-CLV - Customer Lifetime Value
-IQR - Interquartile range
-ESS - error sum of squares
-RMSE - root mean squared error
-FWLS - feature weighted linear stacking
+
+**IT** - Information technology. The use of computers for storing, retrieving and processing information.
+
+**BIM** - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
+
+**ROC** - Reciever Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
+
+**AUC** - Area under the curve.
+In this project, the area refers to the area under an ROC curve.
+Areas greater than 0.5 indicate predictions performing better than random chance while an area of 1 indicates perfect predictions.
+
+**TP** - True positive.
+The case where a case marked 'positive' in a dataset is correctly predicted 'positive' by an algorithm.
+Generally, 'positive' is assigned to categories for which the use case needs to be alerted.
+
+**FP** - False positive. A case marked 'negative' in a dataset is incorrectly predicted 'positive' by an algorithm.
+
+**TN** - True negative. A case marked 'negative' in a dataset is correctly predicted 'negative' by an algorithm.
+
+**FN** - False negative. A case marked 'positive' in a dataset is incorrectly predicted 'negative' by an algorithm.
+
+**TPR** - True positive rate. The proportion of positives that were correctly identified (true positives).
+
+**FPR** - False positive rate. The proportion of positives that were incorrectly identified (false positives).
+
+**DA** - Discriminant analysis. Statistical analysis to predict a categorical response variable.
+
+**CLV** - Customer Lifetime Value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
+
+**IQR** - Interquartile range.
+A measurement of spread or variability for a continuous variable.
+When the data is ordered numerically and divided into four equally sized portions, the values that divide the portions are called quartiles (first, second, third and fourth quartiles).
+The first quartile has the highest value of the lowest quarter.
+The interquartile range is the third quartile value minus the first quartile.
+
+**ESS** - Error sum of squares. The sum of squared errors from a set of predictions.
+
+**RMSE** - Root mean squared error. The square root of the mean of the ESS.
+
+**FWLS** - Feature weighted linear stacking.
+A linear combination of model results which are interacted with original variables.
+This allows certain models to be weighted heavier for certain values of the original features.
 
@@
 
-# Introduction - 2 pages
+# Chapter 1: Introduction
 
 Predicting how long it will take a person to do something is a notoriously tricky task.
-People make mental calculations of these sort every day- guessing how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish.
+People make mental calculations of this sort every day - estimating how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish.
 As most people have experienced, these estimates can be particularly inaccurate if the individual is attempting the task for the first time.
 These errors are trivial in everyday activities, but they have more meaning for businesses that sell their time as their chief revenue.
-Consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time a constultant spent on the problem.
-The delivery of 'expert advice' is generally in the form of a document (which could be anything from a report to hundreds of drawings) and can range from an hours advice to thousands of hours work on a substantial project.
+
+Consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time a consultant spent on the problem.
+The delivery of 'expert advice' is generally in the form of a document (which could be anything from a report to hundreds of drawings) and can range from an hours advice to thousands of hours work.
 Before a project commences, the client and consultant must agree on a fee for the expert advice, or at least a fee structure.
 The nature of how a fee is formulated can be as creative as the engaged parties wish, however a couple of simple examples are as follows: the client agrees to pay for the consultants time by the hour until the task is complete, or the consultant provides a fixed fee to complete the project in full, regardless of their hours spent.
-Both contracts assign financial risk to one party which they must then mitigate to the best of their abilities.
-
-This thesis focuses on the risk taken upon by a consulting company when agreeing to a fixed fee.
+Both contracts assign financial risk to one party which must then mitigate the risk to ensure profitability.
+
+This thesis focuses on the risk taken by a consulting company when agreeing to a fixed fee.
 Clients commonly collect fixed quotes from several consultants before settling with one or negotiating further so fees must be competitive.
 The way a consulting manager calculates their fixed fee varies from industry to industry and even company to company.
-Typically a consulting manager has experience in the type of project he is quoting and can use a combination of intuition, rules of thumb (such as a fraction of the entire project cost), and comparison to past projects.
+Typically a consulting manager has experience in the type of project they are quoting and can use a combination of intuition, rules of thumb (such as a fraction of the entire project cost), and comparison to past projects.
 Such methods can result in a range of profit margins which ultimately depend on how well the manager predicts the amount of time a project demands.
+
 This thesis examines the problem of cost prediction via a case study on a company's internal project performance data for the previous twelve years.
-The study tests whether a statistical model based on historical project data can predict the profitability of a project using characteristics available at the beginning of a project.
-Furthermore, this thesis investigates how the model's predictions could impact a consulting manager's determination of a fixed fee and the resulting impact on the company's bottom line.
-
-# Problem Description
+Statistical models based on historical project data were tested for their ability to predict profitability of projects using characteristics available at the beginning of a project.
+Furthermore, the potential for a model's predictions to improve a company's bottom line is investigated.
+This could be achieved with a strategy for how the model's predictions should change managers' fee proposals for risky projects.
+Due to the commercially sensitive nature of the data used, the source is not disclosed in this dissertation.
+Variables have also been renamed to ensure confidentiality is maintained, however, no such obfuscation has been applied in the methods used – which are presented with integrity.
+
+# Chapter 2: Problem Description
 
@@
 
-## Research Motivation - 4.5 pages
+## Research Motivation
 
@@
 A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average [@Flyvbjerg2011].
-These statistics reveal a concerning glimpse at projects' hidden financial stories which can cause job losses, ruin private businesses, and churn through government balancesheets.
-Complex long term projects are necessary for modern society producing infrastructure, houses, websites, and movies.
-This research focuses on the companies who complete portions of these complex projects and the financial risk that is contractually transferred to them.
-Specifically, the risk of the project owner is transferred to their sub-consultant by mandating fixed flat fee structures to complete a given amount of work.
-The subconsultant commits to completing their role in a project for a nominated price, regardless of whether the work takes less or more time than intially anticipated.
-In a similar way to infrastructure and IT project budgets, consultants are susceptible to underestimating the amount of time a project will demand, and eventually wear the cost of extra employee hours spent.
+These statistics cast a concerning light on projects' hidden financial stories which can cause job losses, ruin private businesses, and shrink government balance sheets.
+
+Although complex projects are more likely to run over budget, they reflect the complexities and priveleges of modern society - from major infrastructure, to housing, websites, and movies.
+All have functional and/or statutory requirements which are defined by a client, and are discharged through contracts to companies via a financially competitive process.
+Consulting companies who complete portions of these complex projects accept the financial risk that is contractually transferred to them.
+More specifically, the risk of the project owner is often transferred to their sub-consultant by mandating fixed flat fee structures to complete a given amount of work.
+The subconsultant commits to completing their role in a project for a nominated price, regardless of whether the work takes less or more time than initially anticipated.
 For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and succeeding.
@@
 The ramifications of unprofitable projects are multifold.
-Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery.
-Employees also experience stress attempting to complete the project within a disappearing budget and may produce lower quality work in exchange for speed.
+Employee morale deteriorates and employees may view themselves incompetent if they are responsible for the project's delivery.
+Employees also experience elevated stress attempting to complete underestimated projects within a disappearing budget and may compromise quality of work in exchange for speed.
 This experience is unfortunate if in reality, the project fee was under-estimated.
-Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would foster growth and futher profits.
-The financial risk taken on fixed flat fee projects may also discourage business' from taking other, more thought-out risks, such as expanding the business into a new area.
-Logically, predicting fixed fee project costs more accurately could positively affect business growth opportunities and staff morale.
+Unprofitable projects also limit a business' ability to invest money into marketing, training, and business development which would foster growth.
+The financial risk taken on fixed flat fee projects may also discourage businesses from taking other, more calculated risks, such as expanding the business into a new area.
+Logically, improving the prediction of project profitability could positively affect business growth opportunities and staff morale.
 
@@
 The challenge for consultants to reduce budget blowouts, in terms of hours spent, is not easily solved.
-Complex projects always differ from one another, and similar work may demand significantly different amounts of time.
-For example, an ad filmed in a remote location will cost significantly more than an ad filmed on an existing set, even though the effect of the final advert could be identical.
-Therefore, a constant fixed price to produce an advert is clearly inappropriate for film projects, but some kind of fixed price is often demanded by the client before a project is awarded.
+Complex projects always differ from one another - similar work may demand significantly different amounts of time.
+For example, an ad filmed in a remote location will cost significantly more than an ad filmed on an existing set, even though the effect of the final advertisement could be identical.
+Therefore, a constant fixed price to produce an advert is clearly inappropriate, but some kind of fixed price is often demanded by the client before a project is awarded.
 Charging clients by the hour is not competitive enough in many industries.
@@
 
-Another strategy to mitigate risks could be to track employee hours, and stop work once the fixed consulting fee has been depleted.
-Then, negotiate further fees, or variations, with the client and persuade them to accept.
-This satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective.
-However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or endure marginable profits to preserve the relationship with their client.
+Another strategy to mitigate risk is to agree to track employee hours, and stop work once the fixed consulting fee has been depleted.
+Then, negotiate further fees, or variations, with the client.
+This approach may satisfactorily reduce a consultant’s financial risk, however, may cause friction between the client and the consultant and affect potential repeat custom.
+Especially if the project followed a predictable process but the hours inflated beyond expectation (a common situation), consultants may prefer to wear losses or endure marginable profits to preserve their client relationship.
 Consultants may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
 Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects.
-In many industries, the reality is fixed flat fee structures are expected and business' must conform to survive.
+In many industries, the reality of fixed flat fee structures are unavoidable businesses conform to survive.
 
@@
 However, two surveys are available from 2000: one on construction projects, the other on IT projects.
-The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practises.
+The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practices.
 They found that the main method for cost estimation was breaking the project into detailed parts and adding up the cost of each item.
@@
 In the same survey, 91% of software companies cited cost estimation as a problem.
-Although limited recent literature is available on industry cost estimation practises, historically it can be observed that for complex construction and software projects, cost predictions are chiefly made with a mixture of experience, project comparisons, and an analysis of the project details and tasks.
-It is also acknowledged that cost estimation is a problem and hypothesised that inexperience and lack of time are chief contributors [@Akintoye2000].
+Although recent literature is limited,  cost estimation was identified as a problem and it was hypothesised that inexperience and lack of time were chief contributors [@Akintoye2000].
 
@@
 For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate.
-According to @Locallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project are ignored in favour of analysing outcomes of several similar projects.
+According to @Lovallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project are ignored in favour of analysing outcomes of several similar projects.
 Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003].
 One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
-The correlation for the current project can then be estimated via a statistical model which is then used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011].
+The correlation for the current project can then be estimated via a statistical model, which is then used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011].
 Managers rarely analyse data from similar projects even though it could significantly improve cost predictions [@Flyvbjerg2011].
@@
 An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].
-
-![This method was implemented by the British government for a rail project which is still under completion](images/lit_review/UKrail.JPG)
-
-The visualisation presents a powerful communication tool for influencing decision makers and improving forecast cost accuracy, however access to credible data to a sufficient number of projects can be a challenge [@Flyvbjerg2011].
+An example reading from this chart would be "If the chance of a cost overrun of 20% is acceptable, the estimated cost should be increased by 55% from original calculations.".
+This builds past experiences of cost overruns into the initial cost estimation.
+
+![This model was implemented by the British government for a rail project that is still under completion](images/lit_review/UKrail.JPG)
+
+The visualisation presents a powerful communication tool for influencing decision makers and improving forecast cost accuracy, however, access to credible data for a sufficient number of projects can be a challenge [@Flyvbjerg2011].
 Overall, the theory of optimism bias is a credible explanation for the problem of persistant underestimation of project costs.
@@
 
-The theory of reference class forecasting was first definitively published in 2003, however the idea of predicting project costs from previous data has been the subject of research for many preceding years and continues to this day.
+The theory of reference class forecasting was first definitively published in 2003, however, the idea of predicting project costs from previous data has been the subject of research for many preceding years and continues to this day.
 Although research results have often been promising, industry uptake of the idea is not yet established.
@@
 A single dataset may consist of projects completed in different countries with different clients and project teams.
-The mathematical methods used to predict project costs vary from statistical regressions to deep-learning neural networks [@Love2005][@Kim2004].
-Most studies report their predictive models perform well and that generally neural networks are more accurate than linear regression.
+The mathematical methods used to predict project costs vary from statistical regressions to deep-learning Neural Networks [@Love2005][@Kim2004].
+Most studies report their predictive models perform well and that generally Neural Networks are more accurate than Linear Regression.
 
@@
 
-Statistical research projects on reference forecasting have shown promising results in IT, construction, and film projects.
-However, persuading industry to adopt these methods into everyday business is a separate challenge.
+The existing models have shown predictive success, however the data and intention from this case study differs from previous studies.
+Prediction of construction cost is a different problem from consulting effort estimation and uses different variables (except the variables describing the construction project since the consulting company works in the construction industry).
+Software effort estimation projects are estimating time, however past studies have compared projects across many companies (obtained by surveys etc) while the case study company has access to their data only.
+Additionally, their data contains a rich variety of variables which cannot be exctracted from surveys.
+The use of their own data would also improve trust in model outputs.
 
@@
 A survey of the construction industry in the UK in 2000 found that a manager's experience and intuition is still the dominant method for construction cost calculation [@Akintoye2000].
-Surveys on current industry practise in cost estimation are scarce, however another study on the IT industry by @Jorgensen2007 argued that previously developed parametric models were not comprehensive enough to be used in industry and that expert judgment was not mobilised enough in the process of developing these tools.
+Surveys on current industry practice in cost estimation are scarce, however, another study on the IT industry by @Jorgensen2007 argued that previously developed parametric models were not comprehensive enough to be used in industry and that expert judgment was not mobilised enough in the process of developing these tools.
 
@@
 A Customer Relationship Management software package (CRM) was employed by this company to collect and store project data.
-It is available to all employees over the company intranet, and each employee completes daily timesheets alotting their hours to certain projects.
+A CRM is a popular type of software used by businesses to record client and project details as well as employee timesheet records.
+In the case study business, the CRM is available to all employees over the company intranet, and each employee completes daily timesheets alotting their hours to certain projects.
 Additionally, information is recorded against each project as well as each client, client contact, and each invoice.
@@
 However, it was important to be aware of limitations when reviewing outcomes and framing the problem.
-The most obvious limitation is that a single company is being disected in a specific industry, which limits the ability to state conclusions applicable to a broad range of cases.
-This can be overcome with future studies.
+The most obvious limitation is that a single company is being disected in a specific industry, which limits the ability to state conclusions applicable to a broad range of companies.
+This can be overcome with future studies, once a range of optimal methods and variables have been determined.
 Other constraining factors include various forms of missing data and human data entry error.
@@
 
-The project outcomes are intended to assist the business in  predicting project costs before project engagement.
-Several statistical and machine learning techniques will be tested, compared and refined.
+The project outcomes are intended to assist the business in predicting project profitability before engagement. Several statistical and machine learning techniques were tested, compared and refined.
 
@@
 
-This work contributes to the body of research in numerous ways.
-There is a gap in the literature of applying statistical and machine learning methods to internally generated business data to facilitate internal business decisions.
-Studies have modelled the commercial performance of movies, customer churn, the share market, and bankruptcies but few have modelled more day-to-day data that a business generates which could actually assist day-to-day decisions.
-Improving business decisions would reduce risk and improve profits in scenarios that could be applicable in thousands of businesses around the world.
-
-In the case of this study, the problem of cost estimation is tackled.
-This issue has been worked on extensively in past research, however the problem has rarely been addressed using a single company's internal data.
+Past research has tackled cost estimation in various industries, from different angles and with a range of data, however the problem has rarely been addressed using a single company's internal data.
 In the literature, dozens of projects from other companys and even countries are collected and analysed.
@@
 In addition to addressing cost estimation with a different kind of dataset and different industry sector, ensemble tree methods will be applied which have been minimally tested on this problem.
-To date, linear regression, neural networks, case based reasoning, and SVM's have been used to predict cost/effort even though ensemble tree methods can perform as well as neural networks in some cases and generally outperform linear regression.
+To date, Linear Regression, Neural Networks, case based reasoning, and SVM's have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks in some cases and generally outperform Linear Regression.
 Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots provide insight into the model's calculations.
-This is in contrast to neural networks and support vector machines (SVM's) which are 'black-box' predictors.
+This is in contrast to Neural Networks and support vector machines (SVM's) which are 'black-box' predictors.
 Providing insight to a user is a major benefit in this case as past research has demonstrated issues with user trust and uptake of models, even if the models were successful.
@@
 These were compared against individual models and the best constructs were selected.
-Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision making scenarios and further applications were considered.
+Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision making scenarios.
+Further applications were also considered.
 
@@
 * Chapter 7: Binary Classification Models
-* Chapter 8: Extended Research including model blending, overall profit analysis, research limitations and future work
+* Chapter 8: Model blending,
+* Chapter 9: Extended Analysis including overall profit analysis, research limitations and future work
 
@@
 Then, a discussion finalises each chapter, where the results are examined deeply and analysed in a broader context.
-The nature of the results are also debated in terms of how surprising they were, their impact on the research hypothesis and limitations as well as suggested future work.
+The nature of the results are also debated in terms of how surprising they were, their impact on the research hypotheses and limitations as well as suggested future work.
 The final chapter, the Conclusion, summarises the findings of the project, how the research answered the hypotheses and how the work could be viably applied in industry.
 
-# Literature Review - 25 pages for pHd
+# Chapter 3: Literature Review
 
@@
 
-**Detailed Analysis**
+#### Detailed Analysis in the Construction Industry
 
@@
 
-**Statistical and Machine Learning Models**
+#### Statistical and Machine Learning Models in the Construction Industry
 
@@
 @Elfaki2014 completed a study on cost estimation techniques and research from 2004 to 2014.
-They found that artificial neural networks and SVM's were the most common machine learning techniques used.
+They found that artificial Neural Networks and SVM's were the most common machine learning techniques used.
 He highlighted that these two machine learning techniques deal well with uncertainty, however lack technical justification to the decision maker (known as a black box).
 Neural networks are also time consuming to train, and must be re-trained and re-tested with each additional piece of data [@Kim2004].
-Nevertheless, neural networks and SVM's recieved a significant amount of attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
-
-@Shin2015 pioneered the application of boosted trees to cost estimation in construction projects.
-This is surprisingly late given the rapid uptake of boosted trees over the past decade.
+Nevertheless, Neural Networks and SVM's recieved a significant amount of attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
+
+@Shin2015 pioneered the application of Boosted Trees to cost estimation in construction projects.
+This is surprisingly late given the rapid uptake of Boosted Trees over the past decade.
 In the study, data from 234 school building construction projects in Korea were used.
-Boosted trees were compared to neural networks, which are one of the most popular algorithms to apply to this problem.
-Boosted trees were found to predict costs slightly more accurately than neural networks, but not statstically significantly (p value < 0.5).
-Even though the two algorithms performed similarly, the output from boosted tree models provide insight into the structure of the model such as variable importance and variable partial dependence plots.
-This additional insight may mean boosted trees are preferable over neural networks because of their potential to engage decision makers.
-
-A notable amount of literature has studied the predictive power of multiple linear regression in the construction cost estimation problem.
-Often, linear regression is the only model assessed, without comparison to other methods such as neural networks, which first started appearing in literature in the 1990's [@Kim2004].
-However, even post-1990, many studies focused on linear regression only.
+Boosted trees were compared to Neural Networks, which are one of the most popular algorithms to apply to this problem.
+Boosted trees were found to predict costs slightly more accurately than Neural Networks, but not statstically significantly (p value < 0.05).
+Even though the two algorithms performed similarly, the output from Boosted Tree models provide insight into the structure of the model such as variable importance and variable partial dependence plots.
+This additional insight may mean Boosted Trees are preferable over Neural Networks because of their potential to engage decision makers.
+
+A notable amount of literature has studied the predictive power of Multiple Linear Regression in the construction cost estimation problem.
+Often, Linear Regression is the only model assessed, without comparison to other methods such as Neural Networks, which first started appearing in literature in the 1990's [@Kim2004].
+However, even post-1990, many studies focused on Linear Regression only.
 This may be because the technique is straightforward, easy to use and widely available in statistical packages [@Chan2005].
-Out of those that have compared regression to neural networks, the studies have shown that neural networks outperform regression, however other studies established they are approximately equal [@Kim2004][@Attalla2003].
-@Dissanayaka1999 found that neural networks outperformed multiple linear regression models, however regression models could be performed first to determine variable importance and condense the number of variables.
-It can be argued that boosted trees could perform as well as neural networks while providing additional insights (even more so than regression models) such as variable importance and partial dependence of variables against the independent variable, cost.
-
-Despite neural networks typically outperforming linear regression models, it is interesting to note some of the outcomes of regression studies in cost estimation.
+Out of those that have compared regression to Neural Networks, the studies have shown that Neural Networks outperform regression, however other studies established they are approximately equal [@Kim2004][@Attalla2003].
+@Dissanayaka1999 found that Neural Networks outperformed multiple Linear Regression models, however regression models could be performed first to determine variable importance and condense the number of variables.
+It can be argued that Boosted Trees could perform as well as Neural Networks while providing additional insights (even more so than regression models) such as variable importance and partial dependence of variables against the independent variable, cost.
+
+Despite Neural Networks typically outperforming Linear Regression models, it is interesting to note some of the outcomes of regression studies in cost estimation.
 Different studies tended to produce a wide range of variables that contribute at fluctuating levels in the models.
@@
 
-**Case Based Reasoning**
+#### Case Based Reasoning in the Construction Industry
 
@@
 
-A study by @Kim2004 compared CBR to neural networks and linear regression in the construction cost estimation problem.
+A study by @Kim2004 compared CBR to Neural Networks and Linear Regression in the construction cost estimation problem.
 For CBR, the ESTEEM software package was used which algorithmically calculated the similarity of variables, weighted the variables using gradient descent, and deduced the cost using the most similar cases (the specific method for this last step was not explained).
-The discussion compared mean absolute error rates of the three methods and showed that CBR was more accurate than regression, but less accurate than neural networks in cost estimation.
-Despite neural networks outperforming CBR, CBR maintained significant advantages over neural networks which have been referenced in other studies as well.
+The discussion compared mean absolute error rates of the three methods and showed that CBR was more accurate than regression, but less accurate than Neural Networks in cost estimation.
+Despite Neural Networks outperforming CBR, CBR maintained significant advantages over Neural Networks which have been referenced in other studies as well.
 
 Advantages of CBR include algorithmic efficiency as well as user engagement.
-In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to neural networks, which must be slowly re-trained and re-tuned.
+In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to Neural Networks, which must be slowly re-trained and re-tuned.
 New cases also did not need to have every variable completed which can occur often in real, messy data.
 Neural networks on the other hand require complete data.
-Another significant advantage of CBR is the ability for users to review chosen similar cases and make sense of the predicted cost, as opposed to neural networks which are a black box [@Kim2004][@Elfaki2014][@Kumar2007].
+Another significant advantage of CBR is the ability for users to review chosen similar cases and make sense of the predicted cost, as opposed to Neural Networks which are a black box [@Kim2004][@Elfaki2014][@Kumar2007].
 One disadvantage is that accuracy can be highly dependent on the number of selected cases [@Elfaki2014].
-Overall, CBR may provide prediction accuracies between regression and neural networks but the results can be justified by the user - a valuable asset.
+Overall, CBR may provide prediction accuracies between regression and Neural Networks but the results can be justified by the user - a valuable asset.
 
@@
 
-**Summary**
+#### Summary of Cost Estimation in the Construction Industry
 
@@
 Neural networks generally outperformed all other systematic methods, such as regression and case based reasoning, but had the significant disadvantage of being a black box.
-Alternative methods such as boosted trees have the potential to perform as well as neural networks and provide insight into the structure of the algorithm.
-This is true for CBR as well, although it has not been shown to predict as accurately as neural networks.
+Alternative methods such as Boosted Trees have the potential to perform as well as Neural Networks and provide insight into the structure of the algorithm.
+This is true for CBR as well, although it has not been shown to predict as accurately as Neural Networks.
 There exists a gap in the research in ways to intelligently combine CBR with machine learning methods that could predict well, provide insight to the structure, and engage the user in reviewing similar projects.
@@
 
-**Detailed Analysis**
+#### Detailed Analysis in the IT Industry
 
@@
 
-**Algorithmic methods**
+#### Algorithmic Methods in the IT Industry
 
@@
 
-Multiple studies have shown that neural networks definitively outperform regression models in effort estimation, although regression is the most popular method in the literature [@Finnie1997][@Pai2013][@Matson1993].
+Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although regression is the most popular method in the literature [@Finnie1997][@Pai2013][@Matson1993].
 Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the models accuracy: size [@Shepperd1996][@Finnie1997][@Pai2013].
@@
 This may be because no model has proved to be outstandingly successful at consistently predicting required effort [@Finnie1997].
-Also, neural networks (the most accurate algorithm) lack the ability to capably explain its results and instill trust in decision makers which is another parallel to research from the construction industry [@Finnie1997].
+Also, Neural Networks (the most accurate algorithm) lack the ability to capably explain its results and instill trust in decision makers which is another parallel to research from the construction industry [@Finnie1997].
 
@@
 
-**Case Based Reasoning**
+#### Case Based Reasoning in the IT Industry
 
@@
 Then, the similar cases are used to predict effort for the new case.
-In @Shepperd1996's study, linear regression was employed.
-
-The literature found CBR to perform approximately equally to algorithmic methods such as neural networks which contrasts findings in the construction industry [@Shepperd1996][@Finnie1997].
-It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects and provides insight into the prediction that neural networks do not [@Finnie1997].
-
-**Summary**
+In @Shepperd1996's study, Linear Regression was employed.
+
+The literature found CBR to perform approximately equally to algorithmic methods such as Neural Networks which contrasts findings in the construction industry [@Shepperd1996][@Finnie1997].
+It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects and provides insight into the prediction that Neural Networks do not [@Finnie1997].
+
+#### Summary of Effort Estimation in the IT Industry
 
@@
 Expert judgment via detailed analysis is still the predominant method for effort/cost estimation.
-Case based reasoning and algorithmic methods such as neural networks have found some success in effort estimation (at similar levels) but have not been successfully adopted by industry.
+Case based reasoning and algorithmic methods such as Neural Networks have found some success in effort estimation (at similar levels) but have not been successfully adopted by industry.
 It was noted that case based reasoning resonated more with decision makers because its predcitions could be better justified by the model.
@@
 These vary from simple methods such as Linear Regression to complex, deep learning Neural Networks.
-The previous section highlighted the use of linear regression, neural networks, SVMs and in one case boosted trees for prediction, however research on other business problems utilised a wider range of methods.
+The previous section highlighted the use of Linear Regression, Neural Networks, SVMs and in one case Boosted Trees for prediction, however research on other business problems utilised a wider range of methods.
 These include Naiive Bayes, Random Forests, and machine learned Bayesian Networks.
@@
 The reason this method is popular is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
-The integration of categorical variables is called Anaylis of Variance (ANOVA) linear regression, and the algorithm compares group mean variances within the categorical variables.
-
-There are several drawbacks to linear regression.
+The integration of categorical variables is called Anaylis of Variance (ANOVA) Linear Regression, and the algorithm compares group mean variances within the categorical variables.
+
+There are several drawbacks to Linear Regression.
 For example assumptions must be made about the structure of the data.
@@
 On the other hand, correlated variables do not impact the predictive accuracy of the model.
-A final drawback is that the variable coefficients cannot be used to compare variable importance unless the variables have all been standardised to one another [@Putler2012].
-This limits the amount of insight that can be drawn from a linear regression model.
-
-Despite limitations of linear regression models, they have widespread business applications due to their simplicity and ease of application.
+A final drawback is that the variable coefficients cannot be used to compare variable importance unless the variables have all been standardised to one another (i.e.
+scaled to the same mean and standard deviation) [@Putler2012].
+This limits the amount of insight that can be drawn from a Linear Regression model.
+
+Despite limitations of Linear Regression models, they have widespread business applications due to their simplicity and ease of application.
 Regression is used to predict risk in the finance and insurance industries, predict who to target in marketing exercises, and predict consumption spending in the field of economics [@Harrell2013].
-Its simplicity and widespread application make linear regression a good starting point for analysis and can serve as a performance benchmark against complex models.
+Its simplicity and widespread application make Linear Regression a good starting point for analysis and can serve as a performance benchmark against complex models.
 
@@
 
-The method behind linear regression can be adapted to predict a binary response variable (1 or 0).
+The method behind Linear Regression can be adapted to predict a binary response variable (1 or 0).
 A linear equation is fit to the log of the odds of the binary response variable against each of the explanatory variables.
@@
 $probability = \frac{odds}{1 + odds}$
+
 [@Lowry2016]
@@
 
-![](images/lit_review/logreg.JPG)
-[@Lowry2016]
+![Plot of log odds vs.
+an 'x' variable and probability vs.
+the 'x' variable.
+To convert log odds to probability, an exponential transformation is first made followed by the transformation from odds to probabilitiy [@Lowry2016].](images/lit_review/logreg.JPG)
 
@@
 success or failure) [@Moore1989].
-Similar pros and cons exist for logistic regression as linear regression, and it can also be used as a benchmark to compare other binary predictive models due to its speed and simplicity.
+Similar pros and cons exist for Logistic Regression as Linear Regression, and it can also be used as a benchmark to compare other binary predictive models due to its speed and simplicity.
 Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survival/death [@Moore1989].
@@
 
-The naive bayes method works by making conditional independence assumptions about the explanatory variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
+The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
 More specifically, it is assumed that each variable is conditionally independent given the respose class, without considering other predictor variables.
@@
 Decision trees are one of the simplest and most intuitive machine learning methods.
-There are several brands of basic decision tree algorithms including ID3, C4.5, CART, and CHAID with CART and C4.5 being the most popular [@Kabra2001].
+There are several brands of basic decision tree algorithms including ID3, C4.5, CART, and CHAID with CART and C4.5 being the most popular [@Kabra2011].
 Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
 Single decision trees tend to overfit the data and provide low predictive accuracy but have favourable characteristics such as creating intuitive rules that a decision maker can follow in real-life scenarios.
-In addition, their results do not depend heavily on the skill of the data analyst [@Kabra2001][@Putler2012].
+In addition, their results do not depend heavily on the skill of the data analyst [@Kabra2011][@Putler2012].
 
 To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
-Three examples of these ensemble decision tree methods are boostrap aggregatiing (bagging), boosting, and random forests.
+Three examples of these ensemble decision tree methods are boostrap aggregatiing (bagging), boosting, and Random Forests.
 Bagging creates multiple trees by sampling a different training data set (*with* replacement) for each tree and then combining tree results [@Breiman1996].
@@
 This allows dominant variables to be supressed for a fraction of the splits, allowing the algorithm to explore signals in weaker variables.
-It also prevents the trees from becoming too correlated [@Breiman2001].
+It also prevents the trees from becoming too correlated [@Breiman2001a].
 The boosted decision tree approach applies gradient descent theory to a series of decision trees.
@@
 
-![](images/lit_review/dec_tree.JPG)
-[@Jeevan2015]
-
-In contrast to linear regression, the predictor variables do not need to be transformed as no assumptions are made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
-Other advantages are that the ensemble methods are very fast in comparison to SVM's and neural networks but perform just as well and provide insights such as variable importance and variable relationships [@Sealfon2012].
+![Graphical illustration of non-linear splits derived from a decision tree with two variables [@Jeevan2015].](images/lit_review/dec_tree.JPG)
+
+In contrast to Linear Regression, the predictor variables do not need to be transformed as no assumptions are made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
+Other advantages are that the ensemble methods are very fast in comparison to SVM's and Neural Networks but perform just as well and provide insights such as variable importance and variable relationships [@Sealfon2012].
 Random forests in particular have the benefit being unaffected by noise in the data as well as the capacity to handle a large number of variables [@Sealfon2012].
-Caruana [@Caruana2006] tested boosted trees, random forests, neural networks, SVM's, logistic regression and naive bayes on 11 binary classification problems and found that boosted trees performed best, followed by random forests.
+Caruana [@Caruana2006] tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
 This demonstrates ensemble tree methods are capable of competing with high level machine learning algorithms.
@@
 This may or may not be required depending on the nature of the problem.
-Also, variable importance tables provided by random forests and boosted trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
+Also, variable importance tables provided by Random Forests and Boosted Trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
 The methods can also be prone to overfitting if not tuned carefully.
@@
 
-A bayesian network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
+A Bayesian Network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
 The model is visually represented by a DAG (directed acyclic graph) and is capable of linking the conditional dependency between any variable to another variable or via other variables.
@@
 Bayesian networks are excellent at representing knowledge from different origins, and then visually communicating results well to decision makers [@Kragt2009].
-Therefore bayesian networks have become an excellent decision support tool.
+Therefore Bayesian Networks have become an excellent decision support tool.
 Network relationships can be machine learned, however this is not widely included as part of the suite of machine learning methods and is usually recommeneded for modelling complex networks.
-Drawbacks of bayesian networks include their inability to handle continuous numeric variables (all continuous variables must be discretised).
+Drawbacks of Bayesian Networks include their inability to handle continuous numeric variables (all continuous variables must be discretised).
 This ensures all variable relationships are categorical [@Kragt2009].
-Popular applications of bayesian networks are modelling uncertainty in natural resource management and modelling complex business network structures such as airports [@Wu2013].
+Popular applications of Bayesian Networks are modelling uncertainty in natural resource management and modelling complex business network structures such as airports [@Wu2013].
 
@@
 
-Neural networks were inspired by biological neural networks of the human nervous system and require a lot of training data relative to other methods.
+Neural networks were inspired by biological Neural Networks of the human nervous system and require a lot of training data relative to other methods.
 They are flexible in the range of relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
@@
 
-![](images/lit_review/svm.JPG)
-[@Provost2013]
+![Linear boundary with the widest clear margin between data points of different categories [@Provost2013].](images/lit_review/svm.JPG)
 
@@
 They are also known to be robust over different samples and perform well in high-dimensional feature space [@Auria2008].
-On the other hand, the output is in the form of distance to the boundary (as opposed to probability) and results are not transparent (similar to neural networks) [@Caruana2006].
+On the other hand, the output is in the form of distance to the boundary (as opposed to probability) and results are not transparent (similar to Neural Networks) [@Caruana2006].
 Also, SVM's are known to be very slow to train and therefore often not suitable for industry purposes [@Auria2008].
@@
 This section reviewed the advantages and disadvantages of statistical and machine learning techniques that have been successfully applied to business problems.
-These included linear regression, logistic regression, naive bayes, decision trees and ensemble trees, bayesian networks, neural networks, and SVM's.
-Linear/logistic regression and naive bayes are computationally very fast with a simple concept that make broad assumptions about the data, but work well as baseline models against which to compare complex models.
+These included Linear Regression, Logistic Regression, Naive Bayes, decision trees and ensemble trees, Bayesian Networks, Neural Networks, and SVM's.
+Linear/Logistic Regression and Naive Bayes are computationally very fast with a simple concept that make broad assumptions about the data, but work well as baseline models against which to compare complex models.
 In certain cases these simpler models are significantly outperformed by complex models however it is important to compare both.
@@
 
-On the other hand, ensemble tree methods are capable of performing as well as neural networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
+On the other hand, ensemble tree methods are capable of performing as well as Neural Networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
 Bayesian networks are excellent for presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally been excluded from the machine learning literature in comparison to the other methods.
-This may be due to low awareness of learned bayesian networks or lower performance.
-In conclusion, simple models such as linear/logistic regression and naive bayes are good baseline models to implement, while ensemble tree methods and bayesian networks may handle messy 'real world' data better than the simpler models while still providing insight to decision makers.
-
-## Statistical and Machine Learning Methods Applied to Business Problems
+This may be due to low awareness of learned Bayesian Networks or lower performance.
+In conclusion, simple models such as linear/Logistic Regression and Naive Bayes are good baseline models to implement, while ensemble tree methods and Bayesian Networks may handle messy 'real world' data better than the simpler models while still providing insight to decision makers.
+
+## Review of Statistical and Machine Learning Applications to Business Problems
 
@@
 The literature pertaining closely to the cost prediction topic has been reviewed in the first section of this chapter, however it is relevant to broaden the critique of the literature on advanced machine learning techniques that have been applied to general business problems.
+This is to research whether the broader literature found success in methods that cost estimation had not put to the test.
 The simpler statistical methods have been covered already, so the focus of this section are the machine learning prediction methods developed for business decisions.
@@
 
-According to Breiman [@Breiman2001b], simple theoretical models such as linear regression are informative but often provide unsatisfactory accuracy for modelling real life data.
+According to Breiman [@Breiman2001b], simple theoretical models such as Linear Regression are informative but often provide unsatisfactory accuracy for modelling real life data.
 This data is often messy and does not necessarily align with the rules of statistical theory.
@@
 Ravi Kumar and Ravi [@Kumar2007] performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
-The most widely used model was neural networks, however logistic and linear regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
-Supporting @Breiman2001b's statements, it was found that DA and linear regression techniques were not preferred due to their low accuracy.
-The overall assessment was that SVM's outperformed neural networks (back propagation neural networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
+The most widely used model was Neural Networks, however logistic and Linear Regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
+Supporting @Breiman2001b's statements, it was found that DA and Linear Regression techniques were not preferred due to their low accuracy.
+The overall assessment was that SVM's outperformed Neural Networks (back propagation Neural Networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
 Although SVM's performed the best, as discussed, they are often extremely complex and slow, requiring a great deal of memory [@Kumar2007].
@@
 
-As SVM's are often computationally too expensive and advanced for smaller scale business applications, which this research project would fall into, it is useful to further compare neural networks and decision trees which are simpler than SVM's but are also capable of achieving very accurate predictive models.
-Ravi Kumar and Ravi [@Kumar2007] found that neural networks and decision trees were both capable of out-performing the other, depending on the context - which is limited to bankruptcy prediction in Kumar's study.
-Both have advantages and disadvantages to consider, however as derived in the previous section, @Tsai2009's study on earnings management prediction found the most notable drawback for neural networks is the difficulty in interpreting the reasoning behind the algorithm.
+As SVM's are often computationally too expensive and advanced for smaller scale business applications, which this research project would fall into, it is useful to further compare Neural Networks and decision trees which are simpler than SVM's but are also capable of achieving very accurate predictive models.
+Ravi Kumar and Ravi [@Kumar2007] found that Neural Networks and decision trees were both capable of out-performing the other, depending on the context - which is limited to bankruptcy prediction in Kumar's study.
+Both have advantages and disadvantages to consider, however as derived in the previous section, @Tsai2009's study on earnings management prediction found the most notable drawback for Neural Networks is the difficulty in interpreting the reasoning behind the algorithm.
 Earning's management is influenced by managers' choices and insight into the model structure could provide explanations for human behaviour - a valuable output in business problems.
@@
 Ravi Kumar and Ravi's [@Kumar2007] study concludes that ensemble techniques, which refers to combinations of two completely different algorithms, can often outperform individual techniques.
-For example, combining the contrasting advantages of neural networks and decision trees is a worthwhile ensemble technique.
+For example, combining the contrasting advantages of Neural Networks and decision trees is a worthwhile ensemble technique.
 @Tsai2009's study eventually combined these two methods after trialling each one.
-They were run successively and the neural network ran first because it had a higher rate of prediction accuracy via its complex non-linear learning properties.
-Then, to resolve the lack of explanatory qualities that neural networks are notorious for, decision trees were employed.
-The 81% of cases that were correctly predicted by the neural network were used to generate decision trees, and in turn useful decision rules [@Tsai2009].
+They were run successively and the Neural Network ran first because it had a higher rate of prediction accuracy via its complex non-linear learning properties.
+Then, to resolve the lack of explanatory qualities that Neural Networks are notorious for, decision trees were employed.
+The 81% of cases that were correctly predicted by the Neural Network were used to generate decision trees, and in turn useful decision rules [@Tsai2009].
 This strategy is a promising way to benefit from the strengths of well performing complementary techniques.
@@
 Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
-All three models performed at similar levels for overall accuracy at around 80% (the general ranking was random forests, followed by SVM then Naive Bayes).
-However, when the true positive rates (accuracy in  predicting of employee resignations) of the models were compared, the SVM model far out performed random forests and Naive Bayes by achieving 81% TP vs.
+All three models performed at similar levels for overall accuracy at around 80% (the general ranking was Random Forests, followed by SVM then Naive Bayes).
+However, when the true positive rates (accuracy in  predicting of employee resignations) of the models were compared, the SVM model far out performed Random Forests and Naive Bayes by achieving 81% TP vs.
 51% and 55% respectively.
@@
 This means that SVM's had an effective method of increasing the importance of 'positive' cases in the data set which was skewed towards the majority of employees who kept their job versus the minority who left (25% churn).
-It should be noted that the boosted tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVMs.
-Furthermore, boosted trees provide insight into the predictive model which could build trust in the company's decision makers.
+It should be noted that the Boosted Tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVMs.
+Furthermore, Boosted Trees provide insight into the predictive model which could build trust in the company's decision makers.
 Nevertheless, it is noteworthy that their study proved employee churn models for customers could be translated to predicting internal churn of employees.
@@
 In this research, internal consulting project data will be used to predict a new project's profitability in terms of internal employee structure as well as client characteristics.
-A review of the literature in machine learning applied to business problems revealed that SVM's were found to be the most accurate technique followed by neural networks and ensemble tree methods.
-As discussed in the previous section, neural networks and SVM's do not provide insight into the predictions to the decision maker.
+A review of the literature in machine learning applied to business problems revealed that SVM's were found to be the most accurate technique followed by Neural Networks and ensemble tree methods.
+As discussed in the previous section, Neural Networks and SVM's do not provide insight into the predictions to the decision maker.
 A gap exists in applying these cutting edge machine learning techniques to the wide range of internal decisions that businesses make.
@@
 The problem of cost prediction has been addressed in many studies beforehand which have generally been limited to prediction of construction costs for buildings and infrastructure as well as effort estimation for IT projects.
-The prediction models consisted of case based reasoning, linear regression, neural networks and in one or two instances an ensemble tree or SVM.
+The prediction models consisted of case based reasoning, Linear Regression, Neural Networks and in one or two instances an ensemble tree or SVM.
 This leaves a gap in the literature in effort estimation for consulting companies in the construction industry in combination with trialling a broader range of prediction methods.
 
-An exhaustive list of machine learning and statistical models were reviewed theoretically as well as practically in the broader scope of business problem prediction.
-It was determined that bayesian networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as neural networks while providing the additional benefit of insights into the reasoning behind model predictions.
+A comprehensive list of machine learning and statistical models were reviewed theoretically as well as practically in the broader scope of business problem prediction.
+It was determined that Bayesian Networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as Neural Networks while providing the additional benefit of insights into the reasoning behind model predictions.
 These insights are very important as they address the lack of industry uptake - a major issue identified by numerous studies in both industries.
@@
 
-# Method - 2 to 8 pages
+# Chapter 4: Method
 
 The aim of this project is to determine whether the profitability of projects for consulting businesses can be predicted using statistical and machine learning techniques.
-As stated in the literature review, a couple of gaps exist in this research space including working with internal CRM data from a consulting company in the construction industry, trialling ensemble tree methods against the more popular regression and neural networks, and analysing the impact model results could have on business decisions.
+As stated in the literature review, a couple of gaps exist in this research space including working with internal CRM data from a consulting company in the construction industry, trialling ensemble tree methods against the more popular regression and Neural Networks, and analysing the impact model results could have on business decisions.
 This section will explain the chosen methodology for fulfilling the research aim using a case study engineering consulting company that charges for their time spent on project delivery.
@@
 * Invoicing records: dated invoiced amounts and sizes for each project and whether the invoice had been paid or written off
+
+    **Example Invoicing Data Structure** - Note data is fabricated
+
+| Job Number | Date Issued | Amount Invoiced | Amount Paid | Invoice Status |
+|------------|-------------|-----------------|-------------|----------------|
+| 1.2.300    | 2014/09/01  | \$5,000         | \$5,000     | Paid           |
+| 1.2.300    | 2014/10/02  | \$2,000         | \$2,000     | Paid           |
+| 1.2.300    | 2014/11/04  | \$3,000         | \$3,000     | Paid           |
+| 1.2.300    | 2014/12/02  | \$4,000         | \$4,000     | Paid           |
+| 1.2.300    | 2015/02/02  | \$5,000         | \$0         | Outstanding    |
+| 1.2.400    | 2014/08/05  | \$20,000        | \$20,000    | Paid           |
+| 1.2.400    | 2014/09/01  | \$2,100         | \$0         | Outstanding    |
+| ...        | ...         | ...             | ...         | ...            |\
+\\
+\
+
 * Timesheet entries: the number of hours spent on each project and on which day.
 Entries were input by individual employees, but names for each entry had been replaced with employee positions (such as mid-level technical)
+
+    **Example Timesheet Data Structure** - Note data is fabricated
+
+| Date    | Hours | Charge Amount | Cost Amount | Created By | Discipline | Job No. | Reconciled | Scope       |
+|---------|-------|---------------|-------------|------------|------------|---------|------------|-------------|
+| 1.2.300 | 4     | 640           | 400         | P12        | Civil      | 1.2.300 | Yes        | Concept     |
+| 1.2.400 | 4.5   | 720           | 450         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 1.2.500 | 2     | 480           | 320         | P34        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 1.2.600 | 7     | 1120          | 700         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 1.2.700 | 1     | 160           | 100         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 1.2.800 | 3     | 390           | 240         | P3         | Structural | 1.2.400 | Yes        | Site        |
+| ...     | ...   | ...           | ...         | ...        | ...        |         | ...        | ...         |\
+\\
+\
+
 * Project summary data: information describing each project such as client code, client contact code, discipline, subject, post code, director code, and suburb
+
+    **Example Project Summary Data Structure** - Note data is fabricated
+
+| Job Number | Job Address      | Director | Job Name    | Discipline | Post Code | Client | Project Engineer | Further Job Details.. |
+|------------|------------------|----------|-------------|------------|-----------|--------|------------------|-----------------------|
+| 1.2.300    | 1 Apple St.      | D1       | "Disguised" | Civil      | 4002      | C5030  | P12              | ...                   |
+| 1.2.400    | 33 Rogue Rd.     | D2       | "Disguised" | Structural | 4122      | C2000  | P24              | ...                   |
+| 1.2.500    | 4 Window Dr.     | D3       | "Disguised" | Water      | 4230      | C2045  | P30              | ...                   |
+| 1.2.600    | 864 Hamilton St. | D2       | "Disguised" | Structural | 4004      | C3001  | P1               | ...                   |
+| 1.2.700    | 2 Shark St.      | D3       | "Disguised" | Water      | 4570      | C3020  | P5               | ...                   |
+| 1.2.800    | 24 Campbell Rd.  | D3       | "Disguised" | Water      | 4222      | C4010  | P1               | ...                   |
+| ...        | ...              | ...      | ...         | ...        | ...       | ...    | ...              | ...                   |\
+\\
+\
 
@@
 * number of disciplines active in a project 
+\ 
+\
 
@@
 * client invoice frequency which gives an indication of how much work the company does for that client. 
+\ 
+\
 
@@
 
-$project\ profit = total\ invoiced\ amount - cost\ from\ recorded\ employee\ hours$
+    $project\ profit = total\ invoiced\ amount - cost\ from\ employee\ hours$
 
@@
 Different types of predictive models have different methods of calculating variable importance and since it was not known which model would best predict return per dollar, a few models' variable outputs were reviewed.
-These included ANOVA linear regression and two ensemble tree methods.
+These included ANOVA Linear Regression and two ensemble tree methods.
 Before models assessing variable importance were built, outlier cases were deleted to prevent distortion of valuable explanatory variables.
@@
 This was anticipated as each method models data using distinct theoretical foundations.
-The tested methods were ANOVA linear regression, conditional inference forests (cforests), and random forests.
-It was expected that random forests would favour variables that had more categories, a well known bias; which is also a bias that cforests have overcome [@Strobl2007].
-ANOVA is based on linear theory which contrasts conditional forest and random forest's free-form structure.
+The tested methods were ANOVA Linear Regression, conditional inference forests (cForests), and Random Forests.
+It was expected that Random Forests would favour variables that had more categories, a well known bias; which is also a bias that cForests have overcome [@Strobl2007].
+ANOVA is based on linear theory which contrasts conditional forest and Random Forest's free-form structure.
 Therefore, it is benefical to compare the methods' assessments of variable importance.
@@
 
-The random forest algorithm can produce a permutation variable importance for each variable as part of its output.
+The Random Forest algorithm can produce a permutation variable importance for each variable as part of its output.
 Variable importance is represented as a score, which is calculated by performing permutations on each variable in the test data for each tree and re-running the test data through the trees.
 More specifically, consider a single variable.
-The values of this variable in the out-of-bag sample (the test data that the random forest algorithm partitions before building each tree) are randomly permuted (reordered) for each tree.
+The values of this variable in the out-of-bag sample (the test data that the Random Forest algorithm partitions before building each tree) are randomly permuted (reordered) for each tree.
 The purpose of permuting is to mimic the absence of that variable.
-The trees in the random forest are then run again.
-Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted to the original random forest.
+The trees in the Random Forest are then run again.
+Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted to the original Random Forest.
 If the accuracy of the premuted forest is much less than the original forest, then that variable was important.
-Permutation importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the random forest [@Breiman2005].
-
-The random forest variable importance provides valuable insight as
+Permutation importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the Random Forest [@Breiman2005].
+
+The Random Forest variable importance provides valuable insight as
 it addresses the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables [@Strobl2007].
-However, a severe disadvantage of random forest variable importance is that it is not reliable for variable selection where explanatory variables vary in their scale of measurement or their count of categories.
+However, a severe disadvantage of Random Forest variable importance is that it is not reliable for variable selection where explanatory variables vary in their scale of measurement or their count of categories.
 Also, the importance of correlated predictors is overestimated and the algorithm tends to favour variables that have many possible splits or many missing values [@Strobl2007].
 The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably, however this does not necessarily indicate the variable impacts the response variable most.
-Overall, the random forest permutation importance is useful as it is able to compare variables in a machine learning environment and can include multivariate interactions, but the current random forest importance measure overestimates the importance of variables with many categories and some numeric variables.
+Overall, the Random Forest permutation importance is useful as it is able to compare variables in a machine learning environment and can include multivariate interactions, but the current Random Forest importance measure overestimates the importance of variables with many categories and some numeric variables.
 
@@
 
-Cforests are an alterantive ensemble tree method to random forests that overcome the aforementioned shortfalls of random forest variable importance measures.
+Cforests are an alterantive ensemble tree method to Random Forests that overcome the aforementioned shortfalls of Random Forest variable importance measures.
 It is built from ctrees which are decision trees based on a conditional inference framework.
@@
 The predictor variable with the lowest p-value is then selected as the splitting variable [@Hothorn2006].
-Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as random forests, via permuting a variable, re-running the forest and comparing the decrease in response variable accuracy.
-Several sources recommend cforests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as random forests (as discussed) as well as boosted trees [@Strobl2007][@Strobl2009][@Hothorn2006].
+Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable, re-running the forest and comparing the decrease in response variable accuracy.
+Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) as well as Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].
 
@@
 Less variables means that a simpler model is being used for prediction which in turn is easier for stakeholders to understand [@Weisberg2005].
-For these reasons, a subset of the most important variables were chosen before modelling which entailed first eliminating outliers, then comparing important variables from ANOVA as well as random forests and cforests.
-It can be inferred from the literature that cforests would suit the case study data best as it is unbiased and not limited to linear theory [@Strobl2007].
+For these reasons, a subset of the most important variables were chosen before modelling which entailed first eliminating outliers, then comparing important variables from ANOVA as well as Random Forests and cForests.
+It can be inferred from the literature that cForests would suit the case study data best as it is unbiased and not limited to linear theory [@Strobl2007].
 
@@
 
-Previous studies predicting project profitability in software and construction projects were predominantly limited to case-based reasoning, regression and neural networks.
-There is a need to test other sophisticated machine learning algorithms that are as powerful as neural networks but provide insight into the reasoning behind predictions.
-Ensemble trees (boosted trees and random forests) and bayesian networks fit these criteria.
+Previous studies predicting project profitability in software and construction projects were predominantly limited to case-based reasoning, regression and Neural Networks.
+There is a need to test other sophisticated machine learning algorithms that are as powerful as Neural Networks but provide insight into the reasoning behind predictions.
+Ensemble trees (Boosted Trees and Random Forests) and Bayesian Networks fit these criteria.
 Regression and naiive bayes models were also included as simple baseline models.
@@
 * Random forest
-* Gradient boosted trees
+* Gradient Boosted Trees
 
@@
 
-![](images/method/dendrogram.png)
+![Hierarchical dendrogram of the Timespan variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
 
@@
 Once the number of clusters had been decided, each case was assigned a cluster label to replace its numeric variable.
-It was decided that for timespans and invoiced amounts, the discretised variables should be applied to all models (not just bayesian networks) because manager's must guess these values initially when operating the model.
+It was decided that for timespans and invoiced amounts, the discretised variables should be applied to all models (not just Bayesian Networks) because manager's must guess these values initially when operating the model.
 It is easier for a manager to predict a timespan category than the exact number of days a project will last.
@@
 
-| Timespan Categories  | Total Invoiced Amount Categories |
-|----------------------|----------------------------------|
-| 1 day - 3 weeks      | \$100 - \$600                    |
-| 3 weeks - 2.5 months | \$600 - \$2,500                  |
-| 2.5 - 9 months       | \$2,500 - \$8,000                |
-| 9 months - 1.5 years | \$8,000 - \$60,000               |
-| 1.5 - 3 years        | \$60,000 - \$1.8m                |
-| more than 3 years    |                                  |
+Timespan Categories
+
+* 1 day - 3 weeks
+* 3 weeks - 2.5 months
+* 2.5 - 9 months
+* 9 months - 1.5 years
+* 1.5 - 3 years
+* more than 3 years
+
+Total Invoiced Amount Categories
+
+* \$100 - \$600
+* \$600 - \$2,500
+* \$2,500 - \$8,000
+* \$8,000 - \$60,000
+* \$60,000 - \$1.8m
 
@@
 
-All methods except gradient boosted trees could not handle missing data.
+All methods except gradient Boosted Trees could not handle missing data.
 Therefore, preliminary runs of each method used subsets of the data which had complete data.
-Then, imputation methods were trialled against gradient boosted tree results since boosted trees can handle missing values and should perform equally well with imputed data.
-
-The MICE random forest method was chosen because it has been proven to work well with complex data sets [@Shah2014].
-It first performs a standard random forest imputation of missing values which are treated as initial 'place holders'.
+It was possile that if missing data was imputed, predictions from the models that were limited by missing data could improve.
+Also, a complete data set allows for complete sets of predictions from each method, which can then be blended to further improve results.
+First, imputation methods were trialled against gradient Boosted Tree results since Boosted Trees can handle missing values and should perform equally well with imputed data.
+
+The MICE Random Forest method was chosen because it has been proven to work well with complex data sets [@Shah2014].
+It first performs a standard Random Forest imputation of missing values which are treated as initial 'place holders'.
 Random forest imputation is a complex process which works by first doing a rough imputation of the missing values.
@@
 
-After the initial random forest imputation is complete, the MICE process continues.
-The imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable again using random forest imputation.
+After the initial Random Forest imputation is complete, the MICE process continues.
+The imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable again using Random Forest imputation.
 This is repeated for each variable and their 'place-holder' imputations are replaced by an imputation targeted at the single variable.
@@
 
-The final imputed dataset was fed into a boosted tree algorithm and if similar predictive results were obtained using boosted trees imputed data and unimputed data, the imputed data must be reasonable.
+The final imputed dataset was fed into a Boosted Tree algorithm and if similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
 The imputed dataset was then trialled on the remaining methods and compared to un-imputed trials.
@@
 That is, if a probability threshold is chosen to be 0.6, each case in the test set can then be classified as 0 or 1 depending on whether the probability calculated by the model is greater or less than 0.6.
-Once the classifications have been made, the true positive rate and false positive rate can be calculated for that model and threshold.
+Once the classifications have been made, the TPR and FPR can be calculated for that model and threshold.
 This is plotted as a single point on the ROC curve, say point C.
@@
 
-![](images/method/ROC1.JPG)
-
-[@Provost2013]
+![Points with TPR's and FPR's as coordinates for different probability thresholds [@Provost2013]](images/method/ROC1.JPG)
 
@@
 
-![](images/method/ROC2.JPG)
-
-[@Provost2013]
+![An ROC defined by data points calculated from 5 probability thresholds [@Provost2013]](images/method/ROC2.JPG)
 
@@
 Six blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
-These included simple averaging of the individual model results, building a regression model using the individual model results only, a boosted tree model using individual model results only, feature weighted linear stacking (FWLS), random forests, and boosted trees.
+These included simple averaging of the individual model results, building a regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
 The first three models are simply the average, weighted average, of the three individual model predictions.
 The last three methods take advantage of potentially meaningful interactions between an individual model's predictions and meta-features (the original variables).
-For example, if the boosted tree model predicted profitability better than the other models for Water projects, the blending model would take advantage of this interaction.
+For example, if the Boosted Tree model predicted profitability better than the other models for Water projects, the blending model would take advantage of this interaction.
 Boosted tree predictions would be weighted higher than the other model predictions for Water projects.
 
-A simplified explanation of FWLS is as a linear regression where meta-features as well as model predictions from individual models are included as explanatory variables.
+A simplified explanation of FWLS is as a Linear Regression where meta-features as well as model predictions from individual models are included as explanatory variables.
 Then, each meta-feature is interacted with each set of model results [@Sill2009].
-A linear regression is performed to weight each term's contribution to the final predictive accuracy.
-Random forests and boosted trees were also used to blend meta-features with output from the best models.
+A Linear Regression is performed to weight each term's contribution to the final predictive accuracy.
+Random forests and Boosted Trees were also used to blend meta-features with output from the best models.
 Feature interaction is performed passively due to the nature of how trees are built.
@@
 
-## Method Conclusion
+<!-- ## Method Conclusion -->
 
@@
 
-# Chapter _ Variable Selection
+# Chapter 5: Variable Selection
 
@@
 As stated in the previous chapter, three predictive methods were selected which output variable importance scores which differ due to their unique theoretical foundations.
-These were linear regression, random forests, and cforests.
+These were Linear Regression, Random Forests, and cForests.
 
@@
 If another incomplete variable was added, the complete set of data would shrink again.
-Therefore, to make the most of the available data, each incomplete variable was added to the core variables one at a time, where a linear regression model was built for each.
+Therefore, to make the most of the available data, each incomplete variable was added to the core variables one at a time, where a Linear Regression model was built for each.
 That meant a model was made for each incomplete variable, resulting in 21 separate ANOVA models.
-The p-values for the F statistic of each variable is plotted below.
+The p-values for the F-statistic of each variable is plotted below.
 The core variables recieved p-values in all 21 models, while 'add.variable' represents p-values for the additional incomplete variable in each model.
 
-```{r, echo=FALSE, fig.height=5}
+```{r, echo=FALSE, fig.height=5, fig.cap = c("P-values of an ANOVA regression's F-statistics used to interpret variable importance")}
 
@@
         scale_y_continuous(breaks = c(0, 0.05, .25, .5, .75, 1)) +
-        labs(y = "P value of F statistic", x = "Independent Variables",
-             title = "Summary of F-statistic P-values from ANOVA models")
+        labs(y = "P-value of F-statistic", x = "Independent Variables",
+             title = "Summary of variable F-statistic P-values from ANOVA models")
 
@@
 The variables listed above had p-values below 0.05, meaning the null hypothesis, that the continuous explanatory variables have no linear relationship with the reponse variable or that categorical variables have the same mean response values for each group, could be rejected.
-The important variable relationships found in linear ANOVA regression shall be compared to output from random forests and cforests below.
+The important variable relationships found in linear ANOVA regression shall be compared to output from Random Forests and cForests below.
 
@@
 Instead, the completeness of the variables were assessed and it was decided that all variables with at least 2300 overlapping complete cases would be chosen.
-Additionally, the random forest algorithm cannot process categorical variables with more than 53 categories.
+Additionally, the Random Forest algorithm cannot process categorical variables with more than 53 categories.
 This eliminated more variables and the resulting 'core' dataset contained 15 variables and 2364 cases.
-Because of the random forest algorithm's known biases and limitations with numbers of categories, a briefer approach to variable importance analysis was taken in comparison to regression and cforests.
-
-Each tree in a random forest is created with a bootstrapped training sample of the data.
+Because of the Random Forest algorithm's known biases and limitations with numbers of categories, a briefer approach to variable importance analysis was taken in comparison to regression and cForests.
+
+Each tree in a Random Forest is created with a bootstrapped training sample of the data.
 The random sampling means the results of each forest are slightly different.
@@
 Variables with an unlimited number of categories can be included in the cForest function which is an advantage over randomForests [@Hothorn2006; Strobl2007; Strobl2008].
-Cforests compute a variable importance ranking which is a number relating to the reduction in error the variable provides (similar method to the permutation importance in random forests).
+Cforests compute a variable importance ranking which is a number relating to the reduction in error the variable provides (similar method to the permutation importance in Random Forests).
 The variable importance rankings for the core 15 variables are as follows:
 
-```{r, echo = FALSE, fig.depth = 5}
+```{r, echo = FALSE, fig.depth = 5, fig.cap = c("Variable importance output from a cForest built from 15 core variables")}
 
@@
 
-# Chapter _ Regression Models
+# Chapter 6: Regression Models
 
@@
 
-results.core = readRDS("reg_lm_core.rds")
+results.core = readRDS("report_data/reg_lm_core.rds")
 # n.samples(data = results.core$diff)
 
-results.A2 = readRDS("reg_lm.rds")
+results.A2 = readRDS("report_data/reg_lm.rds")
 results.A2 <- dplyr::filter(results.A2, diff>-15)
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
 
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
 
@@
 
-## Random Forest
+## Random Forests
 
 ```{r load_rf_results}
-results.rf.core = readRDS("reg_rf_core.rds")
+results.rf.core = readRDS("report_data/reg_rf_core.rds")
 # n.samples(data = results.rf.core$diff)
 
-results.rf2 = readRDS("reg_rf.rds")
+results.rf2 = readRDS("report_data/reg_rf.rds")
 # n.samples(data = results.rf2$diff)
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
 
@@
 results.rf.core %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 randomForest models- core variables only")
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models- core variables only")
 
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
 
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
 
@@
 results.rf2 %>% ggplot(., aes(x =final.RMSE, y = ..density..))+ geom_histogram() + geom_density() +
-        labs(x = "RMSE", title = "Density Plot of RMSE for 50 randomForest models - all variables sampled")
+        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models - Variables Randomly Sampled")
 
@@
 
-## Regression Models Discussion
+## Regression Models: Discussion
 
@@
 
-```{r}
+```{r, fig.cap = c("Distribution of "return per dollar" for all projects")}
 if(.Platform$OS.type == 'windows'){
@@
 
-# Chapter _ Binary Classification Models
+# Chapter 7: Binary Classification Models
 
@@
 The models were compared using the AUC test statistic calculated from the ROC as this gives a good indication of model performance under binary classification [@pROC][@Putler2012].
-An example plot of five ROC's built from 5-fold cross validation using the boosted tree algorithm is shown below.
+An example plot of five ROC's built from 5-fold cross validation using the Boosted Tree algorithm is shown below.
 
@@
 
-![](images/boosted tree/5_fold_ROC.png)
-
-[@pROC]
+![ROC's for five boosted tree models built on different train/test partitions of the dataset [@pROC]](images/boosted tree/5_fold_ROC.png)
 
@@
 20 models and their resulting AUC statistics from each method were initially determined.
-These were used to calculate how many models were required to achieve a minimum statistical power of 0.8 when compared to the boosted tree results.
+These were used to calculate how many models were required to achieve a minimum statistical power of 0.8 when compared to the Boosted Tree results.
 The comparison of logistic regression required 100 models, while randomForest required over 700 because the AUC distributions were close.
@@
 The violin plot below summarises the AUC values produced by each method.
-The 'violins' are coloured according to whether the distributions significantly vary to boosted trees using a critical value of 0.05.
-
-```{r reduced_violin}
+The 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
+
+```{r reduced_violin, fig.cap = c("Violing plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Subsets of the data were used for Logistic Regression and randomForests in order to provide datasets without missing values.")}
 
@@
 Imputation was done using the MICE method with randomForest imputation [@mice].
-Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for random forest unless over 450,000 models were made.
-
-```{r mice_violin}
-
-mice.totAUC <- readRDS("mice_df_auc.rds")
+Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for Random Forest unless over 450,000 models were made.
+
+```{r mice_violin, fig.cap = c("Violing plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Each method was fed the same imputed full dataset.")}
+
+mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
 
@@
 
-Boosted Trees, Logistic Regression, and randomForest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed boosted trees.
+Boosted Trees, Logistic Regression, and randomForest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed Boosted Trees.
 The results from each method using reduced data sets (to cater for missing values) alongside results using imputed complete data are shown below.
 
-```{r all_violin}
+```{r all_violin, fig.cap = c("Combined violin plots comparing AUC results for each method predicting profit/loss using subsets of data vs. the full imputed data set.")}
 
@@
         geom_violin(aes(fill = Data_Source, colour = Data_Source)) +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+
+        # stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
         scale_fill_manual(values = c('sienna1','navajowhite'),
@@
 
-## Binary Classification Discussion
+## Binary Classification: Discussion
 
@@
 
-# Chapter _ Extended Analysis
+# Chapter 8: Model Blending
 
@@
 
-*Variable Importance: cforest*
-
-```{r}
+*Variable Importance: cForest*
+
+```{r, fig.cap = c("Variable importance output from a cForest blended model. The results from the three best performing models were added as explanatory variables.")}
 
@@
         geom_bar(stat='identity') + 
-                        labs(title = "Cforest variable importance: blend")
+        labs(title = "Cforest Variable Importance: Blended Model",
+             x = "Variable", y = "Variable Importance Score")
 
@@
 * Total amount invoiced * results of randomForest model
-* Client Business type * results of boosted tree model
+* Client Business type * results of Boosted Tree model 
+\ 
+\
 
@@
 * Results from each individual model 
+\ 
+\
 
@@
 
-```{r}
-
-rf.varimp = readRDS("rfblend.rds")
+```{r, fig.cap = c("Variable importance output from a randomForest blended model. The results from the three best performing models were added as explanatory variables.")}
+
+rf.varimp = readRDS("report_data/rfblend.rds")
 rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
@@
         geom_bar(stat='identity') +
-        labs(title = "randomForest variable importance: blend")
+        labs(title = "randomForest Variable Importance: Blended Model")
 
@@
 
-```{r}
-
-boost.varimp = readRDS("boostblend.rds")
+```{r, fig.cap = c("Variable importance output from a Boosted Tree blended model. The results from the three best performing models were added as explanatory variables")}
+
+boost.varimp = readRDS("report_data/boostblend.rds")
 boost.varimp = add_rownames(boost.varimp)
@@
         geom_bar(stat='identity') +
-        labs(title = "Boosted Trees variable importance: blend")
+        labs(title = "Boosted Trees Variable Importance: Blended Model")
 
@@
 
-```{r blend_violin, fig.height=8}
+```{r blend_violin, fig.cap = c("Violing plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
+100 models were built for each method.")}
 
@@
 * the simple average 
-* and the two boosted tree models.
+* and the two Boosted Tree models.
 
@@
 Therefore, 50 more iterations of the simple averaged and simple logistic models were performed.
-The boosted tree models would not achieve a statistical power of 0.8 with only 150 samples when compared to the original logistic regression and were therefore dropped from the next stage of analysis.
+The Boosted Tree models would not achieve a statistical power of 0.8 with only 150 samples when compared to the original logistic regression and were therefore dropped from the next stage of analysis.
 
@@
 
-```{r, fig.width = 8, fig.height=5}
+```{r, fig.cap = c("Violing plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
+150 models were built for each method to achieve a statistical power of 0.8.")}
 
@@
 
-The simple Logistic Regression blend and simple averaging of the individual models (logistic regression, boosted trees and random forests) gave virtually the same distribution of results, where both AUC distributions were signficantly greater than the original logistic regression model.
+The simple Logistic Regression blend and simple averaging of the individual models (logistic regression, Boosted Trees and Random Forests) gave virtually the same distribution of results, where both AUC distributions were signficantly greater than the original logistic regression model.
 The p-values from the two-sample t-tests were `r pvals$p_value[1]` and `r pvals$p_value[2]` for simple average and simple Logistic Regression methods respectively.
@@
 
-## Profit Analysis
+# Chapter 9: Extended Analysis
+
+## Profit Curve Analysis
 
@@
 Two-sampled power calculations (for a power of 0.8) were performed between the results of each method at their optimal threshold against the best performing method (simple average blend).
-For example, the optimal threshold for the simple average blended model was 0.55 while the optimal threshold for the original boosted tree model was 0.60.
+For example, the optimal threshold for the simple average blended model was 0.55 while the optimal threshold for the original Boosted Tree model was 0.60.
 The 10 'total profit' values at these two thresholds were compared in the two-sample power calculation to determine the number of samples required to achieve a power of 0.8.
@@
 
-```{r profit_curve, fig.depth = 10}
-
-finance.plot = readRDS("plot_finance_100_4.rds" )
+```{r profit_curve, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+
+finance.plot = readRDS("report_data/plot_finance_100_4.rds" )
 finance.plot$method <- factor(finance.plot$method, 
@@
 
-```{r av_profit_curve}
+```{r av_profit_curve, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression Blended Method")}
+
 ggplot(finance.plot %>% filter(method == "simp.log"), 
@@
 
-## Profit Curve Analysis - Discussion
-
-```{r load_winner_results}
-
-simp.l <- readRDS("finance_average_model.rds")
+## Profit Analysis: Discussion
+
+```{r}
+simp.l <- readRDS("report_data/finance_average_model.rds")
 
@@
 In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.
+
+The plot below illustrates this proportion of projects as well as the proportion of profits and losses assigned to a range of probability levels by the simple logistic regression blend.
+The red line sits at its optimal threshold point, i.e.
+projects greater than the threshold point would be rejected:
+
+```{r, fig.cap = c("Distribution of projects according to the probability outputs from a typical simple Logistic Regression blended model. A bargraph of counts and relative profits are shown.")}
+
+reject.plot <- readRDS('report_data/reject_plot.rds')
+
+profs <- reject.plot  %>% .$balance.mlsto %>% abs %>% sum
+simp.l2 <- reject.plot %>% filter(compare == 'full') %>% select(simp.log, balance.mlsto, inv.mlsto.log)
+# simp.l2 <- mutate(simp.l2, invoiced.ratio = exp(inv.mlsto.log)/sum(exp(inv.mlsto.log))) 
+simp.l2$profit.ratio <- ifelse(simp.l2$balance.mlsto>0, simp.l2$balance.mlsto/profs,
+                              simp.l2$balance.mlsto/profs)
+simp.l2 <- rbind(simp.l2,simp.l2)
+simp.l2$faceter <- c(rep("hist", 2364), rep("profs",2364))
+simp.l2$prob.bin <- cut(simp.l2$simp.log, seq(0,1,by = 0.05), labels = seq(.05,1,by = 0.05))
+
+ggplot(simp.l2, aes(x = prob.bin)) + 
+        facet_grid(faceter~., scales = "free_y",
+                   labeller = labeller(faceter = c("hist" = "Count of Projects",
+                                       "profs"= "Relative Profits/Losses"
+                                       ))) +
+        geom_bar(data = simp.l2 %>% filter(faceter == "hist")) + 
+        geom_bar(data = simp.l2 %>% filter(faceter == "profs"),
+                 aes(x = prob.bin, y = profit.ratio), stat = "identity") +
+        # geom_bar(data = simp.l2 %>% filter(faceter == "inv"),
+                 # aes(x = prob.bin, y = invoiced.ratio), stat = "identity") +
+        labs(x = "Probability Score", 
+             y = "",
+             title = "Distribution of Project Counts and Project Profits Against Probability") +
+        geom_vline(xintercept = 10.5, colour = hcl(h = 15, l = 65, c = 100), linetype = 1,
+                   alpha = 0.8)
+
+```
 
@@
 
-```{r TP_plot}
-conf.plot <- readRDS("conf_plot.rds")
+```{r TP_plot, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
+conf.plot <- readRDS("report_data/conf_plot.rds")
 
@@
 
-The plot above shows the Complex Boosted Tree Blend has the highest TP rate followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
-All of these TP rates are below 25% which shows it was not worth catching a large percentage of loss-making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
+The plot above shows the Complex Boosted Tree Blend has the highest TPR followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
+All of these TPRs are below 25% which shows it was not worth catching a large percentage of loss-making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
 This implies the models were not able to very cleanly divide loss-making jobs from profitable jobs.
@@
 
-```{r TN_plot}
-
-conf.plot <- readRDS("conf_plot.rds")
+```{r TN_plot, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
+
+conf.plot <- readRDS("report_data/conf_plot.rds")
 
@@
 
-The complex Boosted Tree method had the highest TP rate, but also had a high FP rate.
+The complex Boosted Tree method had the highest TPR, but also had a high FPR.
 The increased number of FP (i.e.
 rejected profitable jobs) offset the more accurate rejection of loss-making jobs so much that the simple averaged blend and simple logisitc regression blend outperformed it.
-The afore-mentioned simple blends performed similarly, with a higher than average TP rate and average FP rate in comparison to other methods.
-The complex Random Forest blend and FLWS both had lower TP rates than the simple blends but higher TN rates.
-In the end, the improved TN rate did not outweight the lower TP rate, i.e.
+The afore-mentioned simple blends performed similarly, with a higher than average TPR and average FPR in comparison to other methods.
+The complex Random Forest blend and FLWS both had lower TPRs than the simple blends but higher TN rates.
+In the end, the improved TN rate did not outweight the lower TPR, i.e.
 the higher rate of accepting profitable jobs did not outweight the lower rate of accurately rejecting loss-making jobs.
 
-All individual original methods used substantially lower TP rates at their optimal thresholds, while their TN rates were spread out.
-This means, as the trialled threshold values moved, and the TP rate improved, the increasing lost profits from FP's was too great to justify a TP rate higher than 4.6%.
+All individual original methods used substantially lower TPRs at their optimal thresholds, while their TN rates were spread out.
+This means, as the trialled threshold values moved, and the TPR improved, the increasing lost profits from FP's was too great to justify a TPR higher than 4.6%.
 Blending the individual methods has made use of their unique strengths and seems to have clustered the probabilities of the actual loss-making jobs more accurately towards 1, and the profitable jobs towards 0.
-This allowed for higher TP rates with lower FP rates at the blended methods' optimal thresholds.
+This allowed for higher TPRs with lower FPRs at the blended methods' optimal thresholds.
 
@@
 
-```{r}
-reject.plot <- readRDS('reject_plot.rds')
+```{r, fig.cap = c("Proportion of profitable and loss-making jobs in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+reject.plot <- readRDS('report_data/reject_plot.rds')
 
@@
 
-```{r}
+```{r, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
+Colours are alternated between different jobs to indicate profit/loss magnitude.
+Predictions were from a typical simple Logistic Regression.")}
 ggplot(reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))), 
@@
 
-```{r}
+```{r, fig.cap = c("Distribution of the percentage of hours completed by a professional level employee in the full data set of projects vs.
+the rejected projects only.
+Predictions from a typical simple Logistic Regression were used.")}
 ggplot(reject.plot, 
@@
 
-```{r}
+```{r, fig.cap = c("Proportion of each timespan category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
 ggplot(reject.plot, 
@@
 
-```{r}
+```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
 ggplot(reject.plot, 
@@
 Industry's lack of uptake of cost estimation models was marked as a significant issue in the literature [@Akintoye2000].
-Decision makers did not sufficiently trust the model output which seemed disconnected to their experiences.
-In many studies, case-based reasoning had more traction than more accurate models because users' could justify the output.
+Decision makers did not sufficiently trust the model output which were not directly connected to their experiences.
+In many studies, case-based reasoning had more traction than more accurate models because decision makers could justify the output.
 Clearly this is an important aspect of the study to address.
@@
 
+User interfaces also allow input values to be tweaked and re-run so the decision maker can test different scenarios.
 A prototype user interface was developed using the Shiny package in R where a strategy was developed for how the input options and results would be communicated [@shiny].
 
-![](images/app/front_page.png)
+![Opening panel and tab of user interface application for the predictive model](images/app/front_page.png)
 
@@
 This output is not intuitive to understand for a manager because the probability score does not represent a statistical probability, rather a score the algorithm generated using its own measures.
-The optimal threshold dictated that scores below 0.6 should be classified as profitable jobs and scores above 0.6 are loss-making.
+The optimal threshold analysis found that scores below 0.6 should be classified as profitable jobs and scores above 0.6 are loss-making.
+
 The scores were re-scaled to a more intuitive numbering system where scores between 0 and 0.6 were scaled between 0 and 10, and scores higher than 0.6 were scaled between 0 and -10.
@@
 6-8).
-This provides an intuitive feel for the results the model's uncertainty.
+This provides an intuitive feel for the model's uncertainty.
 
@@
 Algorithms were originally selected that could reveal their predictive structure by either variable importance and/or variable relationships.
-The first tab in the application displays several variable dependency plots extracted from a typical boosted tree model.
+The first tab in the application displays several variable dependency plots extracted from a typical Boosted Tree model.
 Understanding the plot would require some training.
@@
 
-![](images/app/team_size_dep.png) 
-![](images/app/timespan_dep.png)
+![Graphic for model prediction and Boosted Tree partial dependency plot](images/app/team_size_dep.png)
+
+![Boosted Tree Partial dependency plot for Timespan embedded in the user interface](images/app/timespan_dep.png)
 
@@
 
-![](images/app/NN_top.png) 
-![](images/app/NN_emp.png) 
-![](images/app/NN_client.png)
+![User interface for Nearest Neighbour algorithm input and output chart](images/app/NN_top.png) 
+![Nearest Neighbour chart coloured by the position of the main employee on the projects](images/app/NN_emp.png) 
+![Nearest Neighbour chart coloured by the client for the projects](images/app/NN_client.png)
 
@@
 
-![](images/app/NN_table_top.png) 
-![](images/app/NN_table.png)
+![Tables below the chart in the application give specific details of each of the nearest neighbours](images/app/NN_table_top.png) 
+![The nearest neighbour details are split into sections such as 'Finances' and 'Staff Details'](images/app/NN_table.png)
 
@@
 
-The `r (best.mean-100)`% improvements in profit sound promising, however it is important to consider possible limitations in the data and implementation process, which in turn motivates future research.
-These include how accurately users of the algorithm would be capable of guessing input values before onset of a project and successful user uptake of the model, which was identified as a problem in the literature.
+The `r (best.mean-100)`% improvements in profit sound promising, however it is important to consider limitations in the data and implementation process, which in turn motivates future research.
+This includes how accurately users of the algorithm would be capable of guessing input values and how to integrate new data into the model.
 
 First, succesfully estimating project variables is critical in translating this research into tangible business benefits.
-The research data was obtained from the company's records of *completed* projects and initial expected values of variables were not documented.
-For example, if a project lasted 2 years and it was initially expected to take 1.5 years, the only data the company retains is that it lasted two years.
-This begs the question, how well could a user predict the variables before beginning a project?
+The data represents the company's records of *completed* projects so initially expected variable values were not documented.
+For example, if a project was originally expected to last 1.5 years but instead took 2, the company data only retains the two year timespan.
+This prompts the question, how well could a user predict the variables before beginning a project?
 Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e.
 "1 day - 3 weeks", "3 weeks - 2.5 months", etc).
-However, two remaining categories: team size and percent of project completed by a professional level employees were left as continuous variables.
-It would be interesting to test how well business managers could guess these four variables at the start of a project.
-
-The process of guessing varibles at the beginning of a project and then comparing estimates upon project completion is a long checking process.
-However, it would be worthwhile integrating this practice into the system's pre-existing data-collecting processes.
-When a new project is entered into the database, it could be required that the project manager guesses the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent to be completed by professionals.
-This data itself could feed into a revised version of the prediction algorithm.
-
-**updating the model!**
-A lot of data cleaning in past data
-need to build a pipeline of cleaning - only 11 variables now, not too hard
-* re run algorithm monthly? does not take long, maybe 1 minute.
-* repeat validation every few months
+However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables.
+It would be interesting to test how well business managers could guess the aforementioned four variables at the start of a project.
+
+The process of initially guessing variable values and then checking them upon project completion is time consuming.
+However, it would be worthwhile integrating into the pre-existing data-collection system.
+When a new project is entered into the database, it could be mandated that the project manager guesses the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent by professionals.
+Over time, this estimation data could contribute to a revised version of the algorithm.
+
+<!-- The strategy for updating the algorithm as new data is generated should be developed in future work.  -->
+<!-- A lot of data cleaning in past data -->
+<!-- need to build a pipeline of cleaning - only 11 variables now, not too hard -->
+<!-- * re run algorithm monthly? does not take long, maybe 1 minute. -->
+<!-- * repeat validation every few months -->
 
@@
 
-# Conclusion
-
-**General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
-Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
-
-The project outcomes are intended to assist the business in  predicting project costs before project engagement.
-Several statistical and machine learning techniques will be tested, compared and refined.
-
-**Hypothesis 1**
-A statistical or machine learning model based on historical project data can predict the profitability of a new project.
-
-**Hypothesis 2**
-The predictive model built from Hypothesis 1 can be shown to have a positive impact on the bottom line of the case study business.
+# Chapter 10: Conclusion
 
@@
 This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitablility and as a binary classification problem predicting simply profit or loss.
-Several statistical and machine learning approaches were applied to both problems including naive bayes, bayesian networks, linear regression, random forests, gradient boosted trees and multiple methods of blending the individual models.
+Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees and multiple methods of blending the individual models.
 
@@
 
-Lastly, hypothesis 2 stated the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom l ine) of the case study business.
-This theory was tested on a range of predictive models because their AUC scores were not directly linked to profitability once the models were applied to a business scenario.The models output a probability, so a business scenario was tested where a probability threshold was chosen, and if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
-If a project was accepted, the profits and losses were summed while if a project was rejected the profits were forfeited.
-If a model performed perfectly, all losses
-
-After testing several machine-learning and statistical models, a blended model of logistic regression, random forests, and boosted tree models predicted project profitability best.
-When applied to improving overall profitability of the business, the blended model which simply averaged the raw probability results from the three aforementioned models outperformed the individual models as well as the logistic regression blended  model.
-Final results showed the model could improve profitability of the business by between `r (-1.96*best.sd + best.mean) %>% round(1)`% and `r (1.96*best.sd + best.mean) %>% round(1)`% (the 95% confidence interval) by rejecting projects that have a 'probability' of being a loss-making job over `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).
-
-These results make a significant contribution to literature in this area as they provide a clear framework for how a business can realise tangible profit improvements using a set of algorithms and project data that is passively collected by businesses.
-Previous studies verified algorithms' prediction accuracy in cost estimation but stopped short of how the algorithms could be implemented in every day decision making and the affect on budgets.
-Additionally, this research tested ensemble tree methods and furthermore blended models on the problem of cost estimation.
-Past research had largely experimented with regression models and to some extent neural networks and case based reasoning.
-The advantage of ensemble tree methods is that they provide insight into the model's reasoning while having comparable predictive power to neural networks.
-Having said that, a limitation to the study was that there was not sufficient data to utilise the strength of ensemble tree methods and logistic regression performed simiarly.
-However, the final highest performing model was a blend of regression and two ensemble tree methods.
-
-There are several aspects of this work that should be investigated further including how well managers can estimate project input variables before a project is won.
-Particularly, estimations of the timespan, team size, total invoiced amount, and percentage of hours completed by professionals should be tested.
-These initial estimates could also be included in the each project's data entry for future analysis.
-Finally, research into user uptake of the model should be assessed.
-It is likely that in order to establish a manager's trust with the algorithm, additional model output will be required such as partial dependency plots from the boosted trees.
-Also, a 'nearest neighbour' algorithm could be run concurrently to show the manager outcomes of past similar projects, which should align with the model's probabiltiy prediction.
-Overall, this project has successfully demonstrated a blended statistical and machine learning model has the potential to aid manager's in rejecting loss-making projects to improve their business's profitability.
+Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom l ine) of the case study business.
+This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business scenario.
+The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss-making'.
+A range of threshold were trialled, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
+If a project was accepted, the profits and losses were summed while if a project was rejected the profits and losses were forfeited.
+A perfect model would reject all loss-making jobs and accept all profitable jobs, thereby improving overall profits.
+Final results showed the Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
+The 95% confidence interval for these improvements were between `r best.fin %>% filter(method == "simp.log") %>% .$lower_bound_95CI - 100`% and `r best.fin %>% filter(method == "simp.log") %>% .$upper_bound_95CI - 100`% using a probability threshold of `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).
+
+These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
+Ensemble tree methods and blending have been applied minimally to cost estimation.
+Past research tested regression models and to some extent Neural Networks and case based reasoning.
+The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting as well as Neural Networks.
+Next, a clear framework for how a business can tangibly improve profits was developed and proven successful.
+Previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
+Lastly, the decision framework was progressed even further into a prototype application where user trust was addressed.
+Previous studies had shown that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
+This was addressed by providing intuitive plots of the model structure and a parallel nearest neighbour graph of similar internal projects.
+Further work is required to test user confidence in the output.
+
+Another topic identified for future research was to test how well managers estimate project input variables.
+Estimation of the timespan, team size, total invoiced amount, and percentage of hours completed by professionals in particular should be tested.
+Initial estimates of input variables could also be kept in the database for future analysis.
+Overall, this project has successfully demonstrated a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models could guide managers in rejecting loss-making projects and improve net profitability.
 

### 2016-07-03 finalwriteup/trial_code.rmd
for duncan

--- 2016-06-17 finalwriteup/trial_code.rmd
@@
 output: word_document
+bibliography: Library.bib
+
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, echo=FALSE, include=FALSE}
+
+library('knitr')
+library('ggplot2')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library("rpart")
+library('car')
+library('e1071')
+library('party')
+library('randomForest')
+library('RColorBrewer')
+library('pwr')
+library('scales')
+
+# detect OS
+if(.Platform$OS.type == 'windows'){
+        # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
+} else{
+        # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+        # setwd("~/Documents/github/consult/finalwriteup/report_data")
+        opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/report_data')
+}
+
+opts_chunk$set(fig.width=6.7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
 ```
 
-$effort = \alpha*size^{\beta}$  
-$\alpha = productivity \ coefficient$  
-$\beta = economies \ of\ scale\ coefficient$  
-$size = estimated \ lines \ of \ code$
+```{r load_rf_results}
+results.rf.core = readRDS("reg_rf_core.rds")
+# n.samples(data = results.rf.core$diff)
 
-$$
-\begin{aligned}
- effort &= \alpha*size^{\beta} \\
- \alpha &= productivity \ coefficient \\
- \beta &= economies \ of\ scale\ coefficient \\
- size &= estimated \ lines \ of \ code
-\end{aligned}
-$$
+results.rf2 = readRDS("reg_rf.rds")
+# n.samples(data = results.rf2$diff)
 
-* $project\ profit = total\ invoiced\ amount - cost\ from\ recorded\ employee\ hours$
-* $return\ per\ dollar = \frac{project\ profit}{cost\ from\ hours}$
+```
+
+```{r}
+#load results from 150 iterations
+auc.150 = readRDS("blend_auc150.rds")
+
+```
+
+```{r insert_data}
+# load function for determining number of required samples for 0.8 power
+n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
+        if(type.calc == "one.sample"){
+              d.calc = abs(mean(data))/sd(data)
+              pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
+                                   power =.8, type = c(type.calc), 
+                                   alternative = alt)
+              return(ceiling(pow_pow$n))  
+        }
+
+        if(type.calc == 'two.sample'){
+                d.calc = abs(mean(data[[1]])-mean(data[[2]]))/
+                        (sqrt((sd(data[[1]])^2+sd(data[[2]])^2)/2))
+                pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
+                                   power =.8, type = c(type.calc), 
+                                   alternative = 'two.sided')
+                return(ceiling(pow_pow$n)) 
+        }
+
+}
+
+results.core = readRDS("reg_lm_core.rds")
+# n.samples(data = results.core$diff)
+
+results.A2 = readRDS("reg_lm.rds")
+results.A2 <- dplyr::filter(results.A2, diff>-15)
+# n.samples(data = results.A2$diff)
+
+```
+
+1
+
+```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of an ANOVA regression's F-statistics used to interpret variable importance")}
+
+plot.pvals = readRDS("AOV_varimp_plot.rds")
+
+ggplot(plot.pvals, aes(x = mod.vars, y = p.val)) + geom_boxplot() +
+        theme(axis.text.x = element_text(angle = 45, hjust = 1),
+              panel.background = element_rect(fill = 'grey95')) +
+        geom_hline(yintercept = 0.05, colour = 'red', linetype = 2) +
+        scale_y_continuous(breaks = c(0, 0.05, .25, .5, .75, 1)) +
+        labs(y = "P-value of F-statistic", x = "Independent Variables",
+             title = "Summary of variable F-statistic P-values from ANOVA models")
+
+```
+
+2
+
+```{r, echo = FALSE, fig.height = 4.5, fig.cap = c("Variable importance output from a cForest built from 15 core variables")}
+
+cfor.varimp = readRDS("cfor_varimp.rds")
+# how did pc.director sneak in there?
+cfor.varimp = cfor.varimp %>% filter(! var == "pc.director")
+
+ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        geom_bar(stat='identity') +
+        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance Rankings of Complete Variables") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+3
+
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
+
+#plot histogram to show differences
+results.core %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models- core variables only") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+4
+
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
+
+#plot histogram to show differences
+results.A2 %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models - all variables sampled") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+5
+
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
+
+#plot histogram to show differences
+results.rf.core %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models- core variables only") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+6
+
+```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
+If the ANOVA model is effective, the difference should be greater than 0.")}
+
+#plot histogram to show differences
+results.rf2 %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models - all variables sampled") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+7
+
+```{r, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
+
+#plot histogram of RMSE's
+results.rf2 %>% ggplot(., aes(x =final.RMSE, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models - Variables Randomly Sampled") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+8
+
+```{r, fig.cap = c("Distribution of 'return per dollar' for all projects")}
+if(.Platform$OS.type == 'windows'){
+        all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv' )
+} else{
+        all8a<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all8a.csv')
+}
+ggplot(all8a, aes(x = return.pdol, y = ..density..)) + 
+        geom_histogram() + geom_density() +
+        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+9
+
+```{r reduced_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Subsets of the data were used for Logistic Regression and randomForests in order to provide datasets without missing values.")}
+
+reduc.totAUC <- readRDS("reduc_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(reduc.totAUC[[2]], reduc.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+reduc.totAUC = melt(reduc.totAUC)
+reduc.totAUC$sigdiff = ifelse(reduc.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from boosted trees") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from boosted trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of each Method") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naiive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
+
+```
+
+```{r mice_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Each method was fed the same imputed full dataset.")}
+
+mice.totAUC <- readRDS("mice_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+mice.totAUC = melt(mice.totAUC)
+mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from boosted trees") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from boosted trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of each Method - Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naiive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
+```
+
+```{r all_violin, fig.height = 4.6, fig.cap = c("Combined violin plots comparing AUC results for each method predicting profit/loss using subsets of data vs. the full imputed data set.")}
+
+reduc.totAUC$Data_Source = "Original"
+mice.totAUC$Data_Source = "Imputed"
+combine.micer <- rbind(reduc.totAUC, mice.totAUC)
+
+combine.micer$Data_Source <- factor(combine.micer$Data_Source, 
+                                    levels=c("Original", "Imputed"))
+
+ggplot(combine.micer, aes(x= factor(L1), y= value)) +
+        geom_violin(aes(fill = Data_Source, colour = Data_Source)) +
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+
+        # stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Data Source") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Data Source") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "Comparison of Original Data vs. Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naiive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
+
+```
+
+```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a cForest blended model. The results from the three best performing models were added as explanatory variables.")}
+
+cfor.varimp = readRDS("cfor_blend_varimp.rds")
+
+ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) + 
+        geom_bar(stat='identity') + 
+        labs(title = "Cforest Variable Importance: Blended Model",
+             x = "Variable", y = "Variable Importance Score")
+
+```
+
+```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a randomForest blended model. The results from the three best performing models were added as explanatory variables.")}
+
+rf.varimp = readRDS("rfblend.rds")
+rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
+rf.varimp = add_rownames(rf.varimp)
+rf.varimp = arrange(rf.varimp, desc(MeanDecreaseAccuracy))
+# order levels
+rf.varimp$rowname <- factor(rf.varimp$rowname,
+                               levels= rf.varimp$rowname)
+
+#plot p values
+ggplot(data=rf.varimp, aes(x=rowname, y=MeanDecreaseAccuracy)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) +
+        geom_bar(stat='identity') +
+        labs(title = "randomForest Variable Importance: Blended Model")
+
+```
+
+```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a Boosted Tree blended model. The results from the three best performing models were added as explanatory variables")}
+
+boost.varimp = readRDS("boostblend.rds")
+boost.varimp = add_rownames(boost.varimp)
+# order levels
+boost.varimp$rowname <- factor(boost.varimp$rowname,
+                               levels= boost.varimp$rowname)
+
+#plot p values
+ggplot(data=boost.varimp, aes(x=rowname, y=rel.inf)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) +
+        geom_bar(stat='identity') +
+        labs(title = "Boosted Trees Variable Importance: Blended Model")
+
+```
+
+```{r blend_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
+100 models were built for each method.")}
+
+auc.list <- readRDS("blend_auc2.rds")
+means = lapply(auc.list, mean) %>% unlist
+greater = means[means> means["orig.log"]] %>% names
+auc.melt = melt(auc.list)
+auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)
+
+ggplot(auc.melt, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = greater, colour = greater)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 Blended Models of each Method") +
+        scale_x_discrete(labels = c("boost" = "BT Blend", 
+                         "log" = "FWLS", 
+                         "average" = "Simple Average Blend",
+                         "orig.log" = "Orignal LR", 
+                         "rf" = "RF Blend",
+                         "simp.boost" = "Simple BT Blend",
+                         "simp.log" = "Simple LR Blend")) +
+        scale_fill_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC greater than Original Logistic Regression") +
+        scale_colour_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC greater than Original Logistic Regression") +
+        theme(legend.position = "bottom")
+
+```
+
+```{r, fig.height = 5.5, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
+150 models were built for each method to achieve a statistical power of 0.8.")}
+
+index = c(2:3)
+pvals = data.frame("method" = names(auc.150)[c(2:3)], 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(auc.150[[1]], auc.150[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric }
+
+signif = pvals %>% filter(p_value<0.05) %>% .$method
+auc.150 = melt(auc.150)
+auc.150$signif = ifelse(auc.150$L1 %in% signif, TRUE, FALSE)
+
+ggplot(auc.150, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = signif, colour = signif)) + 
+        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
+        stat_summary(fun.y = "mean", geom="point", colour = "gray60", shape =3) +
+        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 150 Blended Models of Best Methods") +
+        scale_x_discrete(labels = c( 
+                         "average" = "Simple Average Blend",
+                         "orig.log" = "Orignal LR",
+                         "simp.log" = "Simple LR Blend")) +
+        scale_fill_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC significantly greater than Original Logistic Regression") +
+        scale_colour_manual(values = c('navajowhite','sienna1'),
+                          name = "Mean AUC significantly greater than Original Logistic Regression") +
+        theme(legend.position = "bottom")
+
+```
+
+```{r profit_curve, fig.height = 6, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+
+finance.plot = readRDS("plot_finance_100_4.rds" )
+finance.plot$method <- factor(finance.plot$method, 
+                                    levels=c("average", "simp.log","simp.boost",
+                                             "comp.boost","comp.rf", "FWLS","orig.log",
+                                             "orig.rf","orig.boost"))
+
+ggplot(finance.plot, aes(x= threshold, y= profit.ratio, colour = method)) + 
+        # geom_point(alpha = 0.005, shape = 16, size = 1) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) +
+        stat_summary(fun.y = mean, 
+                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
+                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
+                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
+        stat_summary(fun.y = mean, geom="line") +
+        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .1)) +  
+        theme(legend.position = "none",
+              panel.background = element_rect(fill = 'gray95')) +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Profit Curve for Blended and Individual Models") + 
+        facet_wrap(~method, 
+                   labeller = labeller( method = c("average" = "Simple Average Blend",
+                                       "orig.boost"= "Orig. Boosted Trees",
+                                       "orig.log" = "Orig. Logistic Regression",
+                                       "orig.rf" = "Orig. Random Forest",
+                                       "simp.log" = "Simple Log. Reg. Blend",
+                                       "simp.boost" = "Simple Boosted Tree Blend",
+                                       "comp.boost"= "Complex Boosted Tree Blend",
+                                       "comp.rf"= "Complex randomForest Blend",
+                                       "FWLS" = "FWLS"
+                                       )))
+
+```
+
+```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression Blended Method")}
+
+ggplot(finance.plot %>% filter(method == "simp.log"), 
+       aes(x= threshold, y= profit.ratio, colour = method)) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
+        geom_point(alpha = 0.1, shape = 16) +
+        stat_summary(fun.y = mean, 
+                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
+                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
+                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
+        stat_summary(fun.y = mean, geom="line") +
+        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
+        theme(legend.position = "none", 
+              panel.background = element_rect(fill = 'grey95')) +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Profit Curve for Simple Logistic Regression Blended Model")
+
+```
+
+```{r, fig.height = 5.5, fig.cap = c("Distribution of projects according to the probability outputs from a typical simple Logistic Regression blended model.
+A bargraph of counts and relative profits are shown.")}
+
+reject.plot <- readRDS('reject_plot.rds')
+
+profs <- reject.plot  %>% .$balance.mlsto %>% abs %>% sum
+simp.l2 <- reject.plot %>% filter(compare == 'full') %>% select(simp.log, balance.mlsto, inv.mlsto.log)
+# simp.l2 <- mutate(simp.l2, invoiced.ratio = exp(inv.mlsto.log)/sum(exp(inv.mlsto.log))) 
+simp.l2$profit.ratio <- ifelse(simp.l2$balance.mlsto>0, simp.l2$balance.mlsto/profs,
+                              simp.l2$balance.mlsto/profs)
+simp.l2 <- rbind(simp.l2,simp.l2)
+simp.l2$faceter <- c(rep("hist", 2364), rep("profs",2364))
+simp.l2$prob.bin <- cut(simp.l2$simp.log, seq(0,1,by = 0.05), labels = seq(.05,1,by = 0.05))
+
+ggplot(simp.l2, aes(x = prob.bin)) + 
+        facet_grid(faceter~., scales = "free_y",
+                   labeller = labeller(faceter = c("hist" = "Count of Projects",
+                                       "profs"= "Relative Profits/Losses"
+                                       ))) +
+        geom_bar(data = simp.l2 %>% filter(faceter == "hist")) + 
+        geom_bar(data = simp.l2 %>% filter(faceter == "profs"),
+                 aes(x = prob.bin, y = profit.ratio), stat = "identity") +
+        # geom_bar(data = simp.l2 %>% filter(faceter == "inv"),
+                 # aes(x = prob.bin, y = invoiced.ratio), stat = "identity") +
+        labs(x = "Probability Score", 
+             y = "",
+             title = "Distribution of Project Counts and Project Profits Against Probability") +
+        geom_vline(xintercept = 10.5, colour = hcl(h = 15, l = 65, c = 100), linetype = 1,
+                   alpha = 0.8) +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
+```
+
+```{r TP_plot, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
+conf.plot <- readRDS("conf_plot.rds")
+
+#if model.pred = loss, this is TP
+conf.plot$model.pred <- ifelse(conf.plot$actual == "loss" & 
+                                       conf.plot$model.pred == "loss", "TP", "FN")
+# reorder levels in method to suit TP 
+conf.plot$method <- factor(conf.plot$method, 
+                           levels= conf.plot %>% filter(actual == "loss", model.pred == "TP") %>%
+                                   arrange(desc(pc)) %>% .$method)
+
+ggplot(conf.plot %>% filter(actual == "loss"), 
+       aes(x = method, y = pc, fill = model.pred)) +
+        geom_bar(stat = "identity", position = "stack", alpha = 0.7) +
+        scale_fill_manual(values = c('red2','dodgerblue2'),
+                          labels = c("FN: predicted as profit", "TP: predicted as loss"),
+                          name = "Confusion \nMatrix Reference") +
+        labs(x = "Method (Original or Blended)", y = "Percent of All Loss-Making Projects (%)",
+             title = "Breakdown of True Positives and False Negatives for Loss-Making Jobs") +
+        scale_x_discrete(labels = c("average" = "Simple Av",
+                                      "orig.boost"= "Orig. BT",
+                                      "orig.log" = "Orig. LR",
+                                      "orig.rf" = "Orig. RF",
+                                      "simp.log" = "Simple LR",
+                                      "simp.boost" = "Simple BT",
+                                      "comp.boost"= "Complex BT",
+                                      "comp.rf"= "Complex RF",
+                                      "FWLS" = "FWLS")) +
+        theme(axis.text.x=element_text(angle=30, hjust=1))
+
+```
+
+```{r TN_plot, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
+
+conf.plot <- readRDS("conf_plot.rds")
+
+#if model.pred = loss, this is TP
+conf.plot$model.pred <- ifelse(conf.plot$actual == "profit" & 
+                                       conf.plot$model.pred == "profit", "TN", "FP")
+# reorder levels in method to suit TP 
+conf.plot$method <- factor(conf.plot$method, 
+                           levels= conf.plot %>% filter(actual == "profit", model.pred == "TN") %>%
+                                   arrange(desc(pc)) %>% .$method)
+
+conf.plot$model.pred <- factor(conf.plot$model.pred, 
+                           levels= c("TN","FP"))
+
+ggplot(conf.plot %>% filter(actual == "profit"), 
+       aes(x = method, y = pc, fill = model.pred)) +
+        geom_bar(stat = "identity", position = "stack", alpha = 0.7) +
+        scale_fill_manual(values = c('dodgerblue2','red2'),
+                          labels = c("TN: predicted as profit", "FP: predicted as loss"),
+                          name = "Confusion \nMatrix Reference") +
+        labs(x = "Method (Original or Blended)", y = "Percent of All Profitable Projects (%)",
+             title = "Breakdown of True Negatives and False Positives for Profitable Jobs") +
+        scale_x_discrete(labels = c("average" = "Simple Av",
+                                      "orig.boost"= "Orig. BT",
+                                      "orig.log" = "Orig. LR",
+                                      "orig.rf" = "Orig. RF",
+                                      "simp.log" = "Simple LR",
+                                      "simp.boost" = "Simple BT",
+                                      "comp.boost"= "Complex BT",
+                                      "comp.rf"= "Complex RF",
+                                      "FWLS" = "FWLS")) +
+        theme(axis.text.x=element_text(angle=30, hjust=1))
+
+```
+
+```{r, fig.cap = c("Proportion of profitable and loss-making jobs in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+reject.plot <- readRDS('reject_plot.rds')
+
+ggplot(reject.plot, aes(x = f.rpdol)) +
+        geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) + 
+        scale_y_continuous(labels = scales::percent) +
+        labs(x = "Project Performance", 
+             y = "Relative Frequency",
+             title = "Ratio of Profitable vs. Loss-Making Projects")
+
+```
+
+```{r, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
+Colours are alternated between different jobs to indicate profit/loss magnitude.
+Predictions were from a typical simple Logistic Regression.")}
+ggplot(reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))), 
+             aes(x = f.rpdol, y = abs(balance.mlsto), fill = mlsto)) +
+        geom_bar(stat = "identity") + 
+        theme(legend.position="none") +
+        scale_fill_manual(values = rep(c('dodgerblue4', 'darkolivegreen1'), 59)) + 
+        scale_y_continuous(labels = scales::dollar) +
+        labs(x = "Project Performance", 
+             y = "Cumulative Profits/Losses",
+             title = "Magnitude of Profits/Losses from Rejected Projects") +
+        geom_text(data = NULL, x = .5, y = 720000, size = 3, 
+                  colour = "gray60", hjust = 0,
+                  label = "Alternate Colours \nIndicate Different \nProjects")
+
+```
+
+```{r, fig.cap = c("Distribution of the percentage of hours completed by a professional level employee in the full data set of projects vs.
+the rejected projects only.
+Predictions from a typical simple Logistic Regression were used.")}
+ggplot(reject.plot, 
+             aes(x = pc.pro)) +
+        geom_density() + 
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) +
+        labs(x = "% of Hours Completed by Professional Level Employee", 
+             y = "Density",
+             title = "Density of Percentage of Hours by Professional Level Employees") 
+```
+
+```{r, fig.cap = c("Proportion of each timespan category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+ggplot(reject.plot, 
+             aes(x = b.timespan.cbrt)) +
+        geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) +
+        labs(x = "Timespan", 
+             y = "Relative Frequency",
+             title = "Relative Frequency of Project Timespans") + 
+        theme(axis.text.x=element_text(angle=30, hjust=1))
+
+```
+
+```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+ggplot(reject.plot, 
+             aes(x = return.pdol)) +
+        geom_density() + 
+        facet_wrap(~compare,
+                   labeller = labeller(compare = 
+                                        c("full" = "All Projects",
+                                       "reject"= "Rejected Projects Only"))) +
+        labs(x = "Return per Dollar", 
+             y = "Density",
+             title = "Density Plots of Return per Dollar") +
+        scale_x_continuous(labels = scales::dollar)
+```

### 2016-07-03 finalwriteup/trial_code.rmd
trial code for graphs

@@
 
-```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
+```{r, fig.height = 4.3, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
 If the ANOVA model is effective, the difference should be greater than 0.")}
@@
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models- core variables only") +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models \nCore Variables Only") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
 If the ANOVA model is effective, the difference should be greater than 0.")}
@@
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models - all variables sampled") +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models \nAll Variables Sampled") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
 If the ANOVA model is effective, the difference should be greater than 0.")}
@@
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models- core variables only") +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nCore Variables Only") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-```{r, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
 If the ANOVA model is effective, the difference should be greater than 0.")}
@@
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models - all variables sampled") +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nAll Variables Sampled") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-```{r, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
 
@@
         geom_histogram() + geom_density() +
-        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models - Variables Randomly Sampled") +
+        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models \nVariables Randomly Sampled") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-```{r TP_plot, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
+```{r TP_plot, fig.height = 5, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
 conf.plot <- readRDS("conf_plot.rds")
@@
                           labels = c("FN: predicted as profit", "TP: predicted as loss"),
-                          name = "Confusion \nMatrix Reference") +
+                          name = "Confusion Matrix Reference") +
         labs(x = "Method (Original or Blended)", y = "Percent of All Loss-Making Projects (%)",
@@
                                       "FWLS" = "FWLS")) +
-        theme(axis.text.x=element_text(angle=30, hjust=1))
+        theme(axis.text.x=element_text(angle=30, hjust=1),
+              legend.position = "bottom")
 
@@
 
-```{r TN_plot, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
+```{r TN_plot, fig.height = 5, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
 
@@
                           labels = c("TN: predicted as profit", "FP: predicted as loss"),
-                          name = "Confusion \nMatrix Reference") +
+                          name = "Confusion Matrix Reference") +
         labs(x = "Method (Original or Blended)", y = "Percent of All Profitable Projects (%)",
@@
              y = "Relative Frequency",
-             title = "Ratio of Profitable vs. Loss-Making Projects")
+             title = "Ratio of Profitable vs. Loss-Making Projects") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 
-```{r, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
+```{r, fig.height = 5, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
 Colours are alternated between different jobs to indicate profit/loss magnitude.
@@
              y = "Density",
-             title = "Density of Percentage of Hours by Professional Level Employees") 
+             title = "Density of Percentage of Hours by Professional Level Employees") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 ```
@@
              title = "Relative Frequency of Project Timespans") + 
-        theme(axis.text.x=element_text(angle=30, hjust=1))
+        theme(axis.text.x=element_text(angle=30, hjust=1),
+              panel.background = element_rect(fill = 'gray95'))
 
@@
              title = "Density Plots of Return per Dollar") +
-        scale_x_continuous(labels = scales::dollar)
+        scale_x_continuous(labels = scales::dollar) +
+        theme(panel.background = element_rect(fill = 'gray95'))
 ```

### 2016-07-04 finalwriteup/draft.rmd
gave draft to kerrrie

@@
 
-opts_chunk$set(fig.width=8, fig.depth = 8, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
+opts_chunk$set(fig.width=6.7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
 ```
@@
 
-# Contents
-
-# List of Figures
+# Table of Contents
+
+# Table of Figures
 
@@
 
-**ANOVA** - Anaylis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
+**ANOVA** - Analysis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
 
@@
 
-**ROC** - Reciever Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
+**ROC** - receiver Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
 
@@
 @Lovallo2003 call the practise of analysing a project based on project details and its unique complexities the 'inside view'.
-This constrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details.
+This contrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details.
 For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate.
@@
 
-In contrast to construction, IT projects are a relatively new practise where alternative contractural arangements have been trialled.
+In contrast to construction, IT projects are a relatively new practise where alternative contractural arangements have been trialed.
 Through experimentation with contracts, the Agile movement was developed.
@@
 
-A statistical or machine learning model predicting a measure of project financial success was developed using the available project information and by trialling numerous machine learning algorithms and statistical methods.
+A statistical or machine learning model predicting a measure of project financial success was developed using the available project information and by trialing numerous machine learning algorithms and statistical methods.
 The company uses well established CRM software which readily provides simple output statistics, however, analytical capabilities were limited to:
@@
 It has been shown that these models improve the cost estimation, however industry uptake has not been successful.
-It has proven to be a challenging task persuading managers to use reference class forecasting models in practise and remove themselves from summaries based on intricate details of the completed project.
+It has proven to be a challenging task persuading managers to use reference class forecasting models in practice and remove themselves from summaries based on intricate details of the completed project.
 This may be due to a number of reasons including lack of understanding, time, and lack of collaboration with industry in development of these models.
@@
 It is anticipated that using internal data only also improves user trust and creates tangible meaning for users.
-Blending multiple machine learning and statistical models was trialled which has not yet been applied to the cost estimation problem.
+Blending multiple machine learning and statistical models was trialed which has not yet been applied to the cost estimation problem.
 
@@
 
-@Elfaki2014 discusses how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator, which contributes to the lack of accuracy in predciting final cost of a project.
+@Elfaki2014 discusses how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator, which contributes to the lack of accuracy in predicting final cost of a project.
 He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and, in turn, their estimate is prone to subjectivity.
@@
 Neural networks are also time consuming to train, and must be re-trained and re-tested with each additional piece of data [@Kim2004].
-Nevertheless, Neural Networks and SVM's recieved a significant amount of attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
+Nevertheless, Neural Networks and SVM's received a significant amount of attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
 
@@
 Other studies found variables such as project complexity, duration, team experience, information availablility, site requirements, and labour climate to contribute most [@Akintoye2000][@Trost2003][@Pinto1988].
-As these studies obtained information from questionaires, the available variables also largely depended on the questions in the questionaires, and therefore each study produced a range of different important variables.
+As these studies obtained information from questionnaires, the available variables also largely depended on the questions in the questionnaires, and therefore each study produced a range of different important variables.
 Despite the differences in qualitative outcomes, the studies all generally reported their mean absolute error or a similar metric from the tested models.
@@
 The nearest neighbour algorithm has been used to find the distance between cases in previous studies.
-It is programmed to calculate the euclidean distance in n-dimensional space between cases, where each variable is a dimension [@Kumar2007].
+It is programmed to calculate the Euclidean distance in n-dimensional space between cases, where each variable is a dimension [@Kumar2007].
 3.
@@
 
-In summary, research into cost estimation in the construction industry has trialled a variety of methods ranging from detailed manual analysis, algorithmic models, and case based reasoning.
+In summary, research into cost estimation in the construction industry has trialed a variety of methods ranging from detailed manual analysis, algorithmic models, and case based reasoning.
 Neural networks generally outperformed all other systematic methods, such as regression and case based reasoning, but had the significant disadvantage of being a black box.
@@
 The most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%), which falls squarely within the theory of optimism bias, or the 'inside view' [@Flyvbjerg2011].
-Given this widespread problem in the industry, plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailsed analysis, algorithmic modelling, and case based reasoning.
+Given this widespread problem in the industry, plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modelling, and case based reasoning.
 
@@
 Another study by @Heemstra1992 found that there was no evidence estimation accuracy improved when estimation tools were used.
-@Bergeron1992 similarly found jobs that used algorithmic models were actually associated with less accurate estimtes.
+@Bergeron1992 similarly found jobs that used algorithmic models were actually associated with less accurate estimates.
 However, this trend may be coincidental due to a lack of cases where estimation tools were used [@Molokken2003].
@@
 
-As alluded to previosuly, studies on effort estimation as opposed to project cost estimation are concentrated in the software industry.
+As alluded to previously, studies on effort estimation as opposed to project cost estimation are concentrated in the software industry.
 This leaves a gap in studies for effort estimation in other industries such as the construction industry.
@@
 In the IT industry, CBR applications have been researched using similar methodologies to the construction industry.
-The procedure is described as algorithmically finding similar projects using measurements of euclidean distance in n-dimensional space, where each dimension corresponds to a variable [@Shepperd1996].
+The procedure is described as algorithmically finding similar projects using measurements of Euclidean distance in n-dimensional space, where each dimension corresponds to a variable [@Shepperd1996].
 Then, the similar cases are used to predict effort for the new case.
@@
 Case based reasoning and algorithmic methods such as Neural Networks have found some success in effort estimation (at similar levels) but have not been successfully adopted by industry.
-It was noted that case based reasoning resonated more with decision makers because its predcitions could be better justified by the model.
+It was noted that case based reasoning resonated more with decision makers because its predictions could be better justified by the model.
 The visibility of model structure is very important for future work if the results are to be translated to industry.
@@
 In contrast, effort estimation in the construction industry faces a long tradition of fee structures that is difficult to reinvent.
-This leaves a gap in research of algorithimic models that predict effort (and not construction cost) for consulting companies in the construction industry.
+This leaves a gap in research of algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
 From findings in previous literature, the models should reveal their predictive structure to decision makers and be based on internal company data.
 Review of the literature as a whole reveals a lack in analysis of overall profit improvement a model could provide to a business.
-This could pesuade industry adoption and provide a framework around how models should affect decisions.
+This could persuade industry adoption and provide a framework around how models should affect decisions.
 
@@
 The previous section highlighted the use of Linear Regression, Neural Networks, SVMs and in one case Boosted Trees for prediction, however research on other business problems utilised a wider range of methods.
-These include Naiive Bayes, Random Forests, and machine learned Bayesian Networks.
-The following section presents each method's advantages and disadvatanges and an example of a past successful application.
+These include Naive Bayes, Random Forests, and machine learned Bayesian Networks.
+The following section presents each method's advantages and disadvantanges and an example of a past successful application.
 
@@
 The reason this method is popular is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
-The integration of categorical variables is called Anaylis of Variance (ANOVA) Linear Regression, and the algorithm compares group mean variances within the categorical variables.
+The integration of categorical variables is called Analysis of Variance (ANOVA) Linear Regression, and the algorithm compares group mean variances within the categorical variables.
 
@@
 Many real world phenomena do not correspond to these assumptions which can make it difficult, sometimes impossible, to produce useful results [@Hastie2009][@Putler2012].
-Furthermore, response variables must be continous numeric variables and if two or more predictor variables are correlated, the values of their coefficients are highly unstable.
+Furthermore, response variables must be continuous numeric variables and if two or more predictor variables are correlated, the values of their coefficients are highly unstable.
 On the other hand, correlated variables do not impact the predictive accuracy of the model.
@@
 the 'x' variable.
-To convert log odds to probability, an exponential transformation is first made followed by the transformation from odds to probabilitiy [@Lowry2016].](images/lit_review/logreg.JPG)
+To convert log odds to probability, an exponential transformation is first made followed by the transformation from odds to probability [@Lowry2016].](images/lit_review/logreg.JPG)
 
@@
 The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
-More specifically, it is assumed that each variable is conditionally independent given the respose class, without considering other predictor variables.
+More specifically, it is assumed that each variable is conditionally independent given the response class, without considering other predictor variables.
 Mathematically this means that each variable's probability contribution can be simply calculated by looking at the probability of a response class given a single predictor value.
@@
 This is fine for ranking.
-On the other hand, the output probabilities are not realisitic, and the values themselves should not be used [@Caruana2006].
-Another disadvantage is that in order to calculate the class probability from a numeric variable, a distribution must be assumed (often gaussian) which the numeric data may not fit into neatly [@Caruana2006].
-Finally, Naive Bayes classifers are known to not perform well in binary classification [@Caruana2006].
+On the other hand, the output probabilities are not realistic, and the values themselves should not be used [@Caruana2006].
+Another disadvantage is that in order to calculate the class probability from a numeric variable, a distribution must be assumed (often Gaussian) which the numeric data may not fit into neatly [@Caruana2006].
+Finally, Naive Bayes classifiers are known to not perform well in binary classification [@Caruana2006].
 
@@
 To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
-Three examples of these ensemble decision tree methods are boostrap aggregatiing (bagging), boosting, and Random Forests.
+Three examples of these ensemble decision tree methods are boostrap aggregating (bagging), boosting, and Random Forests.
 Bagging creates multiple trees by sampling a different training data set (*with* replacement) for each tree and then combining tree results [@Breiman1996].
 Random forests is an advanced form of bagging and builds on the idea of training multiple trees from the same data by sampling bootstrapped training sets with replacement.
-However, when creating each tree, a random subset of attributes (variables) are considered at each split.
+However, when creating each tree, a random subset of attributes (variables) is considered at each split.
 The reduced subset of attributes is resampled for each split in the tree.
-This allows dominant variables to be supressed for a fraction of the splits, allowing the algorithm to explore signals in weaker variables.
+This allows dominant variables to be suppressed for a fraction of the splits, allowing the algorithm to explore signals in weaker variables.
 It also prevents the trees from becoming too correlated [@Breiman2001a].
@@
 The links are graphically represented by arcs or edges [@Heckerman1998].
-This kind of conditional relationship is bayesian, where the probabilities in one node are conditional upon values in nodes directed to it as well as preceding nodes.
+This kind of conditional relationship is Bayesian, where the probabilities in one node are conditional upon values in nodes directed to it as well as preceding nodes.
 Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.
@@
 Therefore Bayesian Networks have become an excellent decision support tool.
-Network relationships can be machine learned, however this is not widely included as part of the suite of machine learning methods and is usually recommeneded for modelling complex networks.
+Network relationships can be machine learned, however this is not widely included as part of the suite of machine learning methods and is usually recommended for modelling complex networks.
 Drawbacks of Bayesian Networks include their inability to handle continuous numeric variables (all continuous variables must be discretised).
@@
 
-This concept can be adapted to non linear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
-Advantages of SVM's include the ability to model non linear relationships in data, and that no assumptions about the distributions of the data must be made as the model is non-probabilistic.
+This concept can be adapted to nonlinear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
+Advantages of SVM's include the ability to model nonlinear relationships in data, and that no assumptions about the distributions of the data must be made as the model is non-probabilistic.
 They are also known to be robust over different samples and perform well in high-dimensional feature space [@Auria2008].
@@
 This is to research whether the broader literature found success in methods that cost estimation had not put to the test.
-The simpler statistical methods have been covered already, so the focus of this section are the machine learning prediction methods developed for business decisions.
+The simpler statistical methods have been covered already, so the focus of this section is the machine learning prediction methods developed for business decisions.
 Popular applied topics include predicting stock fluctuation, customer churn analysis, fraud prediction, customer classification, market segment analysis, product success prediction, and recommendation systems (Seng & Chen, 2010).
@@
 Although SVM's performed the best, as discussed, they are often extremely complex and slow, requiring a great deal of memory [@Kumar2007].
-In futher support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms generally agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
+In further support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms generally agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
 
@@
 For example, combining the contrasting advantages of Neural Networks and decision trees is a worthwhile ensemble technique.
-@Tsai2009's study eventually combined these two methods after trialling each one.
+@Tsai2009's study eventually combined these two methods after trialing each one.
 They were run successively and the Neural Network ran first because it had a higher rate of prediction accuracy via its complex non-linear learning properties.
@@
 An exception is @Saradhi2011's study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
-@Saradhi2011 applied popular customer churn predictive models to employee churn - a novel application that focussed internally on employees rather than externally on customers.
+@Saradhi2011 applied popular customer churn predictive models to employee churn - a novel application that focused internally on employees rather than externally on customers.
 The project tested three machine learning classification algorithms that were commonly applied to customer churn.
@@
 Another idea @Saradhi2011 adopted from customer churn methods was the value models.
-Company's calcuate customer lifetime values (CLV's) in order to structure which potentially churning customer's should be acted upon.
+Company's calculate customer lifetime values (CLV's) in order to structure which potentially churning customer's should be acted upon.
 Then the optimal number of customers to reach out to (reaching out incurs a cost) that prevents the most financial loss from churn can be determined.
@@
 Manager's can then start brainstorming what they can do to prevent high value, high risk of churn employees from leaving.
-This extension of the study provided a comprehensive analysis of how business manager's could adopt their findings to improve their business operations.
+This extension of the study provided a comprehensive analysis of how business managers could adopt their findings to improve their business operations.
 It was a valuable addition which is absent from most cost estimation and effort estimation research.
@@
 The prediction models consisted of case based reasoning, Linear Regression, Neural Networks and in one or two instances an ensemble tree or SVM.
-This leaves a gap in the literature in effort estimation for consulting companies in the construction industry in combination with trialling a broader range of prediction methods.
+This leaves a gap in the literature in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods.
 
@@
 The aim of this project is to determine whether the profitability of projects for consulting businesses can be predicted using statistical and machine learning techniques.
-As stated in the literature review, a couple of gaps exist in this research space including working with internal CRM data from a consulting company in the construction industry, trialling ensemble tree methods against the more popular regression and Neural Networks, and analysing the impact model results could have on business decisions.
+As stated in the literature review, a couple of gaps exist in this research space including working with internal CRM data from a consulting company in the construction industry, trialing ensemble tree methods against the more popular regression and Neural Networks, and analysing the impact model results could have on business decisions.
 This section will explain the chosen methodology for fulfilling the research aim using a case study engineering consulting company that charges for their time spent on project delivery.
@@
 
-* percent of hours performed by each professional role over the course of a project
-* timespan of the entered project hours
-* percent of hours performed by 'profesional' employees as opposed to 'technical' employees
-* position of the employee who completed the most hours on each project
-* percent of hours done by the majority contributor to a project
-* total cost of employee hours per project
-* total cost of external subcontractors or disbursements per project
-* total number of users that entered hours on each project
-* mean number of hours per day entered on a project
-* number of disciplines active in a project 
+* Percent of hours performed by each professional role over the course of a project
+* Timespan of the entered project hours
+* Percent of hours performed by 'professional' employees as opposed to 'technical' employees
+* Position of the employee who completed the most hours on each project
+* Percent of hours done by the majority contributor to a project
+* Total cost of employee hours per project
+* Total cost of external subcontractors or disbursements per project
+* Total number of users that entered hours on each project
+* Mean number of hours per day entered on a project
+* Number of disciplines active in a project 
 \ 
@@
 
-* total amount invoiced and renumerated per project
-* mean invoice size per project (for example if a project had four invoices totalling \$10,000, the mean invoice size would be \$2,500)
-* mean invoice per client, i.e.
+* Total amount invoiced and remunerated per project
+* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500)
+* Mean invoice per client, i.e.
 the average invoice size across all invoices sent to a client.
 This gives an indication of whether the client generally does fast paced big jobs which would result in large monthly invoices for example or small jobs with small monthly invoiced amounts.
-* invoice frequency per project. This indicates how drawn out a project was.
-* client invoice frequency which gives an indication of how much work the company does for that client. 
+* Invoice frequency per project. This indicates how drawn out a project was.
+* Client invoice frequency which gives an indication of how much work the company does for that client. 
 \ 
@@
 
-* profit
+* Profit
 
@@
 
-* return per dollar as the measure of profitability
+* Return per dollar as the measure of profitability
 
@@
 Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles.
-This rule was intially practised by Tukey, the inventor of the box plot [@Tukey1977].
+This rule was initially practised by Tukey, the inventor of the box plot [@Tukey1977].
 All outlier projects determined by this method were reviewed with an employee from the case study company and based on their domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined.
@@
 
-Three methods which output variable importance were trialled and compared to test whether they highlighted different variables as valuable.
+Three methods which output variable importance were trialed and compared to test whether they highlighted different variables as valuable.
 This was anticipated as each method models data using distinct theoretical foundations.
 The tested methods were ANOVA Linear Regression, conditional inference forests (cForests), and Random Forests.
-It was expected that Random Forests would favour variables that had more categories, a well known bias; which is also a bias that cForests have overcome [@Strobl2007].
+It was expected that Random Forests would favour variables that had more categories, a well-known bias; which is also a bias that cForests have overcome [@Strobl2007].
 ANOVA is based on linear theory which contrasts conditional forest and Random Forest's free-form structure.
-Therefore, it is benefical to compare the methods' assessments of variable importance.
+Therefore, it is beneficial to compare the methods' assessments of variable importance.
 
@@
 For this problem, the response variable could be continuous or categorical i.e.
-*how* profitable a project was (continous) or whether a project was profitable or not (binary classification).
+*how* profitable a project was (continuous) or whether a project was profitable or not (binary classification).
 The continuous response variable was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
@@
 categorical variables) [@Lunney1970].
-The method is simple and fast but assumes variables are normally distributed and linearlly related.
-The variables must therefore be normalised before analysis to achieve a statistical distribution as close to gaussian as possible.
+The method is simple and fast but assumes variables are normally distributed and linearly related.
+The variables must therefore be normalised before analysis to achieve a statistical distribution as close to Gaussian as possible.
 A disadvantage of this method is that linear relationships and normally distributed variables do not necessarily represent real world data accurately [@Breiman2001b].
@@
 
-The p-value output from ANOVA models give a good indication of which variables have a statistically significant relationship with the target variable.
+The p-value output from ANOVA models gives a good indication of which variables have a statistically significant relationship with the target variable.
 However, the values should not be used to rank the variables against one another, and significant p-values can occur by chance.
@@
 Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted to the original Random Forest.
-If the accuracy of the premuted forest is much less than the original forest, then that variable was important.
+If the accuracy of the permuted forest is much less than the original forest, then that variable was important.
 Permutation importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the Random Forest [@Breiman2005].
@@
 
-Cforests are an alterantive ensemble tree method to Random Forests that overcome the aforementioned shortfalls of Random Forest variable importance measures.
-It is built from ctrees which are decision trees based on a conditional inference framework.
-The key difference between ctrees and standard decision trees is that a significance test is used for splitting instead of a purity measure such as the Gini coefficient, which is centered around information gain.
-
-The procedure behind the significance test at each split in a ctree is as follows.
+Cforests are an alternative ensemble tree method to Random Forests that overcome the aforementioned shortfalls of Random Forest variable importance measures.
+It is built from cTrees which are decision trees based on a conditional inference framework.
+The key difference between cTrees and standard decision trees is that a significance test is used for splitting instead of a purity measure such as the Gini coefficient, which is centered around information gain.
+
+The procedure behind the significance test at each split in a cTree is as follows.
 To determine which explanatory variable should be used at each split, each variable is permuted in every possible way, and a correlation value is calculated between the tested variable and response variable, for each permutation.
@@
 Ensemble trees (Boosted Trees and Random Forests) and Bayesian Networks fit these criteria.
-Regression and naiive bayes models were also included as simple baseline models.
+Regression and Naive Bayes models were also included as simple baseline models.
 Complex models should be measured against simple models that can be built at a fraction of the computational cost.
-The complete list of models tested in this study are:
+The complete list of models tested in this study is:
 
 * Regression - baseline model
-* Naiive bayes - baseline model
+* Naive Bayes - baseline model
 * Bayesian network
@@
 For example, the height of the top horizontal bar indicates the ESS for one cluster and the height of the second from top horizontal bar is the ESS for two clusters.
-The hieght of the vertical branch between the two bars indicates the improvement in ESS by adding a cluster [@Putler2012].
+The height of the vertical branch between the two bars indicates the improvement in ESS by adding a cluster [@Putler2012].
 As more clusters are added, decreases in ESS are less significant and the aim is to choose the number of clusters at an optimal level in this process.
-After initial comparisons of prediction methods had been trialled, the number of clusters for each discretised variable was re-assessed by trialling one more or one less cluster and comparing model output.
+After initial comparisons of prediction methods had been trialed, the number of clusters for each discretised variable was re-assessed by trialing one more or one less cluster and comparing model output.
 
 Once the number of clusters had been decided, each case was assigned a cluster label to replace its numeric variable.
-It was decided that for timespans and invoiced amounts, the discretised variables should be applied to all models (not just Bayesian Networks) because manager's must guess these values initially when operating the model.
+It was decided that for timespans and invoiced amounts, the discretised variables should be applied to all models (not just Bayesian Networks) because managers must guess these values initially when operating the model.
 It is easier for a manager to predict a timespan category than the exact number of days a project will last.
@@
 Therefore, preliminary runs of each method used subsets of the data which had complete data.
-It was possile that if missing data was imputed, predictions from the models that were limited by missing data could improve.
+It was possible that if missing data was imputed, predictions from the models that were limited by missing data could improve.
 Also, a complete data set allows for complete sets of predictions from each method, which can then be blended to further improve results.
-First, imputation methods were trialled against gradient Boosted Tree results since Boosted Trees can handle missing values and should perform equally well with imputed data.
+First, imputation methods were trialed against gradient Boosted Tree results since Boosted Trees can handle missing values and should perform equally well with imputed data.
 
@@
 The final imputed dataset was fed into a Boosted Tree algorithm and if similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
-The imputed dataset was then trialled on the remaining methods and compared to un-imputed trials.
+The imputed dataset was then trialed on the remaining methods and compared to un-imputed trials.
 
@@
 
-A model that is perfectly classified would have all positives correctly classified (1.0 True Postiive Rate) and no incorrectly classified positives (0.0 False Positive Rate).
+A model that is perfectly classified would have all positives correctly classified (1.0 True Positive Rate) and no incorrectly classified positives (0.0 False Positive Rate).
 If a model has a 1.0 TPR and 1.0 FPR (top right corner) this means it is correctly classifying all positives at the expense of incorrectly classifying all negatives as positive.
 It is 'dumbly' classifying all cases as positive.
-if a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previous example.
+If a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previous example.
 There is an 80% chance *any* case will be classified as positive.
@@
 Historically, in statistics this was called model averaging and since the 1990's research has shown that averaging results from different methods provides better predictive accuracy than any single model [@Madigan1994].
-New methods of model blending were developed for the Netflix Prize that applied sophisiticated machine learning techniques to big, messy data.
+New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big, messy data.
 The prize-winning solution was a complex blend of sub-blends which interacted results of individual models with the original variables.
@@
 The multiple layers of blending gave incremental improvements in predictive accuracy, but the computational cost of all the complex layers did not justify the benefits for Netflix in practice and a simplified model was adopted.
-Therefore, in this project just a single layer of blending was trialled.
+Therefore, in this project just a single layer of blending was trialed.
 
@@
 A single training set for a blended model must be built from the test results of the individual models.
-Therefore, five fold cross validation was performed using each individual model to create a full dataset as a compilation of test data results.
+Therefore, fivefold cross validation was performed using each individual model to create a full dataset as a compilation of test data results.
 The complete test data predictions were then added to the complete original dataset, and training and testing data could be partitioned to test the blended models.
@@
 Then, a two-sample power calculation was performed to calculate the number of samples required to achieve statistical power of 0.8.
-The specified number of models were then developed so that the highest performing blended model could be tested against the others for statistical significance.
+The specified number of models was then developed so that the highest performing blended model could be tested against the others for statistical significance.
 The next step was to assess the blended models' performance in terms of improving overall profits for the company.
@@
 
-Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the affect on a business' bottom line was tested.
+Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the effect on a business' bottom line was tested.
 This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs.
@@
 A 95% confidence interval also be determined around the highest point on each curve.
-The final expected increase in profit and the percentage of projects to be rejected presented a clear scenario that the case study business manager's could assess in terms of their business strategy.
+The final expected increase in profit and the percentage of projects to be rejected presented a clear scenario that the case study business managers could assess in terms of their business strategy.
 
@@
 the overlapping important variables were chosen first as well as some others -->
-<!-- * 5 models were tested: baseiline: log reg, naiive bayes. -->
-<!-- * sophisticated ML: boosted reg, random frest, bayesian networks -->
+<!-- * 5 models were tested: baseiline: log reg, Naive Bayes. -->
+<!-- * sophisticated ML: boosted reg, random frest, Bayesian networks -->
 <!-- * best were comapred to a model averaged version -->
@@
 The p-values for the F-statistic of each variable is plotted below.
-The core variables recieved p-values in all 21 models, while 'add.variable' represents p-values for the additional incomplete variable in each model.
-
-```{r, echo=FALSE, fig.height=5, fig.cap = c("P-values of an ANOVA regression's F-statistics used to interpret variable importance")}
+The core variables received p-values in all 21 models, while 'add.variable' represents p-values for the additional incomplete variable in each model.
+
+```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of an ANOVA regression's F-statistics used to interpret variable importance")}
 
@@
 ggplot(plot.pvals, aes(x = mod.vars, y = p.val)) + geom_boxplot() +
-        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+        theme(axis.text.x = element_text(angle = 45, hjust = 1),
+              panel.background = element_rect(fill = 'gray95')) +
         geom_hline(yintercept = 0.05, colour = 'red', linetype = 2) +
@@
 
-The variables listed above had p-values below 0.05, meaning the null hypothesis, that the continuous explanatory variables have no linear relationship with the reponse variable or that categorical variables have the same mean response values for each group, could be rejected.
+The variables listed above had p-values below 0.05, meaning the null hypothesis, that the continuous explanatory variables have no linear relationship with the response variable or that categorical variables have the same mean response values for each group, could be rejected.
 The important variable relationships found in linear ANOVA regression shall be compared to output from Random Forests and cForests below.
@@
 This was not desirable in the early stage of variable selection.
-Instead, the completeness of the variables were assessed and it was decided that all variables with at least 2300 overlapping complete cases would be chosen.
+Instead, the completeness of the variables was assessed and it was decided that all variables with at least 2300 overlapping complete cases would be chosen.
 Additionally, the Random Forest algorithm cannot process categorical variables with more than 53 categories.
@@
 
-```{r, echo = FALSE, fig.depth = 5, fig.cap = c("Variable importance output from a cForest built from 15 core variables")}
+```{r, echo = FALSE, fig.height = 4.5, fig.cap = c("Variable importance output from a cForest built from 15 core variables")}
 
@@
 ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) +
         geom_bar(stat='identity') +
-        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance Rankings of Complete Variables")
+        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance Rankings of Complete Variables") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 
-This was not necessarily expected, particularly between ANOVA and the ensemble tree methods as ANOVA measures linear relationships in constrast to non-linear ensemble trees.
+This was not necessarily expected, particularly between ANOVA and the ensemble tree methods as ANOVA measures linear relationships in contrast to non-linear ensemble trees.
 Random forests favour variables that are numeric or have many categories and this held true for the analysis.
@@
 It was expected that cForests would rank the categorical variables with only 4 categories more fairly.
-This was the case with Discipline and it recieved a mid-level ranking.
+This was the case with Discipline and it received a mid-level ranking.
 Business category of the client and Broad business type however, were ranked similarly to randomForests.
@@
 In contrast, linear relationships are examined between continuous explanatory variables and the response variable.
-Since linear relationships are being tested, continous variables are assumed to be normally distributed.
+Since linear relationships are being tested, continuous variables are assumed to be normally distributed.
 Therefore, where appropriate, numeric variables were normalised via logarithmic or cube root transformations.
@@
 
-results.core = readRDS("report_data/reg_lm_core.rds")
+results.core = readRDS("reg_lm_core.rds")
 # n.samples(data = results.core$diff)
 
-results.A2 = readRDS("report_data/reg_lm.rds")
+results.A2 = readRDS("reg_lm.rds")
 results.A2 <- dplyr::filter(results.A2, diff>-15)
@@
 #plot histogram to show differences
-results.core %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models- core variables only")
+results.core %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models \nCore Variables Only") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 #plot histogram to show differences
-results.A2 %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models - all variables sampled")
+results.A2 %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models \nAll Variables Sampled") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 A reason for this could be that the mean number of cases in the dataset was `r results.core$test.nrow[1]*4` for the core variables but averaged only `r (mean(results.A2$test.nrow)*4) %>% round(0)` for the sampled set of all variables.
-Regression results from ANOVA models were poor - the randomForest algorithm was trialled next.
+Regression results from ANOVA models were poor - the randomForest algorithm was trialed next.
 
@@
 #plot histogram to show differences
-results.rf.core %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models- core variables only")
+results.rf.core %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nCore Variables Only") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 #plot histogram to show differences
-results.rf2 %>% ggplot(., aes(x =diff, y = ..density..))+ geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models - all variables sampled")
+results.rf2 %>% ggplot(., aes(x =diff, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nAll Variables Sampled") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 
-```{r, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
 
 #plot histogram of RMSE's
-results.rf2 %>% ggplot(., aes(x =final.RMSE, y = ..density..))+ geom_histogram() + geom_density() +
-        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models - Variables Randomly Sampled")
+results.rf2 %>% ggplot(., aes(x =final.RMSE, y = ..density..))+ 
+        geom_histogram() + geom_density() +
+        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models \nVariables Randomly Sampled") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 It is possible the less complete variables had non-linear relationships with the response variable.
-RandomForest is also an ensemble method while ANOVA is not.
+Random Forest is also an ensemble method while ANOVA is not.
 Each randomForest is the average of 500 single trees, whereas each ANOVA model is a single model only.
@@
 
-```{r, fig.cap = c("Distribution of "return per dollar" for all projects")}
+```{r, fig.cap = c("Distribution of 'return per dollar' for all projects")}
+
 if(.Platform$OS.type == 'windows'){
@@
 }
-ggplot(all8a, aes(x = return.pdol, y = ..density..)) + geom_histogram() + geom_density() +
-        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects")
+ggplot(all8a, aes(x = return.pdol, y = ..density..)) + 
+        geom_histogram() + geom_density() +
+        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects") +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 
-Further regression methods could have been persued, however randomForest is a high-performing algorithm and was far from providing usable results.
+Further regression methods could have been pursued, however randomForest is a high-performing algorithm and was far from providing usable results.
 It is likely that either not enough data was available or that the given variables did not capably explain all the factors affecting return per dollar.
@@
 The original 'return per dollar' response variable was transformed by converting all 'return per dollar' values less than or equal to zero to 'loss making' jobs and 'return per dollar' values greater than 1 to 'profitable' jobs.
-The following section summarises results from the trialled methods.
+The following section summarises results from the trialed methods.
 
@@
 
-The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were intially developed without imputing data, using the maximum amount of data possible depending on the method.
-Boosted Trees and Naiive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
+The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were initially developed without imputing data, using the maximum amount of data possible depending on the method.
+Boosted Trees and Naive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
 Logistic Regression, Random Forest and Bayesian Networks require datasets without missing values so a subset of complete data was used.
@@
 The comparison of logistic regression required 100 models, while randomForest required over 700 because the AUC distributions were close.
-Bayesian Networks and Naiive Bayes did not perform as well and the power calculation revealed only 5 and 8 models were required respectively.
-Bayesian Networks were then dropped as Naiive Bayes perfomed slightly better and were faster and simple enough to work as a baseline comparison to the higher-performing methods.
+Bayesian Networks and Naive Bayes did not perform as well and the power calculation revealed only 5 and 8 models were required respectively.
+Bayesian Networks were then dropped as Naive Bayes performed slightly better and were faster and simple enough to work as a baseline comparison to the higher-performing methods.
 
@@
 
-```{r reduced_violin, fig.cap = c("Violing plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+```{r reduced_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
 Subsets of the data were used for Logistic Regression and randomForests in order to provide datasets without missing values.")}
 
-reduc.totAUC <- readRDS("report_data/reduc_df_auc.rds")
+reduc.totAUC <- readRDS("reduc_df_auc.rds")
 
@@
         scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different \nfrom boosted trees") +
+                          name = "Significantly different from boosted trees") +
         scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different \nfrom boosted trees") +
+                          name = "Significantly different from boosted trees") +
         labs(x = "Method", y = "Area Under Curve (AUC)",
@@
                          "log" = "Logistic Regression",
-                         "naiive" = "Naiive Bayes", "rf" = "Random Forest"))
+                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
 
@@
 
-Next, data imputation methods were trialled which would make the complete data set available to Logistic Regression and randomForests.
+Next, data imputation methods were trialed which would make the complete data set available to Logistic Regression and randomForests.
 It was possible this would improve AUC results and the same process as above could be repeated.
@@
 
-```{r mice_violin, fig.cap = c("Violing plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+```{r mice_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
 Each method was fed the same imputed full dataset.")}
 
-mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
+mice.totAUC <- readRDS("mice_df_auc.rds")
 
@@
                          "log" = "Logistic Regression",
-                         "naiive" = "Naiive Bayes", "rf" = "Random Forest"))
+                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
 ```
@@
 
-```{r all_violin, fig.cap = c("Combined violin plots comparing AUC results for each method predicting profit/loss using subsets of data vs. the full imputed data set.")}
+```{r all_violin, fig.height = 4.6, fig.cap = c("Combined violin plots comparing AUC results for each method predicting profit/loss using subsets of data vs. the full imputed data set.")}
 
@@
 With complete model predictions, the high performing models could be blended to test whether better predictions were possible.
-For this reason, the imputed data models were propogated forward to the development of blended models.
+For this reason, the imputed data models were propagated forward to the development of blended models.
 
 In summary, the individual binary classification models performed well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a new job will be profitable or loss-making.
-Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model actually affects decisions in practise.
+Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model actually affects decisions in practice.
 Results from model blending and overall profit impact are presented and discussed in the next chapter.
@@
 Blended models combine the predictions from each higher-performing individual model to create averaged or 'blended' predictions.
-In this case, the randomForest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialled ranging from simple to complex.
+In this case, the randomForest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialed ranging from simple to complex.
 The simplest way averaged the probability output (a number between 0 and 1) of all three models.
@@
 
-*Variable Importance: cForest*
-
-```{r, fig.cap = c("Variable importance output from a cForest blended model. The results from the three best performing models were added as explanatory variables.")}
-
-cfor.varimp = readRDS("report_data/cfor_blend_varimp.rds")
+#### Variable Importance: cForest
+
+```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a cForest blended model. The results from the three best performing models were added as explanatory variables.")}
+
+cfor.varimp = readRDS("cfor_blend_varimp.rds")
 
 ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) + 
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) + 
         geom_bar(stat='identity') + 
@@
 
-*Variable Importance: Logistic Regression*
+#### Variable Importance: Logistic Regression
 
@@
 
-*Variable Importance: randomForest*
-
-```{r, fig.cap = c("Variable importance output from a randomForest blended model. The results from the three best performing models were added as explanatory variables.")}
-
-rf.varimp = readRDS("report_data/rfblend.rds")
+#### Variable Importance: randomForest
+
+```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a randomForest blended model. The results from the three best performing models were added as explanatory variables.")}
+
+rf.varimp = readRDS("rfblend.rds")
 rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
@@
 ggplot(data=rf.varimp, aes(x=rowname, y=MeanDecreaseAccuracy)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) +
         geom_bar(stat='identity') +
@@
 
-*Variable Importance: Boosted Trees*
-
-```{r, fig.cap = c("Variable importance output from a Boosted Tree blended model. The results from the three best performing models were added as explanatory variables")}
-
-boost.varimp = readRDS("report_data/boostblend.rds")
+#### Variable Importance: Boosted Trees
+
+```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a Boosted Tree blended model. The results from the three best performing models were added as explanatory variables")}
+
+boost.varimp = readRDS("boostblend.rds")
 boost.varimp = add_rownames(boost.varimp)
@@
 ggplot(data=boost.varimp, aes(x=rowname, y=rel.inf)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) +
         geom_bar(stat='identity') +
@@
 
-*Variable Importance Summary*
+#### Variable Importance Summary
 
@@
 
-```{r blend_violin, fig.cap = c("Violing plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
+```{r blend_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
 100 models were built for each method.")}
 
-auc.list <- readRDS("report_data/blend_auc2.rds")
+auc.list <- readRDS("blend_auc2.rds")
 means = lapply(auc.list, mean) %>% unlist
@@
         scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC greater \nthan Original Logistic \nRegression") +
+                          name = "Mean AUC greater than Original Logistic Regression") +
         scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC greater \nthan Original Logistic \nRegression")
+                          name = "Mean AUC greater than Original Logistic Regression") +
+        theme(legend.position = "bottom")
 
@@
 
-```{r, fig.cap = c("Violing plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
+```{r, fig.height = 5.5, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
 150 models were built for each method to achieve a statistical power of 0.8.")}
@@
         scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC significantly \ngreater than Original \nLogistic Regression") +
+                          name = "Mean AUC significantly greater than Original Logistic Regression") +
         scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC significantly \ngreater than Original \nLogistic Regression")
+                          name = "Mean AUC significantly greater than Original Logistic Regression") +
+        theme(legend.position = "bottom")
 
@@
 I.e.
-the total profits at each point in the profit curve depends on the model parameters that were calculated using the training data and which projects were analysed in the test data.
+the total profits at each point in the profit curve depend on the model parameters that were calculated using the training data and which projects were analysed in the test data.
 Therefore, the process of training a blended model, then calculating the profit curve from the test data was repeated multiple times to create a distribution of profit curves for each method (blended and individual).
@@
 
-```{r profit_curve, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
-
-finance.plot = readRDS("report_data/plot_finance_100_4.rds" )
+```{r profit_curve, fig.height = 6, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+
+finance.plot = readRDS("plot_finance_100_4.rds" )
 finance.plot$method <- factor(finance.plot$method, 
@@
         stat_summary(fun.y = mean, geom="line") +
-        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
-        scale_x_continuous(breaks = seq(0,1, by = .1)) +  
-        theme(legend.position = "none") +
+        scale_y_continuous(breaks = seq(0,110, by = 20)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .2)) +  
+        theme(legend.position = "none",
+              panel.background = element_rect(fill = 'gray95')) +
         labs(x = "Probability Threshold", 
@@
 
-```{r av_profit_curve, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression Blended Method")}
+```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression Blended Method")}
 
@@
         scale_x_continuous(breaks = seq(0,1, by = .1)) + 
-        theme(legend.position = "none") +
+        theme(legend.position = "none", 
+              panel.background = element_rect(fill = 'grey95'),
+              panel.grid.minor = element_blank()) +
         labs(x = "Probability Threshold", 
@@
 ```{r}
-simp.l <- readRDS("report_data/finance_average_model.rds")
+simp.l <- readRDS("finance_average_model.rds")
 
@@
 
-```{r, fig.cap = c("Distribution of projects according to the probability outputs from a typical simple Logistic Regression blended model. A bargraph of counts and relative profits are shown.")}
-
-reject.plot <- readRDS('report_data/reject_plot.rds')
+```{r, fig.height = 5.5, fig.cap = c("Distribution of projects according to the probability outputs from a typical simple Logistic Regression blended model.
+A bargraph of counts and relative profits are shown.")}
+
+reject.plot <- readRDS('reject_plot.rds')
 
@@
         geom_vline(xintercept = 10.5, colour = hcl(h = 15, l = 65, c = 100), linetype = 1,
-                   alpha = 0.8)
+                   alpha = 0.8) +
+        theme(panel.background = element_rect(fill = 'gray95'))
 
@@
 
-```{r TP_plot, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
-conf.plot <- readRDS("report_data/conf_plot.rds")
+```{r TP_plot, fig.height = 5, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
+
+conf.plot <- readRDS("conf_plot.rds")
 
@@
 
-```{r TN_plot, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
-
-conf.plot <- readRDS("report_data/conf_plot.rds")
+```{r TN_plot, fig.height = 5, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
+
+conf.plot <- readRDS("conf_plot.rds")
 
@@
 All individual original methods used substantially lower TPRs at their optimal thresholds, while their TN rates were spread out.
-This means, as the trialled threshold values moved, and the TPR improved, the increasing lost profits from FP's was too great to justify a TPR higher than 4.6%.
+This means, as the trialed threshold values moved, and the TPR improved, the increasing lost profits from FP's was too great to justify a TPR higher than 4.6%.
 Blending the individual methods has made use of their unique strengths and seems to have clustered the probabilities of the actual loss-making jobs more accurately towards 1, and the profitable jobs towards 0.
@@
 ```{r, fig.cap = c("Proportion of profitable and loss-making jobs in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
-reject.plot <- readRDS('report_data/reject_plot.rds')
+
+reject.plot <- readRDS('reject_plot.rds')
 
@@
 
-```{r, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
+```{r, fig.height = 5, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
 Colours are alternated between different jobs to indicate profit/loss magnitude.
 Predictions were from a typical simple Logistic Regression.")}
+
 ggplot(reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))), 
@@
 Predictions from a typical simple Logistic Regression were used.")}
+
 ggplot(reject.plot, 
@@
              y = "Density",
-             title = "Density of Percentage of Hours by Professional Level Employees") 
+             title = "Density of Percentage of Hours by Professional Level Employees") +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
 ```
@@
 ```{r, fig.cap = c("Proportion of each timespan category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+
 ggplot(reject.plot, 
@@
              title = "Relative Frequency of Project Timespans") + 
-        theme(axis.text.x=element_text(angle=30, hjust=1))
-
-```
-
-The statistical distribution of return per dollar in the rejected projects are skewed towards negative returns.
+        theme(axis.text.x=element_text(angle=30, hjust=1),
+              panel.background = element_rect(fill = 'gray95'))
+
+```
+
+The statistical distribution of return per dollar in the rejected projects is skewed towards negative returns.
 This means the algorithm has identified a proportionally much higher percentage of loss-making projects which is not surprising given previous results.
@@
 ```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+
 ggplot(reject.plot, 
@@
              title = "Density Plots of Return per Dollar") +
-        scale_x_continuous(labels = scales::dollar)
+        scale_x_continuous(labels = scales::dollar) +
+        theme(panel.background = element_rect(fill = 'gray95'))
+
 ```
@@
 Additional methods built on top of the current model may be able to produce statistical probabilities instead of probability scores.
-This may be achieved using bayesian methods such as bayesian additive regression trees, however it is uncertain whether this would improve practical effectiveness.
+This may be achieved using Bayesian methods such as Bayesian additive regression trees, however it is uncertain whether this would improve practical effectiveness.
 
@@
 ![User interface for Nearest Neighbour algorithm input and output chart](images/app/NN_top.png)
+
 ![Nearest Neighbour chart coloured by the position of the main employee on the projects](images/app/NN_emp.png)
+
 ![Nearest Neighbour chart coloured by the client for the projects](images/app/NN_client.png)
@@
 ![Tables below the chart in the application give specific details of each of the nearest neighbours](images/app/NN_table_top.png)
+
 ![The nearest neighbour details are split into sections such as 'Finances' and 'Staff Details'](images/app/NN_table.png)
@@
 The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss-making'.
-A range of threshold were trialled, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
+A range of threshold were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
 If a project was accepted, the profits and losses were summed while if a project was rejected the profits and losses were forfeited.

### 2016-07-06 finalwriteup/draft.rmd
ran individual folds

@@
 
+* Abstracts allow readers to quickly decide whether the research is relevant to their own interests
+* get at your main argument quickly
+
+*explain the purpose, goal methods, leave out results?
+
+* why did you do the study?, what is the problem
+* is it a general problem or something specific
+* why is it important?
+* why should someone read your essay?
+* what is your main claim?
+
+* discuss variables and your approaches
+* describe evidence to support your claim
+* overview of most important sources
+
+* Results: answer , was hypothesis supported
+
+* implications of your work, meaning
+
 # Table of Contents
@@
 Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects.
-In many industries, the reality of fixed flat fee structures are unavoidable businesses conform to survive.
+In many industries, the reality of fixed flat fee structures are unavoidable and businesses conform to survive.
 
@@
 @Lovallo2003 tackled the psychology behind the high failure rates of executives in predicting costs of projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
-Their theory stems from Kahneman's work on decision-making that won him the nobel prize for economics in 2002.
+Their theory stems from Kahneman's work on decision-making that won him the Nobel Prize for economics in 2002.
 His research argues that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.

### 2016-07-06 functions.R

@@ -362,5 +362,170 @@ iter.auc <- function(mult =2, x.folds.iter = 5,
 
 # dontknow = map(saveit, function(x) spread(x, iteration, pred))
 
+# new function for saving each individual results from 5-fold cross validation and performing
+# profit curve calc on each of the folds
+
+#each curve.fold will create 5 profit curves for the method. one method tackled at a time.
+curve.fold <- function(df.pred = all10mice, method = "log", list.folds = folds){
+        
+
+        #create NULL dataframe with row names from threshold values
+        prof.curvs = data.frame('threshold' = seq(0, 1, by = .05))
+        
+        
+        #define formula
+        if(method %in% c('log','rf')){
+                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
+        } 
+        
+        if(method == 'boost') {
+                #no JD.Second
+                formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
+        }
+        
+        #loop over each fold, and compile predictions
+        for (j in 1:length(list.folds)) {
+                
+                #turn predictors and response into 'ordered' variables for roc() to work
+                test<- df.pred[list.folds[[j]],]
+                levels(test$f.rpdol) <- c('profit', 'loss')
+                test$f.rpdol <- ordered(test$f.rpdol)
+                
+                if(method %in% 'rf'){
+                        fit <- randomForest(as.formula(formula), data = df.pred[-list.folds[[j]],], 
+                                            mtry = 4, ntree = 1000)
+                        pred <- predict(fit, test, type = "prob")
+                        pred <- pred[,2]
+                        
+                }
+                
+                if(method %in% 'log'){
+                        fit<- glm(as.formula(formula), 
+                                  data = df.pred[-list.folds[[j]],], family = binomial())
+                        pred <- predict(fit, test,
+                                        type = "response")
+                } 
+                if(method %in% 'boost'){
+                        fit <- gbm(as.formula(formula), data = df.pred[-list.folds[[j]],],
+                                   distribution = "bernoulli", n.trees = 10000,
+                                   shrinkage = 0.001, interaction.depth = 5,
+                                   n.minobsinnode = 20)
+                        
[…]

### 2016-07-07 finalwriteup/draft.rmd
write some abstract

--- 2016-07-04 finalwriteup/draft.rmd
@@
 
+* Abstracts allow readers to quickly decide whether the research is relevant to their own interests
+* get at your main argument quickly
+* explain the purpose, goal methods, and results
+
+* why did you do the study?, what is the problem
+performing loss-making jobs for fixed fees is a major problem in the consulting, particularly in the competitive construction industry.
+
+* why is it important?
+affects the porfitability of a company which impacts the careers, internal morale, and even viability of a company.
+
+* why should someone read your essay?
+learn if a company's passively collected data could help them avoid loss making jobs or tactfully enforce strictier contracts.
+
+* what is your main claim?
+improve profitability by 9% by avoiding ~4% of projects is a possible way to interpret the agorithm output. other ways are possible.
+
+* discuss variables and your approaches
+variables: main employee, % time by each type of employee, team size, client charactersitcs, project characteristics (size, type of work, length)
+approaches: complex data with mnay variables, thousands of data points. tried regression as well as machine learning algs: randomf roest and boosted trees, bayesian networks, naive bayes.
+
+* describe evidence to support your claim
+hundred of cross validated trials using the 5 methods, as well as 5 blending methods.
+obtained a 95% confidence interval for each method and simple logistic regression of three models outperformed the others, closely followed by simple averaging.
+the 95% confidence interval across all cross validated trials was __ to __ profit improvement.
+
+* overview of most important sources
+
+* Results: answer , was hypothesis supported
+hypothesis was supprted: CAN improve bottom line of business by mobilising their passively collected data into mL algorhtms.
+
+* implications of your work, meaning
+more work to be done on testing the user interface and user ailities further.
+could assist managers in making better decisions about which projects they should reject or develop a more careful contract structure to improve profits.
+
 # Table of Contents

### 2016-07-07 financial.rmd
ran individual folds
write some abstract

@@
 ```
+
+```{r}
+
+wot <- each.fold(methods = c("log", "rf", "boost"), df.assem = all10mice, x.folds.assem = 5,
+                 iterations = 100, seed = 200)
+
+saveRDS(wot, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/indiv_folds.rds")
+
+if(.Platform$OS.type == 'windows'){
+        ind_folds <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/indiv_folds.rds")
+} else{
+        ind_folds <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/indiv_folds.rds")
+}
+```

### 2016-07-08 financial.rmd
plotted each fold

--- 2016-06-29 financial.rmd
@@
 ```
+
+```{r}
+
+wot <- each.fold(methods = c("log", "rf", "boost"), df.assem = all10mice, x.folds.assem = 5,
+                 iterations = 100, seed = 200)
+
+saveRDS(wot, "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/indiv_folds.rds")
+
+if(.Platform$OS.type == 'windows'){
+        ind_folds <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/indiv_folds.rds")
+} else{
+        ind_folds <- readRDS("/Users/yam/Documents/github/consult/finalwriteup/report_data/indiv_folds.rds")
+}
+
+#plot profit curves for each fold
+names(ind_folds) <- c(1:100)
+folds_plot <- bind_rows(ind_folds, .id = "id")
+folds_plot <- gather(folds_plot, key = iteration, value = profit, -threshold, - id)
+
+temp = gsub("([:0-9:])", "\\1~", folds_plot$iteration, ignore.case = TRUE) %>% strsplit(., "~")
+
+folds_plot$iteration = sapply(temp, function(x) x[1])
+folds_plot$method = sapply(temp, function(x) x[2])
+
+ggplot(folds_plot, aes(x = threshold, y = profit, colour = id)) +
+        facet_wrap(~method) +
+        geom_point()+
+        # geom_line(aes(group = id)) +
+        theme(legend.position = "none")
+
+```

### 2016-07-11 finalwriteup/draft.rmd
now have added indiv folds
started dads corrections

@@
 
-Originality Statement
-
-'I hereby declare that this submission .... look up QUT'
-
-Signed .................
-Date....................
+# Keywords
 
@@
 
-# Table of Figures
+# List of Figures
 
@@
 
+# Acknowledgements
+
+# Publications
+
+Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.
+
+# Statement of Original Authorship
+
+The work contained in this thesis has not been previously submitted to meet requirements for an award at this or any other higher education institution.
+To the best of my knowledge and belief, the thesis contains no material previously published or written by another person except where due reference is made.
+
+Signed .................
+Date....................
+
 # Acronyms
@@
 
+**AUC** - Area under the curve.
+In this project, the area refers to the area under an ROC curve.
+Areas greater than 0.5 indicate predictions performing better than random chance while an area of 1 indicates perfect predictions.
+
 **B2B** - Business to Business. Commerce transactions between two businesses.
+
+**BIM** - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
+
+**CLV** - Customer Lifetime Value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
 
@@
 
+**DA** - Discriminant analysis. Statistical analysis to predict a categorical response variable.
+
 **DAG** - Directed acyclic graph.
@@
 The arrangement of the edges is such that a vertex cannot follow a path that loops back to itself.
+**ESS** - Error sum of squares. The sum of squared errors from a set of predictions.
+
+**FP** - False positive. A case marked 'negative' in a dataset is incorrectly predicted 'positive' by an algorithm.
+
+**FPR** - False positive rate. The proportion of positives that were incorrectly identified (false positives).
+
+**FN** - False negative. A case marked 'positive' in a dataset is incorrectly predicted 'negative' by an algorithm.
+
+**FWLS** - Feature weighted linear stacking.
+A linear combination of model results which are interacted with original variables.
+This allows certain models to be weighted heavier for certain values of the original features.
 
 **IT** - Information technology. The use of computers for storing, retrieving and processing information.
-
-**BIM** - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
-
-**ROC** - receiver Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
-
-**AUC** - Area under the curve.
-In this project, the area refers to the area under an ROC curve.
-Areas greater than 0.5 indicate predictions performing better than random chance while an area of 1 indicates perfect predictions.
-
-**TP** - True positive.
-The case where a case marked 'positive' in a dataset is correctly predicted 'positive' by an algorithm.
-Generally, 'positive' is assigned to categories for which the use case needs to be alerted.
-
-**FP** - False positive. A case marked 'negative' in a dataset is incorrectly predicted 'positive' by an algorithm.
-
-**TN** - True negative. A case marked 'negative' in a dataset is correctly predicted 'negative' by an algorithm.
-
-**FN** - False negative. A case marked 'positive' in a dataset is incorrectly predicted 'negative' by an algorithm.
-
-**TPR** - True positive rate. The proportion of positives that were correctly identified (true positives).
-
-**FPR** - False positive rate. The proportion of positives that were incorrectly identified (false positives).
-
-**DA** - Discriminant analysis. Statistical analysis to predict a categorical response variable.
-
-**CLV** - Customer Lifetime Value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
 
@@
 
-**ESS** - Error sum of squares. The sum of squared errors from a set of predictions.
-
 **RMSE** - Root mean squared error. The square root of the mean of the ESS.
 
-**FWLS** - Feature weighted linear stacking.
-A linear combination of model results which are interacted with original variables.
-This allows certain models to be weighted heavier for certain values of the original features.
-
-# Acknowledgements
-
-# Publications
-
-Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.
+**ROC** - receiver Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
+
+**TN** - True negative. A case marked 'negative' in a dataset is correctly predicted 'negative' by an algorithm.
+
+**TP** - True positive.
+The case where a case marked 'positive' in a dataset is correctly predicted 'positive' by an algorithm.
+Generally, 'positive' is assigned to categories for which the use case needs to be alerted.
+
+**TPR** - True positive rate. The proportion of positives that were correctly identified (true positives).
 
@@
 People make mental calculations of this sort every day - estimating how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish.
-As most people have experienced, these estimates can be particularly inaccurate if the individual is attempting the task for the first time.
-These errors are trivial in everyday activities, but they have more meaning for businesses that sell their time as their chief revenue.
-
-Consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time a consultant spent on the problem.
-The delivery of 'expert advice' is generally in the form of a document (which could be anything from a report to hundreds of drawings) and can range from an hours advice to thousands of hours work.
-Before a project commences, the client and consultant must agree on a fee for the expert advice, or at least a fee structure.
-The nature of how a fee is formulated can be as creative as the engaged parties wish, however a couple of simple examples are as follows: the client agrees to pay for the consultants time by the hour until the task is complete, or the consultant provides a fixed fee to complete the project in full, regardless of their hours spent.
-Both contracts assign financial risk to one party which must then mitigate the risk to ensure profitability.
-
-This thesis focuses on the risk taken by a consulting company when agreeing to a fixed fee.
-Clients commonly collect fixed quotes from several consultants before settling with one or negotiating further so fees must be competitive.
-The way a consulting manager calculates their fixed fee varies from industry to industry and even company to company.
-Typically a consulting manager has experience in the type of project they are quoting and can use a combination of intuition, rules of thumb (such as a fraction of the entire project cost), and comparison to past projects.
-Such methods can result in a range of profit margins which ultimately depend on how well the manager predicts the amount of time a project demands.
-
-This thesis examines the problem of cost prediction via a case study on a company's internal project performance data for the previous twelve years.
-Statistical models based on historical project data were tested for their ability to predict profitability of projects using characteristics available at the beginning of a project.
-Furthermore, the potential for a model's predictions to improve a company's bottom line is investigated.
-This could be achieved with a strategy for how the model's predictions should change managers' fee proposals for risky projects.
-Due to the commercially sensitive nature of the data used, the source is not disclosed in this dissertation.
-Variables have also been renamed to ensure confidentiality is maintained, however, no such obfuscation has been applied in the methods used – which are presented with integrity.
+As most people have experienced, these estimates can be particularly inaccurate if the task is new and difficult.
+These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.
+
+Consulting businesses give expert advice to other professionals in exchange for a fee based on the amount of time their consultanting staff spend on the project.
+The deliverable 'expert advice' is generally in the form of a document or a package of documents.
+A document could range from a 5 page report written by a single person to a sequence of hundreds of drawings assembled by a team of 10 over many years.
+Despite the range of possible deliverables, before any project commences, the client and consultant must agree on a fee, or at least a fee structure.
+The fee structure can be as creative as the engaged parties wish, but in fact tend to fit a few common models.
+A couple of simple examples are:
+
+1. The client agrees to pay for the consultants' time by the hour until the task is complete (referred to as hourly rate) or 
+2. the consultant offers a fixed fee to complete the project in full, regardless of time spent (referred to as fixed price or fixed fee).
+
+In each of these contract models, financial risk is assigned to one of the parties.
+In model 1, the client must pay for an unknown quantity of consultants' billable hours and in model 2 the consultants must complete the project for a fixed price while not knowing how much of their own time will be spent.
+
+This thesis focuses on the risk taken by a consulting company when offering a fixed fee.
+Clients commonly seek fixed price quotes from several consultants before selecting one or negotiating further.
+Quoted fees must be competitive.
+
+The way a consulting manager calculates a fixed price varies from industry to industry and even company to company.
+Typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb (such as a set fraction of the entire client budget for a large project).
+Such qualitative methods can result in profit margins which ultimately depend on the skill (or luck) the manager has in predicting the amount of time a project demands.
+
+This thesis focuses on a single case study consulting company in the construction industry that has persevered through many fixed price projects that resulted in losses.
+A source of insight into the loss-making projects was identified - twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
+Was it possible to find trends in what makes a project loss-making and can these characteristics be identified before a fixed price is offered?
+Can these trends be modelled into a predictive algorithm, and if so, how could the outcomes be integrated into business decisions?
+Finally, if the predictive outcomes were integrated into decision-making, would they deinitely improve the business' bottom line?
+These questions are explored in depth and answered over the course of this dissertation.
+
+Due to the commercially sensitive nature of the data used, the source is not disclosed. Variables have been de-identified to ensure confidentiality is maintained.
 
 # Chapter 2: Problem Description
+
+“It’s tough to make predictions, especially about the future” - Source Unknown
 
@@
 
-Project managers have struggled with forecasting project costs accurately for decades, across many industries.
-A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007].
-A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average [@Flyvbjerg2011].
-These statistics cast a concerning light on projects' hidden financial stories which can cause job losses, ruin private businesses, and shrink government balance sheets.
-
-Although complex projects are more likely to run over budget, they reflect the complexities and priveleges of modern society - from major infrastructure, to housing, websites, and movies.
-All have functional and/or statutory requirements which are defined by a client, and are discharged through contracts to companies via a financially competitive process.
-Consulting companies who complete portions of these complex projects accept the financial risk that is contractually transferred to them.
-More specifically, the risk of the project owner is often transferred to their sub-consultant by mandating fixed flat fee structures to complete a given amount of work.
-The subconsultant commits to completing their role in a project for a nominated price, regardless of whether the work takes less or more time than initially anticipated.
-For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and succeeding.
-
-The ramifications of unprofitable projects are multifold.
+Project managers across many industries have always struggled with forecasting project costs.
+A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by an average of 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007].
+A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget [@Flyvbjerg2011].
+These statistics describe destructive conditions that can result in job losses, business bankruptcies, and government budget blow outs.
+
+Smaller consulting companies in competitive industries, such as the construction industry, experience similar difficulties in forecasting their project costs.
+They must offer appealing fixed prices in order to remain competitive and win projects.
+However, when these projects go over budget it negatively impacts the business in many ways.
 Employee morale deteriorates and employees may view themselves incompetent if they are responsible for the project's delivery.
 Employees also experience elevated stress attempting to complete underestimated projects within a disappearing budget and may compromise quality of work in exchange for speed.
-This experience is unfortunate if in reality, the project fee was under-estimated.
-Unprofitable projects also limit a business' ability to invest money into marketing, training, and business development which would foster growth.
-The financial risk taken on fixed flat fee projects may also discourage businesses from taking other, more calculated risks, such as expanding the business into a new area.
-Logically, improving the prediction of project profitability could positively affect business growth opportunities and staff morale.
-
-**Why It's Not Easily Solvable:**
+This is unfortunate if the only issue with the project was an under-estimated fee.
+Unprofitable projects also limit a business financially, hindering investment into marketing, training, and business development.
+The financial risk taken on fixed price projects may also discourage businesses from taking other, more calculated risks, such as expanding the business into a new area.
+Clearly, improving the prediction of project profitability would improve business growth and staff morale.
+
+**A Wicked Problem**
 
 The challenge for consultants to reduce budget blowouts, in terms of hours spent, is not easily solved.
-Complex projects always differ from one another - similar work may demand significantly different amounts of time.
-For example, an ad filmed in a remote location will cost significantly more than an ad filmed on an existing set, even though the effect of the final advertisement could be identical.
-Therefore, a constant fixed price to produce an advert is clearly inappropriate, but some kind of fixed price is often demanded by the client before a project is awarded.
-Charging clients by the hour is not competitive enough in many industries.
-Other consultancies are often willing to offer a fixed flat fee for the same work, luring the client with reduced financial risk on their part.
-
-Another strategy to mitigate risk is to agree to track employee hours, and stop work once the fixed consulting fee has been depleted.
+Complex projects always differ from one another - even similar work may end up demanding significantly different amounts of time.
+Even though this is known, a fixed price is often demanded by clients before a project is awarded.
+If an hourly rate fee structure is proposed, clients in many industries can simply turn to another consultant willing to take the financial risks involved with fixed price contracts **(...ref...)**.
+
+A compromise between hourly rate and fixed price fee structures is to track employee hours, and stop work once the fixed consulting fee has been exhausted.
 Then, negotiate further fees, or variations, with the client.
-This approach may satisfactorily reduce a consultant’s financial risk, however, may cause friction between the client and the consultant and affect potential repeat custom.
-Especially if the project followed a predictable process but the hours inflated beyond expectation (a common situation), consultants may prefer to wear losses or endure marginable profits to preserve their client relationship.
-Consultants may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
-Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects.
-In many industries, the reality of fixed flat fee structures are unavoidable and businesses conform to survive.
-
-**Current Practice:**
-
-Limited research is available on complex-project cost estimation techniques performed by industry.
-However, two surveys are available from 2000: one on construction projects, the other on IT projects.
+This can either be the consultant's internal strategy or a strategy agreed upon by the client and consultant beforehand.
+If this strategy is proposed to the client before the project is awarded, it is likely the client would again turn to another consultant who is willing to offer a fixed price.
+If this approach is the consultant's internal strategy, their financial risk may satisfactorily be reduced.
+However, this behaviour may cause friction between the client and the consultant and reduce the potential for repeat business.
+If the project followed a predictable process but the hours inflated beyond expectation (a common situation), consultants may prefer to wear losses or endure marginable profits to preserve their client relationship.
+They may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
+Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed prices in complex projects.
+In many industries, the reality of fixed price fee structures are unavoidable and businesses conform to survive.
+
+**Current Fixed Price Estimation Methods:**
+
+It has been established that estimating fixed price fees in consulting is a difficult task with unfortunate consequences if the price is estimated too conservatively.
+Thie project aims to mobilise a consultant's historic data describing profitable and loss-making jobs in order to improve their fixed price setting practices.
+But what is the current method for estimating fixed prices or project costs?
+Is past data currently used in other cost estimation practises and to what degree of success?
+
+Unfortunately, limited research is available that docuemnts the methods industry uses to estimate costs or fixed price fees for complex-projects.
+However, two surveys are available: one from 2000 that covers construction project cost estimation, the other is from 1992 and focuses on IT project effort estimation (which can be likened to fixed price estimation for consultants).
 The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practices.
-They found that the main method for cost estimation was breaking the project into detailed parts and adding up the cost of each item.
+They found that the main method was breaking the project into detailed parts and summing up the cost of each item.
 The next two most popular methods were 'comparison with similar projects based on documented facts', and 'comparison with similar projects based on personal experience'.
@@
 
-A survey by Moores and Edwards in 1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project.
+A survey by Moores and Edwards of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project.
 In the same survey, 91% of software companies cited cost estimation as a problem.
-Although recent literature is limited,  cost estimation was identified as a problem and it was hypothesised that inexperience and lack of time were chief contributors [@Akintoye2000].
+
+Both surveys indicated cost and effort estimation were performed via a detailed breakdown of the tasks required to complete the project.
+Analysis of past project data was alluded to (comparison of similar projects) but a rigorous data analysis process was not common.
+Therefore it is worth investigating if a predictive model based on past data could improve the fixed price setting practice.
+
+**you are here**
 

### 2016-07-14 finalwriteup/draft.rmd
finished reviewing lit review

@@
 
-# Acknowledgements
-
-# Publications
-
-Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.
-
-# Statement of Original Authorship
-
-The work contained in this thesis has not been previously submitted to meet requirements for an award at this or any other higher education institution.
-To the best of my knowledge and belief, the thesis contains no material previously published or written by another person except where due reference is made.
-
-Signed .................
-Date....................
-
-# Acronyms
+# List of Abbreviations
 
@@
 **TPR** - True positive rate. The proportion of positives that were correctly identified (true positives).
+
+# Acknowledgements
+
+# Publications
+
+Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.
+
+# Statement of Original Authorship
+
+The work contained in this thesis has not been previously submitted to meet requirements for an award at this or any other higher education institution.
+To the best of my knowledge and belief, the thesis contains no material previously published or written by another person except where due reference is made.
+
+Signed .................
+Date....................
 
@@
 A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget [@Flyvbjerg2011].
-These statistics describe destructive conditions that can result in job losses, business bankruptcies, and government budget blow outs.
+Another review of 6 IT project surveys between 1984 and 1994 by @Molokken2003 revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
+These statistics describe difficult conditions that can result in job losses, business bankruptcies, and government budget blow outs.
+A survey from @Moores1992 asked software managers whether they saw effort estimation as a problem, and 91% responded 'yes'.
 
@@
 The challenge for consultants to reduce budget blowouts, in terms of hours spent, is not easily solved.
-Complex projects always differ from one another - even similar work may end up demanding significantly different amounts of time.
-Even though this is known, a fixed price is often demanded by clients before a project is awarded.
-If an hourly rate fee structure is proposed, clients in many industries can simply turn to another consultant willing to take the financial risks involved with fixed price contracts **(...ref...)**.
-
-A compromise between hourly rate and fixed price fee structures is to track employee hours, and stop work once the fixed consulting fee has been exhausted.
+Complex projects always differ from one another - even similar projects may end up demanding significantly different amounts of time.
+Although this is known, a fixed price is often demanded by clients before a project is awarded.
+If an hourly rate fee structure is proposed, clients in many industries can simply turn to another consultant willing to take the financial risks involved with fixed prices **(...ref...)**.
+Two strategies to guard against fixed price budget blowouts are detailed below along with their limitations.
+
+A compromise between the hourly rate and fixed price structure is to track employee hours, and stop work once the fixed consulting fee has been exhausted.
 Then, negotiate further fees, or variations, with the client.
-This can either be the consultant's internal strategy or a strategy agreed upon by the client and consultant beforehand.
-If this strategy is proposed to the client before the project is awarded, it is likely the client would again turn to another consultant who is willing to offer a fixed price.
+This can either be agreed upon with the client beforehand or the consultant's internal strategy.
+If this strategy is proposed to the client up front, it is likely the client would again turn to another consultant who is willing to offer a fixed price.
 If this approach is the consultant's internal strategy, their financial risk may satisfactorily be reduced.
-However, this behaviour may cause friction between the client and the consultant and reduce the potential for repeat business.
-If the project followed a predictable process but the hours inflated beyond expectation (a common situation), consultants may prefer to wear losses or endure marginable profits to preserve their client relationship.
-They may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
+However, the behaviour may cause friction between the client and the consultant, therefore reducing the potential for repeat business.
+In the event of hours inflating beyond expectation, consultants may choose to wear losses or endure marginable profits to preserve their client relationship.
+They may reason it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
+
+Another strategy for managing this risk is by stipulating a specific one-off variation opportunity in the contract.
+This applies to consulting work that falls under the client's larger project budget.
+The variation would specify a percentage fee of the overall project cost that can overrule the original fixed price.
+The purpose of this rule is to guard against the case where the project creeps in size, and the consultant finds themselves gradually undertaking more work and more work for the same price even though the client was happy to spend more on the overall project.
+This may account for some or in the best case, all, of their overspent time, although again a savvy client may turn to other consultants that agree to omit variation opportunities.
+
 Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed prices in complex projects.
@@
 
-Unfortunately, limited research is available that docuemnts the methods industry uses to estimate costs or fixed price fees for complex-projects.
-However, two surveys are available: one from 2000 that covers construction project cost estimation, the other is from 1992 and focuses on IT project effort estimation (which can be likened to fixed price estimation for consultants).
+Unfortunately, limited research is available that documents the methods industry uses for complex-projects.
+However, two surveys are available: one from 2000 that covers construction project cost estimation, the other from 1992 and focuses on IT project effort estimation (which can be likened to fixed price estimation for consultants).
 The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practices.
@@
 
-A survey by Moores and Edwards of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project.
+A survey by @Moores1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project.
 In the same survey, 91% of software companies cited cost estimation as a problem.
 
-Both surveys indicated cost and effort estimation were performed via a detailed breakdown of the tasks required to complete the project.
+Both surveys indicated cost and effort estimation were performed via a detailed breakdown of the tasks required.
 Analysis of past project data was alluded to (comparison of similar projects) but a rigorous data analysis process was not common.
@@
 
-**you are here**
-
 **Cost Estimation and Human Nature**
 
-@Lovallo2003 tackled the psychology behind the high failure rates of executives in predicting costs of projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
-Their theory stems from Kahneman's work on decision-making that won him the Nobel Prize for economics in 2002.
-His research argues that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
+Although the surveys indicated industry did not analyse past project cost data, the idea of using this data (in the same way) has been studied before.
+Research exists that tested people's estimation accuracy with and without comparitive historic data with interesting results.
+This is summarised below along with the resulting data-driven models.
+
+@Lovallo2003 delved into the psychology behind why executives so often severely underestimate costs of larger projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
+Their theory stemmed from Kahneman's work on decision-making that won him the Nobel Prize for economics in 2002.
+His research argued that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
 A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplays or ignores the risk of problems that are out of the team's control.
-It is for this reason that it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
+This is why it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
 Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].
 
-@Lovallo2003 call the practise of analysing a project based on project details and its unique complexities the 'inside view'.
+@Lovallo2003 call the practice of analysing a project based on project details without considering unforseen risks the 'inside view'.
+A survey by @Moores1992 aligned with this theory and found the most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%).
 This contrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details.
 For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate.
-According to @Lovallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project are ignored in favour of analysing outcomes of several similar projects.
+According to @Lovallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project should be ignored in favour of analysing outcomes of several similar projects.
 Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003].
-One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
-The correlation for the current project can then be estimated via a statistical model, which is then used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011].
-Managers rarely analyse data from similar projects even though it could significantly improve cost predictions [@Flyvbjerg2011].
-This tendency led to a formalised introduction of 'outside view' methods into some institutions.
+One method @Lovallo2003 recommended was to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
+The correlation for the current project can then be estimated via a statistical model, which is then used to adjust the forecasted cost made by detailed analysis (the 'inside view') [@Flyvbjerg2011].
+As demonstrated in the surveys, @Flyvbjerg2011 found that managers rarely analysed data from similar projects even though it could significantly improve cost predictions.
+This discovery led the team to formally introduce 'outside view' methods into some institutions.
 
@@
 An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].
-An example reading from this chart would be "If the chance of a cost overrun of 20% is acceptable, the estimated cost should be increased by 55% from original calculations.".
+An example reading from the chart is as follows: "If the chance of a cost overrun of 20% is acceptable, the estimated cost should be increased by 55% from original calculations.".
 This builds past experiences of cost overruns into the initial cost estimation.
 
-![This model was implemented by the British government for a rail project that is still under completion](images/lit_review/UKrail.JPG)
+![This model was implemented by the British government for a rail project that is still under completion [@Flyvbjerg2011]](images/lit_review/UKrail.JPG)
 
 The visualisation presents a powerful communication tool for influencing decision makers and improving forecast cost accuracy, however, access to credible data for a sufficient number of projects can be a challenge [@Flyvbjerg2011].
-Overall, the theory of optimism bias is a credible explanation for the problem of persistant underestimation of project costs.
-The accompanying solution of reference class forecasting presents a promising solution given enough accurate data.
-
-The theory of reference class forecasting was first definitively published in 2003, however, the idea of predicting project costs from previous data has been the subject of research for many preceding years and continues to this day.
-Although research results have often been promising, industry uptake of the idea is not yet established.
+Overall, the optimism bias theory credibly helps to explain the persistant problem of project cost underestimation.
 
@@
 
-In the following section, an overview of several attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.
-
-#### Existing Models
-
-A substantial amount of research on project cost prediction has been dedicated to the fields of software development and construction projects, along with some work in the film industry.
-These analyses have generally used project information from 15 to 20 large projects.
-A single dataset may consist of projects completed in different countries with different clients and project teams.
-The mathematical methods used to predict project costs vary from statistical regressions to deep-learning Neural Networks [@Love2005][@Kim2004].
-Most studies report their predictive models perform well and that generally Neural Networks are more accurate than Linear Regression.
-
-A McKinsey study outlined a case in the film industry where a motion picture company used reference class forecasting of movie revenue (cases were weighted by similarity) to decide which movies should be heavily promoted.
-This model improved forecasts on financial return by 135% relative to comparison of a movie to a single past case [@Flyvbjerg2014].
-
-The existing models have shown predictive success, however the data and intention from this case study differs from previous studies.
-Prediction of construction cost is a different problem from consulting effort estimation and uses different variables (except the variables describing the construction project since the consulting company works in the construction industry).
-Software effort estimation projects are estimating time, however past studies have compared projects across many companies (obtained by surveys etc) while the case study company has access to their data only.
-Additionally, their data contains a rich variety of variables which cannot be exctracted from surveys.
-The use of their own data would also improve trust in model outputs.
-
-#### Industry Uptake of Existing Models and Current Practices
-
-A survey of the construction industry in the UK in 2000 found that a manager's experience and intuition is still the dominant method for construction cost calculation [@Akintoye2000].
-Surveys on current industry practice in cost estimation are scarce, however, another study on the IT industry by @Jorgensen2007 argued that previously developed parametric models were not comprehensive enough to be used in industry and that expert judgment was not mobilised enough in the process of developing these tools.
-
-Although the surveys reveal reference class models have not gained traction in either the software or construction industry, it is worth outlining their current practices in cost estimation, fee structures, and the reasons behind these.
-In the construction industry, infrastructure and building projects follow a traditional process, where contractual norms have developed over many decades [@Badenfelt2011].
+Although @Flyvbjerg2011's reference class forecasting theory and the psychology behind it was first published in 2003, predicting project costs using previous data has been researched for many preceding years and continues to this day.
+In this section, a brief overview of other reference class forecasting models is presented (they are reviewed in greater detail in the literature review) followed by an analysis of why industry has not yet adopted them.
+
+Most cost estimation or reference forecasting analytical models focused on software development and construction projects, along with some work on film projects.
+These analyses generally used information from 15 to 20 large projects and the dataset summarised projects from around the world with different clients and teams.
+Applied mathematical methods varied from statistical Regressions to deep-learning Neural Networks [@Love2005][@Kim2004].
+Most studies reported that their predictive models performed well and that Neural Networks were generally more accurate than Linear Regression.
+
+#### Industry Uptake of Existing Models
+
+Despite the predictive success in statistical and machine learning cost estimation models as well as the revealing outcomes from 'inside' and 'outside' view studies, reference class forecasting there has been a lack of industry uptake.
+Two reports reviewing cost estimation in the software industry investigated reasons why data-driven forecasting models have not been translated to industry.
+The key causes were:
+
+* A lack of a decision framework to support manager's use of the 'outside view' model [@Moores1992]
+* A failure in marketing of the potential for the models to improve business outcomes [@Moores1992]
+* The products of the research were not comprehensive enough to be used in industry [@Jorgensen2007]
+* Expert judgment was not mobilised enough in the process of developing the tools and the resulting product felt disconnected to actual requirements [@Jorgensen2007].
+
+Reports on other industries were not available, however the insights are general and may explain the lack of adoption in other industries (such as the construction industry).
+
+In the software industry it is also possible the problem of fixed price contract blowouts has been solved by an alternative contract structure which has made cost estimation models unnecessary.
+In contrast to construction, IT projects are a relatively new practice and the industry has had success in trialing alternative contractural arangements.
+Through experimentation, the Agile movement was developed.
+In summary, the method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
+Furthermore, the project outcomes are continually adjusted and revised based on frequent progress review points that assess unexpected problems.
+The flexible upfront arrangement with clients has often been found to forge better long term relationships [Cockburn2001].
+This agile methodology suits the nature of software development work, where curtailing the scope of work is relatively straightforward.
+Also, since the industry is more modern, the new methodology was able to sweep across the field and could be proven to deliver better budget and product outcomes to clients [@Rasmusson2010].
+
+Unfortunately, in other industries such as the construction industry, infrastructure and building projects are steeped in a traditional process, where contractual norms have developed over a much longer time period [@Badenfelt2011].
 The industry is so large, broad and well established that it has inertia in the way projects are estimated, and contractually bound.
-The contractor (builder) first reviews the drawings and if the plans are not detailed enough, they offer a temporary budget price.
-Once they feel that the drawings are sufficiently detailed, they will offer a fixed price [@Badenfelt2011].
-As the job progresses however, if additional work on top of the detailed drawings is required, the builder maintains the right to charge a variation [@Badenfelt2011].
-A dual source of income therefore exists, the initial proposed fee for construction along with variations due to additional materials (and time required to install the materials) that were not explicitly specified on drawings.
-Variations fees are generally calculated after completion the work, but the initial construction fee contains risk.
-Material costs are forseeable but labour and time estimates can still be under calculated, even from detailed drawings.
-For large infrastructure projects, small percentage miscalculations can be exceedingly expensive.
-
-The builders' procedure of calculating a fixed price from detailed drawings and charging for variations does not translate to other consultants in the construction industry involved with the design of structures, such as engineers and architects.
-These consultants develop the detailed plans for the builders, and must therefore offer fixed price contracts based on low-detail preliminary drawings that sum up the client's expectations.
-It is also much harder for a consultant to charge variations for portions of their time spent on items that were not included in the original simple plans.
-Time is less tangible than physical additional materials required, and often is shrugged off by consultants for the sake of client relationships.
-This is despite the real and common instance of budget blow outs due to overspent time [@Harris1999].
-One method of creating a variation opportunity is by stipulating a percentage fee of the final building cost that can overrule the original fixed price contract.
-If a consultant has spent time beyond their fixed price budget, this may account for some or in the best case, all, of their overspent time.
-
-Construction projects generally are completed more on time and on budget when compared to IT projects as the greatest cost is the project is building materials [@Badenfelt2011].
-Unlike time, building materials can be accurately calculated from detailed design drawings.
-Again, this omits data from consultants who are part of the building industry but charge based on their time only.
-
-In contrast to construction, IT projects are a relatively new practise where alternative contractural arangements have been trialed.
-Through experimentation with contracts, the Agile movement was developed.
-In summary, the method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
-Furthermore, the project outcomes are continually adjusted ad revised based on frequent progress review points that assesses unexpected problems.
-This agile attitude has become ingrained in many IT practices and the flexible upfront arrangement with clients often forges better long term relationships [Cockburn2001].
-
-It is not completely clear why industry uptake of cost estimation tools using the outside view has been weak, and it may be the case that in modern industries such as the IT industry, problems associated with cost estimation can be innovatively designed out of the contract.
-Possible reasons that parametric reference class forecasting models developed by researchers have not been incorporated into industry practice include:
-
-* lack of appropriate data
-    * if a project is unique, such as a national infrastructure project, it many not feel appropriate to gather sufficient data points from other countries and time periods.
-* lack of understanding from industry decision makers - the prediction process was not explained well in marketing of the tools
-* models were not developed in collaboration with industry
-* does not feel intuitive to price from an 'outside view' of a project which disregards small project details
-* insufficient time allocated to pricing that makes change burdensome
-
-In fact, a study by @Moores1992 concluded that a lack of framework to support use of the outside view model and a failure in marketing were the two most common explanations for lack of uptake of parametric cost estimating tools.
-This is unfortunate as almost all research in the area indicates an outside view analysis of a job substantially improves cost estimation accuracy for fixed price projects [@Moores1992].
+Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high level of legal responsibility and due diligence.
+
+This project aims to assist a company in the construction industry in managing problematic fixed price projects, since fixed price contracts are unavoidable.
+The blocks preventing industry adoption, such as providing a clear decision framework and incorporating expert judgment will be carefully addressed.
 
@@
 
-This research explored the potential for reference class forecasting to improve cost estimation in the context of an engineering consulting company in the construction industry.
-In this field, labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm.
-In recent years, the size of losses from unprofitable projects in the case study equalled up to 25-30% of the profits from profitable projects.
-This demonstrates substantial room for improvement and the possiblity of multitude positive flow on effects.
-
-Cost estimation is currently performed by first carefully reviewing preliminary drawings for a project where time costs will be calculated using a manager's personal experience with similar projects.
+The case study company is an engineering consulting company in the construction industry.
+In their profession, labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm.
+In recent years, the size of losses from unprofitable projects equalled 25-30% of their profits from profitable projects.
+This demonstrates the magnitude of their cost estimation problem and the substantial room for improvement that a data-driven model may facilitate.
+
+Cost estimation is currently performed in the case study company by first carefully reviewing preliminary drawings where time costs for each task are estimated and summed by an experienced manager (the 'inside view').
 This is often cross-checked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
@@
 * lack of time to interrogate past project data
-* clunky data availability offered by the current database software
+* clunky data availability offered by their database software
 * lack of awareness of similar projects performed by other managers
 
-A Customer Relationship Management software package (CRM) was employed by this company to collect and store project data.
+A Customer Relationship Management software package (CRM) is currently employed to collect and store project data.
 A CRM is a popular type of software used by businesses to record client and project details as well as employee timesheet records.
 In the case study business, the CRM is available to all employees over the company intranet, and each employee completes daily timesheets alotting their hours to certain projects.
-Additionally, information is recorded against each project as well as each client, client contact, and each invoice.
-
-Reference class forecasting in this study differs from previous research in that statistical analysis was performed on internal past company project data that had been progressively recorded in the company's CRM.
-This presented an opportunity to evaluate reference class forecasting in a unique situation, where the 'reference' projects were all internal and a comparatively large number of cases (over 2,000 projects) were available.
-
-The CRM data stood as an untapped source of information within the case study organisation. It stores a rich variety of data including:
+Additionally, technical information is recorded against each project as well and the CRM data stood as an untapped source of information within the case study organisation.
+It stores a rich variety of data including:
 
@@
 
-A statistical or machine learning model predicting a measure of project financial success was developed using the available project information and by trialing numerous machine learning algorithms and statistical methods.
-The company uses well established CRM software which readily provides simple output statistics, however, analytical capabilities were limited to:
+The CRM software readily provides simple output statistics, however, analytical capabilities were limited to:
 
 * simple scatter plots and bar charts of the raw data
-* summaries such as overall hours spent vs. invoiced amount for each project
+* summaries such as overall hours spent vs. the invoiced amount for each project
 
 The company had a wealth of data but limited means to extract insight.
-This study will exploit the CRM data by performing more sophisticated statistical analysis with the intention of building a predictive model for cost estimation.
-The benefits of this study in comparison to previous cases are that thousands of past data points are available, when previous studies used on average 15-20 cases and at most 300 [@Finnie1997][@Pai2013][@Shin2015].
+This study exploits the CRM data by performing more sophisticated statistical analysis with the intention of building a predictive model to improve cost estimation.
+Thousands of past projects are available for analysis, as opposed to previous studies that used on average 15-20 cases and at most 300 [@Finnie1997][@Pai2013][@Shin2015].
 This improves the potential for accurate predictions.
-If successful, the model also has higher potential for managerial uptake as the algorithm will be directly built and trained on company data.
-Managers can interrogate 'similar' data points and relate to the actual projects or speak to a colleague who was involved.
-This has the potential to give the 'outside view' stronger influence on the final decision.
+If successful, the model also has higher potential for managerial uptake as the algorithm will be directly built and trained on internal company data.
+Managers could interrogate how the model outputs were calculated and relate to the actual projects used in the model or speak to a colleague that was involved.
+This may give the presented 'outside view' a stronger meaning and more potential to influence the final fee proposal.
 
@@
 
-The case study provided a framework to test the value of a consulting company's CRM data in assisting project cost estimates.
-However, it was important to be aware of limitations when reviewing outcomes and framing the problem.
-The most obvious limitation is that a single company is being disected in a specific industry, which limits the ability to state conclusions applicable to a broad range of companies.
-This can be overcome with future studies, once a range of optimal methods and variables have been determined.
-Other constraining factors include various forms of missing data and human data entry error.
-
-The data set from this case study includes scarce information on the overall cost of entire construction projects.
-For example, managers commonly price jobs based on the budget for the entire project, of which the consultant is a minor part.
-The manager might typically moderate his fee by say 1% of the entire project cost.
-The value of the entire project has typically not been recorded although there is an allotted space in the project database for this.
-It would be even more beneficial for managers to record the expected total project cost at the beginning of a project versus the concluding completed cost.
-Another piece of data not recorded was the external factors influencing the fixed price quote including competition from similar firms that clients may have played against the case study company.
-
-In addition to the missing data mentioned above that could be entered by workers in the future, the database also excludes the detailed information about projects that contribute to final pricing.
-Such details include the preliminary drawings describing the project (square meterage, number of storeys), and project summaries provided by the client to enable a quote.
-This level of information will never be recorded in a CRM and in many ways contradicts the purpose of the 'outside view', the intention of which is to present similar projects without being over-influenced by the finer details.
-Given this lack of detailed information, it was not reasonable for an algorithm to predict the numeric fee for a project.
-It appeared much more effective to predict the profitability of a project (ie return per dollar) based on available data.
-This profitability is a ratio between amount invoiced and business costs.
-The algorithm therefore would predict the ratio of these two values based on fees and costs generated using the established method of fee proposals and time expenditure in historical cases.
-
-Finally, CRM data is entered by employees manually as time passes, jobs are invoiced, and projects are finally completed.
-Because the data is input by people, it is susceptible to input error.
+The case study provides an opportunity to test whether passively collected CRM data can be mobilised into a predictive model to improve fixed price cost estimation.
+However, there are some limitations in the study including the size of the study, missing data, and errors in data entry.
+This section addresses how the limitations can be addressed either in future work or within the research framework.
+
+The most obvious limitation is that the work focuses on a single company in a specific industry, which limits the application of conclusions to a broader range of companies.
+This can be overcome by first determining the optimum method and variables for this case and then testing them on future case study companies.
+
+Another limitation is the type of unavailable data.
+Information on overall costs of construction projects was scarce.
+This is relevant because managers sometimes price jobs based on the budget for the entire project, of which the consultant is a minor part.
+For example, engineers may moderate their fee by 1% of the entire project cost.
+In the future, it would be beneficial for managers to record the expected total project cost at the beginning of a project versus the concluding completed cost.
+Another piece of data not recorded was the external factors influencing the fixed price quote including known competition from similar firms.
+
+The database also excludes detailed technical information about each project.
+Such details are found in the preliminary drawings describing the project (square meterage, number of storeys), as well as project summary documents provided by the client to enable a quote.
+However, this level of information actually contradicts the purpose of the 'outside view', the intention of which is to present similar projects without being over-influenced by finer details.
+Given the lack of detailed information describing the physical size of projects, it was not reasonable for an algorithm to predict the numeric fee.
+Instead, it was more logical to predict the profitability of a project i.e.
+return per dollar.
+This measure represents the ratio of the amount invoiced to business costs, and can easily be compared between projects of different sizes.
+
+Finally, all CRM data is entered by employees manually.
+Because the data is input by people, it is susceptible to some error.
 Many errors were detectable during data cleaning, however it is possible that a low number of undetectable human errors exist in the data and marginally influence predictions.
 
-Although this case study presented an opportunity to test the predictive power held within the organisations' existing CRM data, limitations exist which must be understood and built into the framework of the problem.
-These include missing data that could have been entered for each project, but historically has not been required, missing data that is very detailed and will never be available, and the potential for a small degree of human error to be embedded within the data.
-
 ### Research Motivation Summary
 
-Fixed price cost estimation for complex projects is a difficult task and when performed incorrectly, can have devastating affects on businesses and governments.
-Many theories have delved into human nature and our tendency to optimistically assess our capabilities with respect to a given task.
-In this case, the optimism often results in negative business consequences affecting employee morale and the capacity of a company to flourish and grow.
-Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project, by statistically comparing a new project to a collection of similar projects and their characteristics.
-It has been shown that these models improve the cost estimation, however industry uptake has not been successful.
-It has proven to be a challenging task persuading managers to use reference class forecasting models in practice and remove themselves from summaries based on intricate details of the completed project.
-This may be due to a number of reasons including lack of understanding, time, and lack of collaboration with industry in development of these models.
-This research examines the potential for a company's internally generated CRM data to be used in creating a reference class forecasting model.
-Familiarity with the data has the potential to influence managers' decisions in the company more, however its limitations must be understood in order to create a valid model.
+Fixed price cost estimation for complex projects is a difficult task and when performed incorrectly, cause substantial financial losses to businesses and governments.
+Studies have explored human nature and our tendency to optimistically assess our capabilities with respect to a given task.
+In the case of cost estimation, the optimism can result in negative business consequences affecting employee morale and the capacity of a company to flourish and grow.
+Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project.
+This was done by statistically comparing a new project to a collection of similar projects and their characteristics.
+It has been shown that these models improve cost estimation accuracy, however industry has not adopted them.
+Managers tend to predict costs based on intricate details of a project in favour of reference class forecasting.
+This may be due to poor marketing of the cost estimation tools, a lack of a decision framework in relation to the tools, and a lack of collaboration with industry during model development.
 
@@
 Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
-
 The project outcomes are intended to assist the business in predicting project profitability before engagement. Several statistical and machine learning techniques were tested, compared and refined.
@@
 
-Past research has tackled cost estimation in various industries, from different angles and with a range of data, however the problem has rarely been addressed using a single company's internal data.
-In the literature, dozens of projects from other companys and even countries are collected and analysed.
-This thesis uses the internal CRM data from a company which contains over 2,000 past job cases - far more than the usual cost estimation study.
-Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry that estimates and must stand by their fixed fees before a project is won.
-This differs from other studies estimating actual construction costs of projects and is likened more to software development effort estimation although the nature of the work differs considerably.
-
-In addition to addressing cost estimation with a different kind of dataset and different industry sector, ensemble tree methods will be applied which have been minimally tested on this problem.
-To date, Linear Regression, Neural Networks, case based reasoning, and SVM's have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks in some cases and generally outperform Linear Regression.
-Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots provide insight into the model's calculations.
-This is in contrast to Neural Networks and support vector machines (SVM's) which are 'black-box' predictors.
-Providing insight to a user is a major benefit in this case as past research has demonstrated issues with user trust and uptake of models, even if the models were successful.
-It is anticipated that using internal data only also improves user trust and creates tangible meaning for users.
-Blending multiple machine learning and statistical models was trialed which has not yet been applied to the cost estimation problem.
-
-Finally, this study takes the model one step further than previous studies and analyses concrete profit improvements for the company if the model output were to be used in certain decision-making scenarios.
-A few possibilities are presented as ways to take model results into account with business decisions and where possible, final profit calcuations are presented.
-It is anticipated that this kind of thought experiment would further push industry to seriously consider the worth in adopting statistical cost estimation models.
+In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the dataset is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
+These points are outlined in more detail below.
+
+In past research, cost estimation datasets generally consisted of dozens of projects from a range of a companys and even countries.
+This thesis case study only had access to the company's internal data but this contained over 2,000 past jobs - far more than the usual cost estimation study and with many more variables.
+Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry which differs from other studies in the construction industry that estimate actual construction costs.
+Consulting in the construction industry can be likened more to software development effort estimation although the nature of the work differs considerably.
+
+In addition to addressing cost estimation with a different kind of dataset, ensemble tree methods are applied which have been minimally tested on this problem.
+To date, Linear Regression, Neural Networks, Case Based Reasoning, and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks in some cases and generally outperform Linear Regression (**ref**).
+Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots that provide insight into the model's calculations.
+This contrasts Neural Networks and SVM's which are 'black-box' predictors.
+Providing insight to a user is a major benefit as past research has demonstrated issues with user trust and uptake of models, even if the models predicted well.
+It is anticipated that using internal data will also improve trust and create tangible meaning for the user.
+Blending multiple machine learning and statistical models has also not yet been applied to the cost estimation problem and may improve predictions.
+
+Finally, this study propels the predictive model one step further than other studies by analysing bottom-line profit improvements that could result from the research.
+A few possibile ways to integrate model results into business decision-making scenarios are presented, and where possible, final profit increases are evaluated.
+This kind of thought experiment is designed to present a clear case for industry when evaluating if and how to adopt cost estimation models.
 
@@
 
-The remainder of this thesis has the following structure: the next chapter is a thorough review of the literature relating to this thesis to date.
+The next chapter, Chapter 3 is a thorough review of the literature to date that relates to this thesis.
 This covers work solving the cost-estimation problem, studies on statistical and machine-learning models in business applications, and a review of effective statistical and machine learning methods.
 
-The following chapter, Chapter 4, describes the method followed throughout the research process.
+Chapter 4 describes the method followed throughout the research process.
 It includes first how the data was obtained, the lengthy cleaning process, followed by variable importance analysis, variable selection, trials of selected predictive algorithms and problem constructs.
@@
 
-The stages of research outlined in the method were divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion:
+The research outlined in the method were divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion:
 
@@
 
-The results section in each chapter presents key charts and tables that progressively answer the research question: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
-Then, a discussion finalises each chapter, where the results are examined deeply and analysed in a broader context.
+The results section in each of the chapters listed above presents key charts and tables that progressively answer the research question: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
+Then, a discussion finalises each chapter, where the results are examined more deeply and analysed in a broader context.
 The nature of the results are also debated in terms of how surprising they were, their impact on the research hypotheses and limitations as well as suggested future work.
@@
 
-This chapter provides an overview of the literature available in current cost estimation methods, use cases of CRM data, as well as machine learning algorithms and statistical business prediction models.
-The most prominent statistical and machine-learning models developed for business applications shall be reviewed along with models applied to business cases that can be related to this project.
+"If I have seen further than others, it is by standing upon the shoulders of giants." - Isaac Newton
+
+Researching outcomes of comparable past studies may impart how to best approach the problem and reveal shortcomings that can be improved upon.
+This chapter provides an overview of the literature describing early and current cost estimation methods.
+The most prominent machine learning and statistical methods shall then be reviewed followed by a summary of papers on more general business applications that have tried a wider variety of methods.
 Gaps in past research will also be highlighted.
@@
 
-The application for this project is the task of estimating fixed fees for complex consulting projects.
-It is therefore prudent to understand what research has been performed to date and how this project contributes to the body of work.
-The task of estimating fixed fees can be likened to the terms 'cost estimation' or 'effort estimation' in the literature.
+The practice of estimating fixed fees is usually represented by the terms 'cost estimation' or 'effort estimation' in the literature.
 The bulk of research to date has been performed with project data in either the construction industry or software development.
-Research into cost estimation in each of these industries is reviewed below, detailing different methods and outcomes.
-Finally a gap in the literature will be highlighted which this research aims to fill.
+The different methods and outcomes in each of these industries is reviewed below.
+Finally a gap in the literature will be highlighted which this research addresses.
 
@@
 
-Research on cost estimation in the construction industry has primarily focused on predicting the final building cost of construction.
-Various studies have surveyed methods for cost estimation while others have investigated new methods such as building information modelling (BIM) analyisis, case based reasoning, and predictive algorithms.
-The cost estimation methods of detailed analysis (inside view), case based reasoning, and predictive algorithms are reviewed in this section.
+Research in the construction industry has primarily focused on predicting the final building cost of construction.
+Some studies reported surveyed cost estimation practices while others tested new methods such as building information modelling (BIM) analyisis, case based reasoning, and predictive algorithms.
+The cost estimation methods of detailed analysis (the 'inside view'), Case Based Reasoning, and predictive algorithms are reviewed in this section.
 
@@
 [@Akintoye2000].
-@Akintoye2000's study based in the UK found that this is the most popular method for pricing amongst contractors.
-The second most popular method was to carry out an additional comparison with similar projects and make judgments based on personal experience.
-Note that this does not involve a mathematical model.
-
-@Elfaki2014 discusses how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator, which contributes to the lack of accuracy in predicting final cost of a project.
+As referenced in the Research Motivation, @Akintoye2000's study found that this was the most popular method amongst contractors.
+
+@Elfaki2014 discusses how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator.
+This contributes to the lack of accuracy in predicting final costs of a project.
 He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and, in turn, their estimate is prone to subjectivity.
@@
 
-Groundbreaking methods into calculating construction costs by @Ma2014 mobilise the data from BIM models, although this method is still technically a detailed analysis of the costs.
-The idea is to quantify the cost of building a structure directly from a three dimensional model created by the design consultants (engineers and architects).
+Groundbreaking methods into calculating construction costs by @Ma2014 mobilise data from building information modeling (BIM) models, although this is still technically a detailed analysis of the costs.
+The idea is to quantify the cost of building a structure directly from a three dimensional BIM model created by the design consultants (engineers and architects).
 @Ma2014's first trial was to automate the cost calculation of a reinforced concrete structure.
-His chief contribution was to program his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
-From the construction technique and material, a cost can be derived for each element.
+He programmed his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
+From the construction technique and material, a cost was derived for each element.
 This was successful and @Ma2014 aims to further his work so that more construction information can be intelligently obtained from BIM models.
-A system like @Ma2014's would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could be prone to innaccurate estimates similarly to detailed analysis by a person.
-The detailed view does not 'step back' and take into account set backs in construction that may have affected similar projects in the past.
-There is potential for this innovation to be combined with an outside-view system but @Ma2014's is still in early stages of development.
-
-Surveyed information on the latest techniques for cost estimation is limited, however what is available indicates that cost estimation through detailed analysis (inside view) remains the most prevalent method.
-This is despite construction having a long history of projects running over time and over budget using the same cost estimation technique [@Shane2009].
-There is even current research on how to automate this way of cost estimation using BIM.
-The method of detailed sums shall be compared to the results of studies using parametric algorithms and case based reasoning ('outside view') in the following sections.
+A system like @Ma2014's would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could still deliver innaccurate estimates similar to detailed analysis by a person.
+The detailed view does not 'step back' and take into account set backs in construction that may have affected past similar projects.
+There is potential for this innovation to be combined with an 'outside view' system.
+
+Detailed analysis remains the most prevalent method for construction cost estimation despite the industry having a long history of projects running over time and budget using the same technique [@Shane2009].
+Next, the method of detailed sums shall be compared to results from parametric algorithms and case based reasoning ('outside view').
 
@@
 
-As the problem of cost estimation in construction projects is well recognised, dozens of studies have built predictive models of these projects using machine learning and statistical models.
-The aim is to test whether a mathematical comparison of many similar projects, using meaningful descriptive variables, can predict the cost of construction.
+Dozens of studies have built machine learning and statistical models to predict construction costs.
 This section will first review the previous studies and most accurate algorithms and finally summarise industry uptake of the methods.
 
-@Elfaki2014 completed a study on cost estimation techniques and research from 2004 to 2014.
-They found that artificial Neural Networks and SVM's were the most common machine learning techniques used.
-He highlighted that these two machine learning techniques deal well with uncertainty, however lack technical justification to the decision maker (known as a black box).
+@Elfaki2014 completed a review of cost estimation research from 2004 to 2014.
+They found that artificial Neural Networks and SVM's were the most common machine learning techniques.
+He highlighted that they dealt well with uncertainty, however lack technical justification to the decision maker (known as a black box predictor).
 Neural networks are also time consuming to train, and must be re-trained and re-tested with each additional piece of data [@Kim2004].
-Nevertheless, Neural Networks and SVM's received a significant amount of attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
+Nevertheless, Neural Networks and SVM's received significant attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
 
@@
 This is surprisingly late given the rapid uptake of Boosted Trees over the past decade.
-In the study, data from 234 school building construction projects in Korea were used.
-Boosted trees were compared to Neural Networks, which are one of the most popular algorithms to apply to this problem.
-Boosted trees were found to predict costs slightly more accurately than Neural Networks, but not statstically significantly (p value < 0.05).
-Even though the two algorithms performed similarly, the output from Boosted Tree models provide insight into the structure of the model such as variable importance and variable partial dependence plots.
-This additional insight may mean Boosted Trees are preferable over Neural Networks because of their potential to engage decision makers.
-
-A notable amount of literature has studied the predictive power of Multiple Linear Regression in the construction cost estimation problem.
-Often, Linear Regression is the only model assessed, without comparison to other methods such as Neural Networks, which first started appearing in literature in the 1990's [@Kim2004].
+In their study, data from 234 school building construction projects in Korea were used.
+Boosted trees were compared to Neural Networks and were found to predict costs slightly more accurately than Neural Networks, but not statstically significantly (p value < 0.05).
+
+A notable amount of literature tested the predictive power of Multiple Linear Regression in the construction cost estimation problem.
+Often, Linear Regression was the only model assessed, without comparison to other methods such as Neural Networks, which first started appearing in literature in the 1990's [@Kim2004].
 However, even post-1990, many studies focused on Linear Regression only.
 This may be because the technique is straightforward, easy to use and widely available in statistical packages [@Chan2005].
-Out of those that have compared regression to Neural Networks, the studies have shown that Neural Networks outperform regression, however other studies established they are approximately equal [@Kim2004][@Attalla2003].
-@Dissanayaka1999 found that Neural Networks outperformed multiple Linear Regression models, however regression models could be performed first to determine variable importance and condense the number of variables.
-It can be argued that Boosted Trees could perform as well as Neural Networks while providing additional insights (even more so than regression models) such as variable importance and partial dependence of variables against the independent variable, cost.
-
-Despite Neural Networks typically outperforming Linear Regression models, it is interesting to note some of the outcomes of regression studies in cost estimation.
-Different studies tended to produce a wide range of variables that contribute at fluctuating levels in the models.
-This is likely because construction projects are multidisciplinary and involve a range of different parties such as the clients, consultants, contractors and suppliers, each of which contribute varying levels of input [@Chan2005].
+Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004][@Attalla2003].
+@Dissanayaka1999 suggested that Regression models could be performed first to determine variable importance and condense the number of variables, then Neural Networks could be used for better estimation.
+Boosted Trees may perform as well as Neural Networks while providing even more insight than Regression models because it is possible to produce partial dependency plots.
+These visualise the relationship between each variable and the response variable (cost).
+
+Regression models from different studies tended to use a wide range of variables, each of which influenced the models to varying degrees.
+This is likely because construction projects are multidisciplinary and involve many parties such as the clients, consultants, contractors and suppliers, [@Chan2005].
 @Chan2005's study on 87 building projects in Singapore included special, complex projects.
@@
 Other studies found variables such as project complexity, duration, team experience, information availablility, site requirements, and labour climate to contribute most [@Akintoye2000][@Trost2003][@Pinto1988].
-As these studies obtained information from questionnaires, the available variables also largely depended on the questions in the questionnaires, and therefore each study produced a range of different important variables.
-Despite the differences in qualitative outcomes, the studies all generally reported their mean absolute error or a similar metric from the tested models.
-Similar result structures are useful, however there is a gap in the literature in how to translate the improved errors from the model into a well justified argument for industry to reduce losses by integrating a parametric model into their business.
-
-Surveys into construction industry uptake of these algorithmic models are sparse, however a survey by @Akintoye2000 showed that parametric estimating methods have not been adopted by contractors, despite the research into this area over the 20 prior years.
-The majority of contractors use experience based methods.
-@Akintoye2000 hypothesises that this could be due to a lack of familiarity and knowledge about the techniques, doubts whether these techniques are applicable to the construction industry, and the availability of sound data to ensure confidence.
-This presents a clear gap in the literature on how to assess the models in terms of the end user's goals and limitations.
-
-The literature on cost estimation for building projects does not break down the cost estimation task into smaller components, even though a construction project is completed by dozens of consultants and sub-contractors.
-Many of the consultants, such as engineers and architects, produce their own cost estimates for their effort that contributes to the project.
-Their contracts are often fixed price, with minimal room for movement, which means they wear the risk of over-spending their time to meet fixed cost contracts [@Harris1999].
-Furthermore, the existing studies collect data from a wide range of projects that have certain characteristics in common, such as country or building type, and information is obtained through surveys.
-This leaves a gap in research exploring a single contractor's history of projects, which would be available in their company database.
+As these studies obtained information from questionnaires, the available variables depended on the questions.
+Despite the differences in qualitative outcomes, the studies all generally reported their mean absolute error or a similar metric from the tested models, and results were generally positive.
+
+Similar result statistics are useful, however there is a gap in the literature rationalising which model industry should adopt, how to adopt it into a business, and how much loss can be reduced.
+This gap may explain why uptake of these algorithmic models are sparse, as indicated by @Akintoye2000's survey, despite 20 years of prior research.
+@Akintoye2000 hypothesised this could be due to a lack of knowledge about the techniques, doubts whether these techniques are applicable to the construction industry, and the availability of sound data to ensure confidence.
+
+The literature on construction cost estimation does not acknowledge that dozens of consultants and sub-contractors contribute to a large project.
+Most consultants, such as engineers and architects, produce their own cost estimates for their effort that contributes to the project.
+Their contracts are often fixed price, with minimal room for movement, and they wear the risk of over-spending their time [@Harris1999].
+Furthermore, the existing studies collected data from many businesses for a single study, which leaves a gap in research exploring a single contractor's history of projects at one time.
 
@@
 
-Case based reasoning (CBR) is a method of estimation that uses the results of similar cases.
+Case based reasoning (CBR) is a systematic method of expert judgment, where the decision maker manually compares similar projects from his or her experience [@Shepperd1996].
 The idea is that in the case of a new project, similar projects are chosen, before generalisations about the data are made.
@@
 
-A study by @Kim2004 compared CBR to Neural Networks and Linear Regression in the construction cost estimation problem.
-For CBR, the ESTEEM software package was used which algorithmically calculated the similarity of variables, weighted the variables using gradient descent, and deduced the cost using the most similar cases (the specific method for this last step was not explained).
-The discussion compared mean absolute error rates of the three methods and showed that CBR was more accurate than regression, but less accurate than Neural Networks in cost estimation.
-Despite Neural Networks outperforming CBR, CBR maintained significant advantages over Neural Networks which have been referenced in other studies as well.
+A study by @Kim2004 compared CBR to Neural Networks and Linear Regression in construction cost estimation.
+For CBR, the ESTEEM software package algorithmically calculated the similarity of variables, weighted the variables using gradient descent, and deduced the cost using the most similar cases (the specific method for this last step was not explained).
+The discussion compared mean absolute error rates of the three methods and showed that CBR was more accurate than Regression, but less accurate than Neural Networks in estimating cost.
+Despite Neural Networks outperforming CBR, CBR maintained advantages over Neural Networks which have been referenced in other studies as well.
 
@@
 In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to Neural Networks, which must be slowly re-trained and re-tuned.
-New cases also did not need to have every variable completed which can occur often in real, messy data.
-Neural networks on the other hand require complete data.
-Another significant advantage of CBR is the ability for users to review chosen similar cases and make sense of the predicted cost, as opposed to Neural Networks which are a black box [@Kim2004][@Elfaki2014][@Kumar2007].
+New cases also did not need to have every variable complete which can occur often in real, messy data.
+Neural Networks on the other hand require complete data.
+Another advantage of CBR is the ability for users to review chosen similar cases and make sense of the prediction, as opposed to Neural Networks [@Kim2004][@Elfaki2014][@Kumar2007].
 One disadvantage is that accuracy can be highly dependent on the number of selected cases [@Elfaki2014].
@@
 
-CBR can be described as a systematic method of expert judgment, where the decision maker manually compares similar projects from his or her experience [@Shepperd1996].
 CBR that is performed manually through personal experience, without the assistance of algorithms, was the second most popular method for construction estimation in @Akintoye2000's survey in the UK.
@@
 
-In summary, research into cost estimation in the construction industry has trialed a variety of methods ranging from detailed manual analysis, algorithmic models, and case based reasoning.
-Neural networks generally outperformed all other systematic methods, such as regression and case based reasoning, but had the significant disadvantage of being a black box.
+In summary, research into cost estimation in the construction industry has tested a variety of methods ranging from detailed manual analysis, algorithmic models, and CBR.
+Neural Networks generally outperformed other methods, such as Regression and CBR, but had the significant disadvantage of being a black box.
 Alternative methods such as Boosted Trees have the potential to perform as well as Neural Networks and provide insight into the structure of the algorithm.
@@
 There exists a gap in the research in ways to intelligently combine CBR with machine learning methods that could predict well, provide insight to the structure, and engage the user in reviewing similar projects.
-Also, as stated previously, there exists a gap in the application of cost estimating in the construction industry to smaller components of the project, such as consulting and contracting companies who each face a similar problem in estimating their fixed fee.
-These companies could mobilise information in their internal project databases, as opposed to the current body of literature that used data from dozens of companies.
+Also, as stated previously, there exists a gap in the application of cost estimating in the construction industry to smaller contributors to projects.
+These contracting companies face a similar problem in estimating their fixed fee.
+They could mobilise information in their internal project databases, as opposed to the current body of literature that used data from dozens of companies in one study.
 
@@
 
-Like the construction industry, the software industry faces the challenging task of estimating the cost of a project before formal engagement by their client.
-However, unlike construction projects, the major cost associated with the project is *effort* as opposed to the cost of building materials.
-A parallel can be drawn between software development effort estimation and consultants' effort estimation in the construction industry - which was highlighted as a gap in the research of the previous section.
-
-Effort estimation is indeed a problem, and was demonstrated by a review of 6 surveys of IT projects between 1984 and 1994 by @Molokken2003.
-It revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
-A survey of software managers asked whether they saw effort estimation as a problem, and 91% responded 'yes', while only 9% answered 'no' [@Moores1992].
-The most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%), which falls squarely within the theory of optimism bias, or the 'inside view' [@Flyvbjerg2011].
-Given this widespread problem in the industry, plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modelling, and case based reasoning.
+In the software industry, the main component of cost is *effort* as opposed to the cost of building materials in the construction industry.
+A parallel can be drawn between software development effort estimation and consultants' effort estimation in the construction industry - which was highlighted as a gap in the previous section.
+Plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modelling, and Case Based Reasoning.
 
@@
 Similar to the construction industry, expert judgment or detailed analysis, is the most widely practised method for effort estimation [@Shepperd1996][@Molokken2003].
-This is despite many years of research being dedicated to developing algorithmic models and industry tools that, in the research context, outperform expert judgment.
+This is despite many years of research being dedicated to developing algorithmic models that, in the research context, outperform expert judgment.
 A survey by @Moores1992 found that software tools assisting a software project were in significant use, but only the project planning tools, not the cost estimating tools.
-This suggests a mental link for project managers to price jobs based on *details* of the software project as opposed to taking an outside view.
-Another study by @Heemstra1992 found that there was no evidence estimation accuracy improved when estimation tools were used.
+This suggests a mental link for project managers to price jobs based on *details* as opposed to an outside view.
+Another study by @Heemstra1992 found there was no evidence estimation accuracy improved when estimation tools were used.
 @Bergeron1992 similarly found jobs that used algorithmic models were actually associated with less accurate estimates.
 However, this trend may be coincidental due to a lack of cases where estimation tools were used [@Molokken2003].
-The scarcity of real project evidence that algorithmic estimation tools improve estimations and an unintuitive mental jump could be reasons contributing to detailed expert analysis remaining the most widespread technique for effort estimation.
-This trend mirrors what was found in the construction industry.
+The scarcity of real project evidence that algorithmic estimation tools improve estimations and an unintuitive mental jump could be reasons contributing to detailed expert analysis remaining the most widespread technique and mirrors what was found in the construction industry.
 
@@
 
-Algorithmic methods to predict effort for IT project costs has been a popular subject of research.
-This section will assess the details and performance of the algorithmic models, important variables, applications and finally gaps in the literature.
-
-Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although regression is the most popular method in the literature [@Finnie1997][@Pai2013][@Matson1993].
-Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the models accuracy: size [@Shepperd1996][@Finnie1997][@Pai2013].
-Size can refer to the expected number of lines of code in the software package or function point.
-Function point refers to the amount of business functionality expected from an information system product.
+This section assesses the details and performance of the algorithmic models, important variables, and gaps in the literature.
+
+Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although Regression is the most popular method [@Finnie1997][@Pai2013][@Matson1993].
+Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the model's accuracy: size [@Shepperd1996][@Finnie1997][@Pai2013].
+Size can refer to the expected number of lines of code in the software package or function point, which refers to the amount of business functionality expected from the product.
 This is easier to guess correctly at the beginning of a project than lines of code, and therefore leads to more accurate predictions [@Finnie1997].
@@
 
-This model was compared to CBR, which outperformed the model [@Shepperd1996].
+This model was compared to CBR, which outperformed the above model [@Shepperd1996].
 
 There are disadvantages to algorithmic methods, similar to the construction industry.
-First, there are often not enough cases to create a good model in the software industry, particularly if the cases must be from within the company [@Finnie1997][@Pai2013].
+First, there are often not enough cases to create a good model, particularly if the cases must be from within the company [@Finnie1997][@Pai2013].
 A study by @Mendes2004 demonstrated using 67 web projects that cross-company models were significantly less accurate than a within company model, so it is in the company's interest to create an in-house model.
-This means that effort estimation tools should use in-house data and must be adapted by statistical experts to each company [@Shepperd1996].
+This means that effort estimation tools should be refined in-house and adapted by statistical experts to each company [@Shepperd1996].
 
@@
 This may be because no model has proved to be outstandingly successful at consistently predicting required effort [@Finnie1997].
-Also, Neural Networks (the most accurate algorithm) lack the ability to capably explain its results and instill trust in decision makers which is another parallel to research from the construction industry [@Finnie1997].
-
-As alluded to previously, studies on effort estimation as opposed to project cost estimation are concentrated in the software industry.
-This leaves a gap in studies for effort estimation in other industries such as the construction industry.
-Infrastructure projects do not have a function point variable and would need to utilise other information such as expected total project cost (which is notoriously hard to predict based on the previous section), client characteristics, and other project characteristics.
+Again, Neural Networks were criticised for their inability to explain their results [@Finnie1997].
+
+in comparison to the construction industry, infrastructure projects do not have a function point variable and need to utilise other information such as expected total project cost (which is notoriously hard to predict based on the previous section), client characteristics, and other project characteristics.
 
@@
 
-In the IT industry, CBR applications have been researched using similar methodologies to the construction industry.
-The procedure is described as algorithmically finding similar projects using measurements of Euclidean distance in n-dimensional space, where each dimension corresponds to a variable [@Shepperd1996].
-Then, the similar cases are used to predict effort for the new case.
-In @Shepperd1996's study, Linear Regression was employed.
-
-The literature found CBR to perform approximately equally to algorithmic methods such as Neural Networks which contrasts findings in the construction industry [@Shepperd1996][@Finnie1997].
-It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects and provides insight into the prediction that Neural Networks do not [@Finnie1997].
+CBR applications have been researched using similar methodologies to the construction industry.
+In a study by @Shepperd1996, once the similar cases were narrowed down, Linear Regression was employed to predict effort for the new case.
+CBR was found to perform approximately equally to algorithmic methods such as Neural Networks.
+This contrasts results in the construction industry where Neural Networks performed better [@Shepperd1996][@Finnie1997].
+It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects  [@Finnie1997].
 
@@
 
-Research into effort estimation in the software industry has followed a similar path to cost prediction studies on construction projects.
+Effort estimation in the software industry has followed a similar path to cost prediction of construction projects.
 Expert judgment via detailed analysis is still the predominant method for effort/cost estimation.
 Case based reasoning and algorithmic methods such as Neural Networks have found some success in effort estimation (at similar levels) but have not been successfully adopted by industry.
-It was noted that case based reasoning resonated more with decision makers because its predictions could be better justified by the model.
-The visibility of model structure is very important for future work if the results are to be translated to industry.
-Effort estimation remains a problem, however agile methods of project management and effort management are changing the contractual approach of consulting in the software industry [@Badenfelt2011].
-This innovative method keeps delivery times short (and therefore more predictable effort) and defines projects as something with a fixed cost and time.
-Therefore, if a change occurs, other less important elements must be discarded to maintain the fixed amount.
-Software development is a relatively new field in comparison to the construction industry and was able to quickly adopt this new model.
-In contrast, effort estimation in the construction industry faces a long tradition of fee structures that is difficult to reinvent.
-This leaves a gap in research of algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
+It was again noted that CBR and visibility of model structure resonated more with decision makers.
+The software industry is unique in that agile methods of delivery are changing the contractual approach of consulting [@Badenfelt2011].
+This innovative method keeps delivery times short and defines projects as something with a fixed cost and time.
+If a change occurs, other less important elements are discarded to maintain the fixed amount.
+Software development is a relatively new field in comparison to construction, where the tradition of fixed fee structures is difficult to reinvent.
+Also, portions of construction projects cannot be simply curtailed.
+
+There is a need for research on algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
 From findings in previous literature, the models should reveal their predictive structure to decision makers and be based on internal company data.
-Review of the literature as a whole reveals a lack in analysis of overall profit improvement a model could provide to a business.
-This could persuade industry adoption and provide a framework around how models should affect decisions.
+Review of the literature revealed a need for an industry focussed decision framework that outlines how to integrate the predictive model in a way that optimises profits.
 
@@
 
-AS stated, the aim of this project is to use statistical and machine learning techniques to model the profitability of projects.
-Therefore, in order to optimise the model, several statistical and machine learning techniques will be tested and it is prudent to review the range of algorithms that have been successfully applied in the literature.
+As stated, the aim of this project is to use statistical and machine learning techniques to model the profitability of projects.
+Several techniques will be tested and this section reviews a range of predictive algorithms that have been successfully applied to real, messy data.
 These vary from simple methods such as Linear Regression to complex, deep learning Neural Networks.
-The previous section highlighted the use of Linear Regression, Neural Networks, SVMs and in one case Boosted Trees for prediction, however research on other business problems utilised a wider range of methods.
+The review of methods applied to the construction and IT industry highlighted the use of Linear Regression, Neural Networks, SVM's and in one case Boosted Trees for prediction, however research on other business problems utilised a wider range of methods.
 These include Naive Bayes, Random Forests, and machine learned Bayesian Networks.
-The following section presents each method's advantages and disadvantanges and an example of a past successful application.
+The following section presents each method's advantages and disadvantanges and an example of a successful application.
 
@@
 
-Linear regression is one of the simplest and most popular statistical prediction methods [@Hastie2009].
-An equation for a line is learned, which is defined by a constant and a coefficient times the value of each variable.
+Linear Regression is one of the simplest and most popular statistical prediction methods [@Hastie2009].
+An equation for a line is learned, which is defined by a constant and a coefficient multiplied by the value of each variable.
 The coefficients are chosen to minimise the residual sum of squares (least squared) [@Hastie2009].
-The reason this method is popular is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
+The reason for this method's popularity is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
 The integration of categorical variables is called Analysis of Variance (ANOVA) Linear Regression, and the algorithm compares group mean variances within the categorical variables.
@@
 
-Despite limitations of Linear Regression models, they have widespread business applications due to their simplicity and ease of application.
+Despite the limitations of Linear Regression models, they have widespread business applications due to their simplicity and ease of application.
 Regression is used to predict risk in the finance and insurance industries, predict who to target in marketing exercises, and predict consumption spending in the field of economics [@Harrell2013].
-Its simplicity and widespread application make Linear Regression a good starting point for analysis and can serve as a performance benchmark against complex models.
+Its simplicity and success in other applications make Linear Regression a good starting point for analysis and can serve as a performance benchmark against complex models.
 
@@
 
-The method behind Linear Regression can be adapted to predict a binary response variable (1 or 0).
+The method behind Linear Regression can be adapted to the prediction of a binary response variable (1 or 0).
 A linear equation is fit to the log of the odds of the binary response variable against each of the explanatory variables.
-The response variable (log odds) in the linear equation can then be translated back to odds by taking the inverse of log, the exponential.
+The response variable (log odds) in the linear equation can then be translated back to odds by taking the inverse log, i.e.
+the exponential:
+
+$odds = e^{\ln\  odds}$
+
 Finally the odds is converted to probability by the following equation:
@@
 success or failure) [@Moore1989].
-Similar pros and cons exist for Logistic Regression as Linear Regression, and it can also be used as a benchmark to compare other binary predictive models due to its speed and simplicity.
+Similar pros and cons exist for Logistic Regression as Linear Regression, and it can be used as a benchmark to compare other binary predictive models for the same reasons.
 Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survival/death [@Moore1989].
@@
 
-The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to greatly simplify probability calculations for the response variable (the response variable must be categorical).
+The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for the response variable (the response variable must be categorical).
 More specifically, it is assumed that each variable is conditionally independent given the response class, without considering other predictor variables.
-Mathematically this means that each variable's probability contribution can be simply calculated by looking at the probability of a response class given a single predictor value.
-A probability of each response class can be calculated for each explanatory variable.
-These are then combined for each possible response class using Bayes theorem, and the response class with the highest probability is chosen [@Provost2013].
+Mathematically this means that each variable's probability contribution can be calculated by looking at the probability of a response class given a single variable's value.
+The probability of each response class can therefore be calculated for each explanatory variable.
+These are then combined for each possible response class using Bayes theorem, and the class with the highest probability is chosen [@Provost2013].
 
 The advantages of this method are that the conditional independence assumption enables very fast calculations and predictions.
-The method can perform very well for real world tasks because the assumption of independence does not damage predictions significantly.
-This is because if multiple variables are related, the variables independently still direct the prediction in the correct direction.
-The correlated variables will double or triple their emphasis on the predicted response variable, however this does not necessarily impact classification accuracy as the final step is to simply choose the class with the *highest* probability [@Provost2013].
-This is fine for ranking.
-On the other hand, the output probabilities are not realistic, and the values themselves should not be used [@Caruana2006].
-Another disadvantage is that in order to calculate the class probability from a numeric variable, a distribution must be assumed (often Gaussian) which the numeric data may not fit into neatly [@Caruana2006].
-Finally, Naive Bayes classifiers are known to not perform well in binary classification [@Caruana2006].
-
-A real world success story for Naive Bayes classifiers hinged on their ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
+The method can perform well in real world tasks because the assumption of independence does not significantly damage predictions.
+If multiple variables are actually related, the variables separately still direct the prediction in the correct direction.
+The correlated variables will double or triple their emphasis on the predicted response variable, however this does not necessarily impact classification accuracy because the final step is to choose the class with the *highest* probability [@Provost2013].
+This is fine for ranking but the output probabilities are not accurate statistical probabilities [@Caruana2006].
+Another disadvantage is that the distribution of the numeric variable must be assumed (often Gaussian) which the data may not fit into neatly [@Caruana2006].
+
+One real world success story for this classifier hinged on its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
 Naive Bayes was used in a complex spam detection system where new spam emails or toxic text themes could be quickly added to the filtering model.
-The Naive Bayes method provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
+Although the example is quite different to cost estimation, the Naive Bayes method generally provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
 
@@
 Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
-Single decision trees tend to overfit the data and provide low predictive accuracy but have favourable characteristics such as creating intuitive rules that a decision maker can follow in real-life scenarios.
+Single decision trees tend to overfit the data and provide low predictive accuracy but on the other hand create intuitive rules that a decision maker can follow in real-life scenarios.
 In addition, their results do not depend heavily on the skill of the data analyst [@Kabra2011][@Putler2012].
@@
 To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
-Three examples of these ensemble decision tree methods are boostrap aggregating (bagging), boosting, and Random Forests.
+Three examples of these methods are boostrap aggregating (bagging), boosting, and Random Forests.
 Bagging creates multiple trees by sampling a different training data set (*with* replacement) for each tree and then combining tree results [@Breiman1996].
@@
 The limited depth of each tree prevents overfitting at each stage and the combined result of up to thousands of trees is very powerful [@Elith2008].
-
-Ensemble decision trees have a different foundation to traditional statistical prediction methods such as regression.
+The three methods of combining hundreds or thousands of trees turn decision trees into high performing, stable predictors.
+
+Ensemble decision trees have a different foundation to traditional prediction methods such as regression.
 For this reason, there are some advantages and disadvantages.
-Decision trees use a progression of binary splits which means the data is first split by one variable at a certain value, then these subsets are split by another variable and so on in order to partition the response variable values into groups with similar values.
+Decision trees use a progression of binary splits which means the data is first split by one variable at a certain value into two subsets.
+Then each subset is split by another variable and so on in order to partition the response variable values into groups with similar values.
 The aim of each split is to achieve the most 'pure' split of the response values in a greedy fashion.
@@
 
-In contrast to Linear Regression, the predictor variables do not need to be transformed as no assumptions are made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
-Other advantages are that the ensemble methods are very fast in comparison to SVM's and Neural Networks but perform just as well and provide insights such as variable importance and variable relationships [@Sealfon2012].
-Random forests in particular have the benefit being unaffected by noise in the data as well as the capacity to handle a large number of variables [@Sealfon2012].
+In contrast to Linear Regression, the predictor variables do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
+Other advantages are that ensemble tree methods are faster than SVM's and Neural Networks but performing just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
+Random forests in particular have the benefit of being unaffected by noise in the data and has the capacity to handle a large number of variables [@Sealfon2012].
 Caruana [@Caruana2006] tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
-This demonstrates ensemble tree methods are capable of competing with high level machine learning algorithms.
-
-Ensemble trees have numerous advantages however it is important to be aware of their limitations.
+This demonstrates how ensemble tree methods are capable of competing with high level machine learning algorithms.
+
+Although ensemble trees have numerous advantages it is understand their limitations.
 Trees are not built on a probabilistic framework, and therefore their results can not be provided in this framework.
@@
 Also, variable importance tables provided by Random Forests and Boosted Trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
-The methods can also be prone to overfitting if not tuned carefully.
-Ensemble trees have many advantages particularly predictive power, but it is important to recognise that the best method for a problem often depends on the type of problem [@Louppe2014][@Caruana2006].
+The methods can also be prone to overfitting if not tuned carefully [@Louppe2014][@Caruana2006].
 
@@
 A Bayesian Network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
-The model is visually represented by a DAG (directed acyclic graph) and is capable of linking the conditional dependency between any variable to another variable or via other variables.
+The model is visually represented by a directed acyclic graph (DAG) and is capable of linking the conditional dependency between any variable to another variable.
 The links are graphically represented by arcs or edges [@Heckerman1998].
@@
 Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.
-The statistical relationships between variables can be determined from separate sources and manually combined into a single model [@Kragt2009].
 
@@
 They are flexible in the range of relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
-A disadvantage is that although the results are accurate, the algorithm is a black box - meaning the data is input and the results output without providing the user with insight into the methodology.
+A disadvantage that has been mentioned earlier is that the algorithm is a black box.
 Neural networks can perform high level tasks as a result of deep learning such as hand writing recognition, vehicle control, face recognition, and cancer detection [@Haykin2004].
@@
 The simplest type of SVM's calculate the linear boundary between two categories in feature space so that the category of future cases may be determined depending on which side of the boundary they fall.
-The location of the boundary is determined by the position of the widest clear margin that can be drawn between data points from the two categories, while allowing a certain number of misclassifications.
+The location of the boundary is determined by the position of the widest clear margin that can be drawn between data points from the two categories, while allowing a certain number of misclassifications (this is usually optimised for the problem).
 The simple plot below demonstrates this concept and has no misclassifications:
@@
 
-This concept can be adapted to nonlinear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
+The concept can be adapted to nonlinear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
 Advantages of SVM's include the ability to model nonlinear relationships in data, and that no assumptions about the distributions of the data must be made as the model is non-probabilistic.
@@
 In certain cases these simpler models are significantly outperformed by complex models however it is important to compare both.
-Neural networks and SVM's are excellent machine learning predictors which can perform well with messy data and no distributional assumptions, however they are slow and do not provide insight into the results of analysis.
-This is not appropriate for the effort estimation problem in this project because communication of the reasoning behind predictions is crucial for uptake of the model by decision makers.
-
-On the other hand, ensemble tree methods are capable of performing as well as Neural Networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
-Bayesian networks are excellent for presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally been excluded from the machine learning literature in comparison to the other methods.
+Neural networks and SVM's are excellent machine learning predictors but are slow black box algorithms and are not appropriate for the effort estimation problem.
+
+Ensemble tree methods are capable of performing as well as Neural Networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
+Bayesian networks are excellent for visually presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally been excluded from the machine learning literature in comparison to the other methods.
 This may be due to low awareness of learned Bayesian Networks or lower performance.
-In conclusion, simple models such as linear/Logistic Regression and Naive Bayes are good baseline models to implement, while ensemble tree methods and Bayesian Networks may handle messy 'real world' data better than the simpler models while still providing insight to decision makers.
+In conclusion, simple models such as linear/Logistic Regression and Naive Bayes are good baseline models to implement, while ensemble tree methods and Bayesian Networks may handle messy 'real world' data better than the simpler models while providing insight to decision makers.
 
@@
 
-The aim of this research project is to create a predictive model for the profitability of a company's consulting projects using their internal CRM data.
-The literature pertaining closely to the cost prediction topic has been reviewed in the first section of this chapter, however it is relevant to broaden the critique of the literature on advanced machine learning techniques that have been applied to general business problems.
-This is to research whether the broader literature found success in methods that cost estimation had not put to the test.
+Literature pertaining closely to the cost prediction topic has been reviewed in the first section of this chapter, however it is relevant to broaden the critique to advanced machine learning methods that have been applied to general business problems.
+This is to find out whether methods that cost estimation had not put to the test succeeded in the broader literature.
 The simpler statistical methods have been covered already, so the focus of this section is the machine learning prediction methods developed for business decisions.
 Popular applied topics include predicting stock fluctuation, customer churn analysis, fraud prediction, customer classification, market segment analysis, product success prediction, and recommendation systems (Seng & Chen, 2010).
-This section will first discuss the use and merits of the advanced machine learning methods applied to these problems, followed by a case study of employee-churn prediction.
+First, the performance of the advanced machine learning methods will be discussed, followed by a case study of employee-churn prediction.
 
@@
 
-According to Breiman [@Breiman2001b], simple theoretical models such as Linear Regression are informative but often provide unsatisfactory accuracy for modelling real life data.
-This data is often messy and does not necessarily align with the rules of statistical theory.
-As this project has a large quantity of messy data, it is worthwhile reviewing the algorithmic machine-learning predictive models that have developed in recent years.
-They have found success because they were able to fit real-life data with more accuracy than theoretical models [@Breiman2001b].
-
 Ravi Kumar and Ravi [@Kumar2007] performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
-The most widely used model was Neural Networks, however logistic and Linear Regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
-Supporting @Breiman2001b's statements, it was found that DA and Linear Regression techniques were not preferred due to their low accuracy.
+The most widely used model was Neural Networks, however Logistic and Linear Regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
+It was found that DA and Linear Regression techniques were not preferred due to their low accuracy.
 The overall assessment was that SVM's outperformed Neural Networks (back propagation Neural Networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
-Although SVM's performed the best, as discussed, they are often extremely complex and slow, requiring a great deal of memory [@Kumar2007].
-In further support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms generally agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
-
-As SVM's are often computationally too expensive and advanced for smaller scale business applications, which this research project would fall into, it is useful to further compare Neural Networks and decision trees which are simpler than SVM's but are also capable of achieving very accurate predictive models.
-Ravi Kumar and Ravi [@Kumar2007] found that Neural Networks and decision trees were both capable of out-performing the other, depending on the context - which is limited to bankruptcy prediction in Kumar's study.
-Both have advantages and disadvantages to consider, however as derived in the previous section, @Tsai2009's study on earnings management prediction found the most notable drawback for Neural Networks is the difficulty in interpreting the reasoning behind the algorithm.
-Earning's management is influenced by managers' choices and insight into the model structure could provide explanations for human behaviour - a valuable output in business problems.
-On the other hand, trees are known to be intuitive to understand and have been argued to function similarly to how the human mind thinks [@Carrizosa2010].
-
-Ravi Kumar and Ravi's [@Kumar2007] study concludes that ensemble techniques, which refers to combinations of two completely different algorithms, can often outperform individual techniques.
+Although SVM's performed the best, as discussed, they are often complex and slow, requiring a great deal of memory [@Kumar2007].
+In further support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
+
+When comparing Neural Networks to ensemble trees, Ravi Kumar and Ravi [@Kumar2007] found that Neural Networks and ensemble trees were both capable of out-performing the other, depending on the context.
+
+Their study also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
 For example, combining the contrasting advantages of Neural Networks and decision trees is a worthwhile ensemble technique.
-@Tsai2009's study eventually combined these two methods after trialing each one.
-They were run successively and the Neural Network ran first because it had a higher rate of prediction accuracy via its complex non-linear learning properties.
-Then, to resolve the lack of explanatory qualities that Neural Networks are notorious for, decision trees were employed.
+@Tsai2009's study combined these two methods after trialing each one.
+They were run successively with the Neural Network running first because it had a higher rate of prediction accuracy.
+Then, to resolve the notorious lack of explanatory qualities, decision trees were employed.
 The 81% of cases that were correctly predicted by the Neural Network were used to generate decision trees, and in turn useful decision rules [@Tsai2009].
@@
 
-The literature on predictive business models is generally limited to applications on product sales, customer behaviour, or overall performance of large businesses.
-There are far fewer studies that assist with internal business decisions dealing with employees, teams, and delivering projects which is what this research project aims to achieve.
+Beside effort estimation, predictive business models in the literature generally focus on response variables external to the inner workings of a business such as product popularity, customer behaviour, or stock performance.
+There are far fewer studies that analyse problems affecting internal business decisions such as dealing with employees, teams, and clients.
 An exception is @Saradhi2011's study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
 @Saradhi2011 applied popular customer churn predictive models to employee churn - a novel application that focused internally on employees rather than externally on customers.
-The project tested three machine learning classification algorithms that were commonly applied to customer churn.
-The reasoning behind this was that the variables and model behaviour of employee churn could be related to customer churn.
-The associated costs of losing customers and finding new customers can also be connected to the costs of losing staff and hiring new staff.
+The project tested three machine learning classification algorithms that have been commonly applied to customer churn.
+The associated costs of losing customers and finding new customers can be connected to the costs of losing staff and hiring new staff.
 
 Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
-All three models performed at similar levels for overall accuracy at around 80% (the general ranking was Random Forests, followed by SVM then Naive Bayes).
-However, when the true positive rates (accuracy in  predicting of employee resignations) of the models were compared, the SVM model far out performed Random Forests and Naive Bayes by achieving 81% TP vs.
+All three models performed at similar levels for overall accuracy at around 80% (Random Forests had the best results, followed by SVM then Naive Bayes).
+However, when the true positive rates (TPR, accuracy in  predicting employee resignations) were compared, the SVM model far out performed Random Forests and Naive Bayes by achieving 81% TPR vs.
 51% and 55% respectively.
-This was attributed to the ability of SVM's to incorporate class penalties whereas the other two methods were limited by the class imbalance problem [@Saradhi2011].
-This means that SVM's had an effective method of increasing the importance of 'positive' cases in the data set which was skewed towards the majority of employees who kept their job versus the minority who left (25% churn).
-It should be noted that the Boosted Tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVMs.
-Furthermore, Boosted Trees provide insight into the predictive model which could build trust in the company's decision makers.
-Nevertheless, it is noteworthy that their study proved employee churn models for customers could be translated to predicting internal churn of employees.
-
-Another idea @Saradhi2011 adopted from customer churn methods was the value models.
+This was attributed to the ability of SVM's to incorporate class penalties whereas the other methods were limited by the class imbalance problem [@Saradhi2011].
+This means that SVM's had an effective method of increasing the importance of 'positive' cases whicih was advantageous because the data set was skewed towards the majority of employees who kept their job versus the minority who left (25% churn).
+It should be noted that the Boosted Tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVM's.
+Furthermore, Boosted Trees provide insight into the predictive model to the decision makers.
+It is noteworthy that their study proved employee churn models for customers could be translated to predicting internal churn of employees.
+
+Another idea @Saradhi2011 adopted from customer churn analysis was the value models.
 Company's calculate customer lifetime values (CLV's) in order to structure which potentially churning customer's should be acted upon.
-Then the optimal number of customers to reach out to (reaching out incurs a cost) that prevents the most financial loss from churn can be determined.
-@Saradhi2011 developed a system for determining the value of each employee in terms of the importance of the projects they were on and their monthly chargeability.
-This allowed them to rank the employees identified as 'high risk of churn' by value and provides a clear direction for manager's to act upon.
-Manager's can then start brainstorming what they can do to prevent high value, high risk of churn employees from leaving.
-This extension of the study provided a comprehensive analysis of how business managers could adopt their findings to improve their business operations.
-It was a valuable addition which is absent from most cost estimation and effort estimation research.
+Then the optimal number of customers to reach out to can be determined by optimising the cost of reaching out against preventing financial loss from churn.
+A system was developed for determining the value of each employee in terms of the importance of the projects they were on and their monthly chargeability.
+This allowed them to rank employees identified as 'high risk of churn' by value and provided a clear ranking for manager's to act upon.
+Manager's could then start brainstorming actions to prevent high value, high risk employees from leaving.
+This extension of the study provided a comprehensive framework for how business managers could adopt their findings to improve business operations.
+It was a valuable addition which is absent from most cost and effort estimation research.
 
@@
 
-The limited literature on analysis similar to @Saradhi2011's Employee Churn study highlights a gap in the application of predictive techniques to model internal performance in businesses.
-This is particularly relevant to consulting companies that tackle complicated discrete projects with set timelines.
-In this research, internal consulting project data will be used to predict a new project's profitability in terms of internal employee structure as well as client characteristics.
-A review of the literature in machine learning applied to business problems revealed that SVM's were found to be the most accurate technique followed by Neural Networks and ensemble tree methods.
-As discussed in the previous section, Neural Networks and SVM's do not provide insight into the predictions to the decision maker.
+The limited literature on work similar to @Saradhi2011's Employee Churn study highlights a gap in the application of predictive techniques that use internal data to model internal decisions.
+This is particularly relevant to consulting companies that have complicated internal processes necessary for each project.
+In this research, internal employee team structure as well as client records shall be used to aid prediction.
+The literature on machine learning applied to general business problems revealed SVM's were usually the most accurate technique followed by Neural Networks and ensemble tree methods.
 A gap exists in applying these cutting edge machine learning techniques to the wide range of internal decisions that businesses make.
 
-## Conclusion
+## Literature Review Conclusion
 
 The literature pertaining to the aim of this research project, cost prediction using internal consulting CRM data, has been extensively reviewed.
-The problem of cost prediction has been addressed in many studies beforehand which have generally been limited to prediction of construction costs for buildings and infrastructure as well as effort estimation for IT projects.
-The prediction models consisted of case based reasoning, Linear Regression, Neural Networks and in one or two instances an ensemble tree or SVM.
-This leaves a gap in the literature in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods.
-
-A comprehensive list of machine learning and statistical models were reviewed theoretically as well as practically in the broader scope of business problem prediction.
-It was determined that Bayesian Networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as Neural Networks while providing the additional benefit of insights into the reasoning behind model predictions.
-These insights are very important as they address the lack of industry uptake - a major issue identified by numerous studies in both industries.
-Other considerations and strategies will be developed to address this issue as well.
+The problem has been addressed in studies beforehand which have generally been limited to prediction of construction costs for buildings/infrastructure and effort estimation for IT projects.
+The prediction models consisted of Case Based Reasoning, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM.
+This leaves a gap in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods.
+
+A comprehensive list of machine learning and statistical models were reviewed in the broader scope of business problem prediction.
+It was determined that Bayesian Networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as Neural Networks while providing insights into model structure.
+Other strategies will be developed to address the problem of industry uptake as well.
 

### 2016-07-14 finalwriteup/trial_code_eq.rmd
finished reviewing lit review

--- 2016-07-14 finalwriteup/draft.rmd
@@
 ---
-title: "draft 1"
-author: "Amy Cook"
-date: "Tuesday, July 07, 2015"
+title: "R Notebook"
 output:
-  word_document:
-    reference_docx: mystyles.docx
-  html_document: default
-bibliography: Library.bib
+  word_document: default
+  html_notebook: default
 ---
@@
 
-# Keywords
-
-# Abstract
-
-* Abstracts allow readers to quickly decide whether the research is relevant to their own interests
-* get at your main argument quickly
-* explain the purpose, goal methods, and results
-
-* why did you do the study?, what is the problem
-performing loss-making jobs for fixed fees is a major problem in the consulting, particularly in the competitive construction industry.
-
-* why is it important?
-affects the porfitability of a company which impacts the careers, internal morale, and even viability of a company.
-
-* why should someone read your essay?
-learn if a company's passively collected data could help them avoid loss making jobs or tactfully enforce strictier contracts.
-
-* what is your main claim?
-improve profitability by 9% by avoiding ~4% of projects is a possible way to interpret the agorithm output. other ways are possible.
-
-* discuss variables and your approaches
-variables: main employee, % time by each type of employee, team size, client charactersitcs, project characteristics (size, type of work, length)
-approaches: complex data with mnay variables, thousands of data points. tried regression as well as machine learning algs: randomf roest and boosted trees, bayesian networks, naive bayes.
-
-* describe evidence to support your claim
-hundred of cross validated trials using the 5 methods, as well as 5 blending methods.
-obtained a 95% confidence interval for each method and simple logistic regression of three models outperformed the others, closely followed by simple averaging.
-the 95% confidence interval across all cross validated trials was __ to __ profit improvement.
-
-* overview of most important sources
-
-* Results: answer , was hypothesis supported
-hypothesis was supprted: CAN improve bottom line of business by mobilising their passively collected data into mL algorhtms.
-
-* implications of your work, meaning
-more work to be done on testing the user interface and user ailities further.
-could assist managers in making better decisions about which projects they should reject or develop a more careful contract structure to improve profits.
-
-# Table of Contents
-
-# List of Figures
-
-# List of Tables
-
-# List of Abbreviations
-
-**ANOVA** - Analysis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
-
-**AUC** - Area under the curve.
-In this project, the area refers to the area under an ROC curve.
-Areas greater than 0.5 indicate predictions performing better than random chance while an area of 1 indicates perfect predictions.
-
-**B2B** - Business to Business. Commerce transactions between two businesses.
-
-**BIM** - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
-
-**CLV** - Customer Lifetime Value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
-
-**CBR** - Case based reasoning. A method of estimating by using the results of similar cases.
-
-**CRM** - Customer Relationship Management.
-Refers to a common type of software used by businesses to record client and project details as well as employee timesheet records.
-Generally accessable to each employee in a company.
-
-**DA** - Discriminant analysis. Statistical analysis to predict a categorical response variable.
-
-**DAG** - Directed acyclic graph.
-A graph consisting of vertices connected by edges, where each edge directs a vertex to another vertex.
-The arrangement of the edges is such that a vertex cannot follow a path that loops back to itself.
-**ESS** - Error sum of squares. The sum of squared errors from a set of predictions.
-
-**FP** - False positive. A case marked 'negative' in a dataset is incorrectly predicted 'positive' by an algorithm.
-
-**FPR** - False positive rate. The proportion of positives that were incorrectly identified (false positives).
-
-**FN** - False negative. A case marked 'positive' in a dataset is incorrectly predicted 'negative' by an algorithm.
-
-**FWLS** - Feature weighted linear stacking.
-A linear combination of model results which are interacted with original variables.
-This allows certain models to be weighted heavier for certain values of the original features.
-
-**IT** - Information technology. The use of computers for storing, retrieving and processing information.
-
-**IQR** - Interquartile range.
-A measurement of spread or variability for a continuous variable.
-When the data is ordered numerically and divided into four equally sized portions, the values that divide the portions are called quartiles (first, second, third and fourth quartiles).
-The first quartile has the highest value of the lowest quarter.
-The interquartile range is the third quartile value minus the first quartile.
-
-**RMSE** - Root mean squared error. The square root of the mean of the ESS.
-
-**ROC** - receiver Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
-
-**TN** - True negative. A case marked 'negative' in a dataset is correctly predicted 'negative' by an algorithm.
-
-**TP** - True positive.
-The case where a case marked 'positive' in a dataset is correctly predicted 'positive' by an algorithm.
-Generally, 'positive' is assigned to categories for which the use case needs to be alerted.
-
-**TPR** - True positive rate. The proportion of positives that were correctly identified (true positives).
-
-# Acknowledgements
-
-# Publications
-
-Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.
-
-# Statement of Original Authorship
-
-The work contained in this thesis has not been previously submitted to meet requirements for an award at this or any other higher education institution.
-To the best of my knowledge and belief, the thesis contains no material previously published or written by another person except where due reference is made.
-
-Signed .................
-Date....................
-
-# Chapter 1: Introduction
-
-Predicting how long it will take a person to do something is a notoriously tricky task.
-People make mental calculations of this sort every day - estimating how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish.
-As most people have experienced, these estimates can be particularly inaccurate if the task is new and difficult.
-These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.
-
-Consulting businesses give expert advice to other professionals in exchange for a fee based on the amount of time their consultanting staff spend on the project.
-The deliverable 'expert advice' is generally in the form of a document or a package of documents.
-A document could range from a 5 page report written by a single person to a sequence of hundreds of drawings assembled by a team of 10 over many years.
-Despite the range of possible deliverables, before any project commences, the client and consultant must agree on a fee, or at least a fee structure.
-The fee structure can be as creative as the engaged parties wish, but in fact tend to fit a few common models.
-A couple of simple examples are:
-
-1. The client agrees to pay for the consultants' time by the hour until the task is complete (referred to as hourly rate) or 
-2. the consultant offers a fixed fee to complete the project in full, regardless of time spent (referred to as fixed price or fixed fee).
-
-In each of these contract models, financial risk is assigned to one of the parties.
-In model 1, the client must pay for an unknown quantity of consultants' billable hours and in model 2 the consultants must complete the project for a fixed price while not knowing how much of their own time will be spent.
-
-This thesis focuses on the risk taken by a consulting company when offering a fixed fee.
-Clients commonly seek fixed price quotes from several consultants before selecting one or negotiating further.
-Quoted fees must be competitive.
-
-The way a consulting manager calculates a fixed price varies from industry to industry and even company to company.
-Typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb (such as a set fraction of the entire client budget for a large project).
-Such qualitative methods can result in profit margins which ultimately depend on the skill (or luck) the manager has in predicting the amount of time a project demands.
-
-This thesis focuses on a single case study consulting company in the construction industry that has persevered through many fixed price projects that resulted in losses.
-A source of insight into the loss-making projects was identified - twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
-Was it possible to find trends in what makes a project loss-making and can these characteristics be identified before a fixed price is offered?
-Can these trends be modelled into a predictive algorithm, and if so, how could the outcomes be integrated into business decisions?
-Finally, if the predictive outcomes were integrated into decision-making, would they deinitely improve the business' bottom line?
-These questions are explored in depth and answered over the course of this dissertation.
-
-Due to the commercially sensitive nature of the data used, the source is not disclosed. Variables have been de-identified to ensure confidentiality is maintained.
-
-# Chapter 2: Problem Description
-
-“It’s tough to make predictions, especially about the future” - Source Unknown
-
-In this chapter the research motivations are explained along with the goals of the project and the specific research hypotheses.
-Finally the contributions of this body of work are outlined along with the structure of the thesis.
-
-## Research Motivation
-
-**Problem Description:**
-
-Project managers across many industries have always struggled with forecasting project costs.
-A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by an average of 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007].
-A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget [@Flyvbjerg2011].
-Another review of 6 IT project surveys between 1984 and 1994 by @Molokken2003 revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
-These statistics describe difficult conditions that can result in job losses, business bankruptcies, and government budget blow outs.
-A survey from @Moores1992 asked software managers whether they saw effort estimation as a problem, and 91% responded 'yes'.
-
-Smaller consulting companies in competitive industries, such as the construction industry, experience similar difficulties in forecasting their project costs.
-They must offer appealing fixed prices in order to remain competitive and win projects.
-However, when these projects go over budget it negatively impacts the business in many ways.
-Employee morale deteriorates and employees may view themselves incompetent if they are responsible for the project's delivery.
-Employees also experience elevated stress attempting to complete underestimated projects within a disappearing budget and may compromise quality of work in exchange for speed.
-This is unfortunate if the only issue with the project was an under-estimated fee.
-Unprofitable projects also limit a business financially, hindering investment into marketing, training, and business development.
-The financial risk taken on fixed price projects may also discourage businesses from taking other, more calculated risks, such as expanding the business into a new area.
-Clearly, improving the prediction of project profitability would improve business growth and staff morale.
-
-**A Wicked Problem**
-
-The challenge for consultants to reduce budget blowouts, in terms of hours spent, is not easily solved.
-Complex projects always differ from one another - even similar projects may end up demanding significantly different amounts of time.
-Although this is known, a fixed price is often demanded by clients before a project is awarded.
-If an hourly rate fee structure is proposed, clients in many industries can simply turn to another consultant willing to take the financial risks involved with fixed prices **(...ref...)**.
-Two strategies to guard against fixed price budget blowouts are detailed below along with their limitations.
-
-A compromise between the hourly rate and fixed price structure is to track employee hours, and stop work once the fixed consulting fee has been exhausted.
-Then, negotiate further fees, or variations, with the client.
-This can either be agreed upon with the client beforehand or the consultant's internal strategy.
-If this strategy is proposed to the client up front, it is likely the client would again turn to another consultant who is willing to offer a fixed price.
-If this approach is the consultant's internal strategy, their financial risk may satisfactorily be reduced.
-However, the behaviour may cause friction between the client and the consultant, therefore reducing the potential for repeat business.
-In the event of hours inflating beyond expectation, consultants may choose to wear losses or endure marginable profits to preserve their client relationship.
-They may reason it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
-
-Another strategy for managing this risk is by stipulating a specific one-off variation opportunity in the contract.
-This applies to consulting work that falls under the client's larger project budget.
-The variation would specify a percentage fee of the overall project cost that can overrule the original fixed price.
-The purpose of this rule is to guard against the case where the project creeps in size, and the consultant finds themselves gradually undertaking more work and more work for the same price even though the client was happy to spend more on the overall project.
-This may account for some or in the best case, all, of their overspent time, although again a savvy client may turn to other consultants that agree to omit variation opportunities.
-
-Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed prices in complex projects.
-In many industries, the reality of fixed price fee structures are unavoidable and businesses conform to survive.
-
-**Current Fixed Price Estimation Methods:**
-
-It has been established that estimating fixed price fees in consulting is a difficult task with unfortunate consequences if the price is estimated too conservatively.
-Thie project aims to mobilise a consultant's historic data describing profitable and loss-making jobs in order to improve their fixed price setting practices.
-But what is the current method for estimating fixed prices or project costs?
-Is past data currently used in other cost estimation practises and to what degree of success?
-
-Unfortunately, limited research is available that documents the methods industry uses for complex-projects.
-However, two surveys are available: one from 2000 that covers construction project cost estimation, the other from 1992 and focuses on IT project effort estimation (which can be likened to fixed price estimation for consultants).
-The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practices.
-They found that the main method was breaking the project into detailed parts and summing up the cost of each item.
-The next two most popular methods were 'comparison with similar projects based on documented facts', and 'comparison with similar projects based on personal experience'.
-These can all be classified as experience based models [@Akintoye2000].
-
-A survey by @Moores1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project.
-In the same survey, 91% of software companies cited cost estimation as a problem.
-
-Both surveys indicated cost and effort estimation were performed via a detailed breakdown of the tasks required.
-Analysis of past project data was alluded to (comparison of similar projects) but a rigorous data analysis process was not common.
-Therefore it is worth investigating if a predictive model based on past data could improve the fixed price setting practice.
-
-**Cost Estimation and Human Nature**
-
-Although the surveys indicated industry did not analyse past project cost data, the idea of using this data (in the same way) has been studied before.
-Research exists that tested people's estimation accuracy with and without comparitive historic data with interesting results.
-This is summarised below along with the resulting data-driven models.
-
-@Lovallo2003 delved into the psychology behind why executives so often severely underestimate costs of larger projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
-Their theory stemmed from Kahneman's work on decision-making that won him the Nobel Prize for economics in 2002.
-His research argued that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
-A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplays or ignores the risk of problems that are out of the team's control.
-This is why it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
-Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].
-
-@Lovallo2003 call the practice of analysing a project based on project details without considering unforseen risks the 'inside view'.
-A survey by @Moores1992 aligned with this theory and found the most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%).
-This contrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details.
-For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate.
-According to @Lovallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project should be ignored in favour of analysing outcomes of several similar projects.
-Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003].
-One method @Lovallo2003 recommended was to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
-The correlation for the current project can then be estimated via a statistical model, which is then used to adjust the forecasted cost made by detailed analysis (the 'inside view') [@Flyvbjerg2011].
-As demonstrated in the surveys, @Flyvbjerg2011 found that managers rarely analysed data from similar projects even though it could significantly improve cost predictions.
-This discovery led the team to formally introduce 'outside view' methods into some institutions.
-
-The implementation of reference class forecasting began in project management for the first time in 2004.
-It was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, and Australia.
-An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].
-An example reading from the chart is as follows: "If the chance of a cost overrun of 20% is acceptable, the estimated cost should be increased by 55% from original calculations.".
-This builds past experiences of cost overruns into the initial cost estimation.
-
-![This model was implemented by the British government for a rail project that is still under completion [@Flyvbjerg2011]](images/lit_review/UKrail.JPG)
-
-The visualisation presents a powerful communication tool for influencing decision makers and improving forecast cost accuracy, however, access to credible data for a sufficient number of projects can be a challenge [@Flyvbjerg2011].
-Overall, the optimism bias theory credibly helps to explain the persistant problem of project cost underestimation.
-
-### Further Reference Class Forecasting Models and Industry Uptake
-
-Although @Flyvbjerg2011's reference class forecasting theory and the psychology behind it was first published in 2003, predicting project costs using previous data has been researched for many preceding years and continues to this day.
-In this section, a brief overview of other reference class forecasting models is presented (they are reviewed in greater detail in the literature review) followed by an analysis of why industry has not yet adopted them.
-
-Most cost estimation or reference forecasting analytical models focused on software development and construction projects, along with some work on film projects.
-These analyses generally used information from 15 to 20 large projects and the dataset summarised projects from around the world with different clients and teams.
-Applied mathematical methods varied from statistical Regressions to deep-learning Neural Networks [@Love2005][@Kim2004].
-Most studies reported that their predictive models performed well and that Neural Networks were generally more accurate than Linear Regression.
-
-#### Industry Uptake of Existing Models
-
-Despite the predictive success in statistical and machine learning cost estimation models as well as the revealing outcomes from 'inside' and 'outside' view studies, reference class forecasting there has been a lack of industry uptake.
-Two reports reviewing cost estimation in the software industry investigated reasons why data-driven forecasting models have not been translated to industry.
-The key causes were:
-
-* A lack of a decision framework to support manager's use of the 'outside view' model [@Moores1992]
-* A failure in marketing of the potential for the models to improve business outcomes [@Moores1992]
-* The products of the research were not comprehensive enough to be used in industry [@Jorgensen2007]
-* Expert judgment was not mobilised enough in the process of developing the tools and the resulting product felt disconnected to actual requirements [@Jorgensen2007].
-
-Reports on other industries were not available, however the insights are general and may explain the lack of adoption in other industries (such as the construction industry).
-
-In the software industry it is also possible the problem of fixed price contract blowouts has been solved by an alternative contract structure which has made cost estimation models unnecessary.
-In contrast to construction, IT projects are a relatively new practice and the industry has had success in trialing alternative contractural arangements.
-Through experimentation, the Agile movement was developed.
-In summary, the method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
-Furthermore, the project outcomes are continually adjusted and revised based on frequent progress review points that assess unexpected problems.
-The flexible upfront arrangement with clients has often been found to forge better long term relationships [Cockburn2001].
-This agile methodology suits the nature of software development work, where curtailing the scope of work is relatively straightforward.
-Also, since the industry is more modern, the new methodology was able to sweep across the field and could be proven to deliver better budget and product outcomes to clients [@Rasmusson2010].
-
-Unfortunately, in other industries such as the construction industry, infrastructure and building projects are steeped in a traditional process, where contractual norms have developed over a much longer time period [@Badenfelt2011].
-The industry is so large, broad and well established that it has inertia in the way projects are estimated, and contractually bound.
-Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high level of legal responsibility and due diligence.
-
-This project aims to assist a company in the construction industry in managing problematic fixed price projects, since fixed price contracts are unavoidable.
-The blocks preventing industry adoption, such as providing a clear decision framework and incorporating expert judgment will be carefully addressed.
-
-### Case Study
-
-The case study company is an engineering consulting company in the construction industry.
-In their profession, labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm.
-In recent years, the size of losses from unprofitable projects equalled 25-30% of their profits from profitable projects.
-This demonstrates the magnitude of their cost estimation problem and the substantial room for improvement that a data-driven model may facilitate.
-
-Cost estimation is currently performed in the case study company by first carefully reviewing preliminary drawings where time costs for each task are estimated and summed by an experienced manager (the 'inside view').
-This is often cross-checked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
-No formal mathematical method of comparison to similar projects is performed.
-This is the case for a number of possible reasons:
-
-* lack of time to interrogate past project data
-* clunky data availability offered by their database software
-* lack of awareness of similar projects performed by other managers
-
-A Customer Relationship Management software package (CRM) is currently employed to collect and store project data.
-A CRM is a popular type of software used by businesses to record client and project details as well as employee timesheet records.
-In the case study business, the CRM is available to all employees over the company intranet, and each employee completes daily timesheets alotting their hours to certain projects.
-Additionally, technical information is recorded against each project as well and the CRM data stood as an untapped source of information within the case study organisation.
-It stores a rich variety of data including:
-
-* employee timesheet hours with dates
-* other project costs (taxis, printing)
-* client information/characteristics
-* client identification code
-* invoiced amounts for each project and dates
-* employee costs
-* employee charge out rates
-* project description
-
-The CRM software readily provides simple output statistics, however, analytical capabilities were limited to:
-
-* simple scatter plots and bar charts of the raw data
-* summaries such as overall hours spent vs. the invoiced amount for each project
-
-The company had a wealth of data but limited means to extract insight.
-This study exploits the CRM data by performing more sophisticated statistical analysis with the intention of building a predictive model to improve cost estimation.
-Thousands of past projects are available for analysis, as opposed to previous studies that used on average 15-20 cases and at most 300 [@Finnie1997][@Pai2013][@Shin2015].
-This improves the potential for accurate predictions.
-If successful, the model also has higher potential for managerial uptake as the algorithm will be directly built and trained on internal company data.
-Managers could interrogate how the model outputs were calculated and relate to the actual projects used in the model or speak to a colleague that was involved.
-This may give the presented 'outside view' a stronger meaning and more potential to influence the final fee proposal.
-
-### Case Study Limitations
-
-The case study provides an opportunity to test whether passively collected CRM data can be mobilised into a predictive model to improve fixed price cost estimation.
-However, there are some limitations in the study including the size of the study, missing data, and errors in data entry.
-This section addresses how the limitations can be addressed either in future work or within the research framework.
-
-The most obvious limitation is that the work focuses on a single company in a specific industry, which limits the application of conclusions to a broader range of companies.
-This can be overcome by first determining the optimum method and variables for this case and then testing them on future case study companies.
-
-Another limitation is the type of unavailable data.
-Information on overall costs of construction projects was scarce.
-This is relevant because managers sometimes price jobs based on the budget for the entire project, of which the consultant is a minor part.
-For example, engineers may moderate their fee by 1% of the entire project cost.
-In the future, it would be beneficial for managers to record the expected total project cost at the beginning of a project versus the concluding completed cost.
-Another piece of data not recorded was the external factors influencing the fixed price quote including known competition from similar firms.
-
-The database also excludes detailed technical information about each project.
-Such details are found in the preliminary drawings describing the project (square meterage, number of storeys), as well as project summary documents provided by the client to enable a quote.
-However, this level of information actually contradicts the purpose of the 'outside view', the intention of which is to present similar projects without being over-influenced by finer details.
-Given the lack of detailed information describing the physical size of projects, it was not reasonable for an algorithm to predict the numeric fee.
-Instead, it was more logical to predict the profitability of a project i.e.
-return per dollar.
-This measure represents the ratio of the amount invoiced to business costs, and can easily be compared between projects of different sizes.
-
-Finally, all CRM data is entered by employees manually.
-Because the data is input by people, it is susceptible to some error.
-Many errors were detectable during data cleaning, however it is possible that a low number of undetectable human errors exist in the data and marginally influence predictions.
-
-### Research Motivation Summary
-
-Fixed price cost estimation for complex projects is a difficult task and when performed incorrectly, cause substantial financial losses to businesses and governments.
-Studies have explored human nature and our tendency to optimistically assess our capabilities with respect to a given task.
-In the case of cost estimation, the optimism can result in negative business consequences affecting employee morale and the capacity of a company to flourish and grow.
-Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project.
-This was done by statistically comparing a new project to a collection of similar projects and their characteristics.
-It has been shown that these models improve cost estimation accuracy, however industry has not adopted them.
-Managers tend to predict costs based on intricate details of a project in favour of reference class forecasting.
-This may be due to poor marketing of the cost estimation tools, a lack of a decision framework in relation to the tools, and a lack of collaboration with industry during model development.
-
-## Thesis aims
-
-For the reasons explained above, the aim of this research is as follows:
-
-**General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
-Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
-The project outcomes are intended to assist the business in predicting project profitability before engagement. Several statistical and machine learning techniques were tested, compared and refined.
-
-**Hypothesis 1**
-A statistical or machine learning model based on historical project data can predict the profitability of a new project.
-
-**Hypothesis 2**
-The predictive model built from Hypothesis 1 can be shown to have a positive impact on the bottom line of the case study business.
-
-## Thesis Contributions
-
-In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the dataset is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
-These points are outlined in more detail below.
-
-In past research, cost estimation datasets generally consisted of dozens of projects from a range of a companys and even countries.
-This thesis case study only had access to the company's internal data but this contained over 2,000 past jobs - far more than the usual cost estimation study and with many more variables.
-Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry which differs from other studies in the construction industry that estimate actual construction costs.
-Consulting in the construction industry can be likened more to software development effort estimation although the nature of the work differs considerably.
-
-In addition to addressing cost estimation with a different kind of dataset, ensemble tree methods are applied which have been minimally tested on this problem.
-To date, Linear Regression, Neural Networks, Case Based Reasoning, and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks in some cases and generally outperform Linear Regression (**ref**).
-Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots that provide insight into the model's calculations.
-This contrasts Neural Networks and SVM's which are 'black-box' predictors.
-Providing insight to a user is a major benefit as past research has demonstrated issues with user trust and uptake of models, even if the models predicted well.
-It is anticipated that using internal data will also improve trust and create tangible meaning for the user.
-Blending multiple machine learning and statistical models has also not yet been applied to the cost estimation problem and may improve predictions.
-
-Finally, this study propels the predictive model one step further than other studies by analysing bottom-line profit improvements that could result from the research.
-A few possibile ways to integrate model results into business decision-making scenarios are presented, and where possible, final profit increases are evaluated.
-This kind of thought experiment is designed to present a clear case for industry when evaluating if and how to adopt cost estimation models.
-
-## Thesis Structure
-
-The next chapter, Chapter 3 is a thorough review of the literature to date that relates to this thesis.
-This covers work solving the cost-estimation problem, studies on statistical and machine-learning models in business applications, and a review of effective statistical and machine learning methods.
-
-Chapter 4 describes the method followed throughout the research process.
-It includes first how the data was obtained, the lengthy cleaning process, followed by variable importance analysis, variable selection, trials of selected predictive algorithms and problem constructs.
-The cost estimation problem was attempted first as a regression problem predicting return per dollar, followed by a simpler binary classification problem predicting profit or loss.
-Once the best methods were selected, they were blended in numerous ways, using both simple averaging techniques and sophisticated machine-learning algorithms.
-These were compared against individual models and the best constructs were selected.
-Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision making scenarios.
-Further applications were also considered.
-
-The research outlined in the method were divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion:
-
-* Chapter 5: Variable Selection
-* Chapter 6: Regression Models
-* Chapter 7: Binary Classification Models
-* Chapter 8: Model blending,
-* Chapter 9: Extended Analysis including overall profit analysis, research limitations and future work
-
-The results section in each of the chapters listed above presents key charts and tables that progressively answer the research question: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
-Then, a discussion finalises each chapter, where the results are examined more deeply and analysed in a broader context.
-The nature of the results are also debated in terms of how surprising they were, their impact on the research hypotheses and limitations as well as suggested future work.
-The final chapter, the Conclusion, summarises the findings of the project, how the research answered the hypotheses and how the work could be viably applied in industry.
-
-# Chapter 3: Literature Review
-
-"If I have seen further than others, it is by standing upon the shoulders of giants." - Isaac Newton
-
-Researching outcomes of comparable past studies may impart how to best approach the problem and reveal shortcomings that can be improved upon.
-This chapter provides an overview of the literature describing early and current cost estimation methods.
-The most prominent machine learning and statistical methods shall then be reviewed followed by a summary of papers on more general business applications that have tried a wider variety of methods.
-Gaps in past research will also be highlighted.
-
-## Project Cost Estimation Methods
-
-The practice of estimating fixed fees is usually represented by the terms 'cost estimation' or 'effort estimation' in the literature.
-The bulk of research to date has been performed with project data in either the construction industry or software development.
-The different methods and outcomes in each of these industries is reviewed below.
-Finally a gap in the literature will be highlighted which this research addresses.
-
-### Cost Estimation in the Construction Industry
-
-Research in the construction industry has primarily focused on predicting the final building cost of construction.
-Some studies reported surveyed cost estimation practices while others tested new methods such as building information modelling (BIM) analyisis, case based reasoning, and predictive algorithms.
-The cost estimation methods of detailed analysis (the 'inside view'), Case Based Reasoning, and predictive algorithms are reviewed in this section.
-
-#### Detailed Analysis in the Construction Industry
-
-Detailed analysis refers to the process of an engineer or builder in the construction team carefully reviewing construction drawings to sum the cost of materials, labour, machinery hire, overheads, and profit etc.
-[@Akintoye2000].
-As referenced in the Research Motivation, @Akintoye2000's study found that this was the most popular method amongst contractors.
-
-@Elfaki2014 discusses how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator.
-This contributes to the lack of accuracy in predicting final costs of a project.
-He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and, in turn, their estimate is prone to subjectivity.
-@Shane2009 theorises that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues, that it is almost impossible to achieve an accurate cost prediction manually.
-
-Groundbreaking methods into calculating construction costs by @Ma2014 mobilise data from building information modeling (BIM) models, although this is still technically a detailed analysis of the costs.
-The idea is to quantify the cost of building a structure directly from a three dimensional BIM model created by the design consultants (engineers and architects).
-@Ma2014's first trial was to automate the cost calculation of a reinforced concrete structure.
-He programmed his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
-From the construction technique and material, a cost was derived for each element.
-This was successful and @Ma2014 aims to further his work so that more construction information can be intelligently obtained from BIM models.
-A system like @Ma2014's would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could still deliver innaccurate estimates similar to detailed analysis by a person.
-The detailed view does not 'step back' and take into account set backs in construction that may have affected past similar projects.
-There is potential for this innovation to be combined with an 'outside view' system.
-
-Detailed analysis remains the most prevalent method for construction cost estimation despite the industry having a long history of projects running over time and budget using the same technique [@Shane2009].
-Next, the method of detailed sums shall be compared to results from parametric algorithms and case based reasoning ('outside view').
-
-#### Statistical and Machine Learning Models in the Construction Industry
-
-Dozens of studies have built machine learning and statistical models to predict construction costs.
-This section will first review the previous studies and most accurate algorithms and finally summarise industry uptake of the methods.
-
-@Elfaki2014 completed a review of cost estimation research from 2004 to 2014.
-They found that artificial Neural Networks and SVM's were the most common machine learning techniques.
-He highlighted that they dealt well with uncertainty, however lack technical justification to the decision maker (known as a black box predictor).
-Neural networks are also time consuming to train, and must be re-trained and re-tested with each additional piece of data [@Kim2004].
-Nevertheless, Neural Networks and SVM's received significant attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
-
-@Shin2015 pioneered the application of Boosted Trees to cost estimation in construction projects.
-This is surprisingly late given the rapid uptake of Boosted Trees over the past decade.
-In their study, data from 234 school building construction projects in Korea were used.
-Boosted trees were compared to Neural Networks and were found to predict costs slightly more accurately than Neural Networks, but not statstically significantly (p value < 0.05).
-
-A notable amount of literature tested the predictive power of Multiple Linear Regression in the construction cost estimation problem.
-Often, Linear Regression was the only model assessed, without comparison to other methods such as Neural Networks, which first started appearing in literature in the 1990's [@Kim2004].
-However, even post-1990, many studies focused on Linear Regression only.
-This may be because the technique is straightforward, easy to use and widely available in statistical packages [@Chan2005].
-Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004][@Attalla2003].
-@Dissanayaka1999 suggested that Regression models could be performed first to determine variable importance and condense the number of variables, then Neural Networks could be used for better estimation.
-Boosted Trees may perform as well as Neural Networks while providing even more insight than Regression models because it is possible to produce partial dependency plots.
-These visualise the relationship between each variable and the response variable (cost).
-
-Regression models from different studies tended to use a wide range of variables, each of which influenced the models to varying degrees.
-This is likely because construction projects are multidisciplinary and involve many parties such as the clients, consultants, contractors and suppliers, [@Chan2005].
-@Chan2005's study on 87 building projects in Singapore included special, complex projects.
-Therefore, variables such as contractor's specialised skills, who the client was (public vs.
-private), the client level of experience, and the contractor's financial management highly influenced final cost of the project.
-Other studies found variables such as project complexity, duration, team experience, information availablility, site requirements, and labour climate to contribute most [@Akintoye2000][@Trost2003][@Pinto1988].
-As these studies obtained information from questionnaires, the available variables depended on the questions.
-Despite the differences in qualitative outcomes, the studies all generally reported their mean absolute error or a similar metric from the tested models, and results were generally positive.
-
-Similar result statistics are useful, however there is a gap in the literature rationalising which model industry should adopt, how to adopt it into a business, and how much loss can be reduced.
-This gap may explain why uptake of these algorithmic models are sparse, as indicated by @Akintoye2000's survey, despite 20 years of prior research.
-@Akintoye2000 hypothesised this could be due to a lack of knowledge about the techniques, doubts whether these techniques are applicable to the construction industry, and the availability of sound data to ensure confidence.
-
-The literature on construction cost estimation does not acknowledge that dozens of consultants and sub-contractors contribute to a large project.
-Most consultants, such as engineers and architects, produce their own cost estimates for their effort that contributes to the project.
-Their contracts are often fixed price, with minimal room for movement, and they wear the risk of over-spending their time [@Harris1999].
-Furthermore, the existing studies collected data from many businesses for a single study, which leaves a gap in research exploring a single contractor's history of projects at one time.
-
-#### Case Based Reasoning in the Construction Industry
-
-Case based reasoning (CBR) is a systematic method of expert judgment, where the decision maker manually compares similar projects from his or her experience [@Shepperd1996].
-The idea is that in the case of a new project, similar projects are chosen, before generalisations about the data are made.
-This contrasts to algorithmic models that are built from the entire set of data [@Elfaki2014].
-Research into the application of case based reasoning to construction cost estimation began in the 1980's [@Kim2004].
-The general procedure involves:
-
-1. Storing a collection of projects with key variable values
-2.
-Once a new project or case arrives, similar cases are retrieved.
-This can be achieved either algorithmically with a distance function applied to the variables, or manually by users reviewing past cases.
-The nearest neighbour algorithm has been used to find the distance between cases in previous studies.
-It is programmed to calculate the Euclidean distance in n-dimensional space between cases, where each variable is a dimension [@Kumar2007].
-3.
-The cost of the new project is estimated by extrapolating characteristics of the similar cases to the new case.
-This can also be done either algorithmically or manually by an experienced decision maker.
-
-A study by @Kim2004 compared CBR to Neural Networks and Linear Regression in construction cost estimation.
-For CBR, the ESTEEM software package algorithmically calculated the similarity of variables, weighted the variables using gradient descent, and deduced the cost using the most similar cases (the specific method for this last step was not explained).
-The discussion compared mean absolute error rates of the three methods and showed that CBR was more accurate than Regression, but less accurate than Neural Networks in estimating cost.
-Despite Neural Networks outperforming CBR, CBR maintained advantages over Neural Networks which have been referenced in other studies as well.
-
-Advantages of CBR include algorithmic efficiency as well as user engagement.
-In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to Neural Networks, which must be slowly re-trained and re-tuned.
-New cases also did not need to have every variable complete which can occur often in real, messy data.
-Neural Networks on the other hand require complete data.
-Another advantage of CBR is the ability for users to review chosen similar cases and make sense of the prediction, as opposed to Neural Networks [@Kim2004][@Elfaki2014][@Kumar2007].
-One disadvantage is that accuracy can be highly dependent on the number of selected cases [@Elfaki2014].
-Overall, CBR may provide prediction accuracies between regression and Neural Networks but the results can be justified by the user - a valuable asset.
-
-CBR that is performed manually through personal experience, without the assistance of algorithms, was the second most popular method for construction estimation in @Akintoye2000's survey in the UK.
-This indicates it is quite an intuitive method that allows decision makers to cross check their detailed cost estimates, although an algorithmic version of CBR has not found success in industry.
-
-#### Summary of Cost Estimation in the Construction Industry
-
-In summary, research into cost estimation in the construction industry has tested a variety of methods ranging from detailed manual analysis, algorithmic models, and CBR.
-Neural Networks generally outperformed other methods, such as Regression and CBR, but had the significant disadvantage of being a black box.
-Alternative methods such as Boosted Trees have the potential to perform as well as Neural Networks and provide insight into the structure of the algorithm.
-This is true for CBR as well, although it has not been shown to predict as accurately as Neural Networks.
-There exists a gap in the research in ways to intelligently combine CBR with machine learning methods that could predict well, provide insight to the structure, and engage the user in reviewing similar projects.
-Also, as stated previously, there exists a gap in the application of cost estimating in the construction industry to smaller contributors to projects.
-These contracting companies face a similar problem in estimating their fixed fee.
-They could mobilise information in their internal project databases, as opposed to the current body of literature that used data from dozens of companies in one study.
-
-### Effort Estimation in the IT Industry
-
-In the software industry, the main component of cost is *effort* as opposed to the cost of building materials in the construction industry.
-A parallel can be drawn between software development effort estimation and consultants' effort estimation in the construction industry - which was highlighted as a gap in the previous section.
-Plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modelling, and Case Based Reasoning.
-
-#### Detailed Analysis in the IT Industry
-
-Similar to the construction industry, expert judgment or detailed analysis, is the most widely practised method for effort estimation [@Shepperd1996][@Molokken2003].
-This is despite many years of research being dedicated to developing algorithmic models that, in the research context, outperform expert judgment.
-A survey by @Moores1992 found that software tools assisting a software project were in significant use, but only the project planning tools, not the cost estimating tools.
-This suggests a mental link for project managers to price jobs based on *details* as opposed to an outside view.
-Another study by @Heemstra1992 found there was no evidence estimation accuracy improved when estimation tools were used.
-@Bergeron1992 similarly found jobs that used algorithmic models were actually associated with less accurate estimates.
-However, this trend may be coincidental due to a lack of cases where estimation tools were used [@Molokken2003].
-The scarcity of real project evidence that algorithmic estimation tools improve estimations and an unintuitive mental jump could be reasons contributing to detailed expert analysis remaining the most widespread technique and mirrors what was found in the construction industry.
-
-#### Algorithmic Methods in the IT Industry
-
-This section assesses the details and performance of the algorithmic models, important variables, and gaps in the literature.
-
-Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although Regression is the most popular method [@Finnie1997][@Pai2013][@Matson1993].
-Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the model's accuracy: size [@Shepperd1996][@Finnie1997][@Pai2013].
-Size can refer to the expected number of lines of code in the software package or function point, which refers to the amount of business functionality expected from the product.
-This is easier to guess correctly at the beginning of a project than lines of code, and therefore leads to more accurate predictions [@Finnie1997].
-One study used a single variable approach (size) with a linear coefficient and an exponential coefficient as follows:
-
-$$
-\begin{aligned}
- effort &= \alpha*size^{\beta} \\
- \alpha &= productivity \ coefficient \\
- \beta &= economies \ of\ scale\ coefficient \\
- size &= estimated \ lines \ of \ code
-\end{aligned}
-$$
-
-This model was compared to CBR, which outperformed the above model [@Shepperd1996].
-
-There are disadvantages to algorithmic methods, similar to the construction industry.
-First, there are often not enough cases to create a good model, particularly if the cases must be from within the company [@Finnie1997][@Pai2013].
-A study by @Mendes2004 demonstrated using 67 web projects that cross-company models were significantly less accurate than a within company model, so it is in the company's interest to create an in-house model.
-This means that effort estimation tools should be refined in-house and adapted by statistical experts to each company [@Shepperd1996].
-
-Algorithmic estimation tools have not been successfully adopted by the IT industry.
-This may be because no model has proved to be outstandingly successful at consistently predicting required effort [@Finnie1997].
-Again, Neural Networks were criticised for their inability to explain their results [@Finnie1997].
-
-in comparison to the construction industry, infrastructure projects do not have a function point variable and need to utilise other information such as expected total project cost (which is notoriously hard to predict based on the previous section), client characteristics, and other project characteristics.
-
-#### Case Based Reasoning in the IT Industry
-
-CBR applications have been researched using similar methodologies to the construction industry.
-In a study by @Shepperd1996, once the similar cases were narrowed down, Linear Regression was employed to predict effort for the new case.
-CBR was found to perform approximately equally to algorithmic methods such as Neural Networks.
-This contrasts results in the construction industry where Neural Networks performed better [@Shepperd1996][@Finnie1997].
-It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects  [@Finnie1997].
-
-#### Summary of Effort Estimation in the IT Industry
-
-Effort estimation in the software industry has followed a similar path to cost prediction of construction projects.
-Expert judgment via detailed analysis is still the predominant method for effort/cost estimation.
-Case based reasoning and algorithmic methods such as Neural Networks have found some success in effort estimation (at similar levels) but have not been successfully adopted by industry.
-It was again noted that CBR and visibility of model structure resonated more with decision makers.
-The software industry is unique in that agile methods of delivery are changing the contractual approach of consulting [@Badenfelt2011].
-This innovative method keeps delivery times short and defines projects as something with a fixed cost and time.
-If a change occurs, other less important elements are discarded to maintain the fixed amount.
-Software development is a relatively new field in comparison to construction, where the tradition of fixed fee structures is difficult to reinvent.
-Also, portions of construction projects cannot be simply curtailed.
-
-There is a need for research on algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
-From findings in previous literature, the models should reveal their predictive structure to decision makers and be based on internal company data.
-Review of the literature revealed a need for an industry focussed decision framework that outlines how to integrate the predictive model in a way that optimises profits.
-
-## Statistical and Machine Learning Methods
-
-### Introduction
-
-As stated, the aim of this project is to use statistical and machine learning techniques to model the profitability of projects.
-Several techniques will be tested and this section reviews a range of predictive algorithms that have been successfully applied to real, messy data.
-These vary from simple methods such as Linear Regression to complex, deep learning Neural Networks.
-The review of methods applied to the construction and IT industry highlighted the use of Linear Regression, Neural Networks, SVM's and in one case Boosted Trees for prediction, however research on other business problems utilised a wider range of methods.
-These include Naive Bayes, Random Forests, and machine learned Bayesian Networks.
-The following section presents each method's advantages and disadvantanges and an example of a successful application.
-
-### Linear Regression
-
-Linear Regression is one of the simplest and most popular statistical prediction methods [@Hastie2009].
-An equation for a line is learned, which is defined by a constant and a coefficient multiplied by the value of each variable.
-The coefficients are chosen to minimise the residual sum of squares (least squared) [@Hastie2009].
-The reason for this method's popularity is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
-The integration of categorical variables is called Analysis of Variance (ANOVA) Linear Regression, and the algorithm compares group mean variances within the categorical variables.
-
-There are several drawbacks to Linear Regression.
-For example assumptions must be made about the structure of the data.
-Predictor variables are assumed to be linearly related to the response variable and the variables to be normally distributed.
-Many real world phenomena do not correspond to these assumptions which can make it difficult, sometimes impossible, to produce useful results [@Hastie2009][@Putler2012].
-Furthermore, response variables must be continuous numeric variables and if two or more predictor variables are correlated, the values of their coefficients are highly unstable.
-On the other hand, correlated variables do not impact the predictive accuracy of the model.
-A final drawback is that the variable coefficients cannot be used to compare variable importance unless the variables have all been standardised to one another (i.e.
-scaled to the same mean and standard deviation) [@Putler2012].
-This limits the amount of insight that can be drawn from a Linear Regression model.
-
-Despite the limitations of Linear Regression models, they have widespread business applications due to their simplicity and ease of application.
-Regression is used to predict risk in the finance and insurance industries, predict who to target in marketing exercises, and predict consumption spending in the field of economics [@Harrell2013].
-Its simplicity and success in other applications make Linear Regression a good starting point for analysis and can serve as a performance benchmark against complex models.
-
-### Logistic Regression
-
-The method behind Linear Regression can be adapted to the prediction of a binary response variable (1 or 0).
-A linear equation is fit to the log of the odds of the binary response variable against each of the explanatory variables.
-The response variable (log odds) in the linear equation can then be translated back to odds by taking the inverse log, i.e.
-the exponential:
-
 $odds = e^{\ln\  odds}$
-
-Finally the odds is converted to probability by the following equation:
-
-$probability = \frac{odds}{1 + odds}$
-
-[@Lowry2016]
-
-This translates the original log odds response variable to a probability S curve bounded by 0 and 1. An example of this translation is graphically shown below:
-
-![Plot of log odds vs.
-an 'x' variable and probability vs.
-the 'x' variable.
-To convert log odds to probability, an exponential transformation is first made followed by the transformation from odds to probability [@Lowry2016].](images/lit_review/logreg.JPG)
-
-The result of this linear fit and logistic translation is a model that predicts the probability (between 0 and 1) that an input case will result in a 0 or 1 (i.e.
-success or failure) [@Moore1989].
-Similar pros and cons exist for Logistic Regression as Linear Regression, and it can be used as a benchmark to compare other binary predictive models for the same reasons.
-Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survival/death [@Moore1989].
-
-### Naive Bayes
-
-The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for the response variable (the response variable must be categorical).
-More specifically, it is assumed that each variable is conditionally independent given the response class, without considering other predictor variables.
-Mathematically this means that each variable's probability contribution can be calculated by looking at the probability of a response class given a single variable's value.
-The probability of each response class can therefore be calculated for each explanatory variable.
-These are then combined for each possible response class using Bayes theorem, and the class with the highest probability is chosen [@Provost2013].
-
-The advantages of this method are that the conditional independence assumption enables very fast calculations and predictions.
-The method can perform well in real world tasks because the assumption of independence does not significantly damage predictions.
-If multiple variables are actually related, the variables separately still direct the prediction in the correct direction.
-The correlated variables will double or triple their emphasis on the predicted response variable, however this does not necessarily impact classification accuracy because the final step is to choose the class with the *highest* probability [@Provost2013].
-This is fine for ranking but the output probabilities are not accurate statistical probabilities [@Caruana2006].
-Another disadvantage is that the distribution of the numeric variable must be assumed (often Gaussian) which the data may not fit into neatly [@Caruana2006].
-
-One real world success story for this classifier hinged on its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
-Naive Bayes was used in a complex spam detection system where new spam emails or toxic text themes could be quickly added to the filtering model.
-Although the example is quite different to cost estimation, the Naive Bayes method generally provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
-
-### Decision Trees/ Ensemble Trees
-
-Decision trees are one of the simplest and most intuitive machine learning methods.
-There are several brands of basic decision tree algorithms including ID3, C4.5, CART, and CHAID with CART and C4.5 being the most popular [@Kabra2011].
-Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
-Single decision trees tend to overfit the data and provide low predictive accuracy but on the other hand create intuitive rules that a decision maker can follow in real-life scenarios.
-In addition, their results do not depend heavily on the skill of the data analyst [@Kabra2011][@Putler2012].
-
-To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
-Three examples of these methods are boostrap aggregating (bagging), boosting, and Random Forests.
-Bagging creates multiple trees by sampling a different training data set (*with* replacement) for each tree and then combining tree results [@Breiman1996].
-Random forests is an advanced form of bagging and builds on the idea of training multiple trees from the same data by sampling bootstrapped training sets with replacement.
-However, when creating each tree, a random subset of attributes (variables) is considered at each split.
-The reduced subset of attributes is resampled for each split in the tree.
-This allows dominant variables to be suppressed for a fraction of the splits, allowing the algorithm to explore signals in weaker variables.
-It also prevents the trees from becoming too correlated [@Breiman2001a].
-The boosted decision tree approach applies gradient descent theory to a series of decision trees.
-The trees are limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) of the preceding tree.
-The limited depth of each tree prevents overfitting at each stage and the combined result of up to thousands of trees is very powerful [@Elith2008].
-The three methods of combining hundreds or thousands of trees turn decision trees into high performing, stable predictors.
-
-Ensemble decision trees have a different foundation to traditional prediction methods such as regression.
-For this reason, there are some advantages and disadvantages.
-Decision trees use a progression of binary splits which means the data is first split by one variable at a certain value into two subsets.
-Then each subset is split by another variable and so on in order to partition the response variable values into groups with similar values.
-The aim of each split is to achieve the most 'pure' split of the response values in a greedy fashion.
-If this process is visualised graphically, it is clear that non-linear relationships can be captured.
-
-![Graphical illustration of non-linear splits derived from a decision tree with two variables [@Jeevan2015].](images/lit_review/dec_tree.JPG)
-
-In contrast to Linear Regression, the predictor variables do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
-Other advantages are that ensemble tree methods are faster than SVM's and Neural Networks but performing just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
-Random forests in particular have the benefit of being unaffected by noise in the data and has the capacity to handle a large number of variables [@Sealfon2012].
-Caruana [@Caruana2006] tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
-This demonstrates how ensemble tree methods are capable of competing with high level machine learning algorithms.
-
-Although ensemble trees have numerous advantages it is understand their limitations.
-Trees are not built on a probabilistic framework, and therefore their results can not be provided in this framework.
-For example, statistical confidence intervals for predictions are not available for standard ensemble methods [@Louppe2014].
-This may or may not be required depending on the nature of the problem.
-Also, variable importance tables provided by Random Forests and Boosted Trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
-The methods can also be prone to overfitting if not tuned carefully [@Louppe2014][@Caruana2006].
-
-### Bayesian Networks
-
-A Bayesian Network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
-The model is visually represented by a directed acyclic graph (DAG) and is capable of linking the conditional dependency between any variable to another variable.
-The links are graphically represented by arcs or edges [@Heckerman1998].
-This kind of conditional relationship is Bayesian, where the probabilities in one node are conditional upon values in nodes directed to it as well as preceding nodes.
-Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.
-
-Bayesian networks are excellent at representing knowledge from different origins, and then visually communicating results well to decision makers [@Kragt2009].
-Therefore Bayesian Networks have become an excellent decision support tool.
-Network relationships can be machine learned, however this is not widely included as part of the suite of machine learning methods and is usually recommended for modelling complex networks.
-Drawbacks of Bayesian Networks include their inability to handle continuous numeric variables (all continuous variables must be discretised).
-This ensures all variable relationships are categorical [@Kragt2009].
-Popular applications of Bayesian Networks are modelling uncertainty in natural resource management and modelling complex business network structures such as airports [@Wu2013].
-
-### SVM's and Neural Networks
-
-Neural networks were inspired by biological Neural Networks of the human nervous system and require a lot of training data relative to other methods.
-They are flexible in the range of relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
-A disadvantage that has been mentioned earlier is that the algorithm is a black box.
-Neural networks can perform high level tasks as a result of deep learning such as hand writing recognition, vehicle control, face recognition, and cancer detection [@Haykin2004].
-
-The simplest type of SVM's calculate the linear boundary between two categories in feature space so that the category of future cases may be determined depending on which side of the boundary they fall.
-The location of the boundary is determined by the position of the widest clear margin that can be drawn between data points from the two categories, while allowing a certain number of misclassifications (this is usually optimised for the problem).
-The simple plot below demonstrates this concept and has no misclassifications:
-
-![Linear boundary with the widest clear margin between data points of different categories [@Provost2013].](images/lit_review/svm.JPG)
-
-The concept can be adapted to nonlinear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
-Advantages of SVM's include the ability to model nonlinear relationships in data, and that no assumptions about the distributions of the data must be made as the model is non-probabilistic.
-They are also known to be robust over different samples and perform well in high-dimensional feature space [@Auria2008].
-On the other hand, the output is in the form of distance to the boundary (as opposed to probability) and results are not transparent (similar to Neural Networks) [@Caruana2006].
-Also, SVM's are known to be very slow to train and therefore often not suitable for industry purposes [@Auria2008].
-Because SVM's perform well in high-dimensional space and can form non-linear rules they have been applied to protein classification (medicine) and text and image recognition [@Byun2002].
-
-### Summary
-
-This section reviewed the advantages and disadvantages of statistical and machine learning techniques that have been successfully applied to business problems.
-These included Linear Regression, Logistic Regression, Naive Bayes, decision trees and ensemble trees, Bayesian Networks, Neural Networks, and SVM's.
-Linear/Logistic Regression and Naive Bayes are computationally very fast with a simple concept that make broad assumptions about the data, but work well as baseline models against which to compare complex models.
-In certain cases these simpler models are significantly outperformed by complex models however it is important to compare both.
-Neural networks and SVM's are excellent machine learning predictors but are slow black box algorithms and are not appropriate for the effort estimation problem.
-
-Ensemble tree methods are capable of performing as well as Neural Networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
-Bayesian networks are excellent for visually presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally been excluded from the machine learning literature in comparison to the other methods.
-This may be due to low awareness of learned Bayesian Networks or lower performance.
-In conclusion, simple models such as linear/Logistic Regression and Naive Bayes are good baseline models to implement, while ensemble tree methods and Bayesian Networks may handle messy 'real world' data better than the simpler models while providing insight to decision makers.
-
-## Review of Statistical and Machine Learning Applications to Business Problems
-
-Literature pertaining closely to the cost prediction topic has been reviewed in the first section of this chapter, however it is relevant to broaden the critique to advanced machine learning methods that have been applied to general business problems.
-This is to find out whether methods that cost estimation had not put to the test succeeded in the broader literature.
-The simpler statistical methods have been covered already, so the focus of this section is the machine learning prediction methods developed for business decisions.
-Popular applied topics include predicting stock fluctuation, customer churn analysis, fraud prediction, customer classification, market segment analysis, product success prediction, and recommendation systems (Seng & Chen, 2010).
-First, the performance of the advanced machine learning methods will be discussed, followed by a case study of employee-churn prediction.
-
-### Summary of Advanced Methods
-
-Ravi Kumar and Ravi [@Kumar2007] performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
-The most widely used model was Neural Networks, however Logistic and Linear Regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
-It was found that DA and Linear Regression techniques were not preferred due to their low accuracy.
-The overall assessment was that SVM's outperformed Neural Networks (back propagation Neural Networks were used most extensively), which sometimes outperformed decision trees, and the rest of the methods were generally inferior.
-Although SVM's performed the best, as discussed, they are often complex and slow, requiring a great deal of memory [@Kumar2007].
-In further support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
-
-When comparing Neural Networks to ensemble trees, Ravi Kumar and Ravi [@Kumar2007] found that Neural Networks and ensemble trees were both capable of out-performing the other, depending on the context.
-
-Their study also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
-For example, combining the contrasting advantages of Neural Networks and decision trees is a worthwhile ensemble technique.
-@Tsai2009's study combined these two methods after trialing each one.
-They were run successively with the Neural Network running first because it had a higher rate of prediction accuracy.
-Then, to resolve the notorious lack of explanatory qualities, decision trees were employed.
-The 81% of cases that were correctly predicted by the Neural Network were used to generate decision trees, and in turn useful decision rules [@Tsai2009].
-This strategy is a promising way to benefit from the strengths of well performing complementary techniques.
-
-### Employee Churn Case Study
-
-Beside effort estimation, predictive business models in the literature generally focus on response variables external to the inner workings of a business such as product popularity, customer behaviour, or stock performance.
-There are far fewer studies that analyse problems affecting internal business decisions such as dealing with employees, teams, and clients.
-An exception is @Saradhi2011's study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
-@Saradhi2011 applied popular customer churn predictive models to employee churn - a novel application that focused internally on employees rather than externally on customers.
-The project tested three machine learning classification algorithms that have been commonly applied to customer churn.
-The associated costs of losing customers and finding new customers can be connected to the costs of losing staff and hiring new staff.
-
-Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
-All three models performed at similar levels for overall accuracy at around 80% (Random Forests had the best results, followed by SVM then Naive Bayes).
-However, when the true positive rates (TPR, accuracy in  predicting employee resignations) were compared, the SVM model far out performed Random Forests and Naive Bayes by achieving 81% TPR vs.
-51% and 55% respectively.
-This was attributed to the ability of SVM's to incorporate class penalties whereas the other methods were limited by the class imbalance problem [@Saradhi2011].
-This means that SVM's had an effective method of increasing the importance of 'positive' cases whicih was advantageous because the data set was skewed towards the majority of employees who kept their job versus the minority who left (25% churn).
-It should be noted that the Boosted Tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVM's.
-Furthermore, Boosted Trees provide insight into the predictive model to the decision makers.
-It is noteworthy that their study proved employee churn models for customers could be translated to predicting internal churn of employees.
-
-Another idea @Saradhi2011 adopted from customer churn analysis was the value models.
-Company's calculate customer lifetime values (CLV's) in order to structure which potentially churning customer's should be acted upon.
-Then the optimal number of customers to reach out to can be determined by optimising the cost of reaching out against preventing financial loss from churn.
-A system was developed for determining the value of each employee in terms of the importance of the projects they were on and their monthly chargeability.
-This allowed them to rank employees identified as 'high risk of churn' by value and provided a clear ranking for manager's to act upon.
-Manager's could then start brainstorming actions to prevent high value, high risk employees from leaving.
-This extension of the study provided a comprehensive framework for how business managers could adopt their findings to improve business operations.
-It was a valuable addition which is absent from most cost and effort estimation research.
-
-### Gap
-
-The limited literature on work similar to @Saradhi2011's Employee Churn study highlights a gap in the application of predictive techniques that use internal data to model internal decisions.
-This is particularly relevant to consulting companies that have complicated internal processes necessary for each project.
-In this research, internal employee team structure as well as client records shall be used to aid prediction.
-The literature on machine learning applied to general business problems revealed SVM's were usually the most accurate technique followed by Neural Networks and ensemble tree methods.
-A gap exists in applying these cutting edge machine learning techniques to the wide range of internal decisions that businesses make.
-
-## Literature Review Conclusion
-
-The literature pertaining to the aim of this research project, cost prediction using internal consulting CRM data, has been extensively reviewed.
-The problem has been addressed in studies beforehand which have generally been limited to prediction of construction costs for buildings/infrastructure and effort estimation for IT projects.
-The prediction models consisted of Case Based Reasoning, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM.
-This leaves a gap in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods.
-
-A comprehensive list of machine learning and statistical models were reviewed in the broader scope of business problem prediction.
-It was determined that Bayesian Networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as Neural Networks while providing insights into model structure.
-Other strategies will be developed to address the problem of industry uptake as well.
-
-# Chapter 4: Method
-
-The aim of this project is to determine whether the profitability of projects for consulting businesses can be predicted using statistical and machine learning techniques.
-As stated in the literature review, a couple of gaps exist in this research space including working with internal CRM data from a consulting company in the construction industry, trialing ensemble tree methods against the more popular regression and Neural Networks, and analysing the impact model results could have on business decisions.
-This section will explain the chosen methodology for fulfilling the research aim using a case study engineering consulting company that charges for their time spent on project delivery.
-
-## Obtaining data
-
-An assessment of the literature revealed that the capacity of internal timesheet data (extracted from a company's CRM software) to predict project profitability has not yet been tested in the context of consulting companies in the construction industry.
-Project data can be extracted from the CRM, interviewing project team members, emails, accounting records, and project drawings and calculations.
-As the internal CRM data in this case study houses a detailed account of time spent on each project (timesheet data) along with client information and invoicing records for the preceding twelve years, it was clear that this source of data was not only the most accessible and suitable for statistical analysis, but had the potential to reveal information not typically investigated by the company.
-
-The complete twelve years of project timesheets and invoices was extracted via the CRM software interface, which described 4169 projects.
-The projects varied from total invoiced amounts of \$500 to over \$1,000,000 and were divided between four internal disciplines in the company.
-During data extraction, it was not clear whether the full twelve years were necessary and relevant to predicting future project profitability but this was determined later in the analysis.
-
-Once the data was extracted from the CRM, an employee from the case study company performed the task of de-identification.
-This is an important step before research commences as the privacy of project employees, clients, and project names must be removed for ethical reasons.
-In the case of employee names, professional titles replaced the names, and client codes replaced client names.
-Project names and addresses were removed, and each project was identified by a number.
-The resulting de-identified datasets were three .csv files of different data structures which are described in the next section.
-
-## Cleaning and variable engineering
-
-The data contained a rich source of project information and a lengthy process of cleaning the data followed by engineering potentially useful variables was then required.
-The three initial datasets described:
-
-* Invoicing records: dated invoiced amounts and sizes for each project and whether the invoice had been paid or written off
-
-    **Example Invoicing Data Structure** - Note data is fabricated
-
-| Job Number | Date Issued | Amount Invoiced | Amount Paid | Invoice Status |
-|------------|-------------|-----------------|-------------|----------------|
-| 1.2.300    | 2014/09/01  | \$5,000         | \$5,000     | Paid           |
-| 1.2.300    | 2014/10/02  | \$2,000         | \$2,000     | Paid           |
-| 1.2.300    | 2014/11/04  | \$3,000         | \$3,000     | Paid           |
-| 1.2.300    | 2014/12/02  | \$4,000         | \$4,000     | Paid           |
-| 1.2.300    | 2015/02/02  | \$5,000         | \$0         | Outstanding    |
-| 1.2.400    | 2014/08/05  | \$20,000        | \$20,000    | Paid           |
-| 1.2.400    | 2014/09/01  | \$2,100         | \$0         | Outstanding    |
-| ...        | ...         | ...             | ...         | ...            |\
-\\
-\
-
-* Timesheet entries: the number of hours spent on each project and on which day.
-Entries were input by individual employees, but names for each entry had been replaced with employee positions (such as mid-level technical)
-
-    **Example Timesheet Data Structure** - Note data is fabricated
-
-| Date    | Hours | Charge Amount | Cost Amount | Created By | Discipline | Job No. | Reconciled | Scope       |
-|---------|-------|---------------|-------------|------------|------------|---------|------------|-------------|
-| 1.2.300 | 4     | 640           | 400         | P12        | Civil      | 1.2.300 | Yes        | Concept     |
-| 1.2.400 | 4.5   | 720           | 450         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.500 | 2     | 480           | 320         | P34        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.600 | 7     | 1120          | 700         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.700 | 1     | 160           | 100         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.800 | 3     | 390           | 240         | P3         | Structural | 1.2.400 | Yes        | Site        |
-| ...     | ...   | ...           | ...         | ...        | ...        |         | ...        | ...         |\
-\\
-\
-
-* Project summary data: information describing each project such as client code, client contact code, discipline, subject, post code, director code, and suburb
-
-    **Example Project Summary Data Structure** - Note data is fabricated
-
-| Job Number | Job Address      | Director | Job Name    | Discipline | Post Code | Client | Project Engineer | Further Job Details.. |
-|------------|------------------|----------|-------------|------------|-----------|--------|------------------|-----------------------|
-| 1.2.300    | 1 Apple St.      | D1       | "Disguised" | Civil      | 4002      | C5030  | P12              | ...                   |
-| 1.2.400    | 33 Rogue Rd.     | D2       | "Disguised" | Structural | 4122      | C2000  | P24              | ...                   |
-| 1.2.500    | 4 Window Dr.     | D3       | "Disguised" | Water      | 4230      | C2045  | P30              | ...                   |
-| 1.2.600    | 864 Hamilton St. | D2       | "Disguised" | Structural | 4004      | C3001  | P1               | ...                   |
-| 1.2.700    | 2 Shark St.      | D3       | "Disguised" | Water      | 4570      | C3020  | P5               | ...                   |
-| 1.2.800    | 24 Campbell Rd.  | D3       | "Disguised" | Water      | 4222      | C4010  | P1               | ...                   |
-| ...        | ...              | ...      | ...         | ...        | ...       | ...    | ...              | ...                   |\
-\\
-\
-
-These three sources needed to be compiled into a single dataset that detailed one project per row as the model was to predict the overall profitability of a project.
-Before compilation could begin however, thorough cleaning was required where the data was plotted and statistically assessed so that outliers could be visually or statistically discovered.
-Outliers were then investigated for data entry errors.
-Many such errors were encountered.
-For example, a common glitch in the data collection process translated a user entering a 123 kilometer drive in a car as 123 hours.
-Or an expense printing claim of $9.50 would be translated to 9.5 hours spent on that job.
-Once these errors were detected, they were discussed with the company directors and appropriate corrective action was taken.
-Mistakes in data entry were often able to be converted into correct data by notes that a user had entered into the system.
-It is possible that some data entry errors were not detected and this is a problem that must be dealt with in implementation of the model as well.
-
-### Variable Engineering
-
-Although dozens of variables were available from the initial dataset, it was advantageous to engineer further descriptive variables to trial in the models.
-This was particularly true for the invoicing dataset and timesheet dataset where a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
-All variables would eventually be tested for variable importance and predictive power with respect to project profitability.
-Examples of these engineered variables include:
-
-**Timesheet Data**
-
-* Percent of hours performed by each professional role over the course of a project
-* Timespan of the entered project hours
-* Percent of hours performed by 'professional' employees as opposed to 'technical' employees
-* Position of the employee who completed the most hours on each project
-* Percent of hours done by the majority contributor to a project
-* Total cost of employee hours per project
-* Total cost of external subcontractors or disbursements per project
-* Total number of users that entered hours on each project
-* Mean number of hours per day entered on a project
-* Number of disciplines active in a project 
-\ 
-\
-
-**Invoicing data**
-
-* Total amount invoiced and remunerated per project
-* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500)
-* Mean invoice per client, i.e.
-the average invoice size across all invoices sent to a client.
-This gives an indication of whether the client generally does fast paced big jobs which would result in large monthly invoices for example or small jobs with small monthly invoiced amounts.
-* Invoice frequency per project. This indicates how drawn out a project was.
-* Client invoice frequency which gives an indication of how much work the company does for that client. 
-\ 
-\
-
-**Project Data**
-
-* text analysis of project descriptions detected a list of key words that could classify projects into 16 categories.
-This key word analysis was done in conjunction with a company employee and the resulting classifications were reviewed by the employee to ensure accuracy.
-* number of projects completed with each client and client contact
-
-With the engineered variables, all three data sources could be represented in the format of a single row per project.
-They were hten combined and further variables were engineered using combinations of variables across the three data sets:
-
-* Profit
-
-    $project\ profit = total\ invoiced\ amount - cost\ from\ employee\ hours$
-
-* Return per dollar as the measure of profitability
-
-    $return\ per\ dollar = \frac{project\ profit}{cost\ from\ hours}$
-
-Besides the engineered variables, additional variables included in the project summary dataset were client industry, internal company discipline, job description, and post code.
-Once numerous potentially important variables were engineered, the most important ones could then be narrowed down.
-This practice improves the accuracy of a model because unnecessary or irrelevant variables add noise to the prediction of target values.
-Variable selection also increases computational efficiency by reducing the number of calculations and enhances understanding of the prediction structure [@Weisberg2005].
-
-## Variable Selection
-
-Once variables were engineered, the value each variable could contribute to the predictive model was assessed.
-As previously stated, variable selection can enhance model accuracy by reducing noise from irrelevant variables and increase simplicity which in turn improves interpretability for the user.
-To assess variable importance, models predicting 'return per dollar' were built using all available explanatory variables.
-The models then output which variables contributed significantly or most improved results.
-Different types of predictive models have different methods of calculating variable importance and since it was not known which model would best predict return per dollar, a few models' variable outputs were reviewed.
-These included ANOVA Linear Regression and two ensemble tree methods.
-Before models assessing variable importance were built, outlier cases were deleted to prevent distortion of valuable explanatory variables.
-
-### Outlier Deletion
-
-Case study projects with extreme 'return per dollar' values were likely the result of special scenarios which the model is not intended to predict.
-For example, many projects appeared wildly profitable but on closer inspection these were all very small jobs.
-After discussion with the case study company, it was concluded that for those outliers, the small jobs most likely required so little time that busy employees did not bother writing down their hours for that job.
-Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles.
-This rule was initially practised by Tukey, the inventor of the box plot [@Tukey1977].
-All outlier projects determined by this method were reviewed with an employee from the case study company and based on their domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined.
-In summary, projects with 'return per dollar' values greater than 3 or below -2 were removed.
-
-### Variable Selection Methods
-
-Three methods which output variable importance were trialed and compared to test whether they highlighted different variables as valuable.
-This was anticipated as each method models data using distinct theoretical foundations.
-The tested methods were ANOVA Linear Regression, conditional inference forests (cForests), and Random Forests.
-It was expected that Random Forests would favour variables that had more categories, a well-known bias; which is also a bias that cForests have overcome [@Strobl2007].
-ANOVA is based on linear theory which contrasts conditional forest and Random Forest's free-form structure.
-Therefore, it is beneficial to compare the methods' assessments of variable importance.
-
-When running variable importance tests, the dependent variable must be clearly defined.
-For this problem, the response variable could be continuous or categorical i.e.
-*how* profitable a project was (continuous) or whether a project was profitable or not (binary classification).
-The continuous response variable was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
-
-#### ANOVA
-
-ANOVA is closely related to linear fit models but incorporates the analysis of differences in group means (i.e.
-categorical variables) [@Lunney1970].
-The method is simple and fast but assumes variables are normally distributed and linearly related.
-The variables must therefore be normalised before analysis to achieve a statistical distribution as close to Gaussian as possible.
-A disadvantage of this method is that linear relationships and normally distributed variables do not necessarily represent real world data accurately [@Breiman2001b].
-
-In order to compare variable importance, two types of output from the model are assessed which include the linear coefficients for each variable and the p-values for the coefficients [@Markham2016].
-The magnitude of a coefficient does not indicate importance relative to the other variables because its value is directly reflects the values within the variable.
-The sign of the coefficient however indicates whether the variable is directly or inversely related to the response variable.
-A p-values for each variable coefficient is output as the result of a significance test for whether the coefficient equals zero.
-Therefore, if the null hypothesis is rejected (p value < 0.05), it indicates that changes in the explanatory variable are related to changes in the response variable.
-P-values are a good indication of whether variables have a relationship with the target variable, however they do not necessarily rank importance.
-If a variable is measured more precisely, it will have a smaller p-value, whereas a variable measured roughly will have a higher p-value.
-This does not necessarily make the more precisely measure variable is more important than the other.
-It is also important to keep in mind that, with a dozen variables, chance alone can produce a variable with a p-value < 0.001 7% of the time [@Rice1989].
-
-The p-value output from ANOVA models gives a good indication of which variables have a statistically significant relationship with the target variable.
-However, the values should not be used to rank the variables against one another, and significant p-values can occur by chance.
-
-#### Random Forest
-
-The Random Forest algorithm can produce a permutation variable importance for each variable as part of its output.
-Variable importance is represented as a score, which is calculated by performing permutations on each variable in the test data for each tree and re-running the test data through the trees.
-More specifically, consider a single variable.
-The values of this variable in the out-of-bag sample (the test data that the Random Forest algorithm partitions before building each tree) are randomly permuted (reordered) for each tree.
-The purpose of permuting is to mimic the absence of that variable.
-The trees in the Random Forest are then run again.
-Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted to the original Random Forest.
-If the accuracy of the permuted forest is much less than the original forest, then that variable was important.
-Permutation importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the Random Forest [@Breiman2005].
-
-The Random Forest variable importance provides valuable insight as
-it addresses the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables [@Strobl2007].
-However, a severe disadvantage of Random Forest variable importance is that it is not reliable for variable selection where explanatory variables vary in their scale of measurement or their count of categories.
-Also, the importance of correlated predictors is overestimated and the algorithm tends to favour variables that have many possible splits or many missing values [@Strobl2007].
-The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably, however this does not necessarily indicate the variable impacts the response variable most.
-Overall, the Random Forest permutation importance is useful as it is able to compare variables in a machine learning environment and can include multivariate interactions, but the current Random Forest importance measure overestimates the importance of variables with many categories and some numeric variables.
-
-#### Cforests
-
-Cforests are an alternative ensemble tree method to Random Forests that overcome the aforementioned shortfalls of Random Forest variable importance measures.
-It is built from cTrees which are decision trees based on a conditional inference framework.
-The key difference between cTrees and standard decision trees is that a significance test is used for splitting instead of a purity measure such as the Gini coefficient, which is centered around information gain.
-
-The procedure behind the significance test at each split in a cTree is as follows.
-To determine which explanatory variable should be used at each split, each variable is permuted in every possible way, and a correlation value is calculated between the tested variable and response variable, for each permutation.
-The unchanged variable correlation is then compared with the correlation values for all permutations of that variable.
-From this, a p-value for the true correlation value compared to the permuted correlation values can be calculated.
-The predictor variable with the lowest p-value is then selected as the splitting variable [@Hothorn2006].
-Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable, re-running the forest and comparing the decrease in response variable accuracy.
-Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) as well as Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].
-
-#### Variable Selection Summary
-
-Limiting the predictive model to a concise set of meaningful variables reduces noise and improves predictions.
-Less variables means that a simpler model is being used for prediction which in turn is easier for stakeholders to understand [@Weisberg2005].
-For these reasons, a subset of the most important variables were chosen before modelling which entailed first eliminating outliers, then comparing important variables from ANOVA as well as Random Forests and cForests.
-It can be inferred from the literature that cForests would suit the case study data best as it is unbiased and not limited to linear theory [@Strobl2007].
-
-## Model selection
-
-Previous studies predicting project profitability in software and construction projects were predominantly limited to case-based reasoning, regression and Neural Networks.
-There is a need to test other sophisticated machine learning algorithms that are as powerful as Neural Networks but provide insight into the reasoning behind predictions.
-Ensemble trees (Boosted Trees and Random Forests) and Bayesian Networks fit these criteria.
-Regression and Naive Bayes models were also included as simple baseline models.
-Complex models should be measured against simple models that can be built at a fraction of the computational cost.
-The complete list of models tested in this study is:
-
-* Regression - baseline model
-* Naive Bayes - baseline model
-* Bayesian network
-* Random forest
-* Gradient Boosted Trees
-
-Some models have limitations and required additional data processing steps such as normalising numeric variables (Bayesian Networks and Regression) and discretising continuous variables (Bayesian Networks).
-Discretising numeric variables was performed by generating a hierarchical dendrogram of each variable to visualise the clusters.
-Between four and six clusters were chosen and summarised to find the maximum and minimum values within each cluster.
-For example, the diagram below illustrates the hierarchical dendrogram for timespan with 6 clusters boxed.
-
-![Hierarchical dendrogram of the Timespan variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
-
-The number of clusters was determined both visually from the dendrogram and experimentally.
-Visually, the height of the dendrogram 'branches' indicates the change in 'tightness' of the data points to their cluster centroids as the number of clusters increases.
-Tightness in the case of Ward's method is the sum of squared distances of each data point to the centroid of its respective cluster (ESS).
-For example, the height of the top horizontal bar indicates the ESS for one cluster and the height of the second from top horizontal bar is the ESS for two clusters.
-The height of the vertical branch between the two bars indicates the improvement in ESS by adding a cluster [@Putler2012].
-As more clusters are added, decreases in ESS are less significant and the aim is to choose the number of clusters at an optimal level in this process.
-After initial comparisons of prediction methods had been trialed, the number of clusters for each discretised variable was re-assessed by trialing one more or one less cluster and comparing model output.
-
-Once the number of clusters had been decided, each case was assigned a cluster label to replace its numeric variable.
-It was decided that for timespans and invoiced amounts, the discretised variables should be applied to all models (not just Bayesian Networks) because managers must guess these values initially when operating the model.
-It is easier for a manager to predict a timespan category than the exact number of days a project will last.
-For example, a small job could be confidently assigned to less than three weeks and a large job could be assigned to 1.5 - 3 years.
-
-The revised discretised categories for timespan and total amount invoiced were as follows:
-
-Timespan Categories
-
-* 1 day - 3 weeks
-* 3 weeks - 2.5 months
-* 2.5 - 9 months
-* 9 months - 1.5 years
-* 1.5 - 3 years
-* more than 3 years
-
-Total Invoiced Amount Categories
-
-* \$100 - \$600
-* \$600 - \$2,500
-* \$2,500 - \$8,000
-* \$8,000 - \$60,000
-* \$60,000 - \$1.8m
-
-### Missing Data Imputation
-
-All methods except gradient Boosted Trees could not handle missing data.
-Therefore, preliminary runs of each method used subsets of the data which had complete data.
-It was possible that if missing data was imputed, predictions from the models that were limited by missing data could improve.
-Also, a complete data set allows for complete sets of predictions from each method, which can then be blended to further improve results.
-First, imputation methods were trialed against gradient Boosted Tree results since Boosted Trees can handle missing values and should perform equally well with imputed data.
-
-The MICE Random Forest method was chosen because it has been proven to work well with complex data sets [@Shah2014].
-It first performs a standard Random Forest imputation of missing values which are treated as initial 'place holders'.
-Random forest imputation is a complex process which works by first doing a rough imputation of the missing values.
-Then a ten-tree forest is run to calculate proximities between all cases.
-Proximities between two cases represent the number of trees in a forest where the two cases were assigned to the same end node in a tree.
-Once proximities are calculated, missing values for a variable (say variable m) are filled by an average of the other complete m values weighted by the proximity of the case representing each m value.
-
-After the initial Random Forest imputation is complete, the MICE process continues.
-The imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable again using Random Forest imputation.
-This is repeated for each variable and their 'place-holder' imputations are replaced by an imputation targeted at the single variable.
-This cycle is repeated five times by default [@Azur2011].
-
-The final imputed dataset was fed into a Boosted Tree algorithm and if similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
-The imputed dataset was then trialed on the remaining methods and compared to un-imputed trials.
-
-## Model Comparison
-
-To compare the models, root mean squared error (RMSE) statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to measure binary classification.
-For binary classification, AUC is a more useful statistic than classification accuracy when output is available as a probability which can be applied to the problem.
-This is pertinent when it is useful to rank the outcomes and trial probability thresholds that will optimise the rates of true positives and true negatives [@Huang2005].
-Note the 'probability' outputs from the models are not true statistical probabilities, but a score the model has assigned to a case using its own measure of certainty.
-
-An ROC graph visualises the curve from which an AUC score is calculated.
-Its two axes are the false positive rate (FPR) on the x-axis and true positive rate (TPR) on the y-axis.
-In this problem a 'positive' is a loss-making job so a true positive (TP) is a case when the model predicts a job will be loss making, and it did indeed lose money.
-A false positive (FP) is when the model predicts a job will be loss making but in reality it was profitable.
-These values are calculated as follows:
-
-$$TPR = \frac{Count of TP}{Count of Positives}$$
-
-$$FPR = \frac{Count of FP}{Count of Negatives}$$
-
-Only predictions made on the test data should be entered into the equations above, therefore after a model must be first created using a training set, then the test data is input.
-From the test data, the model outputs probability values between 0 and 1 for each case, which cannot be entered into the equations above.
-In order to classify each case in the test set as 0 or 1, a probability threshold must be chosen.
-That is, if a probability threshold is chosen to be 0.6, each case in the test set can then be classified as 0 or 1 depending on whether the probability calculated by the model is greater or less than 0.6.
-Once the classifications have been made, the TPR and FPR can be calculated for that model and threshold.
-This is plotted as a single point on the ROC curve, say point C.
-Refer the diagram below:
-
-![Points with TPR's and FPR's as coordinates for different probability thresholds [@Provost2013]](images/method/ROC1.JPG)
-
-A model that is perfectly classified would have all positives correctly classified (1.0 True Positive Rate) and no incorrectly classified positives (0.0 False Positive Rate).
-If a model has a 1.0 TPR and 1.0 FPR (top right corner) this means it is correctly classifying all positives at the expense of incorrectly classifying all negatives as positive.
-It is 'dumbly' classifying all cases as positive.
-If a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previous example.
-There is an 80% chance *any* case will be classified as positive.
-Therefore, models that lie on the diagonal line classify cases as well as random chance.
-
-To plot the curve in an ROC, a series of points and their coordinates are first calculated.
-This is done by choosing numerous threshold probability values between 0 and 1, then applying each threshold point to the probability outcomes in the test set, and calculating the TPR and FPR for that threshold point.
-Once enough points are plotted, a curve can be drawn.
-The closer the curve reaches toward the top left hand corner, the closer the algorithm is to perfectly predicting positive and negative cases at some optimal threshold.
-This would give an AUC of 1.
-An AUC of between 0.5 and 1 means the model is performing better than random chance (as previously stated the diagonal line represents a model randomly assigning positives to all cases at a certain rate).
-
-![An ROC defined by data points calculated from 5 probability thresholds [@Provost2013]](images/method/ROC2.JPG)
-
-In order to compare which models performed significantly better than others, an adequate sample size of results statistics was required.
-Multiple models could be made by using different data in the training vs.
-testing sets, each providing a resulting test statistic (RMSE or AUC).
-5 fold cross validation was used, which created 5 models, then the division of 5 folds was repeated using a different random seed to create another set of 5 training/testing data sets.
-Initially, 20 models of each method were created in this fashion.
-Then a two-sample power calculation was run using the two sets of 20 results to determine the sample size to achieve a statistical power of 0.8.
-Once the number of required samples was determined, more models were built via 5 fold cross validation to obtain additional results statistics.
-The highest performing models were advanced to the next stage where various methods of combining, or blending, methods were tested.
-
-## Model Blending
-
-Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009].
-In this document, the term 'blending' will be used to avoid confusion with ensemble tree methods.
-The idea behind model blending is to combine predictions from several models derived from different methods that have different theoretical foundations.
-In this way, the strengths of each method can be combined.
-Historically, in statistics this was called model averaging and since the 1990's research has shown that averaging results from different methods provides better predictive accuracy than any single model [@Madigan1994].
-New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big, messy data.
-The prize-winning solution was a complex blend of sub-blends which interacted results of individual models with the original variables.
-When original variables are used in blending they are called meta-features [@Sill2009].
-The multiple layers of blending gave incremental improvements in predictive accuracy, but the computational cost of all the complex layers did not justify the benefits for Netflix in practice and a simplified model was adopted.
-Therefore, in this project just a single layer of blending was trialed.
-
-Six blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
-These included simple averaging of the individual model results, building a regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
-The first three models are simply the average, weighted average, of the three individual model predictions.
-The last three methods take advantage of potentially meaningful interactions between an individual model's predictions and meta-features (the original variables).
-For example, if the Boosted Tree model predicted profitability better than the other models for Water projects, the blending model would take advantage of this interaction.
-Boosted tree predictions would be weighted higher than the other model predictions for Water projects.
-
-A simplified explanation of FWLS is as a Linear Regression where meta-features as well as model predictions from individual models are included as explanatory variables.
-Then, each meta-feature is interacted with each set of model results [@Sill2009].
-A Linear Regression is performed to weight each term's contribution to the final predictive accuracy.
-Random forests and Boosted Trees were also used to blend meta-features with output from the best models.
-Feature interaction is performed passively due to the nature of how trees are built.
-A split in a node determined by one variable is conditional upon the preceding split which was based on another variable and so on.
-
-The five blended model types were built on training and testing data sets multiple times in a similar fashion to the individual models.
-A single training set for a blended model must be built from the test results of the individual models.
-Therefore, fivefold cross validation was performed using each individual model to create a full dataset as a compilation of test data results.
-The complete test data predictions were then added to the complete original dataset, and training and testing data could be partitioned to test the blended models.
-A maximum of five blended models were created from a complete set of test results (built from the individual models), then a new set of test results were created for the next five blended models.
-This ensured the blended models were trained on a thorough mix of the complete data set.
-
-20 models of each blending method were initially created.
-Then, a two-sample power calculation was performed to calculate the number of samples required to achieve statistical power of 0.8.
-The specified number of models was then developed so that the highest performing blended model could be tested against the others for statistical significance.
-The next step was to assess the blended models' performance in terms of improving overall profits for the company.
-This analysis differed from accuracy in predicting profitability, so several blended models as well as individual models were carried forward for this analysis.
-
-## Profit Curve
-
-Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the effect on a business' bottom line was tested.
-This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs.
-the probability threshold on the x-axis.
-
-A simple approach was taken for this analysis, where projects with a probability to be a loss-making job greater than the threshold were rejected entirely.
-Therefore all profits and losses from jobs above the threshold were discounted.
-If the threshold was zero, all jobs were rejected and the profit would be $0.
-If the threshold was 1.0, all jobs were accepted and the profit would be the same as the profit the company actually experienced since the data is a sample of historic projects.
-The aim was to find the optimal threshold point where saying 'no' to a job above that level would result in higher profits, because jobs that were likely to make a loss were being rejected.
-This chart will clarified what percentage of profit increase the company could expect by integrating the algorithm into decision making.
-It also provides a clear way to implement the algorithm and promote industry uptake of the research.
-
-Finally, since a profit curve is made from a single training/testing instance of the data, the curve varies with different divisions of the data.
-Therefore, in order to understand the uncertainty around the profit curve and to determine which blended or individual model performed statistically better than others, a large sample size of curve was required.
-Again, 20 curves of each method were built and a power calculation was performed to achieve a power of 0.8.
-This determined the required sample size which was achieved by repeating 5-fold training/testing splits.
-This enabled the highest points of each curve to be statistically compared to one another.
-A 95% confidence interval also be determined around the highest point on each curve.
-The final expected increase in profit and the percentage of projects to be rejected presented a clear scenario that the case study business managers could assess in terms of their business strategy.
-
-<!-- ## Method Conclusion -->
-
-<!-- In summary, to determine whether a  -->
-<!-- * obtained data: 4169 jobs from a single case study company over 12 years. Data was de-dentified by an employee of the company -->
-<!-- * the data was all from the company CRM. There were three chief sources: -->
-<!--     * invoicing data -->
-<!--     * timesheet data -->
-<!--     * project summary data -->
-<!-- * the data was cleaned from mistakes which were generally discovered as outliers  -->
-<!-- * many new variables were engineered from the data, often by summarising time series data into a value that is relevant to a final data set of a single job per row -->
-<!-- * variable selection: dozens of variables available which were condensed to the most important 10-12.
-variable importance methods were ANOVA, random forest, and cforests.
-the overlapping important variables were chosen first as well as some others -->
-<!-- * 5 models were tested: baseiline: log reg, Naive Bayes. -->
-<!-- * sophisticated ML: boosted reg, random frest, Bayesian networks -->
-<!-- * best were comapred to a model averaged version -->
-<!-- * Analysis of impact on bottom line for business if model is integrated into decision making -->
-<!-- * link: results -->
-
-# Chapter 5: Variable Selection
-
-To create the best possible model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
-Initially there were 34 explanatory variables for the single response variable, return per dollar, i.e.
-the chosen measure of profitability for a project.
-By the end of variable selection, only 11 explanatory variables remained.
-As stated in the previous chapter, three predictive methods were selected which output variable importance scores which differ due to their unique theoretical foundations.
-These were Linear Regression, Random Forests, and cForests.
-
-## Linear Regression
-
-Linear Regression must be fed complete datasets but the case study data set only had a core set of 13 complete variables (including the response variable).
-If an incomplete variable was added to the core variables, the 'complete' data set would shrink to the size of the incomplete variable.
-If another incomplete variable was added, the complete set of data would shrink again.
-Therefore, to make the most of the available data, each incomplete variable was added to the core variables one at a time, where a Linear Regression model was built for each.
-That meant a model was made for each incomplete variable, resulting in 21 separate ANOVA models.
-The p-values for the F-statistic of each variable is plotted below.
-The core variables received p-values in all 21 models, while 'add.variable' represents p-values for the additional incomplete variable in each model.
-
-```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of an ANOVA regression's F-statistics used to interpret variable importance")}
-
-plot.pvals = readRDS("AOV_varimp_plot.rds")
-
-ggplot(plot.pvals, aes(x = mod.vars, y = p.val)) + geom_boxplot() +
-        theme(axis.text.x = element_text(angle = 45, hjust = 1),
-              panel.background = element_rect(fill = 'gray95')) +
-        geom_hline(yintercept = 0.05, colour = 'red', linetype = 2) +
-        scale_y_continuous(breaks = c(0, 0.05, .25, .5, .75, 1)) +
-        labs(y = "P-value of F-statistic", x = "Independent Variables",
-             title = "Summary of variable F-statistic P-values from ANOVA models")
-
-```
-
-The ANOVA models indicated that the following variables have a median p-value below 0.05 and therefore significantly contribute to the rejection of the null hypothesis:
-
-* number of employees on the project team (no.users)
-* Percent of hours completed by a professional as opposed to technical employee (pc.pro)
-* Business category of the client (Business) 
-* Internal discipline (Discipline)
-* Number of internal disciplines involved in the project (Num.disc)
-* Timespan of the project from first hours to final hours entered in timesheets (timespan)
-* Total amount invoiced for the project (inv.mlsto)
-
-Four of the added incomplete variables (represented as 'add.variable') had F-statistic p-values below 0.05 in their ANOVA models:
-
-* Client id (code.client)
-* Internal project category (JD.Second)
-* Position of the main professional working on the project, ie mid-level, senior etc. (ProjEng.Pos)
-* Id of the main professional working on the project (code.ProjEng)
-
-The variables listed above had p-values below 0.05, meaning the null hypothesis, that the continuous explanatory variables have no linear relationship with the response variable or that categorical variables have the same mean response values for each group, could be rejected.
-The important variable relationships found in linear ANOVA regression shall be compared to output from Random Forests and cForests below.
-
-## Random Forests
-
-The randomForest algorithm can process datasets with missing values, however it does so by automatically imputing missing values [@randomForest].
-This was not desirable in the early stage of variable selection.
-Instead, the completeness of the variables was assessed and it was decided that all variables with at least 2300 overlapping complete cases would be chosen.
-Additionally, the Random Forest algorithm cannot process categorical variables with more than 53 categories.
-This eliminated more variables and the resulting 'core' dataset contained 15 variables and 2364 cases.
-Because of the Random Forest algorithm's known biases and limitations with numbers of categories, a briefer approach to variable importance analysis was taken in comparison to regression and cForests.
-
-Each tree in a Random Forest is created with a bootstrapped training sample of the data.
-The random sampling means the results of each forest are slightly different.
-To cater for the variation in results, 10 forests were run and the mean rank of each variable was recorded.
-Variable importance output from a single forest is simply a ranked list of the variables where rank 1 indicates the most important variable:
-
-```{r, echo = FALSE}
-
-#dataset created in var_select.rmd
-rf.rank = readRDS("RF_var_rank_subbed.rds")
-kable(rf.rank, format = 'markdown')
-
-```
-
-## Cforest
-
-Cforests were also run with the core 15 complete variables with 2364 complete cases.
-Incomplete variables were added to the core variables individually, each with their own separate run.
-Variables with an unlimited number of categories can be included in the cForest function which is an advantage over randomForests [@Hothorn2006; Strobl2007; Strobl2008].
-Cforests compute a variable importance ranking which is a number relating to the reduction in error the variable provides (similar method to the permutation importance in Random Forests).
-The variable importance rankings for the core 15 variables are as follows:
-
-```{r, echo = FALSE, fig.height = 4.5, fig.cap = c("Variable importance output from a cForest built from 15 core variables")}
-
-cfor.varimp = readRDS("cfor_varimp.rds")
-# how did pc.director sneak in there?
-cfor.varimp = cfor.varimp %>% filter(! var == "pc.director")
-
-ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1),
-              panel.background = element_rect(fill = 'gray95')) +
-        geom_bar(stat='identity') +
-        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance Rankings of Complete Variables") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-Four incomplete variables that were added indpendently ranked in the top 6 most important variables in their model:
-
-* Percent of Hours completed by the main employee on the project (pc.majpos)
-* Client id (code.client)
-* Client contact id (code.contact)
-* Project category (JD.Second)
-
-Once several methods of variable importance analysis had been completed, the separate results were rationalised and final variables were selected for the predictive models.
-
-## Variable Selection Results Summary
-
-Based on the results of the three methods for variable importance, 11 explanatory variables were selected for predictive modelling.
-They are listed in the table below along with the methods that highlighted each variable as important (randomForest = RF, cForest = CF):
-
-| Variable Description                                                 | ANOVA | RF | CF |
-|----------------------------------------------------------------------|-------|----|----|
-| % of Hours Completed by a Professional-level Employee                | x     | x  | x  |
-| Timespan                                                             | x     | x  | x  |
-| Number of Employees on Project                                       | x     | x  | x  |
-| Amount Invoiced for Project                                          | x     | x  | x  |
-| Business Category of Client                                          | x     | x  | x  |
-| Project Category                                                     | x     |    | x  |
-| Total Amount Invoiced from the Client - Past Jobs                    |       | x  |    |
-| Discipline                                                           | x     |    |    |
-| Position of Main Project Employee - new variable from ANOVA findings | x     |    |    |
-| % of Hours by Main Project Employee                                  |       |    | x  |
-| Billing Type (requested by case study company)                       |       |    |    |
-
-## Variable Selection Discussion
-
-The variable importance results from the three types of models: ANOVA regression, randomForest, and cForest showed more similarities than differences. All three models ranked five variables highly:
-
-* percent of hours by a professional
-* project timespan
-* number of employees in project team
-* total amount invoiced
-* business category of client
-
-This was not necessarily expected, particularly between ANOVA and the ensemble tree methods as ANOVA measures linear relationships in contrast to non-linear ensemble trees.
-Random forests favour variables that are numeric or have many categories and this held true for the analysis.
-The 6 highest-ranked variables were numeric with one categorical variable, Business category of client, which had 28 categories.
-The two lowest ranking variables had only 4 categories.
-It was expected that cForests would rank the categorical variables with only 4 categories more fairly.
-This was the case with Discipline and it received a mid-level ranking.
-Business category of the client and Broad business type however, were ranked similarly to randomForests.
-
-The qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, general client characteristics, project timespan, and project category.
-It was expected that Billing Type would play a more important role in predicting profitability as it describes the structure of the incoming fees.
-For this reason, and with advice from the case study business, the Billing Type was kept in the model.
-
-For construction projects and IT software projects, the literature indicated size was the most important variable predicting effort.
-Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996][@Finnie1997][@Pai2013].
-In this case study, profitability is the dependent variable, not total effort and the size variable plays a different role.
-Instead of contributing to predicting the total effort, it is describing whether the size of a project correlates with how profitable the project is.
-The closest estimate of size in this case study is the invoiced amount variable which did rank as very important.
-The exact invoiced amount is difficult to guess before a project begins, and is part of the problem this study is addressing.
-Therefore, the invoiced amount was binned into categories that a manager could much more confidently choose as an estimate.
-The other variables indicated as important in this study  (including the number of employees on a project, client business type and percent of hours by certain types of employees) have not been strongly represented in the literature to date.
-This is because the variables were easily calculated from internal CRM data but would not be easily obtained externally from projects across many companies or regions.
-
-The 11 explanatory variables from table __ were progressed to the next stage where they were used to build predictive models.
-The outcomes of the regression and categorical models are summarised and discussed in the following two chapters.
-
-# Chapter 6: Regression Models
-
-Prediction of 'return per dollar' as a regression problem was attempted first followed by prediction of profit or loss as a binary classification problem.
-Linear Regression, Random Forests and Boosted Tree methods were applied to both problems with varying success.
-
-## ANOVA
-
-In ANOVA, categorical and continuous explanatory variables are treated differently.
-For categorical variables, differences in the mean values of the response variable in each group are tested.
-In contrast, linear relationships are examined between continuous explanatory variables and the response variable.
-Since linear relationships are being tested, continuous variables are assumed to be normally distributed.
-Therefore, where appropriate, numeric variables were normalised via logarithmic or cube root transformations.
-Trials of various variable interactions revealed that interactions between:
-
-* total amount invoiced * percent of job by professional level employee
-* total amount invoiced * client id
-* project category * number of employees on the job
-
-were statistically significant. Two types of ANOVA models were run. The first used only a 'core' list of 6 variables that were largely complete and any relevant interaction terms.
-
-* Total amount invoiced 
-* Discipline 
-* Timespan
-* Team size
-* % of hours completed by a professional level employee
-* Client business category
-
-The second method was developed to overcome the problem of missing data.
-When all variables were combined, only 15% of the project cases had complete data and therefore the ANOVA test would eliminate 85% of the data upon execution.
-Imputing the missing data was not desirable at that stage due to the volume of missing categorical data and an inability to ensure the imputed data was not hampering results.
-The second method was a more complicated procedure that required dozens of models but allowed variables with missing data to be included.
-First, a project was randomly selected (a row in the dataset) which would have certain variables complete.
-The data was then filtered for all projects that had at least the same columns complete, and an ANOVA model was built on this reduced data set.
-This method allowed variables with a larger proportion of missing data to be included while not attempting to analyse all variables at once.
-
-```{r insert_data}
-# load function for determining number of required samples for 0.8 power
-n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
-        if(type.calc == "one.sample"){
-              d.calc = abs(mean(data))/sd(data)
-              pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
-                                   power =.8, type = c(type.calc), 
-                                   alternative = alt)
-              return(ceiling(pow_pow$n))  
-        }
-
-        if(type.calc == 'two.sample'){
-                d.calc = abs(mean(data[[1]])-mean(data[[2]]))/
-                        (sqrt((sd(data[[1]])^2+sd(data[[2]])^2)/2))
-                pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05,
-                                   power =.8, type = c(type.calc), 
-                                   alternative = 'two.sided')
-                return(ceiling(pow_pow$n)) 
-        }
-
-}
-
-results.core = readRDS("reg_lm_core.rds")
-# n.samples(data = results.core$diff)
-
-results.A2 = readRDS("reg_lm.rds")
-results.A2 <- dplyr::filter(results.A2, diff>-15)
-# n.samples(data = results.A2$diff)
-
-```
-
-The results of the ANOVA models were assessed using the root mean squared error (RMSE) statistic of the return per dollar predictions.
-These were compared against a baseline RMSE which was the RMSE of the predicted return per dollar values against the mean return per dollar of all projects.
-Then, the RMSE from each model was subtracted from the RMSE using mean return per dollar.
-If the models were effective, the RMSE should be lower for ANOVA model predictions, so the difference would be greater than 0.
-Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only (method 1 using ANOVA).
-Note, only `r n.samples(data = results.A2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
-
-```{r, fig.height = 4.3, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
-If the ANOVA model is effective, the difference should be greater than 0.")}
-
-#plot histogram to show differences
-results.core %>% ggplot(., aes(x =diff, y = ..density..))+ 
-        geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models \nCore Variables Only") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-A student t-test indicates the difference is significantly above 0 with a p-value of `r t.test(results.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
-Therefore, the null hypothesis that the difference is less than or equal to zero can be rejected.
-In other words, the ANOVA model improved estimates of return per dollar over using the mean return per dollar to a statistically significant degree.
-However, the mean difference is only \$`r mean(results.core$diff) %>% format(digits = 1)`.
-Furthermore, the mean RMSE in return per dollar is \$`r mean(results.core$final.RMSE) %>% format(digits = 2)` which is unacceptably high.
-Almost 50 cents is a large mean error when the value is estimating the profit from \$1 spent.
-
-Results from 100 models using method 2, which captures incomplete variables, are shown below.
-
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
-If the ANOVA model is effective, the difference should be greater than 0.")}
-
-#plot histogram to show differences
-results.A2 %>% ggplot(., aes(x =diff, y = ..density..))+ 
-        geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models \nAll Variables Sampled") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-A student t-test indicates the difference is not significantly above 0 which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
-This method clearly performed poorly in comparison to the core variables.
-A reason for this could be that the mean number of cases in the dataset was `r results.core$test.nrow[1]*4` for the core variables but averaged only `r (mean(results.A2$test.nrow)*4) %>% round(0)` for the sampled set of all variables.
-Regression results from ANOVA models were poor - the randomForest algorithm was trialed next.
-
-## Random Forests
-
-```{r load_rf_results}
-results.rf.core = readRDS("reg_rf_core.rds")
-# n.samples(data = results.rf.core$diff)
-
-results.rf2 = readRDS("reg_rf.rds")
-# n.samples(data = results.rf2$diff)
-
-```
-
-Numeric predictions of return per dollar were attempted first with core variables that were largely complete followed by a method of sampling a mix of core and less complete variables in every run (similar to the two ANOVA methods).
-The randomForest parameters were first tuned using the caret package and optimal values mtry = 5 (the number of randomly selected explanatory variables to consider at each split) and ntree = 500 (the number of decision trees to ensemble) were determined [@caret].
-A density plot of 50 models of the core variables using different training/test sets is shown below.
-The plot was created using the same procedure used in the ANOVA models where the x-axis measures the RMSE obtained by using the randomForest model subtracted from the RMSE using the mean of all project return per dollar values.
-Note, only `r n.samples(data = results.rf2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
-
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
-If the ANOVA model is effective, the difference should be greater than 0.")}
-
-#plot histogram to show differences
-results.rf.core %>% ggplot(., aes(x =diff, y = ..density..))+ 
-        geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nCore Variables Only") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-This plot shows a mean return per dollar RMSE improvement of \$`r mean(results.core$diff) %>% format(digits = 1)` which is the same as the ANOVA model.
-The p-value for the null hypothesis that the difference is less than or equal to zero is `r t.test(results.rf.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)` and could be rejected.
-Results from 100 randomForest models using method 2 are shown below:
-
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
-If the ANOVA model is effective, the difference should be greater than 0.")}
-
-#plot histogram to show differences
-results.rf2 %>% ggplot(., aes(x =diff, y = ..density..))+ 
-        geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nAll Variables Sampled") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-The p-value for these models and the afore-mentioned null hypothesis is `r t.test(results.rf2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
-The null hypothesis was rejected which contrasted the matching ANOVA test.
-On the right hand tail of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar on the left tail.
-A density plot of the RMSE for the models is shown below:
-
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
-
-#plot histogram of RMSE's
-results.rf2 %>% ggplot(., aes(x =final.RMSE, y = ..density..))+ 
-        geom_histogram() + geom_density() +
-        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models \nVariables Randomly Sampled") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-Although the randomForest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for return per dollar was wide and the mean RMSE was high (\$`r mean(results.rf2$final.RMSE) %>% round(2)`).
-
-The two regression models predicting 'return per dollar' resulted in high mean RMSE's ranging from \$`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
-The next section discusses possible reasons for the poor results and the decision to halt progression of regression models.
-
-## Regression Models: Discussion
-
-Both ANOVA and randomForest models produced a similar lowest RMSE value, but performed differently when directly comparing models built on the same list of variables.
-Both methods predicted at best within an average of \$0.48 for the response variable, 'return per dollar', using either core variables or a sample of core and less complete variables.
-Random Forest predictictions were much more accurate than ANOVA's with a mix of core (almost complete) and less variables while randomForest and ANOVA performed similarly with the core variables only.
-There was much less data available for training the models that used less-complete variables so it was interesting that randomForest performed considerably better (RMSE of \$0.50 for randomForest vs.
-\$2 for ANOVA).
-The main difference between ensemble tree methods and ANOVA is that ANOVA assumes linear relationships between variables (some of which are normalised) whereas trees are capable of capturing non-linear relationships.
-It is possible the less complete variables had non-linear relationships with the response variable.
-Random Forest is also an ensemble method while ANOVA is not.
-Each randomForest is the average of 500 single trees, whereas each ANOVA model is a single model only.
-50 randomForest models produce 500 times more models than 50 ANOVA models so the averaging effect may have eliminated outliers for better overall performance.
-
-If the best RMSE achievable at that point was \$0.48, what does that imply in terms of the business decision?
-The magnitude of the RMSE can be evaluated from the end-user's perspective (the manager who is proposing fees for new projects).
-To give a feel for typical project performance, refer the plot below of historical 'return per dollar' values:
-
-```{r, fig.cap = c("Distribution of 'return per dollar' for all projects")}
-
-if(.Platform$OS.type == 'windows'){
-        all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv' )
-} else{
-        all8a<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all8a.csv')
-}
-ggplot(all8a, aes(x = return.pdol, y = ..density..)) + 
-        geom_histogram() + geom_density() +
-        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-The 95% confidence interval for this data is \$`r (mean(all8a$return.pdol)- qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)` to \$`r (mean(all8a$return.pdol) + qnorm(0.975)*sd(all8a$return.pdol)) %>% round(2)`.
-If a project returns \$0 for every dollar spent (return per dollar = \$0), the project has earned only enough to cover the cost of doing the project (salary costs and business costs).
-If the return per dollar is -\$0.50, the project has earned enough to cover half of the costs of doing the job.
-Finally, if the return per dollar is \$1, the project earned double what it cost to perform.
-Using this reasoning, if model predictions have an average error of \$.50 for return per dollar, this represents an error in revenue of half what the project cost to complete.
-Unfortunately, this is clearly not informative enough to bring into the business decision-making process.
-
-Further regression methods could have been pursued, however randomForest is a high-performing algorithm and was far from providing usable results.
-It is likely that either not enough data was available or that the given variables did not capably explain all the factors affecting return per dollar.
-Each row of data is a project, which can take the company from a few weeks to sometimes years to complete.
-Therefore, more data cannot be easily accessed or created.
-It was decided categorical predictions should be pursued next as predicting whether a job will be profitable or not is a simpler task and could more effectively leverage the limited data.
-
-This conclusion differs to case studies in the literature which tend to predict total cost/effort as a regression problem [@Kim2004][@Chan2005][@Akintoye2000].
-This is likely because the literature dealt with substantial projects which are rarer and impact governments and large companies.
-A simple profit or loss prediction would not be useful.
-The case study company on the other hand may find a profit/loss binary prediction useful because they complete hundreds of small projects a year.
-If binary classification is successful, applications of the predictions must be explored along with contributions to the company decisions and profits.
-
-In summary, because of the poor results when predicting return per dollar, further development of regression models was deferred in favour of binary classification models.
-Binary classification is a simpler task than regression because it requires less precision.
-The challenge is how to gainfully apply the results to the business scenario.
-
-# Chapter 7: Binary Classification Models
-
-The cost estimation problem analysed in the literature review had been converted to a profitability estimation problem and was finally converted to a binary profit/loss prediction.
-The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable project (or not).
-The original 'return per dollar' response variable was transformed by converting all 'return per dollar' values less than or equal to zero to 'loss making' jobs and 'return per dollar' values greater than 1 to 'profitable' jobs.
-The following section summarises results from the trialed methods.
-
-## Results from Five Methods
-
-The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were initially developed without imputing data, using the maximum amount of data possible depending on the method.
-Boosted Trees and Naive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
-Logistic Regression, Random Forest and Bayesian Networks require datasets without missing values so a subset of complete data was used.
-It was expected that Boosted Trees would perform best as they are known to have powerful predictive capabilities and can use incomplete data.
-For this reason, 4 x 5-fold runs of each method were performed (20 models each) and statistically compared against Boosted Trees' results.
-A power calculation was then used to calculate the number of runs required to achieve a statistical power of 0.8.
-
-The models were compared using the AUC test statistic calculated from the ROC as this gives a good indication of model performance under binary classification [@pROC][@Putler2012].
-An example plot of five ROC's built from 5-fold cross validation using the Boosted Tree algorithm is shown below.
-
-Before the plot could be made, the Boosted Tree algorithm parameters were tuned to optimise the AUC outcomes. The caret package was used to tune the following parameters [@caret]:
-
-* shrinkage = 0.001
-* number of trees = 10,000
-* minimum terminal node size = 20
-* maximum tree depth = 5
-
-<!-- image below generated from file rf_log_naiive_auc.rmd -->
-<!-- lines 683 to 689 -->
-
-![ROC's for five boosted tree models built on different train/test partitions of the dataset [@pROC]](images/boosted tree/5_fold_ROC.png)
-
-In the plot above, the average AUC was 0.763 which indicated that altogether, taking into account a range of possible thresholds, the model was predicting profit/loss better than random chance (AUC = 0.5) but not perfectly (AUC = 1).
-20 models and their resulting AUC statistics from each method were initially determined.
-These were used to calculate how many models were required to achieve a minimum statistical power of 0.8 when compared to the Boosted Tree results.
-The comparison of logistic regression required 100 models, while randomForest required over 700 because the AUC distributions were close.
-Bayesian Networks and Naive Bayes did not perform as well and the power calculation revealed only 5 and 8 models were required respectively.
-Bayesian Networks were then dropped as Naive Bayes performed slightly better and were faster and simple enough to work as a baseline comparison to the higher-performing methods.
-
-It was finally decided that randomForest and Boosted Trees would likely not be distinguishable since the power calculation indicated that 700 models were required.
-Hence, 100 models of each method were built as calculated for logistic regression.
-10-fold cross validation was performed 10 times.
-The violin plot below summarises the AUC values produced by each method.
-The 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
-
-```{r reduced_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
-Subsets of the data were used for Logistic Regression and randomForests in order to provide datasets without missing values.")}
-
-reduc.totAUC <- readRDS("reduc_df_auc.rds")
-
-#significant difference??
-index = c(1,3,4)
-pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
-for(i in seq_along(index)){
-        pvals$p_value[i] = t.test(reduc.totAUC[[2]], reduc.totAUC[[index[i]]])$p.value %>%
-                format(digits = 2) %>% as.numeric
-}
-
-signif = pvals %>% filter(p_value < 0.05) %>% .$method
-reduc.totAUC = melt(reduc.totAUC)
-reduc.totAUC$sigdiff = ifelse(reduc.totAUC$L1 %in% signif, TRUE, FALSE)
-
-ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
-        scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
-        scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method") +
-        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
-                         "log" = "Logistic Regression",
-                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
-        theme(legend.position = "bottom")
-
-```
-
-The AUC performance of Logistic Regression and the randomForest algorithms cannot be statistically differentiated from Boosted Trees.
-
-Next, data imputation methods were trialed which would make the complete data set available to Logistic Regression and randomForests.
-It was possible this would improve AUC results and the same process as above could be repeated.
-Imputation was done using the MICE method with randomForest imputation [@mice].
-Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for Random Forest unless over 450,000 models were made.
-
-```{r mice_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
-Each method was fed the same imputed full dataset.")}
-
-mice.totAUC <- readRDS("mice_df_auc.rds")
-
-#significant difference??
-index = c(1,3,4)
-pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
-for(i in seq_along(index)){
-        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
-                format(digits = 2) %>% as.numeric
-}
-
-signif = pvals %>% filter(p_value < 0.05) %>% .$method
-mice.totAUC = melt(mice.totAUC)
-mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)
-
-ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
-        scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
-        scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method - Imputed Data") +
-        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
-                         "log" = "Logistic Regression",
-                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
-        theme(legend.position = "bottom")
-```
-
-Boosted Trees, Logistic Regression, and randomForest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed Boosted Trees.
-The results from each method using reduced data sets (to cater for missing values) alongside results using imputed complete data are shown below.
-
-```{r all_violin, fig.height = 4.6, fig.cap = c("Combined violin plots comparing AUC results for each method predicting profit/loss using subsets of data vs. the full imputed data set.")}
-
-reduc.totAUC$Data_Source = "Original"
-mice.totAUC$Data_Source = "Imputed"
-combine.micer <- rbind(reduc.totAUC, mice.totAUC)
-
-combine.micer$Data_Source <- factor(combine.micer$Data_Source, 
-                                    levels=c("Original", "Imputed"))
-
-ggplot(combine.micer, aes(x= factor(L1), y= value)) +
-        geom_violin(aes(fill = Data_Source, colour = Data_Source)) +
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-
-        # stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
-        scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Data Source") +
-        scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Data Source") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "Comparison of Original Data vs. Imputed Data") +
-        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
-                         "log" = "Logistic Regression",
-                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
-        theme(legend.position = "bottom")
-
-```
-
-## Binary Classification: Discussion
-
-Testing Boosted Trees, Random Forests, Naive Bayes, Logistic Regression and Bayesian Networks on the binary classification problem showed that Boosted Trees, Logistic Regression and Random Forests performed best (according to AUC).
-Bayesian Networks are not particularly popular and require discretised variables so it was expected that this algorithm would not perform as well as highly regarded methods such as Boosted Trees and Random Forests.
-Naive Bayes and Logistic Regression were included as baseline models and it was expected that the more complex models would outperform these.
-Therefore it was surprising that results from 100 Logistic Regression models were not statistically lower than Boosted Trees.
-
-A possible explanation for this, according to the literature, is that there was not enough data for the ensemble trees to learn the complex decision rules at which they excel.
-Trees tend to overfit the patterns in a smaller training set.
-Logistic regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
-This may explain Logistic Regression's comparitively high performance on the case study's smaller dataset with many variables.
-
-Boosted trees, Random Forests and Logistic Regression all had mean AUC values around 0.775.
-The tests were repeated using imputed data and again the means sat around 0.775.
-The variables were not particularly well suited for imputing so this result is logical and could have even diminished the mean AUC.
-For example, many of the imputed variables were categorical such as job category, percent of job completed by a professional level employee, and the position of the majority worker on a job.
-The benefit of having an imputed dataset that performs acceptably well is that each model is fed the full data set and therefore produces predictions for the full data.
-Beforehand, randomForest and Logistic Regression models were limited to a complete set of original data which implied a reduced data set.
-With complete model predictions, the high performing models could be blended to test whether better predictions were possible.
-For this reason, the imputed data models were propagated forward to the development of blended models.
-
-In summary, the individual binary classification models performed well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a new job will be profitable or loss-making.
-Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model actually affects decisions in practice.
-Results from model blending and overall profit impact are presented and discussed in the next chapter.
-
-# Chapter 8: Model Blending
-
-## Blended Models
-
-Blended models combine the predictions from each higher-performing individual model to create averaged or 'blended' predictions.
-In this case, the randomForest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialed ranging from simple to complex.
-The simplest way averaged the probability output (a number between 0 and 1) of all three models.
-The next two simplest methods consisted of building a Logistic Regression and Boosted Trees model from the three individual model outputs only.
-Feature Weighted Linear Stacking (FWLS), randomForest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (meta-features).
-The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature conditions could be weighted as such.
-Some meta-features may have been dispensable in the blending process, so in order to maintain simple comprehensible models, a variable importance study was conducted again.
-
-### Variable Selection: Blended Model
-
-Variable assessments were completed and compared using cForests, Logistic Regression, randomForests and Boosted Trees.
-Cforests have the favourable characteristic of being unbiased while the remaining methods were used for the blending process.
-The results are summarised in the sections below.
-
-#### Variable Importance: cForest
-
-```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a cForest blended model. The results from the three best performing models were added as explanatory variables.")}
-
-cfor.varimp = readRDS("cfor_blend_varimp.rds")
-
-ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1),
-              panel.background = element_rect(fill = 'gray95')) + 
-        geom_bar(stat='identity') + 
-        labs(title = "Cforest Variable Importance: Blended Model",
-             x = "Variable", y = "Variable Importance Score")
-
-```
-
-The cForest highlighted the results from the Logistic Regression and Boosted Trees as the most important variables.
-This was followed by the results from the randomForest model and the project team size, project category, total invoiced amount (category) and timespan (category).
-
-#### Variable Importance: Logistic Regression
-
-```{r}
-
-log.varimp = readRDS("logblend.rds")
-
-log.varimp = summary(log.varimp)$coefficients %>% as.data.frame
-colnames(log.varimp)[names(log.varimp) %in% 'Pr(>|z|)']<-'P_value'
-log.varimp = add_rownames(log.varimp)
-log.varimp = log.varimp %>% filter(P_value <= 0.05) %>% 
-        select(rowname, Estimate, P_value) %>% arrange(P_value)
-
-# order levels
-log.varimp$rowname <- factor(log.varimp$rowname,
-                               levels= log.varimp$rowname)
-
-#plot p values
-# ggplot(data=log.varimp, aes(x=rowname, y=P_value)) +
-#         theme(axis.text.x=element_text(angle=45,hjust=1)) + 
-#         geom_bar(stat='identity') +
-#         labs(title = "LogReg variable importance: blend")
-
-```
-
-An ANOVA regression found the F-statistics from the following variables were significant in rejecting the null hypothesis for p-value < 0.10:
-
-Interaction term between:
-
-* Timespan * results of logistic regression model
-* Total amount invoiced * results of logistic regression model
-* Total amount invoiced * results of randomForest model
-* Client Business type * results of Boosted Tree model 
-\ 
-\
-
-Single variables that were not also in interaction terms:
-
-* Total amount invoiced for the job
-* Timespan
-* Team size
-* Client business category
-* Project category
-* Results from each individual model 
-\ 
-\
-
-#### Variable Importance: randomForest
-
-```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a randomForest blended model. The results from the three best performing models were added as explanatory variables.")}
-
-rf.varimp = readRDS("rfblend.rds")
-rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
-rf.varimp = add_rownames(rf.varimp)
-rf.varimp = arrange(rf.varimp, desc(MeanDecreaseAccuracy))
-# order levels
-rf.varimp$rowname <- factor(rf.varimp$rowname,
-                               levels= rf.varimp$rowname)
-
-#plot p values
-ggplot(data=rf.varimp, aes(x=rowname, y=MeanDecreaseAccuracy)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1),
-              panel.background = element_rect(fill = 'gray95')) +
-        geom_bar(stat='identity') +
-        labs(title = "randomForest Variable Importance: Blended Model")
-
-```
-
-Permutation importance by randomForest rated all three prediction models as most important followed by team size for the project, percent of hours done by the chief employee on the project, timespan (category), project category, and client business type.
-
-#### Variable Importance: Boosted Trees
-
-```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a Boosted Tree blended model. The results from the three best performing models were added as explanatory variables")}
-
-boost.varimp = readRDS("boostblend.rds")
-boost.varimp = add_rownames(boost.varimp)
-# order levels
-boost.varimp$rowname <- factor(boost.varimp$rowname,
-                               levels= boost.varimp$rowname)
-
-#plot p values
-ggplot(data=boost.varimp, aes(x=rowname, y=rel.inf)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1),
-              panel.background = element_rect(fill = 'gray95')) +
-        geom_bar(stat='identity') +
-        labs(title = "Boosted Trees Variable Importance: Blended Model")
-
-```
-Boosted Trees ranked output from all three individual prediction models highest (Logistic Regression ranked highest) followed by Project category, Client Business type, and timespan (category).
-
-#### Variable Importance Summary
-
-Upon reviewing important variables ranked by cForest and each method of model blending, the following variables were deemed important by at least two methods and were used in the development of blended models:
-
-* All three results of prediction models, particularly logistic regression
-* Project category
-* Client business type
-* Total amount invoiced for a project (category)
-* Team size on a project
-* Timespan
-
-Random Forests and Boosted Trees passively interact explanatory variables, but for logistic regression, these must be manually input.
-Only interaction terms with statistically significant F-statistics from the ANOVA test were included in order to reduce noise.
-
-### Comparison of Blended Models
-
-Only three of the model blending techniques used the 8 variables determined above.
-These were FWLS, randomForest, and Boosted Trees.
-The remaining 3 techniques used only the predictions of the three individual models.
-All six methods were compared against the original logistic regression model.
-Initially 100 train/test runs were performed for each blending method by running 20 iterations of 5-fold cross validation.
-The AUC results of these iterations are displayed below:
-
-```{r blend_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
-100 models were built for each method.")}
-
-auc.list <- readRDS("blend_auc2.rds")
-means = lapply(auc.list, mean) %>% unlist
-greater = means[means> means["orig.log"]] %>% names
-auc.melt = melt(auc.list)
-auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)
-
-ggplot(auc.melt, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = greater, colour = greater)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 Blended Models of each Method") +
-        scale_x_discrete(labels = c("boost" = "BT Blend", 
-                         "log" = "FWLS", 
-                         "average" = "Simple Average Blend",
-                         "orig.log" = "Orignal LR", 
-                         "rf" = "RF Blend",
-                         "simp.boost" = "Simple BT Blend",
-                         "simp.log" = "Simple LR Blend")) +
-        scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC greater than Original Logistic Regression") +
-        scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC greater than Original Logistic Regression") +
-        theme(legend.position = "bottom")
-
-```
-
-The above plot shows that 4 methods had a higher mean AUC than the original logistic regression model:
-
-* the simple blended logistic regression 
-* the simple average 
-* and the two Boosted Tree models.
-
-To achieve a power of 0.8, a two-sample power analysis was performed on the 100 AUC results from each blended model against the AUC results from the original logistic regression.
-The boosted models required over 1000 samples while the simple averaged and simple logistic regression models required only 120.
-Therefore, 50 more iterations of the simple averaged and simple logistic models were performed.
-The Boosted Tree models would not achieve a statistical power of 0.8 with only 150 samples when compared to the original logistic regression and were therefore dropped from the next stage of analysis.
-
-```{r}
-#load results from 150 iterations
-auc.150 = readRDS("blend_auc150.rds")
-
-```
-
-The results of the 150 iterations are shown in the plot below.
-Simple model averaging resulted in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logisitc regression blended model had a two-sample t-test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
-
-```{r, fig.height = 5.5, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
-150 models were built for each method to achieve a statistical power of 0.8.")}
-
-index = c(2:3)
-pvals = data.frame("method" = names(auc.150)[c(2:3)], 'p_value' = NA)
-for(i in seq_along(index)){
-        pvals$p_value[i] = t.test(auc.150[[1]], auc.150[[index[i]]])$p.value %>%
-                format(digits = 2) %>% as.numeric }
-
-signif = pvals %>% filter(p_value<0.05) %>% .$method
-auc.150 = melt(auc.150)
-auc.150$signif = ifelse(auc.150$L1 %in% signif, TRUE, FALSE)
-
-ggplot(auc.150, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = signif, colour = signif)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y = "mean", geom="point", colour = "gray60", shape =3) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 150 Blended Models of Best Methods") +
-        scale_x_discrete(labels = c( 
-                         "average" = "Simple Average Blend",
-                         "orig.log" = "Orignal LR",
-                         "simp.log" = "Simple LR Blend")) +
-        scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC significantly greater than Original Logistic Regression") +
-        scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC significantly greater than Original Logistic Regression") +
-        theme(legend.position = "bottom")
-
-```
-
-The simple Logistic Regression blend and simple averaging of the individual models (logistic regression, Boosted Trees and Random Forests) gave virtually the same distribution of results, where both AUC distributions were signficantly greater than the original logistic regression model.
-The p-values from the two-sample t-tests were `r pvals$p_value[1]` and `r pvals$p_value[2]` for simple average and simple Logistic Regression methods respectively.
-It was encouraging that blended models produced higher AUC values than the original logistic regression model, however for the case study business, the real measure of success is the model's impact on the business' overall profits.
-It was not guaranteed higher AUC values directly related to improvements in profit so several methods, including original individual models as well as blended methods, were progressed to the profit analysis.
-
-## Blended Models Discussion
-
-Trials of different methods of blending model results demonstrated that again, the simplest methods worked best with the data.
-In this case, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
-It was plausible that simple blending models would perform better than complex methods because the data was not big enough for complex models to learn intricate patterns.
-This was observed in the single method models (previous chapter) when Logistic Regression performed as well as Random Forests and Boosted Trees.
-The literature does not extensively cover blended machine learning methods however their success in the 2009 Netflix competition generated some publications.
-The Netflix data set comprised of almost 3,000,000 observations, so the size of the data could have enabled the success of complex blending methods such as FWLS [@Sill2009].
-That being the case, it was still surprising that simple averaging of probabilities from the individual models performed almost as well as blending results using Logistic Regression.
-It was expected that Logistic Regression would have an advantage over model averaging due to its ability to weight the prediction variables while maintaining a simple model structure.
-
-The Logistic Regression blending model did indeed weight predictions from the original models differently.
-A summary of the resulting coefficients from an example blended Logistic Regression model is shown in the table below.
-
-| Model Results as Variable | Coefficient (Log Odds) | Odds  | Increase in Variable Input | Increases in Odds |
-|---------------------------|------------------------|-------|----------------------------|-------------------|
-| Logistic Regression       | 1.7223                 | 5.60  | 0.1                        | 56%               |
-| Random Forest             | 1.4753                 | 4.37  | 0.1                        | 43.7%             |
-| Boosted Trees             | 2.7943                 | 16.35 | 0.1                        | 163.5%            |
-
-As shown in the table, if the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model prediciting a loss-making job increased by 56.0%.
-The weightings contrasted the other simple model that averaged the original probabilities (i.e.
-weighted them equally) but still managed to produce similar AUC results.
-It was not clear why this was the case, however both simple models were compared again in the analysis of the impact each model could have on the company's overall profits.
-The analysis of overall profits did not directly relate to AUC measurements, so a range of models were carried forward for comparison.
-These included the three original individual models, three simple blended models and three complex blended models.
-
-# Chapter 9: Extended Analysis
-
-## Profit Curve Analysis
-
-The first research question, whether machine learning and statistical techniques can predict project profitability based on historic data, has been answered.
-The measure of profitability could not be modelled but simplifying the profitability prediction to a binary (profit/loss) model was successful.
-To answer the second research question, whether the model would improve the business' profitability, a profit analysis was performed.
-
-The models output a 'probability' between 0 and 1 that each project will be a loss-making job (probability = 1 indicates a loss-making job).
-This leaves the question, at what probability would a decision-maker round the probability up to 1 or down to 0?
-And what business decision would then be made?
-To find the threshold point for rounding, an experimental business-scenario was tested.
-At a given threshold, all projects calculated above the threshold were considered too risky, and were rejected.
-All profits and losses from these projects were forfeited, while the profits from the remaining jobs (below the threshold) were summed to give a revised total profit.
-
-```{r power_profit_curve}
-
-# How to achieve statistical power of 0.8?
-finance.plot = readRDS("plot_finance.rds" )
-fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
-        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
-                            sd_prof_ratio = round(sd(.), 1)),
-                       profit.ratio)
-
-#for each method find max mean 
-best.fin.1 = fin.summary.1 %>% group_by(method) %>% 
-        filter(mean_profit_ratio == max(mean_profit_ratio)) %>% 
-        arrange(desc(mean_profit_ratio))
-
-#two sample power test between each method and its % data points at its threshold
-finance.plot$threshold <- as.factor(finance.plot$threshold)
-power.fin = finance.plot %>% filter(threshold == '0.55' & method == "average"|
-                                    threshold == '0.6' & method == "orig.boost"|
-                                    threshold == '0.75' & method == "orig.log"|
-                                    threshold == '0.55' & method == "orig.rf"|
-                                    threshold == '0.6' & method == "simp.log")
-
-# test each sample against "average" because it's the best
-# n.samples function is defined in 'draft.rmd'
-index = c("orig.boost","orig.log", "orig.rf","simp.log")
-ns = NULL
-for(i in seq_along(index)){
-        temp = n.samples(data = list(power.fin %>% filter(method == index[i]) %>% .$profit.ratio, power.fin %>% filter(method == 'average') %>% .$profit.ratio), type.calc = 'two.sample', alt = 'two.sided')
-        ns[i] = temp
-}
-# ns
-
-# orig.boost needs 6, orig.log needs 4, orig.rf needs 9 and simp.log needs 64 : ) do 100!
-
-```
-
-Using this logic, if the threshold was 0, all projects were rejected and the total profit would equal \$0.
-If the threshold was 1, all projects were accepted and the total profits would be the same as historical figures.
-This profit calculation was made on a range of thresholds between 0 and 1 at 0.05 increments.
-The total profits were plotted for each threshold and joined to make a profit curve where at an optimal threshold, the total profits were highest (refer plot below).
-
-Profit curves for the three simple blended models as well as the three individual models were calculated.
-However, the total profit coordinates were based on probability outputs in test data sets.
-This means that the profit curve for the simple logistic regression blend (for example) would be different for different training/testing data divisions.
-I.e.
-the total profits at each point in the profit curve depend on the model parameters that were calculated using the training data and which projects were analysed in the test data.
-Therefore, the process of training a blended model, then calculating the profit curve from the test data was repeated multiple times to create a distribution of profit curves for each method (blended and individual).
-
-To start, 10 curves for each method were calculated (2 x 5-fold cross validation).
-Two-sampled power calculations (for a power of 0.8) were performed between the results of each method at their optimal threshold against the best performing method (simple average blend).
-For example, the optimal threshold for the simple average blended model was 0.55 while the optimal threshold for the original Boosted Tree model was 0.60.
-The 10 'total profit' values at these two thresholds were compared in the two-sample power calculation to determine the number of samples required to achieve a power of 0.8.
-The maximum required number of samples was 64 so 100 models were of each method were built (20 x 5 fold cross validation).
-
-The plot below illustrates the results of each model at each threshold with a line joining the mean value at each threshold for each method.
-The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold for the 100 models.
-
-```{r profit_curve, fig.height = 6, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
-
-finance.plot = readRDS("plot_finance_100_4.rds" )
-finance.plot$method <- factor(finance.plot$method, 
-                                    levels=c("average", "simp.log","simp.boost",
-                                             "comp.boost","comp.rf", "FWLS","orig.log",
-                                             "orig.rf","orig.boost"))
-
-ggplot(finance.plot, aes(x= threshold, y= profit.ratio, colour = method)) + 
-        # geom_point(alpha = 0.005, shape = 16, size = 1) + 
-        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
-        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) +
-        stat_summary(fun.y = mean, 
-                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
-                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
-                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
-        stat_summary(fun.y = mean, geom="line") +
-        scale_y_continuous(breaks = seq(0,110, by = 20)) + 
-        scale_x_continuous(breaks = seq(0,1, by = .2)) +  
-        theme(legend.position = "none",
-              panel.background = element_rect(fill = 'gray95')) +
-        labs(x = "Probability Threshold", 
-             y = "Ratio of Model Profit to Historical Profit (%)",
-             title = "Profit Curve for Blended and Individual Models") + 
-        facet_wrap(~method, 
-                   labeller = labeller( method = c("average" = "Simple Average Blend",
-                                       "orig.boost"= "Orig. Boosted Trees",
-                                       "orig.log" = "Orig. Logistic Regression",
-                                       "orig.rf" = "Orig. Random Forest",
-                                       "simp.log" = "Simple Log. Reg. Blend",
-                                       "simp.boost" = "Simple Boosted Tree Blend",
-                                       "comp.boost"= "Complex Boosted Tree Blend",
-                                       "comp.rf"= "Complex randomForest Blend",
-                                       "FWLS" = "FWLS"
-                                       )))
-
-```
-
-Two blended models clearly outperformed the individual models as shown in the plot above with higher profit ratios as well as tighter confidence intervals.
-The table below denotes each models' optimal thresholds and the corresponding increase in total profits:
-
-```{r profit_table}
-
-finance.plot$method = as.factor(finance.plot$method)
-# want mean and sd for each method at each threshold
-
-fin.summary = finance.plot %>% group_by(threshold, method) %>% 
-        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
-                            sd_prof_ratio = round(sd(.), 2)),
-                       profit.ratio)
-#for each method find max mean
-
-best.fin = fin.summary %>% group_by(method) %>% 
-        filter(mean_profit_ratio == max(mean_profit_ratio)) %>% 
-        arrange(desc(mean_profit_ratio), sd_prof_ratio) %>% 
-        slice(1) %>% ungroup() %>% 
-        arrange(desc(mean_profit_ratio), sd_prof_ratio)
-
-best.fin <- best.fin %>% mutate(lower_bound_95CI = (mean_profit_ratio - 1.96*sd_prof_ratio)%>% round(1))
-best.fin <- best.fin %>% mutate(upper_bound_95CI = (mean_profit_ratio + 1.96*sd_prof_ratio)%>% round(1))
-
-kable(best.fin)
-# saveRDS(best.fin, "best_fin.rds")
-
-best.mean = best.fin %>% filter(method == "simp.log") %>% .$mean_profit_ratio
-best.sd = best.fin %>% filter(method == "simp.log") %>% .$sd_prof_ratio
-best.t = best.fin %>% filter(method == "simp.log") %>% .$threshold
-
-```
-
-The simple logistic regression blend performed best with the highest mean profit ratio at `r best.mean`% with a standard deviation of `r best.sd`%.
-This means that for the simple blended logistic regression model, if all jobs above the probability threshold `r best.t` were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profits.
-
-For clarity, the plot for the simple blended logistic regression model is shown below:
-
-```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression Blended Method")}
-
-ggplot(finance.plot %>% filter(method == "simp.log"), 
-       aes(x= threshold, y= profit.ratio, colour = method)) + 
-        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
-        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
-        geom_point(alpha = 0.1, shape = 16) +
-        stat_summary(fun.y = mean, 
-                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
-                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
-                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
-        stat_summary(fun.y = mean, geom="line") +
-        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
-        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
-        theme(legend.position = "none", 
-              panel.background = element_rect(fill = 'grey95'),
-              panel.grid.minor = element_blank()) +
-        labs(x = "Probability Threshold", 
-             y = "Ratio of Model Profit to Historical Profit (%)",
-             title = "Profit Curve for Simple Logistic Regression Blended Model")
-
-```
-
-```{r}
-# Results - over 5 pages, up to 20 pages
-# 
-# Guidelines:
-# * Outlines what you found out in relation to your research questions or hypotheses, presented in figures and in written text.
-# * Results contain the facts of your research.
-Often you will include a brief comment on the significance of key results, with the expectation that more generalised comments about results will be made in the Discussion section.
-# 
-# Which results to discuss
-# * restate research questions and hypotheses. can we predict profitability? ofocurse you can make a prediction, 
-# real question might be is it worthwhile for the business to engage in this activity?
-we will look at this as a bottom line analysis.
-acknowledge that it may be worthwhile for other reasons such as confidence and transparency in historic projects.
-# 
-# * bottom line analysis - run 1000 times to get confidence interval around the line.
-#     * should the business adopt the model?
-#     
-# * to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the case based reasoning approach
-```
-
-## Profit Analysis: Discussion
-
-```{r}
-simp.l <- readRDS("finance_average_model.rds")
-
-```
-
-### Highest Profit Curves
-
-The simple average blend and logistic regression blended models outperformed the complex blended models, which is rational because their AUC values were significantly higher than the others.
-The simple average of Logistic Regression, Random Forests, and Boosted Trees results produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression that blended the same three variables.
-The blended Simple Logistic Regression produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`.
-To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r comma_format() ((best.mean-100)/100*1000000)` by rejecting all jobs with a probability higher than `r best.t`.
-In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.
-
-The plot below illustrates this proportion of projects as well as the proportion of profits and losses assigned to a range of probability levels by the simple logistic regression blend.
-The red line sits at its optimal threshold point, i.e.
-projects greater than the threshold point would be rejected:
-
-```{r, fig.height = 5.5, fig.cap = c("Distribution of projects according to the probability outputs from a typical simple Logistic Regression blended model.
-A bargraph of counts and relative profits are shown.")}
-
-reject.plot <- readRDS('reject_plot.rds')
-
-profs <- reject.plot  %>% .$balance.mlsto %>% abs %>% sum
-simp.l2 <- reject.plot %>% filter(compare == 'full') %>% select(simp.log, balance.mlsto, inv.mlsto.log)
-# simp.l2 <- mutate(simp.l2, invoiced.ratio = exp(inv.mlsto.log)/sum(exp(inv.mlsto.log))) 
-simp.l2$profit.ratio <- ifelse(simp.l2$balance.mlsto>0, simp.l2$balance.mlsto/profs,
-                              simp.l2$balance.mlsto/profs)
-simp.l2 <- rbind(simp.l2,simp.l2)
-simp.l2$faceter <- c(rep("hist", 2364), rep("profs",2364))
-simp.l2$prob.bin <- cut(simp.l2$simp.log, seq(0,1,by = 0.05), labels = seq(.05,1,by = 0.05))
-
-ggplot(simp.l2, aes(x = prob.bin)) + 
-        facet_grid(faceter~., scales = "free_y",
-                   labeller = labeller(faceter = c("hist" = "Count of Projects",
-                                       "profs"= "Relative Profits/Losses"
-                                       ))) +
-        geom_bar(data = simp.l2 %>% filter(faceter == "hist")) + 
-        geom_bar(data = simp.l2 %>% filter(faceter == "profs"),
-                 aes(x = prob.bin, y = profit.ratio), stat = "identity") +
-        # geom_bar(data = simp.l2 %>% filter(faceter == "inv"),
-                 # aes(x = prob.bin, y = invoiced.ratio), stat = "identity") +
-        labs(x = "Probability Score", 
-             y = "",
-             title = "Distribution of Project Counts and Project Profits Against Probability") +
-        geom_vline(xintercept = 10.5, colour = hcl(h = 15, l = 65, c = 100), linetype = 1,
-                   alpha = 0.8) +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-### Comparison of All Profit Curves
-
-The tested profit curve models can be divided into three categories: original individual methods, simple blends of the individual methods, and complex blends of the individual methods that employed interactions with original features.
-All original methods did achieve improvements of profit at their optimal thresholds, but did not necessarily consistently improve profits across all cross-folded models.
-The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
-That is, it cannot be said with 95% confidence that the original methods' would produce a profit higher than historical profits if projects higher than the optimal thresholds (jobs that are likely to be loss-making) were rejected.
-Clearly this is not helpful for the case-study business to adopt.
-
-The complex blended models had mixed results.
-FWLS and randomForests performed more poorly than the individual methods while Boosted Trees produced a marginally higher profitability ratio (`r best.fin %>% filter(method == "comp.boost") %>% .$mean_profit_ratio`) but the lower bound of its confidence interval was `r best.fin %>% filter(method == "comp.boost") %>% .$lower_bound_95CI`, i.e.
-not above original profits.
-
-The simple blended models performed more favourably, as already discussed, particularly the simple average and Logistic Regression.
-The Boosted Tree's lower bound confidence interval was `r best.fin %>% filter(method == "simp.boost") %>% .$lower_bound_95CI`% at its optimal threshold point.
-This is not above 100%, and again does not make a convincing argument for the case-study to adopt it.
-
-It is not clear why simple linear models and the averaging model outperformed ensemble tree blending methods.
-It may be that there was not enough data for the ensemble trees to adequately learn the complex series of rules they develop as previously stated.
-It was also surprising FWLS was ineffective given the success it had in the Netflix Prize competition [@Sill2009].
-The only differences between FWLS and the simple Logistic Regression (that performed best) were two additional explanatory variables and four interaction terms.
-The number of variables was not high for the amount of data.
-According to @Peduzzi1996, 10 events per explanatory variable or more avoids the risk of biased estimation of variable coefficients in logistic regression.
-The data contained 315 events per explanatory variable and did not pose that risk.
-Nevertheless, the additional variables must have added misleading noise to the model.
-The performance of the models so far has been concerned with AUC and profit, however insight into the differences in model performance may be gained by reviewing confusion matrix statistics such as TPR and FPR.
-
-### Comparison of Confusion Matrix Statistics
-
-To investigate method performance further, confusion matrix statistics were scrutinised.
-For each method, 100 full data sets of results were generated via 5-fold cross validation.
-Each data set of probability results differed to the next because of the random division into the five-fold data sets used for training and testing.
-Nine full data sets of results was chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
-These were deemed a 'typical' set of predictions for each method.
-Then, the confusion matrix for each set of results at their respective thresholds could be calculated.
-The plot below shows the breakdown of predictions by each method on actual loss-making projects.
-If the project was predicted correctly as loss-making, this was a true positive, and the job was rejected (in the fabricated scenario).
-If it was predicted incorrectly as a profit-making job, this was a false negative, and the project was completed, resulting in a loss for the business.
-
-```{r TP_plot, fig.height = 5, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
-
-conf.plot <- readRDS("conf_plot.rds")
-
-#if model.pred = loss, this is TP
-conf.plot$model.pred <- ifelse(conf.plot$actual == "loss" & 
-                                       conf.plot$model.pred == "loss", "TP", "FN")
-# reorder levels in method to suit TP 
-conf.plot$method <- factor(conf.plot$method, 
-                           levels= conf.plot %>% filter(actual == "loss", model.pred == "TP") %>%
-                                   arrange(desc(pc)) %>% .$method)
-
-ggplot(conf.plot %>% filter(actual == "loss"), 
-       aes(x = method, y = pc, fill = model.pred)) +
-        geom_bar(stat = "identity", position = "stack", alpha = 0.7) +
-        scale_fill_manual(values = c('red2','dodgerblue2'),
-                          labels = c("FN: predicted as profit", "TP: predicted as loss"),
-                          name = "Confusion Matrix Reference") +
-        labs(x = "Method (Original or Blended)", y = "Percent of All Loss-Making Projects (%)",
-             title = "Breakdown of True Positives and False Negatives for Loss-Making Jobs") +
-        scale_x_discrete(labels = c("average" = "Simple Av",
-                                      "orig.boost"= "Orig. BT",
-                                      "orig.log" = "Orig. LR",
-                                      "orig.rf" = "Orig. RF",
-                                      "simp.log" = "Simple LR",
-                                      "simp.boost" = "Simple BT",
-                                      "comp.boost"= "Complex BT",
-                                      "comp.rf"= "Complex RF",
-                                      "FWLS" = "FWLS")) +
-        theme(axis.text.x=element_text(angle=30, hjust=1),
-              legend.position = "bottom")
-
-```
-
-Choosing the right probability threshold point is about finding the right balance point in an imperfect model.
-In this case, projects above a threshold point will include profitable and loss-making jobs.
-The optimal point is where the savings from rejecting loss-making jobs (TP) outweigh the lost profits from rejecting profitable jobs (FP) the most.
-All models were capable of different TP and FN ratios at alternate threshold points.
-
-The plot above shows the Complex Boosted Tree Blend has the highest TPR followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
-All of these TPRs are below 25% which shows it was not worth catching a large percentage of loss-making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
-This implies the models were not able to very cleanly divide loss-making jobs from profitable jobs.
-
-A summary of predicted results on actual profitable jobs is shown below at each method's optimal threshold point.
-If a job was predicted to be profitable, and was therefore accepted, it was classified TN.
-If a job was loss-making but labelled profitable, the business would complete the job and make a loss (FN).
-
-```{r TN_plot, fig.height = 5, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
-
-conf.plot <- readRDS("conf_plot.rds")
-
-#if model.pred = loss, this is TP
-conf.plot$model.pred <- ifelse(conf.plot$actual == "profit" & 
-                                       conf.plot$model.pred == "profit", "TN", "FP")
-# reorder levels in method to suit TP 
-conf.plot$method <- factor(conf.plot$method, 
-                           levels= conf.plot %>% filter(actual == "profit", model.pred == "TN") %>%
-                                   arrange(desc(pc)) %>% .$method)
-
-conf.plot$model.pred <- factor(conf.plot$model.pred, 
-                           levels= c("TN","FP"))
-
-ggplot(conf.plot %>% filter(actual == "profit"), 
-       aes(x = method, y = pc, fill = model.pred)) +
-        geom_bar(stat = "identity", position = "stack", alpha = 0.7) +
-        scale_fill_manual(values = c('dodgerblue2','red2'),
-                          labels = c("TN: predicted as profit", "FP: predicted as loss"),
-                          name = "Confusion Matrix Reference") +
-        labs(x = "Method (Original or Blended)", y = "Percent of All Profitable Projects (%)",
-             title = "Breakdown of True Negatives and False Positives for Profitable Jobs") +
-        scale_x_discrete(labels = c("average" = "Simple Av",
-                                      "orig.boost"= "Orig. BT",
-                                      "orig.log" = "Orig. LR",
-                                      "orig.rf" = "Orig. RF",
-                                      "simp.log" = "Simple LR",
-                                      "simp.boost" = "Simple BT",
-                                      "comp.boost"= "Complex BT",
-                                      "comp.rf"= "Complex RF",
-                                      "FWLS" = "FWLS")) +
-        theme(axis.text.x=element_text(angle=30, hjust=1),
-              legend.position = "bottom")
-
-```
-
-The complex Boosted Tree method had the highest TPR, but also had a high FPR.
-The increased number of FP (i.e.
-rejected profitable jobs) offset the more accurate rejection of loss-making jobs so much that the simple averaged blend and simple logisitc regression blend outperformed it.
-The afore-mentioned simple blends performed similarly, with a higher than average TPR and average FPR in comparison to other methods.
-The complex Random Forest blend and FLWS both had lower TPRs than the simple blends but higher TN rates.
-In the end, the improved TN rate did not outweight the lower TPR, i.e.
-the higher rate of accepting profitable jobs did not outweight the lower rate of accurately rejecting loss-making jobs.
-
-All individual original methods used substantially lower TPRs at their optimal thresholds, while their TN rates were spread out.
-This means, as the trialed threshold values moved, and the TPR improved, the increasing lost profits from FP's was too great to justify a TPR higher than 4.6%.
-Blending the individual methods has made use of their unique strengths and seems to have clustered the probabilities of the actual loss-making jobs more accurately towards 1, and the profitable jobs towards 0.
-This allowed for higher TPRs with lower FPRs at the blended methods' optimal thresholds.
-
-### Review of Jobs Rejected by Simple Logistic Regression
-
-The profit curves and TP/FP rates give a good indication of overall performance of the methods, however an analysis of the characteristics of the predicted loss-making projects can provide more detailed insight.
-
-All below plots and comments summarise results from one Simple Blended Logistic Regression Model.
-The results from this model were chosen because it represented the mean for improvement in profits output by this blending method.
-The plot below illustrates how the ratio of profitable to unprofitable jobs compare in the full project population vs.
-the rejected projects.
-
-```{r, fig.cap = c("Proportion of profitable and loss-making jobs in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
-
-reject.plot <- readRDS('reject_plot.rds')
-
-ggplot(reject.plot, aes(x = f.rpdol)) +
-        geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +
-        facet_wrap(~compare,
-                   labeller = labeller(compare = 
-                                        c("full" = "All Projects",
-                                       "reject"= "Rejected Projects Only"))) + 
-        scale_y_continuous(labels = scales::percent) +
-        labs(x = "Project Performance", 
-             y = "Relative Frequency",
-             title = "Ratio of Profitable vs. Loss-Making Projects") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-Originally, 25% of projects were loss-making.
-The model rejects a population of projects which are over 60% loss-making.
-This represents a significant swing in proportions, but is not perfect.
-The class imbalance towards profitable jobs in the initial dataset may have weakened the model's ability to classify losses.
-@Saradhi2011 took advantage of SVM's ability to adjust class penalties to address class imbalance, however SVM's were avoided because of their black box predictions in this work and the remaining methods cannot accommodate class penalties.
-
-The plot below illustrates the monetary size of profits and losses from rejected projects.
-The different coloured bands indicate different projects and it can be observed that the magnitude of rejected profits and losses vary greatly.
-They range from profits/losses of just \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% tail(1) %>% round(0)` to losses of \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% head(1) %>% round(-3)`.
-
-```{r, fig.height = 5, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
-Colours are alternated between different jobs to indicate profit/loss magnitude.
-Predictions were from a typical simple Logistic Regression.")}
-
-ggplot(reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))), 
-             aes(x = f.rpdol, y = abs(balance.mlsto), fill = mlsto)) +
-        geom_bar(stat = "identity") + 
-        theme(legend.position="none") +
-        scale_fill_manual(values = rep(c('dodgerblue4', 'darkolivegreen1'), 59)) + 
-        scale_y_continuous(labels = scales::dollar) +
-        labs(x = "Project Performance", 
-             y = "Cumulative Profits/Losses",
-             title = "Magnitude of Profits/Losses from Rejected Projects") +
-        geom_text(data = NULL, x = .5, y = 720000, size = 3, 
-                  colour = "gray60", hjust = 0,
-                  label = "Alternate Colours \nIndicate Different \nProjects")
-
-```
-
-The variable 'percentage of hours completed by a professional level employee' had an interesting contrast in the statistical distribution of all projects vs. the rejected projects (refer plot below).
-
-```{r, fig.cap = c("Distribution of the percentage of hours completed by a professional level employee in the full data set of projects vs.
-the rejected projects only.
-Predictions from a typical simple Logistic Regression were used.")}
-
-ggplot(reject.plot, 
-             aes(x = pc.pro)) +
-        geom_density() + 
-        facet_wrap(~compare,
-                   labeller = labeller(compare = 
-                                        c("full" = "All Projects",
-                                       "reject"= "Rejected Projects Only"))) +
-        labs(x = "% of Hours Completed by Professional Level Employee", 
-             y = "Density",
-             title = "Density of Percentage of Hours by Professional Level Employees") +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-The algorithm tended to accept projects which were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
-This could either indicate that professional level employees complete projects more efficiently or that the types of projects that do not require technical employees are more profitable.
-
-The distribution of timespans of rejected projects differed significantly to the timespan distribution of all projects.
-The chart below shows that proportionally very few short jobs were rejectd, i.e.
-projects that lasted between 1 day and 2.5 months.
-These projects are likely small and may be easier to estimate fixed fees.
-A much higher relative proportion of projects lasting 1.5 - 3 years were rejected.
-They may face an opposite challenge to the small projects.
-On the other hand, large projects can provide long term security for a small business and it is unlikely the company would want to reject them.
-It may make more sense to adjust the contract structure of large projects that have been highlighted as a risk.
-Alternate uses of the algorithm results are discussed in the next section.
-
-```{r, fig.cap = c("Proportion of each timespan category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
-
-ggplot(reject.plot, 
-             aes(x = b.timespan.cbrt)) +
-        geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
-        facet_wrap(~compare,
-                   labeller = labeller(compare = 
-                                        c("full" = "All Projects",
-                                       "reject"= "Rejected Projects Only"))) +
-        labs(x = "Timespan", 
-             y = "Relative Frequency",
-             title = "Relative Frequency of Project Timespans") + 
-        theme(axis.text.x=element_text(angle=30, hjust=1),
-              panel.background = element_rect(fill = 'gray95'))
-
-```
-
-The statistical distribution of return per dollar in the rejected projects is skewed towards negative returns.
-This means the algorithm has identified a proportionally much higher percentage of loss-making projects which is not surprising given previous results.
-
-```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
-
-ggplot(reject.plot, 
-             aes(x = return.pdol)) +
-        geom_density() + 
-        facet_wrap(~compare,
-                   labeller = labeller(compare = 
-                                        c("full" = "All Projects",
-                                       "reject"= "Rejected Projects Only"))) +
-        labs(x = "Return per Dollar", 
-             y = "Density",
-             title = "Density Plots of Return per Dollar") +
-        scale_x_continuous(labels = scales::dollar) +
-        theme(panel.background = element_rect(fill = 'gray95'))
-
-```
-
-There were a couple more revealing differences between the full set of projects and the rejected projects.
-The first was that rejected projects had disproportionately more projects with mid-level and senior-level technical employees as the main worker on a project.
-The second was that a comparatively high percentage of rejected projects were subdivisions, sewerage, or civil building works.
-
-The detailed analysis of rejected project characteristics gave insight into the characteristics of projects that are likely to be loss-making as well as the accuracy of the algorithm.
-While it is interesting to observe characeristics of rejected projects, a complex combination of all the features contribute to the final probability of loss prediction.
-
-### Alternate Interpretations of Blended Model Results
-
-Alternative strategies exist for how to deal with projects that are marked with a probability of loss greater than `r best.t`.
-For example, contracts for these projects could be changed to hourly rate contracts without exception.
-This means that each hour booked on the project will definitely be charged to the client for each stage of the project.
-Another example would be to increase fixed fees for projects above the threshold by a nominal percentage.
-However, this method seems less practical as the case study projects above `r best.t` would need to increase fees by 46% in order to generate a 15% profit.
-A proposed fee increase of that magnitude may not be as palatable as an hourly rate contract depending on the project and client.
-Contracts could be adjusted in many other creative ways to reduce risk in high probability projects.
-The threshold of `r best.t` was optimised based on the accept/reject strategy but different courses of actions could be developed for specific clients or project sizes via similar analyses of threshold scenarios.
-
-To conclude, because of the promising AUC results, it was logical that the models translated into financial benefits for the company.
-The `r (best.mean-100)`% improvement in profits looks reasonable, and may be high enough to trigger further cost-benefit analyses before finally introducing the model into business decisons.
-Development of the user interface as well as some further data analyses are recommended before implementation of the model and are described in the 'User Interface and Engagement' and 'Limitations and Future Work' sections below.
-
-### User Interface and Engagement
-
-Industry's lack of uptake of cost estimation models was marked as a significant issue in the literature [@Akintoye2000].
-Decision makers did not sufficiently trust the model output which were not directly connected to their experiences.
-In many studies, case-based reasoning had more traction than more accurate models because decision makers could justify the output.
-Clearly this is an important aspect of the study to address.
-The user interface should do two things well:
-
-* Communicate the model results practically and clearly
-* Create trust via a connection to the user's experiences
-
-User interfaces also allow input values to be tweaked and re-run so the decision maker can test different scenarios.
-A prototype user interface was developed using the Shiny package in R where a strategy was developed for how the input options and results would be communicated [@shiny].
-
-![Opening panel and tab of user interface application for the predictive model](images/app/front_page.png)
-
-The manager selects the charactersitics that describe their new job in the top panel.
-Then once the 'Calculate' button is clicked, the blended Logistic Regression model works behind the scenes to generate a probability score between 0 and 1 where 1 represents a loss-making job.
-This output is not intuitive to understand for a manager because the probability score does not represent a statistical probability, rather a score the algorithm generated using its own measures.
-The optimal threshold analysis found that scores below 0.6 should be classified as profitable jobs and scores above 0.6 are loss-making.
-
-The scores were re-scaled to a more intuitive numbering system where scores between 0 and 0.6 were scaled between 0 and 10, and scores higher than 0.6 were scaled between 0 and -10.
-The shaded confidence interval was calculated from the distributed spread of probability outputs from the 100 models calculated in the Profit Curve Analysis.
-The 100 models gave 100 probability outputs for each project, and the 95% confidence interval for these outputs was calculated.
-A typical spread for the 95% confidence interval is shaded in the graphic which is relative to the 'bin' that the score fell into (i.e.
-6-8).
-This provides an intuitive feel for the model's uncertainty.
-
-This graphic is an initial suggestion for how to communicate model results to industry.
-Many variations are possible which could depend on how the business wants to act on the output; for example whether they plan to reject risky projects outright or adjust the fee structure.
-Additional methods built on top of the current model may be able to produce statistical probabilities instead of probability scores.
-This may be achieved using Bayesian methods such as Bayesian additive regression trees, however it is uncertain whether this would improve practical effectiveness.
-
-Two strategies were adopted to facilitate user trust in the model's output: illustration of the model structure and a nearest neighbour graph.
-Algorithms were originally selected that could reveal their predictive structure by either variable importance and/or variable relationships.
-The first tab in the application displays several variable dependency plots extracted from a typical Boosted Tree model.
-Understanding the plot would require some training.
-The y-axis represents the probability a project will be loss-making after taking into account average effects from variables other than the x-axis variable.
-
-![Graphic for model prediction and Boosted Tree partial dependency plot](images/app/team_size_dep.png)
-
-![Boosted Tree Partial dependency plot for Timespan embedded in the user interface](images/app/timespan_dep.png)
-
-Secondly, the nearest neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
-The aim was to provide the user with a cross check of the algorithm's results by finding projects with the most similar features to the one currently being assessed.
-This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'nearest neighbours'.
-Some screen shots of the nearest neigbhour graphic are presented below.
-
-![User interface for Nearest Neighbour algorithm input and output chart](images/app/NN_top.png)
-
-![Nearest Neighbour chart coloured by the position of the main employee on the projects](images/app/NN_emp.png)
-
-![Nearest Neighbour chart coloured by the client for the projects](images/app/NN_client.png)
-
-Tables detailing specific details of the most similar projects are also provided below the chart:
-
-![Tables below the chart in the application give specific details of each of the nearest neighbours](images/app/NN_table_top.png)
-
-![The nearest neighbour details are split into sections such as 'Finances' and 'Staff Details'](images/app/NN_table.png)
-
-The similar projects are calculated using a different method from the blended algorithm but it is important to have a form of cross-checking and engage the user's own experiences.
-Further development and testing is required to assess how intuitive the results are for users to interpret, how much users trust the algorithm output given the variable plots and nearest neighbour charts, and finally how trustworthiness may be further improved.
-
-## Limitations and Future Work
-
-The `r (best.mean-100)`% improvements in profit sound promising, however it is important to consider limitations in the data and implementation process, which in turn motivates future research.
-This includes how accurately users of the algorithm would be capable of guessing input values and how to integrate new data into the model.
-
-First, succesfully estimating project variables is critical in translating this research into tangible business benefits.
-The data represents the company's records of *completed* projects so initially expected variable values were not documented.
-For example, if a project was originally expected to last 1.5 years but instead took 2, the company data only retains the two year timespan.
-This prompts the question, how well could a user predict the variables before beginning a project?
-Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e.
-"1 day - 3 weeks", "3 weeks - 2.5 months", etc).
-However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables.
-It would be interesting to test how well business managers could guess the aforementioned four variables at the start of a project.
-
-The process of initially guessing variable values and then checking them upon project completion is time consuming.
-However, it would be worthwhile integrating into the pre-existing data-collection system.
-When a new project is entered into the database, it could be mandated that the project manager guesses the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent by professionals.
-Over time, this estimation data could contribute to a revised version of the algorithm.
-
-<!-- The strategy for updating the algorithm as new data is generated should be developed in future work.  -->
-<!-- A lot of data cleaning in past data -->
-<!-- need to build a pipeline of cleaning - only 11 variables now, not too hard -->
-<!-- * re run algorithm monthly? does not take long, maybe 1 minute. -->
-<!-- * repeat validation every few months -->
-
-<!-- The Discussion section: -->
-
-<!-- comments on your results; -->
-<!-- explains what your results mean; -->
-<!-- interprets your results in a wider context; indicates which results were expected or unexpected; -->
-<!-- provides explanations for unexpected results. -->
-<!-- The Discussion should also relate your specific results to previous research or theory.
-You should point out what the limitations were of your study, and note any questions that remain unanswered.
-The Discussion CAN also include Conclusions/Future Research.
--->
-
-# Chapter 10: Conclusion
-
-The objectives of this research consisted of a general aim and two research hypotheses, all of which have been fulfilled and tested.
-The findings made a significant contribution to the present body of literature and highlighted areas of future work that would refine the outcomes.
-
-The general aim was to use statistical techniques to model the profitability of projects for a case study consulting business using their internal CRM data.
-This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitablility and as a binary classification problem predicting simply profit or loss.
-Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees and multiple methods of blending the individual models.
-
-Hypothesis 1 stated a statistical or machine learning model based on the CRM data could predict the profitability of a new project.
-Testing this hypothesis revealed that the degree of profitability (the regression problem) could not be accurately predicted but whether a job would be profitable or not (binary classification) could be predicted to a level significantly better than random assignment (AUC = 0.76).
-This was achieved by 3 individual methods while various techniques that blended the individual methods improved results further (AUC = 0.77).
-It is likely the simpler task of binary classification suited the complex yet small data set better.
-
-Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom l ine) of the case study business.
-This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business scenario.
-The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss-making'.
-A range of threshold were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
-If a project was accepted, the profits and losses were summed while if a project was rejected the profits and losses were forfeited.
-A perfect model would reject all loss-making jobs and accept all profitable jobs, thereby improving overall profits.
-Final results showed the Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
-The 95% confidence interval for these improvements were between `r best.fin %>% filter(method == "simp.log") %>% .$lower_bound_95CI - 100`% and `r best.fin %>% filter(method == "simp.log") %>% .$upper_bound_95CI - 100`% using a probability threshold of `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).
-
-These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
-Ensemble tree methods and blending have been applied minimally to cost estimation.
-Past research tested regression models and to some extent Neural Networks and case based reasoning.
-The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting as well as Neural Networks.
-Next, a clear framework for how a business can tangibly improve profits was developed and proven successful.
-Previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
-Lastly, the decision framework was progressed even further into a prototype application where user trust was addressed.
-Previous studies had shown that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
-This was addressed by providing intuitive plots of the model structure and a parallel nearest neighbour graph of similar internal projects.
-Further work is required to test user confidence in the output.
-
-Another topic identified for future research was to test how well managers estimate project input variables.
-Estimation of the timespan, team size, total invoiced amount, and percentage of hours completed by professionals in particular should be tested.
-Initial estimates of input variables could also be kept in the database for future analysis.
-Overall, this project has successfully demonstrated a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models could guide managers in rejecting loss-making projects and improve net profitability.
-
-# References

### 2016-07-15 finalwriteup/draft.rmd
finished editing method

@@
 The literature pertaining to the aim of this research project, cost prediction using internal consulting CRM data, has been extensively reviewed.
-The problem has been addressed in studies beforehand which have generally been limited to prediction of construction costs for buildings/infrastructure and effort estimation for IT projects.
+Previous studies addressing the problem have generally been limited to prediction of construction costs for buildings/infrastructure and effort estimation for IT projects.
 The prediction models consisted of Case Based Reasoning, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM.
@@
 
-The aim of this project is to determine whether the profitability of projects for consulting businesses can be predicted using statistical and machine learning techniques.
-As stated in the literature review, a couple of gaps exist in this research space including working with internal CRM data from a consulting company in the construction industry, trialing ensemble tree methods against the more popular regression and Neural Networks, and analysing the impact model results could have on business decisions.
-This section will explain the chosen methodology for fulfilling the research aim using a case study engineering consulting company that charges for their time spent on project delivery.
+This section outlines the chosen methodology for fulfilling the research aim - predicting project profitability using internal company data and testing how much it improves the business' bottom line.
+The process of how the data was collected, cleaned, and how the predictive models were progressively built and tested is explained along with why the steps were taken.
+This section should enable an experienced practioner to replicate the study.
 
@@
 An assessment of the literature revealed that the capacity of internal timesheet data (extracted from a company's CRM software) to predict project profitability has not yet been tested in the context of consulting companies in the construction industry.
-Project data can be extracted from the CRM, interviewing project team members, emails, accounting records, and project drawings and calculations.
-As the internal CRM data in this case study houses a detailed account of time spent on each project (timesheet data) along with client information and invoicing records for the preceding twelve years, it was clear that this source of data was not only the most accessible and suitable for statistical analysis, but had the potential to reveal information not typically investigated by the company.
-
-The complete twelve years of project timesheets and invoices was extracted via the CRM software interface, which described 4169 projects.
-The projects varied from total invoiced amounts of \$500 to over \$1,000,000 and were divided between four internal disciplines in the company.
-During data extraction, it was not clear whether the full twelve years were necessary and relevant to predicting future project profitability but this was determined later in the analysis.
+Alternate sources of project data were considered including interviewing project team members, emails, accounting records, and project drawings and calculations.
+However the CRM data from this case study could be extracted in a spreadsheet format and housed a detailed account of time spent on each project (timesheet data) along with client information and invoicing records for the preceding twelve years.
+This source of data was not only the most accessible and suitable for statistical analysis, but had the potential to reveal information not typically investigated by the company.
+
+The twelve years of project data was extracted via the CRM software interface, which described 4169 projects.
+Projects varied from total invoiced amounts of \$500 to over \$1,000,000 and were divided between four internal company disciplines.
+During data extraction, it was not clear whether the full twelve years were relevant to predicting future project profitability but this was determined later in the analysis.
 
@@
 Project names and addresses were removed, and each project was identified by a number.
-The resulting de-identified datasets were three .csv files of different data structures which are described in the next section.
-
-## Cleaning and variable engineering
-
-The data contained a rich source of project information and a lengthy process of cleaning the data followed by engineering potentially useful variables was then required.
+The resulting de-identified data consisted of three .csv files of different data structures which are described in the next section.
+
+## Cleaning and Variable Engineering
+
+The data contained a rich source of project information and a lengthy process of cleaning the data was required.
+Then, potentially useful variables were engineered which could be tested alongside original variables.
 The three initial datasets described:
 
-* Invoicing records: dated invoiced amounts and sizes for each project and whether the invoice had been paid or written off
+* Invoicing records: dated invoiced amounts for each project and whether the invoice had been paid or alternatively, written off.
 
@@
 
-| Date    | Hours | Charge Amount | Cost Amount | Created By | Discipline | Job No. | Reconciled | Scope       |
+| Date    | Hours | Charge Amount (\$)| Cost Amount (\$)| Created By | Discipline | Job No. | Reconciled | Scope       |
 |---------|-------|---------------|-------------|------------|------------|---------|------------|-------------|
-| 1.2.300 | 4     | 640           | 400         | P12        | Civil      | 1.2.300 | Yes        | Concept     |
-| 1.2.400 | 4.5   | 720           | 450         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.500 | 2     | 480           | 320         | P34        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.600 | 7     | 1120          | 700         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.700 | 1     | 160           | 100         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
-| 1.2.800 | 3     | 390           | 240         | P3         | Structural | 1.2.400 | Yes        | Site        |
+| 2014/09/01 | 4     | 640           | 400         | P12        | Civil      | 1.2.300 | Yes        | Concept     |
+| 2014/09/03 | 4.5   | 720           | 450         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/03 | 2     | 480           | 320         | P34        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/08 | 7     | 1120          | 700         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/09 | 1     | 160           | 100         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/15 | 3     | 390           | 240         | P3         | Structural | 1.2.400 | Yes        | Site        |
 | ...     | ...   | ...           | ...         | ...        | ...        |         | ...        | ...         |\
@@
 Or an expense printing claim of $9.50 would be translated to 9.5 hours spent on that job.
-Once these errors were detected, they were discussed with the company directors and appropriate corrective action was taken.
-Mistakes in data entry were often able to be converted into correct data by notes that a user had entered into the system.
-It is possible that some data entry errors were not detected and this is a problem that must be dealt with in implementation of the model as well.
+Once these errors were detected, they were discussed with the company directors and appropriate corrective action to the data was taken.
+Mistakes in data entry were often able to be corrected by notes that a user had entered into the system.
+It is possible that some data entry errors were not detected and this is a problem that also must be dealt with during implementation of the model.
 
@@
 
-Although dozens of variables were available from the initial dataset, it was advantageous to engineer further descriptive variables to trial in the models.
-This was particularly true for the invoicing dataset and timesheet dataset where a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
+Although dozens of variables were available from the initial dataset, further descriptive variables were engineered to trial in the models.
+This was particularly advantageous in the invoicing dataset and timesheet dataset where a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
 All variables would eventually be tested for variable importance and predictive power with respect to project profitability.
@@
 * Percent of hours performed by 'professional' employees as opposed to 'technical' employees
-* Position of the employee who completed the most hours on each project
+* Position of the employee that completed the most hours on each project
 * Percent of hours done by the majority contributor to a project
@@
 * Total cost of external subcontractors or disbursements per project
-* Total number of users that entered hours on each project
+* Total number of employees that entered hours on each project
 * Mean number of hours per day entered on a project
@@
 * Total amount invoiced and remunerated per project
-* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500)
-* Mean invoice per client, i.e.
-the average invoice size across all invoices sent to a client.
+* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500)over all invoices sent to a client accross all their projects.
 This gives an indication of whether the client generally does fast paced big jobs which would result in large monthly invoices for example or small jobs with small monthly invoiced amounts.
-* Invoice frequency per project. This indicates how drawn out a project was.
 * Client invoice frequency which gives an indication of how much work the company does for that client. 
@@
 This key word analysis was done in conjunction with a company employee and the resulting classifications were reviewed by the employee to ensure accuracy.
+For example, project descriptions containing the words 'residence', 'house', 'apartment', 'home', 'townhouse', 'unit', and 'dwelling' were classified as 'residential'
 * number of projects completed with each client and client contact
 
-With the engineered variables, all three data sources could be represented in the format of a single row per project.
-They were hten combined and further variables were engineered using combinations of variables across the three data sets:
+With the engineered variables, all three data sources could be represented in the single-row-per-project format.
+The three datasets were then combined and further variables were engineered using combinations of variables across the original data sets:
 
@@
 
-Besides the engineered variables, additional variables included in the project summary dataset were client industry, internal company discipline, job description, and post code.
-Once numerous potentially important variables were engineered, the most important ones could then be narrowed down.
+Besides the engineered variables, original variables included in the project summary dataset were client industry, internal company discipline, job description, and post code.
+Once numerous potentially important variables were engineered, the most important ones could then be selected.
 This practice improves the accuracy of a model because unnecessary or irrelevant variables add noise to the prediction of target values.
-Variable selection also increases computational efficiency by reducing the number of calculations and enhances understanding of the prediction structure [@Weisberg2005].
+Variable selection also increases computational efficiency and simplifies understanding of the prediction structure [@Weisberg2005].
 
@@
 
-Once variables were engineered, the value each variable could contribute to the predictive model was assessed.
-As previously stated, variable selection can enhance model accuracy by reducing noise from irrelevant variables and increase simplicity which in turn improves interpretability for the user.
-To assess variable importance, models predicting 'return per dollar' were built using all available explanatory variables.
+After variable engineering, the contribution of each variable to the predictive model was assessed.
+To evaluate variable importance, several models predicting 'return per dollar' were built using all explanatory variables.
 The models then output which variables contributed significantly or most improved results.
-Different types of predictive models have different methods of calculating variable importance and since it was not known which model would best predict return per dollar, a few models' variable outputs were reviewed.
-These included ANOVA Linear Regression and two ensemble tree methods.
-Before models assessing variable importance were built, outlier cases were deleted to prevent distortion of valuable explanatory variables.
+Different types of predictive models have different methods of calculating variable importance and since the best predictive model was unknown, a few importance outputs were reviewed.
+These included one linear and two ensemble tree methods.
+However, before variable importance models were built, outlier cases were deleted because they are often the result of unusual events or errors in data entry.
 
@@
 After discussion with the case study company, it was concluded that for those outliers, the small jobs most likely required so little time that busy employees did not bother writing down their hours for that job.
-Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles.
+Outliers were initially defined as 'return per dollar' values sitting outside 1.5 x the interquartile range (IQR) from the upper and lower quartiles.
 This rule was initially practised by Tukey, the inventor of the box plot [@Tukey1977].
 All outlier projects determined by this method were reviewed with an employee from the case study company and based on their domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined.
-In summary, projects with 'return per dollar' values greater than 3 or below -2 were removed.
+In summary, projects with 'return per dollar' values greater than 3 were removed and one project with a very low 'return per dollar' was removed because it contained incorrect invoiced amounts.
 
@@
 
-Three methods which output variable importance were trialed and compared to test whether they highlighted different variables as valuable.
-This was anticipated as each method models data using distinct theoretical foundations.
+The three methods were anticipated to output different levels of influence for each variable.
 The tested methods were ANOVA Linear Regression, conditional inference forests (cForests), and Random Forests.
 It was expected that Random Forests would favour variables that had more categories, a well-known bias; which is also a bias that cForests have overcome [@Strobl2007].
-ANOVA is based on linear theory which contrasts conditional forest and Random Forest's free-form structure.
+ANOVA is based on linear theory which contrasts conditional forest and Random Forest's more free-form structure.
 Therefore, it is beneficial to compare the methods' assessments of variable importance.
@@
 *how* profitable a project was (continuous) or whether a project was profitable or not (binary classification).
-The continuous response variable was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
-
-#### ANOVA
+The continuous response variable, 'reurn per dollar', was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
+
+#### ANOVA Variable Importance
 
@@
 categorical variables) [@Lunney1970].
-The method is simple and fast but assumes variables are normally distributed and linearly related.
-The variables must therefore be normalised before analysis to achieve a statistical distribution as close to Gaussian as possible.
-A disadvantage of this method is that linear relationships and normally distributed variables do not necessarily represent real world data accurately [@Breiman2001b].
-
-In order to compare variable importance, two types of output from the model are assessed which include the linear coefficients for each variable and the p-values for the coefficients [@Markham2016].
-The magnitude of a coefficient does not indicate importance relative to the other variables because its value is directly reflects the values within the variable.
+Since the method assumes variables are normally distributed, numeric variables must be normalised before analysis to achieve a statistical distribution as close to Gaussian as possible.
+Therefore a disadvantage of this method is that linear relationships and normally distributed variables do not necessarily represent real world data accurately [@Breiman2001b].
+
+In order to compare variable importance, two ANOVA outputs were assessed: the linear coefficients for each variable and the p-values for the coefficients [@Markham2016].
+The coefficient does not indicate importance relative to the other variables because its value reflects values within the variable.
 The sign of the coefficient however indicates whether the variable is directly or inversely related to the response variable.
-A p-values for each variable coefficient is output as the result of a significance test for whether the coefficient equals zero.
-Therefore, if the null hypothesis is rejected (p value < 0.05), it indicates that changes in the explanatory variable are related to changes in the response variable.
+A significance test for whether the coefficient equals zero is performed on each coefficient and the p-value is reported.
+Therefore, if the null hypothesis is rejected (p value < 0.05), changes in the explanatory variable are indeed related to changes in the response variable.
 P-values are a good indication of whether variables have a relationship with the target variable, however they do not necessarily rank importance.
-If a variable is measured more precisely, it will have a smaller p-value, whereas a variable measured roughly will have a higher p-value.
-This does not necessarily make the more precisely measure variable is more important than the other.
+Precision of variable values can influence p-values, for example a variable with high precision will have a smaller p-value, whereas a variable measured roughly will have a higher p-value.
+This does not necessarily make the more precisely measured variable more important than the other.
 It is also important to keep in mind that, with a dozen variables, chance alone can produce a variable with a p-value < 0.001 7% of the time [@Rice1989].
@@
 
-#### Random Forest
+#### Random Forest Variable Importance
 
 The Random Forest algorithm can produce a permutation variable importance for each variable as part of its output.
-Variable importance is represented as a score, which is calculated by performing permutations on each variable in the test data for each tree and re-running the test data through the trees.
+Variable importance is represented as a score, which is calculated by performing permutations on each variable in the test data for each tree, then re-running the test data through the trees.
 More specifically, consider a single variable.
@@
 The trees in the Random Forest are then run again.
-Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted to the original Random Forest.
+Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted against the original Random Forest.
 If the accuracy of the permuted forest is much less than the original forest, then that variable was important.
-Permutation importance scores are calculated as the mean decrease in accuracy over all trees for that permuted variable in the Random Forest [@Breiman2005].
-
-The Random Forest variable importance provides valuable insight as
-it addresses the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables [@Strobl2007].
-However, a severe disadvantage of Random Forest variable importance is that it is not reliable for variable selection where explanatory variables vary in their scale of measurement or their count of categories.
+Permutation importance scores are calculated as the mean decrease in accuracy over all trees for the permuted variable in the model [@Breiman2005].
+
+These scores provide valuable insight because they represent the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables [@Strobl2007].
+However, a severe disadvantage is that the scores are not reliable when the variables differ in their numeric scales or their counts of categories.
 Also, the importance of correlated predictors is overestimated and the algorithm tends to favour variables that have many possible splits or many missing values [@Strobl2007].
-The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably, however this does not necessarily indicate the variable impacts the response variable most.
-Overall, the Random Forest permutation importance is useful as it is able to compare variables in a machine learning environment and can include multivariate interactions, but the current Random Forest importance measure overestimates the importance of variables with many categories and some numeric variables.
-
-#### Cforests
-
-Cforests are an alternative ensemble tree method to Random Forests that overcome the aforementioned shortfalls of Random Forest variable importance measures.
+The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably.
+However this does not necessarily indicate the variable impacts the response variable the most.
+Overall, the permutation importance is useful to compare variables in a non-linear environment and can include multivariate interactions, but overestimates the importance of variables with with many categories and some numeric variables.
+
+#### Cforest Variable Importance
+
+Cforests are an alternative ensemble tree method that overcomes the aforementioned shortfalls of Random Forest scores.
 It is built from cTrees which are decision trees based on a conditional inference framework.
-The key difference between cTrees and standard decision trees is that a significance test is used for splitting instead of a purity measure such as the Gini coefficient, which is centered around information gain.
+The key difference between cTrees and standard decision trees is that a significance test is used for splitting instead of a purity measure such as the Gini coefficient (which is centered around information gain).
 
@@
 Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable, re-running the forest and comparing the decrease in response variable accuracy.
-Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) as well as Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].
+Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) and Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].
 
@@
 Limiting the predictive model to a concise set of meaningful variables reduces noise and improves predictions.
-Less variables means that a simpler model is being used for prediction which in turn is easier for stakeholders to understand [@Weisberg2005].
-For these reasons, a subset of the most important variables were chosen before modelling which entailed first eliminating outliers, then comparing important variables from ANOVA as well as Random Forests and cForests.
-It can be inferred from the literature that cForests would suit the case study data best as it is unbiased and not limited to linear theory [@Strobl2007].
+Less variables and a simpler model is easier for stakeholders to understand [@Weisberg2005].
+For these reasons, a subset of the most important variables were chosen before modelling.
+This was done by first eliminating outliers, then comparing important variables from ANOVA as well as Random Forests and cForests.
+It can be inferred from the literature that cForests would suit this project's data best as it is unbiased and not limited to linear theory [@Strobl2007].
 
@@
 
-Previous studies predicting project profitability in software and construction projects were predominantly limited to case-based reasoning, regression and Neural Networks.
-There is a need to test other sophisticated machine learning algorithms that are as powerful as Neural Networks but provide insight into the reasoning behind predictions.
+Previous studies predicting profitability in software and construction projects were predominantly limited to case-based reasoning, regression and Neural Networks.
+There is a gap in testing other sophisticated machine learning algorithms that are as powerful as Neural Networks but provide insight into the reasoning behind predictions.
 Ensemble trees (Boosted Trees and Random Forests) and Bayesian Networks fit these criteria.
@@
 * Naive Bayes - baseline model
-* Bayesian network
-* Random forest
+* Bayesian Network
+* Random Forest
 * Gradient Boosted Trees
@@
 Some models have limitations and required additional data processing steps such as normalising numeric variables (Bayesian Networks and Regression) and discretising continuous variables (Bayesian Networks).
-Discretising numeric variables was performed by generating a hierarchical dendrogram of each variable to visualise the clusters.
+Discretisation was performed by generating a hierarchical dendrogram of each variable to visualise the clusters.
 Between four and six clusters were chosen and summarised to find the maximum and minimum values within each cluster.
@@
 For example, the height of the top horizontal bar indicates the ESS for one cluster and the height of the second from top horizontal bar is the ESS for two clusters.
-The height of the vertical branch between the two bars indicates the improvement in ESS by adding a cluster [@Putler2012].
+The difference in the two heights indicates the improvement in ESS by adding a cluster [@Putler2012].
 As more clusters are added, decreases in ESS are less significant and the aim is to choose the number of clusters at an optimal level in this process.
-After initial comparisons of prediction methods had been trialed, the number of clusters for each discretised variable was re-assessed by trialing one more or one less cluster and comparing model output.
+After initial comparisons of predictive methods had been trialed, the number of clusters for each discretised variable was re-assessed by trialing one more or one less cluster and comparing model output.
 
@@
 
-### Missing Data Imputation
-
-All methods except gradient Boosted Trees could not handle missing data.
-Therefore, preliminary runs of each method used subsets of the data which had complete data.
-It was possible that if missing data was imputed, predictions from the models that were limited by missing data could improve.
-Also, a complete data set allows for complete sets of predictions from each method, which can then be blended to further improve results.
-First, imputation methods were trialed against gradient Boosted Tree results since Boosted Trees can handle missing values and should perform equally well with imputed data.
-
-The MICE Random Forest method was chosen because it has been proven to work well with complex data sets [@Shah2014].
-It first performs a standard Random Forest imputation of missing values which are treated as initial 'place holders'.
-Random forest imputation is a complex process which works by first doing a rough imputation of the missing values.
-Then a ten-tree forest is run to calculate proximities between all cases.
-Proximities between two cases represent the number of trees in a forest where the two cases were assigned to the same end node in a tree.
-Once proximities are calculated, missing values for a variable (say variable m) are filled by an average of the other complete m values weighted by the proximity of the case representing each m value.
-
-After the initial Random Forest imputation is complete, the MICE process continues.
-The imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable again using Random Forest imputation.
-This is repeated for each variable and their 'place-holder' imputations are replaced by an imputation targeted at the single variable.
-This cycle is repeated five times by default [@Azur2011].
-
-The final imputed dataset was fed into a Boosted Tree algorithm and if similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
-The imputed dataset was then trialed on the remaining methods and compared to un-imputed trials.
+Although five methods were being tested, the response variable (profitability) could be numeric or categorical, i.e.
+
+* the 'return per dollar' of each project (numerical) or 
+* whether each project made a 'profit' or 'loss' (categorical)
+
+This means the problem could be framed as either a regression or binary classification problem, and each predictive method could be applied to both.
+Since it was uncertain which framework would provide the most benefit, the models were built in both.
 
@@
 
-To compare the models, root mean squared error (RMSE) statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to measure binary classification.
+**you are here**
+
+To compare the models, the root mean squared error (RMSE) statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to measure binary classification.
+
+An ROC graph visualises the curve from which an AUC score is calculated.
+Its two axes are the false positive rate (FPR) on the x-axis and true positive rate (TPR) on the y-axis.
+In this problem a 'positive' is a loss-making job so a TP is a case when the model predicts a job will be loss making, and it did indeed lose money.
+A false positive (FP) is when the model predicts a job will be loss making but in reality it was profitable.
+These values are calculated as follows:
+
+$$TPR = \frac{Count of TP}{Count of Positives}$$
+
+$$FPR = \frac{Count of FP}{Count of Negatives}$$
+
+Only predictions made on the test data should be entered into the equations above, therefore a model must be first created using a training set.
+The model can then make predictions on the test set.
+Binary class model predictions are probability values between 0 and 1, which cannot be entered into the equations above.
+In order to classify each case in the test set as 0 or 1, a probability threshold must be chosen.
+That is, if a probability threshold is chosen to be 0.6, each case in the test set can then be classified as 0 or 1 depending on whether its predicted probability is greater or less than 0.6.
+Once the 0 or 1 classifications have been assigned for a certain threshold, the TPR and FPR can be calculated for that threshold.
+This is plotted as a single point on the ROC curve, say point C.
+Refer the diagram below:
+
+![Points with TPR's and FPR's as coordinates for different probability thresholds [@Provost2013]](images/method/ROC1.JPG)
+
+A model that is perfectly classified would have all positives correctly classified (1.0 True Positive Rate) and no incorrectly classified positives (0.0 False Positive Rate).
+If a model has a 1.0 TPR and 1.0 FPR (top right corner) it has correctly classified all positives at the expense of incorrectly classifying all negatives as positive.
+It is 'dumbly' classifying all cases as positive.
+If a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previous example.
+There is an 80% chance *any* case will be classified as positive.
+Therefore, a point that lies on the diagonal line classifies cases with the same capability as random chance.
+
+To plot the curve in an ROC, a series of points and their coordinates must first be calculated.
+This is done by choosing numerous threshold probability values between 0 and 1, then classifying each probability outcome in the test set as 0 or 1 based on the threshold value.
+Finally the TPR and FPR for that threshold point can be calculated which are the coordinates.
+Once enough points are plotted, a curve can be drawn.
+The closer the curve reaches toward the top left hand corner, the closer the algorithm is to perfectly predicting positive and negative cases at some optimal threshold.
+A curve that squarely reaches the top left hand corner would have an area under its curve (AUC) of 1.
+An AUC between 0.5 and 1 means the model is performing better than random chance (since the diagonal line represents a model randomly assigning positives to all cases at a certain rate and would have an AUC of 0.5).
+
+![An ROC defined by data points calculated from 5 probability thresholds [@Provost2013]](images/method/ROC2.JPG)
+
 For binary classification, AUC is a more useful statistic than classification accuracy when output is available as a probability which can be applied to the problem.
@@
 
-An ROC graph visualises the curve from which an AUC score is calculated.
-Its two axes are the false positive rate (FPR) on the x-axis and true positive rate (TPR) on the y-axis.
-In this problem a 'positive' is a loss-making job so a true positive (TP) is a case when the model predicts a job will be loss making, and it did indeed lose money.
-A false positive (FP) is when the model predicts a job will be loss making but in reality it was profitable.
-These values are calculated as follows:
-
-$$TPR = \frac{Count of TP}{Count of Positives}$$
-
-$$FPR = \frac{Count of FP}{Count of Negatives}$$
-
-Only predictions made on the test data should be entered into the equations above, therefore after a model must be first created using a training set, then the test data is input.
-From the test data, the model outputs probability values between 0 and 1 for each case, which cannot be entered into the equations above.
-In order to classify each case in the test set as 0 or 1, a probability threshold must be chosen.
-That is, if a probability threshold is chosen to be 0.6, each case in the test set can then be classified as 0 or 1 depending on whether the probability calculated by the model is greater or less than 0.6.
-Once the classifications have been made, the TPR and FPR can be calculated for that model and threshold.
-This is plotted as a single point on the ROC curve, say point C.
-Refer the diagram below:
-
-![Points with TPR's and FPR's as coordinates for different probability thresholds [@Provost2013]](images/method/ROC1.JPG)
-
-A model that is perfectly classified would have all positives correctly classified (1.0 True Positive Rate) and no incorrectly classified positives (0.0 False Positive Rate).
-If a model has a 1.0 TPR and 1.0 FPR (top right corner) this means it is correctly classifying all positives at the expense of incorrectly classifying all negatives as positive.
-It is 'dumbly' classifying all cases as positive.
-If a model classifies 0.8 of its positives as TP but also 0.8 of its negative cases as positive it is 'dumb' in a similar way to the previous example.
-There is an 80% chance *any* case will be classified as positive.
-Therefore, models that lie on the diagonal line classify cases as well as random chance.
-
-To plot the curve in an ROC, a series of points and their coordinates are first calculated.
-This is done by choosing numerous threshold probability values between 0 and 1, then applying each threshold point to the probability outcomes in the test set, and calculating the TPR and FPR for that threshold point.
-Once enough points are plotted, a curve can be drawn.
-The closer the curve reaches toward the top left hand corner, the closer the algorithm is to perfectly predicting positive and negative cases at some optimal threshold.
-This would give an AUC of 1.
-An AUC of between 0.5 and 1 means the model is performing better than random chance (as previously stated the diagonal line represents a model randomly assigning positives to all cases at a certain rate).
-
-![An ROC defined by data points calculated from 5 probability thresholds [@Provost2013]](images/method/ROC2.JPG)
-
-In order to compare which models performed significantly better than others, an adequate sample size of results statistics was required.
+In order to compare which models performed significantly better than others, an adequate sample size of results statistics is required.
 Multiple models could be made by using different data in the training vs.
@@
 The highest performing models were advanced to the next stage where various methods of combining, or blending, methods were tested.
+
+### Missing Data Imputation
+
+All methods except gradient Boosted Trees and Naive Bayes could not handle missing data.
+Therefore, preliminary runs of each method used subsets of the data which had complete data.
+It was possible that if missing data was imputed, predictions from the models that were limited by missing data could improve.
+Also, a complete data set allows for complete sets of predictions from each method, and these predictions could then be blended to further improve results.
+The procedure behind model blending is addressed further in the next section.
+The imputed data set was compared to gradient Boosted Tree results since Boosted Trees can handle missing values and should perform equally well with imputed data.
+
+The MICE Random Forest method was chosen for imputation because it has been proven to work well with complex data sets [@Shah2014].
+It first performs a standard Random Forest imputation of missing values which are treated as initial 'place holders'.
+Random forest imputation is a complex process which works by first doing a rough imputation of the missing values.
+Then a ten-tree forest is run to calculate proximities between all cases.
+Proximities between two cases represent the number of trees in a forest where the two cases were assigned to the same end node in a tree.
+Once proximities are calculated, missing values for a variable (say variable 'm') are filled by an average of the other complete 'm' values weighted by the proximity of the case representing each 'm' value.
+
+After the initial Random Forest imputation is complete, the MICE process continues.
+The imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable again using Random Forest imputation.
+This is repeated for each variable and their 'place-holder' imputations are replaced by an imputation targeted at the single variable.
+This cycle is repeated five times by default [@Azur2011].
+
+The final imputed dataset was fed into a Boosted Tree algorithm and if similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
+The imputed dataset was then trialed on the remaining methods and compared to un-imputed trials.
 
@@
 These included simple averaging of the individual model results, building a regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
-The first three models are simply the average, weighted average, of the three individual model predictions.
-The last three methods take advantage of potentially meaningful interactions between an individual model's predictions and meta-features (the original variables).
-For example, if the Boosted Tree model predicted profitability better than the other models for Water projects, the blending model would take advantage of this interaction.
+The first three models are simply the average, weighted average of the three individual model predictions.
+The last three methods take advantage of potentially meaningful interactions between an individual model's predictions and meta-features.
+For example, if the Boosted Tree model predicted profitability better than the other models for projects in the "Water" discipline, the blending model would take advantage of this interaction.
 Boosted tree predictions would be weighted higher than the other model predictions for Water projects.
@@
 A Linear Regression is performed to weight each term's contribution to the final predictive accuracy.
-Random forests and Boosted Trees were also used to blend meta-features with output from the best models.
+Random Forests and Boosted Trees were also used to blend meta-features with output from the best models.
 Feature interaction is performed passively due to the nature of how trees are built.
@@
 A single training set for a blended model must be built from the test results of the individual models.
-Therefore, fivefold cross validation was performed using each individual model to create a full dataset as a compilation of test data results.
+Therefore, five-fold cross validation was performed using each individual model to create a full dataset as a compilation of test data results.
 The complete test data predictions were then added to the complete original dataset, and training and testing data could be partitioned to test the blended models.
@@
 
-Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the effect on a business' bottom line was tested.
+Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the effect on a business' bottom line was examined.
 This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs.
@@
 The aim was to find the optimal threshold point where saying 'no' to a job above that level would result in higher profits, because jobs that were likely to make a loss were being rejected.
-This chart will clarified what percentage of profit increase the company could expect by integrating the algorithm into decision making.
+This chart will clarify what percentage of profit increase the company could expect by integrating the algorithm into decision making.
 It also provides a clear way to implement the algorithm and promote industry uptake of the research.
@@
 Finally, since a profit curve is made from a single training/testing instance of the data, the curve varies with different divisions of the data.
-Therefore, in order to understand the uncertainty around the profit curve and to determine which blended or individual model performed statistically better than others, a large sample size of curve was required.
+Therefore, in order to understand the uncertainty around the profit curve and to determine which blended or individual model performed statistically better than others, a large sample size of curves was required.
 Again, 20 curves of each method were built and a power calculation was performed to achieve a power of 0.8.
 This determined the required sample size which was achieved by repeating 5-fold training/testing splits.
-This enabled the highest points of each curve to be statistically compared to one another.
-A 95% confidence interval also be determined around the highest point on each curve.
+This enabled the highest (optimal) threshold points of each curve to be statistically compared to one another.
+A 95% confidence interval could also be determined around the highest point on each curve.
 The final expected increase in profit and the percentage of projects to be rejected presented a clear scenario that the case study business managers could assess in terms of their business strategy.
 
-<!-- ## Method Conclusion -->
+## Method Conclusion
+
+The procedure outlined in this section described how to statistically determine the best methods, including blended methods, for predicting project profitability.
+Furthermore, it was explained how the models could be statistically compared to each other in terms of improving the company's overall profits and to what certainty.
+The following chapters present results and the discussion from these steps.
 
@@
 
-To create the best possible model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
+To create the be st possible model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
 Initially there were 34 explanatory variables for the single response variable, return per dollar, i.e.

### 2016-07-15 finalwriteup/trial_code_eq.rmd
finished editing method

@@
 $odds = e^{\ln\  odds}$
+
+| Date    | Hours | Charge Amount (\$)| Cost Amount (\$)| Created By | Discipline | Job No. | Reconciled | Scope       |
+|---------|-------|---------------|-------------|------------|------------|---------|------------|-------------|
+| 2014/09/01 | 4     | 640           | 400         | P12        | Civil      | 1.2.300 | Yes        | Concept     |
+| 2014/09/03 | 4.5   | 720           | 450         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/03 | 2     | 480           | 320         | P34        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/08 | 7     | 1120          | 700         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/09 | 1     | 160           | 100         | P12        | Civil      | 1.2.300 | Yes        | Preliminary |
+| 2014/09/15 | 3     | 390           | 240         | P3         | Structural | 1.2.400 | Yes        | Site        |
+| ...     | ...   | ...           | ...         | ...        | ...        |         | ...        | ...         |\
+\\

### 2016-07-18 finalwriteup/draft.rmd
started kerries changes

@@
 
+- don't mention bligh tanner!
+
 # Publications
@@
 These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.
+Furthermore,
 
@@
 The fee structure can be as creative as the engaged parties wish, but in fact tend to fit a few common models.
-A couple of simple examples are:
+A couple of examples are:
 
@@
 For example, engineers may moderate their fee by 1% of the entire project cost.
-In the future, it would be beneficial for managers to record the expected total project cost at the beginning of a project versus the concluding completed cost.
 Another piece of data not recorded was the external factors influencing the fixed price quote including known competition from similar firms.
@@
 **Hypothesis 1**
-A statistical or machine learning model based on historical project data can predict the profitability of a new project.
+A statistical or machine learning model based on historical project data can predict the profitability of a new project with greater accuracy than a baseline predictor.
+A baseline predictor for this project is one that predicts the average of a numeric response variable for all cases or, if the response variable is categorical, randomly assigned categories for each case, where the proportion of assigned categories matches the true categorical proportions.
 
 **Hypothesis 2**
-The predictive model built from Hypothesis 1 can be shown to have a positive impact on the bottom line of the case study business.
+The predictive model built from Hypothesis 1 can be shown to have a positive impact on the overall profit earned by the case study business.
+The overall profit is represented by the following equation:
+
+$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee + Business\ Costs)$
 
@@
 In addition to addressing cost estimation with a different kind of dataset, ensemble tree methods are applied which have been minimally tested on this problem.
+Ensemble tree methods combine hundreds or thousands of decision trees to make predictions.
+A single decision tree is a non-parametric machine-learned model built by progressively determining the best binary split to partition the data according to the response variable.
+For further explanation of the method behind decision trees and ensemble three excellent sources are @Hastie2009, @Elith2008, and @Breiman2005.
 To date, Linear Regression, Neural Networks, Case Based Reasoning, and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks in some cases and generally outperform Linear Regression (**ref**).
@@
 The next chapter, Chapter 3 is a thorough review of the literature to date that relates to this thesis.
-This covers work solving the cost-estimation problem, studies on statistical and machine-learning models in business applications, and a review of effective statistical and machine learning methods.
+This covers work solving the cost-estimation problem, studies on statistical and machine-learning models in business applications, and a review of potential statistical and machine learning methods.
 
@@
 
-The research outlined in the method were divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion:
-
-* Chapter 5: Variable Selection
-* Chapter 6: Regression Models
-* Chapter 7: Binary Classification Models
-* Chapter 8: Model blending,
-* Chapter 9: Extended Analysis including overall profit analysis, research limitations and future work
-
-The results section in each of the chapters listed above presents key charts and tables that progressively answer the research question: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project and have a positive impact on the bottom line of the business?
+The research outlined in the method were divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion.
+Chapter 5 describes the variable selection process.
+Chapter 6 and 7 cover the results of the two ways of framing the cost estimation problem: using regression models and binary classification models.
+Chapter 8 shows the outcomes of model blending, then Chapter 9  extends the predictive analyses into an overall profit analysis, as well as describing research limitations and future work.
+
+The results section in each of the chapters listed above presents key charts and tables that progressively answer the research hypotheses: can a statistical or machine learning model based on a company's internal historical project data predict the profitability of a new project better than a baseline model?
+And would these predictions have a positive impact on the overall profits earned by the business?
 Then, a discussion finalises each chapter, where the results are examined more deeply and analysed in a broader context.
@@
 
-Linear Regression is one of the simplest and most popular statistical prediction methods [@Hastie2009].
-An equation for a line is learned, which is defined by a constant and a coefficient multiplied by the value of each variable.
-The coefficients are chosen to minimise the residual sum of squares (least squared) [@Hastie2009].
+Linear Regression is one of the simplest and most popular statistical prediction methods.
+An equation for a line is learned, which is defined by a constant and coefficients multiplied by each explanatory variable which are assumed to be independent.
+This can be mathematically represented by the equation below:
+
+$f(x) = \beta_{0} + \sum_{j = 1}^{p} X_{j} \beta_{j}$
+
+[@Hastie2009]
+
+where $X_{j}$ represents the jth input variable, $\beta_{j}$ is the coefficient or parameter being solved for, and $\beta_{0}$ is the y-intercept.
+The most popular method to find the coefficients is to minimise the residual sum of squares (ordinary least squared method).
+The residuals represent the Y distance of each point from the line (predicted values), and the residual sum of squares is the sum of the squared value of each point's residual.
+It can be represented by the equation below:
+
+$RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$
+[@Hastie2009]
+
 The reason for this method's popularity is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
@@
 
-There are several drawbacks to Linear Regression.
-For example assumptions must be made about the structure of the data.
-Predictor variables are assumed to be linearly related to the response variable and the variables to be normally distributed.
-Many real world phenomena do not correspond to these assumptions which can make it difficult, sometimes impossible, to produce useful results [@Hastie2009][@Putler2012].
-Furthermore, response variables must be continuous numeric variables and if two or more predictor variables are correlated, the values of their coefficients are highly unstable.
-On the other hand, correlated variables do not impact the predictive accuracy of the model.
-A final drawback is that the variable coefficients cannot be used to compare variable importance unless the variables have all been standardised to one another (i.e.
-scaled to the same mean and standard deviation) [@Putler2012].
-This limits the amount of insight that can be drawn from a Linear Regression model.
+If structured correctly, linear regression results can show the importance of the explanatory variables.
+The variables must be standardised to one another (i.e.
+scaled to the same mean and standard deviation), and then the coefficients (the slopes) indicate how *much* a variable changes with the response variable.
+This relationship can then be compared against the other explanatory variables [@Putler2012].
+However, if two or more predictor variables are correlated, the values of their coefficients are unstable.
+More specifically, this means that if the a model has two correlated variables, and is re-run with a slightly different set of data, the coefficients for the two variables may be quite different because the assumption of independent variables is not true.
+On the other hand, correlated variables do not impact the predictive accuracy of the model [@Putler2012].
+
+**http://stats.stackexchange.com/questions/65912/precise-meaning-of-and-comparison-between-influential-point-high-leverage-point**
+
+In linear regression, no assumptions are made about the distributions of the explanatory variables or response variable.
+Skewed explanatory variable distributions are acceptable but it is generally better to normalise them by, for example, taking the logarithm or square root of the variable.
+The reason for this is that the more extreme values from the skewed distribution may create high leverage points and influence the slope (variable coeffiecients) too strongly [@Chatterjee1986].
+The distribution of the residuals on the other hand, should always be Gaussian with a mean of 0.
+This is because if modeled well, the regression equation should follow the shape of the data [@Hastie2009].
+
+Unfortunately many real world world phenomena do not have linear relationships which can make it difficult to produce accurate results using this method [@Breiman2001b].
+There are methods to overcome this limitation such as including polynomial terms and interaction terms.
+For example, in the linear equation, as well as $X_{1}$, include the terms $X_{2} = X_1^2$ and $X_{3} = X_1^3$ to build a third degree polynomial.
+Interactions between explanatory variables can be modelled by including a term such as $X_5$ where $X_5 = X_4*X_1$.
+This versatility is advantageous, however it can be challenging to know which non-linear terms of which variables to include.
+Sometimes data cannot be closely modelled by any equation and this is where other non-linear non-parametric methods can outperform linear regression.
 
@@
 
-The method behind Linear Regression can be adapted to the prediction of a binary response variable (1 or 0).
-A linear equation is fit to the log of the odds of the binary response variable against each of the explanatory variables.
-The response variable (log odds) in the linear equation can then be translated back to odds by taking the inverse log, i.e.
-the exponential:
-
-$odds = e^{\ln\  odds}$
-
-Finally the odds is converted to probability by the following equation:
+The ordinary least squared regression method described in the previous section cannot be applied to binary classification problems, where the objective is to predict a probability of an event occurring.
+Probability is a continuous response variable, however predictions from a linear regression would not be bound by 0 and 1.
+Instead, predicted values could reach further along the line outputting probabilities greater than 1 and less than 0 which is difficult to interpret.
+Also, residuals from linear regression should be normally distributed, but this is not possible when the response variable consists of only two values (**refappstate**).
+
+Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable (1 or 0).
+First, a sigmoid function is assumed as the functional form of the probabilties against variable X.
+This is because it is pragmatic and works well: sigmoid functions can be bound by 0 and 1, and a logit transformation of the probability results in a linear relationship between the log odds and X.
+Refer to the equation below where *p* represents probability:
+
+$\operatorname{logit}(p) = \log\left(\frac{p}{1-p}\right) = \log(odds)$
+
+[@Macdonald1975]
+
+Because of the linear relationship, an equation can be fit to the log odds of the binary response variable against each of the explanatory variables using linear regression.
+Then, the linear equation describing the log odds can be transformed back to probability by taking the inverse log, i.e.
+the exponential and rearranging:
+
+http://www.appstate.edu/~whiteheadjc/service/logit/intro.htm
+https://en.wikipedia.org/wiki/Logit
+
+$$
+\begin{aligned}
+e^{\log(odds)} &= e^{\log\left(\frac{p}{1-p}\right)} \\ 
+&= e^{-\log\left(\frac{1-p}{p}\right)} \\ 
+&= e^{-\log\left(\frac{1}{p} - 1\right)} \\ 
+&= 1 - \frac{1}{p}
+\end{aligned}
+$$
 
@@
 
-This translates the original log odds response variable to a probability S curve bounded by 0 and 1. An example of this translation is graphically shown below:
+This results in a sigmoidal function bound by 0 and 1 for the probability of the binary response variable against X.
+The sigmoidal function has been fit to the log odds in the data.
+This transformation is visualised graphically below:
 
@@
 
-The result of this linear fit and logistic translation is a model that predicts the probability (between 0 and 1) that an input case will result in a 0 or 1 (i.e.
+The result of this linear fit and non linear transformation is a model that predicts the probability (between 0 and 1) that an input case will result in a 0 or 1 (i.e.
 success or failure) [@Moore1989].
-Similar pros and cons exist for Logistic Regression as Linear Regression, and it can be used as a benchmark to compare other binary predictive models for the same reasons.
+Similar pros and cons exist for Logistic Regression as Linear Regression, however the interpretation of the linear coefficients is more complex in logistic regression.
+A coefficient represents the change in the log odds of the response variable for each unit increase in an explanatory variable.
+Therefore, taking the exponential of the coefficient is the change in odds of the response variable.
+If the change in odds was 2, then if the explanatory variable increased by 1, the response variable event would be twice as likely to occur.
+Similarly to linear regression, logistic regression is a good  benchmark to compare other binary predictive models due to its simplicity and speed.
 Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survival/death [@Moore1989].
@@
 Then, to resolve the notorious lack of explanatory qualities, decision trees were employed.
-The 81% of cases that were correctly predicted by the Neural Network were used to generate decision trees, and in turn useful decision rules [@Tsai2009].
+The 81% of cases that were correctly predicted by the Neural Network were used to generate decision trees, and in turn decision rules that could be understood by decision makers [@Tsai2009].
 This strategy is a promising way to benefit from the strengths of well performing complementary techniques.
@@
 The data contained a rich source of project information and a lengthy process of cleaning the data was required.
-Then, potentially useful variables were engineered which could be tested alongside original variables.
+Then, potentially meaningful variables were engineered which could be tested alongside original variables.
 The three initial datasets described:
@@
 
-**you are here**
-
 To compare the models, the root mean squared error (RMSE) statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to measure binary classification.
@@
 
-For binary classification, AUC is a more useful statistic than classification accuracy when output is available as a probability which can be applied to the problem.
-This is pertinent when it is useful to rank the outcomes and trial probability thresholds that will optimise the rates of true positives and true negatives [@Huang2005].
+For binary classification, AUC is a more meaningful statistic than classification accuracy when the output is a probability which can be applied to the problem.
+The AUC indicates model performance across many thresholds while classification accuracy represents a single threshold.
+The AUC statistic is therefore useful in problems where cases can be ranked and a cut-off threshold that will optimise the rates of true positives and true negatives can be found [@Huang2005].
 Note the 'probability' outputs from the models are not true statistical probabilities, but a score the model has assigned to a case using its own measure of certainty.

### 2016-07-19 finalwriteup/draft.rmd
half thru kerries comments

@@
 
-opts_chunk$set(fig.width=6.7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE)
+opts_chunk$set(fig.width=6.7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)
 ```
@@
 In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to Neural Networks, which must be slowly re-trained and re-tuned.
-New cases also did not need to have every variable complete which can occur often in real, messy data.
+New cases also did not need to have every variable complete which can occur often in 'real world' data.
 Neural Networks on the other hand require complete data.
@@
 As stated, the aim of this project is to use statistical and machine learning techniques to model the profitability of projects.
-Several techniques will be tested and this section reviews a range of predictive algorithms that have been successfully applied to real, messy data.
+Several techniques will be tested and this section reviews a range of predictive algorithms that have been successfully applied to business and social data.
 These vary from simple methods such as Linear Regression to complex, deep learning Neural Networks.
@@
 
-**http://stats.stackexchange.com/questions/65912/precise-meaning-of-and-comparison-between-influential-point-high-leverage-point**
-
 In linear regression, no assumptions are made about the distributions of the explanatory variables or response variable.
@@
 
-http://www.appstate.edu/~whiteheadjc/service/logit/intro.htm
-https://en.wikipedia.org/wiki/Logit
-
 $$
 \begin{aligned}
-e^{\log(odds)} &= e^{\log\left(\frac{p}{1-p}\right)} \\ 
-&= e^{-\log\left(\frac{1-p}{p}\right)} \\ 
-&= e^{-\log\left(\frac{1}{p} - 1\right)} \\ 
-&= 1 - \frac{1}{p}
+\log(odds) &= \beta_0 + X_1\beta_1 \\ 
+\log\left(\frac{p}{1-p}\right) &= \beta_0 + X_1\beta_1 \\ 
+-\log\left(\frac{1-p}{p}\right) &= \beta_0 + X_1\beta_1 \\ 
+-\log\left(\frac{1}{p} - 1\right) &= \beta_0 + X_1\beta_1 \\ 
+\log\left(\frac{1}{p} - 1\right) &= -\beta_0 - X_1\beta_1 \\ 
+\exp\left(\log\left(\frac{1}{p} - 1\right)\right) &= \exp(-\beta_0 - X_1\beta_1) \\ 
+\frac{1}{p} - 1 &= \exp(-\beta_0 - X_1\beta_1) \\ 
+\frac{1}{p} &= 1 + \exp(-\beta_0 - X_1\beta_1) \\ 
+p &= \frac{1}{1 + \exp(-\beta_0 - X_1\beta_1)}
 \end{aligned}
@@
 
-$probability = \frac{odds}{1 + odds}$
-
-[@Lowry2016]
-
-This results in a sigmoidal function bound by 0 and 1 for the probability of the binary response variable against X.
-The sigmoidal function has been fit to the log odds in the data.
+[@Macdonald1975]
+
+The result in a sigmoidal function bound by 0 and 1 that describes the probability of the binary response variable against X.
+The sigmoidal function originates from a linear fit of the log odds in the data.
 This transformation is visualised graphically below:
@@
 success or failure) [@Moore1989].
-Similar pros and cons exist for Logistic Regression as Linear Regression, however the interpretation of the linear coefficients is more complex in logistic regression.
+Further tests must then be done to determine at what probability threshold a case would be predicted as 0 vs.
+1.
+This would depend on the purpose of the predictive model, but an example would be the threshold that results in the highest number of correctly predicted 0's and 1's.
+
+If two variables are being used to predict a binary response variable, the linear equation for the log odds can be transformed into a linear boundary in the two variables' feature space.
+First, a probability threshold must be chosen; say this threshold is 0.5.
+The linear equation for log odds can then be rearranged as follows:
+
+$$
+\begin{aligned} 
+\log(odds) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+\log\left(\frac{p}{1-p}\right) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+\log\left(\frac{0.5}{0.5}\right) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+\log\(1) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+0 &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+X_2\beta_2 &= -\beta_0 - X_1\beta_1 \\ 
+X_2 &= -\frac{\beta_0}{beta_2} - X_1\frac{\beta_1}{\beta_2} \\ 
+\end{aligned}
+$$
+
+This results in an equation for a line in the $X_2$ vs.
+$X_1$ feature space, where points are classified in the response variable according to whether they fall above or below the line (above or below the threshold).
+An example of what this looks like graphically is shown below:
+
+```{r log_reg_decboundary, fig.cap = c("Example Bivariate Decision Boundary for Probabability Threshold = 0.5")}
+set.seed(1234)
+
+X1 <- rnorm(20, 1, 2)
+X2 <- rnorm(20)
+
+y <- sign(-1 - 2 * X1 + 4 * X2 )
+
+y[ y == -1] <- 0
+
+df <- cbind.data.frame( y, X1, X2)
+
+mdl <- glm( y ~ . , data = df , family=binomial)
+
+slope <- coef(mdl)[2]/(-coef(mdl)[3])
+intercept <- coef(mdl)[1]/(-coef(mdl)[3])
+
+ggplot(df, aes(x = X1, y = X2, colour = y %>% as.factor)) +
+        geom_point(size = 2) +
+        geom_abline(slope = slope, intercept = intercept, colour = "darkgray", size = 0.8) +
+        theme(legend.position = "none") +
+        labs(title = "Logistic Regression Decision Boundary in Bivariate Space")
+
+```
+
+Similar pros and cons exist for Logistic Regression as Linear Regression, however the interpretation of the linear coefficients is more complex in Logistic Regression.
 A coefficient represents the change in the log odds of the response variable for each unit increase in an explanatory variable.
@@
 If the change in odds was 2, then if the explanatory variable increased by 1, the response variable event would be twice as likely to occur.
-Similarly to linear regression, logistic regression is a good  benchmark to compare other binary predictive models due to its simplicity and speed.
+Similarly to Linear Regression, Logistic Regression is a good  benchmark to compare other binary predictive models due to its simplicity and speed.
 Many business problems have binary response variables, such as yes/no, male/female, buy/do not buy, success/failure, or survival/death [@Moore1989].
@@
 The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for the response variable (the response variable must be categorical).
-More specifically, it is assumed that each variable is conditionally independent given the response class, without considering other predictor variables.
-Mathematically this means that each variable's probability contribution can be calculated by looking at the probability of a response class given a single variable's value.
-The probability of each response class can therefore be calculated for each explanatory variable.
-These are then combined for each possible response class using Bayes theorem, and the class with the highest probability is chosen [@Provost2013].
+To demonstrate the simplified calculation, consider a problem where Y is the response variable, and there are two explanatory variables, $X_1$ and $X_2$.
+If conditional independence is *not* assumed, the probability of $X_1$ and $X_2$ given Y is as follows:
+
+$p(X_1 X_2|Y) = p(X_1|Y) \cdot p(X_2|X_1 Y)$
+
+If conditional independence is assumed, the equation becomes:
+
+$p(X_1 X_2|Y) = p(X_1|Y) \cdot p(X_2|Y)$
+
+Therefore, if conditional independence is assumed, each variable's probability contribution can be calculated by looking at the probability of a variable's value given the response class only.
+This simplification can be favourably integrated into Bayes Theorem which is shown below:
+
+$p(A|B) = \frac{p(B|A) \cdot p(A)}{p(B)}$
+
+Say the probability of 'yes' or 'no' (in a binary reponse variable Y) given the explanatory variables $X_1$ and $X_2$ is being sought.
+The equation above could be rewritten for both possible response categories:
+
+$p(yes|X_1 X_2) = \frac{p(X_1 X_2|yes) \cdot p(yes)}{p(X_1 X_2)}$
+
+$p(no|X_1 X_2) = \frac{p(X_1 X_2|no) \cdot p(no)}{p(X_1 X_2)}$
+
+The aim is to find which category has a higher probability, so to simplify calcualtions, the denominator can be dropped.
+Because it is the same for both equations, the denominator does not affect the final ratio of probability of 'yes' to probability of 'no'.
+Integrating the assumption of conditional independence, the equations can be written as follows:
+
+$p(yes|X_1 X_2) = p(X_1|yes) \cdot p(X_2|yes) \cdot p(yes)$
+
+$p(no|X_1 X_2) = p(X_1|no) \cdot p(X_2|no) \cdot p(no)$
+
+These equations are simple to compute given the data, and the class with the highest probability is chosen [@Provost2013].
 
@@
 This is fine for ranking but the output probabilities are not accurate statistical probabilities [@Caruana2006].
-Another disadvantage is that the distribution of the numeric variable must be assumed (often Gaussian) which the data may not fit into neatly [@Caruana2006].
+Another disadvantage is that the distribution of the numeric variable must be assumed (often Gaussian) in order to calculate the liklihood terms which the data may not fit into neatly [@Caruana2006].
+
+The Gaussian probability densities of two response classes can be visualised on a bivariate feature space.
+The two normal distrubtions represent the probability of each class given values of the two explanatory variables (x and y axes):
+
+![Plot of Gaussian probability density distributions of two response classes given two explanatory variables (represented by the x and y axes) [@Bulatov2010].](images/lit_review/NB_boundary2.JPG)
+
+Where the density functions intersect becomes the decision boundary between the two classes.
 
@@
 Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
-Single decision trees tend to overfit the data and provide low predictive accuracy but on the other hand create intuitive rules that a decision maker can follow in real-life scenarios.
-In addition, their results do not depend heavily on the skill of the data analyst [@Kabra2011][@Putler2012].
-
-To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's with success [@Breiman1996].
+Single decision trees tend to overfit the data and provide low predictive accuracy but on the other hand create intuitive rules that a decision maker can follow in real-life scenarios [@Kabra2011][@Putler2012].
+
+To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's [@Breiman1996].
 Three examples of these methods are boostrap aggregating (bagging), boosting, and Random Forests.
@@
 
-In contrast to Linear Regression, the predictor variables do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
-Other advantages are that ensemble tree methods are faster than SVM's and Neural Networks but performing just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
+Similarly to Linear Regression, the predictor variables do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
+However, it is often wise to transform skewed distributions to achieve more even segmentation.
+Other advantages are that ensemble tree methods are faster than SVM's and Neural Networks but perform just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
 Random forests in particular have the benefit of being unaffected by noise in the data and has the capacity to handle a large number of variables [@Sealfon2012].
@@
 
-Although ensemble trees have numerous advantages it is understand their limitations.
+Although ensemble trees have numerous advantages it is just as important to understand their limitations.
 Trees are not built on a probabilistic framework, and therefore their results can not be provided in this framework.
 For example, statistical confidence intervals for predictions are not available for standard ensemble methods [@Louppe2014].
-This may or may not be required depending on the nature of the problem.
+This will be addressed again in the summary for this section.
 Also, variable importance tables provided by Random Forests and Boosted Trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
@@
 These included Linear Regression, Logistic Regression, Naive Bayes, decision trees and ensemble trees, Bayesian Networks, Neural Networks, and SVM's.
-Linear/Logistic Regression and Naive Bayes are computationally very fast with a simple concept that make broad assumptions about the data, but work well as baseline models against which to compare complex models.
+Linear/Logistic Regression and Naive Bayes are computationally very fast with a simple concept that make some parametric assumptions about the variable relationships, but work well as baseline models against which to compare complex models.
 In certain cases these simpler models are significantly outperformed by complex models however it is important to compare both.
@@
 This may be due to low awareness of learned Bayesian Networks or lower performance.
-In conclusion, simple models such as linear/Logistic Regression and Naive Bayes are good baseline models to implement, while ensemble tree methods and Bayesian Networks may handle messy 'real world' data better than the simpler models while providing insight to decision makers.
+
+None of the methods output statistical uncertainty or a confidence interval associated with each prediction.
+Understanding uncertainty should always be sought in decision making be it from a confidence interval or sensitivity testing.
+One way to achieve this would be to build hundreds of models using the same method but based on different training divisions of the data.
+This would produce a range of results for each case, and a statistical confidence interval or classification probability can be calculated for each prediction.
+Another way for a decision maker to get a feel for uncertainty is to perform sensitivity testing on a prediction.
+This means the user would run a case through a model several times, each time altering one of the explanatory variable to test the degree of variation in the model predictions.
+Both strategies would give an idea of the certainty with which a model is predicting a response value or category.
+
+In conclusion, simple models such as Linear/Logistic Regression and Naive Bayes are good baseline models to implement, while ensemble tree methods and Bayesian Networks may handle bigger unstructured data better than the simpler models while providing insight to decision makers.
 
@@
 Case study projects with extreme 'return per dollar' values were likely the result of special scenarios which the model is not intended to predict.
-For example, many projects appeared wildly profitable but on closer inspection these were all very small jobs.
-After discussion with the case study company, it was concluded that for those outliers, the small jobs most likely required so little time that busy employees did not bother writing down their hours for that job.
+For example, many projects appeared wildly profitable but on closer inspection these were very small jobs.
 Outliers were initially defined as 'return per dollar' values sitting outside 1.5 x the interquartile range (IQR) from the upper and lower quartiles.
 This rule was initially practised by Tukey, the inventor of the box plot [@Tukey1977].
-All outlier projects determined by this method were reviewed with an employee from the case study company and based on their domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined.
+All outlier projects determined by this method were reviewed with an employee from the case study company.
+It was concluded that outliers that were small jobs most likely required so little time that busy employees did not bother writing down their hours for that job.
+Other bigger outlier projects had unrealistically low hours logged against them which could have been the result of data entry errors or errors from when the data was transferred across databases.
+Based on the employee's domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined.
 In summary, projects with 'return per dollar' values greater than 3 were removed and one project with a very low 'return per dollar' was removed because it contained incorrect invoiced amounts.
+
+26 of the 36 outlier cases with a high 'return per dollar' were small projects worth less than \$4,000.
+This may indicate small projects are not suitable to predict due to timesheet entry habits on these projects, however after outlier deletion, 1,159 projects below \$4,000 remained in the dataset.
+The prediction model would therefore still be applicable to small projects under \$4,000.
 
@@
 Historically, in statistics this was called model averaging and since the 1990's research has shown that averaging results from different methods provides better predictive accuracy than any single model [@Madigan1994].
-New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big, messy data.
+New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big socially generated data.
 The prize-winning solution was a complex blend of sub-blends which interacted results of individual models with the original variables.
@@
 
-The qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, general client characteristics, project timespan, and project category.
+The qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, client characteristics, project timespan, and project category.
 It was expected that Billing Type would play a more important role in predicting profitability as it describes the structure of the incoming fees.

### 2016-07-19 finalwriteup/trial_code_eq.rmd
started kerries changes
half thru kerries comments

@@
 \\
+
+$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee + Business\ Costs)$
+
+$f(x) = \beta_{0} + \sum_{j = 1}^{p} X_{j} \beta_{j}$
+
+$RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$
+
+where $X_{j}$ is simply something
+
+$X_{1}$, include the terms $X_{2} = X_1^2$ and $X_{3} = X_1^3$ to build a third degree polynomial.
+Interactions between explanatory variables can be modelled by including a terms such as $X_5 = X_4*X_1$
+
+$\operatorname{logit}(p) = \log\left(\frac{p}{1-p}\right) = \log(odds)$
+
+$$
+\begin{aligned} 
+\log(odds) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+\log\left(\frac{p}{1-p}\right) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+\log\left(\frac{0.5}{0.5}\right) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+\log(1) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+0 &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+X_2\beta_2 &= -\beta_0 - X_1\beta_1 \\ 
+X_2 &= -\frac{\beta_0}{\nbeta_2} - X_1\frac{\beta_1}{\beta_2} \\ 
+\end{aligned}
+$$
+
+$p(X_1 X_2|Y) = p(X_1|Y) \cdot p(X_2|X_1 Y)$
+
+If conditional independence is assumed, the equation becomes:
+
+$p(yes|X_1 X_2) = p(X_1|yes) \cdot p(X_2|yes) \cdot p(yes)$
+
+![Plot of Gaussian probability density distributions of two response classes given x and y explanatory values [@Bulatov2010].](images/lit_review/NB_boundary2.JPG)

### 2016-07-22 finalwriteup/draft.rmd
finished lit review kerrie comments

@@
 Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
-Single decision trees tend to overfit the data and provide low predictive accuracy but on the other hand create intuitive rules that a decision maker can follow in real-life scenarios [@Kabra2011][@Putler2012].
-
-To combat the problems of low predictive accuracy and instability, ensemble tree methods were pioneered in the 1990's [@Breiman1996].
-Three examples of these methods are boostrap aggregating (bagging), boosting, and Random Forests.
-Bagging creates multiple trees by sampling a different training data set (*with* replacement) for each tree and then combining tree results [@Breiman1996].
+Single decision trees create intuitive rules that a decision maker can follow in real-life scenarios but on the other hand, they tend to overfit the data and provide low predictive accuracy [@Kabra2011][@Putler2012].
+Ensemble trees were developed to solve this problem and several ensemble techniques are described in this section.
+
+Decision trees have a different theoretical foundation to traditional prediction methods such as regression.
+They are created by a progression of binary splits of the dataset based on response variable values.
+Before a split is made, each explanatory variable and each value within the explanatory variables is tested for how 'purely' it partitions the response variable into two subsets.
+For example, the explanatory variable 'gender' might split a response variable 'hair length' of a class of students into two subsets of students with more similar hair lengths.
+This similarity is called 'purity' and can be measured using several formulas.
+An example for regression is:
+
+$SSN - (SSL + SSR)$
+
+$SSN = \sum(y_i - \bar{y})^2$
+
+where SSN is the sum of squared errors of a node (dataset before the split).
+$y_i$ are the response variable values and $\bar{y}$ is the mean response variable value in the node.
+SSL and SSR are the sum of squared errors in the left and right datasets that resulted from the split.
+Other purity formulas exist for categorical response variables such as Gini purity and information gain [@Hastie2009].
+After the first split is made on the dataset, each subset is split by another variable (or the same explanatory variable again) and so on resulting in final subsets of data with similar response variable values.
+If this process is visualised graphically, it is clear that non-linear relationships can be captured.
+
+![Graphical illustration of non-linear splits derived from a decision tree with two variables [@Jeevan2015].](images/lit_review/dec_tree.JPG)
+
+Once the splitting rules have been learned, a new case is run through the tree from the top.
+It follows the path provided by which side of the binary split its explanatory variable values fall under and ends up in a final node.
+The mean response variable value from the initial data that ended up in that node is the predicted response value for the new case.
+Decision trees are easy to overfit on the training data.
+This means that if the splitting rules are continued until no more splits can be possibly made, when the test data is run through the tree, the predictions will be too closely aligned to the exact structure of the training dataset.
+Prediction rules need to generalise trends found in the training data.
+This can be managed in a single tree by using 'stopping rules' such as a minimum number of data in a final node or a minimum purity increase for a split to be valid (**ref**).
+
+As stated previously, single decision trees also suffer from low predictive accuracy and instability, so to combat these problems, ensemble tree methods were pioneered in the 1990's [@Breiman1996].
+Three examples of these methods are boostrap aggregating (bagging), Random Forests, and boosting.
+Bagging builds multiple trees from different training sample data sets (wich are sampled from the full data set *with* replacement) and then combines tree results [@Breiman1996].
+The method of combination is generally averaging numeric predictions or taking the majority rule for categorical variables.
+
 Random forests is an advanced form of bagging and builds on the idea of training multiple trees from the same data by sampling bootstrapped training sets with replacement.
@@
 It also prevents the trees from becoming too correlated [@Breiman2001a].
+
 The boosted decision tree approach applies gradient descent theory to a series of decision trees.
 The trees are limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) of the preceding tree.
+By modeling the errors, misclassified cases are weighted higher than correctly classified cases and influence the structure of the tree more [@Hastie2009].
+This increased weighting on errors is why the algorithm is called boosting.
+
+This idea can be visualised in a simple scenario consisting of a single explanatory variable (shown on x-axis) and response variable (y-axis).
+Each tree is limited to one split, and the first tree is shown below, where the vertical line is the splitting point:
+
+![First tree in short progression of boosted trees [@Ihler2012].](images/lit_review/boost_tree1_.png)
+
+The horizontal lines indicate the mean response variable value in each subset after the split.
+It is the value the first tree would predict for data that falls into each subset.
+The next tree is built on the residuals of the preceding tree.
+
+![Second tree modeling residuals in short progression of boosted trees [@Ihler2012].](images/lit_review/boost_tree2_.png)
+
+The y-axis represents the distance each point was from the predicted response value (horizontal lines) in the preceding tree.
+The new split is represented by the vertical line, which creates two subsets of data with more similar residual values.
+Again, the horizontal lines are mean residuals from the previous tree, which become the predicted residual for each point.
+The boosting algorithm combines the series of tree predictions.
+In this case, this is done by adding the predicted residuals from the second tree to the predicted response variable values of the first tree.
+Visually, the y values of the horizontal lines are added:
+
+![Graphically combined sequence of two boosted trees [@Ihler2012].](images/lit_review/boost_add2.png)
+
+The combination of tree predictions and residual predictions becomes the final predicted response values in boosted trees.
 The limited depth of each tree prevents overfitting at each stage and the combined result of up to thousands of trees is very powerful [@Elith2008].
@@
 
-Ensemble decision trees have a different foundation to traditional prediction methods such as regression.
-For this reason, there are some advantages and disadvantages.
-Decision trees use a progression of binary splits which means the data is first split by one variable at a certain value into two subsets.
-Then each subset is split by another variable and so on in order to partition the response variable values into groups with similar values.
-The aim of each split is to achieve the most 'pure' split of the response values in a greedy fashion.
-If this process is visualised graphically, it is clear that non-linear relationships can be captured.
-
-![Graphical illustration of non-linear splits derived from a decision tree with two variables [@Jeevan2015].](images/lit_review/dec_tree.JPG)
-
-Similarly to Linear Regression, the predictor variables do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
+The advantages of ensemble tree methods are numerous.
+Similarly to Linear Regression, the predictor variables in trees do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
 However, it is often wise to transform skewed distributions to achieve more even segmentation.
-Other advantages are that ensemble tree methods are faster than SVM's and Neural Networks but perform just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
+Additionally, ensemble tree methods are faster than SVM's and Neural Networks but perform just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
 Random forests in particular have the benefit of being unaffected by noise in the data and has the capacity to handle a large number of variables [@Sealfon2012].
@@
 This kind of conditional relationship is Bayesian, where the probabilities in one node are conditional upon values in nodes directed to it as well as preceding nodes.
+Refer to the simple example below:
+
+![Simple 4 node Bayesian Network](images/lit_review/BN_1.png)
+
+If each variable, A, B, C, and D are binary categories, there are $2^4$, or 16 combinations of results and each variable is linked to each of the other variables with a conditionally dependency.
+The probability of each combination could be calculated as per below using the chain rule for probabilities:
+
+$p(ABCD) = p(C|ABD)\cdot p(A|BD) \cdot p(B|D) \cdot p(D)
+
+However, if it is known through domain knowledge or can be reasoned through the data that some variables are conditionally independent from one another, the calculations can be simplified [@Barber2012].
+For example, if B and D, A and D, and A and C are conditionally independent, the DAG would be:
+
+![4 node Bayesian Network with conditional independencies](images/lit_review/BN_2.png)
+
+Now, the dependencies that no longer exist can be deleted from the equation:
+
+$p(ABCD) = p(C|BD)\cdot p(A|B) \cdot p(B) \cdot p(D)
+
+In this way, Bayesian Networks reduce computations required to find the probability of an event given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes cannot [@Barber2012].
+Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes though (as Naive Bayes is the simplest form of a Bayesian Netwoork) and variables with many categories dramatically increases the expense [@Zhang2004].
+
+Network relationships can be learned from the data, however this is not widely included as part of the suite of machine learning methods.
+The conditional variable dependencies are calculated from the data which in turn can define the graph structure.
+Some conditional dependencies can be set before the structure is learned [@Barber2012].
+
+Drawbacks of Bayesian Networks include the difficulty for them to process continuous variables.
+These demand sampling from conditional density models or complex algebra whereas categorical variables require comparitively simple calculations [@Hofmann1996].
+It is therefore common to discretise continuous variables [@Kragt2009].
+
 Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.
-
-Bayesian networks are excellent at representing knowledge from different origins, and then visually communicating results well to decision makers [@Kragt2009].
+They are excellent at combining knowledge from different origins, and then visually communicating results well to decision makers [@Kragt2009].
 Therefore Bayesian Networks have become an excellent decision support tool.
-Network relationships can be machine learned, however this is not widely included as part of the suite of machine learning methods and is usually recommended for modelling complex networks.
-Drawbacks of Bayesian Networks include their inability to handle continuous numeric variables (all continuous variables must be discretised).
-This ensures all variable relationships are categorical [@Kragt2009].
+Another benefit of Bayesian Networks as decision support tools is they allow the user to fix certain variable values, re-run the network and observe changes in probabilities.
+Thisis a kind of sensitivity testing that gives users a feel for whether the model makes sense and the degree of uncertainty.
 Popular applications of Bayesian Networks are modelling uncertainty in natural resource management and modelling complex business network structures such as airports [@Wu2013].
 
-### SVM's and Neural Networks
-
-Neural networks were inspired by biological Neural Networks of the human nervous system and require a lot of training data relative to other methods.
-They are flexible in the range of relationships they can mimic and are not greatly dependent on the skills of the analyst (Putler & Krider, 2012).
-A disadvantage that has been mentioned earlier is that the algorithm is a black box.
-Neural networks can perform high level tasks as a result of deep learning such as hand writing recognition, vehicle control, face recognition, and cancer detection [@Haykin2004].
-
-The simplest type of SVM's calculate the linear boundary between two categories in feature space so that the category of future cases may be determined depending on which side of the boundary they fall.
-The location of the boundary is determined by the position of the widest clear margin that can be drawn between data points from the two categories, while allowing a certain number of misclassifications (this is usually optimised for the problem).
-The simple plot below demonstrates this concept and has no misclassifications:
+### Neural Networks
+
+Algorithmic Neural Networks were inspired by Neural Networks of the human nervous system.
+Biological Neurons recieve electro-chemical input from the terminals of adjacent neurons into its dendrites.
+If enough electro-chemical input is recieved, the neuron will fire, which means a signal will be sent from the dendrite end, through the axon to its terminal.
+Its terminal then sends signals to adjacent dendrites of other neurons and the process propagates.
+Neurons are able to learn over time by strengthening the connections of adjacent neurons that frequently cause the next one to fire.
+This means the required firing threshold for the recieving neuron lowers [@Putler2012].
+
+Algorithmic Neural Networks replace neurons with layers of nodes set out in the following structure:
+
+![Layer structure of a Neural Network [@Karpathy2016]](images/lit_review/NN_structure.PNG)
+
+In fact, a wide variety of neural network structures and dynamics exist so a simple popular version is explained here.
+The first input layer is passive, and each input variable is passed through one input node.
+These nodes are passive and the variable values are simply passed to each of the nodes in the next layer, called a hidden layer.
+Neural networks may have any number of hidden layers with any number of nodes as selected by the analyst.
+Often however, a single hidden layer is adequate.
+Calibration for predictions is done in the hidden nodes.
+At these nodes, each input variable is weighted then combined (similar to linear regression) to produce a single value.
+This is then transformed and assessed by a threshold function to determine if the hidden node should 'fire' to the next nodes.
+Once the signals and values have propogated to the final layer they are recombined to produce an output, or a prediction [@Putler2012].
+
+The algorithm learns from the data by starting with trial weights for each node.
+This produces a set of predictions from which errors are calculated.
+Then optimisation techniques are used to adjust the weights and reduce the error.
+This process can result in overfitting, so stopping rules such as regularisation strength should be used [@Karpathy2016].
+
+Because of the interwoven network of nodes, Neural Networks can model  complex non-linear relationships. An example of a possible decision boundary in bivariate space is shown below:
+
+![Non linear decision boundary created by a Neural Network[@Karpathy2016]](images/lit_review/NN_bivariate.PNG)
+
+Clearly, a flexible range of relationships can be mimicked by Neural Networks, and if more nodes are added, more flexibility is available [@Putler2012].
+A previously mentioned disadvantage is that the algorithm is a black box because the internal structure is too complex for interpretation.
+They also require a lot of training data relative to other methods.
+Despite these disadvantages, Neural Networks have found success in high level tasks where the model structure does not need to be understood.
+These include hand writing recognition, vehicle control, and face recognition [@Haykin2004].
+
+### SVM's
+
+SVM's are non-probabalistic learning algorithms that are typically used for binary classification, although multi-class and regression problems can be solved.
+It works by geometrically determining the best decision boundary, called a hyperplane, between two classes in multi-dimensional space [@Byun2002] .
+A hyperplane is simply a plane in more than 3 dimensions, i.e.
+more than 3 explanatory variables.
+The simplest type of SVM resembles Logistic Regression because the hyperplane is a linear boundary, and like Logistic Regression, the category of future cases depends on which side of the boundary they fall.
+
+The method for finding the boundary, however, is different to Logistic Regression.
+The aim is to find the hyperplane with the widest clear margin between data points from the two categories, while allowing a certain number of misclassifications (this is usually optimised for the problem).
+To do this, the equation for the hyperplane is expressed as follows:
+
+$\boldsymbol{w} \cdot \boldsymbol{x} + \beta_0 = 0$
+
+where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables.
+For two variables, $\boldsymbol{w}$ would be $\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$ [@Fletcher2009].
+Expanding the dot product of these two vectors gives the equation for a line:
+
+$\beta_1 X_1 + \beta_2 X_2 + \beta_0 = 0$
+
+The plane is expressed with vectors because they can be defined such that $\boldsymbol{w}$ is perpendicular to $\boldsymbol{x}$.
+This aids in optimisation calculations to find the hyperplane with the widest margin, i.e.
+two hyperplanes with the largest perpendicular distance between them and less than the allowable misclassified points [@Fletcher2009].
+The plot below demonstrates this concept and has no misclassifications:
 
@@
 
-The concept can be adapted to nonlinear boundaries by performing kernel transformations on prior variables as well as regression problems by introducing loss functions [@Auria2008].
+The hyperplanes at the boundaries of the margin are defined by the points closest to the margin.
+These points are called Support Vectors and during optimisation, these are the only points that matter in defining the final hyperplane position and direction [Auria2008].
+
+The SVM concept can be adapted to nonlinear relationships by performing kernel transformations on the explanatory variables [@Auria2008].
+There are several types of kernel transformations that can be applied, but the common aim is to add another dimension to the data points.
+This can enable a linear hyperplane to be found in multi dimensional space which separates the response variable classes better [@Fletcher2009].
+A simple example of this is shown below.
+The left plot shows two classes that cannot be cleanly separated with a plane.
+After a radial basis kernel transformation, the binary data in the right hand plot could be separated with a tilted linear plane:
+
+![Data set before and after a radial basis kernel transformation for an SVM model [@Fletcher2009]](images/lit_review/SVM_kernel.PNG)
+
 Advantages of SVM's include the ability to model nonlinear relationships in data, and that no assumptions about the distributions of the data must be made as the model is non-probabilistic.
@@
 On the other hand, the output is in the form of distance to the boundary (as opposed to probability) and results are not transparent (similar to Neural Networks) [@Caruana2006].
-Also, SVM's are known to be very slow to train and therefore often not suitable for industry purposes [@Auria2008].
+Also, SVM's can be very slow to train and therefore not suitable for industry purposes [@Auria2008].
 Because SVM's perform well in high-dimensional space and can form non-linear rules they have been applied to protein classification (medicine) and text and image recognition [@Byun2002].
@@
 Linear/Logistic Regression and Naive Bayes are computationally very fast with a simple concept that make some parametric assumptions about the variable relationships, but work well as baseline models against which to compare complex models.
-In certain cases these simpler models are significantly outperformed by complex models however it is important to compare both.
-Neural networks and SVM's are excellent machine learning predictors but are slow black box algorithms and are not appropriate for the effort estimation problem.
+In certain cases these simpler models are outperformed by complex models however it is important to compare both.
 
 Ensemble tree methods are capable of performing as well as Neural Networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
-Bayesian networks are excellent for visually presenting the structure and reasoning behind the predictive model, however are very slow to train and have generally been excluded from the machine learning literature in comparison to the other methods.
+Bayesian networks are excellent for visually presenting the structure and reasoning behind the predictive model, however are very slow to train and have been minimally included in the machine learning literature compared to the other methods.
 This may be due to low awareness of learned Bayesian Networks or lower performance.
+Neural Networks and SVM's are excellent machine learning predictors but are slow black box algorithms.
+Because it is so important to engage decision makers with a model that can explain its results, these two methods are not appropriate for the effort estimation problem.
 
@@
 This means the user would run a case through a model several times, each time altering one of the explanatory variable to test the degree of variation in the model predictions.
-Both strategies would give an idea of the certainty with which a model is predicting a response value or category.
+Both strategies would give an idea of the confidence with which a model is outputting response values.
 

### 2016-07-22 finalwriteup/trial_code_eq.rmd
finished lit review kerrie comments

@@
 
-![Plot of Gaussian probability density distributions of two response classes given x and y explanatory values [@Bulatov2010].](images/lit_review/NB_boundary2.JPG)
+$\boldsymbol{w} \cdot \boldsymbol{x} + \beta_0 = 0$
+
+where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables.
+For two variables, $\boldsymbol{w}$ would be $w\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $x\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$.
+Expanding the dot product of these two vectors gives the equation for a line:
+
+$\beta_1 X_1 + \beta_2 X_2 + \beta_0 = 0$

### 2016-07-25 finalwriteup/draft.rmd
started kerrie table

@@
 
-Ravi Kumar and Ravi [@Kumar2007] performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
+@Kumar2007 performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
 The most widely used model was Neural Networks, however Logistic and Linear Regression, decision trees, SVM's, discriminant analysis (DA), and statistical clustering techniques (such as K Nearest Neighbour) were also popular.
@@
 
-Case study projects with extreme 'return per dollar' values were likely the result of special scenarios which the model is not intended to predict.
-For example, many projects appeared wildly profitable but on closer inspection these were very small jobs.
-Outliers were initially defined as 'return per dollar' values sitting outside 1.5 x the interquartile range (IQR) from the upper and lower quartiles.
-This rule was initially practised by Tukey, the inventor of the box plot [@Tukey1977].
-All outlier projects determined by this method were reviewed with an employee from the case study company.
-It was concluded that outliers that were small jobs most likely required so little time that busy employees did not bother writing down their hours for that job.
-Other bigger outlier projects had unrealistically low hours logged against them which could have been the result of data entry errors or errors from when the data was transferred across databases.
-Based on the employee's domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined.
+Case study projects with extreme 'return per dollar' values could be the result of mistakes or special scenarios which the model is not intended to predict.
+Several methods could have been used for outlier detection including the distribution of the response variable, an analysis of high leverage cases, and cluster based methods.
+Guidelines for the first method were intially defined by Tukey, the inventor of the box plot [@Tukey1977].
+He recommended investigating cases with response variable values sitting outside 1.5 x the interquartile range (IQR) from the upper and lower quartiles.
+
+Alternatively, defining outliers in terms of high leverage requires solving for statistical measures such as Cook's distance or DFFITS for each case.
+Cook's Distance is the scaled change of all predicted values resulting from the deletion of an observation using linear or logistic regression [@Nurunnabi2009].
+In contrast, DFFITS measures the change in the predicted value of a single case when the case is left out of the regression model [@Nurunnabi2009].
+Cluster based outlier detection involves clustering your data via unsupervised learning, and inspecting the cases with the highest distance to the centroid of their cluster [@Schubert2014].
+For this project, Tukey's method was initially chosen so that the case study company would have a simple intuitive method to follow for excluding future outlier data points.
+A simple method was also appealing because the outliers would be reviewed with an employee so that their domain knowledge could interpret causes for outliers and refine the rules.
+
+Upon review of the outliers with an employee, many projects which appeared wildly profitable, were actually small jobs and high profitability ratios only indicated extra profits of a few hundred dollars.
+It was concluded that small job outliers most likely required so little time that busy employees did not bother writing down their hours for that job.
+Other bigger outlier projects had unrealistically low hours logged against them which could have been the result of data entry errors or errors from when the data was transferred across databases about 4 years prior to data extraction.
+Based on the employee's domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were revised.
+
 In summary, projects with 'return per dollar' values greater than 3 were removed and one project with a very low 'return per dollar' was removed because it contained incorrect invoiced amounts.
-
 26 of the 36 outlier cases with a high 'return per dollar' were small projects worth less than \$4,000.

### 2016-07-25 finalwriteup/trial_code_eq.rmd
started kerrie table

@@
 where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables.
-For two variables, $\boldsymbol{w}$ would be $w\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $x\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$.
+For two variables, $\boldsymbol{w}$ would be $\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$.
 Expanding the dot product of these two vectors gives the equation for a line:

### 2016-07-26 finalwriteup/draft.rmd
finished comparison table

@@
 
+### Summary Comparison between the Construction and IT Industry
+
+As previously stated, there are many parallels and some contrasts in cost estimation research in the Construction industry vs.
+the IT industry.
+These have been summarised into a table below, which is coloured according to similarities and differences.
+
+![Summary table of similarities and differences between cost estimation research in the IT and Construction industry](images/lit_review/compare_tbl.png)
+
 ## Statistical and Machine Learning Methods
@@
 categorical variables) [@Lunney1970].
-Since the method assumes variables are normally distributed, numeric variables must be normalised before analysis to achieve a statistical distribution as close to Gaussian as possible.
-Therefore a disadvantage of this method is that linear relationships and normally distributed variables do not necessarily represent real world data accurately [@Breiman2001b].
+It is good practice to normalise numeric variables before analysis to minimise effects of any high leverage data points.
+ANOVA assumes linear relationships between explanatory variables and the response variable, which can be a disadvantage as real world data is not necessarily linear [@Breiman2001b].
 
 In order to compare variable importance, two ANOVA outputs were assessed: the linear coefficients for each variable and the p-values for the coefficients [@Markham2016].
-The coefficient does not indicate importance relative to the other variables because its value reflects values within the variable.
+The coefficient does not indicate importance relative to the other variables because its magnitude reflects values within the variable.
 The sign of the coefficient however indicates whether the variable is directly or inversely related to the response variable.
@@
 P-values are a good indication of whether variables have a relationship with the target variable, however they do not necessarily rank importance.
-Precision of variable values can influence p-values, for example a variable with high precision will have a smaller p-value, whereas a variable measured roughly will have a higher p-value.
+Precision of variable values can influence p-values; for example a variable with high precision will have a smaller p-value, whereas a variable measured roughly will have a higher p-value.
 This does not necessarily make the more precisely measured variable more important than the other.
-It is also important to keep in mind that, with a dozen variables, chance alone can produce a variable with a p-value < 0.001 7% of the time [@Rice1989].
+It is also true that with a dozen variables, chance alone can produce a variable with a p-value < 0.001 7% of the time [@Rice1989].
 
@@
 The Random Forest algorithm can produce a permutation variable importance for each variable as part of its output.
-Variable importance is represented as a score, which is calculated by performing permutations on each variable in the test data for each tree, then re-running the test data through the trees.
-More specifically, consider a single variable.
-The values of this variable in the out-of-bag sample (the test data that the Random Forest algorithm partitions before building each tree) are randomly permuted (reordered) for each tree.
+Variable importance is represented as a score, which is calculated sequentially for each variable by comparing predictive accuracy between out-of-bag (OOB) data run through the trees without, then with a variable permutation (reordering).
+Since each tree is a bootstrapped sample of the complete dataset, approximately 1/3 of the data is held aside as the OOB sample for each tree.
+It is passed through its partner tree and produces an unbiased estimate of error and in this case, is also used to estimate variable importance.
+After the original OOB sample has been run through the trees, a variable is permuted in all OOB samples and are run through all of the trees again.
+
 The purpose of permuting is to mimic the absence of that variable.
-The trees in the Random Forest are then run again.
-Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted against the original Random Forest.
+Now the accuracy of the trees where the covariate has been permuted can be compared against the original Random Forest.
+For a categorical response variable, refer to the equation describing variable importance score below:
+
+$Raw\ VI(X_i) = \frac{\sum_{n = 1}^{ntrees} \left(C_{n} - C_{n,X_1} \right)}{ntrees}$
+
+Where
+
+$$
+\begin{aligned}
+Raw VI(X_i) &= \textrm{raw variable importance for variable}\ X_i \\
+ntrees &= \textrm{number of trees specified in the Random Forest} \\
+C_n &= \textrm{Number of correctly classified cases in the nth tree using the OOB sample} \\
+C_{n,X_1} &= \textrm{Number of correctly classified cases in the nth tree using the OOB sample with permuted} \\
+&\qquad \textrm{values for} X_i
+\end{aligned}
+$$
+
 If the accuracy of the permuted forest is much less than the original forest, then that variable was important.
-Permutation importance scores are calculated as the mean decrease in accuracy over all trees for the permuted variable in the model [@Breiman2005].
+As shown in the equation, permutation importance scores are calculated from the mean decrease in accuracy over all trees for the permuted variable in the OOB samples [@Breiman2005].
 
@@
 The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably.
-However this does not necessarily indicate the variable impacts the response variable the most.
-Overall, the permutation importance is useful to compare variables in a non-linear environment and can include multivariate interactions, but overestimates the importance of variables with with many categories and some numeric variables.
+However this does not necessarily indicate the variable predicts the response variable the most.
+Overall, the permutation importance is useful to compare variables in a non-linear environment and can include multivariate interactions, but overestimates the importance of variables with many categories and some numeric variables.
 
@@
 The predictor variable with the lowest p-value is then selected as the splitting variable [@Hothorn2006].
-Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable, re-running the forest and comparing the decrease in response variable accuracy.
+Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable in the OOB sample, re-running the forest and comparing the decrease in response variable accuracy.
 Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) and Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].
@@
 To compare the models, the root mean squared error (RMSE) statistic was used for regression models and the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to measure binary classification.
+
+The RMSE is a popular method for gauging numeric predictions and is represented by the equation below:
+
+$RMSE =  \sqrt{\frac{\sum_{t = 1}^{n} (\widehat{y}_t - y_t)^2}{n}}$
+
+Where
+
+$$
+\begin{aligned}
+n &= \textrm{number of data points} \\
+\widehat{y} &= \textrm{predicted values of the response variable} \\
+y &= \textrm{observed values of the response variable} \\
+\end{aligned}
+$$
 
@@
 Historically, in statistics this was called model averaging and since the 1990's research has shown that averaging results from different methods provides better predictive accuracy than any single model [@Madigan1994].
+
 New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big socially generated data.

### 2016-07-26 finalwriteup/draft.rmd
draft.rmd clash

--- 2016-07-25 finalwriteup/draft.rmd
@@
 When comparing Neural Networks to ensemble trees, Ravi Kumar and Ravi [@Kumar2007] found that Neural Networks and ensemble trees were both capable of out-performing the other, depending on the context.
-
-Their study also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
+For example in financial credit scoring, a study by @Brown2012 found that Random Forests and Boosted Trees consistently outperformed Neural Networks in classification.
+Logistic regression also outperformed Neural Networks which may align with a theory that most credit scoring datasets are only weakly non-linear [@Brown2012].
+This demonstrates how method performance can depend on the dataset and the problem.
+
+@Kumar2007's review also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
 For example, combining the contrasting advantages of Neural Networks and decision trees is a worthwhile ensemble technique.
 @Tsai2009's study combined these two methods after trialing each one.
-They were run successively with the Neural Network running first because it had a higher rate of prediction accuracy.
+They were run successively with the Neural Network running first because it had a higher predictive accuracy in this case.
 Then, to resolve the notorious lack of explanatory qualities, decision trees were employed.

### 2016-07-27 finalwriteup/draft.rmd
finished ading maths..

@@
 The MICE Random Forest method was chosen for imputation because it has been proven to work well with complex data sets [@Shah2014].
-It first performs a standard Random Forest imputation of missing values which are treated as initial 'place holders'.
-Random forest imputation is a complex process which works by first doing a rough imputation of the missing values.
+It first performs a standard Random Forest imputation of missing values in the full dataset which are treated as initial 'place holders'.
+The Random Forest imputation process will be explained in more detail later in this section.
+Next, the imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable.
+Random Forest is again used for the imputation.
+This is repeated for each variable and their initial 'place-holder' imputations are replaced by an imputation targeted at the single variable.
+This cycle is programmed to repeat five times for the full set of variables when using the 'mice' package, but can be increased [@Azur2011; @mice].
+
+Random forest imputation is a complex process which works by first doing a rough imputation of the missing values (using averages and majority votes).
 Then a ten-tree forest is run to calculate proximities between all cases.
-Proximities between two cases represent the number of trees in a forest where the two cases were assigned to the same end node in a tree.
-Once proximities are calculated, missing values for a variable (say variable 'm') are filled by an average of the other complete 'm' values weighted by the proximity of the case representing each 'm' value.
-
-After the initial Random Forest imputation is complete, the MICE process continues.
-The imputations for one variable are deleted and the remaining full variables are used to impute the missing values from the single variable again using Random Forest imputation.
-This is repeated for each variable and their 'place-holder' imputations are replaced by an imputation targeted at the single variable.
-This cycle is repeated five times by default [@Azur2011].
-
-The final imputed dataset was fed into a Boosted Tree algorithm and if similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
+Proximity between two cases is a measure of similarity and represents the number of trees in a forest where the two cases were assigned to the same end node in a tree.
+For each tree, the points can be run down the tree from the training or OOB sample.
+It can be represented by the equation below:
+
+$Px_{i,j}  = \frac{1}{10}\ \sum_{n = 1}^{10} I(E_{n,i} = E_{n,j})$
+
+Where
+
+$$
+\begin{aligned}
+I(\cdot)\  \textrm{is the indicator function} \\
+Px_{i,j} = \textrm{proximity between datapoints}\ i\  \textrm{and}\ j \\
+E_{n,i} = \textrm{is the end node that case}\ i\ \textrm{falls into in tree}\ n
+\end{aligned}
+$$
+
+Once proximities are calculated for each pair of datapoints, missing values for a variable (say variable 'm') are filled.
+This is done for each datapoint by averaging the other complete 'm' values weighted by the proximity of the case representing each 'm' value to the current point.
+
+The final imputed dataset, which was created using the mice package and Random Forest imputation, was fed into a Boosted Tree algorithm.
+If similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
 The imputed dataset was then trialed on the remaining methods and compared to un-imputed trials.
@@
 Six blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
-These included simple averaging of the individual model results, building a regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
-The first three models are simply the average, weighted average of the three individual model predictions.
-The last three methods take advantage of potentially meaningful interactions between an individual model's predictions and meta-features.
-For example, if the Boosted Tree model predicted profitability better than the other models for projects in the "Water" discipline, the blending model would take advantage of this interaction.
-Boosted tree predictions would be weighted higher than the other model predictions for Water projects.
+These included simple averaging of the individual model results, building a Logistic Regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
+Where possible, these methods shall be expressed as equations in an example problem with three original explanatory variables, $X_1$, $X_2$, $X_3$, and two full sets of predictions from individual predictive models, $M_1$ and $M_2$, with a response variable $Y$.
+Values in $M_1$ and $M_2$ are probabilities between 0 and 1.
+
+The first two models are simply the average, and weighted average of the three individual model predictions.
+
+Simple Average:
+
+$Y = \frac{1}{2} \cdot (M_1 + M_2)$
+
+Simple Logistic Regression:
+
+$Y = \beta_0 + \beta_1 M_1 + \beta_2 M_2$
+
+The third method builds Boosted Trees from the individual models, i.e.
+using $M_1$ and $M_2$ only to predict $Y$ in a Boosted Tree model.
+The last three methods take advantage of potentially meaningful interactions between the individual models' predictions and meta-features.
+For example, if the $M_1$ model predicted profitability better than $M_2$ for projects in category $a$ of $X_1$, the blending model would take advantage of this interaction.
+$M_1$ would be weighted higher than $M_2$ for projects where $X_1 = a$.
 
@@
 Then, each meta-feature is interacted with each set of model results [@Sill2009].
-A Linear Regression is performed to weight each term's contribution to the final predictive accuracy.
+A Logistic Regression is performed to weight each term's contribution to the final predictive accuracy and can be expressed for the example problem by the equation below:
+
+$$
+\begin{aligned} 
+Y &= \beta_0 + \beta_1 M_1 + \beta_2 M_2 + \beta_3 X_1 + \beta_4 X_2 + \beta_5 X_3 + \\
+\qquad \beta_6(M_1 \cdot X_1) + \beta_7(M_1 \cdot X_2) + \beta_8(M_1 \cdot X_3) +\\
+\qquad \beta_9(M_2 \cdot X_1) + \beta_10(M_2 \cdot X_2) + \beta_11(M_2 \cdot X_3)
+\end{aligned}
+$$
+
 Random Forests and Boosted Trees were also used to blend meta-features with output from the best models.
@@
 A split in a node determined by one variable is conditional upon the preceding split which was based on another variable and so on.
-
-The five blended model types were built on training and testing data sets multiple times in a similar fashion to the individual models.
+Random Forests and Boosted Tree models are difficult to express as equations, but for the example problem, $X_1$, $X_2$, $X_3$, $M_1$, and $M_2$ would all be included as explanatory variables to predict $Y$.
+
+The six blended model types were built on training and testing data sets multiple times using the same procedure used with the individual models.
 A single training set for a blended model must be built from the test results of the individual models.
@@
 Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the effect on a business' bottom line was examined.
-This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs.
+This was done using a profit curve - a chart that plots the change in profit the company earns on the y-axis vs.
 the probability threshold on the x-axis.
@@
 Therefore all profits and losses from jobs above the threshold were discounted.
+An equation defining the change in profit as a percentage of the original profit using threshold rejection is shown below:
+
+$\Delta\ Overall\ Profit\ \textrm{(%)} = \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100$
+
+Where
+
+$$
+\begin{aligned}
+I(\cdot) &= \textrm{the indicator function} \\
+N\ &= \textrm{the number of individual projects that are being included in the analysis} \\
+Pr_{p} &= \textrm{probability output from the algorithm for project}\ p \textrm{. Values are}\\
+& \qquad \textrm{betweeen 0 and 1 where 1 is loss making)} \\
+threshold &= \textrm{a chosen value between 0 and 1.}\ \Delta\ Overall\ Profit\ \textrm{is calculated for} \\
+& \qquad \textrm{several}\ threshold\ \textrm{values which defines the profit curve} \\
+Profit_p &= \textrm{profit for individual project}\ p
+\end{aligned}
+$$
+
 If the threshold was zero, all jobs were rejected and the profit would be $0.

### 2016-07-27 finalwriteup/trial_code_eq.rmd
finished comparison table
finished ading maths..

--- 2016-07-19 finalwriteup/trial_code_eq.rmd
@@
 
-If conditional independence is assumed, the equation becomes:
+Simple Average:
 
-$p(yes|X_1 X_2) = p(X_1|yes) \cdot p(X_2|yes) \cdot p(yes)$
+$Y = \frac{1}{2} \cdot (M_1 + M_2)$
 
-![Plot of Gaussian probability density distributions of two response classes given x and y explanatory values [@Bulatov2010].](images/lit_review/NB_boundary2.JPG)
+Simple Logistic Regression:
+
+$Y = \beta_0 + \beta_1 M_1 + \beta_2 M_2$
+
+Six blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
+These included simple averaging of the individual model results, building a Logistic Regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
+To explain further, these methods shall be expressed as equations in a problem with three original explanatory variables, $X_1$, $X_2$, $X_3$, and two full sets of predictions from individual predictive models, $M_1$ and $M_2$, with a response variable $Y$.
+Values in $M_1$ and $M_2$ are probabilities between 0 and 1.
+
+The first two models are simply the average, and weighted average of the three individual model predictions.
+
+Simple Average:
+
+$Y = \frac{1}{2} \cdot (M_1 + M_2)$
+
+Simple Logistic Regression:
+
+$Y = \beta_0 + \beta_1 M_1 + \beta_2 M_2$
+
+The third method builds Boosted Trees from the individual models, i.e.
+using $M_1$ and $M_2$ only to predict $Y$ in a Boosted Tree model.
+The last three methods take advantage of potentially meaningful interactions between the individual models' predictions and meta-features.
+For example, if the $M_1$ model predicted profitability better than $M_2$ for projects in category $a$ of $X_1$, the blending model would take advantage of this interaction.
+$M_1$ would be weighted higher than $M_2$ for projects where $X_1 = a$.
+
+A simplified explanation of FWLS is as a Linear Regression where meta-features as well as model predictions from individual models are included as explanatory variables.
+Then, each meta-feature is interacted with each set of model results [@Sill2009].
+A Logistic Regression is performed to weight each term's contribution to the final predictive accuracy and can be expressed for the example problem by the equation below:
+
+$$
+\begin{aligned} 
+Y &= \beta_0 + \beta_1 M_1 + \beta_2 M_2 + \beta_3 X_1 + \beta_4 X_2 + \beta_5 X_3 + \\
+\qquad \beta_6(M_1 \cdot X_1) + \beta_7(M_1 \cdot X_2) + \beta_8(M_1 \cdot X_3) +\\
+\qquad \beta_9(M_2 \cdot X_1) + \beta_10(M_2 \cdot X_2) + \beta_11(M_2 \cdot X_3)
+\end{aligned}
+$$
+
+Random Forests and Boosted Trees were also used to blend meta-features with output from the best models.
+Feature interaction is performed passively due to the nature of how trees are built.
+A split in a node determined by one variable is conditional upon the preceding split which was based on another variable and so on.
+Random Forests and Boosted Tree models are difficult to write as equations, but for the example problem, $X_1$, $X_2$, $X_3$, $M_1$, and $M_2$ would all be included as explanatory variables to predict $Y$.

### 2016-07-27 finalwriteup/draft.rmd
abstract

@@
 
-* why did you do the study?, what is the problem
-performing loss-making jobs for fixed fees is a major problem in the consulting, particularly in the competitive construction industry.
+Engaging in loss-making jobs for fixed fees is a major problem in consulting[,]
+"businesses."
+"This is especially" 
+[particularly] 
+"true" in the competitive construction industry. This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help a 
+[company] avoid loss-making jobs or 
+"facilitate improved commercial strategy." 
+[tactfully enforce stricter contracts.] 
+It was found that in a specific decision framework, a case study's profits could be improved 9% by avoiding approximately 4% of projects.
+"THIS IS AWESOME!!!! WHERE CAN I GET ONE?"
+Alternative decision frameworks are also 
+"proposed and evaluated" 
+[debated and proposed].
+The internal data collected by the case study company described each of their projects in terms of employee positions, fraction of project time completed by each employee, client characteristics, project characteristics (timespan, hours, technical details) and invoicing history.
+Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naïve Bayes, and Bayesian Networks were applied as well as blended combinations of these methods.
+A decision scenario
+[that] 
+"which" rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
+This process was repeated on hundreds of models built from different training subsets of the data in order to obtain a 95% statistical confidence interval of profit improvements.
+The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% profit improvements.
+Further work is recommended to test user input estimation and user interface designs.
+The findings from this research
+"has the potential to" 
+[could] assist managers in reducing losses by highlighting risky projects and 
+"guiding" 
+[facilitating specific changes] "improvements" to their fee structures.
 
 * why is it important?
-affects the porfitability of a company which impacts the careers, internal morale, and even viability of a company.
+affects the porfitability of a company which impacts the careers, company morale, and even viability of a company.
 
 * why should someone read your essay?
-learn if a company's passively collected data could help them avoid loss making jobs or tactfully enforce strictier contracts.
 
 * what is your main claim?
-improve profitability by 9% by avoiding ~4% of projects is a possible way to interpret the agorithm output. other ways are possible.
 
 * discuss variables and your approaches
-variables: main employee, % time by each type of employee, team size, client charactersitcs, project characteristics (size, type of work, length)
-approaches: complex data with mnay variables, thousands of data points. tried regression as well as machine learning algs: randomf roest and boosted trees, bayesian networks, naive bayes.
 
@@
 the 95% confidence interval across all cross validated trials was __ to __ profit improvement.
-
-* overview of most important sources
 
@@
 As stated, the aim of this project is to use statistical and machine learning techniques to model the profitability of projects.
-Several techniques will be tested and this section reviews a range of predictive algorithms that have been successfully applied to business and social data.
+Several techniques will be tested and this section reviews a range of predictive algorithms that have been applied to business and social data.
 These vary from simple methods such as Linear Regression to complex, deep learning Neural Networks.
-The review of methods applied to the construction and IT industry highlighted the use of Linear Regression, Neural Networks, SVM's and in one case Boosted Trees for prediction, however research on other business problems utilised a wider range of methods.
-These include Naive Bayes, Random Forests, and machine learned Bayesian Networks.
-The following section presents each method's advantages and disadvantanges and an example of a successful application.
+The review of methods applied to the construction and IT industry highlighted the use of Linear Regression, Neural Networks, SVM's and in one case Boosted Trees, however research on other business problems utilised a wider range of methods.
+These included Naive Bayes, Random Forests, and machine learned Bayesian Networks.
+The following section presents an explanation fo each method, their advantages and disadvantanges and an example of a successful application.
 
@@
 Linear Regression is one of the simplest and most popular statistical prediction methods.
-An equation for a line is learned, which is defined by a constant and coefficients multiplied by each explanatory variable which are assumed to be independent.
+An equation for a line is learned, which is defined by a constant and coefficients multiplied by each explanatory variable.
 This can be mathematically represented by the equation below:
@@
 
-where $X_{j}$ represents the jth input variable, $\beta_{j}$ is the coefficient or parameter being solved for, and $\beta_{0}$ is the y-intercept.
+where
+
+$$
+\begin{aligned}
+X_{j}\ \textrm{represents the}\ j\textrm{th input variable} \\
+\beta_{j}\ \textrm{is the coefficient or parameter being solved for} \\
+\beta_{0}\ \textrm{is the y intercept}
+\end{aligned}
+$$
+
 The most popular method to find the coefficients is to minimise the residual sum of squares (ordinary least squared method).
@@
 The reason for this method's popularity is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
-The integration of categorical variables is called Analysis of Variance (ANOVA) Linear Regression, and the algorithm compares group mean variances within the categorical variables.
-
-If structured correctly, linear regression results can show the importance of the explanatory variables.
-The variables must be standardised to one another (i.e.
-scaled to the same mean and standard deviation), and then the coefficients (the slopes) indicate how *much* a variable changes with the response variable.
-This relationship can then be compared against the other explanatory variables [@Putler2012].
+The integration of categorical variables is called Analysis of Variance (ANOVA) Linear Regression, which compares group mean variances within the categorical variables.
+
+If structured correctly, Linear Regression results can show the importance of the explanatory variables.
+The variables must be first standardised to one another (i.e.
+scaled to the same mean and standard deviation).
+Then, their coefficients can be directly compared [@Putler2012].
+The magnitude of the coefficients (the slopes) indicate how *much* a variable changes with the response variable.
 However, if two or more predictor variables are correlated, the values of their coefficients are unstable.
-More specifically, this means that if the a model has two correlated variables, and is re-run with a slightly different set of data, the coefficients for the two variables may be quite different because the assumption of independent variables is not true.
+More specifically, this means that if the a model has two correlated variables, and is re-run with a slightly different set of data, the coefficients for the two variables may be quite different.
+This is because explanatory variables are assumed to be independent, and if variables are correlated, it breaks this assumption.
 On the other hand, correlated variables do not impact the predictive accuracy of the model [@Putler2012].
 
-In linear regression, no assumptions are made about the distributions of the explanatory variables or response variable.
+In linear regression, no assumptions are made about the distributions of the variables.
 Skewed explanatory variable distributions are acceptable but it is generally better to normalise them by, for example, taking the logarithm or square root of the variable.
@@
 Unfortunately many real world world phenomena do not have linear relationships which can make it difficult to produce accurate results using this method [@Breiman2001b].
-There are methods to overcome this limitation such as including polynomial terms and interaction terms.
+There are methods to model non-linear relationships such as including polynomial terms and interaction terms.
 For example, in the linear equation, as well as $X_{1}$, include the terms $X_{2} = X_1^2$ and $X_{3} = X_1^3$ to build a third degree polynomial.
-Interactions between explanatory variables can be modelled by including a term such as $X_5$ where $X_5 = X_4*X_1$.
+Interactions between explanatory variables can be modeled by including a term such as $X_5$ where $X_5 = X_4*X_1$.
 This versatility is advantageous, however it can be challenging to know which non-linear terms of which variables to include.
@@
 
-The ordinary least squared regression method described in the previous section cannot be applied to binary classification problems, where the objective is to predict a probability of an event occurring.
+The ordinary least squared regression method described in the previous section cannot be applied to binary classification problems, where the objective is to predict the probability of an event occurring.
 Probability is a continuous response variable, however predictions from a linear regression would not be bound by 0 and 1.
-Instead, predicted values could reach further along the line outputting probabilities greater than 1 and less than 0 which is difficult to interpret.
-Also, residuals from linear regression should be normally distributed, but this is not possible when the response variable consists of only two values (**refappstate**).
-
-Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable (1 or 0).
-First, a sigmoid function is assumed as the functional form of the probabilties against variable X.
+Instead, predicted values could reach further along the fitted line, outputting probabilities greater than 1 and less than 0 which is difficult to interpret.
+Also, residuals from linear regression should be normally distributed, but this is not possible when the response variable consists of only two values [@Lowry2016].
+
+Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable (represented by 1 or 0).
+First, a sigmoid function is assumed as the functional form of the probabilties against an example variable X.
 This is because it is pragmatic and works well: sigmoid functions can be bound by 0 and 1, and a logit transformation of the probability results in a linear relationship between the log odds and X.
@@
 The aim is to find which category has a higher probability, so to simplify calcualtions, the denominator can be dropped.
-Because it is the same for both equations, the denominator does not affect the final ratio of probability of 'yes' to probability of 'no'.
-Integrating the assumption of conditional independence, the equations can be written as follows:
+Because the denominator is the same for both equations, it does not affect the final ratio of the probability of 'yes' to the probability of 'no'.
+Also integrating the assumption of conditional independence, the equations can be written as follows:
 
@@
 
-These equations are simple to compute given the data, and the class with the highest probability is chosen [@Provost2013].
+These equations are simple to compute given the data, and the class with the highest probability can then be chosen [@Provost2013].
 
@@
 The method can perform well in real world tasks because the assumption of independence does not significantly damage predictions.
-If multiple variables are actually related, the variables separately still direct the prediction in the correct direction.
+If multiple variables are actually related, the variables still separately direct the prediction in the correct direction.
 The correlated variables will double or triple their emphasis on the predicted response variable, however this does not necessarily impact classification accuracy because the final step is to choose the class with the *highest* probability [@Provost2013].
 This is fine for ranking but the output probabilities are not accurate statistical probabilities [@Caruana2006].
-Another disadvantage is that the distribution of the numeric variable must be assumed (often Gaussian) in order to calculate the liklihood terms which the data may not fit into neatly [@Caruana2006].
-
-The Gaussian probability densities of two response classes can be visualised on a bivariate feature space.
+Another disadvantage is that the distribution of numeric variables must be assumed (often Gaussian) in order to calculate the liklihood terms, and the data may not fit into these distributions neatly [@Caruana2006].
+
+Example Gaussian probability densities of two response classes can be visualised on a bivariate feature space.
 The two normal distrubtions represent the probability of each class given values of the two explanatory variables (x and y axes):
@@
 Naive Bayes was used in a complex spam detection system where new spam emails or toxic text themes could be quickly added to the filtering model.
-Although the example is quite different to cost estimation, the Naive Bayes method generally provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
+Although the given example is different to cost estimation, the Naive Bayes method generally provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
 
@@
 Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
-Single decision trees create intuitive rules that a decision maker can follow in real-life scenarios but on the other hand, they tend to overfit the data and provide low predictive accuracy [@Kabra2011][@Putler2012].
+Single decision trees create intuitive rules that a decision maker can follow in real-life scenarios but also tend to overfit the data and provide low predictive accuracy [@Kabra2011][@Putler2012].
 Ensemble trees were developed to solve this problem and several ensemble techniques are described in this section.
@@
 Decision trees have a different theoretical foundation to traditional prediction methods such as regression.
-They are created by a progression of binary splits of the dataset based on response variable values.
+They are created by a progression of binary splits of the dataset that attempt to group similar response variable values.
 Before a split is made, each explanatory variable and each value within the explanatory variables is tested for how 'purely' it partitions the response variable into two subsets.
@@
 This similarity is called 'purity' and can be measured using several formulas.
-An example for regression is:
-
-$SSN - (SSL + SSR)$
+An example for regression (a problem with a numeric response variable) is:
+
+$Purity = SSN - (SSL + SSR)$
+
+and
 
@@
 
-where SSN is the sum of squared errors of a node (dataset before the split).
-$y_i$ are the response variable values and $\bar{y}$ is the mean response variable value in the node.
-SSL and SSR are the sum of squared errors in the left and right datasets that resulted from the split.
-Other purity formulas exist for categorical response variables such as Gini purity and information gain [@Hastie2009].
-After the first split is made on the dataset, each subset is split by another variable (or the same explanatory variable again) and so on resulting in final subsets of data with similar response variable values.
+where
+
+$$
+\begin{aligned}
+SSN\ \textrm{is the sum of squared errors of a node (dataset before the split)} \\
+y_i\ \textrm{is the}\ i \textrm{th response variable value} \\
+\bar{y}\ \textrm{is the mean response variable value in the node} \\
+SSL\ \textrm{is the sum of squared errors in the left dataset that resulted from the split} \\ 
+SSR\ \textrm{is the sum of squared errors in the right dataset that resulted from the split} 
+\end{aligned}
+$$
+
+Other purity formulas also exist for categorical response variables such as Gini purity and information gain [@Hastie2009].
+After the first split is made on the dataset, each subset is split by another variable (or the same explanatory variable again) and so on resulting in final subsets of data with similar response variable values (high purity).
 If this process is visualised graphically, it is clear that non-linear relationships can be captured.
@@
 Once the splitting rules have been learned, a new case is run through the tree from the top.
-It follows the path provided by which side of the binary split its explanatory variable values fall under and ends up in a final node.
-The mean response variable value from the initial data that ended up in that node is the predicted response value for the new case.
+It follows the path provided by which side of the binary split its explanatory variable values fall under and finds its way to an end node.
+The mean response variable value from the training data that ended up in that node is the predicted response value for the new case.
 Decision trees are easy to overfit on the training data.
-This means that if the splitting rules are continued until no more splits can be possibly made, when the test data is run through the tree, the predictions will be too closely aligned to the exact structure of the training dataset.
+This means that if the splitting rules are continued until no more splits can be possibly made, when the test data is run through the tree, the predictions will be too closely aligned to the exact structure of the training data.
 Prediction rules need to generalise trends found in the training data.
-This can be managed in a single tree by using 'stopping rules' such as a minimum number of data in a final node or a minimum purity increase for a split to be valid (**ref**).
+This can be managed in a single tree by using 'stopping rules' such as a minimum number of data in a final node or a minimum purity increase for a split to occur [@Perlich2003].
 
 As stated previously, single decision trees also suffer from low predictive accuracy and instability, so to combat these problems, ensemble tree methods were pioneered in the 1990's [@Breiman1996].
-Three examples of these methods are boostrap aggregating (bagging), Random Forests, and boosting.
+Three examples of these methods are boostrap aggregating (bagging), Random Forests, and Boosted Trees.
 Bagging builds multiple trees from different training sample data sets (wich are sampled from the full data set *with* replacement) and then combines tree results [@Breiman1996].
@@
 
-The boosted decision tree approach applies gradient descent theory to a series of decision trees.
+Lastly, the boosted decision tree approach applies gradient descent theory to a series of decision trees.
 The trees are limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) of the preceding tree.
-By modeling the errors, misclassified cases are weighted higher than correctly classified cases and influence the structure of the tree more [@Hastie2009].
+By modeling the errors, misclassified cases are weighted higher than correctly classified cases and influence the structure of the latest tree more [@Hastie2009].
 This increased weighting on errors is why the algorithm is called boosting.
 
-This idea can be visualised in a simple scenario consisting of a single explanatory variable (shown on x-axis) and response variable (y-axis).
+The boosting idea can be visualised in a simple scenario consisting of a single explanatory variable (shown on the x-axis) and response variable (y-axis).
 Each tree is limited to one split, and the first tree is shown below, where the vertical line is the splitting point:
@@
 
-The horizontal lines indicate the mean response variable value in each subset after the split.
+The horizontal lines indicate the mean response variable value in the left and right subset after the split.
 It is the value the first tree would predict for data that falls into each subset.
 The next tree is built on the residuals of the preceding tree.
-
 ![Second tree modeling residuals in short progression of boosted trees [@Ihler2012].](images/lit_review/boost_tree2_.png)
@@
 
-The advantages of ensemble tree methods are numerous.
+Further advantages of ensemble tree methods are numerous.
 Similarly to Linear Regression, the predictor variables in trees do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
 However, it is often wise to transform skewed distributions to achieve more even segmentation.
-Additionally, ensemble tree methods are faster than SVM's and Neural Networks but perform just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
-Random forests in particular have the benefit of being unaffected by noise in the data and has the capacity to handle a large number of variables [@Sealfon2012].
-Caruana [@Caruana2006] tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
+Additionally, ensemble tree methods are faster than SVM's and Neural Networks but can perform just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
+Random Forests in particular have the benefit of being unaffected by noise in the data and have the capacity to handle a large number of variables [@Sealfon2012].
+@Caruana2006 tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
 This demonstrates how ensemble tree methods are capable of competing with high level machine learning algorithms.
@@
 For example, statistical confidence intervals for predictions are not available for standard ensemble methods [@Louppe2014].
-This will be addressed again in the summary for this section.
-Also, variable importance tables provided by Random Forests and Boosted Trees can be biased towards variables with many categories and another type of ensemble method, conditional forests, should be used [@Radenkovic2010].
-The methods can also be prone to overfitting if not tuned carefully [@Louppe2014][@Caruana2006].
+Solutions for this problem will be addressed in the summary of this section as the problem occurs in several methods.
+Also, variable importance tables can be output by Random Forests and Boosted Trees, but are often biased towards variables with many categories.
+Another type of ensemble method, conditional forests, should be used to check variable importance [@Radenkovic2010].
+Finally, the methods can also be prone to overfitting if not tuned carefully [@Louppe2014][@Caruana2006].
 
@@
 The model is visually represented by a directed acyclic graph (DAG) and is capable of linking the conditional dependency between any variable to another variable.
-The links are graphically represented by arcs or edges [@Heckerman1998].
-This kind of conditional relationship is Bayesian, where the probabilities in one node are conditional upon values in nodes directed to it as well as preceding nodes.
+arcs or edges represent links in the DAG [@Heckerman1998].
+The conditional relationships are Bayesian, where the probabilities in one node are conditional upon values in nodes directed towards it as well as preceding nodes.
 Refer to the simple example below:
@@
 
-If each variable, A, B, C, and D are binary categories, there are $2^4$, or 16 combinations of results and each variable is linked to each of the other variables with a conditionally dependency.
-The probability of each combination could be calculated as per below using the chain rule for probabilities:
+If each variable, A, B, C, and D are binary categories, there are $2^4$, or 16 unique possible combinations of results and each variable is linked to each of the other variables with a conditionally dependency.
+The probability of each combination could be calculated using the chain rule for probabilities:
 
@@
 
-However, if it is known through domain knowledge or can be reasoned through the data that some variables are conditionally independent from one another, the calculations can be simplified [@Barber2012].
+However, domain knowledge or reasoning through the data could indicate that some variables are conditionally independent from one another.
+Then the calculations could be simplified [@Barber2012].
 For example, if B and D, A and D, and A and C are conditionally independent, the DAG would be:
@@
 
-In this way, Bayesian Networks reduce computations required to find the probability of an event given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes cannot [@Barber2012].
-Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes though (as Naive Bayes is the simplest form of a Bayesian Netwoork) and variables with many categories dramatically increases the expense [@Zhang2004].
+In this way, Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes, for example, cannot [@Barber2012].
+Also, some equations for a comination are now clones of another combination and do not require recalculation.
+Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes (as Naive Bayes is the simplest form of a Bayesian Netwoork) and variables with many categories dramatically increases the expense [@Zhang2004].
 
@@
 These demand sampling from conditional density models or complex algebra whereas categorical variables require comparitively simple calculations [@Hofmann1996].
-It is therefore common to discretise continuous variables [@Kragt2009].
+It is therefore common to discretise continuous variables which simultaneously decreases precision of that variable [@Kragt2009].
 
 Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.
-They are excellent at combining knowledge from different origins, and then visually communicating results well to decision makers [@Kragt2009].
-Therefore Bayesian Networks have become an excellent decision support tool.
+Conditional probability dependencies can be calculated from different data sources and separately input into the network.
+They are therefore excellent at combining knowledge from different origins, and then visually communicating results to decision makers [@Kragt2009].
 Another benefit of Bayesian Networks as decision support tools is they allow the user to fix certain variable values, re-run the network and observe changes in probabilities.
-Thisis a kind of sensitivity testing that gives users a feel for whether the model makes sense and the degree of uncertainty.
-Popular applications of Bayesian Networks are modelling uncertainty in natural resource management and modelling complex business network structures such as airports [@Wu2013].
+This is a kind of sensitivity testing that gives users a feel for whether the model makes sense.
+Popular applications of Bayesian Networks are modeling uncertainty in natural resource management and modeling complex business network structures such as airports [@Wu2013].
 
@@
 Algorithmic Neural Networks were inspired by Neural Networks of the human nervous system.
-Biological Neurons recieve electro-chemical input from the terminals of adjacent neurons into its dendrites.
-If enough electro-chemical input is recieved, the neuron will fire, which means a signal will be sent from the dendrite end, through the axon to its terminal.
+In the human body, biological neurons recieve electro-chemical input into their dendrites from the terminals of adjacent neurons.
+If enough electro-chemical input is recieved, the neuron will 'fire', which means a signal will be sent from the dendrite end, through the axon to its terminal.
 Its terminal then sends signals to adjacent dendrites of other neurons and the process propagates.
 Neurons are able to learn over time by strengthening the connections of adjacent neurons that frequently cause the next one to fire.
-This means the required firing threshold for the recieving neuron lowers [@Putler2012].
+This means the required firing threshold for the recieving neuron lowers in strengthened connections [@Putler2012].
 
@@
 
-In fact, a wide variety of neural network structures and dynamics exist so a simple popular version is explained here.
+A wide variety of neural network structures and dynamics exist, so a simple popular version is explained here.
 The first input layer is passive, and each input variable is passed through one input node.
-These nodes are passive and the variable values are simply passed to each of the nodes in the next layer, called a hidden layer.
+These nodes simply pass the input variable values to each of the nodes in the next layer, called a hidden layer.
 Neural networks may have any number of hidden layers with any number of nodes as selected by the analyst.
 Often however, a single hidden layer is adequate.
-Calibration for predictions is done in the hidden nodes.
-At these nodes, each input variable is weighted then combined (similar to linear regression) to produce a single value.
+Calibration for predictions is done in the hidden nodes, where each input variable is weighted then combined (similar to linear regression) to produce a single value.
 This is then transformed and assessed by a threshold function to determine if the hidden node should 'fire' to the next nodes.
-Once the signals and values have propogated to the final layer they are recombined to produce an output, or a prediction [@Putler2012].
+Once the signals and values have propogated to the final layer they are recombined to produce an output value, or a prediction [@Putler2012].
 
@@
 Then optimisation techniques are used to adjust the weights and reduce the error.
-This process can result in overfitting, so stopping rules such as regularisation strength should be used [@Karpathy2016].
+This process can result in overfitting, so stopping rules such as regularisation strength should be used [@Karpathy2016], however these details are outside the scope of this review.
 
@@
 SVM's are non-probabalistic learning algorithms that are typically used for binary classification, although multi-class and regression problems can be solved.
-It works by geometrically determining the best decision boundary, called a hyperplane, between two classes in multi-dimensional space [@Byun2002] .
+The method works by geometrically determining the best decision boundary, called a hyperplane, between two classes in multi-dimensional space [@Byun2002] .
 A hyperplane is simply a plane in more than 3 dimensions, i.e.
@@
 This aids in optimisation calculations to find the hyperplane with the widest margin, i.e.
-two hyperplanes with the largest perpendicular distance between them and less than the allowable misclassified points [@Fletcher2009].
+two hyperplanes with the largest perpendicular distance between them that contains less than the allowable number of misclassified points [@Fletcher2009].
 The plot below demonstrates this concept and has no misclassifications:
@@
 This section reviewed the advantages and disadvantages of statistical and machine learning techniques that have been successfully applied to business problems.
-These included Linear Regression, Logistic Regression, Naive Bayes, decision trees and ensemble trees, Bayesian Networks, Neural Networks, and SVM's.
+These included Linear Regression, Logistic Regression, Naive Bayes, Decision Trees and Ensemble Trees, Bayesian Networks, Neural Networks, and SVM's.
 Linear/Logistic Regression and Naive Bayes are computationally very fast with a simple concept that make some parametric assumptions about the variable relationships, but work well as baseline models against which to compare complex models.
@@
 
-Ensemble tree methods are capable of performing as well as Neural Networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
-Bayesian networks are excellent for visually presenting the structure and reasoning behind the predictive model, however are very slow to train and have been minimally included in the machine learning literature compared to the other methods.
+Ensemble Tree methods are capable of performing as well as Neural Networks and SVM's and make no assumptions about the structure of the data, but provide insights into the model such as variable importance and variable relationships.
+Bayesian Networks are excellent for visually presenting the structure and reasoning behind the predictive model, however are very slow to train and have been minimally included in the machine learning literature compared to the other methods.
 This may be due to low awareness of learned Bayesian Networks or lower performance.
@@
 
-None of the methods output statistical uncertainty or a confidence interval associated with each prediction.
+None of the methods output statistical uncertainty or statistical confidence interval associated with each prediction.
 Understanding uncertainty should always be sought in decision making be it from a confidence interval or sensitivity testing.
@@
 Another way for a decision maker to get a feel for uncertainty is to perform sensitivity testing on a prediction.
-This means the user would run a case through a model several times, each time altering one of the explanatory variable to test the degree of variation in the model predictions.
+This means the user would run a case through a model several times, each time altering one of the explanatory variables to test the degree of variation in the model predictions.
 Both strategies would give an idea of the confidence with which a model is outputting response values.
 
-In conclusion, simple models such as Linear/Logistic Regression and Naive Bayes are good baseline models to implement, while ensemble tree methods and Bayesian Networks may handle bigger unstructured data better than the simpler models while providing insight to decision makers.
+In conclusion, simple models such as Linear/Logistic Regression and Naive Bayes are good baseline models to implement, while Ensemble Tree methods and Bayesian Networks may handle bigger unstructured data better than the simpler models while providing insight to decision makers.
 

### 2016-07-27 finalwriteup/trial_code_eq.rmd
abstract

--- 2016-07-25 finalwriteup/trial_code_eq.rmd
@@
 
-If conditional independence is assumed, the equation becomes:
-
-$p(yes|X_1 X_2) = p(X_1|yes) \cdot p(X_2|yes) \cdot p(yes)$
-
 $\boldsymbol{w} \cdot \boldsymbol{x} + \beta_0 = 0$
@@
 where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables.
-For two variables, $\boldsymbol{w}$ would be $\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$.
+For two variables, $\boldsymbol{w}$ would be $\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$ [@Fletcher2009].
 Expanding the dot product of these two vectors gives the equation for a line:

### 2016-07-28 finalwriteup/draft.rmd
finished kerrie editing

@@
 ---
-title: "draft 1"
+title: "Profitability Prediction of Consulting Projects in the Construction Industry"
 author: "Amy Cook"
@@
 bibliography: Library.bib
+csl: apa.csl
 ---
@@
 
-* Abstracts allow readers to quickly decide whether the research is relevant to their own interests
-* get at your main argument quickly
-* explain the purpose, goal methods, and results
-
-Engaging in loss-making jobs for fixed fees is a major problem in consulting[,]
-"businesses."
-"This is especially" 
-[particularly] 
-"true" in the competitive construction industry. This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help a 
-[company] avoid loss-making jobs or 
-"facilitate improved commercial strategy." 
-[tactfully enforce stricter contracts.] 
+Engaging in loss making jobs for fixed fees is a major problem in consulting, particularly in the competitive construction industry.
+This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs tactfully choose when to enforce stricter contracts.
 It was found that in a specific decision framework, a case study's profits could be improved 9% by avoiding approximately 4% of projects.
-"THIS IS AWESOME!!!! WHERE CAN I GET ONE?"
-Alternative decision frameworks are also 
-"proposed and evaluated" 
-[debated and proposed].
-The internal data collected by the case study company described each of their projects in terms of employee positions, fraction of project time completed by each employee, client characteristics, project characteristics (timespan, hours, technical details) and invoicing history.
-Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naïve Bayes, and Bayesian Networks were applied as well as blended combinations of these methods.
-A decision scenario
-[that] 
-"which" rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
+Alternative decision frameworks are also proposed and evaluated.
+The internal data collected by the case study company described each of their projects in terms of employee positions, fraction of project time completed by each employee, client characteristics, project characteristics (time span, hours, technical details) and invoicing history.
+Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods.
+A decision scenario which" rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
 This process was repeated on hundreds of models built from different training subsets of the data in order to obtain a 95% statistical confidence interval of profit improvements.
@@
 Further work is recommended to test user input estimation and user interface designs.
-The findings from this research
-"has the potential to" 
-[could] assist managers in reducing losses by highlighting risky projects and 
-"guiding" 
-[facilitating specific changes] "improvements" to their fee structures.
-
-* why is it important?
-affects the porfitability of a company which impacts the careers, company morale, and even viability of a company.
-
-* why should someone read your essay?
-
-* what is your main claim?
-
-* discuss variables and your approaches
-
-* describe evidence to support your claim
-hundred of cross validated trials using the 5 methods, as well as 5 blending methods.
-obtained a 95% confidence interval for each method and simple logistic regression of three models outperformed the others, closely followed by simple averaging.
-the 95% confidence interval across all cross validated trials was __ to __ profit improvement.
-
-* Results: answer , was hypothesis supported
-hypothesis was supprted: CAN improve bottom line of business by mobilising their passively collected data into mL algorhtms.
-
-* implications of your work, meaning
-more work to be done on testing the user interface and user ailities further.
-could assist managers in making better decisions about which projects they should reject or develop a more careful contract structure to improve profits.
+The findings from this research have the potential to assist managers in reducing losses by highlighting risky projects and guiding project-based changes to fee structures.
 
@@
 
-**BIM** - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
+**BIM** - Building Information modeling. A digital representation of physical characteristics of an object or structure.
+
+**CBR** - Case based reasoning. A method of estimating by using the results of similar cases.
 
@@
 
-**CBR** - Case based reasoning. A method of estimating by using the results of similar cases.
-
 **CRM** - Customer Relationship Management.
-Refers to a common type of software used by businesses to record client and project details as well as employee timesheet records.
-Generally accessable to each employee in a company.
+Refers to a common type of software used by businesses to record client and project details as well as employee time sheet records.
+Generally accessible to each employee in a company.
 
@@
 
-**FP** - False positive. A case marked 'negative' in a dataset is incorrectly predicted 'positive' by an algorithm.
+**FP** - False positive. A case marked 'negative' in a data set is incorrectly predicted 'positive' by an algorithm.
 
@@
 
-**FN** - False negative. A case marked 'positive' in a dataset is incorrectly predicted 'negative' by an algorithm.
+**FN** - False negative. A case marked 'positive' in a data set is incorrectly predicted 'negative' by an algorithm.
 
 **FWLS** - Feature weighted linear stacking.
-A linear combination of model results which are interacted with original variables.
+A linear combination of model results that is interacted with original variables.
 This allows certain models to be weighted heavier for certain values of the original features.
-
-**IT** - Information technology. The use of computers for storing, retrieving and processing information.
 
@@
 
+**IT** - Information technology. The use of computers for storing, retrieving and processing information.
+
 **RMSE** - Root mean squared error. The square root of the mean of the ESS.
@@
 
-**TN** - True negative. A case marked 'negative' in a dataset is correctly predicted 'negative' by an algorithm.
+**TN** - True negative. A case marked 'negative' in a data set is correctly predicted 'negative' by an algorithm.
 
 **TP** - True positive.
-The case where a case marked 'positive' in a dataset is correctly predicted 'positive' by an algorithm.
+The case where a case marked 'positive' in a data set is correctly predicted 'positive' by an algorithm.
 Generally, 'positive' is assigned to categories for which the use case needs to be alerted.
@@
 
-- don't mention bligh tanner!
+I would like to express special appreciation to my Principal Supervisor Professor Kerrie Mengersen.
+Thanks for extending a hand and giving me this opportunity as well as deftly guiding my work and welcoming me into your excellent research environment.
+Next, I would like to thank my Associate Supervisor Dr.
+Paul Wu for kindly and cheerfully working with me on the finer details of my research, giving up many hours of your time.
+Lastly, I would not have learned half as much without my colleagues in the BRAG group and ACEMS team.
 
@@
 
-Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.
+Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In *Big Data Visual Analytics (BDVA), 2015* (pp. 1-2). IEEE.
 
@@
 
-Signed .................
+Signed ................................
 Date....................
 
-# Chapter 1: Introduction
+# Introduction
 
@@
 
-Consulting businesses give expert advice to other professionals in exchange for a fee based on the amount of time their consultanting staff spend on the project.
+Consulting businesses give expert advice to other professionals in exchange for a fee based on the amount of time their consulting staff spends on the project.
 The deliverable 'expert advice' is generally in the form of a document or a package of documents.
-A document could range from a 5 page report written by a single person to a sequence of hundreds of drawings assembled by a team of 10 over many years.
+A document could range from a 5-page report written by a single person to a sequence of hundreds of drawings assembled by a team of 10 over many years.
 Despite the range of possible deliverables, before any project commences, the client and consultant must agree on a fee, or at least a fee structure.
@@
 1. The client agrees to pay for the consultants' time by the hour until the task is complete (referred to as hourly rate) or 
-2. the consultant offers a fixed fee to complete the project in full, regardless of time spent (referred to as fixed price or fixed fee).
+2. The consultant offers a fixed fee to complete the project in full, regardless of time spent (referred to as fixed price or fixed fee).
 
@@
 Typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb (such as a set fraction of the entire client budget for a large project).
-Such qualitative methods can result in profit margins which ultimately depend on the skill (or luck) the manager has in predicting the amount of time a project demands.
+Such qualitative methods can result in profit margins, which ultimately depend on the skill (or luck) the manager has in predicting the amount of time a project demands.
 
 This thesis focuses on a single case study consulting company in the construction industry that has persevered through many fixed price projects that resulted in losses.
-A source of insight into the loss-making projects was identified - twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
-Was it possible to find trends in what makes a project loss-making and can these characteristics be identified before a fixed price is offered?
-Can these trends be modelled into a predictive algorithm, and if so, how could the outcomes be integrated into business decisions?
-Finally, if the predictive outcomes were integrated into decision-making, would they deinitely improve the business' bottom line?
+A source of insight into the loss making projects was identified - twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
+Was it possible to find trends in what makes a project loss making and can these characteristics be identified before a fixed price is offered?
+Can these trends be modeled into a predictive algorithm, and if so, how could the outcomes be integrated into business decisions?
+Finally, if the predictive outcomes were integrated into decision-making, would they definitely improve the business' bottom line?
 These questions are explored in depth and answered over the course of this dissertation.
@@
 
-# Chapter 2: Problem Description
+# Problem Description
 
@@
 
-**Problem Description:**
+### Problem Description
 
@@
 Another review of 6 IT project surveys between 1984 and 1994 by @Molokken2003 revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
-These statistics describe difficult conditions that can result in job losses, business bankruptcies, and government budget blow outs.
+These statistics describe difficult conditions that can result in job losses, business bankruptcies, and government budget blowouts.
 A survey from @Moores1992 asked software managers whether they saw effort estimation as a problem, and 91% responded 'yes'.
@@
 
-**A Wicked Problem**
+### A Wicked Problem
 
@@
 Complex projects always differ from one another - even similar projects may end up demanding significantly different amounts of time.
-Although this is known, a fixed price is often demanded by clients before a project is awarded.
-If an hourly rate fee structure is proposed, clients in many industries can simply turn to another consultant willing to take the financial risks involved with fixed prices **(...ref...)**.
+Although this is known, clients often request a fixed price before a project is awarded.
+If an hourly rate fee structure is proposed, clients in many industries can simply turn to another consultant willing to take the financial risks involved with fixed prices [@Pai2013].
 Two strategies to guard against fixed price budget blowouts are detailed below along with their limitations.
@@
 However, the behaviour may cause friction between the client and the consultant, therefore reducing the potential for repeat business.
-In the event of hours inflating beyond expectation, consultants may choose to wear losses or endure marginable profits to preserve their client relationship.
+In the event of hours inflating beyond expectation, consultants may choose to wear losses or endure marginal profits to preserve their client relationship.
 They may reason it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction.
@@
 Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed prices in complex projects.
-In many industries, the reality of fixed price fee structures are unavoidable and businesses conform to survive.
-
-**Current Fixed Price Estimation Methods:**
+In many industries, the reality of fixed price fee structures is unavoidable and businesses conform to survive.
+
+### Current Fixed Price Estimation Methods
 
 It has been established that estimating fixed price fees in consulting is a difficult task with unfortunate consequences if the price is estimated too conservatively.
-Thie project aims to mobilise a consultant's historic data describing profitable and loss-making jobs in order to improve their fixed price setting practices.
+This project aims to mobilise a consultant's historic data describing profitable and loss making jobs in order to improve their fixed price setting practices.
 But what is the current method for estimating fixed prices or project costs?
-Is past data currently used in other cost estimation practises and to what degree of success?
+Is past data currently used in other cost estimation practices and to what degree of success?
 
@@
 
-**Cost Estimation and Human Nature**
+### Cost Estimation and Human Nature
 
 Although the surveys indicated industry did not analyse past project cost data, the idea of using this data (in the same way) has been studied before.
-Research exists that tested people's estimation accuracy with and without comparitive historic data with interesting results.
+Research exists that tested people's estimation accuracy with and without comparative historic data with interesting results.
 This is summarised below along with the resulting data-driven models.
@@
 A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplays or ignores the risk of problems that are out of the team's control.
-This is why it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
+This is why it does not matter if the project is broken down to the highest level of detail for cost prediction; all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
 Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].
 
-@Lovallo2003 call the practice of analysing a project based on project details without considering unforseen risks the 'inside view'.
+@Lovallo2003 called the practice of analysing a project based on project details without considering unforeseen risks the 'inside view'.
 A survey by @Moores1992 aligned with this theory and found the most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%).
@@
 One method @Lovallo2003 recommended was to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost.
-The correlation for the current project can then be estimated via a statistical model, which is then used to adjust the forecasted cost made by detailed analysis (the 'inside view') [@Flyvbjerg2011].
+The correlation for the current project can then be estimated via a statistical model, which is then used to adjust the forecast cost made by detailed analysis (the 'inside view') [@Flyvbjerg2011].
 As demonstrated in the surveys, @Flyvbjerg2011 found that managers rarely analysed data from similar projects even though it could significantly improve cost predictions.
@@
 It was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, and Australia.
-An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].
+An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecast cost [@Flyvbjerg2011].
 An example reading from the chart is as follows: "If the chance of a cost overrun of 20% is acceptable, the estimated cost should be increased by 55% from original calculations.".
@@
 The visualisation presents a powerful communication tool for influencing decision makers and improving forecast cost accuracy, however, access to credible data for a sufficient number of projects can be a challenge [@Flyvbjerg2011].
-Overall, the optimism bias theory credibly helps to explain the persistant problem of project cost underestimation.
+Overall, the optimism bias theory credibly helps to explain the persistent problem of project cost underestimation.
 
@@
 
-Although @Flyvbjerg2011's reference class forecasting theory and the psychology behind it was first published in 2003, predicting project costs using previous data has been researched for many preceding years and continues to this day.
+Although Flyvbjerg's (2011) reference class forecasting theory and the psychology behind it was first published in 2003, predicting project costs using previous data has been researched for many preceding years and continues to this day.
 In this section, a brief overview of other reference class forecasting models is presented (they are reviewed in greater detail in the literature review) followed by an analysis of why industry has not yet adopted them.
 
-Most cost estimation or reference forecasting analytical models focused on software development and construction projects, along with some work on film projects.
-These analyses generally used information from 15 to 20 large projects and the dataset summarised projects from around the world with different clients and teams.
-Applied mathematical methods varied from statistical Regressions to deep-learning Neural Networks [@Love2005][@Kim2004].
+Most cost estimation or reference forecasting analytic models focused on software development and construction projects, along with some work on film projects.
+These analyses generally used information from 15 to 20 large projects and the data set summarised projects from around the world with different clients and teams.
+Applied mathematical methods varied from statistical Regressions to deep-learning Neural Networks [@Love2005; @Kim2004].
 Most studies reported that their predictive models performed well and that Neural Networks were generally more accurate than Linear Regression.
@@
 
-In the software industry it is also possible the problem of fixed price contract blowouts has been solved by an alternative contract structure which has made cost estimation models unnecessary.
-In contrast to construction, IT projects are a relatively new practice and the industry has had success in trialing alternative contractural arangements.
+In the software industry it is also possible the problem of fixed price contract blowouts has been solved by an alternative contract structure, which has made cost estimation models unnecessary.
+In contrast to construction, IT projects are a relatively new practice and the industry has had success in trialing alternative contractual arrangements.
 Through experimentation, the Agile movement was developed.
@@
 Furthermore, the project outcomes are continually adjusted and revised based on frequent progress review points that assess unexpected problems.
-The flexible upfront arrangement with clients has often been found to forge better long term relationships [Cockburn2001].
+The flexible upfront arrangement with clients has often been found to forge better long-term relationships [Cockburn2001].
 This agile methodology suits the nature of software development work, where curtailing the scope of work is relatively straightforward.
@@
 The industry is so large, broad and well established that it has inertia in the way projects are estimated, and contractually bound.
-Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high level of legal responsibility and due diligence.
+Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high-level of legal responsibility and due diligence.
 
@@
 In their profession, labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm.
-In recent years, the size of losses from unprofitable projects equalled 25-30% of their profits from profitable projects.
+In recent years, the size of losses from unprofitable projects equaled 25-30% of their profits from profitable projects.
 This demonstrates the magnitude of their cost estimation problem and the substantial room for improvement that a data-driven model may facilitate.
@@
 Cost estimation is currently performed in the case study company by first carefully reviewing preliminary drawings where time costs for each task are estimated and summed by an experienced manager (the 'inside view').
-This is often cross-checked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
+This is often crosschecked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
 No formal mathematical method of comparison to similar projects is performed.
@@
 
-* lack of time to interrogate past project data
-* clunky data availability offered by their database software
-* lack of awareness of similar projects performed by other managers
+* Lack of time to interrogate past project data
+* Clunky data availability offered by their database software
+* Lack of awareness of similar projects performed by other managers
 
 A Customer Relationship Management software package (CRM) is currently employed to collect and store project data.
-A CRM is a popular type of software used by businesses to record client and project details as well as employee timesheet records.
-In the case study business, the CRM is available to all employees over the company intranet, and each employee completes daily timesheets alotting their hours to certain projects.
+A CRM is a popular type of software used by businesses to record client and project details as well as employee time sheet records.
+In the case study business, the CRM is available to all employees over the company intranet, and each employee completes daily time sheets allotting their hours to certain projects.
 Additionally, technical information is recorded against each project as well and the CRM data stood as an untapped source of information within the case study organisation.
@@
 
-* employee timesheet hours with dates
-* other project costs (taxis, printing)
-* client information/characteristics
-* client identification code
-* invoiced amounts for each project and dates
-* employee costs
-* employee charge out rates
-* project description
-
-The CRM software readily provides simple output statistics, however, analytical capabilities were limited to:
-
-* simple scatter plots and bar charts of the raw data
-* summaries such as overall hours spent vs. the invoiced amount for each project
+* Employee time sheet hours with dates
+* Other project costs (taxis, printing)
+* Client information/characteristics
+* Client identification code
+* Invoiced amounts for each project and dates
+* Employee costs
+* Employee charge out rates
+* Project description
+
+The CRM software readily provides simple output statistics, however, analytic capabilities were limited to:
+
+* Simple scatter plots and bar charts of the raw data
+* Summaries such as overall hours spent vs. the invoiced amount for each project
 
@@
 This study exploits the CRM data by performing more sophisticated statistical analysis with the intention of building a predictive model to improve cost estimation.
-Thousands of past projects are available for analysis, as opposed to previous studies that used on average 15-20 cases and at most 300 [@Finnie1997][@Pai2013][@Shin2015].
+Thousands of past projects are available for analysis, as opposed to previous studies that used on average 15-20 cases and at most 300 [@Finnie1997; @Pai2013; @Shin2015].
 This improves the potential for accurate predictions.
@@
 The most obvious limitation is that the work focuses on a single company in a specific industry, which limits the application of conclusions to a broader range of companies.
-This can be overcome by first determining the optimum method and variables for this case and then testing them on future case study companies.
+First determining the optimum method and variables for this case and then testing them on future case study companies can overcome this.
 
@@
 
-Finally, all CRM data is entered by employees manually.
+Finally, employees enter all CRM data manually.
 Because the data is input by people, it is susceptible to some error.
@@
 
-In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the dataset is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
+In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the data set is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
 These points are outlined in more detail below.
 
-In past research, cost estimation datasets generally consisted of dozens of projects from a range of a companys and even countries.
+In past research, cost estimation data sets generally consisted of dozens of projects from a range of a companies and even countries.
 This thesis case study only had access to the company's internal data but this contained over 2,000 past jobs - far more than the usual cost estimation study and with many more variables.
-Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry which differs from other studies in the construction industry that estimate actual construction costs.
+Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry, which differs from other studies in the construction industry that estimate actual construction costs.
 Consulting in the construction industry can be likened more to software development effort estimation although the nature of the work differs considerably.
 
-In addition to addressing cost estimation with a different kind of dataset, ensemble tree methods are applied which have been minimally tested on this problem.
+In addition to addressing cost estimation with a different kind of data set, ensemble tree methods are applied which have been minimally tested on this problem.
 Ensemble tree methods combine hundreds or thousands of decision trees to make predictions.
 A single decision tree is a non-parametric machine-learned model built by progressively determining the best binary split to partition the data according to the response variable.
-For further explanation of the method behind decision trees and ensemble three excellent sources are @Hastie2009, @Elith2008, and @Breiman2005.
-To date, Linear Regression, Neural Networks, Case Based Reasoning, and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks in some cases and generally outperform Linear Regression (**ref**).
+For further explanation of the method behind decision trees and ensemble trees, three excellent sources are @Hastie2009, @Elith2008, and @Breiman2005.
+To date, Linear Regression, Neural Networks, Case Based Reasoning, and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks and generally outperform Linear Regression [@Caruana2006].
 Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots that provide insight into the model's calculations.
-This contrasts Neural Networks and SVM's which are 'black-box' predictors.
+This contrasts Neural Networks and SVM's, which are 'black-box' predictors.
 Providing insight to a user is a major benefit as past research has demonstrated issues with user trust and uptake of models, even if the models predicted well.
@@
 Finally, this study propels the predictive model one step further than other studies by analysing bottom-line profit improvements that could result from the research.
-A few possibile ways to integrate model results into business decision-making scenarios are presented, and where possible, final profit increases are evaluated.
+A few possible ways to integrate model results into business decision-making scenarios are presented, and where possible, final profit increases are evaluated.
 This kind of thought experiment is designed to present a clear case for industry when evaluating if and how to adopt cost estimation models.
@@
 These were compared against individual models and the best constructs were selected.
-Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision making scenarios.
+Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision-making scenarios.
 Further applications were also considered.
 
-The research outlined in the method were divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion.
+The research outlined in the method was divided into four stages, each of which was assigned a chapter containing the results and discussion of that portion.
 Chapter 5 describes the variable selection process.
 Chapter 6 and 7 cover the results of the two ways of framing the cost estimation problem: using regression models and binary classification models.
-Chapter 8 shows the outcomes of model blending, then Chapter 9  extends the predictive analyses into an overall profit analysis, as well as describing research limitations and future work.
+Chapter 8 shows the outcomes of model blending, and then Chapter 9 extends the predictive analyses into an overall profit analysis, as well as describing research limitations and future work.
 
@@
 
-# Chapter 3: Literature Review
-
-"If I have seen further than others, it is by standing upon the shoulders of giants." - Isaac Newton
+# Literature Review
 
@@
 The bulk of research to date has been performed with project data in either the construction industry or software development.
-The different methods and outcomes in each of these industries is reviewed below.
+The different methods and outcomes in each of these industries are reviewed below.
 Finally a gap in the literature will be highlighted which this research addresses.
@@
 Research in the construction industry has primarily focused on predicting the final building cost of construction.
-Some studies reported surveyed cost estimation practices while others tested new methods such as building information modelling (BIM) analyisis, case based reasoning, and predictive algorithms.
+Some studies reported surveyed cost estimation practices while others tested new methods such as building information modeling (BIM) analysis, case based reasoning, and predictive algorithms.
 The cost estimation methods of detailed analysis (the 'inside view'), Case Based Reasoning, and predictive algorithms are reviewed in this section.
@@
 [@Akintoye2000].
-As referenced in the Research Motivation, @Akintoye2000's study found that this was the most popular method amongst contractors.
-
-@Elfaki2014 discusses how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator.
+As referenced in the Research Motivation, Akintoye and Fitzgerald's [-@Akintoye2000] study found that this was the most popular method amongst contractors.
+
+@Elfaki2014 discuss how much detailed cost estimates for construction projects can vary from estimator (engineer) to estimator.
 This contributes to the lack of accuracy in predicting final costs of a project.
-He argues that an engineer's expertise is not documented or measured in any way and therefore their expertise and, in turn, their estimate is prone to subjectivity.
-@Shane2009 theorises that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues, that it is almost impossible to achieve an accurate cost prediction manually.
-
-Groundbreaking methods into calculating construction costs by @Ma2014 mobilise data from building information modeling (BIM) models, although this is still technically a detailed analysis of the costs.
+They argue that an engineer's expertise is not documented or measured in any way and therefore their expertise and, in turn, their estimate is prone to subjectivity.
+@Shane2009 theorise that final cost can be subject to so many different unpredictable parameters, such as weather, unexpected ground conditions, duration or sub-contractor issues, that it is almost impossible to achieve an accurate cost prediction manually.
+
+Groundbreaking methods into calculating construction costs by @Ma2014 mobilised data from building information modeling (BIM) models, although this is still technically a detailed analysis of the costs.
 The idea is to quantify the cost of building a structure directly from a three dimensional BIM model created by the design consultants (engineers and architects).
-@Ma2014's first trial was to automate the cost calculation of a reinforced concrete structure.
-He programmed his algorithm to intelligently establish construction techniques for each element as this is not provided by architectural or engineering models.
+Ma and Liu's [-@Ma2014] first trial was to automate the cost calculation of a reinforced concrete structure.
+They programmed their algorithm to intelligently establish construction techniques for each element, as architectural or engineering models do not provide this.
 From the construction technique and material, a cost was derived for each element.
-This was successful and @Ma2014 aims to further his work so that more construction information can be intelligently obtained from BIM models.
-A system like @Ma2014's would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could still deliver innaccurate estimates similar to detailed analysis by a person.
+This was successful and @Ma2014 aim to further their work so that more construction information can be intelligently obtained from BIM models.
+A system like Ma and Liu's [-@Ma2014] would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could still deliver inaccurate estimates similar to detailed analysis by a person.
 The detailed view does not 'step back' and take into account set backs in construction that may have affected past similar projects.
@@
 They found that artificial Neural Networks and SVM's were the most common machine learning techniques.
-He highlighted that they dealt well with uncertainty, however lack technical justification to the decision maker (known as a black box predictor).
+These techniques also deal well with uncertainty, however lack technical justification for the decision maker (known as a black box predictor).
 Neural networks are also time consuming to train, and must be re-trained and re-tested with each additional piece of data [@Kim2004].
-Nevertheless, Neural Networks and SVM's received significant attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015][@Kim2004].
+Nevertheless, Neural Networks and SVM's received significant attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015; @Kim2004].
 
@@
 In their study, data from 234 school building construction projects in Korea were used.
-Boosted trees were compared to Neural Networks and were found to predict costs slightly more accurately than Neural Networks, but not statstically significantly (p value < 0.05).
+Boosted trees were compared to Neural Networks and were found to predict costs slightly more accurately than Neural Networks, but not statistically significantly (p-value < 0.05).
 
@@
 This may be because the technique is straightforward, easy to use and widely available in statistical packages [@Chan2005].
-Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004][@Attalla2003].
+Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004; @Attalla2003].
 @Dissanayaka1999 suggested that Regression models could be performed first to determine variable importance and condense the number of variables, then Neural Networks could be used for better estimation.
@@
 This is likely because construction projects are multidisciplinary and involve many parties such as the clients, consultants, contractors and suppliers, [@Chan2005].
-@Chan2005's study on 87 building projects in Singapore included special, complex projects.
+Chan and Park's [-@Chan2005] study on 87 building projects in Singapore included special, complex projects.
 Therefore, variables such as contractor's specialised skills, who the client was (public vs.
 private), the client level of experience, and the contractor's financial management highly influenced final cost of the project.
-Other studies found variables such as project complexity, duration, team experience, information availablility, site requirements, and labour climate to contribute most [@Akintoye2000][@Trost2003][@Pinto1988].
+Other studies found variables such as project complexity, duration, team experience, information availability, site requirements, and labour climate to contribute most [@Akintoye2000; @Trost2003; @Pinto1988].
 As these studies obtained information from questionnaires, the available variables depended on the questions.
@@
 Similar result statistics are useful, however there is a gap in the literature rationalising which model industry should adopt, how to adopt it into a business, and how much loss can be reduced.
-This gap may explain why uptake of these algorithmic models are sparse, as indicated by @Akintoye2000's survey, despite 20 years of prior research.
+This gap may explain why uptake of these algorithmic models is sparse, as indicated by Akintoye and Fitzgerald's [-@Akintoye2000] survey, despite 20 years of prior research.
 @Akintoye2000 hypothesised this could be due to a lack of knowledge about the techniques, doubts whether these techniques are applicable to the construction industry, and the availability of sound data to ensure confidence.
@@
 It is programmed to calculate the Euclidean distance in n-dimensional space between cases, where each variable is a dimension [@Kumar2007].
-3.
-The cost of the new project is estimated by extrapolating characteristics of the similar cases to the new case.
-This can also be done either algorithmically or manually by an experienced decision maker.
+3. The cost of the new project is estimated by extrapolating characteristics of the similar cases to the new case. An experienced decision maker can also do this either algorithmically or manually.
 
@@
 The discussion compared mean absolute error rates of the three methods and showed that CBR was more accurate than Regression, but less accurate than Neural Networks in estimating cost.
-Despite Neural Networks outperforming CBR, CBR maintained advantages over Neural Networks which have been referenced in other studies as well.
+Despite Neural Networks outperforming CBR, CBR maintained advantages over Neural Networks, which have been referenced in other studies as well.
 
 Advantages of CBR include algorithmic efficiency as well as user engagement.
-In @Kim2004's study, it was noted that CBR models were simple to update with new data in comparison to Neural Networks, which must be slowly re-trained and re-tuned.
-New cases also did not need to have every variable complete which can occur often in 'real world' data.
+In Kim, An, and Kang's [-@Kim2004] study, it was noted that CBR models were simple to update with new data in comparison to Neural Networks, which must be slowly re-trained and re-tuned.
+New cases also did not need to have every variable complete, which can occur often in 'real world' data.
 Neural Networks on the other hand require complete data.
-Another advantage of CBR is the ability for users to review chosen similar cases and make sense of the prediction, as opposed to Neural Networks [@Kim2004][@Elfaki2014][@Kumar2007].
+Another advantage of CBR is the ability for users to review chosen similar cases and make sense of the prediction, as opposed to Neural Networks [@Kim2004; @Elfaki2014; @Kumar2007].
 One disadvantage is that accuracy can be highly dependent on the number of selected cases [@Elfaki2014].
@@
 
-CBR that is performed manually through personal experience, without the assistance of algorithms, was the second most popular method for construction estimation in @Akintoye2000's survey in the UK.
+CBR that is performed manually through personal experience, without the assistance of algorithms, was the second most popular method for construction estimation in Akintoye and Fitzgerald's [-@Akintoye2000] survey in the UK.
 This indicates it is quite an intuitive method that allows decision makers to cross check their detailed cost estimates, although an algorithmic version of CBR has not found success in industry.
@@
 A parallel can be drawn between software development effort estimation and consultants' effort estimation in the construction industry - which was highlighted as a gap in the previous section.
-Plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modelling, and Case Based Reasoning.
+Plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modeling, and Case Based Reasoning.
 
@@
 
-Similar to the construction industry, expert judgment or detailed analysis, is the most widely practised method for effort estimation [@Shepperd1996][@Molokken2003].
+Similar to the construction industry, expert judgment or detailed analysis, is the most widely practised method for effort estimation [@Shepperd1996; @Molokken2003].
 This is despite many years of research being dedicated to developing algorithmic models that, in the research context, outperform expert judgment.
@@
 
-Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although Regression is the most popular method [@Finnie1997][@Pai2013][@Matson1993].
-Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the model's accuracy: size [@Shepperd1996][@Finnie1997][@Pai2013].
+Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although Regression is the most popular method [@Finnie1997; @Pai2013; @Matson1993].
+Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the model's accuracy: size [@Shepperd1996; @Finnie1997; @Pai2013].
 Size can refer to the expected number of lines of code in the software package or function point, which refers to the amount of business functionality expected from the product.
@@
 There are disadvantages to algorithmic methods, similar to the construction industry.
-First, there are often not enough cases to create a good model, particularly if the cases must be from within the company [@Finnie1997][@Pai2013].
+First, there are often not enough cases to create a good model, particularly if the cases must be from within the company [@Finnie1997; @Pai2013].
 A study by @Mendes2004 demonstrated using 67 web projects that cross-company models were significantly less accurate than a within company model, so it is in the company's interest to create an in-house model.
@@
 
-in comparison to the construction industry, infrastructure projects do not have a function point variable and need to utilise other information such as expected total project cost (which is notoriously hard to predict based on the previous section), client characteristics, and other project characteristics.
+In comparison to the construction industry, infrastructure projects do not have a function point variable and need to utilise other information such as expected total project cost (which is notoriously hard to predict based on the previous section), client characteristics, and other project characteristics.
 
@@
 CBR was found to perform approximately equally to algorithmic methods such as Neural Networks.
-This contrasts results in the construction industry where Neural Networks performed better [@Shepperd1996][@Finnie1997].
+This contrasts results in the construction industry where Neural Networks performed better [@Shepperd1996; @Finnie1997].
 It was also highlighted that CBR is intuitively similar to how an expert thinks about pricing projects  [@Finnie1997].
@@
 From findings in previous literature, the models should reveal their predictive structure to decision makers and be based on internal company data.
-Review of the literature revealed a need for an industry focussed decision framework that outlines how to integrate the predictive model in a way that optimises profits.
+Review of the literature revealed a need for an industry focused decision framework that outlines how to integrate the predictive model in a way that optimises profits.
 
@@
 These included Naive Bayes, Random Forests, and machine learned Bayesian Networks.
-The following section presents an explanation fo each method, their advantages and disadvantanges and an example of a successful application.
+The following section presents an explanation of each method, their advantages and disadvantages and an example of a successful application.
 
@@
 
-where
+Where
 
@@
 $RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$
+
 [@Hastie2009]
@@
 Skewed explanatory variable distributions are acceptable but it is generally better to normalise them by, for example, taking the logarithm or square root of the variable.
-The reason for this is that the more extreme values from the skewed distribution may create high leverage points and influence the slope (variable coeffiecients) too strongly [@Chatterjee1986].
+The reason for this is that the more extreme values from the skewed distribution may create high leverage points and influence the slope (variable coefficients) too strongly [@Chatterjee1986].
 The distribution of the residuals on the other hand, should always be Gaussian with a mean of 0.
@@
 
-Unfortunately many real world world phenomena do not have linear relationships which can make it difficult to produce accurate results using this method [@Breiman2001b].
+Unfortunately many real world phenomena do not have linear relationships, which can make it difficult to produce accurate results using this method [@Breiman2001b].
 There are methods to model non-linear relationships such as including polynomial terms and interaction terms.
@@
 This versatility is advantageous, however it can be challenging to know which non-linear terms of which variables to include.
-Sometimes data cannot be closely modelled by any equation and this is where other non-linear non-parametric methods can outperform linear regression.
+Sometimes data cannot be closely modeled by any equation and this is where other non-linear non-parametric methods can outperform linear regression.
 
@@
 Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable (represented by 1 or 0).
-First, a sigmoid function is assumed as the functional form of the probabilties against an example variable X.
+First, a sigmoid function is assumed as the functional form of the probabilities against an example variable X.
 This is because it is pragmatic and works well: sigmoid functions can be bound by 0 and 1, and a logit transformation of the probability results in a linear relationship between the log odds and X.
@@
 
-The result of this linear fit and non linear transformation is a model that predicts the probability (between 0 and 1) that an input case will result in a 0 or 1 (i.e.
+The result of this linear fit and non-linear transformation is a model that predicts the probability (between 0 and 1) that an input case will result in a 0 or 1 (i.e.
 success or failure) [@Moore1989].
@@
 \log\left(\frac{0.5}{0.5}\right) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
-\log\(1) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
+\log(1) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
 0 &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
@@
 
-```{r log_reg_decboundary, fig.cap = c("Example Bivariate Decision Boundary for Probabability Threshold = 0.5")}
+```{r log_reg_decboundary, fig.cap = c("Example Bivariate Decision Boundary for Probability Threshold = 0.5")}
 set.seed(1234)
@@
 Therefore, if conditional independence is assumed, each variable's probability contribution can be calculated by looking at the probability of a variable's value given the response class only.
-This simplification can be favourably integrated into Bayes Theorem which is shown below:
+This simplification can be favourably integrated into Bayes Theorem, which is shown below:
 
@@
 
-Say the probability of 'yes' or 'no' (in a binary reponse variable Y) given the explanatory variables $X_1$ and $X_2$ is being sought.
+Say the probability of 'yes' or 'no' (in a binary response variable Y) given the explanatory variables $X_1$ and $X_2$ is being sought.
 The equation above could be rewritten for both possible response categories:
@@
 
-The aim is to find which category has a higher probability, so to simplify calcualtions, the denominator can be dropped.
+The aim is to find which category has a higher probability, which means the denominator can be dropped to simplify calculations.
 Because the denominator is the same for both equations, it does not affect the final ratio of the probability of 'yes' to the probability of 'no'.
@@
 This is fine for ranking but the output probabilities are not accurate statistical probabilities [@Caruana2006].
-Another disadvantage is that the distribution of numeric variables must be assumed (often Gaussian) in order to calculate the liklihood terms, and the data may not fit into these distributions neatly [@Caruana2006].
+Another disadvantage is that the distribution of numeric variables must be assumed (often Gaussian) in order to calculate the likelihood terms, and the data may not fit into these distributions neatly [@Caruana2006].
 
 Example Gaussian probability densities of two response classes can be visualised on a bivariate feature space.
-The two normal distrubtions represent the probability of each class given values of the two explanatory variables (x and y axes):
+The two normal distributions represent the probability of each class given values of the two explanatory variables (x and y axes):
 
@@
 
-One real world success story for this classifier hinged on its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
+In one case, this classifier succeeded because of its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
 Naive Bayes was used in a complex spam detection system where new spam emails or toxic text themes could be quickly added to the filtering model.
@@
 There are several brands of basic decision tree algorithms including ID3, C4.5, CART, and CHAID with CART and C4.5 being the most popular [@Kabra2011].
-Ravi Kumar and Ravi [@Kumar2007] recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
-Single decision trees create intuitive rules that a decision maker can follow in real-life scenarios but also tend to overfit the data and provide low predictive accuracy [@Kabra2011][@Putler2012].
+@Kumar2007 recommended the CART algorithm as it is capable of solving both classification and regression problems whereas the remaining decision trees solve classification problems only.
+Single decision trees create intuitive rules that a decision maker can follow in real-life scenarios but also tend to overfit the data and provide low predictive accuracy [@Kabra2011; @Putler2012].
 Ensemble trees were developed to solve this problem and several ensemble techniques are described in this section.
@@
 Decision trees have a different theoretical foundation to traditional prediction methods such as regression.
-They are created by a progression of binary splits of the dataset that attempt to group similar response variable values.
+They are created by a progression of binary splits of the data set that attempt to group similar response variable values.
 Before a split is made, each explanatory variable and each value within the explanatory variables is tested for how 'purely' it partitions the response variable into two subsets.
@@
 
-and
+And
 
@@
 Other purity formulas also exist for categorical response variables such as Gini purity and information gain [@Hastie2009].
-After the first split is made on the dataset, each subset is split by another variable (or the same explanatory variable again) and so on resulting in final subsets of data with similar response variable values (high purity).
+After the first split is made on the data set, each subset is split by another variable (or the same explanatory variable again) and so on resulting in final subsets of data with similar response variable values (high purity).
 If this process is visualised graphically, it is clear that non-linear relationships can be captured.
@@
 As stated previously, single decision trees also suffer from low predictive accuracy and instability, so to combat these problems, ensemble tree methods were pioneered in the 1990's [@Breiman1996].
-Three examples of these methods are boostrap aggregating (bagging), Random Forests, and Boosted Trees.
-Bagging builds multiple trees from different training sample data sets (wich are sampled from the full data set *with* replacement) and then combines tree results [@Breiman1996].
+Three examples of these methods are bootstrap aggregating (bagging), Random Forests, and Boosted Trees.
+Bagging builds multiple trees from different training sample data sets (which are sampled from the full data set *with* replacement) and then combines tree results [@Breiman1996].
 The method of combination is generally averaging numeric predictions or taking the majority rule for categorical variables.
@@
 Further advantages of ensemble tree methods are numerous.
-Similarly to Linear Regression, the predictor variables in trees do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014][@Radenkovic2010].
+Similarly to Linear Regression, the predictor variables in trees do not need to be transformed as no assumptions need to be made about the data's statistical distributions [@Louppe2014; @Radenkovic2010].
 However, it is often wise to transform skewed distributions to achieve more even segmentation.
@@
 @Caruana2006 tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
-This demonstrates how ensemble tree methods are capable of competing with high level machine learning algorithms.
+This demonstrates how ensemble tree methods are capable of competing with high-level machine learning algorithms.
 
 Although ensemble trees have numerous advantages it is just as important to understand their limitations.
-Trees are not built on a probabilistic framework, and therefore their results can not be provided in this framework.
+Trees are not built on a probabilistic framework, and therefore their results cannot be provided in this framework.
 For example, statistical confidence intervals for predictions are not available for standard ensemble methods [@Louppe2014].
@@
 Another type of ensemble method, conditional forests, should be used to check variable importance [@Radenkovic2010].
-Finally, the methods can also be prone to overfitting if not tuned carefully [@Louppe2014][@Caruana2006].
+Finally, the methods can also be prone to overfitting if not tuned carefully [@Louppe2014; @Caruana2006].
 
@@
 The model is visually represented by a directed acyclic graph (DAG) and is capable of linking the conditional dependency between any variable to another variable.
-arcs or edges represent links in the DAG [@Heckerman1998].
+Arcs or edges represent links in the DAG [@Heckerman1998].
 The conditional relationships are Bayesian, where the probabilities in one node are conditional upon values in nodes directed towards it as well as preceding nodes.
@@
 
-$p(ABCD) = p(C|ABD)\cdot p(A|BD) \cdot p(B|D) \cdot p(D)
+$p(ABCD) = p(C|ABD)\cdot p(A|BD) \cdot p(B|D) \cdot p(D)$
 
@@
 
-$p(ABCD) = p(C|BD)\cdot p(A|B) \cdot p(B) \cdot p(D)
+$p(ABCD) = p(C|BD)\cdot p(A|B) \cdot p(B) \cdot p(D)$
 
 In this way, Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes, for example, cannot [@Barber2012].
-Also, some equations for a comination are now clones of another combination and do not require recalculation.
-Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes (as Naive Bayes is the simplest form of a Bayesian Netwoork) and variables with many categories dramatically increases the expense [@Zhang2004].
+Also, some equations for a combination are now clones of another combination and do not require recalculation.
+Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes (as Naive Bayes is the simplest form of a Bayesian Network) and variables with many categories dramatically increases the expense [@Zhang2004].
 
 Network relationships can be learned from the data, however this is not widely included as part of the suite of machine learning methods.
-The conditional variable dependencies are calculated from the data which in turn can define the graph structure.
+The conditional variable dependencies are calculated from the data, which in turn can define the graph structure.
 Some conditional dependencies can be set before the structure is learned [@Barber2012].
@@
 Drawbacks of Bayesian Networks include the difficulty for them to process continuous variables.
-These demand sampling from conditional density models or complex algebra whereas categorical variables require comparitively simple calculations [@Hofmann1996].
-It is therefore common to discretise continuous variables which simultaneously decreases precision of that variable [@Kragt2009].
+These demand sampling from conditional density models or complex algebra whereas categorical variables require comparatively simple calculations [@Hofmann1996].
+It is therefore common to discretise continuous variables, which simultaneously decreases precision of that variable [@Kragt2009].
 
@@
 Algorithmic Neural Networks were inspired by Neural Networks of the human nervous system.
-In the human body, biological neurons recieve electro-chemical input into their dendrites from the terminals of adjacent neurons.
-If enough electro-chemical input is recieved, the neuron will 'fire', which means a signal will be sent from the dendrite end, through the axon to its terminal.
+In the human body, biological neurons receive electro-chemical input into their dendrites from the terminals of adjacent neurons.
+If enough electro-chemical input is received, the neuron will 'fire', which means a signal will be sent from the dendrite end, through the axon to its terminal.
 Its terminal then sends signals to adjacent dendrites of other neurons and the process propagates.
 Neurons are able to learn over time by strengthening the connections of adjacent neurons that frequently cause the next one to fire.
-This means the required firing threshold for the recieving neuron lowers in strengthened connections [@Putler2012].
+This means the required firing threshold for the receiving neuron lowers in strengthened connections [@Putler2012].
 
@@
 This is then transformed and assessed by a threshold function to determine if the hidden node should 'fire' to the next nodes.
-Once the signals and values have propogated to the final layer they are recombined to produce an output value, or a prediction [@Putler2012].
+Once the signals and values have propagated to the final layer they are recombined to produce an output value, or a prediction [@Putler2012].
 
@@
 Then optimisation techniques are used to adjust the weights and reduce the error.
-This process can result in overfitting, so stopping rules such as regularisation strength should be used [@Karpathy2016], however these details are outside the scope of this review.
+This process can result in overfitting, so stopping rules such as regularisation strength should be used, however these details are outside the scope of this review [@Karpathy2016].
 
@@
 
-![Non linear decision boundary created by a Neural Network[@Karpathy2016]](images/lit_review/NN_bivariate.PNG)
-
-Clearly, a flexible range of relationships can be mimicked by Neural Networks, and if more nodes are added, more flexibility is available [@Putler2012].
+![Non-linear decision boundary created by a Neural Network [@Karpathy2016]](images/lit_review/NN_bivariate.PNG)
+
+Clearly, Neural Networks can mimick a flexible range of relationships, and if more nodes are added, more flexibility is available [@Putler2012].
 A previously mentioned disadvantage is that the algorithm is a black box because the internal structure is too complex for interpretation.
 They also require a lot of training data relative to other methods.
-Despite these disadvantages, Neural Networks have found success in high level tasks where the model structure does not need to be understood.
+Despite these disadvantages, Neural Networks have found success in high-level tasks where the model structure does not need to be understood.
 These include hand writing recognition, vehicle control, and face recognition [@Haykin2004].
@@
 
-SVM's are non-probabalistic learning algorithms that are typically used for binary classification, although multi-class and regression problems can be solved.
-The method works by geometrically determining the best decision boundary, called a hyperplane, between two classes in multi-dimensional space [@Byun2002] .
+SVM's are non-probabilistic learning algorithms that are typically used for binary classification, although multi-class and regression problems can be solved.
+The method works by geometrically determining the best decision boundary, called a hyperplane, between two classes in multi-dimensional space [@Byun2002].
 A hyperplane is simply a plane in more than 3 dimensions, i.e.
@@
 
-where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables.
+Where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables.
 For two variables, $\boldsymbol{w}$ would be $\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$ [@Fletcher2009].
@@
 There are several types of kernel transformations that can be applied, but the common aim is to add another dimension to the data points.
-This can enable a linear hyperplane to be found in multi dimensional space which separates the response variable classes better [@Fletcher2009].
+This can enable a linear hyperplane to be found in multi-dimensional space that separates the response variable classes better [@Fletcher2009].
 A simple example of this is shown below.
@@
 Linear/Logistic Regression and Naive Bayes are computationally very fast with a simple concept that make some parametric assumptions about the variable relationships, but work well as baseline models against which to compare complex models.
-In certain cases these simpler models are outperformed by complex models however it is important to compare both.
+In certain cases complex models outperform these simpler models, however it is important to compare both.
 
@@
 None of the methods output statistical uncertainty or statistical confidence interval associated with each prediction.
-Understanding uncertainty should always be sought in decision making be it from a confidence interval or sensitivity testing.
+Understanding uncertainty should always be sought in decision-making be it from a confidence interval or sensitivity testing.
 One way to achieve this would be to build hundreds of models using the same method but based on different training divisions of the data.
@@
 Although SVM's performed the best, as discussed, they are often complex and slow, requiring a great deal of memory [@Kumar2007].
-In further support of SVM predictive performance, a study by Davenport and Harris [@Davenport2007] concluded that many statisticians experienced in predictive machine learning algorithms agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
-
-When comparing Neural Networks to ensemble trees, Ravi Kumar and Ravi [@Kumar2007] found that Neural Networks and ensemble trees were both capable of out-performing the other, depending on the context.
+In further support of SVM predictive performance, a study by @Davenport2007 concluded that many statisticians experienced in predictive machine learning algorithms agree that SVM's yield the highest predictive accuracy compared to other machine learning algorithms.
+
+When comparing Neural Networks to ensemble trees, @Kumar2007 found that Neural Networks and ensemble trees were both capable of out-performing the other, depending on the context.
 For example in financial credit scoring, a study by @Brown2012 found that Random Forests and Boosted Trees consistently outperformed Neural Networks in classification.
-Logistic regression also outperformed Neural Networks which may align with a theory that most credit scoring datasets are only weakly non-linear [@Brown2012].
-This demonstrates how method performance can depend on the dataset and the problem.
-
-@Kumar2007's review also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
+Logistic regression also outperformed Neural Networks which may align with a theory that most credit scoring data sets are only weakly non-linear [@Brown2012].
+This demonstrates how method performance can depend on the data set and the problem.
+
+Kumar and Ravi's [-@Kumar2007] review also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
 For example, combining the contrasting advantages of Neural Networks and decision trees is a worthwhile ensemble technique.
-@Tsai2009's study combined these two methods after trialing each one.
+Tsai & Chiou's [-@Tsai2009] study combined these two methods after trialing each one.
 They were run successively with the Neural Network running first because it had a higher predictive accuracy in this case.
@@
 There are far fewer studies that analyse problems affecting internal business decisions such as dealing with employees, teams, and clients.
-An exception is @Saradhi2011's study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
+An exception is Saradhi & Palshikar's [-@Saradhi2011] study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
 @Saradhi2011 applied popular customer churn predictive models to employee churn - a novel application that focused internally on employees rather than externally on customers.
@@
 This was attributed to the ability of SVM's to incorporate class penalties whereas the other methods were limited by the class imbalance problem [@Saradhi2011].
-This means that SVM's had an effective method of increasing the importance of 'positive' cases whicih was advantageous because the data set was skewed towards the majority of employees who kept their job versus the minority who left (25% churn).
+This means that SVM's had an effective method of increasing the importance of 'positive' cases which was advantageous because the data set was skewed towards the majority of employees who kept their job versus the minority who left (25% churn).
 It should be noted that the Boosted Tree ensemble method performs a similar task of weighting misclassified cases or outliers and it would have been worthwhile comparing this method to SVM's.
@@
 Another idea @Saradhi2011 adopted from customer churn analysis was the value models.
-Company's calculate customer lifetime values (CLV's) in order to structure which potentially churning customer's should be acted upon.
+Companies calculate customer lifetime values (CLV's) in order to structure which potentially churning customer's should be acted upon.
 Then the optimal number of customers to reach out to can be determined by optimising the cost of reaching out against preventing financial loss from churn.
@@
 This allowed them to rank employees identified as 'high risk of churn' by value and provided a clear ranking for manager's to act upon.
-Manager's could then start brainstorming actions to prevent high value, high risk employees from leaving.
+Managers could then start brainstorming actions to prevent high value, high risk employees from leaving.
 This extension of the study provided a comprehensive framework for how business managers could adopt their findings to improve business operations.
-It was a valuable addition which is absent from most cost and effort estimation research.
+It was a valuable addition that is absent from most cost and effort estimation research.
 
@@
 
-The limited literature on work similar to @Saradhi2011's Employee Churn study highlights a gap in the application of predictive techniques that use internal data to model internal decisions.
+The limited literature on work similar to Saradhi & Palshikar's [-@Saradhi2011] Employee Churn study highlights a gap in the application of predictive techniques that use internal data to model internal decisions.
 This is particularly relevant to consulting companies that have complicated internal processes necessary for each project.
@@
 
-# Chapter 4: Method
+# Method
 
@@
 The process of how the data was collected, cleaned, and how the predictive models were progressively built and tested is explained along with why the steps were taken.
-This section should enable an experienced practioner to replicate the study.
+This section should enable an experienced practitioner to replicate the study.
 
@@
 
-An assessment of the literature revealed that the capacity of internal timesheet data (extracted from a company's CRM software) to predict project profitability has not yet been tested in the context of consulting companies in the construction industry.
+An assessment of the literature revealed that the capacity of internal time sheet data (extracted from a company's CRM software) to predict project profitability has not yet been tested in the context of consulting companies in the construction industry.
 Alternate sources of project data were considered including interviewing project team members, emails, accounting records, and project drawings and calculations.
-However the CRM data from this case study could be extracted in a spreadsheet format and housed a detailed account of time spent on each project (timesheet data) along with client information and invoicing records for the preceding twelve years.
+However the CRM data from this case study could be extracted in a spreadsheet format and housed a detailed account of time spent on each project (time sheet data) along with client information and invoicing records for the preceding twelve years.
 This source of data was not only the most accessible and suitable for statistical analysis, but had the potential to reveal information not typically investigated by the company.
 
-The twelve years of project data was extracted via the CRM software interface, which described 4169 projects.
+The twelve years of project data was extracted via the CRM software interface, which described 4,169 projects.
 Projects varied from total invoiced amounts of \$500 to over \$1,000,000 and were divided between four internal company disciplines.
@@
 Once the data was extracted from the CRM, an employee from the case study company performed the task of de-identification.
-This is an important step before research commences as the privacy of project employees, clients, and project names must be removed for ethical reasons.
+This is an important step before research commences, as the privacy of project employees, clients, and project names must be removed for ethical reasons.
 In the case of employee names, professional titles replaced the names, and client codes replaced client names.
 Project names and addresses were removed, and each project was identified by a number.
-The resulting de-identified data consisted of three .csv files of different data structures which are described in the next section.
+The resulting de-identified data consisted of three .csv files of different data structures, which are described in the next section.
 
@@
 Then, potentially meaningful variables were engineered which could be tested alongside original variables.
-The three initial datasets described:
+The three initial data sets described:
 
@@
 
-* Timesheet entries: the number of hours spent on each project and on which day.
+* Time sheet entries: the number of hours spent on each project and on which day.
 Entries were input by individual employees, but names for each entry had been replaced with employee positions (such as mid-level technical)
@@
 
-These three sources needed to be compiled into a single dataset that detailed one project per row as the model was to predict the overall profitability of a project.
+These three sources needed to be compiled into a single data set that detailed one project per row, as the model was to predict the overall profitability of a project.
 Before compilation could begin however, thorough cleaning was required where the data was plotted and statistically assessed so that outliers could be visually or statistically discovered.
@@
 Once these errors were detected, they were discussed with the company directors and appropriate corrective action to the data was taken.
-Mistakes in data entry were often able to be corrected by notes that a user had entered into the system.
+Mistakes in data entry could often be corrected by notes that a user had entered into the system.
 It is possible that some data entry errors were not detected and this is a problem that also must be dealt with during implementation of the model.
@@
 
-Although dozens of variables were available from the initial dataset, further descriptive variables were engineered to trial in the models.
-This was particularly advantageous in the invoicing dataset and timesheet dataset where a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
+Although dozens of variables were available from the initial data set, further descriptive variables were engineered to trial in the models.
+This was particularly advantageous in the invoicing data set and time sheet data set where a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
 All variables would eventually be tested for variable importance and predictive power with respect to project profitability.
@@
 * Percent of hours performed by each professional role over the course of a project
-* Timespan of the entered project hours
+* Time span of the entered project hours
 * Percent of hours performed by 'professional' employees as opposed to 'technical' employees
@@
 * Total amount invoiced and remunerated per project
-* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500)over all invoices sent to a client accross all their projects.
-This gives an indication of whether the client generally does fast paced big jobs which would result in large monthly invoices for example or small jobs with small monthly invoiced amounts.
-* Client invoice frequency which gives an indication of how much work the company does for that client. 
+* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500) over all invoices sent to a client across all their projects.
+This gives an indication of whether the client generally does fast paced big jobs, which would result in large monthly invoices for example, or small jobs with small monthly invoiced amounts.
+* Client invoice frequency, which gives an indication of how much work the company does for that client. 
 \ 
@@
 
-* text analysis of project descriptions detected a list of key words that could classify projects into 16 categories.
+* Text analysis of project descriptions detected a list of key words that could classify projects into 16 categories.
 This key word analysis was done in conjunction with a company employee and the resulting classifications were reviewed by the employee to ensure accuracy.
 For example, project descriptions containing the words 'residence', 'house', 'apartment', 'home', 'townhouse', 'unit', and 'dwelling' were classified as 'residential'
-* number of projects completed with each client and client contact
+* Number of projects completed with each client and client contact
 
 With the engineered variables, all three data sources could be represented in the single-row-per-project format.
-The three datasets were then combined and further variables were engineered using combinations of variables across the original data sets:
+The three data sets were then combined and further variables were engineered using combinations of variables across the original data sets:
 
@@
 
-Besides the engineered variables, original variables included in the project summary dataset were client industry, internal company discipline, job description, and post code.
+Besides the engineered variables, original variables included in the project summary data set were client industry, internal company discipline, job description, and post code.
 Once numerous potentially important variables were engineered, the most important ones could then be selected.
@@
 
-Case study projects with extreme 'return per dollar' values could be the result of mistakes or special scenarios which the model is not intended to predict.
+Case study projects with extreme 'return per dollar' values could be the result of mistakes or special scenarios that the model is not intended to predict.
 Several methods could have been used for outlier detection including the distribution of the response variable, an analysis of high leverage cases, and cluster based methods.
-Guidelines for the first method were intially defined by Tukey, the inventor of the box plot [@Tukey1977].
+Guidelines for the first method were initially defined by Tukey, the inventor of the box plot [@Tukey1977].
 He recommended investigating cases with response variable values sitting outside 1.5 x the interquartile range (IQR) from the upper and lower quartiles.
@@
 
-Upon review of the outliers with an employee, many projects which appeared wildly profitable, were actually small jobs and high profitability ratios only indicated extra profits of a few hundred dollars.
+Upon review of the outliers with an employee, many projects that appeared wildly profitable were actually small jobs and high profitability ratios only indicated extra profits of a few hundred dollars.
 It was concluded that small job outliers most likely required so little time that busy employees did not bother writing down their hours for that job.
@@
 26 of the 36 outlier cases with a high 'return per dollar' were small projects worth less than \$4,000.
-This may indicate small projects are not suitable to predict due to timesheet entry habits on these projects, however after outlier deletion, 1,159 projects below \$4,000 remained in the dataset.
+This may indicate small projects are not suitable to predict due to time sheet entry habits on these projects, however after outlier deletion, 1,159 projects below \$4,000 remained in the data set.
 The prediction model would therefore still be applicable to small projects under \$4,000.
@@
 It was expected that Random Forests would favour variables that had more categories, a well-known bias; which is also a bias that cForests have overcome [@Strobl2007].
-ANOVA is based on linear theory which contrasts conditional forest and Random Forest's more free-form structure.
+ANOVA is based on linear theory, which contrasts conditional forest and Random Forest's more free-form structure.
 Therefore, it is beneficial to compare the methods' assessments of variable importance.
@@
 *how* profitable a project was (continuous) or whether a project was profitable or not (binary classification).
-The continuous response variable, 'reurn per dollar', was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
+The continuous response variable, 'return per dollar', was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
 
@@
 A significance test for whether the coefficient equals zero is performed on each coefficient and the p-value is reported.
-Therefore, if the null hypothesis is rejected (p value < 0.05), changes in the explanatory variable are indeed related to changes in the response variable.
+Therefore, if the null hypothesis is rejected (p-value < 0.05), changes in the explanatory variable are indeed related to changes in the response variable.
 P-values are a good indication of whether variables have a relationship with the target variable, however they do not necessarily rank importance.
@@
 Variable importance is represented as a score, which is calculated sequentially for each variable by comparing predictive accuracy between out-of-bag (OOB) data run through the trees without, then with a variable permutation (reordering).
-Since each tree is a bootstrapped sample of the complete dataset, approximately 1/3 of the data is held aside as the OOB sample for each tree.
+Since each tree is a bootstrapped sample of the complete data set, approximately 1/3 of the data is held aside as the OOB sample for each tree.
 It is passed through its partner tree and produces an unbiased estimate of error and in this case, is also used to estimate variable importance.
@@
 C_n &= \textrm{Number of correctly classified cases in the nth tree using the OOB sample} \\
-C_{n,X_1} &= \textrm{Number of correctly classified cases in the nth tree using the OOB sample with permuted} \\
-&\qquad \textrm{values for} X_i
+C_{n,X_1} &= \textrm{Number of correctly classified cases in the nth tree using the OOB sample} \\
+&\qquad \textrm{with permuted values for} X_i
 \end{aligned}
@@
 
-#### Cforest Variable Importance
-
-Cforests are an alternative ensemble tree method that overcomes the aforementioned shortfalls of Random Forest scores.
-It is built from cTrees which are decision trees based on a conditional inference framework.
+#### CForest Variable Importance
+
+CForests are an alternative ensemble tree method that overcomes the aforementioned shortfalls of Random Forest scores.
+It is built from cTrees, which are decision trees based on a conditional inference framework.
 The key difference between cTrees and standard decision trees is that a significance test is used for splitting instead of a purity measure such as the Gini coefficient (which is centered around information gain).
@@
 Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable in the OOB sample, re-running the forest and comparing the decrease in response variable accuracy.
-Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) and Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].
+Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) and Boosted Trees [@Strobl2007; @Strobl2009; @Hothorn2006].
 
@@
 Less variables and a simpler model is easier for stakeholders to understand [@Weisberg2005].
-For these reasons, a subset of the most important variables were chosen before modelling.
+For these reasons, a subset of the most important variables was chosen before modeling.
 This was done by first eliminating outliers, then comparing important variables from ANOVA as well as Random Forests and cForests.
@@
 Between four and six clusters were chosen and summarised to find the maximum and minimum values within each cluster.
-For example, the diagram below illustrates the hierarchical dendrogram for timespan with 6 clusters boxed.
-
-![Hierarchical dendrogram of the Timespan variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
+For example, the diagram below illustrates the hierarchical dendrogram for time span with 6 clusters boxed.
+
+![Hierarchical dendrogram of the Time span variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
 
@@
 Once the number of clusters had been decided, each case was assigned a cluster label to replace its numeric variable.
-It was decided that for timespans and invoiced amounts, the discretised variables should be applied to all models (not just Bayesian Networks) because managers must guess these values initially when operating the model.
-It is easier for a manager to predict a timespan category than the exact number of days a project will last.
+It was decided that for time spans and invoiced amounts, the discretised variables should be applied to all models (not just Bayesian Networks) because managers must guess these values initially when operating the model.
+It is easier for a manager to predict a time span category than the exact number of days a project will last.
 For example, a small job could be confidently assigned to less than three weeks and a large job could be assigned to 1.5 - 3 years.
 
-The revised discretised categories for timespan and total amount invoiced were as follows:
-
-Timespan Categories
+The revised discretised categories for time span and total amount invoiced were as follows:
+
+Time span Categories
 
@@
 * 1.5 - 3 years
-* more than 3 years
+* More than 3 years
 
@@
 
-* the 'return per dollar' of each project (numerical) or 
-* whether each project made a 'profit' or 'loss' (categorical)
+* The 'return per dollar' of each project (numerical) or 
+* Whether each project made a 'profit' or 'loss' (categorical)
 
@@
 Its two axes are the false positive rate (FPR) on the x-axis and true positive rate (TPR) on the y-axis.
-In this problem a 'positive' is a loss-making job so a TP is a case when the model predicts a job will be loss making, and it did indeed lose money.
+In this problem a 'positive' is a loss making job so a TP is a case when the model predicts a job will be loss making, and it did indeed lose money.
 A false positive (FP) is when the model predicts a job will be loss making but in reality it was profitable.
@@
 
-For binary classification, AUC is a more meaningful statistic than classification accuracy when the output is a probability which can be applied to the problem.
+For binary classification, AUC is a more meaningful statistic than classification accuracy when the output is a probability that can be applied to the problem.
 The AUC indicates model performance across many thresholds while classification accuracy represents a single threshold.
@@
 All methods except gradient Boosted Trees and Naive Bayes could not handle missing data.
-Therefore, preliminary runs of each method used subsets of the data which had complete data.
+Therefore, preliminary runs of each method used subsets of the data that had complete data.
 It was possible that if missing data was imputed, predictions from the models that were limited by missing data could improve.
@@
 The MICE Random Forest method was chosen for imputation because it has been proven to work well with complex data sets [@Shah2014].
-It first performs a standard Random Forest imputation of missing values in the full dataset which are treated as initial 'place holders'.
+It first performs a standard Random Forest imputation of missing values in the full data set, which are treated as initial 'place holders'.
 The Random Forest imputation process will be explained in more detail later in this section.
@@
 
-Random forest imputation is a complex process which works by first doing a rough imputation of the missing values (using averages and majority votes).
+Random forest imputation is a complex process that works by first doing a rough imputation of the missing values (using averages and majority votes).
 Then a ten-tree forest is run to calculate proximities between all cases.
@@
 
-Once proximities are calculated for each pair of datapoints, missing values for a variable (say variable 'm') are filled.
-This is done for each datapoint by averaging the other complete 'm' values weighted by the proximity of the case representing each 'm' value to the current point.
-
-The final imputed dataset, which was created using the mice package and Random Forest imputation, was fed into a Boosted Tree algorithm.
-If similar predictive results were obtained using Boosted Trees imputed data and unimputed data, the imputed data must be reasonable.
-The imputed dataset was then trialed on the remaining methods and compared to un-imputed trials.
+Once proximities are calculated for each pair of data points, missing values for a variable (say variable 'm') are filled.
+This is done for each data point by averaging the other complete 'm' values weighted by the proximity of the case representing each 'm' value to the current point.
+
+The final imputed data set, which was created using the mice package and Random Forest imputation, was fed into a Boosted Tree algorithm.
+If similar predictive results were obtained using Boosted Trees imputed data and imputed data, the imputed data must be reasonable.
+The imputed data set was then trialed on the remaining methods and compared to unimputed trials.
 
@@
 New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big socially generated data.
-The prize-winning solution was a complex blend of sub-blends which interacted results of individual models with the original variables.
+The prize-winning solution was a complex blend of sub-blends that interacted results of individual models with the original variables.
 When original variables are used in blending they are called meta-features [@Sill2009].
@@
 Feature interaction is performed passively due to the nature of how trees are built.
-A split in a node determined by one variable is conditional upon the preceding split which was based on another variable and so on.
+A split in a node determined by one variable is conditional upon the preceding split, which was based on another variable and so on.
 Random Forests and Boosted Tree models are difficult to express as equations, but for the example problem, $X_1$, $X_2$, $X_3$, $M_1$, and $M_2$ would all be included as explanatory variables to predict $Y$.
@@
 A single training set for a blended model must be built from the test results of the individual models.
-Therefore, five-fold cross validation was performed using each individual model to create a full dataset as a compilation of test data results.
-The complete test data predictions were then added to the complete original dataset, and training and testing data could be partitioned to test the blended models.
+Therefore, five-fold cross validation was performed using each individual model to create a full data set as a compilation of test data results.
+The complete test data predictions were then added to the complete original data set, and training and testing data could be partitioned to test the blended models.
 A maximum of five blended models were created from a complete set of test results (built from the individual models), then a new set of test results were created for the next five blended models.
@@
 
-A simple approach was taken for this analysis, where projects with a probability to be a loss-making job greater than the threshold were rejected entirely.
+A simple approach was taken for this analysis, where projects with a probability to be a loss making job greater than the threshold were rejected entirely.
 Therefore all profits and losses from jobs above the threshold were discounted.
@@
 The aim was to find the optimal threshold point where saying 'no' to a job above that level would result in higher profits, because jobs that were likely to make a loss were being rejected.
-This chart will clarify what percentage of profit increase the company could expect by integrating the algorithm into decision making.
+This chart will clarify what percentage of profit increase the company could expect by integrating the algorithm into decision-making.
 It also provides a clear way to implement the algorithm and promote industry uptake of the research.
@@
 Again, 20 curves of each method were built and a power calculation was performed to achieve a power of 0.8.
-This determined the required sample size which was achieved by repeating 5-fold training/testing splits.
+This determined the required sample size, which was achieved by repeating 5-fold training/testing splits.
 This enabled the highest (optimal) threshold points of each curve to be statistically compared to one another.
@@
 <!-- * best were comapred to a model averaged version -->
-<!-- * Analysis of impact on bottom line for business if model is integrated into decision making -->
+<!-- * Analysis of impact on bottom line for business if model is integrated into decision-making -->
 <!-- * link: results -->
 
-# Chapter 5: Variable Selection
-
-To create the be st possible model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
+# Variable Selection
+
+To create the best possible model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
 Initially there were 34 explanatory variables for the single response variable, return per dollar, i.e.
@@
 
-Linear Regression must be fed complete datasets but the case study data set only had a core set of 13 complete variables (including the response variable).
+Linear Regression must be fed complete data sets but the case study data set only had a core set of 13 complete variables (including the response variable).
 If an incomplete variable was added to the core variables, the 'complete' data set would shrink to the size of the incomplete variable.
@@
 That meant a model was made for each incomplete variable, resulting in 21 separate ANOVA models.
-The p-values for the F-statistic of each variable is plotted below.
+The p-values for the F-statistic of each variable are plotted below.
 The core variables received p-values in all 21 models, while 'add.variable' represents p-values for the additional incomplete variable in each model.
@@
 
-* number of employees on the project team (no.users)
+* Number of employees on the project team (no.users)
 * Percent of hours completed by a professional as opposed to technical employee (pc.pro)
@@
 * Number of internal disciplines involved in the project (Num.disc)
-* Timespan of the project from first hours to final hours entered in timesheets (timespan)
+* Time span of the project from first hours to final hours entered in time sheets (time span)
 * Total amount invoiced for the project (inv.mlsto)
@@
 * Internal project category (JD.Second)
-* Position of the main professional working on the project, ie mid-level, senior etc. (ProjEng.Pos)
+* Position of the main professional working on the project, i.e. mid-level, senior etc. (ProjEng.Pos)
 * Id of the main professional working on the project (code.ProjEng)
@@
 
-The randomForest algorithm can process datasets with missing values, however it does so by automatically imputing missing values [@randomForest].
+The Random Forest algorithm can process data sets with missing values, however it does so by automatically imputing missing values [@randomForest].
 This was not desirable in the early stage of variable selection.
@@
 Additionally, the Random Forest algorithm cannot process categorical variables with more than 53 categories.
-This eliminated more variables and the resulting 'core' dataset contained 15 variables and 2364 cases.
+This eliminated more variables and the resulting 'core' data set contained 15 variables and 2364 cases.
 Because of the Random Forest algorithm's known biases and limitations with numbers of categories, a briefer approach to variable importance analysis was taken in comparison to regression and cForests.
@@
 
-Cforests were also run with the core 15 complete variables with 2364 complete cases.
+CForests were also run with the core 15 complete variables with 2364 complete cases.
 Incomplete variables were added to the core variables individually, each with their own separate run.
-Variables with an unlimited number of categories can be included in the cForest function which is an advantage over randomForests [@Hothorn2006; Strobl2007; Strobl2008].
-Cforests compute a variable importance ranking which is a number relating to the reduction in error the variable provides (similar method to the permutation importance in Random Forests).
+Variables with an unlimited number of categories can be included in the cForest function, which is an advantage over Random Forests [@Hothorn2006; @Strobl2007; @Strobl2008].
+CForests compute a variable importance ranking which is a number relating to the reduction in error the variable provides (similar method to the permutation importance in Random Forests).
 The variable importance rankings for the core 15 variables are as follows:
@@
 
-Four incomplete variables that were added indpendently ranked in the top 6 most important variables in their model:
+Four incomplete variables that were added independently ranked in the top 6 most important variables in their model:
 
@@
 
-Based on the results of the three methods for variable importance, 11 explanatory variables were selected for predictive modelling.
-They are listed in the table below along with the methods that highlighted each variable as important (randomForest = RF, cForest = CF):
+Based on the results of the three methods for variable importance, 11 explanatory variables were selected for predictive modeling.
+They are listed in the table below along with the methods that highlighted each variable as important (Random Forest = RF, cForest = CF):
 
@@
 | % of Hours Completed by a Professional-level Employee                | x     | x  | x  |
-| Timespan                                                             | x     | x  | x  |
+| Time span                                                             | x     | x  | x  |
 | Number of Employees on Project                                       | x     | x  | x  |
@@
 
-The variable importance results from the three types of models: ANOVA regression, randomForest, and cForest showed more similarities than differences. All three models ranked five variables highly:
+The variable importance results from the three types of models: ANOVA regression, Random Forest, and cForest showed more similarities than differences. All three models ranked five variables highly:
 
 * percent of hours by a professional
-* project timespan
+* project time span
 * number of employees in project team
@@
 This was the case with Discipline and it received a mid-level ranking.
-Business category of the client and Broad business type however, were ranked similarly to randomForests.
-
-The qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, client characteristics, project timespan, and project category.
+Business category of the client and Broad business type however, were ranked similarly to Random Forests.
+
+The qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, client characteristics, project time span, and project category.
 It was expected that Billing Type would play a more important role in predicting profitability as it describes the structure of the incoming fees.
@@
 For construction projects and IT software projects, the literature indicated size was the most important variable predicting effort.
-Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996][@Finnie1997][@Pai2013].
+Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996; @Finnie1997; @Pai2013].
 In this case study, profitability is the dependent variable, not total effort and the size variable plays a different role.
 Instead of contributing to predicting the total effort, it is describing whether the size of a project correlates with how profitable the project is.
-The closest estimate of size in this case study is the invoiced amount variable which did rank as very important.
+The closest estimate of size in this case study is the invoiced amount variable, which did rank as very important.
 The exact invoiced amount is difficult to guess before a project begins, and is part of the problem this study is addressing.
@@
 
-# Chapter 6: Regression Models
+# Regression Models
 
@@
 * Discipline 
-* Timespan
+* Time span
 * Team size
@@
 The second method was a more complicated procedure that required dozens of models but allowed variables with missing data to be included.
-First, a project was randomly selected (a row in the dataset) which would have certain variables complete.
+First, a project was randomly selected (a row in the data set) which would have certain variables complete.
 The data was then filtered for all projects that had at least the same columns complete, and an ANOVA model was built on this reduced data set.
@@
 The results of the ANOVA models were assessed using the root mean squared error (RMSE) statistic of the return per dollar predictions.
-These were compared against a baseline RMSE which was the RMSE of the predicted return per dollar values against the mean return per dollar of all projects.
+These were compared against a baseline RMSE, which was the RMSE of the predicted return per dollar values against the mean return per dollar of all projects.
 Then, the RMSE from each model was subtracted from the RMSE using mean return per dollar.
@@
 However, the mean difference is only \$`r mean(results.core$diff) %>% format(digits = 1)`.
-Furthermore, the mean RMSE in return per dollar is \$`r mean(results.core$final.RMSE) %>% format(digits = 2)` which is unacceptably high.
+Furthermore, the mean RMSE in return per dollar is \$`r mean(results.core$final.RMSE) %>% format(digits = 2)`, which is unacceptably high.
 Almost 50 cents is a large mean error when the value is estimating the profit from \$1 spent.
@@
 
-A student t-test indicates the difference is not significantly above 0 which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
+A student t-test indicates the difference is not significantly above 0, which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
 This method clearly performed poorly in comparison to the core variables.
-A reason for this could be that the mean number of cases in the dataset was `r results.core$test.nrow[1]*4` for the core variables but averaged only `r (mean(results.A2$test.nrow)*4) %>% round(0)` for the sampled set of all variables.
-Regression results from ANOVA models were poor - the randomForest algorithm was trialed next.
+A reason for this could be that the mean number of cases in the data set was `r results.core$test.nrow[1]*4` for the core variables but averaged only `r (mean(results.A2$test.nrow)*4) %>% round(0)` for the sampled set of all variables.
+Regression results from ANOVA models were poor - the Random Forest algorithm was trialed next.
 
@@
 Numeric predictions of return per dollar were attempted first with core variables that were largely complete followed by a method of sampling a mix of core and less complete variables in every run (similar to the two ANOVA methods).
-The randomForest parameters were first tuned using the caret package and optimal values mtry = 5 (the number of randomly selected explanatory variables to consider at each split) and ntree = 500 (the number of decision trees to ensemble) were determined [@caret].
+The Random Forest parameters were first tuned using the caret package and optimal values mtry = 5 (the number of randomly selected explanatory variables to consider at each split) and ntree = 500 (the number of decision trees to ensemble) were determined [@caret].
 A density plot of 50 models of the core variables using different training/test sets is shown below.
-The plot was created using the same procedure used in the ANOVA models where the x-axis measures the RMSE obtained by using the randomForest model subtracted from the RMSE using the mean of all project return per dollar values.
+The plot was created using the same procedure used in the ANOVA models where the x-axis measures the RMSE obtained by using the Random Forest model subtracted from the RMSE using the mean of all project return per dollar values.
 Note, only `r n.samples(data = results.rf2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
@@
 
-This plot shows a mean return per dollar RMSE improvement of \$`r mean(results.core$diff) %>% format(digits = 1)` which is the same as the ANOVA model.
+This plot shows a mean return per dollar RMSE improvement of \$`r mean(results.core$diff) %>% format(digits = 1)`, which is the same as the ANOVA model.
 The p-value for the null hypothesis that the difference is less than or equal to zero is `r t.test(results.rf.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)` and could be rejected.
-Results from 100 randomForest models using method 2 are shown below:
+Results from 100 Random Forest models using method 2 are shown below:
 
@@
 
-The p-value for these models and the afore-mentioned null hypothesis is `r t.test(results.rf2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
+The p-value for these models and the aforementioned null hypothesis is `r t.test(results.rf2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
 The null hypothesis was rejected which contrasted the matching ANOVA test.
@@
 
-Although the randomForest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for return per dollar was wide and the mean RMSE was high (\$`r mean(results.rf2$final.RMSE) %>% round(2)`).
+Although the Random Forest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for return per dollar was wide and the mean RMSE was high (\$`r mean(results.rf2$final.RMSE) %>% round(2)`).
 
@@
 
-Both ANOVA and randomForest models produced a similar lowest RMSE value, but performed differently when directly comparing models built on the same list of variables.
+Both ANOVA and Random Forest models produced a similar lowest RMSE value, but performed differently when directly comparing models built on the same list of variables.
 Both methods predicted at best within an average of \$0.48 for the response variable, 'return per dollar', using either core variables or a sample of core and less complete variables.
-Random Forest predictictions were much more accurate than ANOVA's with a mix of core (almost complete) and less variables while randomForest and ANOVA performed similarly with the core variables only.
-There was much less data available for training the models that used less-complete variables so it was interesting that randomForest performed considerably better (RMSE of \$0.50 for randomForest vs.
+Random Forest predictions were much more accurate than ANOVA's with a mix of core (almost complete) and less variables while Random Forest and ANOVA performed similarly with the core variables only.
+There was much less data available for training the models that used less-complete variables so it was interesting that Random Forest performed considerably better (RMSE of \$0.50 for Random Forest vs.
 \$2 for ANOVA).
@@
 Random Forest is also an ensemble method while ANOVA is not.
-Each randomForest is the average of 500 single trees, whereas each ANOVA model is a single model only.
-50 randomForest models produce 500 times more models than 50 ANOVA models so the averaging effect may have eliminated outliers for better overall performance.
+Each Random Forest is the average of 500 single trees, whereas each ANOVA model is a single model only.
+50 Random Forest models produce 500 times more models than 50 ANOVA models so the averaging effect may have eliminated outliers for better overall performance.
 
@@
 
-Further regression methods could have been pursued, however randomForest is a high-performing algorithm and was far from providing usable results.
+Further regression methods could have been pursued, however Random Forest is a high-performing algorithm and was far from providing usable results.
 It is likely that either not enough data was available or that the given variables did not capably explain all the factors affecting return per dollar.
@@
 
-This conclusion differs to case studies in the literature which tend to predict total cost/effort as a regression problem [@Kim2004][@Chan2005][@Akintoye2000].
-This is likely because the literature dealt with substantial projects which are rarer and impact governments and large companies.
+This conclusion differs to case studies in the literature, which tend to predict total cost/effort as a regression problem [@Kim2004; @Chan2005; @Akintoye2000].
+This is likely because the literature dealt with substantial projects that are rarer and impact governments and large companies.
 A simple profit or loss prediction would not be useful.
@@
 
-# Chapter 7: Binary Classification Models
+# Binary Classification Models
 
@@
 Boosted Trees and Naive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
-Logistic Regression, Random Forest and Bayesian Networks require datasets without missing values so a subset of complete data was used.
+Logistic Regression, Random Forest and Bayesian Networks require data sets without missing values so a subset of complete data was used.
 It was expected that Boosted Trees would perform best as they are known to have powerful predictive capabilities and can use incomplete data.
@@
 
-The models were compared using the AUC test statistic calculated from the ROC as this gives a good indication of model performance under binary classification [@pROC][@Putler2012].
+The models were compared using the AUC test statistic calculated from the ROC as this gives a good indication of model performance under binary classification [@pROC; @Putler2012].
 An example plot of five ROC's built from 5-fold cross validation using the Boosted Tree algorithm is shown below.
@@
 
-* shrinkage = 0.001
-* number of trees = 10,000
-* minimum terminal node size = 20
-* maximum tree depth = 5
+* Shrinkage = 0.001
+* Number of trees = 10,000
+* Minimum terminal node size = 20
+* Maximum tree depth = 5
 
@@
 
-![ROC's for five boosted tree models built on different train/test partitions of the dataset [@pROC]](images/boosted tree/5_fold_ROC.png)
-
-In the plot above, the average AUC was 0.763 which indicated that altogether, taking into account a range of possible thresholds, the model was predicting profit/loss better than random chance (AUC = 0.5) but not perfectly (AUC = 1).
+![ROC's for five boosted tree models built on different train/test partitions of the data set [@pROC]](images/boosted tree/5_fold_ROC.png)
+
+In the plot above, the average AUC was 0.763, which indicated that altogether, taking into account a range of possible thresholds, the model was predicting profit/loss better than random chance (AUC = 0.5) but not perfectly (AUC = 1).
 20 models and their resulting AUC statistics from each method were initially determined.
 These were used to calculate how many models were required to achieve a minimum statistical power of 0.8 when compared to the Boosted Tree results.
-The comparison of logistic regression required 100 models, while randomForest required over 700 because the AUC distributions were close.
-Bayesian Networks and Naive Bayes did not perform as well and the power calculation revealed only 5 and 8 models were required respectively.
+The comparison of logistic regression required 100 models, while Random Forest required over 700 because the AUC distributions were close.
+Bayesian Networks and Naive Bayes did not perform as well and the power calculations revealed only 5 and 8 models were required respectively.
 Bayesian Networks were then dropped as Naive Bayes performed slightly better and were faster and simple enough to work as a baseline comparison to the higher-performing methods.
 
-It was finally decided that randomForest and Boosted Trees would likely not be distinguishable since the power calculation indicated that 700 models were required.
+It was finally decided that Random Forest and Boosted Trees would likely not be distinguishable since the power calculation indicated that 700 models were required.
 Hence, 100 models of each method were built as calculated for logistic regression.
@@
 
-The AUC performance of Logistic Regression and the randomForest algorithms cannot be statistically differentiated from Boosted Trees.
-
-Next, data imputation methods were trialed which would make the complete data set available to Logistic Regression and randomForests.
+The AUC performance of Logistic Regression and the Random Forest algorithms cannot be statistically differentiated from Boosted Trees.
+
+Next, data imputation methods were trialed which would make the complete data set available to Logistic Regression and Random Forests.
 It was possible this would improve AUC results and the same process as above could be repeated.
-Imputation was done using the MICE method with randomForest imputation [@mice].
+Imputation was done using the MICE method with Random Forest imputation [@mice].
 Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for Random Forest unless over 450,000 models were made.
@@
 
-Boosted Trees, Logistic Regression, and randomForest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed Boosted Trees.
+Boosted Trees, Logistic Regression, and Random Forest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed Boosted Trees.
 The results from each method using reduced data sets (to cater for missing values) alongside results using imputed complete data are shown below.
@@
 Logistic regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
-This may explain Logistic Regression's comparitively high performance on the case study's smaller dataset with many variables.
+This may explain Logistic Regression's comparatively high performance on the case study's smaller data set with many variables.
 
@@
 For example, many of the imputed variables were categorical such as job category, percent of job completed by a professional level employee, and the position of the majority worker on a job.
-The benefit of having an imputed dataset that performs acceptably well is that each model is fed the full data set and therefore produces predictions for the full data.
-Beforehand, randomForest and Logistic Regression models were limited to a complete set of original data which implied a reduced data set.
+The benefit of having an imputed data set that performs acceptably well is that each model is fed the full data set and therefore produces predictions for the full data.
+Beforehand, Random Forest and Logistic Regression models were limited to a complete set of original data, which implied a reduced data set.
 With complete model predictions, the high performing models could be blended to test whether better predictions were possible.
@@
 
-In summary, the individual binary classification models performed well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a new job will be profitable or loss-making.
+In summary, the individual binary classification models performed well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a new job will be profitable or loss making.
 Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model actually affects decisions in practice.
@@
 
-# Chapter 8: Model Blending
+# Model Blending
 
@@
 Blended models combine the predictions from each higher-performing individual model to create averaged or 'blended' predictions.
-In this case, the randomForest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialed ranging from simple to complex.
+In this case, the Random Forest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialed ranging from simple to complex.
 The simplest way averaged the probability output (a number between 0 and 1) of all three models.
 The next two simplest methods consisted of building a Logistic Regression and Boosted Trees model from the three individual model outputs only.
-Feature Weighted Linear Stacking (FWLS), randomForest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (meta-features).
+Feature Weighted Linear Stacking (FWLS), Random Forest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (meta-features).
 The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature conditions could be weighted as such.
@@
 
-Variable assessments were completed and compared using cForests, Logistic Regression, randomForests and Boosted Trees.
-Cforests have the favourable characteristic of being unbiased while the remaining methods were used for the blending process.
+Variable assessments were completed and compared using cForests, Logistic Regression, Random Forests and Boosted Trees.
+CForests have the favourable characteristic of being unbiased while the remaining methods were used for the blending process.
 The results are summarised in the sections below.
@@
 The cForest highlighted the results from the Logistic Regression and Boosted Trees as the most important variables.
-This was followed by the results from the randomForest model and the project team size, project category, total invoiced amount (category) and timespan (category).
+This was followed by the results from the Random Forest model and the project team size, project category, total invoiced amount (category) and time span (category).
 
@@
 
-#plot p values
+#plot p-values
 # ggplot(data=log.varimp, aes(x=rowname, y=P_value)) +
@@
 
-* Timespan * results of logistic regression model
+* Time span * results of logistic regression model
 * Total amount invoiced * results of logistic regression model
-* Total amount invoiced * results of randomForest model
+* Total amount invoiced * results of Random Forest model
 * Client Business type * results of Boosted Tree model 
@@
 * Total amount invoiced for the job
-* Timespan
+* Time span
 * Team size
@@
 
-#plot p values
+#plot p-values
 ggplot(data=rf.varimp, aes(x=rowname, y=MeanDecreaseAccuracy)) +
@@
 
-Permutation importance by randomForest rated all three prediction models as most important followed by team size for the project, percent of hours done by the chief employee on the project, timespan (category), project category, and client business type.
+Permutation importance by Random Forest rated all three prediction models as most important followed by team size for the project, percent of hours done by the chief employee on the project, time span (category), project category, and client business type.
 
@@
 
-#plot p values
+#plot p-values
 ggplot(data=boost.varimp, aes(x=rowname, y=rel.inf)) +
@@
 ```
-Boosted Trees ranked output from all three individual prediction models highest (Logistic Regression ranked highest) followed by Project category, Client Business type, and timespan (category).
+Boosted Trees ranked output from all three individual prediction models highest (Logistic Regression ranked highest) followed by Project category, Client Business type, and time span (category).
 
@@
 * Team size on a project
-* Timespan
+* Time span
 
@@
 Only three of the model blending techniques used the 8 variables determined above.
-These were FWLS, randomForest, and Boosted Trees.
+These were FWLS, Random Forest, and Boosted Trees.
 The remaining 3 techniques used only the predictions of the three individual models.
@@
 The results of the 150 iterations are shown in the plot below.
-Simple model averaging resulted in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logisitc regression blended model had a two-sample t-test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
+Simple model averaging resulted in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logistic regression blended model had a two-sample t-test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
 
@@
 
-The simple Logistic Regression blend and simple averaging of the individual models (logistic regression, Boosted Trees and Random Forests) gave virtually the same distribution of results, where both AUC distributions were signficantly greater than the original logistic regression model.
+The simple Logistic Regression blend and simple averaging of the individual models (logistic regression, Boosted Trees and Random Forests) gave virtually the same distribution of results, where both AUC distributions were significantly greater than the original logistic regression model.
 The p-values from the two-sample t-tests were `r pvals$p_value[1]` and `r pvals$p_value[2]` for simple average and simple Logistic Regression methods respectively.
@@
 
-As shown in the table, if the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model prediciting a loss-making job increased by 56.0%.
+As shown in the table, if the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model predicting a loss making job increased by 56.0%.
 The weightings contrasted the other simple model that averaged the original probabilities (i.e.
@@
 
-# Chapter 9: Extended Analysis
+# Extended Analysis
 
@@
 The first research question, whether machine learning and statistical techniques can predict project profitability based on historic data, has been answered.
-The measure of profitability could not be modelled but simplifying the profitability prediction to a binary (profit/loss) model was successful.
+The measure of profitability could not be modeled but simplifying the profitability prediction to a binary (profit/loss) model was successful.
 To answer the second research question, whether the model would improve the business' profitability, a profit analysis was performed.
 
-The models output a 'probability' between 0 and 1 that each project will be a loss-making job (probability = 1 indicates a loss-making job).
+The models output a 'probability' between 0 and 1 that each project will be a loss making job (probability = 1 indicates a loss making job).
 This leaves the question, at what probability would a decision-maker round the probability up to 1 or down to 0?
@@
 ```{r}
-# Results - over 5 pages, up to 20 pages
+# Results - over 5-pages, up to 20 pages
 # 
@@
 The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
-That is, it cannot be said with 95% confidence that the original methods' would produce a profit higher than historical profits if projects higher than the optimal thresholds (jobs that are likely to be loss-making) were rejected.
+That is, it cannot be said with 95% confidence that the original methods' would produce a profit higher than historical profits if projects higher than the optimal thresholds (jobs that are likely to be loss making) were rejected.
 Clearly this is not helpful for the case-study business to adopt.
@@
 The complex blended models had mixed results.
-FWLS and randomForests performed more poorly than the individual methods while Boosted Trees produced a marginally higher profitability ratio (`r best.fin %>% filter(method == "comp.boost") %>% .$mean_profit_ratio`) but the lower bound of its confidence interval was `r best.fin %>% filter(method == "comp.boost") %>% .$lower_bound_95CI`, i.e.
+FWLS and Random Forests performed more poorly than the individual methods while Boosted Trees produced a marginally higher profitability ratio (`r best.fin %>% filter(method == "comp.boost") %>% .$mean_profit_ratio`) but the lower bound of its confidence interval was `r best.fin %>% filter(method == "comp.boost") %>% .$lower_bound_95CI`, i.e.
 not above original profits.
@@
 Then, the confusion matrix for each set of results at their respective thresholds could be calculated.
-The plot below shows the breakdown of predictions by each method on actual loss-making projects.
-If the project was predicted correctly as loss-making, this was a true positive, and the job was rejected (in the fabricated scenario).
+The plot below shows the breakdown of predictions by each method on actual loss making projects.
+If the project was predicted correctly as loss making, this was a true positive, and the job was rejected (in the fabricated scenario).
 If it was predicted incorrectly as a profit-making job, this was a false negative, and the project was completed, resulting in a loss for the business.
@@
                           name = "Confusion Matrix Reference") +
-        labs(x = "Method (Original or Blended)", y = "Percent of All Loss-Making Projects (%)",
-             title = "Breakdown of True Positives and False Negatives for Loss-Making Jobs") +
+        labs(x = "Method (Original or Blended)", y = "Percent of All Loss Making Projects (%)",
+             title = "Breakdown of True Positives and False Negatives for Loss Making Jobs") +
         scale_x_discrete(labels = c("average" = "Simple Av",
@@
 Choosing the right probability threshold point is about finding the right balance point in an imperfect model.
-In this case, projects above a threshold point will include profitable and loss-making jobs.
-The optimal point is where the savings from rejecting loss-making jobs (TP) outweigh the lost profits from rejecting profitable jobs (FP) the most.
+In this case, projects above a threshold point will include profitable and loss making jobs.
+The optimal point is where the savings from rejecting loss making jobs (TP) outweigh the lost profits from rejecting profitable jobs (FP) the most.
 All models were capable of different TP and FN ratios at alternate threshold points.
@@
 The plot above shows the Complex Boosted Tree Blend has the highest TPR followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
-All of these TPRs are below 25% which shows it was not worth catching a large percentage of loss-making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
-This implies the models were not able to very cleanly divide loss-making jobs from profitable jobs.
+All of these TPR's are below 25% which shows it was not worth catching a large percentage of loss making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
+This implies the models were not able to very cleanly divide loss making jobs from profitable jobs.
 
@@
 If a job was predicted to be profitable, and was therefore accepted, it was classified TN.
-If a job was loss-making but labelled profitable, the business would complete the job and make a loss (FN).
+If a job was loss making but labelled profitable, the business would complete the job and make a loss (FN).
 
@@
 The increased number of FP (i.e.
-rejected profitable jobs) offset the more accurate rejection of loss-making jobs so much that the simple averaged blend and simple logisitc regression blend outperformed it.
-The afore-mentioned simple blends performed similarly, with a higher than average TPR and average FPR in comparison to other methods.
-The complex Random Forest blend and FLWS both had lower TPRs than the simple blends but higher TN rates.
-In the end, the improved TN rate did not outweight the lower TPR, i.e.
-the higher rate of accepting profitable jobs did not outweight the lower rate of accurately rejecting loss-making jobs.
-
-All individual original methods used substantially lower TPRs at their optimal thresholds, while their TN rates were spread out.
+rejected profitable jobs) offset the more accurate rejection of loss making jobs so much that the simple averaged blend and simple logistic regression blend outperformed it.
+The aforementioned simple blends performed similarly, with a higher than average TPR and average FPR in comparison to other methods.
+The complex Random Forest blend and FLWS both had lower TPR's than the simple blends but higher TN rates.
+In the end, the improved TN rate did not outweigh the lower TPR, i.e.
+the higher rate of accepting profitable jobs did not outweigh the lower rate of accurately rejecting loss making jobs.
+
+All individual original methods used substantially lower TPR's at their optimal thresholds, while their TN rates were spread out.
 This means, as the trialed threshold values moved, and the TPR improved, the increasing lost profits from FP's was too great to justify a TPR higher than 4.6%.
-Blending the individual methods has made use of their unique strengths and seems to have clustered the probabilities of the actual loss-making jobs more accurately towards 1, and the profitable jobs towards 0.
-This allowed for higher TPRs with lower FPRs at the blended methods' optimal thresholds.
+Blending the individual methods has made use of their unique strengths and seems to have clustered the probabilities of the actual loss making jobs more accurately towards 1, and the profitable jobs towards 0.
+This allowed for higher TPR's with lower FPR's at the blended methods' optimal thresholds.
 
@@
 
-The profit curves and TP/FP rates give a good indication of overall performance of the methods, however an analysis of the characteristics of the predicted loss-making projects can provide more detailed insight.
+The profit curves and TP/FP rates give a good indication of overall performance of the methods, however an analysis of the characteristics of the predicted loss making projects can provide more detailed insight.
 
@@
 
-```{r, fig.cap = c("Proportion of profitable and loss-making jobs in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+```{r, fig.cap = c("Proportion of profitable and loss making jobs in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
 
@@
              y = "Relative Frequency",
-             title = "Ratio of Profitable vs. Loss-Making Projects") +
+             title = "Ratio of Profitable vs. Loss Making Projects") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-Originally, 25% of projects were loss-making.
-The model rejects a population of projects which are over 60% loss-making.
+Originally, 25% of projects were loss making.
+The model rejects a population of projects which are over 60% loss making.
 This represents a significant swing in proportions, but is not perfect.
-The class imbalance towards profitable jobs in the initial dataset may have weakened the model's ability to classify losses.
+The class imbalance towards profitable jobs in the initial data set may have weakened the model's ability to classify losses.
 @Saradhi2011 took advantage of SVM's ability to adjust class penalties to address class imbalance, however SVM's were avoided because of their black box predictions in this work and the remaining methods cannot accommodate class penalties.
@@
 
-```{r, fig.height = 5, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss-making jobs in the rejected projects only.
+```{r, fig.height = 5, fig.cap = c("Proportion of the absolute value of profits and losses from profitable and loss making jobs in the rejected projects only.
 Colours are alternated between different jobs to indicate profit/loss magnitude.
@@
 
-The distribution of timespans of rejected projects differed significantly to the timespan distribution of all projects.
-The chart below shows that proportionally very few short jobs were rejectd, i.e.
+The distribution of time spans of rejected projects differed significantly to the time span distribution of all projects.
+The chart below shows that proportionally very few short jobs were rejected, i.e.
 projects that lasted between 1 day and 2.5 months.
@@
 They may face an opposite challenge to the small projects.
-On the other hand, large projects can provide long term security for a small business and it is unlikely the company would want to reject them.
+On the other hand, large projects can provide long-term security for a small business and it is unlikely the company would want to reject them.
 It may make more sense to adjust the contract structure of large projects that have been highlighted as a risk.
@@
 The statistical distribution of return per dollar in the rejected projects is skewed towards negative returns.
-This means the algorithm has identified a proportionally much higher percentage of loss-making projects which is not surprising given previous results.
+This means the algorithm has identified a proportionally much higher percentage of loss making projects which is not surprising given previous results.
 
@@
 
-The detailed analysis of rejected project characteristics gave insight into the characteristics of projects that are likely to be loss-making as well as the accuracy of the algorithm.
-While it is interesting to observe characeristics of rejected projects, a complex combination of all the features contribute to the final probability of loss prediction.
+The detailed analysis of rejected project characteristics gave insight into the characteristics of projects that are likely to be loss making as well as the accuracy of the algorithm.
+While it is interesting to observe characteristics of rejected projects, a complex combination of all the features contribute to the final probability of loss prediction.
 
@@
 To conclude, because of the promising AUC results, it was logical that the models translated into financial benefits for the company.
-The `r (best.mean-100)`% improvement in profits looks reasonable, and may be high enough to trigger further cost-benefit analyses before finally introducing the model into business decisons.
+The `r (best.mean-100)`% improvement in profits looks reasonable, and may be high enough to trigger further cost-benefit analyses before finally introducing the model into business decisions.
 Development of the user interface as well as some further data analyses are recommended before implementation of the model and are described in the 'User Interface and Engagement' and 'Limitations and Future Work' sections below.
@@
 
-The manager selects the charactersitics that describe their new job in the top panel.
-Then once the 'Calculate' button is clicked, the blended Logistic Regression model works behind the scenes to generate a probability score between 0 and 1 where 1 represents a loss-making job.
+The manager selects the characteristics that describe their new job in the top panel.
+Then once the 'Calculate' button is clicked, the blended Logistic Regression model works behind the scenes to generate a probability score between 0 and 1 where 1 represents a loss making job.
 This output is not intuitive to understand for a manager because the probability score does not represent a statistical probability, rather a score the algorithm generated using its own measures.
-The optimal threshold analysis found that scores below 0.6 should be classified as profitable jobs and scores above 0.6 are loss-making.
+The optimal threshold analysis found that scores below 0.6 should be classified as profitable jobs and scores above 0.6 are loss making.
 
@@
 Understanding the plot would require some training.
-The y-axis represents the probability a project will be loss-making after taking into account average effects from variables other than the x-axis variable.
+The y-axis represents the probability a project will be loss making after taking into account average effects from variables other than the x-axis variable.
 
@@
 
-First, succesfully estimating project variables is critical in translating this research into tangible business benefits.
+First, successfully estimating project variables is critical in translating this research into tangible business benefits.
 The data represents the company's records of *completed* projects so initially expected variable values were not documented.
-For example, if a project was originally expected to last 1.5 years but instead took 2, the company data only retains the two year timespan.
+For example, if a project was originally expected to last 1.5 years but instead took 2, the company data only retains the two year time span.
 This prompts the question, how well could a user predict the variables before beginning a project?
-Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e.
+Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'time span' to 5 or 6 broad categories (i.e.
 "1 day - 3 weeks", "3 weeks - 2.5 months", etc).
@@
 
-# Chapter 10: Conclusion
+# Conclusion
 
@@
 The general aim was to use statistical techniques to model the profitability of projects for a case study consulting business using their internal CRM data.
-This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitablility and as a binary classification problem predicting simply profit or loss.
+This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitability and as a binary classification problem predicting simply profit or loss.
 Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees and multiple methods of blending the individual models.
@@
 
-Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom l ine) of the case study business.
+Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
 This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business scenario.
-The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss-making'.
+The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss making'.
 A range of threshold were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
 If a project was accepted, the profits and losses were summed while if a project was rejected the profits and losses were forfeited.
-A perfect model would reject all loss-making jobs and accept all profitable jobs, thereby improving overall profits.
+A perfect model would reject all loss making jobs and accept all profitable jobs, thereby improving overall profits.
 Final results showed the Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
@@
 Another topic identified for future research was to test how well managers estimate project input variables.
-Estimation of the timespan, team size, total invoiced amount, and percentage of hours completed by professionals in particular should be tested.
+Estimation of the time span, team size, total invoiced amount, and percentage of hours completed by professionals in particular should be tested.
 Initial estimates of input variables could also be kept in the database for future analysis.
-Overall, this project has successfully demonstrated a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models could guide managers in rejecting loss-making projects and improve net profitability.
+Overall, this project has successfully demonstrated a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models could guide managers in rejecting loss making projects and improve net profitability.
 

### 2016-07-28 finalwriteup/trial_code_eq.rmd
finished kerrie editing

@@
 
-$$
-\begin{aligned} 
-\log(odds) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
-\log\left(\frac{p}{1-p}\right) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
-\log\left(\frac{0.5}{0.5}\right) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
-\log(1) &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
-0 &= \beta_0 + X_1\beta_1 + X_2\beta_2 \\ 
-X_2\beta_2 &= -\beta_0 - X_1\beta_1 \\ 
-X_2 &= -\frac{\beta_0}{\nbeta_2} - X_1\frac{\beta_1}{\beta_2} \\ 
-\end{aligned}
-$$
+$p(ABCD) = p(C|ABD)\cdot p(A|BD) \cdot p(B|D) \cdot p(D)$
 
-$p(X_1 X_2|Y) = p(X_1|Y) \cdot p(X_2|X_1 Y)$
-
-$\boldsymbol{w} \cdot \boldsymbol{x} + \beta_0 = 0$
-
-where $\boldsymbol{w}$ and $\boldsymbol{x}$ are vectors with lengths corresponding to the number of explanatory variables.
-For two variables, $\boldsymbol{w}$ would be $\begin{pmatrix}\beta_1 \\ \beta_2\end{pmatrix}$ and $\boldsymbol{w}$ would be $\begin{pmatrix}X_1 \\ X_2\end{pmatrix}$ [@Fletcher2009].
-Expanding the dot product of these two vectors gives the equation for a line:
-
-$\beta_1 X_1 + \beta_2 X_2 + \beta_0 = 0$
+$p(ABCD) = p(C|BD)\cdot p(A|B) \cdot p(B) \cdot p(D)$

### 2016-08-01 finalwriteup/draft.rmd
almost finished reading results and method

@@
 Instead, it was more logical to predict the profitability of a project i.e.
-return per dollar.
+'return per dollar'.
 This measure represents the ratio of the amount invoiced to business costs, and can easily be compared between projects of different sizes.
@@
 For further explanation of the method behind decision trees and ensemble trees, three excellent sources are @Hastie2009, @Elith2008, and @Breiman2005.
-To date, Linear Regression, Neural Networks, Case Based Reasoning, and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks and generally outperform Linear Regression [@Caruana2006].
+To date, Linear Regression, Neural Networks, Case Based Reasoning (CBR), and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks and generally outperform Linear Regression [@Caruana2006].
 Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots that provide insight into the model's calculations.
@@
 It includes first how the data was obtained, the lengthy cleaning process, followed by variable importance analysis, variable selection, trials of selected predictive algorithms and problem constructs.
-The cost estimation problem was attempted first as a regression problem predicting return per dollar, followed by a simpler binary classification problem predicting profit or loss.
+The cost estimation problem was attempted first as a regression problem predicting 'return per dollar', followed by a simpler binary classification problem predicting profit or loss.
 Once the best methods were selected, they were blended in numerous ways, using both simple averaging techniques and sophisticated machine-learning algorithms.
@@
 Research in the construction industry has primarily focused on predicting the final building cost of construction.
-Some studies reported surveyed cost estimation practices while others tested new methods such as building information modeling (BIM) analysis, case based reasoning, and predictive algorithms.
-The cost estimation methods of detailed analysis (the 'inside view'), Case Based Reasoning, and predictive algorithms are reviewed in this section.
+Some studies reported surveyed cost estimation practices while others tested new methods such as building information modeling (BIM) analysis, CBR, and predictive algorithms.
+The cost estimation methods of detailed analysis (the 'inside view'), CBR, and predictive algorithms are reviewed in this section.
 
@@
 Detailed analysis remains the most prevalent method for construction cost estimation despite the industry having a long history of projects running over time and budget using the same technique [@Shane2009].
-Next, the method of detailed sums shall be compared to results from parametric algorithms and case based reasoning ('outside view').
+Next, the method of detailed sums shall be compared to results from parametric algorithms and CBR ('outside view').
 
@@
 
-#### Case Based Reasoning in the Construction Industry
-
-Case based reasoning (CBR) is a systematic method of expert judgment, where the decision maker manually compares similar projects from his or her experience [@Shepperd1996].
+#### CBR in the Construction Industry
+
+CBR is a systematic method of expert judgment, where the decision maker manually compares similar projects from his or her experience [@Shepperd1996].
 The idea is that in the case of a new project, similar projects are chosen, before generalisations about the data are made.
 This contrasts to algorithmic models that are built from the entire set of data [@Elfaki2014].
-Research into the application of case based reasoning to construction cost estimation began in the 1980's [@Kim2004].
+Research into the application of CBR to construction cost estimation began in the 1980's [@Kim2004].
 The general procedure involves:
@@
 A parallel can be drawn between software development effort estimation and consultants' effort estimation in the construction industry - which was highlighted as a gap in the previous section.
-Plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modeling, and Case Based Reasoning.
+Plenty of research has been dedicated to the problem of effort prediction, with methods that can be categorised similarly to the construction industry: detailed analysis, algorithmic modeling, and CBR.
 
@@
 
-#### Case Based Reasoning in the IT Industry
+#### CBR in the IT Industry
 
@@
 Previous studies addressing the problem have generally been limited to prediction of construction costs for buildings/infrastructure and effort estimation for IT projects.
-The prediction models consisted of Case Based Reasoning, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM.
+The prediction models consisted of CBR, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM.
 This leaves a gap in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods.
@@
 
-* Return per dollar as the measure of profitability
+* 'Return per dollar' as the measure of profitability
 
@@
 Other bigger outlier projects had unrealistically low hours logged against them which could have been the result of data entry errors or errors from when the data was transferred across databases about 4 years prior to data extraction.
-Based on the employee's domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were revised.
+Based on the employee's domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of 'return per dollar' values were revised.
 
@@
 
-To create the best possible model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
-Initially there were 34 explanatory variables for the single response variable, return per dollar, i.e.
-the chosen measure of profitability for a project.
-By the end of variable selection, only 11 explanatory variables remained.
-As stated in the previous chapter, three predictive methods were selected which output variable importance scores which differ due to their unique theoretical foundations.
-These were Linear Regression, Random Forests, and cForests.
+To create a high-performing model, only important variables were selected in order to reduce the effect of noisy explanatory variables and enhance simplicity and model comprehension.
+Initially there were 34 explanatory variables for the single response variable, 'return per dollar'.
+By the end of variable selection, only 11 remained.
+As stated in the previous chapter, three predictive methods were selected that output variable importance scores.
+These were Linear Regression, Random Forests, and cForests, which have unique theoretical foundations.
 
@@
 
-Linear Regression must be fed complete data sets but the case study data set only had a core set of 13 complete variables (including the response variable).
-If an incomplete variable was added to the core variables, the 'complete' data set would shrink to the size of the incomplete variable.
+Linear Regression must be fed complete data sets but the case study data had a core set of only 13 complete variables (including the response variable).
+If an incomplete variable was added to the core variables, the 'complete' data set would shrink to the complete rows of the incomplete variable.
 If another incomplete variable was added, the complete set of data would shrink again.
@@
 The p-values for the F-statistic of each variable are plotted below.
-The core variables received p-values in all 21 models, while 'add.variable' represents p-values for the additional incomplete variable in each model.
-
-```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of an ANOVA regression's F-statistics used to interpret variable importance")}
+The core variables received p-values in all 21 models, while 'add.variable' represents the p-values for the unique additional incomplete variable in each model.
+
+```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of ANOVA regressions' F-statistics which were used to interpret variable importance")}
 
@@
         scale_y_continuous(breaks = c(0, 0.05, .25, .5, .75, 1)) +
-        labs(y = "P-value of F-statistic", x = "Independent Variables",
-             title = "Summary of variable F-statistic P-values from ANOVA models")
-
-```
-
-The ANOVA models indicated that the following variables have a median p-value below 0.05 and therefore significantly contribute to the rejection of the null hypothesis:
+        labs(y = "P-Value of F-Statistic", x = "Independent Variables",
+             title = "Summary of Variable F-Statistic P-Values from ANOVA Models")
+
+```
+
+The ANOVA models indicated that the following variables have a median p-value below 0.05 and therefore significantly contribute to the rejection of the regression null hypothesis:
 
@@
 
-The Random Forest algorithm can process data sets with missing values, however it does so by automatically imputing missing values [@randomForest].
+The Random Forest algorithm can process data sets with missing values, however it does so by automatically imputing those missing values [@randomForest].
 This was not desirable in the early stage of variable selection.
@@
 Additionally, the Random Forest algorithm cannot process categorical variables with more than 53 categories.
-This eliminated more variables and the resulting 'core' data set contained 15 variables and 2364 cases.
-Because of the Random Forest algorithm's known biases and limitations with numbers of categories, a briefer approach to variable importance analysis was taken in comparison to regression and cForests.
-
-Each tree in a Random Forest is created with a bootstrapped training sample of the data.
+This eliminated more variables and the resulting 'core' data set contained 15 variables and 2364 complete cases.
+Because of the Random Forest algorithm's known biases and limitations with numbers of categories, a briefer approach to variable importance analysis was taken in comparison to Regression and cForests.
+
+As explained in the Literature Review, each tree in a Random Forest is created with a bootstrapped training sample of the data.
 The random sampling means the results of each forest are slightly different.
@@
 
-## Cforest
-
-CForests were also run with the core 15 complete variables with 2364 complete cases.
+## CForest
+
+CForests were also run using the core 15 complete variables with 2364 complete cases.
 Incomplete variables were added to the core variables individually, each with their own separate run.
 Variables with an unlimited number of categories can be included in the cForest function, which is an advantage over Random Forests [@Hothorn2006; @Strobl2007; @Strobl2008].
-CForests compute a variable importance ranking which is a number relating to the reduction in error the variable provides (similar method to the permutation importance in Random Forests).
-The variable importance rankings for the core 15 variables are as follows:
+CForests compute a variable importance ranking which is a score relating to the reduction in error that the variable provides (a similar method to the permutation importance in Random Forests).
+The variable importance rankings for the core 15 variables were as follows:
 
@@
         geom_bar(stat='identity') +
-        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "cForest Variable Importance Rankings of Complete Variables") +
+        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "CForest Variable Importance Rankings of Complete Variables") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-Four incomplete variables that were added independently ranked in the top 6 most important variables in their model:
+Four incomplete variables that were added individually ranked in the top 6 most important variables in their model:
 
@@
 
-The variable importance results from the three types of models: ANOVA regression, Random Forest, and cForest showed more similarities than differences. All three models ranked five variables highly:
-
-* percent of hours by a professional
-* project time span
-* number of employees in project team
-* total amount invoiced
-* business category of client
-
-This was not necessarily expected, particularly between ANOVA and the ensemble tree methods as ANOVA measures linear relationships in contrast to non-linear ensemble trees.
-Random forests favour variables that are numeric or have many categories and this held true for the analysis.
-The 6 highest-ranked variables were numeric with one categorical variable, Business category of client, which had 28 categories.
+The variable importance results from the three types of models: ANOVA Regression, Random Forest, and cForest showed more similarities than differences. All three models ranked five variables highly:
+
+* Percent of hours by a professional
+* Project time span
+* Number of employees in project team
+* Total amount invoiced
+* Business category of client
+
+This was not necessarily expected, particularly between ANOVA and the ensemble tree methods, because ANOVA measures linear relationships in contrast to non-linear ensemble trees.
+
+Random Forests favour variables that are numeric or have many categories and this held true for the analysis.
+Its 6 highest-ranked variables were numeric with one categorical variable, Business category of the client, which had 28 categories.
 The two lowest ranking variables had only 4 categories.
@@
 
-The qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, client characteristics, project time span, and project category.
+The overall qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, client characteristics, project time span, and project category.
 It was expected that Billing Type would play a more important role in predicting profitability as it describes the structure of the incoming fees.
@@
 
-For construction projects and IT software projects, the literature indicated size was the most important variable predicting effort.
+For IT software projects, the literature indicated size was the most important variable predicting effort, and in Construction projects size was also often an explanatory variable.
 Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996; @Finnie1997; @Pai2013].
 In this case study, profitability is the dependent variable, not total effort and the size variable plays a different role.
-Instead of contributing to predicting the total effort, it is describing whether the size of a project correlates with how profitable the project is.
+Instead of predicting total effort, it is describing whether the size of a project correlates with how profitable the project is.
 The closest estimate of size in this case study is the invoiced amount variable, which did rank as very important.
@@
 The other variables indicated as important in this study (including the number of employees on a project, client business type and percent of hours by certain types of employees) have not been strongly represented in the literature to date.
-This is because the variables were easily calculated from internal CRM data but would not be easily obtained externally from projects across many companies or regions.
-
-The 11 explanatory variables from table __ were progressed to the next stage where they were used to build predictive models.
-The outcomes of the regression and categorical models are summarised and discussed in the following two chapters.
+This is because the variables were easily calculated from internal CRM data but would not be easily obtained from an external standpoint for projects across many companies or regions.
+
+The 11 explanatory variables from table __ were progressed to the next stage where they were used to build the first set of predictive models.
+The outcomes of the regression and categorical models are summarised and discussed in the next two chapters.
 
@@
 Prediction of 'return per dollar' as a regression problem was attempted first followed by prediction of profit or loss as a binary classification problem.
-Linear Regression, Random Forests and Boosted Tree methods were applied to both problems with varying success.
+Linear Regression, Random Forests and Boosted Tree methods were applied to both problems.
 
@@
 In ANOVA, categorical and continuous explanatory variables are treated differently.
-For categorical variables, differences in the mean values of the response variable in each group are tested.
+For categorical variables, differences in the mean values and variances of the response variable in each group are tested.
 In contrast, linear relationships are examined between continuous explanatory variables and the response variable.
-Since linear relationships are being tested, continuous variables are assumed to be normally distributed.
-Therefore, where appropriate, numeric variables were normalised via logarithmic or cube root transformations.
+The continuous variables were normalised to kerb the effects of possible high leverage points, and logarithmic or cube root transformations were used for this.
 Trials of various variable interactions revealed that interactions between:
@@
 
-were statistically significant. Two types of ANOVA models were run. The first used only a 'core' list of 6 variables that were largely complete and any relevant interaction terms.
+were statistically significant. Two types of ANOVA models were built. The first used only a 'core' list of 6 variables that were largely complete along with relevant interaction terms:
 
@@
 When all variables were combined, only 15% of the project cases had complete data and therefore the ANOVA test would eliminate 85% of the data upon execution.
-Imputing the missing data was not desirable at that stage due to the volume of missing categorical data and an inability to ensure the imputed data was not hampering results.
-The second method was a more complicated procedure that required dozens of models but allowed variables with missing data to be included.
+Imputing the missing data was not desirable at that stage due to the volume of missing categorical data and an inability to ensure the imputed data did not adversely influence results.
+The second method was a complicated procedure that required dozens of models but allowed variables with missing data to be included.
 First, a project was randomly selected (a row in the data set) which would have certain variables complete.
@@
 
-The results of the ANOVA models were assessed using the root mean squared error (RMSE) statistic of the return per dollar predictions.
-These were compared against a baseline RMSE, which was the RMSE of the predicted return per dollar values against the mean return per dollar of all projects.
-Then, the RMSE from each model was subtracted from the RMSE using mean return per dollar.
+The results of the ANOVA models were assessed using the RMSE statistic of the 'return per dollar' predictions.
+These were compared against a baseline RMSE, which was the RMSE of the predicted 'return per dollar' values less the mean 'return per dollar' of all projects.
+
+$RMSE_b = \sqrt{\frac{\sum_{t = 1}^{n} (\widehat{y}_t - \bar y)^2}{n}}$
+
+RMSE_b = baseline root mean squared error
+y = predicted 'return per dollar'
+y = mean 'return per dollar' for al projects
+
+Where
+
+$$
+\begin{aligned}
+RMSE_b = \textrm{baseline root mean squared error} \\
+n &= \textrm{number of data points} \\
+\widehat{y} &= \textrm{predicted values of the response variable} \\
+\bar y &= \textrm{mean 'return per dollar' for all projects} \\
+\end{aligned}
+$$
+
+Then, the RMSE from each model was subtracted from the baseline RMSE:
+
+$\Delta\ RMSE = RMSE_b - RMSE$
+
 If the models were effective, the RMSE should be lower for ANOVA model predictions, so the difference would be greater than 0.
-Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only (method 1 using ANOVA).
+Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only (method 1 ANOVA).
 Note, only `r n.samples(data = results.A2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
 
-```{r, fig.height = 4.3, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
-If the ANOVA model is effective, the difference should be greater than 0.")}
+```{r, fig.height = 4.3, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on 6 core variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
+If the ANOVA model was effective, the difference should be greater than 0.")}
 
 #plot histogram to show differences
+
 results.core %>% ggplot(., aes(x =diff, y = ..density..))+ 
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 100 ANOVA models \nCore Variables Only") +
+        labs(x = expression(Delta~RMSE), 
+             y = "Density", title = expression(atop(paste("Density Plot of")~Delta~RMSE~paste("for 50 ANOVA Models"),Core~Variables~Only))) +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-A student t-test indicates the difference is significantly above 0 with a p-value of `r t.test(results.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
+A student t-test indicated $\Delta RMSE$ was significantly above 0 with a p-value of `r t.test(results.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`.
 Therefore, the null hypothesis that the difference is less than or equal to zero can be rejected.
-In other words, the ANOVA model improved estimates of return per dollar over using the mean return per dollar to a statistically significant degree.
-However, the mean difference is only \$`r mean(results.core$diff) %>% format(digits = 1)`.
-Furthermore, the mean RMSE in return per dollar is \$`r mean(results.core$final.RMSE) %>% format(digits = 2)`, which is unacceptably high.
-Almost 50 cents is a large mean error when the value is estimating the profit from \$1 spent.
-
-Results from 100 models using method 2, which captures incomplete variables, are shown below.
-
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
+In other words, the ANOVA model improved estimates of 'return per dollar' over using the mean 'return per dollar' to a statistically significant degree.
+However, the mean difference was only \$`r mean(results.core$diff) %>% format(digits = 1)`.
+Furthermore, the mean RMSE in 'return per dollar' was \$`r mean(results.core$final.RMSE) %>% format(digits = 2)`, which is unacceptably high.
+Almost 50 cents is a large error when the value is estimating the profit after spending \$1.
+
+Results from 100 models using method 2, which captures incomplete variables, are shown below:
+
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between ANOVA model predictions (built on all variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
 If the ANOVA model is effective, the difference should be greater than 0.")}
@@
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 ANOVA models \nAll Variables Sampled") +
+        labs(x = expression(Delta~RMSE), 
+             y = "Density", title = expression(atop(paste("Density Plot of")~Delta~RMSE~paste("for 100 ANOVA Models"),All~Variables~Sampled))) +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-A student t-test indicates the difference is not significantly above 0, which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
+A student t-test indicated the difference was not significantly above 0, which is clear from the plot (p-value = `r t.test(results.A2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 3)`).
 This method clearly performed poorly in comparison to the core variables.
 A reason for this could be that the mean number of cases in the data set was `r results.core$test.nrow[1]*4` for the core variables but averaged only `r (mean(results.A2$test.nrow)*4) %>% round(0)` for the sampled set of all variables.
-Regression results from ANOVA models were poor - the Random Forest algorithm was trialed next.
+Overall, the Regression results from Linear ANOVA models were poor and the Random Forest algorithm was trialed next.
 
@@
 ```{r load_rf_results}
-results.rf.core = readRDS("reg_rf_core.rds")
+results.rf.core = readRDS("report_data/reg_rf_core.rds")
 # n.samples(data = results.rf.core$diff)
 
-results.rf2 = readRDS("reg_rf.rds")
+results.rf2 = readRDS("report_data/reg_rf.rds")
 # n.samples(data = results.rf2$diff)
@@
 
-Numeric predictions of return per dollar were attempted first with core variables that were largely complete followed by a method of sampling a mix of core and less complete variables in every run (similar to the two ANOVA methods).
-The Random Forest parameters were first tuned using the caret package and optimal values mtry = 5 (the number of randomly selected explanatory variables to consider at each split) and ntree = 500 (the number of decision trees to ensemble) were determined [@caret].
+Numeric predictions of 'return per dollar' were attempted first with core variables that were largely complete followed by a method of sampling a mix of core and less complete variables in every run (similar to the two ANOVA methods).
+The Random Forest parameters were first tuned using the caret package and the optimal values mtry = 5 (the number of randomly selected explanatory variables to consider at each split) and ntree = 500 (the number of decision trees to ensemble) were determined [@caret].
 A density plot of 50 models of the core variables using different training/test sets is shown below.
-The plot was created using the same procedure used in the ANOVA models where the x-axis measures the RMSE obtained by using the Random Forest model subtracted from the RMSE using the mean of all project return per dollar values.
+The plot was created using the same procedure used in the ANOVA models where the x-axis measures $\Delta RMSE$.
 Note, only `r n.samples(data = results.rf2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
 
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a base line predictor that uses the mean return per dollar as its prediction.
-If the ANOVA model is effective, the difference should be greater than 0.")}
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
+If the Random Forest model was effective, the difference would be greater than 0.")}
 
@@
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nCore Variables Only") +
+        labs(x = expression(Delta~RMSE), 
+             y = "Density", title = expression(atop(paste("Density Plot of")~Delta~RMSE~paste("for 50 Random Forest Models"),Core~Variables~Only))) +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-This plot shows a mean return per dollar RMSE improvement of \$`r mean(results.core$diff) %>% format(digits = 1)`, which is the same as the ANOVA model.
+This plot shows a mean 'return per dollar' RMSE improvement of \$`r mean(results.core$diff) %>% format(digits = 1)`, which is the same as the ANOVA model.
 The p-value for the null hypothesis that the difference is less than or equal to zero is `r t.test(results.rf.core$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)` and could be rejected.
@@
 
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a base line predictor that uses the mean return per dollar as its prediction.
-If the ANOVA model is effective, the difference should be greater than 0.")}
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
+If the Random Forest model is effective, the difference should be greater than 0.")}
 
@@
         geom_histogram() + geom_density() +
-        labs(x = "Difference in RMSE (Mean RMSE - Model RMSE)", title = "Density Plot of change in RMSE for 50 randomForest models \nAll Variables Sampled") +
+        labs(x = expression(Delta~RMSE), 
+             y = "Density", title = expression(atop(paste("Density Plot of")~Delta~RMSE~paste("for 100 Random Forest Models"),All~Variables~Sampled)))+
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 On the right hand tail of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar on the left tail.
-A density plot of the RMSE for the models is shown below:
+A density plot of the unaltered RMSE values for the models are shown below:
 
@@
         geom_histogram() + geom_density() +
-        labs(x = "RMSE", title = "Density Plot of RMSE for 100 randomForest models \nVariables Randomly Sampled") +
+        labs(x = "RMSE", title = "Density Plot of RMSE for 100 Random Forest Models \nVariables Randomly Sampled",
+             y = "Density") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 
-Although the Random Forest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for return per dollar was wide and the mean RMSE was high (\$`r mean(results.rf2$final.RMSE) %>% round(2)`).
+Although the Random Forest algorithm performed better than ANOVA when incorporating less complete variables, the spread of RMSE for 'return per dollar' was wide and the mean RMSE was high (\$`r mean(results.rf2$final.RMSE) %>% round(2)`).
 
 The two regression models predicting 'return per dollar' resulted in high mean RMSE's ranging from \$`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
-The next section discusses possible reasons for the poor results and the decision to halt progression of regression models.
+The next section discusses possible reasons for the poor results and the decision to halt development of regression models.
 
@@
 Both ANOVA and Random Forest models produced a similar lowest RMSE value, but performed differently when directly comparing models built on the same list of variables.
-Both methods predicted at best within an average of \$0.48 for the response variable, 'return per dollar', using either core variables or a sample of core and less complete variables.
+Both methods predicted at best within an average of \$0.48 for the response variable, 'return per dollar', using either core variables or a sample of core and incomplete variables.
 Random Forest predictions were much more accurate than ANOVA's with a mix of core (almost complete) and less variables while Random Forest and ANOVA performed similarly with the core variables only.
@@
 Each Random Forest is the average of 500 single trees, whereas each ANOVA model is a single model only.
-50 Random Forest models produce 500 times more models than 50 ANOVA models so the averaging effect may have eliminated outliers for better overall performance.
+50 Random Forest models produce improved overall performance.
 
@@
 The magnitude of the RMSE can be evaluated from the end-user's perspective (the manager who is proposing fees for new projects).
-To give a feel for typical project performance, refer the plot below of historical 'return per dollar' values:
+To give a feel for typical project performance, refer to the plot below of historic 'return per dollar' values:
 
@@
         geom_histogram() + geom_density() +
-        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects") +
+        labs(x = "Return per Dollar ($)", title = "Density Plot of Return per Dollar - All Projects",
+             y = "Density") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 If a project returns \$0 for every dollar spent (return per dollar = \$0), the project has earned only enough to cover the cost of doing the project (salary costs and business costs).
-If the return per dollar is -\$0.50, the project has earned enough to cover half of the costs of doing the job.
-Finally, if the return per dollar is \$1, the project earned double what it cost to perform.
-Using this reasoning, if model predictions have an average error of \$.50 for return per dollar, this represents an error in revenue of half what the project cost to complete.
+If the 'return per dollar' is -\$0.50, the project has earned enough to cover half of the costs of doing the job.
+Finally, if the 'return per dollar' is \$1, the project earned double what it cost to perform.
+Using this reasoning, if model predictions have an average error of \$.50 for 'return per dollar', this represents an error in revenue of half of the project cost.
 Unfortunately, this is clearly not informative enough to bring into the business decision-making process.
@@
 Further regression methods could have been pursued, however Random Forest is a high-performing algorithm and was far from providing usable results.
-It is likely that either not enough data was available or that the given variables did not capably explain all the factors affecting return per dollar.
-Each row of data is a project, which can take the company from a few weeks to sometimes years to complete.
-Therefore, more data cannot be easily accessed or created.
-It was decided categorical predictions should be pursued next as predicting whether a job will be profitable or not is a simpler task and could more effectively leverage the limited data.
-
-This conclusion differs to case studies in the literature, which tend to predict total cost/effort as a regression problem [@Kim2004; @Chan2005; @Akintoye2000].
-This is likely because the literature dealt with substantial projects that are rarer and impact governments and large companies.
-A simple profit or loss prediction would not be useful.
-The case study company on the other hand may find a profit/loss binary prediction useful because they complete hundreds of small projects a year.
+It is likely that either not enough data was available or that the given variables did not capably explain all the factors affecting 'return per dollar'.
+Each row of data is a project, which can take a few weeks or in some cases, years, to complete.
+Therefore, more data could not be easily accessed or created.
+It was decided binary classification should be pursued next as predicting whether a job will be profitable or not is a simpler task and could more effectively leverage the limited data.
+
+This conclusion differs to case studies in the literature, which tended to predict total cost or effort as a regression problem [@Kim2004; @Chan2005; @Akintoye2000].
+This is likely because a simple profit or loss prediction would not be useful for the way the problem was framed in the literature.
+The problem was to estimate costs better for project budgeting, without considering contract structure or the choice to not do a project at all.
+The case study company on the other hand may find a profit/loss binary prediction useful because they complete hundreds of small projects a year, and if risky projects are identified, the project could be rejected or have a conservative contract structure.
 If binary classification is successful, applications of the predictions must be explored along with contributions to the company decisions and profits.
 
-In summary, because of the poor results when predicting return per dollar, further development of regression models was deferred in favour of binary classification models.
+In summary, because of the poor results when predicting 'return per dollar', further development of regression models was deferred in favour of binary classification models.
 Binary classification is a simpler task than regression because it requires less precision.
@@
 
-The cost estimation problem analysed in the literature review had been converted to a profitability estimation problem and was finally converted to a binary profit/loss prediction.
-The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable project (or not).
+The cost estimation problem analysed in the literature review had been translated to a profitability estimation problem and was finally converted to a binary profit/loss prediction.
+The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable/unprofitable projects.
 The original 'return per dollar' response variable was transformed by converting all 'return per dollar' values less than or equal to zero to 'loss making' jobs and 'return per dollar' values greater than 1 to 'profitable' jobs.
@@
 
-The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were initially developed without imputing data, using the maximum amount of data possible depending on the method.
+The five prediction methods, Boosted Trees, Naive Bayes, Logistic Regression, Random Forests and Bayesian Networks, were initially built without imputing data, by using the maximum amount of data possible depending on the method.
 Boosted Trees and Naive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
 Logistic Regression, Random Forest and Bayesian Networks require data sets without missing values so a subset of complete data was used.
-It was expected that Boosted Trees would perform best as they are known to have powerful predictive capabilities and can use incomplete data.
+It was expected that Boosted Trees would perform best as they are known to have powerful predictive capabilities and can use incomplete data (i.e.
+could make use of all the data).
 For this reason, 4 x 5-fold runs of each method were performed (20 models each) and statistically compared against Boosted Trees' results.
@@
 It was finally decided that Random Forest and Boosted Trees would likely not be distinguishable since the power calculation indicated that 700 models were required.
-Hence, 100 models of each method were built as calculated for logistic regression.
-10-fold cross validation was performed 10 times.
-The violin plot below summarises the AUC values produced by each method.
-The 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
+Hence, 100 models of each method were built as calculated for Logistic Regression.
+This was achieved by performing 10-fold cross validation 10 times.
+The violin plot below summarises the AUC values produced by each method and the 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
 
@@
         scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
+                          name = "Significantly Different from Boosted Trees") +
         labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method") +
+             title = "AUC Distribution from 100 models of Each Method") +
         scale_x_discrete(labels = c("boost" = "Boosted Trees", 
@@
 Imputation was done using the MICE method with Random Forest imputation [@mice].
-Again, 100 models were required to achieve a power of 0.8 although this could not be achieved for Random Forest unless over 450,000 models were made.
+Again, 100 models were required to achieve a power of 0.8 with respect to Boosted Trees, although this could not be achieved for Random Forest unless over 450,000 models were made.
 
@@
 
-mice.totAUC <- readRDS("mice_df_auc.rds")
+mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
 
@@
         labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of each Method - Imputed Data") +
+             title = "AUC Distribution from 100 Models of Each Method - Imputed Data") +
         scale_x_discrete(labels = c("boost" = "Boosted Trees", 
@@
 Testing Boosted Trees, Random Forests, Naive Bayes, Logistic Regression and Bayesian Networks on the binary classification problem showed that Boosted Trees, Logistic Regression and Random Forests performed best (according to AUC).
-Bayesian Networks are not particularly popular and require discretised variables so it was expected that this algorithm would not perform as well as highly regarded methods such as Boosted Trees and Random Forests.
+Bayesian Networks were not popular in the Literature and required discretised variables, so this algorithm was not expected to perform as well as more highly regarded methods (Boosted Trees and Random Forests).
 Naive Bayes and Logistic Regression were included as baseline models and it was expected that the more complex models would outperform these.
-Therefore it was surprising that results from 100 Logistic Regression models were not statistically lower than Boosted Trees.
+Therefore it was surprising that results from 100 Logistic Regression models were not statistically significantly lower than Boosted Trees.
 
@@
 Trees tend to overfit the patterns in a smaller training set.
-Logistic regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
-This may explain Logistic Regression's comparatively high performance on the case study's smaller data set with many variables.
+Logistic Regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
+This may explain Logistic Regression's comparatively high performance on the case study's small but complex data set.
 
@@
 The benefit of having an imputed data set that performs acceptably well is that each model is fed the full data set and therefore produces predictions for the full data.
-Beforehand, Random Forest and Logistic Regression models were limited to a complete set of original data, which implied a reduced data set.
-With complete model predictions, the high performing models could be blended to test whether better predictions were possible.
-For this reason, the imputed data models were propagated forward to the development of blended models.
-
-In summary, the individual binary classification models performed well above random chance (AUC = 0.5) and it can be expected that implementing a model would improve a manager's ability to predict whether a new job will be profitable or loss making.
-Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model actually affects decisions in practice.
+Beforehand, Random Forest and Logistic Regression models were limited to complete original data, which implied a reduced final data set.
+With complete sets of model predictions, the high performing models could be blended to test whether better predictions were possible.
+For this reason, the imputed data models were chosen to be progressed to the development of blended models.
+
+In summary, the individual binary classification models performed well above random chance (AUC = 0.5) and implementing a model is expected to improve a manager's ability to predict whether a new job will be profitable or loss making.
+Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model can affect decisions in practice.
 Results from model blending and overall profit impact are presented and discussed in the next chapter.
@@
 #     
-# * to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the case based reasoning approach
+# * to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the CBR approach
 ```
@@
 
-The statistical distribution of return per dollar in the rejected projects is skewed towards negative returns.
+The statistical distribution of 'return per dollar' in the rejected projects is skewed towards negative returns.
 This means the algorithm has identified a proportionally much higher percentage of loss making projects which is not surprising given previous results.
@@
 Ensemble tree methods and blending have been applied minimally to cost estimation.
-Past research tested regression models and to some extent Neural Networks and case based reasoning.
+Past research tested regression models and to some extent Neural Networks and CBR.
 The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting as well as Neural Networks.

### 2016-08-01 finalwriteup/trial_code_eq.rmd
almost finished reading results and method

@@
 
-where $X_{j}$ is simply something
+$RMSE_b = \sqrt{\frac{\sum_{t = 1}^{n} (\widehat{y}_t - \bar y)^2}{n}}$
 
-$X_{1}$, include the terms $X_{2} = X_1^2$ and $X_{3} = X_1^3$ to build a third degree polynomial.
-Interactions between explanatory variables can be modelled by including a terms such as $X_5 = X_4*X_1$
+$$
+\begin{aligned}
+RMSE_b = \textrm{baseline root mean squared error} \\
+n &= \textrm{number of data points} \\
+\widehat{y} &= \textrm{predicted values of the response variable} \\
+\bar y &= \textrm{mean 'return per dollar' for all projects} \\
+\end{aligned}
+$$
 
-$\operatorname{logit}(p) = \log\left(\frac{p}{1-p}\right) = \log(odds)$
-
-$p(ABCD) = p(C|ABD)\cdot p(A|BD) \cdot p(B|D) \cdot p(D)$
-
-$p(ABCD) = p(C|BD)\cdot p(A|B) \cdot p(B) \cdot p(D)$
+$\Delta\ error = RMSE_b - RMSE$

### 2016-08-01 finalwriteup/draft.rmd
started reading model blending

@@
 
-Blended models combine the predictions from each higher-performing individual model to create averaged or 'blended' predictions.
+Blended models combine the predictions from each high-performing individual model to create averaged or 'blended' predictions.
 In this case, the Random Forest, Boosted Trees, and Logistic Regression models were chosen and as outlined in the method, 6 techniques for model blending were trialed ranging from simple to complex.
-The simplest way averaged the probability output (a number between 0 and 1) of all three models.
-The next two simplest methods consisted of building a Logistic Regression and Boosted Trees model from the three individual model outputs only.
-Feature Weighted Linear Stacking (FWLS), Random Forest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (meta-features).
-The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature conditions could be weighted as such.
+The simplest method averaged the probability output (a number between 0 and 1) of all three models.
+The next two simplest methods consisted of building a Logistic Regression and Boosted Trees model from the three prediction model outputs only.
+Feature Weighted Linear Stacking (FWLS), Random Forest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (called meta-features in this context).
+The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature states could be weighted as such.
 Some meta-features may have been dispensable in the blending process, so in order to maintain simple comprehensible models, a variable importance study was conducted again.
@@
 
-Variable assessments were completed and compared using cForests, Logistic Regression, Random Forests and Boosted Trees.
-CForests have the favourable characteristic of being unbiased while the remaining methods were used for the blending process.
+Explanatory variable assessments were completed and compared using cForests, Logistic Regression, Random Forests and Boosted Trees.
+CForests have the favourable characteristic of being unbiased, and the remaining methods are known to perform well with the data and were chosen for blending.
 The results are summarised in the sections below.
@@
 
-cfor.varimp = readRDS("cfor_blend_varimp.rds")
+cfor.varimp = readRDS("report_data/cfor_blend_varimp.rds")
+cfor.varimp <- mutate(cfor.varimp, var = revalue(var, c("norm.log" = "log.pred",
+                                         "norm.rf" = "rf.pred",
+                                         "norm.boost" = "boost.pred")))
 
@@
         geom_bar(stat='identity') + 
-        labs(title = "Cforest Variable Importance: Blended Model",
-             x = "Variable", y = "Variable Importance Score")
-
-```
-
-The cForest highlighted the results from the Logistic Regression and Boosted Trees as the most important variables.
+        labs(title = "CForest Variable Importance: Blended Model",
+             x = "", y = "Variable Importance Score")
+```
+
+The cForest highlighted predictions from the Logistic Regression and Boosted Trees as the most important variables.
 This was followed by the results from the Random Forest model and the project team size, project category, total invoiced amount (category) and time span (category).
@@
 
-log.varimp = readRDS("logblend.rds")
-
+log.varimp = readRDS("report_data/logblend.rds")
 log.varimp = summary(log.varimp)$coefficients %>% as.data.frame
@@
 
-An ANOVA regression found the F-statistics from the following variables were significant in rejecting the null hypothesis for p-value < 0.10:
-
-Interaction term between:
+A logistic regression found the p-values from the following variables were significant in rejecting the null hypothesis for a p-value < 0.05:
+
+Interaction terms between:
 
@@
 
-Single variables that were not also in interaction terms:
+As well as the following variables as individual terms:
 
@@
 
-#### Variable Importance: randomForest
-
-```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a randomForest blended model. The results from the three best performing models were added as explanatory variables.")}
-
-rf.varimp = readRDS("rfblend.rds")
+#### Variable Importance: Random Forest
+
+```{r, fig.height = 4.5, fig.cap = c("Variable importance output from a Random Forest blended model. The results from the three best performing models were added as explanatory variables.")}
+
+rf.varimp = readRDS("report_data/rfblend.rds")
 rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
@@
 rf.varimp = arrange(rf.varimp, desc(MeanDecreaseAccuracy))
+rf.varimp <- mutate(rf.varimp, rowname = revalue(rowname, c("norm.log" = "log.pred",
+                                         "norm.rf" = "rf.pred",
+                                         "norm.boost" = "boost.pred")))
 # order levels
@@
         geom_bar(stat='identity') +
-        labs(title = "randomForest Variable Importance: Blended Model")
-
-```
-
-Permutation importance by Random Forest rated all three prediction models as most important followed by team size for the project, percent of hours done by the chief employee on the project, time span (category), project category, and client business type.
+        labs(title = "Random Forest Variable Importance: Blended Model",
+             x ="", y = "Variable Importance Score")
+
+```
+
+Permutation importance by Random Forest rated all three prediction models highest followed by team size for the project, percent of hours done by the chief employee on the project, time span (category), project category, and client business type.
 
@@
 
-boost.varimp = readRDS("boostblend.rds")
+boost.varimp = readRDS("report_data/boostblend.rds")
 boost.varimp = add_rownames(boost.varimp)
+boost.varimp <- mutate(boost.varimp, rowname = revalue(rowname, 
+                                                       c("norm.log" = "log.pred",
+                                         "norm.rf" = "rf.pred",
+                                         "norm.boost" = "boost.pred")))
 # order levels
@@
         geom_bar(stat='identity') +
-        labs(title = "Boosted Trees Variable Importance: Blended Model")
-
-```
-Boosted Trees ranked output from all three individual prediction models highest (Logistic Regression ranked highest) followed by Project category, Client Business type, and time span (category).
+        labs(title = "Boosted Trees Variable Importance: Blended Model",
+             x ="", y = "Variable Importance Score")
+
+```
+Boosted Trees ranked output from all three individual prediction models highest (Logistic Regression was positioned first) followed by Project category, Client Business type, and time span (categorical).
 
@@
 
-Upon reviewing important variables ranked by cForest and each method of model blending, the following variables were deemed important by at least two methods and were used in the development of blended models:
+Upon reviewing important variables ranked by cForest, Logistic Regression, and each method of model blending, the following variables were deemed important by at least two methods and were used in the development of blended models:
 
@@
 
-Random Forests and Boosted Trees passively interact explanatory variables, but for logistic regression, these must be manually input.
-Only interaction terms with statistically significant F-statistics from the ANOVA test were included in order to reduce noise.
+Random Forests and Boosted Trees passively interact explanatory variables, but for Logistic Regression, these must be manually input.
+Only interaction terms with statistically significant p-values from the Logisitic Regression were included in order to reduce noise.
 
@@
 The remaining 3 techniques used only the predictions of the three individual models.
-All six methods were compared against the original logistic regression model.
+All six methods were compared against the original Logistic Regression model.
 Initially 100 train/test runs were performed for each blending method by running 20 iterations of 5-fold cross validation.
@@
 
-auc.list <- readRDS("blend_auc2.rds")
+auc.list <- readRDS("report_data/blend_auc2.rds")
 means = lapply(auc.list, mean) %>% unlist
@@
         scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC greater than Original Logistic Regression") +
+                          name = "Mean AUC Greater than Original Logistic Regression") +
         scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC greater than Original Logistic Regression") +
+                          name = "Mean AUC Greater than Original Logistic Regression") +
         theme(legend.position = "bottom")
@@
 
-The above plot shows that 4 methods had a higher mean AUC than the original logistic regression model:
-
-* the simple blended logistic regression 
+The above plot shows that 4 methods had a higher mean AUC than the original Logistic Regression model:
+
+* the simple blended Logistic Regression 
 * the simple average 
@@
 
-To achieve a power of 0.8, a two-sample power analysis was performed on the 100 AUC results from each blended model against the AUC results from the original logistic regression.
-The boosted models required over 1000 samples while the simple averaged and simple logistic regression models required only 120.
-Therefore, 50 more iterations of the simple averaged and simple logistic models were performed.
-The Boosted Tree models would not achieve a statistical power of 0.8 with only 150 samples when compared to the original logistic regression and were therefore dropped from the next stage of analysis.
+To achieve a power of 0.8, a two-sample power analysis was performed on the 100 AUC results from each blended model against the AUC results from the original Logistic Regression.
+The boosted models required over 1000 samples while the simple averaged and simple Logistic Regression models required only 120.
+Therefore, 50 more iterations of the simple averaged and simple Logistic Regression models were performed.
+The Boosted Tree models would not achieve a statistical power of 0.8 with only 150 samples when compared to the original Logistic Regression and were therefore dropped from the next stage of analysis.
 
@@
 The results of the 150 iterations are shown in the plot below.
-Simple model averaging resulted in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple logistic regression blended model had a two-sample t-test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
+Simple model averaging resulted in a p-value of `r t.test(auc.150[['orig.log']], auc.150[['average']])$p.value %>% format(digits = 2) %>% as.numeric` while the simple Logistic Regression blended model had a two-sample t-test p-value of `r t.test(auc.150[['orig.log']], auc.150[['simp.log']])$p.value %>% format(digits = 2) %>% as.numeric`.
 
@@
         scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC significantly greater than Original Logistic Regression") +
+                          name = "Mean AUC Significantly Greater than Original Logistic Regression") +
         scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC significantly greater than Original Logistic Regression") +
+                          name = "Mean AUC Significantly Greater than Original Logistic Regression") +
         theme(legend.position = "bottom")
@@
 
-The simple Logistic Regression blend and simple averaging of the individual models (logistic regression, Boosted Trees and Random Forests) gave virtually the same distribution of results, where both AUC distributions were significantly greater than the original logistic regression model.
+The simple Logistic Regression blend and simple averaging of the individual models (Logistic Regression, Boosted Trees and Random Forests) gave virtually the same distribution of results, where both AUC distributions were significantly greater than the original Logistic Regression model.
 The p-values from the two-sample t-tests were `r pvals$p_value[1]` and `r pvals$p_value[2]` for simple average and simple Logistic Regression methods respectively.
-It was encouraging that blended models produced higher AUC values than the original logistic regression model, however for the case study business, the real measure of success is the model's impact on the business' overall profits.
-It was not guaranteed higher AUC values directly related to improvements in profit so several methods, including original individual models as well as blended methods, were progressed to the profit analysis.
+It was encouraging that blended models produced higher AUC values than the original Logistic Regression model, however for the case study business, the real measure of success is the model's impact on the business' overall profits.
+Higher AUC values did not guaruntee more lucrative profit curves so several methods, including the original individual models as well as blended methods, were progressed to the profit analysis.
 

### 2016-08-02 finalwriteup/draft.rmd
just conclusion left now

@@
 library('scales')
+library('purrr')
 
@@
 
-A model that is perfectly classified would have all positives correctly classified (1.0 True Positive Rate) and no incorrectly classified positives (0.0 False Positive Rate).
+A model that is perfectly classified would have all positives correctly classified (1.0 TPR) and no incorrectly classified positives (0.0 FPR).
 If a model has a 1.0 TPR and 1.0 FPR (top right corner) it has correctly classified all positives at the expense of incorrectly classifying all negatives as positive.
@@
 The AUC indicates model performance across many thresholds while classification accuracy represents a single threshold.
-The AUC statistic is therefore useful in problems where cases can be ranked and a cut-off threshold that will optimise the rates of true positives and true negatives can be found [@Huang2005].
+The AUC statistic is therefore useful in problems where cases can be ranked and a cut-off threshold that will optimise the rates of true positives (TP's) and true negatives (TN's) can be found [@Huang2005].
 Note the 'probability' outputs from the models are not true statistical probabilities, but a score the model has assigned to a case using its own measure of certainty.
@@
 ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        geom_violin(aes(fill = sigdiff, colour = sigdiff), alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "gray20", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
         scale_fill_manual(values = c('sienna1','navajowhite'),
@@
 ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = sigdiff, colour = sigdiff)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        geom_violin(aes(fill = sigdiff, colour = sigdiff), 
+                    alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "gray20", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
         scale_fill_manual(values = c('sienna1','navajowhite'),
@@
         geom_violin(aes(fill = Data_Source, colour = Data_Source)) +
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-
         # stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
@@
 
-Boosted trees, Random Forests and Logistic Regression all had mean AUC values around 0.775.
-The tests were repeated using imputed data and again the means sat around 0.775.
-The variables were not particularly well suited for imputing so this result is logical and could have even diminished the mean AUC.
-For example, many of the imputed variables were categorical such as job category, percent of job completed by a professional level employee, and the position of the majority worker on a job.
-The benefit of having an imputed data set that performs acceptably well is that each model is fed the full data set and therefore produces predictions for the full data.
-Beforehand, Random Forest and Logistic Regression models were limited to complete original data, which implied a reduced final data set.
+Boosted trees, Random Forests and Logistic Regression all had mean AUC values between 0.756 and 0.767.
+The tests were repeated using imputed data and again the means sat in a similar range from 0.755 to 0.764.
+Many variables were categorical which are not particularly well suited for imputation.
+These included variables such as job category, percent of job completed by a professional level employee, and the position of the majority worker on a job.
+Therefore it is logical the AUC's did not improve, in fact they marginally reduced.
+The benefit of having an imputed data set that performs acceptably well is that each model is fed the full data set and therefore produces predictions for the full data set.
+Before, Random Forest and Logistic Regression models were limited to complete original data, which implied a reduced final data set.
 With complete sets of model predictions, the high performing models could be blended to test whether better predictions were possible.
@@
 ggplot(auc.melt, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = greater, colour = greater)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        geom_violin(aes(fill = greater, colour = greater), 
+                    alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "gray30", 
+                     shape =4) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "gray20", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        theme(axis.text.x=element_text(angle=40,hjust=1)) +
         labs(x = "Method", y = "Area Under Curve (AUC)",
@@
         scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC Greater than Original Logistic Regression") +
+                          name = "AUC Greater than Original Logistic Regression") +
         scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC Greater than Original Logistic Regression") +
+                          name = "AUC Greater than Original Logistic Regression") +
         theme(legend.position = "bottom")
@@
 for(i in seq_along(index)){
-        pvals$p_value[i] = t.test(auc.150[[1]], auc.150[[index[i]]])$p.value %>%
+        pvals$p_value[i] = t.test(auc.150[[1]],
+                                  auc.150[[index[i]]])$p.value %>%
                 format(digits = 2) %>% as.numeric }
@@
 ggplot(auc.150, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = signif, colour = signif)) + 
-        geom_jitter(alpha = 0.2, shape = 16, colour= 'gray70') +
-        stat_summary(fun.y = "mean", geom="point", colour = "gray60", shape =3) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        geom_violin(aes(fill = signif, colour = signif),
+                    alpha = .8) + 
+        stat_summary(fun.y = "mean", geom="point", 
+                     colour = "gray30", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "gray20", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        theme(axis.text.x=element_text(angle=40,hjust=1)) +
         labs(x = "Method", y = "Area Under Curve (AUC)",
@@
         scale_fill_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC Significantly Greater than Original Logistic Regression") +
+                          name = "AUC Significantly Greater than Original Logistic Regression") +
         scale_colour_manual(values = c('navajowhite','sienna1'),
-                          name = "Mean AUC Significantly Greater than Original Logistic Regression") +
+                          name = "AUC Significantly Greater than Original Logistic Regression") +
         theme(legend.position = "bottom")
@@
 
-Trials of different methods of blending model results demonstrated that again, the simplest methods worked best with the data.
-In this case, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
-It was plausible that simple blending models would perform better than complex methods because the data was not big enough for complex models to learn intricate patterns.
-This was observed in the single method models (previous chapter) when Logistic Regression performed as well as Random Forests and Boosted Trees.
-The literature does not extensively cover blended machine learning methods however their success in the 2009 Netflix competition generated some publications.
+```{r blend_vs_initial}
+auc.150 = readRDS("report_data/blend_auc150.rds")
+mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
+compare.blends <- append(auc.150, mice.totAUC)
+
+index = c(3,4,5,6)
+pvals = data.frame("method" = names(compare.blends)[index], 
+                   "mean_AUC" = map_dbl(compare.blends, function(x) round(mean(x),3))[index],
+                   'p_value' = NA)
+
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(compare.blends[[3]],
+                                  compare.blends[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+pvals[1,'p_value'] <- NA
+pvals <- pvals %>% arrange(desc(mean_AUC))
+
+new.levels = c("simp.log" = "SBLR",
+               "log" = "Original Logistic Regression",
+               "boost" = "Original Boosted Trees",
+               "rf" = "Original Random Forest")
+
+pvals <- mutate(pvals, method = revalue(method, new.levels))
+names(pvals) = c("Method", "Mean AUC", "P-Value for t-test against SBLR")
+
+```
+
+The AUC distributions from the two best blended models, simple Logistic Regression and simple averaging, were statistically significantly higher than the AUC distribution of the original Logistic Regression model.
+Blended models improved the mean AUC from 0.759 to 0.77, which is an increase of `r ((0.77-0.759)/0.759) %>% "*"(100) %>% round(1)`%.
+The original Logistic Regression model was chosen for comparison because it was a fast baseline model that had an AUC distribution not statistically different to the other high performing original models.
+It was expected that model blending would improve predictions because it combined the strengths of individual models with different theoretical foundations.
+Next, the performance of different blended models is discussed.
+
+The trials of different blending methods demonstrated that again, the simplest methods worked best with the case study data.
+In this incident, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
+
+Simple blended models could perform better than complex methods if the data is not big enough for the complex models to learn the more intricate patterns at which they excel.
+This was observed in the single model methods (previous chapter) when Logistic Regression performed as well as Random Forests and Boosted Trees.
+There is limited literature on blended machine learning methods, however their success in the 2009 Netflix competition generated some publications.
 The Netflix data set comprised of almost 3,000,000 observations, so the size of the data could have enabled the success of complex blending methods such as FWLS [@Sill2009].
-That being the case, it was still surprising that simple averaging of probabilities from the individual models performed almost as well as blending results using Logistic Regression.
-It was expected that Logistic Regression would have an advantage over model averaging due to its ability to weight the prediction variables while maintaining a simple model structure.
-
-The Logistic Regression blending model did indeed weight predictions from the original models differently.
+That being the case, it was still surprising that simple averaging of probabilities from the individual models performed almost as well as Logistic Regression.
+Logistic Regression's ability to weight the prediction variables while maintaining a simple model structure was expected to give an advantage over averaging.
+
+The blended Logistic Regression model did indeed weight predictions from the original models differently.
 A summary of the resulting coefficients from an example blended Logistic Regression model is shown in the table below.
@@
 
-As shown in the table, if the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the model predicting a loss making job increased by 56.0%.
-The weightings contrasted the other simple model that averaged the original probabilities (i.e.
-weighted them equally) but still managed to produce similar AUC results.
+As displayed in the table, if the predicted probability from the original Logistic Regression model increased by 0.1, the odds of the blended model predicting a loss making job increased by 56.0%.
+This same interpretation can be made for the Random Forest and Boosted Trees values, which means the results from Boosted Tree models were weighted highest.
+Clearly, the weightings contrasted the equal weightings from the simple averaged model, yet their AUC results were similar.
 It was not clear why this was the case, however both simple models were compared again in the analysis of the impact each model could have on the company's overall profits.
@@
 
+This chapter first presents the full range of method results in terms of improvements to the case study's bottom line.
+A business decision making scenario was created so that profit curves based on the decision rule could be built and analysed.
+From the profit curves, optimal probability thresholds could be derived for each method (individual and blended methods).
+These threshold points allowed the probability output from each model to be converted to categorical output (profit or loss) which in turn could generate confusion matrix statistics for each method.
+The confusion matrix statistics are discussed after the profit curve results and discussion.
+Then, output from the method which produced the best profit curve was used to create two subsets of profitable and loss mkaing projects.
+The characteristics of these subsets are compared and contrasted to gauge any overall trends.
+Finally, alternate ways of framing the business decision scenario are proposed.
+
 ## Profit Curve Analysis
@@
 The first research question, whether machine learning and statistical techniques can predict project profitability based on historic data, has been answered.
-The measure of profitability could not be modeled but simplifying the profitability prediction to a binary (profit/loss) model was successful.
+The measure of profitability could not be modeled well, but simplifying the profitability prediction to a binary (profit/loss) model was successful and a mean AUC of 0.77 was achieved for the best blended model.
+This falls midway between a model that predicts perfectly (AUC = 1) and a model that predicts as well as random chance (AUC = 0.5).
 To answer the second research question, whether the model would improve the business' profitability, a profit analysis was performed.
 
-The models output a 'probability' between 0 and 1 that each project will be a loss making job (probability = 1 indicates a loss making job).
-This leaves the question, at what probability would a decision-maker round the probability up to 1 or down to 0?
+The predictive models output a 'probability' between 0 and 1 that each project will be a loss making job (where probability = 1 indicates a loss making job).
+The question the arises, at what probability would a decision-maker round the probability up to 1 or down to 0?
 And what business decision would then be made?
 To find the threshold point for rounding, an experimental business-scenario was tested.
-At a given threshold, all projects calculated above the threshold were considered too risky, and were rejected.
+At a given threshold, all projects with probability outputs above the threshold were considered too risky, and were rejected.
 All profits and losses from these projects were forfeited, while the profits from the remaining jobs (below the threshold) were summed to give a revised total profit.
@@
 # How to achieve statistical power of 0.8?
-finance.plot = readRDS("plot_finance.rds" )
+finance.plot = readRDS("report_data/plot_finance.rds" )
 fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
@@
 This profit calculation was made on a range of thresholds between 0 and 1 at 0.05 increments.
-The total profits were plotted for each threshold and joined to make a profit curve where at an optimal threshold, the total profits were highest (refer plot below).
+The total profits were plotted for each threshold and joined to make a profit curve where at the optimal threshold, the total profits were highest (refer plot below).
 
@@
 However, the total profit coordinates were based on probability outputs in test data sets.
-This means that the profit curve for the simple logistic regression blend (for example) would be different for different training/testing data divisions.
-I.e.
-the total profits at each point in the profit curve depend on the model parameters that were calculated using the training data and which projects were analysed in the test data.
-Therefore, the process of training a blended model, then calculating the profit curve from the test data was repeated multiple times to create a distribution of profit curves for each method (blended and individual).
-
-To start, 10 curves for each method were calculated (2 x 5-fold cross validation).
-Two-sampled power calculations (for a power of 0.8) were performed between the results of each method at their optimal threshold against the best performing method (simple average blend).
-For example, the optimal threshold for the simple average blended model was 0.55 while the optimal threshold for the original Boosted Tree model was 0.60.
+This means that the profit curve for the simple Logistic Regression blend (for example) would be different for different training/testing data divisions.
+In other words, the total profits at each point in the profit curve depend on the model parameters that were calculated using the training data and which projects were analysed in the test data.
+Therefore, the process of training a blended model, then calculating the profit curve from the test data, was repeated multiple times to create a distribution of profit curves for each method (blended and individual).
+
+To start, 10 curves for each method were calculated (using 2 x 5-fold cross validation).
+Two-sampled power calculations (for a power of 0.8) were performed between the results of each method (at their optimal threshold points) against the best performing method (simple Logistic Regression blend).
+For example, the optimal threshold for the simple Logistic Regression blended model was 0.55 while the optimal threshold for the original Boosted Tree model was 0.60.
 The 10 'total profit' values at these two thresholds were compared in the two-sample power calculation to determine the number of samples required to achieve a power of 0.8.
-The maximum required number of samples was 64 so 100 models were of each method were built (20 x 5 fold cross validation).
-
-The plot below illustrates the results of each model at each threshold with a line joining the mean value at each threshold for each method.
-The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold for the 100 models.
+Out of all the methods compared against the Blended Logistic Regression, the maximum required number of samples was 64, so 100 models of each method were built (20 x 5-fold cross validation).
+
+The plot below illustrates the distrubtion of profit curves for each method, where the solid lines join the mean values at each threshold point.
+The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold point for the 100 models.
 
@@
 
-Two blended models clearly outperformed the individual models as shown in the plot above with higher profit ratios as well as tighter confidence intervals.
+Two blended models clearly outperformed the individual models as shown in the above plot with higher profit ratios as well as tighter confidence intervals.
 The table below denotes each models' optimal thresholds and the corresponding increase in total profits:
@@
 
-The simple logistic regression blend performed best with the highest mean profit ratio at `r best.mean`% with a standard deviation of `r best.sd`%.
-This means that for the simple blended logistic regression model, if all jobs above the probability threshold `r best.t` were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profits.
-
-For clarity, the plot for the simple blended logistic regression model is shown below:
+The simple Logistic Regression blend performed best with the highest mean profit ratio of `r best.mean`% with a standard deviation of `r best.sd`%.
+This means that for the simple blended Logistic Regression model, if all jobs above the probability threshold `r best.t` were rejected, the profits would increase on average by `r best.mean - 100`% in comparison to historical profits.
+
+For clarity, the plot for the simple blended Logistic Regression model is shown below:
 
@@
 ```
-
-### Highest Profit Curves
-
-The simple average blend and logistic regression blended models outperformed the complex blended models, which is rational because their AUC values were significantly higher than the others.
-The simple average of Logistic Regression, Random Forests, and Boosted Trees results produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression that blended the same three variables.
-The blended Simple Logistic Regression produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`.
-To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r comma_format() ((best.mean-100)/100*1000000)` by rejecting all jobs with a probability higher than `r best.t`.
+<!-- comments on your results; -->
+<!-- explains what your results mean; -->
+<!-- interprets your results in a wider context; indicates which results were expected or unexpected; -->
+<!-- provides explanations for unexpected results. -->
+<!-- The Discussion should also relate your specific results to previous research or theory.
+You should point out what the limitations were of your study, and note any questions that remain unanswered.
+The Discussion CAN also include Conclusions/Future Research.
+-->
+
+In this section, the results from the models that delivered the highest performing profit curves are reviewed first and analysed in a broader business context.
+Then, all the methods that produced profit curves are compared and possible explanations for differences are evaluated.
+
+### Examination of the Highest Profit Curves
+
+Profit curves from the simple average blend and Logistic Regression blended models outperformed the complex blended models, which follows logically from their significantly higher AUC distributions.
+The simple average blended model produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression blend.
+As summarised in the previous section, the blended Simple Logistic Regression produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`.
+In a business context, this means that if the company's yearly profits were \$1,000,000, the mean increased profit could be \$`r comma_format() ((best.mean-100)/100*1000000)` by rejecting all jobs with a probability higher than `r best.t`.
 In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.
 
-The plot below illustrates this proportion of projects as well as the proportion of profits and losses assigned to a range of probability levels by the simple logistic regression blend.
-The red line sits at its optimal threshold point, i.e.
-projects greater than the threshold point would be rejected:
+The plot below illustrates the relative profits and losses that sit within the range of probability levels, as assigned by the simple Logistic Regression blend.
+
+$Relative\ Profit_t = \frac{Profit_t}{\sum_{t = 1}^{n}\left|Profit_t\right|}$
+
+Where
+
+$$
+\begin{aligned}
+Profit_t &= \textrm{Profit from project}\ t\textrm{. Negative values indicate loss making projects}
+\end{aligned}
+$$
+
+The red line sits at the profit curve's optimal threshold point, 0.6, so projects greater than the threshold point would be rejected:
 
@@
 
-reject.plot <- readRDS('reject_plot.rds')
+reject.plot <- readRDS('report_data/reject_plot.rds')
 
@@
                    labeller = labeller(faceter = c("hist" = "Count of Projects",
-                                       "profs"= "Relative Profits/Losses"
+                                       "profs"= "Relative Profits and Losses"
                                        ))) +
@@
              title = "Distribution of Project Counts and Project Profits Against Probability") +
-        geom_vline(xintercept = 10.5, colour = hcl(h = 15, l = 65, c = 100), linetype = 1,
+        geom_vline(xintercept = 10.5, 
+                   colour = hcl(h = 15, l = 65, c = 100), linetype = 1,
                    alpha = 0.8) +
+        geom_hline(yintercept = 0, colour = "white", linetype = 1)
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 ```
+
+As shown in the upper plot, a low count of projects fall above the optimal threshold for the simple Logistic Regression blend.
+The lower plot illustrates the relative profits (and losses for negative values) within each probability score bin which are normalised by the sum of the absolute value of all profits and losses.
+The model has assigned low probability scores to profitable projects and high scores to unprofitable jobs well.
+The heights of the bars reduce from left to right, and loss making jobs dominate projects on the right hand side of the optimal threshold point.
+This is advantageous as these jobs would be rejected.
 
@@
 The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
-That is, it cannot be said with 95% confidence that the original methods' would produce a profit higher than historical profits if projects higher than the optimal thresholds (jobs that are likely to be loss making) were rejected.
-Clearly this is not helpful for the case-study business to adopt.
+That is, it cannot be said with 95% confidence that the original methods' would produce a mean profit higher than historical profits in the given decision framework.
+This framework stated that projects higher than the optimal thresholds (jobs that are likely to be loss making) should be rejected.
+Clearly this level of certainty is not beneficial for the case-study business to adopt.
 
 The complex blended models had mixed results.
-FWLS and Random Forests performed more poorly than the individual methods while Boosted Trees produced a marginally higher profitability ratio (`r best.fin %>% filter(method == "comp.boost") %>% .$mean_profit_ratio`) but the lower bound of its confidence interval was `r best.fin %>% filter(method == "comp.boost") %>% .$lower_bound_95CI`, i.e.
-not above original profits.
-
-The simple blended models performed more favourably, as already discussed, particularly the simple average and Logistic Regression.
-The Boosted Tree's lower bound confidence interval was `r best.fin %>% filter(method == "simp.boost") %>% .$lower_bound_95CI`% at its optimal threshold point.
-This is not above 100%, and again does not make a convincing argument for the case-study to adopt it.
+FWLS and Random Forests performed more poorly than the individual methods while Boosted Trees produced a marginally higher profitability ratio (`r best.fin %>% filter(method == "comp.boost") %>% .$mean_profit_ratio`%) but the lower bound of its confidence interval was `r best.fin %>% filter(method == "comp.boost") %>% .$lower_bound_95CI`%, i.e.
+not an improvement on original profits.
+
+The simple blended models performed much more favourably, as already discussed, particularly the simple average and Logistic Regression blends.
+The blended Boosted Tree's lower bound confidence interval was `r best.fin %>% filter(method == "simp.boost") %>% .$lower_bound_95CI`% at its optimal threshold point.
+This is not above 100%, and again does not make a convincing argument to the case study business.
 
 It is not clear why simple linear models and the averaging model outperformed ensemble tree blending methods.
-It may be that there was not enough data for the ensemble trees to adequately learn the complex series of rules they develop as previously stated.
+As previously stated, the ensemble trees may not have received enough data to adequately learn the complex series of rules they develop.
 It was also surprising FWLS was ineffective given the success it had in the Netflix Prize competition [@Sill2009].
 The only differences between FWLS and the simple Logistic Regression (that performed best) were two additional explanatory variables and four interaction terms.
-The number of variables was not high for the amount of data.
-According to @Peduzzi1996, 10 events per explanatory variable or more avoids the risk of biased estimation of variable coefficients in logistic regression.
+The number of variables was not high for the amount of data since according to @Peduzzi1996, 10 events per explanatory variable or more avoids the risk of biased estimation of variable coefficients in Logistic Regression.
 The data contained 315 events per explanatory variable and did not pose that risk.
 Nevertheless, the additional variables must have added misleading noise to the model.
-The performance of the models so far has been concerned with AUC and profit, however insight into the differences in model performance may be gained by reviewing confusion matrix statistics such as TPR and FPR.
-
-### Comparison of Confusion Matrix Statistics
-
-To investigate method performance further, confusion matrix statistics were scrutinised.
+
+The performance of the models so far has been evaluated in terms of the AUC statistics and profit improvements, however further insight into may be gained by reviewing confusion matrix statistics such as TPR's and FPR's.
+
+### Profit Analysis Conclusion
+
+To conclude, because of the promising AUC results, it was logical that the models translated into financial benefits for the company.
+The `r (best.mean-100)`% improvement in profits produced by the simple blended Logistic Regression is reasonable, and may be high enough to trigger further cost-benefit analyses and the development of a more comprehensive framework describing different decision scenarios.
+
+## Categorical Predictions Analysis and Discussion
+
+The probability thresholds derived from the profit curves enabled the numeric model probability outputs to be converted to categorical output: profitable and loss making projects.
+The final classification of projects drove alternative ways to review method performance.
+These include an analysis of TPR's and FPR's for the various methods as well as a review of project characteristics of rejected vs.
+accepted projects.
+Finally, alternate decision rules are proposed and considered.
+
+### Comparison of Confusion Matrix Statistics from All Methods
+
+To review method performance further, confusion matrix statistics were scrutinised.
 For each method, 100 full data sets of results were generated via 5-fold cross validation.
 Each data set of probability results differed to the next because of the random division into the five-fold data sets used for training and testing.
-Nine full data sets of results was chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
+For this analysis, nine full data sets of results was chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
 These were deemed a 'typical' set of predictions for each method.
 Then, the confusion matrix for each set of results at their respective thresholds could be calculated.
-The plot below shows the breakdown of predictions by each method on actual loss making projects.
-If the project was predicted correctly as loss making, this was a true positive, and the job was rejected (in the fabricated scenario).
-If it was predicted incorrectly as a profit-making job, this was a false negative, and the project was completed, resulting in a loss for the business.
+The plot below shows the breakdown of FN's and TP's on true loss making projects.
+If the project was predicted correctly as loss making, this was a TP, and the job was rejected (in the fabricated decision scenario).
+If it was predicted incorrectly as a profit-making job, this was a false negative (FN), and the project was completed, resulting in a loss for the business.
 
@@
 
-conf.plot <- readRDS("conf_plot.rds")
+conf.plot <- readRDS("report_data/conf_plot.rds")
 
@@
 In this case, projects above a threshold point will include profitable and loss making jobs.
-The optimal point is where the savings from rejecting loss making jobs (TP) outweigh the lost profits from rejecting profitable jobs (FP) the most.
+The optimal point is where the savings from rejecting loss making jobs (TP's) most outweigh the lost profits from rejecting profitable jobs (FP's).
 All models were capable of different TP and FN ratios at alternate threshold points.
@@
 The plot above shows the Complex Boosted Tree Blend has the highest TPR followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
-All of these TPR's are below 25% which shows it was not worth catching a large percentage of loss making jobs in comparison to the losses in rejected profitable jobs (FP's) that accompanied the TP's.
-This implies the models were not able to very cleanly divide loss making jobs from profitable jobs.
-
-A summary of predicted results on actual profitable jobs is shown below at each method's optimal threshold point.
-If a job was predicted to be profitable, and was therefore accepted, it was classified TN.
-If a job was loss making but labelled profitable, the business would complete the job and make a loss (FN).
+All of these TPR's are below 25% which shows it was not worth rejecting a large percentage of loss making jobs in comparison to the accompanying losses in rejected profitable jobs (FP's).
+This also implies the models were not able to cleanly divide loss making jobs from profitable jobs.
+
+A summary of predicted results on true profitable jobs is shown below at each method's optimal threshold point.
+If a job was predicted to be profitable, and was therefore accepted, it was classified as a TN.
+If a job was profitable but labelled loss making, the business would reject the job and forfeit the profits (FP).
 
@@
 The complex Boosted Tree method had the highest TPR, but also had a high FPR.
-The increased number of FP (i.e.
-rejected profitable jobs) offset the more accurate rejection of loss making jobs so much that the simple averaged blend and simple logistic regression blend outperformed it.
+The increased number of FP's (i.e.
+rejected profitable jobs) offset the more accurate rejection of loss making jobs so much that the simple averaged blend and simple Logistic Regression blend outperformed it.
 The aforementioned simple blends performed similarly, with a higher than average TPR and average FPR in comparison to other methods.
@@
 All below plots and comments summarise results from one Simple Blended Logistic Regression Model.
-The results from this model were chosen because it represented the mean for improvement in profits output by this blending method.
+The results from this model were chosen because it represented the mean improvement in profits output by this blending method.
 The plot below illustrates how the ratio of profitable to unprofitable jobs compare in the full project population vs.
@@
 The class imbalance towards profitable jobs in the initial data set may have weakened the model's ability to classify losses.
-@Saradhi2011 took advantage of SVM's ability to adjust class penalties to address class imbalance, however SVM's were avoided because of their black box predictions in this work and the remaining methods cannot accommodate class penalties.
+@Saradhi2011 took advantage of SVM's ability to adjust class penalties to address class imbalance, however SVM's were avoided because of their black box predictions in this work and the remaining methods could not accommodate class penalties.
 
 The plot below illustrates the monetary size of profits and losses from rejected projects.
-The different coloured bands indicate different projects and it can be observed that the magnitude of rejected profits and losses vary greatly.
-They range from profits/losses of just \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% tail(1) %>% round(0)` to losses of \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% head(1) %>% round(-3)`.
+The coloured bands indicate different projects and it can be observed that the financial magnitude of rejected profits and losses vary greatly.
+They range from profits/losses of just \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% tail(1) %>% round(0)` to losses of \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% head(1) %>% round(-3) %>% format(digits = 2)`.
 
@@
 
-The algorithm tended to accept projects which were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
+The model tended to accept projects which were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
 This could either indicate that professional level employees complete projects more efficiently or that the types of projects that do not require technical employees are more profitable.
@@
 ```{r, fig.cap = c("Proportion of each timespan category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+
+reject.plot <- within(reject.plot, 
+                         b.timespan.cbrt <- factor(b.timespan.cbrt, 
+                                               levels= c("1d-3wk",
+                                                         "3wk-2.5m",
+                                                         "2.5m-9m",
+                                                         "9m-1.5y",
+                                                         "1.5y-3y",
+                                                         ">3y")))
 
@@
 The statistical distribution of 'return per dollar' in the rejected projects is skewed towards negative returns.
-This means the algorithm has identified a proportionally much higher percentage of loss making projects which is not surprising given previous results.
+This means the algorithm has identified a proportionally higher percentage of loss making projects which is not surprising given previous results.
 
@@
 
-The detailed analysis of rejected project characteristics gave insight into the characteristics of projects that are likely to be loss making as well as the accuracy of the algorithm.
-While it is interesting to observe characteristics of rejected projects, a complex combination of all the features contribute to the final probability of loss prediction.
+The detailed analysis of rejected project characteristics gave insight into the nature of projects that are likely to be loss making as well as the accuracy of the algorithm.
+While it is interesting to observe characteristics of rejected projects, a complex combination of all the features (which is learned by the algorithms) contributes to the final probability of loss prediction.
 
@@
 However, this method seems less practical as the case study projects above `r best.t` would need to increase fees by 46% in order to generate a 15% profit.
-A proposed fee increase of that magnitude may not be as palatable as an hourly rate contract depending on the project and client.
+A proposed fee increase of that magnitude may not be as palatable as an hourly rate contract, but this depends on the project and client.
 Contracts could be adjusted in many other creative ways to reduce risk in high probability projects.
-The threshold of `r best.t` was optimised based on the accept/reject strategy but different courses of actions could be developed for specific clients or project sizes via similar analyses of threshold scenarios.
-
-To conclude, because of the promising AUC results, it was logical that the models translated into financial benefits for the company.
-The `r (best.mean-100)`% improvement in profits looks reasonable, and may be high enough to trigger further cost-benefit analyses before finally introducing the model into business decisions.
-Development of the user interface as well as some further data analyses are recommended before implementation of the model and are described in the 'User Interface and Engagement' and 'Limitations and Future Work' sections below.
-
-### User Interface and Engagement
+The threshold of `r best.t` was optimised based on the accept/reject strategy but different actions could be developed for certain clients or project sizes via similar analyses of thresholds.
+
+### Categorical Predictions Conclusion
+
+The confusion matrix statistics revealed that the most profitable method, the simple Logistic Regression blended model, was optimised at a threshold with a TPR of only `r conf.plot %>% filter(method == "simp.log", model.pred == "TP") %>% .$pc`% and a TNR of `r conf.plot %>% filter(method == "simp.log", model.pred == "FN", actual == "profit") %>% .$pc %>% .[2]`%.
+Despite the low TPR, it was higher than almost all other methods at their optimal threshold and produced notable improvements in profits to the business.
+Review of the rejected project characteristics were logical, and development of the user interface as well as some further data analyses must be completed before the model can be finally introduced into business decisions.
+These are described in the 'User Interface and Engagement' and 'Limitations and Future Work' sections below.
+
+## User Interface and Engagement
 
@@
 Decision makers did not sufficiently trust the model output which were not directly connected to their experiences.
-In many studies, case-based reasoning had more traction than more accurate models because decision makers could justify the output.
-Clearly this is an important aspect of the study to address.
-The user interface should do two things well:
+In many studies, CBR had more traction than more accurate models because decision makers could justify the output.
+Clearly this is an important aspect of the study to address and could be managed at the point where the decision maker engages with the algorithm: the user interface.
+This interface should do two things well:
 
@@
 
-User interfaces also allow input values to be tweaked and re-run so the decision maker can test different scenarios.
+User interfaces also allow input values to be tweaked and re-run so the decision maker can test different scenarios (active sensitivity analysis).
 A prototype user interface was developed using the Shiny package in R where a strategy was developed for how the input options and results would be communicated [@shiny].
@@
 The data represents the company's records of *completed* projects so initially expected variable values were not documented.
-For example, if a project was originally expected to last 1.5 years but instead took 2, the company data only retains the two year time span.
+For example, if a project was originally expected to last 1.5 years but instead took 2 years, the company data only retains the two year time span.
 This prompts the question, how well could a user predict the variables before beginning a project?
@@
 The general aim was to use statistical techniques to model the profitability of projects for a case study consulting business using their internal CRM data.
-This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitability and as a binary classification problem predicting simply profit or loss.
-Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees and multiple methods of blending the individual models.
+This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitability and as a binary classification problem predicting either profit or loss.
+Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models.
 
@@
 Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
-This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business scenario.
+This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business decision scenario.
 The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss making'.
-A range of threshold were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
+A range of threshold values were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
 If a project was accepted, the profits and losses were summed while if a project was rejected the profits and losses were forfeited.
 A perfect model would reject all loss making jobs and accept all profitable jobs, thereby improving overall profits.
-Final results showed the Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
-The 95% confidence interval for these improvements were between `r best.fin %>% filter(method == "simp.log") %>% .$lower_bound_95CI - 100`% and `r best.fin %>% filter(method == "simp.log") %>% .$upper_bound_95CI - 100`% using a probability threshold of `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).
+Final results showed the simple Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
+The 95% confidence interval for these improvements was between `r best.fin %>% filter(method == "simp.log") %>% .$lower_bound_95CI - 100`% and `r best.fin %>% filter(method == "simp.log") %>% .$upper_bound_95CI - 100`% using a probability threshold of `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).
 
 These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
-Ensemble tree methods and blending have been applied minimally to cost estimation.
+Ensemble tree methods and blending had been applied minimally to cost estimation to date.
 Past research tested regression models and to some extent Neural Networks and CBR.
-The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting as well as Neural Networks.
-Next, a clear framework for how a business can tangibly improve profits was developed and proven successful.
+The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting at a similar level to Neural Networks.
+Next, a clear framework for how a business could tangibly improve profits was developed and proven successful.
 Previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
@@
 Previous studies had shown that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
-This was addressed by providing intuitive plots of the model structure and a parallel nearest neighbour graph of similar internal projects.
+This was addressed by providing intuitive plots of the model structure and a parallel nearest neighbour graph of similar internal projects to provide a clear and intuitive 'outside view'.
 Further work is required to test user confidence in the output.
@@
 Initial estimates of input variables could also be kept in the database for future analysis.
-Overall, this project has successfully demonstrated a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models could guide managers in rejecting loss making projects and improve net profitability.
+
+Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models, built on a consulting company's internal project data.
+This blended model can predict whether a project will be profitable or not and in a reasonable business decision framework, can guide managers in rejecting financially risky projects and improving final profitability.
 

### 2016-08-05 finalwriteup/draft.rmd
can't remember what I changed -- some grammar

@@
 Engaging in loss making jobs for fixed fees is a major problem in consulting, particularly in the competitive construction industry.
-This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs tactfully choose when to enforce stricter contracts.
+This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs or help tactfully choose when to enforce stricter contracts.
 It was found that in a specific decision framework, a case study's profits could be improved 9% by avoiding approximately 4% of projects.
@@
         scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
+                          name = "Significantly different from Boosted Trees") +
         scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly Different from Boosted Trees") +
+                          name = "Significantly different from Boosted Trees") +
         labs(x = "Method", y = "Area Under Curve (AUC)",
@@
 
-mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
+mice.totAUC <- readRDS("mice_df_auc.rds")
 
@@
 
-cfor.varimp = readRDS("report_data/cfor_blend_varimp.rds")
+cfor.varimp = readRDS("cfor_blend_varimp.rds")
 cfor.varimp <- mutate(cfor.varimp, var = revalue(var, c("norm.log" = "log.pred",
@@
 
-log.varimp = readRDS("report_data/logblend.rds")
+log.varimp = readRDS("logblend.rds")
 log.varimp = summary(log.varimp)$coefficients %>% as.data.frame
@@
 
-auc.list <- readRDS("report_data/blend_auc2.rds")
+auc.list <- readRDS("blend_auc2.rds")
 means = lapply(auc.list, mean) %>% unlist
@@
 ```{r blend_vs_initial}
-auc.150 = readRDS("report_data/blend_auc150.rds")
-mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
+auc.150 = readRDS("blend_auc150.rds")
+mice.totAUC <- readRDS("mice_df_auc.rds")
 compare.blends <- append(auc.150, mice.totAUC)
@@
 
-reject.plot <- readRDS('report_data/reject_plot.rds')
+reject.plot <- readRDS('reject_plot.rds')
 

### 2016-08-05 finalwriteup/trial_code_eq.rmd
just conclusion left now
can't remember what I changed -- some grammar

@@
 
-$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee + Business\ Costs)$
-
-$f(x) = \beta_{0} + \sum_{j = 1}^{p} X_{j} \beta_{j}$
-
-$RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$
-
-$RMSE_b = \sqrt{\frac{\sum_{t = 1}^{n} (\widehat{y}_t - \bar y)^2}{n}}$
+$Relative\ Profit_t = \frac{Profit_t}{\sum_{t = 1}^{n}\left|Profit_t\right|}$
 
@@
 \begin{aligned}
-RMSE_b = \textrm{baseline root mean squared error} \\
-n &= \textrm{number of data points} \\
-\widehat{y} &= \textrm{predicted values of the response variable} \\
-\bar y &= \textrm{mean 'return per dollar' for all projects} \\
+Profit_t &= \textrm{Profit from project}\ t\textrm{. Negative values indicate loss making projects}
 \end{aligned}
@@
 $\Delta\ error = RMSE_b - RMSE$
+
+to losses of \$`r 156100 %>% format(digits = 2)`
+
+```{r}
+finance.plot = readRDS("plot_finance_100_4.rds" )
+
+finance.plot$method <- factor(finance.plot$method, 
+                                    levels=c("average", "simp.log","simp.boost",
+                                             "comp.boost","comp.rf", "FWLS","orig.log",
+                                             "orig.rf","orig.boost"))
+
+ggplot(finance.plot %>% filter(method == "average") %>% slice(43:63), 
+       aes(x= threshold, y= profit.ratio, colour = method)) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
+        geom_line() +
+        geom_point() +
+        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
+        theme(legend.position = "none", 
+              panel.background = element_rect(fill = 'grey95'),
+              panel.grid.minor = element_blank()) +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Example Profit Curve for Simple Average Blended Model")
+
+```

### 2016-08-10 finalwriteup/trial_code_eq.rmd
not ssure

@@
 ```
+
+```{r reduced_violin, fig.height = 4.4, fig.width = 5.2, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Subsets of the data were used for Logistic Regression and randomForests in order to provide datasets without missing values.")}
+
+C1 = "darkred"
+C2 = "brown1"
+
+reduc.totAUC <- readRDS("reduc_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(reduc.totAUC[[2]], reduc.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+reduc.totAUC = melt(reduc.totAUC)
+reduc.totAUC$sigdiff = ifelse(reduc.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff), alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "white", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "white", 
+                     hjust = -.2, angle = 40, vjust = -.2, size = 3
+                     ) +
+        scale_fill_manual(values = c(C1, C2),
+                          name = "Significantly different from Boosted Trees") +
+        scale_colour_manual(values = c(C1, C2),
+                          name = "Significantly different from Boosted Trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of Each Method") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
+
+```
+
+```{r mice_violin, fig.height = 4.4, fig.width = 5.2, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Each method was fed the same imputed full dataset.")}
+
+mice.totAUC <- readRDS("mice_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+mice.totAUC = melt(mice.totAUC)
+mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff), 
+                    alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "white", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "white", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        scale_fill_manual(values = c(C1, C2),
+                          name = "Significantly different from boosted trees") +
+        scale_colour_manual(values = c(C1, C2),
+                          name = "Significantly different from boosted trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 Models of Each Method - Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom", plot.title = element_text(size = rel(1)))
+```
+
+```{r all_violin, fig.height = 4.2, fig.width = 6, fig.cap = c("Combined violin plots comparing AUC results for each method predicting profit/loss using subsets of data vs.
+the full imputed data set.")}
+
+reduc.totAUC$Data_Source = "Original"
+mice.totAUC$Data_Source = "Imputed"
+combine.micer <- rbind(reduc.totAUC, mice.totAUC)
+
+combine.micer$Data_Source <- factor(combine.micer$Data_Source, 
+                                    levels=c("Original", "Imputed"))
+
+ggplot(combine.micer, aes(x= factor(L1), y= value)) +
+        geom_violin(aes(fill = Data_Source, colour = Data_Source)) +
+        # stat_summary(fun.y="mean", geom="point", colour = "gray60", shape =3) +
+        scale_fill_manual(values = c(C1, C2),
+                          name = "Data Source") +
+        scale_colour_manual(values = c(C1, C2),
+                          name = "Data Source") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "Comparison of Original Data vs. Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
+
+```
+
+```{r, fig.height = 3.8, fig.width = 5.5, fig.cap = c("Variable importance output from a Boosted Tree blended model.
+The results from the three best performing models were added as explanatory variables")}
+
+boost.varimp = readRDS("boostblend.rds")
+boost.varimp = add_rownames(boost.varimp)
+boost.varimp <- mutate(boost.varimp, rowname = revalue(rowname, 
+                                                       c("norm.log" = "log.pred",
+                                         "norm.rf" = "rf.pred",
+                                         "norm.boost" = "boost.pred")))
+# order levels
+boost.varimp$rowname <- factor(boost.varimp$rowname,
+                               levels= boost.varimp$rowname)
+
+#plot p-values
+ggplot(data=boost.varimp, aes(x=rowname, y=rel.inf)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) +
+        geom_bar(stat='identity') +
+        labs(title = "Boosted Trees Variable Importance: Blended Model",
+             x ="", y = "Variable Importance Score")
+
+```
+
+```{r blend_violin, fig.height = 4.6, fig.width = 6, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
+100 models were built for each method.")}
+
+auc.list <- readRDS("blend_auc2.rds")
+means = lapply(auc.list, mean) %>% unlist
+greater = means[means> means["orig.log"]] %>% names
+auc.melt = melt(auc.list)
+auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)
+
+ggplot(auc.melt, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = greater, colour = greater), 
+                    alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "white", 
+                     shape =4) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "white", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        theme(axis.text.x=element_text(angle=40,hjust=1)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 Blended Models of each Method") +
+        scale_x_discrete(labels = c("boost" = "BT Blend", 
+                         "log" = "FWLS", 
+                         "average" = "Simple Average Blend",
+                         "orig.log" = "Orignal LR", 
+                         "rf" = "RF Blend",
+                         "simp.boost" = "Simple BT Blend",
+                         "simp.log" = "Simple LR Blend")) +
+        scale_fill_manual(values = c(C2, C1),
+                          name = "AUC Greater than Original Logistic Regression") +
+        scale_colour_manual(values = c(C2, C1),
+                          name = "AUC Greater than Original Logistic Regression") +
+        theme(legend.position = "bottom")
+
+```
+
+```{r, fig.height = 4.5, fig.width = 6, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
+150 models were built for each method to achieve a statistical power of 0.8.")}
+
+auc.150 = readRDS("blend_auc150.rds")
+
+index = c(2:3)
+pvals = data.frame("method" = names(auc.150)[c(2:3)], 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(auc.150[[1]],
+                                  auc.150[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric }
+
+signif = pvals %>% filter(p_value<0.05) %>% .$method
+auc.150 = melt(auc.150)
+auc.150$signif = ifelse(auc.150$L1 %in% signif, TRUE, FALSE)
+
+ggplot(auc.150, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = signif, colour = signif),
+                    alpha = .8) + 
+        stat_summary(fun.y = "mean", geom="point", 
+                     colour = "white", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "white", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        theme(axis.text.x=element_text(angle=40,hjust=1)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 150 Blended Models of Best Methods") +
+        scale_x_discrete(labels = c( 
+                         "average" = "Simple Average Blend",
+                         "orig.log" = "Orignal LR",
+                         "simp.log" = "Simple LR Blend")) +
+        scale_fill_manual(values = c(C2, C1),
+                          name = "AUC Significantly Greater than Original Logistic Regression") +
+        scale_colour_manual(values = c(C2, C1),
+                          name = "AUC Significantly Greater than Original Logistic Regression") +
+        theme(legend.position = "bottom")
+
+```
+
+```{r profit_curve, fig.height = 4.7, fig.width = 5.4, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+
+finance.plot = readRDS("plot_finance_100_4.rds" )
+finance.plot$method <- factor(finance.plot$method, 
+                                    levels=c("average", "simp.log","simp.boost",
+                                             "comp.boost","comp.rf", "FWLS","orig.log",
+                                             "orig.rf","orig.boost"))
+
+ggplot(finance.plot, aes(x= threshold, y= profit.ratio, colour = method)) + 
+        # geom_point(alpha = 0.005, shape = 16, size = 1) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) +
+        stat_summary(fun.y = mean, 
+                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
+                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
+                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
+        stat_summary(fun.y = mean, geom="line") +
+        scale_y_continuous(breaks = seq(0,110, by = 20)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .2)) +  
+        theme(legend.position = "none",
+              panel.background = element_rect(fill = 'gray95')) +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Profit Curve for Blended and Individual Models") + 
+        facet_wrap(~method, 
+                   labeller = labeller( method = c("average" = "Simple Average Blend",
+                                       "orig.boost"= "Orig. Boosted Trees",
+                                       "orig.log" = "Orig. Logistic Regression",
+                                       "orig.rf" = "Orig. Random Forest",
+                                       "simp.log" = "Simple LR Blend",
+                                       "simp.boost" = "Simple BT Blend",
+                                       "comp.boost"= "Complex BT Blend",
+                                       "comp.rf"= "Complex RF Blend",
+                                       "FWLS" = "FWLS"
+                                       )))
+
+```

### 2016-08-18 finalwriteup/trial_code_eq.rmd
it's been a while

@@
 library('scales')
+library(visdat)
 
@@
 ```
+
+```{r}
+miss <- read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all9c.csv')
+
+miss <- rename(miss, Project_type = JD.Second,
+               Client_Business = Business,
+               Timespan_cat = b.timespan.cbrt,
+               Invoiced_cat = b.inv.log,
+               Team_size = no.users,
+               Loss_making = b.rpdol,
+               Client_inv = client.totinv.log, 
+               pc.majpos = pc.majpos.log)
+
+# summary(miss)
+vis_dat(miss %>% select(-code.client, -code.contact, -mlsto,
+                        -X, -Year, -timespan.cbrt, -inv.mlsto.log,
+                        -return.pdol, -Billing.Type)) +
+        scale_fill_brewer(palette ='Paired')
+
+```
+```{r}
+test <- read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
+
+head(test)
+
+test %>% filter(mlsto == "2010.165")
+
+```

### 2016-08-23 finalwriteup/Appendix.rmd
Appendix

---
title: "Appendix"
output: word_document
---

```{r, echo=FALSE, include=FALSE}

library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('party')
library('randomForest')
library('RColorBrewer')
library('pwr')
library('scales')
library('caret')
library('pROC')
library('gbm')

# detect OS
if(.Platform$OS.type == 'windows'){
        # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
} else{
        # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
        # setwd("~/Documents/github/consult/finalwriteup/report_data")
        opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/report_data')
}

opts_chunk$set(fig.width=6.7, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)
```

# Appendix

Example results from:

## Regression models

###Linear regression

Example summary of model built with core variables only:
```{r}

all8a = read.csv("report_data/all8a.csv")[,-1]

#function below outputs summary of model
final.output<- function(df=all8a, no.tests = 50, core = TRUE) {

        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)

        for(i in 1:no.tests){

        if(core == FALSE){
        samp = sample.int(nrow(df), 1)                        
        #create sample df - use all variables that are complete from a sample row of data
        sample = df[samp,]
        sample = sample[,!is.na(sample)]
        vars= names(sample)
        vars = vars[! vars %in% c('mlsto','Year','return.pdol')]
        variables = vars
        } else{
                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
                              'no.users', 'pc.pro', 'Business')
        }

        #make reduced dataset with complete cases for all variables we want to use
        reduced = df[complete.cases(df[,variables]),]

        formula = paste("return.pdol ~ ")

        for(j in 1:length(variables)){
                formula = paste(formula, variables[j], sep= " + ")
        }

        if('inv.mlsto.log' %in% variables &
           'pc.pro' %in% variables){
                formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
        }

        if('inv.mlsto.log' %in% variables &
           'code.client' %in% variables){
                formula = paste(formula, 'inv.mlsto.log*code.client', sep= " + ")
        }

        if('JD.Second' %in% variables &
           'no.users' %in% variables){
                formula = paste(formula, 'JD.Second*no.users', sep= " + ")
        }

        samp_size = floor(0.75*nrow(reduced))
        create.train = base::sample.int(nrow(reduced), samp_size)
        train.df = reduced[create.train,]
        test.df = reduced[-create.train,]

        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
        for(k in seq_along(cats)){
                if(cats[k] %in% variables){
                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
        }
        }

        final = aov(as.formula(formula), data = train.df)

        fit= summary(final)

        }

        return(fit)
}

final.output(df=all8a, no.tests = 1, core = TRUE)

```

Example summary of model built with all variables:

```{r}

final.output(df=all8a, no.tests = 1, core = FALSE)

```

### Random Forest

Example summary of model built with core variables only:

```{r}

final.rf.output<- function(df=all8a, no.tests = 50, core = TRUE) {

        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)

        for(i in 1:no.tests){

        if(core == FALSE){
        samp = sample.int(nrow(df), 1)                        
        #create sample df - use all variables that are complete from a sample row of data
        sample = df[samp,]
        sample = sample[,!is.na(sample)]
        vars= names(sample)
        vars = vars[! vars %in% c('mlsto','Year','return.pdol')]
        variables = vars
        } else{
                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
                              'no.users', 'pc.pro', 'Business')
        }

        #make reduced dataset with complete cases for all variables we want to use
        reduced = df[complete.cases(df[,variables]),]

        if('code.client' %in% variables){
        sample = factor(sample(unique(reduced$code.client),size=53,
                                     replace=FALSE))
        reduced= reduced[reduced$code.client %in% sample,]

        reduced$code.client= factor(reduced$code.client)
        }

        if('code.contact' %in% variables){
        sample1 = factor(sample(unique(reduced$code.contact),size=53,
                                     replace=FALSE))
        reduced= reduced[reduced$code.contact %in% sample1,]

        reduced$code.contact= factor(reduced$code.contact)
        }

        formula = paste("return.pdol ~ ")

        for(j in 1:length(variables)){
                formula = paste(formula, variables[j], sep= " + ")
        }

        samp_size = floor(0.75*nrow(reduced))
        create.train = base::sample.int(nrow(reduced), samp_size)
        train.df = reduced[create.train,]
        test.df = reduced[-create.train,]

        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
        for(k in seq_along(cats)){
                if(cats[k] %in% variables){
                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
        }
        }

        final = randomForest(as.formula(formula), data= train.df, mtry=5,ntree=500, importance=FALSE)

        }
        return(final)
}

rf.core <- final.rf.output(df=all8a, no.tests = 1, core = TRUE)

print(rf.core)
getTree(rf.core, 1) %>% head(20)
```

All variables:

```{r}
rf.all <- final.rf.output(df=all8a, no.tests = 1, core = FALSE)

print(rf.all)
getTree(rf.all, 1) %>% head(20)

```

## Classification models

### Logistic Regression

Example summary output imputed data:

```{r}

# load imputed data
all10mice_f = read.csv("report_data/all10mice_May16.csv")[,-1]

#function for running models
test.roc <- function(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
                     extra.var = NULL, mult = 1, x.folds = 10){

        auc<- rep(list(rep(0, length(dfs))), length(dfs))
        colours <- topo.colors(x.folds*mult)

        for(i in 1:length(dfs)){
                df = get(dfs[i])

                # mlsto is spinning errors, get rid of
                df = df[, !(names(df) %in% 'mlsto')]
                set.seed(seed)

                # create f.rpdol
                df$f.rpdol <- as.factor(df$b.rpdol)
                levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
                levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"

                #set up 10 fold cv using caret function
                folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
                #now have folds[[1]] through folds[[10]] list of rows to exclude

                #for Appendix just want one fold
                folds <- folds[[1]]

                if(mult > 1){
                        for(k in 2:mult){
                                set.seed(sample(1:1000,1))
                                folds.temp = createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
                                folds = append(folds, folds.temp)
                        }
                }

                #if extra.var not found in names(df), change to NULL
                if(!is.null(extra.var)){
                        if(!(extra.var %in% names(df))) extra.var = NULL
                }

                #define formula for algorith methods, either rf, log, boost

                if(nrow(df) == 2364){
                        if(is.null(extra.var)){
                                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

                        }}

                if(method %in% 'boost') {
                        #no JD.Second
                        formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"

                }

                if(method %in% 'naiive') {
                        formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type + JD.Second"

                }

                        #turn predictors and response into 'ordered' variables for roc() to work
                        test<- df[folds,]
                        levels(test$f.rpdol) <- c('profit', 'loss')
                        test$f.rpdol <- ordered(test$f.rpdol)

                        if(method %in% 'rf'){
                                fit <- randomForest(as.formula(formula), data = df[-folds,], 
                                            mtry = mtrys[i], ntree = 1000)
                                pred <- predict(fit, test, type = "prob")
                                pred <- pred[,2]

                        }

                        if(method %in% 'log'){
                                fit<- glm(as.formula(formula), 
                                          data = df[-folds,], family = binomial())
                                pred <- predict(fit, test,
                                                type = "response")
                        } 
                        if(method %in% 'boost'){
                                fit <- gbm(as.formula(formula), data = df[-folds,],
                                           distribution = "bernoulli", n.trees = 10000,
                                           shrinkage = 0.001, interaction.depth = 5,
                                           n.minobsinnode = 20)

                                pred <- predict(fit, test, n.tree = 10000, type = "response")
                        }

                        if(method %in% 'naiive'){
                                fit<- naiveBayes( as.formula(formula), df[-folds,])
                                pred <- predict(fit, test, type = 'raw')
                                pred<- pred[,2]
                        }

                        #pROC
                        pred.p <- pROC::roc(test$f.rpdol, pred)

                        if (i == 1) {
                        plot.roc(pred.p, col = colours[i], print.thres = F)

                        } else{
                        plot.roc(
                        pred.p, col = colours, add = TRUE , print.thres = F)
                        }

                        auc<- auc(pred.p)
                        names(auc) = method

        }

        assign("auc.func", auc, envir = .GlobalEnv)
        assign("formula", formula, envir = .GlobalEnv)
        return(fit)

}

Log.Reg <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
                     extra.var = NULL, mult = 1, x.folds = 10)
auc.func

summary(Log.Reg)

```

### Random Forest

```{r}

Rand.F <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "rf",
                     extra.var = NULL, mult = 1, x.folds = 10)
auc.func

Rand.F
summary(Rand.F)

```

### Boosted Trees

```{r}

Boost.T <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "boost",
                     extra.var = NULL, mult = 1, x.folds = 10)
auc.func

# Summary of the first tree in the boosted tree
pretty.gbm.tree(Boost.T) %>% kable

summary(Boost.T)

```

### Naive Bayes

```{r}

Naive <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "naiive",
                     extra.var = NULL, mult = 1, x.folds = 10)
auc.func

Naive

```

### Bayesian Network

## Blended Models

### Logistic Regression

#### Simple Blend

```{r}

if(.Platform$OS.type == 'windows'){

        source("C:/Users/n9232371/Documents/github/consult/functions.R")
} else{

        source("/Users/yam/Documents/github/consult/functions.R")

}

df.blend <- readRDS("report_data/dfblend.rds")

iter.auc <- function(mult =1, x.folds.iter = 5, 
                     methods = c("simp.log"),
                     df.iter = df.blend) {

        #create test and train set: 0.75/0.25
        pred.list <- vector(mode = "list", length = length(methods))
        names(pred.list) = methods

        #create model averaged column
        if("average" %in% methods){
        df.iter = mutate(df.iter, model.av = (log + rf + boost)/3) 
        }

        #set up 10 fold cv using caret function
        folds <- createFolds(df.iter$f.rpdol, k = x.folds.iter, list = T, returnTrain = F)
        #now have folds[[1]] through folds[[10]] list of rows to exclude
        folds <- folds[[1]]

                train = df.iter[-folds,]
                test = df.iter[folds,]

                if ("simp.log" %in% methods) {
                        print("simp.log")
                        fit <- glm(f.rpdol ~ log + rf + boost,
                                   data = train,
                                   family = binomial())

                        pred <- predict(fit, test, type = "response")
                        pred.list[["simp.log"]] = rbind(pred.list[["simp.log"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))

                }

                if ("simp.boost" %in% methods) {
                        print("simp.boost")
                        fit <- gbm(b.rpdol ~ norm.log + norm.rf + norm.boost, data = train,
                                   distribution = "bernoulli", n.trees = 6000,
                                   shrinkage = 0.0005, interaction.depth = 6,
                                   n.minobsinnode = 10)

                        pred <- predict(fit, test, n.tree = 6000, type = "response")
                        pred.list[["simp.boost"]] = rbind(pred.list[["simp.boost"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))

                }

                if ("orig.log" %in% methods) { 
                        print("orig.log")
                        pred <- test$log
                        pred.list[["orig.log"]] = rbind(pred.list[["orig.log"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))

                }

                if ("orig.rf" %in% methods) { 
                        print("orig.rf")
                        pred <- test$rf
                        pred.list[["orig.rf"]] = rbind(pred.list[["orig.rf"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))

                }

                if ("orig.boost" %in% methods) { 
                        print("orig.boost")
                        pred <- test$boost
                        pred.list[["orig.boost"]] = rbind(pred.list[["orig.boost"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))

                }

                if ("average" %in% methods) {
                        print("average")
                        pred <- test$model.av
                        pred.list[["average"]] = rbind(pred.list[["average"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))
                }

                if ("FWLS" %in% methods) {
                        print("FWLS")
                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + b.inv.log*norm.log + b.inv.log*norm.rf + Business*norm.boost,
                                   data = train,
                                   family = binomial())
                        pred <- predict(fit, test, type = "response")
                        pred.list[["FWLS"]] = rbind(pred.list[["FWLS"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))
                }

                if ("comp.boost" %in% methods) {
                        print("comp.boost")
                        fit <- gbm(b.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost, 
                                   data = train,
                                   distribution = "bernoulli", n.trees = 6000,
                                   shrinkage = 0.0005, interaction.depth = 3,
                                   n.minobsinnode = 10)

                        pred <- predict(fit, test, n.tree = 6000, type = "response")
                        pred.list[["comp.boost"]] = rbind(pred.list[["comp.boost"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))
                }

                if ("comp.rf" %in% methods) {
                        print("comp.rf")
                        fit <- randomForest(f.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost,
                                            data = train, mtry = 3, 
                                            ntree = 1000, importance = TRUE
                        )
                        pred <- predict(fit, test, type = "prob")
                        pred <- pred[, 2]
                        pred.list[["comp.rf"]] = rbind(pred.list[["comp.rf"]],
                                                        data.frame("index" = folds,
                                                                   "pred" = pred,
                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))
                }

        #pROC
                        pred.p <- pROC::roc(test$f.rpdol, pred)

                        plot.roc(pred.p, col = colours[i], print.thres = F)

                        auc<- auc(pred.p)  
                        assign("auc.func", auc, envir = .GlobalEnv)

        return(fit)

}

simp.LR <- iter.auc(mult =1, x.folds.iter = 5, 
                     methods = c("simp.log"),
                     df.iter = df.blend)
auc.func
summary(simp.LR)

```

#### Complex Logistic Regression Blend

```{r}
complex.LR <- iter.auc(mult =1, x.folds.iter = 5, 
                     methods = c("FWLS"),
                     df.iter = df.blend)
auc.func
summary(complex.LR)

```

### Boosted Trees

#### Simple Blend

```{r}
simp.boost <- iter.auc(mult =1, x.folds.iter = 5, 
                     methods = c("simp.boost"),
                     df.iter = df.blend)
auc.func
summary(simp.boost)

# Summary of the first tree in the boosted tree
pretty.gbm.tree(simp.boost) %>% kable

```

#### Complex Blend

```{r}
comp.boost <- iter.auc(mult =1, x.folds.iter = 5, 
                     methods = c("comp.boost"),
                     df.iter = df.blend)
auc.func
summary(comp.boost)

# Summary of the first tree in the boosted tree
pretty.gbm.tree(simp.boost) %>% kable

```

### Random Forest Complex Blend

```{r}
comp.rf <- iter.auc(mult =1, x.folds.iter = 5, 
                     methods = c("comp.rf"),
                     df.iter = df.blend)
auc.func
comp.rf
summary(comp.rf)

```

### 2016-08-24 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd
started paper and checked appendix

---
title: Prediction of Consulting Project Profitability
author:
  - name: Amy Cook, Paul Wu, Kerrie Mengersen
    email: a21.cook@qut.edu.au
    affiliation: Queensland University of Technology
    footnote: Corresponding Author
address:
  - code: Queensland University of Technology
    address: School of Mathematical Sciences, George Street, Brisbane, QLD, 4000
abstract: |
  A concise abstract is required. limit to 250 words. clearly state purpose of research, principal resutls and major conclusions. no refs

  Engaging in loss making jobs for fixed fees is a major problem in consulting, particularly in the competitive construction industry. This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs or help tactfully choose when to enforce stricter contracts. It was found that in a specific decision framework, a case study's profits could be improved 9% by declining approximately 4% of projects. Alternative decision frameworks are also proposed and evaluated. Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods. A decision scenario which rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements. The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% profit improvements. The findings from this research have the potential to assist managers in reducing losses by highlighting risky projects and guiding project-based changes to fee structures.
keywords: |
  consulting; machine learning; profitability; predictive model; construction industry; data mining
abbreviations: 
  - name: Customer Relationship Management
    footnote: Abbreviations

bibliography: mybibfile.bib
output: rticles::elsevier_article
---

_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_

1. Introduction
==========================

Clearly state the research question and objectives of the work. Briefly provide any necessary
background to frame the research question. Concisely summarize the major findings/results.
Summary of Key Related Research This section should include a brief summary of key related research.
Emphasis should be on demonstrating the foundation for the current investigation. Specifically, the
goal is to clearly delineate a gap or missing link that the current research fills. Authors should avoid
presenting a litany of past research and should focus on prior work necessary to demonstrate the
existence of the research gap addressed in the manuscript.

1 page

Intro - pick out bits from thesis

#### 1.1 Problem motivation

- amount of losses, sources

- 'outside' view, that kind of thing

#### 1.2 Case Study

2. Literature Review
====================

1 page

#### 2.1 Cost estimation in the Construction Industry and IT Industry

#### 2.2 Methods used in other business applications

3. Methods
==========

Should provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant method modifications should be described.

#### 3.1 Predictive methods

- description of each of the methods: refer thesis

#### 3.2 Procedure

- Regression, classification, blending, bottom line analysis

4. Results and Discussion
=========================

Present results clearly and concisely
discussion should explore the signficance of the results of the work, not repeat them.

#### 4.1 Predictive Analysis
- regression was attempted - failed

- binary classification - results of 5 methods

- blended models - improved results

#### 4.2 Bottom line analysis

- decision scenario

- profit curves

7. Decision Support Tool
========================

- how to generate actions from predictive results

- alternative decision scenarios

- user trust

9. Conclusions and Future Work
==============================

3/4 page

Acknowledgments
===============

This research was supported by a scholarship/ ACEMS?

References {#references .unnumbered}
==========

#### Installation

If the document class *elsarticle* is not available on your computer,
you can download and install the system package *texlive-publishers*
(Linux) or install the LaTeX package *elsarticle* using the package
manager of your TeX installation, which is typically TeX Live or MikTeX.

The author names and affiliations could be formatted in two ways:

(1) Group the authors per affiliation.

(2) Use footnotes to indicate the affiliations.

Bullet points.

-   document style

-   baselineskip

-   front matter

-   keywords and MSC codes

Here are two sample references: @Feynman1963118 [@Dirac1953888].

### 2016-08-24 finalwriteup/draft.rmd
Appendix
started paper and checked appendix

@@
 * total amount invoiced * percent of job by professional level employee
-* total amount invoiced * client id
 * project category * number of employees on the job
@@
 When all variables were combined, only 15% of the project cases had complete data and therefore the ANOVA test would eliminate 85% of the data upon execution.
-Imputing the missing data was not desirable at that stage due to the volume of missing categorical data and an inability to ensure the imputed data did not adversely influence results.
+Imputing the missing data was undesirable at that stage due to the volume of missing categorical data and an inability to ensure the imputed data did not adversely influence results.
 The second method was a complicated procedure that required dozens of models but allowed variables with missing data to be included.
@@
 These were compared against a baseline RMSE, which was the RMSE of the predicted 'return per dollar' values less the mean 'return per dollar' of all projects.
+For the core complete variables, the baseline RMSE using the mean 'return per dollar' was \$0.53 and for the random subset of projects the baseline RMSE ranged between \$0.51 and \$0.56.
 

### 2016-08-24 finalwriteup/draft.rmd
merge with home sick work

--- 2016-08-05 finalwriteup/draft.rmd
@@
 This can be achieved either algorithmically with a distance function applied to the variables, or manually by users reviewing past cases.
-The nearest neighbour algorithm has been used to find the distance between cases in previous studies.
+The Nearest Neighbour algorithm has been used to find the distance between cases in previous studies.
 It is programmed to calculate the Euclidean distance in n-dimensional space between cases, where each variable is a dimension [@Kumar2007].
@@
 An equation for a line is learned, which is defined by a constant and coefficients multiplied by each explanatory variable.
-This can be mathematically represented by the equation below:
+This can be mathematically represented by the equation below [@Hastie2009]:
 
 $f(x) = \beta_{0} + \sum_{j = 1}^{p} X_{j} \beta_{j}$
-
-[@Hastie2009]
 
@@
 The residuals represent the Y distance of each point from the line (predicted values), and the residual sum of squares is the sum of the squared value of each point's residual.
-It can be represented by the equation below:
+It can be represented by the equation below [@Hastie2009]:
 
 $RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$
-
-[@Hastie2009]
 
@@
 Then, their coefficients can be directly compared [@Putler2012].
-The magnitude of the coefficients (the slopes) indicate how *much* a variable changes with the response variable.
+The magnitude of the coefficients (the slopes) indicates how *much* a variable changes with the response variable.
 However, if two or more predictor variables are correlated, the values of their coefficients are unstable.
-More specifically, this means that if the a model has two correlated variables, and is re-run with a slightly different set of data, the coefficients for the two variables may be quite different.
+More specifically, this means that if a model has two correlated variables, and is re-run with a slightly different set of data, the coefficients for the two variables may be quite different.
 This is because explanatory variables are assumed to be independent, and if variables are correlated, it breaks this assumption.
@@
 This is because it is pragmatic and works well: sigmoid functions can be bound by 0 and 1, and a logit transformation of the probability results in a linear relationship between the log odds and X.
-Refer to the equation below where *p* represents probability:
+Refer to the equation below where *p* represents probability [@Macdonald1975]:
 
 $\operatorname{logit}(p) = \log\left(\frac{p}{1-p}\right) = \log(odds)$
-
-[@Macdonald1975]
 
@@
 Then, the linear equation describing the log odds can be transformed back to probability by taking the inverse log, i.e.
-the exponential and rearranging:
+the exponential and rearranging [@Macdonald1975]:
 
@@
 
-[@Macdonald1975]
-
 The result in a sigmoidal function bound by 0 and 1 that describes the probability of the binary response variable against X.
@@
 This similarity is called 'purity' and can be measured using several formulas.
-An example for regression (a problem with a numeric response variable) is:
+An example for regression (a problem with a numeric response variable) is [@Hastie2009]:
 
@@
 
-![Simple 4 node Bayesian Network](images/lit_review/BN_1.png)
+![Simple 4-Node Bayesian Network](images/lit_review/BN_1.png)
 
@@
 
-![4 node Bayesian Network with conditional independencies](images/lit_review/BN_2.png)
+![4-Node Bayesian Network with conditional independencies](images/lit_review/BN_2.png)
 
@@
 Also, some equations for a combination are now clones of another combination and do not require recalculation.
-Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes (as Naive Bayes is the simplest form of a Bayesian Network) and variables with many categories dramatically increases the expense [@Zhang2004].
+Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes (as Naive Bayes is the simplest form of a Bayesian Network) and variables with many categories dramatically increase the expense [@Zhang2004].
 
@@
 The aim is to find the hyperplane with the widest clear margin between data points from the two categories, while allowing a certain number of misclassifications (this is usually optimised for the problem).
-To do this, the equation for the hyperplane is expressed as follows:
+To do this, the equation for the hyperplane is expressed as follows [@Fletcher2009]:
 
@@
 The plane is expressed with vectors because they can be defined such that $\boldsymbol{w}$ is perpendicular to $\boldsymbol{x}$.
-This aids in optimisation calculations to find the hyperplane with the widest margin, i.e.
+This aids optimisation calculations to find the hyperplane with the widest margin, i.e.
 two hyperplanes with the largest perpendicular distance between them that contains less than the allowable number of misclassified points [@Fletcher2009].
@@
 Now the accuracy of the trees where the covariate has been permuted can be compared against the original Random Forest.
-For a categorical response variable, refer to the equation describing variable importance score below:
+For a categorical response variable, refer to the equation describing variable importance score below [@Breiman2005]:
 
@@
 
-RMSE_b = baseline root mean squared error
-y = predicted 'return per dollar'
-y = mean 'return per dollar' for al projects
-
 Where
@@
 \begin{aligned}
-RMSE_b = \textrm{baseline root mean squared error} \\
+RMSE_b &= \textrm{baseline root mean squared error} \\
 n &= \textrm{number of data points} \\
@@
 
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between Random Forest predictions (built on 6 core variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
 If the Random Forest model was effective, the difference would be greater than 0.")}
@@
 
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between Random Forest predictions (built on all variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
 If the Random Forest model is effective, the difference should be greater than 0.")}
@@
 On the right hand tail of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar on the left tail.
-A density plot of the unaltered RMSE values for the models are shown below:
-
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
+A density plot of the unaltered RMSE values for the models is shown below:
+
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 Random Forest models built on random subsamples of the variables")}
 
@@
 Both methods predicted at best within an average of \$0.48 for the response variable, 'return per dollar', using either core variables or a sample of core and incomplete variables.
-Random Forest predictions were much more accurate than ANOVA's with a mix of core (almost complete) and less variables while Random Forest and ANOVA performed similarly with the core variables only.
+Random Forest predictions were much more accurate than ANOVA's with a mix of core (almost complete) and less complete variables while Random Forest and ANOVA performed similarly with the core variables only.
 There was much less data available for training the models that used less-complete variables so it was interesting that Random Forest performed considerably better (RMSE of \$0.50 for Random Forest vs.
@@
 ```{r reduced_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
-Subsets of the data were used for Logistic Regression and randomForests in order to provide datasets without missing values.")}
+Subsets of the data were used for Logistic Regression and Random Forests in order to provide datasets without missing values.")}
 
@@
 The confusion matrix statistics are discussed after the profit curve results and discussion.
-Then, output from the method which produced the best profit curve was used to create two subsets of profitable and loss mkaing projects.
+Then, output from the method which produced the best profit curve was used to create two subsets of profitable and loss making projects.
 The characteristics of these subsets are compared and contrasted to gauge any overall trends.
@@
 
-```{r profit_curve, fig.height = 6, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+```{r profit_curve, fig.height = 6, fig.cap = c("Profit curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
 
@@
                                        "comp.boost"= "Complex Boosted Tree Blend",
-                                       "comp.rf"= "Complex randomForest Blend",
+                                       "comp.rf"= "Complex Random Forest Blend",
                                        "FWLS" = "FWLS"
@@
 
-```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression Blended Method")}
+```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit curve of the best performing method: the simple Logistic Regression Blended Method")}
 
@@
 #     
-# * to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the CBR approach
+# * to address the problem of trust and create insight to the past data - Nearest Neighbour algorithm. This also mimics the CBR approach
 ```
@@
 Each data set of probability results differed to the next because of the random division into the five-fold data sets used for training and testing.
-For this analysis, nine full data sets of results was chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
+For this analysis, nine full data sets of results were chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
 These were deemed a 'typical' set of predictions for each method.
@@
 The results from this model were chosen because it represented the mean improvement in profits output by this blending method.
-The plot below illustrates how the ratio of profitable to unprofitable jobs compare in the full project population vs.
+The plot below illustrates how the ratio of profitable to unprofitable jobs compares in the full project population vs.
 the rejected projects.
@@
 Industry's lack of uptake of cost estimation models was marked as a significant issue in the literature [@Akintoye2000].
-Decision makers did not sufficiently trust the model output which were not directly connected to their experiences.
+Decision makers did not sufficiently trust the model output which was not directly connected to their experiences.
 In many studies, CBR had more traction than more accurate models because decision makers could justify the output.
@@
 
-Two strategies were adopted to facilitate user trust in the model's output: illustration of the model structure and a nearest neighbour graph.
+Two strategies were adopted to facilitate user trust in the model's output: illustration of the model structure and a Nearest Neighbour graph.
 Algorithms were originally selected that could reveal their predictive structure by either variable importance and/or variable relationships.
@@
 
-![Boosted Tree Partial dependency plot for Timespan embedded in the user interface](images/app/timespan_dep.png)
-
-Secondly, the nearest neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
+![Boosted Tree partial dependency plot for 'Timespan' embedded in the user interface](images/app/timespan_dep.png)
+
+Secondly, the Nearest Neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
 The aim was to provide the user with a cross check of the algorithm's results by finding projects with the most similar features to the one currently being assessed.
-This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'nearest neighbours'.
+This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'Nearest Neighbours'.
 Some screen shots of the nearest neigbhour graphic are presented below.
@@
 
-![Tables below the chart in the application give specific details of each of the nearest neighbours](images/app/NN_table_top.png)
-
-![The nearest neighbour details are split into sections such as 'Finances' and 'Staff Details'](images/app/NN_table.png)
-
-The similar projects are calculated using a different method from the blended algorithm but it is important to have a form of cross-checking and engage the user's own experiences.
-Further development and testing is required to assess how intuitive the results are for users to interpret, how much users trust the algorithm output given the variable plots and nearest neighbour charts, and finally how trustworthiness may be further improved.
+![Tables below the chart in the application give specific details of each of the Nearest Neighbours](images/app/NN_table_top.png)
+
+![The Nearest Neighbour details are split into sections such as 'Finances' and 'Staff Details'](images/app/NN_table.png)
+
+The similar projects are calculated using a different method from the blended algorithm but it is important to have a form of cross-checking and to engage the user's own experiences.
+Further development and testing is required to assess how intuitive the results are for users to interpret, how much users trust the algorithm output given the variable plots and Nearest Neighbour charts, and finally how trustworthiness may be further improved.
 
@@
 Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'time span' to 5 or 6 broad categories (i.e.
-"1 day - 3 weeks", "3 weeks - 2.5 months", etc).
+"1 day - 3 weeks", "3 weeks - 2.5 months", etc.).
 However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables.
@@
 This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitability and as a binary classification problem predicting either profit or loss.
-Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models.
+Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models' output.
 
@@
 Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
-This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business decision scenario.
-The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss making'.
+This theory was tested using a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business decision scenario.
+The models output probability scores, so an optimal probability threshold needed to be found that would dictate whether a job was classified as 'profitable' or 'loss making'.
 A range of threshold values were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
@@
 These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
-Ensemble tree methods and blending had been applied minimally to cost estimation to date.
-Past research tested regression models and to some extent Neural Networks and CBR.
-The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting at a similar level to Neural Networks.
+Ensemble tree methods and blending had been applied minimally to cost estimation previously, and past research had favoured Linear Regression, Neural Networks and CBR.
+Ensemble tree methods are advantageous because they provide insight into the model's structure while predicting at a similar level to Neural Networks.
 Next, a clear framework for how a business could tangibly improve profits was developed and proven successful.
@@
 Lastly, the decision framework was progressed even further into a prototype application where user trust was addressed.
-Previous studies had shown that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
-This was addressed by providing intuitive plots of the model structure and a parallel nearest neighbour graph of similar internal projects to provide a clear and intuitive 'outside view'.
+Past surveys have reported that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
+This was addressed by displaying intuitive plots of the model structure as well as a Nearest Neighbour graph of similar internal projects that provide an 'outside view'.
 Further work is required to test user confidence in the output.
 
-Another topic identified for future research was to test how well managers estimate project input variables.
-Estimation of the time span, team size, total invoiced amount, and percentage of hours completed by professionals in particular should be tested.
+Another topic identified for future research was to test how well managers estimate some numeric project input variables.
+In particular, time span, team size, total invoiced amount, and percentage of hours completed by professionals should be tested as these variables were calculated post project completion.
+Time span and total invoiced amount were discretised into wide categories, which should be easier for a manager to choose between.
 Initial estimates of input variables could also be kept in the database for future analysis.
 
-Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models, built on a consulting company's internal project data.
-This blended model can predict whether a project will be profitable or not and in a reasonable business decision framework, can guide managers in rejecting financially risky projects and improving final profitability.
+Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, built on a consulting company's internal project data.
+This blended model can predict whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business.
 

### 2016-08-24 finalwriteup/Appendix.rmd

@@
 
-Example results from:
+Example results from a single model from each method. This is broken into the three main types of methods that were trialled:
+
+1. Regression models
+2. Classification models
+3. Blended classification models
 
@@
 
-###Linear regression
+### ANOVA regression
 
@@
 
-all8a = read.csv("report_data/all8a.csv")[,-1]
+all8a = read.csv("all8a.csv")[,-1]
 
@@
         vars= names(sample)
-        vars = vars[! vars %in% c('mlsto','Year','return.pdol')]
+        vars = vars[! vars %in% c('mlsto','Year','return.pdol',
+                                  'code.client', 'code.contact')]
         variables = vars
@@
 
-Example summary of model built with all variables:
+Example summary of model built with all variables. Note that each model used a subset of all variables based on the available variables for a randomly selected project:
 
@@
 # load imputed data
-all10mice_f = read.csv("report_data/all10mice_May16.csv")[,-1]
+all10mice_f = read.csv("all10mice_May16.csv")[,-1]
 
@@
 
-### Logistic Regression
-
-#### Simple Blend
+### Simple Average
 
@@
 
-df.blend <- readRDS("report_data/dfblend.rds")
+df.blend <- readRDS("dfblend.rds")
 
@@
                         print("simp.log")
-                        fit <- glm(f.rpdol ~ log + rf + boost,
+                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost,
                                    data = train,
@@
                                                                    "pred" = pred,
-                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
 
@@
                                                                    "pred" = pred,
-                                                                   "iteration" = paste("iter",ceiling(k/x.folds.iter),m, sep = "")))
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
                 }
@@
                         print("FWLS")
-                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + b.inv.log*norm.log + b.inv.log*norm.rf + Business*norm.boost,
+                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + b.inv.log*norm.log + b.inv.log*norm.rf + Business*norm.boost + JD.Second + no.users,
                                    data = train,
@@
 
-                        plot.roc(pred.p, col = colours[i], print.thres = F)
+                        plot.roc(pred.p, col = "darkblue", print.thres = F)
 
@@
 }
+
+simp.av <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("average"),
+                     df.iter = df.blend)
+auc.func
+
+```
+
+### Logistic Regression
+
+#### Simple Blend
+
+```{r}
 

### 2016-08-24 finalwriteup/Consulting Profitability Paper/Consulting_Profitability_Paper.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
%\makeatletter
%\def\ps@pprintTitle{%
% \let\@oddhead\@empty
% \let\@evenhead\@empty
% \def\@oddfoot{\it \hfill\today}%
% \let\@evenfoot\@oddfoot}
%\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Prediction of Consulting Project Profitability},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\usepackage[nomarkers]{endfloat}

\begin{document}
\begin{frontmatter}

  \title{Prediction of Consulting Project Profitability}
    \author[Queensland University of Technology]{Amy Cook, Paul Wu, Kerrie Mengersen\corref{c1}}
   \ead{a21.cook@qut.edu.au} 
   \cortext[c1]{Corresponding Author}
      \address[Queensland University of Technology]{School of Mathematical Sciences, George Street, Brisbane, QLD, 4000}

  \begin{abstract}
  A concise abstract is required. limit to 250 words. clearly state
  purpose of research, principal resutls and major conclusions. no refs

  Engaging in loss making jobs for fixed fees is a major problem in
  consulting, particularly in the competitive construction industry. This
  thesis investigates whether machine learning techniques applied to a
  company's passively collected internal data could help avoid loss making
  jobs or help tactfully choose when to enforce stricter contracts. It was
  found that in a specific decision framework, a case study's profits
  could be improved 9\% by declining approximately 4\% of projects.
  Alternative decision frameworks are also proposed and evaluated.
  Algorithmic methods such as Logistic Regression, Random Forests, Boosted
  Trees, Naive Bayes, and Bayesian Networks were applied as well as
  blended combinations of these methods. A decision scenario which
  rejected projects above a sequence of tested thresholds was run in order
  to find the optimal threshold for profit improvements. The blended
  Logistic Regression model outperformed other methods and produced a 95\%
  confidence interval of 6.5 - 11.5\% profit improvements. The findings
  from this research have the potential to assist managers in reducing
  losses by highlighting risky projects and guiding project-based changes
  to fee structures.
  \end{abstract}
   \begin{keyword} consulting; machine learning; profitability; predictive model;
construction industry; data mining \sep \end{keyword}
 \end{frontmatter}

\emph{Text based on elsarticle sample manuscript, see
\url{http://www.elsevier.com/author-schemas/latex-instructions\#elsarticle}}

\section{1. Introduction}\label{introduction}

Clearly state the research question and objectives of the work. Briefly
provide any necessary background to frame the research question.
Concisely summarize the major findings/results. Summary of Key Related
Research This section should include a brief summary of key related
research. Emphasis should be on demonstrating the foundation for the
current investigation. Specifically, the goal is to clearly delineate a
gap or missing link that the current research fills. Authors should
avoid presenting a litany of past research and should focus on prior
work necessary to demonstrate the existence of the research gap
addressed in the manuscript.

1 page

Intro - pick out bits from thesis

\paragraph{1.1 Problem motivation}\label{problem-motivation}

\begin{itemize}
\item
  amount of losses, sources
\item
  `outside' view, that kind of thing
\end{itemize}

\paragraph{1.2 Case Study}\label{case-study}

\section{2. Literature Review}\label{literature-review}

1 page

\paragraph{2.1 Cost estimation in the Construction Industry and IT
Industry}\label{cost-estimation-in-the-construction-industry-and-it-industry}

\paragraph{2.2 Methods used in other business
applications}\label{methods-used-in-other-business-applications}

\section{3. Methods}\label{methods}

Should provide sufficient detail to allow the work to be reproduced.
Methods already published should be indicated by a reference: only
relevant method modifications should be described.

\paragraph{3.1 Predictive methods}\label{predictive-methods}

\begin{itemize}
\tightlist
\item
  description of each of the methods: refer thesis
\end{itemize}

\paragraph{3.2 Procedure}\label{procedure}

\begin{itemize}
\tightlist
\item
  Regression, classification, blending, bottom line analysis
\end{itemize}

\section{4. Results and Discussion}\label{results-and-discussion}

Present results clearly and concisely discussion should explore the
signficance of the results of the work, not repeat them.

\paragraph{4.1 Predictive Analysis}\label{predictive-analysis}

\begin{itemize}
\item
  regression was attempted - failed
\item
  binary classification - results of 5 methods
\item
  blended models - improved results
\end{itemize}

\paragraph{4.2 Bottom line analysis}\label{bottom-line-analysis}

\begin{itemize}
\item
  decision scenario
\item
  profit curves
\end{itemize}

\section{7. Decision Support Tool}\label{decision-support-tool}

\begin{itemize}
\item
  how to generate actions from predictive results
\item
  alternative decision scenarios
\item
  user trust
\end{itemize}

\section{9. Conclusions and Future
Work}\label{conclusions-and-future-work}

3/4 page

\section{Acknowledgments}\label{acknowledgments}

This research was supported by a scholarship/ ACEMS?

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\paragraph{Installation}\label{installation}

If the document class \emph{elsarticle} is not available on your
computer, you can download and install the system package
\emph{texlive-publishers} (Linux) or install the LaTeX~package
\emph{elsarticle} using the package manager of your TeX~installation,
which is typically TeX~Live or MikTeX.

The author names and affiliations could be formatted in two ways:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  Group the authors per affiliation.
\item
  Use footnotes to indicate the affiliations.
\end{enumerate}

Bullet points.

\begin{itemize}
\item
  document style
\item
  baselineskip
\item
  front matter
\item
  keywords and MSC codes
\end{itemize}

Here are two sample references: Feynman and Vernon Jr. (1963; Dirac
1953).

\hypertarget{refs}{}
\hypertarget{ref-Dirac1953888}{}
Dirac, P.A.M. 1953. ``The Lorentz Transformation and Absolute Time.''
\emph{Physica} 19 (1---12): 888--96.
doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}.

\hypertarget{ref-Feynman1963118}{}
Feynman, R.P, and F.L Vernon Jr. 1963. ``The Theory of a General Quantum
System Interacting with a Linear Dissipative System.'' \emph{Annals of
Physics} 24: 118--73.
doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}.

\end{document}

### 2016-08-25 finalwriteup/draft.rmd
done!

@@
 These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.
-Furthermore,
 
@@
 The fee structure can be as creative as the engaged parties wish, but in fact tend to fit a few common models.
-A couple of examples are:
+Two examples are:
 
@@
 Clients commonly seek fixed price quotes from several consultants before selecting one or negotiating further.
-Quoted fees must be competitive.
+Hence, quoted fees must be competitive.
 
@@
 Typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb (such as a set fraction of the entire client budget for a large project).
-Such qualitative methods can result in profit margins, which ultimately depend on the skill (or luck) the manager has in predicting the amount of time a project demands.
+Such qualitative methods can result in profit margins, which ultimately depend on the skill (or luck) of the manager in predicting the amount of time a project demands.
 
@@
 Was it possible to find trends in what makes a project loss making and can these characteristics be identified before a fixed price is offered?
-Can these trends be modeled into a predictive algorithm, and if so, how could the outcomes be integrated into business decisions?
+Can these trends be modeled into a predictive algorithm, and if so, how could the outcomes become integrated into business decisions?
 Finally, if the predictive outcomes were integrated into decision-making, would they definitely improve the business' bottom line?
@@
 A system like Ma and Liu's [-@Ma2014] would definitely accelerate the cost estimation process and reduce human error, however it is still an 'inside view' and could still deliver inaccurate estimates similar to detailed analysis by a person.
-The detailed view does not 'step back' and take into account set backs in construction that may have affected past similar projects.
+The detailed view does not 'step back' and take into account setbacks in construction that may have affected past similar projects.
 There is potential for this innovation to be combined with an 'outside view' system.
@@
 
-Similar to the construction industry, expert judgment or detailed analysis, is the most widely practised method for effort estimation [@Shepperd1996; @Molokken2003].
+Similar to the construction industry, expert judgment or detailed analysis, is the most widely practiced method for effort estimation [@Shepperd1996; @Molokken2003].
 This is despite many years of research being dedicated to developing algorithmic models that, in the research context, outperform expert judgment.
@@
 
-Clearly, Neural Networks can mimick a flexible range of relationships, and if more nodes are added, more flexibility is available [@Putler2012].
+Clearly, Neural Networks can mimic a flexible range of relationships, and if more nodes are added, more flexibility is available [@Putler2012].
 A previously mentioned disadvantage is that the algorithm is a black box because the internal structure is too complex for interpretation.
@@
 In contrast, linear relationships are examined between continuous explanatory variables and the response variable.
-The continuous variables were normalised to kerb the effects of possible high leverage points, and logarithmic or cube root transformations were used for this.
+The continuous variables were normalised to curb the effects of possible high leverage points, and logarithmic or cube root transformations were used for this.
 Trials of various variable interactions revealed that interactions between:
@@
 The two regression models predicting 'return per dollar' resulted in high mean RMSE's ranging from \$`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`.
+Example individual results from the ANOVA Linear Regression and Random Forest regression models can be found in the Appendix, section 11.1.
 The next section discusses possible reasons for the poor results and the decision to halt development of regression models.
@@
 
+Summaries of an example model from each of the individual classification methods can be found in the Appendix, section 11.2.
+
 ## Binary Classification: Discussion
@@
 Random Forests and Boosted Trees passively interact explanatory variables, but for Logistic Regression, these must be manually input.
-Only interaction terms with statistically significant p-values from the Logisitic Regression were included in order to reduce noise.
+Only interaction terms with statistically significant p-values from the Logistic Regression were included in order to reduce noise.
 
@@
 Initially 100 train/test runs were performed for each blending method by running 20 iterations of 5-fold cross validation.
+Summaries of a single example model from each of the blending methods can be found in the Appendix, section 11.3.
 The AUC results of these iterations are displayed below:
@@
 It was encouraging that blended models produced higher AUC values than the original Logistic Regression model, however for the case study business, the real measure of success is the model's impact on the business' overall profits.
-Higher AUC values did not guaruntee more lucrative profit curves so several methods, including the original individual models as well as blended methods, were progressed to the profit analysis.
+Higher AUC values did not guarantee more lucrative profit curves so several methods, including the original individual models as well as blended methods, were progressed to the profit analysis.
 
@@
 
-The plot below illustrates the distrubtion of profit curves for each method, where the solid lines join the mean values at each threshold point.
+The plot below illustrates the distribution of profit curves for each method, where the solid lines join the mean values at each threshold point.
 The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold point for the 100 models.
@@
 ```{r, fig.height = 5.5, fig.cap = c("Distribution of projects according to the probability outputs from a typical simple Logistic Regression blended model.
-A bargraph of counts and relative profits are shown.")}
+A bar graph of counts and relative profits are shown.")}
 
@@
 
-The general aim was to use statistical techniques to model the profitability of projects for a case study consulting business using their internal CRM data.
+The general aim was to use statistical techniques to predict the profitability of projects for a case study consulting business using their internal CRM data.
 This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitability and as a binary classification problem predicting either profit or loss.
@@
 
-Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
-This theory was tested using a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business decision scenario.
-The models output probability scores, so an optimal probability threshold needed to be found that would dictate whether a job was classified as 'profitable' or 'loss making'.
-A range of threshold values were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
-If a project was accepted, the profits and losses were summed while if a project was rejected the profits and losses were forfeited.
-A perfect model would reject all loss making jobs and accept all profitable jobs, thereby improving overall profits.
+Lastly, hypothesis 2 proposed that predictive models developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
+A range of probability threshold values were trialed for each method using a decision framework where projects scored below the threshold were accepted, while projects above the threshold were rejected.
+If a project was rejected, the profits and losses were forfeited, and the remaining accepted profits and losses were summed.
 Final results showed the simple Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
@@
 These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
-Ensemble tree methods and blending had been applied minimally to cost estimation previously, and past research had favoured Linear Regression, Neural Networks and CBR.
-Ensemble tree methods are advantageous because they provide insight into the model's structure while predicting at a similar level to Neural Networks.
-Next, a clear framework for how a business could tangibly improve profits was developed and proven successful.
-Previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
-Lastly, the decision framework was progressed even further into a prototype application where user trust was addressed.
+Ensemble tree methods and blending had been applied minimally to cost estimation previously, even though ensemble trees provide insight into model structure while predicting at a similar level to Neural Networks.
+Next, previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
+This study presented a clear framework for how a business could improve profits by applying the algorithm.
+Lastly, a prototype application was developed where insight into the model structure encouraged user trust  and a Nearest Neighbour chart presented an ‘outside view’ based on the company’s projects.
 Past surveys have reported that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
-This was addressed by displaying intuitive plots of the model structure as well as a Nearest Neighbour graph of similar internal projects that provide an 'outside view'.
 Further work is required to test user confidence in the output.
@@
 Time span and total invoiced amount were discretised into wide categories, which should be easier for a manager to choose between.
-Initial estimates of input variables could also be kept in the database for future analysis.
-
-Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, built on a consulting company's internal project data.
+
+Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, from a consulting company's internal project data.
 This blended model can predict whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business.
 
+# Appendix
+
+Example results from a single model from each method. This is broken into the three main types of methods that were trialled:
+
+1. Regression models
+2. Classification models
+3. Blended classification models
+
+## Regression models
+
+ANOVA Linear Regression and Random Forests were applied to profitability as a regression problem.
+
+### ANOVA regression
+
+Example summary of model built with core variables only:
+```{r}
+
+all8a = read.csv("all8a.csv")[,-1]
+
+#function below outputs summary of model
+final.output<- function(df=all8a, no.tests = 50, core = TRUE) {
+
+        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)
+
+        for(i in 1:no.tests){
+
+        if(core == FALSE){
+        samp = sample.int(nrow(df), 1)                        
+        #create sample df - use all variables that are complete from a sample row of data
+        sample = df[samp,]
+        sample = sample[,!is.na(sample)]
+        vars= names(sample)
+        vars = vars[! vars %in% c('mlsto','Year','return.pdol',
+                                  'code.client', 'code.contact')]
+        variables = vars
+        } else{
+                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
+                              'no.users', 'pc.pro', 'Business')
+        }
+
+        #make reduced dataset with complete cases for all variables we want to use
+        reduced = df[complete.cases(df[,variables]),]
+
+        formula = paste("return.pdol ~ ")
+
+        for(j in 1:length(variables)){
+                formula = paste(formula, variables[j], sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'pc.pro' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'code.client' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*code.client', sep= " + ")
+        }
+
+        if('JD.Second' %in% variables &
+           'no.users' %in% variables){
+                formula = paste(formula, 'JD.Second*no.users', sep= " + ")
+        }
+
+        samp_size = floor(0.75*nrow(reduced))
+        create.train = base::sample.int(nrow(reduced), samp_size)
+        train.df = reduced[create.train,]
+        test.df = reduced[-create.train,]
+
+        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
+        for(k in seq_along(cats)){
+                if(cats[k] %in% variables){
+                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
+        }
+        }
+
+        final = aov(as.formula(formula), data = train.df)
+
+        fit= summary(final)
+
+        }
+
+        return(fit)
+}
+
+final.output(df=all8a, no.tests = 1, core = TRUE)
+
+```
+
+Example summary of model built with all variables. Note that each model used a subset of all variables based on the available variables for a randomly selected project:
+
+```{r}
+
+final.output(df=all8a, no.tests = 1, core = FALSE)
+
+```
+
+### Random Forest
+
+Example summary of model built with core variables only:
+
+```{r}
+
+final.rf.output<- function(df=all8a, no.tests = 50, core = TRUE) {
+
+        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)
+
+        for(i in 1:no.tests){
+
+        if(core == FALSE){
+        samp = sample.int(nrow(df), 1)                        
+        #create sample df - use all variables that are complete from a sample row of data
+        sample = df[samp,]
+        sample = sample[,!is.na(sample)]
+        vars= names(sample)
+        vars = vars[! vars %in% c('mlsto','Year','return.pdol')]
+        variables = vars
+        } else{
+                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
+                              'no.users', 'pc.pro', 'Business')
+        }
+
+        #make reduced dataset with complete cases for all variables we want to use
+        reduced = df[complete.cases(df[,variables]),]
+
+        if('code.client' %in% variables){
+        sample = factor(sample(unique(reduced$code.client),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.client %in% sample,]
+
+        reduced$code.client= factor(reduced$code.client)
+        }
+
+        if('code.contact' %in% variables){
+        sample1 = factor(sample(unique(reduced$code.contact),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.contact %in% sample1,]
+
+        reduced$code.contact= factor(reduced$code.contact)
+        }
+
+        formula = paste("return.pdol ~ ")
+
+        for(j in 1:length(variables)){
+                formula = paste(formula, variables[j], sep= " + ")
+        }
+
+        samp_size = floor(0.75*nrow(reduced))
+        create.train = base::sample.int(nrow(reduced), samp_size)
+        train.df = reduced[create.train,]
+        test.df = reduced[-create.train,]
+
+        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
+        for(k in seq_along(cats)){
+                if(cats[k] %in% variables){
+                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
+        }
+        }
+
+        final = randomForest(as.formula(formula), data= train.df, mtry=5,ntree=500, importance=FALSE)
+
+        }
+        return(final)
+}
+
+rf.core <- final.rf.output(df=all8a, no.tests = 1, core = TRUE)
+
+print(rf.core)
+getTree(rf.core, 1) %>% head(20)
+```
+
+All variables:
+
+```{r}
+rf.all <- final.rf.output(df=all8a, no.tests = 1, core = FALSE)
+
+print(rf.all)
+getTree(rf.all, 1) %>% head(20)
+
+```
+
+## Classification models
+
+Logistic Regression, Random Forests, Boosted Trees, Naïve Bayes, and Bayesian Networks were applied to profitability prediction as a classification problem (will a project be profitable or not?)
+
+### Logistic Regression
+
+Example summary output imputed data:
+
+```{r}
+
+# load imputed data
+all10mice_f = read.csv("all10mice_May16.csv")[,-1]
+
+#function for running models
+test.roc <- function(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
+                     extra.var = NULL, mult = 1, x.folds = 10){
+
+        auc<- rep(list(rep(0, length(dfs))), length(dfs))
+        colours <- topo.colors(x.folds*mult)
+
+        for(i in 1:length(dfs)){
+                df = get(dfs[i])
+
+                # mlsto is spinning errors, get rid of
+                df = df[, !(names(df) %in% 'mlsto')]
+                set.seed(seed)
+
+                # create f.rpdol
+                df$f.rpdol <- as.factor(df$b.rpdol)
+                levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
+                levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"
+
+                #set up 10 fold cv using caret function
+                folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
+                #now have folds[[1]] through folds[[10]] list of rows to exclude
+
+                #for Appendix just want one fold
+                folds <- folds[[1]]
+
+                if(mult > 1){
+                        for(k in 2:mult){
+                                set.seed(sample(1:1000,1))
+                                folds.temp = createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
+                                folds = append(folds, folds.temp)
+                        }
+                }
+
+                #if extra.var not found in names(df), change to NULL
+                if(!is.null(extra.var)){
+                        if(!(extra.var %in% names(df))) extra.var = NULL
+                }
+
+                #define formula for algorith methods, either rf, log, boost
+
+                if(nrow(df) == 2364){
+                        if(is.null(extra.var)){
+                                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
+
+                        }}
+
+                if(method %in% 'boost') {
+                        #no JD.Second
+                        formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
+
+                }
+
+                if(method %in% 'naiive') {
+                        formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type + JD.Second"
+
+                }
+
+                        #turn predictors and response into 'ordered' variables for roc() to work
+                        test<- df[folds,]
+                        levels(test$f.rpdol) <- c('profit', 'loss')
+                        test$f.rpdol <- ordered(test$f.rpdol)
+
+                        if(method %in% 'rf'){
+                                fit <- randomForest(as.formula(formula), data = df[-folds,], 
+                                            mtry = mtrys[i], ntree = 1000)
+                                pred <- predict(fit, test, type = "prob")
+                                pred <- pred[,2]
+
+                        }
+
+                        if(method %in% 'log'){
+                                fit<- glm(as.formula(formula), 
+                                          data = df[-folds,], family = binomial())
+                                pred <- predict(fit, test,
+                                                type = "response")
+                        } 
+                        if(method %in% 'boost'){
+                                fit <- gbm(as.formula(formula), data = df[-folds,],
+                                           distribution = "bernoulli", n.trees = 10000,
+                                           shrinkage = 0.001, interaction.depth = 5,
+                                           n.minobsinnode = 20)
+
+                                pred <- predict(fit, test, n.tree = 10000, type = "response")
+                        }
+
+                        if(method %in% 'naiive'){
+                                fit<- naiveBayes( as.formula(formula), df[-folds,])
+                                pred <- predict(fit, test, type = 'raw')
+                                pred<- pred[,2]
+                        }
+
+                        #pROC
+                        pred.p <- pROC::roc(test$f.rpdol, pred)
+
+                        if (i == 1) {
+                        plot.roc(pred.p, col = colours[i], print.thres = F)
+
+                        } else{
+                        plot.roc(
+                        pred.p, col = colours, add = TRUE , print.thres = F)
+                        }
+
+                        auc<- auc(pred.p)
+                        names(auc) = method
+
+        }
+
+        assign("auc.func", auc, envir = .GlobalEnv)
+        assign("formula", formula, envir = .GlobalEnv)
+        return(fit)
+
+}
+
+Log.Reg <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+summary(Log.Reg)
+
+```
+
+### Random Forest
+
+```{r}
+
+Rand.F <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "rf",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+Rand.F
+summary(Rand.F)
+
+```
+
+### Boosted Trees
+
+```{r}
+
+Boost.T <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "boost",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+# Summary of the first tree in the boosted tree
+pretty.gbm.tree(Boost.T) %>% kable
+
+summary(Boost.T)
+
+```
+
+### Naive Bayes
+
+```{r}
+
+Naive <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "naiive",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+Naive
+
+```
+
+### Bayesian Network
+
+[@Genie]
+
+## Blended Models
+
+6 different methods for blending were applied. Three were simple and 3 were complex as listed below:
+Simple:
+
+* Averaging
+* Logistic Regression
+* Boosted Trees
+
+Complex:
+
+* Logistic Regression
+* Boosted Trees 
+* Random Forests
+
+### Simple Average
+
+```{r}
+
+if(.Platform$OS.type == 'windows'){
+
+        source("C:/Users/n9232371/Documents/github/consult/functions.R")
+} else{
+
+        source("/Users/yam/Documents/github/consult/functions.R")
+
+}
+
+df.blend <- readRDS("dfblend.rds")
+
+iter.auc <- function(mult =1, x.folds.iter = 5, 
+                     methods = c("simp.log"),
+                     df.iter = df.blend) {
+
+        #create test and train set: 0.75/0.25
+        pred.list <- vector(mode = "list", length = length(methods))
+        names(pred.list) = methods
+
+        #create model averaged column
+        if("average" %in% methods){
+        df.iter = mutate(df.iter, model.av = (log + rf + boost)/3) 
+        }
+
+        #set up 10 fold cv using caret function
+        folds <- createFolds(df.iter$f.rpdol, k = x.folds.iter, list = T, returnTrain = F)
+        #now have folds[[1]] through folds[[10]] list of rows to exclude
+        folds <- folds[[1]]
+
+                train = df.iter[-folds,]
+                test = df.iter[folds,]
+
+                if ("simp.log" %in% methods) {
+                        print("simp.log")
+                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost,
+                                   data = train,
+                                   family = binomial())
+
+                        pred <- predict(fit, test, type = "response")
+                        pred.list[["simp.log"]] = rbind(pred.list[["simp.log"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("simp.boost" %in% methods) {
+                        print("simp.boost")
+                        fit <- gbm(b.rpdol ~ norm.log + norm.rf + norm.boost, data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 6,
+                                   n.minobsinnode = 10)
+
+                        pred <- predict(fit, test, n.tree = 6000, type = "response")
+                        pred.list[["simp.boost"]] = rbind(pred.list[["simp.boost"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("orig.log" %in% methods) { 
+                        print("orig.log")
+                        pred <- test$log
+                        pred.list[["orig.log"]] = rbind(pred.list[["orig.log"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("orig.rf" %in% methods) { 
+                        print("orig.rf")
+                        pred <- test$rf
+                        pred.list[["orig.rf"]] = rbind(pred.list[["orig.rf"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("orig.boost" %in% methods) { 
+                        print("orig.boost")
+                        pred <- test$boost
+                        pred.list[["orig.boost"]] = rbind(pred.list[["orig.boost"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("average" %in% methods) {
+                        print("average")
+                        pred <- test$model.av
+                        pred.list[["average"]] = rbind(pred.list[["average"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+                if ("FWLS" %in% methods) {
+                        print("FWLS")
+                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + b.inv.log*norm.log + b.inv.log*norm.rf + Business*norm.boost + JD.Second + no.users,
+                                   data = train,
+                                   family = binomial())
+                        pred <- predict(fit, test, type = "response")
+                        pred.list[["FWLS"]] = rbind(pred.list[["FWLS"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+                if ("comp.boost" %in% methods) {
+                        print("comp.boost")
+                        fit <- gbm(b.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost, 
+                                   data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 3,
+                                   n.minobsinnode = 10)
+
+                        pred <- predict(fit, test, n.tree = 6000, type = "response")
+                        pred.list[["comp.boost"]] = rbind(pred.list[["comp.boost"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+                if ("comp.rf" %in% methods) {
+                        print("comp.rf")
+                        fit <- randomForest(f.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost,
+                                            data = train, mtry = 3, 
+                                            ntree = 1000, importance = TRUE
+                        )
+                        pred <- predict(fit, test, type = "prob")
+                        pred <- pred[, 2]
+                        pred.list[["comp.rf"]] = rbind(pred.list[["comp.rf"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+        #pROC
+                        pred.p <- pROC::roc(test$f.rpdol, pred)
+
+                        plot.roc(pred.p, col = "darkblue", print.thres = F)
+
+                        auc<- auc(pred.p)  
+                        assign("auc.func", auc, envir = .GlobalEnv)
+
+        return(fit)
+
+}
+
+simp.av <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("average"),
+                     df.iter = df.blend)
+auc.func
+
+```
+
+### Logistic Regression
+
+#### Simple Blend
+
+```{r}
+
+simp.LR <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("simp.log"),
+                     df.iter = df.blend)
+auc.func
+summary(simp.LR)
+
+```
+
+#### Complex Logistic Regression Blend
+
+```{r}
+complex.LR <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("FWLS"),
+                     df.iter = df.blend)
+auc.func
+summary(complex.LR)
+
+```
+
+### Boosted Trees
+
+#### Simple Blend
+
+```{r}
+simp.boost <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("simp.boost"),
+                     df.iter = df.blend)
+auc.func
+summary(simp.boost)
+
+# Summary of the first tree in the boosted tree
+pretty.gbm.tree(simp.boost) %>% kable
+
+```
+
+#### Complex Blend
+
+```{r}
+comp.boost <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("comp.boost"),
+                     df.iter = df.blend)
+auc.func
+summary(comp.boost)
+
+# Summary of the first tree in the boosted tree
+pretty.gbm.tree(simp.boost) %>% kable
+
+```
+
+### Random Forest Complex Blend
+
+```{r}
+comp.rf <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("comp.rf"),
+                     df.iter = df.blend)
+auc.func
+comp.rf
+summary(comp.rf)
+
+```
+
 # References

### 2016-08-26 finalwriteup/draft.rmd
getting paper going

@@
 
-Only three of the model blending techniques used the 8 variables determined above.
+Only three of the model blending techniques used the 6 variables determined above.
 These were FWLS, Random Forest, and Boosted Trees.
@@
 Summaries of a single example model from each of the blending methods can be found in the Appendix, section 11.3.
-The AUC results of these iterations are displayed below:
+The AUC results of the 100 iterations are displayed below:
 
@@
 All original methods did achieve improvements of profit at their optimal thresholds, but did not necessarily consistently improve profits across all cross-folded models.
-The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
-That is, it cannot be said with 95% confidence that the original methods' would produce a mean profit higher than historical profits in the given decision framework.
+The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods were above 100% on the y-axis.
+That is, it cannot be said with 95% confidence that the original methods would produce a mean profit higher than historical profits in the given decision framework.
 This framework stated that projects higher than the optimal thresholds (jobs that are likely to be loss making) should be rejected.

### 2016-08-30 finalwriteup/draft.rmd
finished minor edits

--- 2016-08-24 finalwriteup/draft.rmd
@@
 Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods.
-A decision scenario which" rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
+A decision scenario that rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
 This process was repeated on hundreds of models built from different training subsets of the data in order to obtain a 95% statistical confidence interval of profit improvements.
@@
 
-**ANOVA** - Analysis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
+**ANOVA** - Analysis of variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
 
@@
 
-**B2B** - Business to Business. Commerce transactions between two businesses.
-
-**BIM** - Building Information modeling. A digital representation of physical characteristics of an object or structure.
+**B2B** - Business to business. Commerce transactions between two businesses.
+
+**BIM** - Building information modeling. A digital representation of physical characteristics of an object or structure.
 
@@
 
-**CLV** - Customer Lifetime Value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
-
-**CRM** - Customer Relationship Management.
+**CLV** - Customer lifetime value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
+
+**CRM** - Customer relationship management.
 Refers to a common type of software used by businesses to record client and project details as well as employee time sheet records.
@@
 
+**FN** - False negative. A case marked 'positive' in a data set is incorrectly predicted 'negative' by an algorithm.
+
+**FNR** - False negative rate. The proportion of negatives that were incorrectly identified (false negatives).
+
 **FP** - False positive. A case marked 'negative' in a data set is incorrectly predicted 'positive' by an algorithm.
@@
 **FPR** - False positive rate. The proportion of positives that were incorrectly identified (false positives).
-
-**FN** - False negative. A case marked 'positive' in a data set is incorrectly predicted 'negative' by an algorithm.
 
@@
 
-**ROC** - receiver Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
+**ROC** - Receiver operating characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
 
 **TN** - True negative. A case marked 'negative' in a data set is correctly predicted 'negative' by an algorithm.
+
+**TNR** - True negative rate. The proportion of negatives that were correctly identified (true negatives).
 
@@
 As most people have experienced, these estimates can be particularly inaccurate if the task is new and difficult.
-These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.
-Furthermore,
+These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well being.
 
@@
 
-“It’s tough to make predictions, especially about the future” - Source Unknown
+“It’s tough to make predictions, especially about the future” - Danish Proverb
 
@@
 The purpose of this rule is to guard against the case where the project creeps in size, and the consultant finds themselves gradually undertaking more work and more work for the same price even though the client was happy to spend more on the overall project.
-This may account for some or in the best case, all, of their overspent time, although again a savvy client may turn to other consultants that agree to omit variation opportunities.
+This may account for some, or in the best case all, of their overspent time, although again a savvy client may turn to other consultants that agree to omit variation opportunities.
 
@@
 A survey by @Moores1992 aligned with this theory and found the most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%).
-This contrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details.
-For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate.
-According to @Lovallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project should be ignored in favour of analysing outcomes of several similar projects.
+This contrasts with an 'outside view' where a subject is analysed by comparing it to other subjects and disregarding specific subject details.
+For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate [@Lovallo2003].
+This 'outside view' can be applied to complex projects, and stipulates that the details of the current project should be ignored in favour of analysing outcomes of several similar projects.
 Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003].
@@
 
-![This model was implemented by the British government for a rail project that is still under completion [@Flyvbjerg2011]](images/lit_review/UKrail.JPG)
+![This model was implemented by the British government for a rail project that was under completion at the time of publication [@Flyvbjerg2011]](images/lit_review/UKrail.JPG)
 
@@
 
-Despite the predictive success in statistical and machine learning cost estimation models as well as the revealing outcomes from 'inside' and 'outside' view studies, reference class forecasting there has been a lack of industry uptake.
+Despite the predictive success in statistical and machine learning cost estimation models as well as the revealing outcomes from 'inside' and 'outside' view studies, there has been a lack of industry uptake.
 Two reports reviewing cost estimation in the software industry investigated reasons why data-driven forecasting models have not been translated to industry.
@@
 
-Reports on other industries were not available, however the insights are general and may explain the lack of adoption in other industries (such as the construction industry).
+Reports on other industries were unavailable, however the insights are general and may explain the lack of adoption in other industries (such as the construction industry).
 
@@
 Through experimentation, the Agile movement was developed.
-In summary, the method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
+The method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
 Furthermore, the project outcomes are continually adjusted and revised based on frequent progress review points that assess unexpected problems.
@@
 This agile methodology suits the nature of software development work, where curtailing the scope of work is relatively straightforward.
-Also, since the industry is more modern, the new methodology was able to sweep across the field and could be proven to deliver better budget and product outcomes to clients [@Rasmusson2010].
+Also, since the industry is relatively young, the new methodology was able to sweep across the field and was shown to deliver better budget and product outcomes to clients [@Rasmusson2010].
 
@@
 The industry is so large, broad and well established that it has inertia in the way projects are estimated, and contractually bound.
-Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high-level of legal responsibility and due diligence.
-
-This project aims to assist a company in the construction industry in managing problematic fixed price projects, since fixed price contracts are unavoidable.
-The blocks preventing industry adoption, such as providing a clear decision framework and incorporating expert judgment will be carefully addressed.
+Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high-level of time, legal responsibility and due diligence.
+
+This project aims to assist a company in the construction industry in managing problematic fixed price projects, since fixed price contracts can seem unavoidable.
+The blocks preventing industry adoption, such as providing a clear decision framework for how to implement the model will be carefully addressed.
 
@@
 Cost estimation is currently performed in the case study company by first carefully reviewing preliminary drawings where time costs for each task are estimated and summed by an experienced manager (the 'inside view').
-This is often crosschecked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
+This is often crosschecked with a value that is based off a percentage of the estimated final cost of construction (generally the client budget).
 No formal mathematical method of comparison to similar projects is performed.
@@
 Managers could interrogate how the model outputs were calculated and relate to the actual projects used in the model or speak to a colleague that was involved.
-This may give the presented 'outside view' a stronger meaning and more potential to influence the final fee proposal.
+This may make the presented 'outside view' more relatable and have more potential to influence the final fee proposal.
 
@@
 'return per dollar'.
-This measure represents the ratio of the amount invoiced to business costs, and can easily be compared between projects of different sizes.
+This measure represents the ratio of the project profit to business costs, and can easily be compared between projects of different sizes.
 
@@
 
-$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee + Business\ Costs)$
+$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee\ Costs + Business\ Costs)$
 
@@
 
-In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the data set is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
+In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the data set is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending were tested, and the prediction results were integrated into a decision framework for the business.
 These points are outlined in more detail below.
@@
 In past research, cost estimation data sets generally consisted of dozens of projects from a range of a companies and even countries.
-This thesis case study only had access to the company's internal data but this contained over 2,000 past jobs - far more than the usual cost estimation study and with many more variables.
+This thesis' case study only had access to the company's internal data but this contained over 2,000 past jobs - far more than the usual cost estimation study and with many more variables.
 Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry, which differs from other studies in the construction industry that estimate actual construction costs.
@@
 
-The next chapter, Chapter 3 is a thorough review of the literature to date that relates to this thesis.
+The next chapter, Chapter 3, is a thorough review of the literature to date that relates to this thesis.
 This covers work solving the cost-estimation problem, studies on statistical and machine-learning models in business applications, and a review of potential statistical and machine learning methods.
@@
 
-Researching outcomes of comparable past studies may impart how to best approach the problem and reveal shortcomings that can be improved upon.
+Assessing outcomes of comparable past studies may impart how to best approach the problem and reveal shortcomings that can be improved upon.
 This chapter provides an overview of the literature describing early and current cost estimation methods.
@@
 Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004; @Attalla2003].
-@Dissanayaka1999 suggested that Regression models could be performed first to determine variable importance and condense the number of variables, then Neural Networks could be used for better estimation.
+@Dissanayaka1999 suggested that Regression models could be performed first to determine variable importance and condense the number of variables, and then Neural Networks could be used for better estimation.
 Boosted Trees may perform as well as Neural Networks while providing even more insight than Regression models because it is possible to produce partial dependency plots.
@@
 This can be achieved either algorithmically with a distance function applied to the variables, or manually by users reviewing past cases.
-The Nearest Neighbour algorithm has been used to find the distance between cases in previous studies.
+The nearest neighbour algorithm has been used to find the distance between cases in previous studies.
 It is programmed to calculate the Euclidean distance in n-dimensional space between cases, where each variable is a dimension [@Kumar2007].
@@
 This is despite many years of research being dedicated to developing algorithmic models that, in the research context, outperform expert judgment.
-A survey by @Moores1992 found that software tools assisting a software project were in significant use, but only the project planning tools, not the cost estimating tools.
-This suggests a mental link for project managers to price jobs based on *details* as opposed to an outside view.
 Another study by @Heemstra1992 found there was no evidence estimation accuracy improved when estimation tools were used.
@@
 
-This section assesses the details and performance of the algorithmic models, important variables, and gaps in the literature.
+This section assesses the details and performance of the algorithmic models in the IT industry, including their important variables, and gaps in the literature.
 
@@
 The software industry is unique in that agile methods of delivery are changing the contractual approach of consulting [@Badenfelt2011].
-This innovative method keeps delivery times short and defines projects as something with a fixed cost and time.
-If a change occurs, other less important elements are discarded to maintain the fixed amount.
-Software development is a relatively new field in comparison to construction, where the tradition of fixed fee structures is difficult to reinvent.
-Also, portions of construction projects cannot be simply curtailed.
-
-There is a need for research on algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
-From findings in previous literature, the models should reveal their predictive structure to decision makers and be based on internal company data.
-Review of the literature revealed a need for an industry focused decision framework that outlines how to integrate the predictive model in a way that optimises profits.
+
+There is a need for research into algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
+For industry adoption, the models should reveal their predictive structure to decision makers and provide a decision framework that outlines how to integrate the predictive model in a way that optimises profits.
 
@@
 
+Where
+
+$$
+\begin{aligned}
+\beta\ \textrm{represents the set of coefficients for a model} \\
+N\ \textrm{is the number of points in a data set} \\
+y_{i}\ \textrm{is the response variable value for the}\ i\textrm{th case in the data set} \\
+f(x_{i})\ \textrm{is the predicted response variable value for the}\ i\textrm{th case in the data set}
+\end{aligned}
+$$
+
 The reason for this method's popularity is that it is simple, provides stable predictions, and can be adapted for categorical variables [@Seng2010].
@@
 Then, their coefficients can be directly compared [@Putler2012].
-The magnitude of the coefficients (the slopes) indicates how *much* a variable changes with the response variable.
+The magnitude of the coefficients (the slopes) indicate how *much* a variable changes with the response variable.
 However, if two or more predictor variables are correlated, the values of their coefficients are unstable.
-More specifically, this means that if a model has two correlated variables, and is re-run with a slightly different set of data, the coefficients for the two variables may be quite different.
+More specifically, this means that if the a model has two correlated variables, and is re-run with a slightly different set of data, the coefficients for the two variables may be quite different.
 This is because explanatory variables are assumed to be independent, and if variables are correlated, it breaks this assumption.
@@
 
-In linear regression, no assumptions are made about the distributions of the variables.
+In Linear Regression, no assumptions are made about the distributions of the variables.
 Skewed explanatory variable distributions are acceptable but it is generally better to normalise them by, for example, taking the logarithm or square root of the variable.
@@
 
-Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable (represented by 1 or 0).
+Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable.
 First, a sigmoid function is assumed as the functional form of the probabilities against an example variable X.
 This is because it is pragmatic and works well: sigmoid functions can be bound by 0 and 1, and a logit transformation of the probability results in a linear relationship between the log odds and X.
-Refer to the equation below where *p* represents probability [@Macdonald1975]:
+Refer to the equation below where *p* represents probability:
 
 $\operatorname{logit}(p) = \log\left(\frac{p}{1-p}\right) = \log(odds)$
+
+[@Macdonald1975]
 
@@
 Then, the linear equation describing the log odds can be transformed back to probability by taking the inverse log, i.e.
-the exponential and rearranging [@Macdonald1975]:
+the exponential and rearranging:
 
@@
 
+[@Macdonald1975]
+
 The result in a sigmoidal function bound by 0 and 1 that describes the probability of the binary response variable against X.
@@
 ![Plot of log odds vs.
-an 'x' variable and probability vs.
-the 'x' variable.
+an 'X' variable and probability vs.
+the 'X' variable.
 To convert log odds to probability, an exponential transformation is first made followed by the transformation from odds to probability [@Lowry2016].](images/lit_review/logreg.JPG)
@@
 
-In one case, this classifier succeeded because of its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
+In one industry application, this classifier succeeded because of its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
 Naive Bayes was used in a complex spam detection system where new spam emails or toxic text themes could be quickly added to the filtering model.
@@
 This similarity is called 'purity' and can be measured using several formulas.
-An example for regression (a problem with a numeric response variable) is [@Hastie2009]:
+An example for regression (a problem with a numeric response variable) is:
 
@@
 SSN\ \textrm{is the sum of squared errors of a node (dataset before the split)} \\
+SSL\ \textrm{is the sum of squared errors in the left dataset that resulted from the split} \\ 
+SSR\ \textrm{is the sum of squared errors in the right dataset that resulted from the split} \\
 y_i\ \textrm{is the}\ i \textrm{th response variable value} \\
-\bar{y}\ \textrm{is the mean response variable value in the node} \\
-SSL\ \textrm{is the sum of squared errors in the left dataset that resulted from the split} \\ 
-SSR\ \textrm{is the sum of squared errors in the right dataset that resulted from the split} 
+\bar{y}\ \textrm{is the mean response variable value in the node}
 \end{aligned}
@@
 
-![Simple 4-Node Bayesian Network](images/lit_review/BN_1.png)
+![Simple 4 node Bayesian Network](images/lit_review/BN_1.png)
 
@@
 
-![4-Node Bayesian Network with conditional independencies](images/lit_review/BN_2.png)
+![4 node Bayesian Network with conditional independencies](images/lit_review/BN_2.png)
 
@@
 
-In this way, Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes, for example, cannot [@Barber2012].
+In this way, Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values, while still taking into account many conditional dependencies that Naive Bayes, for example, cannot [@Barber2012].
 Also, some equations for a combination are now clones of another combination and do not require recalculation.
-Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes (as Naive Bayes is the simplest form of a Bayesian Network) and variables with many categories dramatically increase the expense [@Zhang2004].
+Calculating the conditional probability tables for each variable is still much more computationally expensive than Naive Bayes (as Naive Bayes is the simplest form of a Bayesian Network) and variables with many categories dramatically increases the expense [@Zhang2004].
 
@@
 
-Algorithmic Neural Networks were inspired by Neural Networks of the human nervous system.
+Neural Networks of the human nervous system inspired Algorithmic Neural Networks.
 In the human body, biological neurons receive electro-chemical input into their dendrites from the terminals of adjacent neurons.
@@
 The aim is to find the hyperplane with the widest clear margin between data points from the two categories, while allowing a certain number of misclassifications (this is usually optimised for the problem).
-To do this, the equation for the hyperplane is expressed as follows [@Fletcher2009]:
+To do this, the equation for the hyperplane is expressed as follows:
 
@@
 The plane is expressed with vectors because they can be defined such that $\boldsymbol{w}$ is perpendicular to $\boldsymbol{x}$.
-This aids optimisation calculations to find the hyperplane with the widest margin, i.e.
+This aids in optimisation calculations to find the hyperplane with the widest margin, i.e.
 two hyperplanes with the largest perpendicular distance between them that contains less than the allowable number of misclassified points [@Fletcher2009].
@@
 On the other hand, the output is in the form of distance to the boundary (as opposed to probability) and results are not transparent (similar to Neural Networks) [@Caruana2006].
-Also, SVM's can be very slow to train and therefore not suitable for industry purposes [@Auria2008].
+Also, SVM's can be very slow to train and sometimes not suitable for industry purposes [@Auria2008].
 Because SVM's perform well in high-dimensional space and can form non-linear rules they have been applied to protein classification (medicine) and text and image recognition [@Byun2002].
@@
 Literature pertaining closely to the cost prediction topic has been reviewed in the first section of this chapter, however it is relevant to broaden the critique to advanced machine learning methods that have been applied to general business problems.
-This is to find out whether methods that cost estimation had not put to the test succeeded in the broader literature.
+This is to discover whether other methods have succeeded in the business literature that were excluded from cost estimation.
 The simpler statistical methods have been covered already, so the focus of this section is the machine learning prediction methods developed for business decisions.
@@
 
-Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
+Naive Bayes, Random forests, and SVM's were built.
 All three models performed at similar levels for overall accuracy at around 80% (Random Forests had the best results, followed by SVM then Naive Bayes).
@@
 
-### Gap
-
+### Research Gap
+
+Prediction studies on business problems found success in a similar suite of methods as cost estimation, but with much more attention given to ensemble tree methods.
+This study aims to explore this gap.
 The limited literature on work similar to Saradhi & Palshikar's [-@Saradhi2011] Employee Churn study highlights a gap in the application of predictive techniques that use internal data to model internal decisions.
@@
 In this research, internal employee team structure as well as client records shall be used to aid prediction.
-The literature on machine learning applied to general business problems revealed SVM's were usually the most accurate technique followed by Neural Networks and ensemble tree methods.
-A gap exists in applying these cutting edge machine learning techniques to the wide range of internal decisions that businesses make.
 
@@
 Previous studies addressing the problem have generally been limited to prediction of construction costs for buildings/infrastructure and effort estimation for IT projects.
-The prediction models consisted of CBR, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM.
-This leaves a gap in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods.
+The prediction models consisted of CBR, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM's.
+This leaves a gap in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods, and propagating the research into business adoption methods.
 
@@
 It was determined that Bayesian Networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as Neural Networks while providing insights into model structure.
-Other strategies will be developed to address the problem of industry uptake as well.
+Strategies and research into how a business could integrate the model into decisions will be developed.
 
@@
 In the case of employee names, professional titles replaced the names, and client codes replaced client names.
-Project names and addresses were removed, and each project was identified by a number.
+Project names and addresses were removed, and a number identified each project.
 The resulting de-identified data consisted of three .csv files of different data structures, which are described in the next section.
@@
 
-| Job Number | Date Issued | Amount Invoiced | Amount Paid | Invoice Status |
+| Job Number | Date Issued | Amount Invoiced (\$) | Amount Paid (\$) | Invoice Status |
 |------------|-------------|-----------------|-------------|----------------|
-| 1.2.300    | 2014/09/01  | \$5,000         | \$5,000     | Paid           |
-| 1.2.300    | 2014/10/02  | \$2,000         | \$2,000     | Paid           |
-| 1.2.300    | 2014/11/04  | \$3,000         | \$3,000     | Paid           |
-| 1.2.300    | 2014/12/02  | \$4,000         | \$4,000     | Paid           |
-| 1.2.300    | 2015/02/02  | \$5,000         | \$0         | Outstanding    |
-| 1.2.400    | 2014/08/05  | \$20,000        | \$20,000    | Paid           |
-| 1.2.400    | 2014/09/01  | \$2,100         | \$0         | Outstanding    |
+| 1.2.300    | 2014/09/01  | 5,000         | 5,000     | Paid           |
+| 1.2.300    | 2014/10/02  | 2,000         | 2,000     | Paid           |
+| 1.2.300    | 2014/11/04  | 3,000         | 3,000     | Paid           |
+| 1.2.300    | 2014/12/02  | 4,000         | 4,000     | Paid           |
+| 1.2.300    | 2015/02/02  | 5,000         | 0         | Outstanding    |
+| 1.2.400    | 2014/08/05  | 20,000        | 20,000    | Paid           |
+| 1.2.400    | 2014/09/01  | 2,100         | 0         | Outstanding    |
 | ...        | ...         | ...           | ...       | ...            |\
@@
 
-* Project summary data: information describing each project such as client code, client contact code, discipline, subject, post code, director code, and suburb
+* Project summary data: information describing each project such as client code, client contact code, discipline, subject, postcode, director code, and suburb
 
@@
 |------------|------------------|----------|-------------|------------|-----------|--------|------------------|-----------------------|
-| 1.2.300    | 1 Apple St.      | D1       | "Disguised" | Civil      | 4002      | C5030  | P12              | ...                   |
-| 1.2.400    | 33 Rogue Rd.     | D2       | "Disguised" | Structural | 4122      | C2000  | P24              | ...                   |
-| 1.2.500    | 4 Window Dr.     | D3       | "Disguised" | Water      | 4230      | C2045  | P30              | ...                   |
-| 1.2.600    | 864 Hamilton St. | D2       | "Disguised" | Structural | 4004      | C3001  | P1               | ...                   |
-| 1.2.700    | 2 Shark St.      | D3       | "Disguised" | Water      | 4570      | C3020  | P5               | ...                   |
-| 1.2.800    | 24 Campbell Rd.  | D3       | "Disguised" | Water      | 4222      | C4010  | P1               | ...                   |
+| 1.2.300    | "Disguised"      | D1       | "Disguised" | Civil      | 4002      | C5030  | P12              | ...                   |
+| 1.2.400    | "Disguised"     | D2       | "Disguised" | Structural | 4122      | C2000  | P24              | ...                   |
+| 1.2.500    | "Disguised"     | D3       | "Disguised" | Water      | 4230      | C2045  | P30              | ...                   |
+| 1.2.600    | "Disguised" | D2       | "Disguised" | Structural | 4004      | C3001  | P1               | ...                   |
+| 1.2.700    | "Disguised"      | D3       | "Disguised" | Water      | 4570      | C3020  | P5               | ...                   |
+| 1.2.800    | "Disguised"  | D3       | "Disguised" | Water      | 4222      | C4010  | P1               | ...                   |
 | ...        | ...              | ...      | ...         | ...        | ...       | ...    | ...              | ...                   |\
@@
 These three sources needed to be compiled into a single data set that detailed one project per row, as the model was to predict the overall profitability of a project.
-Before compilation could begin however, thorough cleaning was required where the data was plotted and statistically assessed so that outliers could be visually or statistically discovered.
+Before compilation could begin however, thorough cleaning was required where the data was plotted and assessed so that outliers could be visually or statistically discovered.
 Outliers were then investigated for data entry errors.
@@
 Once these errors were detected, they were discussed with the company directors and appropriate corrective action to the data was taken.
-Mistakes in data entry could often be corrected by notes that a user had entered into the system.
+Notes that a user had entered into the system could often correct mistakes in data entry.
 It is possible that some data entry errors were not detected and this is a problem that also must be dealt with during implementation of the model.
@@
 * Total amount invoiced and remunerated per project
-* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500) over all invoices sent to a client across all their projects.
+* Mean invoice size per project.
+For example, if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500 for that project.
+This was done for all invoices sent to a client across all their projects and the mean was taken.
 This gives an indication of whether the client generally does fast paced big jobs, which would result in large monthly invoices for example, or small jobs with small monthly invoiced amounts.
@@
 
-* Profit
+* Profit. Note the cost of employee hours includes a margin to account for the business overheads.
 
@@
 Now the accuracy of the trees where the covariate has been permuted can be compared against the original Random Forest.
-For a categorical response variable, refer to the equation describing variable importance score below [@Breiman2005]:
+For a categorical response variable, refer to the equation describing variable importance score below:
 
@@
 
-![Hierarchical dendrogram of the Time span variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
+![Hierarchical dendrogram of the 'time span' variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
 
@@
 
-The revised discretised categories for time span and total amount invoiced were as follows:
-
-Time span Categories
+The revised discretised categories for 'time span' and 'total amount invoiced' were as follows:
+
+'Time span' Categories
 
@@
 
-Total Invoiced Amount Categories
+'Total Invoiced Amount' Categories
 
@@
 n &= \textrm{number of data points} \\
-\widehat{y} &= \textrm{predicted values of the response variable} \\
-y &= \textrm{observed values of the response variable} \\
+\widehat{y}_t &= \textrm{predicted values of the response variable} \\
+y_t &= \textrm{observed values of the response variable} \\
 \end{aligned}
@@
 In order to compare which models performed significantly better than others, an adequate sample size of results statistics is required.
-Multiple models could be made by using different data in the training vs.
-testing sets, each providing a resulting test statistic (RMSE or AUC).
+Using different data in the training vs.
+testing sets could produce multiple models, each providing a resulting test statistic (RMSE or AUC).
 5 fold cross validation was used, which created 5 models, then the division of 5 folds was repeated using a different random seed to create another set of 5 training/testing data sets.
@@
 \begin{aligned}
+Px_{i,j} = \textrm{proximity between datapoints}\ i\  \textrm{and}\ j \\
 I(\cdot)\  \textrm{is the indicator function} \\
-Px_{i,j} = \textrm{proximity between datapoints}\ i\  \textrm{and}\ j \\
 E_{n,i} = \textrm{is the end node that case}\ i\ \textrm{falls into in tree}\ n
@@
 Linear Regression must be fed complete data sets but the case study data had a core set of only 13 complete variables (including the response variable).
-If an incomplete variable was added to the core variables, the 'complete' data set would shrink to the complete rows of the incomplete variable.
-If another incomplete variable was added, the complete set of data would shrink again.
+If an incomplete variable were added to the core variables, the 'complete' data set would shrink to the complete rows of the incomplete variable.
+If another incomplete variable were added, the complete set of data would shrink again.
 Therefore, to make the most of the available data, each incomplete variable was added to the core variables one at a time, where a Linear Regression model was built for each.
@@
 
-```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of ANOVA regressions' F-statistics which were used to interpret variable importance")}
+```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of ANOVA regression F-statistics which were used to interpret variable importance")}
 
@@
         scale_y_continuous(breaks = c(0, 0.05, .25, .5, .75, 1)) +
-        labs(y = "P-Value of F-Statistic", x = "Independent Variables",
+        labs(y = "P-Value of F-Statistic", x = "Explanatory Variables",
              title = "Summary of Variable F-Statistic P-Values from ANOVA Models")
@@
         geom_bar(stat='identity') +
-        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "CForest Variable Importance Rankings of Complete Variables") +
+        labs(y = "Relative Variable Importance Value", x = "Explanatory Variables", title = "CForest Variable Importance Rankings of Complete Variables") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 Random Forests favour variables that are numeric or have many categories and this held true for the analysis.
-Its 6 highest-ranked variables were numeric with one categorical variable, Business category of the client, which had 28 categories.
+Its 6 highest-ranked variables were numeric with one categorical variable, 'business category of the client', which had 28 categories.
 The two lowest ranking variables had only 4 categories.
 It was expected that cForests would rank the categorical variables with only 4 categories more fairly.
-This was the case with Discipline and it received a mid-level ranking.
-Business category of the client and Broad business type however, were ranked similarly to Random Forests.
+This was the case with 'discipline' and it received a mid-level ranking.
+'Business category of the client' and 'broad business type' however, were ranked similarly to Random Forests.
 
 The overall qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, client characteristics, project time span, and project category.
-It was expected that Billing Type would play a more important role in predicting profitability as it describes the structure of the incoming fees.
-For this reason, and with advice from the case study business, the Billing Type was kept in the model.
-
-For IT software projects, the literature indicated size was the most important variable predicting effort, and in Construction projects size was also often an explanatory variable.
+It was expected that 'billing type' would play a more important role in predicting profitability as it describes the structure of the incoming fees.
+For this reason, and with advice from the case study business, the 'billing type' was kept in the model.
+
+For IT software projects, the literature indicated size was the most important variable predicting effort, and in construction projects size was also often an explanatory variable.
 Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996; @Finnie1997; @Pai2013].
@@
 Instead of predicting total effort, it is describing whether the size of a project correlates with how profitable the project is.
-The closest estimate of size in this case study is the invoiced amount variable, which did rank as very important.
+The closest estimate of size in this case study is the 'invoiced amount' variable, which did rank as very important.
 The exact invoiced amount is difficult to guess before a project begins, and is part of the problem this study is addressing.
-Therefore, the invoiced amount was binned into categories that a manager could much more confidently choose as an estimate.
+Therefore, the 'invoiced amount' was binned into categories that a manager could much more confidently choose as an estimate.
 The other variables indicated as important in this study (including the number of employees on a project, client business type and percent of hours by certain types of employees) have not been strongly represented in the literature to date.
@@
 
+RMSE_b = baseline root mean squared error
+y = predicted 'return per dollar'
+y = mean 'return per dollar' for al projects
+
 Where
@@
 \begin{aligned}
-RMSE_b &= \textrm{baseline root mean squared error} \\
+RMSE_b = \textrm{baseline root mean squared error} \\
 n &= \textrm{number of data points} \\
@@
 
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between Random Forest predictions (built on 6 core variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on 6 core variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
 If the Random Forest model was effective, the difference would be greater than 0.")}
@@
 
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between Random Forest predictions (built on all variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of the difference in RMSE between randomForest predictions (built on all variables) and a baseline predictor that uses the mean 'return per dollar' as its prediction.
 If the Random Forest model is effective, the difference should be greater than 0.")}
@@
 On the right hand tail of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar on the left tail.
-A density plot of the unaltered RMSE values for the models is shown below:
-
-```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 Random Forest models built on random subsamples of the variables")}
+A density plot of the unaltered RMSE values for the models are shown below:
+
+```{r, fig.height = 4.3, echo = FALSE, fig.cap = c("Distribution of RMSE's of 'return per dollar' from 100 randomForest models built on random subsamples of the variables")}
 
@@
 Both methods predicted at best within an average of \$0.48 for the response variable, 'return per dollar', using either core variables or a sample of core and incomplete variables.
-Random Forest predictions were much more accurate than ANOVA's with a mix of core (almost complete) and less complete variables while Random Forest and ANOVA performed similarly with the core variables only.
+Random Forest predictions were much more accurate than ANOVA's with a mix of core (almost complete) and less variables while Random Forest and ANOVA performed similarly with the core variables only.
 There was much less data available for training the models that used less-complete variables so it was interesting that Random Forest performed considerably better (RMSE of \$0.50 for Random Forest vs.
@@
 ```{r reduced_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
-Subsets of the data were used for Logistic Regression and Random Forests in order to provide datasets without missing values.")}
+Subsets of the data were used for Logistic Regression and randomForests in order to provide datasets without missing values.")}
 
@@
                           name = "Significantly different from Boosted Trees") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
+        labs(x = "", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 100 models of Each Method") +
@@
         scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
+                          name = "Significantly different from Boosted Trees") +
         scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from boosted trees") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
+                          name = "Significantly different from Boosted Trees") +
+        labs(x = "", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 100 Models of Each Method - Imputed Data") +
@@
                           name = "Data Source") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
+        labs(x = "", y = "Area Under Curve (AUC)",
              title = "Comparison of Original Data vs. Imputed Data") +
@@
         theme(axis.text.x=element_text(angle=40,hjust=1)) +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
+        labs(x = "", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 100 Blended Models of each Method") +
@@
 
-```{r, fig.height = 5.5, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
+```{r, fig.height = 5.5, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best two blending methods when predicting profit/loss.
 150 models were built for each method to achieve a statistical power of 0.8.")}
@@
         theme(axis.text.x=element_text(angle=40,hjust=1)) +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
+        labs(x = "", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 150 Blended Models of Best Methods") +
@@
 It was not clear why this was the case, however both simple models were compared again in the analysis of the impact each model could have on the company's overall profits.
-The analysis of overall profits did not directly relate to AUC measurements, so a range of models were carried forward for comparison.
+The analysis of overall profits did not directly relate to AUC measurements, so a range of models was carried forward for comparison.
 These included the three original individual models, three simple blended models and three complex blended models.
@@
 From the profit curves, optimal probability thresholds could be derived for each method (individual and blended methods).
-These threshold points allowed the probability output from each model to be converted to categorical output (profit or loss) which in turn could generate confusion matrix statistics for each method.
+These threshold points allowed the probability output from each model to be converted to categorical output (profit or loss), which in turn could generate confusion matrix statistics for each method.
 The confusion matrix statistics are discussed after the profit curve results and discussion.
-Then, output from the method which produced the best profit curve was used to create two subsets of profitable and loss making projects.
+Then, output from the method which produced the best profit curve was used to create two subsets of profitable and loss mkaing projects.
 The characteristics of these subsets are compared and contrasted to gauge any overall trends.
@@
 The predictive models output a 'probability' between 0 and 1 that each project will be a loss making job (where probability = 1 indicates a loss making job).
-The question the arises, at what probability would a decision-maker round the probability up to 1 or down to 0?
+The question then arises, at what probability would a decision-maker round the probability up to 1 or down to 0?
 And what business decision would then be made?
@@
 
-```{r profit_curve, fig.height = 6, fig.cap = c("Profit curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+```{r profit_curve, fig.height = 6, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
 
@@
                                        "comp.boost"= "Complex Boosted Tree Blend",
-                                       "comp.rf"= "Complex Random Forest Blend",
+                                       "comp.rf"= "Complex randomForest Blend",
                                        "FWLS" = "FWLS"
@@
 
-```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit curve of the best performing method: the simple Logistic Regression Blended Method")}
+```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression blended method")}
 
@@
 #     
-# * to address the problem of trust and create insight to the past data - Nearest Neighbour algorithm. This also mimics the CBR approach
+# * to address the problem of trust and create insight to the past data - nearest neighbour algorithm. This also mimics the CBR approach
 ```
@@
 \begin{aligned}
-Profit_t &= \textrm{Profit from project}\ t\textrm{. Negative values indicate loss making projects}
+Profit_t &= \textrm{Profit from project}\ t\textrm{. Negative values indicate loss making projects} \\
+n &= \textrm{Number of projects}
 \end{aligned}
@@
 As shown in the upper plot, a low count of projects fall above the optimal threshold for the simple Logistic Regression blend.
-The lower plot illustrates the relative profits (and losses for negative values) within each probability score bin which are normalised by the sum of the absolute value of all profits and losses.
+The lower plot illustrates the relative profits (and losses for negative values) within each probability score bin, which are normalised by the sum of the absolute value of all profits and losses.
 The model has assigned low probability scores to profitable projects and high scores to unprofitable jobs well.
@@
 All original methods did achieve improvements of profit at their optimal thresholds, but did not necessarily consistently improve profits across all cross-folded models.
-The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
+The shaded grey 95% confidence intervals in the profit curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
 That is, it cannot be said with 95% confidence that the original methods' would produce a mean profit higher than historical profits in the given decision framework.
@@
 Each data set of probability results differed to the next because of the random division into the five-fold data sets used for training and testing.
-For this analysis, nine full data sets of results were chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
+For this analysis, nine full data sets of results was chosen that produced the mean profitability ratio at each of the nine method's optimal threshold points.
 These were deemed a 'typical' set of predictions for each method.
@@
 
-```{r TP_plot, fig.height = 5, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
-
-conf.plot <- readRDS("conf_plot.rds")
+```{r TP_plot, fig.height = 5, fig.cap = c("True positive and false negative rates for each method using a typical model from each - displayed as a bar graph.")}
+
+conf.plot1 <- readRDS("report_data/conf_plot.rds")
 
 #if model.pred = loss, this is TP
-conf.plot$model.pred <- ifelse(conf.plot$actual == "loss" & 
-                                       conf.plot$model.pred == "loss", "TP", "FN")
+conf.plot1$model.pred <- ifelse(conf.plot1$actual == "loss" & 
+                                       conf.plot1$model.pred == "loss", "TP", "FN")
 # reorder levels in method to suit TP 
-conf.plot$method <- factor(conf.plot$method, 
-                           levels= conf.plot %>% filter(actual == "loss", model.pred == "TP") %>%
+conf.plot1$method <- factor(conf.plot1$method, 
+                           levels= conf.plot1 %>% filter(actual == "loss", model.pred == "TP") %>%
                                    arrange(desc(pc)) %>% .$method)
 
-ggplot(conf.plot %>% filter(actual == "loss"), 
+ggplot(conf.plot1 %>% filter(actual == "loss"), 
        aes(x = method, y = pc, fill = model.pred)) +
@@
 
-The plot above shows the Complex Boosted Tree Blend has the highest TPR followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
-All of these TPR's are below 25% which shows it was not worth rejecting a large percentage of loss making jobs in comparison to the accompanying losses in rejected profitable jobs (FP's).
+The plot above shows the complex Boosted Tree blended model has the highest TPR followed by the simple Logistic Regression blend and the simple averaged blend.
+All of these TPR's are below 25%, which shows it was not worth rejecting a large percentage of loss making jobs in comparison to the accompanying losses in rejected profitable jobs (FP's).
 This also implies the models were not able to cleanly divide loss making jobs from profitable jobs.
@@
 
-```{r TN_plot, fig.height = 5, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
+```{r TN_plot, fig.height = 5, fig.cap = c("TNR and FPR for each method using a typical model from each - displayed as a bar graph.")}
 
@@
 
-### Review of Jobs Rejected by Simple Logistic Regression
+### Review of Projects Rejected by Simple Logistic Regression
 
@@
 
-All below plots and comments summarise results from one Simple Blended Logistic Regression Model.
+All below plots and comments summarise results from one simple blended Logistic Regression model.
 The results from this model were chosen because it represented the mean improvement in profits output by this blending method.
-The plot below illustrates how the ratio of profitable to unprofitable jobs compares in the full project population vs.
+The plot below illustrates how the ratio of profitable to unprofitable jobs compare in the full project population vs.
 the rejected projects.
@@
 Originally, 25% of projects were loss making.
-The model rejects a population of projects which are over 60% loss making.
+The model rejects a population of projects that are over 60% loss making.
 This represents a significant swing in proportions, but is not perfect.
@@
 Colours are alternated between different jobs to indicate profit/loss magnitude.
-Predictions were from a typical simple Logistic Regression.")}
+Predictions were from a typical simple Logistic Regression blended model.")}
 
@@
 the rejected projects only.
-Predictions from a typical simple Logistic Regression were used.")}
+Predictions from a typical simple Logistic Regression blend were used.")}
 
@@
 
-The model tended to accept projects which were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
+The model tended to accept projects that were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
 This could either indicate that professional level employees complete projects more efficiently or that the types of projects that do not require technical employees are more profitable.
@@
 
-```{r, fig.cap = c("Proportion of each timespan category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+```{r, fig.cap = c("Proportion of each 'timespan' category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression blend were used.")}
 
@@
 The statistical distribution of 'return per dollar' in the rejected projects is skewed towards negative returns.
-This means the algorithm has identified a proportionally higher percentage of loss making projects which is not surprising given previous results.
-
-```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
+This means the algorithm has identified a proportionally higher percentage of loss making projects, which is not surprising given previous results.
+
+```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression blend were used.")}
 
@@
 
-The confusion matrix statistics revealed that the most profitable method, the simple Logistic Regression blended model, was optimised at a threshold with a TPR of only `r conf.plot %>% filter(method == "simp.log", model.pred == "TP") %>% .$pc`% and a TNR of `r conf.plot %>% filter(method == "simp.log", model.pred == "FN", actual == "profit") %>% .$pc %>% .[2]`%.
+The confusion matrix statistics revealed that the most profitable method, the simple Logistic Regression blended model, was optimised at a threshold with a TPR of only `r conf.plot1 %>% filter(method == "simp.log", model.pred == "TP") %>% .$pc`% and a TNR of `r conf.plot1 %>% filter(method == "simp.log", model.pred == "FN", actual == "profit") %>% .$pc %>% .[2]`%.
 Despite the low TPR, it was higher than almost all other methods at their optimal threshold and produced notable improvements in profits to the business.
 Review of the rejected project characteristics were logical, and development of the user interface as well as some further data analyses must be completed before the model can be finally introduced into business decisions.
-These are described in the 'User Interface and Engagement' and 'Limitations and Future Work' sections below.
-
-## User Interface and Engagement
+These are described in the 'User Interface and User Engagement' and 'Limitations and Future Work' sections below.
+
+## User Interface and User Engagement
 
 Industry's lack of uptake of cost estimation models was marked as a significant issue in the literature [@Akintoye2000].
-Decision makers did not sufficiently trust the model output which was not directly connected to their experiences.
+Decision makers did not sufficiently trust the model output that were not directly connected to their experiences.
 In many studies, CBR had more traction than more accurate models because decision makers could justify the output.
@@
 The 100 models gave 100 probability outputs for each project, and the 95% confidence interval for these outputs was calculated.
-A typical spread for the 95% confidence interval is shaded in the graphic which is relative to the 'bin' that the score fell into (i.e.
+A typical spread for the 95% confidence interval is shaded in the graphic, which is relative to the 'bin' that the score fell into (i.e.
 6-8).
@@
 This graphic is an initial suggestion for how to communicate model results to industry.
-Many variations are possible which could depend on how the business wants to act on the output; for example whether they plan to reject risky projects outright or adjust the fee structure.
+Many variations are possible, which could depend on how the business wants to act on the output; for example whether they plan to reject risky projects outright or adjust the fee structure.
 Additional methods built on top of the current model may be able to produce statistical probabilities instead of probability scores.
@@
 
-Two strategies were adopted to facilitate user trust in the model's output: illustration of the model structure and a Nearest Neighbour graph.
+Two strategies were adopted to facilitate user trust in the model's output: illustration of the model structure and a nearest neighbour graph.
 Algorithms were originally selected that could reveal their predictive structure by either variable importance and/or variable relationships.
@@
 
-![Graphic for model prediction and Boosted Tree partial dependency plot](images/app/team_size_dep.png)
-
-![Boosted Tree partial dependency plot for 'Timespan' embedded in the user interface](images/app/timespan_dep.png)
-
-Secondly, the Nearest Neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
+![Graphic for model prediction and Boosted Tree partial dependency plot embedded in the user interface](images/app/team_size_dep.png)
+
+![Boosted Tree Partial dependency plot for 'timespan' embedded in the user interface](images/app/timespan_dep.png)
+
+Secondly, the nearest neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
 The aim was to provide the user with a cross check of the algorithm's results by finding projects with the most similar features to the one currently being assessed.
-This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'Nearest Neighbours'.
+This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'nearest neighbours'.
 Some screen shots of the nearest neigbhour graphic are presented below.
 
-![User interface for Nearest Neighbour algorithm input and output chart](images/app/NN_top.png)
-
-![Nearest Neighbour chart coloured by the position of the main employee on the projects](images/app/NN_emp.png)
+![User interface for Nearest Neighbour algorithm input options and chart](images/app/NN_top.png)
+
+![User interface Nearest Neighbour chart coloured by the position of the main employee on the projects](images/app/NN_emp.png)
 
@@
 
-![Tables below the chart in the application give specific details of each of the Nearest Neighbours](images/app/NN_table_top.png)
-
-![The Nearest Neighbour details are split into sections such as 'Finances' and 'Staff Details'](images/app/NN_table.png)
-
-The similar projects are calculated using a different method from the blended algorithm but it is important to have a form of cross-checking and to engage the user's own experiences.
-Further development and testing is required to assess how intuitive the results are for users to interpret, how much users trust the algorithm output given the variable plots and Nearest Neighbour charts, and finally how trustworthiness may be further improved.
+![Tables below the chart in the application give specific details of each of the nearest neighbours](images/app/NN_table_top.png)
+
+![The nearest neighbour details are split into sections such as 'Finances' and 'Staff Details'](images/app/NN_table.png)
+
+The similar projects are calculated using a different method from the blended algorithm but it is important to have a form of cross-checking and engage the user's own experiences.
+Further development and testing is required to assess how intuitive the results are for users to interpret, how much users trust the algorithm output given the variable plots and nearest neighbour charts, and finally how trustworthiness may be further improved.
 
@@
 Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'time span' to 5 or 6 broad categories (i.e.
-"1 day - 3 weeks", "3 weeks - 2.5 months", etc.).
-However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables.
+"1 day - 3 weeks", "3 weeks - 2.5 months", etc).
+However, two remaining categories: 'team size' and 'percent of project completed by professional level employees' were left as continuous variables.
 It would be interesting to test how well business managers could guess the aforementioned four variables at the start of a project.
@@
 However, it would be worthwhile integrating into the pre-existing data-collection system.
-When a new project is entered into the database, it could be mandated that the project manager guesses the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent by professionals.
+When a new project is entered into the database, it could be mandated that the project manager guess the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent by professionals.
 Over time, this estimation data could contribute to a revised version of the algorithm.
@@
 This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitability and as a binary classification problem predicting either profit or loss.
-Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models' output.
+Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models.
 
@@
 Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
-This theory was tested using a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business decision scenario.
-The models output probability scores, so an optimal probability threshold needed to be found that would dictate whether a job was classified as 'profitable' or 'loss making'.
+This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business decision scenario.
+The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss making'.
 A range of threshold values were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
@@
 These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
-Ensemble tree methods and blending had been applied minimally to cost estimation previously, and past research had favoured Linear Regression, Neural Networks and CBR.
-Ensemble tree methods are advantageous because they provide insight into the model's structure while predicting at a similar level to Neural Networks.
+Ensemble tree methods and blending had been applied minimally to cost estimation to date.
+Past research tested regression models and to some extent Neural Networks and CBR.
+The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting at a similar level to Neural Networks.
 Next, a clear framework for how a business could tangibly improve profits was developed and proven successful.
@@
 Lastly, the decision framework was progressed even further into a prototype application where user trust was addressed.
-Past surveys have reported that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
-This was addressed by displaying intuitive plots of the model structure as well as a Nearest Neighbour graph of similar internal projects that provide an 'outside view'.
+Previous studies had shown that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
+This was addressed by providing intuitive plots of the model structure and a parallel nearest neighbour graph of similar internal projects to provide a clear and intuitive 'outside view'.
 Further work is required to test user confidence in the output.
 
-Another topic identified for future research was to test how well managers estimate some numeric project input variables.
-In particular, time span, team size, total invoiced amount, and percentage of hours completed by professionals should be tested as these variables were calculated post project completion.
-Time span and total invoiced amount were discretised into wide categories, which should be easier for a manager to choose between.
+Another topic identified for future research was to test how well managers estimate project input variables.
+Estimation of the time span, team size, total invoiced amount, and percentage of hours completed by professionals in particular should be tested.
 Initial estimates of input variables could also be kept in the database for future analysis.
 
-Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, built on a consulting company's internal project data.
-This blended model can predict whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business.
+Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models, built on a consulting company's internal project data.
+This blended model can predict whether a project will be profitable or not and in a reasonable business decision framework, can guide managers in rejecting financially risky projects and improving final profitability.
 

### 2016-08-30 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd

@@
 ---
-title: Prediction of Consulting Project Profitability
+title: Predictive Models to Support Quoting of Fixed Fee Consulting Projects
 author:
@@
 
-bibliography: mybibfile.bib
+bibliography: Library.bib
 output: rticles::elsevier_article
@@
 _Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_
+
+```{r, echo=FALSE, include=FALSE}
+
+library('knitr')
+library('ggplot2')
+library("plyr")
+library("dplyr")
+library('magrittr')
+library('reshape2')
+library("rpart")
+library('car')
+library('e1071')
+library('party')
+library('randomForest')
+library('RColorBrewer')
+library('pwr')
+library('scales')
+library('purrr')
+
+# detect OS
+if(.Platform$OS.type == 'windows'){
+        # setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+        opts_knit$set(root.dir= "C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
+} else{
+        # setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+        # setwd("~/Documents/github/consult/finalwriteup/report_data")
+        opts_knit$set(root.dir= '~/Documents/github/consult/finalwriteup/report_data')
+}
+
+opts_chunk$set(fig.width=6.7, dpi = 200, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)
+```
 
@@
 background to frame the research question. Concisely summarize the major findings/results.
-Summary of Key Related Research This section should include a brief summary of key related research.
-Emphasis should be on demonstrating the foundation for the current investigation. Specifically, the
-goal is to clearly delineate a gap or missing link that the current research fills. Authors should avoid
-presenting a litany of past research and should focus on prior work necessary to demonstrate the
-existence of the research gap addressed in the manuscript.
 
@@
 
+for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.  
+This thesis focuses on the risk taken by a consulting company when offering a fixed fee.
+Clients commonly seek fixed price quotes from several consultants before selecting one or negotiating further.
+Hence, quoted fees must be competitive.
+The way a consulting manager calculates a fixed price varies from industry to industry and even company to company.
+Typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb
 - amount of losses, sources
+A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget [@Flyvbjerg2011].
+A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by an average of 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007].
+Smaller consulting companies in competitive industries, such as the construction industry, experience similar difficulties in forecasting their project costs.
+losses = low employee morale, elevated stress, low work quality to improve speed, possibly redundancies
+@Lovallo2003 delved into the psychology behind why executives so often severely underestimate costs of larger projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
+Their theory stemmed from Kahneman's work on decision-making that won him the Nobel Prize for economics in 2002.
+His research argued that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
+A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplays or ignores the risk of problems that are out of the team's control.
+all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
+Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].
 
-- 'outside' view, that kind of thing
+Unfortunately, limited research is available that documents the methods industry uses for complex-projects.
+The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors - detailed analysis and experience based models
+A survey by @Moores1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools
+
+Detailed analysis remains the most prevalent method for construction cost estimation despite the industry having a long history of projects running over time and budget using the same technique [@Shane2009].
 
@@
 
+This thesis focuses on a single case study consulting company in the construction industry.
+twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
+labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm. 20% of projects are loss making.
+Fees calculated via detailed analysis
+A Customer Relationship Management software package (CRM) is currently employed to collect and store project data.
+A CRM is a popular type of software used by businesses to record client and project details as well as employee time sheet records.
+In the case study business, the CRM is available to all employees over the company intranet, and each employee completes daily time sheets allotting their hours to certain projects.
+Additionally, technical information is recorded against each project
+
+* Employee time sheet hours with dates
+* Other project costs (taxis, printing)
+* Client information/characteristics
+* Client identification code
+* Invoiced amounts for each project and dates
+* Employee costs
+* Employee charge out rates
+* Project description
+
+2364 past projects are available.
+missing data due to how the data collection changed over time and some fields were not mandatory.
+
+- aim
+**General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
+Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
+The project outcomes are intended to assist the business in predicting project profitability before engagement. Several statistical and machine learning techniques were tested, compared and refined.
+
+**Hypothesis 1**
+A statistical or machine learning model based on historical project data can predict the profitability of a new project with greater accuracy than a baseline predictor.
+A baseline predictor for this project is one that predicts the average of a numeric response variable for all cases or, if the response variable is categorical, randomly assigned categories for each case, where the proportion of assigned categories matches the true categorical proportions.
+
+**Hypothesis 2**
+The predictive model built from Hypothesis 1 can be shown to have a positive impact on the overall profit earned by the case study business.
+The overall profit is represented by the following equation:
+
+- brief summary of major findings/results
+
 2. Literature Review
 ====================
+
+Summary of Key Related Research 
+This section should include a brief summary of key related research. Emphasis should be on demonstrating the foundation for the current investigation. Specifically, the goal is to 
+clearly delineate a gap or missing link that the current 
+research fills. Authors should avoid presenting a litany of past research and should focus
+on prior work necessary to demonstrate the
+existence of the research gap addressed in the manuscript.
 
@@
 
+The bulk of research to date has been performed with project data in either the construction industry or software development.
+Research in the construction industry has primarily focused on predicting the final building cost of construction.
+Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project.
+This was done by statistically comparing a new project to a collection of similar projects and their characteristics.
+It has been shown that these models improve cost estimation accuracy, however industry has not adopted them.
+
+##### 2.1.1 Construction Industry
+@Elfaki2014 completed a review of cost estimation research from 2004 to 2014. They found that artificial Neural Networks and SVM's were the most common machine learning techniques.
+Neural Networks and SVM's received significant attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015; @Kim2004].
+@Shin2015 pioneered the application of Boosted Trees to cost estimation in construction projects.
+Often, Linear Regression was the only model assessed, without comparison to other methods such as Neural Networks, which first started appearing in literature in the 1990's [@Kim2004]
+Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004; @Attalla2003].
+
+Variety of variables used: [@Chan2005][@Akintoye2000; @Trost2003; @Pinto1988]
+the studies all generally reported their mean absolute error or a similar metric from the tested models, and results were generally positive. 
+Furthermore, the existing studies collected data from many businesses for a single study
+
+##### 2.1.2 Software Industry
+
+In the software industry, the main component of cost is *effort* as opposed to the cost of building materials in the construction industry.
+
+Similar to the construction industry, expert judgment or detailed analysis, is the most widely practiced method for effort estimation [@Shepperd1996; @Molokken2003]
+
+Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although Regression is the most popular method [@Finnie1997; @Pai2013; @Matson1993].
+
+ Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the model's accuracy: size (or funciton point) [@Shepperd1996; @Finnie1997; @Pai2013]
+
+ A study by @Mendes2004 demonstrated using 67 web projects that cross-company models were significantly less accurate than a within company model [@Shepperd1996]
+ Again, Neural Networks were criticised for their inability to explain their results [@Finnie1997].
+ The software industry is unique in that agile methods of delivery are changing the contractual approach of consulting [@Badenfelt2011].
+Software development is a relatively new field in comparison to construction, where the tradition of fixed fee structures is difficult to reinvent.
+
 #### 2.2 Methods used in other business applications
 
-3. Methods
-==========
+The review of methods applied to the construction and IT industry highlighted the use of Linear Regression, Neural Networks, SVM's and in one case Boosted Trees, however research on other business problems utilised a wider range of methods.
+These included Naive Bayes, Random Forests, and machine learned Bayesian Networks.
+
+Although SVM's performed the best, as discussed, they are often complex and slow, requiring a great deal of memory [@Kumar2007].
+@Kumar2007 performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
+
+For example in financial credit scoring, a study by @Brown2012 found that Random Forests and Boosted Trees consistently outperformed Neural Networks in classification.
+Kumar and Ravi's [-@Kumar2007] review also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
+
+An exception is Saradhi & Palshikar's [-@Saradhi2011] study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
+Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
+determining the value of each employee in terms of the importance of the projects they were on and their monthly chargeability.
+This allowed them to rank employees identified as 'high risk of churn' by value and provided a clear ranking for manager's to act upon.
+This extension of the study provided a comprehensive framework for how business managers could adopt their findings to improve business operations.
+It was a valuable addition that is absent from most cost and effort estimation research.
+
+#### 2.3 Gaps
+
+In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the data set is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
+In past research, cost estimation data sets generally consisted of <300 of projects from a range of a companies and even countries.
+This thesis case study only had access to the company's internal data but this contained over 2,000 past jobs
+Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry, which differs from other studies in the construction industry that estimate actual construction costs.
+Consulting in the construction industry can be likened more to software development effort estimation although the nature of the work differs considerably.
+
+ensemble tree methods and blended methods are applied which have been minimally tested on this problem.
+To date, Linear Regression, Neural Networks, Case Based Reasoning (CBR), and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks and generally outperform Linear Regression [@Caruana2006].
+Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots that provide insight into the model's calculations.
+This contrasts Neural Networks and SVM's, which are 'black-box' predictors.
+
+Finally, this study propels the predictive model one step further than other studies by analysing bottom-line profit improvements that could result from the research.
+final profit increases are evaluated. This kind of thought experiment is designed to present a clear case for industry when evaluating if and how to adopt cost estimation models.
+
+3. Prediction Methods
+=====================
 
@@
 
+Neural Netowrks and SVM were not purusued because black box
+A previously mentioned disadvantage is that the NN algorithm is a black box because the internal structure is too complex for interpretation.
+They also require a lot of training data relative to other methods.
+
+Also, SVM's can be very slow to train and therefore not suitable for industry purposes [@Auria2008].
+Because it is so important to engage decision makers with a model that can explain its results, these two methods are not appropriate for the effort estimation problem.
+
 #### 3.1 Predictive methods
 
-- description of each of the methods: refer thesis
+binary classification problems, where the objective is to predict the probability of an event occurring.
+Probability is a continuous response variable, however predictions from a linear regression would not be bound by 0 and 1.
+
+##### 3.1.1 Logistic Regression
+
+Because of the linear relationship, an equation can be fit to the log odds of the binary response variable against each of the explanatory variables using linear regression.
+Then, the linear equation describing the log odds can be transformed back to probability by taking the inverse log i.e.
+the exponential and rearranging [@Macdonald1975]:
+
+The result in a sigmoidal function bound by 0 and 1.
+The sigmoidal function originates from a linear fit of the log odds in the data.
+Assume linear relationship between covariates and response variable
+
+ A coefficient represents the change in the log odds of the response variable for each unit increase in an explanatory variable.
+Therefore, taking the exponential of the coefficient is the change in odds of the response variable.
+If the change in odds was 2, then if the explanatory variable increased by 1, the response variable event would be twice as likely to occur.
+
+ Logistic Regression is a good  benchmark to compare other binary predictive models due to its simplicity and speed [@Moore1989]. single linear relationship means Not prone to overfitting
+
+##### 3.1.2 Random Forests
+
+single decision trees also suffer from low predictive accuracy and instability, so to combat these problems, ensemble tree methods were pioneered in the 1990's [@Breiman1996].
+
+The three methods of combining hundreds or thousands of trees turn decision trees into high performing, stable predictors.
+
+training multiple trees from the same data by sampling bootstrapped training sets with replacement.
+However, when creating each tree, a random subset of attributes (variables) is considered at each split.
+The reduced subset of attributes is resampled for each split in the tree.
+This allows dominant variables to be suppressed for a fraction of the splits, allowing the algorithm to explore signals in weaker variables [@Breiman2001a].
+
+Additionally, ensemble tree methods are faster than SVM's and Neural Networks but can perform just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
+
+##### 3.1.3 Gradient Boosted Trees
+
+Lastly, the boosted decision tree approach applies gradient descent theory to a series of decision trees.
+The trees are limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) of the preceding tree.
+By modeling the errors, misclassified cases are weighted higher than correctly classified cases and influence the structure of the latest tree more [@Hastie2009].
+This increased weighting on errors is why the algorithm is called boosting.
+The limited depth of each tree prevents overfitting at each stage and the combined result of up to thousands of trees is very powerful [@Elith2008].
+
+@Caruana2006 tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
+This demonstrates how ensemble tree methods are capable of competing with high-level machine learning algorithms.
+
+Trees are not built on a probabilistic framework, and therefore their results cannot be provided in this framework [@Louppe2014].
+
+##### 3.1.4 Naive Bayes
+
+The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for the response variable (the response variable must be categorical).
+
+These equations are simple to compute given the data, and the class with the highest probability can then be chosen [@Provost2013].
+
+The advantages of this method are that the conditional independence assumption enables very fast calculations and predictions.
+The method can perform well in real world tasks because the assumption of independence does not significantly damage predictions [@Provost2013].
+Assumptions of independence genearlly not true (correlated variables).
+This is fine for ranking but the output probabilities are not accurate statistical probabilities [@Caruana2006].
+
+Naive Bayes method generally provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
+
+##### 3.1.5 Bayesian Networks
+
+A Bayesian Network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
+The model is visually represented by a directed acyclic graph (DAG) and is capable of linking the conditional dependency between any variable to another variable [@Heckerman1998].
+The conditional relationships are Bayesian, where the probabilities in one node are conditional upon values in nodes directed towards it as well as preceding nodes
+
+Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes [@Barber2012]
+
+Naive Bayes is the simplest form of a Bayesian Network [@Zhang2004]
+
+Network relationships can be learned from the data, however this is not widely included as part of the suite of machine learning methods.
+The conditional variable dependencies are calculated from the data, which in turn can define the graph structure.
+Some conditional dependencies can be set before the structure is learned [@Barber2012].
+
+Drawbacks of Bayesian Networks include the difficulty for them to process continuous variables.
+
+Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.[@Kragt2009]
 
@@
 
-- Regression, classification, blending, bottom line analysis
+ It includes first how the data was obtained, the lengthy cleaning process, 
+ Projects varied from total invoiced amounts of \$500 to over \$1,000,000
+ a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
+ 10 engineered variables describing timesheets: Percent of hours performed by 'professional' employees as opposed to 'technical' employees, Position of the employee that completed the most hours on each project
+ 3 engineered from invoicing data: 
+ 4 from project data: number of projects completed with each client
+ text analysis of project titles to develop project classifications
 
-4. Results and Discussion
-=========================
+ Discretisation was performed by generating a hierarchical dendrogram of each variable to visualise the clusters.
+
+Limiting the predictive model to a concise set of meaningful variables reduces noise and improves predictions.
+Less variables and a simpler model is easier for stakeholders to understand [@Weisberg2005].
+followed by variable importance analysis, variable selection: varialbe importance aalyses done with linear regression, random forest, cforest to represent the range of algroithms that were trialled.
+11 variables chosen:
+
+ trials of selected predictive algorithms.
+binary classification problem predicting profit or loss.
+Complex models should be measured against simple models that can be built at a fraction of the computational cost.
+
+ * Regression - baseline model
+* Naive Bayes - baseline model
+* Bayesian Network - grabage.
+* Random Forest
+* Gradient Boosted Trees
+
+To compare the models, the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used.
+An ROC graph visualises the curve from which an AUC score is calculated.
+In this problem a 'positive' is a loss making job.
+For binary classification, AUC is a more meaningful statistic than classification accuracy when the output is a probability that can be applied to the problem.
+The AUC indicates model performance across many thresholds while classification accuracy represents a single threshold.
+
+a model must be first created using a training set. The model can then make predictions on the test set. this is used to build the ROC
+A model that is perfectly classified would have an AUC = 1.
+A model that predicts as well as random chance would have an AUC = 0.5. An AUC between 0.5 and 1 means the model is performing better than random chance.
+
+In order to compare which models performed significantly better than others, an adequate sample size of results statistics is required.
+Multiple models could be made by using different data in the training vs.
+testing sets, each providing a resulting test statistic (RMSE or AUC).
+Initially, 20 models of each method were created in this fashion.
+Then a two-sample power calculation was run using the two sets of 20 results to determine the sample size to achieve a statistical power of 0.8 (100 to achieve power of 0.8 between boosted trees and naive bayes)
+
+Missing Data Imputation.
+All methods except gradient Boosted Trees and Naive Bayes could not handle missing data.
+a complete data set allows for complete sets of predictions from each method, and these predictions could then be blended to further improve results.
+The MICE Random Forest method was chosen for imputation because it has been proven to work well with complex data sets [@Shah2014].
+If similar predictive results were obtained using Boosted Trees imputed data and imputed data, the imputed data must be reasonable.
+The imputed data set was then trialed on the remaining methods and compared to unimputed trials.
+
+ Once the best methods were selected, they were blended in numerous ways, using both simple averaging techniques and sophisticated machine-learning algorithms.
+These were compared against individual models and the best constructs were selected.
+Six blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
+These included simple averaging of the individual model results, building a Logistic Regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
+A simplified explanation of FWLS is as a Linear Regression where meta-features as well as model predictions from individual models are included as explanatory variables.
+Then, each meta-feature is interacted with each set of model results [@Sill2009].
+ Feature interaction is performed passively due to the nature of how trees are built.
+A split in a node determined by one variable is conditional upon the preceding split, which was based on another variable and so on.
+ The simplest method averaged the probability output (a number between 0 and 1) of all three models.
+The next two simplest methods consisted of building a Logistic Regression and Boosted Trees model from the three prediction model outputs only.
+Feature Weighted Linear Stacking (FWLS), Random Forest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (called meta-features in this context).
+The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature states could be weighted as such.
+
+Again, 100 models were built to achieve a power of 0.8 for the variation in AUC test statistics across different divisions of training/testing data.
+A maximum of five blended models were created from a complete set of test results (built from the individual models), then a new set of test results were created for the next five blended models.
+
+ Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision-making scenarios.
+This analysis differed from accuracy in predicting profitability, so several blended models as well as individual models were carried forward for this analysis.
+This was done using a profit curve - a chart that plots the change in profit the company earns on the y-axis vs.
+the probability threshold on the x-axis.
+A simple approach was taken for this analysis, where projects with a probability to be a loss making job greater than the threshold were rejected entirely.
+Therefore all profits and losses from jobs above the threshold were discounted.
+An equation defining the change in profit as a percentage of the original profit using threshold rejection is shown below:
+
+\begin{equation}
+\Delta\ Overall\ Profit\ \textrm{(\%)} = \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100
+\end{equation}
+
+<!-- \Delta\ Overall\ Profit\ \textrm{(%)} = \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100 -->
+
+Where
+
+$$
+\begin{aligned}
+I(\cdot) &= \textrm{the indicator function} \\
+N\ &= \textrm{the number of individual projects that are being included in the analysis} \\
+Pr_{p} &= \textrm{probability output from the algorithm for project}\ p \textrm{. Values are}\\
+& \qquad \textrm{betweeen 0 and 1 where 1 is loss making)} \\
+threshold &= \textrm{a chosen value between 0 and 1.}\ \Delta\ Overall\ Profit\ \textrm{is calculated for} \\
+& \qquad \textrm{several}\ threshold\ \textrm{values which defines the profit curve} \\
+Profit_p &= \textrm{profit for individual project}\ p
+\end{aligned}
+$$
+
+If the threshold was zero, all jobs were rejected and the profit would be $0.
+If the threshold was 1.0, all jobs were accepted and the profit would be the same as the profit the company actually experienced since the data is a sample of historic projects.
+The aim was to find the optimal threshold point where saying 'no' to a job above that level would result in higher profits, because jobs that were likely to make a loss were being rejected.
+This chart will clarify what percentage of profit increase the company could expect by integrating the algorithm into decision-making.
+
+the curve varies with different divisions of the data.
+Therefore, in order to understand the uncertainty around the profit curve and to determine which blended or individual model performed statistically better than others, a large sample size of curves was required.
+100 to achieve a power of 0.8.
+which was achieved by repeating 5-fold training/testing splits.
+A 95% confidence interval could also be determined around the highest point on each curve.
+The final expected increase in profit and the percentage of projects to be rejected presented a clear scenario that the case study business managers could assess in terms of their business strategy.
+
+4. Prediction Results and Discussion
+====================================
 
@@
 
-#### 4.1 Predictive Analysis
-- regression was attempted - failed
+#### 4.1 Individual Models
 
-- binary classification - results of 5 methods
+The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable/unprofitable projects.
+The five prediction methods, were initially built without imputing data, by using the maximum amount of data possible depending on the method.
+Boosted Trees and Naive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
+100 models of each method were built.
+The violin plot below summarises the AUC values produced by each method and the 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
 
-- blended models - improved results
+```{r reduced_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Subsets of the data were used for Logistic Regression and Random Forests in order to provide datasets without missing values.")}
 
-#### 4.2 Bottom line analysis
+reduc.totAUC <- readRDS("reduc_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(reduc.totAUC[[2]], reduc.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+reduc.totAUC = melt(reduc.totAUC)
+reduc.totAUC$sigdiff = ifelse(reduc.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff), alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "gray20", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from Boosted Trees") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from Boosted Trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 models of Each Method") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
+
+```
+
+The AUC performance of Logistic Regression and the Random Forest algorithms cannot be statistically differentiated from Boosted Trees.
+
+Next, data imputation methods were trialed which would make the complete data set available to Logistic Regression and Random Forests.
+Again, 100 models were required to achieve a power of 0.8 with respect to Boosted Trees
+
+```{r mice_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
+Each method was fed the same imputed full dataset.")}
+
+mice.totAUC <- readRDS("mice_df_auc.rds")
+
+#significant difference??
+index = c(1,3,4)
+pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
+for(i in seq_along(index)){
+        pvals$p_value[i] = t.test(mice.totAUC[[2]], mice.totAUC[[index[i]]])$p.value %>%
+                format(digits = 2) %>% as.numeric
+}
+
+signif = pvals %>% filter(p_value < 0.05) %>% .$method
+mice.totAUC = melt(mice.totAUC)
+mice.totAUC$sigdiff = ifelse(mice.totAUC$L1 %in% signif, TRUE, FALSE)
+
+ggplot(mice.totAUC, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = sigdiff, colour = sigdiff), 
+                    alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "gray20", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        scale_fill_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from boosted trees") +
+        scale_colour_manual(values = c('sienna1','navajowhite'),
+                          name = "Significantly different from boosted trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 Models of Each Method - Imputed Data") +
+        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
+                         "log" = "Logistic Regression",
+                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
+        theme(legend.position = "bottom")
+```
+
+Boosted Trees, Logistic Regression, and Random Forest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed Boosted Trees.
+
+Testing Boosted Trees, Random Forests, Naive Bayes, Logistic Regression and Bayesian Networks on the binary classification problem showed that Boosted Trees, Logistic Regression and Random Forests performed best (according to AUC).
+Naive Bayes and Logistic Regression were included as baseline models and it was expected that the more complex models would outperform these.
+Therefore it was surprising that results from 100 Logistic Regression models were not statistically significantly lower than Boosted Trees.
+
+A possible explanation for this, according to the literature, is that there was not enough data for the ensemble trees to learn the complex decision rules at which they excel.
+Trees tend to overfit the patterns in a smaller training set.
+Logistic Regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
+This may explain Logistic Regression's comparatively high performance on the case study's small but complex data set.
+
+Boosted trees, Random Forests and Logistic Regression all had mean AUC values between 0.755 to 0.764.
+In summary, the individual binary classification models performed well above random chance (AUC = 0.5).
+Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model can affect decisions in practice.
+
+- NB and BN were bad - one sentence
+
+#### 4.2 Blended Models
+
+Blended models combine the predictions from each high-performing individual model to create averaged or 'blended' predictions.
+In this case, the Random Forest, Boosted Trees, and Logistic Regression models were chosen.
+Through variable importance studies, 6 variables were included as metafeatures.
+All six methods were compared against the original Logistic Regression model's AUC distribution (as it was not statistically different from the ensemble trees)
+
+```{r blend_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
+100 models were built for each method.")}
+
+auc.list <- readRDS("blend_auc2.rds")
+means = lapply(auc.list, mean) %>% unlist
+greater = means[means> means["orig.log"]] %>% names
+auc.melt = melt(auc.list)
+auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)
+
+ggplot(auc.melt, aes(x= factor(L1), y= value)) + 
+        geom_violin(aes(fill = greater, colour = greater), 
+                    alpha = 0.8) + 
+        stat_summary(fun.y="mean", geom="point", colour = "gray30", 
+                     shape =4) +
+        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
+                     geom="text", colour = "gray20", size =3, 
+                     hjust = -.2, angle = 40, vjust = -.2
+                     ) +
+        theme(axis.text.x=element_text(angle=40,hjust=1)) +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
+             title = "AUC Distribution from 100 Blended Models of each Method") +
+        scale_x_discrete(labels = c("boost" = "BT Blend", 
+                         "log" = "FWLS", 
+                         "average" = "Simple Average Blend",
+                         "orig.log" = "Orignal LR", 
+                         "rf" = "RF Blend",
+                         "simp.boost" = "Simple BT Blend",
+                         "simp.log" = "Simple LR Blend")) +
+        scale_fill_manual(values = c('navajowhite','sienna1'),
+                          name = "AUC Greater than Original Logistic Regression") +
+        scale_colour_manual(values = c('navajowhite','sienna1'),
+                          name = "AUC Greater than Original Logistic Regression") +
+        theme(legend.position = "bottom")
+
+```
+
+The above plot shows that 4 methods had a higher mean AUC than the original Logistic Regression model:
+
+* the simple blended Logistic Regression 
+* the simple average 
+* and the two Boosted Tree models.
+
+The simple averaged method and simple logistic regression could achieve a statistical power of 0.8 with 150 samples (models), whereas the boosted trees required thousands.
+The additional 50 models were run for the simple logistic regression and simple average to reveal similar distributions that were indeed statistically significantly higher than the individual logistic regression AUC distribution.
+P-values from two sample t-test were 0.0019 and 0.0021 respectively.
+
+ Blended models improved the mean AUC from 0.759 to 0.77, which is an increase of `r ((0.77-0.759)/0.759) %>% "*"(100) %>% round(1)`%.
+It was expected that model blending would improve predictions because it combined the strengths of individual models with different theoretical foundations.
+
+ The trials of different blending methods demonstrated that again, the simplest methods worked best with the case study data.
+In this incident, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
+Simple blended models could perform better than complex methods if the data is not big enough for the complex models to learn the more intricate patterns at which they excel.
+This was observed in the single model methods.
+their success in the 2009 Netflix competition generated some publications.
+The Netflix data set comprised of almost 3,000,000 observations, so the size of the data could have enabled the success of complex blending methods such as FWLS [@Sill2009]
+
+5. Business Impact
+==================
+
+This chapter first presents the full range of method results in terms of improvements to the case study's bottom line.
+A business decision making scenario was created so that profit curves based on the decision rule could be built and analysed.
+From the profit curves, optimal probability thresholds could be derived for each method (individual and blended methods).
+The predictive models output a 'probability' between 0 and 1 that each project will be a loss making job (where probability = 1 indicates a loss making job).
+The question the arises, at what probability would a decision-maker round the probability up to 1 or down to 0?
+And what business decision would then be made?
+To find the threshold point for rounding, an experimental business-scenario was tested.
+At a given threshold, all projects with probability outputs above the threshold were considered too risky, and were rejected.
+All profits and losses from these projects were forfeited, while the profits from the remaining jobs (below the threshold) were summed to give a revised total profit.
+This profit calculation was made on a range of thresholds between 0 and 1 at 0.05 increments.
+The total profits were plotted for each threshold and joined to make a profit curve
+
+The plot below illustrates the distribution of profit curves for each method, where the solid lines join the mean values at each threshold point.
+The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold point for the 100 models.
+
+```{r profit_curve, fig.height = 6, fig.cap = c("Profit curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+
+finance.plot = readRDS("plot_finance_100_4.rds" )
+finance.plot$method <- factor(finance.plot$method, 
+                                    levels=c("average", "simp.log","simp.boost",
+                                             "comp.boost","comp.rf", "FWLS","orig.log",
+                                             "orig.rf","orig.boost"))
+
+ggplot(finance.plot, aes(x= threshold, y= profit.ratio, colour = method)) + 
+        # geom_point(alpha = 0.005, shape = 16, size = 1) + 
+        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
+        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) +
+        stat_summary(fun.y = mean, 
+                     fun.ymin = function(x) mean(x) - 1.96*sd(x),
+                     fun.ymax = function(x) mean(x) + 1.96*sd(x),
+                     fill = "gray", geom= "ribbon", alpha = 0.7, colour = NA) +
+        stat_summary(fun.y = mean, geom="line") +
+        scale_y_continuous(breaks = seq(0,110, by = 20)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .2)) +  
+        theme(legend.position = "none",
+              panel.background = element_rect(fill = 'gray95')) +
+        labs(x = "Probability Threshold", 
+             y = "Ratio of Model Profit to Historical Profit (%)",
+             title = "Profit Curve for Blended and Individual Models") + 
+        facet_wrap(~method, 
+                   labeller = labeller( method = c("average" = "Simple Average Blend",
+                                       "orig.boost"= "Orig. Boosted Trees",
+                                       "orig.log" = "Orig. Logistic Regression",
+                                       "orig.rf" = "Orig. Random Forest",
+                                       "simp.log" = "Simple Log. Reg. Blend",
+                                       "simp.boost" = "Simple Boosted Tree Blend",
+                                       "comp.boost"= "Complex Boosted Tree Blend",
+                                       "comp.rf"= "Complex Random Forest Blend",
+                                       "FWLS" = "FWLS"
+                                       )))
+
+```
+
+Two blended models clearly outperformed the individual models as shown in the above plot with higher profit ratios as well as tighter confidence intervals.
+The simple Logistic Regression blend performed best with the highest mean profit ratio of 109% with a standard deviation of 1.28%.
+%.
+This means that for the simple blended Logistic Regression model, if all jobs above the probability threshold 0.6 were rejected, the profits would increase on average by 9% in comparison to historical profits.
+
+Profit curves from the simple average blend and Logistic Regression blended models outperformed the complex blended models, which follows logically from their significantly higher AUC distributions.
+The simple average blended model produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression blend.
+In the data, projects assigned a probability higher than 0.6 represent only 4.3% of all projects.
+
+The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' or complex blending methods were above 100% on the y-axis.
+That is, it cannot be said with 95% confidence that the original methods would produce a mean profit higher than historical profits in the given decision framework.
+Clearly this level of certainty is not beneficial for the case-study business to adopt.
+
+It is not clear why simple linear models and the averaging model outperformed ensemble tree blending methods.
+As previously stated, the ensemble trees may not have received enough data to adequately learn the complex series of rules they develop.
+The only differences between FWLS and the simple Logistic Regression (that performed best) were two additional explanatory variables and four interaction terms.
+The number of variables was not high for the amount of data since according to @Peduzzi1996, 10 events per explanatory variable or more avoids the risk of biased estimation of variable coefficients in Logistic Regression.
+The data contained 315 events per explanatory variable and did not pose that risk.
+Nevertheless, the additional variables must have added misleading noise to the model.
+
+To conclude, because of the promising AUC results, it was logical that the models translated into financial benefits for the company.
+The 9% improvement in profits produced by the simple blended Logistic Regression is reasonable, and may be high enough to trigger further cost-benefit analyses and the development of a more comprehensive framework describing different decision scenarios.
 
@@
 
-7. Decision Support Tool
-========================
+- prototype decision support interface
 
-- how to generate actions from predictive results
-
-- alternative decision scenarios
-
-- user trust
-
-9. Conclusions and Future Work
+6. Conclusions and Future Work
 ==============================
@@
 3/4 page
+
+The general aim was to use statistical techniques to predict the profitability of projects for a case study consulting business using their internal CRM data.
+This was rigorously completed by approaching the prediction of 'profitability' as a binary classification problem predicting either profit or loss.
+Several statistical and machine learning approaches were applied including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models' output.
+
+AUC = 0.76). This was achieved by 3 individual methods while various techniques that blended the individual methods improved results further (AUC = 0.77).
+
+predictive models could be shown to improve the overall profitability (bottom line) of the case study business.
+A range of probability threshold values were trialed for each method using a decision framework where projects scored below the threshold were accepted, while projects above the threshold were rejected.
+If a project was rejected, the profits and losses were forfeited, and the remaining accepted profits and losses were summed.
+
+Final results showed the simple Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
+The 95% confidence interval for these improvements was between 6.5% and 11.5% using a probability threshold of 0.6 (approximately 4.3% of projects).
+
+These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
+Ensemble tree methods and blending had been applied minimally to cost estimation previously, even though ensemble trees provide insight into model structure while predicting at a similar level to Neural Networks.
+Next, previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
+This study presented a clear framework for how a business could improve profits by applying the algorithm.
+
+Further work is required to test user confidence in the output.
+Another topic identified for future research was to test how well managers estimate some numeric project input variables.
+In particular, time span, team size, total invoiced amount, and percentage of hours completed by professionals should be tested as these variables were calculated post project completion.
+Time span and total invoiced amount were discretised into wide categories, which should be easier for a manager to choose between.
+
+future work: user interface.
+
+Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, from a consulting company's internal project data.
+This blended model can predict whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business.
 
@@
 -   keywords and MSC codes
-
-Here are two sample references: @Feynman1963118 [@Dirac1953888].

### 2016-08-30 finalwriteup/draft.rmd

@@
 
-[@Macdonald1975]
-
 Because of the linear relationship, an equation can be fit to the log odds of the binary response variable against each of the explanatory variables using linear regression.
 Then, the linear equation describing the log odds can be transformed back to probability by taking the inverse log, i.e.
-the exponential and rearranging:
+the exponential and rearranging [@Macdonald1975]:
 
@@
 Initially 100 train/test runs were performed for each blending method by running 20 iterations of 5-fold cross validation.
-The AUC results of these iterations are displayed below:
+Summaries of a single example model from each of the blending methods can be found in the Appendix, section 11.3.
+The AUC results of the 100 iterations are displayed below:
 
@@
 From the profit curves, optimal probability thresholds could be derived for each method (individual and blended methods).
-These threshold points allowed the probability output from each model to be converted to categorical output (profit or loss), which in turn could generate confusion matrix statistics for each method.
+These threshold points allowed the probability output from each model to be converted to categorical output (profit or loss) which in turn could generate confusion matrix statistics for each method.
 The confusion matrix statistics are discussed after the profit curve results and discussion.
-Then, output from the method which produced the best profit curve was used to create two subsets of profitable and loss mkaing projects.
+Then, output from the method which produced the best profit curve was used to create two subsets of profitable and loss making projects.
 The characteristics of these subsets are compared and contrasted to gauge any overall trends.
@@
 
-The plot below illustrates the distrubtion of profit curves for each method, where the solid lines join the mean values at each threshold point.
+The plot below illustrates the distribution of profit curves for each method, where the solid lines join the mean values at each threshold point.
 The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold point for the 100 models.
 
-```{r profit_curve, fig.height = 6, fig.cap = c("Profit Curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+```{r profit_curve, fig.height = 6, fig.cap = c("Profit curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
 
@@
 
-Secondly, the nearest neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
+Secondly, the Nearest Neighbour algorithm is automatically run with the predictive model and its results are displayed in the second main tab.
 The aim was to provide the user with a cross check of the algorithm's results by finding projects with the most similar features to the one currently being assessed.
-This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'nearest neighbours'.
+This triggers memories from the user's past experiences on similar projects and promotes communication with other managers who delivered the 'Nearest Neighbours'.
 Some screen shots of the nearest neigbhour graphic are presented below.
@@
 
-The general aim was to use statistical techniques to model the profitability of projects for a case study consulting business using their internal CRM data.
+The general aim was to use statistical techniques to predict the profitability of projects for a case study consulting business using their internal CRM data.
 This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitability and as a binary classification problem predicting either profit or loss.
-Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models.
+Several statistical and machine learning approaches were applied to both problems including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models' output.
 
@@
 
-Lastly, hypothesis 2 proposed that the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
-This theory was tested on a range of predictive models because AUC scores were not directly linked to profitability once the models were applied to a business decision scenario.
-The models output a probability, so an optimal probability threshold needed to be found that would dictate whether a job was classified 'profitable' or 'loss making'.
-A range of threshold values were trialed, and a decision framework was created where if a project was scored below the threshold it was accepted while projects above the threshold were rejected.
-If a project was accepted, the profits and losses were summed while if a project was rejected the profits and losses were forfeited.
-A perfect model would reject all loss making jobs and accept all profitable jobs, thereby improving overall profits.
+Lastly, hypothesis 2 proposed that predictive models developed from hypothesis 1 could be shown to improve the overall profitability (bottom line) of the case study business.
+A range of probability threshold values were trialed for each method using a decision framework where projects scored below the threshold were accepted, while projects above the threshold were rejected.
+If a project was rejected, the profits and losses were forfeited, and the remaining accepted profits and losses were summed.
 Final results showed the simple Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
@@
 These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
-Ensemble tree methods and blending had been applied minimally to cost estimation to date.
-Past research tested regression models and to some extent Neural Networks and CBR.
-The advantage of ensemble tree methods is that they provide insight into the model's structure while predicting at a similar level to Neural Networks.
-Next, a clear framework for how a business could tangibly improve profits was developed and proven successful.
-Previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
-Lastly, the decision framework was progressed even further into a prototype application where user trust was addressed.
-Previous studies had shown that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
-This was addressed by providing intuitive plots of the model structure and a parallel nearest neighbour graph of similar internal projects to provide a clear and intuitive 'outside view'.
+Ensemble tree methods and blending had been applied minimally to cost estimation previously, even though ensemble trees provide insight into model structure while predicting at a similar level to Neural Networks.
+Next, previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
+This study presented a clear framework for how a business could improve profits by applying the algorithm.
+Lastly, a prototype application was developed where insight into the model structure encouraged user trust  and a Nearest Neighbour chart presented an ‘outside view’ based on the company’s projects.
+Past surveys have reported that despite decades of promising research in cost estimation, decision maker's still mainly use an 'inside view' during fee estimation.
 Further work is required to test user confidence in the output.
 
-Another topic identified for future research was to test how well managers estimate project input variables.
-Estimation of the time span, team size, total invoiced amount, and percentage of hours completed by professionals in particular should be tested.
-Initial estimates of input variables could also be kept in the database for future analysis.
-
-Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forest, and Boosted Tree models, built on a consulting company's internal project data.
-This blended model can predict whether a project will be profitable or not and in a reasonable business decision framework, can guide managers in rejecting financially risky projects and improving final profitability.
+Another topic identified for future research was to test how well managers estimate some numeric project input variables.
+In particular, time span, team size, total invoiced amount, and percentage of hours completed by professionals should be tested as these variables were calculated post project completion.
+Time span and total invoiced amount were discretised into wide categories, which should be easier for a manager to choose between.
+
+Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, from a consulting company's internal project data.
+This blended model can predict whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business.
+
+# Appendix
+
+Example results from a single model from each method. This is broken into the three main types of methods that were trialled:
+
+1. Regression models
+2. Classification models
+3. Blended classification models
+
+## Regression models
+
+ANOVA Linear Regression and Random Forests were applied to profitability as a regression problem.
+
+### ANOVA regression
+
+Example summary of model built with core variables only:
+```{r}
+
+all8a = read.csv("all8a.csv")[,-1]
+
+#function below outputs summary of model
+final.output<- function(df=all8a, no.tests = 50, core = TRUE) {
+
+        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)
+
+        for(i in 1:no.tests){
+
+        if(core == FALSE){
+        samp = sample.int(nrow(df), 1)                        
+        #create sample df - use all variables that are complete from a sample row of data
+        sample = df[samp,]
+        sample = sample[,!is.na(sample)]
+        vars= names(sample)
+        vars = vars[! vars %in% c('mlsto','Year','return.pdol',
+                                  'code.client', 'code.contact')]
+        variables = vars
+        } else{
+                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
+                              'no.users', 'pc.pro', 'Business')
+        }
+
+        #make reduced dataset with complete cases for all variables we want to use
+        reduced = df[complete.cases(df[,variables]),]
+
+        formula = paste("return.pdol ~ ")
+
+        for(j in 1:length(variables)){
+                formula = paste(formula, variables[j], sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'pc.pro' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
+        }
+
+        if('inv.mlsto.log' %in% variables &
+           'code.client' %in% variables){
+                formula = paste(formula, 'inv.mlsto.log*code.client', sep= " + ")
+        }
+
+        if('JD.Second' %in% variables &
+           'no.users' %in% variables){
+                formula = paste(formula, 'JD.Second*no.users', sep= " + ")
+        }
+
+        samp_size = floor(0.75*nrow(reduced))
+        create.train = base::sample.int(nrow(reduced), samp_size)
+        train.df = reduced[create.train,]
+        test.df = reduced[-create.train,]
+
+        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
+        for(k in seq_along(cats)){
+                if(cats[k] %in% variables){
+                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
+        }
+        }
+
+        final = aov(as.formula(formula), data = train.df)
+
+        fit= summary(final)
+
+        }
+
+        return(fit)
+}
+
+final.output(df=all8a, no.tests = 1, core = TRUE)
+
+```
+
+Example summary of model built with all variables. Note that each model used a subset of all variables based on the available variables for a randomly selected project:
+
+```{r}
+
+final.output(df=all8a, no.tests = 1, core = FALSE)
+
+```
+
+### Random Forest
+
+Example summary of model built with core variables only:
+
+```{r}
+
+final.rf.output<- function(df=all8a, no.tests = 50, core = TRUE) {
+
+        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)
+
+        for(i in 1:no.tests){
+
+        if(core == FALSE){
+        samp = sample.int(nrow(df), 1)                        
+        #create sample df - use all variables that are complete from a sample row of data
+        sample = df[samp,]
+        sample = sample[,!is.na(sample)]
+        vars= names(sample)
+        vars = vars[! vars %in% c('mlsto','Year','return.pdol')]
+        variables = vars
+        } else{
+                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
+                              'no.users', 'pc.pro', 'Business')
+        }
+
+        #make reduced dataset with complete cases for all variables we want to use
+        reduced = df[complete.cases(df[,variables]),]
+
+        if('code.client' %in% variables){
+        sample = factor(sample(unique(reduced$code.client),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.client %in% sample,]
+
+        reduced$code.client= factor(reduced$code.client)
+        }
+
+        if('code.contact' %in% variables){
+        sample1 = factor(sample(unique(reduced$code.contact),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.contact %in% sample1,]
+
+        reduced$code.contact= factor(reduced$code.contact)
+        }
+
+        formula = paste("return.pdol ~ ")
+
+        for(j in 1:length(variables)){
+                formula = paste(formula, variables[j], sep= " + ")
+        }
+
+        samp_size = floor(0.75*nrow(reduced))
+        create.train = base::sample.int(nrow(reduced), samp_size)
+        train.df = reduced[create.train,]
+        test.df = reduced[-create.train,]
+
+        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
+        for(k in seq_along(cats)){
+                if(cats[k] %in% variables){
+                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
+        }
+        }
+
+        final = randomForest(as.formula(formula), data= train.df, mtry=5,ntree=500, importance=FALSE)
+
+        }
+        return(final)
+}
+
+rf.core <- final.rf.output(df=all8a, no.tests = 1, core = TRUE)
+
+print(rf.core)
+getTree(rf.core, 1) %>% head(20)
+```
+
+All variables:
+
+```{r}
+rf.all <- final.rf.output(df=all8a, no.tests = 1, core = FALSE)
+
+print(rf.all)
+getTree(rf.all, 1) %>% head(20)
+
+```
+
+## Classification models
+
+Logistic Regression, Random Forests, Boosted Trees, Naïve Bayes, and Bayesian Networks were applied to profitability prediction as a classification problem (will a project be profitable or not?)
+
+### Logistic Regression
+
+Example summary output imputed data:
+
+```{r}
+
+# load imputed data
+all10mice_f = read.csv("all10mice_May16.csv")[,-1]
+
+#function for running models
+test.roc <- function(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
+                     extra.var = NULL, mult = 1, x.folds = 10){
+
+        auc<- rep(list(rep(0, length(dfs))), length(dfs))
+        colours <- topo.colors(x.folds*mult)
+
+        for(i in 1:length(dfs)){
+                df = get(dfs[i])
+
+                # mlsto is spinning errors, get rid of
+                df = df[, !(names(df) %in% 'mlsto')]
+                set.seed(seed)
+
+                # create f.rpdol
+                df$f.rpdol <- as.factor(df$b.rpdol)
+                levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
+                levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"
+
+                #set up 10 fold cv using caret function
+                folds <- createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
+                #now have folds[[1]] through folds[[10]] list of rows to exclude
+
+                #for Appendix just want one fold
+                folds <- folds[[1]]
+
+                if(mult > 1){
+                        for(k in 2:mult){
+                                set.seed(sample(1:1000,1))
+                                folds.temp = createFolds(df$f.rpdol, k = x.folds, list = T, returnTrain = F)
+                                folds = append(folds, folds.temp)
+                        }
+                }
+
+                #if extra.var not found in names(df), change to NULL
+                if(!is.null(extra.var)){
+                        if(!(extra.var %in% names(df))) extra.var = NULL
+                }
+
+                #define formula for algorith methods, either rf, log, boost
+
+                if(nrow(df) == 2364){
+                        if(is.null(extra.var)){
+                                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
+
+                        }}
+
+                if(method %in% 'boost') {
+                        #no JD.Second
+                        formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
+
+                }
+
+                if(method %in% 'naiive') {
+                        formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + JD.Second + pc.majpos.log + Billing.Type + JD.Second"
+
+                }
+
+                        #turn predictors and response into 'ordered' variables for roc() to work
+                        test<- df[folds,]
+                        levels(test$f.rpdol) <- c('profit', 'loss')
+                        test$f.rpdol <- ordered(test$f.rpdol)
+
+                        if(method %in% 'rf'){
+                                fit <- randomForest(as.formula(formula), data = df[-folds,], 
+                                            mtry = mtrys[i], ntree = 1000)
+                                pred <- predict(fit, test, type = "prob")
+                                pred <- pred[,2]
+
+                        }
+
+                        if(method %in% 'log'){
+                                fit<- glm(as.formula(formula), 
+                                          data = df[-folds,], family = binomial())
+                                pred <- predict(fit, test,
+                                                type = "response")
+                        } 
+                        if(method %in% 'boost'){
+                                fit <- gbm(as.formula(formula), data = df[-folds,],
+                                           distribution = "bernoulli", n.trees = 10000,
+                                           shrinkage = 0.001, interaction.depth = 5,
+                                           n.minobsinnode = 20)
+
+                                pred <- predict(fit, test, n.tree = 10000, type = "response")
+                        }
+
+                        if(method %in% 'naiive'){
+                                fit<- naiveBayes( as.formula(formula), df[-folds,])
+                                pred <- predict(fit, test, type = 'raw')
+                                pred<- pred[,2]
+                        }
+
+                        #pROC
+                        pred.p <- pROC::roc(test$f.rpdol, pred)
+
+                        if (i == 1) {
+                        plot.roc(pred.p, col = colours[i], print.thres = F)
+
+                        } else{
+                        plot.roc(
+                        pred.p, col = colours, add = TRUE , print.thres = F)
+                        }
+
+                        auc<- auc(pred.p)
+                        names(auc) = method
+
+        }
+
+        assign("auc.func", auc, envir = .GlobalEnv)
+        assign("formula", formula, envir = .GlobalEnv)
+        return(fit)
+
+}
+
+Log.Reg <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "log",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+summary(Log.Reg)
+
+```
+
+### Random Forest
+
+```{r}
+
+Rand.F <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "rf",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+Rand.F
+summary(Rand.F)
+
+```
+
+### Boosted Trees
+
+```{r}
+
+Boost.T <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "boost",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+# Summary of the first tree in the boosted tree
+pretty.gbm.tree(Boost.T) %>% kable
+
+summary(Boost.T)
+
+```
+
+### Naive Bayes
+
+```{r}
+
+Naive <- test.roc(dfs = c('all10mice_f'), seed = 100, mtrys = c(5), method = "naiive",
+                     extra.var = NULL, mult = 1, x.folds = 10)
+auc.func
+
+Naive
+
+```
+
+### Bayesian Network
+
+[@Genie]
+
+## Blended Models
+
+6 different methods for blending were applied. Three were simple and 3 were complex as listed below:
+Simple:
+
+* Averaging
+* Logistic Regression
+* Boosted Trees
+
+Complex:
+
+* Logistic Regression
+* Boosted Trees 
+* Random Forests
+
+### Simple Average
+
+```{r}
+
+if(.Platform$OS.type == 'windows'){
+
+        source("C:/Users/n9232371/Documents/github/consult/functions.R")
+} else{
+
+        source("/Users/yam/Documents/github/consult/functions.R")
+
+}
+
+df.blend <- readRDS("dfblend.rds")
+
+iter.auc <- function(mult =1, x.folds.iter = 5, 
+                     methods = c("simp.log"),
+                     df.iter = df.blend) {
+
+        #create test and train set: 0.75/0.25
+        pred.list <- vector(mode = "list", length = length(methods))
+        names(pred.list) = methods
+
+        #create model averaged column
+        if("average" %in% methods){
+        df.iter = mutate(df.iter, model.av = (log + rf + boost)/3) 
+        }
+
+        #set up 10 fold cv using caret function
+        folds <- createFolds(df.iter$f.rpdol, k = x.folds.iter, list = T, returnTrain = F)
+        #now have folds[[1]] through folds[[10]] list of rows to exclude
+        folds <- folds[[1]]
+
+                train = df.iter[-folds,]
+                test = df.iter[folds,]
+
+                if ("simp.log" %in% methods) {
+                        print("simp.log")
+                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost,
+                                   data = train,
+                                   family = binomial())
+
+                        pred <- predict(fit, test, type = "response")
+                        pred.list[["simp.log"]] = rbind(pred.list[["simp.log"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("simp.boost" %in% methods) {
+                        print("simp.boost")
+                        fit <- gbm(b.rpdol ~ norm.log + norm.rf + norm.boost, data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 6,
+                                   n.minobsinnode = 10)
+
+                        pred <- predict(fit, test, n.tree = 6000, type = "response")
+                        pred.list[["simp.boost"]] = rbind(pred.list[["simp.boost"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("orig.log" %in% methods) { 
+                        print("orig.log")
+                        pred <- test$log
+                        pred.list[["orig.log"]] = rbind(pred.list[["orig.log"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("orig.rf" %in% methods) { 
+                        print("orig.rf")
+                        pred <- test$rf
+                        pred.list[["orig.rf"]] = rbind(pred.list[["orig.rf"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("orig.boost" %in% methods) { 
+                        print("orig.boost")
+                        pred <- test$boost
+                        pred.list[["orig.boost"]] = rbind(pred.list[["orig.boost"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+
+                }
+
+                if ("average" %in% methods) {
+                        print("average")
+                        pred <- test$model.av
+                        pred.list[["average"]] = rbind(pred.list[["average"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+                if ("FWLS" %in% methods) {
+                        print("FWLS")
+                        fit <- glm(f.rpdol ~ norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + b.inv.log*norm.log + b.inv.log*norm.rf + Business*norm.boost + JD.Second + no.users,
+                                   data = train,
+                                   family = binomial())
+                        pred <- predict(fit, test, type = "response")
+                        pred.list[["FWLS"]] = rbind(pred.list[["FWLS"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+                if ("comp.boost" %in% methods) {
+                        print("comp.boost")
+                        fit <- gbm(b.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost, 
+                                   data = train,
+                                   distribution = "bernoulli", n.trees = 6000,
+                                   shrinkage = 0.0005, interaction.depth = 3,
+                                   n.minobsinnode = 10)
+
+                        pred <- predict(fit, test, n.tree = 6000, type = "response")
+                        pred.list[["comp.boost"]] = rbind(pred.list[["comp.boost"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+                if ("comp.rf" %in% methods) {
+                        print("comp.rf")
+                        fit <- randomForest(f.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost,
+                                            data = train, mtry = 3, 
+                                            ntree = 1000, importance = TRUE
+                        )
+                        pred <- predict(fit, test, type = "prob")
+                        pred <- pred[, 2]
+                        pred.list[["comp.rf"]] = rbind(pred.list[["comp.rf"]],
+                                                        data.frame("index" = folds,
+                                                                   "pred" = pred,
+                                                                   "iteration" = paste("iter",ceiling(x.folds.iter), sep = "")))
+                }
+
+        #pROC
+                        pred.p <- pROC::roc(test$f.rpdol, pred)
+
+                        plot.roc(pred.p, col = "darkblue", print.thres = F)
+
+                        auc<- auc(pred.p)  
+                        assign("auc.func", auc, envir = .GlobalEnv)
+
+        return(fit)
+
+}
+
+simp.av <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("average"),
+                     df.iter = df.blend)
+auc.func
+
+```
+
+### Logistic Regression
+
+#### Simple Blend
+
+```{r}
+
+simp.LR <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("simp.log"),
+                     df.iter = df.blend)
+auc.func
+summary(simp.LR)
+
+```
+
+#### Complex Logistic Regression Blend
+
+```{r}
+complex.LR <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("FWLS"),
+                     df.iter = df.blend)
+auc.func
+summary(complex.LR)
+
+```
+
+### Boosted Trees
+
+#### Simple Blend
+
+```{r}
+simp.boost <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("simp.boost"),
+                     df.iter = df.blend)
+auc.func
+summary(simp.boost)
+
+# Summary of the first tree in the boosted tree
+pretty.gbm.tree(simp.boost) %>% kable
+
+```
+
+#### Complex Blend
+
+```{r}
+comp.boost <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("comp.boost"),
+                     df.iter = df.blend)
+auc.func
+summary(comp.boost)
+
+# Summary of the first tree in the boosted tree
+pretty.gbm.tree(simp.boost) %>% kable
+
+```
+
+### Random Forest Complex Blend
+
+```{r}
+comp.rf <- iter.auc(mult =1, x.folds.iter = 5, 
+                     methods = c("comp.rf"),
+                     df.iter = df.blend)
+auc.func
+comp.rf
+summary(comp.rf)
+
+```
 

### 2016-08-31 finalwriteup/draft.rmd
paul mark ups

@@
 This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs or help tactfully choose when to enforce stricter contracts.
-It was found that in a specific decision framework, a case study's profits could be improved 9% by avoiding approximately 4% of projects.
+It was found that in a model-informed decision framework, a case study's profits could be improved 9% by avoiding approximately 4% of projects.
 Alternative decision frameworks are also proposed and evaluated.
@@
 This process was repeated on hundreds of models built from different training subsets of the data in order to obtain a 95% statistical confidence interval of profit improvements.
-The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% profit improvements.
+The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% improvement in profit.
 Further work is recommended to test user input estimation and user interface designs.
@@
 As most people have experienced, these estimates can be particularly inaccurate if the task is new and difficult.
-These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well being.
+These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial wellbeing.
 
@@
 Despite the range of possible deliverables, before any project commences, the client and consultant must agree on a fee, or at least a fee structure.
-The fee structure can be as creative as the engaged parties wish, but in fact tend to fit a few common models.
+The fee structure can be as creative as the engaged parties wish, but in fact tend to fit a few common models [@Malhotra2009].
 Two examples are:
@@
 In each of these contract models, financial risk is assigned to one of the parties.
-In model 1, the client must pay for an unknown quantity of consultants' billable hours and in model 2 the consultants must complete the project for a fixed price while not knowing how much of their own time will be spent.
+In model 1, the client must pay for an unknown quantity of consultants' billable hours and in model 2 the consultants must complete the project for a fixed price while not knowing how much of their own time will be spent [@Malhotra2009].
 
@@
 Typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb (such as a set fraction of the entire client budget for a large project).
-Such qualitative methods can result in profit margins, which ultimately depend on the skill (or luck) of the manager in predicting the amount of time a project demands.
+Such qualitative methods can impact profit margins, which ultimately depend on the skill (or luck) of the manager in predicting the amount of time a project demands.
 
 This thesis focuses on a single case study consulting company in the construction industry that has persevered through many fixed price projects that resulted in losses.
-A source of insight into the loss making projects was identified - twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
+A source of insight could be identified from twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
 Was it possible to find trends in what makes a project loss making and can these characteristics be identified before a fixed price is offered?
@@
 
-The three methods were anticipated to output different levels of influence for each variable.
-The tested methods were ANOVA Linear Regression, conditional inference forests (cForests), and Random Forests.
+Once outliers were removed, the next step was to determine a succinct list of variables that contributed most to the prediction.
+The three chosen variable importance methods were anticipated to output different levels of influence for each variable.
+These methods were ANOVA Linear Regression, conditional inference forests (cForests), and Random Forests.
 It was expected that Random Forests would favour variables that had more categories, a well-known bias; which is also a bias that cForests have overcome [@Strobl2007].
-ANOVA is based on linear theory, which contrasts conditional forest and Random Forest's more free-form structure.
+ANOVA is based on linear theory, which contrasts with conditional forest and Random Forest's more free-form structure.
 Therefore, it is beneficial to compare the methods' assessments of variable importance.
@@
 These included simple averaging of the individual model results, building a Logistic Regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
-Where possible, these methods shall be expressed as equations in an example problem with three original explanatory variables, $X_1$, $X_2$, $X_3$, and two full sets of predictions from individual predictive models, $M_1$ and $M_2$, with a response variable $Y$.
+Where possible, these methods shall be expressed as equations in an example problem with three original explanatory variables, $X_1$, $X_2$, $X_3$, and two full sets of predictions from individual predictive models, $M_1$ and $M_2$, with a response variable $p (probability)$.
 Values in $M_1$ and $M_2$ are probabilities between 0 and 1.
@@
 
-$Y = \frac{1}{2} \cdot (M_1 + M_2)$
+$p = \frac{1}{2} \cdot (M_1 + M_2)$
 
@@
 
-$Y = \beta_0 + \beta_1 M_1 + \beta_2 M_2$
+$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1 M_1 + \beta_2 M_2$
 
 The third method builds Boosted Trees from the individual models, i.e.
-using $M_1$ and $M_2$ only to predict $Y$ in a Boosted Tree model.
+using $M_1$ and $M_2$ only to predict $p$ in a Boosted Tree model.
 The last three methods take advantage of potentially meaningful interactions between the individual models' predictions and meta-features.
@@
 \begin{aligned} 
-Y &= \beta_0 + \beta_1 M_1 + \beta_2 M_2 + \beta_3 X_1 + \beta_4 X_2 + \beta_5 X_3 + \\
+\log\left(\frac{p}{1-p}\right) &= \beta_0 + \beta_1 M_1 + \beta_2 M_2 + \beta_3 X_1 + \beta_4 X_2 + \beta_5 X_3 + \\
 \qquad \beta_6(M_1 \cdot X_1) + \beta_7(M_1 \cdot X_2) + \beta_8(M_1 \cdot X_3) +\\
@@
 A split in a node determined by one variable is conditional upon the preceding split, which was based on another variable and so on.
-Random Forests and Boosted Tree models are difficult to express as equations, but for the example problem, $X_1$, $X_2$, $X_3$, $M_1$, and $M_2$ would all be included as explanatory variables to predict $Y$.
+Random Forests and Boosted Tree models are difficult to express as equations, but for the example problem, $X_1$, $X_2$, $X_3$, $M_1$, and $M_2$ would all be included as explanatory variables to predict $p$.
 
@@
 
-$RMSE_b = \sqrt{\frac{\sum_{t = 1}^{n} (\widehat{y}_t - \bar y)^2}{n}}$
+$RMSE_b = \sqrt{\frac{\sum_{t = 1}^{n} (\bar y - y_t)^2}{n}}$
 
@@
 n &= \textrm{number of data points} \\
-\widehat{y} &= \textrm{predicted values of the response variable} \\
-\bar y &= \textrm{mean 'return per dollar' for all projects} \\
+\bar y &= \textrm{mean 'return per dollar' as predicted value of the response variable} \\
+y_t &= \textrm{observed 'return per dollar' values} \\
 \end{aligned}
@@
 If the models were effective, the RMSE should be lower for ANOVA model predictions, so the difference would be greater than 0.
-Below is a histogram of this difference across 50 models run on randomly sampled 75% train and 25% test sets using core variables only (method 1 ANOVA).
+Because the size of the data set was limited, different divisions of the data into training and test sets would yield different RMSE statistics from the test set.
+Therefore, multiple models were built for each method in order to understand the distribution of resulting test statistics.
+Below is a histogram of $\Delta\ RMSE$ across 50 models run on randomly sampled 75% train and 25% test sets using core variables only (method 1 ANOVA).
 Note, only `r n.samples(data = results.A2$diff)` samples (models) were required for a statistical power of 80% [@pwr].
@@
 This was achieved by performing 10-fold cross validation 10 times.
-The violin plot below summarises the AUC values produced by each method and the 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
+The violin plot below summarises the distributions of the AUC values, where each violin is built from a pair of rotated kernel density plots of the subset of AUC values.
+The 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
 
@@
 The data contained 315 events per explanatory variable and did not pose that risk.
-Nevertheless, the additional variables must have added misleading noise to the model.
+Nevertheless, the additional variables might have added misleading noise to the model.
 

### 2016-12-13 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd
just paper, sort out pngs later

@@
 library('e1071')
-library('party')
 library('randomForest')
@@
 
+Depending on the industry, some consulting projects can be financialy risky endeavors for the consultant.
+Engineering consultanting companies in the construction industry complete a wide range of complex projects in a competitive enviroment.
+Because of the way contracts are structured, many of these projects result in losses.
+The purpose of this study was to test whether these loss making projects could be predicted by statistical and machine learning algorithms using the information available at project conception.
+The second question is whether the predictive algorithm could improve a company's profitability.
+Project data from a case study engineering company was used to answer these questions.
+
 1 page
-
-Intro - pick out bits from thesis
 
@@
 
-for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.  
-This thesis focuses on the risk taken by a consulting company when offering a fixed fee.
-Clients commonly seek fixed price quotes from several consultants before selecting one or negotiating further.
-Hence, quoted fees must be competitive.
-The way a consulting manager calculates a fixed price varies from industry to industry and even company to company.
-Typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb
-- amount of losses, sources
-A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget [@Flyvbjerg2011].
+Financial risk most commonly falls upon a consulting company when they must offer a fixed fee for a project.
+In competitive climates, clients request fixed price quotes for project completion from several consultants before a project is awarded.
+Hence, initial quoted fees must be competitive but also accurately cover consultant salaries and business costs.
+
+How do consultants make this accurate yet competitive fee estimate?
+This varies from industry to industry, but typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb.
+Unfortunately, some projects priced using this method can blow out of budget, causing significant losses to the consultant.
+
+A review of 6 IT project surveys between 1984 and 1994 by Moløkken & Jørgensen (2003) revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
 A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by an average of 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007].
-Smaller consulting companies in competitive industries, such as the construction industry, experience similar difficulties in forecasting their project costs.
-losses = low employee morale, elevated stress, low work quality to improve speed, possibly redundancies
-@Lovallo2003 delved into the psychology behind why executives so often severely underestimate costs of larger projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development.
-Their theory stemmed from Kahneman's work on decision-making that won him the Nobel Prize for economics in 2002.
-His research argued that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project.
-A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplays or ignores the risk of problems that are out of the team's control.
-all complex projects are at risk of encountering a multitude of problems that the manager could never foresee.
-Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].
 
-Unfortunately, limited research is available that documents the methods industry uses for complex-projects.
-The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors - detailed analysis and experience based models
-A survey by @Moores1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools
+Kahneman won a nobel prize in 2002 for his research into decision making on this topic.
+He argued that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project [@Lovallo2003].
+To counter this phenomenon, research by Flyvbjerg [-Flyvbjerg2011] advised that managers should be confronted by the financial performance of similar projects to the one being costed so that they take on a more realistic view of their abilities.
+This was based off a body of findings where people made better predictions about themselves after being exposed to facts about other people.
+For example, subjects were asked to predict their skill levels before and after being exposed to a summary of other people's skills, and their predictions after were significantly more accurate [@Lovallo2003].
 
-Detailed analysis remains the most prevalent method for construction cost estimation despite the industry having a long history of projects running over time and budget using the same technique [@Shane2009].
+This work aims to provide a manager with this more realistic view by predicting whether the new project will be profitable or not, based on a suite of previous project data from the same company.
+The manager could then take multiple actions including not tendering for a project predicted to be loss-making, providing a high fixed price, or restructuring the contract so that it is not 100% fixed price.
+This research tests the simple case of rejecting all projects predicted as loss-making.
 
@@
 
+To assess whether the model would improve the case study company's profitability, a simple managerial decision framework was applied: if the algorithm predicted the project to be loss making, the project was considered 'rejected' and all profits or *losses* of rejected projects were discarded.
+The remaining projects' profits and losses were summed to give a revised overall profit over the last 12 years of projects.
+This was compared to the summed profits and losses from all projects completed over the last 12 years.
+The best algorithm rejected enough loss-making projects to improve overall profits by 9%.
+
 - aim
@@
 The overall profit is represented by the following equation:
-
-- brief summary of major findings/results
 

### 2016-12-13 1st_stats.R

@@ -1,344 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
-
-all6b<- read.csv('all6b.csv')[,-1]
-all6b$Start.Date<- as.Date(all6b$Start.Date)
-
-#delete milestone rows! sort by highest hours - choose this discipline.!
-#sort by hours first
-all6b<- all6b %>% arrange(mlsto, -hours)
-dup<- all6b[,names(all6b) %in% c('mlsto')]
-all7<-all6b[!(all6b$inv.mlsto==0 | duplicated(dup)),]
-
-write.csv(all7, 'all7.csv')
-all7<- read.csv('all7.csv')[,-1]
-all7$Start.Date<- as.Date(all7$Start.Date)
-
-#is return.pdol and inv.vs.cost normally distributed??????
-#use qqnorm and qqline
-library('car')
-library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
-
-qqnorm(all7$return.pdol)
-qqline(all7$return.pdol)
-qqPlot(all7$return.pdol)
-
-# up tail at end. try log of variable
-qqnorm(log(abs(all7$return.pdol+.001)))
-qqline(log(abs(all7$return.pdol+.001)))
-#this looks s shaped
-
-qqnorm(log(all7$return.pdol+.8))
-qqline(log(all7$return.pdol+.8))
-
-#find sd and mean of return.pdol
-sd(all7$return.pdol)
-mean(all7$return.pdol)
-skewness(all7$return.pdol)
-skewness(log(all7$return.pdol+2.15))
-#skewness is 27.2 and 2.12 for logged data!!!
-
-#have a look at the shape of return.pdol as bar graph
-ggplot(all7, aes(x=return.pdol)) + geom_bar(stat='bin')
-ggplot(all7, aes(x=log(all7$return.pdol+.8))) + geom_bar(stat='bin')
[…]

### 2016-12-13 BNprep.R

@@ -1,243 +1,0 @@
-## prep for bayesian network
-
-#discretise numeric variables and change all 'NA' to blank
-
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')[,-1]
-head(all8a)
-
-#discretise numeric columns!!!! pc.pro, timespan.cbrt, inv.mlsto.log, client.totinv.log,
-#pc.majpos.log, return.pdol
-
-###
-#pc.pro
-x<- dist(all8a$pc.pro, method = 'euclidean')
-fit<- hclust(x, method = 'ward.D2')
-plot(fit)
-rect.hclust(fit, k = 5)
-
-#assign a cluster number to all rows in data frame
-all8a$b.pc.pro<- as.factor(cutree(fit, k=5))
-head(all8a$b.pc.pro)
-
-for(i in 1:5){
-        print(
-        summary(all8a %>% select(b.pc.pro, pc.pro) %>% filter(b.pc.pro ==i))
-        )
-}
-
-#create clusters based on summary:
-all8a[all8a$pc.pro>100,'pc.pro']<- 100
-all8a$b.pc.pro<- cut(all8a$pc.pro, breaks=c(0,15,40,75,95,100), include.lowest=TRUE,
-                     labels= c("p0_15", "p15_40", "p40_75", "p75_95", "p100"))
-
-###
-#timespan.cbrt
-# all timespan == 0 should be timespan ==1. Checked all7a and they all have at least 0.5 hour booked, max 18 hours.
-all8a$timespan.cbrt <- ifelse(all8a$timespan.cbrt ==0, 1, all8a$timespan.cbrt)
-first.vec<- all8a$timespan.cbrt
-x<- dist(first.vec, method = 'euclidean')
-fit<- hclust(x, method = 'ward.D2')
-plot(fit)
-k.choose=3
-rect.hclust(fit, k = k.choose)
-
[…]

### 2016-12-13 BNprep_all9c.R

@@ -1,117 +1,0 @@
-## prep for bayesian network
-
-#discretise numeric variables and change all 'NA' to blank
-
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
-head(all9c)
-
-# write a function to
-# discretise numeric columns!!!! pc.pro, timespan.cbrt, inv.mlsto.log, client.totinv.log,
-# pc.majpos.log, return.pdol
-
-discretise<- function(df = all9c, extra.var = ('code.client', 'code.contact')){
-        
-        #pc.pro
-        df[df$pc.pro>100 & !is.na(df$pc.pro),'pc.pro']<- 100
-        df$b.pc.pro<- cut(df$pc.pro, breaks=c(0,15,40,75,95,100), include.lowest=TRUE,
-                             labels= c("p0_15", "p15_40", "p40_75", "p75_95", "p100"))
-        
-        ###
-        #timespan.cbrt
-        #done already
-        
-        #create clusters based on summary:
-        man.clust<- function(first.vec=all9c$timespan.cbrt, breaks.man=c(0,4.16,6.558,15),
-                             labs = c("t0_70", "t70_280", "t280_")){
-                final = cut(first.vec, breaks=breaks.man, include.lowest=TRUE, labels= labs)
-                return(final)
-        }
-        
-        ###
-        #inv.mlsto.log
-        
-        all9c$b.inv.log<- man.clust(first.vec=all9c$inv.mlsto.log, 
-                                    breaks.man=c(4.5, 7.81, 9, 11, 14.4),
-                                    labs = c("i_0_2_5k", "i_2_5k_8_1k", "i_8_1k_60k", "i_60k_1_6m"))
-        
-        ###
-        #client.totinv.log
-        
-        all9c$b.client.totinv <- man.clust(first.vec=all9c$client.totinv.log, 
[…]

### 2016-12-13 Business_compact.R

@@ -1,56 +1,0 @@
-# re-jig Business
-
-str(all9a)
-
-B.invest<- table(all9a$b.rpdol, all9a$Business) %>% t %>% as.data.frame.matrix()
-colnames(B.invest) <- c("profit", "loss")
-B.invest<- transform(B.invest, ratio = round(loss/profit, 2))
-B.invest$JD <- rownames(B.invest) 
-B.invest<- arrange(B.invest, ratio) %>% select(JD, profit, loss, ratio)
-
-#group:
-
-# university, school
-# lawyer/body corp
-# signs/manufacturer
-# environment/water
-# roads/rail/gov/util
-# artist/landscape arch
-# hospital/institution/health
-# utilities
-# engineer
-# water
-# business/building services/internal
-# builder
-# architect
-# townplanner
-# developer/realestate
-# resources
-# person
-
-
-
-#code to compact:
-
-all9a$test<- as.character(all9a$Business)
-all9a$test <- ifelse(grepl("lawyer|body", all9a$Business), 
-                     "lawyer_bodyC", all9a$test)
-all9a$test <- ifelse(grepl("univ|school", all9a$Business), 
-                     "uni_school", all9a$test)
-all9a$test <- ifelse(grepl("sign|manuf", all9a$Business), 
-                     "sign_manufac", all9a$test)
-all9a$test <- ifelse(grepl("environ|water", all9a$Business), 
-                     "enviro_water", all9a$test)
-all9a$test <- ifelse(grepl("road|gov|util", all9a$Business), 
-                     "road_rail_gov_util", all9a$test)
-all9a$test <- ifelse(grepl("artist|landscape", all9a$Business), 
-                     "artist_landarch", all9a$test)
-all9a$test <- ifelse(grepl("hospit|institut|health", all9a$Business), 
-                     "hosp_institut_health", all9a$test)
-all9a$test <- ifelse(grepl("business|building|internal", all9a$Business), 
-                     "biz_bldgserv_internal", all9a$test)
-
-summary(as.factor(all9a$test))
-
-
-

### 2016-12-13 JD.Second_compact.R

@@ -1,70 +1,0 @@
-
-# re-jog JD.Second and Business
-
-str(all9a)
-
-JD.invest<- table(all9a$b.rpdol, all9a$JD.Second) %>% t %>% as.data.frame.matrix()
-colnames(JD.invest) <- c("profit", "loss")
-JD.invest<- transform(JD.invest, ratio = round(loss/profit, 2))
-JD.invest$JD <- rownames(JD.invest) 
-JD.invest<- arrange(JD.invest, ratio) %>% select(JD, profit, loss, ratio)
-
-#group:
-
-# heritage/ expert opinion
-# wharf/port, bridges
-# membrane shade structure/art/awning/facade
-# hospital/emergency services/health/carpark
-# refurb/renovation
-# sign, product development
-# hotel/resort, offices, dining/retail
-# report/investigation
-# waste water/water management
-# university/school/community/extension
-# multi-dwelling/house
-# industrial
-# flood studies, water harvesting, recycled water
-# parks and open spaces
-# civil works to bldg, Subdivision, sewerage
-
-
-
-# delete - aged care, erosion and sediment
-
-#code to compact:
-
-all9a$test<- as.character(all9a$JD.Second)
-all9a$test <- ifelse(grepl("heritage|expert", all9a$JD.Second), 
-                          "heritage_expert", all9a$test)
-all9a$test <- ifelse(grepl("wharf|bridges", all9a$JD.Second), 
-                          "wharf_bridge", all9a$test)
-all9a$test <- ifelse(grepl("membrane|art|awning|facade", all9a$JD.Second), 
-                          "art_facade_awn_memb", all9a$test)
-all9a$test <- ifelse(grepl("hosp|emerg|health|carpark", all9a$JD.Second), 
-                          "hosp_health_carpark", all9a$test)
-all9a$test <- ifelse(grepl("sign|product", all9a$JD.Second), 
-                          "sign_product", all9a$test)
-all9a$test <- ifelse(grepl("hotel|office|dining", all9a$JD.Second), 
-                          "hotel_office_dining", all9a$test)
-all9a$test <- ifelse(grepl("report|investigation", all9a$JD.Second), 
-                          "report", all9a$test)
-all9a$test <- ifelse(grepl("waste water|management", all9a$JD.Second), 
-                          "waste_wat_manage", all9a$test)
-all9a$test <- ifelse(grepl("univ|commun|school|exten", all9a$JD.Second), 
[…]

### 2016-12-13 SS/1stplots.R

@@ -1,191 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-
-all6<- read.csv('all6.csv')[,-1]
-all6$Start.Date<- as.Date(all6$Start.Date)
-#check extremes
-all6 %>% arrange(-return.pdol) %>% select(return.pdol, Tot.Invoiced, mlsto, hours) %>% head(20)
-#check specific job numbers
-all6 %>% select(return.pdol, Tot.Invoiced, mlsto, hours, cost.mlsto, dis.sc.mlsto) %>% filter(mlsto == '2013.430.4')
-
-#all7 has deleted all milestone duplicates and deleted outliers
-all7<- read.csv('all7.csv')[,-1]
-all7$Start.Date<- as.Date(all7$Start.Date)
-all7<- all7 %>% filter(!(inv.vs.cost<0.2), !(inv.vs.cost>4))
-
-
-#profitability
-#discipline vs profit
-#subset out invoiced=0
-dup<- all6[,names(all6) %in% c('mlsto', 'Discipline')]
-subd1<-all6[!all6$Tot.Invoiced==0 & !is.na(all6$Discipline) & !duplicated(dup),] 
-#make outlier points match variable colour
-update_geom_defaults("point", list(colour = NULL))
-bal.discplot<- ggplot(subd1, aes(x=Discipline, y= balance.mlsto,
-                                     colour= Discipline)) + 
-        geom_boxplot(alpha=0) +geom_jitter(alpha= I(1/2))+
-        scale_y_continuous(breaks = seq(round(min(subd1$balance.mlsto),-5), round(max(subd1$balance.mlsto),-5), by = 20000))
-
-prof.discplot<- ggplot(subd1, aes(x=Discipline, y= profit.mlsto,
-                                 colour= Discipline)) + 
-        geom_boxplot(alpha=0) +geom_jitter(alpha= I(1/2))+
[…]

### 2016-12-13 boost_prep.R

@@ -1,29 +1,0 @@
-
-## prep for boosted trees
-
-#discretise numeric variables and change all 'NA' to blank
-
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')[,-1]
-head(all8a)
-
-#try to predict profit/loss for return per dollar!
-###
-#return.pdol
-# split into profit/loss
-
-all8a$b.rpdol<- all8a$return.pdol 
-all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, "loss", "profit")
-all8a$b.rpdol<- as.factor(all8a$b.rpdol)
-
-summary(all8a$b.rpdol)
-
-#build boosted tree
-
-

### 2016-12-13 cleaning.R

@@ -1,695 +1,0 @@
- setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-year13<-read.csv('2013projectdata.csv')
-year14<-read.csv('2014projectdata.csv',
-                 na.strings=c(""," ","Number of Blue Files","JobInvCount","Total Costs (AUD)",
-                                                    "Timesheet Hours","Timesheet Total Charge (AUD)",
-                                                    "Timesheet Total Cost (AUD)","Total Disbursements (AUD)",
-                                                    "Total Fee (AUD)","Total QB Invoiced (AUD)",
-                                                    "Total Sub-Contractors (AUD)","Project Value (AUD)",
-                              "Job Number","NA"))
-year02to12<-read.csv('project_data_all.csv',
-                     na.strings=c(""," ","Number of Blue Files","JobInvCount","Total Costs (AUD)",
-                                  "Timesheet Hours","Timesheet Total Charge (AUD)",
-                                  "Timesheet Total Cost (AUD)","Total Disbursements (AUD)",
-                                  "Total Fee (AUD)","Total QB Invoiced (AUD)",
-                                  "Total Sub-Contractors (AUD)","Project Value (AUD)","Job Number","NA"))
-#delete columns
-year13<-year13[,-c(4,5)]
-year13<-year13[,-c(15:17)]
-year13<-year13[,-c(25:28)]
-year13<-year13[,-c(37:38)]
-head(year13)
-dim(year14)
-dim(year02to12)
-#create separate file for hours of file year02to12
-hours0212<-year02to12[,c(1,41:52)]
-year02to12<-year02to12[,-c(41:52)]
-write.csv(hours0212,file='hours0212.csv')
-c.names<-data.frame("13"=colnames(year13)[1:40],"14"=colnames(year14)[1:40],
-                    "all"=colnames(year02to12)[1:40])
-c.names
-
-
[…]

### 2016-12-13 combine.R

@@ -1,75 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-
-##### COMBINE CLEANED DATA SETS BEFORE DE-IDENTIFICATION #########
-
-#combining all dataset with Job hours data set and Inv_sum data set.
-#Job.Hours data set columns need to replace some of the all columns because they are more accurate. 
-# these include: Start.Date, tot.TS.Charge, tot.TS.cost, tot.TS.hours, Year
-
-all4e<- read.csv('all4e.csv')[,-1]
-str(all4e)
-all4e<- all4e %>% filter(!(Job.Number == '2010.041.1' & dist == 1362.8))
-#delete Start.Date, tot.TS.Charge, tot.TS.cost, tot.TS.hours, Year
-all4e<- all4e %>% select(-Start.Date, -Tot.TS.Cost, -Tot.TS.Charge, -TS.Hours, -Year, -Profit, -Tot.Invoiced)
-
-#combine Inv.Sum
-Inv.sum<- read.csv('inv_sum.csv')[,-1]
-all4e<- merge(all4e, Inv.sum %>% select(Job.Number, Tot.Invoiced), by= 'Job.Number', all.x=TRUE, all.y=FALSE)
-#set NA = 0
-all4e[is.na(all4e$Tot.Invoiced),]$Tot.Invoiced <- 0
-
-#combine Job.Hours
-Job.Hours<- read.csv('Job_hours.csv')[,-1]
-Job.Hours<- Job.Hours %>% select(-mlsto)
-all5<- merge(all4e, Job.Hours, by.x='Job.Number', by.y= 'Job.num', all.x=TRUE, all.y=FALSE)
-str(all5)
-
-#create new 'profit' and 'balance' column
-#profit column
-all5<- transform(all5, profit= Tot.Invoiced-charge-Dis.subcon)
-#balance column - subtract real costs
-all5<- transform(all5, balance= Tot.Invoiced-cost-Dis.subcon)
[…]

### 2016-12-13 combine2.R

@@ -1,168 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library('rCharts',lib='C:/Progra~1/R/R-3.2.0/library')
-
-invoices<- read.csv('invoices_clean.csv')[,-1]
-hours<- read.csv('hours_clean.csv')[,-1]
-
-#### CREATING SHINY PRODUCT - MONTHLY REVIEW
-
-### STEP 0
-# subset invoices df and hours df to 2009 onwards 
-invoices$Invoice.Date <- as.Date(invoices$Invoice.Date)
-invoices<- arrange(invoices, Invoice.Date)
-invoices<- invoices %>% filter(Invoice.Date>= as.Date('2009-01-01'))
-
-hours$Work.Date <- as.Date(hours$Work.Date, "%d/%m/%Y")
-hours<- hours %>% filter(Work.Date>= as.Date('2009-01-01'))
-
-### STEP 0.5
-# assign tag name to all rows between dates
-# use last date in month
-invoices$MonthYear<- format(invoices$Invoice.Date,"%Y-%b")
-tag<- ddply(invoices, .(MonthYear), tail, 1)
-tag<-tag %>% arrange(Invoice.Date)
-
-df<- data.frame('start' = c(as.Date('2008-12-31'), tag$Invoice.Date[-nrow(tag)]),'end'= tag[,6], 'tag' = tag[,10])
-
-
-## try ifelse function --> similar to grepping 
-#assign tag to all hour rows based on whether they are between two dates in the df data.frame (see above)
-#these dates are determined by the last date of invoices in each month
-
-btwn.dates<- function(date.df = df, col.to.tag = hours$Work.Date){
-        a= c(rep(NA,length(col.to.tag)))
-        for( i in 1:nrow(date.df)){
-                a = ifelse(col.to.tag >= date.df[i,1] & col.to.tag < date.df[i,2], as.character(date.df$tag[i]), a) 
-        } 
-        a<- as.factor(a)
-        a
-}
[…]

### 2016-12-13 explore.R

@@ -1,12 +1,0 @@
-###EXPLORATION - new file
-#plot invoiced vs hours spent?
-ggplot(all,aes(Invoiced,Timesheet.Hours))+geom_point(aes(colour=Discipline))
-ggplot(all[all$Invoiced<25000 & all$Invoiced>0,],aes(Invoiced,Timesheet.Hours))+geom_point(aes(colour=Discipline))
-#plot invoiced=0 jobs
-ggplot(all,aes(Tot.Invoiced,Timesheet.Hours))+geom_point(aes(colour=Discipline))
-#get a feel for how many jobs we didn't invoice for
-nrow(all[all$Invoiced==0,])
-nrow(all[all$Invoiced>0,])
-nrow(all[all$Invoiced==0 & all$Timesheet.Hours==0,])
-#plot year vs profit
-ggplot(all,aes(x=Year,y=Profit))+geom_point()

### 2016-12-13 extra_role_concat.R

@@ -1,21 +1,0 @@
-Df<-read.csv('extra_info.csv')
-
-concat<-function(Data=Df){
-        Ans=data.frame(matrix(0,nrow=length(na.omit(unique(Data$jobno.shift))),ncol=2))
-        Is.NA = is.na(Data[,'jobno.shift'])
-        Index.1 = data.frame(RN = 1:length(Is.NA), Is.NA)
-        Index.F = Index.1[Index.1$Is.NA == FALSE,]
-        Ans[,1]=Data[Index.F[,1],4]
-        for (i in 1:(nrow(Index.F)-1)){
-                SS1 = Data[Index.F$RN[i]:(Index.F$RN[i+1]-1),]
-                Ans[i,2]=paste(SS1$Role,collapse='. ')
-        SS2=Data[Index.F$RN[nrow(Index.F)]:nrow(Data),]
-        Ans[nrow(Index.F),2]=paste(SS2$Role,collapse='. ')
-        }
-        return(Ans)
-}
-
-concat<-concat(Data=Df)
-colnames(concat)[colnames(concat)=="X1"] <- "Job.Number"
-head(concat)
-#try merging with all

### 2016-12-13 final_app/server.R

@@ -1,270 +1,0 @@
-if(.Platform$OS.type == 'windows'){
-        setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-        all7<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7.csv')[,-1]
-        all7d<- read.csv('all7d.csv')[,-1]
-        
-} else{
-        setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-        all7<- read.csv('all7.csv')[,-1]
-        all7d<- read.csv('all7d.csv')[,-1]
-        
-}
-library('ggplot2')
-library("dplyr")
-library("plyr")
-library('magrittr')
-library('reshape2')
-library('shiny')
-library('shinythemes')
-library('FNN')
-# library('ggplot2')
-# library("dplyr")
-# library("plyr")
-# library('magrittr')
-# library('reshape2')
-# library('shiny')
-# library('shinythemes')
-
-
-all7$Start.Date<- as.Date(all7$Start.Date)
-#delete anything more than 4 for inv.vs.cost
-#delete anything less than .2 for inv.vs.cost
-all7<- all7 %>% filter(!(inv.vs.cost<0.2), !(inv.vs.cost>4))
-
-#data for final plot
-df <- read.table(text = " id  min  max 
-    Sp1     150          151     ", header=TRUE)
-
-#make a new dataframe with information about crossbar sizes
-df2<- data.frame(id=c(rep("Sp1",1)), ymin= c(-8.8), y.mean = c(-7),
-                 ymax=c(-5.2), confidence=c('90%'))
-df2 <- within(df2, 
-              confidence <- factor(confidence, 
-                                    levels=c('90%')))
-prof.labels = data.frame(xx = 0.53, yy = seq(-10,10,by= 10), 
-                         text = c("Loss-Making Job","Neutral","Profitable Job"))
-
-#### CREATE TABLE DATA FRAME FOR NEAREST NEIGHBOURS ####
-
-all7d$mlsto<- as.character(all7d$mlsto)
-
-# test.case<- all7d %>% select(Discipline, Business, Biz.type, code.contact, code.client, JD.Second, Billing.Type)
-
-#function for retrieving nearest neighbours by inv.mlsto:
-inv.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3){
-        invo.knn = knn(df[,predict], df[new.cases,predict],
-                       df$pc.pro, k=k)
[…]

### 2016-12-13 final_app/ui.R

@@ -1,252 +1,0 @@
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shiny')
-
-shinyUI(fluidPage(
-        theme=shinytheme("journal"),
-        
-        #Application title
-        titlePanel("Quote Assist"),
-        
-        # Sidebar with controls to select the variable to plot against positions
-        wellPanel(
-        fluidRow(
-                column(3,
-        
-                selectInput("disc", "Discipline:",
-                            list("Civil" = "Civil", 
-                                 "Structural" = "Structural", 
-                                 "Environmental Planning" = "Environmental Planning",
-                                 "Water" = "Water",
-                                 "All"= "All")),
-                
-                selectInput("billing", "Billing Type:",
-                            list('Hourly Rate' = 'Hourly Rate',
-                                 "Fixed Fee"= 'Fixed Quote'
-                                 )),
-                
-                selectizeInput(
-                        'JD.Second', 'Job Type:', 
-                        choices = levels(all7$JD.Second),
-                        options = list(
-                                placeholder = 'Please start typing',
-                                onInitialize = I('function() { this.setValue(""); }'),
-                                maxItems=1
-                        ))
-                
-                ),
-                
-                column(3,
-                       
-                       selectizeInput(
-                               'code.client', 'Client:', choices = levels(all7$code.client),
-                               options = list(
-                                       placeholder = 'Please start typing',
-                                       onInitialize = I('function() { this.setValue(""); }'),
-                                       maxItems=1
-                               )),
-                       
-                       selectizeInput(
-                               'Business', 'Client Industry:', 
-                               list("Architect" = "architect", 
-                                    "Artist" = "artist", 
-                                    "Builder" = "builder",
-                                    "Building Services"= 'building services',
-                                    "Business" = "business",
-                                    "Council" = 'council',
-                                    "Developer/Real Estate"= 'developer/real estate',
-                                    "Engineer"= 'engineer',
-                                    "Environment" = 'environment',
-                                    "Friend" = 'friend',
[…]

### 2016-12-13 functions.R

@@ -1,531 +1,0 @@
-#function for running caret..
-
-caret.all <- function(method = "boost", df = train, formula = formula.rf, sample.frac = 0.75, seed = 100){
-        # set seeds
-        set.seed(seed)
-        seeds <- vector(mode = "list", length = 51)
-        for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
-        ## For the last model:
-        seeds[[51]] <- sample.int(1000, 1)
-        
-        train.c = sample_frac(df, sample.frac, replace = FALSE)
-        test.c = setdiff(df, train.c)
-        
-        cv.Control <- trainControl(method = "cv",
-                                   number = 5,
-                                   seeds= seeds
-                                   , summaryFunction= twoClassSummary,
-                                   classProbs=TRUE
-                                   )
-        
-        if(method == "rf"){
-                mtry.Grid <- expand.grid(mtry = c(3,4,5,6,7))
-                
-                set.seed(2)
-                gbmFit <- caret::train(as.formula(formula), data= train.c,
-                                       method = "rf", trControl = cv.Control, verbose = FALSE,
-                                       tuneGrid = mtry.Grid, ntree = 1000,
-                                       metric = 'ROC', allowParallel = T)
-        }
-        
-        if(method == "boost"){
-                train.c$b.rpdol<- as.factor(train.c$b.rpdol)
-                levels(train.c$b.rpdol)[levels(train.c$b.rpdol)=="0"] <- "profit"
-                levels(train.c$b.rpdol)[levels(train.c$b.rpdol)=="1"] <- "loss"
-                Grid <- expand.grid(shrinkage = c(.0005),
-                                         n.trees = c(6000),
-                                    interaction.depth = c(3,4,5,6),
-                                    n.minobsinnode = c(10,20,30,40))
-                
-                # , interaction.depth = c(1,2,3,4,5),
-                # n.minobsinnode = c(10,20,30)
-                
-                set.seed(2)
-                gbmFit <- caret::train(as.formula(formula), data= train.c,
-                                method = "gbm", trControl = cv.Control, verbose = FALSE,
-                                bag.fraction = 0.5, tuneGrid = Grid,
-                                metric = 'ROC',
-                                na.action = na.pass)
-        }
-        
-        plot(gbmFit)
-        return(gbmFit)
-        
-}
-
-cfor.plotsave<- function(df= vars, name= 'cfor_blend', formular = formula.rf){
-                
-                #run cforest
[…]

### 2016-12-13 hours_cleaning.R

@@ -1,211 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-hours12<- read.csv('hours0212.csv', na.strings= c(""," ", "NA"))[,-1]
-hours1314<- read.csv('20132014hoursdata.csv', na.strings= c(""," ", "NA"))
-
-#ensure column orders and names are correct
-hours1314<- hours1314[,c(13,1:12)]
-colnames(hours12)[names(hours12) %in% 'Job.Number']<-'Job.num'
-colnames(hours12)[names(hours12) %in% 'Discipline.1']<-'Discipline'
-
-
-#what is WIP only invoice??
-
-#clean up
-# any columns unnecessary?
-summary(hours1314)
-#not really
-#remove NA's from hours1314 first. look into full NA rows
-dim(hours)
-hours1314[rowSums(is.na(hours1314))== 13,] %>% nrow()
-hours1314<-hours1314[rowSums(is.na(hours1314))!=12,]
-hours1314<-hours1314[rowSums(is.na(hours1314))!=13,]
-#now to rows with 3 values
-hours1314[rowSums(is.na(hours1314))== 9,] %>% head()
-#are all the charge/cost values 0?? yes - only details - delete these rows
-hours1314<-hours1314[rowSums(is.na(hours1314))!=9,]
-#investigate last 9 rows of NA in charge
-hours1314[is.na(hours1314$Charge.Amount..B...AUD.),] %>% head()
-#these were full NA rows
-hours1314<-hours1314[rowSums(is.na(hours1314))!=13,]
-
-#remove NA's from hours12 now. look into full NA rows
-summary(hours12)
-hours12[rowSums(is.na(hours12))== 12,] %>% head()
-hours12<-hours12[rowSums(is.na(hours12))!=13,]
-hours12<-hours12[rowSums(is.na(hours12))!=12,]
-#now to rows with 3 values
-hours12[rowSums(is.na(hours12))== 9,] %>% head()
[…]

### 2016-12-13 imputingall9c.R

@@ -1,122 +1,0 @@
-# impute missing values
-
-library('mice')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
-library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
-library('CALIBERrfimpute', lib = 'C:/Progra~1/R/R-3.2.2/library')
-
-all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
-all9c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all9c.csv')[,-1]
-all9c$b.rpdol<- as.factor(all9c$b.rpdol)
-
-all10rf<- rfImpute(b.rpdol ~ Discipline + pc.pro + timespan.cbrt + no.users + inv.mlsto.log +
-                 client.totinv.log + pc.majpos.log + majority.pos + Business + JD.Second + Billing.Type,
-                 data= all9c,
-              ntree= 500)
-all10rf<- cbind(all10rf, all9c[,names(all9c) %in% c('code.client', 'code.contact', 'mlsto',
-                                                    'b.timespan.cbrt', 'b.rpdol')])
-
-
-## test if it helps boosted trees
-
-#convert predictor variable back to numeric 0,1
-
-levels(all10rf$b.rpdol)<- c("0", "1")
-all10rf$b.rpdol<- as.character(all10rf$b.rpdol)
-all10rf$b.rpdol<- as.integer(all10rf$b.rpdol)
-write.csv(all10rf, 'C:/Users/n9232371/Documents/Consultbusiness/data/all10rf.csv')
-
-levels(all10mice$b.rpdol)<- c("0", "1")
-all10mice$b.rpdol<- as.character(all10mice$b.rpdol)
[…]

### 2016-12-13 invoice_cleaning.R

@@ -1,193 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.0/library')
-
-###DELETE VARIABLES###
-
-invoices<-read.csv('invoices.csv')
-head(invoices)
-str(invoices)
-summary(invoices)
-invoices<-invoices[,-c(10,11)]
-invoices<-invoices[,!(names(invoices) %in% 'Created.Date')]
-invoices<-invoices[,!(names(invoices) %in% 'Status')]
-
-###RENAME COLUMNS###
-colnames(invoices)[names(invoices) %in% 'Job']<-'Job.Name'
-colnames(invoices)[names(invoices) %in% 'Amount.Paid..AUD.']<-'Amnt.Paid'
-colnames(invoices)[names(invoices) %in% 'Inv.Amount..E...AUD.']<-'Inv.Amount'
-colnames(invoices)[names(invoices) %in% 'qbin_CreatedBy']<-'Created.by'
-
-###RE-CLASS COLUMNS###
-invoices$Invoice.Date<-as.Date(invoices$Invoice.Date, '%d/%m/%Y')
-
-###DELETE CASES CREATED ON JAN 16 2014 - after original data was harvested
-invoices<-invoices[!(invoices$Invoice.Date > as.Date("2015-01-10")),]
-
-###ENGINEERED VARIABLES###
-
-#delete job number from job.name
-invoices$Job.Name<-as.character(invoices$Job.Name)
-strsplit(invoices$Job.Name[1],split='[:]')[[1]][2]
-invoices$Job.Name<-sapply(invoices$Job.Name,FUN=function(x){strsplit(x,split='[:]')[[1]][2]})
-invoices$Job.Name<-sub(' ','',invoices$Job.Name)
-#year of job
-invoices$Job.Number<-as.character(invoices$Job.Number)
-strsplit(invoices$Job.Number[1],split='[.]')[[1]][1]
-invoices$Year<-sapply(invoices$Job.Number,FUN=function(x){strsplit(x,split='[.]')[[1]][1]})
-invoices$Year<-as.integer(invoices$Year)
[…]

### 2016-12-13 invoice_eng.R

@@ -1,136 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.0/library')
-
-
-Invoices.a<-read.csv('invoices_clean.csv')[,-1]
-Inv.sum<-read.csv('inv_sum.csv')[,-1]
-
-
-#### ENGINEERING VARIABLES ######
-
-#need to do by milestone!! because this is what relates directly to the return.pdol variable which we are measuring as 'success'
-
-#Need to sum up all the profits and balances for jobs with milestones
-# first 
-#create subset of just the things i want to add up - profits, balance, hours in each milestone
-temp<- Invoices.a %>% select(Job.Number, Inv.Amount, Invoice.Date)
-temp<- temp %>% arrange(Job.Number)
-temp$Invoice.Date <- as.character(temp$Invoice.Date)
-temp$Job.Number <- as.character(temp$Job.Number)
-#create column for Date of prev invoice, prev job number, and prev invoiced amount
-temp$testp.Invoice<- c(NA, temp$Invoice.Date[-nrow(temp)])
-temp$prev.JN<- c(NA, temp$Job.Number[-nrow(temp)])
-temp$prev.inv<- c(NA, temp$Inv.Amount[-nrow(temp)])
-#new column which prints 'NA" if prev job.number is not the same as current OR if prevInv <=0
-temp$prev.Invoice<- ifelse(temp$Job.Number == temp$prev.JN & temp$prev.inv>0 & temp$Inv.Amount>0, temp$testp.Invoice, NA)
-temp<- temp %>% select(-testp.Invoice, -prev.JN, -prev.inv)
-temp$Invoice.Date<- as.Date(temp$Invoice.Date)
-temp$prev.Invoice<- as.Date(temp$prev.Invoice)
-
-#create new column for milestone prefix
-temp1<- temp[1:7680,]
[…]

### 2016-12-13 monthly_ggvis/server.R

@@ -1,72 +1,0 @@
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shiny')
-
-
-# month.a<- read.csv('monthly.csv')[,-1]
-# month.a$Date<- as.Date(month.a$Date)
-# month.a$return.cut <- cut_number(month.a$return.mon,15)
-levels(month.a$return.cut) =c("< -$0.51", "-$0.51 to -$0.24", "-$0.24 to -$0.02","-$0.02 to $0.18",
-                              "$0.18 to $0.33", "$0.33 to $0.45", "$0.45 to $0.50", "$0.50 to $0.61",
-                              "$0.61 to $0.75", "$0.75 to $0.92", "$0.92 to $1.20", "$1.20 to $1.68", 
-                              "$1.68 to $2.47", "$2.47 to $5.24", "> $5.24")
-
-#shiny function
-
-shinyServer(
-        function(input,output){
-                
-                # Compute the forumla text in a reactive function since it is 
-                # shared by the output$caption and output$mpgPlot functions
-                formulaText<- reactive({
-                        paste("Monthly Summary", input$disc, sep=" ")
-                })
-                
-                # Return the formula text for printing as a caption
-                output$caption <- renderText( {
-                        formulaText()
-                })
-                
-                
-                #check for the input variable
-                data.r= reactive ({
-                        if(input$disc == 'All') {
-                                posData<- month.a}
-                        
-                        else{
-                                posData<- month.a %>% filter(Discipline == input$disc)
-                        }
-                        
-                })
-        
-                
-                #subset data based on slider inputs for Date.
-
-                data.r2 = reactive({
-                        a = subset(data.r(), Date >= input$date[1] & Date <= input$date[2])
-                        return(a)
-                })
-
-                
-                #generate a plot of requested variable
-                
-                
-                gv<- reactive( {
-
-                                               
-                        dd= data.r2()
-                        
[…]

### 2016-12-13 monthly_review rCharts/server.R

@@ -1,57 +1,0 @@
-
-
-
-shinyServer(function(input, output) {
-        
-        # Compute the forumla text in a reactive function since it is 
-        # shared by the output$caption and output$mpgPlot functions
-        formulaText1<- reactive({
-                paste("Monthly Summary for", input$disc, sep=" ")
-        })
-        
-        # Return the formula text for printing as a caption
-        output$caption <- renderText( {
-                formulaText1()
-        })
-        
-        formulaText <- reactive({
-                paste(input$y, "~", "return.cut")
-        })
-        
-        #check for the input variable
-        data.r= reactive ({
-                if(input$disc == 'All') {
-                        posData<- month.a}
-                
-                else{
-                        posData<- month.a %>% filter(Discipline == input$disc)
-                }
-                return(posData)
-                
-        })
-        
-        #subset data based on slider inputs for Date.
-        
-        data.r2 = reactive({
-                a = subset(data.r(), Date >= input$date[1] & Date <= input$date[2])
-                return(a)
-        })
-
-        
-        output$myChart <- renderChart({
-
-                p1 <- nPlot(as.formula(formulaText()), group = input$variable, data = data.r2(),
-                            type = 'multiBarChart')
-                
-                p1$addParams(dom = 'myChart')
-                p1$xAxis( rotateLabels = -45, axisLabel = "Return per Dollar (Balance/Cost)")
-                p1$chart( xDomain = levels(month.a$return.cut),
-                          reduceXTicks = FALSE)
-                p1$params$width <- 800
-                p1$params$height <- 350
-                
-
-                return(p1)
-                
-        })
-})

### 2016-12-13 monthly_review rCharts/ui.R

@@ -1,54 +1,0 @@
-
-library('rHighcharts',lib='C:/Progra~1/R/R-3.2.0/library')
-
-shinyUI(fluidPage(
-        titlePanel("Monthly Review"),
-        
-        fluidRow(
-                column(3,
-                selectInput("disc", "Discipline:",
-                            list("Civil" = "Civil", 
-                                 "Structural" = "Structural", 
-                                 "Environmental Planning" = "Environmental Planning",
-                                 "Water" = "Water",
-                                 "All"= "All")),
-                
-                selectInput("variable", "Fill Variable:",
-                            list('Client Sector' = 'Biz.type',
-                                 "Job Number Code"= 'code.jobnum',
-                                 "Total Timespan" = "timespan.cut", 
-                                 "Client Code" = "code.client",
-                                 "Director Code" = "code.director",
-                                 "Project Engineer Code"= "code.ProjEng",
-                                 "Billing Type"= "Billing.Type",
-                                 "Discipline"= "Discipline",
-                                 "Primary Job Type"= 'JD.Primary',
-                                 "Client Count"= 'clientcount.cut',
-                                 "Business"= 'Business',
-                                 "Client no. Employees" ='clientemps.cut',
-                                 'Client Size' = 'Biz.size',
-                                 'Secondary Job Type'='JD.Second',
-                                 'Job Type'='Type')),
-                
-                selectInput("y", "Y Axis:",
-                            list("Invoiced Amount ($)" = "Invoiced.Amount", 
-                                 "Balance ($)" = "bal.mon")),
-                
-                dateRangeInput('date', 'Date', start = '2008-12-31', end = '2009-01-31', min = '2008-12-31', max = '2014-12-15', 
-                               format = "dd-mm-yyyy", startview = "month"),
-                
-                submitButton(text="Update")
-                ),
-                
-                column(9,
-                       
-                h3(textOutput("caption")),
-                
-                showOutput("myChart", "nvd3")
-#                 HTML('<style>.rChart {width: 100%; height: 600px}</style>')
-                
-        )
-)))
-
-
-

### 2016-12-13 monthly_review/server.R

@@ -1,79 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shiny')
-
-
-month.a<- read.csv('monthly.csv')[,-1]
-month.a$Date<- as.Date(month.a$Date)
-month.a$return.cut <- cut_number(month.a$return.mon,15)
-levels(month.a$return.cut) =c("< -$0.51", "-$0.51 to -$0.24", "-$0.24 to -$0.02","-$0.02 to $0.18",
-                              "$0.18 to $0.33", "$0.33 to $0.45", "$0.45 to $0.50", "$0.50 to $0.61",
-                              "$0.61 to $0.75", "$0.75 to $0.92", "$0.92 to $1.20", "$1.20 to $1.68", 
-                              "$1.68 to $2.47", "$2.47 to $5.24", "> $5.24")
-
-#shiny function
-
-shinyServer(
-        function(input,output){
-                
-                # Compute the forumla text in a reactive function since it is 
-                # shared by the output$caption and output$mpgPlot functions
-                formulaText<- reactive({
-                        paste("Monthly Summary", input$disc, sep=" ")
-                })
-                
-                # Return the formula text for printing as a caption
-                output$caption <- renderText( {
-                        formulaText()
-                })
-                
-                
-                #check for the input variable
-                data.r= reactive ({
-                        if(input$disc == 'All') {
-                                posData<- month.a}
-                        
-                        else{
-                                posData<- month.a %>% filter(Discipline == input$disc)
-                        }
-                        
-                })
-        
-                
-                #subset data based on slider inputs for Date.
-
-                data.r2 = reactive({
-                        a = subset(data.r(), Date >= input$date[1] & Date <= input$date[2])
-                        return(a)
-                })
-
-                
-                #generate a plot of requested variable
-                
-                
-                output$pos.plot<- renderPlot( {
-
-                                               
-                        dd= data.r2()
-                        
[…]

### 2016-12-13 monthly_review/ui.R

@@ -1,62 +1,0 @@
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shiny')
-
-shinyUI(pageWithSidebar(
-        #theme="amelia.css",
-        
-        #Application title
-        headerPanel("Monthly Summary"),
-        
-        # Sidebar with controls to select the variable to plot against positions
-        sidebarPanel(
-                
-                selectInput("disc", "Discipline:",
-                            list("Civil" = "Civil", 
-                                 "Structural" = "Structural", 
-                                 "Environmental Planning" = "Environmental Planning",
-                                 "Water" = "Water",
-                                 "All"= "All")),
-                
-                selectInput("variable", "Fill Variable:",
-                            list('Client Sector' = 'Biz.type',
-                                 "Job Number Code"= 'code.jobnum',
-                                 "Total Timespan" = "timespan.cut", 
-                                 "Client Code" = "code.client",
-                                 "Director Code" = "code.director",
-                                 "Project Engineer Code"= "code.ProjEng",
-                                 "Billing Type"= "Billing.Type",
-                                 "Discipline"= "Discipline",
-                                 "Primary Job Type"= 'JD.Primary',
-                                 "Client Count"= 'clientcount.cut',
-                                 "Business"= 'Business',
-                                 "Client no. Employees" ='clientemps.cut',
-                                 'Client Size' = 'Biz.size',
-                                 'Secondary Job Type'='JD.Second',
-                                 'Job Type'='Type')),
-                
-                dateRangeInput('date', 'Date', start = '2008-12-31', end = '2009-01-31', min = '2008-12-31', max = '2014-12-15', 
-                               format = "dd-mm-yyyy", startview = "month"),
-                
-                selectInput("y", "Y Axis:",
-                            list("Invoiced Amount" = "Invoiced.Amount", 
-                                 "Balance (invoiced-costs)" = "bal.mon")),
-                
-                submitButton(text="Update")
-                
-        ),
-        
[…]

### 2016-12-13 monthlyreviewplotly/server.R

@@ -1,106 +1,0 @@
-# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
-library('ggplot2')
-library("dplyr")
-library("plyr")
-library('magrittr')
-library('reshape2')
-library('shiny')
-
-
-
-# month.a<- read.csv('monthly.csv')
-month.a$Date<- as.Date(month.a$Date)
-
-levels(month.a$return.cut) =c("< -&#xfe69;0.51", "-&#xfe69;0.51 to -&#xfe69;0.24", "-&#xfe69;0.24 to -&#xfe69;0.02","-&#xfe69;0.02 to &#xfe69;0.18",
-                              "&#xfe69;0.18 to &#xfe69;0.33", "&#xfe69;0.33 to &#xfe69;0.45", "&#xfe69;0.45 to &#xfe69;0.50", "&#xfe69;0.50 to &#xfe69;0.61",
-                              "&#xfe69;0.61 to &#xfe69;0.75", "&#xfe69;0.75 to &#xfe69;0.92", "&#xfe69;0.92 to &#xfe69;1.20", "&#xfe69;1.20 to &#xfe69;1.68", 
-                              "&#xfe69;1.68 to &#xfe69;2.47", "&#xfe69;2.47 to &#xfe69;5.24", "> &#xfe69;5.24")
-#order month.a according to return.cut levels
-month.a<- month.a %>% arrange(return.cut)
-
-
-
-
-#                         gg$layout<- list(fileopt='overwrite', filename='R-Cookbook/axes/reversed ordered axes - 1')
-
-#shiny function
-
-shinyServer(
-        function(input,output, session){
-                
-                # Compute the forumla text in a reactive function since it is 
-                # shared by the output$caption and output$mpgPlot functions
-                formulaText<- reactive({
-                        paste("Monthly Summary", input$disc, sep=" ")
-                })
-                
-                # Return the formula text for printing as a caption
-                output$caption <- renderText( {
-                        formulaText()
-                })
-                
-                #check for the input variable
-                data.r= reactive ({
-                        if(input$disc == 'All') {
-                                posData<- month.a}
-                        
-                        else{
-                                posData<- month.a %>% filter(Discipline == input$disc)
-                        }
-                        
-                })
-        
-                #subset data based on slider inputs for Date.
-
-                data.r2 = reactive({
-                        a = subset(data.r(), Date >= input$date[1] & Date <= input$date[2])
-                        return(a)
-                })
-
-                #generate a plot of requested variable
-                
-                
[…]

### 2016-12-13 monthlyreviewplotly/ui.R

@@ -1,64 +1,0 @@
-# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-# library('shiny')
-
-
-
-shinyUI(pageWithSidebar(
-        #theme="amelia.css",
-        
-        #Application title
-        headerPanel("Monthly Summary"),
-        
-        # Sidebar with controls to select the variable to plot against positions
-        sidebarPanel(
-                
-                selectInput("disc", "Discipline:",
-                            list("Civil" = "Civil", 
-                                 "Structural" = "Structural", 
-                                 "Environmental Planning" = "Environmental Planning",
-                                 "Water" = "Water",
-                                 "All"= "All")),
-                
-                selectInput("variable", "Fill Variable:",
-                            list('Client Sector' = 'Biz.type',
-                                 "Job Number Code"= 'code.jobnum',
-                                 "Total Timespan" = "timespan.cut", 
-                                 "Client Code" = "code.client",
-                                 "Director Code" = "code.director",
-                                 "Project Engineer Code"= "code.ProjEng",
-                                 "Billing Type"= "Billing.Type",
-                                 "Discipline"= "Discipline",
-                                 "Primary Job Type"= 'JD.Primary',
-                                 "Client Count"= 'clientcount.cut',
-                                 "Business"= 'Business',
-                                 "Client no. Employees" ='clientemps.cut',
-                                 'Client Size' = 'Biz.size',
-                                 'Secondary Job Type'='JD.Second',
-                                 'Job Type'='Type')),
-                
-                dateRangeInput('date', 'Date', start = '2008-12-31', end = '2009-01-31', min = '2008-12-31', max = '2014-12-15', 
-                               format = "dd-mm-yyyy", startview = "month"),
-                
-                selectInput("y", "Y Axis:",
-                            list("Invoiced Amount" = "Invoiced.Amount", 
-                                 "Balance (invoiced-costs)" = "bal.mon")),
-                
-                submitButton(text="Update")
-                
-        ),
-        
[…]

### 2016-12-13 percent.pos/ui.R

@@ -1,35 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library('shiny')
-
-shinyUI(pageWithSidebar(
-        #theme="amelia.css",
-        
-        #Application title
-        headerPanel("Team Ratios"),
-        
-        # Sidebar with controls to select the variable to plot against positions
-        sidebarPanel(
-                selectInput("disc", "Discipline:",
-                            list("Civil" = "Civil", 
-                                 "Structural" = "Structural", 
-                                 "Environmental Planning" = "Environmental Planning",
-                                 "Water" = "Water",
-                                 "All"= "All"))
-                
-        ),
-        
-        #main panel. show plot of each position and percentage within selected variable
-        mainPanel(
-                h3(textOutput("caption")),
-                
-                plotOutput("pos.plot")
-                
-        )
-        
-                          
-))

### 2016-12-13 postcodes.R

@@ -1,365 +1,0 @@
-#### create file with distances from Brisbane city using latitude/longitude coords
-
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('fields',lib='C:/Progra~1/R/R-3.1.3/library')
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-all3<- read.csv('all3.csv')
-coords<- read.csv('postcodes.csv')
-
-#merge all3 postcodes with coords
-#create temp file with only post codes and suburbs and states from all3
-str(all3)
-temp<- all3[, names(all3) %in% c('Post.Code', 'Suburb', 'State')] %>% head()
-temp$Post.Code<- as.integer(as.character(temp$Post.Code))
-#very confusing because so many suburb names match the post code
-#need to merge original databases using postcode AND suburb name
-#therefore need suburb names to be in capital letters. use toupper()
-temp$Suburb<- toupper(temp$Suburb)
-temp$Suburb<- as.factor(temp$Suburb)
-#temp<- temp[temp$Post.Code>0,]
-#create temp file with only post codes and suburbs and states from coords, and coords
-temp2<- coords[, names(coords) %in% c('postcode', 'suburb', 'state', 'lat', 'lon')]
-head(temp2)
-#merge the two
-PC<- merge(temp, temp2, by.x= c('Post.Code', 'Suburb'), by.y= c('postcode','suburb'), all.x=TRUE)
-PC[400:500,]
-#cross check with all3 
-all3[all3$Suburb %in% 'Lismore', names(all3) %in% c('Suburb', 'Post.Code')]
-#looks fine
-#delete rows with 3 NA's in a single row AND Post.Code = 0
-PC<-PC[!(PC$Post.Code == 0 & rowSums(is.na(PC[,c(2,4:6)]))==4),]
-#any rows of full NA left?
-PC<- PC[!(rowSums(is.na(PC))==6),]
-ggplot(PC, aes(x=Post.Code)) + geom_point(stat='bin')
-# yes, deleted them
-#want unique cases of each suburb
-PC<- unique(PC)
-# want to trim down to first entry for each post code number
-#create a function
-
[…]

### 2016-12-13 postcodes4c.R

@@ -1,337 +1,0 @@
-#### create file with distances from Brisbane city using latitude/longitude coords
-
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('fields',lib='C:/Progra~1/R/R-3.1.3/library')
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-all4c<- read.csv('all4c.csv')[,-1]
-coords<- read.csv('postcodes.csv')
-
-#merge all3 postcodes with coords
-#create temp file with only post codes and suburbs and states from all3
-temp<- all4c[, names(all4c) %in% c('Post.Code', 'Suburb', 'State')]
-temp$Post.Code<- as.integer(as.character(temp$Post.Code))
-#very confusing because so many suburb names match the post code
-#need to merge original databases using postcode AND suburb name
-#therefore need suburb names to be in capital letters. use toupper()
-temp$Suburb<- toupper(temp$Suburb)
-temp$Suburb<- as.factor(temp$Suburb)
-#temp<- temp[temp$Post.Code>0,]
-#create temp file with only post codes and suburbs and states from coords, and coords
-temp2<- coords[, names(coords) %in% c('postcode', 'suburb', 'state', 'lat', 'lon')]
-head(temp2)
-#merge the two
-PC<- merge(temp, temp2, by.x= c('Post.Code', 'Suburb'), by.y= c('postcode','suburb'), all.x=TRUE)
-PC[400:500,]
-#cross check with all3 
-all4c[all4c$Suburb %in% 'LISMORE', names(all4c) %in% c('Suburb', 'Post.Code')]
-#looks fine
-#delete rows with 3 NA's in a single row AND Post.Code = 0
-PC<-PC[!(PC$Post.Code == 0 & rowSums(is.na(PC[,c(2,4:6)]))==4),]
-#any rows of full NA left?
-PC<- PC[!(rowSums(is.na(PC))==6),]
-ggplot(PC, aes(x=Post.Code)) + geom_point(stat='bin')
-# yes, deleted them
-#want unique cases of each suburb
-PC<- unique(PC)
-# want to trim down to first entry for each post code number
-#create a function
-
-skim<- function(df= PC, col=PC$Post.Code, n=1){
-        ref= unique(col)
[…]

### 2016-12-13 radarplot.R

@@ -1,119 +1,0 @@
-library('plotrix', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-
-all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
-all7d$Post.Code<- as.factor(all7d$Post.Code)
-
-#lets look at case 960
-#most important variables were:
-# no.users, inv.mlsto.log, pc.pro, timespan, majority.pos, Discipline
-#make a radar chart with no.users, inv.mlsto, pc.pro, client.code, majority.pos
-
-#create a new column for profitable or not
-
-all7d$good<- ifelse(all7d$return.pdol<0, 'bad','good')
-all7d$good<- as.factor(all7d$good)
-
-#create invoiced amount categories
-all7d$Inv.cat<- rep(NA, nrow(all7d))
-all7d$Inv.cat<- ifelse(all7d$inv.mlsto <5000 , '<$5k', all7d$Inv.cat)
-all7d$Inv.cat<- ifelse(all7d$inv.mlsto <20000 & all7d$inv.mlsto >=5000, '$5-20k', all7d$Inv.cat)
-all7d$Inv.cat<- ifelse(all7d$inv.mlsto <60000 & all7d$inv.mlsto >=20000, '$20-60k', all7d$Inv.cat)
-all7d$Inv.cat<- ifelse(all7d$inv.mlsto >=60000 , '>$60k', all7d$Inv.cat)
-all7d$Inv.cat<- as.factor(all7d$Inv.cat)
-all7d <- within(all7d, 
-                Inv.cat <- factor(Inv.cat,levels=
-                                          c('<$5k', '$5-20k', '$20-60k', '>$60k')))
-
-#create pc.pro categories, use hierarchical clustering to find groups
-#Ward's clustering
-
-x<- dist(all7d %>% select(pc.pro), method = 'euclidean')
-fit<- hclust(x, method = 'ward.D2')
-plot(fit)
-
[…]

### 2016-12-13 rising_stars/server.R

@@ -1,107 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library('shiny')
-library('rCharts',lib='C:/Progra~1/R/R-3.2.0/library')
-
-#load dataset
-all6<- read.csv('all6.csv')[,-1]
-all6$Start.Date<- as.Date(all6$Start.Date)
-
-
-#rising stars mplot rCharts
-dup<- all6[,names(all6) %in% c('mlsto', 'Discipline')]
-sub11<-all6[!all6$Tot.Invoiced==0 & !duplicated(dup) & all6$client.count>1,]
-
-# need slope column
-mods<- dlply(sub11, .(code.client), lm, formula= return.pdol ~ Start.Date)
-coefs <- ldply(mods, coef)
-colnames(coefs)[names(coefs) %in% 'Start.Date']<-'Risestar.coef'
-#coefs %>% arrange(-Start.Date) %>% filter( !is.na(Start.Date)) %>% tail()
-sub11a<- merge(sub11, coefs %>% select(code.client, Risestar.coef), by='code.client', all.x=TRUE)
-sub11a<- sub11a %>% filter(!is.na(Risestar.coef)) %>% arrange(-Risestar.coef)
-#reorder levels of code.client descending according to Risestar.coef
-sub11a <- within(sub11a, 
-                 code.client <- factor(code.client, 
-                                       levels= sub11a[order(-sub11a$Risestar.coef),]$code.client %>% unique()))
-
-
-
-#shiny function
-pagerange = list(c(1,6),c(7,12),c(13,18))
-
-shinyServer(
-        function(input,output){
-                
-                # Compute the forumla text in a reactive function since it is 
-                # shared by the output$caption and output$mpgPlot functions
-                formulaText<- reactive({
-                        paste("Rising performers in", input$business, sep=" ")
-                })
-                
-                # Return the formula text for printing as a caption
-                output$caption <- renderText( {
-                        formulaText()
-                })
-                
-                #check for the input variable
-                data.r= reactive ({
-                        if(input$business == 'All') {
[…]

### 2016-12-13 rising_stars/ui.R

@@ -1,70 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library('shiny')
-
-shinyUI(pageWithSidebar(
-        #theme="amelia.css",
-        
-        #Application title
-        headerPanel("Rising Stars"),
-        
-        # Sidebar with controls to select the variable to plot against positions
-        sidebarPanel(
-                selectInput("business", "Client Industry:",
-                            list("Architect" = "architect", 
-                                 "Artist" = "artist", 
-                                 "Builder" = "builder",
-                                 "Building Services"= 'building services',
-                                 "Business" = "business",
-                                 "Council" = 'council',
-                                 "Developer/Real Estate"= 'developer/real estate',
-                                 "Engineer"= 'engineer',
-                                 "Environment" = 'environment',
-                                 "Friend" = 'friend',
-                                 "Government"='gov',
-                                 "Health"= 'health',
-                                 "Hospital" = 'hospital',
-                                 "Institution" = 'institution',
-                                 "Interior Design"= 'interior design',
-                                 "Internal" = 'internal',
-                                 "Landscape Architect" = 'landscape arch',
-                                 "Solicitor"= 'lawyer',
-                                 "Manufacturer/Supplier"='manufacturer/supplier',
-                                 'Membrane Fabricator'= 'membrane fabricator',
-                                 "Person" ='person',
-                                 "Resources"= 'resources',
-                                 "Roads/Rail"= 'roads/rail',
-                                 "School"='school',
-                                 "signs"='signs',
-                                 "Town Planner"= 'town planner',
-                                 "University"= 'university',
-                                 'Utilities Provider'= 'utilities provider',
-                                 "Water"= 'water',
-                                 "All"= "All")),
-                
-                dateRangeInput('date', 'Date', start = '2003-10-01', end = NULL, min = '2003-10-01', max = '2014-12-15', 
-                               format = "dd-mm-yyyy", startview = "month"),
-                
-                selectInput("teehee", "Page:",
-                            list("1"= 1,
-                                 "2"= 2,
[…]

### 2016-12-13 team_ratio-pcXaxis/server.R

@@ -1,83 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library('shiny')
-
-#load Data
-all6<- read.csv('all6.csv')[,-1]
-all6$Start.Date<- as.Date(all6$Start.Date)
-dup1<- all6[,names(all6) %in% c('mlsto')]
-sub10<-all6[!all6$Tot.Invoiced==0 & !duplicated(dup1),]
-sub10<- melt(sub10 %>% select(code.jobnum, pc.contracttech, pc.director, pc.gradpro, pc.midpro, pc.midtech, pc.seniorpro,
-                              pc.seniortech, pc.unknown),value.name='percent.hrs') %>% filter(percent.hrs>.5)
-sub10 <- merge(sub10, all6 %>% select(Start.Date, balance.mlsto, Discipline, code.jobnum, return.pdol), by= 'code.jobnum', all.x=TRUE, all.y=FALSE) %>%
-        arrange(Start.Date)
-sub10$Start.Date<- as.Date(sub10$Start.Date)
-sub10$return.cut<- cut_number(sub10$return.pdol, 10, labels= c(1:10)) %>% as.integer()
-
-#shiny function
-
-shinyServer(
-        function(input,output){
-                
-                # Compute the forumla text in a reactive function since it is 
-                # shared by the output$caption and output$mpgPlot functions
-                formulaText<- reactive({
-                        paste("Position percentages in", input$disc, sep=" ")
-                })
-                
-                # Return the formula text for printing as a caption
-                output$caption <- renderText( {
-                        formulaText()
-                })
-                
-                #check for the input variable
-                data.r= reactive ({
-                        if(input$disc == 'All') {
-                                posData<- sub10}
-                        
-                        else{
-                                posData<- sub10 %>% filter(Discipline == input$disc)
-                        }
-                        
-                })
-        
-                
-                #subset data based on slider inputs for Date.
-
-                data.r2 = reactive({
-                        a = subset(data.r(), Start.Date >= input$date[1] & Start.Date <= input$date[2])
[…]

### 2016-12-13 team_ratio-pcXaxis/ui.R

@@ -1,40 +1,0 @@
-setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
-library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
-library('shiny')
-
-shinyUI(pageWithSidebar(
-        #theme="amelia.css",
-        
-        #Application title
-        headerPanel("Team Ratios"),
-        
-        # Sidebar with controls to select the variable to plot against positions
-        sidebarPanel(
-                selectInput("disc", "Discipline:",
-                            list("Civil" = "Civil", 
-                                 "Structural" = "Structural", 
-                                 "Environmental Planning" = "Environmental Planning",
-                                 "Water" = "Water",
-                                 "All"= "All")),
-                
-                dateRangeInput('date', 'Date', start = '2003-10-01', end = NULL, min = '2003-10-01', max = '2014-12-15', 
-                               format = "dd-mm-yyyy", startview = "month"),
-                
-                submitButton(text="Update")
-                
-        ),
-        
-        #main panel. show plot of each position and percentage within selected variable
-        mainPanel(
-                h3(textOutput("caption")),
-                
-                plotOutput("pos.plot")
-                
-        )
-        
-                          
-))

### 2016-12-13 SS/Regression.rmd
reset

--- 2015-10-07 Regression.rmd
@@
 library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('caret',lib='C:/Progra~1/R/R-3.2.1/library')
 # library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
@@
                          client.totinv,code.director,Num.disc,Business,Biz.type,
-                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,
-                         majority.pos,
+                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,majority.pos,
                          pc.majpos)
@@
 
-RMSE(ans$answer, ans$full.lm.fit)
+RMSE(ans$answer, ans$full.lm)
 
@@
 
-RMSE is 0.4768 which isn't great in terms of predicting return per dollar.
-R^2 is 0.1905
-
-Let's try deleting some full variables
-
-```{r, echo=FALSE}
-full.2<- update(full, .~. -client.totinv.log)
-summary(full.2)
-full.3<- update(full.2, .~. -code.director)
-summary(full.3)
-full.4<- update(full.3, .~. -Num.disc)
-summary(full.4)
-full.5<- update(full.4, .~. -code.ProjEng)
-summary(full.5)
-full.6<- update(full.5, .~. -Biz.type + Discipline)
-summary(full.6)
-
-#RMSE check
-
-ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full.6, all7d, interval= 'predict'))
-
-RMSE(ans$answer, ans$full.lm.fit)
-
-```
-
-Feels better. That leaves us with
-
-* inv.mlsto.log
-* Discipline
-* timespan.cbrt
-* no.users
-* pc.pro
-* Discipline
-* Business
-
-RMSE = 0.48266
-
-Going back to full variables, lets try pairwise multiplying..
-
-```{r, echo=FALSE}
-#try diff pairs
-
-pairs.lm<- lm(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
-                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
-              data=all7d)
-
-summary(pairs.lm)
-pairs<- predict(pairs.lm, all7d)
-ans<- data.frame(answer= all7d$return.pdol, pairs.lm = pairs)
-
-RMSE(ans$answer, ans$pairs.lm)
-
-#down to 0.4759
-
-```
+RMSE is 0.53 which isn't great in terms of predicting return per dollar.
+R^2 is 0.2152
 
@@
 ```{r, echo=FALSE}
-#try adding code.client
+#non full variables
+#Post.Code
+PostCode<- lm(return.pdol.cbrt ~ Post.Code, data=all7d)
+summary(PostCode)
+plot(Post.Code ~ return.pdol.cbrt, data=all7d)
+abline(PostCode)
 
-addone<- function(var= 'JD.Second', df= all7d){
-        reduced = df[!(is.na(df[,var])),]
-        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business", var, sep=" + ")
-        add.one = lm(as.formula(formula), data = reduced)
+#write function
 
-        print(summary(add.one))
+col.lm<- function(vars= 'Post.Code', df=all7d, ret= 'return.pdol.cbrt'){
+        #loop through and create lm model's
+        names= NULL
 
-        ans= data.frame(answer= reduced$return.pdol, addone.lm = predict(add.one, reduced, interval= 'predict'))
+        for(i in 1:length(vars)){
+                formula = paste(ret, "~", vars[i], sep = " ")
+                linear = lm(as.formula(formula), data = df)
+                assign(paste(vars[i],"lm",sep="."), linear, envir=.GlobalEnv)
 
-        print(head(ans))
+                names = rbind(names, paste(vars[i],"lm",sep="."))
 
-        print(
-        paste("RMSE=", RMSE(ans$answer, ans$addone.lm.fit) %>% round(4), sep= " ")
-        )
-
+                cat(i," ")
         }
 
-#without additional variables, RMSE sits at 0.4759
+        names = names %>% as.vector
+}
 
-addone(var='JD.Second', df=all7d)
-#RMSE = .4709
-addone(var='Post.Code', df=all7d)
-#RMSE = .4353
-addone(var='majority.pos', df=all7d)
-#RMSE = .4541
-addone(var='pc.majpos.log', df=all7d)
-#RMSE = .4553
-addone(var='code.client', df=all7d)
-#RMSE = .4394
-addone(var='Billing.Type', df=all7d)
-#RMSE = .4579
-addone(var='code.contact', df=all7d)
-#RMSE = .3972
+#now put all incomplete variables through col.lm function
+varis= c('Post.Code', 'majority.pos','pc.majpos.log','code.client', 'JD.Second',
+        'Billing.Type','code.contact','Project.Value.log')
 
-#dont use post code ever. 
-#JD.Second doesn't help that much. acutally makes RMSE worse.
+lm.vec<- col.lm(vars= varis, df=all7d, ret= 'return.pdol.cbrt')
+
+#compile results of all lm's into ans dataframe
+
+ans.complete<- function(lms = 'Post.Code.lm', df=all7d, ans.df= ans){
+        ans.df=ans
+        for(i in 1:length(lms)){
+                column = predict(get(lms[i]), df)
+                #transform from cube root
+                column = column^3
+                ans.df = cbind(ans.df, column)
+
+                #rename column
+                colnames(ans.df)[i+2]<-lms[i]
+
+                cat(i," ")
+        }
+
+        ans.df
+}
+
+ans.full<- ans.complete(lms=lm.vec, df=all7d, ans.df=ans)
+
+# RMSE(ans$answer, ans$majority.pos.lm) --  doesn't work, need to adjust RMSE function
 
@@
 
-Delete Post.Code option
+So now we have our predictions for the full variables as well as a column for each of the incomplete variables.
 
-Billing.Type should be paired with inv.mlsto and maybe Discipline
+We want to recombine them using linear regression again!
 
-maybe client should be paired with inv.mlsto
-maybe JD.Second should be paired with inv.mlsto and code.client and timespan and pc.pro and no.users?? which ever one does something.
+For a check, I will find all cases which have the full variables as well as client.code, code.contact, and Billing.Type
 
-majority.pos pair with inv.mlsto
-pc.majpos pair with inv.mlsto
+```{r, echo=FALSE}
+#reduce ans data frame to only complete cases with variables you are interested in
+ans.red = ans.full %>% select(answer, full.lm, code.client.lm, Billing.Type.lm, code.contact.lm)
+ans.red <- ans.red[complete.cases(ans.red),]
 
-Write a function that lists the extra variable you're trying and write out the interaction.
+#run final lm
+final.lm<- lm(answer~. , data=ans.red)
+final<- predict(final.lm, ans.red)
+ans.red$final.lm<- final
+
+RMSE(ans.red$answer, ans.red$final.lm)
+
+ans.red %>% View
+
+```
+
+Wouldnt you be better off just running a regression with the complete cases?
 
@@
 
-addmult<- function(var= c('JD.Second', 'code.client'), combos= c('JD.Second*inv.mlsto.log'), df= all7d){
-        reduced = df[complete.cases(df[,var]),]
-        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business")
+reduced<- all7d %>% select(inv.mlsto.log , timespan.cbrt , no.users , Discipline , pc.pro , client.totinv.log , 
+                                   code.director , Num.disc , Business , Biz.type , code.ProjEng, Billing.Type,
+                           code.client, code.contact, return.pdol.cbrt)
+reduced<- reduced[complete.cases(reduced),]
 
-        for(i in 1:length(var)){
-                formula = paste(formula, var[i], sep= " + ")
-        }
+test.lm<- lm(return.pdol.cbrt ~ ., data= reduced)
+test<- predict(test.lm, reduced)
+ans.red$test.lm <- test
 
-        for(i in 1:length(combos)){
-                formula = paste(formula, combos[i], sep= " + ")
-        }
-
-        add.one = lm(as.formula(formula), data = reduced)
-
-        print(summary(add.one))
-
-        ans= data.frame(answer= reduced$return.pdol, addone.lm = predict(add.one, reduced, interval= 'predict'))
-
-        print(head(ans))
-
-        print(
-        paste("RMSE=", RMSE(ans$answer, ans$addone.lm.fit) %>% round(4), sep= " ")
-        )
-
-}
-
-#try Billing.Type with inv.mlsto.log
-addmult(var= c('Billing.Type'), combos= c('Billing.Type*inv.mlsto.log'), df= all7d)
-# no
-
-#try code.client*inv.mlsto.log
-addmult(var= c('code.client'), combos= c('code.client*inv.mlsto.log'), df= all7d)
-#made a difference, brought RMSE down to .4076 :)!
-
-#try JD.Second with inv.mlsto.log (no), code.client (too intense), timespan (no), pc.pro (no), no.users (yes)?
-addmult(var= c('JD.Second'), combos= c('JD.Second*no.users'), df= all7d)
-#add JD.Second*no.users
-
-addmult(var= c('majority.pos'), combos= c('majority.pos*inv.mlsto.log'), df= all7d)
-#no
-
-addmult(var= c('pc.majpos.log'), combos= c('pc.majpos.log*inv.mlsto.log'), df= all7d)
-#no
+RMSE(ans.red$answer, ans.red$test.lm)
 
 ```
-
-It is worthwhile to include:
-
-* code.client by inv.mlsto.log because brought down RMSE from 0.4394 (code.client) to 0.4076 (code.client + code.client by inv.mlsto.log)
-* JD.Second by no.users. Brought down RMSE from 0.4709 to 0.4612...
-
-Now look at cross validation for full variable set:
-
-```{r, echo=FALSE}
-
-#using caret package
-tc <- trainControl("cv",10,savePred=F)
-fit<- train(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
-                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
-              data=all7d,
-            method = 'lm', trControl = tc)
-
-#RMSE = 0.49246
-#compared to non-cross validated
-# RMSE = 0.4759
-
-#try using caret package on code.client
-fit<- train(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
-                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business + code.client,
-              data=all7d,
-            method = 'lm', trControl = tc)
-
-```
-
-Not really trying to tune anything via cross validation, however RMSE increased from 0.4759 to 0.49246.
-
-Not THAT bad.
-
-Now, how to get variable importance from lm..
-
-```{r, echo=FALSE}
-
-#variable importance for full variables only - pairs.lm
-#use varImp from caret package - uses t-test statistic
-
-imp<- varImp(pairs.lm, scale=FALSE)
-imp$variable<- rownames(imp)
-imp<- imp %>% select(variable, Overall)
-imp %>% arrange(-Overall) %>% head()
-
-sum(imp$Overall)
-
-```
-
-Trial some scenarios:
-
-Run full variables plus
-
-Make a function to output:
-
-* var importances
-* RMSE
-
-based on a list of variables that you want to use
-
-```{r, echo=FALSE}
-
-sample<- all7d[1000,]
-sample<- sample[,!is.na(sample)]
-vars<- names(sample) 
-vars<- vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc','Biz.type','Post.Code',
-               'code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto','balance.mlsto','client.totinv.log','pc.majpos')]
-vars
-
-#make new dataframe for future use with normalised variables and variables that you want to use only
-all8a<- all7d %>% select(no.users, Discipline, pc.pro, Business, code.client, JD.Second, timespan.cbrt,
-                         inv.mlsto.log,client.totinv.log, pc.majpos.log, code.contact, Billing.Type,
-                         majority.pos, return.pdol, mlsto, Year)
-write.csv(all8a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')
-
-RMSE<- function(answer, guess, print= TRUE){
-        error= answer-guess
-        root.error= sqrt(mean(error^2))
-        if(print == TRUE){
-        print(root.error)
-        }
-}
-
-final.output<- function(df=all7d, samp = 800) {
-        #create sample df
-        sample = df[samp,]
-        sample = sample[,!is.na(sample)]
-        vars= names(sample)
-        vars = vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc',
-                                  'Biz.type','Post.Code','code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto',
-                                  'dis.sc.mlsto',
-                                  'balance.mlsto','client.totinv.log','pc.majpos')]
-        variables = vars
-
-        print(
-                list("sample case", t(sample %>% select(-Year, -timespan.cbrt, -inv.mlsto.log, -client.totinv.log)))
-        )
-
-        #make reduced dataset with complete cases for all variables we want to use
-        reduced = df[-samp,]
-        reduced = reduced[complete.cases(reduced[,variables]),]
-
-        print(
-                paste(nrow(reduced), "cases in full model")
-        )
-
-        formula = paste("return.pdol ~ ")
-
-        for(i in 1:length(variables)){
-                formula = paste(formula, variables[i], sep= " + ")
-        }
-
-        if('inv.mlsto.log' %in% variables &
-           'inv.mlsto.log' %in% variables){
-                formula = paste(formula, 'inv.mlsto.log*pc.pro', sep= " + ")
-        }
-
-        if('inv.mlsto.log' %in% variables &
-           'code.client' %in% variables){
-                formula = paste(formula, 'inv.mlsto.log*code.client', sep= " + ")
-        }
-
-        if('JD.Second' %in% variables &
-           'no.users' %in% variables){
-                formula = paste(formula, 'JD.Second*no.users', sep= " + ")
-        }
-
-        final = aov(as.formula(formula), data = reduced)
-
-        fit= summary(final)
-
-        ans= data.frame(answer= reduced$return.pdol, final.lm = predict(final, reduced, interval= 'predict'))
-
-        # print(head(ans))
-
-        #print variable importances for reduced lm just run
-
-        imp= data.frame('vars'= rownames(fit[[1]]), 'F.val' = fit[[1]]$`F value`)
-        #remove white space from levels in imp$vars
-        levels(imp$vars)<- sub("\\s+$", "", levels(imp$vars))
-        #remove Residuals row
-        imp= imp[!(imp$vars %in% 'Residuals'),]
-
-        #print variable importances
-        print(
-                list("variable importances",
-                     imp %>% arrange(-F.val) %>% head())
-        )
-
-        #print RMSE for cases that match certain categories only
-        #first subset predictions to cases that match the category
-
-        reduced$predict= ans$final.lm.fit
-
-        cats= c('code.client', 'code.contact', 'JD.Second', 'Business', 'no.users', 'majority.pos')
-
-        for(i in 1:length(cats)){
-                if(cats[i] %in% vars){
-                        cats.red = reduced[reduced[,cats[i]] %in% sample[,cats[i]],]
-                        cat.RMSE = RMSE(cats.red$return.pdol, cats.red$predict) %>% round(4)
-                        print(
-                                paste(cats[i], "RMSE =", cat.RMSE, sep= " "))
-                        print(
-                                paste("Number of cases =", nrow(cats.red), sep= " "))
-                        print(
-                                head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto','return.pdol','predict',cats[i])])
-                        )
-                        }
-        }
-
-        num.cats= c('timespan.cbrt', 'inv.mlsto.log', 'pc.pro')
-
-        for(i in 1:length(num.cats)){
-                if(num.cats[i] %in% vars){
-                        span = 0.05*(max(reduced[,num.cats][i])-min(reduced[,num.cats[i]]))
-                        range = c(df[samp,num.cats[i]]-span, 
-                                  df[samp,num.cats[i]]+span)
-                        cats.red = reduced[reduced[,num.cats[i]]<range[2] & reduced[,num.cats[i]]>range[1],]
-                        cat.RMSE = RMSE(cats.red$return.pdol, cats.red$predict) %>% round(4)
-                        print(
-                                paste(num.cats[i], "RMSE =", cat.RMSE, sep= " "))
-                        print(
-                                paste("Number of cases =", nrow(cats.red), sep= " "))
-                        print(
-                                if(num.cats[i] == 'pc.pro'){
-                                        head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto','return.pdol','predict',
-                                                 num.cats[i])])} 
-                                else if(num.cats[i] == 'timespan.cbrt'){
-                                                head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto',
-                                                                 'return.pdol','predict', 'timespan')])}
-                                else {
-                                        head(cats.red[,c('mlsto','inv.mlsto','cost.mlsto',
-                                                                 'return.pdol','predict')])
-                                }
-
-                        )}
-        }
-
-        #random RMSE
-        r= mean(reduced$return.pdol)
-        random.RMSE= RMSE(reduced$return.pdol, r, print=T) %>% round(3)
-
-        print(
-        paste("Random Guess RMSE=", random.RMSE , sep= " ")
-        )
-
-        #final RMSE for all rows in reduced in a regression
-        final.RMSE= RMSE(ans$answer, ans$final.lm.fit) %>% round(3)
-
-        print(
-        paste("Overall predicted RMSE=", final.RMSE , sep= " ")
-        )
-
-        predict.rpdol = predict(final, sample) %>% round(3)
-        predict.rpdol = predict.rpdol[[1]]
-        predict.cost = sample$inv.mlsto/(1+predict.rpdol) %>% round(1)
-        print(
-               paste("return per dollar prediction", predict.rpdol, sep = " ")
-        )
-
-        print(
-               paste("answer", sample$return.pdol %>% round(3), sep = " ")
-        )
-
-        print(
-                paste("Invoiced= $", sample$inv.mlsto %>% round(2), ",   predicted cost= $", predict.cost %>% round(1), sep="")
-        )
-
-        print(
-                paste("Actual cost= $", sample$cost.mlsto %>% round(2), sep="")
-        )
-
-}
-
-final.output(df=all7d, samp = 800)
-
-```
-
-Check what RMSE would be for random guess
-
-```{r, echo=F}
-
-m<- mean(all7d$return.pdol)
-
-RMSE(all7d$return.pdol, m, print=T)
-
-```
-
-RMSE from random guess is 53 cents.
-
-# delete any code.client or contact. client which has less than 3 entries!
-# need to revisit how i got my inv.mlsto, balance.mlsto, cost.mlsto for job 2013.301.2

### 2016-12-13 boost_tree.rmd
reset

@@
                             ))
+# all9c <- read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all9c.csv')[,-1]
 
@@
 #create test train
-sample<- sample(1:nrow(all9c), 2/3*nrow(all9c), replace=F)
+sample<- sample(1:nrow(all9c), 3/4*nrow(all9c), replace=F)
 train<- all9c[sample,]
@@
 gbmGrid <- expand.grid(shrinkage = c(0.001),
-                       n.trees = c(10000), interaction.depth = c(3,4,5,6),
+                       n.trees = c(10000), interaction.depth = c(4),
                        n.minobsinnode = c(20))

### 2016-12-13 finalwriteup/Consulting Profitability Paper/Consulting_Profitability_Paper.tex
reset

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
             pdfauthor={},
-            pdftitle={Prediction of Consulting Project Profitability},
+            pdftitle={Predictive Models to Support Quoting of Fixed Fee Consulting Projects},
             colorlinks=true,
@@
 
-  \title{Prediction of Consulting Project Profitability}
+  \title{Predictive Models to Support Quoting of Fixed Fee Consulting Projects}
     \author[Queensland University of Technology]{Amy Cook, Paul Wu, Kerrie Mengersen\corref{c1}}
@@
 provide any necessary background to frame the research question.
-Concisely summarize the major findings/results. Summary of Key Related
-Research This section should include a brief summary of key related
-research. Emphasis should be on demonstrating the foundation for the
-current investigation. Specifically, the goal is to clearly delineate a
-gap or missing link that the current research fills. Authors should
-avoid presenting a litany of past research and should focus on prior
-work necessary to demonstrate the existence of the research gap
-addressed in the manuscript.
+Concisely summarize the major findings/results.
+
+Depending on the industry, some consulting projects can be financialy
+risky endeavors for the consultant. Engineering consultanting companies
+in the construction industry complete a wide range of complex projects
+in a competitive enviroment. Because of the way contracts are
+structured, many of these projects result in losses. The purpose of this
+study was to test whether these loss making projects could be predicted
+by statistical and machine learning algorithms using the information
+available at project conception. The second question is whether the
+predictive algorithm could improve a company's profitability. Project
+data from a case study engineering company was used to answer these
+questions.
 
@@
 
-Intro - pick out bits from thesis
-
 \paragraph{1.1 Problem motivation}\label{problem-motivation}
 
-\begin{itemize}
-\item
-  amount of losses, sources
-\item
-  `outside' view, that kind of thing
-\end{itemize}
+Financial risk most commonly falls upon a consulting company when they
+must offer a fixed fee for a project. In competitive climates, clients
+request fixed price quotes for project completion from several
+consultants before a project is awarded. Hence, initial quoted fees must
+be competitive but also accurately cover consultant salaries and
+business costs.
+
+How do consultants make this accurate yet competitive fee estimate? This
+varies from industry to industry, but typically a consulting manager has
+experience in the type of project they are quoting and, after reviewing
+the project details, can use a combination of intuition, comparison to
+past projects, and rules of thumb. Unfortunately, some projects priced
+using this method can blow out of budget, causing significant losses to
+the consultant.
+
+A review of 6 IT project surveys between 1984 and 1994 by Moløkken \&
+Jørgensen (2003) revealed that 60-80\% of IT projects encountered effort
+and/or schedule overruns, where the average overrun was by 30 to 40\%. A
+study on large-scale infrastructure projects over the past seventy years
+revealed that cost forecasts consistently underestimated the cost of
+rail projects by an average of 44.7\%, the cost of bridge and tunnel
+projects by 33.8\% and the cost of road projects by 20.4\% (Flyvbjerg
+2007).
+
+Kahneman won a nobel prize in 2002 for his research into decision making
+on this topic. He argued that a person's natural optimistic view of
+their own skills leads to consistent underestimation of the time and
+risks involved in a project (Lovallo and Kahneman 2003). To counter this
+phenomenon, research by Flyvbjerg {[}-Flyvbjerg2011{]} advised that
+managers should be confronted by the financial performance of similar
+projects to the one being costed so that they take on a more realistic
+view of their abilities. This was based off a body of findings where
+people made better predictions about themselves after being exposed to
+facts about other people. For example, subjects were asked to predict
+their skill levels before and after being exposed to a summary of other
+people's skills, and their predictions after were significantly more
+accurate (Lovallo and Kahneman 2003).
+
+This work aims to provide a manager with this more realistic view by
+predicting whether the new project will be profitable or not, based on a
+suite of previous project data from the same company. The manager could
+then take multiple actions including not tendering for a project
+predicted to be loss-making, providing a high fixed price, or
+restructuring the contract so that it is not 100\% fixed price.
 
@@
 
+This thesis focuses on a single case study consulting company in the
+construction industry. twelve years of passively collected data that
+described each project in terms of their clients, invoice history,
+employee hours and technical details. labour is the chief cost and the
+long traditional history of this industry mandates fixed price projects
+as the norm. 20\% of projects are loss making. Fees calculated via
+detailed analysis A Customer Relationship Management software package
+(CRM) is currently employed to collect and store project data. A CRM is
+a popular type of software used by businesses to record client and
+project details as well as employee time sheet records. In the case
+study business, the CRM is available to all employees over the company
+intranet, and each employee completes daily time sheets allotting their
+hours to certain projects. Additionally, technical information is
+recorded against each project
+
+\begin{itemize}
+\tightlist
+\item
+  Employee time sheet hours with dates
+\item
+  Other project costs (taxis, printing)
+\item
+  Client information/characteristics
+\item
+  Client identification code
+\item
+  Invoiced amounts for each project and dates
+\item
+  Employee costs
+\item
+  Employee charge out rates
+\item
+  Project description
+\end{itemize}
+
+2364 past projects are available. missing data due to how the data
+collection changed over time and some fields were not mandatory.
+
+To assess whether the model would improve the case study company's
+profitability, a simple managerial decision framework was applied: if
+the algorithm predicted the project to be loss making, the project was
+considered `rejected' and all profits or \emph{losses} of rejected
+projects were discarded. The remaining projects' profits and losses were
+summed to give a revised overall profit over the last 12 years of
+projects. This was compared to the summed profits and losses from all
+projects completed over the last 12 years. The best algorithm rejected
+enough loss-making projects to improve overall profits by 9\%.
+
+\begin{itemize}
+\tightlist
+\item
+  aim \textbf{General Aim} Use statistical techniques to model the
+  profitability of projects for consulting businesses using their
+  internal CRM data. Research will focus on a case study Engineering
+  consulting company that offers their expert advice (in the currency of
+  time) to business clients. The project outcomes are intended to assist
+  the business in predicting project profitability before engagement.
+  Several statistical and machine learning techniques were tested,
+  compared and refined.
+\end{itemize}
+
+\textbf{Hypothesis 1} A statistical or machine learning model based on
+historical project data can predict the profitability of a new project
+with greater accuracy than a baseline predictor. A baseline predictor
+for this project is one that predicts the average of a numeric response
+variable for all cases or, if the response variable is categorical,
+randomly assigned categories for each case, where the proportion of
+assigned categories matches the true categorical proportions.
+
+\textbf{Hypothesis 2} The predictive model built from Hypothesis 1 can
+be shown to have a positive impact on the overall profit earned by the
+case study business. The overall profit is represented by the following
+equation:
+
 \section{2. Literature Review}\label{literature-review}
+
+Summary of Key Related Research This section should include a brief
+summary of key related research. Emphasis should be on demonstrating the
+foundation for the current investigation. Specifically, the goal is to
+clearly delineate a gap or missing link that the current research fills.
+Authors should avoid presenting a litany of past research and should
+focus on prior work necessary to demonstrate the existence of the
+research gap addressed in the manuscript.
 
@@
 
+The bulk of research to date has been performed with project data in
+either the construction industry or software development. Research in
+the construction industry has primarily focused on predicting the final
+building cost of construction. Over the past few decades, significant
+research has been dedicated to creating predictive models that present
+the `outside view' of a project. This was done by statistically
+comparing a new project to a collection of similar projects and their
+characteristics. It has been shown that these models improve cost
+estimation accuracy, however industry has not adopted them.
+
+\subparagraph{2.1.1 Construction Industry}\label{construction-industry}
+
+Elfaki, Alatawi, and Abushandi (2014) completed a review of cost
+estimation research from 2004 to 2014. They found that artificial Neural
+Networks and SVM's were the most common machine learning techniques.
+Neural Networks and SVM's received significant attention in the 1990's
+for their ability to accurately predict construction costs with limited
+detailed information (Shin 2015; Kim, An, and Kang 2004). Shin (2015)
+pioneered the application of Boosted Trees to cost estimation in
+construction projects. Often, Linear Regression was the only model
+assessed, without comparison to other methods such as Neural Networks,
+which first started appearing in literature in the 1990's (Kim, An, and
+Kang 2004) Some studies showed that Neural Networks outperform
+Regression, however other studies established they are approximately
+equal (Kim, An, and Kang 2004; Attalla and Hegazy 2003).
+
+Variety of variables used: (Chan and Park 2005)(Akintoye and Fitzgerald
+2000; Trost and Oberlender 2003; Pinto and Slevin 1988) the studies all
+generally reported their mean absolute error or a similar metric from
+the tested models, and results were generally positive. Furthermore, the
+existing studies collected data from many businesses for a single study
+
+\subparagraph{2.1.2 Software Industry}\label{software-industry}
+
+In the software industry, the main component of cost is \emph{effort} as
+opposed to the cost of building materials in the construction industry.
+
+Similar to the construction industry, expert judgment or detailed
+analysis, is the most widely practiced method for effort estimation
+(Shepperd, Schofield, and Kitchenham 1996; Moløkken and Jørgensen 2003)
+
+Multiple studies have shown that Neural Networks definitively outperform
+regression models in effort estimation, although Regression is the most
+popular method (Finnie, Wittig, and Desharnais 1997; Pai, McFall, and
+Subramanian 2013; Matson and Mellichamp 1993).
+
+Interestingly, several studies have found that even if 15 or so
+variables are included, often only one variable contributes
+significantly to the model's accuracy: size (or funciton point)
+(Shepperd, Schofield, and Kitchenham 1996; Finnie, Wittig, and
+Desharnais 1997; Pai, McFall, and Subramanian 2013)
+
+A study by Mendes and Kitchenham (2004) demonstrated using 67 web
+projects that cross-company models were significantly less accurate than
+a within company model (Shepperd, Schofield, and Kitchenham 1996) Again,
+Neural Networks were criticised for their inability to explain their
+results (Finnie, Wittig, and Desharnais 1997). The software industry is
+unique in that agile methods of delivery are changing the contractual
+approach of consulting (Badenfelt 2011). Software development is a
+relatively new field in comparison to construction, where the tradition
+of fixed fee structures is difficult to reinvent.
+
 \paragraph{2.2 Methods used in other business
@@
 
-\section{3. Methods}\label{methods}
+The review of methods applied to the construction and IT industry
+highlighted the use of Linear Regression, Neural Networks, SVM's and in
+one case Boosted Trees, however research on other business problems
+utilised a wider range of methods. These included Naive Bayes, Random
+Forests, and machine learned Bayesian Networks.
+
+Although SVM's performed the best, as discussed, they are often complex
+and slow, requiring a great deal of memory (Kumar and Ravi 2007). Kumar
+and Ravi (2007) performed a detailed review of statistical and machine
+learning techniques that were applied over 37 years in the context of
+bankruptcy prediction in banks.
+
+For example in financial credit scoring, a study by Brown and Mues
+(2012) found that Random Forests and Boosted Trees consistently
+outperformed Neural Networks in classification. Kumar and Ravi's (2007)
+review also assessed ensemble techniques, which refers to combinations
+of two completely different algorithms, and found they can often
+outperform individual methods.
+
+An exception is Saradhi \& Palshikar's (2011) study on employee churn,
+where `churn' refers to the number of individuals moving out of a group
+within a certain time. Naive Bayes, Random forests (an ensemble decision
+tree method) and SVM's were built. determining the value of each
+employee in terms of the importance of the projects they were on and
+their monthly chargeability. This allowed them to rank employees
+identified as `high risk of churn' by value and provided a clear ranking
+for manager's to act upon. This extension of the study provided a
+comprehensive framework for how business managers could adopt their
+findings to improve business operations. It was a valuable addition that
+is absent from most cost and effort estimation research.
+
+\paragraph{2.3 Gaps}\label{gaps}
+
+In comparison to previous studies, this project advances the body of
+work on cost estimation in three ways: the data set is larger than other
+studies and sourced from a single company's CRM database, ensemble tree
+methods and model blending was tested, and the prediction results were
+integrated into a decision framework for the business. In past research,
+cost estimation data sets generally consisted of \textless{}300 of
+projects from a range of a companies and even countries. This thesis
+case study only had access to the company's internal data but this
+contained over 2,000 past jobs Furthermore, the problem of cost
+estimation is applied to a consulting company in the construction
+industry, which differs from other studies in the construction industry
+that estimate actual construction costs. Consulting in the construction
+industry can be likened more to software development effort estimation
+although the nature of the work differs considerably.
+
+ensemble tree methods and blended methods are applied which have been
+minimally tested on this problem. To date, Linear Regression, Neural
+Networks, Case Based Reasoning (CBR), and Support Vector Machines
+(SVM's) have been used to predict cost/effort even though ensemble tree
+methods can perform as well as Neural Networks and generally outperform
+Linear Regression (Caruana and Niculescu-Mizil 2006). Another benefit of
+ensemble trees is the output types available, such as partial dependency
+plots and variable importance plots that provide insight into the
+model's calculations. This contrasts Neural Networks and SVM's, which
+are `black-box' predictors.
+
+Finally, this study propels the predictive model one step further than
+other studies by analysing bottom-line profit improvements that could
+result from the research. final profit increases are evaluated. This
+kind of thought experiment is designed to present a clear case for
+industry when evaluating if and how to adopt cost estimation models.
+
+\section{3. Prediction Methods}\label{prediction-methods}
 
@@
 
+Neural Netowrks and SVM were not purusued because black box A previously
+mentioned disadvantage is that the NN algorithm is a black box because
+the internal structure is too complex for interpretation. They also
+require a lot of training data relative to other methods.
+
+Also, SVM's can be very slow to train and therefore not suitable for
+industry purposes (Auria and Moro 2008). Because it is so important to
+engage decision makers with a model that can explain its results, these
+two methods are not appropriate for the effort estimation problem.
+
 \paragraph{3.1 Predictive methods}\label{predictive-methods}
+
+binary classification problems, where the objective is to predict the
+probability of an event occurring. Probability is a continuous response
+variable, however predictions from a linear regression would not be
+bound by 0 and 1.
+
+\subparagraph{3.1.1 Logistic Regression}\label{logistic-regression}
+
+Because of the linear relationship, an equation can be fit to the log
+odds of the binary response variable against each of the explanatory
+variables using linear regression. Then, the linear equation describing
+the log odds can be transformed back to probability by taking the
+inverse log i.e.~the exponential and rearranging (Macdonald 1975):
+
+The result in a sigmoidal function bound by 0 and 1. The sigmoidal
+function originates from a linear fit of the log odds in the data.
+Assume linear relationship between covariates and response variable
+
+A coefficient represents the change in the log odds of the response
+variable for each unit increase in an explanatory variable. Therefore,
+taking the exponential of the coefficient is the change in odds of the
+response variable. If the change in odds was 2, then if the explanatory
+variable increased by 1, the response variable event would be twice as
+likely to occur.
+
+Logistic Regression is a good benchmark to compare other binary
+predictive models due to its simplicity and speed (Moore and McCabe
+1989). single linear relationship means Not prone to overfitting
+
+\subparagraph{3.1.2 Random Forests}\label{random-forests}
+
+single decision trees also suffer from low predictive accuracy and
+instability, so to combat these problems, ensemble tree methods were
+pioneered in the 1990's (Breiman 1996).
+
+The three methods of combining hundreds or thousands of trees turn
+decision trees into high performing, stable predictors.
+
+training multiple trees from the same data by sampling bootstrapped
+training sets with replacement. However, when creating each tree, a
+random subset of attributes (variables) is considered at each split. The
+reduced subset of attributes is resampled for each split in the tree.
+This allows dominant variables to be suppressed for a fraction of the
+splits, allowing the algorithm to explore signals in weaker variables
+(Breiman 2001).
+
+Additionally, ensemble tree methods are faster than SVM's and Neural
+Networks but can perform just as well and even provide insights such as
+variable importance and variable relationships (Sealfon and Gymrek
+2012).
+
+\subparagraph{3.1.3 Gradient Boosted
+Trees}\label{gradient-boosted-trees}
+
+Lastly, the boosted decision tree approach applies gradient descent
+theory to a series of decision trees. The trees are limited to a certain
+depth to maintain simplicity, and each tree models the residuals (or
+errors) of the preceding tree. By modeling the errors, misclassified
+cases are weighted higher than correctly classified cases and influence
+the structure of the latest tree more (Hastie, Tibshirani, and Friedman
+2009). This increased weighting on errors is why the algorithm is called
+boosting. The limited depth of each tree prevents overfitting at each
+stage and the combined result of up to thousands of trees is very
+powerful (Elith, Leathwick, and Hastie 2008).
+
+Caruana and Niculescu-Mizil (2006) tested Boosted Trees, Random Forests,
+Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary
+classification problems and found that Boosted Trees performed best,
+followed by Random Forests. This demonstrates how ensemble tree methods
+are capable of competing with high-level machine learning algorithms.
+
+Trees are not built on a probabilistic framework, and therefore their
+results cannot be provided in this framework (Louppe and Prettenhofer
+2014).
+
+\subparagraph{3.1.4 Naive Bayes}\label{naive-bayes}
+
+The Naive Bayes method works by making conditional independence
+assumptions about the explanatory variables in order to simplify
+probability calculations for the response variable (the response
+variable must be categorical).
+
+These equations are simple to compute given the data, and the class with
+the highest probability can then be chosen (F. Provost and Fawcett
+2013).
+
+The advantages of this method are that the conditional independence
+assumption enables very fast calculations and predictions. The method
+can perform well in real world tasks because the assumption of
+independence does not significantly damage predictions (F. Provost and
+Fawcett 2013). Assumptions of independence genearlly not true
+(correlated variables). This is fine for ranking but the output
+probabilities are not accurate statistical probabilities (Caruana and
+Niculescu-Mizil 2006).
+
+Naive Bayes method generally provides a good benchmark to compare
+against more complex models that should outperform it (Caruana and
+Niculescu-Mizil 2006).
+
+\subparagraph{3.1.5 Bayesian Networks}\label{bayesian-networks}
+
+A Bayesian Network is a graphical probabilistic model that illustrates
+the conditional dependencies between variables in a data set. The model
+is visually represented by a directed acyclic graph (DAG) and is capable
+of linking the conditional dependency between any variable to another
+variable (Heckerman 1998). The conditional relationships are Bayesian,
+where the probabilities in one node are conditional upon values in nodes
+directed towards it as well as preceding nodes
+
+Bayesian Networks reduce computations required to find the probability
+of a unique combination given other explanatory variable values while
+still taking into account many conditional dependencies that Naive Bayes
+(Barber 2012)
+
+Naive Bayes is the simplest form of a Bayesian Network (Zhang 2004)
+
+Network relationships can be learned from the data, however this is not
+widely included as part of the suite of machine learning methods. The
+conditional variable dependencies are calculated from the data, which in
+turn can define the graph structure. Some conditional dependencies can
+be set before the structure is learned (Barber 2012).
+
+Drawbacks of Bayesian Networks include the difficulty for them to
+process continuous variables.
+
+Bayesian networks have found success in combining deterministic models
+with observed data as well as expert knowledge.(Kragt 2009)
+
+\paragraph{3.2 Procedure}\label{procedure}
+
+It includes first how the data was obtained, the lengthy cleaning
+process, Projects varied from total invoiced amounts of \$500 to over
+\$1,000,000 a project could have thousands of rows of relevant data that
+needed to be converted into a single row per project. 10 engineered
+variables describing timesheets: Percent of hours performed by
+`professional' employees as opposed to `technical' employees, Position
+of the employee that completed the most hours on each project 3
+engineered from invoicing data: 4 from project data: number of projects
+completed with each client text analysis of project titles to develop
+project classifications
+
+Discretisation was performed by generating a hierarchical dendrogram of
+each variable to visualise the clusters.
+
+Limiting the predictive model to a concise set of meaningful variables
+reduces noise and improves predictions. Less variables and a simpler
+model is easier for stakeholders to understand (Weisberg 2005). followed
+by variable importance analysis, variable selection: varialbe importance
+aalyses done with linear regression, random forest, cforest to represent
+the range of algroithms that were trialled. 11 variables chosen:
+
+trials of selected predictive algorithms. binary classification problem
+predicting profit or loss. Complex models should be measured against
+simple models that can be built at a fraction of the computational cost.
 
@@
 \item
-  description of each of the methods: refer thesis
+  Regression - baseline model
+\item
+  Naive Bayes - baseline model
+\item
+  Bayesian Network - grabage.
+\item
+  Random Forest
+\item
+  Gradient Boosted Trees
 \end{itemize}
 
-\paragraph{3.2 Procedure}\label{procedure}
+To compare the models, the area under the receiver operating
+characteristic (ROC) curve (AUC) statistic was used. An ROC graph
+visualises the curve from which an AUC score is calculated. In this
+problem a `positive' is a loss making job. For binary classification,
+AUC is a more meaningful statistic than classification accuracy when the
+output is a probability that can be applied to the problem. The AUC
+indicates model performance across many thresholds while classification
+accuracy represents a single threshold.
+
+a model must be first created using a training set. The model can then
+make predictions on the test set. this is used to build the ROC A model
+that is perfectly classified would have an AUC = 1. A model that
+predicts as well as random chance would have an AUC = 0.5. An AUC
+between 0.5 and 1 means the model is performing better than random
+chance.
+
+In order to compare which models performed significantly better than
+others, an adequate sample size of results statistics is required.
+Multiple models could be made by using different data in the training
+vs.~testing sets, each providing a resulting test statistic (RMSE or
+AUC). Initially, 20 models of each method were created in this fashion.
+Then a two-sample power calculation was run using the two sets of 20
+results to determine the sample size to achieve a statistical power of
+0.8 (100 to achieve power of 0.8 between boosted trees and naive bayes)
+
+Missing Data Imputation. All methods except gradient Boosted Trees and
+Naive Bayes could not handle missing data. a complete data set allows
+for complete sets of predictions from each method, and these predictions
+could then be blended to further improve results. The MICE Random Forest
+method was chosen for imputation because it has been proven to work well
+with complex data sets (Shah et al. 2014). If similar predictive results
+were obtained using Boosted Trees imputed data and imputed data, the
+imputed data must be reasonable. The imputed data set was then trialed
+on the remaining methods and compared to unimputed trials.
+
+Once the best methods were selected, they were blended in numerous ways,
+using both simple averaging techniques and sophisticated
+machine-learning algorithms. These were compared against individual
+models and the best constructs were selected. Six blending methods,
+ranging from simple to complex, were tested using predictions from the
+top performing individual models. These included simple averaging of the
+individual model results, building a Logistic Regression model using the
+individual model results only, a Boosted Tree model using individual
+model results only, feature weighted linear stacking (FWLS), Random
+Forests, and Boosted Trees. A simplified explanation of FWLS is as a
+Linear Regression where meta-features as well as model predictions from
+individual models are included as explanatory variables. Then, each
+meta-feature is interacted with each set of model results (Sill et al.
+2009). Feature interaction is performed passively due to the nature of
+how trees are built. A split in a node determined by one variable is
+conditional upon the preceding split, which was based on another
+variable and so on. The simplest method averaged the probability output
+(a number between 0 and 1) of all three models. The next two simplest
+methods consisted of building a Logistic Regression and Boosted Trees
+model from the three prediction model outputs only. Feature Weighted
+Linear Stacking (FWLS), Random Forest and Boosted Trees were also tested
+as blending methods in a more complex scenario where the predictions
+from each model became additional variables to the original explanatory
+variables (called meta-features in this context). The model predictions
+were interacted with each original variable so that models that
+performed more strongly under certain meta-feature states could be
+weighted as such.
+
+Again, 100 models were built to achieve a power of 0.8 for the variation
+in AUC test statistics across different divisions of training/testing
+data. A maximum of five blended models were created from a complete set
+of test results (built from the individual models), then a new set of
+test results were created for the next five blended models.
+
+Finally the impact of the algorithm on the overall profits of the case
+study company was analysed via decision-making scenarios. This analysis
+differed from accuracy in predicting profitability, so several blended
+models as well as individual models were carried forward for this
+analysis. This was done using a profit curve - a chart that plots the
+change in profit the company earns on the y-axis vs.~the probability
+threshold on the x-axis. A simple approach was taken for this analysis,
+where projects with a probability to be a loss making job greater than
+the threshold were rejected entirely. Therefore all profits and losses
+from jobs above the threshold were discounted. An equation defining the
+change in profit as a percentage of the original profit using threshold
+rejection is shown below:
+
+\begin{equation}
+\Delta\ Overall\ Profit\ \textrm{(\%)} = \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100
+\end{equation}
+
+Where
+
+\[
+\begin{aligned}
+I(\cdot) &= \textrm{the indicator function} \\
+N\ &= \textrm{the number of individual projects that are being included in the analysis} \\
+Pr_{p} &= \textrm{probability output from the algorithm for project}\ p \textrm{. Values are}\\
+& \qquad \textrm{betweeen 0 and 1 where 1 is loss making)} \\
+threshold &= \textrm{a chosen value between 0 and 1.}\ \Delta\ Overall\ Profit\ \textrm{is calculated for} \\
+& \qquad \textrm{several}\ threshold\ \textrm{values which defines the profit curve} \\
+Profit_p &= \textrm{profit for individual project}\ p
+\end{aligned}
+\]
+
+If the threshold was zero, all jobs were rejected and the profit would
+be \$0. If the threshold was 1.0, all jobs were accepted and the profit
+would be the same as the profit the company actually experienced since
+the data is a sample of historic projects. The aim was to find the
+optimal threshold point where saying `no' to a job above that level
+would result in higher profits, because jobs that were likely to make a
+loss were being rejected. This chart will clarify what percentage of
+profit increase the company could expect by integrating the algorithm
+into decision-making.
+
+the curve varies with different divisions of the data. Therefore, in
+order to understand the uncertainty around the profit curve and to
+determine which blended or individual model performed statistically
+better than others, a large sample size of curves was required. 100 to
+achieve a power of 0.8. which was achieved by repeating 5-fold
+training/testing splits. A 95\% confidence interval could also be
+determined around the highest point on each curve. The final expected
+increase in profit and the percentage of projects to be rejected
+presented a clear scenario that the case study business managers could
+assess in terms of their business strategy.
+
+\section{4. Prediction Results and
+Discussion}\label{prediction-results-and-discussion}
+
+Present results clearly and concisely discussion should explore the
+signficance of the results of the work, not repeat them.
+
+\paragraph{4.1 Individual Models}\label{individual-models}
+
+The predictive formula was re-structured so that the 11 explanatory
+variables predicted a new response variable: profitable/unprofitable
+projects. The five prediction methods, were initially built without
+imputing data, by using the maximum amount of data possible depending on
+the method. Boosted Trees and Naive Bayes are able to process data with
+missing values, so all data could be input (Ridgeway 2015; Meyer et al.
+2014). 100 models of each method were built. The violin plot below
+summarises the AUC values produced by each method and the `violins' are
+coloured according to whether the distributions significantly vary to
+Boosted Trees using a critical value of 0.05.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Consulting_Profitability_Paper_files/figure-latex/reduced_violin-1.pdf}
+\caption{Violin plot vertically illustrating the distribution of AUC
+values from each of the methods when predicting profit/loss. Subsets of
+the data were used for Logistic Regression and Random Forests in order
+to provide datasets without missing values.}
+\end{figure}
+
+The AUC performance of Logistic Regression and the Random Forest
+algorithms cannot be statistically differentiated from Boosted Trees.
+
+Next, data imputation methods were trialed which would make the complete
+data set available to Logistic Regression and Random Forests. Again, 100
+models were required to achieve a power of 0.8 with respect to Boosted
+Trees
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Consulting_Profitability_Paper_files/figure-latex/mice_violin-1.pdf}
+\caption{Violin plot vertically illustrating the distribution of AUC
+values from each of the methods when predicting profit/loss. Each method
+was fed the same imputed full dataset.}
+\end{figure}
+
+Boosted Trees, Logistic Regression, and Random Forest all performed
+significantly better than the baseline algorithm, Naive Bayes, however
+none outperformed Boosted Trees.
+
+Testing Boosted Trees, Random Forests, Naive Bayes, Logistic Regression
+and Bayesian Networks on the binary classification problem showed that
+Boosted Trees, Logistic Regression and Random Forests performed best
+(according to AUC). Naive Bayes and Logistic Regression were included as
+baseline models and it was expected that the more complex models would
+outperform these. Therefore it was surprising that results from 100
+Logistic Regression models were not statistically significantly lower
+than Boosted Trees.
+
+A possible explanation for this, according to the literature, is that
+there was not enough data for the ensemble trees to learn the complex
+decision rules at which they excel. Trees tend to overfit the patterns
+in a smaller training set. Logistic Regression on the other hand is
+capable of only one decision boundary (which does not have to be
+parallel to the variable axes) and is not prone to overfitting (Perlich,
+Provost, and Simonoff 2003). This may explain Logistic Regression's
+comparatively high performance on the case study's small but complex
+data set.
+
+Boosted trees, Random Forests and Logistic Regression all had mean AUC
+values between 0.755 to 0.764. In summary, the individual binary
+classification models performed well above random chance (AUC = 0.5).
+Whether the model is worth implementing in the work place is dependent
+on the extent to which the algorithm would improve `bottom line' profits
+for the business and if the model can affect decisions in practice.
 
@@
 \item
-  Regression, classification, blending, bottom line analysis
+  NB and BN were bad - one sentence
 \end{itemize}
 
-\section{4. Results and Discussion}\label{results-and-discussion}
+\paragraph{4.2 Blended Models}\label{blended-models}
 
-Present results clearly and concisely discussion should explore the
-signficance of the results of the work, not repeat them.
+Blended models combine the predictions from each high-performing
+individual model to create averaged or `blended' predictions. In this
+case, the Random Forest, Boosted Trees, and Logistic Regression models
+were chosen. Through variable importance studies, 6 variables were
+included as metafeatures. All six methods were compared against the
+original Logistic Regression model's AUC distribution (as it was not
+statistically different from the ensemble trees)
 
-\paragraph{4.1 Predictive Analysis}\label{predictive-analysis}
+\begin{figure}[htbp]
+\centering
+\includegraphics{Consulting_Profitability_Paper_files/figure-latex/blend_violin-1.pdf}
+\caption{Violin plot vertically illustrating the distribution of AUC
+values from each of the blending methods when predicting profit/loss.
+100 models were built for each method.}
+\end{figure}
+
+The above plot shows that 4 methods had a higher mean AUC than the
+original Logistic Regression model:
 
 \begin{itemize}
+\tightlist
 \item
-  regression was attempted - failed
+  the simple blended Logistic Regression
 \item
-  binary classification - results of 5 methods
+  the simple average
 \item
-  blended models - improved results
+  and the two Boosted Tree models.
 \end{itemize}
 
-\paragraph{4.2 Bottom line analysis}\label{bottom-line-analysis}
+The simple averaged method and simple logistic regression could achieve
+a statistical power of 0.8 with 150 samples (models), whereas the
+boosted trees required thousands. The additional 50 models were run for
+the simple logistic regression and simple average to reveal similar
+distributions that were indeed statistically significantly higher than
+the individual logistic regression AUC distribution. P-values from two
+sample t-test were 0.0019 and 0.0021 respectively.
+
+Blended models improved the mean AUC from 0.759 to 0.77, which is an
+increase of 1.4\%. It was expected that model blending would improve
+predictions because it combined the strengths of individual models with
+different theoretical foundations.
+
+The trials of different blending methods demonstrated that again, the
+simplest methods worked best with the case study data. In this incident,
+averaging the results of the best three single models (Logistic
+Regression, Random Forests, and Boosted Trees) or taking a Logistic
+Regression of the three models outperformed more complex Boosted Trees,
+Random Forests, and Logistic Regression blends that facilitated
+interaction of the individual model results and original variables
+(meta-features). Simple blended models could perform better than complex
+methods if the data is not big enough for the complex models to learn
+the more intricate patterns at which they excel. This was observed in
+the single model methods. their success in the 2009 Netflix competition
+generated some publications. The Netflix data set comprised of almost
+3,000,000 observations, so the size of the data could have enabled the
+success of complex blending methods such as FWLS (Sill et al. 2009)
+
+\section{5. Business Impact}\label{business-impact}
+
+This chapter first presents the full range of method results in terms of
+improvements to the case study's bottom line. A business decision making
+scenario was created so that profit curves based on the decision rule
+could be built and analysed. From the profit curves, optimal probability
+thresholds could be derived for each method (individual and blended
+methods). The predictive models output a `probability' between 0 and 1
+that each project will be a loss making job (where probability = 1
+indicates a loss making job). The question the arises, at what
+probability would a decision-maker round the probability up to 1 or down
+to 0? And what business decision would then be made? To find the
+threshold point for rounding, an experimental business-scenario was
+tested. At a given threshold, all projects with probability outputs
+above the threshold were considered too risky, and were rejected. All
+profits and losses from these projects were forfeited, while the profits
+from the remaining jobs (below the threshold) were summed to give a
+revised total profit. This profit calculation was made on a range of
+thresholds between 0 and 1 at 0.05 increments. The total profits were
+plotted for each threshold and joined to make a profit curve
+
+The plot below illustrates the distribution of profit curves for each
+method, where the solid lines join the mean values at each threshold
+point. The grey ribbon illustrates the 95\% confidence interval for the
+profit improvement ratio at each threshold point for the 100 models.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Consulting_Profitability_Paper_files/figure-latex/profit_curve-1.pdf}
+\caption{Profit curves summarising results from 100 models of 9 methods:
+3 simple blends, 3 complex blends, and the original 3 best methods.}
+\end{figure}
+
+Two blended models clearly outperformed the individual models as shown
+in the above plot with higher profit ratios as well as tighter
+confidence intervals. The simple Logistic Regression blend performed
+best with the highest mean profit ratio of 109\% with a standard
+deviation of 1.28\%. \%. This means that for the simple blended Logistic
+Regression model, if all jobs above the probability threshold 0.6 were
+rejected, the profits would increase on average by 9\% in comparison to
+historical profits.
+
+Profit curves from the simple average blend and Logistic Regression
+blended models outperformed the complex blended models, which follows
+logically from their significantly higher AUC distributions. The simple
+average blended model produced a profit curve with an almost identical
+profit improvement (and standard deviation) to the Logistic Regression
+blend. In the data, projects assigned a probability higher than 0.6
+represent only 4.3\% of all projects.
+
+The shaded grey 95\% confidence intervals in the Profit Curve plot shows
+that with the 100 sample models that were analysed, none of the lower
+bounds for the original methods' or complex blending methods were above
+100\% on the y-axis. That is, it cannot be said with 95\% confidence
+that the original methods would produce a mean profit higher than
+historical profits in the given decision framework. Clearly this level
+of certainty is not beneficial for the case-study business to adopt.
+
+It is not clear why simple linear models and the averaging model
+outperformed ensemble tree blending methods. As previously stated, the
+ensemble trees may not have received enough data to adequately learn the
+complex series of rules they develop. The only differences between FWLS
+and the simple Logistic Regression (that performed best) were two
+additional explanatory variables and four interaction terms. The number
+of variables was not high for the amount of data since according to
+Peduzzi et al. (1996), 10 events per explanatory variable or more avoids
+the risk of biased estimation of variable coefficients in Logistic
+Regression. The data contained 315 events per explanatory variable and
+did not pose that risk. Nevertheless, the additional variables must have
+added misleading noise to the model.
+
+To conclude, because of the promising AUC results, it was logical that
+the models translated into financial benefits for the company. The 9\%
+improvement in profits produced by the simple blended Logistic
+Regression is reasonable, and may be high enough to trigger further
+cost-benefit analyses and the development of a more comprehensive
+framework describing different decision scenarios.
 
@@
   profit curves
+\item
+  prototype decision support interface
 \end{itemize}
 
-\section{7. Decision Support Tool}\label{decision-support-tool}
-
-\begin{itemize}
-\item
-  how to generate actions from predictive results
-\item
-  alternative decision scenarios
-\item
-  user trust
-\end{itemize}
-
-\section{9. Conclusions and Future
+\section{6. Conclusions and Future
 Work}\label{conclusions-and-future-work}
@@
 3/4 page
+
+The general aim was to use statistical techniques to predict the
+profitability of projects for a case study consulting business using
+their internal CRM data. This was rigorously completed by approaching
+the prediction of `profitability' as a binary classification problem
+predicting either profit or loss. Several statistical and machine
+learning approaches were applied including Naive Bayes, Bayesian
+Networks, Linear Regression, Random Forests, gradient Boosted Trees as
+well as multiple methods of blending the individual models' output.
+
+AUC = 0.76). This was achieved by 3 individual methods while various
+techniques that blended the individual methods improved results further
+(AUC = 0.77).
+
+predictive models could be shown to improve the overall profitability
+(bottom line) of the case study business. A range of probability
+threshold values were trialed for each method using a decision framework
+where projects scored below the threshold were accepted, while projects
+above the threshold were rejected. If a project was rejected, the
+profits and losses were forfeited, and the remaining accepted profits
+and losses were summed.
+
+Final results showed the simple Logistic Regression blend of the
+individual Logistic Regression, Random Forest and Boosted Tree models
+improved profits the most. The 95\% confidence interval for these
+improvements was between 6.5\% and 11.5\% using a probability threshold
+of 0.6 (approximately 4.3\% of projects).
+
+These results contribute significantly to the research in cost
+estimation in three ways: the applied methods, the decision framework,
+and the appeal to user trust. Ensemble tree methods and blending had
+been applied minimally to cost estimation previously, even though
+ensemble trees provide insight into model structure while predicting at
+a similar level to Neural Networks. Next, previous studies have verified
+predictive accuracy but stopped short of how the algorithm would affect
+decisions and what the measured benefits would be. This study presented
+a clear framework for how a business could improve profits by applying
+the algorithm.
+
+Further work is required to test user confidence in the output. Another
+topic identified for future research was to test how well managers
+estimate some numeric project input variables. In particular, time span,
+team size, total invoiced amount, and percentage of hours completed by
+professionals should be tested as these variables were calculated post
+project completion. Time span and total invoiced amount were discretised
+into wide categories, which should be easier for a manager to choose
+between.
+
+future work: user interface.
+
+Overall, this work has successfully built a mathematical blend of
+Logistic Regression, Random Forests, and Boosted Tree models, from a
+consulting company's internal project data. This blended model can
+predict whether a project will be profitable or not and in a reasonable
+decision framework, can guide managers in rejecting financially risky
+projects and improving profitability of the business.
 
@@
 
-Here are two sample references: Feynman and Vernon Jr. (1963; Dirac
-1953).
+\hypertarget{refs}{}
+\hypertarget{ref-Akintoye2000}{}
+Akintoye, Akintola, and Eamon Fitzgerald. 2000. ``A Survey of Current
+Cost Estimating Practices in the Uk.'' Journal Article.
+\emph{Construction Management and Economics} 18 (2): 161--72.
+doi:\href{https://doi.org/10.1080/014461900370799}{10.1080/014461900370799}.
 
-\hypertarget{refs}{}
-\hypertarget{ref-Dirac1953888}{}
-Dirac, P.A.M. 1953. ``The Lorentz Transformation and Absolute Time.''
-\emph{Physica} 19 (1---12): 888--96.
-doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}.
+\hypertarget{ref-Attalla2003}{}
+Attalla, Mohamed, and Tarek Hegazy. 2003. ``Predicting Cost Deviation in
+Reconstruction Projects: Artificial Neural Networks Versus Regression.''
+Journal Article. \emph{Journal of Construction Engineering and
+Management} 129 (4): 405--11.
 
-\hypertarget{ref-Feynman1963118}{}
-Feynman, R.P, and F.L Vernon Jr. 1963. ``The Theory of a General Quantum
-System Interacting with a Linear Dissipative System.'' \emph{Annals of
-Physics} 24: 118--73.
-doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}.
+\hypertarget{ref-Auria2008}{}
+Auria, Laura, and Rouslan A Moro. 2008. ``Support Vector Machines (Svm)
+as a Technique for Solvency Analysis.'' Journal Article.
+
+\hypertarget{ref-Badenfelt2011}{}
+Badenfelt, Ulrika. 2011. ``Fixing the Contract After the Contract Is
+Fixed: A Study of Incomplete Contracts in It and Construction
+Projects.'' Journal Article. \emph{International Journal of Project
+Management} 29 (5): 568--76.
+doi:\href{https://doi.org/http://dx.doi.org/10.1016/j.ijproman.2010.04.003}{http://dx.doi.org/10.1016/j.ijproman.2010.04.003}.
+
+\hypertarget{ref-Barber2012}{}
+Barber, David. 2012. \emph{Bayesian Reasoning and Machine Learning}.
+Book. Cambridge University Press.
+
+\hypertarget{ref-Breiman1996}{}
+Breiman, Leo. 1996. ``Bagging Predictors.'' Journal Article.
+\emph{Machine Learning} 24 (2): 123--40.
+
+\hypertarget{ref-Breiman2001a}{}
+---------. 2001. ``Random Forests.'' Journal Article. \emph{Machine
+Learning} 45 (1): 5--32.
+doi:\href{https://doi.org/10.1023/A:1010933404324}{10.1023/A:1010933404324}.
+
+\hypertarget{ref-Brown2012}{}
+Brown, Iain, and Christophe Mues. 2012. ``An Experimental Comparison of
+Classification Algorithms for Imbalanced Credit Scoring Data Sets.''
+Journal Article. \emph{Expert Systems with Applications} 39 (3):
+3446--53.
+
+\hypertarget{ref-Caruana2006}{}
+Caruana, Rich, and Alexandru Niculescu-Mizil. 2006. ``An Empirical
+Comparison of Supervised Learning Algorithms.'' Conference Proceedings.
+In, 148:161--68. ACM.
+doi:\href{https://doi.org/10.1145/1143844.1143865}{10.1145/1143844.1143865}.
+
+\hypertarget{ref-Chan2005}{}
+Chan, Swee Lean, and Moonseo Park. 2005. ``Project Cost Estimation Using
+Principal Component Regression.'' Journal Article. \emph{Construction
+Management and Economics} 23 (3): 295--304.
+
+\hypertarget{ref-Elfaki2014}{}
+Elfaki, Abdelrahman Osman, Saleh Alatawi, and Eyad Abushandi. 2014.
+``Using Intelligent Techniques in Construction Project Cost Estimation:
+10-Year Survey.'' Journal Article. \emph{Advances in Civil Engineering}
+2014.
+
+\hypertarget{ref-Elith2008}{}
+Elith, Jane, John R Leathwick, and Trevor Hastie. 2008. ``A Working
+Guide to Boosted Regression Trees.'' Journal Article. \emph{Journal of
+Animal Ecology} 77 (4): 802--13.
+
+\hypertarget{ref-Finnie1997}{}
+Finnie, Gavin R, Gerhard E Wittig, and Jean-Marc Desharnais. 1997. ``A
+Comparison of Software Effort Estimation Techniques: Using Function
+Points with Neural Networks, Case-Based Reasoning and Regression
+Models.'' Journal Article. \emph{Journal of Systems and Software} 39
+(3): 281--89.
+
+\hypertarget{ref-Flyvbjerg2007}{}
+Flyvbjerg, Bent. 2007. ``Cost Overruns and Demand Shortfalls in Urban
+Rail and Other Infrastructure.'' Journal Article. \emph{Transportation
+Planning and Technology} 30 (1): 9--30.
+
+\hypertarget{ref-Hastie2009}{}
+Hastie, Trevor, Robert Tibshirani, and Jerome Friedman. 2009. \emph{The
+Elements of Statistical Learning: Data Mining, Inference, and
+Prediction, Second Edition}. Book. Springer New York.
+\url{http://qut.summon.serialssolutions.com/2.0.0/link/0/eLvHCXMwzV1bS8MwFA7iQBwD3dR6mRAQfNk6trb0Ivi0TQY6EJwgQxi9JLKHddpVf7_nJE269cFnH9NCm_Y7JF--cyPEtnp9s7Im8JCzAWxVjFluEnMWBF4MyyfzE4-HVsR29Tfd9rO89s-BrzBiPSwRH8tocBGfgTxSlGFGrbpQQFAAGIV52JmKxhByhShS_lQg53OGzhsR9IHSKJ6YsfDxcttlP8vYzzrrTEJ4vjYNGaUN8EcoeYfpUofnMCyKACQZ-O6qMEelNAQVpUEpjR1VUGrnKIpucN_BcvnlzqK86ZUNR4cBwrbpes4tFjlf4Yfds9R8fYH9dAA8FGnufK71MnR_ArkqHO7iTW5RQEm_uU4aMEPxsA9subLFFGbHpMYwfaRJ9ljaIkeqaQYt1tAWqU91odzNCXkH0KgCja453QKNKtDuKEJGJWRdqgHrUoCLlnB1qQSLFmCdkpuH8Ww4MXemu_iUBUYW8r_YZ6QRYppDmot0yMQgNQ6myQzkCQZ8l0EO3oKnkT95HMphUw17G5Gz1_vKDaAlwrJNt-edE2pb3PEiL-aRkzh-AKfFqO8nPse2BbFt9S9I-69JXf59-4oclpbTJvt59s2uRQ7tL6t_R8w}.
+
+\hypertarget{ref-Heckerman1998}{}
+Heckerman, David. 1998. \emph{A Tutorial on Learning with Bayesian
+Networks}. Book. Springer.
+
+\hypertarget{ref-Kim2004}{}
+Kim, Gwang-Hee, Sung-Hoon An, and Kyung-In Kang. 2004. ``Comparison of
+Construction Cost Estimating Models Based on Regression Analysis, Neural
+Networks, and Case-Based Reasoning.'' Journal Article. \emph{Building
+and Environment} 39 (10): 1235--42.
+doi:\href{https://doi.org/10.1016/j.buildenv.2004.02.013}{10.1016/j.buildenv.2004.02.013}.
+
+\hypertarget{ref-Kragt2009}{}
+Kragt, Marit E. 2009. \emph{A Beginners Guide to Bayesian Network
+Modelling for Integrated Catchment Management}. Book. Landscape Logic.
+
+\hypertarget{ref-Kumar2007}{}
+Kumar, P Ravi, and Vadlamani Ravi. 2007. ``Bankruptcy Prediction in
+Banks and Firms via Statistical and Intelligent Techniques--A Review.''
+Journal Article. \emph{European Journal of Operational Research} 180
+(1): 1--28.
+
+\hypertarget{ref-Louppe2014}{}
+Louppe, Gilles, and Peter Prettenhofer. 2014. ``Gradient Boosted
+Regression Trees.'' Online Multimedia. PyData.
+
+\hypertarget{ref-Lovallo2003}{}
+Lovallo, Dan, and Daniel Kahneman. 2003. ``Delusions of Success: How
+Optimism Undermines Executives' Decisions.'' Generic. HARVARD BUSINESS
+SCHOOL PUBLISHING CORPORATION.
+\url{http://qut.summon.serialssolutions.com/2.0.0/link/0/eLvHCXMw3V3Pb9MwFLYQQxMSEgxYKAPJBwQHlCqO61-TdoC2aNImcdgmcYscxwGktoMlQfvzeXYSJy1M4swxjZuq_uz3vs95PxCi6TSJd2yCAU-T53pmjbIgoblMCqGYyEtpEpoWs-3zt9Bqb_jsfwB-YVdN1Ye3XTS-IaLT_a553GcwEGvXFsO3O1q7kPf3y1trGmf0fDzFouu5U41Za99AKETJb79ROL_-pbtXOIthsZ3pbxvbna-2iexbJww0RKMGq9l5srHVlGS0OsTIBDI-cqakTczcrnO9439CVCCIZQ6ml751Vc_XxXdTn9hNfHUBDhZ4q1cFH5d3utk_feqIVDxwOT5N9Vd24ZnE5RP0uC_pjT-02B2ge3bzFO33k_sMzQOE-LrEHYTHGADEPYB4ABAPAL7DAb7n6OrT8nJ-GnfNLuKvBChkzFTJcgnkVadGg1XkVllKiXYEDHYLYxKUK3imPCfGklIwA8KxYJQpSsAaSHqIHmmXFLGpffJkEaG9EhayjRyriOBfRmj_izpfyNOzeXt50F9OK5_hN_1ZRzC7fh_EfCpeICxEIg2fcebqtoHtzlOic26UYobKWZFMUOQmPXObor7RJiMJT0BSJGqCDnscsmK1yqgrlQfklMBXWlSyH209lQwoEgPdDnfetDCFO2lWpRmoVBDJqRBOq9a3NTxgZ1i_cOABY3jDfSdkYIhgzEueCSL_MmzeFcl3xSHql3f-6BF6OGyaV-h-fdPY1z5t-DdLboxJ}.
+
+\hypertarget{ref-Macdonald1975}{}
+Macdonald, P. 1975. ``The Logit Transformation: With Special Reference
+to Its Uses in Bio-Assay.'' Journal Article. \emph{Journal of the
+Operational Research Society} 25 (1): 201--2.
+
+\hypertarget{ref-Matson1993}{}
+Matson, Jack E, and Joseph M Mellichamp. 1993. ``An Object‐oriented Tool
+for Function Point Analysis.'' Journal Article. \emph{Expert Systems} 10
+(1): 3--14.
+
+\hypertarget{ref-Mendes2004}{}
+Mendes, Emilia, and Barbara Kitchenham. 2004. ``Further Comparison of
+Cross-Company and Within-Company Effort Estimation Models for Web
+Applications.'' Conference Proceedings. In \emph{Software Metrics, 2004.
+Proceedings. 10th International Symposium on}, 348--57. IEEE.
+
+\hypertarget{ref-e1071}{}
+Meyer, David, Evgenia Dimitriadou, Kurt Hornik, Andreas Weingessel, and
+Friedrich Leisch. 2014. \emph{E1071: Misc Functions of the Department of
+Statistics (E1071), Tu Wien}.
+\url{https://CRAN.R-project.org/package=e1071}.
+
+\hypertarget{ref-Molokken2003}{}
+Moløkken, Kjetil, and Magne Jørgensen. 2003. ``A Review of Software
+Surveys on Software Effort Estimation.'' Conference Proceedings. In
+\emph{Empirical Software Engineering, 2003. Isese 2003. Proceedings.
+2003 International Symposium on}, 223--30. IEEE.
+
+\hypertarget{ref-Moore1989}{}
+Moore, David S, and George P McCabe. 1989. \emph{Introduction to the
+Practice of Statistics}. Book. WH Freeman/Times Books/Henry Holt \& Co.
+
+\hypertarget{ref-Pai2013}{}
+Pai, Dinesh R., Kevin S. McFall, and Girish H. Subramanian. 2013.
+``Software Effort Estimation Using a Neural Network Ensemble.'' Journal
+Article. \emph{Journal of Computer Information Systems} 53 (4): 49--58.
+\url{http://qut.summon.serialssolutions.com/2.0.0/link/0/eLvHCXMwnV1Nb9QwEB0VekFClPLVAJV8gGPAH4kbc4GlyraIbhftpurRysTe06os3VTqz8fjTbawBanqKZHiOIntvBmPPe8BKPmBpxuYgMHsmgALqtHauDrYJRTaZNlB4b3XiBvxty99pGDV2z1IRuR2PxsKmn8UAUm1IcKwz4tfKclI0XJrp6nxALaJVoSG_Ej-WC8rFHnUE4x_Fmlv_e1WUh7I1fKfFiham-EO9Nuf-tjX_FZywW02x3t_xFN40nmlbLAaRruw5S-ewU6v-MA6AHgOn6bjYXU-mJQs-LfjScXKafVtFMNcjAQ8jtiAnZZnk8FJOFTn48l3Vp5Oy9HXk_IFnA3L6vA47dQX0gWlFaWNwjwjOjfDZ43hssFaovNCu7wIZwpnJrhmuSPtK2xE0QhZ68IJn3PnMzxQL-FxTbv0L9qYzef2gKkw-eJZ1BnFzGcKQ_Ua3SxUyRELTGC_byZbI8V6mnZpbxopgVfr624-t2G-QynmSooE3q-6zy5WLB1W2qW03BYBoSTdrbltr9sE9jbKkaChKLgyCbz7s-PXBWjaGJ6guInGPQFxl2KHHcU6UQu0r__73m_gkYwCG7QB-C08bC-v_H5MOv0NwrD2yw}.
+
+\hypertarget{ref-Peduzzi1996}{}
+Peduzzi, Peter, John Concato, Elizabeth Kemper, Theodore R Holford, and
+Alvan R Feinstein. 1996. ``A Simulation Study of the Number of Events
+Per Variable in Logistic Regression Analysis.'' Journal Article.
+\emph{Journal of Clinical Epidemiology} 49 (12): 1373--9.
+
+\hypertarget{ref-Perlich2003}{}
+Perlich, Claudia, Foster Provost, and Jeffrey S Simonoff. 2003. ``Tree
+Induction Vs. Logistic Regression: A Learning-Curve Analysis.'' Journal
+Article. \emph{The Journal of Machine Learning Research} 4: 211--55.
+
+\hypertarget{ref-Pinto1988}{}
+Pinto, Jeffrey K, and Dennis P Slevin. 1988. ``Critical Success Factors
+Across the Project Life Cycle.'' Conference Proceedings. In. Project
+Management Institute.
+
+\hypertarget{ref-Provost2013}{}
+Provost, F., and T. Fawcett. 2013. \emph{Data Science for Business}.
+Book. O'Reilly. \url{https://books.google.com.au/books?id=_1b4nAEACAAJ}.
+
+\hypertarget{ref-gbm}{}
+Ridgeway, Greg. 2015. \emph{Gbm: Generalized Boosted Regression Models}.
+\url{https://CRAN.R-project.org/package=gbm}.
+
+\hypertarget{ref-Saradhi2011}{}
+Saradhi, V Vijaya, and Girish Keshav Palshikar. 2011. ``Employee Churn
+Prediction.'' Journal Article. \emph{Expert Systems with Applications}
+38 (3): 1999--2006.
+
+\hypertarget{ref-Sealfon2012}{}
+Sealfon, Rachel, and Melissa Gymrek. 2012. ``Recitation 6: Random
+Forests and Affinity Propagation.'' Online Multimedia. MIT University.
+\url{https://stellar.mit.edu/S/course/6/fa12/6.047/courseMaterial/topics/topic4/lectureNotes/recitation6/recitation6.pdf}.
+
+\hypertarget{ref-Shah2014}{}
+Shah, Anoop D, Jonathan W Bartlett, James Carpenter, Owen Nicholas, and
+Harry Hemingway. 2014. ``Comparison of Random Forest and Parametric
+Imputation Models for Imputing Missing Data Using Mice: A Caliber
+Study.'' Journal Article. \emph{American Journal of Epidemiology},
+kwt312.
+
+\hypertarget{ref-Shepperd1996}{}
+Shepperd, Martin, Chris Schofield, and Barbara Kitchenham. 1996.
+``Effort Estimation Using Analogy.'' Conference Proceedings. In
+\emph{Proceedings of the 18th International Conference on Software
+Engineering}, 170--78. IEEE Computer Society.
+
+\hypertarget{ref-Shin2015}{}
+Shin, Y. 2015. ``Application of Boosting Regression Trees to Preliminary
+Cost Estimation in Building Construction Projects.'' Journal Article.
+\emph{COMPUTATIONAL INTELLIGENCE AND NEUROSCIENCE} 2015: 149702.
+doi:\href{https://doi.org/10.1155/2015/149702}{10.1155/2015/149702}.
+
+\hypertarget{ref-Sill2009}{}
+Sill, Joseph, Gábor Takács, Lester Mackey, and David Lin. 2009.
+``Feature-Weighted Linear Stacking.'' Journal Article. \emph{ArXiv
+Preprint ArXiv:0911.0460}.
+
+\hypertarget{ref-Trost2003}{}
+Trost, Steven M, and Garold D Oberlender. 2003. ``Predicting Accuracy of
+Early Cost Estimates Using Factor Analysis and Multivariate
+Regression.'' Journal Article. \emph{Journal of Construction Engineering
+and Management} 129 (2): 198--204.
+
+\hypertarget{ref-Weisberg2005}{}
+Weisberg, Sanford. 2005. \emph{Applied Linear Regression}. Book. Vol.
+528. John Wiley \& Sons.
+
+\hypertarget{ref-Zhang2004}{}
+Zhang, Harry. 2004. ``The Optimality of Naive Bayes.'' Journal Article.
+\emph{AA} 1 (2): 3.
 

### 2016-12-13 finalwriteup/Consulting Profitability Paper/test.tex
reset

--- 2016-08-24 finalwriteup/Consulting Profitability Paper/Consulting_Profitability_Paper.tex
@@
 computer, you can download and install the system package
-\emph{texlive-publishers} (Linux) or install the LaTeX~package
-\emph{elsarticle} using the package manager of your TeX~installation,
-which is typically TeX~Live or MikTeX.
+\emph{texlive-publishers} (Linux) or install the LaTeX package
+\emph{elsarticle} using the package manager of your TeX installation,
+which is typically TeX Live or MikTeX.
 

### 2016-12-13 finalwriteup/draft.rmd
reset

--- 2016-08-30 finalwriteup/draft.rmd
@@
 library('e1071')
-library('party')
+# library('party')
 library('randomForest')
@@
 Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods.
-A decision scenario that rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
+A decision scenario which" rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
 This process was repeated on hundreds of models built from different training subsets of the data in order to obtain a 95% statistical confidence interval of profit improvements.
@@
 
-**ANOVA** - Analysis of variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
+**ANOVA** - Analysis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
 
@@
 
-**B2B** - Business to business. Commerce transactions between two businesses.
-
-**BIM** - Building information modeling. A digital representation of physical characteristics of an object or structure.
+**B2B** - Business to Business. Commerce transactions between two businesses.
+
+**BIM** - Building Information modeling. A digital representation of physical characteristics of an object or structure.
 
@@
 
-**CLV** - Customer lifetime value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
-
-**CRM** - Customer relationship management.
+**CLV** - Customer Lifetime Value. Prediction of the profit a customer will bring over the course of a business' relationship with the customer.
+
+**CRM** - Customer Relationship Management.
 Refers to a common type of software used by businesses to record client and project details as well as employee time sheet records.
@@
 
+**FP** - False positive. A case marked 'negative' in a data set is incorrectly predicted 'positive' by an algorithm.
+
+**FPR** - False positive rate. The proportion of positives that were incorrectly identified (false positives).
+
 **FN** - False negative. A case marked 'positive' in a data set is incorrectly predicted 'negative' by an algorithm.
-
-**FNR** - False negative rate. The proportion of negatives that were incorrectly identified (false negatives).
-
-**FP** - False positive. A case marked 'negative' in a data set is incorrectly predicted 'positive' by an algorithm.
-
-**FPR** - False positive rate. The proportion of positives that were incorrectly identified (false positives).
 
@@
 
-**ROC** - Receiver operating characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
+**ROC** - receiver Operating Characteristic. An ROC curve plots the performance of a binary classification algorithm its scoring threshold is varied.
 
 **TN** - True negative. A case marked 'negative' in a data set is correctly predicted 'negative' by an algorithm.
-
-**TNR** - True negative rate. The proportion of negatives that were correctly identified (true negatives).
 
@@
 As most people have experienced, these estimates can be particularly inaccurate if the task is new and difficult.
-These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well being.
+These errors are trivial in everyday activities, but for consulting businesses that solve complex problems, similar time estimation errors determine their financial well-being.
 
@@
 
-“It’s tough to make predictions, especially about the future” - Danish Proverb
+“It’s tough to make predictions, especially about the future” - Source Unknown
 
@@
 The purpose of this rule is to guard against the case where the project creeps in size, and the consultant finds themselves gradually undertaking more work and more work for the same price even though the client was happy to spend more on the overall project.
-This may account for some, or in the best case all, of their overspent time, although again a savvy client may turn to other consultants that agree to omit variation opportunities.
+This may account for some or in the best case, all, of their overspent time, although again a savvy client may turn to other consultants that agree to omit variation opportunities.
 
@@
 A survey by @Moores1992 aligned with this theory and found the most popular response for the supposed reason for cost overruns was over-optimistic estimates (51%).
-This contrasts with an 'outside view' where a subject is analysed by comparing it to other subjects and disregarding specific subject details.
-For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate [@Lovallo2003].
-This 'outside view' can be applied to complex projects, and stipulates that the details of the current project should be ignored in favour of analysing outcomes of several similar projects.
+This contrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details.
+For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate.
+According to @Lovallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project should be ignored in favour of analysing outcomes of several similar projects.
 Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003].
@@
 
-![This model was implemented by the British government for a rail project that was under completion at the time of publication [@Flyvbjerg2011]](images/lit_review/UKrail.JPG)
+![This model was implemented by the British government for a rail project that is still under completion [@Flyvbjerg2011]](images/lit_review/UKrail.JPG)
 
@@
 
-Despite the predictive success in statistical and machine learning cost estimation models as well as the revealing outcomes from 'inside' and 'outside' view studies, there has been a lack of industry uptake.
+Despite the predictive success in statistical and machine learning cost estimation models as well as the revealing outcomes from 'inside' and 'outside' view studies, reference class forecasting there has been a lack of industry uptake.
 Two reports reviewing cost estimation in the software industry investigated reasons why data-driven forecasting models have not been translated to industry.
@@
 
-Reports on other industries were unavailable, however the insights are general and may explain the lack of adoption in other industries (such as the construction industry).
+Reports on other industries were not available, however the insights are general and may explain the lack of adoption in other industries (such as the construction industry).
 
@@
 Through experimentation, the Agile movement was developed.
-The method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
+In summary, the method treats both cost and time as fixed quantities for a project, and any change or additional work can only be accommodated if another, less important requirement is excluded [@Badenfelt2011].
 Furthermore, the project outcomes are continually adjusted and revised based on frequent progress review points that assess unexpected problems.
@@
 This agile methodology suits the nature of software development work, where curtailing the scope of work is relatively straightforward.
-Also, since the industry is relatively young, the new methodology was able to sweep across the field and was shown to deliver better budget and product outcomes to clients [@Rasmusson2010].
+Also, since the industry is more modern, the new methodology was able to sweep across the field and could be proven to deliver better budget and product outcomes to clients [@Rasmusson2010].
 
@@
 The industry is so large, broad and well established that it has inertia in the way projects are estimated, and contractually bound.
-Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high-level of time, legal responsibility and due diligence.
-
-This project aims to assist a company in the construction industry in managing problematic fixed price projects, since fixed price contracts can seem unavoidable.
-The blocks preventing industry adoption, such as providing a clear decision framework for how to implement the model will be carefully addressed.
+Furthermore, the industry would struggle to take on an agile approach as the scope cannot be easily reduced as the project unfolds - the project must result in construction and building certification which implies a high-level of legal responsibility and due diligence.
+
+This project aims to assist a company in the construction industry in managing problematic fixed price projects, since fixed price contracts are unavoidable.
+The blocks preventing industry adoption, such as providing a clear decision framework and incorporating expert judgment will be carefully addressed.
 
@@
 Cost estimation is currently performed in the case study company by first carefully reviewing preliminary drawings where time costs for each task are estimated and summed by an experienced manager (the 'inside view').
-This is often crosschecked with a value that is based off a percentage of the estimated final cost of construction (generally the client budget).
+This is often crosschecked with a value based off a percentage of the estimated final cost of construction (generally the client budget).
 No formal mathematical method of comparison to similar projects is performed.
@@
 Managers could interrogate how the model outputs were calculated and relate to the actual projects used in the model or speak to a colleague that was involved.
-This may make the presented 'outside view' more relatable and have more potential to influence the final fee proposal.
+This may give the presented 'outside view' a stronger meaning and more potential to influence the final fee proposal.
 
@@
 'return per dollar'.
-This measure represents the ratio of the project profit to business costs, and can easily be compared between projects of different sizes.
+This measure represents the ratio of the amount invoiced to business costs, and can easily be compared between projects of different sizes.
 
@@
 
-$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee\ Costs + Business\ Costs)$
+$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee + Business\ Costs)$
 
@@
 
-In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the data set is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending were tested, and the prediction results were integrated into a decision framework for the business.
+In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the data set is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
 These points are outlined in more detail below.
@@
 In past research, cost estimation data sets generally consisted of dozens of projects from a range of a companies and even countries.
-This thesis' case study only had access to the company's internal data but this contained over 2,000 past jobs - far more than the usual cost estimation study and with many more variables.
+This thesis case study only had access to the company's internal data but this contained over 2,000 past jobs - far more than the usual cost estimation study and with many more variables.
 Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry, which differs from other studies in the construction industry that estimate actual construction costs.
@@
 
-The next chapter, Chapter 3, is a thorough review of the literature to date that relates to this thesis.
+The next chapter, Chapter 3 is a thorough review of the literature to date that relates to this thesis.
 This covers work solving the cost-estimation problem, studies on statistical and machine-learning models in business applications, and a review of potential statistical and machine learning methods.
@@
 
-Assessing outcomes of comparable past studies may impart how to best approach the problem and reveal shortcomings that can be improved upon.
+Researching outcomes of comparable past studies may impart how to best approach the problem and reveal shortcomings that can be improved upon.
 This chapter provides an overview of the literature describing early and current cost estimation methods.
@@
 Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004; @Attalla2003].
-@Dissanayaka1999 suggested that Regression models could be performed first to determine variable importance and condense the number of variables, and then Neural Networks could be used for better estimation.
+@Dissanayaka1999 suggested that Regression models could be performed first to determine variable importance and condense the number of variables, then Neural Networks could be used for better estimation.
 Boosted Trees may perform as well as Neural Networks while providing even more insight than Regression models because it is possible to produce partial dependency plots.
@@
 
-Similar to the construction industry, expert judgment or detailed analysis, is the most widely practised method for effort estimation [@Shepperd1996; @Molokken2003].
+Similar to the construction industry, expert judgment or detailed analysis, is the most widely practiced method for effort estimation [@Shepperd1996; @Molokken2003].
 This is despite many years of research being dedicated to developing algorithmic models that, in the research context, outperform expert judgment.
+A survey by @Moores1992 found that software tools assisting a software project were in significant use, but only the project planning tools, not the cost estimating tools.
+This suggests a mental link for project managers to price jobs based on *details* as opposed to an outside view.
 Another study by @Heemstra1992 found there was no evidence estimation accuracy improved when estimation tools were used.
@@
 
-This section assesses the details and performance of the algorithmic models in the IT industry, including their important variables, and gaps in the literature.
+This section assesses the details and performance of the algorithmic models, important variables, and gaps in the literature.
 
@@
 The software industry is unique in that agile methods of delivery are changing the contractual approach of consulting [@Badenfelt2011].
-
-There is a need for research into algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
-For industry adoption, the models should reveal their predictive structure to decision makers and provide a decision framework that outlines how to integrate the predictive model in a way that optimises profits.
+This innovative method keeps delivery times short and defines projects as something with a fixed cost and time.
+If a change occurs, other less important elements are discarded to maintain the fixed amount.
+Software development is a relatively new field in comparison to construction, where the tradition of fixed fee structures is difficult to reinvent.
+Also, portions of construction projects cannot be simply curtailed.
+
+There is a need for research on algorithmic models that predict effort (and not construction cost) for consulting companies in the construction industry.
+From findings in previous literature, the models should reveal their predictive structure to decision makers and be based on internal company data.
+Review of the literature revealed a need for an industry focused decision framework that outlines how to integrate the predictive model in a way that optimises profits.
 
@@
 $RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$
-
-Where
-
-$$
-\begin{aligned}
-\beta\ \textrm{represents the set of coefficients for a model} \\
-N\ \textrm{is the number of points in a data set} \\
-y_{i}\ \textrm{is the response variable value for the}\ i\textrm{th case in the data set} \\
-f(x_{i})\ \textrm{is the predicted response variable value for the}\ i\textrm{th case in the data set}
-\end{aligned}
-$$
 
@@
 
-In Linear Regression, no assumptions are made about the distributions of the variables.
+In linear regression, no assumptions are made about the distributions of the variables.
 Skewed explanatory variable distributions are acceptable but it is generally better to normalise them by, for example, taking the logarithm or square root of the variable.
@@
 
-Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable.
+Nevertheless, the theory behind Linear Regression can be adapted to the prediction of probabilities pertaining to a binary response variable (represented by 1 or 0).
 First, a sigmoid function is assumed as the functional form of the probabilities against an example variable X.
 This is because it is pragmatic and works well: sigmoid functions can be bound by 0 and 1, and a logit transformation of the probability results in a linear relationship between the log odds and X.
-Refer to the equation below where *p* represents probability:
+Refer to the equation below where *p* represents probability [@Macdonald1975]:
 
@@
 ![Plot of log odds vs.
-an 'X' variable and probability vs.
-the 'X' variable.
+an 'x' variable and probability vs.
+the 'x' variable.
 To convert log odds to probability, an exponential transformation is first made followed by the transformation from odds to probability [@Lowry2016].](images/lit_review/logreg.JPG)
@@
 
-In one industry application, this classifier succeeded because of its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
+In one case, this classifier succeeded because of its ability to be refined incrementally with each piece of new data, as opposed to re-calibrating the entire model.
 Naive Bayes was used in a complex spam detection system where new spam emails or toxic text themes could be quickly added to the filtering model.
@@
 SSN\ \textrm{is the sum of squared errors of a node (dataset before the split)} \\
+y_i\ \textrm{is the}\ i \textrm{th response variable value} \\
+\bar{y}\ \textrm{is the mean response variable value in the node} \\
 SSL\ \textrm{is the sum of squared errors in the left dataset that resulted from the split} \\ 
-SSR\ \textrm{is the sum of squared errors in the right dataset that resulted from the split} \\
-y_i\ \textrm{is the}\ i \textrm{th response variable value} \\
-\bar{y}\ \textrm{is the mean response variable value in the node}
+SSR\ \textrm{is the sum of squared errors in the right dataset that resulted from the split} 
 \end{aligned}
@@
 
-In this way, Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values, while still taking into account many conditional dependencies that Naive Bayes, for example, cannot [@Barber2012].
+In this way, Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes, for example, cannot [@Barber2012].
 Also, some equations for a combination are now clones of another combination and do not require recalculation.
@@
 
-Neural Networks of the human nervous system inspired Algorithmic Neural Networks.
+Algorithmic Neural Networks were inspired by Neural Networks of the human nervous system.
 In the human body, biological neurons receive electro-chemical input into their dendrites from the terminals of adjacent neurons.
@@
 On the other hand, the output is in the form of distance to the boundary (as opposed to probability) and results are not transparent (similar to Neural Networks) [@Caruana2006].
-Also, SVM's can be very slow to train and sometimes not suitable for industry purposes [@Auria2008].
+Also, SVM's can be very slow to train and therefore not suitable for industry purposes [@Auria2008].
 Because SVM's perform well in high-dimensional space and can form non-linear rules they have been applied to protein classification (medicine) and text and image recognition [@Byun2002].
@@
 Literature pertaining closely to the cost prediction topic has been reviewed in the first section of this chapter, however it is relevant to broaden the critique to advanced machine learning methods that have been applied to general business problems.
-This is to discover whether other methods have succeeded in the business literature that were excluded from cost estimation.
+This is to find out whether methods that cost estimation had not put to the test succeeded in the broader literature.
 The simpler statistical methods have been covered already, so the focus of this section is the machine learning prediction methods developed for business decisions.
@@
 
-Naive Bayes, Random forests, and SVM's were built.
+Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
 All three models performed at similar levels for overall accuracy at around 80% (Random Forests had the best results, followed by SVM then Naive Bayes).
@@
 
-### Research Gap
-
-Prediction studies on business problems found success in a similar suite of methods as cost estimation, but with much more attention given to ensemble tree methods.
-This study aims to explore this gap.
+### Gap
+
 The limited literature on work similar to Saradhi & Palshikar's [-@Saradhi2011] Employee Churn study highlights a gap in the application of predictive techniques that use internal data to model internal decisions.
@@
 In this research, internal employee team structure as well as client records shall be used to aid prediction.
+The literature on machine learning applied to general business problems revealed SVM's were usually the most accurate technique followed by Neural Networks and ensemble tree methods.
+A gap exists in applying these cutting edge machine learning techniques to the wide range of internal decisions that businesses make.
 
@@
 Previous studies addressing the problem have generally been limited to prediction of construction costs for buildings/infrastructure and effort estimation for IT projects.
-The prediction models consisted of CBR, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM's.
-This leaves a gap in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods, and propagating the research into business adoption methods.
+The prediction models consisted of CBR, Linear Regression, Neural Networks and in one or two instances ensemble trees or SVM.
+This leaves a gap in effort estimation for consulting companies in the construction industry in combination with trialing a broader range of prediction methods.
 
@@
 It was determined that Bayesian Networks and ensemble tree methods have potential to perform estimation as well as complex algorithms such as Neural Networks while providing insights into model structure.
-Strategies and research into how a business could integrate the model into decisions will be developed.
+Other strategies will be developed to address the problem of industry uptake as well.
 
@@
 In the case of employee names, professional titles replaced the names, and client codes replaced client names.
-Project names and addresses were removed, and a number identified each project.
+Project names and addresses were removed, and each project was identified by a number.
 The resulting de-identified data consisted of three .csv files of different data structures, which are described in the next section.
@@
 
-| Job Number | Date Issued | Amount Invoiced (\$) | Amount Paid (\$) | Invoice Status |
+| Job Number | Date Issued | Amount Invoiced | Amount Paid | Invoice Status |
 |------------|-------------|-----------------|-------------|----------------|
-| 1.2.300    | 2014/09/01  | 5,000         | 5,000     | Paid           |
-| 1.2.300    | 2014/10/02  | 2,000         | 2,000     | Paid           |
-| 1.2.300    | 2014/11/04  | 3,000         | 3,000     | Paid           |
-| 1.2.300    | 2014/12/02  | 4,000         | 4,000     | Paid           |
-| 1.2.300    | 2015/02/02  | 5,000         | 0         | Outstanding    |
-| 1.2.400    | 2014/08/05  | 20,000        | 20,000    | Paid           |
-| 1.2.400    | 2014/09/01  | 2,100         | 0         | Outstanding    |
+| 1.2.300    | 2014/09/01  | \$5,000         | \$5,000     | Paid           |
+| 1.2.300    | 2014/10/02  | \$2,000         | \$2,000     | Paid           |
+| 1.2.300    | 2014/11/04  | \$3,000         | \$3,000     | Paid           |
+| 1.2.300    | 2014/12/02  | \$4,000         | \$4,000     | Paid           |
+| 1.2.300    | 2015/02/02  | \$5,000         | \$0         | Outstanding    |
+| 1.2.400    | 2014/08/05  | \$20,000        | \$20,000    | Paid           |
+| 1.2.400    | 2014/09/01  | \$2,100         | \$0         | Outstanding    |
 | ...        | ...         | ...             | ...         | ...            |\
@@
 
-* Project summary data: information describing each project such as client code, client contact code, discipline, subject, postcode, director code, and suburb
+* Project summary data: information describing each project such as client code, client contact code, discipline, subject, post code, director code, and suburb
 
@@
 |------------|------------------|----------|-------------|------------|-----------|--------|------------------|-----------------------|
-| 1.2.300    | "Disguised"      | D1       | "Disguised" | Civil      | 4002      | C5030  | P12              | ...                   |
-| 1.2.400    | "Disguised"     | D2       | "Disguised" | Structural | 4122      | C2000  | P24              | ...                   |
-| 1.2.500    | "Disguised"     | D3       | "Disguised" | Water      | 4230      | C2045  | P30              | ...                   |
-| 1.2.600    | "Disguised" | D2       | "Disguised" | Structural | 4004      | C3001  | P1               | ...                   |
-| 1.2.700    | "Disguised"      | D3       | "Disguised" | Water      | 4570      | C3020  | P5               | ...                   |
-| 1.2.800    | "Disguised"  | D3       | "Disguised" | Water      | 4222      | C4010  | P1               | ...                   |
+| 1.2.300    | 1 Apple St.      | D1       | "Disguised" | Civil      | 4002      | C5030  | P12              | ...                   |
+| 1.2.400    | 33 Rogue Rd.     | D2       | "Disguised" | Structural | 4122      | C2000  | P24              | ...                   |
+| 1.2.500    | 4 Window Dr.     | D3       | "Disguised" | Water      | 4230      | C2045  | P30              | ...                   |
+| 1.2.600    | 864 Hamilton St. | D2       | "Disguised" | Structural | 4004      | C3001  | P1               | ...                   |
+| 1.2.700    | 2 Shark St.      | D3       | "Disguised" | Water      | 4570      | C3020  | P5               | ...                   |
+| 1.2.800    | 24 Campbell Rd.  | D3       | "Disguised" | Water      | 4222      | C4010  | P1               | ...                   |
 | ...        | ...              | ...      | ...         | ...        | ...       | ...    | ...              | ...                   |\
@@
 These three sources needed to be compiled into a single data set that detailed one project per row, as the model was to predict the overall profitability of a project.
-Before compilation could begin however, thorough cleaning was required where the data was plotted and assessed so that outliers could be visually or statistically discovered.
+Before compilation could begin however, thorough cleaning was required where the data was plotted and statistically assessed so that outliers could be visually or statistically discovered.
 Outliers were then investigated for data entry errors.
@@
 Once these errors were detected, they were discussed with the company directors and appropriate corrective action to the data was taken.
-Notes that a user had entered into the system could often correct mistakes in data entry.
+Mistakes in data entry could often be corrected by notes that a user had entered into the system.
 It is possible that some data entry errors were not detected and this is a problem that also must be dealt with during implementation of the model.
@@
 * Total amount invoiced and remunerated per project
-* Mean invoice size per project.
-For example, if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500 for that project.
-This was done for all invoices sent to a client across all their projects and the mean was taken.
+* Mean invoice size per project (for example if a project had four invoices totaling \$10,000, the mean invoice size would be \$2,500) over all invoices sent to a client across all their projects.
 This gives an indication of whether the client generally does fast paced big jobs, which would result in large monthly invoices for example, or small jobs with small monthly invoiced amounts.
@@
 
-* Profit. Note the cost of employee hours includes a margin to account for the business overheads.
+* Profit
 
@@
 
-![Hierarchical dendrogram of the 'time span' variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
+![Hierarchical dendrogram of the Time span variable with 6 clusters highlighted [@stats]](images/method/dendrogram.png)
 
@@
 
-The revised discretised categories for 'time span' and 'total amount invoiced' were as follows:
-
-'Time span' Categories
+The revised discretised categories for time span and total amount invoiced were as follows:
+
+Time span Categories
 
@@
 
-'Total Invoiced Amount' Categories
+Total Invoiced Amount Categories
 
@@
 n &= \textrm{number of data points} \\
-\widehat{y}_t &= \textrm{predicted values of the response variable} \\
-y_t &= \textrm{observed values of the response variable} \\
+\widehat{y} &= \textrm{predicted values of the response variable} \\
+y &= \textrm{observed values of the response variable} \\
 \end{aligned}
@@
 In order to compare which models performed significantly better than others, an adequate sample size of results statistics is required.
-Using different data in the training vs.
-testing sets could produce multiple models, each providing a resulting test statistic (RMSE or AUC).
+Multiple models could be made by using different data in the training vs.
+testing sets, each providing a resulting test statistic (RMSE or AUC).
 5 fold cross validation was used, which created 5 models, then the division of 5 folds was repeated using a different random seed to create another set of 5 training/testing data sets.
@@
 \begin{aligned}
+I(\cdot)\  \textrm{is the indicator function} \\
 Px_{i,j} = \textrm{proximity between datapoints}\ i\  \textrm{and}\ j \\
-I(\cdot)\  \textrm{is the indicator function} \\
 E_{n,i} = \textrm{is the end node that case}\ i\ \textrm{falls into in tree}\ n
@@
 Linear Regression must be fed complete data sets but the case study data had a core set of only 13 complete variables (including the response variable).
-If an incomplete variable were added to the core variables, the 'complete' data set would shrink to the complete rows of the incomplete variable.
-If another incomplete variable were added, the complete set of data would shrink again.
+If an incomplete variable was added to the core variables, the 'complete' data set would shrink to the complete rows of the incomplete variable.
+If another incomplete variable was added, the complete set of data would shrink again.
 Therefore, to make the most of the available data, each incomplete variable was added to the core variables one at a time, where a Linear Regression model was built for each.
@@
 
-```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of ANOVA regression F-statistics which were used to interpret variable importance")}
+```{r, echo=FALSE, fig.height=4.7, fig.cap = c("P-values of ANOVA regressions' F-statistics which were used to interpret variable importance")}
 
@@
         scale_y_continuous(breaks = c(0, 0.05, .25, .5, .75, 1)) +
-        labs(y = "P-Value of F-Statistic", x = "Explanatory Variables",
+        labs(y = "P-Value of F-Statistic", x = "Independent Variables",
              title = "Summary of Variable F-Statistic P-Values from ANOVA Models")
@@
         geom_bar(stat='identity') +
-        labs(y = "Relative Variable Importance Value", x = "Explanatory Variables", title = "CForest Variable Importance Rankings of Complete Variables") +
+        labs(y = "Relative Variable Importance Value", x = "Independent Variables", title = "CForest Variable Importance Rankings of Complete Variables") +
         theme(panel.background = element_rect(fill = 'gray95'))
@@
 Random Forests favour variables that are numeric or have many categories and this held true for the analysis.
-Its 6 highest-ranked variables were numeric with one categorical variable, 'business category of the client', which had 28 categories.
+Its 6 highest-ranked variables were numeric with one categorical variable, Business category of the client, which had 28 categories.
 The two lowest ranking variables had only 4 categories.
 It was expected that cForests would rank the categorical variables with only 4 categories more fairly.
-This was the case with 'discipline' and it received a mid-level ranking.
-'Business category of the client' and 'broad business type' however, were ranked similarly to Random Forests.
+This was the case with Discipline and it received a mid-level ranking.
+Business category of the client and Broad business type however, were ranked similarly to Random Forests.
 
 The overall qualities of the higher ranked variables were not surprising, and were generally centered around who did how much of the work internally, client characteristics, project time span, and project category.
-It was expected that 'billing type' would play a more important role in predicting profitability as it describes the structure of the incoming fees.
-For this reason, and with advice from the case study business, the 'billing type' was kept in the model.
-
-For IT software projects, the literature indicated size was the most important variable predicting effort, and in construction projects size was also often an explanatory variable.
+It was expected that Billing Type would play a more important role in predicting profitability as it describes the structure of the incoming fees.
+For this reason, and with advice from the case study business, the Billing Type was kept in the model.
+
+For IT software projects, the literature indicated size was the most important variable predicting effort, and in Construction projects size was also often an explanatory variable.
 Size could mean size of a building or function point (level of functionality of the software) [@Shepperd1996; @Finnie1997; @Pai2013].
@@
 Instead of predicting total effort, it is describing whether the size of a project correlates with how profitable the project is.
-The closest estimate of size in this case study is the 'invoiced amount' variable, which did rank as very important.
+The closest estimate of size in this case study is the invoiced amount variable, which did rank as very important.
 The exact invoiced amount is difficult to guess before a project begins, and is part of the problem this study is addressing.
-Therefore, the 'invoiced amount' was binned into categories that a manager could much more confidently choose as an estimate.
+Therefore, the invoiced amount was binned into categories that a manager could much more confidently choose as an estimate.
 The other variables indicated as important in this study (including the number of employees on a project, client business type and percent of hours by certain types of employees) have not been strongly represented in the literature to date.
@@
                           name = "Significantly different from Boosted Trees") +
-        labs(x = "", y = "Area Under Curve (AUC)",
+        labs(x = "Method", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 100 models of Each Method") +
@@
         scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from Boosted Trees") +
+                          name = "Significantly different from boosted trees") +
         scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from Boosted Trees") +
-        labs(x = "", y = "Area Under Curve (AUC)",
+                          name = "Significantly different from boosted trees") +
+        labs(x = "Method", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 100 Models of Each Method - Imputed Data") +
@@
                           name = "Data Source") +
-        labs(x = "", y = "Area Under Curve (AUC)",
+        labs(x = "Method", y = "Area Under Curve (AUC)",
              title = "Comparison of Original Data vs. Imputed Data") +
@@
         theme(axis.text.x=element_text(angle=40,hjust=1)) +
-        labs(x = "", y = "Area Under Curve (AUC)",
+        labs(x = "Method", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 100 Blended Models of each Method") +
@@
 
-```{r, fig.height = 5.5, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best two blending methods when predicting profit/loss.
+```{r, fig.height = 5.5, fig.cap = c("Violin plot illustrating the distribution of AUC values from the best three blending methods when predicting profit/loss.
 150 models were built for each method to achieve a statistical power of 0.8.")}
@@
         theme(axis.text.x=element_text(angle=40,hjust=1)) +
-        labs(x = "", y = "Area Under Curve (AUC)",
+        labs(x = "Method", y = "Area Under Curve (AUC)",
              title = "AUC Distribution from 150 Blended Models of Best Methods") +
@@
 It was not clear why this was the case, however both simple models were compared again in the analysis of the impact each model could have on the company's overall profits.
-The analysis of overall profits did not directly relate to AUC measurements, so a range of models was carried forward for comparison.
+The analysis of overall profits did not directly relate to AUC measurements, so a range of models were carried forward for comparison.
 These included the three original individual models, three simple blended models and three complex blended models.
@@
 The predictive models output a 'probability' between 0 and 1 that each project will be a loss making job (where probability = 1 indicates a loss making job).
-The question then arises, at what probability would a decision-maker round the probability up to 1 or down to 0?
+The question the arises, at what probability would a decision-maker round the probability up to 1 or down to 0?
 And what business decision would then be made?
@@
 
-```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit Curve of the best performing method: the simple Logistic Regression blended method")}
+```{r av_profit_curve, fig.height = 4.5, fig.cap = c("Profit curve of the best performing method: the simple Logistic Regression Blended Method")}
 
@@
 ```{r}
-simp.l <- readRDS("finance_average_model.rds")
+simp.l <- readRDS("report_data/finance_average_model.rds")
 
@@
 \begin{aligned}
-Profit_t &= \textrm{Profit from project}\ t\textrm{. Negative values indicate loss making projects} \\
-n &= \textrm{Number of projects}
+Profit_t &= \textrm{Profit from project}\ t\textrm{. Negative values indicate loss making projects}
 \end{aligned}
@@
 As shown in the upper plot, a low count of projects fall above the optimal threshold for the simple Logistic Regression blend.
-The lower plot illustrates the relative profits (and losses for negative values) within each probability score bin, which are normalised by the sum of the absolute value of all profits and losses.
+The lower plot illustrates the relative profits (and losses for negative values) within each probability score bin which are normalised by the sum of the absolute value of all profits and losses.
 The model has assigned low probability scores to profitable projects and high scores to unprofitable jobs well.
@@
 All original methods did achieve improvements of profit at their optimal thresholds, but did not necessarily consistently improve profits across all cross-folded models.
-The shaded grey 95% confidence intervals in the profit curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' were above 100% on the y-axis.
-That is, it cannot be said with 95% confidence that the original methods' would produce a mean profit higher than historical profits in the given decision framework.
+The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods were above 100% on the y-axis.
+That is, it cannot be said with 95% confidence that the original methods would produce a mean profit higher than historical profits in the given decision framework.
 This framework stated that projects higher than the optimal thresholds (jobs that are likely to be loss making) should be rejected.
@@
 
-```{r TP_plot, fig.height = 5, fig.cap = c("True positive and false negative rates for each method using a typical model from each - displayed as a bar graph.")}
-
-conf.plot1 <- readRDS("report_data/conf_plot.rds")
+```{r TP_plot, fig.height = 5, fig.cap = c("True positive and false negative rates for each blended method using a typical model from each - displayed as a bar graph.")}
+
+conf.plot <- readRDS("conf_plot.rds")
 
 #if model.pred = loss, this is TP
-conf.plot1$model.pred <- ifelse(conf.plot1$actual == "loss" & 
-                                       conf.plot1$model.pred == "loss", "TP", "FN")
+conf.plot$model.pred <- ifelse(conf.plot$actual == "loss" & 
+                                       conf.plot$model.pred == "loss", "TP", "FN")
 # reorder levels in method to suit TP 
-conf.plot1$method <- factor(conf.plot1$method, 
-                           levels= conf.plot1 %>% filter(actual == "loss", model.pred == "TP") %>%
+conf.plot$method <- factor(conf.plot$method, 
+                           levels= conf.plot %>% filter(actual == "loss", model.pred == "TP") %>%
                                    arrange(desc(pc)) %>% .$method)
 
-ggplot(conf.plot1 %>% filter(actual == "loss"), 
+ggplot(conf.plot %>% filter(actual == "loss"), 
        aes(x = method, y = pc, fill = model.pred)) +
@@
 
-The plot above shows the complex Boosted Tree blended model has the highest TPR followed by the simple Logistic Regression blend and the simple averaged blend.
-All of these TPR's are below 25%, which shows it was not worth rejecting a large percentage of loss making jobs in comparison to the accompanying losses in rejected profitable jobs (FP's).
+The plot above shows the Complex Boosted Tree Blend has the highest TPR followed by the Simple Logistic Regression Blend and the Simple Averaged Blend.
+All of these TPR's are below 25% which shows it was not worth rejecting a large percentage of loss making jobs in comparison to the accompanying losses in rejected profitable jobs (FP's).
 This also implies the models were not able to cleanly divide loss making jobs from profitable jobs.
@@
 
-```{r TN_plot, fig.height = 5, fig.cap = c("TNR and FPR for each method using a typical model from each - displayed as a bar graph.")}
+```{r TN_plot, fig.height = 5, fig.cap = c("True negative and false positive rates for each blended method using a typical model from each - displayed as a bar graph.")}
 
@@
 
-### Review of Projects Rejected by Simple Logistic Regression
+### Review of Jobs Rejected by Simple Logistic Regression
 
@@
 
-All below plots and comments summarise results from one simple blended Logistic Regression model.
+All below plots and comments summarise results from one Simple Blended Logistic Regression Model.
 The results from this model were chosen because it represented the mean improvement in profits output by this blending method.
-The plot below illustrates how the ratio of profitable to unprofitable jobs compare in the full project population vs.
+The plot below illustrates how the ratio of profitable to unprofitable jobs compares in the full project population vs.
 the rejected projects.
@@
 Originally, 25% of projects were loss making.
-The model rejects a population of projects that are over 60% loss making.
+The model rejects a population of projects which are over 60% loss making.
 This represents a significant swing in proportions, but is not perfect.
@@
 Colours are alternated between different jobs to indicate profit/loss magnitude.
-Predictions were from a typical simple Logistic Regression blended model.")}
+Predictions were from a typical simple Logistic Regression.")}
 
@@
 the rejected projects only.
-Predictions from a typical simple Logistic Regression blend were used.")}
+Predictions from a typical simple Logistic Regression were used.")}
 
@@
 
-The model tended to accept projects that were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
+The model tended to accept projects which were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee.
 This could either indicate that professional level employees complete projects more efficiently or that the types of projects that do not require technical employees are more profitable.
@@
 
-```{r, fig.cap = c("Proportion of each 'timespan' category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression blend were used.")}
+```{r, fig.cap = c("Proportion of each timespan category for projects in the full data set vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
 
@@
 The statistical distribution of 'return per dollar' in the rejected projects is skewed towards negative returns.
-This means the algorithm has identified a proportionally higher percentage of loss making projects, which is not surprising given previous results.
-
-```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression blend were used.")}
+This means the algorithm has identified a proportionally higher percentage of loss making projects which is not surprising given previous results.
+
+```{r, fig.cap = c("Distribution of 'return per dollar' in the full data set of projects vs. the rejected projects only. Predictions from a typical simple Logistic Regression were used.")}
 
@@
 
-The confusion matrix statistics revealed that the most profitable method, the simple Logistic Regression blended model, was optimised at a threshold with a TPR of only `r conf.plot1 %>% filter(method == "simp.log", model.pred == "TP") %>% .$pc`% and a TNR of `r conf.plot1 %>% filter(method == "simp.log", model.pred == "FN", actual == "profit") %>% .$pc %>% .[2]`%.
+The confusion matrix statistics revealed that the most profitable method, the simple Logistic Regression blended model, was optimised at a threshold with a TPR of only `r conf.plot %>% filter(method == "simp.log", model.pred == "TP") %>% .$pc`% and a TNR of `r conf.plot %>% filter(method == "simp.log", model.pred == "FN", actual == "profit") %>% .$pc %>% .[2]`%.
 Despite the low TPR, it was higher than almost all other methods at their optimal threshold and produced notable improvements in profits to the business.
 Review of the rejected project characteristics were logical, and development of the user interface as well as some further data analyses must be completed before the model can be finally introduced into business decisions.
-These are described in the 'User Interface and User Engagement' and 'Limitations and Future Work' sections below.
-
-## User Interface and User Engagement
+These are described in the 'User Interface and Engagement' and 'Limitations and Future Work' sections below.
+
+## User Interface and Engagement
 
 Industry's lack of uptake of cost estimation models was marked as a significant issue in the literature [@Akintoye2000].
-Decision makers did not sufficiently trust the model output that were not directly connected to their experiences.
+Decision makers did not sufficiently trust the model output which was not directly connected to their experiences.
 In many studies, CBR had more traction than more accurate models because decision makers could justify the output.
@@
 The 100 models gave 100 probability outputs for each project, and the 95% confidence interval for these outputs was calculated.
-A typical spread for the 95% confidence interval is shaded in the graphic, which is relative to the 'bin' that the score fell into (i.e.
+A typical spread for the 95% confidence interval is shaded in the graphic which is relative to the 'bin' that the score fell into (i.e.
 6-8).
@@
 This graphic is an initial suggestion for how to communicate model results to industry.
-Many variations are possible, which could depend on how the business wants to act on the output; for example whether they plan to reject risky projects outright or adjust the fee structure.
+Many variations are possible which could depend on how the business wants to act on the output; for example whether they plan to reject risky projects outright or adjust the fee structure.
 Additional methods built on top of the current model may be able to produce statistical probabilities instead of probability scores.
@@
 
-![Graphic for model prediction and Boosted Tree partial dependency plot embedded in the user interface](images/app/team_size_dep.png)
-
-![Boosted Tree Partial dependency plot for 'timespan' embedded in the user interface](images/app/timespan_dep.png)
+![Graphic for model prediction and Boosted Tree partial dependency plot](images/app/team_size_dep.png)
+
+![Boosted Tree partial dependency plot for 'Timespan' embedded in the user interface](images/app/timespan_dep.png)
 
@@
 
-![User interface for Nearest Neighbour algorithm input options and chart](images/app/NN_top.png)
-
-![User interface Nearest Neighbour chart coloured by the position of the main employee on the projects](images/app/NN_emp.png)
+![User interface for Nearest Neighbour algorithm input and output chart](images/app/NN_top.png)
+
+![Nearest Neighbour chart coloured by the position of the main employee on the projects](images/app/NN_emp.png)
 
@@
 Uncertainty has been managed by binning variables such as 'total amount invoiced' and 'time span' to 5 or 6 broad categories (i.e.
-"1 day - 3 weeks", "3 weeks - 2.5 months", etc).
-However, two remaining categories: 'team size' and 'percent of project completed by professional level employees' were left as continuous variables.
+"1 day - 3 weeks", "3 weeks - 2.5 months", etc.).
+However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables.
 It would be interesting to test how well business managers could guess the aforementioned four variables at the start of a project.
@@
 However, it would be worthwhile integrating into the pre-existing data-collection system.
-When a new project is entered into the database, it could be mandated that the project manager guess the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent by professionals.
+When a new project is entered into the database, it could be mandated that the project manager guesses the time-span (1 of 6 categories), total amount invoiced (1 of 5 categories), team size and percent by professionals.
 Over time, this estimation data could contribute to a revised version of the algorithm.

### 2016-12-13 finalwriteup/trial_code.rmd
reset

@@
         theme(axis.text.x = element_text(angle = 45, hjust = 1),
-              panel.background = element_rect(fill = 'grey95')) +
+              panel.background = element_rect(fill = 'gray95')) +
         geom_hline(yintercept = 0.05, colour = 'red', linetype = 2) +
@@
 ggplot(data=cfor.varimp, aes(x=var, y=imp)) +
-        theme(axis.text.x=element_text(angle=45,hjust=1)) +
+        theme(axis.text.x=element_text(angle=45,hjust=1),
+              panel.background = element_rect(fill = 'gray95')) +
         geom_bar(stat='identity') +
@@
         stat_summary(fun.y = mean, geom="line") +
-        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
-        scale_x_continuous(breaks = seq(0,1, by = .1)) +  
+        scale_y_continuous(breaks = seq(0,110, by = 20)) + 
+        scale_x_continuous(breaks = seq(0,1, by = .2)) +  
         theme(legend.position = "none",
@@
         theme(legend.position = "none", 
-              panel.background = element_rect(fill = 'grey95')) +
+              panel.background = element_rect(fill = 'grey95'),
+              panel.grid.minor = element_blank()) +
         labs(x = "Probability Threshold", 

### 2016-12-13 model_blend.rmd
reset

@@
 ```{r}
-fit2 <- aov(b.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + no.users*norm.log + b.inv.log*norm.log + Business*norm.log + JD.Second*norm.log + b.timespan.cbrt*norm.rf + no.users*norm.rf + b.inv.log*norm.rf + Business*norm.rf + JD.Second*norm.rf + b.timespan.cbrt*norm.boost + no.users*norm.boost + b.inv.log*norm.boost + Business*norm.boost + JD.Second*norm.boost + Discipline*norm.boost + Discipline*norm.log + Discipline*norm.rf,
+fit2 <- glm(b.rpdol ~ b.timespan.cbrt + no.users + b.inv.log + Business + JD.Second + norm.log + norm.rf + norm.boost + b.timespan.cbrt*norm.log + no.users*norm.log + b.inv.log*norm.log + Business*norm.log + JD.Second*norm.log + b.timespan.cbrt*norm.rf + no.users*norm.rf + b.inv.log*norm.rf + Business*norm.rf + JD.Second*norm.rf + b.timespan.cbrt*norm.boost + no.users*norm.boost + b.inv.log*norm.boost + Business*norm.boost + JD.Second*norm.boost + Discipline*norm.boost + Discipline*norm.log + Discipline*norm.rf, family = binomial(),
                                    data = all10.blend)
+
+backwards = step(fit2)
+formula(backwards)
 

### 2016-12-13 prob_review.rmd
reset

@@
 ```{r_loadpackages, echo=FALSE, include=FALSE}
-library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
-library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
-library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
-library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('randomForest', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
-library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.3/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.3/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.3/library')
+library('gbm', lib = 'C:/Progra~1/R/R-3.2.3/library')
+library('randomForest', lib = 'C:/Progra~1/R/R-3.2.3/library')
+library('caret', lib = 'C:/Progra~1/R/R-3.2.3/library')
+library('pROC',lib='C:/Progra~1/R/R-3.2.3/library')
 
@@
 
+        #order levels
+        df <- within(df, b.timespan.cbrt <- factor(b.timespan.cbrt, 
+                                               levels=c('1d-3wk', '3wk-2.5m','2.5m-9m','9m-1.5y','1.5y-3y','>3y')))
+        df <- within(df, majority.pos <- factor(majority.pos, 
+                                               levels=c('contracttech', 'midtech','seniortech','gradpro',
+                                                        'midpro','seniorpro','director')))
+
+        levels(df$b.timespan.cbrt) <- c('1d-3wk', '3wk-2.5m','2.5m-9m','9m-1.5y','1.5y-3y','>3y')
+        df$b.timespan.cbrt <- ordered(df$b.timespan.cbrt)
+
         #set up 10 fold cv using caret function
@@
                                 } else {
-                                        fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
+                                        fit <- gbm(as.formula(formula), data = df, #[-folds[[j]],],
                                                    distribution = "bernoulli", n.trees = 10000,
@@
 
+#partial dependence plot for boosted only
+fit$var.names
+i=5
+a<- plot.gbm(fit, i, type="response")
+a
+
 ```

### 2016-12-13 rand_forest.rmd
reset

@@
 ```
+
+New analysis for thesis write up
+
+```{r, echo=FALSE}
+
+sample<- all7d[1000,]
+sample<- sample[,!is.na(sample)]
+vars<- names(sample) 
+vars<- vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc','Biz.type','Post.Code',
+               'code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto','balance.mlsto','client.totinv.log','pc.majpos')]
+vars
+
+#make new dataframe for future use with normalised variables and variables that you want to use only
+all8a<- all7d %>% select(no.users, Discipline, pc.pro, Business, code.client, JD.Second, timespan.cbrt,
+                         inv.mlsto.log,client.totinv.log, pc.majpos.log, code.contact, Billing.Type,
+                         majority.pos, return.pdol, mlsto, Year)
+write.csv(all8a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')
+all8a <- read.csv("~/OneDrive/shared files/Bligh Tanner/masters/data/all8a.csv")
+
+RMSE<- function(answer, guess){
+        error= answer-guess
+        root.error= sqrt(mean(error^2))
+        return(root.error)
+}
+
+final.rf.output<- function(df=all8a, no.tests = 50, core = TRUE) {
+
+        summary.tests = matrix(NA, nrow = no.tests, ncol = 3)
+
+        for(i in 1:no.tests){
+
+        if(core == FALSE){
+        samp = sample.int(nrow(df), 1)                        
+        #create sample df - use all variables that are complete from a sample row of data
+        sample = df[samp,]
+        sample = sample[,!is.na(sample)]
+        vars= names(sample)
+        vars = vars[! vars %in% c('mlsto','Year','return.pdol')]
+        variables = vars
+        } else{
+                variables = c('inv.mlsto.log', 'Discipline', 'timespan.cbrt',
+                              'no.users', 'pc.pro', 'Business')
+        }
+
+        #make reduced dataset with complete cases for all variables we want to use
+        reduced = df[complete.cases(df[,variables]),]
+
+        if('code.client' %in% variables){
+        sample = factor(sample(unique(reduced$code.client),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.client %in% sample,]
+
+        reduced$code.client= factor(reduced$code.client)
+        }
+
+        if('code.contact' %in% variables){
+        sample1 = factor(sample(unique(reduced$code.contact),size=53,
+                                     replace=FALSE))
+        reduced= reduced[reduced$code.contact %in% sample1,]
+
+        reduced$code.contact= factor(reduced$code.contact)
+        }
+
+        formula = paste("return.pdol ~ ")
+
+        for(j in 1:length(variables)){
+                formula = paste(formula, variables[j], sep= " + ")
+        }
+
+        samp_size = floor(0.75*nrow(reduced))
+        create.train = base::sample.int(nrow(reduced), samp_size)
+        train.df = reduced[create.train,]
+        test.df = reduced[-create.train,]
+
+        cats = c('code.client', 'code.contact', 'JD.Second', 'Business', 'majority.pos')
+        for(k in seq_along(cats)){
+                if(cats[k] %in% variables){
+                test.df = test.df[test.df[,cats[k]] %in% unique(train.df[,cats[k]]),]
+        }
+        }
+
+        final = randomForest(as.formula(formula), data= train.df, mtry=5,ntree=500, importance=FALSE)
+
+        ans = data.frame(answer= test.df$return.pdol, final.lm = predict(final, test.df))
+
+        #print RMSE for cases that match certain categories only
+        #first subset predictions to cases that match the category
+
+        test.df$predict= ans$final.lm
+
+        #random RMSE
+        r= mean(reduced$return.pdol)
+        random.RMSE= RMSE(reduced$return.pdol, r) %>% round(3)
+
+        #final RMSE for all rows in reduced in a regression
+        final.RMSE= RMSE(ans$answer, ans$final.lm) %>% round(3)
+
+        summary.tests[i,1]= random.RMSE
+        summary.tests[i,2]= final.RMSE
+        summary.tests[i,3]= nrow(test.df)
+
+        #track progress
+        cat(i," ")
+        }
+        summary.tests = as.data.frame(summary.tests)
+        names(summary.tests) = c('random.RMSE', 'final.RMSE', 'test.nrow')
+        return(summary.tests)
+}
+
+#how many tests required? power calculation
+n.samples = function(data = results.core$diff, type.calc = 'one.sample', alt = 'greater'){
+        d.calc = abs(mean(data))/sd(data)
+        pow_pow = pwr.t.test(n = NULL, d = d.calc , sig.level = 0.05, power = .8, type = c(type.calc), alternative = alt)
+        return(ceiling(pow_pow$n))
+}
+
+n.samples(data = results.core$diff)
+n.samples(data = results$diff)
+
+# 3 tests required for results.core
+
+results = final.rf.output(df=all8a, no.tests = 100, core = FALSE)
+results = results %>% mutate(diff = random.RMSE - final.RMSE)
+results = results %>% filter(diff>-25)
+results.core = final.rf.output(df=all8a, no.tests = 50, core = TRUE)
+results.core = results.core %>% mutate(diff = random.RMSE - final.RMSE)
+saveRDS(results, "~/Documents/github/consult/finalwriteup/report_data/reg_rf.rds")
+saveRDS(results.core, "~/Documents/github/consult/finalwriteup/report_data/reg_rf_core.rds")
+
+results.core %>% ggplot(., aes(x = test.nrow, y = final.RMSE)) + geom_point()
+
+#histogram of difference
+results %>% ggplot(., aes(x =diff, y = ..density..)) + geom_histogram() + geom_density()
+
+# t- test
+t_test = t.test(results$diff, alternative = "greater", mu = 0)
+str(t_test)
+t_test$p.value
+```

### 2016-12-13 rf_log_naiive_auc.rmd
reset

@@
 for(i in seq_along(index)){
-        temp = n.samples(data = tot.auc.m[c(2, index[i])], type.calc = 'two.sample', alt = 'two.sided')
+        temp = n.samples(data = reduc.totAUC[c(2, index[i])], type.calc = 'two.sample', alt = 'two.sided')
         ns[i] = temp

### 2016-12-13 summit_report/summit_report.rmd
reset

---
title: "BT_insights"
author: "Amy Cook"
date: "October 7, 2016"
output:
  word_document: default
  html_document: default
---

```{r loadpackages, echo=FALSE, include=FALSE}
library('knitr')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('randomForest')
library('caret')
library('pROC')
library('testthat')
library('purrr')
library(tidyr)
library(broom)

opts_chunk$set(fig.width=6.7, fig.height = 5.5, dpi = 300, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
```

```{r, include = FALSE}
if(.Platform$OS.type == 'windows'){
        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
        source("C:/Users/n9232371/Documents/github/consult/functions.R")
} else{
        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
        source("/Users/yam/Documents/github/consult/functions.R")

}

```

The output below plots the tendency for a project to be loss making (y-axis) against an explanatory variable (x-axis).

The further up the y-axis a point sits, the more likely the project is to be loss-making is the simplest way of thinking about it.

On the x-axis, some of the category labels are missing, so I've printed out all the categories in the line below.

Also, some of the variable names on the x-axis are complicated, so I've added a comment after each plot which will help you interpret the x axis.

```{r, cache = TRUE, }

formula <-"b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

fit <- gbm(as.formula(formula), data = all10mice,
                                   distribution = "bernoulli", n.trees = 10000,
                                   shrinkage = 0.001, interaction.depth = 5,
                                   n.minobsinnode = 20)

comments <- c(
        Discipline = "striaghtforward",
        pc.pro = "% of hours completed by an engineer on a project",
        b.timespan.cbrt = "timespan category - timespan of a project from first hour logged til last",
        no.users = "number of employees that worked on a project",
        b.inv.log = "total invoiced amount for a project ($) - ignore x-axis title",
        client.totinv.log = "average amount invoiced in total for a client across all jobs with them. 
        You need to take the natural exponent of the numbers on the x axis (i.e. e^6 = $403)",
        Business = "business category of the client, i.e. architect",
        majority.pos = "position of employee that completed the majority number of hours on a project.
        'tech' refers to drafters, 'pro' refers to engineers",
        pc.majpos.log = "% of hours completed by the main employee on the project. You need to take the natural 
        exponent of these values on the x-axis, i.e. e^4 = 55%",
        JD.second = "category of the project, i.e. subdivisions",
        Billing.Type = 'self explanatory'

)

# partial dependency plots
for(i in 1:11){
        plot.gbm(fit, i, type="response")
        print(fit$var.levels[[i]])
        print(paste("\n"))
        print(comments[i])
}

```

### 2016-12-13 1st_stats.R

@@ -1,0 +1,344 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
+
+all6b<- read.csv('all6b.csv')[,-1]
+all6b$Start.Date<- as.Date(all6b$Start.Date)
+
+#delete milestone rows! sort by highest hours - choose this discipline.!
+#sort by hours first
+all6b<- all6b %>% arrange(mlsto, -hours)
+dup<- all6b[,names(all6b) %in% c('mlsto')]
+all7<-all6b[!(all6b$inv.mlsto==0 | duplicated(dup)),]
+
+write.csv(all7, 'all7.csv')
+all7<- read.csv('all7.csv')[,-1]
+all7$Start.Date<- as.Date(all7$Start.Date)
+
+#is return.pdol and inv.vs.cost normally distributed??????
+#use qqnorm and qqline
+library('car')
+library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
+
+qqnorm(all7$return.pdol)
+qqline(all7$return.pdol)
+qqPlot(all7$return.pdol)
+
+# up tail at end. try log of variable
+qqnorm(log(abs(all7$return.pdol+.001)))
+qqline(log(abs(all7$return.pdol+.001)))
+#this looks s shaped
+
+qqnorm(log(all7$return.pdol+.8))
+qqline(log(all7$return.pdol+.8))
+
+#find sd and mean of return.pdol
+sd(all7$return.pdol)
+mean(all7$return.pdol)
+skewness(all7$return.pdol)
+skewness(log(all7$return.pdol+2.15))
+#skewness is 27.2 and 2.12 for logged data!!!
+
+#have a look at the shape of return.pdol as bar graph
+ggplot(all7, aes(x=return.pdol)) + geom_bar(stat='bin')
+ggplot(all7, aes(x=log(all7$return.pdol+.8))) + geom_bar(stat='bin')
[…]

### 2016-12-13 BNprep.R

@@ -1,0 +1,243 @@
+## prep for bayesian network
+
+#discretise numeric variables and change all 'NA' to blank
+
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
+
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+all8a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all8a.csv')[,-1]
+head(all8a)
+
+#discretise numeric columns!!!! pc.pro, timespan.cbrt, inv.mlsto.log, client.totinv.log,
+#pc.majpos.log, return.pdol
+
+###
+#pc.pro
+x<- dist(all8a$pc.pro, method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+rect.hclust(fit, k = 5)
+
+#assign a cluster number to all rows in data frame
+all8a$b.pc.pro<- as.factor(cutree(fit, k=5))
+head(all8a$b.pc.pro)
+
+for(i in 1:5){
+        print(
+        summary(all8a %>% select(b.pc.pro, pc.pro) %>% filter(b.pc.pro ==i))
+        )
+}
+
+#create clusters based on summary:
+all8a[all8a$pc.pro>100,'pc.pro']<- 100
+all8a$b.pc.pro<- cut(all8a$pc.pro, breaks=c(0,15,40,75,95,100), include.lowest=TRUE,
+                     labels= c("p0_15", "p15_40", "p40_75", "p75_95", "p100"))
+
+###
+#timespan.cbrt
+# all timespan == 0 should be timespan ==1. Checked all7a and they all have at least 0.5 hour booked, max 18 hours.
+all8a$timespan.cbrt <- ifelse(all8a$timespan.cbrt ==0, 1, all8a$timespan.cbrt)
+first.vec<- all8a$timespan.cbrt
+x<- dist(first.vec, method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+k.choose=3
+rect.hclust(fit, k = k.choose)
+
[…]

### 2016-12-13 Business_compact.R

@@ -1,0 +1,56 @@
+# re-jig Business
+
+str(all9a)
+
+B.invest<- table(all9a$b.rpdol, all9a$Business) %>% t %>% as.data.frame.matrix()
+colnames(B.invest) <- c("profit", "loss")
+B.invest<- transform(B.invest, ratio = round(loss/profit, 2))
+B.invest$JD <- rownames(B.invest) 
+B.invest<- arrange(B.invest, ratio) %>% select(JD, profit, loss, ratio)
+
+#group:
+
+# university, school
+# lawyer/body corp
+# signs/manufacturer
+# environment/water
+# roads/rail/gov/util
+# artist/landscape arch
+# hospital/institution/health
+# utilities
+# engineer
+# water
+# business/building services/internal
+# builder
+# architect
+# townplanner
+# developer/realestate
+# resources
+# person
+
+
+
+#code to compact:
+
+all9a$test<- as.character(all9a$Business)
+all9a$test <- ifelse(grepl("lawyer|body", all9a$Business), 
+                     "lawyer_bodyC", all9a$test)
+all9a$test <- ifelse(grepl("univ|school", all9a$Business), 
+                     "uni_school", all9a$test)
+all9a$test <- ifelse(grepl("sign|manuf", all9a$Business), 
+                     "sign_manufac", all9a$test)
+all9a$test <- ifelse(grepl("environ|water", all9a$Business), 
+                     "enviro_water", all9a$test)
+all9a$test <- ifelse(grepl("road|gov|util", all9a$Business), 
+                     "road_rail_gov_util", all9a$test)
+all9a$test <- ifelse(grepl("artist|landscape", all9a$Business), 
+                     "artist_landarch", all9a$test)
+all9a$test <- ifelse(grepl("hospit|institut|health", all9a$Business), 
+                     "hosp_institut_health", all9a$test)
+all9a$test <- ifelse(grepl("business|building|internal", all9a$Business), 
+                     "biz_bldgserv_internal", all9a$test)
+
+summary(as.factor(all9a$test))
+
+
+

### 2016-12-13 SS/1stplots.R

@@ -1,0 +1,191 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+
+all6<- read.csv('all6.csv')[,-1]
+all6$Start.Date<- as.Date(all6$Start.Date)
+#check extremes
+all6 %>% arrange(-return.pdol) %>% select(return.pdol, Tot.Invoiced, mlsto, hours) %>% head(20)
+#check specific job numbers
+all6 %>% select(return.pdol, Tot.Invoiced, mlsto, hours, cost.mlsto, dis.sc.mlsto) %>% filter(mlsto == '2013.430.4')
+
+#all7 has deleted all milestone duplicates and deleted outliers
+all7<- read.csv('all7.csv')[,-1]
+all7$Start.Date<- as.Date(all7$Start.Date)
+all7<- all7 %>% filter(!(inv.vs.cost<0.2), !(inv.vs.cost>4))
+
+
+#profitability
+#discipline vs profit
+#subset out invoiced=0
+dup<- all6[,names(all6) %in% c('mlsto', 'Discipline')]
+subd1<-all6[!all6$Tot.Invoiced==0 & !is.na(all6$Discipline) & !duplicated(dup),] 
+#make outlier points match variable colour
+update_geom_defaults("point", list(colour = NULL))
+bal.discplot<- ggplot(subd1, aes(x=Discipline, y= balance.mlsto,
+                                     colour= Discipline)) + 
+        geom_boxplot(alpha=0) +geom_jitter(alpha= I(1/2))+
+        scale_y_continuous(breaks = seq(round(min(subd1$balance.mlsto),-5), round(max(subd1$balance.mlsto),-5), by = 20000))
+
+prof.discplot<- ggplot(subd1, aes(x=Discipline, y= profit.mlsto,
+                                 colour= Discipline)) + 
+        geom_boxplot(alpha=0) +geom_jitter(alpha= I(1/2))+
[…]

### 2016-12-13 combine.R

@@ -1,0 +1,75 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+
+##### COMBINE CLEANED DATA SETS BEFORE DE-IDENTIFICATION #########
+
+#combining all dataset with Job hours data set and Inv_sum data set.
+#Job.Hours data set columns need to replace some of the all columns because they are more accurate. 
+# these include: Start.Date, tot.TS.Charge, tot.TS.cost, tot.TS.hours, Year
+
+all4e<- read.csv('all4e.csv')[,-1]
+str(all4e)
+all4e<- all4e %>% filter(!(Job.Number == '2010.041.1' & dist == 1362.8))
+#delete Start.Date, tot.TS.Charge, tot.TS.cost, tot.TS.hours, Year
+all4e<- all4e %>% select(-Start.Date, -Tot.TS.Cost, -Tot.TS.Charge, -TS.Hours, -Year, -Profit, -Tot.Invoiced)
+
+#combine Inv.Sum
+Inv.sum<- read.csv('inv_sum.csv')[,-1]
+all4e<- merge(all4e, Inv.sum %>% select(Job.Number, Tot.Invoiced), by= 'Job.Number', all.x=TRUE, all.y=FALSE)
+#set NA = 0
+all4e[is.na(all4e$Tot.Invoiced),]$Tot.Invoiced <- 0
+
+#combine Job.Hours
+Job.Hours<- read.csv('Job_hours.csv')[,-1]
+Job.Hours<- Job.Hours %>% select(-mlsto)
+all5<- merge(all4e, Job.Hours, by.x='Job.Number', by.y= 'Job.num', all.x=TRUE, all.y=FALSE)
+str(all5)
+
+#create new 'profit' and 'balance' column
+#profit column
+all5<- transform(all5, profit= Tot.Invoiced-charge-Dis.subcon)
+#balance column - subtract real costs
+all5<- transform(all5, balance= Tot.Invoiced-cost-Dis.subcon)
[…]

### 2016-12-13 combine2.R

@@ -1,0 +1,168 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+library('rCharts',lib='C:/Progra~1/R/R-3.2.0/library')
+
+invoices<- read.csv('invoices_clean.csv')[,-1]
+hours<- read.csv('hours_clean.csv')[,-1]
+
+#### CREATING SHINY PRODUCT - MONTHLY REVIEW
+
+### STEP 0
+# subset invoices df and hours df to 2009 onwards 
+invoices$Invoice.Date <- as.Date(invoices$Invoice.Date)
+invoices<- arrange(invoices, Invoice.Date)
+invoices<- invoices %>% filter(Invoice.Date>= as.Date('2009-01-01'))
+
+hours$Work.Date <- as.Date(hours$Work.Date, "%d/%m/%Y")
+hours<- hours %>% filter(Work.Date>= as.Date('2009-01-01'))
+
+### STEP 0.5
+# assign tag name to all rows between dates
+# use last date in month
+invoices$MonthYear<- format(invoices$Invoice.Date,"%Y-%b")
+tag<- ddply(invoices, .(MonthYear), tail, 1)
+tag<-tag %>% arrange(Invoice.Date)
+
+df<- data.frame('start' = c(as.Date('2008-12-31'), tag$Invoice.Date[-nrow(tag)]),'end'= tag[,6], 'tag' = tag[,10])
+
+
+## try ifelse function --> similar to grepping 
+#assign tag to all hour rows based on whether they are between two dates in the df data.frame (see above)
+#these dates are determined by the last date of invoices in each month
+
+btwn.dates<- function(date.df = df, col.to.tag = hours$Work.Date){
+        a= c(rep(NA,length(col.to.tag)))
+        for( i in 1:nrow(date.df)){
+                a = ifelse(col.to.tag >= date.df[i,1] & col.to.tag < date.df[i,2], as.character(date.df$tag[i]), a) 
+        } 
+        a<- as.factor(a)
+        a
+}
[…]

### 2016-12-13 de_identify.R

@@ -1,0 +1,319 @@
+####### DEIDENTIFYING DATA
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("magrittr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+
+
+
+#have added distances to all4
+all4<- read.csv('all4.csv')
+#de-identify data
+
+#codify client2, Client.contact, job.number, Director, Project.Engineer, Engineer.2, Suburb
+#code client2
+code.client2<- data.frame(client2=unique(all4$client2))
+code.client2$code<- paste('C', sample(2000:3000, nrow(code.client2), replace=F), sep="")
+write.csv(code.client2,'code_client2.csv')
+#code Client.Contact
+code.contact<- data.frame(contact=unique(all4$Client.Contact))
+code.contact$code<- paste('CN', sample(1000:2000, nrow(code.contact), replace=F), sep="")
+write.csv(code.contact,'code_contact.csv')
+#code Job.Number
+# need to separate milestones
+code.job<- data.frame(job=unique(all4$Job.Number))
+#have a look at job numbers in order
+code.job<- code.job[order(code.job$job),]
+#split off the first way of numbering jobs
+code.job1<- data.frame(job1=code.job[1:2426])
+code.job2<- data.frame(job2=code.job[2427:length(code.job)])
+
+#code.job1 - split up job number into 2010.409 and 1 - 2 separate columns
+code.job1$job1<- as.character(code.job1$job1)
+a<-strsplit(code.job1$job1[1],split='[.]')[[1]]
+paste(a[1],a[2],sep='.')
+code.job1$split<-sapply(code.job1$job1,FUN=function(x){strsplit(x,split='[.]')[[1]]})
+code.job1$first<-sapply(code.job1$split,FUN=function(x){paste(x[1],x[2],sep=".")})
+code.job1$second<-sapply(code.job1$split,FUN=function(x){x[3]})
+# need to codify 'first' and 'second' columns separately, then combine
+#first do 'first' column
+First<- data.frame(first=unique(code.job1$first))
+First$code<- paste('J', sample(1000:3000, nrow(First), replace=F), sep="")
+#now code for 'second' column
[…]

### 2016-12-13 final_app/server.R

@@ -1,0 +1,270 @@
+if(.Platform$OS.type == 'windows'){
+        setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+        all7<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7.csv')[,-1]
+        all7d<- read.csv('all7d.csv')[,-1]
+        
+} else{
+        setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+        all7<- read.csv('all7.csv')[,-1]
+        all7d<- read.csv('all7d.csv')[,-1]
+        
+}
+library('ggplot2')
+library("dplyr")
+library("plyr")
+library('magrittr')
+library('reshape2')
+library('shiny')
+library('shinythemes')
+library('FNN')
+# library('ggplot2')
+# library("dplyr")
+# library("plyr")
+# library('magrittr')
+# library('reshape2')
+# library('shiny')
+# library('shinythemes')
+
+
+all7$Start.Date<- as.Date(all7$Start.Date)
+#delete anything more than 4 for inv.vs.cost
+#delete anything less than .2 for inv.vs.cost
+all7<- all7 %>% filter(!(inv.vs.cost<0.2), !(inv.vs.cost>4))
+
+#data for final plot
+df <- read.table(text = " id  min  max 
+    Sp1     150          151     ", header=TRUE)
+
+#make a new dataframe with information about crossbar sizes
+df2<- data.frame(id=c(rep("Sp1",1)), ymin= c(-8.8), y.mean = c(-7),
+                 ymax=c(-5.2), confidence=c('90%'))
+df2 <- within(df2, 
+              confidence <- factor(confidence, 
+                                    levels=c('90%')))
+prof.labels = data.frame(xx = 0.53, yy = seq(-10,10,by= 10), 
+                         text = c("Loss-Making Job","Neutral","Profitable Job"))
+
+#### CREATE TABLE DATA FRAME FOR NEAREST NEIGHBOURS ####
+
+all7d$mlsto<- as.character(all7d$mlsto)
+
+# test.case<- all7d %>% select(Discipline, Business, Biz.type, code.contact, code.client, JD.Second, Billing.Type)
+
+#function for retrieving nearest neighbours by inv.mlsto:
+inv.knn<- function(df= all7d, predict= 'inv.mlsto', new.cases= c(1000), k=3){
+        invo.knn = knn(df[,predict], df[new.cases,predict],
+                       df$pc.pro, k=k)
[…]

### 2016-12-13 final_app/ui.R

@@ -1,0 +1,252 @@
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+# library('shiny')
+
+shinyUI(fluidPage(
+        theme=shinytheme("journal"),
+        
+        #Application title
+        titlePanel("Quote Assist"),
+        
+        # Sidebar with controls to select the variable to plot against positions
+        wellPanel(
+        fluidRow(
+                column(3,
+        
+                selectInput("disc", "Discipline:",
+                            list("Civil" = "Civil", 
+                                 "Structural" = "Structural", 
+                                 "Environmental Planning" = "Environmental Planning",
+                                 "Water" = "Water",
+                                 "All"= "All")),
+                
+                selectInput("billing", "Billing Type:",
+                            list('Hourly Rate' = 'Hourly Rate',
+                                 "Fixed Fee"= 'Fixed Quote'
+                                 )),
+                
+                selectizeInput(
+                        'JD.Second', 'Job Type:', 
+                        choices = levels(all7$JD.Second),
+                        options = list(
+                                placeholder = 'Please start typing',
+                                onInitialize = I('function() { this.setValue(""); }'),
+                                maxItems=1
+                        ))
+                
+                ),
+                
+                column(3,
+                       
+                       selectizeInput(
+                               'code.client', 'Client:', choices = levels(all7$code.client),
+                               options = list(
+                                       placeholder = 'Please start typing',
+                                       onInitialize = I('function() { this.setValue(""); }'),
+                                       maxItems=1
+                               )),
+                       
+                       selectizeInput(
+                               'Business', 'Client Industry:', 
+                               list("Architect" = "architect", 
+                                    "Artist" = "artist", 
+                                    "Builder" = "builder",
+                                    "Building Services"= 'building services',
+                                    "Business" = "business",
+                                    "Council" = 'council',
+                                    "Developer/Real Estate"= 'developer/real estate',
+                                    "Engineer"= 'engineer',
+                                    "Environment" = 'environment',
+                                    "Friend" = 'friend',
[…]

### 2016-12-13 final_plot.R

@@ -1,0 +1,111 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
+
+theData <- data.frame('category' = sample(LETTERS[1:3], 1000, replace = T),
+                      'value' = rnorm(1000))
+theData<-theData %>% filter(category=='A') 
+
+ggplot(theData, aes(x = category, y = value)) +
+        stat_ydensity(geom="segment", aes(xend=..x..+..scaled../3, 
+                                          yend=..y.., alpha=..scaled..), size=1, trim=FALSE) +
+        stat_ydensity(geom="segment", aes(xend=..x..-..scaled../3, 
+                                          yend=..y.., alpha=..scaled..), size=1, trim=FALSE) +
+        scale_alpha_continuous(range= c(0, 1)) +
+        theme_bw()
+
+p <- ggplot(theData, aes(category, value)) + 
+        geom_tile(aes(fill = value)) + 
+        scale_fill_gradient(low = "white", high = "steelblue")
+
+p <- ggplot(theData, aes(category, value)) + 
+        stat_ydensity(aes(xend=..x..+..scaled../3, 
+                          yend=..y.., alpha=..scaled../3), geom="tile", position="identity")+
+        geom_violin(aes(alpha=0))
+
+
+ggplot(mtcars, aes(factor(cyl), mpg))+ 
+        geom_violin(fill = wt, colour = "#3366FF")
+
+### TRY CUSTOMISING BOXPLOT TO CREATE FAN LOOK
+
+
+
+pos.plot3<- ggplot(theData, aes(x=category, y= value,
+                              colour= category)) +
+        
+        stat_summary(fun.y= mean, geom='line') + 
+        stat_summary(fun.y= mean, geom='point') +
+        #stat_summary(fun.data= mean_sdl, geom='crossbar', mult =1) +
+        theme(legend.position="right", axis.text.x=element_text(angle=45,hjust=1),
+              text=element_text(size=12))
[…]

### 2016-12-13 functions.R

@@ -1,0 +1,531 @@
+#function for running caret..
+
+caret.all <- function(method = "boost", df = train, formula = formula.rf, sample.frac = 0.75, seed = 100){
+        # set seeds
+        set.seed(seed)
+        seeds <- vector(mode = "list", length = 51)
+        for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
+        ## For the last model:
+        seeds[[51]] <- sample.int(1000, 1)
+        
+        train.c = sample_frac(df, sample.frac, replace = FALSE)
+        test.c = setdiff(df, train.c)
+        
+        cv.Control <- trainControl(method = "cv",
+                                   number = 5,
+                                   seeds= seeds
+                                   , summaryFunction= twoClassSummary,
+                                   classProbs=TRUE
+                                   )
+        
+        if(method == "rf"){
+                mtry.Grid <- expand.grid(mtry = c(3,4,5,6,7))
+                
+                set.seed(2)
+                gbmFit <- caret::train(as.formula(formula), data= train.c,
+                                       method = "rf", trControl = cv.Control, verbose = FALSE,
+                                       tuneGrid = mtry.Grid, ntree = 1000,
+                                       metric = 'ROC', allowParallel = T)
+        }
+        
+        if(method == "boost"){
+                train.c$b.rpdol<- as.factor(train.c$b.rpdol)
+                levels(train.c$b.rpdol)[levels(train.c$b.rpdol)=="0"] <- "profit"
+                levels(train.c$b.rpdol)[levels(train.c$b.rpdol)=="1"] <- "loss"
+                Grid <- expand.grid(shrinkage = c(.0005),
+                                         n.trees = c(6000),
+                                    interaction.depth = c(3,4,5,6),
+                                    n.minobsinnode = c(10,20,30,40))
+                
+                # , interaction.depth = c(1,2,3,4,5),
+                # n.minobsinnode = c(10,20,30)
+                
+                set.seed(2)
+                gbmFit <- caret::train(as.formula(formula), data= train.c,
+                                method = "gbm", trControl = cv.Control, verbose = FALSE,
+                                bag.fraction = 0.5, tuneGrid = Grid,
+                                metric = 'ROC',
+                                na.action = na.pass)
+        }
+        
+        plot(gbmFit)
+        return(gbmFit)
+        
+}
+
+cfor.plotsave<- function(df= vars, name= 'cfor_blend', formular = formula.rf){
+                
+                #run cforest
[…]

### 2016-12-13 hours_cleaning.R

@@ -1,0 +1,211 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+hours12<- read.csv('hours0212.csv', na.strings= c(""," ", "NA"))[,-1]
+hours1314<- read.csv('20132014hoursdata.csv', na.strings= c(""," ", "NA"))
+
+#ensure column orders and names are correct
+hours1314<- hours1314[,c(13,1:12)]
+colnames(hours12)[names(hours12) %in% 'Job.Number']<-'Job.num'
+colnames(hours12)[names(hours12) %in% 'Discipline.1']<-'Discipline'
+
+
+#what is WIP only invoice??
+
+#clean up
+# any columns unnecessary?
+summary(hours1314)
+#not really
+#remove NA's from hours1314 first. look into full NA rows
+dim(hours)
+hours1314[rowSums(is.na(hours1314))== 13,] %>% nrow()
+hours1314<-hours1314[rowSums(is.na(hours1314))!=12,]
+hours1314<-hours1314[rowSums(is.na(hours1314))!=13,]
+#now to rows with 3 values
+hours1314[rowSums(is.na(hours1314))== 9,] %>% head()
+#are all the charge/cost values 0?? yes - only details - delete these rows
+hours1314<-hours1314[rowSums(is.na(hours1314))!=9,]
+#investigate last 9 rows of NA in charge
+hours1314[is.na(hours1314$Charge.Amount..B...AUD.),] %>% head()
+#these were full NA rows
+hours1314<-hours1314[rowSums(is.na(hours1314))!=13,]
+
+#remove NA's from hours12 now. look into full NA rows
+summary(hours12)
+hours12[rowSums(is.na(hours12))== 12,] %>% head()
+hours12<-hours12[rowSums(is.na(hours12))!=13,]
+hours12<-hours12[rowSums(is.na(hours12))!=12,]
+#now to rows with 3 values
+hours12[rowSums(is.na(hours12))== 9,] %>% head()
[…]

### 2016-12-13 imputingall9c.R

@@ -1,0 +1,122 @@
+# impute missing values
+
+library('mice')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
+library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')
+library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
+library('CALIBERrfimpute', lib = 'C:/Progra~1/R/R-3.2.2/library')
+
+all9c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9c.csv')[,-1]
+all9c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all9c.csv')[,-1]
+all9c$b.rpdol<- as.factor(all9c$b.rpdol)
+
+all10rf<- rfImpute(b.rpdol ~ Discipline + pc.pro + timespan.cbrt + no.users + inv.mlsto.log +
+                 client.totinv.log + pc.majpos.log + majority.pos + Business + JD.Second + Billing.Type,
+                 data= all9c,
+              ntree= 500)
+all10rf<- cbind(all10rf, all9c[,names(all9c) %in% c('code.client', 'code.contact', 'mlsto',
+                                                    'b.timespan.cbrt', 'b.rpdol')])
+
+
+## test if it helps boosted trees
+
+#convert predictor variable back to numeric 0,1
+
+levels(all10rf$b.rpdol)<- c("0", "1")
+all10rf$b.rpdol<- as.character(all10rf$b.rpdol)
+all10rf$b.rpdol<- as.integer(all10rf$b.rpdol)
+write.csv(all10rf, 'C:/Users/n9232371/Documents/Consultbusiness/data/all10rf.csv')
+
+levels(all10mice$b.rpdol)<- c("0", "1")
+all10mice$b.rpdol<- as.character(all10mice$b.rpdol)
[…]

### 2016-12-13 invoice_eng.R

@@ -1,0 +1,136 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.0/library')
+
+
+Invoices.a<-read.csv('invoices_clean.csv')[,-1]
+Inv.sum<-read.csv('inv_sum.csv')[,-1]
+
+
+#### ENGINEERING VARIABLES ######
+
+#need to do by milestone!! because this is what relates directly to the return.pdol variable which we are measuring as 'success'
+
+#Need to sum up all the profits and balances for jobs with milestones
+# first 
+#create subset of just the things i want to add up - profits, balance, hours in each milestone
+temp<- Invoices.a %>% select(Job.Number, Inv.Amount, Invoice.Date)
+temp<- temp %>% arrange(Job.Number)
+temp$Invoice.Date <- as.character(temp$Invoice.Date)
+temp$Job.Number <- as.character(temp$Job.Number)
+#create column for Date of prev invoice, prev job number, and prev invoiced amount
+temp$testp.Invoice<- c(NA, temp$Invoice.Date[-nrow(temp)])
+temp$prev.JN<- c(NA, temp$Job.Number[-nrow(temp)])
+temp$prev.inv<- c(NA, temp$Inv.Amount[-nrow(temp)])
+#new column which prints 'NA" if prev job.number is not the same as current OR if prevInv <=0
+temp$prev.Invoice<- ifelse(temp$Job.Number == temp$prev.JN & temp$prev.inv>0 & temp$Inv.Amount>0, temp$testp.Invoice, NA)
+temp<- temp %>% select(-testp.Invoice, -prev.JN, -prev.inv)
+temp$Invoice.Date<- as.Date(temp$Invoice.Date)
+temp$prev.Invoice<- as.Date(temp$prev.Invoice)
+
+#create new column for milestone prefix
+temp1<- temp[1:7680,]
[…]

### 2016-12-13 monthly_ggvis/server.R

@@ -1,0 +1,72 @@
+# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+# library('shiny')
+
+
+# month.a<- read.csv('monthly.csv')[,-1]
+# month.a$Date<- as.Date(month.a$Date)
+# month.a$return.cut <- cut_number(month.a$return.mon,15)
+levels(month.a$return.cut) =c("< -$0.51", "-$0.51 to -$0.24", "-$0.24 to -$0.02","-$0.02 to $0.18",
+                              "$0.18 to $0.33", "$0.33 to $0.45", "$0.45 to $0.50", "$0.50 to $0.61",
+                              "$0.61 to $0.75", "$0.75 to $0.92", "$0.92 to $1.20", "$1.20 to $1.68", 
+                              "$1.68 to $2.47", "$2.47 to $5.24", "> $5.24")
+
+#shiny function
+
+shinyServer(
+        function(input,output){
+                
+                # Compute the forumla text in a reactive function since it is 
+                # shared by the output$caption and output$mpgPlot functions
+                formulaText<- reactive({
+                        paste("Monthly Summary", input$disc, sep=" ")
+                })
+                
+                # Return the formula text for printing as a caption
+                output$caption <- renderText( {
+                        formulaText()
+                })
+                
+                
+                #check for the input variable
+                data.r= reactive ({
+                        if(input$disc == 'All') {
+                                posData<- month.a}
+                        
+                        else{
+                                posData<- month.a %>% filter(Discipline == input$disc)
+                        }
+                        
+                })
+        
+                
+                #subset data based on slider inputs for Date.
+
+                data.r2 = reactive({
+                        a = subset(data.r(), Date >= input$date[1] & Date <= input$date[2])
+                        return(a)
+                })
+
+                
+                #generate a plot of requested variable
+                
+                
+                gv<- reactive( {
+
+                                               
+                        dd= data.r2()
+                        
[…]

### 2016-12-13 monthly_review rCharts/server.R

@@ -1,0 +1,57 @@
+
+
+
+shinyServer(function(input, output) {
+        
+        # Compute the forumla text in a reactive function since it is 
+        # shared by the output$caption and output$mpgPlot functions
+        formulaText1<- reactive({
+                paste("Monthly Summary for", input$disc, sep=" ")
+        })
+        
+        # Return the formula text for printing as a caption
+        output$caption <- renderText( {
+                formulaText1()
+        })
+        
+        formulaText <- reactive({
+                paste(input$y, "~", "return.cut")
+        })
+        
+        #check for the input variable
+        data.r= reactive ({
+                if(input$disc == 'All') {
+                        posData<- month.a}
+                
+                else{
+                        posData<- month.a %>% filter(Discipline == input$disc)
+                }
+                return(posData)
+                
+        })
+        
+        #subset data based on slider inputs for Date.
+        
+        data.r2 = reactive({
+                a = subset(data.r(), Date >= input$date[1] & Date <= input$date[2])
+                return(a)
+        })
+
+        
+        output$myChart <- renderChart({
+
+                p1 <- nPlot(as.formula(formulaText()), group = input$variable, data = data.r2(),
+                            type = 'multiBarChart')
+                
+                p1$addParams(dom = 'myChart')
+                p1$xAxis( rotateLabels = -45, axisLabel = "Return per Dollar (Balance/Cost)")
+                p1$chart( xDomain = levels(month.a$return.cut),
+                          reduceXTicks = FALSE)
+                p1$params$width <- 800
+                p1$params$height <- 350
+                
+
+                return(p1)
+                
+        })
+})

### 2016-12-13 monthlyreviewplotly/server.R

@@ -1,0 +1,106 @@
+# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
+library('ggplot2')
+library("dplyr")
+library("plyr")
+library('magrittr')
+library('reshape2')
+library('shiny')
+
+
+
+# month.a<- read.csv('monthly.csv')
+month.a$Date<- as.Date(month.a$Date)
+
+levels(month.a$return.cut) =c("< -&#xfe69;0.51", "-&#xfe69;0.51 to -&#xfe69;0.24", "-&#xfe69;0.24 to -&#xfe69;0.02","-&#xfe69;0.02 to &#xfe69;0.18",
+                              "&#xfe69;0.18 to &#xfe69;0.33", "&#xfe69;0.33 to &#xfe69;0.45", "&#xfe69;0.45 to &#xfe69;0.50", "&#xfe69;0.50 to &#xfe69;0.61",
+                              "&#xfe69;0.61 to &#xfe69;0.75", "&#xfe69;0.75 to &#xfe69;0.92", "&#xfe69;0.92 to &#xfe69;1.20", "&#xfe69;1.20 to &#xfe69;1.68", 
+                              "&#xfe69;1.68 to &#xfe69;2.47", "&#xfe69;2.47 to &#xfe69;5.24", "> &#xfe69;5.24")
+#order month.a according to return.cut levels
+month.a<- month.a %>% arrange(return.cut)
+
+
+
+
+#                         gg$layout<- list(fileopt='overwrite', filename='R-Cookbook/axes/reversed ordered axes - 1')
+
+#shiny function
+
+shinyServer(
+        function(input,output, session){
+                
+                # Compute the forumla text in a reactive function since it is 
+                # shared by the output$caption and output$mpgPlot functions
+                formulaText<- reactive({
+                        paste("Monthly Summary", input$disc, sep=" ")
+                })
+                
+                # Return the formula text for printing as a caption
+                output$caption <- renderText( {
+                        formulaText()
+                })
+                
+                #check for the input variable
+                data.r= reactive ({
+                        if(input$disc == 'All') {
+                                posData<- month.a}
+                        
+                        else{
+                                posData<- month.a %>% filter(Discipline == input$disc)
+                        }
+                        
+                })
+        
+                #subset data based on slider inputs for Date.
+
+                data.r2 = reactive({
+                        a = subset(data.r(), Date >= input$date[1] & Date <= input$date[2])
+                        return(a)
+                })
+
+                #generate a plot of requested variable
+                
+                
[…]

### 2016-12-13 percent.pos/ui.R

@@ -1,0 +1,35 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+library('shiny')
+
+shinyUI(pageWithSidebar(
+        #theme="amelia.css",
+        
+        #Application title
+        headerPanel("Team Ratios"),
+        
+        # Sidebar with controls to select the variable to plot against positions
+        sidebarPanel(
+                selectInput("disc", "Discipline:",
+                            list("Civil" = "Civil", 
+                                 "Structural" = "Structural", 
+                                 "Environmental Planning" = "Environmental Planning",
+                                 "Water" = "Water",
+                                 "All"= "All"))
+                
+        ),
+        
+        #main panel. show plot of each position and percentage within selected variable
+        mainPanel(
+                h3(textOutput("caption")),
+                
+                plotOutput("pos.plot")
+                
+        )
+        
+                          
+))

### 2016-12-13 postcodes.R

@@ -1,0 +1,365 @@
+#### create file with distances from Brisbane city using latitude/longitude coords
+
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('fields',lib='C:/Progra~1/R/R-3.1.3/library')
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+all3<- read.csv('all3.csv')
+coords<- read.csv('postcodes.csv')
+
+#merge all3 postcodes with coords
+#create temp file with only post codes and suburbs and states from all3
+str(all3)
+temp<- all3[, names(all3) %in% c('Post.Code', 'Suburb', 'State')] %>% head()
+temp$Post.Code<- as.integer(as.character(temp$Post.Code))
+#very confusing because so many suburb names match the post code
+#need to merge original databases using postcode AND suburb name
+#therefore need suburb names to be in capital letters. use toupper()
+temp$Suburb<- toupper(temp$Suburb)
+temp$Suburb<- as.factor(temp$Suburb)
+#temp<- temp[temp$Post.Code>0,]
+#create temp file with only post codes and suburbs and states from coords, and coords
+temp2<- coords[, names(coords) %in% c('postcode', 'suburb', 'state', 'lat', 'lon')]
+head(temp2)
+#merge the two
+PC<- merge(temp, temp2, by.x= c('Post.Code', 'Suburb'), by.y= c('postcode','suburb'), all.x=TRUE)
+PC[400:500,]
+#cross check with all3 
+all3[all3$Suburb %in% 'Lismore', names(all3) %in% c('Suburb', 'Post.Code')]
+#looks fine
+#delete rows with 3 NA's in a single row AND Post.Code = 0
+PC<-PC[!(PC$Post.Code == 0 & rowSums(is.na(PC[,c(2,4:6)]))==4),]
+#any rows of full NA left?
+PC<- PC[!(rowSums(is.na(PC))==6),]
+ggplot(PC, aes(x=Post.Code)) + geom_point(stat='bin')
+# yes, deleted them
+#want unique cases of each suburb
+PC<- unique(PC)
+# want to trim down to first entry for each post code number
+#create a function
+
[…]

### 2016-12-13 radarplot.R

@@ -1,0 +1,119 @@
+library('plotrix', lib = 'C:/Progra~1/R/R-3.2.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
+library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
+
+all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
+all7d$Post.Code<- as.factor(all7d$Post.Code)
+
+#lets look at case 960
+#most important variables were:
+# no.users, inv.mlsto.log, pc.pro, timespan, majority.pos, Discipline
+#make a radar chart with no.users, inv.mlsto, pc.pro, client.code, majority.pos
+
+#create a new column for profitable or not
+
+all7d$good<- ifelse(all7d$return.pdol<0, 'bad','good')
+all7d$good<- as.factor(all7d$good)
+
+#create invoiced amount categories
+all7d$Inv.cat<- rep(NA, nrow(all7d))
+all7d$Inv.cat<- ifelse(all7d$inv.mlsto <5000 , '<$5k', all7d$Inv.cat)
+all7d$Inv.cat<- ifelse(all7d$inv.mlsto <20000 & all7d$inv.mlsto >=5000, '$5-20k', all7d$Inv.cat)
+all7d$Inv.cat<- ifelse(all7d$inv.mlsto <60000 & all7d$inv.mlsto >=20000, '$20-60k', all7d$Inv.cat)
+all7d$Inv.cat<- ifelse(all7d$inv.mlsto >=60000 , '>$60k', all7d$Inv.cat)
+all7d$Inv.cat<- as.factor(all7d$Inv.cat)
+all7d <- within(all7d, 
+                Inv.cat <- factor(Inv.cat,levels=
+                                          c('<$5k', '$5-20k', '$20-60k', '>$60k')))
+
+#create pc.pro categories, use hierarchical clustering to find groups
+#Ward's clustering
+
+x<- dist(all7d %>% select(pc.pro), method = 'euclidean')
+fit<- hclust(x, method = 'ward.D2')
+plot(fit)
+
[…]

### 2016-12-13 team_ratio-catXaxis/server.R

@@ -1,0 +1,83 @@
+setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
+library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
+library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
+library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
+library('shiny')
+
+#load Data
+all6<- read.csv('all6.csv')[,-1]
+all6$Start.Date<- as.Date(all6$Start.Date)
+dup1<- all6[,names(all6) %in% c('mlsto')]
+sub10<-all6[!all6$Tot.Invoiced==0 & !duplicated(dup1),]
+sub10<- melt(sub10 %>% select(code.jobnum, pc.contracttech, pc.director, pc.gradpro, pc.midpro, pc.midtech, pc.seniorpro,
+                              pc.seniortech, pc.unknown),value.name='percent.hrs') %>% filter(percent.hrs>.5)
+sub10 <- merge(sub10, all6 %>% select(Start.Date, balance.mlsto, Discipline, code.jobnum, return.pdol), by= 'code.jobnum', all.x=TRUE, all.y=FALSE) %>%
+        arrange(Start.Date)
+sub10$Start.Date<- as.Date(sub10$Start.Date)
+sub10$return.cut<- cut_number(sub10$return.pdol, 10, labels= c(1:10)) %>% as.integer()
+
+#shiny function
+
+shinyServer(
+        function(input,output){
+                
+                # Compute the forumla text in a reactive function since it is 
+                # shared by the output$caption and output$mpgPlot functions
+                formulaText<- reactive({
+                        paste("Position percentages in", input$disc, sep=" ")
+                })
+                
+                # Return the formula text for printing as a caption
+                output$caption <- renderText( {
+                        formulaText()
+                })
+                
+                #check for the input variable
+                data.r= reactive ({
+                        if(input$disc == 'All') {
+                                posData<- sub10}
+                        
+                        else{
+                                posData<- sub10 %>% filter(Discipline == input$disc)
+                        }
+                        
+                })
+        
+                
+                #subset data based on slider inputs for Date.
+
+                data.r2 = reactive({
+                        a = subset(data.r(), Start.Date >= input$date[1] & Start.Date <= input$date[2])
[…]

### 2016-12-14 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd
finished lit review paper

@@
 
-Depending on the industry, some consulting projects can be financialy risky endeavors for the consultant.
+Many consulting projects can be financialy risky endeavors for the consultant, often depending on the industry.
 Engineering consultanting companies in the construction industry complete a wide range of complex projects in a competitive enviroment.
@@
 The purpose of this study was to test whether these loss making projects could be predicted by statistical and machine learning algorithms using the information available at project conception.
-The second question is whether the predictive algorithm could improve a company's profitability.
-Project data from a case study engineering company was used to answer these questions.
+Secondly, could the predictive algorithm improve a company's profitability?
+Predictive models and value frameworks were built using project data from a case study engineering company to answer these questions.
 
@@
 
-Financial risk most commonly falls upon a consulting company when they must offer a fixed fee for a project.
-In competitive climates, clients request fixed price quotes for project completion from several consultants before a project is awarded.
+Financial risk is taken on by a consulting company when they offer a fixed fee to complete a project.
+In competitive climates, clients are in the position to request fixed price quotes from several consultants before a project is awarded.
 Hence, initial quoted fees must be competitive but also accurately cover consultant salaries and business costs.
@@
 How do consultants make this accurate yet competitive fee estimate?
-This varies from industry to industry, but typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, comparison to past projects, and rules of thumb.
-Unfortunately, some projects priced using this method can blow out of budget, causing significant losses to the consultant.
+This varies from industry to industry, but typically a consulting manager has experience in the type of project they are quoting and, after reviewing the project details, can use a combination of intuition, analysis of project details, and rules of thumb.
+Unfortunately, some projects priced using this method can blow out of budget, causing significant financial losses to the consultant.
 
+Cost blowouts of complex projects has a history, particularly documented in the It and construction industry.
 A review of 6 IT project surveys between 1984 and 1994 by Moløkken & Jørgensen (2003) revealed that 60-80% of IT projects encountered effort and/or schedule overruns, where the average overrun was by 30 to 40%.
@@
 
-This work aims to provide a manager with this more realistic view by predicting whether the new project will be profitable or not, based on a suite of previous project data from the same company.
-The manager could then take multiple actions including not tendering for a project predicted to be loss-making, providing a high fixed price, or restructuring the contract so that it is not 100% fixed price.
+This work aims to help managers provide better fixed price quotes by predicting whether a new project will be profitable or not, based on an algorithm trained on all previous project data from the same company.
+The manager could take multiple actions based on the algorithms output including not tendering for a project predicted to be loss-making, providing a high fixed price, or restructuring the contract so that it is not 100% fixed price.
 This research tests the simple case of rejecting all projects predicted as loss-making.
@@
 
-This thesis focuses on a single case study consulting company in the construction industry.
-twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details.
-labour is the chief cost and the long traditional history of this industry mandates fixed price projects as the norm. 20% of projects are loss making.
-Fees calculated via detailed analysis
-A Customer Relationship Management software package (CRM) is currently employed to collect and store project data.
-A CRM is a popular type of software used by businesses to record client and project details as well as employee time sheet records.
-In the case study business, the CRM is available to all employees over the company intranet, and each employee completes daily time sheets allotting their hours to certain projects.
-Additionally, technical information is recorded against each project
+Project data from a single case study consulting company in the construction industry was used to build the predictive algorithm, trial several methods, and test their impact on the company's bottom line.
+Twelve years of passively collected data that described each project in terms of their clients, invoice history, employee hours and technical details was scraped from the company's internal database.
+Customer relationship management (CRM) software provided accesses to the database to all employees.
+Employees completed daily timesheets, alotting their hours to certian projects.
+Project managers input technical information and client details for each project, while the finance team tracked invoicing and additional project costs such as printing and transport in the same system.
+On top of this information, the CRM had the hourly cost of each level of employee.
+The cleaned dataset consisted of 2364 projects between 2004 and 2012.
+70% of projects were fixed price and 20% of all projects made a loss after taking into account business costs.
 
-* Employee time sheet hours with dates
-* Other project costs (taxis, printing)
-* Client information/characteristics
-* Client identification code
-* Invoiced amounts for each project and dates
-* Employee costs
-* Employee charge out rates
-* Project description
-
-2364 past projects are available.
-missing data due to how the data collection changed over time and some fields were not mandatory.
-
-To assess whether the model would improve the case study company's profitability, a simple managerial decision framework was applied: if the algorithm predicted the project to be loss making, the project was considered 'rejected' and all profits or *losses* of rejected projects were discarded.
+Once several methods were built, to assess whether the model would improve the case study company's profitability, a simple managerial decision framework was applied: if the algorithm predicted the project to be loss making, the project was considered 'rejected' and all profits or *losses* of rejected projects were discarded.
 The remaining projects' profits and losses were summed to give a revised overall profit over the last 12 years of projects.
@@
 
-- aim
-**General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
+<!-- **General Aim** Use statistical techniques to model the profitability of projects for consulting businesses using their internal CRM data.
 Research will focus on a case study Engineering consulting company that offers their expert advice (in the currency of time) to business clients.
-The project outcomes are intended to assist the business in predicting project profitability before engagement. Several statistical and machine learning techniques were tested, compared and refined.
+-->
+<!-- The project outcomes are intended to assist the business in predicting project profitability before engagement.
+Several statistical and machine learning techniques were tested, compared and refined.
+-->
 
-**Hypothesis 1**
-A statistical or machine learning model based on historical project data can predict the profitability of a new project with greater accuracy than a baseline predictor.
+<!-- **Hypothesis 1** -->
+<!-- A statistical or machine learning model based on historical project data can predict the profitability of a new project with greater accuracy than a baseline predictor.
 A baseline predictor for this project is one that predicts the average of a numeric response variable for all cases or, if the response variable is categorical, randomly assigned categories for each case, where the proportion of assigned categories matches the true categorical proportions.
+-->
 
-**Hypothesis 2**
-The predictive model built from Hypothesis 1 can be shown to have a positive impact on the overall profit earned by the case study business.
-The overall profit is represented by the following equation:
+<!-- **Hypothesis 2** -->
+<!-- The predictive model built from Hypothesis 1 can be shown to have a positive impact on the overall profit earned by the case study business.
+The overall profit is represented by the following equation: -->
 
@@
 
-Summary of Key Related Research 
-This section should include a brief summary of key related research. Emphasis should be on demonstrating the foundation for the current investigation. Specifically, the goal is to 
-clearly delineate a gap or missing link that the current 
-research fills. Authors should avoid presenting a litany of past research and should focus
-on prior work necessary to demonstrate the
-existence of the research gap addressed in the manuscript.
+<!-- Summary of Key Related Research  -->
+<!-- This section should include a brief summary of key related research. Emphasis should be on demonstrating the foundation for the current investigation. Specifically, the goal is to  -->
+<!-- clearly delineate a gap or missing link that the current  -->
+<!-- research fills. Authors should avoid presenting a litany of past research and should focus -->
+<!-- on prior work necessary to demonstrate the -->
+<!-- existence of the research gap addressed in the manuscript. -->
 
-1 page
+For the case study company, training the algorithm to predict actual fee was not possible because the projects were too varied, and there was inadequate information describing the size and scope of each project in the database.
+However, in the literature, similar studies typically did attempmt to predict project costs of IT consutling projects and final construction costs of buildings and infrastructure.
 
@@
 
-The bulk of research to date has been performed with project data in either the construction industry or software development.
-Research in the construction industry has primarily focused on predicting the final building cost of construction.
-Over the past few decades, significant research has been dedicated to creating predictive models that present the 'outside view' of a project.
-This was done by statistically comparing a new project to a collection of similar projects and their characteristics.
-It has been shown that these models improve cost estimation accuracy, however industry has not adopted them.
+Previous studies in the literature have focussed on reporting the predictive accuracy of their cost estimation models for IT or construction projects using data from multiple businesses or governments around the world.
+In contrast, this research is focussed on determining how or if models predicting project profitability can improve the bottom line for a case study business, given data the company has internally generated.
 
-##### 2.1.1 Construction Industry
-@Elfaki2014 completed a review of cost estimation research from 2004 to 2014. They found that artificial Neural Networks and SVM's were the most common machine learning techniques.
-Neural Networks and SVM's received significant attention in the 1990's for their ability to accurately predict construction costs with limited detailed information [@Shin2015; @Kim2004].
-@Shin2015 pioneered the application of Boosted Trees to cost estimation in construction projects.
-Often, Linear Regression was the only model assessed, without comparison to other methods such as Neural Networks, which first started appearing in literature in the 1990's [@Kim2004]
-Some studies showed that Neural Networks outperform Regression, however other studies established they are approximately equal [@Kim2004; @Attalla2003].
+Reviewing a range of individual studies show that cost estimation models have generally been built from at most 530 projects [@Kim2004] [@Finnie1997] - 299 software development[@Pai2013] - 163 software [@Shin2015] - 234 building projects, and often under 100 projects [@Finnie1997] [@Chan2005] 87 buildings.
+Sometimes it is not stated where the project data came from, and some studies specified the information came from surveys, with responses from companies and governments around the world.
+In this study, the dataset originates from a single company's database, which means the data is easily accessible to the company and there are 2364 projects, which is much more than previous studies.
+A study by @Mendes2004 demonstrated that using within company data was also more accurate than across company data.
+Their dataset consisted of 67 across company web projects and 14 within company projects.
 
-Variety of variables used: [@Chan2005][@Akintoye2000; @Trost2003; @Pinto1988]
-the studies all generally reported their mean absolute error or a similar metric from the tested models, and results were generally positive. 
-Furthermore, the existing studies collected data from many businesses for a single study
+Almost all past studies analysed their model performance by predictive accuracy (reported as R^2 or RMSE) of cost predictions [@llist some stff] and compared these statistics bewteen methods.
+However, these results were not translated into a business case for the impact these models could have on an institution or how the model could be practically integrated into decision making.
+For example, if an institution used the model, how would it be applied to their decision making process and what magnitude of benefits could they expect?
+Saradhi & Palshikar's [-@Saradhi2011] work on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time, presented a good framework for implementing their model.
+Theirsupport vector machine (SVM) predictive model identified employees at high risk of churn.
+Then a method for determining the value of each employee in terms of their importance to projects and their monthly chargeability was described, which enabled high risk employees to be ranked.
+Managers then had a clear direction for who to target first to prevent churn.
+This extension provided a comprehensive framework for how business managers could adopt a machine learning algorithm to improve business operations.
+It was a valuable addition that is absent from most cost estimation research, but will be included in this work.
 
-##### 2.1.2 Software Industry
+#### 2.2 Methods used in Cost Estimation and Other Business Applications
 
-In the software industry, the main component of cost is *effort* as opposed to the cost of building materials in the construction industry.
+Cost estimation literature tended to apply linear regression, neural networks and sometimes SVM methods, while predictive methods applied to other business problems investigated a wider range.
+This study aims to compare a wider range of methods to past cost estimation studies.
 
-Similar to the construction industry, expert judgment or detailed analysis, is the most widely practiced method for effort estimation [@Shepperd1996; @Molokken2003]
+In the construction industry, often only one method was assessed, such as linear regression, without comparison to other methods [@Chan2005; @Elfaki2014].
+Neural Networks started appearing in literature in the 1990's and @Elfaki2014's review of cost estimation methods from 2004 to 2014 found that artificial Neural Networks and SVM's were the most popular machine learning techniques, but studies tended to only test one method.
+Some studies showed that Neural Networks outperformed Linear Regression, however other studies established they are approximately equal [@Kim2004; @Attalla2003].
+To the authors' knowledge @Shin2015 documented the only application of ensemble tree methods (Boosted Trees) to cost estimation in construction projects.
+They found Boosted Trees slightly outperformed Neural Networks, but not significanatly.
 
-Multiple studies have shown that Neural Networks definitively outperform regression models in effort estimation, although Regression is the most popular method [@Finnie1997; @Pai2013; @Matson1993].
-
- Interestingly, several studies have found that even if 15 or so variables are included, often only one variable contributes significantly to the model's accuracy: size (or funciton point) [@Shepperd1996; @Finnie1997; @Pai2013]
-
- A study by @Mendes2004 demonstrated using 67 web projects that cross-company models were significantly less accurate than a within company model [@Shepperd1996]
- Again, Neural Networks were criticised for their inability to explain their results [@Finnie1997].
- The software industry is unique in that agile methods of delivery are changing the contractual approach of consulting [@Badenfelt2011].
-Software development is a relatively new field in comparison to construction, where the tradition of fixed fee structures is difficult to reinvent.
-
-#### 2.2 Methods used in other business applications
+In the software cost estimation literature, Linear Regression was the most popular method, however multiple studies showed that Neural Networks definitively outperform regression models  [@Finnie1997; @Pai2013; @Matson1993].
+Despite the higher performance, Neural Networks were criticised for not providing reasoning or structure behind their predictions (black-box predictor) [@Finnie1997].
 
 The review of methods applied to the construction and IT industry highlighted the use of Linear Regression, Neural Networks, SVM's and in one case Boosted Trees, however research on other business problems utilised a wider range of methods.
-These included Naive Bayes, Random Forests, and machine learned Bayesian Networks.
+In financial credit scoring, a study by @Brown2012 found that Random Forests and Boosted Trees consistently outperformed Neural Networks in classification.
+In addition, ensemble trees such as Random Forests and Boosted Trees can output partial dependency plots and variable importance measures that provide insight into the model's calculations.
+This is a valuable asset oer black box predictors.
+@Kumar2007 performed a detailed review of statistical and machine learning techniques applied over 37 years in the context of bankruptcy prediction in banks and reported that although SVM's performed well, they are often complex and slow, requiring a great deal of memory.
+They also assessed method blending techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods such as Linear Regression or Boosted Trees alone.
 
-Although SVM's performed the best, as discussed, they are often complex and slow, requiring a great deal of memory [@Kumar2007].
-@Kumar2007 performed a detailed review of statistical and machine learning techniques that were applied over 37 years in the context of bankruptcy prediction in banks.
-
-For example in financial credit scoring, a study by @Brown2012 found that Random Forests and Boosted Trees consistently outperformed Neural Networks in classification.
-Kumar and Ravi's [-@Kumar2007] review also assessed ensemble techniques, which refers to combinations of two completely different algorithms, and found they can often outperform individual methods.
-
-An exception is Saradhi & Palshikar's [-@Saradhi2011] study on employee churn, where 'churn' refers to the number of individuals moving out of a group within a certain time.
-Naive Bayes, Random forests (an ensemble decision tree method) and SVM's were built.
-determining the value of each employee in terms of the importance of the projects they were on and their monthly chargeability.
-This allowed them to rank employees identified as 'high risk of churn' by value and provided a clear ranking for manager's to act upon.
-This extension of the study provided a comprehensive framework for how business managers could adopt their findings to improve business operations.
-It was a valuable addition that is absent from most cost and effort estimation research.
+To date, Linear Regression, Neural Networks, and Support Vector Machines (SVM's) have been comprehensively applied to cost estimation.
+This research does not test black-box predictors (Neural Networks and SVM's) because insight into the model's predictions is crucial for facilitating user uptake in the case study company.
+Ensemble tree methods such as Random Forests and Boosted Trees will be used instead, as they can perform as well as Neural Networks while providing insight to their output [@Caruana2006].
+Blending of multiple predictive models is also applied in this work and has not yet been done in cost estimation.
 
@@
 
-In comparison to previous studies, this project advances the body of work on cost estimation in three ways: the data set is larger than other studies and sourced from a single company's CRM database, ensemble tree methods and model blending was tested, and the prediction results were integrated into a decision framework for the business.
-In past research, cost estimation data sets generally consisted of <300 of projects from a range of a companies and even countries.
-This thesis case study only had access to the company's internal data but this contained over 2,000 past jobs
-Furthermore, the problem of cost estimation is applied to a consulting company in the construction industry, which differs from other studies in the construction industry that estimate actual construction costs.
-Consulting in the construction industry can be likened more to software development effort estimation although the nature of the work differs considerably.
-
-ensemble tree methods and blended methods are applied which have been minimally tested on this problem.
-To date, Linear Regression, Neural Networks, Case Based Reasoning (CBR), and Support Vector Machines (SVM's) have been used to predict cost/effort even though ensemble tree methods can perform as well as Neural Networks and generally outperform Linear Regression [@Caruana2006].
-Another benefit of ensemble trees is the output types available, such as partial dependency plots and variable importance plots that provide insight into the model's calculations.
-This contrasts Neural Networks and SVM's, which are 'black-box' predictors.
-
-Finally, this study propels the predictive model one step further than other studies by analysing bottom-line profit improvements that could result from the research.
-final profit increases are evaluated. This kind of thought experiment is designed to present a clear case for industry when evaluating if and how to adopt cost estimation models.
+In comparison to previous studies, this project advances the body of work on cost estimation in two ways.
+First, it focusses on building a model that is practical and profitable for a business to implement by using the case study's interenal database as well as testing a managerial decision framework based off the model output.
+This framework presents a clear measure of expected benefits provided by adopting the profitability model into a business' decision process.
+Secondly, advanced predictive methods are trialled which have been minimally tested on this problem.
+These include ensemble decision tree methods, such as Random Forests and Boosted Trees, and blends of multiple models built from different methods.
 

### 2016-12-14 finalwriteup/Consulting Profitability Paper/Consulting_Profitability_Paper.tex
finished lit review paper

@@
 
-Depending on the industry, some consulting projects can be financialy
-risky endeavors for the consultant. Engineering consultanting companies
-in the construction industry complete a wide range of complex projects
-in a competitive enviroment. Because of the way contracts are
+Many consulting projects can be financialy risky endeavors for the
+consultant, often depending on the industry. Engineering consultanting
+companies in the construction industry complete a wide range of complex
+projects in a competitive enviroment. Because of the way contracts are
 structured, many of these projects result in losses. The purpose of this
@@
 by statistical and machine learning algorithms using the information
-available at project conception. The second question is whether the
-predictive algorithm could improve a company's profitability. Project
-data from a case study engineering company was used to answer these
-questions.
+available at project conception. Secondly, could the predictive
+algorithm improve a company's profitability? Predictive models and value
+frameworks were built using project data from a case study engineering
+company to answer these questions.
 
@@
 
-Financial risk most commonly falls upon a consulting company when they
-must offer a fixed fee for a project. In competitive climates, clients
-request fixed price quotes for project completion from several
-consultants before a project is awarded. Hence, initial quoted fees must
-be competitive but also accurately cover consultant salaries and
-business costs.
+Financial risk is taken on by a consulting company when they offer a
+fixed fee to complete a project. In competitive climates, clients are in
+the position to request fixed price quotes from several consultants
+before a project is awarded. Hence, initial quoted fees must be
+competitive but also accurately cover consultant salaries and business
+costs.
 
@@
 experience in the type of project they are quoting and, after reviewing
-the project details, can use a combination of intuition, comparison to
-past projects, and rules of thumb. Unfortunately, some projects priced
-using this method can blow out of budget, causing significant losses to
-the consultant.
-
-A review of 6 IT project surveys between 1984 and 1994 by Moløkken \&
-Jørgensen (2003) revealed that 60-80\% of IT projects encountered effort
-and/or schedule overruns, where the average overrun was by 30 to 40\%. A
-study on large-scale infrastructure projects over the past seventy years
-revealed that cost forecasts consistently underestimated the cost of
-rail projects by an average of 44.7\%, the cost of bridge and tunnel
-projects by 33.8\% and the cost of road projects by 20.4\% (Flyvbjerg
-2007).
+the project details, can use a combination of intuition, analysis of
+project details, and rules of thumb. Unfortunately, some projects priced
+using this method can blow out of budget, causing significant financial
+losses to the consultant.
+
+Cost blowouts of complex projects has a history, particularly documented
+in the It and construction industry. A review of 6 IT project surveys
+between 1984 and 1994 by Moløkken \& Jørgensen (2003) revealed that
+60-80\% of IT projects encountered effort and/or schedule overruns,
+where the average overrun was by 30 to 40\%. A study on large-scale
+infrastructure projects over the past seventy years revealed that cost
+forecasts consistently underestimated the cost of rail projects by an
+average of 44.7\%, the cost of bridge and tunnel projects by 33.8\% and
+the cost of road projects by 20.4\% (Flyvbjerg 2007).
 
@@
 
-This work aims to provide a manager with this more realistic view by
-predicting whether the new project will be profitable or not, based on a
-suite of previous project data from the same company. The manager could
-then take multiple actions including not tendering for a project
-predicted to be loss-making, providing a high fixed price, or
-restructuring the contract so that it is not 100\% fixed price.
+This work aims to help managers provide better fixed price quotes by
+predicting whether a new project will be profitable or not, based on an
+algorithm trained on all previous project data from the same company.
+The manager could take multiple actions based on the algorithms output
+including not tendering for a project predicted to be loss-making,
+providing a high fixed price, or restructuring the contract so that it
+is not 100\% fixed price. This research tests the simple case of
+rejecting all projects predicted as loss-making.
 
@@
 
-This thesis focuses on a single case study consulting company in the
-construction industry. twelve years of passively collected data that
-described each project in terms of their clients, invoice history,
-employee hours and technical details. labour is the chief cost and the
-long traditional history of this industry mandates fixed price projects
-as the norm. 20\% of projects are loss making. Fees calculated via
-detailed analysis A Customer Relationship Management software package
-(CRM) is currently employed to collect and store project data. A CRM is
-a popular type of software used by businesses to record client and
-project details as well as employee time sheet records. In the case
-study business, the CRM is available to all employees over the company
-intranet, and each employee completes daily time sheets allotting their
-hours to certain projects. Additionally, technical information is
-recorded against each project
-
-\begin{itemize}
-\tightlist
-\item
-  Employee time sheet hours with dates
-\item
-  Other project costs (taxis, printing)
-\item
-  Client information/characteristics
-\item
-  Client identification code
-\item
-  Invoiced amounts for each project and dates
-\item
-  Employee costs
-\item
-  Employee charge out rates
-\item
-  Project description
-\end{itemize}
-
-2364 past projects are available. missing data due to how the data
-collection changed over time and some fields were not mandatory.
-
-To assess whether the model would improve the case study company's
-profitability, a simple managerial decision framework was applied: if
-the algorithm predicted the project to be loss making, the project was
-considered `rejected' and all profits or \emph{losses} of rejected
-projects were discarded. The remaining projects' profits and losses were
-summed to give a revised overall profit over the last 12 years of
-projects. This was compared to the summed profits and losses from all
-projects completed over the last 12 years. The best algorithm rejected
-enough loss-making projects to improve overall profits by 9\%.
-
-\begin{itemize}
-\tightlist
-\item
-  aim \textbf{General Aim} Use statistical techniques to model the
-  profitability of projects for consulting businesses using their
-  internal CRM data. Research will focus on a case study Engineering
-  consulting company that offers their expert advice (in the currency of
-  time) to business clients. The project outcomes are intended to assist
-  the business in predicting project profitability before engagement.
-  Several statistical and machine learning techniques were tested,
-  compared and refined.
-\end{itemize}
-
-\textbf{Hypothesis 1} A statistical or machine learning model based on
-historical project data can predict the profitability of a new project
-with greater accuracy than a baseline predictor. A baseline predictor
-for this project is one that predicts the average of a numeric response
-variable for all cases or, if the response variable is categorical,
-randomly assigned categories for each case, where the proportion of
-assigned categories matches the true categorical proportions.
-
-\textbf{Hypothesis 2} The predictive model built from Hypothesis 1 can
-be shown to have a positive impact on the overall profit earned by the
-case study business. The overall profit is represented by the following
-equation:
+Project data from a single case study consulting company in the
+construction industry was used to build the predictive algorithm, trial
+several methods, and test their impact on the company's bottom line.
+Twelve years of passively collected data that described each project in
+terms of their clients, invoice history, employee hours and technical
+details was scraped from the company's internal database. Customer
+relationship management (CRM) software provided accesses to the database
+to all employees. Employees completed daily timesheets, alotting their
+hours to certian projects. Project managers input technical information
+and client details for each project, while the finance team tracked
+invoicing and additional project costs such as printing and transport in
+the same system. On top of this information, the CRM had the hourly cost
+of each level of employee. The cleaned dataset consisted of 2364
+projects between 2004 and 2012. 70\% of projects were fixed price and
+20\% of all projects made a loss after taking into account business
+costs.
+
+Once several methods were built, to assess whether the model would
+improve the case study company's profitability, a simple managerial
+decision framework was applied: if the algorithm predicted the project
+to be loss making, the project was considered `rejected' and all profits
+or \emph{losses} of rejected projects were discarded. The remaining
+projects' profits and losses were summed to give a revised overall
+profit over the last 12 years of projects. This was compared to the
+summed profits and losses from all projects completed over the last 12
+years. The best algorithm rejected enough loss-making projects to
+improve overall profits by 9\%.
 
@@
 
-Summary of Key Related Research This section should include a brief
-summary of key related research. Emphasis should be on demonstrating the
-foundation for the current investigation. Specifically, the goal is to
-clearly delineate a gap or missing link that the current research fills.
-Authors should avoid presenting a litany of past research and should
-focus on prior work necessary to demonstrate the existence of the
-research gap addressed in the manuscript.
-
-1 page
+For the case study company, training the algorithm to predict actual fee
+was not possible because the projects were too varied, and there was
+inadequate information describing the size and scope of each project in
+the database. However, in the literature, similar studies typically did
+attempmt to predict project costs of IT consutling projects and final
+construction costs of buildings and infrastructure.
 
@@
 
-The bulk of research to date has been performed with project data in
-either the construction industry or software development. Research in
-the construction industry has primarily focused on predicting the final
-building cost of construction. Over the past few decades, significant
-research has been dedicated to creating predictive models that present
-the `outside view' of a project. This was done by statistically
-comparing a new project to a collection of similar projects and their
-characteristics. It has been shown that these models improve cost
-estimation accuracy, however industry has not adopted them.
-
-\subparagraph{2.1.1 Construction Industry}\label{construction-industry}
-
-Elfaki, Alatawi, and Abushandi (2014) completed a review of cost
-estimation research from 2004 to 2014. They found that artificial Neural
-Networks and SVM's were the most common machine learning techniques.
-Neural Networks and SVM's received significant attention in the 1990's
-for their ability to accurately predict construction costs with limited
-detailed information (Shin 2015; Kim, An, and Kang 2004). Shin (2015)
-pioneered the application of Boosted Trees to cost estimation in
-construction projects. Often, Linear Regression was the only model
-assessed, without comparison to other methods such as Neural Networks,
-which first started appearing in literature in the 1990's (Kim, An, and
-Kang 2004) Some studies showed that Neural Networks outperform
-Regression, however other studies established they are approximately
-equal (Kim, An, and Kang 2004; Attalla and Hegazy 2003).
-
-Variety of variables used: (Chan and Park 2005)(Akintoye and Fitzgerald
-2000; Trost and Oberlender 2003; Pinto and Slevin 1988) the studies all
-generally reported their mean absolute error or a similar metric from
-the tested models, and results were generally positive. Furthermore, the
-existing studies collected data from many businesses for a single study
-
-\subparagraph{2.1.2 Software Industry}\label{software-industry}
-
-In the software industry, the main component of cost is \emph{effort} as
-opposed to the cost of building materials in the construction industry.
-
-Similar to the construction industry, expert judgment or detailed
-analysis, is the most widely practiced method for effort estimation
-(Shepperd, Schofield, and Kitchenham 1996; Moløkken and Jørgensen 2003)
-
-Multiple studies have shown that Neural Networks definitively outperform
-regression models in effort estimation, although Regression is the most
-popular method (Finnie, Wittig, and Desharnais 1997; Pai, McFall, and
-Subramanian 2013; Matson and Mellichamp 1993).
-
-Interestingly, several studies have found that even if 15 or so
-variables are included, often only one variable contributes
-significantly to the model's accuracy: size (or funciton point)
-(Shepperd, Schofield, and Kitchenham 1996; Finnie, Wittig, and
-Desharnais 1997; Pai, McFall, and Subramanian 2013)
-
-A study by Mendes and Kitchenham (2004) demonstrated using 67 web
-projects that cross-company models were significantly less accurate than
-a within company model (Shepperd, Schofield, and Kitchenham 1996) Again,
-Neural Networks were criticised for their inability to explain their
-results (Finnie, Wittig, and Desharnais 1997). The software industry is
-unique in that agile methods of delivery are changing the contractual
-approach of consulting (Badenfelt 2011). Software development is a
-relatively new field in comparison to construction, where the tradition
-of fixed fee structures is difficult to reinvent.
-
-\paragraph{2.2 Methods used in other business
-applications}\label{methods-used-in-other-business-applications}
+Previous studies in the literature have focussed on reporting the
+predictive accuracy of their cost estimation models for IT or
+construction projects using data from multiple businesses or governments
+around the world. In contrast, this research is focussed on determining
+how or if models predicting project profitability can improve the bottom
+line for a case study business, given data the company has internally
+generated.
+
+Reviewing a range of individual studies show that cost estimation models
+have generally been built from at most 530 projects (Kim, An, and Kang
+2004) (Finnie, Wittig, and Desharnais 1997) - 299 software
+development(Pai, McFall, and Subramanian 2013) - 163 software (Shin
+2015) - 234 building projects, and often under 100 projects (Finnie,
+Wittig, and Desharnais 1997) (Chan and Park 2005) 87 buildings.
+Sometimes it is not stated where the project data came from, and some
+studies specified the information came from surveys, with responses from
+companies and governments around the world. In this study, the dataset
+originates from a single company's database, which means the data is
+easily accessible to the company and there are 2364 projects, which is
+much more than previous studies. A study by Mendes and Kitchenham (2004)
+demonstrated that using within company data was also more accurate than
+across company data. Their dataset consisted of 67 across company web
+projects and 14 within company projects.
+
+Almost all past studies analysed their model performance by predictive
+accuracy (reported as R\^{}2 or RMSE) of cost predictions
+({\textbf{???}} some stff) and compared these statistics bewteen
+methods. However, these results were not translated into a business case
+for the impact these models could have on an institution or how the
+model could be practically integrated into decision making. For example,
+if an institution used the model, how would it be applied to their
+decision making process and what magnitude of benefits could they
+expect? Saradhi \& Palshikar's (2011) work on employee churn, where
+`churn' refers to the number of individuals moving out of a group within
+a certain time, presented a good framework for implementing their model.
+Theirsupport vector machine (SVM) predictive model identified employees
+at high risk of churn. Then a method for determining the value of each
+employee in terms of their importance to projects and their monthly
+chargeability was described, which enabled high risk employees to be
+ranked. Managers then had a clear direction for who to target first to
+prevent churn. This extension provided a comprehensive framework for how
+business managers could adopt a machine learning algorithm to improve
+business operations. It was a valuable addition that is absent from most
+cost estimation research, but will be included in this work.
+
+\paragraph{2.2 Methods used in Cost Estimation and Other Business
+Applications}\label{methods-used-in-cost-estimation-and-other-business-applications}
+
+Cost estimation literature tended to apply linear regression, neural
+networks and sometimes SVM methods, while predictive methods applied to
+other business problems investigated a wider range. This study aims to
+compare a wider range of methods to past cost estimation studies.
+
+In the construction industry, often only one method was assessed, such
+as linear regression, without comparison to other methods (Chan and Park
+2005; Elfaki, Alatawi, and Abushandi 2014). Neural Networks started
+appearing in literature in the 1990's and Elfaki, Alatawi, and Abushandi
+(2014)`s review of cost estimation methods from 2004 to 2014 found that
+artificial Neural Networks and SVM's were the most popular machine
+learning techniques, but studies tended to only test one method. Some
+studies showed that Neural Networks outperformed Linear Regression,
+however other studies established they are approximately equal (Kim, An,
+and Kang 2004; Attalla and Hegazy 2003). To the authors' knowledge Shin
+(2015) documented the only application of ensemble tree methods (Boosted
+Trees) to cost estimation in construction projects. They found Boosted
+Trees slightly outperformed Neural Networks, but not significanatly.
+
+In the software cost estimation literature, Linear Regression was the
+most popular method, however multiple studies showed that Neural
+Networks definitively outperform regression models (Finnie, Wittig, and
+Desharnais 1997; Pai, McFall, and Subramanian 2013; Matson and
+Mellichamp 1993). Despite the higher performance, Neural Networks were
+criticised for not providing reasoning or structure behind their
+predictions (black-box predictor) (Finnie, Wittig, and Desharnais 1997).
 
@@
 one case Boosted Trees, however research on other business problems
-utilised a wider range of methods. These included Naive Bayes, Random
-Forests, and machine learned Bayesian Networks.
-
-Although SVM's performed the best, as discussed, they are often complex
-and slow, requiring a great deal of memory (Kumar and Ravi 2007). Kumar
-and Ravi (2007) performed a detailed review of statistical and machine
-learning techniques that were applied over 37 years in the context of
-bankruptcy prediction in banks.
-
-For example in financial credit scoring, a study by Brown and Mues
-(2012) found that Random Forests and Boosted Trees consistently
-outperformed Neural Networks in classification. Kumar and Ravi's (2007)
-review also assessed ensemble techniques, which refers to combinations
-of two completely different algorithms, and found they can often
-outperform individual methods.
-
-An exception is Saradhi \& Palshikar's (2011) study on employee churn,
-where `churn' refers to the number of individuals moving out of a group
-within a certain time. Naive Bayes, Random forests (an ensemble decision
-tree method) and SVM's were built. determining the value of each
-employee in terms of the importance of the projects they were on and
-their monthly chargeability. This allowed them to rank employees
-identified as `high risk of churn' by value and provided a clear ranking
-for manager's to act upon. This extension of the study provided a
-comprehensive framework for how business managers could adopt their
-findings to improve business operations. It was a valuable addition that
-is absent from most cost and effort estimation research.
+utilised a wider range of methods. In financial credit scoring, a study
+by Brown and Mues (2012) found that Random Forests and Boosted Trees
+consistently outperformed Neural Networks in classification. In
+addition, ensemble trees such as Random Forests and Boosted Trees can
+output partial dependency plots and variable importance measures that
+provide insight into the model's calculations. This is a valuable asset
+oer black box predictors. Kumar and Ravi (2007) performed a detailed
+review of statistical and machine learning techniques applied over 37
+years in the context of bankruptcy prediction in banks and reported that
+although SVM's performed well, they are often complex and slow,
+requiring a great deal of memory. They also assessed method blending
+techniques, which refers to combinations of two completely different
+algorithms, and found they can often outperform individual methods such
+as Linear Regression or Boosted Trees alone.
+
+To date, Linear Regression, Neural Networks, and Support Vector Machines
+(SVM's) have been comprehensively applied to cost estimation. This
+research does not test black-box predictors (Neural Networks and SVM's)
+because insight into the model's predictions is crucial for facilitating
+user uptake in the case study company. Ensemble tree methods such as
+Random Forests and Boosted Trees will be used instead, as they can
+perform as well as Neural Networks while providing insight to their
+output (Caruana and Niculescu-Mizil 2006). Blending of multiple
+predictive models is also applied in this work and has not yet been done
+in cost estimation.
 
@@
 In comparison to previous studies, this project advances the body of
-work on cost estimation in three ways: the data set is larger than other
-studies and sourced from a single company's CRM database, ensemble tree
-methods and model blending was tested, and the prediction results were
-integrated into a decision framework for the business. In past research,
-cost estimation data sets generally consisted of \textless{}300 of
-projects from a range of a companies and even countries. This thesis
-case study only had access to the company's internal data but this
-contained over 2,000 past jobs Furthermore, the problem of cost
-estimation is applied to a consulting company in the construction
-industry, which differs from other studies in the construction industry
-that estimate actual construction costs. Consulting in the construction
-industry can be likened more to software development effort estimation
-although the nature of the work differs considerably.
-
-ensemble tree methods and blended methods are applied which have been
-minimally tested on this problem. To date, Linear Regression, Neural
-Networks, Case Based Reasoning (CBR), and Support Vector Machines
-(SVM's) have been used to predict cost/effort even though ensemble tree
-methods can perform as well as Neural Networks and generally outperform
-Linear Regression (Caruana and Niculescu-Mizil 2006). Another benefit of
-ensemble trees is the output types available, such as partial dependency
-plots and variable importance plots that provide insight into the
-model's calculations. This contrasts Neural Networks and SVM's, which
-are `black-box' predictors.
-
-Finally, this study propels the predictive model one step further than
-other studies by analysing bottom-line profit improvements that could
-result from the research. final profit increases are evaluated. This
-kind of thought experiment is designed to present a clear case for
-industry when evaluating if and how to adopt cost estimation models.
+work on cost estimation in two ways. First, it focusses on building a
+model that is practical and profitable for a business to implement by
+using the case study's interenal database as well as testing a
+managerial decision framework based off the model output. This framework
+presents a clear measure of expected benefits provided by adopting the
+profitability model into a business' decision process. Secondly,
+advanced predictive methods are trialled which have been minimally
+tested on this problem. These include ensemble decision tree methods,
+such as Random Forests and Boosted Trees, and blends of multiple models
+built from different methods.
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Akintoye2000}{}
-Akintoye, Akintola, and Eamon Fitzgerald. 2000. ``A Survey of Current
-Cost Estimating Practices in the Uk.'' Journal Article.
-\emph{Construction Management and Economics} 18 (2): 161--72.
-doi:\href{https://doi.org/10.1080/014461900370799}{10.1080/014461900370799}.
-
 \hypertarget{ref-Attalla2003}{}
@@
 as a Technique for Solvency Analysis.'' Journal Article.
-
-\hypertarget{ref-Badenfelt2011}{}
-Badenfelt, Ulrika. 2011. ``Fixing the Contract After the Contract Is
-Fixed: A Study of Incomplete Contracts in It and Construction
-Projects.'' Journal Article. \emph{International Journal of Project
-Management} 29 (5): 568--76.
-doi:\href{https://doi.org/http://dx.doi.org/10.1016/j.ijproman.2010.04.003}{http://dx.doi.org/10.1016/j.ijproman.2010.04.003}.
 
@@
 
-\hypertarget{ref-Molokken2003}{}
-Moløkken, Kjetil, and Magne Jørgensen. 2003. ``A Review of Software
-Surveys on Software Effort Estimation.'' Conference Proceedings. In
-\emph{Empirical Software Engineering, 2003. Isese 2003. Proceedings.
-2003 International Symposium on}, 223--30. IEEE.
-
 \hypertarget{ref-Moore1989}{}
@@
 Article. \emph{The Journal of Machine Learning Research} 4: 211--55.
-
-\hypertarget{ref-Pinto1988}{}
-Pinto, Jeffrey K, and Dennis P Slevin. 1988. ``Critical Success Factors
-Across the Project Life Cycle.'' Conference Proceedings. In. Project
-Management Institute.
 
@@
 
-\hypertarget{ref-Shepperd1996}{}
-Shepperd, Martin, Chris Schofield, and Barbara Kitchenham. 1996.
-``Effort Estimation Using Analogy.'' Conference Proceedings. In
-\emph{Proceedings of the 18th International Conference on Software
-Engineering}, 170--78. IEEE Computer Society.
-
 \hypertarget{ref-Shin2015}{}
@@
 
-\hypertarget{ref-Trost2003}{}
-Trost, Steven M, and Garold D Oberlender. 2003. ``Predicting Accuracy of
-Early Cost Estimates Using Factor Analysis and Multivariate
-Regression.'' Journal Article. \emph{Journal of Construction Engineering
-and Management} 129 (2): 198--204.
-
 \hypertarget{ref-Weisberg2005}{}

### 2016-12-16 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd
finished method in paper

@@
 
-Neural Netowrks and SVM were not purusued because black box
-A previously mentioned disadvantage is that the NN algorithm is a black box because the internal structure is too complex for interpretation.
+A range of simple and complex, linear and non-linear methods were applied to the problem including Logistic Regression, Random Forests, Gradient Boosted Trees, and Naiive Bayes.
+Furthermore, logistic regression, random forests, boosted trees and simple averaging were used to blend predictions from individual models.
+A brief description of each method is provided below.
+
+<!-- Neural Netowrks and SVM were not purusued because black box -->
+<!-- A previously mentioned disadvantage is that the NN algorithm is a black box because the internal structure is too complex for interpretation.
 They also require a lot of training data relative to other methods.
+-->
 
-Also, SVM's can be very slow to train and therefore not suitable for industry purposes [@Auria2008].
+<!-- Also, SVM's can be very slow to train and therefore not suitable for industry purposes [@Auria2008].
 Because it is so important to engage decision makers with a model that can explain its results, these two methods are not appropriate for the effort estimation problem.
+-->
 
@@
 
-binary classification problems, where the objective is to predict the probability of an event occurring.
-Probability is a continuous response variable, however predictions from a linear regression would not be bound by 0 and 1.
+The aim of this research is to build a model that predicts whether a consulting project will be profitable or not.
+Therefore each method must be capable of binary classifcation.
+Binary classification output from a model is the probability, between 0 and 1, of an event occuring.
+In this case, the event is a project being loss-making.
 
@@
 
-Because of the linear relationship, an equation can be fit to the log odds of the binary response variable against each of the explanatory variables using linear regression.
-Then, the linear equation describing the log odds can be transformed back to probability by taking the inverse log i.e.
-the exponential and rearranging [@Macdonald1975]:
+Logisitic regression models a linear relationship bewteen the log odds of a binary response variable and each of the explanatory variables.
+This is performed similarly to linear regression, but is transformed to probability of an event occuring by taking the inverse log of the log odds and rearranging [@Macdonald1975]:
+
+$$
+\begin{aligned}
+\log(odds) &= \beta_0 + X_1\beta_1 \\
+\log\left(\frac{p}{1-p}\right) &= \beta_0 + X_1\beta_1 \\
+-\log\left(\frac{1-p}{p}\right) &= \beta_0 + X_1\beta_1 \\
+-\log\left(\frac{1}{p} - 1\right) &= \beta_0 + X_1\beta_1 \\
+\log\left(\frac{1}{p} - 1\right) &= -\beta_0 - X_1\beta_1 \\
+\exp\left(\log\left(\frac{1}{p} - 1\right)\right) &= \exp(-\beta_0 - X_1\beta_1) \\
+\frac{1}{p} - 1 &= \exp(-\beta_0 - X_1\beta_1) \\
+\frac{1}{p} &= 1 + \exp(-\beta_0 - X_1\beta_1) \\
+p &= \frac{1}{1 + \exp(-\beta_0 - X_1\beta_1)}
+\end{aligned}
+$$
 
 The result in a sigmoidal function bound by 0 and 1.
-The sigmoidal function originates from a linear fit of the log odds in the data.
-Assume linear relationship between covariates and response variable
-
 A coefficient represents the change in the log odds of the response variable for each unit increase in an explanatory variable.
-Therefore, taking the exponential of the coefficient is the change in odds of the response variable.
-If the change in odds was 2, then if the explanatory variable increased by 1, the response variable event would be twice as likely to occur.
-
- Logistic Regression is a good  benchmark to compare other binary predictive models due to its simplicity and speed [@Moore1989]. single linear relationship means Not prone to overfitting
+Logistic Regression is a good  benchmark to compare other binary predictive models due to its simplicity and speed [@Moore1989].
+Addittionally, because of the simple, single linear relationship, the method is not prone to overfitting [@Perlich2003].
 
@@
 
-single decision trees also suffer from low predictive accuracy and instability, so to combat these problems, ensemble tree methods were pioneered in the 1990's [@Breiman1996].
+Ensemble tree methods were pioneered in the 1990's to combat the low predictive accuracy and instability of individual decision trees.
+Combining hundreds or thousands of trees transform decision trees into high performing, stable predictors [@Breiman1996].
 
-The three methods of combining hundreds or thousands of trees turn decision trees into high performing, stable predictors.
-
-training multiple trees from the same data by sampling bootstrapped training sets with replacement.
-However, when creating each tree, a random subset of attributes (variables) is considered at each split.
-The reduced subset of attributes is resampled for each split in the tree.
-This allows dominant variables to be suppressed for a fraction of the splits, allowing the algorithm to explore signals in weaker variables [@Breiman2001a].
-
-Additionally, ensemble tree methods are faster than SVM's and Neural Networks but can perform just as well and even provide insights such as variable importance and variable relationships [@Sealfon2012].
+The Random Forests algorithm works by training multiple trees from the bootstrapped samples of a dataset.
+Additionally, when creating each tree, a random subset of attributes (covariates) is considered at each split.
+The reduced subset of covariates is resampled for each split in the tree.
+Dominant covariates are then suppressed for a fraction of the splits, allowing the algorithm to explore signals in weaker covariates [@Breiman2001a].
+Performance statistics and tests run within the algorithm provide insights to the modeller, such as variable importance and variable relationships [@Sealfon2012].
 
@@
 
-Lastly, the boosted decision tree approach applies gradient descent theory to a series of decision trees.
-The trees are limited to a certain depth to maintain simplicity, and each tree models the residuals (or errors) of the preceding tree.
-By modeling the errors, misclassified cases are weighted higher than correctly classified cases and influence the structure of the latest tree more [@Hastie2009].
-This increased weighting on errors is why the algorithm is called boosting.
-The limited depth of each tree prevents overfitting at each stage and the combined result of up to thousands of trees is very powerful [@Elith2008].
+The boosted decision tree approach applies gradient descent theory to a series of decision trees.
+Performance tends to be higher if individual trees are limited to a certain depth to maintain simplicity and prevent overfitting [@Elith2008].
+Unlike Random Forests, the trees are sequential, where each tree models the residuals (or errors) of the preceding tree.
 
+By modeling the errors, misclassified cases are weighted higher than correctly classified cases and influence the structure of the next tree more [@Hastie2009].
+This increased weighting of error cases is called boosting.
+The successive results of up to thousands of trees is very powerful [@Elith2008].
 @Caruana2006 tested Boosted Trees, Random Forests, Neural Networks, SVM's, Logistic Regression and Naive Bayes on 11 binary classification problems and found that Boosted Trees performed best, followed by Random Forests.
-This demonstrates how ensemble tree methods are capable of competing with high-level machine learning algorithms.
-
-Trees are not built on a probabilistic framework, and therefore their results cannot be provided in this framework [@Louppe2014].
+This demonstrates how capable ensemble tree methods are of competing with high-level machine learning algorithms.
 
@@
 
-The Naive Bayes method works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for the response variable (the response variable must be categorical).
+Naive Bayes is a simple, fast method included in this work as it provides a good baseline method against which to compare complex methods [@Caruana2006].
+It works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for a categorical response variable.
+For example, if predicting a binary variable, $R$ (with classes 0 or 1) given two covariates $X_1$ and $X_2$, the equations for the probability of both outcomes is as follows:
+
+$p(R = 1|X_1 X_2) = p(X_1|R = 1) \cdot p(X_2|R = 1) \cdot p(R = 1)$
+
+$p(R = 0|X_1 X_2) = p(X_1|R = 0) \cdot p(X_2|R = 0) \cdot p(R = 0)$
 
 These equations are simple to compute given the data, and the class with the highest probability can then be chosen [@Provost2013].
+The method can perform well in real world tasks because the assumption of independence does not significantly damage predictions, but amplifies the magnitude of the probability [@Provost2013].
+This is fine for ranking class 0 against 1 but the output probabilities are not accurate statistical probabilities [@Caruana2006].
 
-The advantages of this method are that the conditional independence assumption enables very fast calculations and predictions.
-The method can perform well in real world tasks because the assumption of independence does not significantly damage predictions [@Provost2013].
-Assumptions of independence genearlly not true (correlated variables).
-This is fine for ranking but the output probabilities are not accurate statistical probabilities [@Caruana2006].
+##### 3.1.5 Model Blending
 
-Naive Bayes method generally provides a good benchmark to compare against more complex models that should outperform it [@Caruana2006].
+Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009].
+In this paper, the term 'blending' will be used to avoid confusion with ensemble tree methods.
+Model blending combines predictions from several models that used methods with different theoretical foundations.
+In this way, the strengths of each method can be combined.
+Since the 1990's research in statistics has shown that averaging results from different methods can provide better predictive accuracy than any single model [@Madigan1994].
 
-##### 3.1.5 Bayesian Networks
-
-A Bayesian Network is a graphical probabilistic model that illustrates the conditional dependencies between variables in a data set.
-The model is visually represented by a directed acyclic graph (DAG) and is capable of linking the conditional dependency between any variable to another variable [@Heckerman1998].
-The conditional relationships are Bayesian, where the probabilities in one node are conditional upon values in nodes directed towards it as well as preceding nodes
-
-Bayesian Networks reduce computations required to find the probability of a unique combination given other explanatory variable values while still taking into account many conditional dependencies that Naive Bayes [@Barber2012]
-
-Naive Bayes is the simplest form of a Bayesian Network [@Zhang2004]
-
-Network relationships can be learned from the data, however this is not widely included as part of the suite of machine learning methods.
-The conditional variable dependencies are calculated from the data, which in turn can define the graph structure.
-Some conditional dependencies can be set before the structure is learned [@Barber2012].
-
-Drawbacks of Bayesian Networks include the difficulty for them to process continuous variables.
-
-Bayesian networks have found success in combining deterministic models with observed data as well as expert knowledge.[@Kragt2009]
+Six blending methods, ranging from simple to complex, were chosen.
+These included simple averaging of the individual model results, building a Logistic Regression model and Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
+The first two models are simply the average, and weighted average of predictions from separate individual models.
+For example, if the Logistic Regression model output a probability of 0.6 for one case, while a Boosted Tree output 0.7, a simple blended average for that case would be a probability of 0.65.
+The last three methods are more complex, and take advantage of potentially meaningful interactions between the individual model predictions and the original covariates describing each case (called meta-features in the model blending literature) [@Sill2009].
+For example, if the Logistic Regression model predicted profitability better than the Boosted Tree model for projects with only one team member, the complex blending methods would take advantage of this interaction.
+In summary, in the complex methods, the probability outputs from each model are interacted with each of the original covariates.
+Model blending has the potential to elevate output from the original methods by combining each of their strengths.
 
@@
 
- It includes first how the data was obtained, the lengthy cleaning process, 
- Projects varied from total invoiced amounts of \$500 to over \$1,000,000
- a project could have thousands of rows of relevant data that needed to be converted into a single row per project.
- 10 engineered variables describing timesheets: Percent of hours performed by 'professional' employees as opposed to 'technical' employees, Position of the employee that completed the most hours on each project
- 3 engineered from invoicing data: 
- 4 from project data: number of projects completed with each client
- text analysis of project titles to develop project classifications
+The raw data from the case study company's database consisted of three separate datasets: one describing the client and technical details for each project, another with each employee's time sheet entries, and another with all past invoices.
+This data was combined, cleaned, and rearranged so that each project was described within one row.
+From the timesheet data, 10 variables were engineered including the percent of hours performed by 'professional' employees as opposed to 'technical' employees, and the position of the employee that completed the most hours on each project.
+3 variables were engineered from the invoicing data such as total amount invoiced for a project and mean invoice size for a client.
+Finally, from the project dataset, 2 variables were engineered including the number of projects completed with a client, and project category.
+Text analysis of the project description determined the project category variable.
+The response variable, whether the project was profitable or not, was calculated by subtracting the cost of employee hours from the timesheets from the sum of the paid invoices for a project.
+The total amount invoiced for projects ranged from \$500 to over \$1,000,000.
 
- Discretisation was performed by generating a hierarchical dendrogram of each variable to visualise the clusters.
+34 covariates made up the full dataset, and these were narrowed down to 11 in order to reduce predictive noise and provide a simple model that is comprehendable for stakeholders [@Weisberg2005].
+Variable selection methods included linear regression, random forests, and cforests and the final variables are listed below.
 
-Limiting the predictive model to a concise set of meaningful variables reduces noise and improves predictions.
-Less variables and a simpler model is easier for stakeholders to understand [@Weisberg2005].
-followed by variable importance analysis, variable selection: varialbe importance aalyses done with linear regression, random forest, cforest to represent the range of algroithms that were trialled.
-11 variables chosen:
+* % of Hours Completed by a Professional-level Employee
+* Time span
+* Number of Employees on Project
+* Amount Invoiced for Project
+* Business Category of Client
+* Project Category
+* Total Amount Invoiced from the Client - Past Jobs
+* Discipline
+* Position of Main Project Employee - new variable from ANOVA findings
+* % of Hours by Main Project Employee
+* Billing Type
 
- trials of selected predictive algorithms.
-binary classification problem predicting profit or loss.
-Complex models should be measured against simple models that can be built at a fraction of the computational cost.
+Timespan and the total amount invoiced are unknowns at before a project commences.
+Therefore, these variables were discretised into 5 or 6 categories from which a manager would be able to select.
+The category boundaries were determined by visualising clusters using hierarchical dendrograms.
 
- * Regression - baseline model
-* Naive Bayes - baseline model
-* Bayesian Network - grabage.
-* Random Forest
-* Gradient Boosted Trees
+##### 3.2.1 Research Question 1
 
-To compare the models, the area under the receiver operating characteristic (ROC) curve (AUC) statistic was used.
-An ROC graph visualises the curve from which an AUC score is calculated.
-In this problem a 'positive' is a loss making job.
-For binary classification, AUC is a more meaningful statistic than classification accuracy when the output is a probability that can be applied to the problem.
-The AUC indicates model performance across many thresholds while classification accuracy represents a single threshold.
+To answer the first research question, if loss making projects can be predicted using statistical and machine learning algorithms, the binary classification and blending methods were built and evaluated.
+The area under the receiver operating characteristic (ROC) curve (AUC) statistic was used to compare models and assess performance.
+Models were built on a training subset of data (80%) and validated on the remaining data.
+Therefore, different training and testing sets produced different AUC statistics.
+In order to compare which models performed significantly better than others, an adequate sample size of AUC statistics was required.
+A two-sample power calculation run on 20 AUC output for each model determined that 100 samples were required to achieve a statistical power of 0.8.
 
-a model must be first created using a training set. The model can then make predictions on the test set. this is used to build the ROC
-A model that is perfectly classified would have an AUC = 1.
-A model that predicts as well as random chance would have an AUC = 0.5. An AUC between 0.5 and 1 means the model is performing better than random chance.
+Then, the model output from the validation sets were combined and fed into the three simple and three complex blending models.
+Again, 100 models were built to achieve a power of 0.8 for the variation in AUC test statistics across different divisions of training/testing data.
+A maximum of five training and validation divisions were created from a complete set of individual model results.
 
-In order to compare which models performed significantly better than others, an adequate sample size of results statistics is required.
-Multiple models could be made by using different data in the training vs.
-testing sets, each providing a resulting test statistic (RMSE or AUC).
-Initially, 20 models of each method were created in this fashion.
-Then a two-sample power calculation was run using the two sets of 20 results to determine the sample size to achieve a statistical power of 0.8 (100 to achieve power of 0.8 between boosted trees and naive bayes)
+##### 3.2.2 Research Question 2
 
-Missing Data Imputation.
-All methods except gradient Boosted Trees and Naive Bayes could not handle missing data.
-a complete data set allows for complete sets of predictions from each method, and these predictions could then be blended to further improve results.
-The MICE Random Forest method was chosen for imputation because it has been proven to work well with complex data sets [@Shah2014].
-If similar predictive results were obtained using Boosted Trees imputed data and imputed data, the imputed data must be reasonable.
-The imputed data set was then trialed on the remaining methods and compared to unimputed trials.
+The second research questions asks whether a predictive algorithm can improve a company's profitability?
+This was analysed using a simple decision-making scenario.
+The model with the highest AUC score may not improve company profits the most, because this depends on *which* loss making and profitable projects are accurately being identified.
+Therefore, all blended models as well as individual models were carried forward for this analysis.
+A profit curve was built to compare methods, which is a chart that plots the change in profit the company earns on the y-axis vs.
+the probability threshold on the x-axis.
+For this decision making scenario, the threshold is the probability output value (a value between 0 and 1 that the model predicts for each case), from which it is decided whether a project will be accepted or rejected.
+For example, if the output threshold is 0.6, and a model predicts a project has probability of being a loss-making job of 0.7, the project is not taken on by the case study company.
+All profits and losses from jobs above the threshold were discounted.
+The y-axis of the profit curve is the ratio of profit after rejecting projects based on model output to historical profit:
 
- Once the best methods were selected, they were blended in numerous ways, using both simple averaging techniques and sophisticated machine-learning algorithms.
-These were compared against individual models and the best constructs were selected.
-Six blending methods, ranging from simple to complex, were tested using predictions from the top performing individual models.
-These included simple averaging of the individual model results, building a Logistic Regression model using the individual model results only, a Boosted Tree model using individual model results only, feature weighted linear stacking (FWLS), Random Forests, and Boosted Trees.
-A simplified explanation of FWLS is as a Linear Regression where meta-features as well as model predictions from individual models are included as explanatory variables.
-Then, each meta-feature is interacted with each set of model results [@Sill2009].
- Feature interaction is performed passively due to the nature of how trees are built.
-A split in a node determined by one variable is conditional upon the preceding split, which was based on another variable and so on.
- The simplest method averaged the probability output (a number between 0 and 1) of all three models.
-The next two simplest methods consisted of building a Logistic Regression and Boosted Trees model from the three prediction model outputs only.
-Feature Weighted Linear Stacking (FWLS), Random Forest and Boosted Trees were also tested as blending methods in a more complex scenario where the predictions from each model became additional variables to the original explanatory variables (called meta-features in this context).
-The model predictions were interacted with each original variable so that models that performed more strongly under certain meta-feature states could be weighted as such.
-
-Again, 100 models were built to achieve a power of 0.8 for the variation in AUC test statistics across different divisions of training/testing data.
-A maximum of five blended models were created from a complete set of test results (built from the individual models), then a new set of test results were created for the next five blended models.
-
- Finally the impact of the algorithm on the overall profits of the case study company was analysed via decision-making scenarios.
-This analysis differed from accuracy in predicting profitability, so several blended models as well as individual models were carried forward for this analysis.
-This was done using a profit curve - a chart that plots the change in profit the company earns on the y-axis vs.
-the probability threshold on the x-axis.
-A simple approach was taken for this analysis, where projects with a probability to be a loss making job greater than the threshold were rejected entirely.
-Therefore all profits and losses from jobs above the threshold were discounted.
-An equation defining the change in profit as a percentage of the original profit using threshold rejection is shown below:
-
-\begin{equation}
-\Delta\ Overall\ Profit\ \textrm{(\%)} = \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100
-\end{equation}
+$$
+\begin{aligned}
+Ratio\ of\ Model\ Profit\ to\ Historical\ Profit\ (\%) &= \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100
+\end{aligned}
+$$
 
@@
 
-If the threshold was zero, all jobs were rejected and the profit would be $0.
-If the threshold was 1.0, all jobs were accepted and the profit would be the same as the profit the company actually experienced since the data is a sample of historic projects.
-The aim was to find the optimal threshold point where saying 'no' to a job above that level would result in higher profits, because jobs that were likely to make a loss were being rejected.
-This chart will clarify what percentage of profit increase the company could expect by integrating the algorithm into decision-making.
-
-the curve varies with different divisions of the data.
-Therefore, in order to understand the uncertainty around the profit curve and to determine which blended or individual model performed statistically better than others, a large sample size of curves was required.
-100 to achieve a power of 0.8.
-which was achieved by repeating 5-fold training/testing splits.
-A 95% confidence interval could also be determined around the highest point on each curve.
-The final expected increase in profit and the percentage of projects to be rejected presented a clear scenario that the case study business managers could assess in terms of their business strategy.
+If the threshold were zero, all jobs were rejected, the profit would be $0, and the ratio to historical profits would be 0%.
+If the threshold was 1.0, all jobs were accepted and the profit would be the same as historical profits, giving a ratio of 100%.
+The aim was to find the optimal threshold point where mostly loss-making projects were rejected, resulting in higher profits.
+100 profit curves were built for each method to achieve a power of 0.8 and a 95% confidence interval could also be determined around each threshold point.
+The final expected increase in profit combined with the percentage of projects to be rejected presents a clear scenario that the case study business managers could strategically assess for implementation.
 

### 2016-12-16 finalwriteup/Consulting Profitability Paper/Consulting_Profitability_Paper.tex
finished method in paper

@@
 
-Neural Netowrks and SVM were not purusued because black box A previously
-mentioned disadvantage is that the NN algorithm is a black box because
-the internal structure is too complex for interpretation. They also
-require a lot of training data relative to other methods.
-
-Also, SVM's can be very slow to train and therefore not suitable for
-industry purposes (Auria and Moro 2008). Because it is so important to
-engage decision makers with a model that can explain its results, these
-two methods are not appropriate for the effort estimation problem.
+A range of simple and complex, linear and non-linear methods were
+applied to the problem including Logistic Regression, Random Forests,
+Gradient Boosted Trees, and Naiive Bayes. Furthermore, logistic
+regression, random forests, boosted trees and simple averaging were used
+to blend predictions from individual models. A brief description of each
+method is provided below.
 
@@
 
-binary classification problems, where the objective is to predict the
-probability of an event occurring. Probability is a continuous response
-variable, however predictions from a linear regression would not be
-bound by 0 and 1.
+The aim of this research is to build a model that predicts whether a
+consulting project will be profitable or not. Therefore each method must
+be capable of binary classifcation. Binary classification output from a
+model is the probability, between 0 and 1, of an event occuring. In this
+case, the event is a project being loss-making.
 
@@
 
-Because of the linear relationship, an equation can be fit to the log
-odds of the binary response variable against each of the explanatory
-variables using linear regression. Then, the linear equation describing
-the log odds can be transformed back to probability by taking the
-inverse log i.e.~the exponential and rearranging (Macdonald 1975):
-
-The result in a sigmoidal function bound by 0 and 1. The sigmoidal
-function originates from a linear fit of the log odds in the data.
-Assume linear relationship between covariates and response variable
-
-A coefficient represents the change in the log odds of the response
-variable for each unit increase in an explanatory variable. Therefore,
-taking the exponential of the coefficient is the change in odds of the
-response variable. If the change in odds was 2, then if the explanatory
-variable increased by 1, the response variable event would be twice as
-likely to occur.
-
-Logistic Regression is a good benchmark to compare other binary
-predictive models due to its simplicity and speed (Moore and McCabe
-1989). single linear relationship means Not prone to overfitting
+Logisitic regression models a linear relationship bewteen the log odds
+of a binary response variable and each of the explanatory variables.
+This is performed similarly to linear regression, but is transformed to
+probability of an event occuring by taking the inverse log of the log
+odds and rearranging (Macdonald 1975):
+
+\[
+\begin{aligned}
+\log(odds) &= \beta_0 + X_1\beta_1 \\
+\log\left(\frac{p}{1-p}\right) &= \beta_0 + X_1\beta_1 \\
+-\log\left(\frac{1-p}{p}\right) &= \beta_0 + X_1\beta_1 \\
+-\log\left(\frac{1}{p} - 1\right) &= \beta_0 + X_1\beta_1 \\
+\log\left(\frac{1}{p} - 1\right) &= -\beta_0 - X_1\beta_1 \\
+\exp\left(\log\left(\frac{1}{p} - 1\right)\right) &= \exp(-\beta_0 - X_1\beta_1) \\
+\frac{1}{p} - 1 &= \exp(-\beta_0 - X_1\beta_1) \\
+\frac{1}{p} &= 1 + \exp(-\beta_0 - X_1\beta_1) \\
+p &= \frac{1}{1 + \exp(-\beta_0 - X_1\beta_1)}
+\end{aligned}
+\]
+
+The result in a sigmoidal function bound by 0 and 1. A coefficient
+represents the change in the log odds of the response variable for each
+unit increase in an explanatory variable. Logistic Regression is a good
+benchmark to compare other binary predictive models due to its
+simplicity and speed (Moore and McCabe 1989). Addittionally, because of
+the simple, single linear relationship, the method is not prone to
+overfitting (Perlich, Provost, and Simonoff 2003).
 
@@
 
-single decision trees also suffer from low predictive accuracy and
-instability, so to combat these problems, ensemble tree methods were
-pioneered in the 1990's (Breiman 1996).
-
-The three methods of combining hundreds or thousands of trees turn
-decision trees into high performing, stable predictors.
-
-training multiple trees from the same data by sampling bootstrapped
-training sets with replacement. However, when creating each tree, a
-random subset of attributes (variables) is considered at each split. The
-reduced subset of attributes is resampled for each split in the tree.
-This allows dominant variables to be suppressed for a fraction of the
-splits, allowing the algorithm to explore signals in weaker variables
-(Breiman 2001).
-
-Additionally, ensemble tree methods are faster than SVM's and Neural
-Networks but can perform just as well and even provide insights such as
-variable importance and variable relationships (Sealfon and Gymrek
-2012).
+Ensemble tree methods were pioneered in the 1990's to combat the low
+predictive accuracy and instability of individual decision trees.
+Combining hundreds or thousands of trees transform decision trees into
+high performing, stable predictors (Breiman 1996).
+
+The Random Forests algorithm works by training multiple trees from the
+bootstrapped samples of a dataset. Additionally, when creating each
+tree, a random subset of attributes (covariates) is considered at each
+split. The reduced subset of covariates is resampled for each split in
+the tree. Dominant covariates are then suppressed for a fraction of the
+splits, allowing the algorithm to explore signals in weaker covariates
+(Breiman 2001). Performance statistics and tests run within the
+algorithm provide insights to the modeller, such as variable importance
+and variable relationships (Sealfon and Gymrek 2012).
 
@@
 
-Lastly, the boosted decision tree approach applies gradient descent
-theory to a series of decision trees. The trees are limited to a certain
-depth to maintain simplicity, and each tree models the residuals (or
-errors) of the preceding tree. By modeling the errors, misclassified
-cases are weighted higher than correctly classified cases and influence
-the structure of the latest tree more (Hastie, Tibshirani, and Friedman
-2009). This increased weighting on errors is why the algorithm is called
-boosting. The limited depth of each tree prevents overfitting at each
-stage and the combined result of up to thousands of trees is very
-powerful (Elith, Leathwick, and Hastie 2008).
-
+The boosted decision tree approach applies gradient descent theory to a
+series of decision trees. Performance tends to be higher if individual
+trees are limited to a certain depth to maintain simplicity and prevent
+overfitting (Elith, Leathwick, and Hastie 2008). Unlike Random Forests,
+the trees are sequential, where each tree models the residuals (or
+errors) of the preceding tree.
+
+By modeling the errors, misclassified cases are weighted higher than
+correctly classified cases and influence the structure of the next tree
+more (Hastie, Tibshirani, and Friedman 2009). This increased weighting
+of error cases is called boosting. The successive results of up to
+thousands of trees is very powerful (Elith, Leathwick, and Hastie 2008).
 Caruana and Niculescu-Mizil (2006) tested Boosted Trees, Random Forests,
@@
 classification problems and found that Boosted Trees performed best,
-followed by Random Forests. This demonstrates how ensemble tree methods
-are capable of competing with high-level machine learning algorithms.
-
-Trees are not built on a probabilistic framework, and therefore their
-results cannot be provided in this framework (Louppe and Prettenhofer
-2014).
+followed by Random Forests. This demonstrates how capable ensemble tree
+methods are of competing with high-level machine learning algorithms.
 
@@
 
-The Naive Bayes method works by making conditional independence
-assumptions about the explanatory variables in order to simplify
-probability calculations for the response variable (the response
-variable must be categorical).
+Naive Bayes is a simple, fast method included in this work as it
+provides a good baseline method against which to compare complex methods
+(Caruana and Niculescu-Mizil 2006). It works by making conditional
+independence assumptions about the explanatory variables in order to
+simplify probability calculations for a categorical response variable.
+For example, if predicting a binary variable, \(R\) (with classes 0 or
+1) given two covariates \(X_1\) and \(X_2\), the equations for the
+probability of both outcomes is as follows:
+
+\(p(R = 1|X_1 X_2) = p(X_1|R = 1) \cdot p(X_2|R = 1) \cdot p(R = 1)\)
+
+\(p(R = 0|X_1 X_2) = p(X_1|R = 0) \cdot p(X_2|R = 0) \cdot p(R = 0)\)
 
@@
 the highest probability can then be chosen (F. Provost and Fawcett
-2013).
-
-The advantages of this method are that the conditional independence
-assumption enables very fast calculations and predictions. The method
-can perform well in real world tasks because the assumption of
-independence does not significantly damage predictions (F. Provost and
-Fawcett 2013). Assumptions of independence genearlly not true
-(correlated variables). This is fine for ranking but the output
+2013). The method can perform well in real world tasks because the
+assumption of independence does not significantly damage predictions,
+but amplifies the magnitude of the probability (F. Provost and Fawcett
+2013). This is fine for ranking class 0 against 1 but the output
 probabilities are not accurate statistical probabilities (Caruana and
@@
 
-Naive Bayes method generally provides a good benchmark to compare
-against more complex models that should outperform it (Caruana and
-Niculescu-Mizil 2006).
-
-\subparagraph{3.1.5 Bayesian Networks}\label{bayesian-networks}
-
-A Bayesian Network is a graphical probabilistic model that illustrates
-the conditional dependencies between variables in a data set. The model
-is visually represented by a directed acyclic graph (DAG) and is capable
-of linking the conditional dependency between any variable to another
-variable (Heckerman 1998). The conditional relationships are Bayesian,
-where the probabilities in one node are conditional upon values in nodes
-directed towards it as well as preceding nodes
-
-Bayesian Networks reduce computations required to find the probability
-of a unique combination given other explanatory variable values while
-still taking into account many conditional dependencies that Naive Bayes
-(Barber 2012)
-
-Naive Bayes is the simplest form of a Bayesian Network (Zhang 2004)
-
-Network relationships can be learned from the data, however this is not
-widely included as part of the suite of machine learning methods. The
-conditional variable dependencies are calculated from the data, which in
-turn can define the graph structure. Some conditional dependencies can
-be set before the structure is learned (Barber 2012).
-
-Drawbacks of Bayesian Networks include the difficulty for them to
-process continuous variables.
-
-Bayesian networks have found success in combining deterministic models
-with observed data as well as expert knowledge.(Kragt 2009)
+\subparagraph{3.1.5 Model Blending}\label{model-blending}
+
+Several research groups involved in the high profile data science
+competition, Netflix Prize, developed sophisticated methods of model
+stacking, otherwise known as ensemble methods or model blending (Sill et
+al. 2009). In this paper, the term `blending' will be used to avoid
+confusion with ensemble tree methods. Model blending combines
+predictions from several models that used methods with different
+theoretical foundations. In this way, the strengths of each method can
+be combined. Since the 1990's research in statistics has shown that
+averaging results from different methods can provide better predictive
+accuracy than any single model (Madigan and Raftery 1994).
+
+Six blending methods, ranging from simple to complex, were chosen. These
+included simple averaging of the individual model results, building a
+Logistic Regression model and Boosted Tree model using individual model
+results only, feature weighted linear stacking (FWLS), Random Forests,
+and Boosted Trees. The first two models are simply the average, and
+weighted average of predictions from separate individual models. For
+example, if the Logistic Regression model output a probability of 0.6
+for one case, while a Boosted Tree output 0.7, a simple blended average
+for that case would be a probability of 0.65. The last three methods are
+more complex, and take advantage of potentially meaningful interactions
+between the individual model predictions and the original covariates
+describing each case (called meta-features in the model blending
+literature) (Sill et al. 2009). For example, if the Logistic Regression
+model predicted profitability better than the Boosted Tree model for
+projects with only one team member, the complex blending methods would
+take advantage of this interaction. In summary, in the complex methods,
+the probability outputs from each model are interacted with each of the
+original covariates. Model blending has the potential to elevate output
+from the original methods by combining each of their strengths.
 
@@
 
-It includes first how the data was obtained, the lengthy cleaning
-process, Projects varied from total invoiced amounts of \$500 to over
-\$1,000,000 a project could have thousands of rows of relevant data that
-needed to be converted into a single row per project. 10 engineered
-variables describing timesheets: Percent of hours performed by
-`professional' employees as opposed to `technical' employees, Position
-of the employee that completed the most hours on each project 3
-engineered from invoicing data: 4 from project data: number of projects
-completed with each client text analysis of project titles to develop
-project classifications
-
-Discretisation was performed by generating a hierarchical dendrogram of
-each variable to visualise the clusters.
-
-Limiting the predictive model to a concise set of meaningful variables
-reduces noise and improves predictions. Less variables and a simpler
-model is easier for stakeholders to understand (Weisberg 2005). followed
-by variable importance analysis, variable selection: varialbe importance
-aalyses done with linear regression, random forest, cforest to represent
-the range of algroithms that were trialled. 11 variables chosen:
-
-trials of selected predictive algorithms. binary classification problem
-predicting profit or loss. Complex models should be measured against
-simple models that can be built at a fraction of the computational cost.
+The raw data from the case study company's database consisted of three
+separate datasets: one describing the client and technical details for
+each project, another with each employee's time sheet entries, and
+another with all past invoices. This data was combined, cleaned, and
+rearranged so that each project was described within one row. From the
+timesheet data, 10 variables were engineered including the percent of
+hours performed by `professional' employees as opposed to `technical'
+employees, and the position of the employee that completed the most
+hours on each project. 3 variables were engineered from the invoicing
+data such as total amount invoiced for a project and mean invoice size
+for a client. Finally, from the project dataset, 2 variables were
+engineered including the number of projects completed with a client, and
+project category. Text analysis of the project description determined
+the project category variable. The response variable, whether the
+project was profitable or not, was calculated by subtracting the cost of
+employee hours from the timesheets from the sum of the paid invoices for
+a project. The total amount invoiced for projects ranged from \$500 to
+over \$1,000,000.
+
+34 covariates made up the full dataset, and these were narrowed down to
+11 in order to reduce predictive noise and provide a simple model that
+is comprehendable for stakeholders (Weisberg 2005). Variable selection
+methods included linear regression, random forests, and cforests and the
+final variables are listed below.
 
@@
 \item
-  Regression - baseline model
-\item
-  Naive Bayes - baseline model
-\item
-  Bayesian Network - grabage.
-\item
-  Random Forest
-\item
-  Gradient Boosted Trees
+  \% of Hours Completed by a Professional-level Employee
+\item
+  Time span
+\item
+  Number of Employees on Project
+\item
+  Amount Invoiced for Project
+\item
+  Business Category of Client
+\item
+  Project Category
+\item
+  Total Amount Invoiced from the Client - Past Jobs
+\item
+  Discipline
+\item
+  Position of Main Project Employee - new variable from ANOVA findings
+\item
+  \% of Hours by Main Project Employee
+\item
+  Billing Type
 \end{itemize}
 
-To compare the models, the area under the receiver operating
-characteristic (ROC) curve (AUC) statistic was used. An ROC graph
-visualises the curve from which an AUC score is calculated. In this
-problem a `positive' is a loss making job. For binary classification,
-AUC is a more meaningful statistic than classification accuracy when the
-output is a probability that can be applied to the problem. The AUC
-indicates model performance across many thresholds while classification
-accuracy represents a single threshold.
-
-a model must be first created using a training set. The model can then
-make predictions on the test set. this is used to build the ROC A model
-that is perfectly classified would have an AUC = 1. A model that
-predicts as well as random chance would have an AUC = 0.5. An AUC
-between 0.5 and 1 means the model is performing better than random
-chance.
-
-In order to compare which models performed significantly better than
-others, an adequate sample size of results statistics is required.
-Multiple models could be made by using different data in the training
-vs.~testing sets, each providing a resulting test statistic (RMSE or
-AUC). Initially, 20 models of each method were created in this fashion.
-Then a two-sample power calculation was run using the two sets of 20
-results to determine the sample size to achieve a statistical power of
-0.8 (100 to achieve power of 0.8 between boosted trees and naive bayes)
-
-Missing Data Imputation. All methods except gradient Boosted Trees and
-Naive Bayes could not handle missing data. a complete data set allows
-for complete sets of predictions from each method, and these predictions
-could then be blended to further improve results. The MICE Random Forest
-method was chosen for imputation because it has been proven to work well
-with complex data sets (Shah et al. 2014). If similar predictive results
-were obtained using Boosted Trees imputed data and imputed data, the
-imputed data must be reasonable. The imputed data set was then trialed
-on the remaining methods and compared to unimputed trials.
-
-Once the best methods were selected, they were blended in numerous ways,
-using both simple averaging techniques and sophisticated
-machine-learning algorithms. These were compared against individual
-models and the best constructs were selected. Six blending methods,
-ranging from simple to complex, were tested using predictions from the
-top performing individual models. These included simple averaging of the
-individual model results, building a Logistic Regression model using the
-individual model results only, a Boosted Tree model using individual
-model results only, feature weighted linear stacking (FWLS), Random
-Forests, and Boosted Trees. A simplified explanation of FWLS is as a
-Linear Regression where meta-features as well as model predictions from
-individual models are included as explanatory variables. Then, each
-meta-feature is interacted with each set of model results (Sill et al.
-2009). Feature interaction is performed passively due to the nature of
-how trees are built. A split in a node determined by one variable is
-conditional upon the preceding split, which was based on another
-variable and so on. The simplest method averaged the probability output
-(a number between 0 and 1) of all three models. The next two simplest
-methods consisted of building a Logistic Regression and Boosted Trees
-model from the three prediction model outputs only. Feature Weighted
-Linear Stacking (FWLS), Random Forest and Boosted Trees were also tested
-as blending methods in a more complex scenario where the predictions
-from each model became additional variables to the original explanatory
-variables (called meta-features in this context). The model predictions
-were interacted with each original variable so that models that
-performed more strongly under certain meta-feature states could be
-weighted as such.
-
-Again, 100 models were built to achieve a power of 0.8 for the variation
-in AUC test statistics across different divisions of training/testing
-data. A maximum of five blended models were created from a complete set
-of test results (built from the individual models), then a new set of
-test results were created for the next five blended models.
-
-Finally the impact of the algorithm on the overall profits of the case
-study company was analysed via decision-making scenarios. This analysis
-differed from accuracy in predicting profitability, so several blended
-models as well as individual models were carried forward for this
-analysis. This was done using a profit curve - a chart that plots the
-change in profit the company earns on the y-axis vs.~the probability
-threshold on the x-axis. A simple approach was taken for this analysis,
-where projects with a probability to be a loss making job greater than
-the threshold were rejected entirely. Therefore all profits and losses
-from jobs above the threshold were discounted. An equation defining the
-change in profit as a percentage of the original profit using threshold
-rejection is shown below:
-
-\begin{equation}
-\Delta\ Overall\ Profit\ \textrm{(\%)} = \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100
-\end{equation}
+Timespan and the total amount invoiced are unknowns at before a project
+commences. Therefore, these variables were discretised into 5 or 6
+categories from which a manager would be able to select. The category
+boundaries were determined by visualising clusters using hierarchical
+dendrograms.
+
+\subparagraph{3.2.1 Research Question 1}\label{research-question-1}
+
+To answer the first research question, if loss making projects can be
+predicted using statistical and machine learning algorithms, the binary
+classification and blending methods were built and evaluated. The area
+under the receiver operating characteristic (ROC) curve (AUC) statistic
+was used to compare models and assess performance. Models were built on
+a training subset of data (80\%) and validated on the remaining data.
+Therefore, different training and testing sets produced different AUC
+statistics. In order to compare which models performed significantly
+better than others, an adequate sample size of AUC statistics was
+required. A two-sample power calculation run on 20 AUC output for each
+model determined that 100 samples were required to achieve a statistical
+power of 0.8.
+
+Then, the model output from the validation sets were combined and fed
+into the three simple and three complex blending models. Again, 100
+models were built to achieve a power of 0.8 for the variation in AUC
+test statistics across different divisions of training/testing data. A
+maximum of five training and validation divisions were created from a
+complete set of individual model results.
+
+\subparagraph{3.2.2 Research Question 2}\label{research-question-2}
+
+The second research questions asks whether a predictive algorithm can
+improve a company's profitability? This was analysed using a simple
+decision-making scenario. The model with the highest AUC score may not
+improve company profits the most, because this depends on \emph{which}
+loss making and profitable projects are accurately being identified.
+Therefore, all blended models as well as individual models were carried
+forward for this analysis. A profit curve was built to compare methods,
+which is a chart that plots the change in profit the company earns on
+the y-axis vs.~the probability threshold on the x-axis. For this
+decision making scenario, the threshold is the probability output value
+(a value between 0 and 1 that the model predicts for each case), from
+which it is decided whether a project will be accepted or rejected. For
+example, if the output threshold is 0.6, and a model predicts a project
+has probability of being a loss-making job of 0.7, the project is not
+taken on by the case study company. All profits and losses from jobs
+above the threshold were discounted. The y-axis of the profit curve is
+the ratio of profit after rejecting projects based on model output to
+historical profit:
+
+\[
+\begin{aligned}
+Ratio\ of\ Model\ Profit\ to\ Historical\ Profit\ (\%) &= \frac{\sum_{p = 1}^{N} I(Pr_p < threshold) \cdot Profit_p}{\sum_{p = 1}^{N} Profit_p} \cdot100
+\end{aligned}
+\]
 
@@
 
-If the threshold was zero, all jobs were rejected and the profit would
-be \$0. If the threshold was 1.0, all jobs were accepted and the profit
-would be the same as the profit the company actually experienced since
-the data is a sample of historic projects. The aim was to find the
-optimal threshold point where saying `no' to a job above that level
-would result in higher profits, because jobs that were likely to make a
-loss were being rejected. This chart will clarify what percentage of
-profit increase the company could expect by integrating the algorithm
-into decision-making.
-
-the curve varies with different divisions of the data. Therefore, in
-order to understand the uncertainty around the profit curve and to
-determine which blended or individual model performed statistically
-better than others, a large sample size of curves was required. 100 to
-achieve a power of 0.8. which was achieved by repeating 5-fold
-training/testing splits. A 95\% confidence interval could also be
-determined around the highest point on each curve. The final expected
-increase in profit and the percentage of projects to be rejected
-presented a clear scenario that the case study business managers could
-assess in terms of their business strategy.
+If the threshold were zero, all jobs were rejected, the profit would be
+\$0, and the ratio to historical profits would be 0\%. If the threshold
+was 1.0, all jobs were accepted and the profit would be the same as
+historical profits, giving a ratio of 100\%. The aim was to find the
+optimal threshold point where mostly loss-making projects were rejected,
+resulting in higher profits. 100 profit curves were built for each
+method to achieve a power of 0.8 and a 95\% confidence interval could
+also be determined around each threshold point. The final expected
+increase in profit combined with the percentage of projects to be
+rejected presents a clear scenario that the case study business managers
+could strategically assess for implementation.
 
@@
 
-\hypertarget{ref-Auria2008}{}
-Auria, Laura, and Rouslan A Moro. 2008. ``Support Vector Machines (Svm)
-as a Technique for Solvency Analysis.'' Journal Article.
-
-\hypertarget{ref-Barber2012}{}
-Barber, David. 2012. \emph{Bayesian Reasoning and Machine Learning}.
-Book. Cambridge University Press.
-
 \hypertarget{ref-Breiman1996}{}
@@
 
-\hypertarget{ref-Heckerman1998}{}
-Heckerman, David. 1998. \emph{A Tutorial on Learning with Bayesian
-Networks}. Book. Springer.
-
 \hypertarget{ref-Kim2004}{}
@@
 
-\hypertarget{ref-Kragt2009}{}
-Kragt, Marit E. 2009. \emph{A Beginners Guide to Bayesian Network
-Modelling for Integrated Catchment Management}. Book. Landscape Logic.
-
 \hypertarget{ref-Kumar2007}{}
@@
 
-\hypertarget{ref-Louppe2014}{}
-Louppe, Gilles, and Peter Prettenhofer. 2014. ``Gradient Boosted
-Regression Trees.'' Online Multimedia. PyData.
-
 \hypertarget{ref-Lovallo2003}{}
@@
 
+\hypertarget{ref-Madigan1994}{}
+Madigan, David, and Adrian E Raftery. 1994. ``Model Selection and
+Accounting for Model Uncertainty in Graphical Models Using Occam's
+Window.'' Journal Article. \emph{Journal of the American Statistical
+Association} 89 (428): 1535--46.
+
 \hypertarget{ref-Matson1993}{}
@@
 
-\hypertarget{ref-Shah2014}{}
-Shah, Anoop D, Jonathan W Bartlett, James Carpenter, Owen Nicholas, and
-Harry Hemingway. 2014. ``Comparison of Random Forest and Parametric
-Imputation Models for Imputing Missing Data Using Mice: A Caliber
-Study.'' Journal Article. \emph{American Journal of Epidemiology},
-kwt312.
-
 \hypertarget{ref-Shin2015}{}
@@
 
-\hypertarget{ref-Zhang2004}{}
-Zhang, Harry. 2004. ``The Optimality of Naive Bayes.'' Journal Article.
-\emph{AA} 1 (2): 3.
-
 \end{document}

### 2016-12-19 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd
finished results and discussion of paper

@@
 
-opts_chunk$set(fig.width=6.7, dpi = 200, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)
+opts_chunk$set(fig.pos = 'h', dpi = 200, warning = FALSE, message = FALSE, echo = FALSE, warning = FALSE)
+
+# setwd("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data")
 ```
@@
 
-Naive Bayes is a simple, fast method included in this work as it provides a good baseline method against which to compare complex methods [@Caruana2006].
+Naive Bayes is a simple, fast method included in this work because, similarly to Logistic Regression, it provides a good baseline method against which to compare complex methods [@Caruana2006].
 It works by making conditional independence assumptions about the explanatory variables in order to simplify probability calculations for a categorical response variable.
@@
 
-The second research questions asks whether a predictive algorithm can improve a company's profitability?
-This was analysed using a simple decision-making scenario.
-The model with the highest AUC score may not improve company profits the most, because this depends on *which* loss making and profitable projects are accurately being identified.
-Therefore, all blended models as well as individual models were carried forward for this analysis.
-A profit curve was built to compare methods, which is a chart that plots the change in profit the company earns on the y-axis vs.
-the probability threshold on the x-axis.
-For this decision making scenario, the threshold is the probability output value (a value between 0 and 1 that the model predicts for each case), from which it is decided whether a project will be accepted or rejected.
-For example, if the output threshold is 0.6, and a model predicts a project has probability of being a loss-making job of 0.7, the project is not taken on by the case study company.
-All profits and losses from jobs above the threshold were discounted.
-The y-axis of the profit curve is the ratio of profit after rejecting projects based on model output to historical profit:
+The second research question asks whether a predictive algorithm can improve a company's profitability.
+The model with the highest AUC score may not improve company profits the most, because the magnitude of the profits and losses from each project have not been analysed.
+Therefore, all blended and individual models were carried forward to be assessed in a value framework.
+
+Once a model outputs scores for the probability of each job being loss-making, the question then arises at what probability would a manager round the probability up to 1?
+And then what action would be taken?
+A threshold point between 0 and 1 can dictate when a project is rounded up, and classified as 'loss-making'.
+For this analysis, a simple decision scenario is tested where a manager rejects all projects with output above the threshold.
+For example, if the threshold is 0.6, and a project is predicted to have a probability of loss-making of 0.7, the project is rejected and profits or losses associated with the project are ignored.
+
+To test each threshold between 0 and 1, profit curves were built for each method.
+This chart plots the probability threshold on the x-axis and the ratio of model-influenced profits over historical  company profits on the y-axis.
+The ratio represented by the y-axis is calculated using the formula below:
 
@@
 If the threshold was 1.0, all jobs were accepted and the profit would be the same as historical profits, giving a ratio of 100%.
+The total profits were plotted for each threshold and joined to make a profit curve.
 The aim was to find the optimal threshold point where mostly loss-making projects were rejected, resulting in higher profits.
@@
 
-The predictive formula was re-structured so that the 11 explanatory variables predicted a new response variable: profitable/unprofitable projects.
-The five prediction methods, were initially built without imputing data, by using the maximum amount of data possible depending on the method.
-Boosted Trees and Naive Bayes are able to process data with missing values, so all data could be input [@gbm; @e1071].
-100 models of each method were built.
-The violin plot below summarises the AUC values produced by each method and the 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
+The violin plot below summarises the distributions of the 100 AUC values produced by each method.
+The 'violins' are coloured according to whether the distributions significantly vary to Boosted Trees using a critical value of 0.05.
+Boosted Trees was chosen as the comparison method because it was expected to perform well.
 
-```{r reduced_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
-Subsets of the data were used for Logistic Regression and Random Forests in order to provide datasets without missing values.")}
-
-reduc.totAUC <- readRDS("reduc_df_auc.rds")
-
-#significant difference??
-index = c(1,3,4)
-pvals = data.frame("method" = c('log','rf','naiive'), 'p_value' = NA)
-for(i in seq_along(index)){
-        pvals$p_value[i] = t.test(reduc.totAUC[[2]], reduc.totAUC[[index[i]]])$p.value %>%
-                format(digits = 2) %>% as.numeric
-}
-
-signif = pvals %>% filter(p_value < 0.05) %>% .$method
-reduc.totAUC = melt(reduc.totAUC)
-reduc.totAUC$sigdiff = ifelse(reduc.totAUC$L1 %in% signif, TRUE, FALSE)
-
-ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
-        geom_violin(aes(fill = sigdiff, colour = sigdiff), alpha = 0.8) + 
-        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
-        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
-                     geom="text", colour = "gray20", size =3, 
-                     hjust = -.2, angle = 40, vjust = -.2
-                     ) +
-        scale_fill_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from Boosted Trees") +
-        scale_colour_manual(values = c('sienna1','navajowhite'),
-                          name = "Significantly different from Boosted Trees") +
-        labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 models of Each Method") +
-        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
-                         "log" = "Logistic Regression",
-                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
-        theme(legend.position = "bottom")
-
-```
-
-The AUC performance of Logistic Regression and the Random Forest algorithms cannot be statistically differentiated from Boosted Trees.
-
-Next, data imputation methods were trialed which would make the complete data set available to Logistic Regression and Random Forests.
-Again, 100 models were required to achieve a power of 0.8 with respect to Boosted Trees
-
-```{r mice_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss.
-Each method was fed the same imputed full dataset.")}
+```{r mice_violin, fig.pos = 'H', fig.height = 5.5}
 
@@
         labs(x = "Method", y = "Area Under Curve (AUC)",
-             title = "AUC Distribution from 100 Models of Each Method - Imputed Data") +
+             title = "AUC Distribution from 100 Models of Each Method") +
         scale_x_discrete(labels = c("boost" = "Boosted Trees", 
@@
 
-Boosted Trees, Logistic Regression, and Random Forest all performed significantly better than the baseline algorithm, Naive Bayes, however none outperformed Boosted Trees.
+The AUC distrubtions of Logistic Regression and the Random Forest algorithms could not be statistically differentiated from Boosted Trees, while the Naive Bayes distribution was significantly lower.
+The three high-performing methods had AUC values between 0.755 and 0.764.
+This indicates predictions were well above random chance, where AUC = 0.5.
 
-Testing Boosted Trees, Random Forests, Naive Bayes, Logistic Regression and Bayesian Networks on the binary classification problem showed that Boosted Trees, Logistic Regression and Random Forests performed best (according to AUC).
 Naive Bayes and Logistic Regression were included as baseline models and it was expected that the more complex models would outperform these.
 Therefore it was surprising that results from 100 Logistic Regression models were not statistically significantly lower than Boosted Trees.
-
 A possible explanation for this, according to the literature, is that there was not enough data for the ensemble trees to learn the complex decision rules at which they excel.
@@
 Logistic Regression on the other hand is capable of only one decision boundary (which does not have to be parallel to the variable axes) and is not prone to overfitting [@Perlich2003].
-This may explain Logistic Regression's comparatively high performance on the case study's small but complex data set.
-
-Boosted trees, Random Forests and Logistic Regression all had mean AUC values between 0.755 to 0.764.
-In summary, the individual binary classification models performed well above random chance (AUC = 0.5).
-Whether the model is worth implementing in the work place is dependent on the extent to which the algorithm would improve 'bottom line' profits for the business and if the model can affect decisions in practice.
-
-- NB and BN were bad - one sentence
+This may explain Logistic Regression's comparatively high performance on the case study's data set, which is small for the range and complexity of the described projects.
 
@@
 
-Blended models combine the predictions from each high-performing individual model to create averaged or 'blended' predictions.
-In this case, the Random Forest, Boosted Trees, and Logistic Regression models were chosen.
-Through variable importance studies, 6 variables were included as metafeatures.
-All six methods were compared against the original Logistic Regression model's AUC distribution (as it was not statistically different from the ensemble trees)
+Six blended models combined the predictions from the three high-performing individual models: Logistic Regression, Random Forests, and Boosted Trees to create 'blended' predictions.
+There were 3 simple blended models, and 3 complex models that interacted original covariates with the individual model predictions.
+Through variable importance analysis on the original covariates and the 3 new variables containing model predictions, only 6 original covariates were included as metafeatures.
+All six blended methods' AUC values were compared against the Logistic Regression AUC distribution, which was chosen because it was the simplest high performing individual model.
 
-```{r blend_violin, fig.height = 5.5, fig.cap = c("Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss.
-100 models were built for each method.")}
+```{r blend_violin, fig.height = 5.5, fig.pos = 'h'}
 
@@
 
-The above plot shows that 4 methods had a higher mean AUC than the original Logistic Regression model:
-
-* the simple blended Logistic Regression 
-* the simple average 
-* and the two Boosted Tree models.
-
-The simple averaged method and simple logistic regression could achieve a statistical power of 0.8 with 150 samples (models), whereas the boosted trees required thousands.
-The additional 50 models were run for the simple logistic regression and simple average to reveal similar distributions that were indeed statistically significantly higher than the individual logistic regression AUC distribution.
+None of the blended methods were significantly higher than the individual Logistic Regression model with data from 100 models.
+However, the simple averaged method and simple logistic regression could achieve a statistical power of 0.8 with 150 samples (models), whereas the other blended methods required thousands.
+50 more models were run for the simple logistic regression and simple average blend to confirm their AUC distributions were significantly higher than the individual Logistic Regression.
 P-values from two sample t-test were 0.0019 and 0.0021 respectively.
@@
 Blended models improved the mean AUC from 0.759 to 0.77, which is an increase of `r ((0.77-0.759)/0.759) %>% "*"(100) %>% round(1)`%.
-It was expected that model blending would improve predictions because it combined the strengths of individual models with different theoretical foundations.
+This confirms the expectation that model blending would improve predictions by combining the strengths of different methods' theoretical foundations.
 
- The trials of different blending methods demonstrated that again, the simplest methods worked best with the case study data.
-In this incident, averaging the results of the best three single models (Logistic Regression, Random Forests, and Boosted Trees) or taking a Logistic Regression of the three models outperformed more complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated interaction of the individual model results and original variables (meta-features).
-Simple blended models could perform better than complex methods if the data is not big enough for the complex models to learn the more intricate patterns at which they excel.
-This was observed in the single model methods.
-their success in the 2009 Netflix competition generated some publications.
-The Netflix data set comprised of almost 3,000,000 observations, so the size of the data could have enabled the success of complex blending methods such as FWLS [@Sill2009]
+Again, the simplest methods worked best.
+In this case, averaging the results of the best three individual models or taking a Logistic Regression of the three models' outputs outperformed the complex Boosted Trees, Random Forests, and Logistic Regression blends that facilitated meta-feature interactions.
+Simiarly to the individual models, Simple methods may outperform complex methods if the data is not big enough for complex methods to learn the intricate patterns at which they excel.
+The Netflix data set comprised of almost 3,000,000 observations, so the size of the data could have enabled the success of their complex blending methods (Feature Weighted Linear Stacking) [@Sill2009]
+
+The outcomes in this section show that statistical and machine learning methods can predict whether a consulting project will be profitable or not to some extent.
+The average, or weighted average (logistic regression) blend of individual Logistic Regression, Boosted Trees, and Random Forests models produced the highest AUC values.
+However, the business value of the model has not been proven.
+The following section explores whether the models would improve overall profits for the case study business.
 
@@
 
-This chapter first presents the full range of method results in terms of improvements to the case study's bottom line.
-A business decision making scenario was created so that profit curves based on the decision rule could be built and analysed.
-From the profit curves, optimal probability thresholds could be derived for each method (individual and blended methods).
-The predictive models output a 'probability' between 0 and 1 that each project will be a loss making job (where probability = 1 indicates a loss making job).
-The question the arises, at what probability would a decision-maker round the probability up to 1 or down to 0?
-And what business decision would then be made?
-To find the threshold point for rounding, an experimental business-scenario was tested.
-At a given threshold, all projects with probability outputs above the threshold were considered too risky, and were rejected.
-All profits and losses from these projects were forfeited, while the profits from the remaining jobs (below the threshold) were summed to give a revised total profit.
-This profit calculation was made on a range of thresholds between 0 and 1 at 0.05 increments.
-The total profits were plotted for each threshold and joined to make a profit curve
+This chapter transports model results into the value framework to assess impact on the case study's overall profitability.
+The plot below illustrates the distribution of profit curves for the 100 models built using each method.
+Solid lines join the mean values at each threshold point and the grey ribbon represents the 95% confidence interval for the profit ratio.
 
-The plot below illustrates the distribution of profit curves for each method, where the solid lines join the mean values at each threshold point.
-The grey ribbon illustrates the 95% confidence interval for the profit improvement ratio at each threshold point for the 100 models.
-
-```{r profit_curve, fig.height = 6, fig.cap = c("Profit curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.")}
+```{r profit_curve, fig.height = 6, fig.pos = 'h'}
 
@@
 
-Two blended models clearly outperformed the individual models as shown in the above plot with higher profit ratios as well as tighter confidence intervals.
-The simple Logistic Regression blend performed best with the highest mean profit ratio of 109% with a standard deviation of 1.28%.
-%.
-This means that for the simple blended Logistic Regression model, if all jobs above the probability threshold 0.6 were rejected, the profits would increase on average by 9% in comparison to historical profits.
+From the profit curves, the optimal probability thresholds are where the profit curves peak, and peaks above 100% indicate the model assisted profits outperformed historical profits.
+The two simple blended models clearly outperformed all other models with higher profit ratios as well as tighter confidence intervals.
+The simple Logistic Regression blend performed best with the highest mean profit ratio of 109% and a standard deviation of 1.28%.
+This means that if all jobs above the probability threshold 0.6 were rejected, the profits would increase on average by 9% in comparison to historical profits.
+Only 4.3% of projects were rejected in this case.
 
-Profit curves from the simple average blend and Logistic Regression blended models outperformed the complex blended models, which follows logically from their significantly higher AUC distributions.
-The simple average blended model produced a profit curve with an almost identical profit improvement (and standard deviation) to the Logistic Regression blend.
-In the data, projects assigned a probability higher than 0.6 represent only 4.3% of all projects.
+Profit curves from the simple blended models outperformed the complex blended models, which follows logically from their significantly higher AUC distributions.
+The shaded grey 95% confidence intervals show that none of the lower bounds for the original methods' or complex blending methods were above 100% on the y-axis.
+That is, it cannot be said with 95% confidence that these methods would produce a mean profit higher than historical profits in the given decision framework.
+It is interesting that despite the reasonable AUC values, the profit curves for the individual and complex blended models do not make a complelling case for adoption into business practice.
 
-The shaded grey 95% confidence intervals in the Profit Curve plot shows that with the 100 sample models that were analysed, none of the lower bounds for the original methods' or complex blending methods were above 100% on the y-axis.
-That is, it cannot be said with 95% confidence that the original methods would produce a mean profit higher than historical profits in the given decision framework.
-Clearly this level of certainty is not beneficial for the case-study business to adopt.
-
-It is not clear why simple linear models and the averaging model outperformed ensemble tree blending methods.
-As previously stated, the ensemble trees may not have received enough data to adequately learn the complex series of rules they develop.
-The only differences between FWLS and the simple Logistic Regression (that performed best) were two additional explanatory variables and four interaction terms.
-The number of variables was not high for the amount of data since according to @Peduzzi1996, 10 events per explanatory variable or more avoids the risk of biased estimation of variable coefficients in Logistic Regression.
-The data contained 315 events per explanatory variable and did not pose that risk.
-Nevertheless, the additional variables must have added misleading noise to the model.
-
-To conclude, because of the promising AUC results, it was logical that the models translated into financial benefits for the company.
-The 9% improvement in profits produced by the simple blended Logistic Regression is reasonable, and may be high enough to trigger further cost-benefit analyses and the development of a more comprehensive framework describing different decision scenarios.
-
-- decision scenario
-
-- profit curves
-
-- prototype decision support interface
+The simple blended Logistic Regression model on the other hand, a mean 9% improvement in profits, with a lower bound of 6.5%.
+In terms of implementing the model, this improvement may be high enough to trigger further development and analysis of a more sophisticated framework describing decision making scenarios.
 
@@
 
-The general aim was to use statistical techniques to predict the profitability of projects for a case study consulting business using their internal CRM data.
-This was rigorously completed by approaching the prediction of 'profitability' as a binary classification problem predicting either profit or loss.
-Several statistical and machine learning approaches were applied including Naive Bayes, Bayesian Networks, Linear Regression, Random Forests, gradient Boosted Trees as well as multiple methods of blending the individual models' output.
+Complex ensemble tree methods did not outpereform simple linear methods, however, the simple blended predictions of all methods
 
-AUC = 0.76). This was achieved by 3 individual methods while various techniques that blended the individual methods improved results further (AUC = 0.77).
+The general aim of this work was twofold.
+First, statistical and machine learning techniques were tested to determine whether the profitability of projects for a case study consulting business could be predicted using their internal CRM data.
+This was rigorously completed by approaching the prediction of 'profitability' as a binary classification problem with classes profititable or loss-making.
+Naive Bayes, Logistic Regression, Random Forests, gradient Boosted Trees as well as simple and complex blending methods were applied to the problem.
+Secondly, this project tested whether predictions from the models could be proven to improve the overall profitability of the case study company and by how much.
 
-predictive models could be shown to improve the overall profitability (bottom line) of the case study business.
-A range of probability threshold values were trialed for each method using a decision framework where projects scored below the threshold were accepted, while projects above the threshold were rejected.
+Predicting whether a job would be profitable or not (binary classification) was possible to a level significantly better than random assignment (AUC = 0.76).
+This mean statistic was achieved by 100 different training/testing divisions of the data, built for each of the 3 individual methods.
+Simple averaging and simple Logistic Regression blending methods improved results further to a mean of AUC = 0.77.
+
+Two models could also be shown to improve the overall profitability of the case study business.
+This was done by converting the binary predictions from the first analysis into profit curves that plotted model-influenced profit improvements over a range of probability threshold values.
+To determine model-influenced profit improvements, a managerial decision framework was used where projects scored below the threshold were accepted, while projects above the threshold were rejected.
 If a project was rejected, the profits and losses were forfeited, and the remaining accepted profits and losses were summed.
 
+100 training/testing divisions of the data were user to produce 100 profit curves so that a 95% confidence interval could be framed around the curves.
 Final results showed the simple Logistic Regression blend of the individual Logistic Regression, Random Forest and Boosted Tree models improved profits the most.
-The 95% confidence interval for these improvements was between 6.5% and 11.5% using a probability threshold of 0.6 (approximately 4.3% of projects).
+The 95% confidence interval for these improvements was between 6.5% and 11.5% using a probability threshold of 0.6 (approximately 4.3% of projects were rejected).
 
-These results contribute significantly to the research in cost estimation in three ways: the applied methods, the decision framework, and the appeal to user trust.
+These results contribute significantly to the research in cost estimation in two ways: the applied methods, and the value assessment based on a managerial decision framework.
 Ensemble tree methods and blending had been applied minimally to cost estimation previously, even though ensemble trees provide insight into model structure while predicting at a similar level to Neural Networks.
-Next, previous studies have verified predictive accuracy but stopped short of how the algorithm would affect decisions and what the measured benefits would be.
+Next, previous studies compared accuracy of their predictive models, but stopped short of computing how the algorithm would affect decisions and what the measured benefits would be.
 This study presented a clear framework for how a business could improve profits by applying the algorithm.
 
-Further work is required to test user confidence in the output.
-Another topic identified for future research was to test how well managers estimate some numeric project input variables.
-In particular, time span, team size, total invoiced amount, and percentage of hours completed by professionals should be tested as these variables were calculated post project completion.
-Time span and total invoiced amount were discretised into wide categories, which should be easier for a manager to choose between.
-
-future work: user interface.
+Further work is required to measure user confidence in the output.
+Another topic identified for future research is to test how well managers estimate input variables at project conception.
+This is paraticularly relevant for time span and total invoiced amount which were discretised into 4 to 5 categories.
 
 Overall, this work has successfully built a mathematical blend of Logistic Regression, Random Forests, and Boosted Tree models, from a consulting company's internal project data.
-This blended model can predict whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business.
+This blended model predicts whether a project will be profitable or not and in a reasonable decision framework, can guide managers in rejecting financially risky projects and improving profitability of the business by a mean of 9%.
 
@@
 This research was supported by a scholarship/ ACEMS?
+
+Figure Captions
+===============
+
+1. mice_violin: Violin plot vertically illustrating the distribution of AUC values from each of the methods when predicting profit/loss. Each method was fed the same imputed full dataset.
+2. blend_violin: Violin plot vertically illustrating the distribution of AUC values from each of the blending methods when predicting profit/loss. 100 models were built for each method.
+3. profit_curve: Profit curves summarising results from 100 models of 9 methods: 3 simple blends, 3 complex blends, and the original 3 best methods.
 

### 2016-12-19 finalwriteup/Consulting Profitability Paper/Consulting_Profitability_Paper.tex
finished results and discussion of paper

@@
 
-Naive Bayes is a simple, fast method included in this work as it
-provides a good baseline method against which to compare complex methods
-(Caruana and Niculescu-Mizil 2006). It works by making conditional
-independence assumptions about the explanatory variables in order to
-simplify probability calculations for a categorical response variable.
-For example, if predicting a binary variable, \(R\) (with classes 0 or
-1) given two covariates \(X_1\) and \(X_2\), the equations for the
-probability of both outcomes is as follows:
+Naive Bayes is a simple, fast method included in this work because,
+similarly to Logistic Regression, it provides a good baseline method
+against which to compare complex methods (Caruana and Niculescu-Mizil
+2006). It works by making conditional independence assumptions about the
+explanatory variables in order to simplify probability calculations for
+a categorical response variable. For example, if predicting a binary
+variable, \(R\) (with classes 0 or 1) given two covariates \(X_1\) and
+\(X_2\), the equations for the probability of both outcomes is as
+follows:
 
@@
 
-The second research questions asks whether a predictive algorithm can
-improve a company's profitability? This was analysed using a simple
-decision-making scenario. The model with the highest AUC score may not
-improve company profits the most, because this depends on \emph{which}
-loss making and profitable projects are accurately being identified.
-Therefore, all blended models as well as individual models were carried
-forward for this analysis. A profit curve was built to compare methods,
-which is a chart that plots the change in profit the company earns on
-the y-axis vs.~the probability threshold on the x-axis. For this
-decision making scenario, the threshold is the probability output value
-(a value between 0 and 1 that the model predicts for each case), from
-which it is decided whether a project will be accepted or rejected. For
-example, if the output threshold is 0.6, and a model predicts a project
-has probability of being a loss-making job of 0.7, the project is not
-taken on by the case study company. All profits and losses from jobs
-above the threshold were discounted. The y-axis of the profit curve is
-the ratio of profit after rejecting projects based on model output to
-historical profit:
+The second research question asks whether a predictive algorithm can
+improve a company's profitability. The model with the highest AUC score
+may not improve company profits the most, because the magnitude of the
+profits and losses from each project have not been analysed. Therefore,
+all blended and individual models were carried forward to be assessed in
+a value framework.
+
+Once a model outputs scores for the probability of each job being
+loss-making, the question then arises at what probability would a
+manager round the probability up to 1? And then what action would be
+taken? A threshold point between 0 and 1 can dictate when a project is
+rounded up, and classified as `loss-making'. For this analysis, a simple
+decision scenario is tested where a manager rejects all projects with
+output above the threshold. For example, if the threshold is 0.6, and a
+project is predicted to have a probability of loss-making of 0.7, the
+project is rejected and profits or losses associated with the project
+are ignored.
+
+To test each threshold between 0 and 1, profit curves were built for
+each method. This chart plots the probability threshold on the x-axis
+and the ratio of model-influenced profits over historical company
+profits on the y-axis. The ratio represented by the y-axis is calculated
+using the formula below:
 
@@
 was 1.0, all jobs were accepted and the profit would be the same as
-historical profits, giving a ratio of 100\%. The aim was to find the
-optimal threshold point where mostly loss-making projects were rejected,
-resulting in higher profits. 100 profit curves were built for each
-method to achieve a power of 0.8 and a 95\% confidence interval could
-also be determined around each threshold point. The final expected
-increase in profit combined with the percentage of projects to be
-rejected presents a clear scenario that the case study business managers
-could strategically assess for implementation.
+historical profits, giving a ratio of 100\%. The total profits were
+plotted for each threshold and joined to make a profit curve. The aim
+was to find the optimal threshold point where mostly loss-making
+projects were rejected, resulting in higher profits. 100 profit curves
+were built for each method to achieve a power of 0.8 and a 95\%
+confidence interval could also be determined around each threshold
+point. The final expected increase in profit combined with the
+percentage of projects to be rejected presents a clear scenario that the
+case study business managers could strategically assess for
+implementation.
 
@@
 
-The predictive formula was re-structured so that the 11 explanatory
-variables predicted a new response variable: profitable/unprofitable
-projects. The five prediction methods, were initially built without
-imputing data, by using the maximum amount of data possible depending on
-the method. Boosted Trees and Naive Bayes are able to process data with
-missing values, so all data could be input (Ridgeway 2015; Meyer et al.
-2014). 100 models of each method were built. The violin plot below
-summarises the AUC values produced by each method and the `violins' are
-coloured according to whether the distributions significantly vary to
-Boosted Trees using a critical value of 0.05.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Consulting_Profitability_Paper_files/figure-latex/reduced_violin-1.pdf}
-\caption{Violin plot vertically illustrating the distribution of AUC
-values from each of the methods when predicting profit/loss. Subsets of
-the data were used for Logistic Regression and Random Forests in order
-to provide datasets without missing values.}
-\end{figure}
-
-The AUC performance of Logistic Regression and the Random Forest
-algorithms cannot be statistically differentiated from Boosted Trees.
-
-Next, data imputation methods were trialed which would make the complete
-data set available to Logistic Regression and Random Forests. Again, 100
-models were required to achieve a power of 0.8 with respect to Boosted
-Trees
-
-\begin{figure}[htbp]
-\centering
+The violin plot below summarises the distributions of the 100 AUC values
+produced by each method. The `violins' are coloured according to whether
+the distributions significantly vary to Boosted Trees using a critical
+value of 0.05. Boosted Trees was chosen as the comparison method because
+it was expected to perform well.
+
 \includegraphics{Consulting_Profitability_Paper_files/figure-latex/mice_violin-1.pdf}
-\caption{Violin plot vertically illustrating the distribution of AUC
-values from each of the methods when predicting profit/loss. Each method
-was fed the same imputed full dataset.}
-\end{figure}
-
-Boosted Trees, Logistic Regression, and Random Forest all performed
-significantly better than the baseline algorithm, Naive Bayes, however
-none outperformed Boosted Trees.
-
-Testing Boosted Trees, Random Forests, Naive Bayes, Logistic Regression
-and Bayesian Networks on the binary classification problem showed that
-Boosted Trees, Logistic Regression and Random Forests performed best
-(according to AUC). Naive Bayes and Logistic Regression were included as
-baseline models and it was expected that the more complex models would
-outperform these. Therefore it was surprising that results from 100
-Logistic Regression models were not statistically significantly lower
-than Boosted Trees.
-
-A possible explanation for this, according to the literature, is that
+
+The AUC distrubtions of Logistic Regression and the Random Forest
+algorithms could not be statistically differentiated from Boosted Trees,
+while the Naive Bayes distribution was significantly lower. The three
+high-performing methods had AUC values between 0.755 and 0.764. This
+indicates predictions were well above random chance, where AUC = 0.5.
+
+Naive Bayes and Logistic Regression were included as baseline models and
+it was expected that the more complex models would outperform these.
+Therefore it was surprising that results from 100 Logistic Regression
+models were not statistically significantly lower than Boosted Trees. A
+possible explanation for this, according to the literature, is that
 there was not enough data for the ensemble trees to learn the complex
@@
 Provost, and Simonoff 2003). This may explain Logistic Regression's
-comparatively high performance on the case study's small but complex
-data set.
-
-Boosted trees, Random Forests and Logistic Regression all had mean AUC
-values between 0.755 to 0.764. In summary, the individual binary
-classification models performed well above random chance (AUC = 0.5).
-Whether the model is worth implementing in the work place is dependent
-on the extent to which the algorithm would improve `bottom line' profits
-for the business and if the model can affect decisions in practice.
-
-\begin{itemize}
-\tightlist
-\item
-  NB and BN were bad - one sentence
-\end{itemize}
+comparatively high performance on the case study's data set, which is
+small for the range and complexity of the described projects.
 
@@
 
-Blended models combine the predictions from each high-performing
-individual model to create averaged or `blended' predictions. In this
-case, the Random Forest, Boosted Trees, and Logistic Regression models
-were chosen. Through variable importance studies, 6 variables were
-included as metafeatures. All six methods were compared against the
-original Logistic Regression model's AUC distribution (as it was not
-statistically different from the ensemble trees)
-
-\begin{figure}[htbp]
-\centering
+Six blended models combined the predictions from the three
+high-performing individual models: Logistic Regression, Random Forests,
+and Boosted Trees to create `blended' predictions. There were 3 simple
+blended models, and 3 complex models that interacted original covariates
+with the individual model predictions. Through variable importance
+analysis on the original covariates and the 3 new variables containing
+model predictions, only 6 original covariates were included as
+metafeatures. All six blended methods' AUC values were compared against
+the Logistic Regression AUC distribution, which was chosen because it
+was the simplest high performing individual model.
+
 \includegraphics{Consulting_Profitability_Paper_files/figure-latex/blend_violin-1.pdf}
-\caption{Violin plot vertically illustrating the distribution of AUC
-values from each of the blending methods when predicting profit/loss.
-100 models were built for each method.}
-\end{figure}
-
-The above plot shows that 4 methods had a higher mean AUC than the
-original Logistic Regression model:
-
-\begin{itemize}
-\tightlist
-\item
-  the simple blended Logistic Regression
-\item
-  the simple average
-\item
-  and the two Boosted Tree models.
-\end{itemize}
-
-The simple averaged method and simple logistic regression could achieve
-a statistical power of 0.8 with 150 samples (models), whereas the
-boosted trees required thousands. The additional 50 models were run for
-the simple logistic regression and simple average to reveal similar
-distributions that were indeed statistically significantly higher than
-the individual logistic regression AUC distribution. P-values from two
-sample t-test were 0.0019 and 0.0021 respectively.
+
+None of the blended methods were significantly higher than the
+individual Logistic Regression model with data from 100 models. However,
+the simple averaged method and simple logistic regression could achieve
+a statistical power of 0.8 with 150 samples (models), whereas the other
+blended methods required thousands. 50 more models were run for the
+simple logistic regression and simple average blend to confirm their AUC
+distributions were significantly higher than the individual Logistic
+Regression. P-values from two sample t-test were 0.0019 and 0.0021
+respectively.
 
 Blended models improved the mean AUC from 0.759 to 0.77, which is an
-increase of 1.4\%. It was expected that model blending would improve
-predictions because it combined the strengths of individual models with
-different theoretical foundations.
-
-The trials of different blending methods demonstrated that again, the
-simplest methods worked best with the case study data. In this incident,
-averaging the results of the best three single models (Logistic
-Regression, Random Forests, and Boosted Trees) or taking a Logistic
-Regression of the three models outperformed more complex Boosted Trees,
-Random Forests, and Logistic Regression blends that facilitated
-interaction of the individual model results and original variables
-(meta-features). Simple blended models could perform better than complex
-methods if the data is not big enough for the complex models to learn
-the more intricate patterns at which they excel. This was observed in
-the single model methods. their success in the 2009 Netflix competition
-generated some publications. The Netflix data set comprised of almost
-3,000,000 observations, so the size of the data could have enabled the
-success of complex blending methods such as FWLS (Sill et al. 2009)
+increase of 1.4\%. This confirms the expectation that model blending
+would improve predictions by combining the strengths of different
+methods' theoretical foundations.
+
+Again, the simplest methods worked best. In this case, averaging the
+results of the best three individual models or taking a Logistic
+Regression of the three models' outputs outperformed the complex Boosted
+Trees, Random Forests, and Logistic Regression blends that facilitated
+meta-feature interactions. Simiarly to the individual models, Simple
+methods may outperform complex methods if the data is not big enough for
+complex methods to learn the intricate patterns at which they excel. The
+Netflix data set comprised of almost 3,000,000 observations, so the size
+of the data could have enabled the success of their complex blending
+methods (Feature Weighted Linear Stacking) (Sill et al. 2009)
+
+The outcomes in this section show that statistical and machine learning
+methods can predict whether a consulting project will be profitable or
+not to some extent. The average, or weighted average (logistic
+regression) blend of individual Logistic Regression, Boosted Trees, and
+Random Forests models produced the highest AUC values. However, the
+business value of the model has not been proven. The following section
+explores whether the models would improve overall profits for the case
+study business.
 
@@
 
-This chapter first presents the full range of method results in terms of
-improvements to the case study's bottom line. A business decision making
-scenario was created so that profit curves based on the decision rule
-could be built and analysed. From the profit curves, optimal probability
-thresholds could be derived for each method (individual and blended
-methods). The predictive models output a `probability' between 0 and 1
-that each project will be a loss making job (where probability = 1
-indicates a loss making job). The question the arises, at what
-probability would a decision-maker round the probability up to 1 or down
-to 0? And what business decision would then be made? To find the
-threshold point for rounding, an experimental business-scenario was
-tested. At a given threshold, all projects with probability outputs
-above the threshold were considered too risky, and were rejected. All
-profits and losses from these projects were forfeited, while the profits
-from the remaining jobs (below the threshold) were summed to give a
-revised total profit. This profit calculation was made on a range of
-thresholds between 0 and 1 at 0.05 increments. The total profits were
-plotted for each threshold and joined to make a profit curve
-
-The plot below illustrates the distribution of profit curves for each
-method, where the solid lines join the mean values at each threshold
-point. The grey ribbon illustrates the 95\% confidence interval for the
-profit improvement ratio at each threshold point for the 100 models.
-
-\begin{figure}[htbp]
-\centering
+This chapter transports model results into the value framework to assess
+impact on the case study's overall profitability. The plot below
+illustrates the distribution of profit curves for the 100 models built
+using each method. Solid lines join the mean values at each threshold
+point and the grey ribbon represents the 95\% confidence interval for
+the profit ratio.
+
 \includegraphics{Consulting_Profitability_Paper_files/figure-latex/profit_curve-1.pdf}
-\caption{Profit curves summarising results from 100 models of 9 methods:
-3 simple blends, 3 complex blends, and the original 3 best methods.}
-\end{figure}
-
-Two blended models clearly outperformed the individual models as shown
-in the above plot with higher profit ratios as well as tighter
-confidence intervals. The simple Logistic Regression blend performed
-best with the highest mean profit ratio of 109\% with a standard
-deviation of 1.28\%. \%. This means that for the simple blended Logistic
-Regression model, if all jobs above the probability threshold 0.6 were
-rejected, the profits would increase on average by 9\% in comparison to
-historical profits.
-
-Profit curves from the simple average blend and Logistic Regression
-blended models outperformed the complex blended models, which follows
-logically from their significantly higher AUC distributions. The simple
-average blended model produced a profit curve with an almost identical
-profit improvement (and standard deviation) to the Logistic Regression
-blend. In the data, projects assigned a probability higher than 0.6
-represent only 4.3\% of all projects.
-
-The shaded grey 95\% confidence intervals in the Profit Curve plot shows
-that with the 100 sample models that were analysed, none of the lower
-bounds for the original methods' or complex blending methods were above
-100\% on the y-axis. That is, it cannot be said with 95\% confidence
-that the original methods would produce a mean profit higher than
-historical profits in the given decision framework. Clearly this level
-of certainty is not beneficial for the case-study business to adopt.
-
-It is not clear why simple linear models and the averaging model
-outperformed ensemble tree blending methods. As previously stated, the
-ensemble trees may not have received enough data to adequately learn the
-complex series of rules they develop. The only differences between FWLS
-and the simple Logistic Regression (that performed best) were two
-additional explanatory variables and four interaction terms. The number
-of variables was not high for the amount of data since according to
-Peduzzi et al. (1996), 10 events per explanatory variable or more avoids
-the risk of biased estimation of variable coefficients in Logistic
-Regression. The data contained 315 events per explanatory variable and
-did not pose that risk. Nevertheless, the additional variables must have
-added misleading noise to the model.
-
-To conclude, because of the promising AUC results, it was logical that
-the models translated into financial benefits for the company. The 9\%
-improvement in profits produced by the simple blended Logistic
-Regression is reasonable, and may be high enough to trigger further
-cost-benefit analyses and the development of a more comprehensive
-framework describing different decision scenarios.
-
-\begin{itemize}
-\item
-  decision scenario
-\item
-  profit curves
-\item
-  prototype decision support interface
-\end{itemize}
+
+From the profit curves, the optimal probability thresholds are where the
+profit curves peak, and peaks above 100\% indicate the model assisted
+profits outperformed historical profits. The two simple blended models
+clearly outperformed all other models with higher profit ratios as well
+as tighter confidence intervals. The simple Logistic Regression blend
+performed best with the highest mean profit ratio of 109\% and a
+standard deviation of 1.28\%. This means that if all jobs above the
+probability threshold 0.6 were rejected, the profits would increase on
+average by 9\% in comparison to historical profits. Only 4.3\% of
+projects were rejected in this case.
+
+Profit curves from the simple blended models outperformed the complex
+blended models, which follows logically from their significantly higher
+AUC distributions. The shaded grey 95\% confidence intervals show that
+none of the lower bounds for the original methods' or complex blending
+methods were above 100\% on the y-axis. That is, it cannot be said with
+95\% confidence that these methods would produce a mean profit higher
+than historical profits in the given decision framework. It is
+interesting that despite the reasonable AUC values, the profit curves
+for the individual and complex blended models do not make a complelling
+case for adoption into business practice.
+
+The simple blended Logistic Regression model on the other hand, a mean
+9\% improvement in profits, with a lower bound of 6.5\%. In terms of
+implementing the model, this improvement may be high enough to trigger
+further development and analysis of a more sophisticated framework
+describing decision making scenarios.
 
@@
 
-The general aim was to use statistical techniques to predict the
-profitability of projects for a case study consulting business using
-their internal CRM data. This was rigorously completed by approaching
-the prediction of `profitability' as a binary classification problem
-predicting either profit or loss. Several statistical and machine
-learning approaches were applied including Naive Bayes, Bayesian
-Networks, Linear Regression, Random Forests, gradient Boosted Trees as
-well as multiple methods of blending the individual models' output.
-
-AUC = 0.76). This was achieved by 3 individual methods while various
-techniques that blended the individual methods improved results further
-(AUC = 0.77).
-
-predictive models could be shown to improve the overall profitability
-(bottom line) of the case study business. A range of probability
-threshold values were trialed for each method using a decision framework
-where projects scored below the threshold were accepted, while projects
-above the threshold were rejected. If a project was rejected, the
-profits and losses were forfeited, and the remaining accepted profits
-and losses were summed.
-
-Final results showed the simple Logistic Regression blend of the
-individual Logistic Regression, Random Forest and Boosted Tree models
-improved profits the most. The 95\% confidence interval for these
+Complex ensemble tree methods did not outpereform simple linear methods,
+however, the simple blended predictions of all methods
+
+The general aim of this work was twofold. First, statistical and machine
+learning techniques were tested to determine whether the profitability
+of projects for a case study consulting business could be predicted
+using their internal CRM data. This was rigorously completed by
+approaching the prediction of `profitability' as a binary classification
+problem with classes profititable or loss-making. Naive Bayes, Logistic
+Regression, Random Forests, gradient Boosted Trees as well as simple and
+complex blending methods were applied to the problem. Secondly, this
+project tested whether predictions from the models could be proven to
+improve the overall profitability of the case study company and by how
+much.
+
+Predicting whether a job would be profitable or not (binary
+classification) was possible to a level significantly better than random
+assignment (AUC = 0.76). This mean statistic was achieved by 100
+different training/testing divisions of the data, built for each of the
+3 individual methods. Simple averaging and simple Logistic Regression
+blending methods improved results further to a mean of AUC = 0.77.
+
+Two models could also be shown to improve the overall profitability of
+the case study business. This was done by converting the binary
+predictions from the first analysis into profit curves that plotted
+model-influenced profit improvements over a range of probability
+threshold values. To determine model-influenced profit improvements, a
+managerial decision framework was used where projects scored below the
+threshold were accepted, while projects above the threshold were
+rejected. If a project was rejected, the profits and losses were
+forfeited, and the remaining accepted profits and losses were summed.
+
+100 training/testing divisions of the data were user to produce 100
+profit curves so that a 95\% confidence interval could be framed around
+the curves. Final results showed the simple Logistic Regression blend of
+the individual Logistic Regression, Random Forest and Boosted Tree
+models improved profits the most. The 95\% confidence interval for these
 improvements was between 6.5\% and 11.5\% using a probability threshold
-of 0.6 (approximately 4.3\% of projects).
+of 0.6 (approximately 4.3\% of projects were rejected).
 
 These results contribute significantly to the research in cost
-estimation in three ways: the applied methods, the decision framework,
-and the appeal to user trust. Ensemble tree methods and blending had
-been applied minimally to cost estimation previously, even though
-ensemble trees provide insight into model structure while predicting at
-a similar level to Neural Networks. Next, previous studies have verified
-predictive accuracy but stopped short of how the algorithm would affect
-decisions and what the measured benefits would be. This study presented
-a clear framework for how a business could improve profits by applying
-the algorithm.
-
-Further work is required to test user confidence in the output. Another
-topic identified for future research was to test how well managers
-estimate some numeric project input variables. In particular, time span,
-team size, total invoiced amount, and percentage of hours completed by
-professionals should be tested as these variables were calculated post
-project completion. Time span and total invoiced amount were discretised
-into wide categories, which should be easier for a manager to choose
-between.
-
-future work: user interface.
+estimation in two ways: the applied methods, and the value assessment
+based on a managerial decision framework. Ensemble tree methods and
+blending had been applied minimally to cost estimation previously, even
+though ensemble trees provide insight into model structure while
+predicting at a similar level to Neural Networks. Next, previous studies
+compared accuracy of their predictive models, but stopped short of
+computing how the algorithm would affect decisions and what the measured
+benefits would be. This study presented a clear framework for how a
+business could improve profits by applying the algorithm.
+
+Further work is required to measure user confidence in the output.
+Another topic identified for future research is to test how well
+managers estimate input variables at project conception. This is
+paraticularly relevant for time span and total invoiced amount which
+were discretised into 4 to 5 categories.
 
@@
 Logistic Regression, Random Forests, and Boosted Tree models, from a
-consulting company's internal project data. This blended model can
-predict whether a project will be profitable or not and in a reasonable
-decision framework, can guide managers in rejecting financially risky
-projects and improving profitability of the business.
+consulting company's internal project data. This blended model predicts
+whether a project will be profitable or not and in a reasonable decision
+framework, can guide managers in rejecting financially risky projects
+and improving profitability of the business by a mean of 9\%.
 
@@
 This research was supported by a scholarship/ ACEMS?
+
+\section{Figure Captions}\label{figure-captions}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  mice\_violin: Violin plot vertically illustrating the distribution of
+  AUC values from each of the methods when predicting profit/loss. Each
+  method was fed the same imputed full dataset.
+\item
+  blend\_violin: Violin plot vertically illustrating the distribution of
+  AUC values from each of the blending methods when predicting
+  profit/loss. 100 models were built for each method.
+\item
+  profit\_curve: Profit curves summarising results from 100 models of 9
+  methods: 3 simple blends, 3 complex blends, and the original 3 best
+  methods.
+\end{enumerate}
 
@@
 Prediction, Second Edition}. Book. Springer New York.
-\url{http://qut.summon.serialssolutions.com/2.0.0/link/0/eLvHCXMwzV1bS8MwFA7iQBwD3dR6mRAQfNk6trb0Ivi0TQY6EJwgQxi9JLKHddpVf7_nJE269cFnH9NCm_Y7JF--cyPEtnp9s7Im8JCzAWxVjFluEnMWBF4MyyfzE4-HVsR29Tfd9rO89s-BrzBiPSwRH8tocBGfgTxSlGFGrbpQQFAAGIV52JmKxhByhShS_lQg53OGzhsR9IHSKJ6YsfDxcttlP8vYzzrrTEJ4vjYNGaUN8EcoeYfpUofnMCyKACQZ-O6qMEelNAQVpUEpjR1VUGrnKIpucN_BcvnlzqK86ZUNR4cBwrbpes4tFjlf4Yfds9R8fYH9dAA8FGnufK71MnR_ArkqHO7iTW5RQEm_uU4aMEPxsA9subLFFGbHpMYwfaRJ9ljaIkeqaQYt1tAWqU91odzNCXkH0KgCja453QKNKtDuKEJGJWRdqgHrUoCLlnB1qQSLFmCdkpuH8Ww4MXemu_iUBUYW8r_YZ6QRYppDmot0yMQgNQ6myQzkCQZ8l0EO3oKnkT95HMphUw17G5Gz1_vKDaAlwrJNt-edE2pb3PEiL-aRkzh-AKfFqO8nPse2BbFt9S9I-69JXf59-4oclpbTJvt59s2uRQ7tL6t_R8w}.
 
@@
 SCHOOL PUBLISHING CORPORATION.
-\url{http://qut.summon.serialssolutions.com/2.0.0/link/0/eLvHCXMw3V3Pb9MwFLYQQxMSEgxYKAPJBwQHlCqO61-TdoC2aNImcdgmcYscxwGktoMlQfvzeXYSJy1M4swxjZuq_uz3vs95PxCi6TSJd2yCAU-T53pmjbIgoblMCqGYyEtpEpoWs-3zt9Bqb_jsfwB-YVdN1Ye3XTS-IaLT_a553GcwEGvXFsO3O1q7kPf3y1trGmf0fDzFouu5U41Za99AKETJb79ROL_-pbtXOIthsZ3pbxvbna-2iexbJww0RKMGq9l5srHVlGS0OsTIBDI-cqakTczcrnO9439CVCCIZQ6ml751Vc_XxXdTn9hNfHUBDhZ4q1cFH5d3utk_feqIVDxwOT5N9Vd24ZnE5RP0uC_pjT-02B2ge3bzFO33k_sMzQOE-LrEHYTHGADEPYB4ABAPAL7DAb7n6OrT8nJ-GnfNLuKvBChkzFTJcgnkVadGg1XkVllKiXYEDHYLYxKUK3imPCfGklIwA8KxYJQpSsAaSHqIHmmXFLGpffJkEaG9EhayjRyriOBfRmj_izpfyNOzeXt50F9OK5_hN_1ZRzC7fh_EfCpeICxEIg2fcebqtoHtzlOic26UYobKWZFMUOQmPXObor7RJiMJT0BSJGqCDnscsmK1yqgrlQfklMBXWlSyH209lQwoEgPdDnfetDCFO2lWpRmoVBDJqRBOq9a3NTxgZ1i_cOABY3jDfSdkYIhgzEueCSL_MmzeFcl3xSHql3f-6BF6OGyaV-h-fdPY1z5t-DdLboxJ}.
 
@@
 
-\hypertarget{ref-e1071}{}
-Meyer, David, Evgenia Dimitriadou, Kurt Hornik, Andreas Weingessel, and
-Friedrich Leisch. 2014. \emph{E1071: Misc Functions of the Department of
-Statistics (E1071), Tu Wien}.
-\url{https://CRAN.R-project.org/package=e1071}.
-
 \hypertarget{ref-Moore1989}{}
@@
 Article. \emph{Journal of Computer Information Systems} 53 (4): 49--58.
-\url{http://qut.summon.serialssolutions.com/2.0.0/link/0/eLvHCXMwnV1Nb9QwEB0VekFClPLVAJV8gGPAH4kbc4GlyraIbhftpurRysTe06os3VTqz8fjTbawBanqKZHiOIntvBmPPe8BKPmBpxuYgMHsmgALqtHauDrYJRTaZNlB4b3XiBvxty99pGDV2z1IRuR2PxsKmn8UAUm1IcKwz4tfKclI0XJrp6nxALaJVoSG_Ej-WC8rFHnUE4x_Fmlv_e1WUh7I1fKfFiham-EO9Nuf-tjX_FZywW02x3t_xFN40nmlbLAaRruw5S-ewU6v-MA6AHgOn6bjYXU-mJQs-LfjScXKafVtFMNcjAQ8jtiAnZZnk8FJOFTn48l3Vp5Oy9HXk_IFnA3L6vA47dQX0gWlFaWNwjwjOjfDZ43hssFaovNCu7wIZwpnJrhmuSPtK2xE0QhZ68IJn3PnMzxQL-FxTbv0L9qYzef2gKkw-eJZ1BnFzGcKQ_Ua3SxUyRELTGC_byZbI8V6mnZpbxopgVfr624-t2G-QynmSooE3q-6zy5WLB1W2qW03BYBoSTdrbltr9sE9jbKkaChKLgyCbz7s-PXBWjaGJ6guInGPQFxl2KHHcU6UQu0r__73m_gkYwCG7QB-C08bC-v_H5MOv0NwrD2yw}.
-
-\hypertarget{ref-Peduzzi1996}{}
-Peduzzi, Peter, John Concato, Elizabeth Kemper, Theodore R Holford, and
-Alvan R Feinstein. 1996. ``A Simulation Study of the Number of Events
-Per Variable in Logistic Regression Analysis.'' Journal Article.
-\emph{Journal of Clinical Epidemiology} 49 (12): 1373--9.
 
@@
 Provost, F., and T. Fawcett. 2013. \emph{Data Science for Business}.
-Book. O'Reilly. \url{https://books.google.com.au/books?id=_1b4nAEACAAJ}.
-
-\hypertarget{ref-gbm}{}
-Ridgeway, Greg. 2015. \emph{Gbm: Generalized Boosted Regression Models}.
-\url{https://CRAN.R-project.org/package=gbm}.
+Book. O'Reilly.
 

### 2016-12-21 finalwriteup/Consulting Profitability Paper/Consulting Profitability Paper.Rmd
new title

@@
 ---
-title: Predictive Models to Support Quoting of Fixed Fee Consulting Projects
+title: Predictive Framework to Support Managerial Decisions around Fixed Fee Consulting Projects
 author:
@@
 
-  Engaging in loss making jobs for fixed fees is a major problem in consulting, particularly in the competitive construction industry.
-This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs or help tactfully choose when to enforce stricter contracts.
-It was found that in a specific decision framework, a case study's profits could be improved 9% by declining approximately 4% of projects.
-Alternative decision frameworks are also proposed and evaluated.
-Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods.
-A decision scenario which rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements.
-The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% profit improvements.
-The findings from this research have the potential to assist managers in reducing losses by highlighting risky projects and guiding project-based changes to fee structures.
+  Fixed fee projects that run overbudget are a major problem for consulting businesses, particularly in competitive industries such as the construction industry.
+This thesis investigates whether machine learning algorithms trained on a company's passively collected internal data could predict whether a job would be profitable or loss making.
+Logistic Regression, Random Forests, Boosted Trees, and Naive Bayes were applied as well as blended combinations of these methods.
+The aim of these models is to help managers determine when to enforce certain contract structures or when to not engage in a project in order to avoid losses.
+Therefore this paper also examines whether the predictive algorithm improves a case study's overall profits if implementing the algorithm within a simple managerial decision framework.
+The framework was to reject all projects predicted as loss making, where loss making projects were defined as projects with a probability of being loss making above a certain threshold.
+The Logistic Regression model that blended predictive output from a Logistic Regression, Random Forest, and Boosted Tree models performed best and produced a 95% confidence interval of 6.5 - 11.5% profit improvements.
+This improvement was achieved by rejecting approximately 4% of projects that were deemed high risk.
+The findings from this research have the potential to assist managers in reducing consulting project losses by highlighting which projects to either avoid or accept given a fee structure that minimises financial risk.
+
 keywords: |
@@
 
-From the profit curves, the optimal probability thresholds are where the profit curves peak, and peaks above 100% indicate the model assisted profits outperformed historical profits.
+From the profit curves, the optimal probability thresholds are where the profit curves peak, and peaks above 100% indicate the model-assisted profits outperformed historical profits.
 The two simple blended models clearly outperformed all other models with higher profit ratios as well as tighter confidence intervals.
@@
 Only 4.3% of projects were rejected in this case.
+
+| Optimal Threshold | Method               | Mean profit ratio  (%) | SD profit ratio  (%) | Lower bound 95% CI (%) | Upper bound 95% CI (%) |
+|-----------|----------------------|------------------------|----------------------|------------------------|------------------------|
+| 0.60      | Simple LR blend      | 109.0                  | 1.28                 | 106.5                  | 111.5                  |
+| 0.55      | Simple average blend | 108.7                  | 1.56                 | 105.6                  | 111.8                  |
+| 0.55      | Simple BT blend      | 105.6                  | 3.21                 | 99.3                   | 111.9                  |
+| 0.50      | Complex BT blend     | 105.4                  | 3.23                 | 99.1                   | 111.7                  |
+| 0.60      | Individual RF        | 103.2                  | 2.58                 | 98.1                   | 108.3                  |
+| 0.60      | Individual BT        | 103.2                  | 2.77                 | 97.8                   | 108.6                  |
+| 0.75      | Individual LR        | 102.5                  | 2.93                 | 96.8                   | 108.2                  |
+| 0.65      | Complex RF blend     | 101.3                  | 2.72                 | 96.0                   | 106.6                  |
+| 0.65      | Complex LR blend     | 101.0                  | 1.18                 | 98.7                   | 103.3                  |
 
@@
 
+Table Captions
+===============
+1. Values and statistics describing each individual model and blended model's profit curves at the optimal position in the curve.
+
 References {#references .unnumbered}

### 2016-12-21 finalwriteup/Consulting Profitability Paper/Consulting_Profitability_Paper.tex
new title

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{longtable}
 \usepackage{graphicx}
@@
 
-  Engaging in loss making jobs for fixed fees is a major problem in
-  consulting, particularly in the competitive construction industry. This
-  thesis investigates whether machine learning techniques applied to a
-  company's passively collected internal data could help avoid loss making
-  jobs or help tactfully choose when to enforce stricter contracts. It was
-  found that in a specific decision framework, a case study's profits
-  could be improved 9\% by declining approximately 4\% of projects.
-  Alternative decision frameworks are also proposed and evaluated.
-  Algorithmic methods such as Logistic Regression, Random Forests, Boosted
-  Trees, Naive Bayes, and Bayesian Networks were applied as well as
-  blended combinations of these methods. A decision scenario which
-  rejected projects above a sequence of tested thresholds was run in order
-  to find the optimal threshold for profit improvements. The blended
-  Logistic Regression model outperformed other methods and produced a 95\%
-  confidence interval of 6.5 - 11.5\% profit improvements. The findings
-  from this research have the potential to assist managers in reducing
-  losses by highlighting risky projects and guiding project-based changes
-  to fee structures.
+  Fixed fee projects that run overbudget are a major problem for
+  consulting businesses, particularly in competitive industries such as
+  the construction industry. This thesis investigates whether machine
+  learning algorithms trained on a company's passively collected internal
+  data could predict whether a job would be profitable or loss making.
+  Logistic Regression, Random Forests, Boosted Trees, and Naive Bayes were
+  applied as well as blended combinations of these methods. The aim of
+  these models is to help managers determine when to enforce certain
+  contract structures or when to not engage in a project in order to avoid
+  losses. Therefore this paper also examines whether the predictive
+  algorithm improves a case study's overall profits if implementing the
+  algorithm within a simple managerial decision framework. The framework
+  was to reject all projects predicted as loss making, where loss making
+  projects were defined as projects with a probability of being loss
+  making above a certain threshold. The Logistic Regression model that
+  blended predictive output from a Logistic Regression, Random Forest, and
+  Boosted Tree models performed best and produced a 95\% confidence
+  interval of 6.5 - 11.5\% profit improvements. This improvement was
+  achieved by rejecting approximately 4\% of projects that were deemed
+  high risk. The findings from this research have the potential to assist
+  managers in reducing consulting project losses by highlighting which
+  projects to either avoid or accept given a fee structure that minimises
+  financial risk.
   \end{abstract}
@@
 From the profit curves, the optimal probability thresholds are where the
-profit curves peak, and peaks above 100\% indicate the model assisted
+profit curves peak, and peaks above 100\% indicate the model-assisted
 profits outperformed historical profits. The two simple blended models
@@
 projects were rejected in this case.
+
+\begin{longtable}[]{@{}llllll@{}}
+\toprule
+\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
+Optimal Threshold\strut
+\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
+Method\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Mean profit ratio (\%)\strut
+\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
+SD profit ratio (\%)\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Lower bound 95\% CI (\%)\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Upper bound 95\% CI (\%)\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.60\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Simple LR blend\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+109.0\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+1.28\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+106.5\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+111.5\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.55\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Simple average blend\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+108.7\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+1.56\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+105.6\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+111.8\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.55\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Simple BT blend\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+105.6\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+3.21\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+99.3\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+111.9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.50\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Complex BT blend\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+105.4\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+3.23\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+99.1\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+111.7\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.60\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Individual RF\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+103.2\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+2.58\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+98.1\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+108.3\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.60\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Individual BT\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+103.2\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+2.77\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+97.8\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+108.6\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.75\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Individual LR\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+102.5\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+2.93\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+96.8\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+108.2\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.65\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Complex RF blend\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+101.3\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+2.72\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+96.0\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+106.6\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
+0.65\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+Complex LR blend\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+101.0\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+1.18\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+98.7\strut
+\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+103.3\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
+\section{Table Captions}\label{table-captions}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Values and statistics describing each individual model and blended
+  model's profit curves at the optimal position in the curve.
+\end{enumerate}
+
 \section*{References}\label{references}
